### 2021-08-25 paper/main.tex
started filling in paper shell

% bare_jrnl_compsoc, texV1.4b, 2015/08/26, Michael Shell

\documentclass[10pt,journal,compsoc]{IEEEtran}

% *** MISC UTILITY PACKAGES ***

% *** CITATION PACKAGES ***
\ifCLASSOPTIONcompsoc
  % IEEE Computer Society needs nocompress option
  % requires cite.sty v4.0 or later (November 2003)
  \usepackage[nocompress]{cite}
\else
  % normal IEEE
  \usepackage{cite}
\fi

% *** GRAPHICS RELATED PACKAGES ***
\ifCLASSINFOpdf
   \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
   \graphicspath{{./figures/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
   \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
   \graphicspath{{./figures/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
   \DeclareGraphicsExtensions{.eps}
\fi

% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.

% *** MATH PACKAGES ***
\usepackage{amsmath}

% *** SPECIALIZED LIST PACKAGES ***
\usepackage{algorithmic}

% *** ALIGNMENT PACKAGES ***
\usepackage{array}
% IEEEtran contains the IEEEeqnarray family of commands

% *** SUBFIGURE PACKAGES ***
\ifCLASSOPTIONcompsoc
  \usepackage[caption=false,font=footnotesize,labelfont=sf,textfont=sf]{subfig}
\else
  \usepackage[caption=false,font=footnotesize]{subfig}
\fi

% *** FLOAT PACKAGES ***
\usepackage{dblfloatfix}

% *** PDF, URL AND HYPERLINK PACKAGES ***
\usepackage{url}

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
\title{Topological Equivariant Artist Model for Visualization Library Architecture}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc 
% journal papers have the author affiliations above the "Manuscript
% received ..."  text while in non-compsoc journals this is reversed. Sigh.
%~\IEEEmembership{Member,~IEEE,}
\author{Hannah~Aizenman, Thomas~Caswell, and~Michael~Grossberg% <-this % stops a space
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem H. Aizenman and M. Grossberg are with the department of Computer Science, City College of New York. 
\protect\\
% note need leading \protect in front of \\ to get a newline within \thanks as
% \\ is fragile and will error, could use \hfil\break instead.
E-mail: haizenman@ccny.cuny.edu, mgrossberg@ccny.cuny.edu 
\IEEEcompsocthanksitem Thomas Caswell is with National Synchrotron Light Source II, Brookhaven National Lab 
\protect \\
E-mail: tcaswell@bnl.gov}% <-this % stops an unwanted space
\thanks{Manuscript received X XX, XXXX; revised X XX, XXXX.}
}

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}
The abstract goes here.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
%Computer Society, IEEE, IEEEtran, journal, \LaTeX, paper, template.
\end{IEEEkeywords}}

% make the title area
\maketitle

% To allow for easy dual compilation without having to reenter the
% abstract/keywords data, the \IEEEtitleabstractindextext text will
% not be used in maketitle, but will appear (i.e., to be "transported")
% here as \IEEEdisplaynontitleabstractindextext when the compsoc 
% or transmag modes are not selected <OR> if conference mode is selected 
% - because all conference papers position the abstract like regular
% papers do.
\IEEEdisplaynontitleabstractindextext
% \IEEEdisplaynontitleabstractindextext has no effect when using
% compsoc or transmag under a non-conference mode.

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
% Computer Society journal (but not conference!) papers do something unusual
% with the very first section heading (almost always called "Introduction").
% They place it ABOVE the main text! IEEEtran.cls does not automatically do
% this for you, but you can achieve this effect with the provided
% \IEEEraisesectionheading{} command. Note the need to keep any \label that
% is to refer to the section immediately after \section in the above as
% \IEEEraisesectionheading puts \section within a raised box.

% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps (small caps for compsoc).
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by the IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{T}{his} demo file is intended to serve as a ``starter file''
for IEEE Computer Society journal papers produced under \LaTeX\ using
IEEEtran.cls version 1.8b and later.
% You must have at least 2 lines in the paragraph with the drop letter
% (should never be an issue)
I wish you the best of success.

\hfill mds

\hfill August 26, 2015

\subsection{Subsection Heading Here}
Subsection text here.

% needed in second column of first page if using \IEEEpubid
%\IEEEpubidadjcol

\subsubsection{Subsubsection Heading Here}
Subsubsection text here.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that the IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.
% However, the Computer Society has been known to put floats at the bottom.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively. 
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

\section{Conclusion}
The conclusion goes here.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

\appendices
\section{Proof of the First Zonklar Equation}
Appendix one text goes here.

% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
\section{}
Appendix two text goes here.

% use section* for acknowledgment
\ifCLASSOPTIONcompsoc
  % The Computer Society usually uses the plural form
  \section*{Acknowledgments}
\else
  % regular IEEE prefers the singular form
  \section*{Acknowledgment}
\fi

The authors would like to thank...

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{bibliography}

% biography section 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

%\begin{IEEEbiography}{Michael Shell}
%\end{IEEEbiography}

% if you will not have a photo at all:
\begin{IEEEbiographynophoto}{Hannah Aizenman}
Biography text here.
\end{IEEEbiographynophoto

\begin{IEEEbiographynophoto}{Thomas Caswell}
  Biography text here.
\end{IEEEbiographynophoto}
% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

\begin{IEEEbiographynophoto}{Michael Grossberg}
Biography text here.
\end{IEEEbiographynophoto}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}

### 2021-08-26 matplottoy/lib/bar.py
bar and barv

@@ -1,0 +1,162 @@
+from __future__ import annotations
+from collections import namedtuple
+from collections.abc import Iterable
+import copy
+from typing import List, Dict, Tuple, Optional, Any, Callable, Union
+
+
+import numpy as np 
+
+from matplotlib import rcParams
+
+
+import matplotlib.artist as martist
+import matplotlib.cbook as cbook
+import matplotlib.path as mpath
+import matplotlib.transforms as mtransforms
+
+from matplottoy.artists import utils
+from matplottoy.encoders import mtypes
+
+
+# mu = \nu \circ \tau 
+#\tau = data.query().projection()
+Mu = namedtuple('Mu', ['nu_i', 'F_i'])
+Vspec= Dict[str, Tuple[Tuple[str,...], Mu]]
+# Spivak's records = sections evaluated on U \subset K
+Records = Dict[str, Union[Iterable, Callable]]
+
+
+class Rho: 
+    # rho: S \rightarrow H
+    # rho = \q \circ \xi^*(\mu) =? \qhat \circ \mu (over^k)
+    def mu(self, tau_restricted: Records):
+        pass
+    @staticmethod
+    def qhat():
+        pass
+
+class TopologicalArtist: 
+# also maybe should be owned by data library? 
+#is bad name for this since this is more than just qhat
+    def __init__(self, graphic:Rho):
+        # is parameterized by the rho
+        # only input is \tau (E)
+        self.graphic = graphic
+        self.data = None
+
+    #section has to be ducktyped w/ a query method
+    def compose_with_tau(self, section): #doesn't need to be there
+        # do I care about bounds yet? or only at draw time?
+        new = copy.copy(self)
+        new.data = section
+        return new
+
+    def query(self, data_bounds:Optional[dict]=None, sampling_rate:Optional[int]=None)->dict:
+        return self.data.query(data_bounds, sampling_rate)
+        # section.view, #xi seperates nicely 
+
+def barH(V):
+    new_key = {'position': 'x', 'height':'y'}
+    return Bar({(new_key[k] if k in new_key else k):v for (k, v) in V.items()})
+
+def barV(V):
+    new_key = {'position': 'y', 'height':'x'}
+    return BarV({(new_key[k] if k in new_key else k):v for (k, v) in V.items()})
+
+class Bar(Rho): 
+    #what are the 
+    P_required = {'x', 'y'}
[…]

### 2021-08-26 matplottoy/lib/color.py

@@ -1,0 +1,28 @@
+import numpy as np
+
+import matplotlib.colors as mcolors
+
+from matplottoy.encoders import mtypes
+
+class Identity:
+    @staticmethod
+    def validate(mtype):
+        return mtype in ['nominal']
+        
+    @staticmethod
+    def __call__(value):
+        values = np.atleast_1d(np.array(value, dtype=object))
+        return [mtypes.RGBA(*mcolors.to_rgba(v)) for v in values]
+
+class Nominal:
+    def __init__(self, mapping):
+        """goal of init is to store parameters that would otherwise be
+        curried higher level function"""
+        assert(mcolors.is_color_like(color) for color in mapping.values())
+        self._mapping = mapping
+
+    def __call__(self, value):
+        values = np.atleast_1d(value)
+        return [mtypes.RGBA(*mcolors.to_rgba(self._mapping[v])) for v in values]
+
+

### 2021-08-26 matplottoy/lib/datawrap.py

@@ -1,0 +1,21 @@
+from dataclasses import dataclass
+from typing import Tuple, Optional, List
+
+import numpy as np
+import pandas as pd
+
+# index is like the sheaf, stalk is either the fiber or the values? 
+# check the applied category theory book 
+# index is subset on K
+# technically this can be wrapped even more
+# 
+# maybe make a Record type Record = {Fiber_Name: values} 
+class DataFrameWrapper:
+    def __init__(self, section: pd.DataFrame):
+        self.global_section = section
+
+    def query(self, data_bounds:Optional[dict] = None, sampling_rate:int=None)->List[dict]:
+        local_section = self.global_section
+        return [{Fi:local_section[Fi].values.squeeze() for Fi in local_section.columns}]
+
+    

### 2021-08-26 matplottoy/lib/position.py

@@ -1,0 +1,34 @@
+import functools
+
+import numpy as np 
+#taus
+
+#TODO: write atleast_1d wrapper
+class Identity:
+    @staticmethod
+    def __call__(value):
+        return np.array(value)
+    @staticmethod
+    def validate(mtype):
+        return mtype in ['nominal', 'ordinal', 'interval', 'ratio']
+
+class Log:
+    def __init__(self, base=10):
+        self.base = base
+
+    def __call__(self, value):
+        return np.log(value)/np.log(self.base)
+
+    def validate(self, mtype):
+        return True
+    
+class Nominal:
+    def __init__(self, mapping):
+        self._mapping = mapping
+        
+    def __call__(self, value):
+        values = np.atleast_1d(np.array(value, dtype=object))
+        return np.array([self._mapping[v] for v in values])
+
+    def validate(self, mtype):
+        return True

### 2021-08-26 matplottoy/lib/bar.py
inverses

@@ -2,6 +2,7 @@ from __future__ import annotations
 from collections import namedtuple
 from collections.abc import Iterable
 import copy
+from dataclasses import dataclass
 from typing import List, Dict, Tuple, Optional, Any, Callable, Union
 
 
@@ -21,17 +22,28 @@ from matplottoy.encoders import mtypes
 
 # mu = \nu \circ \tau 
 #\tau = data.query().projection()
-Mu = namedtuple('Mu', ['nu_i', 'F_i'])
-Vspec= Dict[str, Tuple[Tuple[str,...], Mu]]
+
+@dataclass
+class FtoV:
+    nu: Callable
+    f: Union[Tuple[str, ...], None] = None
+    nu_inv: Union[Callable, None] = None
+
+Vspec= Dict[str, Tuple[Tuple[str,...], FtoV]]
 # Spivak's records = sections evaluated on U \subset K
 Records = Dict[str, Union[Iterable, Callable]]
-
 
 class Rho: 
     # rho: S \rightarrow H
     # rho = \q \circ \xi^*(\mu) =? \qhat \circ \mu (over^k)
+    
+    def __init__(self, V:Optional[Vspec]=None):
+        # F->V maps need to be fixed here, but use fiber name instead of function
+        self.V = {} if V is None else V
+
     def mu(self, tau_restricted: Records):
         pass
+    
     @staticmethod
     def qhat():
         pass
@@ -70,37 +82,31 @@ class Bar(Rho):
     P_optional = {'floor', 'width', 'facecolor', 'edgecolor', 
                   'linewidth', 'linestyle'}
 
-    def __init__(self, V:Optional[Vspec]=None):
-        # F->V maps need to be fixed here, but use fiber name instead of function
-        self.V = {} if V is None else V
-
     # Spivak r: C \rightarrow U_{\sigma}, U_{sigma}=F_{i}=F_{c} 
-    def compose_with_nu(self, pi:str, fi: Tuple[str,...], nu: callable)->Bar:
+    def compose_with_nu(self, p:str, f:Tuple[str,...], nu:callable, nu_inv:callable=None)->Bar:
         #trying to sort out how to not lose existing specs
         # also maybe type checking here? (Is this nu valid for this fiber?)
         new = copy.copy(self)
         # user needs to do nu_i = nu_i_j \circ nu_i_k composition
-        new.V[pi] = Mu(nu_i=nu, F_i=fi) 
[…]

### 2021-08-26 matplottoy/lib/color.py

@@ -20,9 +20,12 @@ class Nominal:
         curried higher level function"""
         assert(mcolors.is_color_like(color) for color in mapping.values())
         self._mapping = mapping
+        self._inverse = {v:k for k, v in mapping.items()}
 
     def __call__(self, value):
         values = np.atleast_1d(value)
         return [mtypes.RGBA(*mcolors.to_rgba(self._mapping[v])) for v in values]
 
+    def inverse(self, value):
+        return self._inverse[value]
 

### 2021-08-26 matplottoy/lib/datawrap.py

@@ -16,6 +16,6 @@ class DataFrameWrapper:
 
     def query(self, data_bounds:Optional[dict] = None, sampling_rate:int=None)->List[dict]:
         local_section = self.global_section
-        return [{Fi:local_section[Fi].values.squeeze() for Fi in local_section.columns}]
-
-    
+        records = {Fi:local_section[Fi].values.squeeze() for Fi in local_section.columns}
+        records['k'] = local_section.index.values
+        return [records]

### 2021-08-26 matplottoy/lib/position.py

@@ -8,6 +8,7 @@ class Identity:
     @staticmethod
     def __call__(value):
         return np.array(value)
+
     @staticmethod
     def validate(mtype):
         return mtype in ['nominal', 'ordinal', 'interval', 'ratio']
@@ -25,10 +26,16 @@ class Log:
 class Nominal:
     def __init__(self, mapping):
         self._mapping = mapping
+        self._inverse = {v:k for k, v in mapping.items()}
         
     def __call__(self, value):
         values = np.atleast_1d(np.array(value, dtype=object))
         return np.array([self._mapping[v] for v in values])
 
+    def inverse(self, value):
+        return self._inverse[value]
+
     def validate(self, mtype):
         return True
+
+    

### 2021-08-27 matplottoy/lib/bar.py
notes from thursday meeting

@@ -19,6 +19,8 @@ import matplotlib.transforms as mtransforms
 from matplottoy.artists import utils
 from matplottoy.encoders import mtypes
 
+# axes artist, line artist, write a fake xlim/ylim 
+# matplotlib can roundtrip V->H, H->V (transforms)
 
 # mu = \nu \circ \tau 
 #\tau = data.query().projection()
@@ -111,6 +113,7 @@ class Bar(Rho):
 
     @staticmethod
     def qhat(x, width, y, floor, facecolor, edgecolor, linewidth, linestyle):
+        # are matplotlib 
         def fake_draw(render, transform=mtransforms.IdentityTransform()):
             for (xi, wd, yi, fl, fc, ec, lw, ls) in zip(x, width, y, floor, facecolor, edgecolor, linewidth, linestyle):
                 gc = render.new_gc()
@@ -137,19 +140,63 @@ class GenericArtist(martist.Artist):
     def __init__(self, artist:TopologicalArtist):
         super().__init__()
         self.artist = artist
+        self.artist.V['x'] # update w/ transforms ax.transdata.transform
+        self.artist.V['y'] # update transform
   
+    # rename, this isn't on screen
+    # proxy for half xi, VtoF
+    # axes is an artist, data in E is bounds
+    # axes has nus for those bounds to screen space
+    # get data in axes bounds
+    # tau_genericArtist_x \subset tau_Axes_x
+    # tau_genericArtist_y \subset tau_Axes_y
+    # this assumes shared nu
+    # mu_genericArtist_x \subset mu_Axes_x
+    # get_xlim()/get_ylim = (mu_x, mu_y)
+    # genericArtist_nu_preimage(mu_x, mu_y) => genericArtst_tau
+    # AxesArtist_nu_preimage(mu_x, mu_y) => AxesArtist_tau
+    # get_xlim(), get_ylim() -> 0,1-AxesNu->[100,400]
+    # data = nu([50, 20, 19, 290])-DataNu->[100-400]
+    # mu [ 100, 200,  400] <- see on screen which is mu
+    # Assume that the artist shares a nu with an underlying axes
+        # cheat about doing the filtering on f rather than mu
+    # can't cheat
+        # invert/clean_preimage: get the mu_Axes, preimage through artist_nu, filter back into data
+        # handwave as I know about this, but it's out of scopeish
+        # no_invertable: no cheat: k = data.query() \subset preimage , data.query(k)
+    # matplotlib does not return tau_Axes
[…]

### 2021-08-30 matplottoy/lib/bar.py
really borked rendering

@@ -123,8 +123,7 @@ class Bar(Rho):
                 path = mpath.Path([(xi, fl), (xi, fl + yi), 
                            (xi + wd, fl + yi), 
                            (xi + wd, fl), (xi, fl)], closed=True)
-                render.draw_path(gc=gc, path=path, 
-                    transform=transform, 
+                render.draw_path(gc=gc, path=path, transform=transform, 
                     rgbFace=(fc.r, fc.g, fc.b, fc.a))
         return fake_draw
 
@@ -140,63 +139,22 @@ class GenericArtist(martist.Artist):
     def __init__(self, artist:TopologicalArtist):
         super().__init__()
         self.artist = artist
-        self.artist.V['x'] # update w/ transforms ax.transdata.transform
-        self.artist.V['y'] # update transform
-  
-    # rename, this isn't on screen
-    # proxy for half xi, VtoF
-    # axes is an artist, data in E is bounds
-    # axes has nus for those bounds to screen space
-    # get data in axes bounds
-    # tau_genericArtist_x \subset tau_Axes_x
-    # tau_genericArtist_y \subset tau_Axes_y
-    # this assumes shared nu
-    # mu_genericArtist_x \subset mu_Axes_x
-    # get_xlim()/get_ylim = (mu_x, mu_y)
-    # genericArtist_nu_preimage(mu_x, mu_y) => genericArtst_tau
-    # AxesArtist_nu_preimage(mu_x, mu_y) => AxesArtist_tau
-    # get_xlim(), get_ylim() -> 0,1-AxesNu->[100,400]
-    # data = nu([50, 20, 19, 290])-DataNu->[100-400]
-    # mu [ 100, 200,  400] <- see on screen which is mu
-    # Assume that the artist shares a nu with an underlying axes
-        # cheat about doing the filtering on f rather than mu
-    # can't cheat
-        # invert/clean_preimage: get the mu_Axes, preimage through artist_nu, filter back into data
-        # handwave as I know about this, but it's out of scopeish
-        # no_invertable: no cheat: k = data.query() \subset preimage , data.query(k)
-    # matplotlib does not return tau_Axes
-    # matplotlib currently gives approx mu_1_Axes 
-    # datetime->datetime
 
-    # data gets selection 
     def get_screen_bounds_to_data_bounds(self, renderer)->(dict, int):
-        """proxy for S->F over K"""
-        # actual limits...scale...projections...
-        # x1, x1
[…]

### 2021-08-30 matplottoy/lib/datawrap.py

@@ -16,6 +16,7 @@ class DataFrameWrapper:
 
     def query(self, data_bounds:Optional[dict] = None, sampling_rate:int=None)->List[dict]:
         local_section = self.global_section
+    
+
         records = {Fi:local_section[Fi].values.squeeze() for Fi in local_section.columns}
-        records['k'] = local_section.index.values
         return [records]

### 2021-08-30 matplottoy/lib/position.py

@@ -34,7 +34,7 @@ class Nominal:
 
     def inverse(self, value):
         values = np.atleast_1d(np.array(value, dtype=object))
-        return [self._inverse[v] for v in values]
+        return [self._inverse[int(v)] for v in values]
 
     def validate(self, mtype):
         return True

### 2021-09-01 paper/main.tex
bar and barv
filled in outline

@@
 %~\IEEEmembership{Member,~IEEE,}
-\author{Hannah~Aizenman, Thomas~Caswell, and~Michael~Grossberg% <-this % stops a space
+\author{Hannah~Aizenman, Thomas~Caswell, and~Michael~Grossberg,~\IEEEmembership{Member,~IEEE,}% <-this % stops a space
 \IEEEcompsocitemizethanks{\IEEEcompsocthanksitem H. Aizenman and M. Grossberg are with the department of Computer Science, City College of New York. 
@@
 \hfill August 26, 2015
+\section{Introduction}
+\section{Related Work}
+\subsection{Continuity}
+\subsection{Equivariance}
+\subsection{Category Theory}
+\section{Topological Equivariant Artist Model}
+- brief intro to artist in most simple form $\A\rightarrow H$
 
-\subsection{Subsection Heading Here}
-Subsection text here.
+\subsection{Data Bundle}
+\subsubsection{Continuity: Base Space K}
+\subsection{Equivariance: Fiber F}
+\subsection{Structured Data: Section \tau}
+\subsubsection{Continuity of Data: sheaf $\mathscr{O}$}
+\subsection{Graphic Bundle} %probably doesn't all need to be subsections, just figure + short description
+\subsubsection{Continuity: Base Space H}
+\subsubsection{Equivariance: Fiber D}
+\subsubsection{Structured Data: Graphic \rho}
+\subsection{Artist}
+%% flesh out categorical framing of artist, walk through how eq 16 is part of implementing the artists
+%% https://github.com/story645/proposal/blob/main/notes/meetings/2021_08_30.md
 
+\subsubsection{Continuity: xi}
 % needed in second column of first page if using \IEEEpubid
 %\IEEEpubidadjcol
-
-\subsubsection{Subsubsection Heading Here}
-Subsubsection text here.
-
-% An example of a floating figure using the graphicx package.
-% Note that \label must occur AFTER (or within) \caption.
-% For figures, \caption should occur after the \includegraphics.
-% Note that IEEEtran v1.7 and later has special internal code that
-% is designed to preserve the operation of \label within \caption
-% even when the captionsoff option is in effect. However, because
-% of issues like this, it may be the safest practice to put all your
-% \label just after \caption rather than within \caption{}.
-%
-% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
-% option should be used if it is desired that the figures are to be
-% displayed while in draft mode.
-%
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[width=2.5in]{myfigure}
-% where an .eps filename suffix will be assumed under latex, 
-% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-% via \DeclareGraphicsExtensions.
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure}
-
-% Note that the IEEE typically puts floats only at the top, even when this
-% results in a large percentage of a column being occupied by floats.
-% However, the Computer Society has been known to put floats at the bottom.
-
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command,
-% and the \label for the overall figure must come after \caption.
-% \hfil is used as a separator to get equal spacing.
-% Watch out that the combined width of all the subfigures on a 
-% line do not exceed the text width or a line break will occur.
-%
-%\begin{figure*}[!t]
-%\centering
-%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
-%\label{fig_first_case}}
-%\hfil
-%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
-%\label{fig_second_case}}
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure*}
-%
-% Note that often IEEE papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat[]), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
-% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
-% labels, the optional argument to \subfloat must be present. If a
-% subcaption is not desired, just leave its contents blank,
-% e.g., \subfloat[].
-
-% An example of a floating table. Note that, for IEEE style tables, the
-% \caption command should come BEFORE the table and, given that table
-% captions serve much like titles, are usually capitalized except for words
-% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
-% and up, which are usually not capitalized unless they are the first or
-% last word of the caption. Table text will default to \footnotesize as
-% the IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
-
-% Note that the IEEE does not put floats in the very first column
-% - or typically anywhere on the first page for that matter. Also,
-% in-text middle ("here") positioning is typically not used, but it
-% is allowed and encouraged for Computer Society conferences (but
-% not Computer Society journals). Most IEEE journals/conferences use
-% top floats exclusively. 
-% Note that, LaTeX2e, unlike IEEE journals/conferences, places
-% footnotes above bottom floats. This can be corrected via the
-% \fnbelowfloat command of the stfloats package.
+\subsubsection{Visual Bundle: V}
+\subsubsection{Visual encodings: nu} %look at what K&S call these stages
+\subsubsection{Visual composition: Q} % look at what K&S call these stages
+\subsection{Combining Artists}
+\section{Case Study}
+\section{Discussion}
+\subsection{Limitations}
+\subsection{future work}
 
@@
 
-% if have a single appendix:
-%\appendix[Proof of the Zonklar Equations]
-% or
-%\appendix  % for no appendix heading
-% do not use \section anymore after \appendix, only \section*
-% is possibly needed
-
-% use appendices with more than one appendix
-% then use \section to start each appendix
-% you must declare a \section before using any
-% \subsection or using \label (\appendices by itself
-% starts a section numbered zero.)
-%
-
 \appendices
-\section{Proof of the First Zonklar Equation}
-Appendix one text goes here.
+\section{Rendering: Rho}
+\section{Manufacturing Q: Q_hat}
 

### 2021-09-01 matplottoy/lib/bar.py

@@ -80,9 +80,10 @@ def barV(V):
 
 class Bar(Rho): 
     #what are the 
-    P_required = {'x', 'y'}
+    P_required = {'position', 'length'}
     P_optional = {'floor', 'width', 'facecolor', 'edgecolor', 
                   'linewidth', 'linestyle'}
+    bounds = {'x': 'position', 'y': 'length'}
 
     # Spivak r: C \rightarrow U_{\sigma}, U_{sigma}=F_{i}=F_{c} 
     def compose_with_nu(self, p:str, f:Tuple[str,...], nu:callable, nu_inv:callable=None)->Bar:
@@ -112,10 +113,10 @@ class Bar(Rho):
         return mus
 
     @staticmethod
-    def qhat(x, width, y, floor, facecolor, edgecolor, linewidth, linestyle):
+    def qhat(position, width, length, floor, facecolor, edgecolor, linewidth, linestyle):
         # are matplotlib 
         def fake_draw(render, transform=mtransforms.IdentityTransform()):
-            for (xi, wd, yi, fl, fc, ec, lw, ls) in zip(x, width, y, floor, facecolor, edgecolor, linewidth, linestyle):
+            for (xi, wd, yi, fl, fc, ec, lw, ls) in zip(position, width, length, floor, facecolor, edgecolor, linewidth, linestyle):
                 gc = render.new_gc()
                 gc.set_foreground((ec.r, ec.g, ec.b, ec.a))
                 gc.set_dashes(*ls)
@@ -128,11 +129,11 @@ class Bar(Rho):
         return fake_draw
 
 class BarH(Bar):
+    bounds = {'x': 'length', 'y': 'position'}
     @staticmethod
-    def qhat(x, width, y, floor, facecolor, edgecolor, linewidth, linestyle):
+    def qhat(length, width, position, floor, facecolor, edgecolor, linewidth, linestyle):
         # horizontal bar, x0 is floor, x1 is height, y is y0, width is y1
-        return Bar.qhat(floor, x, width, y, facecolor, edgecolor, linewidth, linestyle)
-
+        return Bar.qhat(floor, length, width, position, facecolor, edgecolor, linewidth, linestyle)
 
 class GenericArtist(martist.Artist):
     # start w/ working w/ an artist object then stripped down artist
@@ -146,13 +147,15 @@ class GenericArtist(martist.Artist):
         """
         bounds = {}
         for pos, (vmin, vmax) in [('x', self.axes.get_xlim()), ('y', self.axes.get_ylim())]:
[…]

### 2021-09-01 matplottoy/lib/color.py

@@ -20,7 +20,6 @@ class Nominal:
         curried higher level function"""
         assert(mcolors.is_color_like(color) for color in mapping.values())
         self._mapping = mapping
-        self._inverse = {v:k for k, v in mapping.items()}
 
     def __call__(self, value):
         values = np.atleast_1d(value)
@@ -28,5 +27,5 @@ class Nominal:
 
     def inverse(self, value):
         values = np.atleast_1d(np.array(value, dtype=object))
-        return [self._inverse[v] for v in values]
+        return [k for (k, v) in self._mapping.items() if v in values]
 

### 2021-09-01 matplottoy/lib/datawrap.py

@@ -10,13 +10,26 @@ import pandas as pd
 # technically this can be wrapped even more
 # 
 # maybe make a Record type Record = {Fiber_Name: values} 
-class DataFrameWrapper:
+class FruitFrameWrapper:
+    F = {'fruit': 'nominal',
+         'calories': 'ratio', 
+         'juice': 'nominal'}
+
     def __init__(self, section: pd.DataFrame):
         self.global_section = section
 
     def query(self, data_bounds:Optional[dict] = None, sampling_rate:int=None)->List[dict]:
-        local_section = self.global_section
-    
-
-        records = {Fi:local_section[Fi].values.squeeze() for Fi in local_section.columns}
+        data_bounds = {} if data_bounds is None else data_bounds
+        local = self.global_section
+        #iter tools chain from iterables probably
+        for fibers, bounds in data_bounds.items():
+            for Fi, bound in zip(fibers, bounds):
+                if bound is None:
+                    continue # no selection needed
+                if self.F[Fi] == 'nominal':
+                    local = local[local[Fi].isin(bound)]
+                elif self.F[Fi] == 'ratio':
+                    local = local[(min(bound)<=local[Fi]) & (local[Fi]<=max(bound))]
+        
+        records = {Fi:local[Fi].values.squeeze() for Fi in local.columns}
         return [records]

### 2021-09-01 matplottoy/lib/position.py

@@ -26,15 +26,20 @@ class Log:
 class Nominal:
     def __init__(self, mapping):
         self._mapping = mapping
-        self._inverse = {v:k for k, v in mapping.items()}
         
     def __call__(self, value):
         values = np.atleast_1d(np.array(value, dtype=object))
-        return np.array([self._mapping[v] for v in values])
+        return [self._mapping[v] for v in values]
 
     def inverse(self, value):
+        # really needs something more robust, 
+        # but current assumption is that since position is monotonic
+        # must return elements corresponding to the monotonicity
+        # if you want individaual, must ask for it
+
         values = np.atleast_1d(np.array(value, dtype=object))
-        return [self._inverse[int(v)] for v in values]
+        return ([k for (k, v) in self._mapping.items() 
+                if min(values)<=v<=max(values)], )
 
     def validate(self, mtype):
         return True

### 2021-09-02 paper/main.tex
filled in w/ standing figures + sections

@@
 % bare_jrnl_compsoc, texV1.4b, 2015/08/26, Michael Shell
-
 \documentclass[10pt,journal,compsoc]{IEEEtran}
@@
 
-% *** CITATION PACKAGES ***
-\ifCLASSOPTIONcompsoc
-  % IEEE Computer Society needs nocompress option
-  % requires cite.sty v4.0 or later (November 2003)
 \usepackage[nocompress]{cite}
-\else
-  % normal IEEE
-  \usepackage{cite}
-\fi
 
@@
 \usepackage{amsmath}
+%% with other math-related packages, you may want to disable it.
+\usepackage{amsmath,amsfonts,amssymb,eulervm,xspace}
+%\usepackage{mathrsfs} % math script fonts
 
 % *** SPECIALIZED LIST PACKAGES ***
+\usepackage{xcolor}
 \usepackage{algorithmic}
-
+\usepackage[utf8]{inputenc}
+\usepackage{subfig} %ieee does not like subfigure
+\usepackage{multicol}
+\usepackage{tikz}
+\usetikzlibrary{cd} % commutative diagrams
+\newtheorem{prop}{Proposition} %math?
+\usepackage[switch]{lineno}
+\usepackage{minted}
+\setminted[python]{fontsize=\scriptsize, 
+                   linenos,
+                   numbersep=8pt,
+                   autogobble, 
+                   frame=lines,
+                   framesep=3mm} 
 % *** ALIGNMENT PACKAGES ***
 \usepackage{array}
+\usepackage{tabulary}
 % IEEEtran contains the IEEEeqnarray family of commands
@@
 
+\usepackage{notation} %notation conventions
 % correct bad hyphenation here
@@
 \title{Topological Equivariant Artist Model for Visualization Library Architecture}
-%
-%
 % author names and IEEE memberships
-% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
-% a structure at a ~ so this keeps an author's name from being broken across
-% two lines.
-% use \thanks{} to gain access to the first footnote area
-% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
-% was not built to handle multiple paragraphs
-%
-%
-%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
-% lists the Computer Society journals use for "first footnote" author
-% affiliations. Use \IEEEcompsocthanksitem which works much like \item
-% for each affiliation group. When not in compsoc mode,
-% \IEEEcompsocitemizethanks becomes like \thanks and
-% \IEEEcompsocthanksitem becomes a line break with idention. This
-% facilitates dual compilation, although admittedly the differences in the
-% desired content of \author between the different types of papers makes a
-% one-size-fits-all approach a daunting prospect. For instance, compsoc 
-% journal papers have the author affiliations above the "Manuscript
-% received ..."  text while in non-compsoc journals this is reversed. Sigh.
-%~\IEEEmembership{Member,~IEEE,}
 \author{Hannah~Aizenman, Thomas~Caswell, and~Michael~Grossberg,~\IEEEmembership{Member,~IEEE,}% <-this % stops a space
@@
 
-% To allow for easy dual compilation without having to reenter the
-% abstract/keywords data, the \IEEEtitleabstractindextext text will
-% not be used in maketitle, but will appear (i.e., to be "transported")
-% here as \IEEEdisplaynontitleabstractindextext when the compsoc 
-% or transmag modes are not selected <OR> if conference mode is selected 
-% - because all conference papers position the abstract like regular
-% papers do.
-\IEEEdisplaynontitleabstractindextext
-% \IEEEdisplaynontitleabstractindextext has no effect when using
-% compsoc or transmag under a non-conference mode.
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
 \IEEEpeerreviewmaketitle
@@
 \IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
-% Computer Society journal (but not conference!) papers do something unusual
-% with the very first section heading (almost always called "Introduction").
-% They place it ABOVE the main text! IEEEtran.cls does not automatically do
-% this for you, but you can achieve this effect with the provided
-% \IEEEraisesectionheading{} command. Note the need to keep any \label that
-% is to refer to the section immediately after \section in the above as
-% \IEEEraisesectionheading puts \section within a raised box.
 
-% The very first letter is a 2 line initial drop letter followed
-% by the rest of the first word in caps (small caps for compsoc).
-% 
-% form to use if the first word consists of a single letter:
-% \IEEEPARstart{A}{demo} file is ....
-% 
-% form to use if you need the single drop letter followed by
-% normal text (unknown if ever used by the IEEE):
-% \IEEEPARstart{A}{}demo file is ....
-% 
-% Some journals put the first two words in caps:
-% \IEEEPARstart{T}{his demo} file is ....
-% 
-% Here we have the typical use of a "T" for an initial drop letter
-% and "HIS" in caps to complete the first word.
 \IEEEPARstart{T}{his} demo file is intended to serve as a ``starter file''
@@
 I wish you the best of success.
-
-\hfill mds
-
-\hfill August 26, 2015
-\section{Introduction}
 \section{Related Work}
 \subsection{Continuity}
+\begin{figure}[!t]
+  \includegraphics[width=2.5in]{k_different_types.png}
+\end{figure}
+\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}
 \subsection{Equivariance}
+\begin{figure}[!t]
+  \includegraphics[width=2.5in]{equiv.png}
+\end{figure}
+
 \subsection{Category Theory}
+
 \section{Topological Equivariant Artist Model}
-- brief intro to artist in most simple form $\A\rightarrow H$
+%%- brief intro to artist in most simple form $\A\rightarrow H$
 
-\subsection{Data Bundle}
-\subsubsection{Continuity: Base Space K}
-\subsection{Equivariance: Fiber F}
-\subsection{Structured Data: Section \tau}
-\subsubsection{Continuity of Data: sheaf $\mathscr{O}$}
-\subsection{Graphic Bundle} %probably doesn't all need to be subsections, just figure + short description
-\subsubsection{Continuity: Base Space H}
-\subsubsection{Equivariance: Fiber D}
-\subsubsection{Structured Data: Graphic \rho}
+\subsection{Data Bundle \dtotal}
+\begin{figure}[!t]
+
+\end{figure}
+\paragraph{Continuity: Base Space \dbase}
+\paragraph{Equivariance: Fiber \dfiber}
+\paragraph{Structured Data: Section \dsection}
+\subsubsection{Continuity of Data: sheaf $\mathcal{O}$}
+\begin{figure}
+  \includegraphics[width=2.5in]{sheaf.png}
+  \caption{Maybe create a figure that shows what the sheaf does, like 
+  https://www.euclideanspace.com/maths/topology/sheaf/index.htm}
+\end{figure}
+\subsection{Visual Bundle \gtotal} %probably doesn't all need to be subsections, just figure + short description
+\begin{figure}
+  \includegraphics[width=2.5in]{render.png}
+\end{figure}
+\paragraph{Continuity: Base Space \gbase}
+\paragraph{Equivariance: Fiber \gfiber}
+\paragraph{Structured Data: Visual \gsection}
 \subsection{Artist}
@@
 
-\subsubsection{Continuity: xi}
-% needed in second column of first page if using \IEEEpubid
-%\IEEEpubidadjcol
-\subsubsection{Visual Bundle: V}
-\subsubsection{Visual encodings: nu} %look at what K&S call these stages
-\subsubsection{Visual composition: Q} % look at what K&S call these stages
+\subsubsection{Visual to Data: \vindex}
+\begin{figure}[!t]
+  \includegraphics[width=2.5in]{xi.png}
+\end{figure}
+\subsubsection{Representation Bundle: \vtotal}
+
+\subsubsection{Data to Representation: \vchannel} %look at what K&S call these stages
+\begin{figure}[!t]
+  \includegraphics[width=2.5in]{nu.png}
+\end{figure}
+
+\subsubsection{Representation to Visual: \vmark} % look at what K&S call these stages
+\begin{figure}[!t]
+  \includegraphics[width=2.5in]{q.png}
+\end{figure}
+
+\begin{figure}[!t]
+  \includegraphics[width=2.5in]{path_of_q.png}
+  \caption{rework this as a commutative box w/ the r in E row associated w/ this qhat(k)}
+\end{figure}
+
 \subsection{Combining Artists}
+\begin{figure}[!t]
+\centering
+\subfloat[Artists with shared $\mu_i$ renderered correctly]{\includegraphics[width=2.5in]{combined_artist.png}%
+\label{fig_first_case}}
+\hfil
+\subfloat[Artists without shared $\mu_i$]{\includegraphics[width=2.5in]{exploding_artist.png}%
+\label{fig_second_case}}
+\caption{Simulation results for the network.}
+\label{fig_sim}
+\end{figure}
+
 \section{Case Study}
+
 \section{Discussion}
@@
 \appendices
-\section{Rendering: Rho}
-\section{Manufacturing Q: Q_hat}
+\section{Rendering: \gsection}
+\section{Manufacturing $\vmarkd \leftarrow \vmark$}
 
@@
 Biography text here.
-\end{IEEEbiographynophoto
+\end{IEEEbiographynophoto}
 

### 2021-09-03 paper/main.tex
more paper skeleon

@@
 \section{Related Work}
+\subsection{Describe API w/ Category Theory}
+\begin{equation}
+  \label{eq:artist}
+  \mathscr{\vartist}: \mathscr{\dtotal} \rightarrow \mathscr{\gtotal}
+\end{equation}
+
 \subsection{Continuity}
-\begin{figure}[!t]
+Continuity is
+\begin{figure}[!h]
   \includegraphics[width=2.5in]{k_different_types.png}
 \end{figure}
+
+We care about continuity b/c 
+\begin{figure}
+  \includegraphics[width=2.5in]{whycontinuity.png}
+\end{figure}
+
+We express continuity using fiber bundles, which are ...
 \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}
+\begin{equation}
+  \label{eq:fiber_bundle}
+  \begin{tikzcd}
+      \dfiber \arrow[r, hook] & \dtotal \arrow[r, "\pi"] & \dbase
+  \end{tikzcd}
+\end{equation}
+
 \subsection{Equivariance}
-\begin{figure}[!t]
+What is it?
+
+\begin{figure}[!h]
   \includegraphics[width=2.5in]{equiv.png}
@@
 
-\subsection{Category Theory}
+\begin{tikzcd}
+  data \arrow[r] \arrow[d, "function"] & representation \arrow[r] & visual\;stimulus \arrow[d, "visual\;equivalent\;to\;function"] \\
+  data \arrow[r]                       & representation \arrow[r] & visual\; stimulus                                          
+\end{tikzcd}
 
 \section{Topological Equivariant Artist Model}
-%%- brief intro to artist in most simple form $\A\rightarrow H$
+%%- brief intro to artist in most simple form $\vartist \dtotal \rightarrow \gtotal$
+
+\section{Union of Artists}
+\begin{figure}[!h]
+\centering
+\subfloat[Artists with shared $\mu_i$ renderered correctly]{\includegraphics[width=1.2in]{combined_artist.png}%
+\label{fig_first_case}}
+%\hfil
+\subfloat[Artists without shared $\mu_i$]{\includegraphics[width=1.2in]{exploding_artist.png}%
+\label{fig_second_case}}
+\caption{Simulation results for the network.}
+\label{fig_sim}
+\end{figure}
+}
 
 \subsection{Data Bundle \dtotal}
-\begin{figure}[!t]
+\begin{figure}[!h]
+  \includegraphics[width=2.5in]{k_qspace.png}
+  \caption{this is gonna be replaced w/ more concrete}
+\end{figure}
 
-\end{figure}
-\paragraph{Continuity: Base Space \dbase}
-\paragraph{Equivariance: Fiber \dfiber}
 \paragraph{Structured Data: Section \dsection}
-\subsubsection{Continuity of Data: sheaf $\mathcal{O}$}
+\begin{equation}
+  \begin{tikzcd}
+      \dfiber \arrow[r, hook] & \dtotal \arrow[d, "\pi"'] \\
+                        & \dbase \arrow[u, "\dsection"', bend right]
+  \end{tikzcd}
+\end{equation}
+
+\paragraph{Structure: Continuity and Equivariant Actions}
+\begin{equation}
+  \begin{tikzcd}
+      \dfiber_i \arrow[d, "m_j"'] \arrow[rd, "m_j\;\circ\; m_k"] &         \\
+      \dfiber_i \arrow[r, "m_k"']  & \dfiber_i
+  \end{tikzcd}
+\end{equation}
+
+\subsubsection{Continuity of Data: sheaf \sheaf}
 \begin{figure}
@@
 \end{figure}
-\subsection{Visual Bundle \gtotal} %probably doesn't all need to be subsections, just figure + short description
+
+\subsection{Graphic Bundle \gtotal} %probably doesn't all need to be subsections, just figure + short description
 \begin{figure}
@@
 \end{figure}
+\begin{equation}
+  \begin{tikzcd}[ampersand replacement=\&]
+      \gfiber \arrow[r, hook] \& \gtotal \arrow[d, "\pi"'] \\
+                        \& \gbase \arrow[u, "\gsection"', bend right]
+  \end{tikzcd}
+\end{equation}
 \paragraph{Continuity: Base Space \gbase}
@@
 \paragraph{Structured Data: Visual \gsection}
+
 \subsection{Artist}
@@
 %% https://github.com/story645/proposal/blob/main/notes/meetings/2021_08_30.md
+\begin{equation}
+  \vartist: \sheaf(\dtotal) \rightarrow \sheaf(\gtotal)
+  \label{eq:math:artist:artist}
+\end{equation}
 
-\subsubsection{Visual to Data: \vindex}
-\begin{figure}[!t]
+\subsubsection{Graphic to Data: \vindex}
+\begin{figure}[!h]
   \includegraphics[width=2.5in]{xi.png}
 \end{figure}
-\subsubsection{Representation Bundle: \vtotal}
+\begin{equation}
+  \begin{tikzcd}
+      \dtotal \arrow[d, "\pi"'] & \gtotal \arrow[d, "\pi"'] \\
+      \dbase                   & \gbase \arrow[l, "\vindex"']
+  \end{tikzcd}
+  \label{eq:math:graphic:vindex}
+\end{equation}
+
+\begin{equation}
+  \label{eq:math:artist:diagram}
+  \begin{tikzcd}
+      \dtotal \arrow[r, "\vchannel"] \arrow[rd, "\pi"'] & \vtotal \arrow[d, "\pi"] & \vindex^*\vtotal \arrow[r, "\vmark"] \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"'] & \gtotal \arrow[ld, "\pi"] \\
+                                            & \dbase                  & \gbase \arrow[l, "\vindex"']                                              &                    
+      \end{tikzcd}
+\end{equation}
+
+\subsubsection{Representation \vtotal}
+\begin{equation}
+  \begin{tikzcd}[ampersand replacement=\&]
+      \vfiber \arrow[r, hook] \& \vtotal \arrow[d, "\pi"'] \\
+                        \& \dbase \arrow[u, "\vsection"', bend right]
+  \end{tikzcd}
+\end{equation}
 
 \subsubsection{Data to Representation: \vchannel} %look at what K&S call these stages
-\begin{figure}[!t]
+\begin{figure}[!h]
   \includegraphics[width=2.5in]{nu.png}
 \end{figure}
+\begin{equation}
+  \label{eq:math:artist:nu}
+  \{\vchannel_{0}, \ldots, \vchannel_{n}\}: \{\dsection_{0}, \ldots, \dsection_{n}\} \mapsto \{\vsection_{0}, \ldots, \vsection_{n}\}
+\end{equation}
+
+We enforce the equivariance constraint
+\begin{equation}
+  \label{eq:math:artist:nu_commute}
+\begin{tikzcd}
+  \dtotal_i \arrow[r] \arrow[r, "\vchannel_i"] \arrow[d, "m_{\delement}"'] & \vtotal_i \arrow[d, "m_{\velement}"] \\
+  \dtotal_i \arrow[r, "\vchannel_i"]                           & \vtotal_i               
+\end{tikzcd}
+\end{equation}
 
 \subsubsection{Representation to Visual: \vmark} % look at what K&S call these stages
-\begin{figure}[!t]
+
+\begin{figure}[!h]
   \includegraphics[width=2.5in]{q.png}
@@
 
-\begin{figure}[!t]
-  \includegraphics[width=2.5in]{path_of_q.png}
+\begin{figure}[!h]
+  \includegraphics[width=2.5in]{diff_type_q.png}
   \caption{rework this as a commutative box w/ the r in E row associated w/ this qhat(k)}
@@
 
-\subsection{Combining Artists}
-\begin{figure}[!t]
-\centering
-\subfloat[Artists with shared $\mu_i$ renderered correctly]{\includegraphics[width=2.5in]{combined_artist.png}%
-\label{fig_first_case}}
-\hfil
-\subfloat[Artists without shared $\mu_i$]{\includegraphics[width=2.5in]{exploding_artist.png}%
-\label{fig_second_case}}
-\caption{Simulation results for the network.}
-\label{fig_sim}
-\end{figure}
+\subsubsection{Data $\leftrightarrow$ Visual}
+Putting all the pieces together, we have a system for 
+\begin{equation}
+  \label{eq:math:artist:diagram}
+  \begin{tikzcd}
+      \dtotal \arrow[r, "\vchannel"] \arrow[rd, "\pi"'] & \vtotal \arrow[d, "\pi"] & \vindex^*\vtotal \arrow[r, "\vmark"] \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"'] & \gtotal \arrow[ld, "\pi"] \\
+                                            & \dbase                  & \gbase \arrow[l, "\vindex"']                                              &                    
+      \end{tikzcd}
+\end{equation}
 
 \section{Case Study}
+\begin{figure}[t!]
+  \includegraphics[width=2.5in]{path_of_q.png}
+    \caption{add in xi!}
+  \label{fig:api}
+\end{figure}
+We implement the \texbf{arrows} in \autoref{fig:api}
 
@@
 \section{Manufacturing $\vmarkd \leftarrow \vmark$}
-
-% you can choose not to have a title for an appendix
-% if you want by leaving the argument blank
-\section{}
-Appendix two text goes here.
+\begin{equation}
+  \begin{tikzcd}
+      \textcolor{gray!50}{\dtotal} \arrow[r, "\vchannel", color=gray!50] \arrow[rd, "\pi"', color=gray!50] & \vtotal \arrow[d, "\pi"']                  & \vindex^*\vtotal \arrow[r, "\vmark", color=gray!50] \arrow[d, "\vindex^*\pi"'] \arrow[l,  "\vindex^*"'] & \textcolor{gray!50}{\gtotal} \arrow[ld, "\pi", color=gray!50] \\ & \dbase \arrow[u, "\vsection"', bend right] & \gbase \arrow[l, "\vindex"'] \arrow[u, "\vindex^*\vsection"', bend right]               &
+      \end{tikzcd}
+      \label{eq:math:artist:qhat}
+\end{equation}
 

### 2021-09-06 paper/main.tex
intro, start of background + structure

@@
 
+\newcommand{\note}[1]{\textcolor{magenta}{#1}} 
 \usepackage[nocompress]{cite}
-
 % *** GRAPHICS RELATED PACKAGES ***
@@
 
-\IEEEPARstart{T}{his} demo file is intended to serve as a ``starter file''
-for IEEE Computer Society journal papers produced under \LaTeX\ using
-IEEEtran.cls version 1.8b and later.
-% You must have at least 2 lines in the paragraph with the drop letter
-% (should never be an issue)
-I wish you the best of success.
+\IEEEPARstart{T}his paper uses methods from topology and category theory to develop a model of the transformation from data to graphical representation.
+This model provides a language to specify how data is structured and how this structure is carried through in the visualization, and serves as the basis for a functional approach to implementing visualization library components.
+Topology allows us to describe the structure of the data and graphics in a generalizable, scalable, and trackable way.
+Category theory provides a framework for separating the transformations implemented by visualization libraries from the various stages of visualization and therefore can be used to describe the constraints imposed on the library components \cite{wielsManagementEvolvingSpecifications1998,goguenCategoricalManifesto1991}.
+Well constrained modular components are inherently functional\cite{hughesWhyFunctionalProgramming1989}, and a functional framework yields a library implementation that is likely to be shorter, clearer, and more suitd to distributed, concurrent, and on demand tasks\cite{huHowFunctionalProgramming2015}.
+Using this functional approach, this paper contributes a practical framework for decoupling data processing from visualization generation in a way that allows for modular visualization components that are applicable to a variety of data sets in different formats.
+\note{is it OK that this is something reviewer 4 wrote}
+
 \section{Related Work}
-\subsection{Describe API w/ Category Theory}
-\begin{equation}
-  \label{eq:artist}
-  \mathscr{\vartist}: \mathscr{\dtotal} \rightarrow \mathscr{\gtotal}
-\end{equation}
+This work aims to develop a model for describing visualization transformations that can be translated into visualization library architecture.
+In doing so, we describe how visualization libraries attempt this goal and  build on work that formally describes what properties of data should be preserved.
+We restrict the properties of data that should be preserved to
+
+\begin{LaTeXdescription}
+  \item [continuity] how elements in a dataset are connect to each other, e.g. discrete rows in a table, networked nodes, pixels in an image, points on a line
+  \item [equivariance] functions on data that have an equivalent effect on the graphical representation, e.g.
+rotating a matrix has a matching rotation of the image, translating the points on a line has a matching visual shift in the line plot
+\end{LaTeXdescription}
 
 \subsection{Continuity}
-Continuity is
 \begin{figure}[!h]
-  \includegraphics[width=2.5in]{k_different_types.png}
+  \includegraphics[width=\columnwidth]{k_different_types.png}
+  \caption{Continuity is how elements in a data set are connected to each other, which is distinct from how the data is structured.
+The rows in (a) are discrete, therefore they have discrete continuity as illustrated by the discrete dots.
+The gaussian in (b) is a 1D continuous function, therefore the continuity of the elements of the gaussian can be represented as a line on an interval (0,1).
+In (c), every element of the globe is connected to its nearest neighbors, which yields a 2D continuous continuity as illustrated by the square.}
 \end{figure}
 
+Continuity is
+
 We care about continuity b/c 
-\begin{figure}
-  \includegraphics[width=2.5in]{whycontinuity.png}
+\begin{figure}[!h]
+  \includegraphics[width=\columnwidth]{whycontinuity.png}
+  \caption{Continuity is implicit in choice of visualization rather than explicitely in choice of data container.
+The line plots in (b) are generated by a 2D table (a).
+Structurally this table can be identical to the 2D matrix (a) that generates the image in (c).}
 \end{figure}
@@
 \begin{figure}[!h]
-  \includegraphics[width=2.5in]{equiv.png}
+  \includegraphics[width=\columnwidth]{equiv.png}
+  \caption{Equivariance is that a transformation on the data has a corresponding transformation in the graphical representation.
+For example, in this figure the data is scaled by a factor 10.
+Equivalently the line plot is scaled by factor of 10, resulting in a shrunken line plot.
+Either a transformation on the data side can induce a transformation on the visual side, or a transformation on the visual side indicates that there is also a transformation on the data side.
+}
 \end{figure}
@@
 
-\section{Topological Equivariant Artist Model}
+\section{Modeling Visualization Stages: Fiber bundles}
+
 %%- brief intro to artist in most simple form $\vartist \dtotal \rightarrow \gtotal$ 
-
-\section{Union of Artists}
-\begin{figure}[!h]
-\centering
-\subfloat[Artists with shared $\mu_i$ renderered correctly]{\includegraphics[width=1.2in]{combined_artist.png}%
-\label{fig_first_case}}
-%\hfil
-\subfloat[Artists without shared $\mu_i$]{\includegraphics[width=1.2in]{exploding_artist.png}%
-\label{fig_second_case}}
-\caption{Simulation results for the network.}
-\label{fig_sim}
-\end{figure}
-}
+The \textcolor{artist}{Artist $\mathcal{\vartist}$} is a transformation from
 
 \subsection{Data Bundle \dtotal}
-\begin{figure}[!h]
+\begin{figure}[h!]
   \includegraphics[width=2.5in]{k_qspace.png}
@@
 \end{figure}
+We use topology to model
+
+The The continuity of the data is encoded in the base space \dbase.
+
+The properties of the variables are encoded in the fiber space \dfiber. The \textcolor{fiber}{fiber} is a topological space
+
+Spivak provides notation for describing the set of all possible values $U_{\sigma}$ of a single column with name $c$ and type $T$.
+This set of values is the fiber "F" $F = U_{\sigma}(c)$.
+When data is multivariate, the fiber F is the cartesian product of each columns fiber space (5)
 
@@
 \subsubsection{Continuity of Data: sheaf \sheaf}
-\begin{figure}
+\begin{figure}[h!]
   \includegraphics[width=2.5in]{sheaf.png}
-  \caption{Maybe create a figure that shows what the sheaf does, like 
-  https://www.euclideanspace.com/maths/topology/sheaf/index.htm}
+  \caption{}
 \end{figure}
@@
 \subsection{Graphic Bundle \gtotal} %probably doesn't all need to be subsections, just figure + short description
-\begin{figure}
-  \includegraphics[width=2.5in]{render.png}
+\begin{figure}[h!]
+  \includegraphics[width=\columnwidth]{render.png}
 \end{figure}
@@
 
-\subsubsection{Graphic to Data: \vindex}
+\section{Union of Artists}
 \begin{figure}[!h]
-  \includegraphics[width=2.5in]{xi.png}
+\centering
+\subfloat[Artists with shared $\mu_i$ renderered correctly]{\includegraphics[width=1.2in]{combined_artist.png}%
+\label{fig_first_case}}
+%\hfil
+\subfloat[Artists without shared $\mu_i$]{\includegraphics[width=1.2in]{exploding_artist.png}%
+\label{fig_second_case}}
+\caption{Simulation results for the network.}
+\label{fig_sim}
 \end{figure}
-\begin{equation}
-  \begin{tikzcd}
-      \dtotal \arrow[d, "\pi"'] & \gtotal \arrow[d, "\pi"'] \\
-      \dbase                   & \gbase \arrow[l, "\vindex"']
-  \end{tikzcd}
-  \label{eq:math:graphic:vindex}
-\end{equation}
+
+\section{Construction and formal properties of Artists}
+
+\begin{figure*}[]
+  \includegraphics[width=\linewidth]{q.png}
+\end{figure*}
 
@@
 
-\subsubsection{Representation \vtotal}
+\subsection{Graphic to Data: \vindex}
+\begin{equation}
+  \begin{tikzcd}
+      \dtotal \arrow[d, "\pi"'] & \gtotal \arrow[d, "\pi"'] \\
+      \dbase                   & \gbase \arrow[l, "\vindex"']
+  \end{tikzcd}
+  \label{eq:math:graphic:vindex}
+\end{equation}
+
+\subsection{Visual Bundle \vtotal}
 \begin{equation}
@@
 
-\subsubsection{Data to Representation: \vchannel} %look at what K&S call these stages
+\subsection{Data to Representation: \vchannel} %look at what K&S call these stages
 \begin{figure}[!h]
-  \includegraphics[width=2.5in]{nu.png}
+  \centering
+  \subfloat[Artists with shared $\mu_i$ renderered correctly]{\includegraphics[width=1.2in]{partial_fixed.png}%
+  \label{fig_first_case}}
+  %\hfil
+  \subfloat[Artists without shared $\mu_i$]{\includegraphics[width=1.2in]{partial_invalid.png}%
+  \label{fig_second_case}}
+  \caption{Simulation results for the network.}
+  \label{fig_sim}
   \end{figure}
+
 \begin{equation}
@@
 \begin{figure}[!h]
-  \includegraphics[width=2.5in]{q.png}
-\end{figure}
-
-\begin{figure}[!h]
   \includegraphics[width=2.5in]{diff_type_q.png}
@@
 
-\subsubsection{Data $\leftrightarrow$ Visual}
-Putting all the pieces together, we have a system for 
-\begin{equation}
-  \label{eq:math:artist:diagram}
-  \begin{tikzcd}
-      \dtotal \arrow[r, "\vchannel"] \arrow[rd, "\pi"'] & \vtotal \arrow[d, "\pi"] & \vindex^*\vtotal \arrow[r, "\vmark"] \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"'] & \gtotal \arrow[ld, "\pi"] \\
-                                            & \dbase                  & \gbase \arrow[l, "\vindex"']                                              &                    
-      \end{tikzcd}
-\end{equation}
-
 \section{Case Study}
-\begin{figure}[t!]
-  \includegraphics[width=2.5in]{path_of_q.png}
+\begin{figure}[h!]
+  \includegraphics[width=\columnwidth]{path_of_q.png}
     \caption{add in xi!}
@@
 \end{figure}
-We implement the \texbf{arrows} in \autoref{fig:api}
+We implement the \textbf{arrows} in \autoref{fig:api}
+
+\begin{minted}[]{python}
+
+\end{minted}
 

### 2021-09-13 matplottoy/lib/bar.py
where is Q?

@@ -130,9 +130,11 @@ class Bar(Rho):
 
 class BarH(Bar):
     bounds = {'x': 'length', 'y': 'position'}
+
     @staticmethod
     def qhat(length, width, position, floor, facecolor, edgecolor, linewidth, linestyle):
         # horizontal bar, x0 is floor, x1 is height, y is y0, width is y1
+        # alternative is to do this renamin in nu calling parent nu first
         return Bar.qhat(floor, length, width, position, facecolor, edgecolor, linewidth, linestyle)
 
 class GenericArtist(martist.Artist):

### 2021-09-13 paper/main.tex
continuity and topology

@@
 \usepackage[nocompress]{cite}
+\usepackage{hyperref} % autoref
 % *** GRAPHICS RELATED PACKAGES ***
@@
 \section{Related Work}
-This work aims to develop a model for describing visualization transformations that can be translated into visualization library architecture.
-In doing so, we describe how visualization libraries attempt this goal and  build on work that formally describes what properties of data should be preserved.
+This work aims to develop a model for describing visualization transformations that can serve as guidance for how to architecture a general purpose visualization library.
+We define a general purpose visualization library as one that provides non domain specific building block components\cite{wongsuphasawatNavigatingWideWorld2021} for building visualizations, for example functions for converting data to color or encoding data as dots.
+In this section, we describe how visualization libraries attempt this goal and discuss work that formally describes what properties of data should be preserved in a visualization.
 We restrict the properties of data that should be preserved to
@@
 In (c), every element of the globe is connected to its nearest neighbors, which yields a 2D continuous continuity as illustrated by the square.}
+  \label{fig:related-work:continuity}
 \end{figure}
 
-Continuity is
+Continuity is a representation of how the elements in a dataset are connected to each other.
+For example, in \autoref{fig:related-work:continuity}, each station record in the table is independent of the others; therefore, the continuity of the table is discrete.
+The data provided by the gaussian are points sampled along the curve, therefore the continuity of the points on the line is 1D continuous.
+Every point on the globe is connected to its 6 nearest cardinal neighboring points (NW, N, NE, E, SE, S, SW, W).
 
-We care about continuity b/c 
 \begin{figure}[!h]
   \includegraphics[width=\columnwidth]{whycontinuity.png}
-  \caption{Continuity is implicit in choice of visualization rather than explicitely in choice of data container.
+  \caption{Continuity is implicit in choice of visualization rather than explicitly in choice of data container.
 The line plots in (b) are generated by a 2D table (a).
 Structurally this table can be identical to the 2D matrix (a) that generates the image in (c).}
+  \label{fig:related-work:visual-algorithm}
 \end{figure}
 
-We express continuity using fiber bundles, which are ...
-\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}
+Often continuity is expressed in the choice of visual algorithm (visualization type), as explored in taxonomies by Tory and M\"{o}ller \cite{toryRethinkingVisualizationHighlevel2004} and Chi\cite{chiTaxonomyVisualizationTechniques2000}.
+For example, in \autoref{fig:related-work:visual-algorithm} the same table can be interpreted as a set of 1D continuous curves when visualized as a collection of line plots or as a 2D surface when visualized as an image.
+This means that often there is no way to express data continuity independent of visualization type, meaning most visualization libraries will allow, for example, visualizing discrete data as a line plot or an image.
+General purpose visualization libraries-such as Matplotlib\cite{hunterMatplotlib2DGraphics2007}, Vtk\cite{hanwellVisualizationToolkitVTK2015,geveciVTK2012}, and D3 \cite{bostockDataDrivenDocuments2011}-carry distinct data models as part of the implementation of each visual algorithm.
+The lack of unified data model means that each plot in a linked\cite{beckerBrushingScatterplots1987,bujaInteractiveData1991} visualization is treated as independent, as are the transforms converting each field in the data to a visual equivalent.
+
+Domain specific libraries can guarantee consistency because they have a single model of the data in their software design, as discussed in Heer and Agarwal \cite{HeerSoftware2006}'s survey of visualization software design patterns.
+For example, the relational database is core to tools influenced by APT, such as Tableau\cite{StoltePolaris2002,hanrahanVizQL2006,MackinlayShowme2007} and the Grammar of Graphics\cite{wilkinsonGrammarGraphics2005} inspired ggplot\cite{wickhamGgplot2ElegantGraphics2016a}, Vega\cite{satyanarayanDeclarativeInteractionDesign2014} and Altair\cite{vanderplasAltairInteractiveStatistical2018}.
+Images underpin scientific visualization tools such as Napari\cite{nicholas_sofroniew_2021_4533308} and ImageJ\cite{schneiderNIHImageImageJ2012} and the digital humanities oriented ImagePlot\cite{studiesCulturevisImageplot2021} macro; the need to visualize and manipulate graphs has spawned tools like Gephi\cite{bastianGephiOpenSource2009}, Graphviz\cite{ellsonGraphvizOpenSource2002}, and Networkx\cite{HagbergExploringNetwork2008}.
+
+\subsubsection{Fiber Bundles}
+\begin{figure}[h!]
+  \includegraphics[width=\columnwidth]{fiberbundle.png}
+  \caption{A fiber bundle is mathematical construct that allows us to express the relationship between data and continuity.
+The \textcolor{total}{total} space \dtotal is the topological space in which the data is embedded.
+The \textcolor{fiber}{fiber} space \dfiber\ is embedded in \dtotal\ and is the set of all possible values that any
+  \note{add big rectangle E}}
+  \label{fig:related-work:fiber-bundle}
+\end{figure}
+
+A model that allows for expressing data continuity in a general flexible way allows us to express, and therefore be faithful to, the multivariate consistency constraints described by Qu and Hullman\cite{hullmanKeeping2018}.
+A general data model also allows for consistent adaptation to modern data needs, such as complex, metadata rich, distributed, or streaming data.
+The mathematical theory of fiber bundles provides one such abstraction that can express complicated dimensionality and continuity without being tied to any one data container type, as proposed by Butler, Bryson, and Pendley\cite{butlerVisualizationModelBased1989,butlerVectorBundleClassesForm1992}.
+In this paper, we build on their work that proposes using topological spaces to represent different properties of data.
+Here we present a brief summary of topology, for more information see Hatcher\cite{hatcherAlgebraicTopology2002}, Munkres\cite{munkresElementsAlgebraicTopology1984}, and Bradley et.
+al.
+\cite{bradleyTopologyCategoricalApproach2020}.
+
+A topological space a topological space $(X, \mathscr{T})$ is a set $X$ with a topology $\mathscr{T}$. Topologies are collections of open sets
+
+such that the empty set and $X$ are in the collection of open sets $\mathscr{T}$, the union of elements in $\mathscr{T}
+
+Specifically, Butler, Bryson, and Pendley suggest that fiber bundles be the basis of an abstract data model. Fiber bundles are a collection $(\dtotal, \dbase, \dfiber, \pi)$ of topological spaces
+
+\begin{LaTeXdescription}
+  \item[\textcolor{total}{Total Space} \dtotal]
+  \item[\textcolor{fiber}{Fiber Space} \dfiber]
+  \item[\textcolor{base}{Base Space} \dbase]  
+\end{LaTeXdescription}
+
+with a projection map $\pi:\dtotal\rightarrow\dbase$ that connects every point in \dtotal\ to a point in \dbase. 
 \begin{equation}
@@
 \end{equation}
+
+As indicated by $\hookrightarrow$, the fiber space \dfiber\ is embedded inside the total space \dtotal\.
+This is illustrated in \autoref{fig:related-work:fiber-bundle}, wherein the data lives in the manifold \dtotal.
+Each data point has two dimensions, represented as the square fiber \dfiber.
+While the data is embedded in \dtotal, it's continuity is that each point lies along a line, represented by the base space \dbase.
+
+\subsection{Sheaf Maps \sheaf}
+We can use sheafs to ensure continuity even when the data is broken up. Is the glue rules -> we can haz parallism. 
+\begin{equation}
+  \vartist: \sheaf(\dtotal) \rightarrow \sheaf(\gtotal)
+  \label{eq:math:artist:artist}
+\end{equation}
+
+\begin{figure}[h!]
+  \includegraphics[width=\columnwidth]{sheaf.png}
+  \caption{}
+\end{figure}
 
@@
 
-\section{Modeling Visualization Stages: Fiber bundles}
+\subsubsection{Category Theory}
+In this work, we propose that equivariance constraints can be expressed using category theory.
+Vickers et.
+al provide a brief introduction to category theory for visualization practitioners \cite{vickersUnderstandingVisualizationFormal2013}, but their work focuses on data, representation, and evocation, while this paper is aims to provide guidance on how the map from data to representation should be implemented.
 
+\section{Artist} 
 %%- brief intro to artist in most simple form $\vartist \dtotal \rightarrow \gtotal$ 
-The \textcolor{artist}{Artist $\mathcal{\vartist}$} is a transformation from
+The \textcolor{artist}{Artist $\mathcal{\vartist}$} is a transformation from a 
+%% flesh out categorical framing of artist, walk through how eq 16 is part of implementing the artists
+%% https://github.com/story645/proposal/blob/main/notes/meetings/2021_08_30.md
+
+%% domain of A is category E, with sheaves on E7   
+%% range of A is category H, sheaves on H
+%% category bundles, functions are bundles maps, 
+%% subset/restriction is a type of bundle map - bundle over subset is bundle over whole thing, which induces map on section which goes other way which is restriction
+%% bundle map induces a sheaf map, fibers themselves are categories, map from fiber to fiber that's legit in F
+%%% what is the set up?
+category of bundles, bundle maps - fiber have some structure/any structure/no structure - bundle maps should be functors, A is gonna be a functor, H is a category of bundles + sheaves of bundles, A is a functor
+%%% we claim that for visualizations, A decomposes as 
+%%% nu bundle map on bundle + functor on fibers, Q takes you to H
+%%% K to S is pullback of deform retract, A is functor on imageA in H
+
+\begin{equation}
+  \vartist: \dtotal \rightarrow \gtotal
+  \label{eq:math:artist:artist}
+\end{equation}
 
@@
 \begin{figure}[h!]
-  \includegraphics[width=2.5in]{k_qspace.png}
-  \caption{this is gonna be replaced w/ more concrete}
+  \includegraphics[width=\columnwidth]{fiberbundle.png}
+  \caption{
+
+  \note{replace with more concrete}}
 \end{figure}
@@
 
-\paragraph{Structured Data: Section \dsection}
+\subsubsection{Structured Data: Section \dsection}
 \begin{equation}
@@
 
-\paragraph{Structure: Continuity and Equivariant Actions}
+\subsubsection{Structure: Continuity and Equivariant Actions}
 \begin{equation}
@@
 
-\subsubsection{Continuity of Data: sheaf \sheaf}
-\begin{figure}[h!]
-  \includegraphics[width=2.5in]{sheaf.png}
-  \caption{}
-\end{figure}
-
 \subsection{Graphic Bundle \gtotal} %probably doesn't all need to be subsections, just figure + short description
@@
 \end{figure}
+
 \begin{equation}
@@
 
-\subsection{Artist}
-%% flesh out categorical framing of artist, walk through how eq 16 is part of implementing the artists
-%% https://github.com/story645/proposal/blob/main/notes/meetings/2021_08_30.md
-\begin{equation}
-  \vartist: \sheaf(\dtotal) \rightarrow \sheaf(\gtotal)
-  \label{eq:math:artist:artist}
-\end{equation}
-
-\section{Union of Artists}
+\subsection{Union of Artists}
+\label{sec:artist:union}
 \begin{figure}[!h]
 \centering
-\subfloat[Artists with shared $\mu_i$ renderered correctly]{\includegraphics[width=1.2in]{combined_artist.png}%
+\subfloat[Union of H]{\includegraphics[width=.4\columnwidth]{exploding_artist.png}%
 \label{fig_first_case}}
 %\hfil
-\subfloat[Artists without shared $\mu_i$]{\includegraphics[width=1.2in]{exploding_artist.png}%
+\subfloat[Artist with shared/harmonized \vsection]{\includegraphics[width=.4\columnwidth]{combined_artist.png}%
 \label{fig_second_case}}
@@
 
-\section{Construction and formal properties of Artists}
+\section{Constraints: Construction and formal properties of Artists}
+Identify the constraints, category theory makes it easier to spell out these constraints, this map has to be equivariant
+ordered data & data on trees, there is a monoid that acts on them & is not clear that if your map acts on them there will be a nice map on ordered
+why category theory? 
+- F->V->D
+-> F->V is 1:1 equivariance
+-> V->D is glyph equivariance
+
+What are the formal properties and constraints on xi, nu, and qhat
+
+nu = any function f:E\rightarrow E, equiv V\rightarrow V
+f: F_i\rightarrow F_i, v_i: F_i\right P_i,
+
+define category of structure preserving transformations on F
+  - valid maps F->F
+  - user in setting up fibers, sets up what are valid transforms/morphisms
+  - nu is a functor
+
+%xi is a deform retract, continuous, surjective
+
+%Q: \Gamma (V) \Gamma(H) - is defined on functions not on outputs
+
+%Qhat to Q can't break continuity when it goes from Q to Q hat
+
+%% xi is screen data binding,  map of glyph back to ideal platonic mark
+
+%% continuity is in the sheafs -> requiring that we have a sheaf map and that nu induces a sheaf map, nu can also act on individual components
+
+%% nus are a sheaf map, act on indivdual fibers, induce a sheaf map on E to V, 
+%% pull sheafs and sections up to intro, Q is also sheaf map on S 
+%%% Q is a sheaf map, category of structure preserving maps, Q is a functor at least on the image
 
@@
 
-\subsection{Data to Representation: \vchannel} %look at what K&S call these stages
+\subsection{Data to Visual Encodings: \vchannel} %look at what K&S call these stages
+Visual Channel Encodings
 \begin{figure}[!h]
@@
 We enforce the equivariance constraint
+
 \begin{equation}
@@
 
-\subsubsection{Representation to Visual: \vmark} % look at what K&S call these stages
-
+\subsubsection{Visual to Graphic: \vmark} % look at what K&S call these stages
+Visual encodings to something like marks 
 \begin{figure}[!h]
-  \includegraphics[width=2.5in]{diff_type_q.png}
+  \includegraphics[width=\columnwidth]{diff_type_q.png}
   \caption{rework this as a commutative box w/ the r in E row associated w/ this qhat(k)}
@@
   \includegraphics[width=\columnwidth]{path_of_q.png}
-    \caption{add in xi!}
+    \caption{\note{add in xi!}}
   \label{fig:api}
 \end{figure}
-We implement the \textbf{arrows} in \autoref{fig:api}
+We implement the \textbf{arrows} in \autoref{fig:api}. \mintinline{python}{axesArtist} is a parent artist that acts as a screen. This allows for the composition described in \autoref{sec:artist:union}
 
-\begin{minted}[]{python}
+\subsection{\vartist}
+\begin{minted}{python}
+for local_tau in axesArtist.artist.data.query(screen_bounds, dpi):
+    mu = axesArtist.artist.graphic.mu(local_tau)
+    rho = axesArtist.artist.graphic.qhat(**mu)
+    H = rho(renderer)
+\end{minted}
 
-\end{minted}
+where the artist is already parameterized with the \vindex\ functions and which fibers they are associated to:
+
+\begin{min
+
+\subsubsection{\vindex}
+\subsubsection{\vchannel}
+\subsubsection{\vmarkd}
 

### 2021-09-17 paper/main.tex
sections from notes

@@
 % *** MATH PACKAGES ***
-\usepackage{amsmath}
 %% with other math-related packages, you may want to disable it.
-\usepackage{amsmath,amsfonts,amssymb,eulervm,xspace}
-%\usepackage{mathrsfs} % math script fonts
-
+\usepackage{amsmath, amsthm, amsfonts,amssymb,eulervm,xspace, mathtools}
+\usepackage{mathrsfs} % math script fonts
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}[section]
+\theoremstyle{remark}
+\newtheorem{example}{Example}[section]
 % *** SPECIALIZED LIST PACKAGES ***
@@
 \subsubsection{Fiber Bundles}
+\label{sec:related-work:fiber-bundles}
 \begin{figure}[h!]
@@
 
-such that the empty set and $X$ are in the collection of open sets $\mathscr{T}$, the union of elements in $\mathscr{T}
+such that the empty set and $X$ are in the collection of open sets $\mathscr{T}$, the union of elements in $\mathscr{T}$
 
@@
 As indicated by $\hookrightarrow$, the fiber space \dfiber\ is embedded inside the total space \dtotal\.
-This is illustrated in \autoref{fig:related-work:fiber-bundle}, wherein the data lives in the manifold \dtotal.
-Each data point has two dimensions, represented as the square fiber \dfiber.
-While the data is embedded in \dtotal, it's continuity is that each point lies along a line, represented by the base space \dbase.
+This is illustrated in \autoref{fig:related-work:fiber-bundle}, wherein values lives in the fiber $\dfiber\subseteq\dtotal$.
+In this example, the fiber \dfiber\ is the cartesian product of two sets $F_{0}\times F_{1}$ where each fiber $F_{i}$ is ....
+\begin{equation}
+  formal equation of the fiber?
+  \label{eq:related-work:fiber}
+\end{equation}
+
+While the values are embedded in \dfiber, the continuity of the data is modeled as the base space \dbase, which is the quotient space \cite{QuotientSpaceTopology2020}
+\begin{equation}
+formal math of the base space as equivalence class for F_[k]
+\end{equation}
+which means that every point in $\dfiber_{k}$ maps to a point $k \subset K$.
+
+\begin{equation}
+  \begin{tikzcd}
+      \dfiber \arrow[r, hook] & \dtotal \arrow[d, "\pi"'] \\
+                        & \dbase \arrow[u, "\dsection"', bend right]
+  \end{tikzcd}
+\end{equation}
 
@@
 
+We propose that the artist $\mathcal{\vartist}$ is a function
+
 \begin{equation}
@@
 
-\subsection{Data Bundle \dtotal}
+\begin{definition}The category $\mathcal{\dbase}$ consists of 
+  \begin{enumerate}
+    \item \textit{objects} all open sets $U_{i} \in \{U\}$ in \dbase, including the empty set $\varnothing$ and the maximum set \dbase
+    \item \textit{morphisms} inclusion maps $\iota: U_{i} \hookrightarrow U_{j}$ 
+  \end{enumerate} 
+\end{definition}
+
+\begin{definition} The category $\mathcal{\E}$ consists of 
+  \begin{enumerate}
+    \item\textit{objects} bundles \dtotal
+    \item\textit{morphisms} inclusion maps: $\iota: E_i \hookrightarrow E_j$
+  \end{enumerate}
+\end{definition}
+
+\begin{definition} 
+\end{definition}
+
+\subsection{Data}
+We model data as sections of a fiber bundle $(\dtotal, \pi, \dbase, \dfiber)$.
+We encode the continuity of the data as the \textcolor{base}{base space} $\dbase$.
+\note{does this go in intro?
+This is mostly important for data, and we kinda use it for visual intermediary but not really for H} The \textcolor{fiber}{fiber space} \dfiber\ is the space of all possible values of the data \autoref{eq:related-work:fiber}.
+To allow us to map multiple fields of the same datatype to distinct fibers, we adopt Spivak's formulation of the fiber as a (column name, data domain) simple schema \cite{spivakSIMPLICIALDATABASES,spivakDatabasesAreCategories2010}.
+Spivak models the fiber space as a fiber bundle with sections $\sigma:C \rightarrow DT$ that take as input column name $c \in C$ and returns as output the set of values associated with the type of the field $T \in DT$, for example $\mathbb{R}$ for \texttt{floats}.
+In the paper, the $\dfiber$ is the cartesian cross product of these datatype domains \textbf{DT}.
+We model the data as the section \dsection because, as described in \autoref{sec:related-work:fiber-bundles}, the section is the map from the indexing space \dbase to the space of possible data values \dfiber.
+Spivak's notation also allows for associating elements in a section with the field it comes from.
+This allows for field based selection of values, while inclusion allows for continuity (index) based selections.
+
 \begin{figure}[h!]
   \includegraphics[width=\columnwidth]{fiberbundle.png}
-  \caption{
+  \caption{\note{replace with more concrete}}
+  \label{fig:artist:data}
+\end{figure}
 
-  \note{replace with more concrete}}
-\end{figure}
-We use topology to model
+One example of encoding data as a section of a fiber bundle is illustrated in \autoref{fig:artist:data}.
+In this example, the data is a \note{not totally decided yet} table of weather station data.
+Here we are interested in the time series of temperature values in the data, so we encode the continuity as the 1D interval \dbase.
+In this multivariate data set, the fields we want to visualize are \texttt{time}, \texttt{temperature}, and \texttt{station}.
+The fiber space \dfiber\ is the cartesian cross product of the fibers of each field
+\begin{equation*}
+  \dfiber = \dfiber_{time} \times \dfiber_{temperature} \times \dfiber_{station}
+\end{equation*}
+where each field fiber is the set of values that are valid for the field: 
+\begin{align*}
+  \dfiber_{time} &= \mathbb{R} \\
+  \dfiber_{temperature} &= \mathbb{R}\\
+  \dfiber_{station} &= \{s_0, s_1, \cdots, s_i, \cdots, s_n\} 
+\end{align*}
 
-The The continuity of the data is encoded in the base space \dbase.
+The section \dsection\ is the abstraction of the data being visualized. The section at a point $k \in K$ in the base space returns a value from each field in the fiber.
+\begin{equation*}
+  \tau(k) = ((time, t_k); (precipitation, p_k); (station\;name, n_k))
+\end{equation*}
 
-The properties of the variables are encoded in the fiber space \dfiber. The \textcolor{fiber}{fiber} is a topological space
-
-Spivak provides notation for describing the set of all possible values $U_{\sigma}$ of a single column with name $c$ and type $T$.
-This set of values is the fiber "F" $F = U_{\sigma}(c)$.
-When data is multivariate, the fiber F is the cartesian product of each columns fiber space (5)
-
-\subsubsection{Structured Data: Section \dsection}
-\begin{equation}
-  \begin{tikzcd}
-      \dfiber \arrow[r, hook] & \dtotal \arrow[d, "\pi"'] \\
-                        & \dbase \arrow[u, "\dsection"', bend right]
-  \end{tikzcd}
-\end{equation}
-
-\subsubsection{Structure: Continuity and Equivariant Actions}
-\begin{equation}
-  \begin{tikzcd}
-      \dfiber_i \arrow[d, "m_j"'] \arrow[rd, "m_j\;\circ\; m_k"] & \\
-      \dfiber_i \arrow[r, "m_k"']  & \dfiber_i
-  \end{tikzcd}
-\end{equation}
-
-\subsection{Graphic Bundle \gtotal} %probably doesn't all need to be subsections, just figure + short description
-\begin{figure}[h!]
-  \includegraphics[width=\columnwidth]{render.png}
-\end{figure}
+\subsection{Graphic} 
+The object of the graphic category $\mathcal{\gtotal}$ is the fiberbundle \gtotal. The bundle \gtotal\ has the same structure as the data bundle \dtotal
 
@@
 \end{equation}
-\paragraph{Continuity: Base Space \gbase}
-\paragraph{Equivariance: Fiber \gfiber}
-\paragraph{Structured Data: Visual \gsection}
+with a fiber space \gfiber\ embedded in the total space \gtotal\ and a section map $\gsection:\gbase\rightarrow\gtotal$.
+The attributes of the graphic bundle $(\gtotal, \pi, \gfiber, \gbase)$ encode attributes of the graphic and display space
+
+\begin{LaTeXdescription}
+\item [\textcolor{base}{base space} \gbase] continuity of display space (e.g. screen, 3D print)
+\item [\textcolor{fiber}{fiber space} \gfiber] attributes of the display space (e.g a pixel = (x,y,r,g,b,a))
+\end{Latex}
+\item [\textcolor{section}{section} \gsection] graphic generating function
+\end{LaTeXdescription}.
+
+We use a fiber bundle because it allows us to generalize implementation details.
+In this work, \gtotal\ assumes the the display is an idealized 2D screen, but there is no barrier to encoding for example a 3D print surface.
+In the same vein, \gsection\ is an abstraction of rendering; it can be a specification such as PDF\cite{bienz1993portable}, SVG\cite{quintScalable2003}, ...
+or a rendering engine such as cairo or agg,
+
+\begin{figure}[h!]
+  \includegraphics[width=\columnwidth]{render.png}
+  \caption{}
+  \label{fig:artist:graphic}
+\end{figure}
+
+\begin{example}
+As illustrated in \autoref{fig:artist:graphic}, 
+\end{example}
+
+\begin{definition}[Category $\mathcal{\gtotal}$]
+
+\end{definition}
 
@@
 \section{Constraints: Construction and formal properties of Artists}
-Identify the constraints, category theory makes it easier to spell out these constraints, this map has to be equivariant
-ordered data & data on trees, there is a monoid that acts on them & is not clear that if your map acts on them there will be a nice map on ordered
-why category theory? 
-- F->V->D
--> F->V is 1:1 equivariance
--> V->D is glyph equivariance
+\begin{figure*}[h!]
+  \includegraphics[width=\linewidth]{q.png}
+  \caption{}
+  \label{fig:constraints:q-overall}
+\end{figure*}
 
-What are the formal properties and constraints on xi, nu, and qhat
+\begin{definition} Following from \label{def:artist:} the artist \vartist is a special case of  natural transformations
+  \begin{equation}
+    \vartist: \Gamma(\dtotal, \dbase)\nrightarrow \Gamma(\gtotal, \gbase)
+  \end{equation}
+  between the profunctor objects $\Gamma(E),\Gamma(H)$ that preserves continuity and equivariance.
+\end{definition}
+  Natural transformations are functions that map functions between categories (functors) to other functors in a structure preserving way \cite{riehlCategoryTheoryContext, spanier1989algebraic, fongInvitationAppliedCategory2019}.
 
-nu = any function f:E\rightarrow E, equiv V\rightarrow V
-f: F_i\rightarrow F_i, v_i: F_i\right P_i,
-
-define category of structure preserving transformations on F
-  - valid maps F->F
-  - user in setting up fibers, sets up what are valid transforms/morphisms
-  - nu is a functor
-
-%xi is a deform retract, continuous, surjective
-
-%Q: \Gamma (V) \Gamma(H) - is defined on functions not on outputs
-
-%Qhat to Q can't break continuity when it goes from Q to Q hat
-
-%% xi is screen data binding,  map of glyph back to ideal platonic mark
-
-%% continuity is in the sheafs -> requiring that we have a sheaf map and that nu induces a sheaf map, nu can also act on individual components
-
-%% nus are a sheaf map, act on indivdual fibers, induce a sheaf map on E to V, 
-%% pull sheafs and sections up to intro, Q is also sheaf map on S 
-%%% Q is a sheaf map, category of structure preserving maps, Q is a functor at least on the image
-
-\begin{figure*}[]
-  \includegraphics[width=\linewidth]{q.png}
-\end{figure*}
+\begin{definition} We define the \textcolor{artist}{Artist} natural transformation \vartist\ as the tuple $(\vindex, \vchannel, \vmark, \dtotal, \vtotal, \gtotal)$ where
+  \begin{enumerate}
+    \item \vindex\ is a continuity preserving functor
+    \item \vchannel\ and \vmark\ are equivariance preserving functors
+    \item \dtotal, \vtotal, and \gtotal\ are profunctor categories of sheafs (\autoref{def:artist:profunctor})
+  \end{enumerate}
+\end{definition}
 

### 2021-09-20 paper/main.tex
sheaf notes

@@
 
-A model that allows for expressing data continuity in a general flexible way allows us to express, and therefore be faithful to, the multivariate consistency constraints described by Qu and Hullman\cite{hullmanKeeping2018}.
-A general data model also allows for consistent adaptation to modern data needs, such as complex, metadata rich, distributed, or streaming data.
-The mathematical theory of fiber bundles provides one such abstraction that can express complicated dimensionality and continuity without being tied to any one data container type, as proposed by Butler, Bryson, and Pendley\cite{butlerVisualizationModelBased1989,butlerVectorBundleClassesForm1992}.
+The model described in this work provides a method for expressing different types of continuities and nested continuities using the same model.
+We obtain this generality by using the mathematical theory of fiber bundles as the basis of our abstraction, as proposed by Butler, Bryson, and Pendley\cite{butlerVisualizationModelBased1989,butlerVectorBundleClassesForm1992}.
 In this paper, we build on their work that proposes using topological spaces to represent different properties of data.
@@
 \subsection{Equivariance}
-What is it?
+Many visualization theorists have expressed the notion that scales, which are maps from data to visual representation, 
+Hullman and Qu scales \cite{hullmanKeeping2018}
 
@@
 
-We propose that the artist $\mathcal{\vartist}$ is a function
+We propose that the artist $\mathcal{\vartist}$ is a morphism of presheaves defined to be a natural transformation (\autoref{sec:related-work:category-theory})
+\note{ommutative diagram maybe?}
+
+We can then explain inclusion & continuity as special cases, preservation as consequence of this.
+Generalize inclusion of open sets and restriction (how topology is put together) and equivariance on the fiber in terms of natural transformations <- condition we want to satisfy is natural transformation.
+Constraint on artist is that it's a natural transformations, will construct that way, will show
+
+1) bundle E restricted to U is presheaf,
+
+F_1: \mathcal{K_1} \rightarrow Set <-presheaf is the functor F_1
+any map on base spaces will give us a map on functors, sections will pull back
+
+artist w. particular base space w/ particular,
+
+How do you define the category E
+objects: F_1 
+  - presheaf: is it defined on fixed K and E0 (this provides the Equivariance on fibers + continuity), declaring it as E->V is natural transformation
+  - presheaves: is it defined on arbitrary K and Eo
+morphisms: inclusion -> these go back and forth, hence is a profunctor
+
+fiber preserving maps 
+fiber as category -> map preserves the structure, monotonic etc
+
+How are you defining the artist? fix E0 and K and nu and q
+
+same sheaf w/ arrow to itself , requires that all bundles be morphisms of a bundle to itself
+  - morphisms are actions, bundle to bundle which when restricted to bundle are morphisms of bundle to itself
+
+- presheaf is just a functor
+
+F is object in category that just has itself as only object 
+if we have a bundle transfrom F_0 to F_0, then we have a natural transfrom F_0 to F_0
+
+same thing has to happen on V side, declare which bundle maps are OK on which side, nu goes from one to the other, preserving the structure3 , doesn't have to have the same dimension,
+- we don't care about a change in fiber dim - nu defines type it eats, which is static and defined relative to the fibers - can't change K, or E, nu is fixed on F/E, V and E can be different dimensions
+
+category w/ one functor E_0, morphisms are functor to itself, f is data to data, g is graphic to graphic, nu(f) \rightarrow (g), when we get to H bundle, 
+V over K -> H over S,
+
+nu is bundle to bundle (natural transformation)
+Q is a natural transformation - proposition, map that satisifies this condition, v \rightarrow h v,
 
@@
 
-\begin{definition}The category $\mathcal{\dbase}$ consists of 
+\begin{definition} A presheaf is a category with extra structure - definition
+
+\end{definition}
+
+\begin{definition} a morphism of presheaves is defined to be a natural transform
+
+\end{definition}
+
+1) start w/ presheaf definition and explain K etc in context, $\Gamma(U, E)$ is an object in set, presheaf(U)->\Gamma(E,U)
+
+\begin{definition} 
   \begin{enumerate}
@@
 We use a fiber bundle because it allows us to generalize implementation details.
-In this work, \gtotal\ assumes the the display is an idealized 2D screen, but there is no barrier to encoding for example a 3D print surface.
-In the same vein, \gsection\ is an abstraction of rendering; it can be a specification such as PDF\cite{bienz1993portable}, SVG\cite{quintScalable2003}, ...
-or a rendering engine such as cairo or agg,
+In this work, \gtotal\ assumes the the display is an idealized 2D screen and \gsection\ is an abstraction of rendering.
+For example, \gsection\ can be a specification such as PDF\cite{bienz1993portable}, SVG\cite{quintScalable2003} or an OpenGL scene graph\cite{CarsonOpenGL1997}, or a rendering engine such as Cairo\cite{CairographicsOrg} or AGG\cite{shemanarevAntiGrainGeometry}.
 
@@
 \subsection{Data to Visual Encodings: \vchannel} %look at what K&S call these stages
-Visual Channel Encodings
+\note{needs a sentence on why this is broader than scales}
 \begin{figure}[!h]

### 2021-09-23 paper/main.tex
sheafs

@@
 \end{figure}
+We propose that a robust model of continuity, which is how the elements in a dataset are connected to each other, provides a way to develop library components that \note{finish this sentence in a sensible way}
 
-Continuity is a representation of how the elements in a dataset are connected to each other.
 For example, in \autoref{fig:related-work:continuity}, each station record in the table is independent of the others; therefore, the continuity of the table is discrete.
@@
 \subsection{Equivariance}
-Many visualization theorists have expressed the notion that scales, which are maps from data to visual representation, 
-Hullman and Qu scales \cite{hullmanKeeping2018}
+\label{sec:related-work:equivariance}
+When introducing the retinal variables, Bertin informally specifies that continuity is preserved in the mark and defines equivariance constraints in terms of data and visual variables being selective, associative, ordered, or quantitative\cite{bertinSemiologyGraphicsDiagrams2011a}.
+In the \textit{A Presentation Tool}(APT) model, Mackinlay embeds the continuity constraint in the choice of visualization type and generalizes the equivariance constraint to preserving a binary operator from one domain to another.
+The algebraic model of visualization\cite{kindlmannAlgebraicProcessVisualization2014}, proposed by Kindlmann and Scheidegger, restricts equivariance to invertible transformations.
+
+\note{something about how structure of data field serves as basis of equivariance in viz literature}
+Besides expressing continuities, fiber bundles also provide us a way to richly express the structure of the fields in the data.
+Specifically, we do this by adopting Spivak's formulation of the fiber as a (column name, data domain) simple schema \cite{spivakSIMPLICIALDATABASES,spivakDatabasesAreCategories2010}.
+Spivak formally maps column names and field types to the set of values associated with the field type, for example $\mathbb{R}$ for a \texttt{float} column named \textit{temperature}.
+In the paper, the $\dfiber$ is the cartesian cross product of the fibers for each column.
 
@@
 \section{Artist} 
-%%- brief intro to artist in most simple form $\vartist \dtotal \rightarrow \gtotal$ 
-The \textcolor{artist}{Artist $\mathcal{\vartist}$} is a transformation from a 
-%% flesh out categorical framing of artist, walk through how eq 16 is part of implementing the artists
-%% https://github.com/story645/proposal/blob/main/notes/meetings/2021_08_30.md
-
-%% domain of A is category E, with sheaves on E7   
-%% range of A is category H, sheaves on H
-%% category bundles, functions are bundles maps, 
-%% subset/restriction is a type of bundle map - bundle over subset is bundle over whole thing, which induces map on section which goes other way which is restriction
-%% bundle map induces a sheaf map, fibers themselves are categories, map from fiber to fiber that's legit in F
-%%% what is the set up?
-category of bundles, bundle maps - fiber have some structure/any structure/no structure - bundle maps should be functors, A is gonna be a functor, H is a category of bundles + sheaves of bundles, A is a functor
-%%% we claim that for visualizations, A decomposes as 
-%%% nu bundle map on bundle + functor on fibers, Q takes you to H
-%%% K to S is pullback of deform retract, A is functor on imageA in H
-
-We propose that the artist $\mathcal{\vartist}$ is a morphism of presheaves defined to be a natural transformation (\autoref{sec:related-work:category-theory})
-\note{ommutative diagram maybe?}
-
-We can then explain inclusion & continuity as special cases, preservation as consequence of this.
-Generalize inclusion of open sets and restriction (how topology is put together) and equivariance on the fiber in terms of natural transformations <- condition we want to satisfy is natural transformation.
-Constraint on artist is that it's a natural transformations, will construct that way, will show
-
-1) bundle E restricted to U is presheaf,
-
-F_1: \mathcal{K_1} \rightarrow Set <-presheaf is the functor F_1
-any map on base spaces will give us a map on functors, sections will pull back
-
-artist w. particular base space w/ particular,
-
-How do you define the category E
-objects: F_1 
-  - presheaf: is it defined on fixed K and E0 (this provides the Equivariance on fibers + continuity), declaring it as E->V is natural transformation
-  - presheaves: is it defined on arbitrary K and Eo
-morphisms: inclusion -> these go back and forth, hence is a profunctor
-
-fiber preserving maps 
-fiber as category -> map preserves the structure, monotonic etc
-
-How are you defining the artist? fix E0 and K and nu and q
-
-same sheaf w/ arrow to itself , requires that all bundles be morphisms of a bundle to itself
-  - morphisms are actions, bundle to bundle which when restricted to bundle are morphisms of bundle to itself
-
-- presheaf is just a functor
-
-F is object in category that just has itself as only object 
-if we have a bundle transfrom F_0 to F_0, then we have a natural transfrom F_0 to F_0
-
-same thing has to happen on V side, declare which bundle maps are OK on which side, nu goes from one to the other, preserving the structure3 , doesn't have to have the same dimension,
-- we don't care about a change in fiber dim - nu defines type it eats, which is static and defined relative to the fibers - can't change K, or E, nu is fixed on F/E, V and E can be different dimensions
-
-category w/ one functor E_0, morphisms are functor to itself, f is data to data, g is graphic to graphic, nu(f) \rightarrow (g), when we get to H bundle, 
-V over K -> H over S,
-
-nu is bundle to bundle (natural transformation)
-Q is a natural transformation - proposition, map that satisifies this condition, v \rightarrow h v,
+The \textcolor{artist}{Artist $\mathcal{\vartist}$} is a transformation from data space to graphic.
 
@@
 
-\begin{definition} A presheaf is a category with extra structure - definition
+We propose that the artist $\mathcal{\vartist}$ is a morphism of presheaves defined to be a natural transformation (\autoref{sec:related-work:category-theory})
+\note{commutative diagram maybe?}
 
+We can then explain inclusion & continuity as special cases, preservation as consequence of this.
+Generalize inclusion of open sets and restriction (how topology is put together) and equivariance on the fiber in terms of natural transformations <- condition we want to satisfy is natural transformation.
+Constraint on artist is that it's a natural transformations, will construct that way, will show
+
+To show that the artist is a natural transformation, we first define the categories $\mathcal{E}$ and $\mathcal{H}$.
+In order to define these categories, we introduce the category of open sets $\mathcal{K}$
+
+\begin{definition} The category $\mathcal{K}$ consists of 
+  \begin{itemize}
+    \item \textit{objects} all open sets $U_{i} \in \{U\}$ in \dbase, including the empty set $\varnothing$ and the maximum set \dbase
+    \item \textit{morphisms} inclusion maps from a subspace to a larger space $\iota: U_{i} \hookrightarrow U_{j}$, which is equivalent to $U_{i} \subseteq U_{j}$ and  
+  \end{itemize} 
 \end{definition}
 
-\begin{definition} a morphism of presheaves is defined to be a natural transform
+The open set objects $U_{i}$ are used to track subsets of the data defined globally on \dbase. For example,
 
+\begin{definition} For each open set object $U_{i}$ of category $\mathcal{K}$, there is a set $\Gamma(U_i, \dtotal)$.
+The elements of this set are the sections $\{\dsection^{0},\cdots, \dsection^{i}, \dsection^{n}\}$ of \dtotal\ over open set $\U_{i}$.
+The sections $\Gamma(\dbase, \dtotal)$ are called the global sections of $\dtotal$.
+\cite{SheafMathematics2021,spanier1989algebraic}.
 \end{definition}
 
-1) start w/ presheaf definition and explain K etc in context, $\Gamma(U, E)$ is an object in set, presheaf(U)->\Gamma(E,U)
-
-\begin{definition} 
-  \begin{enumerate}
-    \item \textit{objects} all open sets $U_{i} \in \{U\}$ in \dbase, including the empty set $\varnothing$ and the maximum set \dbase
-    \item \textit{morphisms} inclusion maps $\iota: U_{i} \hookrightarrow U_{j}$ 
-  \end{enumerate} 
-\end{definition}
+Each section is how we represent a unique dataset, where the set of sections $\Gamma$ is the set of all datasets that have the same data fields \dfiber\ and continuity \dbase\.
+For example, database tables that have the same schema and discrete index \cite{spivakSIMPLICIALDATABASES}.
 
 \begin{definition} The category $\mathcal{\E}$ consists of 
-  \begin{enumerate}
-    \item\textit{objects} bundles \dtotal
+  \begin{itemize}
+    \item\textit{objects} the presheaf $F:
+
     \item\textit{morphisms} inclusion maps: $\iota: E_i \hookrightarrow E_j$
-  \end{enumerate}
+  \end{itemize}
 \end{definition}
@@
 \note{does this go in intro?
-This is mostly important for data, and we kinda use it for visual intermediary but not really for H} The \textcolor{fiber}{fiber space} \dfiber\ is the space of all possible values of the data \autoref{eq:related-work:fiber}.
-To allow us to map multiple fields of the same datatype to distinct fibers, we adopt Spivak's formulation of the fiber as a (column name, data domain) simple schema \cite{spivakSIMPLICIALDATABASES,spivakDatabasesAreCategories2010}.
-Spivak models the fiber space as a fiber bundle with sections $\sigma:C \rightarrow DT$ that take as input column name $c \in C$ and returns as output the set of values associated with the type of the field $T \in DT$, for example $\mathbb{R}$ for \texttt{floats}.
-In the paper, the $\dfiber$ is the cartesian cross product of these datatype domains \textbf{DT}.
+The \textcolor{fiber}{fiber space} \dfiber\ is the space of all possible values of the data \autoref{eq:related-work:fiber}.
 We model the data as the section \dsection because, as described in \autoref{sec:related-work:fiber-bundles}, the section is the map from the indexing space \dbase to the space of possible data values \dfiber.
-Spivak's notation also allows for associating elements in a section with the field it comes from.
+Spivak's notation, as discussed in \autoref{sec:related-work:equivariance}, also allows for associating elements in a section with the field it comes from.
 This allows for field based selection of values, while inclusion allows for continuity (index) based selections.
@@
 
-We use a fiber bundle because it allows us to generalize implementation details.
+We represent the graphic output as a fiber bundle because it is an abstraction that is generalizable to various output mediums (screens, 3D prints) and types of graphics.
 In this work, \gtotal\ assumes the the display is an idealized 2D screen and \gsection\ is an abstraction of rendering.
@@
 \label{fig_second_case}}
-\caption{Simulation results for the network.}
+\caption{Coherance? }
 \label{fig_sim}
 \end{figure}
+\note{do we define coherant visualizations as one with shared inclusions?}
 
@@
 \subsection{Data to Visual Encodings: \vchannel} %look at what K&S call these stages
-\note{needs a sentence on why this is broader than scales}
+Visual encoding functions \vindex\ are functors because they are expected to preserve structure on the data and visual side, as discussed in \autoref{sec:related-work:equivariance}.
+
+ This constraint is expressed in our construction of \vchannel
+
 \begin{figure}[!h]
@@
 
+One advantage of expressing the data to visual encoding $\vchannel_i$ as a functor from data fiber $\dfiber_i$ to $\vfiber_i$ is that it lets us formally express that encodings are bound to data and therefore should be consistent across views, as proposed by Hullamn and Qu\cite{hullmanKeeping2018}.
+
 \subsubsection{Visual to Graphic: \vmark} % look at what K&S call these stages
@@
 \end{figure}
-We implement the \textbf{arrows} in \autoref{fig:api}. \mintinline{python}{axesArtist} is a parent artist that acts as a screen. This allows for the composition described in \autoref{sec:artist:union}
+We implement the \textbf{arrows} in \autoref{fig:api}.
+\mintinline{python}{axesArtist} is a parent artist that acts as a screen.
+This allows for the composition described in                                           \autoref{sec:artist:union}
 

### 2021-09-27 paper/main.tex
up to fiber space

@@
 \usepackage[nocompress]{cite}
+\usepackage[]{footmisc}
 \usepackage{hyperref} % autoref
@@
 Category theory provides a framework for separating the transformations implemented by visualization libraries from the various stages of visualization and therefore can be used to describe the constraints imposed on the library components \cite{wielsManagementEvolvingSpecifications1998,goguenCategoricalManifesto1991}.
-Well constrained modular components are inherently functional\cite{hughesWhyFunctionalProgramming1989}, and a functional framework yields a library implementation that is likely to be shorter, clearer, and more suitd to distributed, concurrent, and on demand tasks\cite{huHowFunctionalProgramming2015}.
+Well constrained modular components are inherently functional\cite{hughesWhyFunctionalProgramming1989}, and a functional framework yields a library implementation that is likely to be shorter, clearer, and more suited to distributed, concurrent, and on demand tasks\cite{huHowFunctionalProgramming2015}.
 Using this functional approach, this paper contributes a practical framework for decoupling data processing from visualization generation in a way that allows for modular visualization components that are applicable to a variety of data sets in different formats.
@@
 
-Often continuity is expressed in the choice of visual algorithm (visualization type), as explored in taxonomies by Tory and M\"{o}ller \cite{toryRethinkingVisualizationHighlevel2004} and Chi\cite{chiTaxonomyVisualizationTechniques2000}.
+ The preservation of continuity can be made explicit, as in the transformation of table to parallel coordinates in Ruchikachorn and Mueller \cite{ruchikachornLearningVisualizationsAnalogy2015}, but is often expressed implicitly in the choice of visual algorithm (visualization type), as explored in taxonomies by Tory and M\"{o}ller \cite{toryRethinkingVisualizationHighlevel2004} and Chi\cite{chiTaxonomyVisualizationTechniques2000}.
+
 For example, in \autoref{fig:related-work:visual-algorithm} the same table can be interpreted as a set of 1D continuous curves when visualized as a collection of line plots or as a 2D surface when visualized as an image.
@@
 
-Domain specific libraries can guarantee consistency because they have a single model of the data in their software design, as discussed in Heer and Agarwal \cite{HeerSoftware2006}'s survey of visualization software design patterns.
+Domain specific libraries can often guarantee consistency because they have a single model of the data in their software design, as discussed in Heer and Agarwal \cite{HeerSoftware2006}'s survey of visualization software design patterns.
 For example, the relational database is core to tools influenced by APT, such as Tableau\cite{StoltePolaris2002,hanrahanVizQL2006,MackinlayShowme2007} and the Grammar of Graphics\cite{wilkinsonGrammarGraphics2005} inspired ggplot\cite{wickhamGgplot2ElegantGraphics2016a}, Vega\cite{satyanarayanDeclarativeInteractionDesign2014} and Altair\cite{vanderplasAltairInteractiveStatistical2018}.
@@
 In this paper, we build on their work that proposes using topological spaces to represent different properties of data.
-Here we present a brief summary of topology, for more information see Hatcher\cite{hatcherAlgebraicTopology2002}, Munkres\cite{munkresElementsAlgebraicTopology1984}, and Bradley et.
+
+Specifically, Butler, Bryson, and Pendley suggest that fiber bundles be the basis of an abstract data model. Fiber bundles are a collection of topological spaces. 
+\begin{definition}
+A topological space $(X, \mathscr{T})$ is a set $X$ with a topology $\mathscr{T}$.
+Topologies are collections of open sets $U$ such that the empty set and $X$ are in the collection of open sets $\mathscr{T}$, the union of elements in $\mathscr{T}$ \note{finish out this definition}
+\end{definition}
+
+Here we present a very brief summary of topics, for more information see Hatcher\cite{hatcherAlgebraicTopology2002}, Munkres\cite{munkresElementsAlgebraicTopology1984}, and Bradley et.
 al.
 \cite{bradleyTopologyCategoricalApproach2020}.
-
-A topological space a topological space $(X, \mathscr{T})$ is a set $X$ with a topology $\mathscr{T}$. Topologies are collections of open sets
-
-such that the empty set and $X$ are in the collection of open sets $\mathscr{T}$, the union of elements in $\mathscr{T}$
-
-Specifically, Butler, Bryson, and Pendley suggest that fiber bundles be the basis of an abstract data model. Fiber bundles are a collection $(\dtotal, \dbase, \dfiber, \pi)$ of topological spaces
 
@@
 
-\subsection{Sheaf Maps \sheaf}
-We can use sheafs to ensure continuity even when the data is broken up. Is the glue rules -> we can haz parallism. 
-\begin{equation}
-  \vartist: \sheaf(\dtotal) \rightarrow \sheaf(\gtotal)
-  \label{eq:math:artist:artist}
-\end{equation}
-
-\begin{figure}[h!]
-  \includegraphics[width=\columnwidth]{sheaf.png}
-  \caption{}
-\end{figure}
-
 \subsection{Equivariance}
@@
 
-\begin{tikzcd}
-  data \arrow[r] \arrow[d, "function"] & representation \arrow[r] & visual\;stimulus \arrow[d, "visual\;equivalent\;to\;function"] \\
-  data \arrow[r]                       & representation \arrow[r] & visual\; stimulus                                          
-\end{tikzcd}
-
 \subsubsection{Category Theory}
+\label{sec:related-work:equivariance:category}
 In this work, we propose that equivariance constraints can be expressed using category theory.
@@
 \section{Artist} 
-The \textcolor{artist}{Artist $\mathcal{\vartist}$} is a transformation from data space to graphic.
+The \textcolor{artist}{Artist} $\mathcal{\vartist}$\footnote{We call this transformation the artist because the Artist object in Matplotlib\cite{hunterMatplotlib2DGraphics2007} is responsible for converting data to a prerender graphical object} is a transformation from data  $\mathcal{\dtotal}$ to graphic $\mathcal{\gtotal}$.
 
 \begin{equation}
-  \vartist: \dtotal \rightarrow \gtotal
+  \vartist: \mathcal{\dtotal} \rightarrow \mathcal{\gtotal}
   \label{eq:math:artist:artist}
@@
 
-We propose that the artist $\mathcal{\vartist}$ is a morphism of presheaves defined to be a natural transformation (\autoref{sec:related-work:category-theory})
-\note{commutative diagram maybe?}
+We formally define data and graphic as the categories $\mathcal{\dtotal}$ and $\mathcal{\gtotal}$.
+To explain these categories, we first define the category of open sets $\mathcal{\dbase}$ and the category of fiber spaces $\mathcal{\dfiber}$.
 
-We can then explain inclusion & continuity as special cases, preservation as consequence of this.
-Generalize inclusion of open sets and restriction (how topology is put together) and equivariance on the fiber in terms of natural transformations <- condition we want to satisfy is natural transformation.
-Constraint on artist is that it's a natural transformations, will construct that way, will show
-
-To show that the artist is a natural transformation, we first define the categories $\mathcal{E}$ and $\mathcal{H}$.
-In order to define these categories, we introduce the category of open sets $\mathcal{K}$
-
-\begin{definition} The category $\mathcal{K}$ consists of 
+\begin{definition} The category $\mathcal{\dbase}$ consists of 
   \begin{itemize}
@@
 
-The open set objects $U_{i}$ are used to track subsets of the data defined globally on \dbase. For example,
+The objects $U_{i}$ are a way to manage subsets of the data.
+For example, let there be a data value for each point along the interval $\dbase=\right[0,1\left]$.
+Since this function is continuous, there is a subset of the data defined along the open set $U_{a} = \right(0.2, 0.4\left)$.
+The open set $U_{a}$ provides a unique way of identifying that subset.
 
-\begin{definition} For each open set object $U_{i}$ of category $\mathcal{K}$, there is a set $\Gamma(U_i, \dtotal)$.
-The elements of this set are the sections $\{\dsection^{0},\cdots, \dsection^{i}, \dsection^{n}\}$ of \dtotal\ over open set $\U_{i}$.
+\note{maybe a small image here or when we introduce open sets showing what we mean?}
+
+\begin{definition} For each open set object $U_{i}$ of category $\mathcal{\dbase}$, there is a set $\Gamma(U_i, \dtotal)$.
+The elements of this set are the sections $\{\tau^{j}:U_i \rightarrow \dfiber_{U_{i}}\}_{j\in I}$ of \dtotal\ over open set $\U_{i}$.
 The sections $\Gamma(\dbase, \dtotal)$ are called the global sections of $\dtotal$.
@@
 
-Each section is how we represent a unique dataset, where the set of sections $\Gamma$ is the set of all datasets that have the same data fields \dfiber\ and continuity \dbase\.
-For example, database tables that have the same schema and discrete index \cite{spivakSIMPLICIALDATABASES}.
+Each section is how we represent a unique dataset, where the set of sections $\Gamma$ is the set of all datasets that have the same data fields in the \dfiber\ and continuity as encoded in\dbase\.
+For example, database tables that have the same schema and discrete index, as illustrated in Spivak \cite{spivakSIMPLICIALDATABASES}.
 
-\begin{definition} The category $\mathcal{\E}$ consists of 
-  \begin{itemize}
-    \item\textit{objects} the presheaf $F:
-
-    \item\textit{morphisms} inclusion maps: $\iota: E_i \hookrightarrow E_j$
-  \end{itemize}
+\begin{definition} The fiber space category $\mathcal{\dfiber}$ is a category of sets \textbf{Set}.
+This means its objects are sets, and its morphisms are total functions\cite{barrCategoryTheoryComputing}
 \end{definition}
 
-\begin{definition} 
+\begin{definition} The monad $T$ is [a,b,c]
+
 \end{definition}
+
+For example, these functions could be the binary operations, group actions, and measurement scales discussed in \autoref{sec:related-work:equivariance}.
+These functions could also be monoid actions, which are a way of applying partial order relations to data\cite{fongInvitationAppliedCategory2019}, such as to build multi-ranked indicators\cite{bruggemannRankingPrioritizationMultiindicator2011}
+}
+
+\begin{definition} The category $\mathcal{\dtotal}$ consists of 
+  \begin{itemize}
+    \item\textit{object} the space of continuous\footnote{or piecewise continuous}sections $\sheaf: U_{i} \rightarrow \Gamma(U_{i}, \dtotal) $ which is a sheaf\cite{spanier1989algebraic,SheafMathematics2021}.
+    \item\textit{morphisms} bundle maps E, E, fiber to itself and is the fiber map allowable, functions F to itself forms a monoid (one object category), F and it's allowable maps form a category, monoid could be monotonic functions, group actions, extending that to bundle E, interested in E to itself, which when restricted to fiber are category maps of fiber to itself, maps from sheaf to itself, that when restricted to fiber gives you fiber to itself, given bundle map E->E, then you can show that you have a sheaf map...
+Given E->E, and F->F, then there's a map from one sheaf to another.
+bundle map induces on level of sections induces on sheafs -> part of the definition of the presheaf
+
+  \end{itemize}
+\end{definition
+The functions on the data
+
+\note{equivariance is a special case of functions}
+\note{actions are a special case of actions}
+\begin{definition} category of bundle maps, which include category of fiber maps, what are allowable of fiber to itself, ex translation and scales only ..., objects are a fiber, maps are fiber to fiber defined as their own category, E to V fibers changes
+
+zooming is inclusion maps on both sides 
+\end{definition}
+
+The objects of the data category $\mathcal{\dtotal}$ and graphic category $\mathcal{\gtotal}$ are the sheaves $\Gamma(\dbase, \dtotal)$ and $\Gamma(\gbase, \gtotal)$.
+An alternate notation is $\sheaf(\dtotal)$ and $\sheaf(\gtotal)$.
+The artist is a morphism of sheafs $\sheaf(\dtotal)$ to $\sheaf(\gtotal)$, which means it is by definition a natural transform\cite{SheafMathematics2021}.
+As mentioned in \autoref{sec:related-work:equivariance:category}
+
+that the artist $\vartist$ transforms functions that act on data $f:\dtotal\rightarrow \dtotal$ to functions that act on graphics $\g: \gtotal\rightarrow \dtotal$ in a way in which the following diagram commutes
+\begin{tikzcd}
+  \mathcal{\dtotal} \arrow[r, "\vartist"] \arrow[d, "f"] & \mathcal{\gtotal} \arrow[d, "g"] \arrow[d, "\ca"] \\
+  \mathcal{\dtotal} \arrow[r, "\vartist"]                & \mathcal{\gtotal}                                
+\end{tikzcd}
+meaning that, as described in \autoref{sec:related-work:equivariance}, the artist \vartist is constructed such that $\vartist(f(\dtotal)) = g(\vartist(\dtotal))$, meaning a transformation of the data has a consistent equivalent transformation of the graphic.
+In \autoref{sec:artist:construction}, we show a method of constructing the artist such that continuity and equivariance are preserved.
+We then use this formulation to guide the development of visualization library components in \autoref{sec:case-study}.
 
@@
 We encode the continuity of the data as the \textcolor{base}{base space} $\dbase$.
-\note{does this go in intro?
-The \textcolor{fiber}{fiber space} \dfiber\ is the space of all possible values of the data \autoref{eq:related-work:fiber}.
+\note{does this go in intro?} The \textcolor{fiber}{fiber space} \dfiber\ is the space of all possible values of the data \autoref{eq:related-work:fiber}.
 We model the data as the section \dsection because, as described in \autoref{sec:related-work:fiber-bundles}, the section is the map from the indexing space \dbase to the space of possible data values \dfiber.
@@
 \item [\textcolor{fiber}{fiber space} \gfiber] attributes of the display space (e.g a pixel = (x,y,r,g,b,a))
-\end{Latex}
 \item [\textcolor{section}{section} \gsection] graphic generating function
@@
 \section{Constraints: Construction and formal properties of Artists}
+\label{sec:artist:construction}
 \begin{figure*}[h!]
@@
 
-\begin{definition} We define the \textcolor{artist}{Artist} natural transformation \vartist\ as the tuple $(\vindex, \vchannel, \vmark, \dtotal, \vtotal, \gtotal)$ where
+\begin{definition} We define the natural transformation \textcolor{artist}{Artist}  \vartist\ as the tuple $(\vindex, \vchannel, \vmark, \mathcal{\dtotal},\mathcal{\vtotal}, \mathcal{\gtotal})$ where
   \begin{enumerate}
@@
     \item \vchannel\ and \vmark\ are equivariance preserving functors
-    \item \dtotal, \vtotal, and \gtotal\ are profunctor categories of sheafs (\autoref{def:artist:profunctor})
+    \item $\mathcal{\dtotal}$, $\mathcal{\vtotal}$, and $\mathcal{\gtotal}$ are profunctor categories of sheafs
+    \item 
   \end{enumerate}
@@
 \end{equation}
+
+The functor $\vindex$ is a deformation retract, which means....\cite{hatcherAlgebraicTopology2002,spanier1989algebraic}
 
@@
 \section{Case Study}
+\label{sec:case-study}
 \begin{figure}[h!]

### 2021-10-04 paper/main.tex
F, K, M, A,

@@
 
+Throughout this paper, we denote fiber bundles with the tuple $(\textcolor{total}{total space}, \textcolor{base}{base space}, \pi, \textcolor{fiber}{fiber space})$
+
 \subsection{Equivariance}
@@
 \section{Artist} 
-The \textcolor{artist}{Artist} $\mathcal{\vartist}$\footnote{We call this transformation the artist because the Artist object in Matplotlib\cite{hunterMatplotlib2DGraphics2007} is responsible for converting data to a prerender graphical object} is a transformation from data  $\mathcal{\dtotal}$ to graphic $\mathcal{\gtotal}$.
+The \textcolor{artist}{Artist} $\mathcal{\vartist}$\footnote{We call this transformation the artist because the Artist object in Matplotlib\cite{hunterMatplotlib2DGraphics2007} is responsible for converting data to a prerender graphical object} is a functor from data $\mathcal{\dtotal}$ to graphic $\mathcal{\gtotal}$.}
 
@@
 We formally define data and graphic as the categories $\mathcal{\dtotal}$ and $\mathcal{\gtotal}$.
-To explain these categories, we first define the category of open sets $\mathcal{\dbase}$ and the category of fiber spaces $\mathcal{\dfiber}$.
+We develop these categories as analogues to fiber bundle $(\dtotal, \dbase, \pi, \dfiber)$, introduced in \autoref{sec:related-work:fiber-bundles}, and fiber bundle $(\gtotal, \gbase, \pi, \gfiber)$.
+Throughout this section, we assume that $\mathcal{E}$ is defined in terms of a fixed continuity and fiber space, for example a class of images or tables with the same schema.
+Similarly we assume that $\mathcal{H}$ is a stand in for a specific type of graphic, such as a scatter or line plot.
 
-\begin{definition} The category $\mathcal{\dbase}$ consists of 
+As introduced in \autoref{sec:related-work:fiber-bundles}, sets of open sets form the basis of topological spaces.
+We propose that we can express how the elements in our dataset are glued together by modeling that continuity as a topological space \dbase, and we express the constraints on \dbase by defining the analogues category $\mathcal{\dbase}$.
+
+\begin{definition} The category of open sets $\mathcal{\dbase}$ consists of
+  \label{def:category:K} 
   \begin{itemize}
-    \item \textit{objects} all open sets $U_{i} \in \{U\}$ in \dbase, including the empty set $\varnothing$ and the maximum set \dbase
+    \item \textit{objects} all open sets $U_{i} \in \{U\}$ in the topological space\cite{bradleyTopologyCategoricalApproach2020, } $(\dbase, \mathscr{T})$), including the empty set $\varnothing$ and the maximum set \dbase.
     \item \textit{morphisms} inclusion maps from a subspace to a larger space $\iota: U_{i} \hookrightarrow U_{j}$, which is equivalent to $U_{i} \subseteq U_{j}$ and  
@@
 
-\begin{definition} For each open set object $U_{i}$ of category $\mathcal{\dbase}$, there is a set $\Gamma(U_i, \dtotal)$.
-The elements of this set are the sections $\{\tau^{j}:U_i \rightarrow \dfiber_{U_{i}}\}_{j\in I}$ of \dtotal\ over open set $\U_{i}$.
-The sections $\Gamma(\dbase, \dtotal)$ are called the global sections of $\dtotal$.
-\cite{SheafMathematics2021,spanier1989algebraic}.
+\begin{definition} 
+  \label{def:category:F}
+  The category $\mathcal{\dfiber}$ is a subcategory of an arbitrary category $\mathcal{C}$.
+$\mathcal{\dfiber}$ is a monoidal category, meaning it has one object \texttt{\dfiber} and a functor\footnote{The functor $\otimes$ is a bifunctor because it takes as input the product of two categories \cite{nlab:bifunctor}} $\otimes:\mathcal{\dfiber}\times{\mathcal{\dfiber}}\rightarrow \mathcal{\dfiber}$\cite{barrCategoryTheoryComputing}.
 \end{definition}
 
+The object $\texttt{\dfiber}$ can be an object of any arbitrary category $\mathcal{C}$, which allows $\texttt{\dfiber}$ to encode most data fields.
+For example, a lists of strings is an instance of an object in \textbf{Set}, networks are an instance of \textbf{Graph}, and images are vector spaces which are a specific type of topological space \textbf{Top}.
+
+%Monoid b/c arrows form a Monoid, F + Monoid that acts on F
+
+%% F, M action on F, 
+%%% F \->E->K
+
+\begin{definition} For each open set object $U_{i}$ of category $\mathcal{\dbase}$, there is a set $\Gamma(U_i, \dtotal)$.
+The elements of this set are the sections $\{\tau^{j}:U_i \rightarrow \dfiber_{U_{i}}\}_{j\in I}$ of the fiber bundle \dtotal\ over open set $\U_{i}$.
+
+  The sections $\Gamma(\dbase, \dtotal)$ are called the global sections of $\dtotal$. \cite{SheafMathematics2021,spanier1989algebraic}.
+\end{definition}
 Each section is how we represent a unique dataset, where the set of sections $\Gamma$ is the set of all datasets that have the same data fields in the \dfiber\ and continuity as encoded in\dbase\.
@@
 
-\begin{definition} The fiber space category $\mathcal{\dfiber}$ is a category of sets \textbf{Set}.
-This means its objects are sets, and its morphisms are total functions\cite{barrCategoryTheoryComputing}
-\end{definition}
-
-\begin{definition} The monad $T$ is [a,b,c]
-
-\end{definition}
-
 For example, these functions could be the binary operations, group actions, and measurement scales discussed in \autoref{sec:related-work:equivariance}.
 These functions could also be monoid actions, which are a way of applying partial order relations to data\cite{fongInvitationAppliedCategory2019}, such as to build multi-ranked indicators\cite{bruggemannRankingPrioritizationMultiindicator2011}
-}
 
@@
 bundle map induces on level of sections induces on sheafs -> part of the definition of the presheaf
+    \item \textit{morphisms} E->E such that restricted to fiber it's the allowed maps, and restricted to base it's the inclusion map, sheaf to itself so is an endofunctor (category to itself)
+  \end{itemize}
+\end{definition}
 
-  \end{itemize}
-\end{definition
-The functions on the data
+\begin{definition}We define the endofunctor \monoid on E as the structure preserving map which is the monoid. ...
+\end{definition}
 
@@
 \begin{definition} category of bundle maps, which include category of fiber maps, what are allowable of fiber to itself, ex translation and scales only ..., objects are a fiber, maps are fiber to fiber defined as their own category, E to V fibers changes
-
-zooming is inclusion maps on both sides 
 \end{definition}
@@
 As mentioned in \autoref{sec:related-work:equivariance:category}
-
 that the artist $\vartist$ transforms functions that act on data $f:\dtotal\rightarrow \dtotal$ to functions that act on graphics $\g: \gtotal\rightarrow \dtotal$ in a way in which the following diagram commutes
+\begin{equation}
   \begin{tikzcd}
@@
   \end{tikzcd}
+\end{equation}
+
 meaning that, as described in \autoref{sec:related-work:equivariance}, the artist \vartist is constructed such that $\vartist(f(\dtotal)) = g(\vartist(\dtotal))$, meaning a transformation of the data has a consistent equivalent transformation of the graphic.
@@
 We then use this formulation to guide the development of visualization library components in \autoref{sec:case-study}.
+
+\begin{definition}[Category $\mathcal{\gtotal}$]
+
+\end{definition}
+
+% For this sake we don't do polymorphism, but we have a notion of equivalance classes of artists s.t... every parameter you want to fix is a fiber in E 
+% given a nu, 
+% maximal V has all the visual variables 
+% nu_red()->red
+
+\subsection{Union of Artists}
+\label{sec:artist:union}
+\begin{figure}[!h]
+\centering
+\subfloat[Union of H]{\includegraphics[width=.4\columnwidth]{exploding_artist.png}%
+\label{fig_first_case}}
+%\hfil
+\subfloat[Artist with shared/harmonized \vsection]{\includegraphics[width=.4\columnwidth]{combined_artist.png}%
+\label{fig_second_case}}
+\caption{Coherance? }
+\label{fig_sim}
+\end{figure}
+\note{do we define coherant visualizations as one with shared inclusions?}
+
+\section{Constraints: Construction and formal properties of Artists}
+\label{sec:artist:construction}
+\begin{figure*}[h!]
+  \includegraphics[width=\linewidth]{q.png}
+  \caption{}
+  \label{fig:constraints:q-overall}
+\end{figure*}
+
+\begin{definition} Following from \label{def:artist:} the artist \vartist is a special case of  natural transformations
+  \begin{equation}
+    \vartist: \Gamma(\dtotal, \dbase)\nrightarrow \Gamma(\gtotal, \gbase)
+  \end{equation}
+  between the profunctor objects $\Gamma(E),\Gamma(H)$ that preserves continuity and equivariance.
+\end{definition}
+  Natural transformations are functions that map functions between categories (functors) to other functors in a structure preserving way \cite{riehlCategoryTheoryContext, spanier1989algebraic, fongInvitationAppliedCategory2019}.
+
+\begin{definition} We define the natural transformation \textcolor{artist}{Artist}  \vartist\ as the tuple $(\vindex, \vchannel, \vmark, \mathcal{\dtotal},\mathcal{\vtotal}, \mathcal{\gtotal})$ where
+  \begin{enumerate}
+    \item \vindex\ is a continuity preserving functor
+    \item \vchannel\ and \vmark\ are equivariance preserving functors
+    \item $\mathcal{\dtotal}$, $\mathcal{\vtotal}$, and $\mathcal{\gtotal}$ are profunctor categories of sheafs
+    \item 
+  \end{enumerate}
+\end{definition}
+
+\begin{equation}
+  \label{eq:math:artist:diagram}
+  \begin{tikzcd}
+      \dtotal \arrow[r, "\vchannel"] \arrow[rd, "\pi"'] & \vtotal \arrow[d, "\pi"] & \vindex^*\vtotal \arrow[r, "\vmark"] \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"'] & \gtotal \arrow[ld, "\pi"] \\
+                                            & \dbase                  & \gbase \arrow[l, "\vindex"']                                              &                    
+      \end{tikzcd}
+\end{equation}
 
@@
 
-\begin{definition}[Category $\mathcal{\gtotal}$]
-
-\end{definition}
-
-\subsection{Union of Artists}
-\label{sec:artist:union}
-\begin{figure}[!h]
-\centering
-\subfloat[Union of H]{\includegraphics[width=.4\columnwidth]{exploding_artist.png}%
-\label{fig_first_case}}
-%\hfil
-\subfloat[Artist with shared/harmonized \vsection]{\includegraphics[width=.4\columnwidth]{combined_artist.png}%
-\label{fig_second_case}}
-\caption{Coherance? }
-\label{fig_sim}
-\end{figure}
-\note{do we define coherant visualizations as one with shared inclusions?}
-
-\section{Constraints: Construction and formal properties of Artists}
-\label{sec:artist:construction}
-\begin{figure*}[h!]
-  \includegraphics[width=\linewidth]{q.png}
-  \caption{}
-  \label{fig:constraints:q-overall}
-\end{figure*}
-
-\begin{definition} Following from \label{def:artist:} the artist \vartist is a special case of  natural transformations
-  \begin{equation}
-    \vartist: \Gamma(\dtotal, \dbase)\nrightarrow \Gamma(\gtotal, \gbase)
-  \end{equation}
-  between the profunctor objects $\Gamma(E),\Gamma(H)$ that preserves continuity and equivariance.
-\end{definition}
-  Natural transformations are functions that map functions between categories (functors) to other functors in a structure preserving way \cite{riehlCategoryTheoryContext, spanier1989algebraic, fongInvitationAppliedCategory2019}.
-
-\begin{definition} We define the natural transformation \textcolor{artist}{Artist}  \vartist\ as the tuple $(\vindex, \vchannel, \vmark, \mathcal{\dtotal},\mathcal{\vtotal}, \mathcal{\gtotal})$ where
-  \begin{enumerate}
-    \item \vindex\ is a continuity preserving functor
-    \item \vchannel\ and \vmark\ are equivariance preserving functors
-    \item $\mathcal{\dtotal}$, $\mathcal{\vtotal}$, and $\mathcal{\gtotal}$ are profunctor categories of sheafs
-    \item 
-  \end{enumerate}
-\end{definition}
-
-\begin{equation}
-  \label{eq:math:artist:diagram}
-  \begin{tikzcd}
-      \dtotal \arrow[r, "\vchannel"] \arrow[rd, "\pi"'] & \vtotal \arrow[d, "\pi"] & \vindex^*\vtotal \arrow[r, "\vmark"] \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"'] & \gtotal \arrow[ld, "\pi"] \\
-                                            & \dbase                  & \gbase \arrow[l, "\vindex"']                                              &                    
-      \end{tikzcd}
-\end{equation}
-
-\subsection{Graphic to Data: \vindex}
+\subsection{Visualization Library Components}
+\subsubsection{Graphic to Data: \vindex}
 \begin{equation}
@@
 
-\subsection{Visual Bundle \vtotal}
+\subsubsection{Visual Bundle \vtotal}
 \begin{equation}
@@
 
-\subsection{Data to Visual Encodings: \vchannel} %look at what K&S call these stages
+\subsubsection{Data to Visual Encodings: \vchannel} %look at what K&S call these stages
 Visual encoding functions \vindex\ are functors because they are expected to preserve structure on the data and visual side, as discussed in \autoref{sec:related-work:equivariance}.

### 2021-10-06 paper/main.tex
fiber

@@
 
-\note{something about how structure of data field serves as basis of equivariance in viz literature}
-Besides expressing continuities, fiber bundles also provide us a way to richly express the structure of the fields in the data.
-Specifically, we do this by adopting Spivak's formulation of the fiber as a (column name, data domain) simple schema \cite{spivakSIMPLICIALDATABASES,spivakDatabasesAreCategories2010}.
-Spivak formally maps column names and field types to the set of values associated with the field type, for example $\mathbb{R}$ for a \texttt{float} column named \textit{temperature}.
-In the paper, the $\dfiber$ is the cartesian cross product of the fibers for each column.
-
 \begin{figure}[!h]
@@
 \section{Artist} 
-The \textcolor{artist}{Artist} $\mathcal{\vartist}$\footnote{We call this transformation the artist because the Artist object in Matplotlib\cite{hunterMatplotlib2DGraphics2007} is responsible for converting data to a prerender graphical object} is a functor from data $\mathcal{\dtotal}$ to graphic $\mathcal{\gtotal}$.}
+In this section, we use category theory to formally express the implicit assumptions that visualization library components make in transforming data into graphical representations.
+We propose that the visualization transformation can be modeled as a functor, which we call the \textcolor{artist}{Artist} $\mathcal{\vartist}$\footnote{We call this transformation the artist because the \mintinline{python}{Artist} object in Matplotlib\cite{hunterMatplotlib2DGraphics2007}}.
+The artist \vartist converts data, which we model as the catgory $\mathcal{\dtotal}$ to graphics, which we model as the category $\mathcal{\gtotal}$.}
 
@@
 
-We formally define data and graphic as the categories $\mathcal{\dtotal}$ and $\mathcal{\gtotal}$.
-We develop these categories as analogues to fiber bundle $(\dtotal, \dbase, \pi, \dfiber)$, introduced in \autoref{sec:related-work:fiber-bundles}, and fiber bundle $(\gtotal, \gbase, \pi, \gfiber)$.
-Throughout this section, we assume that $\mathcal{E}$ is defined in terms of a fixed continuity and fiber space, for example a class of images or tables with the same schema.
-Similarly we assume that $\mathcal{H}$ is a stand in for a specific type of graphic, such as a scatter or line plot.
+We define the categories $\mathcal{\dtotal}, \mathcal{\gtotal}$ as sheaves of sections\cite{SheafMathematics2021,spanier1989algebraic}, of fiberbundles, as introduced in \autoref{sec:related-work:fiber-bundles}.
+We assume that $\mathcal{E}$ is defined in terms of a fixed base and fiber space, for example a class of images or tables with the same schema.
+The graphic category $\mathcal{\gtotal}$ consists of the set of graphics that an artist \vartist is allowed to generate from a given $\mathcal{E}$, which is discussed further in \autoref{sec:artist:construction}.
 
 As introduced in \autoref{sec:related-work:fiber-bundles}, sets of open sets form the basis of topological spaces.
-We propose that we can express how the elements in our dataset are glued together by modeling that continuity as a topological space \dbase, and we express the constraints on \dbase by defining the analogues category $\mathcal{\dbase}$.
+We propose that we can express how the elements in our data are connected to each other by modeling the data continuity as a topological space \dbase.
+We express subsets of the continuity, which can be mapped into subsets of the data, as open sets $U_{i}$ in \dbase.
+We introduce the category $\mathcal{\dbase}$ to encapsulate the subsets of the continuity and the inclusion maps that glue them together.
 
@@
 
-The objects $U_{i}$ are a way to manage subsets of the data.
-For example, let there be a data value for each point along the interval $\dbase=\right[0,1\left]$.
-Since this function is continuous, there is a subset of the data defined along the open set $U_{a} = \right(0.2, 0.4\left)$.
-The open set $U_{a}$ provides a unique way of identifying that subset.
+Besides expressing continuities, fiber bundles also provide us a way to richly express the structure of the fields in the data.
+Specifically, we do this by adopting Spivak's formulation of the fiber as a (column name, data domain) simple schema \cite{spivakSIMPLICIALDATABASES,spivakDatabasesAreCategories2010}.
+Spivak formally maps column names and field types to the set of values associated with the field type, for example $\mathbb{R}$ for a \texttt{float} column named \textit{temperature}.
+In the paper, the $\dfiber$ is the cartesian cross product of the fibers for each column.
 
-\note{maybe a small image here or when we introduce open sets showing what we mean?}
+The category $\mathcal{\dfiber}$ is a subcategory of an arbitrary category $\mathcal{C}$.
+$\mathcal{\dfiber}$ is, meaning it has one object \texttt{\dfiber} and a functor\footnote{The functor $\otimes$ is a bifunctor because it takes as input the product of two categories \cite{nlab:bifunctor}} $\otimes:\mathcal{\dfiber}\times{\mathcal{\dfiber}}\rightarrow \mathcal{\dfiber}$\cite{barrCategoryTheoryComputing}.
 
@@
   \label{def:category:F} 
-  The category $\mathcal{\dfiber}$ is a subcategory of an arbitrary category $\mathcal{C}$.
-$\mathcal{\dfiber}$ is a monoidal category, meaning it has one object \texttt{\dfiber} and a functor\footnote{The functor $\otimes$ is a bifunctor because it takes as input the product of two categories \cite{nlab:bifunctor}} $\otimes:\mathcal{\dfiber}\times{\mathcal{\dfiber}}\rightarrow \mathcal{\dfiber}$\cite{barrCategoryTheoryComputing}.
+  The category $\mathcal{\dfiber}$ consists of one object $\dfiber$ with maps back to itself $\monoid: \dfiber \rightarrow \dfiber$. 
 \end{definition}
 
-The object $\texttt{\dfiber}$ can be an object of any arbitrary category $\mathcal{C}$, which allows $\texttt{\dfiber}$ to encode most data fields.
+%F is the one object category with maps back to itself M: F\rightarrow F, fiber and all its allowed maps.
+morphsims in categroy C that go back to same object fixed F, we note that this is also a category w/ one object, a one object categry is a monoid
+
+The object $\texttt{\dfiber}$ is an object of some category $\mathcal{C}$, which allows $\texttt{\dfiber}$ to encode most data fields.
 For example, a lists of strings is an instance of an object in \textbf{Set}, networks are an instance of \textbf{Graph}, and images are vector spaces which are a specific type of topological space \textbf{Top}.
+We note that a one object category is a monoid \cite{nlab:mo}
 
@@
 
+%insteaf of trying to define gamma, write this up as a paragaraph...
+we use the convention \Gamma(U,E) to denote, the map from U_i to gamma U_e...defines a presheaf which is also a sheaf....
+continue w/ global sections
 \begin{definition} For each open set object $U_{i}$ of category $\mathcal{\dbase}$, there is a set $\Gamma(U_i, \dtotal)$.
 The elements of this set are the sections $\{\tau^{j}:U_i \rightarrow \dfiber_{U_{i}}\}_{j\in I}$ of the fiber bundle \dtotal\ over open set $\U_{i}$.
+The sections $\Gamma(\dbase, \dtotal)$ are called the global sections of $\dtotal$.
+\cite{SheafMathematics2021,spanier1989algebraic}.
+\end{definition}
 
-  The sections $\Gamma(\dbase, \dtotal)$ are called the global sections of $\dtotal$. \cite{SheafMathematics2021,spanier1989algebraic}.
-\end{definition}
 Each section is how we represent a unique dataset, where the set of sections $\Gamma$ is the set of all datasets that have the same data fields in the \dfiber\ and continuity as encoded in\dbase\.
@@
     \item\textit{object} the space of continuous\footnote{or piecewise continuous}sections $\sheaf: U_{i} \rightarrow \Gamma(U_{i}, \dtotal) $ which is a sheaf\cite{spanier1989algebraic,SheafMathematics2021}.
+    \item\textit{morphisms} allowed maps from sheaf to sheaf, when restricted to a fiber, that should be a map from a fiber to fiber.  
     \item\textit{morphisms} bundle maps E, E, fiber to itself and is the fiber map allowable, functions F to itself forms a monoid (one object category), F and it's allowable maps form a category, monoid could be monotonic functions, group actions, extending that to bundle E, interested in E to itself, which when restricted to fiber are category maps of fiber to itself, maps from sheaf to itself, that when restricted to fiber gives you fiber to itself, given bundle map E->E, then you can show that you have a sheaf map...
@@
 \end{definition}
-
 % For this sake we don't do polymorphism, but we have a notion of equivalance classes of artists s.t... every parameter you want to fix is a fiber in E 

### 2021-10-06 paper/main.tex
artist

@@
 
+The inclusion maps $iota$ can be used to piece together the subsets $U_{i}$, particularly individual points $\dbasepoint \in \dbase$.
+This is how we provide knowledge of continuity to the artist \vartist\ in a way where the artist does not need all the data at once.
+This is what allows the artist to generate graphics where the subset of data on view is dynamically updated, such as pan and zoom navigation\cite{NekrasovskiEvaluationPanZoom2006} and sliding windows on streaming data\cite{crouchDynamicGraphsSlidingwindow2013,chuTimeSeriesSegmentation1995}.
+
 Besides expressing continuities, fiber bundles also provide us a way to richly express the structure of the fields in the data.
@@
 Spivak formally maps column names and field types to the set of values associated with the field type, for example $\mathbb{R}$ for a \texttt{float} column named \textit{temperature}.
-In the paper, the $\dfiber$ is the cartesian cross product of the fibers for each column.
-
-The category $\mathcal{\dfiber}$ is a subcategory of an arbitrary category $\mathcal{C}$.
-$\mathcal{\dfiber}$ is, meaning it has one object \texttt{\dfiber} and a functor\footnote{The functor $\otimes$ is a bifunctor because it takes as input the product of two categories \cite{nlab:bifunctor}} $\otimes:\mathcal{\dfiber}\times{\mathcal{\dfiber}}\rightarrow \mathcal{\dfiber}$\cite{barrCategoryTheoryComputing}.
+We define $\dfiber$ as  the cartesian cross product of the fibers $\dfiber_c$ for each field, where $c$ is the field name.
 
@@
   \label{def:category:F} 
-  The category $\mathcal{\dfiber}$ consists of one object $\dfiber$ with maps back to itself $\monoid: \dfiber \rightarrow \dfiber$. 
+  The category $\mathcal{\dfiber}$ consists of a single object $\dfiber$ and morphisms between the object and itself $\monoid: \dfiber \rightarrow \dfiber$. 
 \end{definition}
 
-%F is the one object category with maps back to itself M: F\rightarrow F, fiber and all its allowed maps.
-morphsims in categroy C that go back to same object fixed F, we note that this is also a category w/ one object, a one object categry is a monoid
+The object $\texttt{\dfiber}$ can be an object of any category $\mathcal{C}$, allowing $\texttt{\dfiber}$ to encode most data fields.
+For example, a lists of strings is an instance of an object in \textbf{Set}, networks are an instance of \textbf{Graph}, and images are vector spaces which are a specific type of topological space \textbf{Top}.
+We note that a one object category is a monoid \cite{nlab:monoid,barrCategoryTheoryComputing}.
+\note{&?
+why is this important} The morphisms $\monoid$ are functions from the data to itself, for example the binary operations, group actions, and measurement scales discussed in \autoref{sec:related-work:equivariance}.
+These functions could also be monoid actions\cite{barrCategoryTheoryComputing}, which provide a way of applying partial order relations to data\cite{fongInvitationAppliedCategory2019}, such as to build multi-ranked indicators\cite{bruggemannRankingPrioritizationMultiindicator2011}
 
-The object $\texttt{\dfiber}$ is an object of some category $\mathcal{C}$, which allows $\texttt{\dfiber}$ to encode most data fields.
-For example, a lists of strings is an instance of an object in \textbf{Set}, networks are an instance of \textbf{Graph}, and images are vector spaces which are a specific type of topological space \textbf{Top}.
-We note that a one object category is a monoid \cite{nlab:mo}
-
-%Monoid b/c arrows form a Monoid, F + Monoid that acts on F
-
-%% F, M action on F, 
-%%% F \->E->K
-
-%insteaf of trying to define gamma, write this up as a paragaraph...
-we use the convention \Gamma(U,E) to denote, the map from U_i to gamma U_e...defines a presheaf which is also a sheaf....
-continue w/ global sections
-\begin{definition} For each open set object $U_{i}$ of category $\mathcal{\dbase}$, there is a set $\Gamma(U_i, \dtotal)$.
-The elements of this set are the sections $\{\tau^{j}:U_i \rightarrow \dfiber_{U_{i}}\}_{j\in I}$ of the fiber bundle \dtotal\ over open set $\U_{i}$.
-The sections $\Gamma(\dbase, \dtotal)$ are called the global sections of $\dtotal$.
-\cite{SheafMathematics2021,spanier1989algebraic}.
-\end{definition}
-
+We encode individual datasets in the fiberbundle as sections, and we follow the convention that $\Gamma(U_i, \dtotal)$ denotes the set of sections $\{\dsection^{j}:U_i \rightarrow \dfiber_{U_{i}}\}_{j\in I}$ of the fiber bundle \dtotal\ over an open set $\U_{i}$.
+The sections $\Gamma(\dbase, \dtotal)$ are called the global sections of $\dtotal$ \cite{SheafMathematics2021,spanier1989algebraic}.
 Each section is how we represent a unique dataset, where the set of sections $\Gamma$ is the set of all datasets that have the same data fields in the \dfiber\ and continuity as encoded in\dbase\.
-For example, database tables that have the same schema and discrete index, as illustrated in Spivak \cite{spivakSIMPLICIALDATABASES}.
-
-For example, these functions could be the binary operations, group actions, and measurement scales discussed in \autoref{sec:related-work:equivariance}.
-These functions could also be monoid actions, which are a way of applying partial order relations to data\cite{fongInvitationAppliedCategory2019}, such as to build multi-ranked indicators\cite{bruggemannRankingPrioritizationMultiindicator2011}
+For example, Spivak describes tables that have the same schema and discrete index as sections of the same fiber bundle\cite{spivakSIMPLICIALDATABASES}.
 
@@
   \begin{itemize}
-    \item\textit{object} the space of continuous\footnote{or piecewise continuous}sections $\sheaf: U_{i} \rightarrow \Gamma(U_{i}, \dtotal) $ which is a sheaf\cite{spanier1989algebraic,SheafMathematics2021}.
-    \item\textit{morphisms} allowed maps from sheaf to sheaf, when restricted to a fiber, that should be a map from a fiber to fiber.  
-    \item\textit{morphisms} bundle maps E, E, fiber to itself and is the fiber map allowable, functions F to itself forms a monoid (one object category), F and it's allowable maps form a category, monoid could be monotonic functions, group actions, extending that to bundle E, interested in E to itself, which when restricted to fiber are category maps of fiber to itself, maps from sheaf to itself, that when restricted to fiber gives you fiber to itself, given bundle map E->E, then you can show that you have a sheaf map...
-Given E->E, and F->F, then there's a map from one sheaf to another.
-bundle map induces on level of sections induces on sheafs -> part of the definition of the presheaf
-    \item \textit{morphisms} E->E such that restricted to fiber it's the allowed maps, and restricted to base it's the inclusion map, sheaf to itself so is an endofunctor (category to itself)
+    \item\textit{object} the functor $\sheaf: \mathcal{K}^{op} \rightarrow \Gamma(\mathcal{K}, \dtotal)$, which is a sheaf\cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf} \note{this might need an appendix/footnote: it is a presheaf b/c $\iota:\Gamma(U_2)\rightarrow \Gamma(U_1)$ and $\iota: \dbase_1\rightarrow \dbase_2$ (smaller subspace in bigger subspace, larger section includes smaller), which means $\mathcal{K}$ is contravariant, and $\Gamma(U_{i}, \dtotal)$ is \textbf{set}}
+    \item \textit{morphisms} the endomorphic\cite{fongInvitationAppliedCategory2019} maps $\cdot: \dtotal\rightarrow \dtotal$, which are the fiber maps $\monoid:\mathcal{\dfiber}\rightarrow \mathcal{\dfiber}$ when the morphisms are restricted to a single point in the base space $\dbasepoint\in \dbase$
   \end{itemize}
@@
 
-\begin{definition}We define the endofunctor \monoid on E as the structure preserving map which is the monoid. ...
-\end{definition}
+The graphic category $\mathcal{\gtotal}$ consists of the functor $\sheaf: \mathcal{S}^{op} \rightarrow \Gamma(\mathcal{S}, \gtotal)$.
+The category $\mathcal{S}$ is the category of open sets over the graphic base space \gbase\, and $\Gamma(\mathcal{S}, \gtotal)$ is the set of sections $\gsection$ over $\mathcal{\gbase}$.
+The artist is a morphism of sheafs $\vartist: \mathcal{\dtotal} \rightarrow \mathcal{\gtotal}$, which means it is by definition a natural transform\cite{SheafMathematics2021}.
+This means that the following diagram must commute
 
-\note{equivariance is a special case of functions}
-\note{actions are a special case of actions}
-\begin{definition} category of bundle maps, which include category of fiber maps, what are allowable of fiber to itself, ex translation and scales only ..., objects are a fiber, maps are fiber to fiber defined as their own category, E to V fibers changes
-\end{definition}
-
-The objects of the data category $\mathcal{\dtotal}$ and graphic category $\mathcal{\gtotal}$ are the sheaves $\Gamma(\dbase, \dtotal)$ and $\Gamma(\gbase, \gtotal)$.
-An alternate notation is $\sheaf(\dtotal)$ and $\sheaf(\gtotal)$.
-The artist is a morphism of sheafs $\sheaf(\dtotal)$ to $\sheaf(\gtotal)$, which means it is by definition a natural transform\cite{SheafMathematics2021}.
-As mentioned in \autoref{sec:related-work:equivariance:category}
-that the artist $\vartist$ transforms functions that act on data $f:\dtotal\rightarrow \dtotal$ to functions that act on graphics $\g: \gtotal\rightarrow \dtotal$ in a way in which the following diagram commutes
 \begin{equation}
   \begin{tikzcd}
-    \mathcal{\dtotal} \arrow[r, "\vartist"] \arrow[d, "f"] & \mathcal{\gtotal} \arrow[d, "g"] \arrow[d, "\ca"] \\
+    \mathcal{\dtotal} \arrow[r, "\vartist"] \arrow[d, "M"] & \mathcal{\gtotal} \arrow[d, "A(M)"] \arrow[d, "\ca"] \\
     \mathcal{\dtotal} \arrow[r, "\vartist"]                & \mathcal{\gtotal}                                
@@
 
-meaning that, as described in \autoref{sec:related-work:equivariance}, the artist \vartist is constructed such that $\vartist(f(\dtotal)) = g(\vartist(\dtotal))$, meaning a transformation of the data has a consistent equivalent transformation of the graphic.
-In \autoref{sec:artist:construction}, we show a method of constructing the artist such that continuity and equivariance are preserved.
+such that the morphisms on the graphic category $\mathcal{H}$ are homomorphic to the morphisms $\monoid$ on the fibers $\mathcal{F}$ \cite{spanier1989algebraic,fongInvitationAppliedCategory2019}.
+In \autoref{sec:artist:construction}, we show a method of constructing the artist such that this constraint can be satisfied, thereby ensuring equivariance.
 We then use this formulation to guide the development of visualization library components in \autoref{sec:case-study}.
-
-\begin{definition}[Category $\mathcal{\gtotal}$]
-
-\end{definition}
-% For this sake we don't do polymorphism, but we have a notion of equivalance classes of artists s.t... every parameter you want to fix is a fiber in E 
-% given a nu, 
-% maximal V has all the visual variables 
-% nu_red()->red
 

### 2021-10-07 paper/main.tex
sheafs

@@
 We encode individual datasets in the fiberbundle as sections, and we follow the convention that $\Gamma(U_i, \dtotal)$ denotes the set of sections $\{\dsection^{j}:U_i \rightarrow \dfiber_{U_{i}}\}_{j\in I}$ of the fiber bundle \dtotal\ over an open set $\U_{i}$.
+As mentioned in \autoref{sec:related-work:fiber-bundles}, the fiber space is embedded in the total space $\dfiber \hookrightarrow \dtotal$.
 The sections $\Gamma(\dbase, \dtotal)$ are called the global sections of $\dtotal$ \cite{SheafMathematics2021,spanier1989algebraic}.
@@
   \begin{itemize}
-    \item\textit{object} the functor $\sheaf: \mathcal{K}^{op} \rightarrow \Gamma(\mathcal{K}, \dtotal)$, which is a sheaf\cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf} \note{this might need an appendix/footnote: it is a presheaf b/c $\iota:\Gamma(U_2)\rightarrow \Gamma(U_1)$ and $\iota: \dbase_1\rightarrow \dbase_2$ (smaller subspace in bigger subspace, larger section includes smaller), which means $\mathcal{K}$ is contravariant, and $\Gamma(U_{i}, \dtotal)$ is \textbf{set}}
+    \item\textit{object} the contravariant functor $\sheaf: \mathcal{K}^{op} \rightarrow \Gamma(\mathcal{K}, \dtotal)$
     \item \textit{morphisms} the endomorphic\cite{fongInvitationAppliedCategory2019} maps $\cdot: \dtotal\rightarrow \dtotal$, which are the fiber maps $\monoid:\mathcal{\dfiber}\rightarrow \mathcal{\dfiber}$ when the morphisms are restricted to a single point in the base space $\dbasepoint\in \dbase$
@@
 \end{definition}
+
+The functor object of $\mathcal{E}$ is a presheaf because the definition of a presheaf is that it is a contravariant functor from a category $\mathcal{K}$ to a set \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}.
+As illustrated in the diagram,
+\begin{equation}
+  \label{eq:artist:[presheaf}
+  \begin{tikzcd}[row sep=small]
+    U_1 \arrow[dd, "\iota"'] & {\Gamma(U_1, \dtotal)}                              \\
+    {} \arrow[r, "\sheaf"]   & {}                                                  \\
+    U_2                      & {\Gamma(U_2, \dtotal)} \arrow[uu, "\iota^*"']
+    \end{tikzcd}
+\end{equation}
+the functor \sheaf\ is contravariant because the inclusion map $\iota$ between objects in $\mathcal{K}$ goes the in the opposite direction from the map $\iota$ between open set objects in $\mathcal{K}$ \cite{riehlCategoryTheoryContext}.
+The open set $U_1$ is a subset of $U_2$, therefore $\iota: U_1 \rightarrow U_2$.
+On the other hand, a set of functions defined on a space $U_2$ must be defined on the subspace $\iota: \Gamma(U_2, \dbase) \rightarrow \Gamma(U_1, \dbase)$ but the reverse does not need to be true.
+For example, a continuous function defined over the interval $\left[.25,.5\right]$ need not be defined over $\left[0, 1\right]$ but must exist over  $\left[.3, .45\right]$.
+We use the notation $\mathcal{K}^{op}$ to signify that the functor $\sheaf$ is contravariant and $\sheaf$ to denote that it is a sheaf.
+A sheaf is a presheaf where sections are defined over unions of open sets \cite{nlab:sheaf,ghristElementaryAppliedTopology2014}
 
@@
 \begin{equation}
+  \label{eq:artist:natural-transform}
   \begin{tikzcd}
-    \mathcal{\dtotal} \arrow[r, "\vartist"] \arrow[d, "M"] & \mathcal{\gtotal} \arrow[d, "A(M)"] \arrow[d, "\ca"] \\
+  \mathcal{\dtotal} \arrow[r, "\vartist"] \arrow[d, "\monoid"'] & \mathcal{\gtotal} \arrow[d, "\vartist(\monoid)"] \\
   \mathcal{\dtotal} \arrow[r, "\vartist"]                       & \mathcal{\gtotal}                               

### 2021-10-07 paper/main.tex
union

@@
 \centering
-\subfloat[Union of H]{\includegraphics[width=.4\columnwidth]{exploding_artist.png}%
-\label{fig_first_case}}
+\subfloat[Addition]{\includegraphics[width=.4\columnwidth]{exploding_artist.png}%
+\label{fig:artist:union:glue}}
 %\hfil
-\subfloat[Artist with shared/harmonized \vsection]{\includegraphics[width=.4\columnwidth]{combined_artist.png}%
-\label{fig_second_case}}
-\caption{Coherance? }
-\label{fig_sim}
+\subfloat[Inclusion]{\includegraphics[width=.4\columnwidth]{combined_artist.png}%
+\label{fig:artist:union:continuity}}
+\caption{In \autoref{fig:artist:union:glue}, the outputs of various artist are independently rendered to the same graphic space \gtotal.
+The inputs to the artists in \autoref{fig:artist:union:continuity} are sections over the same fiber bundle base space \dbase; therefore, they should share transformation functions to produce a consistent image.}
 \end{figure}
-\note{do we define coherant visualizations as one with shared inclusions?}
+
+Visualizations generally consist of more than one graphic, for example two line plots, a legend, axis ticks, spines, and labels.
+One form of composition is to render the family of artists $(\dtotal_i: i\in I)$  targeting the same graphical space $\gtotal$ as a set of independent artists, as illustrated by \autoref{fig:artist:union:glue}.
+We define this composition as the addition of artists
+\begin{equation}
+  \label{eq:artist:union:glue}
++ \coloneqq \underset{i\in I}{\sqcup} \dtotal_{i}
+\end{equation}
+
+Many times, the graphics are generated from the same dataset, or from datasets that share continuity and fibers.
+This is implicit in figures such as \autoref{fig:artist:union:continuity} where the line plots are two timeseries for the same period, the x and y axis are shared with the timeseries, and the legend reflects the timeseries.
+In these cases, we define a composition that incorporates the inclusion $\iota: \dtotal_i \rightarrow \dtotal_j$ between the data input to the artists.
+Doing so provides a way to keep track of which fiber spaces are the same, and therefore which fibers share transformation functions to visual elements.
+\cite{this is really about shared $\vchannel$ so not sure it should go here} This type of composition is important for visualizations where elements update together in a consistent way, such as multiple views \cite{alboRadarComparativeEvaluation2016a, hullmanKeeping2018} and brush-linked views \cite{beckerBrushingScatterplots1987,bujaInteractiveData1991}.
 

### 2021-10-07 paper/main.tex
component bundle union

@@
     \item\textit{object} the contravariant functor $\sheaf: \mathcal{K}^{op} \rightarrow \Gamma(\mathcal{K}, \dtotal)$
-    \item \textit{morphisms} the endomorphic\cite{fongInvitationAppliedCategory2019} maps $\cdot: \dtotal\rightarrow \dtotal$, which are the fiber maps $\monoid:\mathcal{\dfiber}\rightarrow \mathcal{\dfiber}$ when the morphisms are restricted to a single point in the base space $\dbasepoint\in \dbase$
+    \item \textit{morphisms} the endomorphic\cite{fongInvitationAppliedCategory2019} bundle maps $\bullet: \dtotal\rightarrow \dtotal$, which are the fiber maps $\monoid:\mathcal{\dfiber}\rightarrow \mathcal{\dfiber}$ when the sheaf is over a single point in the base space $\dbasepoint\in \dbase$
   \end{itemize}
@@
 %\hfil
-\subfloat[Inclusion]{\includegraphics[width=.4\columnwidth]{combined_artist.png}%
+\subfloat[Component Bundles]{\includegraphics[width=.4\columnwidth]{combined_artist.png}%
 \label{fig:artist:union:continuity}}
@@
 Visualizations generally consist of more than one graphic, for example two line plots, a legend, axis ticks, spines, and labels.
-One form of composition is to render the family of artists $(\dtotal_i: i\in I)$  targeting the same graphical space $\gtotal$ as a set of independent artists, as illustrated by \autoref{fig:artist:union:glue}.
+One form of composition is to render the family of artists $(\dtotal_i: i\in I)$  targeting the same graphical space $\gtotal$ as a set of independent artists, one example being the disjointed image in \autoref{fig:artist:union:glue}.
 We define this composition as the addition of artists
@@
 
-Many times, the graphics are generated from the same dataset, or from datasets that share continuity and fibers.
+\note{this isn't really composition, cause the actual composition is still the + above, this is about consistency/coherancy/E}
+Many times though, graphics are generated from subsets of a larger data set.
 This is implicit in figures such as \autoref{fig:artist:union:continuity} where the line plots are two timeseries for the same period, the x and y axis are shared with the timeseries, and the legend reflects the timeseries.
-In these cases, we define a composition that incorporates the inclusion $\iota: \dtotal_i \rightarrow \dtotal_j$ between the data input to the artists.
-Doing so provides a way to keep track of which fiber spaces are the same, and therefore which fibers share transformation functions to visual elements.
-\cite{this is really about shared $\vchannel$ so not sure it should go here} This type of composition is important for visualizations where elements update together in a consistent way, such as multiple views \cite{alboRadarComparativeEvaluation2016a, hullmanKeeping2018} and brush-linked views \cite{beckerBrushingScatterplots1987,bujaInteractiveData1991}.
+In these situations, we consider each of the data inputs to the artist $\mathcal{\dtotal}$ as component bundles of a larger bundle \dtotal.
+
+\begin{equation}
+  \label{eq:math:data:base:decompose}
+  \pi:\dtotal_1\oplus\ldots\oplus \dtotal_i \oplus\ldots \oplus \dtotal_n \rightarrow \dbase
+\end{equation}
+
+These component bundles all share the same continuity, which means that sections $\dsection$ on each component bundle over a point $\dbasepoint$ are all sets of data values associated with that point.
+We use this property to establish which of these points are the same, because they come from the same fiber,and therefore should share transformation functions to visual elements.
+This type of composition is important for visualizations where elements update together in a consistent way, such as multiple views \cite{alboRadarComparativeEvaluation2016a, hullmanKeeping2018} and brush-linked views \cite{beckerBrushingScatterplots1987,bujaInteractiveData1991}.
 

### 2021-10-07 paper/main.tex
wilkenson calls continuity topological properties

--- 2021-10-07 paper/main.tex
@@
 \begin{LaTeXdescription}
-  \item [continuity] how elements in a dataset are connect to each other, e.g. discrete rows in a table, networked nodes, pixels in an image, points on a line
+  \item [continuity] how elements in a dataset are organized, e.g. discrete rows in a table, networked nodes, pixels in an image, points on a line
   \item [equivariance] functions on data that have an equivalent effect on the graphical representation, e.g.
@@
 \end{figure}
-We propose that a robust model of continuity, which is how the elements in a dataset are connected to each other, provides a way to develop library components that \note{finish this sentence in a sensible way}
+Continuity, which Wilkinson terms topological properties \cite{wilkinsonGrammarGraphics2005}, describes how elements in a data set are organized.
+Wilkenson provides the examples of values that are isolated from each other, and therefore discrete, and values lying on a continuum or in a compact region.
+We propose that a robust model of continuity, which is how the elements in a dataset are connected to each other, provides a way to develop library components that can
 
@@
 
-The model described in this work provides a method for expressing different types of continuities and nested continuities using the same model.
+The model described in this work provides a modol for expressing data sets with different topological properties. 
 We obtain this generality by using the mathematical theory of fiber bundles as the basis of our abstraction, as proposed by Butler, Bryson, and Pendley\cite{butlerVisualizationModelBased1989,butlerVectorBundleClassesForm1992}.
@@
 
-Specifically, Butler, Bryson, and Pendley suggest that fiber bundles be the basis of an abstract data model. Fiber bundles are a collection of topological spaces. 
+Specifically,Butler, Bryson, and Pendley suggest that fiber bundles be the basis of an abstract data model. Fiber bundles are a collection of topological spaces. 
 \begin{definition}
@@
   \begin{itemize}
-    \item \textit{objects} all open sets $U_{i} \in \{U\}$ in the topological space\cite{bradleyTopologyCategoricalApproach2020, } $(\dbase, \mathscr{T})$), including the empty set $\varnothing$ and the maximum set \dbase.
+    \item \textit{objects} all open sets $U_{i} \in \{U\}$ in the topological space $(\dbase, \mathscr{T})$), including the empty set $\varnothing$ and the maximum set \dbase. 
     \item \textit{morphisms} inclusion maps from a subspace to a larger space $\iota: U_{i} \hookrightarrow U_{j}$, which is equivalent to $U_{i} \subseteq U_{j}$ and  
@@
 \centering
-\subfloat[Union of H]{\includegraphics[width=.4\columnwidth]{exploding_artist.png}%
-\label{fig_first_case}}
+\subfloat[+]{\includegraphics[width=.4\columnwidth]{exploding_artist.png}%
+\label{fig:artist:union:glue}}
 %\hfil
-\subfloat[Artist with shared/harmonized \vsection]{\includegraphics[width=.4\columnwidth]{combined_artist.png}%
-\label{fig_second_case}}
-\caption{Coherance? }
-\label{fig_sim}
+\subfloat[$\oplus$]{\includegraphics[width=.4\columnwidth]{combined_artist.png}%
+\label{fig:artist:union:continuity}}
+\caption{In \autoref{fig:artist:union:glue}, the outputs of various artist are independently rendered to the same graphic space \gtotal.
+The inputs to the artists in \autoref{fig:artist:union:continuity} are sections over the same fiber bundle base space \dbase; therefore, they should share transformation functions to produce a consistent image.
+\note{add a border/make the same size/etc so 1 doesn't run over into two}}
 \end{figure}
-\note{do we define coherant visualizations as one with shared inclusions?}
+
+Visualizations generally consist of more than one graphic, for example two line plots, a legend, axis ticks, spines, and labels.
+\note{formalize superposition, than break down what that means for composition}
+
+These graphics can be generated by disjoint data inputs  $(\dtotal_i: i\in I)$ into different artist functions $\vartist$ with the same output target space $\gtotal$.
+We define this composition via the disjoint union of the inputs
+\begin{equation}
+  \label{eq:artist:union:glue}
++ \coloneqq \underset{i\in I}{\sqcup} \dtotal_{i}
+\end{equation}
+This approach does not guarantee a coherent image, as illustrated in \autoref{fig:artist:union:glue}, because each data is explicitly considered as distinct from the others.
+
+Many times though, graphics are generated from subsets of a larger data set, such as when using multiple views \cite{alboRadarComparativeEvaluation2016a, hullmanKeeping2018} and brush-linked views \cite{beckerBrushingScatterplots1987,bujaInteractiveData1991}.
+In these situations, we consider each of the data inputs as component bundles over a shared continuity \dbase.
+
+\begin{equation}
+  \label{eq:math:data:base:decompose}
+  \pi:\dtotal_1\oplus\ldots\oplus \dtotal_i \oplus\ldots \oplus \dtotal_n \rightarrow \dbase
+\end{equation}
+
+The sections $\dsection$ on each component bundle $\dtotal_i$ over a point $\dbasepoint \in \dbase$ are the elements of the data associated with that point.
+We use this property to establish which of these elements are the same, because they come from the same fiber, and therefore should share transformation functions to visual elements.
+This bookkeeping yields graphics with consistent visual encoding of data elements, as shown in \autoref{fig:artist:union:continuity}.
 

### 2021-10-07 paper/main.tex
retweak continuity intro

@@
 \end{figure}
-Continuity, which Wilkinson terms topological properties \cite{wilkinsonGrammarGraphics2005}, describes how elements in a data set are organized.
-Wilkenson provides the examples of values that are isolated from each other, and therefore discrete, and values lying on a continuum or in a compact region.
-We propose that a robust model of continuity, which is how the elements in a dataset are connected to each other, provides a way to develop library components that can
-
+Continuity describes elements in a data set are organized; this concept is termed topological properties by Wilkinson\cite{wilkinsonGrammarGraphics2005}.
+Wilkinson provides the examples of values that are isolated from each other, and therefore discrete, and values lying on a continuum or in a compact region.
 For example, in \autoref{fig:related-work:continuity}, each station record in the table is independent of the others; therefore, the continuity of the table is discrete.
@@
 Every point on the globe is connected to its 6 nearest cardinal neighboring points (NW, N, NE, E, SE, S, SW, W).
+We propose that a robust model of continuity provides a way to develop library components that can work on the data in small pieces in a manner where the overall continuity of the data is preserved in the visual transformation.
 

### 2021-10-07 paper/main.tex
paper

@@
 
-Here we present a very brief summary of topics, for more information see Hatcher\cite{hatcherAlgebraicTopology2002}, Munkres\cite{munkresElementsAlgebraicTopology1984}, and Bradley et.
-al.
-\cite{bradleyTopologyCategoricalApproach2020}.
+Here we present a very brief summary of topics, for more information Hatcher\cite{hatcherAlgebraicTopology2002}, Munkres\cite{munkresElementsAlgebraicTopology1984}, and Spanier \cite{spanier1989algebraic}.
 
@@
 al provide a brief introduction to category theory for visualization practitioners \cite{vickersUnderstandingVisualizationFormal2013}, but their work focuses on data, representation, and evocation, while this paper is aims to provide guidance on how the map from data to representation should be implemented.
+
+For more information on category theory, see Barr and Wells\cite{barrCategoryTheoryComputing}, Fong and Spivak \cite{fongInvitationAppliedCategory2019}, Riehl\cite{riehlCategoryTheoryContext} and Bradley et.
+al.
+\cite{bradleyTopologyCategoricalApproach2020}.
 

### 2021-10-11 paper/main.tex
thursday meeting on stalks + notes
sheaves

@@
 
-We define the categories $\mathcal{\dtotal}, \mathcal{\gtotal}$ as sheaves of sections\cite{SheafMathematics2021,spanier1989algebraic}, of fiberbundles, as introduced in \autoref{sec:related-work:fiber-bundles}.
+We define the categories $\mathcal{\dtotal}$ and $\mathcal{\gtotal}$ as sheaves of sections\cite{SheafMathematics2021,spanier1989algebraic}, of fiberbundles, as introduced in \autoref{sec:related-work:fiber-bundles}.
+Sheaves are mathematical abstractions of how data over different parts of a continuous base space are glued together \cite{ghristElementaryAppliedTopology2014, ghri}, whether the data is distributed, streaming, or in disk; therefore sheaves are a way to express the continuity that the artist preserves.
 We assume that $\mathcal{E}$ is defined in terms of a fixed base and fiber space, for example a class of images or tables with the same schema.
-The graphic category $\mathcal{\gtotal}$ consists of the set of graphics that an artist \vartist is allowed to generate from a given $\mathcal{E}$, which is discussed further in \autoref{sec:artist:construction}.
+The graphic category $\mathcal{\gtotal}$ consists of the set of graphics that an artist \vartist\ is allowed to generate from a given $\mathcal{E}$, which is discussed further in \autoref{sec:artist:construction}.
 
@@
   \begin{itemize}
-    \item\textit{object} the contravariant functor $\sheaf: \mathcal{K}^{op} \rightarrow \Gamma(\mathcal{K}, \dtotal)$
-    \item \textit{morphisms} the endomorphic\cite{fongInvitationAppliedCategory2019} bundle maps $\bullet: \dtotal\rightarrow \dtotal$, which are the fiber maps $\monoid:\mathcal{\dfiber}\rightarrow \mathcal{\dfiber}$ when the sheaf is over a single point in the base space $\dbasepoint\in \dbase$
+    \item\textit{object} the sheaf\cite{SheafMathematics2021,ghristElementaryAppliedTopology2014} $\sheaf: \mathcal{K}^{op} \rightarrow \Gamma(\mathcal{K}, \dtotal)$
+    \item \textit{morphisms} the endomorphic\cite{fongInvitationAppliedCategory2019} bundle maps $\bullet: \dtotal\rightarrow \dtotal$, which are the fiber maps $Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ when the sheaf is restricted to the stalk\cite{StalkSheaf2019} $\mathscr{F}_{k}$ over a point $\dbasepoint\in \dbase$
   \end{itemize}
@@
 \end{equation}
-the functor \sheaf\ is contravariant because the inclusion map $\iota$ between objects in $\mathcal{K}$ goes the in the opposite direction from the map $\iota$ between open set objects in $\mathcal{K}$ \cite{riehlCategoryTheoryContext}.
+the functor \sheaf\ is contravariant because the inclusion map $\iota$ between open sets in $\mathcal{K}$ goes in the opposite direction from the map $\iota^*$ between the sections in $\Gamma(U_{i}, \dtotal)$ \cite{riehlCategoryTheoryContext}.
 The open set $U_1$ is a subset of $U_2$, therefore $\iota: U_1 \rightarrow U_2$.
@@
 We use the notation $\mathcal{K}^{op}$ to signify that the functor $\sheaf$ is contravariant and $\sheaf$ to denote that it is a sheaf.
-A sheaf is a presheaf where sections are defined over unions of open sets \cite{nlab:sheaf,ghristElementaryAppliedTopology2014}
+A sheaf is a presheaf where sections are defined over unions of open sets \cite{nlab:sheaf,ghristElementaryAppliedTopology2014}.
+The sheaf over a limit of open sets that contain a point \dbasepoint\ is approximately the same as a sheaf over a point \dbase\ and is called the stalk $\mathscr{F}_{k}\coloneqq \lim\limits_{U\ni k} \mathscr{F}(U)$ \cite{StalkSheaf2019}.
+The fiber space over a point is embedded in the stalk over that point $\dfiber_{k}\hookrightarrow \mathscr{F}_{k}$.
+Our criteria for equivariance is that the set of maps from fiber to fiber $Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ is preserved, since these maps include the types of data transformations described in \autoref{sec:related-work:equivariance}.
+When a fiber bundle is trivial, meaning $\dtotal=\dbase \times \dfiber$, then we can dispense with localization and directly consider the fiber maps $Hom(F_{k}, F_{k})$.
+In \autoref{sec:artist:construction}, we assume that the fiber bundles are trivial.
 
@@
 The artist is a morphism of sheafs $\vartist: \mathcal{\dtotal} \rightarrow \mathcal{\gtotal}$, which means it is by definition a natural transform\cite{SheafMathematics2021}.
-This means that the following diagram must commute
+\note{which means?
+why is to good for it to be a natural transformation->}
 
+%% 
+Extra property that implies equivariance is preserved is illustrated in the diagram which must commute
+%% note Im_A in \H, A_M (lifting/pushforward)
 \begin{equation}
@@
 
+%H is the codomain, ImA is the range, K & S -> 
+% clipping problem: shift lines and stuff w/o shifting frame, when we shift the frame, only shifting the bits of H that belongs to the artist  
 such that the morphisms on the graphic category $\mathcal{H}$ are homomorphic to the morphisms $\monoid$ on the fibers $\mathcal{F}$ \cite{spanier1989algebraic,fongInvitationAppliedCategory2019}.
+\note{equivariant not homomorphic, this isn't saying anything, change Hom(F,F)}
+
 In \autoref{sec:artist:construction}, we show a method of constructing the artist such that this constraint can be satisfied, thereby ensuring equivariance.
@@
 
-\begin{min
+\begin{minted}{python}
+\end{minted}
 

### 2021-10-12 paper/main.tex
nat transforms

@@
 We use the notation $\mathcal{K}^{op}$ to signify that the functor $\sheaf$ is contravariant and $\sheaf$ to denote that it is a sheaf.
+We also denote the sheaf over the fiberbundle \dtotal\ as $\sheaf(\dtotal)$.
 A sheaf is a presheaf where sections are defined over unions of open sets \cite{nlab:sheaf,ghristElementaryAppliedTopology2014}.
@@
 
-The graphic category $\mathcal{\gtotal}$ consists of the functor $\sheaf: \mathcal{S}^{op} \rightarrow \Gamma(\mathcal{S}, \gtotal)$.
+The graphic category $\mathcal{\gtotal}$ consists of the functor $\sheaf: \mathcal{S}^{op} \rightarrow \Gamma(\mathcal{S}, \gtotal)$, which we also denote as $\sheaf(\gtotal)$.
 The category $\mathcal{S}$ is the category of open sets over the graphic base space \gbase\, and $\Gamma(\mathcal{S}, \gtotal)$ is the set of sections $\gsection$ over $\mathcal{\gbase}$.
-The artist is a morphism of sheafs $\vartist: \mathcal{\dtotal} \rightarrow \mathcal{\gtotal}$, which means it is by definition a natural transform\cite{SheafMathematics2021}.
-\note{which means?
-why is to good for it to be a natural transformation->}
+The artist is a morphism of sheafs $\vartist: \mathcal{\dtotal} \rightarrow \mathcal{\gtotal}$; therefore it is a map between functors, which means that the artist is a natural transformation\cite{SheafMathematics2021,bradleyWhatNaturalTransformation}.
 
-%% 
 Extra property that implies equivariance is preserved is illustrated in the diagram which must commute
@@
   \begin{tikzcd}
-  \mathcal{\dtotal} \arrow[r, "\vartist"] \arrow[d, "\monoid"'] & \mathcal{\gtotal} \arrow[d, "\vartist(\monoid)"] \\
+    \mathcal{\dtotal} \arrow[r, "\vartist"] \arrow[d, "{Hom_{\mathcal{\dfiber}}(\dfiber, \dfiber)}"'] & \mathcal{\gtotal} \arrow[d, "\vartist_{Hom_{\mathcal{\dfiber}}}"] \\
     \mathcal{\dtotal} \arrow[r, "\vartist"]                                                           & \mathcal{\gtotal}                                                

### 2021-10-12 paper/main.tex
natural transfroms

--- 2021-10-11 paper/main.tex
@@
 We propose that the visualization transformation can be modeled as a functor, which we call the \textcolor{artist}{Artist} $\mathcal{\vartist}$\footnote{We call this transformation the artist because the \mintinline{python}{Artist} object in Matplotlib\cite{hunterMatplotlib2DGraphics2007}}.
-The artist \vartist converts data, which we model as the catgory $\mathcal{\dtotal}$ to graphics, which we model as the category $\mathcal{\gtotal}$.}
+The artist \vartist\ converts data, which we model as the catgory $\mathcal{\dtotal}$ to graphics, which we model as the category $\mathcal{\gtotal}$.
 
@@
   \vartist: \mathcal{\dtotal} \rightarrow \mathcal{\gtotal}
-  \label{eq:math:artist:artist}
+  \label{eq:artist:artist}
 \end{equation}
@@
   \label{def:category:F} 
-  The category $\mathcal{\dfiber}$ consists of a single object $\dfiber$ and morphisms between the object and itself $\monoid: \dfiber \rightarrow \dfiber$. 
+  The category $\mathcal{\dfiber}$ consists of a single object $\dfiber$ and the morphisms between the object and itself $Hom_{\mathcal{\dfiber}}(\dfiber, \dfiber)$, which is the hom-set \cite{maclaneCategoriesWorkingMathematician2013,milewskiCategoryTheoryProgrammers}.
 \end{definition}
@@
 For example, a lists of strings is an instance of an object in \textbf{Set}, networks are an instance of \textbf{Graph}, and images are vector spaces which are a specific type of topological space \textbf{Top}.
-We note that a one object category is a monoid \cite{nlab:monoid,barrCategoryTheoryComputing}.
-\note{&?
-why is this important} The morphisms $\monoid$ are functions from the data to itself, for example the binary operations, group actions, and measurement scales discussed in \autoref{sec:related-work:equivariance}.
+We note that a one object category is a monoid \cite{nlab:monoid,barrCategoryTheoryComputing}, which is an algebraic structure that lends itself to computer library design because they are an abstraction of function composition \cite{yorgeyMonoidsThemeVariations}.
+
+The morphisms $Hom_{\mathcal{\dfiber}}(\dfiber, \dfiber)$ are functions from the data to itself, for example the binary operations, group actions, and measurement scales discussed in \autoref{sec:related-work:equivariance}.
 These functions could also be monoid actions\cite{barrCategoryTheoryComputing}, which provide a way of applying partial order relations to data\cite{fongInvitationAppliedCategory2019}, such as to build multi-ranked indicators\cite{bruggemannRankingPrioritizationMultiindicator2011}
 
-We encode individual datasets in the fiberbundle as sections, and we follow the convention that $\Gamma(U_i, \dtotal)$ denotes the set of sections $\{\dsection^{j}:U_i \rightarrow \dfiber_{U_{i}}\}_{j\in I}$ of the fiber bundle \dtotal\ over an open set $\U_{i}$.
+We encode individual datasets in the fiberbundle as sections of the bundle.
+We follow the convention that $\Gamma(U_i, \dtotal)$ denotes the set of sections $\{\dsection^{j}:U_i \rightarrow \dfiber_{U_{i}}\}_{j\in I}$ of the fiber bundle \dtotal\ over an open set $U_{i}\subseteq \dbase$.
 As mentioned in \autoref{sec:related-work:fiber-bundles}, the fiber space is embedded in the total space $\dfiber \hookrightarrow \dtotal$.
@@
   \begin{itemize}
-    \item\textit{object} the sheaf\cite{SheafMathematics2021,ghristElementaryAppliedTopology2014} $\sheaf: \mathcal{K}^{op} \rightarrow \Gamma(\mathcal{K}, \dtotal)$
+    \item\textit{object} the sheaf\cite{SheafMathematics2021,ghristElementaryAppliedTopology2014} $\sheaf_{\mathcal{\dtotal}}: \mathcal{K}^{op} \rightarrow \Gamma(\mathcal{K}, \dtotal)$
     \item \textit{morphisms} the endomorphic\cite{fongInvitationAppliedCategory2019} bundle maps $\bullet: \dtotal\rightarrow \dtotal$, which are the fiber maps $Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ when the sheaf is restricted to the stalk\cite{StalkSheaf2019} $\mathscr{F}_{k}$ over a point $\dbasepoint\in \dbase$
@@
 \begin{equation}
-  \label{eq:artist:[presheaf}
+  \label{eq:artist:presheaf}
   \begin{tikzcd}[row sep=small]
@@
 
-The graphic category $\mathcal{\gtotal}$ consists of the functor $\sheaf: \mathcal{S}^{op} \rightarrow \Gamma(\mathcal{S}, \gtotal)$.
+The graphic category $\mathcal{\gtotal}$ consists of the functor $\sheaf_{\mathcal{\gtotal}}: \mathcal{S}^{op} \rightarrow \Gamma(\mathcal{S}, \gtotal)$, which we also denote as $\sheaf(\gtotal)$.
 The category $\mathcal{S}$ is the category of open sets over the graphic base space \gbase\, and $\Gamma(\mathcal{S}, \gtotal)$ is the set of sections $\gsection$ over $\mathcal{\gbase}$.
-The artist is a morphism of sheafs $\vartist: \mathcal{\dtotal} \rightarrow \mathcal{\gtotal}$, which means it is by definition a natural transform\cite{SheafMathematics2021}.
-\note{which means?
-why is to good for it to be a natural transformation->}
+The artist is a morphism of sheafs
+\begin{equation}
+  \label{eq:artist:morphism-of-sheaf}
+  \begin{tikzcd}
+    & \Gamma(\dtotal) & & & \Gamma(\gtotal) & \\
+{} \arrow[rr, "f", Rightarrow, shift right, shorten=3em] & 
+& 
+{} \arrow[r, "\vartist", shorten=-1.25em, bold] & {} \arrow[rr, "g", Rightarrow, shift right, shorten=3em] & & {} \\
+& \mathcal{\dbase}^{op} \arrow[uu, "\mathcal{\dtotal}\coloneqq\sheaf_{\mathcal{\dtotal}}", bend left] 
+  \arrow[uu, "\sheaf_{\mathcal{\dtotal}}"', bend right] 
+  & & 
+& \mathcal{\gbase}^{op} \arrow[uu, "\sheaf_{\mathcal{\gtotal}}", bend left] 
+  \arrow[uu, "\mathcal{\gtotal}\coloneqq \sheaf_{\mathcal{\gtotal}}"', bend right] &   
+  \end{tikzcd}
+\end{equation}
 
-%% 
-Extra property that implies equivariance is preserved is illustrated in the diagram which must commute
-%% note Im_A in \H, A_M (lifting/pushforward)
+therefore the artist is a map between functors
 \begin{equation}
-  \label{eq:artist:natural-transform}
+  \tag{\ref{eq:artist:artist}}
+  \vartist:\mathcal{\dtotal}\rightarrow{\mathcal{\gtotal}} 
+\end{equation} which means that the artist is a natural transformation\cite{SheafMathematics2021,bradleyWhatNaturalTransformation}.
+One of the properties of a natural transformation is that the following diagram commutes
+\begin{equation}
+  \label{eg:artist:natural-transformation}
   \begin{tikzcd}
-  \mathcal{\dtotal} \arrow[r, "\vartist"] \arrow[d, "\monoid"'] & \mathcal{\gtotal} \arrow[d, "\vartist(\monoid)"] \\
+    \mathcal{\dtotal} \arrow[d, "f"'] \arrow[r, "\vartist"] & \mathcal{\gtotal} \arrow[d, "\vartist_{f} \coloneqq g"] \\
     \mathcal{\dtotal} \arrow[r, "\vartist"]                 & \mathcal{\gtotal}                                      
@@
 
-%H is the codomain, ImA is the range, K & S -> 
-% clipping problem: shift lines and stuff w/o shifting frame, when we shift the frame, only shifting the bits of H that belongs to the artist  
+meaning that functions $f: \mathcal{\dtotal}\rightarrow \mathcal{\dtotal}$
+
+Extra property that implies equivariance is preserved is illustrated in the diagram which must commute
+%% note Im_A in \H, A_M (lifting/pushforward)
+\begin{equation}
+  \label{eq:artist:equivariance}
+  \begin{tikzcd}
+    \mathcal{\dtotal} \arrow[r, "\vartist"] \arrow[d, "{Hom_{\mathcal{\dfiber}}(\dfiber, \dfiber)}"'] & Im_{\vartist} \in \mathcal{\gtotal} \arrow[d, "\vartist_{Hom_{\mathcal{\dfiber}}}"] \\
+    \mathcal{\dtotal} \arrow[r, "\vartist"]                                                           & Im_{\vartist} \in \mathcal{\gtotal}                                                
+  \end{tikzcd}
+\end{equation}
+
 such that the morphisms on the graphic category $\mathcal{H}$ are homomorphic to the morphisms $\monoid$ on the fibers $\mathcal{F}$ \cite{spanier1989algebraic,fongInvitationAppliedCategory2019}.
-\note{equivariant not homomorphic, this isn't saying anything, change Hom(F,F)}
 
-In \autoref{sec:artist:construction}, we show a method of constructing the artist such that this constraint can be satisfied, thereby ensuring equivariance.
+In \autoref{sec:artist:construction}, we show a method of constructing the artist such that the constraints of \autoref{eq:artist:morphism-of-sheaf} and \autoref{eq:artist:equivariance} can be satisfied, thereby ensuring equivariance and the preservation of continuity.
 We then use this formulation to guide the development of visualization library components in \autoref{sec:case-study}.

### 2021-10-14 paper/main.tex
notes

@@
 
+\note{roadmap the rest of this section}
+
+\subsection{Categorical Artist}
 We define the categories $\mathcal{\dtotal}$ and $\mathcal{\gtotal}$ as sheaves of sections\cite{SheafMathematics2021,spanier1989algebraic}, of fiberbundles, as introduced in \autoref{sec:related-work:fiber-bundles}.
@@
 
+%%define a category of sections here
 We encode individual datasets in the fiberbundle as sections of the bundle.
@@
     \item\textit{object} the sheaf\cite{SheafMathematics2021,ghristElementaryAppliedTopology2014} $\sheaf_{\mathcal{\dtotal}}: \mathcal{K}^{op} \rightarrow \Gamma(\mathcal{K}, \dtotal)$
-    \item \textit{morphisms} the endomorphic\cite{fongInvitationAppliedCategory2019} bundle maps $\bullet: \dtotal\rightarrow \dtotal$, which are the fiber maps $Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ when the sheaf is restricted to the stalk\cite{StalkSheaf2019} $\mathscr{F}_{k}$ over a point $\dbasepoint\in \dbase$
+    \item \textit{morphisms} arbitrary endomorphic\cite{fongInvitationAppliedCategory2019} bundle maps $f: \dtotal\rightarrow \dtotal$, which are the fiber maps $Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ when the sheaf is restricted to the stalk\cite{StalkSheaf2019} $\mathscr{F}_{k}$ over a point $\dbasepoint\in \dbase$
   \end{itemize}
@@
 
-The functor object of $\mathcal{E}$ is a presheaf because the definition of a presheaf is that it is a contravariant functor from a category $\mathcal{K}$ to a set \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}.
+The functor object of $\mathcal{E}$ is a presheaf because the definition of a presheaf is that it is a contravariant functor from a category to a category of sets \textbf{Set} \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}.
 As illustrated in the diagram,
@@
 \end{equation}
-the functor \sheaf\ is contravariant because the inclusion map $\iota$ between open sets in $\mathcal{K}$ goes in the opposite direction from the map $\iota^*$ between the sections in $\Gamma(U_{i}, \dtotal)$ \cite{riehlCategoryTheoryContext}.
+%% sheaf maps over the inclusion rules which is why we care about them ->
+the functor \sheaf\ is contravariant because the inclusion map $\iota$ between open sets $U_i \in \mathcal{K}$ goes in the opposite direction from the map $\iota^*$ between the sections in $\Gamma(U_{i}, \dtotal)$ \cite{riehlCategoryTheoryContext}.
 The open set $U_1$ is a subset of $U_2$, therefore $\iota: U_1 \rightarrow U_2$.
@@
 A sheaf is a presheaf where sections are defined over unions of open sets \cite{nlab:sheaf,ghristElementaryAppliedTopology2014}.
-The sheaf over a limit of open sets that contain a point \dbasepoint\ is approximately the same as a sheaf over a point \dbase\ and is called the stalk $\mathscr{F}_{k}\coloneqq \lim\limits_{U\ni k} \mathscr{F}(U)$ \cite{StalkSheaf2019}.
+The sheaf over a limit of open sets that contain a point \dbasepoint\ is approximately the same as a sheaf over a point \dbase\ and is called the stalk $\mathscr{F}_{k}\coloneqq \lim\limits_{U\ni k} \Gamma(U, \dtotal)$ \cite{StalkSheaf2019}.
 The fiber space over a point is embedded in the stalk over that point $\dfiber_{k}\hookrightarrow \mathscr{F}_{k}$.
@@
 The artist is a morphism of sheafs
+
+%is not \Gamma(E), make a category out of it
 \begin{equation}
@@
 
+% f is a natural transform, all transforms on the sheafs commute
+% g is a natural transform ...
+% space of g->g is too big, so we look at Im...
+% b/c it's a natural transform, - f carries i over b/c it's a natural transformation
+
 therefore the artist is a map between functors
@@
   \vartist:\mathcal{\dtotal}\rightarrow{\mathcal{\gtotal}} 
-\end{equation} which means that the artist is a natural transformation\cite{SheafMathematics2021,bradleyWhatNaturalTransformation}.
-One of the properties of a natural transformation is that the following diagram commutes
-\begin{equation}
-  \label{eg:artist:natural-transformation}
-  \begin{tikzcd}
-    \mathcal{\dtotal} \arrow[d, "f"'] \arrow[r, "\vartist"] & \mathcal{\gtotal} \arrow[d, "\vartist_{f} \coloneqq g"] \\
-    \mathcal{\dtotal} \arrow[r, "\vartist"]                 & \mathcal{\gtotal}                                      
-  \end{tikzcd}
 \end{equation} 
+which means that the artist is a natural transformation\cite{SheafMathematics2021,bradleyWhatNaturalTransformation}.
 
-meaning that functions $f: \mathcal{\dtotal}\rightarrow \mathcal{\dtotal}$
+% such that (9)
 
@@
 such that the morphisms on the graphic category $\mathcal{H}$ are homomorphic to the morphisms $\monoid$ on the fibers $\mathcal{F}$ \cite{spanier1989algebraic,fongInvitationAppliedCategory2019}. 
+%% such that monoid on this side is monoid on that side
 
 In \autoref{sec:artist:construction}, we show a method of constructing the artist such that the constraints of \autoref{eq:artist:morphism-of-sheaf} and \autoref{eq:artist:equivariance} can be satisfied, thereby ensuring equivariance and the preservation of continuity.
-We then use this formulation to guide the development of visualization library components in \autoref{sec:case-study}.
+We then use this formulation to guide the development of visualization
+library components in \autoref{sec:case-study}.
 
-\subsection{Union of Artists}
+\subsection{Artist Composition}
 \label{sec:artist:union}
@@
 \end{figure}
+%% artist composition is fiber products/sums of artists on sheafs -> closed cartesian category maybe?
 
@@
 
-\section{Constraints: Construction and formal properties of Artists}
+\section{Construction and formal properties of Artists}
 \label{sec:artist:construction}

### 2021-10-14 paper/main.tex
section categories and preimages

@@
 
-%%define a category of sections here
 We encode individual datasets in the fiberbundle as sections of the bundle.
 We follow the convention that $\Gamma(U_i, \dtotal)$ denotes the set of sections $\{\dsection^{j}:U_i \rightarrow \dfiber_{U_{i}}\}_{j\in I}$ of the fiber bundle \dtotal\ over an open set $U_{i}\subseteq \dbase$.
+
+\begin{definition} The category $\textcolor{section}{\mathcal{T}}$ consists of 
+  \label{def:category:Tau}
+  \begin{itemize}
+  \item \textit{object} the set of sections $\Gamma(U_i)$ where $U_i$ can be any open set object in $\mathcal{\dbase}$
+  \item \textit{morphisms} fiber maps $Hom(F, F)$, because sections are maps from a point in the base space to a point in the fiber $\tau: \dbase \rightarrow \dfiber$
+  \end{itemize}
+\end{definition}
+
 As mentioned in \autoref{sec:related-work:fiber-bundles}, the fiber space is embedded in the total space $\dfiber \hookrightarrow \dtotal$.
@@
     \item\textit{object} the sheaf\cite{SheafMathematics2021,ghristElementaryAppliedTopology2014} $\sheaf_{\mathcal{\dtotal}}: \mathcal{K}^{op} \rightarrow \Gamma(\mathcal{K}, \dtotal)$
-    \item \textit{morphisms} arbitrary endomorphic\cite{fongInvitationAppliedCategory2019} bundle maps $f: \dtotal\rightarrow \dtotal$, which are the fiber maps $Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ when the sheaf is restricted to the stalk\cite{StalkSheaf2019} $\mathscr{F}_{k}$ over a point $\dbasepoint\in \dbase$
+    \item \textit{morphisms} arbitrary endomorphic\cite{fongInvitationAppliedCategory2019} bundle maps $f: \dtotal\rightarrow \dtotal$.
+When the sheaf is restricted to the stalk\cite{StalkSheaf2019} $\mathscr{F}_{k}$ over a point $\dbasepoint\in \dbase$, then allowable morphisms are $f \in Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$.
   \end{itemize}
@@
 
-The functor object of $\mathcal{E}$ is a presheaf because the definition of a presheaf is that it is a contravariant functor from a category to a category of sets \textbf{Set} \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}.
-As illustrated in the diagram,
+ The artist takes as input the sheaf object $\mathcal{E}\coloneqq\sheaf_{\mathcal{E}}$ because the sheaf is a way of keeping track of how data is glued together \cite{ghristElementaryAppliedTopology2014}.
+This bookkeeping is done via the inclusion maps that are preserved by the presheaf, as illustrated in \autoref{eq:artist:presheaf}.
+Sheafs are presheafs where sections are defined over unions of open sets $U \cup_{U\in \mathscr{U}}$ \cite{spanier1989algebraic, nlab:sheaf}.
+As illustrated in \autoref{eq:artist:presheaf}, the functor $\sheaf_{\mathcal{\dtotal}}$ is a contravariant functor from a category to a category of sets \textbf{Set} \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}, which means it is by definition a presheaf.
+
 \begin{equation}
@@
     U_1 \arrow[dd, "\iota"'] & {\Gamma(U_1, \dtotal)}                              \\
-    {} \arrow[r, "\sheaf"]   & {}                                                  \\
+    {} \arrow[r, "\mathcal{\dtotal}\coloneqq\sheaf_{\mathcal{\dtotal}}"]   & {}                                                  \\
     U_2                      & {\Gamma(U_2, \dtotal)} \arrow[uu, "\iota^*"']
@@
 \end{equation}
-%% sheaf maps over the inclusion rules which is why we care about them ->
-the functor \sheaf\ is contravariant because the inclusion map $\iota$ between open sets $U_i \in \mathcal{K}$ goes in the opposite direction from the map $\iota^*$ between the sections in $\Gamma(U_{i}, \dtotal)$ \cite{riehlCategoryTheoryContext}.
+
+ The functor $\sheaf_{\mathcal{\dtotal}}$ is contravariant because the inclusion map $\iota$ between open sets $U_i \in \mathcal{K}$ goes in the opposite direction from the map $\iota^*$ between the sections in $\Gamma(U_{i}, \dtotal)$ \cite{riehlCategoryTheoryContext}.
 The open set $U_1$ is a subset of $U_2$, therefore $\iota: U_1 \rightarrow U_2$.
@@
 We use the notation $\mathcal{K}^{op}$ to signify that the functor $\sheaf$ is contravariant and $\sheaf$ to denote that it is a sheaf.
-We also denote the sheaf over the fiberbundle \dtotal\ as $\sheaf(\dtotal)$.
-A sheaf is a presheaf where sections are defined over unions of open sets \cite{nlab:sheaf,ghristElementaryAppliedTopology2014}.
-The sheaf over a limit of open sets that contain a point \dbasepoint\ is approximately the same as a sheaf over a point \dbase\ and is called the stalk $\mathscr{F}_{k}\coloneqq \lim\limits_{U\ni k} \Gamma(U, \dtotal)$ \cite{StalkSheaf2019}.
+We also denote the sheaf over the total space \dbase\ of the fiber bundle \dtotal\ as $\sheaf(\dtotal)$.
+
+Arbitrary morphisms on sheafs $f:\mathcal{\dtotal} \rightarrow \mathcal{\dtotal}$ are natural transforms \cite{SheafMathematics2021,bradleyWhatNaturalTransformation}, meaning that the morphisms $f$ compose with the inclusion maps $\iota$ \cite{fongInvitationAppliedCategory2019}.
+As defined in \autoref{def:category:Tau}, the sets of sections $\Gamma(U_{i},\dtotal)$ are objects of the category $\mathcal{T}$.
+
+\begin{equation}
+  \label{eq:artist:morphism-of-sheaf-data}
+  \begin{tikzcd}
+    & \mathcal{T} & \\
+{} \arrow[rr, "f", Rightarrow, shift right, shorten=3em] & & {} \\
+    & \mathcal{\dbase}^{op} \arrow[uu, "\sheaf_{\mathcal{\dtotal}}", bend left] \arrow[uu, "\sheaf_{\mathcal{\dtotal}}"', bend right] &   
+\end{tikzcd}
+\end{equation}
+
+The sheaf $\sheaf_{\mathcal{\dtotal}}$ over a limit of open sets that contain a point $\dbasepoint \in \dbase$ is approximately the same as a sheaf over a point \dbasepoint\ and is called the stalk $\sheaf_{\mathcal{\dtotal}}|_{k}\coloneqq \lim\limits_{U\ni k} \Gamma(U, \dtotal)$ \cite{StalkSheaf2019}.
 The fiber space over a point is embedded in the stalk over that point $\dfiber_{k}\hookrightarrow \mathscr{F}_{k}$.
-Our criteria for equivariance is that the set of maps from fiber to fiber $Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ is preserved, since these maps include the types of data transformations described in \autoref{sec:related-work:equivariance}.
-When a fiber bundle is trivial, meaning $\dtotal=\dbase \times \dfiber$, then we can dispense with localization and directly consider the fiber maps $Hom(F_{k}, F_{k})$.
+Our criteria for equivariance is that bundle maps $f$ that are in the set of fiber to fiber maps $Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ are preserved.
+The maps $f\in Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ include the types of data transformations described in \autoref{sec:related-work:equivariance}.
+When a fiber bundle is trivial, meaning $\dtotal=\dbase \times \dfiber$, then we can dispense with localization and directly consider the fiber maps $Hom(\dfiber_{k}, \dfiber_{k})$ over a point \dbasepoint.
 In \autoref{sec:artist:construction}, we assume that the fiber bundles are trivial.
 
-The graphic category $\mathcal{\gtotal}$ consists of the functor $\sheaf_{\mathcal{\gtotal}}: \mathcal{S}^{op} \rightarrow \Gamma(\mathcal{S}, \gtotal)$, which we also denote as $\sheaf(\gtotal)$.
-The category $\mathcal{S}$ is the category of open sets over the graphic base space \gbase\, and $\Gamma(\mathcal{S}, \gtotal)$ is the set of sections $\gsection$ over $\mathcal{\gbase}$.
-The artist is a morphism of sheafs
+The graphic category $\mathcal{\gtotal}$ consists of the functor object $\sheaf_{\mathcal{\gtotal}}: \mathcal{S}^{op} \rightarrow \Gamma(\mathcal{S}, \gtotal)$ and the sheaf morphisms $g:\sheaf_{\mathcal{\gtotal}} \rightarrow \sheaf_{\mathcal{gtotal}}$.
+The category $\mathcal{S}$ is the category of open sets over the graphic base space \gbase, and the category $\mathcal{P}$ is the set of sections $\gsection$ over $\mathcal{\gbase}$.
+The category of possible graphics $\mathcal{P}$ has as objects all graphics with a shared continuity and display space; bar plots and scatter plots on a 2D screen are two examples of objects in $\mathcal{P}$.
+As will be discussed in \autoref{sec:artist:construction}, the  $\vartist$ is constructed in such a way that it generates a single type of graphic.
+Therefore, we define the object as a morphism from data to the sheaf mage of the artist in the graphic space $\vartist: \mathcal{\dtotal} \rightarrow Im_A \in $
 
-%is not \Gamma(E), make a category out of it
 \begin{equation}
@@
   \begin{tikzcd}
-    & \Gamma(\dtotal) & & & \Gamma(\gtotal) & \\
+    & \mathcal{T} & & & \mathcal{P} & \\
 {} \arrow[rr, "f", Rightarrow, shift right, shorten=3em] & 
@@
 {} \arrow[r, "\vartist", shorten=-1.25em, bold] & {} \arrow[rr, "g", Rightarrow, shift right, shorten=3em] & & {} \\
-& \mathcal{\dbase}^{op} \arrow[uu, "\mathcal{\dtotal}\coloneqq\sheaf_{\mathcal{\dtotal}}", bend left] 
-  \arrow[uu, "\sheaf_{\mathcal{\dtotal}}"', bend right] 
+& \mathcal{\dbase}^{op} \arrow[uu, "\mathcal{\dtotal}", bend left] 
+  \arrow[uu, "\mathcal{\dtotal}"', bend right] 
   & & 
-& \mathcal{\gbase}^{op} \arrow[uu, "\sheaf_{\mathcal{\gtotal}}", bend left] 
-  \arrow[uu, "\mathcal{\gtotal}\coloneqq \sheaf_{\mathcal{\gtotal}}"', bend right] &   
+& \mathcal{\gbase}^{op} \arrow[uu, "\mathcal{\gtotal}", bend left] 
+  \arrow[uu, "\mathcal{\gtotal}"', bend right] &   
   \end{tikzcd}
@@
 \note{formalize superposition, than break down what that means for composition}
+\subsubsection{Superposition: union}
 
@@
 
+\subsection{Shared properties: intersection}
 Many times though, graphics are generated from subsets of a larger data set, such as when using multiple views \cite{alboRadarComparativeEvaluation2016a, hullmanKeeping2018} and brush-linked views \cite{beckerBrushingScatterplots1987,bujaInteractiveData1991}.

### 2021-10-15 paper/main.tex
notes from thursday + some changes

@@
 
-The inclusion maps $iota$ can be used to piece together the subsets $U_{i}$, particularly individual points $\dbasepoint \in \dbase$.
+The inclusion maps $\iota$ can be used to piece together the subsets $U_{i}$, particularly individual points $\dbasepoint \in \dbase$.
 This is how we provide knowledge of continuity to the artist \vartist\ in a way where the artist does not need all the data at once.
@@
   \begin{itemize}
-  \item \textit{object} the set of sections $\Gamma(U_i)$ where $U_i$ can be any open set object in $\mathcal{\dbase}$
-  \item \textit{morphisms} fiber maps $Hom(F, F)$, because sections are maps from a point in the base space to a point in the fiber $\tau: \dbase \rightarrow \dfiber$
+  \item \textit{object} the set of sections $\Gamma(U, \dtotal)$ where $U$ can be any object in $\mathcal{\dbase}$
+  \item \textit{morphisms} restrictions \note{\iota which we will use in (6) so 7 is cleanly abstracted}
   \end{itemize}

### 2021-10-18 paper/main.tex
removed tau reworked with var phi

@@
 
-\begin{definition} The category $\textcolor{section}{\mathcal{T}}$ consists of 
-  \label{def:category:Tau}
-  \begin{itemize}
-  \item \textit{object} the set of sections $\Gamma(U, \dtotal)$ where $U$ can be any object in $\mathcal{\dbase}$
-  \item \textit{morphisms} restrictions \note{\iota which we will use in (6) so 7 is cleanly abstracted}
-  \end{itemize}
+\begin{definition} The category of sets of sections $\Gamma(U_i,\dtotal)$, where $U_i$ are objects in $\mathcal{\dfiber}$, form a subcatagory of the category of sets \textbf{Set}.
+The morphisms between objects of this subcategory are restriction maps $\iota: \Gamma(U_j, \dtotal) \rightarrow \Gamma(U_i, \dtotal)$ where $U_{i} \subseteq U_{j}$.
 \end{definition}
@@
 The sections $\Gamma(\dbase, \dtotal)$ are called the global sections of $\dtotal$ \cite{SheafMathematics2021,spanier1989algebraic}.
-Each section is how we represent a unique dataset, where the set of sections $\Gamma$ is the set of all datasets that have the same data fields in the \dfiber\ and continuity as encoded in\dbase\.
+Each section is how we represent a unique dataset, where the sets of sections $\Gamma$ is the space of all datasets that have the same data fields in the \dfiber\ and continuity as encoded in\dbase\.
 For example, Spivak describes tables that have the same schema and discrete index as sections of the same fiber bundle\cite{spivakSIMPLICIALDATABASES}.
@@
     \item\textit{object} the sheaf\cite{SheafMathematics2021,ghristElementaryAppliedTopology2014} $\sheaf_{\mathcal{\dtotal}}: \mathcal{K}^{op} \rightarrow \Gamma(\mathcal{K}, \dtotal)$
-    \item \textit{morphisms} arbitrary endomorphic\cite{fongInvitationAppliedCategory2019} bundle maps $f: \dtotal\rightarrow \dtotal$.
-When the sheaf is restricted to the stalk\cite{StalkSheaf2019} $\mathscr{F}_{k}$ over a point $\dbasepoint\in \dbase$, then allowable morphisms are $f \in Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$.
+    \item \textit{morphisms} arbitrary endomorphic\cite{fongInvitationAppliedCategory2019} bundle maps $\varphi: \dtotal\rightarrow \dtotal$.
+When the sheaf is restricted to the stalk\cite{StalkSheaf2019} $\mathscr{F}_{k}$ over a point $\dbasepoint\in \dbase$, then allowable morphisms are $\varphi \in Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$.
   \end{itemize}
@@
 
- The artist takes as input the sheaf object $\mathcal{E}\coloneqq\sheaf_{\mathcal{E}}$ because the sheaf is a way of keeping track of how data is glued together \cite{ghristElementaryAppliedTopology2014}.
-This bookkeeping is done via the inclusion maps that are preserved by the presheaf, as illustrated in \autoref{eq:artist:presheaf}.
-Sheafs are presheafs where sections are defined over unions of open sets $U \cup_{U\in \mathscr{U}}$ \cite{spanier1989algebraic, nlab:sheaf}.
-As illustrated in \autoref{eq:artist:presheaf}, the functor $\sheaf_{\mathcal{\dtotal}}$ is a contravariant functor from a category to a category of sets \textbf{Set} \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}, which means it is by definition a presheaf.
+ The artist takes as input the sheaf object $\mathcal{E}\coloneqq\sheaf_{\mathcal{E}}$ because the sheaf is a way of keeping track of how data over topological spaces is glued together \cite{ghristElementaryAppliedTopology2014}.
+This bookkeeping is  illustrated in \autoref{eq:artist:presheaf} as the preservation of inclusion maps $\iota, \iota^{*}$ in presheaf maps between open sets $U_i$ and their corresponding set of sections $\Gamma(U_i, \dtotal)$.
+A presheaf is a contravariant functor from an arbitrary category $\mathcal{C}$ to \textbf{Set} \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf} and sheafs are presheafs where sets of sections $\Gamma$ are defined over unions of open sets $U \cup_{U\in \mathscr{U}}$ \cite{spanier1989algebraic, nlab:sheaf}.
 
@@
 
-Arbitrary morphisms on sheafs $f:\mathcal{\dtotal} \rightarrow \mathcal{\dtotal}$ are natural transforms \cite{SheafMathematics2021,bradleyWhatNaturalTransformation}, meaning that the morphisms $f$ compose with the inclusion maps $\iota$ \cite{fongInvitationAppliedCategory2019}.
+Arbitrary morphisms on sheafs $\varphi:\mathcal{\dtotal} \rightarrow \mathcal{\dtotal}$ are natural transforms \cite{SheafMathematics2021,bradleyWhatNaturalTransformation}, meaning that the morphisms $\varphi$ compose with the inclusion maps $\iota$ \cite{fongInvitationAppliedCategory2019}.
 As defined in \autoref{def:category:Tau}, the sets of sections $\Gamma(U_{i},\dtotal)$ are objects of the category $\mathcal{T}$.
@@
   \begin{tikzcd}
-    & \mathcal{T} & \\
-{} \arrow[rr, "f", Rightarrow, shift right, shorten=3em] & & {} \\
+    & \Gamma(\mathcal{K}, \dfiber) & \\
+{} \arrow[rr, "\varphi", Rightarrow, shift right, shorten=3.5em] & & {} \\
     & \mathcal{\dbase}^{op} \arrow[uu, "\sheaf_{\mathcal{\dtotal}}", bend left] \arrow[uu, "\sheaf_{\mathcal{\dtotal}}"', bend right] &   
@@
 Our criteria for equivariance is that bundle maps $f$ that are in the set of fiber to fiber maps $Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ are preserved.
-The maps $f\in Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ include the types of data transformations described in \autoref{sec:related-work:equivariance}.
+The maps $\varphi \in Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ include the types of data transformations described in \autoref{sec:related-work:equivariance}.
 When a fiber bundle is trivial, meaning $\dtotal=\dbase \times \dfiber$, then we can dispense with localization and directly consider the fiber maps $Hom(\dfiber_{k}, \dfiber_{k})$ over a point \dbasepoint.
@@
 
-The graphic category $\mathcal{\gtotal}$ consists of the functor object $\sheaf_{\mathcal{\gtotal}}: \mathcal{S}^{op} \rightarrow \Gamma(\mathcal{S}, \gtotal)$ and the sheaf morphisms $g:\sheaf_{\mathcal{\gtotal}} \rightarrow \sheaf_{\mathcal{gtotal}}$.
-The category $\mathcal{S}$ is the category of open sets over the graphic base space \gbase, and the category $\mathcal{P}$ is the set of sections $\gsection$ over $\mathcal{\gbase}$.
-The category of possible graphics $\mathcal{P}$ has as objects all graphics with a shared continuity and display space; bar plots and scatter plots on a 2D screen are two examples of objects in $\mathcal{P}$.
-As will be discussed in \autoref{sec:artist:construction}, the  $\vartist$ is constructed in such a way that it generates a single type of graphic.
-Therefore, we define the object as a morphism from data to the sheaf mage of the artist in the graphic space $\vartist: \mathcal{\dtotal} \rightarrow Im_A \in $
+The graphic category $\mathcal{\gtotal}$ consists of the functor object $\sheaf_{\mathcal{\gtotal}}: \mathcal{S}^{op} \rightarrow \Gamma(\mathcal{S}, \gtotal)$ and morphisms between sheaves.
+The category $\mathcal{S}$ is the category of open sets $\{U\}$ over the graphic base space \gbase.
+The set of sets of sections $\Gamma(\mathcal{S}, \gtotal)$ is a subcatagory of \texbf{Set}.
+A set $\Gamma(U_i, \gtotal)$ contains functions to generate arbitrary graphics with a shared continuity and display space; bar plots and scatter plots on a 2D screen are two such objects.
+
+As will be discussed in \autoref{sec:artist:construction}, the  $\vartist$ is constructed in such a way that it generates a single type of graphic;
+
+Therefore, we define the object as a morphism from data to the sheaf mage of the artist in the graphic space $\vartist: \mathcal{\dtotal} \rightarrow Im_A \subset \mathcal{H}$
 

### 2021-10-20 paper/main.tex
3.1

@@
 The artist \vartist\ converts data, which we model as the catgory $\mathcal{\dtotal}$ to graphics, which we model as the category $\mathcal{\gtotal}$.
+We formulate this association as
 
@@
 \end{equation}
-
-\note{roadmap the rest of this section}
+In this section we describe the structure of the data and graphic spaces, which we formulate as objects in categories and the morphisms between those objects.
+We then use these definitions to express the structure that an artist must preserve.
+Finally, we propose that artists can be composed by manufacturing new artists based on various types of input data.
 
@@
 The set of sets of sections $\Gamma(\mathcal{S}, \gtotal)$ is a subcatagory of \texbf{Set}.
-A set $\Gamma(U_i, \gtotal)$ contains functions to generate arbitrary graphics with a shared continuity and display space; bar plots and scatter plots on a 2D screen are two such objects.
+A set $\Gamma(U_i, \gtotal)$ contains functions to generate arbitrary graphics with a shared continuity and display space, while the artist $\vartist$ is constructed such that it generates a single type of graphic, as described in \autoref{sec:artist:construction}.
+Therefore, the range of the the artist functor, which we denote $Im_{\vartist}$, is a sheaf map $Im_{\vartist}$.
+The sheaf $Im_{\vartist}$ is a map between the category of open sets $\gtotal^{op}$ and a subcategory of \textbf{Set}.
+In this subcatagory, each set of sections in the sheaf $Im_{\vartist}$ over an arbitrary open set $U_{i}$ is a subset of the corresponding full set of sections $\Gamma(U_i, \gtotal)$.
 
-As will be discussed in \autoref{sec:artist:construction}, the  $\vartist$ is constructed in such a way that it generates a single type of graphic;
+Since the artist is a morphism of sheafs
+\begin{equation}
+  \vartist: \mathcal{\dtotal} \rightarrow Im_A
+\end{equation}
+it is a natural transformation\cite{SheafMathematics2021,bradleyWhatNaturalTransformation}.
+This means that, by definition, all morphisms $\varphi$ on $\mathcal{\dtotal}$ are expected to commute \cite{spanier1989algebraic,fongInvitationAppliedCategory2019}.
+This is the basis of the equivariance illustrated in \autoref{eq:artist:equivariance}
 
-Therefore, we define the object as a morphism from data to the sheaf mage of the artist in the graphic space $\vartist: \mathcal{\dtotal} \rightarrow Im_A \subset \mathcal{H}$
-
-\begin{equation}
-  \label{eq:artist:morphism-of-sheaf}
-  \begin{tikzcd}
-    & \mathcal{T} & & & \mathcal{P} & \\
-{} \arrow[rr, "f", Rightarrow, shift right, shorten=3em] & 
-& 
-{} \arrow[r, "\vartist", shorten=-1.25em, bold] & {} \arrow[rr, "g", Rightarrow, shift right, shorten=3em] & & {} \\
-& \mathcal{\dbase}^{op} \arrow[uu, "\mathcal{\dtotal}", bend left] 
-  \arrow[uu, "\mathcal{\dtotal}"', bend right] 
-  & & 
-& \mathcal{\gbase}^{op} \arrow[uu, "\mathcal{\gtotal}", bend left] 
-  \arrow[uu, "\mathcal{\gtotal}"', bend right] &   
-  \end{tikzcd}
-\end{equation}
-
-% f is a natural transform, all transforms on the sheafs commute
-% g is a natural transform ...
-% space of g->g is too big, so we look at Im...
-% b/c it's a natural transform, - f carries i over b/c it's a natural transformation
-
-therefore the artist is a map between functors
-\begin{equation}
-  \tag{\ref{eq:artist:artist}}
-  \vartist:\mathcal{\dtotal}\rightarrow{\mathcal{\gtotal}} 
-\end{equation} 
-which means that the artist is a natural transformation\cite{SheafMathematics2021,bradleyWhatNaturalTransformation}.
-
-% such that (9)
-
-Extra property that implies equivariance is preserved is illustrated in the diagram which must commute
-%% note Im_A in \H, A_M (lifting/pushforward)
 \begin{equation}
@@
   \begin{tikzcd}
-    \mathcal{\dtotal} \arrow[r, "\vartist"] \arrow[d, "{Hom_{\mathcal{\dfiber}}(\dfiber, \dfiber)}"'] & Im_{\vartist} \in \mathcal{\gtotal} \arrow[d, "\vartist_{Hom_{\mathcal{\dfiber}}}"] \\
-    \mathcal{\dtotal} \arrow[r, "\vartist"]                                                           & Im_{\vartist} \in \mathcal{\gtotal}                                                
+    \mathcal{\dtotal} \arrow[r, "\vartist"] \arrow[d, "\varphi \in {Hom_{\mathcal{\dfiber}}(\dfiber, \dfiber)}"'] & Im_{\vartist} \arrow[d, "\vartist_{\varphi \in Hom_{\mathcal{\dfiber}}}"] \\
+    \mathcal{\dtotal} \arrow[r, "\vartist"]                                                           & Im_{\vartist}                                                 
   \end{tikzcd}
 \end{equation}
-
-such that the morphisms on the graphic category $\mathcal{H}$ are homomorphic to the morphisms $\monoid$ on the fibers $\mathcal{F}$ \cite{spanier1989algebraic,fongInvitationAppliedCategory2019}. 
-%% such that monoid on this side is monoid on that side
-
+which states that transforms on the data side $\varphi \in Hom_{\mathcal{\dfiber}}(\dfiber, \dfiber)$ have an approximately equivalent transform in the data side $\vartist_{\varphi \in Hom_{\mathcal{\dfiber}}}$.
+This commutativity is described as the matching of data and visual structure by Kindlmann and Scheidegger when introducing their analougous correspondence principal.
 In \autoref{sec:artist:construction}, we show a method of constructing the artist such that the constraints of \autoref{eq:artist:morphism-of-sheaf} and \autoref{eq:artist:equivariance} can be satisfied, thereby ensuring equivariance and the preservation of continuity.
-We then use this formulation to guide the development of visualization
-library components in \autoref{sec:case-study}.
+We then use this formulation to guide the development of visualization library components in \autoref{sec:case-study}.
 

### 2021-10-21 paper/main.tex
composition outline + notes

@@
 
-\subsection{Artist Composition}
+\subsection{Multiple Graphics: Superposition and Composition}
 \label{sec:artist:union}
@@
 \note{formalize superposition, than break down what that means for composition}
-\subsubsection{Superposition: union}
+\subsubsection{Superposition: Multiple Datasets}
+\note{far as we care, this can technically be A = A1 + A2....}
 
-These graphics can be generated by disjoint data inputs  $(\dtotal_i: i\in I)$ into different artist functions $\vartist$ with the same output target space $\gtotal$.
-We define this composition via the disjoint union of the inputs
-\begin{equation}
-  \label{eq:artist:union:glue}
-+ \coloneqq \underset{i\in I}{\sqcup} \dtotal_{i}
-\end{equation}
-This approach does not guarantee a coherent image, as illustrated in \autoref{fig:artist:union:glue}, because each data is explicitly considered as distinct from the others.
+\subsubsection{Composite Artist: Multivariate Datasets}
+\note{This is A = A_1 \circ A_2 b\c it's sharing nus or xis or all of the above}
 
-\subsection{Shared properties: intersection}
-Many times though, graphics are generated from subsets of a larger data set, such as when using multiple views \cite{alboRadarComparativeEvaluation2016a, hullmanKeeping2018} and brush-linked views \cite{beckerBrushingScatterplots1987,bujaInteractiveData1991}.
-In these situations, we consider each of the data inputs as component bundles over a shared continuity \dbase.
+\paragraph{Overlapping Base Spaces}
 
-\begin{equation}
-  \label{eq:math:data:base:decompose}
-  \pi:\dtotal_1\oplus\ldots\oplus \dtotal_i \oplus\ldots \oplus \dtotal_n \rightarrow \dbase
-\end{equation}
-
-The sections $\dsection$ on each component bundle $\dtotal_i$ over a point $\dbasepoint \in \dbase$ are the elements of the data associated with that point.
-We use this property to establish which of these elements are the same, because they come from the same fiber, and therefore should share transformation functions to visual elements.
-This bookkeeping yields graphics with consistent visual encoding of data elements, as shown in \autoref{fig:artist:union:continuity}.
+\paragraph{Shared Fibers}
 

### 2021-10-22 paper/main.tex
composition

@@
 \subsection{Categorical Artist}
+\label{sec:artist:category}
 We define the categories $\mathcal{\dtotal}$ and $\mathcal{\gtotal}$ as sheaves of sections\cite{SheafMathematics2021,spanier1989algebraic}, of fiberbundles, as introduced in \autoref{sec:related-work:fiber-bundles}.
@@
 This is what allows the artist to generate graphics where the subset of data on view is dynamically updated, such as pan and zoom navigation\cite{NekrasovskiEvaluationPanZoom2006} and sliding windows on streaming data\cite{crouchDynamicGraphsSlidingwindow2013,chuTimeSeriesSegmentation1995}.
+\note{does this actually go down in the composition section?}
 
@@
 \begin{equation}
-  \vartist: \mathcal{\dtotal} \rightarrow Im_A
+  \vartist: \mathcal{\dtotal} \rightarrow Im_{\vartist}
 \end{equation}
-it is a natural transformation\cite{SheafMathematics2021,bradleyWhatNaturalTransformation}.
-This means that, by definition, all morphisms $\varphi$ on $\mathcal{\dtotal}$ are expected to commute \cite{spanier1989algebraic,fongInvitationAppliedCategory2019}.
-This is the basis of the equivariance illustrated in \autoref{eq:artist:equivariance}
+it is by definition a natural transformation\cite{SheafMathematics2021,bradleyWhatNaturalTransformation}.
+This means that all morphisms $\varphi$ on $\mathcal{\dtotal}$ are expected to commute \cite{spanier1989algebraic,fongInvitationAppliedCategory2019}.
+This is the basis of the equivariance condition illustrated in \autoref{eq:artist:equivariance}
 
@@
 which states that transforms on the data side $\varphi \in Hom_{\mathcal{\dfiber}}(\dfiber, \dfiber)$ have an approximately equivalent transform in the data side $\vartist_{\varphi \in Hom_{\mathcal{\dfiber}}}$.
-This commutativity is described as the matching of data and visual structure by Kindlmann and Scheidegger when introducing their analougous correspondence principal.
+This commutativity property states that visualization library components must be implemented such that changes in data induce equivalent changes in the output, which is how visualization libraries can produce visualizations that satisfy Kindlmann and Scheidegger's correspondence principal that data and visual structure must match.
+
 In \autoref{sec:artist:construction}, we show a method of constructing the artist such that the constraints of \autoref{eq:artist:morphism-of-sheaf} and \autoref{eq:artist:equivariance} can be satisfied, thereby ensuring equivariance and the preservation of continuity.
@@
 
-\subsection{Multiple Graphics: Superposition and Composition}
+\subsection{Composition of Artists}
 \label{sec:artist:union}
@@
 \label{fig:artist:union:continuity}}
-\caption{In \autoref{fig:artist:union:glue}, the outputs of various artist are independently rendered to the same graphic space \gtotal.
+\caption{In \autoref{fig:artist:union:glue}, the outputs of various artist are independently rendered to the same graphic space.
 The inputs to the artists in \autoref{fig:artist:union:continuity} are sections over the same fiber bundle base space \dbase; therefore, they should share transformation functions to produce a consistent image.
@@
 \end{figure}
-%% artist composition is fiber products/sums of artists on sheafs -> closed cartesian category maybe?
 
-Visualizations generally consist of more than one graphic, for example two line plots, a legend, axis ticks, spines, and labels.
-\note{formalize superposition, than break down what that means for composition}
-\subsubsection{Superposition: Multiple Datasets}
-\note{far as we care, this can technically be A = A1 + A2....}
+\note{is about building artist that preserves constraint A1 out of A2 + A3}
+% generalizing sheaf to collection of sheafs w/ inclusion maps so can tell you how to assemble A1 + A2 such that that push out is preserved 
+% when we build an artist out of nus and qs, this is how to preserve this 
+% if the artist is a black box, what are the pre and post conditionsg
+Visualizations generally consist of more than one visual element, for example line plots, a legend, axis ticks, spines, and labels.
+Many times, visualizations are generated from subsets of a larger data set, such as when using multiple views \cite{alboRadarComparativeEvaluation2016a} and brush-linked views \cite{beckerBrushingScatterplots1987,bujaInteractiveData1991}, and there is an expectation of consistency in how this data is visualized\cite{hullmanKeeping2018}.
+In \autoref{fig:artist:union:glue}, the disjointedness of the graphic lends itself to an interpretation that the data is disjoint, whereas the consistency of the visual encodings in \autoref{fig:artist:union:continuity} indicates that visual elements share data attributes.
+In this section, we describe how to express the structure of the data input in terms of  whether the sheaf objects $\sheaf_{\mathcal{\dtotal}}$ are disjoint or share attributes.
+We then propose that a topologically equivariant artist will preserve this structure in its output.
 
-\subsubsection{Composite Artist: Multivariate Datasets}
-\note{This is A = A_1 \circ A_2 b\c it's sharing nus or xis or all of the above}
+\subsection{Union of Base Spaces}
+\begin{figure}[h!]
+  \includegraphics[width=\columnwidth]{union_k.png}
+  \caption{}
+  \label{fig:artist:compose:union_base}
+\end{figure}
 
-\paragraph{Overlapping Base Spaces}
+\begin{figure}[h!]
+  \includegraphics[width=\columnwidth]{disjoint_k.png}
+  \caption{}
+  \label{fig:artist:compose:disjoint_base}
+\end{figure}
 
-\paragraph{Shared Fibers}
+\paragraph{Shared fibers}
+
+\begin{figure}[h!]
+  \includegraphics[width=\columnwidth]{intersection_f.png}
+  \caption{}
+  \label{fig:artist:compose:union_fiber}
+\end{figure}
 

### 2021-10-25 paper/main.tex
inclusion of base spaces
clean up object E notation

@@
   \begin{itemize}
-    \item\textit{object} the sheaf\cite{SheafMathematics2021,ghristElementaryAppliedTopology2014} $\sheaf_{\mathcal{\dtotal}}: \mathcal{K}^{op} \rightarrow \Gamma(\mathcal{K}, \dtotal)$
+    \item\textit{object} the sheaf\cite{SheafMathematics2021,ghristElementaryAppliedTopology2014} $\sheaf_{\dtotal}: \mathcal{K}^{op} \rightarrow \Gamma(\mathcal{K}, \dtotal)$
     \item \textit{morphisms} arbitrary endomorphic\cite{fongInvitationAppliedCategory2019} bundle maps $\varphi: \dtotal\rightarrow \dtotal$.
@@
 
- The artist takes as input the sheaf object $\mathcal{E}\coloneqq\sheaf_{\mathcal{E}}$ because the sheaf is a way of keeping track of how data over topological spaces is glued together \cite{ghristElementaryAppliedTopology2014}.
+ The artist takes as input the sheaf object $\sheaf_{\dtotal}$ because the sheaf is a way of keeping track of how data over topological spaces is glued together \cite{ghristElementaryAppliedTopology2014}.
 This bookkeeping is  illustrated in \autoref{eq:artist:presheaf} as the preservation of inclusion maps $\iota, \iota^{*}$ in presheaf maps between open sets $U_i$ and their corresponding set of sections $\Gamma(U_i, \dtotal)$.
@@
     U_1 \arrow[dd, "\iota"'] & {\Gamma(U_1, \dtotal)}                              \\
-    {} \arrow[r, "\mathcal{\dtotal}\coloneqq\sheaf_{\mathcal{\dtotal}}"]   & {}                                                  \\
+    {} \arrow[r, "\sheaf_{\dtotal}"]   & {}                                                  \\
     U_2                      & {\Gamma(U_2, \dtotal)} \arrow[uu, "\iota^*"']
@@
 
- The functor $\sheaf_{\mathcal{\dtotal}}$ is contravariant because the inclusion map $\iota$ between open sets $U_i \in \mathcal{K}$ goes in the opposite direction from the map $\iota^*$ between the sections in $\Gamma(U_{i}, \dtotal)$ \cite{riehlCategoryTheoryContext}.
+ The functor $\sheaf_{\dtotal}$ is contravariant because the inclusion map $\iota$ between open sets $U_i \in \mathcal{K}$ goes in the opposite direction from the map $\iota^*$ between the sections in $\Gamma(U_{i}, \dtotal)$ \cite{riehlCategoryTheoryContext}.
 The open set $U_1$ is a subset of $U_2$, therefore $\iota: U_1 \rightarrow U_2$.
@@
 {} \arrow[rr, "\varphi", Rightarrow, shift right, shorten=3.5em] & & {} \\
-    & \mathcal{\dbase}^{op} \arrow[uu, "\sheaf_{\mathcal{\dtotal}}", bend left] \arrow[uu, "\sheaf_{\mathcal{\dtotal}}"', bend right] &   
+    & \mathcal{\dbase}^{op} \arrow[uu, "\sheaf_{\dtotal}", bend left] \arrow[uu, "\sheaf_{\dtotal}"', bend right] &   
 \end{tikzcd}
@@
 
-The sheaf $\sheaf_{\mathcal{\dtotal}}$ over a limit of open sets that contain a point $\dbasepoint \in \dbase$ is approximately the same as a sheaf over a point \dbasepoint\ and is called the stalk $\sheaf_{\mathcal{\dtotal}}|_{k}\coloneqq \lim\limits_{U\ni k} \Gamma(U, \dtotal)$ \cite{StalkSheaf2019}.
+The sheaf $\sheaf_{\dtotal}$ over a limit of open sets that contain a point $\dbasepoint \in \dbase$ is approximately the same as a sheaf over a point \dbasepoint\ and is called the stalk $\sheaf_{\mathcal{\dtotal}}|_{k}\coloneqq \lim\limits_{U\ni k} \Gamma(U, \dtotal)$ \cite{StalkSheaf2019}.
 The fiber space over a point is embedded in the stalk over that point $\dfiber_{k}\hookrightarrow \mathscr{F}_{k}$.
@@
 
-The graphic category $\mathcal{\gtotal}$ consists of the functor object $\sheaf_{\mathcal{\gtotal}}: \mathcal{S}^{op} \rightarrow \Gamma(\mathcal{S}, \gtotal)$ and morphisms between sheaves.
+The graphic category $\mathcal{\gtotal}$ consists of the functor object $\sheaf_{\gtotal}: \mathcal{S}^{op} \rightarrow \Gamma(\mathcal{S}, \gtotal)$ and morphisms between sheaves.
 The category $\mathcal{S}$ is the category of open sets $\{U\}$ over the graphic base space \gbase.
@@
 \label{sec:artist:union}
-\begin{figure}[!h]
-\centering
-\subfloat[+]{\includegraphics[width=.4\columnwidth]{exploding_artist.png}%
-\label{fig:artist:union:glue}}
-%\hfil
-\subfloat[$\oplus$]{\includegraphics[width=.4\columnwidth]{combined_artist.png}%
-\label{fig:artist:union:continuity}}
-\caption{In \autoref{fig:artist:union:glue}, the outputs of various artist are independently rendered to the same graphic space.
-The inputs to the artists in \autoref{fig:artist:union:continuity} are sections over the same fiber bundle base space \dbase; therefore, they should share transformation functions to produce a consistent image.
-\note{add a border/make the same size/etc so 1 doesn't run over into two}}
-\end{figure}
+Visualizations generally consist of more than one visual element, for example line plots, a legend, axis ticks, spines, and labels.
+Many times, visualizations are generated from subsets of a larger data set, such as when using multiple views \cite{alboRadarComparativeEvaluation2016a} and brush-linked views \cite{beckerBrushingScatterplots1987,bujaInteractiveData1991}, and there is an expectation of consistency in how this data is visualized\cite{hullmanKeeping2018}.
+In this section, we generalize the sheaf input object $\sheaf_{\dtotal}$ to a collection of sheaves $E^\prime$ with inclusion maps that dictate the how the sheaves are glued together.
+For example, given$\mathcal{E}_1, \mathcal{E}_2 \in \mathcal{E^{prime}}
 
-\note{is about building artist that preserves constraint A1 out of A2 + A3}
+ (\note{need citation/explainationn of why I can just do this})
+
 % generalizing sheaf to collection of sheafs w/ inclusion maps so can tell you how to assemble A1 + A2 such that that push out is preserved 
@@
 % if the artist is a black box, what are the pre and post conditionsg
-Visualizations generally consist of more than one visual element, for example line plots, a legend, axis ticks, spines, and labels.
-Many times, visualizations are generated from subsets of a larger data set, such as when using multiple views \cite{alboRadarComparativeEvaluation2016a} and brush-linked views \cite{beckerBrushingScatterplots1987,bujaInteractiveData1991}, and there is an expectation of consistency in how this data is visualized\cite{hullmanKeeping2018}.
-In \autoref{fig:artist:union:glue}, the disjointedness of the graphic lends itself to an interpretation that the data is disjoint, whereas the consistency of the visual encodings in \autoref{fig:artist:union:continuity} indicates that visual elements share data attributes.
-In this section, we describe how to express the structure of the data input in terms of  whether the sheaf objects $\sheaf_{\mathcal{\dtotal}}$ are disjoint or share attributes.
-We then propose that a topologically equivariant artist will preserve this structure in its output.
 
 \subsection{Union of Base Spaces}
-\begin{figure}[h!]
-  \includegraphics[width=\columnwidth]{union_k.png}
-  \caption{}
-  \label{fig:artist:compose:union_base}
+
+\begin{figure}[!h]
+  \centering
+  \subfloat[$K_{k}\neq \varnothing$]{\includegraphics[width=.4\columnwidth]{union_k.png}%
+  \label{fig:artist:union:k}}
+  %\hfil
+  \subfloat[$K_{k}=\varnothing$]{\includegraphics[width=.4\columnwidth]{disjoint_k.png}%
+  \label{fig:artist:union:disjoint_k}}
+  \caption{}}
   \end{figure}
 
-\begin{figure}[h!]
-  \includegraphics[width=\columnwidth]{disjoint_k.png}
-  \caption{}
-  \label{fig:artist:compose:disjoint_base}
-\end{figure}
+\begin{tikzcd}
+  \dbase_{k} \arrow[d, "\iota", hook] \arrow[r, "\iota", hook] & \dbase_{j} \arrow[d, "i_{\dbase_j}"]      \\
+  \dbase_{i} \arrow[r, "i_{\dbase_i}"]                         & \dbase_{i} \bigsqcup_{\dbase_{k}}\dbase_{j}
+  \end{tikzcd}
+where the projection functions $i_{\dbase_i},i_{\dbase_j}$ behave such that $i_{\dbase_{i}}(\dbasepoint)|_{\dbasepoint \in \dbase_i} = \[\dbasepoint\]$,  $i_{\dbase_{j}}(\dbasepoint)|_{\dbasepoint \in \dbase_j} = \[\dbasepoint\]$, meaning the projection functions yield equivalent points in the disjoint union $\dbase_i \bigsqcup \dbase_j$.
+The special case of
 

### 2021-10-25 paper/main.tex
fiber caption

@@
   \begin{tikzcd}
-    \mathcal{\dtotal} \arrow[r, "\vartist"] \arrow[d, "\varphi \in {Hom_{\mathcal{\dfiber}}(\dfiber, \dfiber)}"'] & Im_{\vartist} \arrow[d, "\vartist_{\varphi \in Hom_{\mathcal{\dfiber}}}"] \\
+    \mathcal{\dtotal} \arrow[r, "\vartist"] \arrow[d, "\varphi \in {Hom_{\mathcal{\dfiber}}(\dfiber, \dfiber)}"'] & Im_{\vartist} \arrow[d, "\vartist_{\varphi}"] \\
     \mathcal{\dtotal} \arrow[r, "\vartist"] & Im_{\vartist}                                                 
@@
 Visualizations generally consist of more than one visual element, for example line plots, a legend, axis ticks, spines, and labels.
-Many times, visualizations are generated from subsets of a larger data set, such as when using multiple views \cite{alboRadarComparativeEvaluation2016a} and brush-linked views \cite{beckerBrushingScatterplots1987,bujaInteractiveData1991}, and there is an expectation of consistency in how this data is visualized\cite{hullmanKeeping2018}.
-In this section, we generalize the sheaf input object $\sheaf_{\dtotal}$ to a collection of sheaves $E^\prime$ with inclusion maps that dictate the how the sheaves are glued together.
-For example, given$\mathcal{E}_1, \mathcal{E}_2 \in \mathcal{E^{prime}}
+We propose that we can express expected consistency between these elements as a preservation of morphisms between objects of different data categories $\mathcal{\dtotal}$.
 
- (\note{need citation/explainationn of why I can just do this})
-
-% generalizing sheaf to collection of sheafs w/ inclusion maps so can tell you how to assemble A1 + A2 such that that push out is preserved 
-% when we build an artist out of nus and qs, this is how to preserve this 
-% if the artist is a black box, what are the pre and post conditionsg
-
-\subsection{Union of Base Spaces}
-
+\subsection{Shared Index}
 \begin{figure}[!h]
   \centering
-  \subfloat[$K_{k}\neq \varnothing$]{\includegraphics[width=.4\columnwidth]{union_k.png}%
+  \subfloat[$\dbase_c = \dbase_a \cap \dbase_b$]{\includegraphics[width=.4\columnwidth]{union_k.png}%
   \label{fig:artist:union:k}}
-  %\hfil
-  \subfloat[$K_{k}=\varnothing$]{\includegraphics[width=.4\columnwidth]{disjoint_k.png}%
+  %\hfill
+  \subfloat[$\dbase_c = \varnothing$]{\includegraphics[width=.4\columnwidth]{disjoint_k.png}%
   \label{fig:artist:union:disjoint_k}}
-  \caption{}}
+  \caption{In \autoref{fig:artist:union:k}, the input object $\sheaf_{\dtotal}$ encodes a continuous function over spaces $\dtotal_a$, $\dtotal_b$ 
+  and overlapping space $\dtotal_c$.
+In this figure, the visual transformation is to a point on the line on screen.
+An artist that preserves the continuity of the function must generate graphics such that functions evaluted  $\dbasepoint \in \dbase_a \cap \dbase_b$
 \end{figure}
 
+An example of overlapping indexing is a parallelized verion of a sliding window algorithm where window overlaps must resolve to the same value at the same position \cite{chuTimeSeriesSegmentation1995};
+
+\begin{equation}
+\label{eq:artist:union:k}
 \begin{tikzcd}
-  \dbase_{k} \arrow[d, "\iota", hook] \arrow[r, "\iota", hook] & \dbase_{j} \arrow[d, "i_{\dbase_j}"]      \\
-  \dbase_{i} \arrow[r, "i_{\dbase_i}"]                         & \dbase_{i} \bigsqcup_{\dbase_{k}}\dbase_{j}
+  \dbase_{c} \arrow[d, "\iota", hook] \arrow[r, "\iota", hook] & \dbase_{b} \arrow[d, "i_{\dbase_b}"]      \\
+  \dbase_{a} \arrow[r, "i_{\dbase_a}"]                         & \dbase_{a} \bigsqcup_{\dbase_{c}}\dbase_{b}
   \end{tikzcd}
-where the projection functions $i_{\dbase_i},i_{\dbase_j}$ behave such that $i_{\dbase_{i}}(\dbasepoint)|_{\dbasepoint \in \dbase_i} = \[\dbasepoint\]$,  $i_{\dbase_{j}}(\dbasepoint)|_{\dbasepoint \in \dbase_j} = \[\dbasepoint\]$, meaning the projection functions yield equivalent points in the disjoint union $\dbase_i \bigsqcup \dbase_j$.
-The special case of
+\end{equation}
+where the projection functions $i_{\dbase_a},i_{\dbase_b}$ behave such that $i_{\dbase_a}(\dbasepoint)|_{\dbasepoint \in \dbase_b} = \left[\dbasepoint\right]$, $i_{\dbase_b}(\dbasepoint)|_{\dbasepoint \in \dbase_b} = \left[\dbasepoint\right]$, meaning the projection functions yield equivalent points in the disjoint union $\dbase_a \bigsqcup_{\dbase_c} \dbase_b$.
+Data that is completely disjoint in the index, as shown in \autoref{fig:artist:union:disjoint_k}, is a special case where $\dbase_c=\varnothing$ and therefore there is no constraint on the artist with regards to how this data is rendered.
 
-\paragraph{Shared fibers}
+\paragraph{Shared Fields}
 
@@
   \includegraphics[width=\columnwidth]{intersection_f.png}
-  \caption{}
+  \caption{The fiber spaces $F_a, F_b$ have a shared fiber space $F_c$.
+If they are input into the same artist $A_1$, then an equivariant transformation is one where $F_c$ is transformed to a visual element in a consistent matter.
+In this figure, $F_c$ is mapped to the x-axis for both $F_a$ and $F_b$.}
   \label{fig:artist:compose:union_fiber}
 \end{figure}
+an example of expecting consistency in overlapping fields are the multiple view constraints described by Qu and Hullman \cite{hullmanKeeping2018}.
 

### 2021-10-27 paper/main.tex
k caption (borked) but fixes latex
shared nu

@@
 
-\IEEEPARstart{T}his paper uses methods from topology and category theory to develop a model of the transformation from data to graphical representation.
+\IEEEPARstart{T}his paper uses methods from algebraic topology and category theory to develop a model of the transformation from data to graphical representation.
 This model provides a language to specify how data is structured and how this structure is carried through in the visualization, and serves as the basis for a functional approach to implementing visualization library components.
@@
 In this figure, the visual transformation is to a point on the line on screen.
-An artist that preserves the continuity of the function must generate graphics such that functions evaluted  $\dbasepoint \in \dbase_a \cap \dbase_b$
+An artist that preserves the continuity of the function must generate graphics such that functions evaluted  $\dbasepoint \in \dbase_a \cap \dbase_b$}
 \end{figure}
@@
 \begin{equation}
-\label{eq:artist:union:k}
+\label{eq:artist:intersection:basespace}
 \begin{tikzcd}
-  \dbase_{c} \arrow[d, "\iota", hook] \arrow[r, "\iota", hook] & \dbase_{b} \arrow[d, "i_{\dbase_b}"]      \\
-  \dbase_{a} \arrow[r, "i_{\dbase_a}"]                         & \dbase_{a} \bigsqcup_{\dbase_{c}}\dbase_{b}
+  \dfiber_a \arrow[rrrr, "\vartist_a"] \arrow[rrd, "proj_a" description] &  &             &  & {\Gamma(\gbase_a, \gfiber_a)} \arrow[lld, "ga" description, dashed]  \\
+                                                                         &  & \dfiber_{c} &  &                                                                      \\
+  \dfiber_b \arrow[rrrr, "\vartist_b"] \arrow[rru, "proj_b" description] &  &             &  & {\Gamma(\gbase_b, \gfiber_b)} \arrow[llu, "g_b" description, dashed]
   \end{tikzcd}
@@
 an example of expecting consistency in overlapping fields are the multiple view constraints described by Qu and Hullman \cite{hullmanKeeping2018}. 
+\begin{equation}
+  \label{eq:artist:intersection:fiber}
+  \begin{tikzcd}
+    \dfiber_a \arrow[rrd, "proj_a" description] \arrow[rrrr, "\vartist_a"] &  &             &  & {\Gamma(\gbase_a, \gfiber_a)} \arrow[lld, "g_a" description, dashed] \\
+                                                                           &  & \dfiber_{c} &  &                                                                      \\
+    \dfiber_b \arrow[rru, "proj_b" description] \arrow[rrrr, "\vartist_b"] &  &             &  & {\Gamma(\gbase_a, \gfiber_a)} \arrow[llu, "g_b" description, dashed]
+\end{tikzcd}
+\end{equation}
+
+% introduce notion of equivariant fiber, chasing diagram of F is different Gamma - Kosara and Zeminiskwy differentiable glyph 
+% case 1: F3 \subset F1, {JFK} \in {JFK, EWR}
+% case 2: F1 = F3 \times F1, F2= F3 \times F2, 
+  % F1 ->F3 is projection -> Fiber consistency
+% in that case, we can form the plus operator, disjoint union of A1 + A2 going in is consistent on subspace
+
+We can compose these diagrams together to generate an \vartist\ to handle any of the things.
 
@@
 
-We enforce the equivariance constraint
-
+We enforce the equivariance constraint specified in \autoref{eq:artist:construction:multi_nu} by constructing artists that apply \vchannel\ encoders such that 
 \begin{equation}
-  \label{eq:math:artist:nu_commute}
+  \label{eq:artist:construction:multi_nu}
   \begin{tikzcd}
-  \dtotal_i \arrow[r] \arrow[r, "\vchannel_i"] \arrow[d, "m_{\delement}"'] & \vtotal_i \arrow[d, "m_{\velement}"] \\
-  \dtotal_i \arrow[r, "\vchannel_i"]                           & \vtotal_i               
+    \dfiber_a \arrow[rrd, "proj_a" description] \arrow[rrrr, "\vchannel^{a}_{\dfiber_c}"] &  &                                                                                                                       &  & \vfiber^{a}_{\dfiber_c} \\
+                                                                                          &  & \dfiber_{c} \arrow[rru, "\vchannel^{a}_{\dfiber_c}" description] \arrow[rrd, "\vchannel^{b}_{\dfiber_c}" description] &  &                         \\
+    \dfiber_b \arrow[rru, "proj_b" description] \arrow[rrrr, "\vchannel^{a}_{\dfiber_c}"] &  &                                                                                                                       &  & \vfiber^{b}_{\dfiber_c}
     \end{tikzcd}
@@
 
-One advantage of expressing the data to visual encoding $\vchannel_i$ as a functor from data fiber $\dfiber_i$ to $\vfiber_i$ is that it lets us formally express that encodings are bound to data and therefore should be consistent across views, as proposed by Hullamn and Qu\cite{hullmanKeeping2018}.
+the same fiber in two different data sets run through the same $\vchannel_{i}$ encoder will output to the same set of visual values in the corresponding visual fiber $\vfiber_i$.
 

### 2021-10-27 paper/main.tex
restored deleted shared k diagram

@@
 
-\subsection{Shared Index}
+\subsubsection{Shared Index}
 \begin{figure}[!h]
@@
 \begin{tikzcd}
-  \dfiber_a \arrow[rrrr, "\vartist_a"] \arrow[rrd, "proj_a" description] &  &             &  & {\Gamma(\gbase_a, \gfiber_a)} \arrow[lld, "ga" description, dashed]  \\
-                                                                         &  & \dfiber_{c} &  &                                                                      \\
-  \dfiber_b \arrow[rrrr, "\vartist_b"] \arrow[rru, "proj_b" description] &  &             &  & {\Gamma(\gbase_b, \gfiber_b)} \arrow[llu, "g_b" description, dashed]
+  \dbase_{c} \arrow[d, "\iota", hook] \arrow[r, "\iota", hook] & \dbase_{b} \arrow[d, "i_{\dbase_b}"]      \\
+  \dbase_{a} \arrow[r, "i_{\dbase_a}"]                         & \dbase_{a} \bigsqcup_{\dbase_{c}}\dbase_{b}
   \end{tikzcd}
@@
 
-\paragraph{Shared Fields}
+\subsubsection{}{Shared Fields}
 

### 2021-10-28 paper/main.tex
imA

@@
 The set of sets of sections $\Gamma(\mathcal{S}, \gtotal)$ is a subcatagory of \texbf{Set}.
+
+\begin{definition} The sheaf $Im_{\vartist}$ is a map between the category of open sets $\gtotal^{op}$ and a subcategory of \textbf{Set}.
+This subcategory is the set of subsets of sections $\{\gsection^{k}:U_i \rightarrow \gfiber_{U_{i}}\} \subset \Gamma(U_i, \gtotal)$ over the space of all open sets in $\mathcal{\gbase}$.
+\end{definition}
+
 A set $\Gamma(U_i, \gtotal)$ contains functions to generate arbitrary graphics with a shared continuity and display space, while the artist $\vartist$ is constructed such that it generates a single type of graphic, as described in \autoref{sec:artist:construction}.
 Therefore, the range of the the artist functor, which we denote $Im_{\vartist}$, is a sheaf map $Im_{\vartist}$.
-The sheaf $Im_{\vartist}$ is a map between the category of open sets $\gtotal^{op}$ and a subcategory of \textbf{Set}.
-In this subcatagory, each set of sections in the sheaf $Im_{\vartist}$ over an arbitrary open set $U_{i}$ is a subset of the corresponding full set of sections $\Gamma(U_i, \gtotal)$.
-
 Since the artist is a morphism of sheafs
@@
                                                                            &  & \dfiber_{c} &  &                                                                      \\
-    \dfiber_b \arrow[rru, "proj_b" description] \arrow[rrrr, "\vartist_b"] &  &             &  & {\Gamma(\gbase_a, \gfiber_a)} \arrow[llu, "g_b" description, dashed]
+    \dfiber_b \arrow[rru, "proj_b" description] \arrow[rrrr, "\vartist_b"] &  &             &  & Im_{\vartist_b} \in \gtotal \arrow[llu, "g_b" description, dashed]
 \end{tikzcd}
@@
 \subsection{Graphic} 
-The object of the graphic category $\mathcal{\gtotal}$ is the fiberbundle \gtotal. The bundle \gtotal\ has the same structure as the data bundle \dtotal
+The object of the graphic category $\mathcal{\gtotal}$ is the fiber bundle \gtotal. The bundle \gtotal\ has the same structure as the data bundle \dtotal
 

### 2021-10-28 paper/main.tex
q equations

@@
   \label{eq:artist:intersection:fiber}
-  \begin{tikzcd}
-    \dfiber_a \arrow[rrd, "proj_a" description] \arrow[rrrr, "\vartist_a"] &  &             &  & {\Gamma(\gbase_a, \gfiber_a)} \arrow[lld, "g_a" description, dashed] \\
+  \begin{tikzcd}[row sep = .5em]
+    \dfiber_a \arrow[rrd, "proj_a" description] \arrow[rrrr, "\vartist_a"] &  &             &  & Im_{\vartist_a} \arrow[lld, "g_a" description, dashed] \\
                                                                            &  & \dfiber_{c} &  &                                                                      \\
-    \dfiber_b \arrow[rru, "proj_b" description] \arrow[rrrr, "\vartist_b"] &  &             &  & Im_{\vartist_b} \in \gtotal \arrow[llu, "g_b" description, dashed]
+    \dfiber_b \arrow[rru, "proj_b" description] \arrow[rrrr, "\vartist_b"] &  &             &  & Im_{\vartist_b} \in \arrow[llu, "g_b" description, dashed]
 \end{tikzcd}
@@
 \begin{equation}
-  \label{eq:math:artist:diagram}
+  \label{eq:artist:construction:all}
   \begin{tikzcd}
@@
 \subsection{Visualization Library Components}
-\subsubsection{Graphic to Data: \vindex}
-\begin{equation}
-  \begin{tikzcd}
-      \dtotal \arrow[d, "\pi"'] & \gtotal \arrow[d, "\pi"'] \\
-      \dbase                   & \gbase \arrow[l, "\vindex"']
-  \end{tikzcd}
-  \label{eq:math:graphic:vindex}
-\end{equation}
-
-The functor $\vindex$ is a deformation retract, which means....\cite{hatcherAlgebraicTopology2002,spanier1989algebraic}
 
@@
 
-\subsubsection{Data to Visual Encodings: \vchannel} %look at what K&S call these stages
+\subsubsection{Graphic to Data: \vindex}
+\begin{equation}
+  \label{eq:artist:construction:xi}
+  \begin{tikzcd}
+      \dtotal \arrow[d, "\pi"'] & \gtotal \arrow[d, "\pi"'] \\
+      \dbase                   & \gbase \arrow[l, "\vindex"']
+  \end{tikzcd}
+\end{equation}
+
+The functor $\vindex$ is a deformation retract, which means....\cite{hatcherAlgebraicTopology2002,spanier1989algebraic}
+
+\begin{equation}
+  \label{eq:artist:construction:xi:multi_base}
+  \begin{tikzcd}
+    \dbase_{c} \arrow[d, hook] \arrow[r, hook] & \dbase_{b}                 & \gbase_b \arrow[l, "\vindex"]                                                \\
+    \dbase_{a}                                 & \gbase_a \arrow[l, "\vindex"'] & \gbase_c \arrow[l, hook'] \arrow[u, hook] \arrow[llu, "\vindex" description]
+  \end{tikzcd}
+\end{equation}
+
+Can be implemented in V
+
+\begin{equation}
+  \label{eq:artist:construction:xi:v}
+  \begin{tikzcd}
+  \textcolor{color=gray!50}{\dtotal}\arrow[rd, "\pi"', color=gray!50] & \vtotal \arrow[d, "\pi"] & \vindex^*\vtotal \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"', dashed] & \textcolor{color=gray!50}{\gtotal} \arrow[ld, "\pi", color=gray!50] \\
+    & \dbase                   & \gbase \arrow[l, "\vindex"']                                                &                          
+    \end{tikzcd}
+\end{equation}
+
+\subsubsection{Data to Visual Encodings: \vchannel} 
+\begin{equation}
+  \label{eq:artist:construction:nu}
+\begin{tikzcd}
+  \textcolor{gray!50}{\dtotal} \arrow[rd, "\pi"'] \arrow[r, "\vchannel"] & \vtotal \arrow[d, "\pi"] & \vindex^*\vtotal \arrow[d, "\vindex^*\pi"', color=gray!50] \arrow[l, "\vindex^*", color=gray!50] & \textcolor{gray!50}{\gtotal} \arrow[ld, "\pi", color=gray!50] \\
+& \dbase                   & \textcolor{gray!50}{\gbase} \arrow[l, "\vindex"', color=gray!50]                                       &                          
+  \end{tikzcd}
+\end{equation}
+
 Visual encoding functions \vindex\ are functors because they are expected to preserve structure on the data and visual side, as discussed in \autoref{sec:related-work:equivariance}.
@@
 \begin{equation}
+\label{eq:artist:construction:nu:fiber_map}
+\begin{tikzcd}[column sep=.5em]
+  \dfiber \arrow[rr, "\vchannel"] &                                                         & \vfiber \\
+                                  & \dbase \arrow[lu, "\dsection"] \arrow[ru, "\vsection"'] &        
+  \end{tikzcd}
+\end{equation}
+
+which means equivariance of 
+\begin{equation}
+\label{eq:artist:construction:nu:equivariance}
+\begin{tikzcd}
+  \dsection \arrow[d, "\varphi"'] \arrow[r, "\vchannel"] & \vsection \arrow[d, "\vchannel_{\varphi}"] \\
+  \dsection^{\prime} \arrow[r, "\vchannel"]                       & \vsection^{\prime}                                 
+  \end{tikzcd}
+\end{equation}
+
+But actually multiple variables please
+\begin{equation}
   \label{eq:math:artist:nu}
@@
   \label{eq:artist:construction:multi_nu}
-  \begin{tikzcd}
-    \dfiber_a \arrow[rrd, "proj_a" description] \arrow[rrrr, "\vchannel^{a}_{\dfiber_c}"] &  &                                                                                                                       &  & \vfiber^{a}_{\dfiber_c} \\
-                                                                                          &  & \dfiber_{c} \arrow[rru, "\vchannel^{a}_{\dfiber_c}" description] \arrow[rrd, "\vchannel^{b}_{\dfiber_c}" description] &  &                         \\
-    \dfiber_b \arrow[rru, "proj_b" description] \arrow[rrrr, "\vchannel^{a}_{\dfiber_c}"] &  &                                                                                                                       &  & \vfiber^{b}_{\dfiber_c}
+  \begin{tikzcd}[row sep=.5em]
+    \dfiber_a \arrow[rrd, "proj_a" description] \arrow[rrrr, "\vchannel_{\dfiber_c}"]          &  &                                                                                                                        &  & \vfiber_{\dfiber_c}          \\
+    &  & \dfiber_{c} \arrow[rru, "\vchannel_{\dfiber_c}" description] \arrow[rrd, "\vchannel^{\prime}_{\dfiber_c}" description] &  &                              \\
+\dfiber_b \arrow[rru, "proj_b" description] \arrow[rrrr, "\vchannel^{\prime}_{\dfiber_c}"] &  &                                                                                                                        &  & \vfiber^{\prime}_{\dfiber_c}
     \end{tikzcd}
@@
 
-the same fiber in two different data sets run through the same $\vchannel_{i}$ encoder will output to the same set of visual values in the corresponding visual fiber $\vfiber_i$.
+\subsubsection{Visual to Graphic: \vmark} 
+\begin{equation}
+  \label{eq:artist:construction:q}
+  \begin{tikzcd}
+    \textcolor{gray!50}{\dtotal} \arrow[rd, "\pi"', color=gray!50] \arrow[r, "\vchannel", color=gray!50] & \textcolor{gray!50}{\vtotal} \arrow[d, "\pi", color=gray!50] & \vindex^*\vtotal \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"', color=gray!50] \arrow[r, "\vmark"] & \gtotal \arrow[ld, "\pi"] \\
+    & \dbase & \gbase \arrow[l, "\vindex"', color=gray!50]                                            &                          
+    \end{tikzcd}
+\end{equation}
 
-\subsubsection{Visual to Graphic: \vmark} % look at what K&S call these stages
+\begin{equation}
+  \label{eq:artist:construction:q:xi}
+  \begin{tikzcd}[column sep= 1em, row sep=.7em]
+    \vtotal \arrow[rrd, "\hat{\vmark}"]                             &  &         \\
+                                                                    &  & \gtotal \\
+    \vindex^*\vtotal \arrow[uu, "\vindex^*"] \arrow[rru, "\vmark"'] &  &        
+    \end{tikzcd}
+\end{equation}
+
+\begin{equation}
+  \label{eq:artist:construction:q:fiber}
+  \begin{tikzcd}[column sep=.5em]
+    \vfiber \arrow[rr, "\vmark"] &                                                         & \gfiber \\
+    & \gbase \arrow[ru, "\gsection"'] \arrow[lu, "\vsection"] &        
+    \end{tikzcd}
+\end{equation}
 Visual encodings to something like marks
+
+\begin{equation}
+  \label{eq:artist:construction:q:equivariance}
+  \begin{tikzcd}
+    \vsection \arrow[d, "\vchannel_\varphi"'] \arrow[r, "\vmark"] & \gsection \arrow[d, "\vmark_\varphi"] \\
+    \vsection^{\prime} \arrow[r, "\vmark"]                                 & \gsection^{\prime}                            
+    \end{tikzcd}
+\end{equation}
+
 \begin{figure}[!h]
@@
 \appendices
-\section{Rendering: \gsection}
-\section{Manufacturing $\vmarkd \leftarrow \vmark$}
-\begin{equation}
-  \begin{tikzcd}
-      \textcolor{gray!50}{\dtotal} \arrow[r, "\vchannel", color=gray!50] \arrow[rd, "\pi"', color=gray!50] & \vtotal \arrow[d, "\pi"']                  & \vindex^*\vtotal \arrow[r, "\vmark", color=gray!50] \arrow[d, "\vindex^*\pi"'] \arrow[l,  "\vindex^*"'] & \textcolor{gray!50}{\gtotal} \arrow[ld, "\pi", color=gray!50] \\ & \dbase \arrow[u, "\vsection"', bend right] & \gbase \arrow[l, "\vindex"'] \arrow[u, "\vindex^*\vsection"', bend right]               &
-      \end{tikzcd}
-      \label{eq:math:artist:qhat}
-\end{equation}
 

### 2021-10-28 paper/main.tex
changed V V* diagram

@@
 \begin{tikzcd}
-  \textcolor{gray!50}{\dtotal} \arrow[rd, "\pi"'] \arrow[r, "\vchannel"] & \vtotal \arrow[d, "\pi"] & \vindex^*\vtotal \arrow[d, "\vindex^*\pi"', color=gray!50] \arrow[l, "\vindex^*", color=gray!50] & \textcolor{gray!50}{\gtotal} \arrow[ld, "\pi", color=gray!50] \\
+  \dtotal \arrow[rd, "\pi"'] \arrow[r, "\vchannel"] & \vtotal \arrow[d, "\pi"] & \textcolor{fade}{\vindex^*\vtotal} \arrow[d, "\vindex^*\pi"', color=fade] \arrow[l, "\vindex^*", color=fade] & \textcolor{fade}{\gtotal} \arrow[ld, "\pi", color=fade] \\
 & \dbase                   & \textcolor{gray!50}{\gbase} \arrow[l, "\vindex"', color=gray!50]                                       &                          
@@
      &  & \dfiber_{c} \arrow[rru, "\vchannel_{\dfiber_c}" description] \arrow[rrd, "\vchannel^{\prime}_{\dfiber_c}" description] &  &                              \\
-\dfiber_b \arrow[rru, "proj_b" description] \arrow[rrrr, "\vchannel^{\prime}_{\dfiber_c}"] &  &                                                                                                                        &  & \vfiber^{\prime}_{\dfiber_c}
+    \dfiber_b \arrow[rru, "proj_b" description] \arrow[rrrr, "\vchannel^{\prime}_{\dfiber_c}"'] &  &                                                                                                                        &  & \vfiber^{\prime}_{\dfiber_c}
     \end{tikzcd}
@@
   \begin{tikzcd}
-    \textcolor{gray!50}{\dtotal} \arrow[rd, "\pi"', color=gray!50] \arrow[r, "\vchannel", color=gray!50] & \textcolor{gray!50}{\vtotal} \arrow[d, "\pi", color=gray!50] & \vindex^*\vtotal \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"', color=gray!50] \arrow[r, "\vmark"] & \gtotal \arrow[ld, "\pi"] \\
-    & \dbase & \gbase \arrow[l, "\vindex"', color=gray!50]                                            &                          
+    \textcolor{fade}{\dtotal} \arrow[rd, "\pi"', color=fade] \arrow[r, "\vchannel", color=fade] & \textcolor{gray!50}{\vtotal} \arrow[d, "\pi", color=fade] & \vindex^*\vtotal \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"', color=fade] \arrow[r, "\vmark"] & \gtotal \arrow[ld, "\pi"] \\
+    & \textcolor{fade}{\dbase} & \gbase \arrow[l, "\vindex"', color=fade]                                            &                          
     \end{tikzcd}
@@
   \label{eq:artist:construction:q:xi}
-  \begin{tikzcd}[column sep= 1em, row sep=.7em]
-    \vtotal \arrow[rrd, "\hat{\vmark}"]                             &  &         \\
-                                                                    &  & \gtotal \\
-    \vindex^*\vtotal \arrow[uu, "\vindex^*"] \arrow[rru, "\vmark"'] &  &        
+  \begin{tikzcd}
+    \vtotalc \arrow[rr, "\vmarkdc", bend left] & \vtotalpullc \arrow[l, "\vindexc^*"] \arrow[r, "\vmarkc"'] & \gtotalc
   \end{tikzcd}

### 2021-10-29 paper/main.tex
artist diag

@@
   \label{eq:artist:intersection:fiber}
-  \begin{tikzcd}[row sep = .5em]
-    \dfiber_a \arrow[rrd, "proj_a" description] \arrow[rrrr, "\vartist_a"] &  &             &  & Im_{\vartist_a} \arrow[lld, "g_a" description, dashed] \\
-                                                                           &  & \dfiber_{c} &  &                                                                      \\
-    \dfiber_b \arrow[rru, "proj_b" description] \arrow[rrrr, "\vartist_b"] &  &             &  & Im_{\vartist_b} \in \arrow[llu, "g_b" description, dashed]
+  \begin{tikzcd}
+    F_a\times{_{F_c}}F_b     & F_a \arrow[d, "proj_a"] \\
+    F_b \arrow[r, "proj_b"'] & F_c                    
     \end{tikzcd}
@@
   \begin{tikzcd}[row sep=.5em]
-    \dfiber_a \arrow[rrd, "proj_a" description] \arrow[rrrr, "\vchannel_{\dfiber_c}"]           &  &  &  & \vfiber_{\dfiber_c}          \\
-     &  & \dfiber_{c} \arrow[rru, "\vchannel_{\dfiber_c}" description] \arrow[rrd, "\vchannel^{\prime}_{\dfiber_c}" description] &  &                              \\
-    \dfiber_b \arrow[rru, "proj_b" description] \arrow[rrrr, "\vchannel^{\prime}_{\dfiber_c}"'] &  &                                                                                                                        &  & \vfiber^{\prime}_{\dfiber_c}
+    \dfiberc_a \arrow[rrd, "proj_a" description] \arrow[rrrr, "\vchannelc_c"]  &  &  &  & \vfiberc_c \\
+    &  & \dfiberc_{c} \arrow[rru, "\vchannelc_c" description] \arrow[rrd, "\vchannelc^{\prime}_c" description] &  & \\
+    \dfiberc_b \arrow[rru, "proj_b" description] \arrow[rrrr, "{\vchannelc_c}^{\prime}"'] &  &                                                                                                       &  & {\vfiberc_c}^{\prime}
     \end{tikzcd}
@@
 \end{figure}
+
+\begin{equation}
+  \label{eq:artist:construction:q:multifiber}
+\begin{tikzcd}
+  \dfiberc_a \arrow[rr, "\vchannelc"] \arrow[rd, "proj_a" description]          &            & \vfiberc_a \arrow[rr, "\vmarkd"]         &  & Im_{\vmarkdc} \subset \gtotalc \arrow[llld, "g_a" description, dashed]                  \\
+ & \dfiberc_c &                                          &  &\\
+  \dfiberc_b \arrow[rr, "\vchannelc^{\prime}"] \arrow[ru, "proj_b" description] &            & \vfiber_b \arrow[rr, "\vmarkd^{\prime}"] &  & Im_{\vmarkdc^{\prime}}\subset \gtotalc^{\prime} \arrow[lllu, "g_b" description, dashed]
+  \end{tikzcd}
+\end{equation}
 

### 2021-10-29 paper/main.tex
fixing equation

@@
  & \dfiberc_c &                                          &  &\\
-  \dfiberc_b \arrow[rr, "\vchannelc^{\prime}"] \arrow[ru, "proj_b" description] &            & \vfiber_b \arrow[rr, "\vmarkd^{\prime}"] &  & Im_{\vmarkdc^{\prime}}\subset \gtotalc^{\prime} \arrow[lllu, "g_b" description, dashed]
+  \dfiberc_b \arrow[rr, "\vchannelc^{\prime}"] \arrow[ru, "proj_b" description] &            & \vfiberc_b \arrow[rr, "\vmarkd^{\prime}"] &  & Im_{\vmarkdc^{\prime}}\subset \gtotalc^{\prime} \arrow[lllu, "g_b" description, dashed]
   \end{tikzcd}

### 2021-11-08 paper/main.tex
composition diagrams

@@
 
-Here we present a very brief summary of topics, for more information Hatcher\cite{hatcherAlgebraicTopology2002}, Munkres\cite{munkresElementsAlgebraicTopology1984}, and Spanier \cite{spanier1989algebraic}.
+Here we present a very brief summary of fiber bundles, for more information see Hatcher\cite{hatcherAlgebraicTopology2002}, Munkres\cite{munkresElementsAlgebraicTopology1984}, and Spanier \cite{spanier1989algebraic}.
 
@@
   \label{eq:artist:presheaf}
-  \begin{tikzcd}[row sep=small]
-    U_1 \arrow[dd, "\iota"'] & {\Gamma(U_1, \dtotal)}                              \\
-    {} \arrow[r, "\sheaf_{\dtotal}"]   & {}                                                  \\
-    U_2                      & {\Gamma(U_2, \dtotal)} \arrow[uu, "\iota^*"']
+  \begin{tikzcd}[column sep=small]
+    {\Gamma(U_1, \dtotalc)}                                       & {}                                                    & {\Gamma(U_2, \dtotalc)} \arrow[ll, "\iota^*"', hook'] \\
+                                                                  &                                                       &                                                       \\
+    U_1 \arrow[uu, "\sheaf_{\dtotal}"] \arrow[rr, "\iota"', hook] & {} \arrow[uu, "\sheaf_{\dtotal}" description, dotted] & U_2 \arrow[uu, "\sheaf_\dtotal"']                    
     \end{tikzcd}
@@
   \begin{tikzcd}
-    & \Gamma(\mathcal{K}, \dfiber) & \\
-{} \arrow[rr, "\varphi", Rightarrow, shift right, shorten=3.5em] & & {} \\
-    & \mathcal{\dbase}^{op} \arrow[uu, "\sheaf_{\dtotal}", bend left] \arrow[uu, "\sheaf_{\dtotal}"', bend right] &   
+    & \mathcal{\dtotalc} & \\
+{} \arrow[rr, "\dfuncc", Rightarrow, shift right, shorten=3em] & & {} \\
+    & \mathcal{\dbasec}^{op} \arrow[uu, "\sheaf_{\dtotalc}", bend left] \arrow[uu, "\sheaf_{\dtotalc}"', bend right] &   
 \end{tikzcd}
@@
 
-The graphic category $\mathcal{\gtotal}$ consists of the functor object $\sheaf_{\gtotal}: \mathcal{S}^{op} \rightarrow \Gamma(\mathcal{S}, \gtotal)$ and morphisms between sheaves.
-The category $\mathcal{S}$ is the category of open sets $\{U\}$ over the graphic base space \gbase.
-The set of sets of sections $\Gamma(\mathcal{S}, \gtotal)$ is a subcatagory of \texbf{Set}.
+\note{rewrite paragraph}
+The graphic category $\mathcal{\gtotalc}$ consists of the functor object $\sheaf_{\gtotalc}: \mathcal{\gbasec}^{op} \rightarrow \Gamma(\mathcal{\gbasec}, \gtotal)$ and morphisms between sheaves.
+The category $\mathcal{\gbasec}$ is the category of open sets $\{U\}$ over the graphic base space \gbase.
+The set of sets of sections $\Gamma(\mathcal{\gbase}, \gtotal)$ is a subcatagory of \textbf{Set}.
 
@@
 \begin{figure}[!h]
-  \centering
-  \subfloat[$\dbase_c = \dbase_a \cap \dbase_b$]{\includegraphics[width=.4\columnwidth]{union_k.png}%
-  \label{fig:artist:union:k}}
-  %\hfill
-  \subfloat[$\dbase_c = \varnothing$]{\includegraphics[width=.4\columnwidth]{disjoint_k.png}%
+  \includegraphics[width=\columnwidth]{k_union.png}
   \label{fig:artist:union:disjoint_k}}
@@
 In this figure, the visual transformation is to a point on the line on screen.
-An artist that preserves the continuity of the function must generate graphics such that functions evaluted  $\dbasepoint \in \dbase_a \cap \dbase_b$}
+An artist that preserves the continuity of the function must generate graphics such that functions evaluated  $\dbasepoint \in \dbase_a \cap \dbase_b$}
 \end{figure}
 
-An example of overlapping indexing is a parallelized verion of a sliding window algorithm where window overlaps must resolve to the same value at the same position \cite{chuTimeSeriesSegmentation1995};
+An example of overlapping indexing is a parallelized version of a sliding window algorithm where window overlaps must resolve to the same value at the same position \cite{chuTimeSeriesSegmentation1995};
 
@@
 \begin{tikzcd}
-  \dbase_{c} \arrow[d, "\iota", hook] \arrow[r, "\iota", hook] & \dbase_{b} \arrow[d, "i_{\dbase_b}"]      \\
-  \dbase_{a} \arrow[r, "i_{\dbase_a}"]                         & \dbase_{a} \bigsqcup_{\dbase_{c}}\dbase_{b}
+  \dbasec_{c} \arrow[d, "\iota", hook] \arrow[r, "\iota", hook] & \dbasec_{b} \arrow[d, "i_{\dbasec_b}"]      \\
+  \dbasec_{a} \arrow[r, "i_{\dbasec_a}"]                         & \dbasec_{a} \bigsqcup_{\dbasec_{c}}\dbasec_{b}
   \end{tikzcd}
 \end{equation}
+
 where the projection functions $i_{\dbase_a},i_{\dbase_b}$ behave such that $i_{\dbase_a}(\dbasepoint)|_{\dbasepoint \in \dbase_b} = \left[\dbasepoint\right]$, $i_{\dbase_b}(\dbasepoint)|_{\dbasepoint \in \dbase_b} = \left[\dbasepoint\right]$, meaning the projection functions yield equivalent points in the disjoint union $\dbase_a \bigsqcup_{\dbase_c} \dbase_b$.
@@
 
-\subsubsection{}{Shared Fields}
+\subsubsection{Shared Fields}
 
@@
 \end{figure}
-an example of expecting consistency in overlapping fields are the multiple view constraints described by Qu and Hullman \cite{hullmanKeeping2018}. 
+
 \begin{equation}
@@
   \begin{tikzcd}
-    F_a\times{_{F_c}}F_b     & F_a \arrow[d, "proj_a"] \\
-    F_b \arrow[r, "proj_b"'] & F_c                    
+    \dfiberc_{a}\times \dfiberc_{b} \arrow[d, "proj_b"] \arrow[rr, "proj_a"] &  & \dfiberc_{a} \arrow[d, "proj_c"] \\
+    \dfiberc_{b} \arrow[rr, "proj_b"]                                                  &  & \dfiberc_{c}                    
     \end{tikzcd}
 \end{equation}
-
-% introduce notion of equivariant fiber, chasing diagram of F is different Gamma - Kosara and Zeminiskwy differentiable glyph 
-% case 1: F3 \subset F1, {JFK} \in {JFK, EWR}
-% case 2: F1 = F3 \times F1, F2= F3 \times F2, 
-  % F1 ->F3 is projection -> Fiber consistency
-% in that case, we can form the plus operator, disjoint union of A1 + A2 going in is consistent on subspace
 
@@
 \begin{definition} Following from \label{def:artist:} the artist \vartist is a special case of  natural transformations
+
+  %% this equation is actually on sheafs and is not a map 
   \begin{equation}
@@
 
+%%% no arrow V->H
+%%% Q is defined at the level of Gamma\begin{equation}
 \begin{equation}
@@
   \begin{tikzcd}
-      \dtotal \arrow[r, "\vchannel"] \arrow[rd, "\pi"'] & \vtotal \arrow[d, "\pi"] & \vindex^*\vtotal \arrow[r, "\vmark"] \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"'] & \gtotal \arrow[ld, "\pi"] \\
-                                            & \dbase                  & \gbase \arrow[l, "\vindex"']                                              &                    
+    \dtotalc \arrow[rr, "\vchannel"] \arrow[rdd, "\pi"] &         & \dtotalc \arrow[ldd, "\pi"'] & \mathcal{\vtotalc}                                                        & Im_{\vmarkc} \subset \gtotalc                     \\
+    &         &                              & {} \arrow[r, "\vmarkc"]                                                   & {}                                                \\
+    & \dbasec &                              & \mathcal{\dbasec} \arrow[uu, "\sheaf_{\vtotalc}"] \arrow[r, "\vindexc^*"] & \mathcal{\gbasec} \arrow[uu, "\sheaf_{\vmarkc}"']
     \end{tikzcd}
@@
 
-But actually multiple variables please
+\paragraph{Multiview Constraints}
+\begin{figure}[!h]
+  \centering
+  \subfloat[No shared fields]]{\includegraphics[width=.4\columnwidth]{exploding_artist.png}%
+  \label{fig:artist:union:k}}
+  %\hfill
+  \subfloat[Shared fields]{\includegraphics[width=.4\columnwidth]{combined_artist.png}%
+  \label{fig:artist:union:disjoint_k}}
+  \caption{In \autoref{fig:artist:union:k}, the input object $\sheaf_{\dtotal}$ encodes a continuous function over spaces $\dtotal_a$, $\dtotal_b$ 
+  and overlapping space $\dtotal_c$.
+In this figure, the visual transformation is to a point on the line on screen.
+An artist that preserves the continuity of the function must generate graphics such that functions evaluted  $\dbasepoint \in \dbase_a \cap \dbase_b$}
+\end{figure}
+
+The concept that shared data fields should be encoded visually in a consistent is formally expressed by Qu and Hullman \cite{hullmanKeeping2018} as the notion that the same field should have the same scale.
+
 \begin{equation}

### 2021-11-09 paper/main.tex
rearranging

@@
 
+\begin{figure}[!h]
+  \centering
+  \subfloat[\dbasec_a \sqcup_{\dbasec_c} \dase_b]{\includegraphics[width=1.2in]{union_k.png}%
+  \label{fig_first_case}}
+  %\hfil
+  \subfloat[\dbasec_c=\varnothing]{\includegraphics[width=1.2in]{disjoint_k.png}%
+  \label{fig_second_case}}
+  \caption{\note{probably doesn't need the k=0}}
+  \label{fig_sim}
+\end{figure}
+
 \begin{equation}
@@
 In this figure, the visual transformation is to a point on the line on screen.
-An artist that preserves the continuity of the function must generate graphics such that functions evaluted  $\dbasepoint \in \dbase_a \cap \dbase_b$}
+An artist that preserves the continuity of the function must generate graphics such that functions evaluted  $\dbasepoint \in \dbase_a \cap \dbase_b$
+  \note{replace a) w/ fiber cross diagram? - use first half introduced in \autoref{fig:artist:compose:union_fiber}}}
 \end{figure}

### 2021-11-10 paper/main.tex
fb defintions

@@
 
-The model described in this work provides a modol for expressing data sets with different topological properties. 
-We obtain this generality by using the mathematical theory of fiber bundles as the basis of our abstraction, as proposed by Butler, Bryson, and Pendley\cite{butlerVisualizationModelBased1989,butlerVectorBundleClassesForm1992}.
-In this paper, we build on their work that proposes using topological spaces to represent different properties of data.
+The model described in this work provides a model for expressing data sets with different topological properties. 
+We obtain this generality by using a mathematical structure called a fiber bundles as the basis of our abstraction, as proposed by Butler, Bryson, and Pendley\cite{butlerVisualizationModelBased1989,
+butlerVectorBundleClassesForm1992}.
+As described by them, a fiber bundle is a formal model of the mapping between data points and the underlying topological space they lie in.
+For example, nodes on a network and the graph of the network, an image and the underlying grid at which the image is sampled, or table columns and the table index.
 
-Specifically,Butler, Bryson, and Pendley suggest that fiber bundles be the basis of an abstract data model. Fiber bundles are a collection of topological spaces. 
-\begin{definition}
-A topological space $(X, \mathscr{T})$ is a set $X$ with a topology $\mathscr{T}$.
-Topologies are collections of open sets $U$ such that the empty set and $X$ are in the collection of open sets $\mathscr{T}$, the union of elements in $\mathscr{T}$ \note{finish out this definition}
-\end{definition}
-
-Here we present a very brief summary of fiber bundles, for more information see Hatcher\cite{hatcherAlgebraicTopology2002}, Munkres\cite{munkresElementsAlgebraicTopology1984}, and Spanier \cite{spanier1989algebraic}.
-
-\begin{LaTeXdescription}
-  \item[\textcolor{total}{Total Space} \dtotal]
-  \item[\textcolor{fiber}{Fiber Space} \dfiber]
-  \item[\textcolor{base}{Base Space} \dbase]  
-\end{LaTeXdescription}
-
-with a projection map $\pi:\dtotal\rightarrow\dbase$ that connects every point in \dtotal\ to a point in \dbase. 
+Formally, a fiber bundle is a mathematical structure $(\dtotalc, \dbasec, \pi, \dfiberc)$
 \begin{equation}
@@
 
-As indicated by $\hookrightarrow$, the fiber space \dfiber\ is embedded inside the total space \dtotal\.
+\begin{definition}
+The \textcolor{base}{base space} \dbasec\ is a topological space, which means it is a a set $\dbase$ with a collection of open sets surrounding each element in the set.
+For each point $\dbasepoint in \dbase$, there is a collection of subsets $\{U\} \subset \dbase$ such that $x \in U \subseteq \dbase$ \cite{spanier1989algebraic, bradleyTopologyCategoricalApproach2020}
+\end{definition}
+\note{intuition about base space}
+
+\begin{definition} The \textcolor{fiber}{fiber space} \dfiberc\ is a topological space such that for every $\dbasepoint \in \dbase$, the fiber over that point is isomorphic to the preimage of that point $\dfiber_{\dbasepoint} \cong \pi^{-1}(\{k\})$.
+All fibers $\dfiber_{\dbasepoint}$ are homemorphic to each other.
+\cite{spanier1989algebraic, hatcherAlgebraicTopology2002}
+\end{definition}
+\note{intuition about fiber}
+
+\begin{definition} The \textcolor{total}{total space} \dtotalc\ is the space reachable via the projection map $\pi$.
+The total space is always locally trivial, meaning $\dtotal = \dbase \times \dfiber$ over any open neighborhood $U_{k}$.
+The bundle is trivial when $\dtotal = \dbase \times \dfiber$ is true globally; a non-trivial bundle can be thought of as  the twisted product.
+\cite{hatcherAlgebraicTopology2002,nlab:twisting_function}
+\end{definition}
+
 This is illustrated in \autoref{fig:related-work:fiber-bundle}, wherein values lives in the fiber $\dfiber\subseteq\dtotal$.
@@
 
-Throughout this paper, we denote fiber bundles with the tuple $(\textcolor{total}{total space}, \textcolor{base}{base space}, \pi, \textcolor{fiber}{fiber space})$
+Here we present a very brief summary of fiber bundles and sheaves, for more information see Hatcher\cite{hatcherAlgebraicTopology2002}, Munkres\cite{munkresElementsAlgebraicTopology1984}, Spanier \cite{spanier1989algebraic} and Ghrist\cite{ghristElementaryAppliedTopology2014,ghristHomologicalAlgebraData2018}.
 
@@
   \includegraphics[width=\columnwidth]{k_union.png}
-  \label{fig:artist:union:disjoint_k}}
+  \label{fig:artist:union:disjoint_k}
   \caption{In \autoref{fig:artist:union:k}, the input object $\sheaf_{\dtotal}$ encodes a continuous function over spaces $\dtotal_a$, $\dtotal_b$ 

### 2021-11-10 paper/main.tex
sheaves pulled to intro and shortened

@@
 
-This is illustrated in \autoref{fig:related-work:fiber-bundle}, wherein values lives in the fiber $\dfiber\subseteq\dtotal$.
+How these spaces fit together is illustrated in \autoref{fig:related-work:fiber-bundle}, wherein values lives in the fiber $\dfiber\subseteq\dtotal$.
 In this example, the fiber \dfiber\ is the cartesian product of two sets $F_{0}\times F_{1}$ where each fiber $F_{i}$ is ....
@@
 \end{equation}
-
-While the values are embedded in \dfiber, the continuity of the data is modeled as the base space \dbase, which is the quotient space \cite{QuotientSpaceTopology2020}
-\begin{equation}
-formal math of the base space as equivalence class for F_[k]
-\end{equation}
-which means that every point in $\dfiber_{k}$ maps to a point $k \subset K$.
 
@@
 \end{equation}
+
+The set of all sections sections $\{\dsection^{i}: \U_{j}\rightarrow \dtotal\}|_{i\in I}$ over an arbitrary open set $U_{j}$ is denoted $\Gamma(U_{\dbasepoint}, \dtotal)$.
+The set of all sections over the total space \dbase\ is denoted $\Gamma(\dbase, \dtotal)$.
+
+The map from the open set to the set of sections is called a presheaf $\sheaf: U_j \rightarrow \Gamma(U_j, \dtotal)$ \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}.
+The presheaf preserves inclusion maps between the open sets and the sets of sections
+
+\begin{equation}
+  \label{eq:artist:presheaf}
+  \begin{tikzcd}[column sep=small]
+    {\Gamma(U_1, \dtotalc)}                                       & {}                                                    & {\Gamma(U_2, \dtotalc)} \arrow[ll, "\iota^*"', hook'] \\
+                                                                  &                                                       &                                                       \\
+    U_1 \arrow[uu, "\sheaf_{\dtotal}"] \arrow[rr, "\iota"', hook] & {} \arrow[uu, "\sheaf_{\dtotal}" description, dotted] & U_2 \arrow[uu, "\sheaf_\dtotal"']                    
+    \end{tikzcd}
+\end{equation}
+
+meaning that the presheaf ensures that given an inclusion map between open sets $\iota$, there is a corresponding map $\iota^*$ between the sets of sections.
+This means that a smaller space $U_1$ is included in a larger space $U_2$ and a function that is continuous over a larger space $U_2$ is continuous over a subspace $U_1 \subspace U_2$.
+Sheaves are presheafs where sets of sections $\Gamma$ are defined over unions of open sets $U \cup_{U\in \mathscr{U}}$ \cite{spanier1989algebraic, nlab:sheaf}.
+Because they preserve inclusion maps, sheaves are often used as an abstraction for keeping track of how data over topological spaces is glued together \cite{ghristElementaryAppliedTopology2014}.
 
@@
 \end{definition}
-
- The artist takes as input the sheaf object $\sheaf_{\dtotal}$ because the sheaf is a way of keeping track of how data over topological spaces is glued together \cite{ghristElementaryAppliedTopology2014}.
-This bookkeeping is  illustrated in \autoref{eq:artist:presheaf} as the preservation of inclusion maps $\iota, \iota^{*}$ in presheaf maps between open sets $U_i$ and their corresponding set of sections $\Gamma(U_i, \dtotal)$.
-A presheaf is a contravariant functor from an arbitrary category $\mathcal{C}$ to \textbf{Set} \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf} and sheafs are presheafs where sets of sections $\Gamma$ are defined over unions of open sets $U \cup_{U\in \mathscr{U}}$ \cite{spanier1989algebraic, nlab:sheaf}.
-
-\begin{equation}
-  \label{eq:artist:presheaf}
-  \begin{tikzcd}[column sep=small]
-    {\Gamma(U_1, \dtotalc)}                                       & {}                                                    & {\Gamma(U_2, \dtotalc)} \arrow[ll, "\iota^*"', hook'] \\
-                                                                  &                                                       &                                                       \\
-    U_1 \arrow[uu, "\sheaf_{\dtotal}"] \arrow[rr, "\iota"', hook] & {} \arrow[uu, "\sheaf_{\dtotal}" description, dotted] & U_2 \arrow[uu, "\sheaf_\dtotal"']                    
-    \end{tikzcd}
-\end{equation}
-
- The functor $\sheaf_{\dtotal}$ is contravariant because the inclusion map $\iota$ between open sets $U_i \in \mathcal{K}$ goes in the opposite direction from the map $\iota^*$ between the sections in $\Gamma(U_{i}, \dtotal)$ \cite{riehlCategoryTheoryContext}.
-The open set $U_1$ is a subset of $U_2$, therefore $\iota: U_1 \rightarrow U_2$.
-On the other hand, a set of functions defined on a space $U_2$ must be defined on the subspace $\iota: \Gamma(U_2, \dbase) \rightarrow \Gamma(U_1, \dbase)$ but the reverse does not need to be true.
-For example, a continuous function defined over the interval $\left[.25,.5\right]$ need not be defined over $\left[0, 1\right]$ but must exist over  $\left[.3, .45\right]$.
-We use the notation $\mathcal{K}^{op}$ to signify that the functor $\sheaf$ is contravariant and $\sheaf$ to denote that it is a sheaf.
-We also denote the sheaf over the total space \dbase\ of the fiber bundle \dtotal\ as $\sheaf(\dtotal)$.
 

### 2021-11-10 paper/main.tex
new category E that is sets of gamma

@@
   \begin{tikzcd}
-      \dfiber \arrow[r, hook] & \dtotal \arrow[r, "\pi"] & \dbase
+      \dfiberc \arrow[r, hook] & \dtotalc \arrow[r, "\pi"] & \dbasec
   \end{tikzcd}
@@
 \begin{definition}
-The \textcolor{base}{base space} \dbasec\ is a topological space, which means it is a a set $\dbase$ with a collection of open sets surrounding each element in the set.
+The \textcolor{base}{base space} \dbasec\ is a topological space, which means it is a set $\dbase$ with a collection of open sets \cite{TopologyVsTopology} surrounding each element in the set.
+Open sets are a collection of subsets in a set $X$, including the empty set, the whole set $X$, and every union and intersection of its member subsets.
 For each point $\dbasepoint in \dbase$, there is a collection of subsets $\{U\} \subset \dbase$ such that $x \in U \subseteq \dbase$ \cite{spanier1989algebraic, bradleyTopologyCategoricalApproach2020}
@@
 
-How these spaces fit together is illustrated in \autoref{fig:related-work:fiber-bundle}, wherein values lives in the fiber $\dfiber\subseteq\dtotal$.
-In this example, the fiber \dfiber\ is the cartesian product of two sets $F_{0}\times F_{1}$ where each fiber $F_{i}$ is ....
-\begin{equation}
-  formal equation of the fiber?
-  \label{eq:related-work:fiber}
-\end{equation}
-
 \begin{equation}
   \begin{tikzcd}
-      \dfiber \arrow[r, hook] & \dtotal \arrow[d, "\pi"'] \\
-                        & \dbase \arrow[u, "\dsection"', bend right]
+      \dfiberc \arrow[r, hook] & \dtotalc \arrow[d, "\pi"'] \\
+                        & \dbasec \arrow[u, "\dsectionc"', bend right]
   \end{tikzcd}
@@
 
+How these spaces fit together is illustrated in \autoref{fig:related-work:fiber-bundle}, wherein values lives in the fiber $\dfiber\subseteq\dtotal$.
+In this example, the fiber \dfiber\ is the cartesian product of two sets $F_{0}\times F_{1}$ where each fiber $F_{i}$ is ....
+
+\subsubsection{Sheaves}
 The set of all sections sections $\{\dsection^{i}: \U_{j}\rightarrow \dtotal\}|_{i\in I}$ over an arbitrary open set $U_{j}$ is denoted $\Gamma(U_{\dbasepoint}, \dtotal)$.
@@
 We propose that the visualization transformation can be modeled as a functor, which we call the \textcolor{artist}{Artist} $\mathcal{\vartist}$\footnote{We call this transformation the artist because the \mintinline{python}{Artist} object in Matplotlib\cite{hunterMatplotlib2DGraphics2007}}.
-The artist \vartist\ converts data, which we model as the catgory $\mathcal{\dtotal}$ to graphics, which we model as the category $\mathcal{\gtotal}$.
+The artist \vartist\ converts data, which we model as the sheaf $\sheaf_{\dtotal}$ to graphics, which we model as the sheaf $\sheaf_{\vartist}$.
 We formulate this association as
@@
 \begin{equation}
-  \vartist: \mathcal{\dtotal} \rightarrow \mathcal{\gtotal}
+  \vartist: \sheaf_{\dtotal} \rightarrow \sheaf_{\vartist}
   \label{eq:artist:artist}
 \end{equation}
+
 In this section we describe the structure of the data and graphic spaces, which we formulate as objects in categories and the morphisms between those objects.
@@
 \label{sec:artist:category}
-We define the categories $\mathcal{\dtotal}$ and $\mathcal{\gtotal}$ as sheaves of sections\cite{SheafMathematics2021,spanier1989algebraic}, of fiberbundles, as introduced in \autoref{sec:related-work:fiber-bundles}.
-Sheaves are mathematical abstractions of how data over different parts of a continuous base space are glued together \cite{ghristElementaryAppliedTopology2014, ghri}, whether the data is distributed, streaming, or in disk; therefore sheaves are a way to express the continuity that the artist preserves.
-We assume that $\mathcal{E}$ is defined in terms of a fixed base and fiber space, for example a class of images or tables with the same schema.
-The graphic category $\mathcal{\gtotal}$ consists of the set of graphics that an artist \vartist\ is allowed to generate from a given $\mathcal{E}$, which is discussed further in \autoref{sec:artist:construction}.
 
@@
   \begin{itemize}
-    \item \textit{objects} all open sets $U_{i} \in \{U\}$ in the topological space $(\dbase, \mathscr{T})$), including the empty set $\varnothing$ and the maximum set \dbase. 
-    \item \textit{morphisms} inclusion maps from a subspace to a larger space $\iota: U_{i} \hookrightarrow U_{j}$, which is equivalent to $U_{i} \subseteq U_{j}$ and  
+    \item \textit{objects} all open sets $U_{j} \in \{U\}$ in the topological space $(\dbase, \mathscr{T})$), including the empty set $\varnothing$ and the maximum set \dbase. 
+    \item \textit{morphisms} inclusion maps from a subspace to a larger space $\iota: U_{j} \hookrightarrow U_{j}$, which is equivalent to $U_{j} \subseteq U_{j}$ and  
   \end{itemize} 
@@
 This is what allows the artist to generate graphics where the subset of data on view is dynamically updated, such as pan and zoom navigation\cite{NekrasovskiEvaluationPanZoom2006} and sliding windows on streaming data\cite{crouchDynamicGraphsSlidingwindow2013,chuTimeSeriesSegmentation1995}.
-\note{does this actually go down in the composition section?}
 
@@
 
-\begin{definition} The category of sets of sections $\Gamma(U_i,\dtotal)$, where $U_i$ are objects in $\mathcal{\dfiber}$, form a subcatagory of the category of sets \textbf{Set}.
-The morphisms between objects of this subcategory are restriction maps $\iota: \Gamma(U_j, \dtotal) \rightarrow \Gamma(U_i, \dtotal)$ where $U_{i} \subseteq U_{j}$.
+\begin{definition} The category \mathcal{\dtotalc} is the set of sets of sections $\{\Gamma(U_j,\dtotal)\}|U_j \in \mathcal{dbase}$.
+The category \mathcal{\dtotal} is a subcatagory of the category of sets \textbf{Set}.
+The morphisms between objects of $\mathcal{\dtotal}$ are restriction maps $\iota: \Gamma(U_j, \dtotal) \rightarrow \Gamma(U_i, \dtotal)$ where $U_{i} \subseteq U_{j}$.
 \end{definition}
 
-As mentioned in \autoref{sec:related-work:fiber-bundles}, the fiber space is embedded in the total space $\dfiber \hookrightarrow \dtotal$.
-The sections $\Gamma(\dbase, \dtotal)$ are called the global sections of $\dtotal$ \cite{SheafMathematics2021,spanier1989algebraic}.
 Each section is how we represent a unique dataset, where the sets of sections $\Gamma$ is the space of all datasets that have the same data fields in the \dfiber\ and continuity as encoded in\dbase\.
@@
 
-\begin{definition} The category $\mathcal{\dtotal}$ consists of 
-  \begin{itemize}
-    \item\textit{object} the sheaf\cite{SheafMathematics2021,ghristElementaryAppliedTopology2014} $\sheaf_{\dtotal}: \mathcal{K}^{op} \rightarrow \Gamma(\mathcal{K}, \dtotal)$
-    \item \textit{morphisms} arbitrary endomorphic\cite{fongInvitationAppliedCategory2019} bundle maps $\varphi: \dtotal\rightarrow \dtotal$.
-When the sheaf is restricted to the stalk\cite{StalkSheaf2019} $\mathscr{F}_{k}$ over a point $\dbasepoint\in \dbase$, then allowable morphisms are $\varphi \in Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$.
-  \end{itemize}
-\end{definition}
-
-Arbitrary morphisms on sheafs $\varphi:\mathcal{\dtotal} \rightarrow \mathcal{\dtotal}$ are natural transforms \cite{SheafMathematics2021,bradleyWhatNaturalTransformation}, meaning that the morphisms $\varphi$ compose with the inclusion maps $\iota$ \cite{fongInvitationAppliedCategory2019}.
+Arbitrary morphisms on sheafs are natural transforms \cite{SheafMathematics2021,bradleyWhatNaturalTransformation}, meaning that the morphisms $\varphi$ compose with the inclusion maps $\iota$ \cite{fongInvitationAppliedCategory2019}.
 As defined in \autoref{def:category:Tau}, the sets of sections $\Gamma(U_{i},\dtotal)$ are objects of the category $\mathcal{T}$.
@@
   \centering
-  \subfloat[\dbasec_a \sqcup_{\dbasec_c} \dase_b]{\includegraphics[width=1.2in]{union_k.png}%
+  \subfloat[\dbasec_a \sqcup_{\dbasec_c} \dbase_b]{\includegraphics[width=1.2in]{union_k.png}%
   \label{fig_first_case}}

### 2021-11-10 paper/main.tex
sheaf artist
cat section

@@
                                                                   &                                                       &                                                       \\
-    U_1 \arrow[uu, "\sheaf_{\dtotal}"] \arrow[rr, "\iota"', hook] & {} \arrow[uu, "\sheaf_{\dtotal}" description, dotted] & U_2 \arrow[uu, "\sheaf_\dtotal"']                    
+    U_1 \arrow[uu, "\sheafc_{\dtotal}"] \arrow[rr, "\iota"', hook] & {} \arrow[uu, "\sheafc_{\dtotal}" description, dotted] & U_2 \arrow[uu, "\sheafc_\dtotal"']                    
     \end{tikzcd}
@@
 \begin{equation}
-  \vartist: \sheaf_{\dtotal} \rightarrow \sheaf_{\vartist}
+  \vartistc: \sheafc_{\dtotalc} \rightarrow \sheafc_{\vartistc}
   \label{eq:artist:artist}
@@
 {} \arrow[rr, "\dfuncc", Rightarrow, shift right, shorten=3em] & & {} \\
-    & \mathcal{\dbasec}^{op} \arrow[uu, "\sheaf_{\dtotalc}", bend left] \arrow[uu, "\sheaf_{\dtotalc}"', bend right] &   
+    & \mathcal{\dbasec}^{op} \arrow[uu, "\sheafc_{\dtotalc}", bend left] \arrow[uu, "\sheafc_{\dtotalc}"', bend right] &   
 \end{tikzcd}
@@
 Therefore, the range of the the artist functor, which we denote $Im_{\vartist}$, is a sheaf map $Im_{\vartist}$.
-Since the artist is a morphism of sheafs
-\begin{equation}
-  \vartist: \mathcal{\dtotal} \rightarrow Im_{\vartist}
-\end{equation}
-it is by definition a natural transformation\cite{SheafMathematics2021,bradleyWhatNaturalTransformation}.
+
+Since the artist is a morphism of sheafs it is by definition a natural transformation\cite{SheafMathematics2021,bradleyWhatNaturalTransformation}.
 This means that all morphisms $\varphi$ on $\mathcal{\dtotal}$ are expected to commute \cite{spanier1989algebraic,fongInvitationAppliedCategory2019}.
@@
   \begin{tikzcd}
-    \mathcal{\dtotal} \arrow[r, "\vartist"] \arrow[d, "\varphi \in {Hom_{\mathcal{\dfiber}}(\dfiber, \dfiber)}"'] & Im_{\vartist} \arrow[d, "\vartist_{\varphi}"] \\
-    \mathcal{\dtotal} \arrow[r, "\vartist"] & Im_{\vartist}                                                 
+    \sheafc_{\dtotalc} \arrow[r, "\vartistc"] \arrow[d, "\dfuncc_{\dtotalc}"'] & \sheafc_{\vartistc} \arrow[d, "\dfuncc_{\vartistc}"] \\
+    \sheafc_{\dtotalc} \arrow[r, "\vartistc"] & \sheafc_{\vartistc}                                              
   \end{tikzcd}

### 2021-11-11 paper/main.tex
fiber bundle intro

@@
 The \textcolor{base}{base space} \dbasec\ is a topological space, which means it is a set $\dbase$ with a collection of open sets \cite{TopologyVsTopology} surrounding each element in the set.
-Open sets are a collection of subsets in a set $X$, including the empty set, the whole set $X$, and every union and intersection of its member subsets.
-For each point $\dbasepoint in \dbase$, there is a collection of subsets $\{U\} \subset \dbase$ such that $x \in U \subseteq \dbase$ \cite{spanier1989algebraic, bradleyTopologyCategoricalApproach2020}
+Open sets are a collection of subsets $\{\opensetc\}$ in a set \dbase, including the empty set, the whole set \dbase, and every union and intersection of its member subsets.
+For each point $\dbasepoint in \dbase$, there is a collection of subsets $\\{\openset\} \subset \dbase$ such that $\dbasepoint \in \openset \subseteq \dbase$ \cite{spanier1989algebraic, bradleyTopologyCategoricalApproach2020}
 \end{definition}
@@
 
-\begin{definition} The \textcolor{fiber}{fiber space} \dfiberc\ is a topological space such that for every $\dbasepoint \in \dbase$, the fiber over that point is isomorphic to the preimage of that point $\dfiber_{\dbasepoint} \cong \pi^{-1}(\{k\})$.
-All fibers $\dfiber_{\dbasepoint}$ are homemorphic to each other.
-\cite{spanier1989algebraic, hatcherAlgebraicTopology2002}
+\begin{definition} The \textcolor{fiber}{fiber space} \dfiberc\ is a topological space such that for every $\dbasepoint \in \dbase$, the fiber over that point is isomorphic to the preimage of that point $\dfiber_{\dbasepoint} \cong \pi^{-1}(\{\dbasepoint\})$.
+All fibers $\dfiber_{\dbasepoint}$ are homeomorphic to each other \cite{spanier1989algebraic, hatcherAlgebraicTopology2002}
 \end{definition}
@@
 \begin{definition} The \textcolor{total}{total space} \dtotalc\ is the space reachable via the projection map $\pi$.
-The total space is always locally trivial, meaning $\dtotal = \dbase \times \dfiber$ over any open neighborhood $U_{k}$.
-The bundle is trivial when $\dtotal = \dbase \times \dfiber$ is true globally; a non-trivial bundle can be thought of as  the twisted product.
-\cite{hatcherAlgebraicTopology2002,nlab:twisting_function}
+The total space is always locally trivial, meaning $\dtotal = \dbase \times \dfiber$ over any open neighborhood $\openset_{\dbasepoint}$.
 \end{definition}
+The fiber bundle is trivial when $\dtotal = \dbase \times \dfiber$ is true globally; a non-trivial bundle can be thought of as the twisted fiber product $\dtotal=\dbase times_{T} \dfiber$\cite{hatcherAlgebraicTopology2002,nlab:twisting_function}.
+The twisting refers to the fiber \dfiber\ not aligning in the same direction over all $\dopenspace \subseteq \dbase$.
 
@@
 \subsubsection{Sheaves}
-The set of all sections sections $\{\dsection^{i}: \U_{j}\rightarrow \dtotal\}|_{i\in I}$ over an arbitrary open set $U_{j}$ is denoted $\Gamma(U_{\dbasepoint}, \dtotal)$.
+The set of all sections sections $\{\dsection^{i}: \openset_{j}\rightarrow \dtotal\}|_{i\in I}$ over an arbitrary open set $U_{j}$ is denoted $\Gamma(\openset_{\dbasepoint}, \dtotal)$.
 The set of all sections over the total space \dbase\ is denoted $\Gamma(\dbase, \dtotal)$.
 
-The map from the open set to the set of sections is called a presheaf $\sheaf: U_j \rightarrow \Gamma(U_j, \dtotal)$ \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}.
-The presheaf preserves inclusion maps between the open sets and the sets of sections
+The map from the open set to the set of sections is called a presheaf $\sheaf: U_j \rightarrow \Gamma(\openset_j, \dtotal)$ \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}.
+The presheaf preserves inclusion maps between the open sets open sets $\iota$ and the inclusion map $\iota^*$ between the sets of sections.
 
@@
   \begin{tikzcd}[column sep=small]
-    {\Gamma(U_1, \dtotalc)}                                       & {}                                                    & {\Gamma(U_2, \dtotalc)} \arrow[ll, "\iota^*"', hook'] \\
+    {\Gamma(\opensetc_1, \dtotalc)}                                       & {}                                                    & {\Gamma(\opensetc_2, \dtotalc)} \arrow[ll, "\iota^*"', hook'] \\
                                                                   &                                                       &                                                       \\
-    U_1 \arrow[uu, "\sheafc_{\dtotal}"] \arrow[rr, "\iota"', hook] & {} \arrow[uu, "\sheafc_{\dtotal}" description, dotted] & U_2 \arrow[uu, "\sheafc_\dtotal"']                    
+    \opensetc_1 \arrow[uu, "\sheafc_{\dtotal}"] \arrow[rr, "\iota"', hook] & {} \arrow[uu, "\sheafc_{\dtotal}" description, dotted] & \opensetc_2 \arrow[uu, "\sheafc_\dtotal"']                    
     \end{tikzcd}
@@
 
-meaning that the presheaf ensures that given an inclusion map between open sets $\iota$, there is a corresponding map $\iota^*$ between the sets of sections.
-This means that a smaller space $U_1$ is included in a larger space $U_2$ and a function that is continuous over a larger space $U_2$ is continuous over a subspace $U_1 \subspace U_2$.
-Sheaves are presheafs where sets of sections $\Gamma$ are defined over unions of open sets $U \cup_{U\in \mathscr{U}}$ \cite{spanier1989algebraic, nlab:sheaf}.
-Because they preserve inclusion maps, sheaves are often used as an abstraction for keeping track of how data over topological spaces is glued together \cite{ghristElementaryAppliedTopology2014}.
+This means that a smaller space $\openset_1$ is included in a larger space $\openset_2$ and a function that is continuous over a larger space $\openset_2$ is continuous over a subspace $\openset_1 \subspace \openset_2$.
+Sheaves are presheafs where sets of sections $\Gamma$ are defined over unions of open sets over a topology $\bigcup_{j \in I} \openset)_j \in \dtotal$ \cite{spanier1989algebraic, nlab:sheaf}.
+Sheaves are often used as an abstraction for keeping track of how data over topological spaces is glued together \cite{ghristElementaryAppliedTopology2014} because they model the data as sets of sections, continuity of the data in the base, and how subsets fit together through inclusion maps.
 
-Here we present a very brief summary of fiber bundles and sheaves, for more information see Hatcher\cite{hatcherAlgebraicTopology2002}, Munkres\cite{munkresElementsAlgebraicTopology1984}, Spanier \cite{spanier1989algebraic} and Ghrist\cite{ghristElementaryAppliedTopology2014,ghristHomologicalAlgebraData2018}.
+\note{I've seen this convention before, but I also cite all these people above}
+For more information on fiber bundles and sheaves, see Hatcher\cite{hatcherAlgebraicTopology2002}, Munkres\cite{munkresElementsAlgebraicTopology1984}, Spanier \cite{spanier1989algebraic} and Ghrist\cite{ghristElementaryAppliedTopology2014,ghristHomologicalAlgebraData2018}.
 
@@
 Vickers et.
-al provide a brief introduction to category theory for visualization practitioners \cite{vickersUnderstandingVisualizationFormal2013}, but their work focuses on data, representation, and evocation, while this paper is aims to provide guidance on how the map from data to representation should be implemented.
+al provide a brief introduction to category theory for visualization practitioners \cite{vickersUnderstandingVisualizationFormal2013}, but their work focuses on visualization deisgn, while this paper aims to provide guidance on designing visualization library components.
+\begin{definition} A category is a 
+\end{definition}
 

### 2021-11-11 paper/main.tex
sheaves

@@
 \begin{equation}
-  \label{eq:artist:presheaf}
+  \label{eq:related-work:continuity:presheaf}
   \begin{tikzcd}[column sep=small]
@@
 Vickers et.
-al provide a brief introduction to category theory for visualization practitioners \cite{vickersUnderstandingVisualizationFormal2013}, but their work focuses on visualization deisgn, while this paper aims to provide guidance on designing visualization library components.
-\begin{definition} A category is a 
+al provide a brief introduction to category theory for visualization practitioners \cite{vickersUnderstandingVisualizationFormal2013}, but their work focuses on reasoning about visualization design, while this paper aims to provide guidance on designing visualization library components.
+\begin{definition} A category \mathcal{C} consists of a collection of objects $C$ with identity $id_c: C\rightarrow C$ and the set of morphisms between every two objects $f:C_1 \rightarrow C_2$, \cite{fongInvitationAppliedCategory2019,maclaneCategoriesWorkingMathematician2013}.
 \end{definition}
+The set of morphisms between objects is termed the hom set $hom_{\mathcal{C}}(C_1, C_2)$. The morphisms on the category  compose
+\begin{equation*}
+  \begin{tikzcd}
+    C_1 \arrow[r, "f"] \arrow[rd, "g\circ f"'] \arrow["id_{C_1}"', loop, distance=2em, in=215, out=145] & C_2 \arrow[d, "g"] \arrow["id_{C_2}"', loop, distance=2em, in=125, out=55] \\
+  & C_3 \arrow["id_{C_3}"', loop, distance=2em, in=305, out=235]              
+  \end{tikzcd}
+\end{equation*}
+ and are constructed such that the following axioms hold \cite{riehlCategoryTheoryContext}:
+ \begin{LaTeXdescription}
+   \item[associativity] if $f: C_1 \rightarrow C_2$, $g: C_2 \rightarrow C_3$ and $h: C_3 \rightarrow C_4$ then $h\circ (g \circ f) = (h \circ g) \circ f$
+   \item[identity] for every $f: C_1 \rightarrow C_2$ there exists identity morphisms $f \circ id_{C_1} = f = id_{C_2} \circ f$
+ \end{LaTeXdescription}
+
+ \begin{definition} An opposite category $\mathcal{C}^{op}$ is a category with all the same objects of category $\mathcal{C}$ and but the morphisms are reversed.
+For example $f:C_1 \rightarrow C_2$ in $\mathcal{C}$ is $f:C_2 \rightarrow C_1$ in $\mathcal{C}^{op}$.
+ \end{definition}
+
+ \begin{definition} A functor is a morphism between categories $F: \mathcal{C} \rightarrow \mathcal{D}$.
+A functor has the properties of identity and composition \cite{riehlCategoryTheoryContext}\end{definition}
+A functor preserves the structure of the category, namely morphisms and the source and target objects of those morphisms, composition of morphisms, and identity morphisms \cite{riehlCategoryTheoryContext}.
+Contravariant functors are functors where the morphisms in $\mathcal{C}$ go in the opposite direction from the morphisms in $\mathcal{D}$.
+
+ \begin{definition} A presheaf \sheafc\, as introduced in \autoref{eq:related-work:continuity:presheaf}, is generally a functor $\sheaf: \mathcal{C}^{op} \rightarrow \textbf{Set}$.\cite{nlab:presheaf} \end{definition}
+
+ \begin{definition} A natural transformation is a morphism of functors $\alpha: F \rightarrow G$ where $F, G$ are functors from $\mathcal{C}$ to $\mathcal{D}$.
+\cite{riehlCategoryTheoryContext, bradleyWhatNaturalTransformation}
+ \end{definition}
+The natural transformation acts as a wrapper of sorts
+\begin{equation}
+  \label{eq:related-work:equivariance:natural-transform}
+\begin{tikzcd}
+  & {} \arrow[dd, "\alpha", Rightarrow] &             \\
+\mathcal{C} \arrow[rr, "F", bend left] \arrow[rr, "G"', bend right] &                                     & \mathcal{D} \\
+  & {}                                  &            
+\end{tikzcd}
+\end{equation}
+such that $\alpha(F: C \rightarrow D) = G: C \rightarrow D$ commutes for any morphism $f:c\rightarrow c^{\prime}$ \cite{fongInvitationAppliedCategory2019}.
+Arbitrary morphisms on sheaves are natural transforms \cite{SheafMathematics2021,bradleyWhatNaturalTransformation}.
 
@@
 In this section, we use category theory to formally express the implicit assumptions that visualization library components make in transforming data into graphical representations.
-We propose that the visualization transformation can be modeled as a functor, which we call the \textcolor{artist}{Artist} $\mathcal{\vartist}$\footnote{We call this transformation the artist because the \mintinline{python}{Artist} object in Matplotlib\cite{hunterMatplotlib2DGraphics2007}}.
-The artist \vartist\ converts data, which we model as the sheaf $\sheaf_{\dtotal}$ to graphics, which we model as the sheaf $\sheaf_{\vartist}$.
+We propose that the visualization transformation can be modeled as a functor, which we call the \textcolor{artist}{Artist} \vartist \footnote{We call this transformation the artist because the \mintinline{python}{Artist} object in Matplotlib\cite{hunterMatplotlib2DGraphics2007}}.
+The artist \vartist\ converts data, which we model as the sheaf $\sheaf_{\dtotal}$ to graphics in the sheaf $\sheaf_{\gtotal}$.
 We formulate this association as
@@
 
-Arbitrary morphisms on sheafs are natural transforms \cite{SheafMathematics2021,bradleyWhatNaturalTransformation}, meaning that the morphisms $\varphi$ compose with the inclusion maps $\iota$ \cite{fongInvitationAppliedCategory2019}.
+.meaning that the morphisms $\varphi$ compose with the inclusion maps $\iota$ \cite{fongInvitationAppliedCategory2019}.
 As defined in \autoref{def:category:Tau}, the sets of sections $\Gamma(U_{i},\dtotal)$ are objects of the category $\mathcal{T}$.

### 2021-11-11 paper/main.tex
fiber and sheaf rewrite

@@
 Vickers et.
-al provide a brief introduction to category theory for visualization practitioners \cite{vickersUnderstandingVisualizationFormal2013}, but their work focuses on reasoning about visualization design, while this paper aims to provide guidance on designing visualization library components.
+al \cite{vickersUnderstandingVisualizationFormal2013} provide a brief introduction to category theory for visualization practitioners, but their work focuses on reasoning about visualization design, while this paper aims to provide guidance on designing visualization library components.
+Briefly we introduce concepts in category theory that we use to describe our model in \autoref{sec:artist:category} and \autoref{sec:artist:construction}.
+
 \begin{definition} A category \mathcal{C} consists of a collection of objects $C$ with identity $id_c: C\rightarrow C$ and the set of morphisms between every two objects $f:C_1 \rightarrow C_2$, \cite{fongInvitationAppliedCategory2019,maclaneCategoriesWorkingMathematician2013}.
@@
 
- \begin{definition} A presheaf \sheafc\, as introduced in \autoref{eq:related-work:continuity:presheaf}, is generally a functor $\sheaf: \mathcal{C}^{op} \rightarrow \textbf{Set}$.\cite{nlab:presheaf} \end{definition}
+ \begin{definition} A presheaf \sheafc\, as introduced in \autoref{eq:related-work:continuity:presheaf}, is a functor $\sheaf: \mathcal{C}^{op} \rightarrow \textbf{Set}$.\cite{nlab:presheaf} \end{definition}
 
@@
 \begin{tikzcd}
-  & {} \arrow[dd, "\alpha", Rightarrow] &             \\
+  & {} \arrow[dd, "\alpha", Rightarrow, shorten=1.5em] &             \\
 \mathcal{C} \arrow[rr, "F", bend left] \arrow[rr, "G"', bend right] &                                     & \mathcal{D} \\
@@
 \end{equation}
-such that $\alpha(F: C \rightarrow D) = G: C \rightarrow D$ commutes for any morphism $f:c\rightarrow c^{\prime}$ \cite{fongInvitationAppliedCategory2019}.
+such that $\alpha(F: \mathcal{C} \rightarrow \mathcal{D}) = G: \mathcal{C} \rightarrow \mathcal{D}$ commutes for any morphism $f:c\rightarrow c^{\prime}$ \cite{fongInvitationAppliedCategory2019}.
 Arbitrary morphisms on sheaves are natural transforms \cite{SheafMathematics2021,bradleyWhatNaturalTransformation}.
@@
 
-As introduced in \autoref{sec:related-work:fiber-bundles}, sets of open sets form the basis of topological spaces.
-We propose that we can express how the elements in our data are connected to each other by modeling the data continuity as a topological space \dbase.
-We express subsets of the continuity, which can be mapped into subsets of the data, as open sets $U_{i}$ in \dbase.
+Fiber bundles, as introduced in \autoref{sec:related-work:fiber-bundles}, serve as the model of data and graphic spaces in out model.
+In this section, we introduce categorical formulations of the topological spaces that make up a fiber bundle.
+
 We introduce the category $\mathcal{\dbase}$ to encapsulate the subsets of the continuity and the inclusion maps that glue them together.
 
-\begin{definition} The category of open sets $\mathcal{\dbase}$ consists of
-  \label{def:category:K} 
+\begin{definition} \label{def:category:K} The category of open sets $\mathcal{\dbasec}$ consists of
   \begin{itemize}
-    \item \textit{objects} all open sets $U_{j} \in \{U\}$ in the topological space $(\dbase, \mathscr{T})$), including the empty set $\varnothing$ and the maximum set \dbase. 
-    \item \textit{morphisms} inclusion maps from a subspace to a larger space $\iota: U_{j} \hookrightarrow U_{j}$, which is equivalent to $U_{j} \subseteq U_{j}$ and  
+    \item \textit{objects} open sets $\openset_{j} \in \{openset\}$ in the topological space $(\dbase, \mathscr{T})$), including the empty set $\varnothing$ and the maximum set \dbase. 
+    \item \textit{morphisms} $\iota: U_{j} \hookrightarrow U_{k}$ for all  $U_j, U_k \subset \dtotal$
   \end{itemize} 
@@
 
-The inclusion maps $\iota$ can be used to piece together the subsets $U_{i}$, particularly individual points $\dbasepoint \in \dbase$.
-This is how we provide knowledge of continuity to the artist \vartist\ in a way where the artist does not need all the data at once.
-This is what allows the artist to generate graphics where the subset of data on view is dynamically updated, such as pan and zoom navigation\cite{NekrasovskiEvaluationPanZoom2006} and sliding windows on streaming data\cite{crouchDynamicGraphsSlidingwindow2013,chuTimeSeriesSegmentation1995}.
-
-Besides expressing continuities, fiber bundles also provide us a way to richly express the structure of the fields in the data.
-Specifically, we do this by adopting Spivak's formulation of the fiber as a (column name, data domain) simple schema \cite{spivakSIMPLICIALDATABASES,spivakDatabasesAreCategories2010}.
-Spivak formally maps column names and field types to the set of values associated with the field type, for example $\mathbb{R}$ for a \texttt{float} column named \textit{temperature}.
-We define $\dfiber$ as  the cartesian cross product of the fibers $\dfiber_c$ for each field, where $c$ is the field name.
-
-\begin{definition} 
-  \label{def:category:F} 
-  The category $\mathcal{\dfiber}$ consists of a single object $\dfiber$ and the morphisms between the object and itself $Hom_{\mathcal{\dfiber}}(\dfiber, \dfiber)$, which is the hom-set \cite{maclaneCategoriesWorkingMathematician2013,milewskiCategoryTheoryProgrammers}.
+\begin{definition} \label{def:category:E} The category \mathcal{\dtotalc} is a subcategory of \textbf{Set} and consists of 
+  \begin{itemize}
+    \item \textit{objects} sets of sections $\Gamma(\openset_{j}, \dtotal)$ for all $U_j \subset \dtotal$
+    \item \textit{morphisms}  $\iota^*: \Gamma(\openset_{k}, \dtotal) \hookrightarrow \Gamma(\openset_{j}, \dtotal)$ for all  $\openset_j, \openset_k \subset \dtotal$
+  \end{itemize} 
 \end{definition}
 
-The object $\texttt{\dfiber}$ can be an object of any category $\mathcal{C}$, allowing $\texttt{\dfiber}$ to encode most data fields.
-For example, a lists of strings is an instance of an object in \textbf{Set}, networks are an instance of \textbf{Graph}, and images are vector spaces which are a specific type of topological space \textbf{Top}.
-We note that a one object category is a monoid \cite{nlab:monoid,barrCategoryTheoryComputing}, which is an algebraic structure that lends itself to computer library design because they are an abstraction of function composition \cite{yorgeyMonoidsThemeVariations}.
-
-The morphisms $Hom_{\mathcal{\dfiber}}(\dfiber, \dfiber)$ are functions from the data to itself, for example the binary operations, group actions, and measurement scales discussed in \autoref{sec:related-work:equivariance}.
-These functions could also be monoid actions\cite{barrCategoryTheoryComputing}, which provide a way of applying partial order relations to data\cite{fongInvitationAppliedCategory2019}, such as to build multi-ranked indicators\cite{bruggemannRankingPrioritizationMultiindicator2011}
-
-We encode individual datasets in the fiberbundle as sections of the bundle.
-We follow the convention that $\Gamma(U_i, \dtotal)$ denotes the set of sections $\{\dsection^{j}:U_i \rightarrow \dfiber_{U_{i}}\}_{j\in I}$ of the fiber bundle \dtotal\ over an open set $U_{i}\subseteq \dbase$.
-
-\begin{definition} The category \mathcal{\dtotalc} is the set of sets of sections $\{\Gamma(U_j,\dtotal)\}|U_j \in \mathcal{dbase}$.
-The category \mathcal{\dtotal} is a subcatagory of the category of sets \textbf{Set}.
-The morphisms between objects of $\mathcal{\dtotal}$ are restriction maps $\iota: \Gamma(U_j, \dtotal) \rightarrow \Gamma(U_i, \dtotal)$ where $U_{i} \subseteq U_{j}$.
-\end{definition}
-
-Each section is how we represent a unique dataset, where the sets of sections $\Gamma$ is the space of all datasets that have the same data fields in the \dfiber\ and continuity as encoded in\dbase\.
-For example, Spivak describes tables that have the same schema and discrete index as sections of the same fiber bundle\cite{spivakSIMPLICIALDATABASES}.
-
-.meaning that the morphisms $\varphi$ compose with the inclusion maps $\iota$ \cite{fongInvitationAppliedCategory2019}.
-As defined in \autoref{def:category:Tau}, the sets of sections $\Gamma(U_{i},\dtotal)$ are objects of the category $\mathcal{T}$.
-
+We model data as the sheaf $\sheaf_{\dtotal}: \mathcal{\dbase}^{op}\rightarrow \mathcal{\dtotal}$
 \begin{equation}
@@
 \end{equation}
+where the map \dfunc\ is any arbitrary function between sheafs.
+The sheaf $\sheaf_{\dtotal}$ over a limit of open sets that contain a point $\dbasepoint \in \dbase$ is approximately the same as a sheaf over a point \dbasepoint\ and is called the stalk $\sheaf_{\mathcal{\dtotal}}|_{k}\coloneqq \lim\limits_{U\ni k} \Gamma(U, \dtotal)$ \cite{StalkSheaf2019}.
 
-The sheaf $\sheaf_{\dtotal}$ over a limit of open sets that contain a point $\dbasepoint \in \dbase$ is approximately the same as a sheaf over a point \dbasepoint\ and is called the stalk $\sheaf_{\mathcal{\dtotal}}|_{k}\coloneqq \lim\limits_{U\ni k} \Gamma(U, \dtotal)$ \cite{StalkSheaf2019}.
+\begin{definition} 
+  \label{def:category:F} 
+  The fiber category $\mathcal{\dfiber}$ consists of 
+  \begin{itemize}
+    \item \textit{object} a single object \dtotal, where \dtotal\ can be an object of any arbitrary category $\mathcal{C}$ 
+    \item \textit{morphisms} $\dfunc \in hom_{\mathcal{F}}(F,F)$
+\end{definition}
+
 The fiber space over a point is embedded in the stalk over that point $\dfiber_{k}\hookrightarrow \mathscr{F}_{k}$.
-Our criteria for equivariance is that bundle maps $f$ that are in the set of fiber to fiber maps $Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ are preserved.
+The sheaf maps \dfunc\ restricted to the stalk are members of $hom_{\mathcal{F}}(F,F)$.
 The maps $\varphi \in Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ include the types of data transformations described in \autoref{sec:related-work:equivariance}.
-When a fiber bundle is trivial, meaning $\dtotal=\dbase \times \dfiber$, then we can dispense with localization and directly consider the fiber maps $Hom(\dfiber_{k}, \dfiber_{k})$ over a point \dbasepoint.
+When a fiber bundle is trivial, we can dispense with localization and directly consider the fiber maps $Hom(\dfiber_{k}, \dfiber_{k})$ over a point \dbasepoint.
 In \autoref{sec:artist:construction}, we assume that the fiber bundles are trivial.

### 2021-11-11 paper/main.tex
product categories

@@
 
-We model data as the sheaf $\sheaf_{\dtotal}: \mathcal{\dbase}^{op}\rightarrow \mathcal{\dtotal}$
+We model data as the sheaf $\sheafc_{\dtotalc}: \mathcal{\dbasec}^{op}\rightarrow \mathcal{\dtotalc}$
 \begin{equation}
@@
   \label{def:category:F} 
-  The fiber category $\mathcal{\dfiber}$ consists of 
+  The fiber category $\mathcal{\dfiberc}$ is a product category\cite{maclaneCategoriesWorkingMathematician2013,nlab:product_category} of zero or more arbitrary categories.
+In the two category case, the category $\mathcal{\dfiberc}$ consists of
   \begin{itemize}
-    \item \textit{object} a single object \dtotal, where \dtotal\ can be an object of any arbitrary category $\mathcal{C}$ 
-    \item \textit{morphisms} $\dfunc \in hom_{\mathcal{F}}(F,F)$
+    \item \textit{objects} ordered pairs $(x,y)$,  $x$ is an object of $\mathcal{X}$ and  $y$ is an object of $\mathcal{Y}$
+    \item \textit{morphisms} ordered pairs $(f:x\rightarrow x^{\prime}, g:y\rightarrow y^{\prime})$
+    \end{itemize}
+The coomposition of morphisms is defined component wise by composition in $\mathcal{X}$ and $\mathcal{Y}$.ss
 \end{definition}
 
-The fiber space over a point is embedded in the stalk over that point $\dfiber_{k}\hookrightarrow \mathscr{F}_{k}$.
-The sheaf maps \dfunc\ restricted to the stalk are members of $hom_{\mathcal{F}}(F,F)$.
+The fiber space over a point is embedded in the stalk over that point $\dfiber_{\dbasepoint}\hookrightarrow \mathscr{\dfiber}_{\dbasepoint}$.
+The sheaf maps \dfunc\ restricted to the stalk are members of $hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$.
 The maps $\varphi \in Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ include the types of data transformations described in \autoref{sec:related-work:equivariance}.
-When a fiber bundle is trivial, we can dispense with localization and directly consider the fiber maps $Hom(\dfiber_{k}, \dfiber_{k})$ over a point \dbasepoint.
+When a fiber bundle is trivial, we can dispense with localization and directly consider the fiber maps $Hom(\dfiber_{\dbasepoint}, \dfiber_{\dbasepoint})$ over a point \dbasepoint.
 In \autoref{sec:artist:construction}, we assume that the fiber bundles are trivial.
@@
 \note{rewrite paragraph}
-The graphic category $\mathcal{\gtotalc}$ consists of the functor object $\sheaf_{\gtotalc}: \mathcal{\gbasec}^{op} \rightarrow \Gamma(\mathcal{\gbasec}, \gtotal)$ and morphisms between sheaves.
+
+The domain of the artist function is the space of all possible graphics. We model this space using categories that are analogous to the data categories
+
+We model the space of all possible graphics as a sheaf $\sheafc_{\mathcal{H}}:\mathcal{S} \rightarrow \mathcal{H}$
+
+The graphic category $\mathcal{\gtotalc}$ is analogous to the $\
+
+consists of the functor object $\sheaf_{\gtotalc}: \mathcal{\gbasec}^{op} \rightarrow \Gamma(\mathcal{\gbasec}, \gtotal)$ and morphisms between sheaves.
 The category $\mathcal{\gbasec}$ is the category of open sets $\{U\}$ over the graphic base space \gbase.

### 2021-11-11 paper/main.tex
equivariance w/ tie back to rw

@@
 \begin{figure}[!h]
+  \label{fig:related-work:equivariance}
   \includegraphics[width=\columnwidth]{equiv.png}
@@
 
-\note{rewrite paragraph}
+The domain of the artist function is the space of all possible graphics. As with the domain $\sheaf_{E}$, the grapic space is modeled with fiber bundle equivalent categories:
+\begin{table}[!h]
+  \centering
+  \label{tab:artist:graphic}
+  \begin{tabular}{l l l}
+    fiber bundle & data & graphic \\
+    \textcolor{base}{base space} & \mathcal{\dbasec} & \mathcal{\gbasec}\\
+    \textcolor{total}{total space} & \mathcal{\dtotalc} & \mathcal{\gtotalc}\\
+    \textcolor{fiber}{fiber space} & \mathcal{\dfiberc} & \mathcal{\gfiberc}\\
+    \pi^{-1} & \sheaf_{\mathcal{\dtotalc}} & \sheaf_{\mathcal{\gtotalc}}\\
+  \end{tabular}
+\end{table}
 
-The domain of the artist function is the space of all possible graphics. We model this space using categories that are analogous to the data categories
-
-We model the space of all possible graphics as a sheaf $\sheafc_{\mathcal{H}}:\mathcal{S} \rightarrow \mathcal{H}$
-
-The graphic category $\mathcal{\gtotalc}$ is analogous to the $\
-
-consists of the functor object $\sheaf_{\gtotalc}: \mathcal{\gbasec}^{op} \rightarrow \Gamma(\mathcal{\gbasec}, \gtotal)$ and morphisms between sheaves.
-The category $\mathcal{\gbasec}$ is the category of open sets $\{U\}$ over the graphic base space \gbase.
-The set of sets of sections $\Gamma(\mathcal{\gbase}, \gtotal)$ is a subcatagory of \textbf{Set}.
-
-\begin{definition} The sheaf $Im_{\vartist}$ is a map between the category of open sets $\gtotal^{op}$ and a subcategory of \textbf{Set}.
-This subcategory is the set of subsets of sections $\{\gsection^{k}:U_i \rightarrow \gfiber_{U_{i}}\} \subset \Gamma(U_i, \gtotal)$ over the space of all open sets in $\mathcal{\gbase}$.
-\end{definition}
-
-A set $\Gamma(U_i, \gtotal)$ contains functions to generate arbitrary graphics with a shared continuity and display space, while the artist $\vartist$ is constructed such that it generates a single type of graphic, as described in \autoref{sec:artist:construction}.
-Therefore, the range of the the artist functor, which we denote $Im_{\vartist}$, is a sheaf map $Im_{\vartist}$.
-
-Since the artist is a morphism of sheafs it is by definition a natural transformation\cite{SheafMathematics2021,bradleyWhatNaturalTransformation}.
-This means that all morphisms $\varphi$ on $\mathcal{\dtotal}$ are expected to commute \cite{spanier1989algebraic,fongInvitationAppliedCategory2019}.
-This is the basis of the equivariance condition illustrated in \autoref{eq:artist:equivariance}
-
+The category $\mathcal{H}$ has as objects the sets of functions that generate every arbitrary graphic over the open sets in $\mathcal{\gtotal}$.
+While $\mathcal{H}$ is the domain of the artist \vartist, the range of \vartist\ is the subset of graphics $\sheaf_{\vartist} \subset \sheaf_{\mathcal{\dtotal}}$ such that structure imposed by $\dfunc_{\dtotal}$ commutes
 \begin{equation}
@@
 \end{equation}
-which states that transforms on the data side $\varphi \in Hom_{\mathcal{\dfiber}}(\dfiber, \dfiber)$ have an approximately equivalent transform in the data side $\vartist_{\varphi \in Hom_{\mathcal{\dfiber}}}$.
-This commutativity property states that visualization library components must be implemented such that changes in data induce equivalent changes in the output, which is how visualization libraries can produce visualizations that satisfy Kindlmann and Scheidegger's correspondence principal that data and visual structure must match.
+The equivariance condition expressed in \label{eq:artist:equivariance} is that a graphic generated by  transforming and then visualizing data $\vartist(\dfunc_{\dtotal}(\sheaf_{\mathcal{dtotal}}))$ can equivalently be generated by visualizing the data and then transforming the graphic $\dfunc_{\vartist}(\vartist(\sheaf_{\mathcal{\dtotal}}))$.
+This expectation of equivariance is illustrated in \autoref{fig:related-work:equivariance}
 

### 2021-11-11 paper/main.tex
reworked defining artist section

@@
 Open sets are a collection of subsets $\{\opensetc\}$ in a set \dbase, including the empty set, the whole set \dbase, and every union and intersection of its member subsets.
-For each point $\dbasepoint in \dbase$, there is a collection of subsets $\\{\openset\} \subset \dbase$ such that $\dbasepoint \in \openset \subseteq \dbase$ \cite{spanier1989algebraic, bradleyTopologyCategoricalApproach2020}
+For each point $\dbasepoint \in \dbase$, there is a collection of subsets $\\{\openset\} \subset \dbase$ such that $\dbasepoint \in \openset \subseteq \dbase$ \cite{spanier1989algebraic, bradleyTopologyCategoricalApproach2020}
 \end{definition}
@@
 
+\note{the notation here may need to be X and Y}
+
 \begin{definition} A category \mathcal{C} consists of a collection of objects $C$ with identity $id_c: C\rightarrow C$ and the set of morphisms between every two objects $f:C_1 \rightarrow C_2$, \cite{fongInvitationAppliedCategory2019,maclaneCategoriesWorkingMathematician2013}.
@@
 
- \begin{definition} A functor is a morphism between categories $F: \mathcal{C} \rightarrow \mathcal{D}$.
+ \begin{definition} A functor is a morphism between categories $\mathfrak{F}: \mathcal{C} \rightarrow \mathcal{D}$.
 A functor has the properties of identity and composition \cite{riehlCategoryTheoryContext}\end{definition}
@@
 We propose that the visualization transformation can be modeled as a functor, which we call the \textcolor{artist}{Artist} \vartist \footnote{We call this transformation the artist because the \mintinline{python}{Artist} object in Matplotlib\cite{hunterMatplotlib2DGraphics2007}}.
-The artist \vartist\ converts data, which we model as the sheaf $\sheaf_{\dtotal}$ to graphics in the sheaf $\sheaf_{\gtotal}$.
+The artist \vartist\ converts data, which we model as the sheaf $\sheafc_{\dtotalc}$ to graphics in the sheaf $\sheafc_{\gtotalc}$.
 We formulate this association as
@@
 \begin{equation}
-  \vartistc: \sheafc_{\dtotalc} \rightarrow \sheafc_{\vartistc}
   \label{eq:artist:artist}
+  \vartistc: \sheafc_{\dtotalc} \rightarrow \sheafc_{\gtotal}
 \end{equation}
@@
 \end{equation}
-The equivariance condition expressed in \label{eq:artist:equivariance} is that a graphic generated by  transforming and then visualizing data $\vartist(\dfunc_{\dtotal}(\sheaf_{\mathcal{dtotal}}))$ can equivalently be generated by visualizing the data and then transforming the graphic $\dfunc_{\vartist}(\vartist(\sheaf_{\mathcal{\dtotal}}))$.
-This expectation of equivariance is illustrated in \autoref{fig:related-work:equivariance}
+The equivariance condition expressed in \autoref{eq:artist:equivariance} is that a graphic generated by  transforming and then visualizing data $\vartist(\dfunc_{\dtotal}(\sheaf_{\mathcal{dtotal}}))$ can equivalently be generated by visualizing the data and then transforming the graphic $\dfunc_{\vartist}(\vartist(\sheaf_{\mathcal{\dtotal}}))$.
+This is a formal statement of the equivariance illustrated in \autoref{fig:related-work:equivariance}
 
@@
 
-We can compose these diagrams together to generate an \vartist\ to handle any of the things.
+\autoref{eq:artist:intersection:basespace} and \autoref{eq:artist:intersection:fiber} can be composed to specify how different aspects of the structure of a multivariate nested dimensional dataset, such as a spatio-temporal weather dataset, should be preserved in a visualization.
+\note{this is probably the power of this model and should maybe get a figure?}
 
@@
 
-\begin{definition} Following from \label{def:artist:} the artist \vartist is a special case of  natural transformations
+Following from \autoref{eq:artist:equivariance}, the artist \vartistc\ is a natural transformation
+  \begin{equation}
+    \vartist: \sheaf_{\mathcal{\dtotal}}
+  \end{equation}
+that we construct to preserve continuity and equivariance.
 
-  %% this equation is actually on sheafs and is not a map 
-  \begin{equation}
-    \vartist: \Gamma(\dtotal, \dbase)\nrightarrow \Gamma(\gtotal, \gbase)
-  \end{equation}
-  between the profunctor objects $\Gamma(E),\Gamma(H)$ that preserves continuity and equivariance.
-\end{definition}
-  Natural transformations are functions that map functions between categories (functors) to other functors in a structure preserving way \cite{riehlCategoryTheoryContext, spanier1989algebraic, fongInvitationAppliedCategory2019}.
-
-\begin{definition} We define the natural transformation \textcolor{artist}{Artist}  \vartist\ as the tuple $(\vindex, \vchannel, \vmark, \mathcal{\dtotal},\mathcal{\vtotal}, \mathcal{\gtotal})$ where
+\begin{definition} We define the natural transformation \textcolor{artist}{Artist}  \vartist\ as the tuple $(\vindexc, \vchannelc, \vmarkc, \mathcal{\dtotalc},\mathcal{\vtotalc}, \mathcal{\gtotalc})$ where
   \begin{enumerate}
-    \item \vindex\ is a continuity preserving functor
-    \item \vchannel\ and \vmark\ are equivariance preserving functors
-    \item $\mathcal{\dtotal}$, $\mathcal{\vtotal}$, and $\mathcal{\gtotal}$ are profunctor categories of sheafs
-    \item 
+    \item $\vchannelc: \dtotalc \rightarrow \vtotalc$ is a bundle map from data values to the visual variables they are mapped to
+    \item $\vmarkc: \sheafc_{\mathcal{\vtotalc} \rightarrow \sheafc_{\mathcal{\vartistc}}$ is a sheaf map that builds a graphic generating function parameterized by the visual variables 
+    \item $\vindexc: \gbasec \rightarrow \dbasec$ is a surjective map from the graphic topological base to the data topological base
   \end{enumerate}
 \end{definition}
+and $\mathcal{\dtotalc}$, $\mathcal{\vtotalc}$, and $\mathcal{\gtotalc}$ are the categorical representations of fiber bundles that model the data, visual variable, and graphic space of the data to visual transformation.
 
-%%% no arrow V->H
-%%% Q is defined at the level of Gamma\begin{equation}
+\note{figure out how to format notes} Usage of the bundle directly in an equation or diagram - $\dotal, \vtotal, \gtotal$ - denotes that the function can be evaluated pointwise over $\dbasepoint \in \dbase$.
+Use of the categorical form -- $\mathcal{\dtotalc}, \mathcal{\vtotalc}, \mathcal{\gtotalc}$ -- denotes that the function is evaluated over an openset object, meaning that the function needs knowledge of both a value over a point and some information about neighboring values.
+
 \begin{equation}
@@
   \begin{tikzcd}
-    \dtotalc \arrow[rr, "\vchannel"] \arrow[rdd, "\pi"] &         & \dtotalc \arrow[ldd, "\pi"'] & \mathcal{\vtotalc}                                                        & Im_{\vmarkc} \subset \gtotalc                     \\
+    \dtotalc \arrow[rr, "\vchannel"] \arrow[rdd, "\pi"'] &         & \dtotalc \arrow[ldd, "\pi"] & \mathcal{\vtotalc}                                                        & Im_{\vmarkc} \subset \gtotalc                     \\
   &         &                             & {} \arrow[r, "\vmarkc"]                                                   & {}                                \\
@@
 
-\subsection{Data}
+\subsection{Data Domain}
 We model data as sections of a fiber bundle $(\dtotal, \pi, \dbase, \dfiber)$.
@@
 
-\subsection{Graphic} 
+\subsection{Graphic Codomain} 
 The object of the graphic category $\mathcal{\gtotal}$ is the fiber bundle \gtotal. The bundle \gtotal\ has the same structure as the data bundle \dtotal

### 2021-11-11 paper/main.tex
move a note to be part of a paragraph

@@
 and $\mathcal{\dtotalc}$, $\mathcal{\vtotalc}$, and $\mathcal{\gtotalc}$ are the categorical representations of fiber bundles that model the data, visual variable, and graphic space of the data to visual transformation.
-
-\note{figure out how to format notes} Usage of the bundle directly in an equation or diagram - $\dotal, \vtotal, \gtotal$ - denotes that the function can be evaluated pointwise over $\dbasepoint \in \dbase$.
-Use of the categorical form -- $\mathcal{\dtotalc}, \mathcal{\vtotalc}, \mathcal{\gtotalc}$ -- denotes that the function is evaluated over an openset object, meaning that the function needs knowledge of both a value over a point and some information about neighboring values.
+We construct the artist to have two stages
 
@@
 \end{equation}
+which are the data to visual variable map $\vchannel$ at the bundle level and the visual variable to graphic map $\vmark$ at the sheaf level.
+Usage of the bundle directly in an equation or diagram - $\dotal, \vtotal, \gtotal$ - denotes that the function can be evaluated pointwise over $\dbasepoint \in \dbase$.
+Use of the categorical form -- $\mathcal{\dtotalc}, \mathcal{\vtotalc}, \mathcal{\gtotalc}$ -- denotes that the function is evaluated over an openset object, meaning that the function needs knowledge of both a value over a point and some information about neighboring values.
 
@@
 We represent the graphic output as a fiber bundle because it is an abstraction that is generalizable to various output mediums (screens, 3D prints) and types of graphics.
-In this work, \gtotal\ assumes the the display is an idealized 2D screen and \gsection\ is an abstraction of rendering.
-For example, \gsection\ can be a specification such as PDF\cite{bienz1993portable}, SVG\cite{quintScalable2003} or an OpenGL scene graph\cite{CarsonOpenGL1997}, or a rendering engine such as Cairo\cite{CairographicsOrg} or AGG\cite{shemanarevAntiGrainGeometry}.
 
@@
 
-\begin{example}
-As illustrated in \autoref{fig:artist:graphic}, 
-\end{example}
+As illustrated in \autoref{fig:artist:graphic}, in this work, \gtotal\ assumes the the display is an idealized 2D screen.
+\note{include some more about the figure} The graphic section \gsection\ is an abstraction of rendering.
+For example, \gsection\ can be a specification such as PDF\cite{bienz1993portable}, SVG\cite{quintScalable2003} or an OpenGL scene graph\cite{CarsonOpenGL1997}, or a rendering engine such as Cairo\cite{CairographicsOrg} or AGG\cite{shemanarevAntiGrainGeometry}.
 

### 2021-11-11 paper/main.tex
q

@@
 \newtheorem{prop}{Proposition} %math?
-\usepackage[switch]{lineno}
 \usepackage{minted}
@@
 \hyphenation{op-tical net-works semi-conduc-tor}
+
+\usepackage[switch]{lineno}
+\linenumbers
 
@@
     \end{itemize}
-The coomposition of morphisms is defined component wise by composition in $\mathcal{X}$ and $\mathcal{Y}$.ss
+The composition of morphisms is defined component wise by composition in $\mathcal{X}$ and $\mathcal{Y}$.ss
 \end{definition}
+The fiber category $\mathcal{\dfiber}$ can encode a large variety of collections of data fields because it is a product category of arbitrary categories.
+For example, a lists of strings is an instance of an object in \textbf{Set}, networks are an instance of \textbf{Graph}, and images are vector spaces which are a specific type of topological space \textbf{Top}.
+The morphisms $Hom_{\mathcal{\dfiber}}$ are functions from the data to itself, for example the binary operations, group actions, and measurement scales discussed in \autoref{sec:related-work:equivariance}.
+These functions could also be monoid actions\cite{barrCategoryTheoryComputing}, which provide a way of applying partial order relations to data\cite{fongInvitationAppliedCategory2019}, such as to build multi-ranked indicators\cite{bruggemannRankingPrioritizationMultiindicator2011}.
+
+\note{note sure this note applies anymore: We note that a one object category is a monoid \cite{nlab:monoid,barrCategoryTheoryComputing}, which is an algebraic structure that lends itself to computer library design because they are an abstraction of function composition \cite{yorgeyMonoidsThemeVariations}.}
 
@@
 
-The domain of the artist function is the space of all possible graphics. As with the domain $\sheaf_{E}$, the grapic space is modeled with fiber bundle equivalent categories:
+The domain of the artist function is the space of all possible graphics. As with the domain $\sheaf_{E}$, the graphic space is modeled with fiber bundle equivalent categories:
 \begin{table}[!h]
@@
   \begin{tikzcd}
-    \dtotalc \arrow[rr, "\vchannel"] \arrow[rdd, "\pi"'] &         & \dtotalc \arrow[ldd, "\pi"] & \mathcal{\vtotalc}                                                        & Im_{\vmarkc} \subset \gtotalc                     \\
-  &         &                             & {} \arrow[r, "\vmarkc"]                                                   & {}                                \\
-   & \dbasec &                             & \mathcal{\dbasec} \arrow[uu, "\sheaf_{\vtotalc}"] \arrow[r, "\vindexc^*"] & \mathcal{\gbasec} \arrow[uu, "\sheaf_{\vmarkc}"']
+    \dtotalc \arrow[r, "\vindexc"] \arrow[dd, "\pi"'] & \vtotalc \arrow[ldd, "\pi"] & \mathcal{\vtotalc}                                                        & \gtotalc                                           \\
+    &                             & {} \arrow[r, "\vmarkc"]                                                   & {}                                                 \\
+    \dbasec                                           &                             & \mathcal{\dbasec} \arrow[uu, "\sheaf_{\vtotalc}"] \arrow[r, "\vindexc^*"] & \mathcal{\gbasec} \arrow[uu, "\sheaf_{\vartist}"']
     \end{tikzcd}
@@
 We encode the continuity of the data as the \textcolor{base}{base space} $\dbase$.
-\note{does this go in intro?} The \textcolor{fiber}{fiber space} \dfiber\ is the space of all possible values of the data \autoref{eq:related-work:fiber}.
-We model the data as the section \dsection because, as described in \autoref{sec:related-work:fiber-bundles}, the section is the map from the indexing space \dbase to the space of possible data values \dfiber.
-Spivak's notation, as discussed in \autoref{sec:related-work:equivariance}, also allows for associating elements in a section with the field it comes from.
+We model the data as the section \dsection\ because, as described in \autoref{sec:related-work:fiber-bundles}, the section is the map from the indexing space \dbase\ to the space of possible data values \dfiber.
+We adopt Spivak's formulation of the fiber as a (column name, data domain) simple schema \cite{spivakSIMPLICIALDATABASES,spivakDatabasesAreCategories2010}.
+Spivak formally maps column names and field types to the set of values associated with the field type, for example $\mathbb{R}$ for a \texttt{float} column named \textit{temperature}.
 This allows for field based selection of values, while inclusion allows for continuity (index) based selections.
@@
   \begin{tikzcd}[ampersand replacement=\&]
-      \vfiber \arrow[r, hook] \& \vtotal \arrow[d, "\pi"'] \\
-                        \& \dbase \arrow[u, "\vsection"', bend right]
-  \end{tikzcd}
-\end{equation}
-
-\subsubsection{Graphic to Data: \vindex}
-\begin{equation}
-  \label{eq:artist:construction:xi}
-  \begin{tikzcd}
-      \dtotal \arrow[d, "\pi"'] & \gtotal \arrow[d, "\pi"'] \\
-      \dbase                   & \gbase \arrow[l, "\vindex"']
-  \end{tikzcd}
-\end{equation}
-
-The functor $\vindex$ is a deformation retract, which means....\cite{hatcherAlgebraicTopology2002,spanier1989algebraic}
-
-\begin{figure}[!h]
-  \centering
-  \subfloat[\dbasec_a \sqcup_{\dbasec_c} \dbase_b]{\includegraphics[width=1.2in]{union_k.png}%
-  \label{fig_first_case}}
-  %\hfil
-  \subfloat[\dbasec_c=\varnothing]{\includegraphics[width=1.2in]{disjoint_k.png}%
-  \label{fig_second_case}}
-  \caption{\note{probably doesn't need the k=0}}
-  \label{fig_sim}
-\end{figure}
-
-\begin{equation}
-  \label{eq:artist:construction:xi:multi_base}
-  \begin{tikzcd}
-    \dbase_{c} \arrow[d, hook] \arrow[r, hook] & \dbase_{b}                 & \gbase_b \arrow[l, "\vindex"]                                                \\
-    \dbase_{a}                                 & \gbase_a \arrow[l, "\vindex"'] & \gbase_c \arrow[l, hook'] \arrow[u, hook] \arrow[llu, "\vindex" description]
-  \end{tikzcd}
-\end{equation}
-
-Can be implemented in V
-
-\begin{equation}
-  \label{eq:artist:construction:xi:v}
-  \begin{tikzcd}
-  \textcolor{color=gray!50}{\dtotal}\arrow[rd, "\pi"', color=gray!50] & \vtotal \arrow[d, "\pi"] & \vindex^*\vtotal \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"', dashed] & \textcolor{color=gray!50}{\gtotal} \arrow[ld, "\pi", color=gray!50] \\
-    & \dbase                   & \gbase \arrow[l, "\vindex"']                                                &                          
+      \vfiberc \arrow[r, hook] \& \vtotalc \arrow[d, "\pi"'] \\
+                        \& \dbasec \arrow[u, "\vsectionc"', bend right]
   \end{tikzcd}
@@
 \subsubsection{Data to Visual Encodings: \vchannel} 
+\note{maybe figure out how to put these side by side nicely}
+
 \begin{equation}
@@
   \begin{tikzcd}
-  \dtotal \arrow[rd, "\pi"'] \arrow[r, "\vchannel"] & \vtotal \arrow[d, "\pi"] & \textcolor{fade}{\vindex^*\vtotal} \arrow[d, "\vindex^*\pi"', color=fade] \arrow[l, "\vindex^*", color=fade] & \textcolor{fade}{\gtotal} \arrow[ld, "\pi", color=fade] \\
-& \dbase                   & \textcolor{gray!50}{\gbase} \arrow[l, "\vindex"', color=gray!50]                                       &                          
+    \dtotalc \arrow[d, "\pi"'] \arrow[r, "\vchannelc"] & \vtotalc \arrow[ld, "\pi"] &  & \dfiberc \arrow[r, "\vchannelc"]                          & \vfiberc \\
+    \dbasec                                            &                            &  & \dbasec \arrow[u, "\dsectionc"] \arrow[ru, "\vsectionc"'] &         
     \end{tikzcd}
 \end{equation}
-
-Visual encoding functions \vindex\ are functors because they are expected to preserve structure on the data and visual side, as discussed in \autoref{sec:related-work:equivariance}.
-
+Since the functor \vindex\ is bundle wise, it can be evaluated at the bundle at a point, which is the fiber space; therefore \vindex\ is a functors from the product category $\mathcal{\dfiber}$ to the product categrory $\mathcal{\vfiber}$.
 This constraint is expressed in our construction of \vchannel
-
-\begin{figure}[!h]
-  \centering
-  \subfloat[Artists with shared $\mu_i$ renderered correctly]{\includegraphics[width=1.2in]{partial_fixed.png}%
-  \label{fig_first_case}}
-  %\hfil
-  \subfloat[Artists without shared $\mu_i$]{\includegraphics[width=1.2in]{partial_invalid.png}%
-  \label{fig_second_case}}
-  \caption{Simulation results for the network.}
-  \label{fig_sim}
-  \end{figure}
-
-\begin{equation}
-\label{eq:artist:construction:nu:fiber_map}
-\begin{tikzcd}[column sep=.5em]
-  \dfiber \arrow[rr, "\vchannel"] &                                                         & \vfiber \\
-                                  & \dbase \arrow[lu, "\dsection"] \arrow[ru, "\vsection"'] &        
-  \end{tikzcd}
-\end{equation}
 
@@
 \begin{tikzcd}
-  \dsection \arrow[d, "\varphi"'] \arrow[r, "\vchannel"] & \vsection \arrow[d, "\vchannel_{\varphi}"] \\
-  \dsection^{\prime} \arrow[r, "\vchannel"]                       & \vsection^{\prime}                                 
+  \dsectionc \arrow[d, "\dfuncc"'] \arrow[r, "\vchannelc"] & \vsectionc \arrow[d, "\vchannelc_{\dfuncc}"] \\
+  \dsectionc^{\prime} \arrow[r, "\vchannelc"]                       & \vsectionc^{\prime}                                 
   \end{tikzcd}
 \end{equation}
+
+Since the functor $\nu$ acts on product categories, it can be decomposed into $nu_i$ component functors that act on corresponding sections $\dsection_i$ of the fiber $\dfiber_i$.
+
+\begin{equation}
+  \label{eq:math:artist:nu}
+  \{\vchannelc_{0}, \ldots, \vchannelc_{n}\}: \{\dsectionc_{0}, \ldots, \dsectionc_{n}\} \mapsto \{\vsectionc_{0}, \ldots, \vsectionc_{n}\}
+\end{equation}
+
+One example of this is
+
+\begin{figure}[!h]
+  \centering
+  \subfloat[\note{insert title}]{\includegraphics[width=1.2in]{partial_fixed.png}%
+  \label{fig_first_case}}
+  %\hfill
+  \subfloat[\note{insert title}]{\includegraphics[width=1.2in]{partial_invalid.png}%
+  \label{fig_second_case}}
+  \caption{is a weird nu, colors are a but much}
+  \label{fig_sim}
+\end{figure}
 
@@
 
-\begin{equation}
-  \label{eq:math:artist:nu}
-  \{\vchannel_{0}, \ldots, \vchannel_{n}\}: \{\dsection_{0}, \ldots, \dsection_{n}\} \mapsto \{\vsection_{0}, \ldots, \vsection_{n}\}
-\end{equation}
-
 We enforce the equivariance constraint specified in \autoref{eq:artist:construction:multi_nu} by constructing artists that apply \vchannel\ encoders such that 
@@
 \subsubsection{Visual to Graphic: \vmark}
+
 \begin{equation}
@@
   \begin{tikzcd}
-    \textcolor{fade}{\dtotal} \arrow[rd, "\pi"', color=fade] \arrow[r, "\vchannel", color=fade] & \textcolor{gray!50}{\vtotal} \arrow[d, "\pi", color=fade] & \vindex^*\vtotal \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"', color=fade] \arrow[r, "\vmark"] & \gtotal \arrow[ld, "\pi"] \\
-    & \textcolor{fade}{\dbase} & \gbase \arrow[l, "\vindex"', color=fade]                                            &                          
-    \end{tikzcd}
-\end{equation}
-
-\begin{equation}
-  \label{eq:artist:construction:q:xi}
-  \begin{tikzcd}
-    \vtotalc \arrow[rr, "\vmarkdc", bend left] & \vtotalpullc \arrow[l, "\vindexc^*"] \arrow[r, "\vmarkc"'] & \gtotalc
-  \end{tikzcd}
-\end{equation}
-
-\begin{equation}
-  \label{eq:artist:construction:q:fiber}
-  \begin{tikzcd}[column sep=.5em]
-    \vfiber \arrow[rr, "\vmark"] &                                                         & \gfiber \\
-    & \gbase \arrow[ru, "\gsection"'] \arrow[lu, "\vsection"] &        
-    \end{tikzcd}
-\end{equation}
-Visual encodings to something like marks
-
-\begin{equation}
-  \label{eq:artist:construction:q:equivariance}
-  \begin{tikzcd}
-    \vsection \arrow[d, "\vchannel_\varphi"'] \arrow[r, "\vmark"] & \gsection \arrow[d, "\vmark_\varphi"] \\
-    \vsection^{\prime} \arrow[r, "\vmark"]                                 & \gsection^{\prime}                            
+    \mathcal{\vtotalc}                                & \gtotalc                                           \\
+    {} \arrow[r, "\vmarkc"]                           & {}                                                 \\
+    \mathcal{\dbasec} \arrow[uu, "\sheaf_{\vtotalc}"] & \mathcal{\gbasec} \arrow[uu, "\sheaf_{\vartist}"']
     \end{tikzcd}
@@
 \begin{equation} 
-  \label{eq:artist:construction:q:multifiber}
+  \label{eq:artist:construction:latent_vars}
 \begin{tikzcd}
-  \dfiberc_a \arrow[rr, "\vchannelc"] \arrow[rd, "proj_a" description]          &            & \vfiberc_a \arrow[rr, "\vmarkd"]         &  & Im_{\vmarkdc} \subset \gtotalc \arrow[llld, "g_a" description, dashed]                  \\
- & \dfiberc_c &                                          &  &\\
-  \dfiberc_b \arrow[rr, "\vchannelc^{\prime}"] \arrow[ru, "proj_b" description] &            & \vfiberc_b \arrow[rr, "\vmarkd^{\prime}"] &  & Im_{\vmarkdc^{\prime}}\subset \gtotalc^{\prime} \arrow[lllu, "g_b" description, dashed]
+  \mathcal{\vfiberc_a} \arrow[rrr, "\vmarkc_a"] \arrow[rd, "p_c"]          &                      &  & \sheafc_{\vartist} \arrow[lld, "l_c", dashed]                     \\
+                                                                           & \mathcal{\vfiberc_c} &  &                                                                   \\
+  \mathcal{\vfiberc_b} \arrow[rrr, "\vmarkc_b"] \arrow[ru, "p^{\prime}_c"] &                      &  & \sheafc_{\vartistc^{\prime}} \arrow[llu, "l^{\prime}_c"', dashed]
 \end{tikzcd}
 \end{equation}
+
+\subsubsection{Graphic to Data: \vindex}
+\begin{equation}
+  \label{eq:artist:construction:xi}
+  \begin{tikzcd}
+    \textcolor{fade}{\dtotal} \arrow[rd, "\pi"', color=fade] &         & \textcolor{fade}{\vtotal} \arrow[ld, "\pi", color=fade] &  & \textcolor{fade}{\gtotal} \arrow[d, "\pi", color=fade] \\
+  & \dbasec &                                             &  & \gbasec \arrow[lll, "\vindexc"']          
+    \end{tikzcd}
+\end{equation}
+
+The functor $\vindex$ is a deformation retract, which means....\cite{hatcherAlgebraicTopology2002,spanier1989algebraic}
+
+By \note{homotopy lifting?}
+\begin{equation}
+  \label{eq:artist:construction:xi:lifted}
+  \begin{tikzcd}
+    & \opensetc_j \arrow[d, "\iota", hook] \\
+\dbasec \arrow[ru, "\xi^*", dashed] & \gbasec \arrow[l, "\xi"]            
+\end{tikzcd}
+\end{equation}
+
+\begin{figure}[!h]
+  \centering
+  \subfloat[\dbasec_a \sqcup_{\dbasec_c} \dbase_b]{\includegraphics[width=1.2in]{union_k.png}%
+  \label{fig_first_case}}
+  %\hfil
+  \subfloat[\dbasec_c=\varnothing]{\includegraphics[width=1.2in]{disjoint_k.png}%
+  \label{fig_second_case}}
+  \caption{\note{probably doesn't need the k=0}}
+  \label{fig_sim}
+\end{figure}
+
+\begin{equation}
+  \label{eq:artist:construction:xi:multi_base}
+  \begin{tikzcd}
+    \dbasec_{c} \arrow[d, hook] \arrow[r, hook] & \dbasec_{b}                 & \gbasec_b \arrow[l, "\vindexc"]                                                \\
+    \dbasec_{a}                                 & \gbasec_a \arrow[l, "\vindexc"'] & \gbasec_c \arrow[l, hook'] \arrow[u, hook] \arrow[llu, "\vindexc" description]
+  \end{tikzcd}
+\end{equation}
+
+\note{talk about xi even though is not explicitely implemented}
+The mapping between graphic and data space expresses how...
+This is what allows the artist to generate graphics where the subset of data on view is dynamically updated, such as pan and zoom navigation\cite{NekrasovskiEvaluationPanZoom2006} and sliding windows on streaming data\cite{crouchDynamicGraphsSlidingwindow2013,chuTimeSeriesSegmentation1995}.
 
@@
   \includegraphics[width=\columnwidth]{path_of_q.png}
-    \caption{\note{add in xi!}}
+    \caption{\vchannel \& \mark are what we're implementing}
   \label{fig:api}

### 2021-11-17 paper/main.tex
monoidal categories and functors

@@
 \begin{definition} 
-  \label{def:category:F} 
-  The fiber category $\mathcal{\dfiberc}$ is a product category\cite{maclaneCategoriesWorkingMathematician2013,nlab:product_category} of zero or more arbitrary categories.
-In the two category case, the category $\mathcal{\dfiberc}$ consists of
+  \label{def:category:F} The fiber category $\mathcal{\dfiber}$ is a monoidal category\cite{milewskiCategoryTheoryProgrammers}, which means it is a category equipped with a bifunctur $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$
   \begin{itemize}
-    \item \textit{objects} ordered pairs $(x,y)$,  $x$ is an object of $\mathcal{X}$ and  $y$ is an object of $\mathcal{Y}$
-    \item \textit{morphisms} ordered pairs $(f:x\rightarrow x^{\prime}, g:y\rightarrow y^{\prime})$
+    \item \textit{object} a monoid object \dfiber. A monoid is a set with a binary operator that is associative, closed, and for which the set contains an identity element $i$ \cite{nlan:monoid}.
+    \item \textit{morphisms} $\dfiber \otimes \dfiber \rightarrow \dfiber$ and unit $i \rightarrow \dfiber$
   \end{itemize}
-The composition of morphisms is defined component wise by composition in $\mathcal{X}$ and $\mathcal{Y}$.ss
 \end{definition}
-The fiber category $\mathcal{\dfiber}$ can encode a large variety of collections of data fields because it is a product category of arbitrary categories.
+The bifunctor $\otimes$ preserves identity and composition {fongInvitationAppliedCategory2019} for fibers that consist of many fields.
+Given a pair of objects $X, Y$, the functor yeilds a set of pairs $\{(x,y) | x \in X, y \in Y\}$.
+For the associated morphisms $f: X\rightarrow X$ and $g: Y\rightarrow Y$, the function $(f\times g): (X\times Y) \rightarrow (X \times Y)$ is pointwise $(f\times g)(x, y)\coloneqq (f(x), g(y))$.
+This bookkeeping means that the fiber category $\mathcal{\dfiber}$ can encode a large variety of collections of data fields because it is a product category of arbitrary categories.
+
 For example, a lists of strings is an instance of an object in \textbf{Set}, networks are an instance of \textbf{Graph}, and images are vector spaces which are a specific type of topological space \textbf{Top}.
@@
 These functions could also be monoid actions\cite{barrCategoryTheoryComputing}, which provide a way of applying partial order relations to data\cite{fongInvitationAppliedCategory2019}, such as to build multi-ranked indicators\cite{bruggemannRankingPrioritizationMultiindicator2011}.
-
-\note{note sure this note applies anymore: We note that a one object category is a monoid \cite{nlab:monoid,barrCategoryTheoryComputing}, which is an algebraic structure that lends itself to computer library design because they are an abstraction of function composition \cite{yorgeyMonoidsThemeVariations}.}
 
@@
 \end{equation}
-
+%%% the fiber objects need to be composable into product categories, coproduct? -> OR on return values 
 \autoref{eq:artist:intersection:basespace} and \autoref{eq:artist:intersection:fiber} can be composed to specify how different aspects of the structure of a multivariate nested dimensional dataset, such as a spatio-temporal weather dataset, should be preserved in a visualization.

### 2021-11-18 paper/main.tex
xi

@@
 
-\begin{definition} A category \mathcal{C} consists of a collection of objects $C$ with identity $id_c: C\rightarrow C$ and the set of morphisms between every two objects $f:C_1 \rightarrow C_2$, \cite{fongInvitationAppliedCategory2019,maclaneCategoriesWorkingMathematician2013}.
+\begin{definition} A category \mathcal{C} consists of a collection of objects $c$ with identity $id_c: C\rightarrow C$ and the set of morphisms between every two objects $f:C_1 \rightarrow C_2$, \cite{fongInvitationAppliedCategory2019,maclaneCategoriesWorkingMathematician2013}.
 \end{definition}
@@
 
- \begin{definition} A functor is a morphism between categories $\mathfrak{F}: \mathcal{C} \rightarrow \mathcal{D}$.
+ \begin{definition} A functor is a morphism between any two  categories $\mathfrak{F}: \mathcal{C} \rightarrow \mathcal{D}$.
 A functor has the properties of identity and composition \cite{riehlCategoryTheoryContext}\end{definition}
@@
   \begin{itemize}
-    \item \textit{objects} open sets $\openset_{j} \in \{openset\}$ in the topological space $(\dbase, \mathscr{T})$), including the empty set $\varnothing$ and the maximum set \dbase. 
-    \item \textit{morphisms} $\iota: U_{j} \hookrightarrow U_{k}$ for all  $U_j, U_k \subset \dtotal$
+    \item \textit{objects} open sets $\openset_{i} \in \{openset\}$ in the topological space $(\dbase, \mathscr{T})$), including the empty set $\varnothing$ and the maximum set \dbase. 
+    \item \textit{morphisms} $\iota: \openset_{i} \hookrightarrow \openset_{j}$ for all  $\openset_i, \openset_j \subset \dtotal$
   \end{itemize} 
 \end{definition}
+
+We also introduce a category of open sets called $\mathcal{\gbase}$ that encapsulates the topology of the graphic display space.
+The category $\mathcal{\gtotal}$ has  inclusion morphisms $\iota: {\openset_{i}}^{\prime} \hookrightarrow {\openset_{j}}^{\prime}$, including into the total set \gtotal.
+We introduce a surject map $\vindex: \gtotal \rightarrow \dtotal$ from the the graphic base space to the data base space.
+Given the morphism $\iota$
+\begin{equation}
+  \label{eq:artist:vindex}
+  \begin{tikzcd}
+    & \opensetc^{\prime}_i \arrow[d, "\iota"] \\
+  \dbasec \arrow[ru, "\vindexc^*", dashed] & \gbasec \arrow[l, "\vindexc"]   
+\end{tikzcd}
+\end{equation}
+we propose that the map \vindex\ can be pulled back to a map between open sets in \dbase and open sets in \gbase. This map is the functor
+\begin{equation}
+  \vindex^{*}:\mathcal{\dtotal} \rightarrow \mathcal{\gtotal}
+\end{equation}
+that maps from the data base space to the graphic base space. The map $\vindex^{*}$ is a lookup map between where a point is located in the data and where it is located on screen.
 

### 2021-11-18 paper/main.tex
xi and bib:
rewrote cat intro

@@
 
-\usepackage[switch]{lineno}
+\usepackage[switch,columnwise]{lineno}
+
+\begin{document}
 \linenumbers
 
-\begin{document}
-%
 \title{Topological Equivariant Artist Model for Visualization Library Architecture}
@@
 
-\note{I've seen this convention before, but I also cite all these people above}
-For more information on fiber bundles and sheaves, see Hatcher\cite{hatcherAlgebraicTopology2002}, Munkres\cite{munkresElementsAlgebraicTopology1984}, Spanier \cite{spanier1989algebraic} and Ghrist\cite{ghristElementaryAppliedTopology2014,ghristHomologicalAlgebraData2018}.
-
 \subsection{Equivariance}
@@
 
-For more information on category theory, see Barr and Wells\cite{barrCategoryTheoryComputing}, Fong and Spivak \cite{fongInvitationAppliedCategory2019}, Riehl\cite{riehlCategoryTheoryContext} and Bradley et.
-al.
-\cite{bradleyTopologyCategoricalApproach2020}.
+\section{Artist} 
+In this section we describe the properties a data to visualization transform must satisfy to be considered equivariant and continuity preserving.
+We propose that by explicitly defining these properties, we can better incorporate them into visualization library design.
 
-\section{Artist} 
-In this section, we use category theory to formally express the implicit assumptions that visualization library components make in transforming data into graphical representations.
-We propose that the visualization transformation can be modeled as a functor, which we call the \textcolor{artist}{Artist} \vartist \footnote{We call this transformation the artist because the \mintinline{python}{Artist} object in Matplotlib\cite{hunterMatplotlib2DGraphics2007}}.
-The artist \vartist\ converts data, which we model as the sheaf $\sheafc_{\dtotalc}$ to graphics in the sheaf $\sheafc_{\gtotalc}$.
+We formulate the visualization transformation as a functor, which we call the \textcolor{artist}{Artist} \vartist \footnote{We call this transformation the artist because the \mintinline{python}{Artist} object in Matplotlib\cite{hunterMatplotlib2DGraphics2007}}.
+The artist \vartist\ converts\textcolor{total}{data}, which we denote $\mathcal{\dtotalc}$ to \textcolor{total}{graphics} $\mathcal{\gtotalc}$.
 We formulate this association as
@@
   \label{eq:artist:artist}
-  \vartistc: \sheafc_{\dtotalc} \rightarrow \sheafc_{\gtotal}
+  \vartistc: \mathcal{\dtotalc} \rightarrow \mathcal{\gtotalc}
 \end{equation}
 
-In this section we describe the structure of the data and graphic spaces, which we formulate as objects in categories and the morphisms between those objects.
-We then use these definitions to express the structure that an artist must preserve.
-Finally, we propose that artists can be composed by manufacturing new artists based on various types of input data.
+In this section we encapsulate the structure of the data and graphic spaces as categories.
+We then use these definitions to express the structure that an artist preserves.
+Finally, we introduce rules for composing artists in a way where the structure between the inputs to the individual artists is preserved.
 

### 2021-11-18 paper/main.tex
cleanup functor and NT language

@@
 \usepackage{mathrsfs} % math script fonts
+\usepackage{bm}
 \theoremstyle{definition}
@@
 
- \begin{definition} A functor is a morphism between any two  categories $\mathfrak{F}: \mathcal{C} \rightarrow \mathcal{D}$.
+ \begin{definition} A functor is a morphism between any two  categories $\bm{F}: \mathcal{C} \rightarrow \mathcal{D}$.
 A functor has the properties of identity and composition \cite{riehlCategoryTheoryContext}\end{definition}
+
 A functor preserves the structure of the category, namely morphisms and the source and target objects of those morphisms, composition of morphisms, and identity morphisms \cite{riehlCategoryTheoryContext}.
-Contravariant functors are functors where the morphisms in $\mathcal{C}$ go in the opposite direction from the morphisms in $\mathcal{D}$.
+Functors $\bm{F}: \mathcal{C}^{op} \rightarrow \mathcal{D}$ are called contravariant functors.
 
@@
  \end{definition}
-The natural transformation acts as a wrapper of sorts
+
+The natural transformation composes with a functor such that the output is a compatible functor between the same categories 
 \begin{equation}
@@
 \end{equation}
-such that $\alpha(F: \mathcal{C} \rightarrow \mathcal{D}) = G: \mathcal{C} \rightarrow \mathcal{D}$ commutes for any morphism $f:c\rightarrow c^{\prime}$ \cite{fongInvitationAppliedCategory2019}.
+This means that $\alpha(F: \mathcal{C} \rightarrow \mathcal{D}) = G: \mathcal{C} \rightarrow \mathcal{D}$ commutes for any morphism $f:c\rightarrow c^{\prime}$ \cite{fongInvitationAppliedCategory2019}.
+A natural transform is a way to specify which functors between categories preserve the structure of the categories in the same manner.
 Arbitrary morphisms on sheaves are natural transforms \cite{SheafMathematics2021,bradleyWhatNaturalTransformation}.
@@
 We formulate the visualization transformation as a functor, which we call the \textcolor{artist}{Artist} \vartist \footnote{We call this transformation the artist because the \mintinline{python}{Artist} object in Matplotlib\cite{hunterMatplotlib2DGraphics2007}}.
-The artist \vartist\ converts\textcolor{total}{data}, which we denote $\mathcal{\dtotalc}$ to \textcolor{total}{graphics} $\mathcal{\gtotalc}$.
+The artist \vartist\ converts \textcolor{total}{data}, which we denote $\mathcal{\dtotalc}$ to \textcolor{total}{graphics} $\mathcal{\gtotalc}$.
 We formulate this association as

### 2021-11-22 paper/main.tex
sections

@@
 \end{definition}
-\note{intuition about fiber}
+\note{intuition about fiber + spivak}
 
@@
 \begin{equation}
+  \label{eq:related-work:continuity:fiber-bundle}
   \begin{tikzcd}
@@
 
+\note{this needs to be reworked into a coherent sentence}
 How these spaces fit together is illustrated in \autoref{fig:related-work:fiber-bundle}, wherein values lives in the fiber $\dfiber\subseteq\dtotal$.
@@
 
+As proposed by Butler et al, data can be represented as a map from the base space \dbase\ to the fiber space \dfiber.
+This map is called a section $\dsection: \dbase \rightarrow \dtotal$ of the fiber bundle.
+A fiber bundle has many sections, and the set of all sections
+\begin{equation}
+\label{eq:related-work:continuity:sections}
+\{\dsection^{i}: \openset_{j}\rightarrow \dtotal\}|_{i\in I} = \Gamma(\openset_{j}, \dtotal)
+\end{equation}
+over a subset \openset\ is the set of all data sets with a shared fiber \dfiber\ and continuity \dbase. The set of all sections defined over the total base space is denoted $\Gamma(\dbase, \dtotal)$.
+
 \subsubsection{Sheaves}
-The set of all sections sections $\{\dsection^{i}: \openset_{j}\rightarrow \dtotal\}|_{i\in I}$ over an arbitrary open set $U_{j}$ is denoted $\Gamma(\openset_{\dbasepoint}, \dtotal)$.
-The set of all sections over the total space \dbase\ is denoted $\Gamma(\dbase, \dtotal)$.
-
 The map from the open set to the set of sections is called a presheaf $\sheaf: U_j \rightarrow \Gamma(\openset_j, \dtotal)$ \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}.
@@
 
-\subsection{Categorical Artist}
+\subsection{Natural Transformation of Data to Graphic}
 \label{sec:artist:category}
@@
 
-We also introduce a category of open sets called $\mathcal{\gbase}$ that encapsulates the topology of the graphic display space.
-The category $\mathcal{\gtotal}$ has  inclusion morphisms $\iota: {\openset_{i}}^{\prime} \hookrightarrow {\openset_{j}}^{\prime}$, including into the total set \gtotal.
-We introduce a surject map $\vindex: \gtotal \rightarrow \dtotal$ from the the graphic base space to the data base space.
-Given the morphism $\iota$
-\begin{equation}
-  \label{eq:artist:vindex}
-  \begin{tikzcd}
-    & \opensetc^{\prime}_i \arrow[d, "\iota"] \\
-  \dbasec \arrow[ru, "\vindexc^*", dashed] & \gbasec \arrow[l, "\vindexc"]   
-\end{tikzcd}
-\end{equation}
-we propose that the map \vindex\ can be pulled back to a map between open sets in \dbase and open sets in \gbase. This map is the functor
-\begin{equation}
-  \vindex^{*}:\mathcal{\dtotal} \rightarrow \mathcal{\gtotal}
-\end{equation}
-that maps from the data base space to the graphic base space. The map $\vindex^{*}$ is a lookup map between where a point is located in the data and where it is located on screen.
+Over these open sets $\openset_i$ are sets of local sections
 
@@
 
-The category $\mathcal{H}$ has as objects the sets of functions that generate every arbitrary graphic over the open sets in $\mathcal{\gtotal}$.
-While $\mathcal{H}$ is the domain of the artist \vartist, the range of \vartist\ is the subset of graphics $\sheaf_{\vartist} \subset \sheaf_{\mathcal{\dtotal}}$ such that structure imposed by $\dfunc_{\dtotal}$ commutes
+We introduce a surjective map $\vindex: \gtotal \rightarrow \dtotal$ from the the graphic base space to the data base space. Given the morphism $\iota$
 \begin{equation}
-  \label{eq:artist:equivariance}
+  \label{eq:artist:vindex}
   \begin{tikzcd}
-    \sheafc_{\dtotalc} \arrow[r, "\vartistc"] \arrow[d, "\dfuncc_{\dtotalc}"'] & \sheafc_{\vartistc} \arrow[d, "\dfuncc_{\vartistc}"] \\
-    \sheafc_{\dtotalc} \arrow[r, "\vartistc"] & \sheafc_{\vartistc}                                              
+    & \opensetc^{\prime}_i \arrow[d, "\iota"] \\
+  \dbasec \arrow[ru, "\vindexc^*", dashed] & \gbasec \arrow[l, "\vindexc"]   
 \end{tikzcd}
 \end{equation}
-The equivariance condition expressed in \autoref{eq:artist:equivariance} is that a graphic generated by  transforming and then visualizing data $\vartist(\dfunc_{\dtotal}(\sheaf_{\mathcal{dtotal}}))$ can equivalently be generated by visualizing the data and then transforming the graphic $\dfunc_{\vartist}(\vartist(\sheaf_{\mathcal{\dtotal}}))$.
-This is a formal statement of the equivariance illustrated in \autoref{fig:related-work:equivariance}
+we propose that the map \vindex\ can be pushed forward to a map between open sets in \dbase and open sets in \gbase. This map is the functor
+\begin{equation}
+  \vindex_{*}:\mathcal{\dtotal} \rightarrow \mathcal{\gtotal}
+\end{equation}
+that maps from open sets $\openset_i \subset \dbase$ in data space to the open sets in visual space $\openset^{\prime}_i \subset \gbase$. The graphic space is
 
-In \autoref{sec:artist:construction}, we show a method of constructing the artist such that the constraints of \autoref{eq:artist:morphism-of-sheaf} and \autoref{eq:artist:equivariance} can be satisfied, thereby ensuring equivariance and the preservation of continuity.
-We then use this formulation to guide the development of visualization library components in \autoref{sec:case-study}.
+\begin{figure}[h!]
+  \includegraphics[width=\columnwidth]{pushforward.png}
+  \label{artist:pushforward_xi}
+  \caption{The pushforward map $\vindex_*$ repeats the 1D fiber over every point $\dbasepoint \in \dbase$ in data space, turning the fiber into a 2D plane of repeating values over the corresponding region $\dbasepoint = \vindex(\gbasepoint)$ in the graphic space \gbase.
+}
+\end{figure}
 

### 2021-11-22 paper/main.tex
spivak fiber

@@
 \end{definition}
-\note{intuition about fiber + spivak}
+
+The fibers can be thought of as encapsulating data types according to Spivak, who proposes that databases can be represented as fiber bundles \cite{spivakSIMPLICIALDATABASES,spivakDatabasesAreCategories2010}.
+Spivak formulates the fiber as a (field name, data domain) simple schema where field names and field types can be formally mapped to the set of values associated with the field type, for example $\mathbb{R}$ for a \texttt{float} column named \textit{temperature}.
 
@@
 We model the data as the section \dsection\ because, as described in \autoref{sec:related-work:fiber-bundles}, the section is the map from the indexing space \dbase\ to the space of possible data values \dfiber.
-We adopt Spivak's formulation of the fiber as a (column name, data domain) simple schema \cite{spivakSIMPLICIALDATABASES,spivakDatabasesAreCategories2010}.
-Spivak formally maps column names and field types to the set of values associated with the field type, for example $\mathbb{R}$ for a \texttt{float} column named \textit{temperature}.
-This allows for field based selection of values, while inclusion allows for continuity (index) based selections.
 

### 2021-11-22 paper/main.tex
sheaf equation w/ object notation

@@
 
-\note{the notation here may need to be X and Y}
-
 \begin{definition} A category \mathcal{C} consists of a collection of objects $c$ with identity $id_c: C\rightarrow C$ and the set of morphisms between every two objects $f:C_1 \rightarrow C_2$, \cite{fongInvitationAppliedCategory2019,maclaneCategoriesWorkingMathematician2013}.
@@
 
- \begin{definition} A functor is a morphism between any two  categories $\bm{F}: \mathcal{C} \rightarrow \mathcal{D}$.
+ \begin{definition} A functor is a morphism between objects of any two  categories $\bm{F}: \mathcal{C} \rightarrow \mathcal{D}$.
 A functor has the properties of identity and composition \cite{riehlCategoryTheoryContext}\end{definition}
@@
 
-Over these open sets $\openset_i$ are sets of local sections
-
-\begin{definition} \label{def:category:E} The category \mathcal{\dtotalc} is a subcategory of \textbf{Set} and consists of 
-  \begin{itemize}
-    \item \textit{objects} sets of sections $\Gamma(\openset_{j}, \dtotal)$ for all $U_j \subset \dtotal$
-    \item \textit{morphisms}  $\iota^*: \Gamma(\openset_{k}, \dtotal) \hookrightarrow \Gamma(\openset_{j}, \dtotal)$ for all  $\openset_j, \openset_k \subset \dtotal$
-  \end{itemize} 
-\end{definition}
-
-We model data as the sheaf $\sheafc_{\dtotalc}: \mathcal{\dbasec}^{op}\rightarrow \mathcal{\dtotalc}$
+We model data as the sheaf $\sheafc({\dtotalc}): \mathcal{\dbasec}^{op}\rightarrow \bm{Set}$.
+The objects of \textbf{Set} in this sheaf are the sets of sections introduced in \autoref{eq:related-work:continuity:sections} and the morphisms are the inclusion maps in \autoref{eq:related-work:continuity:presheaf}.
 \begin{equation}
@@
   \begin{tikzcd}
-    & \mathcal{\dtotalc} & \\
-{} \arrow[rr, "\dfuncc", Rightarrow, shift right, shorten=3em] & & {} \\
-    & \mathcal{\dbasec}^{op} \arrow[uu, "\sheafc_{\dtotalc}", bend left] \arrow[uu, "\sheafc_{\dtotalc}"', bend right] &   
+    & \Gamma(\opensetc_i, \dtotalc) \in Ob(\bm{Set}) & \\
+{} \arrow[rr, "\dfuncc", Rightarrow, pos=.3, shorten <=3.5em, shorten >=8.85em] & & {} \\
+    & \opensetc_i \in Ob(\mathcal{\dbasec}^{op}) \arrow[uu, "\sheafc(\dtotalc)", bend left, shift left = 10.5] \arrow[uu, "\sheafc(\dtotalc)"', bend right, shift left=10.5] &   
 \end{tikzcd}
 \end{equation}
-where the map \dfunc\ is any arbitrary function between sheafs.
-The sheaf $\sheaf_{\dtotal}$ over a limit of open sets that contain a point $\dbasepoint \in \dbase$ is approximately the same as a sheaf over a point \dbasepoint\ and is called the stalk $\sheaf_{\mathcal{\dtotal}}|_{k}\coloneqq \lim\limits_{U\ni k} \Gamma(U, \dtotal)$ \cite{StalkSheaf2019}.
+ The map $\dfunc: \sheaf(\dtotal) \rightarrow \sheaf(\dtotal)$ is any arbitrary function between sheafs.
+The sheaf $\sheaf(\dtotal)$ over a limit of open sets that contain a point $\dbasepoint \in \dbase$ is approximately the same as a sheaf over a point \dbasepoint\ and is called the stalk $\sheaf(\dtotal)|_{\dbasepoint}\coloneqq \lim\limits_{\openset\ni \dbasepoint} \Gamma(\openset, \dtotal)$ \cite{StalkSheaf2019}.
 

### 2021-11-22 paper/main.tex
artist sheaf

@@
 \begin{definition} 
-  \label{def:category:F} The fiber category $\mathcal{\dfiber}$ is a monoidal category\cite{milewskiCategoryTheoryProgrammers}, which means it is a category equipped with a bifunctur $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$
+  \label{def:category:F} The fiber category $\mathcal{\dfiberc}$ is a monoidal category\cite{milewskiCategoryTheoryProgrammers}, which means it is a category equipped with a bifunctur $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$
   \begin{itemize}
@@
 \end{definition}
-The bifunctor $\otimes$ preserves identity and composition {fongInvitationAppliedCategory2019} for fibers that consist of many fields.
-Given a pair of objects $X, Y$, the functor yeilds a set of pairs $\{(x,y) | x \in X, y \in Y\}$.
-For the associated morphisms $f: X\rightarrow X$ and $g: Y\rightarrow Y$, the function $(f\times g): (X\times Y) \rightarrow (X \times Y)$ is pointwise $(f\times g)(x, y)\coloneqq (f(x), g(y))$.
-This bookkeeping means that the fiber category $\mathcal{\dfiber}$ can encode a large variety of collections of data fields because it is a product category of arbitrary categories.
+The bifunctor $\otimes$ combines categories in a way that  preserves identity and composition \cite{\fongInvitationAppliedCategory2019}.
+For example, given a pair of categories $\mathcal{\dfiber_a},\mathcal{\dfiber_b}$, the bifunctor yields a set of pairs $\{(\dfiber_a, \dfiber_b) | \dfiber_a \in \mathcal{\dfiber_a}, \dfiber_b \in \mathcal{\dfiber_b}\}$.
+For the associated morphisms $g: \mathcal{\dfiber_a}\rightarrow \mathcal{\dfiber_a}$ and $h: \mathcal{\dfiber_b}\rightarrow \mathcal{\dfiber_b}$, the function $(g\times h): (\mathcal{\dfiber_a}\times \mathcal{\dfiber_b}) \rightarrow (\mathcal{\dfiber_a}\times \mathcal{\dfiber_b})$ is pointwise $(g\times h)(\dfiber_a, \dfiber_b)\coloneqq (g(\dfiber_a), h(\dfiber_b))$.
+Because of the bifunctor, objects of the category $\mathcal{\dfiber}$ can be the product of any arbitrary number and types of categories, including \textbf{Set}, \textbf{Graph}, and topological spaces \textbf{Top}.
 
-For example, a lists of strings is an instance of an object in \textbf{Set}, networks are an instance of \textbf{Graph}, and images are vector spaces which are a specific type of topological space \textbf{Top}.
-The morphisms $Hom_{\mathcal{\dfiber}}$ are functions from the data to itself, for example the binary operations, group actions, and measurement scales discussed in \autoref{sec:related-work:equivariance}.
-These functions could also be monoid actions\cite{barrCategoryTheoryComputing}, which provide a way of applying partial order relations to data\cite{fongInvitationAppliedCategory2019}, such as to build multi-ranked indicators\cite{bruggemannRankingPrioritizationMultiindicator2011}.
+The set of all morphism of the fiber category $Hom_{\mathcal{\dfiber}}(\dfiber, \dfiber)$ are all possible morphisms in a given category.
+This includes any of the functions that serve as the basis of equivariance, such as the binary operations, group actions, and measurement scale discussed in  \autoref{sec:related-work:equivariance}.
+The generalization to $Hom_\mathcal{\dfiber}$ also allows for the inclusion of structures such as monoid actions\cite{barrCategoryTheoryComputing}, which provide a way of applying partial order relations to data\cite{fongInvitationAppliedCategory2019}, such as to build multi-ranked indicators\cite{bruggemannRankingPrioritizationMultiindicator2011}.
 
@@
 The sheaf maps \dfunc\ restricted to the stalk are members of $hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$.
-The maps $\varphi \in Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ include the types of data transformations described in \autoref{sec:related-work:equivariance}.
+The maps $\dfunc \in Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ include the types of data transformations described in \autoref{sec:related-work:equivariance}.
 When a fiber bundle is trivial, we can dispense with localization and directly consider the fiber maps $Hom(\dfiber_{\dbasepoint}, \dfiber_{\dbasepoint})$ over a point \dbasepoint.
@@
 
-The domain of the artist function is the space of all possible graphics. As with the domain $\sheaf_{E}$, the graphic space is modeled with fiber bundle equivalent categories:
-\begin{table}[!h]
-  \centering
-  \label{tab:artist:graphic}
-  \begin{tabular}{l l l}
-    fiber bundle & data & graphic \\
-    \textcolor{base}{base space} & \mathcal{\dbasec} & \mathcal{\gbasec}\\
-    \textcolor{total}{total space} & \mathcal{\dtotalc} & \mathcal{\gtotalc}\\
-    \textcolor{fiber}{fiber space} & \mathcal{\dfiberc} & \mathcal{\gfiberc}\\
-    \pi^{-1} & \sheaf_{\mathcal{\dtotalc}} & \sheaf_{\mathcal{\gtotalc}}\\
-  \end{tabular}
-\end{table}
+The continuity of the graphic is encapsulated as a category $\mathcal{\mathcal{\gbase}}$ containing open set objects $\opensetg$, the complete set of opensets $\gbase$ that is the full space of the graphic, and the empty set.
+As with $\mathcal{\dbase}$, the morphisms between the open sets  $\opensetg$ are inclusion morphisms $\iota$.
+Over the graphic space $\opensetg \subseteq \gbase$ are a set of sections
+\begin{equation}
+  \label{eq:artist:graphic_sections}
+  \{\gsectionc^{i}:\opensetgc_i\rightarrow \gtotalc\}|_{i\in I}
+\end{equation}
+which are the set of all graphics with a shared base space and fiber and are a subcategory of \textbf{Set}. This allows us to represent the full space of graphics as a sheaf $\sheaf(\gtotal)$
+\begin{equation}
+  \label{eq:artist:artist_sheaf}
+  \begin{tikzcd}
+    {Ob(\bm{Set}) \ni \Gamma(\opensetgc_i, \gtotalc)}                                                                                                  &  & \{\gsectionc^{j}:\opensetgc_i\rightarrow \gtotalc\}|_{j\in J} \arrow[ll, "\iota"', hook'] \\ &  & \\
+    \opensetgc_i \arrow[uu, "\sheafc(\gtotalc)"] \arrow[rruu, "\vartistc(\sheafc(\dtotalc))\coloneqq \sheaf_{\vartist}(\dtotalc)" description, dashed] &  & 
+    \end{tikzcd}
+\end{equation}
+The set of graphics reachable through an artist function is a subset $\{\gsection^{j}\}|_{j\in J}$ of the full set of graphics.
+These sets of reachable graphics over opensets \opensetg\ are also objects of \textbf{Set}.
+This allows us to formulate the graphic output by the artist as a sheaf $\sheaf_{\vartist}$ from the graphic base space to the set of reachable graphics.
 
-We introduce a surjective map $\vindex: \gtotal \rightarrow \dtotal$ from the the graphic base space to the data base space. Given the morphism $\iota$
+To establish a mapping between graphic and data base spaces, we introduce the surjective map $\vindex: \gtotal \rightarrow \dtotal$. Given the morphism $\iota$
 \begin{equation}
@@
   \begin{tikzcd}
-    & \opensetc^{\prime}_i \arrow[d, "\iota"] \\
+    & \opensetgc_i \arrow[d, "\iota"] \\
   \dbasec \arrow[ru, "\vindexc^*", dashed] & \gbasec \arrow[l, "\vindexc"]   
@@
 \end{equation}
-we propose that the map \vindex\ can be pushed forward to a map between open sets in \dbase and open sets in \gbase. This map is the functor
+we propose that a pushforward functor $\vindex_*$ 
 \begin{equation}
-  \vindex_{*}:\mathcal{\dtotal} \rightarrow \mathcal{\gtotal}
+  \vindexc_{*}:\mathcal{\dbasec} \rightarrow \mathcal{\gbasec}
 \end{equation}
-that maps from open sets $\openset_i \subset \dbase$ in data space to the open sets in visual space $\openset^{\prime}_i \subset \gbase$. The graphic space is
+can be used to map between open sets $\openset_i \subset \dbase$ in data space and the open sets in visual space $\opensetg_i \subset \gbase$.
 
@@
   \includegraphics[width=\columnwidth]{pushforward.png}
-  \label{artist:pushforward_xi}
+  \label{fig:artist:pushforward_xi}
   \caption{The pushforward map $\vindex_*$ repeats the 1D fiber over every point $\dbasepoint \in \dbase$ in data space, turning the fiber into a 2D plane of repeating values over the corresponding region $\dbasepoint = \vindex(\gbasepoint)$ in the graphic space \gbase.
@@
 \end{figure}
+
+We compose the functor $\vindex_*$ with the sheaf $\sheaf$ to move the sets of sections from $\Gamma(\openset, \dtotal)$ to $\Gamma(\opensetg, \vindex_* \dtotal)$.
+As illustrated in \autoref{fig:artist:pushforward_xi}, this operation copies a section and a fiber over one point $\dbasepoint \in \dbase$ such that it repeats over the corresponding region $\gbasepoint \subset \gbase$ in the graphic space.
 

### 2021-11-22 paper/main.tex
natural transform

@@
     {Ob(\bm{Set}) \ni \Gamma(\opensetgc_i, \gtotalc)}                                                                                                  &  & \{\gsectionc^{j}:\opensetgc_i\rightarrow \gtotalc\}|_{j\in J} \arrow[ll, "\iota"', hook'] \\ &  & \\
-    \opensetgc_i \arrow[uu, "\sheafc(\gtotalc)"] \arrow[rruu, "\vartistc(\sheafc(\dtotalc))\coloneqq \sheaf_{\vartist}(\dtotalc)" description, dashed] &  & 
+    \opensetgc_i \arrow[uu, "\sheafc(\gtotalc)"] \arrow[rruu, "\vartistc(\sheafc(\dtotalc))\coloneqq \sheaf_{\vartist}(\gtotalc)" description, dashed] &  & 
     \end{tikzcd}
@@
 can be used to map between open sets $\openset_i \subset \dbase$ in data space and the open sets in visual space $\opensetg_i \subset \gbase$.
+We can then compose the functor $\vindex_*$ with the sheaf $\sheaf(\dtotal)$
+\begin{equation*}
+\vindex_*(\openset) \rightarrow \Gamma(\vindex_*(\openset), \vindex_* \dtotal) \coloneqq \opensetg \rightarrow \gamma(\opensetg, \vindex_* \dtotal) 
+\end{equation*}
+which allows for moving the sets of sections from $\Gamma(\openset, \dtotal)$ to $\Gamma(\opensetg, \vindex_* \dtotal)$. We denote the pulled forward sheaf as $\sheaf_{\vindex_*}(\dtotal)$.
 
@@
 
-We compose the functor $\vindex_*$ with the sheaf $\sheaf$ to move the sets of sections from $\Gamma(\openset, \dtotal)$ to $\Gamma(\opensetg, \vindex_* \dtotal)$.
 As illustrated in \autoref{fig:artist:pushforward_xi}, this operation copies a section and a fiber over one point $\dbasepoint \in \dbase$ such that it repeats over the corresponding region $\gbasepoint \subset \gbase$ in the graphic space.
+The pull forward functor allows us to formulate the data sheaf as
+\begin{equation}
+  \label{eq:artist:data_pushforward_sheaf}
+  \sheaf_{\vindex_*}: \mathcal{\gbase^{op}} \rightarrow \bm{Set}
+\end{equation}
+
+Following from \autoref{eq:artist:data_pushforward_sheaf} and the construction of the artist sheaf in \autoref{eq:artist:artist_sheaf}, we propose that the artist function $\vartist$ is a natural transformation
+
+\begin{equation}
+  \begin{tikzcd}
+    & {} \arrow[dd, "\vartistc", Rightarrow, shorten=1.25em] & \\
+\mathcal{\gbasec}^{op} \arrow[rr, "\sheafc(\dtotalc)", bend left] \arrow[rr, "\sheafc_{\vartistc}(\gtotalc)"', bend right] &  & \textcolor{section}{\bm{Set}}\\
+    & {} &                            
+\end{tikzcd}
+\end{equation}
 

### 2021-11-22 paper/main.tex
nt

@@
     & {} \arrow[dd, "\vartistc", Rightarrow, shorten=1.25em] & \\
-\mathcal{\gbasec}^{op} \arrow[rr, "\sheafc(\dtotalc)", bend left] \arrow[rr, "\sheafc_{\vartistc}(\gtotalc)"', bend right] &  & \textcolor{section}{\bm{Set}}\\
+\mathcal{\gbasec}^{op} \arrow[rr, "\sheafc_{\vindex_*}(\dtotalc)", bend left] \arrow[rr, "\sheafc_{\vartistc}(\gtotalc)"', bend right] &  & \bm{Set}\\
     & {} &                            
@@
 \end{equation}
+
+of sheaf functors 
+\begin{equation}
+  \label{eq:artist:sheaf_map}
+  \vartist: \sheafc_{vindex_*}(\dtotal) \rightarrow \sheafc_{\vartist}(\gtotal)
+\end{equation}
+from the graphic base space category $\mathcal{\gbase}^{op}$ to different objects in \bm{Set}. This means that the following diagram commutes
 

### 2021-11-24 paper/main.tex
rw: bundle and sheaves

@@
 
-\usepackage[switch,columnwise]{lineno}
+\usepackage[]{lineno}
 
@@
 Open sets are a collection of subsets $\{\opensetc\}$ in a set \dbase, including the empty set, the whole set \dbase, and every union and intersection of its member subsets.
-For each point $\dbasepoint \in \dbase$, there is a collection of subsets $\\{\openset\} \subset \dbase$ such that $\dbasepoint \in \openset \subseteq \dbase$ \cite{spanier1989algebraic, bradleyTopologyCategoricalApproach2020}
+For each point $\dbasepoint \in \dbase$, there is a collection of subsets $\{\openset\} \subset \dbase$ such that $\dbasepoint \in \openset \subseteq \dbase$ \cite{spanier1989algebraic, bradleyTopologyCategoricalApproach2020}
 \end{definition}
@@
 \end{definition}
+
 The fiber bundle is trivial when $\dtotal = \dbase \times \dfiber$ is true globally; a non-trivial bundle can be thought of as the twisted fiber product $\dtotal=\dbase times_{T} \dfiber$\cite{hatcherAlgebraicTopology2002,nlab:twisting_function}.
-The twisting refers to the fiber \dfiber\ not aligning in the same direction over all $\dopenspace \subseteq \dbase$.
+The twisting refers to the fiber \dfiber\ not aligning in the same direction over all $\openset \subseteq \dbase$.
+
+As proposed by Butler et al \cite{butlerVisualizationModelBased1989,butlerVectorBundleClassesForm1992}, data can be represented as a map from the base space \dbase\ to the fiber space \dfiber.
+This map is called a section $\dsection: \dbase \rightarrow \dtotal$ of the fiber bundle.
 
@@
   \label{eq:related-work:continuity:fiber-bundle}
-  \begin{tikzcd}
+  \begin{tikzcd}[row sep=large]
       \dfiberc \arrow[r, hook] & \dtotalc \arrow[d, "\pi"'] \\
-                        & \dbasec \arrow[u, "\dsectionc"', bend right]
+                        & \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
   \end{tikzcd}
@@
 
-\note{this needs to be reworked into a coherent sentence}
-How these spaces fit together is illustrated in \autoref{fig:related-work:fiber-bundle}, wherein values lives in the fiber $\dfiber\subseteq\dtotal$.
-In this example, the fiber \dfiber\ is the cartesian product of two sets $F_{0}\times F_{1}$ where each fiber $F_{i}$ is ....
-
-As proposed by Butler et al, data can be represented as a map from the base space \dbase\ to the fiber space \dfiber.
-This map is called a section $\dsection: \dbase \rightarrow \dtotal$ of the fiber bundle.
 A fiber bundle has many sections, and the set of all sections
@@
 \label{eq:related-work:continuity:sections}
-\{\dsection^{i}: \openset_{j}\rightarrow \dtotal\}|_{i\in I} = \Gamma(\openset_{j}, \dtotal)
+\{\dsectionc^{i}: \opensetc\rightarrow \dtotalc\}|_{i\in I} = \cgamma{\opensetc}{\dtotalc}
 \end{equation}
-over a subset \openset\ is the set of all data sets with a shared fiber \dfiber\ and continuity \dbase. The set of all sections defined over the total base space is denoted $\Gamma(\dbase, \dtotal)$.
+over an open set $openset \subseteq \dtotal$ is the set of all data sets with a shared fiber \dfiber\ and continuity \dbase.
+The set of all sections defined over the total base space is denoted $\Gamma(\dbase, \dtotal)$.
 
 \subsubsection{Sheaves}
-The map from the open set to the set of sections is called a presheaf $\sheaf: U_j \rightarrow \Gamma(\openset_j, \dtotal)$ \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}.
-The presheaf preserves inclusion maps between the open sets open sets $\iota$ and the inclusion map $\iota^*$ between the sets of sections.
+Sheaves can be thought of as an "algebraic data structure" \cite{ghristElementaryAppliedTopology2014} for data that lives over topological spaces.
+A presheaf $\sheafc$ is a map from the open set to the set of sections $\sheafc: \opensetc \rightarrow \gammac{\openset}{ \dtotal}$ \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}.
+The presheaf preserves inclusion maps $\iota$ between open sets and the inclusion map $\iota^*$ between the sets of sections.
 
@@
   \begin{tikzcd}[column sep=small]
-    {\Gamma(\opensetc_1, \dtotalc)}                                       & {}                                                    & {\Gamma(\opensetc_2, \dtotalc)} \arrow[ll, "\iota^*"', hook'] \\
+    {\cgamma{\opensetc_1}{\dtotalc}} & {} & {\cgamma{\opensetc_2}{\dtotalc}} \arrow[ll, "\iota^*"', hook'] \\
                                                                   &                                                       &                                                       \\
@@
 
-This means that a smaller space $\openset_1$ is included in a larger space $\openset_2$ and a function that is continuous over a larger space $\openset_2$ is continuous over a subspace $\openset_1 \subspace \openset_2$.
+This means that a smaller space $\openset_1$ is included in a larger space $\openset_2$ and a function that is continuous over a larger space $\openset_2$ is continuous over a subspace $\openset_1 \subset\openset_2$.
 Sheaves are presheafs where sets of sections $\Gamma$ are defined over unions of open sets over a topology $\bigcup_{j \in I} \openset)_j \in \dtotal$ \cite{spanier1989algebraic, nlab:sheaf}.
-Sheaves are often used as an abstraction for keeping track of how data over topological spaces is glued together \cite{ghristElementaryAppliedTopology2014} because they model the data as sets of sections, continuity of the data in the base, and how subsets fit together through inclusion maps.
+Sheaves provide a mathematical abstraction of data, the space it lives over, and relationships between subsets.
 
@@
 \end{definition}
-The bifunctor $\otimes$ combines categories in a way that  preserves identity and composition \cite{\fongInvitationAppliedCategory2019}.
+The bifunctor $\otimes$ combines categories in a way that  preserves identity and composition \cite{fongInvitationAppliedCategory2019}.
 For example, given a pair of categories $\mathcal{\dfiber_a},\mathcal{\dfiber_b}$, the bifunctor yields a set of pairs $\{(\dfiber_a, \dfiber_b) | \dfiber_a \in \mathcal{\dfiber_a}, \dfiber_b \in \mathcal{\dfiber_b}\}$.
@@
 The set of graphics reachable through an artist function is a subset $\{\gsection^{j}\}|_{j\in J}$ of the full set of graphics.
-These sets of reachable graphics over opensets \opensetg\ are also objects of \textbf{Set}.
+These sets of reachable graphics over open sets \opensetg\ are also objects of \textbf{Set}.
 This allows us to formulate the graphic output by the artist as a sheaf $\sheaf_{\vartist}$ from the graphic base space to the set of reachable graphics.
@@
 \begin{equation*}
-\vindex_*(\openset) \rightarrow \Gamma(\vindex_*(\openset), \vindex_* \dtotal) \coloneqq \opensetg \rightarrow \gamma(\opensetg, \vindex_* \dtotal) 
+\vindex_*(\openset) \rightarrow \Gamma(\vindex_*(\openset), \vindex_* \dtotal) \coloneqq \opensetg \rightarrow \gamma(\opensetg, \vindex_*(\dtotal) 
 \end{equation*}
@@
 \end{equation}
-from the graphic base space category $\mathcal{\gbase}^{op}$ to different objects in \bm{Set}. This means that the following diagram commutes
+where each sheaf is a map from an open set in the graphic base space $\opensetg \subset \gbase$ into a set of sections of a fiber bundle over that open set.
+
+Because the artist is a natural transform, it also transforms morphisms between functors\cite{bradleyWhatNaturalTransformation, maclaneCategoriesWorkingMathematician2013}.
+As illustrated in \autoref{eq:artist:morphism-of-sheaf-data}, the morphism between data sheaves is $\dfunc$.
+By definition of a natural transform, the following diagram commutes
 

### 2021-11-24 paper/main.tex
pullback formulas

@@
   \begin{tikzcd}[column sep=small]
-    {\cgamma{\opensetc_1}{\dtotalc}} & {} & {\cgamma{\opensetc_2}{\dtotalc}} \arrow[ll, "\iota^*"', hook'] \\
-                                                                  &                                                       &                                                       \\
+    {\cgamma{\opensetc_1}{\dtotalc}} & {} & {\cgamma{\opensetc_2}{\dtotalc}} \arrow[ll, "\iota^*"', hook'] \\& &\\
     \opensetc_1 \arrow[uu, "\sheafc_{\dtotal}"] \arrow[rr, "\iota"', hook] & {} \arrow[uu, "\sheafc_{\dtotal}" description, dotted] & \opensetc_2 \arrow[uu, "\sheafc_\dtotal"']                    
@@
 
-To establish a mapping between graphic and data base spaces, we introduce the surjective map $\vindex: \gtotal \rightarrow \dtotal$. Given the morphism $\iota$
+To establish a mapping between graphic and data base spaces, we introduce the surjective map $\vindex: \gbase \rightarrow \dbase$. 
 \begin{equation}
@@
   \begin{tikzcd}
-    & \opensetgc_i \arrow[d, "\iota"] \\
-  \dbasec \arrow[ru, "\vindexc^*", dashed] & \gbasec \arrow[l, "\vindexc"]   
+    \dtotalc                        & \vindexc^*\dtotal \arrow[l, "\vindexc^*"', dashed]                                                                           \\
+    \dbasec \arrow[u, "\dsectionc"] & \gbasec \arrow[l, "\vindexc"] \arrow[lu, "\dsectionc \circ \vindexc" description] \arrow[u, "\vindexc^*\dsectionc"', dashed]
     \end{tikzcd}
 \end{equation}
-we propose that a pushforward functor $\vindex_*$ 
-\begin{equation}
-  \vindexc_{*}:\mathcal{\dbasec} \rightarrow \mathcal{\gbasec}
-\end{equation}
-can be used to map between open sets $\openset_i \subset \dbase$ in data space and the open sets in visual space $\opensetg_i \subset \gbase$.
-We can then compose the functor $\vindex_*$ with the sheaf $\sheaf(\dtotal)$
-\begin{equation*}
-\vindex_*(\openset) \rightarrow \Gamma(\vindex_*(\openset), \vindex_* \dtotal) \coloneqq \opensetg \rightarrow \gamma(\opensetg, \vindex_*(\dtotal) 
-\end{equation*}
-which allows for moving the sets of sections from $\Gamma(\openset, \dtotal)$ to $\Gamma(\opensetg, \vindex_* \dtotal)$. We denote the pulled forward sheaf as $\sheaf_{\vindex_*}(\dtotal)$.
+As shown in \autoref{eq:artist:vindex}, the map between spaces $\vindex$ can be composed with section map $\dsection$ such that there is a map from the graphic base space $\gbase$ to the fiber bundle in which the data lives $\dtotal$.
 
 \begin{figure}[h!]
-  \includegraphics[width=\columnwidth]{pushforward.png}
+  \includegraphics[width=\columnwidth]{pullback.png}
   \label{fig:artist:pushforward_xi}
-  \caption{The pushforward map $\vindex_*$ repeats the 1D fiber over every point $\dbasepoint \in \dbase$ in data space, turning the fiber into a 2D plane of repeating values over the corresponding region $\dbasepoint = \vindex(\gbasepoint)$ in the graphic space \gbase.
+  \caption{The map $\vindex^*$ is a map from a repeation of the 1D fiber over every point $\dbasepoint \in \dbase$ in data space, turning the fiber into a 2D plane of repeating values over the corresponding region $\dbasepoint = \vindex(\gbasepoint)$ in the graphic space \gbase.
 }
 \end{figure}
-
-As illustrated in \autoref{fig:artist:pushforward_xi}, this operation copies a section and a fiber over one point $\dbasepoint \in \dbase$ such that it repeats over the corresponding region $\gbasepoint \subset \gbase$ in the graphic space.
-The pull forward functor allows us to formulate the data sheaf as
+Since there is a mapping from the graphic space to the total bundle $\dsection \circ \vindex: \gbase \rightarrow \dtotal$, the mapping $\vindex$ can be pulled through the composition such that there is a map $\vindex^*$ between a total space over graphics $\vindex^*\dtotal$ and its equivalent space over data $\dtotal$.
+As illustrated in the example given in \autoref{fig:artist:pushforward_xi}, a bundle $\dtotal$ over a point $\dbasepoint \in \dbase$ is pulled back through \vindex\ such that it is copied over every point in the corresponding region in graphic space $\vindex(\gbasepoint) = \dbasepoint, \gbasepoint \subset \gbase$.
+The pullback $\vindex^*$ can then be composed with the elements of the set of sections
 \begin{equation}
-  \label{eq:artist:data_pushforward_sheaf}
-  \sheaf_{\vindex_*}: \mathcal{\gbase^{op}} \rightarrow \bm{Set}
+  \label{eq:artist:data_gamma_pullback}
+ \Gamma(\dbase, \dtotal) \xrightarrow{\vindex^*} \Gamma(\gbase, \vindex^*\dtotal)
+\end{equation}
+to generate a set of equivalent sections over the graphic $\Gamma(\gtotal, \vindex^*{\dtotal})$.
+As a consequence of \autoref{eq:artist:data_gamma_pullback}, there exists a data sheaf over graphic space $\sheaf: \mathcal{\gbase^{op}} \rightarrow \Gamma(\gtotal, \vindex^*{\dtotal})$ which is equivalent to
+\begin{equation}
+  \label{eq:artist:data_pullback_sheaf}
+  \sheafc_{\vindexc^*}: \mathcal{\gbasec^{op}} \rightarrow \textcolor{section}{\bm{Set}}
 \end{equation}
 
-Following from \autoref{eq:artist:data_pushforward_sheaf} and the construction of the artist sheaf in \autoref{eq:artist:artist_sheaf}, we propose that the artist function $\vartist$ is a natural transformation
+Following from \autoref{eq:artist:data_pullback_sheaf} and the construction of the artist sheaf in \autoref{eq:artist:artist_sheaf}, we propose that the artist function $\vartist$ is a natural transformation
 
@@
     & {} \arrow[dd, "\vartistc", Rightarrow, shorten=1.25em] & \\
-\mathcal{\gbasec}^{op} \arrow[rr, "\sheafc_{\vindex_*}(\dtotalc)", bend left] \arrow[rr, "\sheafc_{\vartistc}(\gtotalc)"', bend right] &  & \bm{Set}\\
+\mathcal{\gbasec}^{op} \arrow[rr, "\sheafc_{\vindex_*}(\dtotalc)", bend left] \arrow[rr, "\sheafc_{\vartistc}(\gtotalc)"', bend right] &  & \textcolor{section}{\bm{Set}}\\
     & {} &                            
@@
   \label{eq:artist:sheaf_map}
-  \vartist: \sheafc_{vindex_*}(\dtotal) \rightarrow \sheafc_{\vartist}(\gtotal)
+  \vartist: \sheafc_{\vindex^*}(\dtotal) \rightarrow \sheafc_{\vartist}(\gtotal)
 \end{equation}

### 2021-11-24 paper/main.tex
gamma coloring and sheave/fiber rework

@@
 
-A fiber bundle has many sections, and the set of all sections
+A fiber bundle has many sections, and the set of all sections over an open set $openset \subseteq \dtotal$
 \begin{equation}
 \label{eq:related-work:continuity:sections}
-\{\dsectionc^{i}: \opensetc\rightarrow \dtotalc\}|_{i\in I} = \cgamma{\opensetc}{\dtotalc}
+\cgamma{\opensetc}{\dtotalc} \coloneqq \{\dsectionc^{i}: \opensetc\rightarrow \dtotalc\}|_{i\in I}
 \end{equation}
-over an open set $openset \subseteq \dtotal$ is the set of all data sets with a shared fiber \dfiber\ and continuity \dbase.
-The set of all sections defined over the total base space is denoted $\Gamma(\dbase, \dtotal)$.
+ is the set of all data sets with a shared fiber \dfiber\ and continuity \dbase. The set of all sections defined over the total base space is denoted $\Gamma(\dbase, \dtotal)$.
 
@@
 Sheaves can be thought of as an "algebraic data structure" \cite{ghristElementaryAppliedTopology2014} for data that lives over topological spaces.
-A presheaf $\sheafc$ is a map from the open set to the set of sections $\sheafc: \opensetc \rightarrow \gammac{\openset}{ \dtotal}$ \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}.
+A presheaf $\sheafc$ is a map from the open set to the set of sections $\sheafc: \opensetc \rightarrow \cgamma{\openset}{ \dtotal}$ \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}.
 The presheaf preserves inclusion maps $\iota$ between open sets and the inclusion map $\iota^*$ between the sets of sections.
@@
 This means that a smaller space $\openset_1$ is included in a larger space $\openset_2$ and a function that is continuous over a larger space $\openset_2$ is continuous over a subspace $\openset_1 \subset\openset_2$.
-Sheaves are presheafs where sets of sections $\Gamma$ are defined over unions of open sets over a topology $\bigcup_{j \in I} \openset)_j \in \dtotal$ \cite{spanier1989algebraic, nlab:sheaf}.
+Sheaves are presheafs where sets of sections are defined over unions of open sets over a topology $\bigcup_{j \in I} \openset)_j \in \dtotal$ \cite{spanier1989algebraic, nlab:sheaf}.
 Sheaves provide a mathematical abstraction of data, the space it lives over, and relationships between subsets.
@@
 
-\begin{definition} A category \mathcal{C} consists of a collection of objects $c$ with identity $id_c: C\rightarrow C$ and the set of morphisms between every two objects $f:C_1 \rightarrow C_2$, \cite{fongInvitationAppliedCategory2019,maclaneCategoriesWorkingMathematician2013}.
+\begin{definition} A category $\mathcal{C}$ consists of a collection of objects $c$ with identity $id_c: C\rightarrow C$ and the set of morphisms between every two objects $f:C_1 \rightarrow C_2$, \cite{fongInvitationAppliedCategory2019,maclaneCategoriesWorkingMathematician2013}.
 \end{definition}
@@
 
- \begin{definition} An opposite category $\mathcal{C}^{op}$ is a category with all the same objects of category $\mathcal{C}$ and but the morphisms are reversed.
+ \begin{definition} An opposite category $\mathcal{C}^{op}$ is a category with all the same objects of category $\mathcal{C}$ and morphisms that are reversed.
 For example $f:C_1 \rightarrow C_2$ in $\mathcal{C}$ is $f:C_2 \rightarrow C_1$ in $\mathcal{C}^{op}$.
@@
 
-We model data as the sheaf $\sheafc({\dtotalc}): \mathcal{\dbasec}^{op}\rightarrow \bm{Set}$.
+We model data as the sheaf $\sheafc({\dtotalc}): \mathcal{\dbasec}^{op}\rightarrow \textcolor{section}{\bm{Set}}$.
 The objects of \textbf{Set} in this sheaf are the sets of sections introduced in \autoref{eq:related-work:continuity:sections} and the morphisms are the inclusion maps in \autoref{eq:related-work:continuity:presheaf}.
@@
   \begin{tikzcd}
-    & \Gamma(\opensetc_i, \dtotalc) \in Ob(\bm{Set}) & \\
+    & \cgamma{\opensetc_i}{\dtotalc} \in Ob(\textcolor{section}{\bm{Set}}) & \\
 {} \arrow[rr, "\dfuncc", Rightarrow, pos=.3, shorten <=3.5em, shorten >=8.85em] & & {} \\
@@
 \end{equation}
- The map $\dfunc: \sheaf(\dtotal) \rightarrow \sheaf(\dtotal)$ is any arbitrary function between sheafs.
-The sheaf $\sheaf(\dtotal)$ over a limit of open sets that contain a point $\dbasepoint \in \dbase$ is approximately the same as a sheaf over a point \dbasepoint\ and is called the stalk $\sheaf(\dtotal)|_{\dbasepoint}\coloneqq \lim\limits_{\openset\ni \dbasepoint} \Gamma(\openset, \dtotal)$ \cite{StalkSheaf2019}.
+The map $\dfunc: \sheaf(\dtotal) \rightarrow \sheaf(\dtotal)$ is any arbitrary morphism between sheaf functors $\sheaf(\dtotal)$.
+Following from \autoref{eq:related-work:equivariance:natural-transform}, \dfunc\ is a natural transform, which means all inclusions $\iota, \iota^*$ are preserved as part of any $\dfunc$ transformation.
 
@@
 
+When a fiber bundle is trivial, we can dispense with localization and directly consider the fiber maps $Hom(\dfiber_{\dbasepoint}, \dfiber_{\dbasepoint})$ over a point \dbasepoint\ as the basis of our equivariance.
+This is because a sheaf $\sheaf(\dtotal)$ over a limit of open sets that contain a point $\dbasepoint \in \dbase$ is approximately the same as a sheaf over a point \dbasepoint\ and is called the stalk $\sheaf(\dtotal)|_{\dbasepoint}\coloneqq \lim\limits_{\openset\ni \dbasepoint} \Gamma(\openset, \dtotal)$ \cite{StalkSheaf2019}.
 The fiber space over a point is embedded in the stalk over that point $\dfiber_{\dbasepoint}\hookrightarrow \mathscr{\dfiber}_{\dbasepoint}$.
 The sheaf maps \dfunc\ restricted to the stalk are members of $hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$.
-The maps $\dfunc \in Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ include the types of data transformations described in \autoref{sec:related-work:equivariance}.
-When a fiber bundle is trivial, we can dispense with localization and directly consider the fiber maps $Hom(\dfiber_{\dbasepoint}, \dfiber_{\dbasepoint})$ over a point \dbasepoint.
-In \autoref{sec:artist:construction}, we assume that the fiber bundles are trivial.
+The maps $\dfunc \in Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ include the types of data transformations described in \autoref{sec:related-work:equivariance}In \autoref{sec:artist:construction}, we assume that the fiber bundles are trivial.
 
@@
   \label{eq:artist:graphic_sections}
-  \{\gsectionc^{i}:\opensetgc_i\rightarrow \gtotalc\}|_{i\in I}
+  \cgamma{\opensetgc}{\gtotalc} \coloneqq \{\gsectionc^{i}:\opensetgc\rightarrow \gtotalc\}|_{i\in I}
 \end{equation}
@@
   \label{eq:artist:artist_sheaf}
-  \begin{tikzcd}
-    {Ob(\bm{Set}) \ni \Gamma(\opensetgc_i, \gtotalc)}                                                                                                  &  & \{\gsectionc^{j}:\opensetgc_i\rightarrow \gtotalc\}|_{j\in J} \arrow[ll, "\iota"', hook'] \\ &  & \\
+  \begin{tikzcd}[row sep=huge]
+    {Ob(\textcolor{section}{\bm{Set}}) \ni \cgamma{\opensetgc_i}{\gtotalc}}                                                                                                  &  & \{\gsectionc^{j}:\opensetgc_i\rightarrow \gtotalc\}|_{j\in J} \arrow[ll, "\iota"', hook'] \\ &  & \\
     \opensetgc_i \arrow[uu, "\sheafc(\gtotalc)"] \arrow[rruu, "\vartistc(\sheafc(\dtotalc))\coloneqq \sheaf_{\vartist}(\gtotalc)" description, dashed] &  & 
@@
   \label{eq:artist:sheaf_map}
-  \vartist: \sheafc_{\vindex^*}(\dtotal) \rightarrow \sheafc_{\vartist}(\gtotal)
+  \vartistc: \sheafc_{\vindexc^*}(\dtotalc) \rightarrow \sheafc_{\vartistc}(\gtotalc)
 \end{equation}

### 2021-11-25 paper/main.tex
reworking up to artist
nlab pullback bundle

@@
 Since there is a mapping from the graphic space to the total bundle $\dsection \circ \vindex: \gbase \rightarrow \dtotal$, the mapping $\vindex$ can be pulled through the composition such that there is a map $\vindex^*$ between a total space over graphics $\vindex^*\dtotal$ and its equivalent space over data $\dtotal$.
-As illustrated in the example given in \autoref{fig:artist:pushforward_xi}, a bundle $\dtotal$ over a point $\dbasepoint \in \dbase$ is pulled back through \vindex\ such that it is copied over every point in the corresponding region in graphic space $\vindex(\gbasepoint) = \dbasepoint, \gbasepoint \subset \gbase$.
+As illustrated in the example given in \autoref{fig:artist:pushforward_xi}, a bundle $\dtotal$ over a point $\dbasepoint \in \dbase$ is pulled back through \vindex\ \cite{nlab:pullback_bundle} such that it is copied over every point in the corresponding region in graphic space $\vindex(\gbasepoint) = \dbasepoint, \gbasepoint \subset \gbase$.
 The pullback $\vindex^*$ can then be composed with the elements of the set of sections

### 2021-11-26 paper/main.tex
added functor commutative diagram

@@
 A functor preserves the structure of the category, namely morphisms and the source and target objects of those morphisms, composition of morphisms, and identity morphisms \cite{riehlCategoryTheoryContext}.
+For example, the functor $F:\mathcal{C} \rightarrow \mathcal{D}$ maps an object in $\mathcal{C}$ into an object in $\mathcal{D}$
+\begin{equation}
+\label{eq:related-work:equivariance:functor_diagram}
+\begin{tikzcd}
+  c \in Ob(\mathcal{C}) \arrow[d, "f"'] \arrow[r, "F"] & F(c) \in Ob(\mathcal{D}) \arrow[d, "F(f)"] \\
+  c^{\prime} \in Ob(\mathcal{C}) \arrow[r, "F"]        & F(c^{\prime}) \in Ob(\mathcal{D})         
+  \end{tikzcd} 
+\end{equation}
+ such that every morphism $f \in Hom(\mathcal{C})$ there is a compatible morphism $F(f) in Hom(\mathcal{D})$.
 Functors $\bm{F}: \mathcal{C}^{op} \rightarrow \mathcal{D}$ are called contravariant functors.
@@
 \end{equation}
-This means that $\alpha(F: \mathcal{C} \rightarrow \mathcal{D}) = G: \mathcal{C} \rightarrow \mathcal{D}$ commutes for any morphism $f:c\rightarrow c^{\prime}$ \cite{fongInvitationAppliedCategory2019}.
+This means that $\alpha(F: \mathcal{C} \rightarrow \mathcal{D}) = G: \mathcal{C} \rightarrow \mathcal{D}$ commutes for any morphism $f:c\rightarrow c^{\prime}$ \cite{fongInvitationAppliedCategory2019}.Given a morphism $f: c\rightarrow c^{\prime}$ for $c, c^{\prime} \in Ob(\mathcal{C})$, the natural transformation decomposes into the components of the morphism $\alpha_{c}, \alpha_{c^{\prime}}$, which yields the naturality condition \cite{milewskiCategoryTheoryProgrammers}
+\begin{equation}
+  \label{eq:related-work:equivariance:naturality-condition}
+  G(f) \circ \alpha_{c} = \alpha_{c^\prime} \circ F(f)
+\end{equation}
+
+which can be expressed as the diagram commuting
+
+\begin{equation}
+  \label{eq:related-work:equivariance:natural-transform-components}
+  \begin{tikzcd}
+    c \arrow[dd, "f"'] &  & F(c) \arrow[dd, "F(f)"'] \arrow[rr, "\alpha_{c}"] &  & G(c) \arrow[dd, "G(f)"] \\
+    {} \arrow[rr, shorten > = 5]      &  & {}                                                &  &                         \\
+    c^{\prime}         &  & F(c^{\prime}) \arrow[rr, "\alpha_{c^{\prime}}"]   &  & G(c^{\prime})          
+    \end{tikzcd}
+\end{equation}
+
 A natural transform is a way to specify which functors between categories preserve the structure of the categories in the same manner.
@@
 where each sheaf is a map from an open set in the graphic base space $\opensetg \subset \gbase$ into a set of sections of a fiber bundle over that open set.
+As defined in \autoref{eq:related-work:equivariance:natural-transform}, the artist is a natural transformation.
+The artist has two components, a sheaf morphism $\vartist_{\openset}$ and a morphism $\vartist_{\Gamma}$ between sets of sections.
 
-Because the artist is a natural transform, it also transforms morphisms between functors\cite{bradleyWhatNaturalTransformation, maclaneCategoriesWorkingMathematician2013}.
-As illustrated in \autoref{eq:artist:morphism-of-sheaf-data}, the morphism between data sheaves is $\dfunc$.
-By definition of a natural transform, the following diagram commutes
+\begin{equation}
+  \label{eq:artist:artist_gamma}
+\end{equation}
+
+As illustrated in \autoref{eq:artist:morphism-of-sheaf-data}, the morphism between data sheaves is $\dfunc$. By definition of a natural transform, the following diagram commutes
 

### 2021-11-26 paper/main.tex
nt commutative diagram

--- 2021-11-25 paper/main.tex
@@
 A functor preserves the structure of the category, namely morphisms and the source and target objects of those morphisms, composition of morphisms, and identity morphisms \cite{riehlCategoryTheoryContext}.
+For example, the functor $F:\mathcal{C} \rightarrow \mathcal{D}$ maps an object in $\mathcal{C}$ into an object in $\mathcal{D}$
+\begin{equation}
+\label{eq:related-work:equivariance:functor_diagram}
+\begin{tikzcd}
+  c \in Ob(\mathcal{C}) \arrow[d, "f"', shift right=8] \arrow[r, "F"] & F(c) \in Ob(\mathcal{D}) \arrow[d, "F(f)", shift right=8] \\
+  c^{\prime} \in Ob(\mathcal{C}) \arrow[r, "F"']                       & F(c^{\prime}) \in Ob(\mathcal{D})                        
+  \end{tikzcd}
+\end{equation}
+ such that for every morphism $f: c \rightarrow c^{\prime} \in Hom(\mathcal{C})$ there is a compatible morphism $F(f): F(c) \rightarrow F(c^{\prime}) \in Hom(\mathcal{D})$.
 Functors $\bm{F}: \mathcal{C}^{op} \rightarrow \mathcal{D}$ are called contravariant functors.
@@
 
-The natural transformation composes with a functor such that the output is a compatible functor between the same categories 
 \begin{equation}
@@
 \end{equation}
-This means that $\alpha(F: \mathcal{C} \rightarrow \mathcal{D}) = G: \mathcal{C} \rightarrow \mathcal{D}$ commutes for any morphism $f:c\rightarrow c^{\prime}$ \cite{fongInvitationAppliedCategory2019}.
-A natural transform is a way to specify which functors between categories preserve the structure of the categories in the same manner.
-Arbitrary morphisms on sheaves are natural transforms \cite{SheafMathematics2021,bradleyWhatNaturalTransformation}.
+The natural transformation composes with a functor such that the output is a compatible functor between the same categories.
+This means that for every morphism $f: c \rightarrow c^{\prime}$ between objects in $\mathcal{C}$, there is a commutating naturality square for the corresponding objects in $\mathcal{D}$ \cite{milewskiCategoryTheoryProgrammers}.
+\begin{equation}
+  \label{eq:related-work:equivariance:natural_transform_commute}
+  \begin{tikzcd}
+    c \arrow[rr, "F"] \arrow[rrrr, "G", bend left] \arrow[dd, "f"'] &  & F(c) \arrow[dd, "F(f)"'] \arrow[rr, "\alpha_{c}"] &  & G(c) \arrow[dd, "G(f)"] \\
+                                                                    &  & {} \arrow[rr, "\alpha" description, dashed]       &  & {}                      \\
+    c^{\prime} \arrow[rr, "F"] \arrow[rrrr, "G", bend right]        &  & F(c^{\prime}) \arrow[rr, "\alpha_{c^{\prime}}"]   &  & G(c^{\prime})          
+    \end{tikzcd}
+\end{equation}
+\note{unpack this a drop more}
 

### 2021-11-26 paper/main.tex
pullback figure

@@
 As defined in \autoref{eq:related-work:equivariance:natural-transform}, the artist is a natural transformation.
-The artist has two components, a sheaf morphism $\vartist_{\openset}$ and a morphism $\vartist_{\Gamma}$ between sets of sections.
-
+Following from \autoref{eq:related-work:equivariance:natural_transform_commute}, the artist specifies that there are compatible inclusion maps in the data and graphic spaces
 \begin{equation}
-  \label{eq:artist:artist_gamma}
+  \label{eq:artist:inclusion_morphisms}
+  \begin{tikzcd}
+    {\Gamma(\opensetc_1, \dtotalc)} \arrow[dd, "\vartistc_{\opensetc_1}"'] &                                                                                                                                   &  &                                                                                                         & {\Gamma(\opensetc_2,\dtotal)} \arrow[dd, "\vartistc_{\opensetc_2}"] \arrow[llll, "\iota^*"', hook'] \\
+                                                                           & \opensetc_1 \arrow[lu, "\sheafc_{\vindex^*}" description] \arrow[ld, "\sheafc_{\vartistc}" description] \arrow[rr, "\iota", hook] &  & \opensetc_2 \arrow[ru, "\sheafc_{\vindex^*}" description] \arrow[rd, "\sheafc_{\vartistc}" description] & \\
+    {\Gamma(\opensetc_1,\gtotalc)} & &  && {\Gamma(\opensetc_2,\gtotalc)} \arrow[llll, "\iota^*", hook']                                       
+    \end{tikzcd}
 \end{equation}
 
-As illustrated in \autoref{eq:artist:morphism-of-sheaf-data}, the morphism between data sheaves is $\dfunc$. By definition of a natural transform, the following diagram commutes
+In \autoref{eq:artist:morphism-of-sheaf-data}, $\dfuncc$ was introduced as a morphism between sheaf functors $\sheafc^{\vindex^*}$.
+Since the artist is a functor $\vartist: \sheafc^{\vindex^*} \rightarrow \sheaf_{\vartist}$, it follows from \autoref{eq:related-work:equivariance:functor_diagram} that
+\begin{equation}
+  \begin{tikzcd}
+    \sheafc_{\vindex^{*}} \arrow[d, "\dfuncc"'] \arrow[r, "\vartistc"] & \sheafc_{\vartist} \arrow[d, "\dfuncc^{\prime} \coloneqq\vartist(\dfuncc)"] \\
+    \sheafc_{\vindex^{*}} \arrow[r, "\vartistc"']                      & \sheafc_{\vartist}                                                         
+    \end{tikzcd}
+\end{equation}
+an arbitrary morphism $\dfunc$ on the data sheaf has a compatible morphism $\dfunc^{\prime}$ on the graphic sheaf.
 

### 2021-11-26 paper/main.tex
A_u, A_phi

@@
 In \autoref{eq:artist:morphism-of-sheaf-data}, $\dfuncc$ was introduced as a morphism between sheaf functors $\sheafc^{\vindex^*}$.
-Since the artist is a functor $\vartist: \sheafc^{\vindex^*} \rightarrow \sheaf_{\vartist}$, it follows from \autoref{eq:related-work:equivariance:functor_diagram} that
+Since the artist is a functor $\vartist: \sheaf^{\vindex^*} \rightarrow \sheaf_{\vartist}$, it follows from \autoref{eq:related-work:equivariance:functor_diagram} that
 \begin{equation}

### 2021-11-29 paper/main.tex
color coded intro category things

@@
 
-\begin{definition} A category $\mathcal{C}$ consists of a collection of objects $c$ with identity $id_c: C\rightarrow C$ and the set of morphisms between every two objects $f:C_1 \rightarrow C_2$, \cite{fongInvitationAppliedCategory2019,maclaneCategoriesWorkingMathematician2013}.
+\begin{definition} A \textcolor{source}{category} $\textcolor{source}{\mathcal{C}}$ consists of a collection of objects $c$ with identity $id_c: C\rightarrow C$ and the set of morphisms between every two objects $f:C_1 \rightarrow C_2$, \cite{fongInvitationAppliedCategory2019,maclaneCategoriesWorkingMathematician2013}.
 \end{definition}
@@
   \begin{tikzcd}
-    C_1 \arrow[r, "f"] \arrow[rd, "g\circ f"'] \arrow["id_{C_1}"', loop, distance=2em, in=215, out=145] & C_2 \arrow[d, "g"] \arrow["id_{C_2}"', loop, distance=2em, in=125, out=55] \\
-  & C_3 \arrow["id_{C_3}"', loop, distance=2em, in=305, out=235]              
+    \textcolor{source}{C_1} \arrow[r, "f", color=source] \arrow[rd, "g\circ f"', color=source] \arrow["id_{C_1}"', loop, distance=2em, in=215, out=145, color=source] & \textcolor{source}{C_2} \arrow[d, "g", color=source] \arrow["id_{C_2}"', loop, distance=2em, in=125, out=55, color=source] \\
+  & \textcolor{source}{C_3} \arrow["id_{C_3}"', loop, distance=2em, in=305, out=235, color=source]              
   \end{tikzcd}
@@
 
- \begin{definition} A functor is a morphism between objects of any two  categories $\bm{F}: \mathcal{C} \rightarrow \mathcal{D}$.
+ \begin{definition} A functor is a morphism between objects of any two  categories $\textcolor{sheaf}{\bm{F}}: \textcolor{source}{\mathcal{C}} \rightarrow \textcolor{target}{\mathcal{D}}$.
 A functor has the properties of identity and composition \cite{riehlCategoryTheoryContext}\end{definition}
@@
 A functor preserves the structure of the category, namely morphisms and the source and target objects of those morphisms, composition of morphisms, and identity morphisms \cite{riehlCategoryTheoryContext}.
-For example, the functor $F:\mathcal{C} \rightarrow \mathcal{D}$ maps an object in $\mathcal{C}$ into an object in $\mathcal{D}$
+For example, the functor $\textcolor{functor}{F}:\mathcal{C} \textcolor{functor}{\rightarrow} \mathcal{D}$ maps an object in $\textcolor{source}{\mathcal{C}}$ into an object in $\textcolor{target}{\mathcal{D}}$
 \begin{equation}
@@
 \begin{tikzcd}
-  c \in Ob(\mathcal{C}) \arrow[d, "f"', shift right=8] \arrow[r, "F"] & F(c) \in Ob(\mathcal{D}) \arrow[d, "F(f)", shift right=8] \\
-  c^{\prime} \in Ob(\mathcal{C}) \arrow[r, "F"']                       & F(c^{\prime}) \in Ob(\mathcal{D})                        
+  \textcolor{source}{c} \arrow[d, "f"', shift right=0, color=source] \arrow[r, "F", color=functor] & \textcolor{target}{F(c)} \arrow[d, "F(f)", shift right=0, color=target] \\
+  \textcolor{source}{c^{\prime}} \arrow[r, "F"', color=functor]                       & \textcolor{target}{F(c^{\prime})}                      
   \end{tikzcd}
@@
 
- \begin{definition} A natural transformation is a morphism of functors $\alpha: F \rightarrow G$ where $F, G$ are functors from $\mathcal{C}$ to $\mathcal{D}$.
+ \begin{definition} A \textcolor{nattran}{natural transformation} is a morphism of functors $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\rightarrow} \textcolor{functor}{G}$ where $F, G$ are functors from $\mathcal{C}$ to $\mathcal{D}$.
 \cite{riehlCategoryTheoryContext, bradleyWhatNaturalTransformation}
@@
 \begin{tikzcd}
-  & {} \arrow[dd, "\alpha", Rightarrow, shorten=1.5em] &             \\
-\mathcal{C} \arrow[rr, "F", bend left] \arrow[rr, "G"', bend right] &                                     & \mathcal{D} \\
+  & {} \arrow[dd, "\alpha", Rightarrow, shorten=1.5em, color=nattran] &             \\
+\textcolor{source}{\mathcal{C}} \arrow[rr, "F", bend left, color=functor] \arrow[rr, "G"', bend right, color=functor] &                                     & \textcolor{target}{\mathcal{D}} \\
   & {}                                  &            
@@
   \begin{tikzcd}
-    c \arrow[rr, "F"] \arrow[rrrr, "G", bend left] \arrow[dd, "f"'] &  & F(c) \arrow[dd, "F(f)"'] \arrow[rr, "\alpha_{c}"] &  & G(c) \arrow[dd, "G(f)"] \\
-                                                                    &  & {} \arrow[rr, "\alpha" description, dashed]       &  & {}                      \\
-    c^{\prime} \arrow[rr, "F"] \arrow[rrrr, "G", bend right]        &  & F(c^{\prime}) \arrow[rr, "\alpha_{c^{\prime}}"]   &  & G(c^{\prime})          
+    \textcolor{source}{c} \arrow[rr, "F", color=functor, alpha=.5] \arrow[rrrr, "G", bend left, color=functor, alpha=.5] \arrow[dd, "f"', color=source] &  & \textcolor{target}{F(c)} \arrow[dd, "F(f)"', color=target] \arrow[rr, "\alpha_{c}", color=artist] &  & \textcolor{target}{G(c)}\arrow[dd, "G(f)", color=target] \\
+                                                                    &  & {} \arrow[rr, "\alpha" description, dashed, color=artist]       &  & {}                      \\
+    \textcolor{source}{c^{\prime}} \arrow[rr, "F", color=functor, alpha=.5] \arrow[rrrr, "G", bend right, color=functor, alpha=.5]        &  & \textcolor{target}{F(c^{\prime})} \arrow[rr, "\alpha_{c^{\prime}}", color=artist]   &  & \textcolor{target}{G(c^{\prime})}
     \end{tikzcd}

### 2021-11-29 paper/main.tex
varphi eqiuv

@@
   \begin{tikzcd}
-    \textcolor{source}{c} \arrow[rr, "F", color=functor, alpha=.5] \arrow[rrrr, "G", bend left, color=functor, alpha=.5] \arrow[dd, "f"', color=source] &  & \textcolor{target}{F(c)} \arrow[dd, "F(f)"', color=target] \arrow[rr, "\alpha_{c}", color=artist] &  & \textcolor{target}{G(c)}\arrow[dd, "G(f)", color=target] \\
-                                                                    &  & {} \arrow[rr, "\alpha" description, dashed, color=artist]       &  & {}                      \\
-    \textcolor{source}{c^{\prime}} \arrow[rr, "F", color=functor, alpha=.5] \arrow[rrrr, "G", bend right, color=functor, alpha=.5]        &  & \textcolor{target}{F(c^{\prime})} \arrow[rr, "\alpha_{c^{\prime}}", color=artist]   &  & \textcolor{target}{G(c^{\prime})}
+    \textcolor{source}{c} \arrow[rr, "F", color=functor, alpha=.5] \arrow[rrrr, "G", bend left, color=functor, alpha=.5] \arrow[dd, "f"', color=source] &  & \textcolor{target}{F(c)} \arrow[dd, "F(f)"', color=target] \arrow[rr, "\alpha_{c}", color=nattran] &  & \textcolor{target}{G(c)}\arrow[dd, "G(f)", color=target] \\
+                                                                    &  & {} \arrow[rr, "\alpha" description, dashed, color=nattran]       &  & {}                      \\
+    \textcolor{source}{c^{\prime}} \arrow[rr, "F", color=functor, alpha=.5] \arrow[rrrr, "G", bend right, color=functor, alpha=.5]        &  & \textcolor{target}{F(c^{\prime})} \arrow[rr, "\alpha_{c^{\prime}}", color=nattran]   &  & \textcolor{target}{G(c^{\prime})}
     \end{tikzcd}
@@
 
-We model data as the sheaf $\sheafc({\dtotalc}): \mathcal{\dbasec}^{op}\rightarrow \textcolor{section}{\bm{Set}}$.
+We model data as the sheaf $\sheafc({\dtotalc}): \mathcal{\dbasec}^{op} \textcolor{sheaf}{\rightarrow} \setc$.
 The objects of \textbf{Set} in this sheaf are the sets of sections introduced in \autoref{eq:related-work:continuity:sections} and the morphisms are the inclusion maps in \autoref{eq:related-work:continuity:presheaf}.
@@
   \begin{tikzcd}
-    & \cgamma{\opensetc_i}{\dtotalc} \in Ob(\textcolor{section}{\bm{Set}}) & \\
-{} \arrow[rr, "\dfuncc", Rightarrow, pos=.3, shorten <=3.5em, shorten >=8.85em] & & {} \\
-    & \opensetc_i \in Ob(\mathcal{\dbasec}^{op}) \arrow[uu, "\sheafc(\dtotalc)", bend left, shift left = 10.5] \arrow[uu, "\sheafc(\dtotalc)"', bend right, shift left=10.5] &   
+    & \cgamma{\opensetc_i}{\dtotalc} \in Ob(\setc) & \\
+{} \arrow[rr, "\dfuncc", Rightarrow, pos=.3, shorten <=3.5em, shorten >=8.85em, color=action] & & {} \\
+    & \opensetc_i \in Ob(\mathcal{\dbasec}^{\textcolor{base}{op}}) \arrow[uu, "\sheafc(\dtotalc)", bend left, shift left = 10.5, color=sheaf] \arrow[uu, "\sheafc(\dtotalc)"', bend right, shift left=10.5, color=sheaf] &
 \end{tikzcd}
@@
   \label{eq:artist:graphic_sections}
-  \cgamma{\opensetgc}{\gtotalc} \coloneqq \{\gsectionc^{i}:\opensetgc\rightarrow \gtotalc\}|_{i\in I}
+  \cgamma{\opensetgc}{\gtotalc} \coloneqq \textcolor{set}{\{}\gsectionc^{i}:\opensetgc\rightarrow \gtotalc\textcolor{set}{\}}|_{i\in I}
 \end{equation}
@@
   \begin{tikzcd}[row sep=huge]
-    {Ob(\textcolor{section}{\bm{Set}}) \ni \cgamma{\opensetgc_i}{\gtotalc}}                                                                                                  &  & \{\gsectionc^{j}:\opensetgc_i\rightarrow \gtotalc\}|_{j\in J} \arrow[ll, "\iota"', hook'] \\ &  & \\
-    \opensetgc_i \arrow[uu, "\sheafc(\gtotalc)"] \arrow[rruu, "\vartistc(\sheafc(\dtotalc))\coloneqq \sheaf_{\vartist}(\gtotalc)" description, dashed] &  & 
+    {\cgamma{\opensetgc_i}{\gtotalc}}                                                                                                  &  & \textcolor{set}{\{}\gsectionc^{j}:\opensetgc_i\rightarrow \gtotalc\textcolor{set}{\}}|_{j\in J} \arrow[ll, "\iota"', hook', color=set] \\ &  & \\
+    \opensetgc_i \arrow[uu, "\sheafc(\gtotalc)", color=sheaf] \arrow[rruu, "\vartistc(\sheafc(\dtotalc))\coloneqq \sheaf_{\vartist}(\gtotalc)" description, dashed, color=sheaf] &  & 
     \end{tikzcd}
@@
   \begin{tikzcd}
-    \dtotalc                        & \vindexc^*\dtotal \arrow[l, "\vindexc^*"', dashed]                                                                           \\
-    \dbasec \arrow[u, "\dsectionc"] & \gbasec \arrow[l, "\vindexc"] \arrow[lu, "\dsectionc \circ \vindexc" description] \arrow[u, "\vindexc^*\dsectionc"', dashed]
+    \dtotalc                        & \vindexc^*\dtotalc \arrow[l, "\vindexc^*"', dashed, color=artist]                                                                           \\
+    \dbasec \arrow[u, "\dsectionc", color=section] & \gbasec \arrow[l, "\vindexc", color=artist] \arrow[lu, "\dsectionc \circ \vindexc" description] \arrow[u, "\vindexc^*\dsectionc"', dashed, color=section]
     \end{tikzcd}
@@
   \label{eq:artist:data_gamma_pullback}
- \Gamma(\dbase, \dtotal) \xrightarrow{\vindex^*} \Gamma(\gbase, \vindex^*\dtotal)
+ \cgamma{\dbasec}{\dtotalc} \textcolor{artist}{\xrightarrow{\vindex^*}} \cgamma{\gbasec}{\vindexc^*\dtotalc}
 \end{equation}
 to generate a set of equivalent sections over the graphic $\Gamma(\gtotal, \vindex^*{\dtotal})$.
-As a consequence of \autoref{eq:artist:data_gamma_pullback}, there exists a data sheaf over graphic space $\sheaf: \mathcal{\gbase^{op}} \rightarrow \Gamma(\gtotal, \vindex^*{\dtotal})$ which is equivalent to
+As a consequence of \autoref{eq:artist:data_gamma_pullback}, there exists a data sheaf over graphic space
+\begin{equation}
+  \label{eq:artist:data_pullback_sheaf_gamma}
+\sheafc: \mathcal{\gbasec^{\textcolor{base}{op}}} \textcolor{sheaf}{\rightarrow} \cgamma{\gtotalc}{\vindexc^*{\dtotalc}}
+\end{equation}
+which is equivalent to 
 \begin{equation}
   \label{eq:artist:data_pullback_sheaf}
-  \sheafc_{\vindexc^*}: \mathcal{\gbasec^{op}} \rightarrow \textcolor{section}{\bm{Set}}
+  \sheafc_{\vindexc^*}: \mathcal{\gbasec^{\textcolor{base}{op}}} \textcolor{sheaf}{\rightarrow} \setc
 \end{equation}
@@
   \begin{tikzcd}
-    & {} \arrow[dd, "\vartistc", Rightarrow, shorten=1.25em] & \\
-\mathcal{\gbasec}^{op} \arrow[rr, "\sheafc_{\vindex_*}(\dtotalc)", bend left] \arrow[rr, "\sheafc_{\vartistc}(\gtotalc)"', bend right] &  & \textcolor{section}{\bm{Set}}\\
+    & {} \arrow[dd, "\vartistc", Rightarrow, shorten=1.25em, color=artist] & \\
+\mathcal{\gbasec}^{\textcolor{base}{op}} \arrow[rr, "\sheafc_{\vindexc_*}(\dtotalc)", bend left, color=sheaf] \arrow[rr, "\sheafc_{\vartistc}(\gtotalc)"', bend right, color=sheaf] &  & {\setc}\\
     & {} &                            
@@
   \label{eq:artist:sheaf_map}
-  \vartistc: \sheafc_{\vindexc^*}(\dtotalc) \rightarrow \sheafc_{\vartistc}(\gtotalc)
+  \vartistc: \sheafc_{\vindexc^*}(\dtotalc) \textcolor{artist}{\rightarrow} \sheafc_{\vartistc}(\gtotalc)
 \end{equation}
@@
   \label{eq:artist:inclusion_morphisms}
-  \begin{tikzcd}
-    {\Gamma(\opensetc_1, \dtotalc)} \arrow[dd, "\vartistc_{\opensetc_1}"'] &                                                                                                                                   &  &                                                                                                         & {\Gamma(\opensetc_2,\dtotal)} \arrow[dd, "\vartistc_{\opensetc_2}"] \arrow[llll, "\iota^*"', hook'] \\
-                                                                           & \opensetc_1 \arrow[lu, "\sheafc_{\vindex^*}" description] \arrow[ld, "\sheafc_{\vartistc}" description] \arrow[rr, "\iota", hook] &  & \opensetc_2 \arrow[ru, "\sheafc_{\vindex^*}" description] \arrow[rd, "\sheafc_{\vartistc}" description] & \\
-    {\Gamma(\opensetc_1,\gtotalc)} & &  && {\Gamma(\opensetc_2,\gtotalc)} \arrow[llll, "\iota^*", hook']                                       
+  \begin{tikzcd}[row sep=huge]
+    {\cgamma{\opensetgc_1}{\dtotalc}}\arrow[dd, "\vartistc_{\opensetgc_1}"', color=artist] &                                                                                                                                   &  & & {\cgamma{\opensetgc_2}{\dtotalc}} \arrow[dd, "\vartistc_{\opensetgc_2}", color=artist] \arrow[llll, "\iota^*"', hook', color=set] \\
+    & \opensetgc_1 \arrow[lu, "\sheafc_{\vindexc^*}" description, color=sheaf] \arrow[ld, "\sheafc_{\vartistc}" description, color=sheaf] \arrow[rr, "\iota", hook, color=base] &  & \opensetgc_2 \arrow[ru, "\sheafc_{\vindexc^*}" description, color=sheaf] \arrow[rd, "\sheafc_{\vartistc}" description, color=sheaf] & \\
+    {\cgamma{\opensetgc_1}{\gtotalc}} & &  && {\cgamma{\opensetgc_2}{\gtotalc}} \arrow[llll, "\iota^*", hook', color=set]                                       
     \end{tikzcd}
 \end{equation}
+Given that $\opensetg \subseteqq \gbasec$, \autoref{eq:artist:inclusion_morphisms} can be generalized to sections over $\gbase$.
+This fact can be combined with \autoref{eq:artist:data_gamma_pullback} to describe the path from data sections over data space to graphic sections over graphic space
+\begin{equation}
+  \label{eq:artist:gamma_map}
+  \cgamma{\dbasec}{\dtotalc} \textcolor{artist}{\xrightarrow{\vindexc^*}} \cgamma{\gbasec}{\dtotalc} \textcolor{artist}{\xrightarrow{\vartist}} \cgamma{\gbasec}{\gtotalc}
+\end{equation}
 
-In \autoref{eq:artist:morphism-of-sheaf-data}, $\dfuncc$ was introduced as a morphism between sheaf functors $\sheafc^{\vindex^*}$.
-Since the artist is a functor $\vartist: \sheaf^{\vindex^*} \rightarrow \sheaf_{\vartist}$, it follows from \autoref{eq:related-work:equivariance:functor_diagram} that
+which means that the artist, pulled back through $\vindex^*$ is a map of sections 
 \begin{equation}
+  \label{eq:artist:section_map}
+  \vartistc_{\vindexc^*}: \dsectionc \textcolor{artist}{\mapsto}\gsectionc
+\end{equation}
+such that for every dataset, the artists constructs a corresponding graphic generating function.
+Besides preserving inclusions, as described in \autoref{eq:artist:inclusion_morphisms}, the artist function is constructed such that transformations on $\dsection$ and $\gsection$ are equivariant with respect to the morphism between data sheaf functors $\dfunc$ which was introduced in \autoref{eq:artist:morphism-of-sheaf-data}.
+
+\note{is artist functor by virtue of being natural transform or do we define artist as functor between functors}
+
+$\vartist: \sheaf^{\vindex^*} \rightarrow \sheaf_{\vartist}$, it follows from \autoref{eq:related-work:equivariance:functor_diagram} that an arbitrary morphism $\dfunc$ on the data sheaf has a compatible morphism $\dfunc^{\prime}$ on the graphic sheaf.
+\begin{equation}
+  \label{eq:artist:morphism_of_functors}
   \begin{tikzcd}
-    \sheafc_{\vindex^{*}} \arrow[d, "\dfuncc"'] \arrow[r, "\vartistc"] & \sheafc_{\vartist} \arrow[d, "\dfuncc^{\prime} \coloneqq\vartist(\dfuncc)"] \\
-    \sheafc_{\vindex^{*}} \arrow[r, "\vartistc"']                      & \sheafc_{\vartist}                                                         
+    \sheafc_{\vindexc^{*}} \arrow[d, "\dfuncc"', color=action] \arrow[r, "\vartistc", color=artist] & \sheafc_{\vartistc} \arrow[d, "\dfuncpc \coloneqq\vartistc(\dfuncc)", color=action] \\
+    \sheafc_{\vindexc^{*}} \arrow[r, "\vartistc"', color=artist]                      & \sheafc_{\vartistc}     
     \end{tikzcd}
 \end{equation}
-an arbitrary morphism $\dfunc$ on the data sheaf has a compatible morphism $\dfunc^{\prime}$ on the graphic sheaf.
+As a consequence of \autoref{eq:artist:morphism_of_functors},
 

### 2021-11-30 paper/main.tex
put in varphi
nat transform diagram

@@
   \label{eq:related-work:equivariance:natural_transform_commute}
-  \begin{tikzcd}
-    \textcolor{source}{c} \arrow[rr, "F", color=functor, alpha=.5] \arrow[rrrr, "G", bend left, color=functor, alpha=.5] \arrow[dd, "f"', color=source] &  & \textcolor{target}{F(c)} \arrow[dd, "F(f)"', color=target] \arrow[rr, "\alpha_{c}", color=nattran] &  & \textcolor{target}{G(c)}\arrow[dd, "G(f)", color=target] \\
-                                                                    &  & {} \arrow[rr, "\alpha" description, dashed, color=nattran]       &  & {}                      \\
-    \textcolor{source}{c^{\prime}} \arrow[rr, "F", color=functor, alpha=.5] \arrow[rrrr, "G", bend right, color=functor, alpha=.5]        &  & \textcolor{target}{F(c^{\prime})} \arrow[rr, "\alpha_{c^{\prime}}", color=nattran]   &  & \textcolor{target}{G(c^{\prime})}
+  \begin{tikzcd}[column sep=huge]
+    \textcolor{target}{F(c)} \arrow[dd, "F(f)"', color=target] \arrow[rr, "\alpha_{c}", bend left, color=nattran] & \textcolor{source}{c} \arrow[dd, "f" description, color=source] \arrow[r, "G"', color=functor] \arrow[l, "F", color=functor] & \textcolor{target}{G(c)} \arrow[dd, "G(f)", color=target] \\
+    {} \arrow[rr, dotted, color=nattran] & & {}                      \\
+    \textcolor{target}{F(c^{\prime})} \arrow[rr, "\alpha_{c^{\prime}}"', bend right, color=nattran] & \textcolor{source}{c^{\prime}} \arrow[l, "F"', color=functor] \arrow[r, "G", color=functor] & \textcolor{target}{G(c^{\prime})}
     \end{tikzcd}
@@
 
+\note{Should this move to down where we describe equivariance? It's not used at all currently in this section}
 \begin{definition} 
@@
   \begin{tikzcd}[row sep=huge]
-    {\cgamma{\opensetgc_i}{\gtotalc}}                                                                                                  &  & \textcolor{set}{\{}\gsectionc^{j}:\opensetgc_i\rightarrow \gtotalc\textcolor{set}{\}}|_{j\in J} \arrow[ll, "\iota"', hook', color=set] \\ &  & \\
+    {\cgamma{\opensetgc_i}{\gtotalc}}&  & \textcolor{set}{\{}\gsectionc^{j}:\opensetgc_i\rightarrow \gtotalc\textcolor{set}{\}}|_{j\in J} \arrow[ll, "\iota"', hook', color=set] \\ &  & \\
     \opensetgc_i \arrow[uu, "\sheafc(\gtotalc)", color=sheaf] \arrow[rruu, "\vartistc(\sheafc(\dtotalc))\coloneqq \sheaf_{\vartist}(\gtotalc)" description, dashed, color=sheaf] &  & 
@@
     \dtotalc                        & \vindexc^*\dtotalc \arrow[l, "\vindexc^*"', dashed, color=artist]                                                                           \\
-    \dbasec \arrow[u, "\dsectionc", color=section] & \gbasec \arrow[l, "\vindexc", color=artist] \arrow[lu, "\dsectionc \circ \vindexc" description] \arrow[u, "\vindexc^*\dsectionc"', dashed, color=section]
+    \dbasec \arrow[u, "\dsectionc", color=section] & \gbasec \arrow[l, "\vindexc", color=artist] \arrow[lu, "\dsectionc \circ \vindexc" description, color=section] \arrow[u, "\vindexc^*\dsectionc"', dashed, color=section]
     \end{tikzcd}
@@
   \label{fig:artist:pushforward_xi}
-  \caption{The map $\vindex^*$ is a map from a repeation of the 1D fiber over every point $\dbasepoint \in \dbase$ in data space, turning the fiber into a 2D plane of repeating values over the corresponding region $\dbasepoint = \vindex(\gbasepoint)$ in the graphic space \gbase.
+  \caption{The map $\vindex^*$ is a map from a repetition of the 1D fiber over every point $\dbasepoint \in \dbase$ in data space, turning the fiber into a 2D plane of repeating values over the corresponding region $\dbasepoint = \vindex(\gbasepoint)$ in the graphic space \gbase.
 }

### 2021-11-30 paper/main.tex
reworked artist d"

@@
 Following from \autoref{eq:related-work:equivariance:natural_transform_commute}, the artist specifies that there are compatible inclusion maps in the data and graphic spaces
+
 \begin{equation}
-  \label{eq:artist:inclusion_morphisms}
-  \begin{tikzcd}[row sep=huge]
-    {\cgamma{\opensetgc_1}{\dtotalc}}\arrow[dd, "\vartistc_{\opensetgc_1}"', color=artist] &                                                                                                                                   &  & & {\cgamma{\opensetgc_2}{\dtotalc}} \arrow[dd, "\vartistc_{\opensetgc_2}", color=artist] \arrow[llll, "\iota^*"', hook', color=set] \\
-    & \opensetgc_1 \arrow[lu, "\sheafc_{\vindexc^*}" description, color=sheaf] \arrow[ld, "\sheafc_{\vartistc}" description, color=sheaf] \arrow[rr, "\iota", hook, color=base] &  & \opensetgc_2 \arrow[ru, "\sheafc_{\vindexc^*}" description, color=sheaf] \arrow[rd, "\sheafc_{\vartistc}" description, color=sheaf] & \\
-    {\cgamma{\opensetgc_1}{\gtotalc}} & &  && {\cgamma{\opensetgc_2}{\gtotalc}} \arrow[llll, "\iota^*", hook', color=set]                                       
+  \begin{tikzcd}
+    \cgamma{\opensetgc_1}{\dtotalc} \arrow[dd, "\iota^*"', hook, color=set] \arrow[rrrr, "\vartistc_{\opensetgc_1}", bend left, color=artist] &  & \opensetgc_1 \arrow[ll, "\sheafc_{\vindexc^*}", color=sheaf] \arrow[rr, "\sheafc_{\vartistc}"', color=sheaf] &  & \cgamma{\opensetgc_1}{\gtotalc} \arrow[dd, "\iota^*", hook', color=set] \\
+    {} \arrow[rrrr, dotted, color=artist] &&&& {} \\
+    \cgamma{\opensetgc_2}{\dtotalc} \arrow[rrrr, "\vartistc_{\opensetgc_2}"', bend right, color=artist]                            &  & \opensetgc_2 \arrow[ll, "\sheafc_{\vindexc^*}"', color=sheaf] \arrow[rr, "\sheafc_{\vartistc}", color=sheaf] \arrow[uu, "\iota" description, hook, color=base] &  & \cgamma{\opensetgc_2}{\gtotalc}
     \end{tikzcd}

### 2021-12-01 paper/main.tex
reworking sheaf diagram

@@
 We model data as the sheaf  $\sheafc({\dtotalc}): \mathcal{\dbasec}^{op} \textcolor{sheaf}{\rightarrow} \setc$.
-The objects of \textbf{Set} in this sheaf are the sets of sections introduced in \autoref{eq:related-work:continuity:sections} and the morphisms are the inclusion maps in \autoref{eq:related-work:continuity:presheaf}.
+\begin{equation}
+  \label{eq:artist:morphism-of-sheaf-data}
+  \begin{tikzcd}
+    & \cgamma{\opensetc_i}{\dtotalc} \in Ob(\setc) & \\
+    {} & & {} \\
+    & \opensetc_i \in Ob(\mathcal{\dbasec}^{\textcolor{base}{op}}) \arrow[uu, "\sheafc(\dtotalc)", shift left = 10.5, color=sheaf] &   
+\end{tikzcd}
+\end{equation}
+
+The objects of \textbf{Set} in this sheaf are the sets of sections $\Gamma(\dbase, \dtotal)$ introduced in \autoref{eq:related-work:continuity:sections} and the morphisms are the inclusion maps $(\iota, \iota^*)$ introduced in \autoref{eq:related-work:continuity:presheaf}.
+
 \begin{equation}
@@
 \end{equation}
+
 The map $\dfunc: \sheaf(\dtotal) \rightarrow \sheaf(\dtotal)$ is any arbitrary morphism between sheaf functors $\sheaf(\dtotal)$.

### 2021-12-01 paper/main.tex
splitting out continuity and equivariance

@@
 
-\subsection{Natural Transformation of Data to Graphic}
+\subsection{Continuity}
 \label{sec:artist:category}
@@
 \end{equation}
-
 The objects of \textbf{Set} in this sheaf are the sets of sections $\Gamma(\dbase, \dtotal)$ introduced in \autoref{eq:related-work:continuity:sections} and the morphisms are the inclusion maps $(\iota, \iota^*)$ introduced in \autoref{eq:related-work:continuity:presheaf}.
-
-\begin{equation}
-  \label{eq:artist:morphism-of-sheaf-data}
-  \begin{tikzcd}
-    & \cgamma{\opensetc_i}{\dtotalc} \in Ob(\setc) & \\
-{} \arrow[rr, "\dfuncc", Rightarrow, pos=.3, shorten <=3.5em, shorten >=8.85em, color=action] & & {} \\
-    & \opensetc_i \in Ob(\mathcal{\dbasec}^{\textcolor{base}{op}}) \arrow[uu, "\sheafc(\dtotalc)", bend left, shift left = 10.5, color=sheaf] \arrow[uu, "\sheafc(\dtotalc)"', bend right, shift left=10.5, color=sheaf] &
-\end{tikzcd}
-\end{equation}
-
-The map $\dfunc: \sheaf(\dtotal) \rightarrow \sheaf(\dtotal)$ is any arbitrary morphism between sheaf functors $\sheaf(\dtotal)$.
-Following from \autoref{eq:related-work:equivariance:natural-transform}, \dfunc\ is a natural transform, which means all inclusions $\iota, \iota^*$ are preserved as part of any $\dfunc$ transformation.
-
-\note{Should this move to down where we describe equivariance? It's not used at all currently in this section}
-\begin{definition} 
-  \label{def:category:F} The fiber category $\mathcal{\dfiberc}$ is a monoidal category\cite{milewskiCategoryTheoryProgrammers}, which means it is a category equipped with a bifunctur $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$
-  \begin{itemize}
-    \item \textit{object} a monoid object \dfiber. A monoid is a set with a binary operator that is associative, closed, and for which the set contains an identity element $i$ \cite{nlan:monoid}.
-    \item \textit{morphisms} $\dfiber \otimes \dfiber \rightarrow \dfiber$ and unit $i \rightarrow \dfiber$
-  \end{itemize}
-\end{definition}
-The bifunctor $\otimes$ combines categories in a way that  preserves identity and composition \cite{fongInvitationAppliedCategory2019}.
-For example, given a pair of categories $\mathcal{\dfiber_a},\mathcal{\dfiber_b}$, the bifunctor yields a set of pairs $\{(\dfiber_a, \dfiber_b) | \dfiber_a \in \mathcal{\dfiber_a}, \dfiber_b \in \mathcal{\dfiber_b}\}$.
-For the associated morphisms $g: \mathcal{\dfiber_a}\rightarrow \mathcal{\dfiber_a}$ and $h: \mathcal{\dfiber_b}\rightarrow \mathcal{\dfiber_b}$, the function $(g\times h): (\mathcal{\dfiber_a}\times \mathcal{\dfiber_b}) \rightarrow (\mathcal{\dfiber_a}\times \mathcal{\dfiber_b})$ is pointwise $(g\times h)(\dfiber_a, \dfiber_b)\coloneqq (g(\dfiber_a), h(\dfiber_b))$.
-Because of the bifunctor, objects of the category $\mathcal{\dfiber}$ can be the product of any arbitrary number and types of categories, including \textbf{Set}, \textbf{Graph}, and topological spaces \textbf{Top}.
-
-The set of all morphism of the fiber category $Hom_{\mathcal{\dfiber}}(\dfiber, \dfiber)$ are all possible morphisms in a given category.
-This includes any of the functions that serve as the basis of equivariance, such as the binary operations, group actions, and measurement scale discussed in  \autoref{sec:related-work:equivariance}.
-The generalization to $Hom_\mathcal{\dfiber}$ also allows for the inclusion of structures such as monoid actions\cite{barrCategoryTheoryComputing}, which provide a way of applying partial order relations to data\cite{fongInvitationAppliedCategory2019}, such as to build multi-ranked indicators\cite{bruggemannRankingPrioritizationMultiindicator2011}.
-
-When a fiber bundle is trivial, we can dispense with localization and directly consider the fiber maps $Hom(\dfiber_{\dbasepoint}, \dfiber_{\dbasepoint})$ over a point \dbasepoint\ as the basis of our equivariance.
-This is because a sheaf $\sheaf(\dtotal)$ over a limit of open sets that contain a point $\dbasepoint \in \dbase$ is approximately the same as a sheaf over a point \dbasepoint\ and is called the stalk $\sheaf(\dtotal)|_{\dbasepoint}\coloneqq \lim\limits_{\openset\ni \dbasepoint} \Gamma(\openset, \dtotal)$ \cite{StalkSheaf2019}.
-The fiber space over a point is embedded in the stalk over that point $\dfiber_{\dbasepoint}\hookrightarrow \mathscr{\dfiber}_{\dbasepoint}$.
-The sheaf maps \dfunc\ restricted to the stalk are members of $hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$.
-The maps $\dfunc \in Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ include the types of data transformations described in \autoref{sec:related-work:equivariance}In \autoref{sec:artist:construction}, we assume that the fiber bundles are trivial.
 
@@
 
-\note{is artist functor by virtue of being natural transform or do we define artist as functor between functors}
+\subsection{Equivariance}
+
+\begin{equation}
+  \label{eq:artist:morphism-of-sheaf-data}
+  \begin{tikzcd}
+    & \cgamma{\opensetc_i}{\dtotalc} \in Ob(\setc) & \\
+{} \arrow[rr, "\dfuncc", Rightarrow, pos=.3, shorten <=3.5em, shorten >=8.85em, color=action] & & {} \\
+    & \opensetc_i \in Ob(\mathcal{\dbasec}^{\textcolor{base}{op}}) \arrow[uu, "\sheafc(\dtotalc)", bend left, shift left = 10.5, color=sheaf] \arrow[uu, "\sheafc(\dtotalc)"', bend right, shift left=10.5, color=sheaf] &
+\end{tikzcd}
+\end{equation}
+
+The map $\dfunc: \sheaf(\dtotal) \rightarrow \sheaf(\dtotal)$ is any arbitrary morphism between sheaf functors $\sheaf(\dtotal)$.
+Following from \autoref{eq:related-work:equivariance:natural-transform}, \dfunc\ is a natural transform, which means all inclusions $\iota, \iota^*$ are preserved as part of any $\dfunc$ transformation.
+
+\begin{definition} 
+  \label{def:category:F} The fiber category $\mathcal{\dfiberc}$ is a monoidal category\cite{milewskiCategoryTheoryProgrammers}, which means it is a category equipped with a bifunctur $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$
+  \begin{itemize}
+    \item \textit{object} a monoid object \dfiber. A monoid is a set with a binary operator that is associative, closed, and for which the set contains an identity element $i$ \cite{nlan:monoid}.
+    \item \textit{morphisms} $\dfiber \otimes \dfiber \rightarrow \dfiber$ and unit $i \rightarrow \dfiber$
+  \end{itemize}
+\end{definition}
+The bifunctor $\otimes$ combines categories in a way that  preserves identity and composition \cite{fongInvitationAppliedCategory2019}.
+For example, given a pair of categories $\mathcal{\dfiber_a},\mathcal{\dfiber_b}$, the bifunctor yields a set of pairs $\{(\dfiber_a, \dfiber_b) | \dfiber_a \in \mathcal{\dfiber_a}, \dfiber_b \in \mathcal{\dfiber_b}\}$.
+For the associated morphisms $g: \mathcal{\dfiber_a}\rightarrow \mathcal{\dfiber_a}$ and $h: \mathcal{\dfiber_b}\rightarrow \mathcal{\dfiber_b}$, the function $(g\times h): (\mathcal{\dfiber_a}\times \mathcal{\dfiber_b}) \rightarrow (\mathcal{\dfiber_a}\times \mathcal{\dfiber_b})$ is pointwise $(g\times h)(\dfiber_a, \dfiber_b)\coloneqq (g(\dfiber_a), h(\dfiber_b))$.
+Because of the bifunctor, objects of the category $\mathcal{\dfiber}$ can be the product of any arbitrary number and types of categories, including \textbf{Set}, \textbf{Graph}, and topological spaces \textbf{Top}.
+
+The set of all morphism of the fiber category $Hom_{\mathcal{\dfiber}}(\dfiber, \dfiber)$ are all possible morphisms in a given category.
+This includes any of the functions that serve as the basis of equivariance, such as the binary operations, group actions, and measurement scale discussed in  \autoref{sec:related-work:equivariance}.
+The generalization to $Hom_\mathcal{\dfiber}$ also allows for the inclusion of structures such as monoid actions\cite{barrCategoryTheoryComputing}, which provide a way of applying partial order relations to data\cite{fongInvitationAppliedCategory2019}, such as to build multi-ranked indicators\cite{bruggemannRankingPrioritizationMultiindicator2011}.
+
+When a fiber bundle is trivial, we can dispense with localization and directly consider the fiber maps $Hom(\dfiber_{\dbasepoint}, \dfiber_{\dbasepoint})$ over a point \dbasepoint\ as the basis of our equivariance.
+This is because a sheaf $\sheaf(\dtotal)$ over a limit of open sets that contain a point $\dbasepoint \in \dbase$ is approximately the same as a sheaf over a point \dbasepoint\ and is called the stalk $\sheaf(\dtotal)|_{\dbasepoint}\coloneqq \lim\limits_{\openset\ni \dbasepoint} \Gamma(\openset, \dtotal)$ \cite{StalkSheaf2019}.
+The fiber space over a point is embedded in the stalk over that point $\dfiber_{\dbasepoint}\hookrightarrow \mathscr{\dfiber}_{\dbasepoint}$.
+The sheaf maps \dfunc\ restricted to the stalk are members of $hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$.
+The maps $\dfunc \in Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ include the types of data transformations described in \autoref{sec:related-work:equivariance}In \autoref{sec:artist:construction}, we assume that the fiber bundles are trivial.
+
+\note{}
 

### 2021-12-01 paper/main.tex
reshuffled related-work section

@@
 We define a general purpose visualization library as one that provides non domain specific building block components\cite{wongsuphasawatNavigatingWideWorld2021} for building visualizations, for example functions for converting data to color or encoding data as dots.
-In this section, we describe how visualization libraries attempt this goal and discuss work that formally describes what properties of data should be preserved in a visualization.
-We restrict the properties of data that should be preserved to
+We propose that visualization library components should preserve continuity and equivariance, which we define as
 
@@
   \item [continuity] how elements in a dataset are organized, e.g. discrete rows in a table, networked nodes, pixels in an image, points on a line
-  \item [equivariance] functions on data that have an equivalent effect on the graphical representation, e.g.
+  \item [equivariance] data transformations that have an equivalent effect on the graphical representation, e.g.
 rotating a matrix has a matching rotation of the image, translating the points on a line has a matching visual shift in the line plot
@@
 
+In \autoref{sec:related-work:continuity} we discuss how data continuity influences visualization library architecture and in \autoref{sec:related-work:equivariance} we summarize how visualization researchers have approached equivariance.
+
 \subsection{Continuity}
+\label{sec:related-work:continuity}
 \begin{figure}[!h]
@@
 Images underpin scientific visualization tools such as Napari\cite{nicholas_sofroniew_2021_4533308} and ImageJ\cite{schneiderNIHImageImageJ2012} and the digital humanities oriented ImagePlot\cite{studiesCulturevisImageplot2021} macro; the need to visualize and manipulate graphs has spawned tools like Gephi\cite{bastianGephiOpenSource2009}, Graphviz\cite{ellsonGraphvizOpenSource2002}, and Networkx\cite{HagbergExploringNetwork2008}.
+
+\subsection{Equivariance}
+\label{sec:related-work:equivariance}
+When introducing the retinal variables, Bertin informally specifies that continuity is preserved in the mark and defines equivariance constraints in terms of data and visual variables being selective, associative, ordered, or quantitative\cite{bertinSemiologyGraphicsDiagrams2011a}.
+In the \textit{A Presentation Tool}(APT) model, Mackinlay embeds the continuity constraint in the choice of visualization type and generalizes the equivariance constraint to preserving a binary operator from one domain to another.
+The algebraic model of visualization\cite{kindlmannAlgebraicProcessVisualization2014}, proposed by Kindlmann and Scheidegger, restricts equivariance to invertible transformations.
+
+\begin{figure}[!h]
+  \label{fig:related-work:equivariance}
+  \includegraphics[width=\columnwidth]{equiv.png}
+  \caption{Equivariance is that a transformation on the data has a corresponding transformation in the graphical representation.
+For example, in this figure the data is scaled by a factor 10.
+Equivalently the line plot is scaled by factor of 10, resulting in a shrunken line plot.
+Either a transformation on the data side can induce a transformation on the visual side, or a transformation on the visual side indicates that there is also a transformation on the data side.
+}
+\end{figure}
+
+\section{Algebraic Topology And Category Theory}
 
@@
 
-A fiber bundle has many sections, and the set of all sections over an open set $openset \subseteq \dtotal$
+A fiber bundle has many sections, and the set of all sections over an open set $\openset \subseteq \dtotal$
 \begin{equation}
   \label{eq:related-work:continuity:sections}
-\cgamma{\opensetc}{\dtotalc} \coloneqq \{\dsectionc^{i}: \opensetc\rightarrow \dtotalc\}|_{i\in I}
+  \begin{split}
+  \cgamma{\opensetc}{\dtotalc} \coloneqq & \{\dsectionc^{0} \cdots \dsectionc^{i} \cdots \dsectionc^{n}\} 
+  & \dsectionc: \opensetc\rightarrow \dtotalc
+  \end{split}
 \end{equation}
@@
 Sheaves provide a mathematical abstraction of data, the space it lives over, and relationships between subsets.
-
-\subsection{Equivariance}
-\label{sec:related-work:equivariance}
-When introducing the retinal variables, Bertin informally specifies that continuity is preserved in the mark and defines equivariance constraints in terms of data and visual variables being selective, associative, ordered, or quantitative\cite{bertinSemiologyGraphicsDiagrams2011a}.
-In the \textit{A Presentation Tool}(APT) model, Mackinlay embeds the continuity constraint in the choice of visualization type and generalizes the equivariance constraint to preserving a binary operator from one domain to another.
-The algebraic model of visualization\cite{kindlmannAlgebraicProcessVisualization2014}, proposed by Kindlmann and Scheidegger, restricts equivariance to invertible transformations.
-
-\begin{figure}[!h]
-  \label{fig:related-work:equivariance}
-  \includegraphics[width=\columnwidth]{equiv.png}
-  \caption{Equivariance is that a transformation on the data has a corresponding transformation in the graphical representation.
-For example, in this figure the data is scaled by a factor 10.
-Equivalently the line plot is scaled by factor of 10, resulting in a shrunken line plot.
-Either a transformation on the data side can induce a transformation on the visual side, or a transformation on the visual side indicates that there is also a transformation on the data side.
-}
-\end{figure}
 

### 2021-12-01 paper/main.tex
new intro to intro math section

@@
 \section{Algebraic Topology And Category Theory}
+Category theory can be used to express software specifications because it provides a method of expressing structure and how that structure can be composed \cite{wielsManagementEvolvingSpecifications1998} algebraic topology provides a rich language for expressing data continuity.
+In this section, we introduce concepts from algebraic topology and category theory that we will use in \autoref{sec:artist:category} to develop a formal model of specifying visualization component.
 

### 2021-12-01 paper/main.tex
rework gamma tau to remove indexing

--- 2021-12-01 paper/main.tex
@@
 \newtheorem{prop}{Proposition} %math?
+\usepackage[switch]{lineno}
+\renewcommand{\linenumberfont}{\normalfont\bfseries\small\color{lightgray}}
 \usepackage{minted}
@@
 \hyphenation{op-tical net-works semi-conduc-tor}
-
-\usepackage[]{lineno}
 
@@
 
-\section{Algebraic Topology And Category Theory}
+\section{Algebraic Topology \& Category Theory}
+\note{potentially move this to appendix, if so push Butler & Vickers up to RW}
 
-\subsubsection{Fiber Bundles}
+Category theory can be used to express software specifications because it provides a method of expressing structure and how that structure can be composed \cite{wielsManagementEvolvingSpecifications1998} algebraic topology provides a rich language for expressing data continuity.
+In this section, we introduce concepts from algebraic topology and category theory that we will use in \autoref{sec:artist:category} to develop a formal model of specifying visualization component.
+
+\subsection{Fiber Bundles}
 \label{sec:related-work:fiber-bundles}
@@
 As proposed by Butler et al \cite{butlerVisualizationModelBased1989,butlerVectorBundleClassesForm1992}, data can be represented as a map from the base space \dbase\ to the fiber space \dfiber.
-This map is called a section $\dsection: \dbase \rightarrow \dtotal$ of the fiber bundle.
+This map is called a section $\dsection: \dbase \rightarrow \dtotal$ of the fiber bundle
 
@@
 
-A fiber bundle has many sections, and the set of all sections over an open set $\openset \subseteq \dtotal$
+and the set of all sections over \dbase\ is denoted $\Gamma(\dbase, \dtotal)$.
+Sections can also be defined locally, meaning over an open set $\openset \subseteq \dtotal$, and the set of all sections over an arbitrary open set is
+
 \begin{equation}
@@
   \begin{split}
-  \cgamma{\opensetc}{\dtotalc} \coloneqq & \{\dsectionc^{0} \cdots \dsectionc^{i} \cdots \dsectionc^{n}\} 
-  & \dsectionc: \opensetc\rightarrow \dtotalc
+  \cgamma{\opensetc}{\dtotalc|_{\opensetc}} \coloneqq \{&\dsectionc^{0}, \cdots, \dsectionc^{i}, \cdots, \dsectionc^{n}\} \\
+   &\dsectionc: \opensetc\rightarrow \dtotalc|_{\openset}
   \end{split}
@@
 
-\subsubsection{Sheaves}
+\subsection{Sheaves}
 Sheaves can be thought of as an "algebraic data structure" \cite{ghristElementaryAppliedTopology2014} for data that lives over topological spaces.
@@
 
-\subsubsection{Category Theory}
+\subsection{Category Theory}
 \label{sec:related-work:equivariance:category}

### 2021-12-01 paper/main.tex
E restricted to U in presheaf equation

@@
   \begin{tikzcd}[column sep=small]
-    {\cgamma{\opensetc_1}{\dtotalc}} & {} & {\cgamma{\opensetc_2}{\dtotalc}} \arrow[ll, "\iota^*"', hook'] \\& &\\
+    {\cgamma{\opensetc_1}{\dtotalc|_{\opensetc_1}}} & {} & {\cgamma{\opensetc_2}{\dtotalc|_{\opensetc_2}}} \arrow[ll, "\iota^*"', hook'] \\& &\\
     \opensetc_1 \arrow[uu, "\sheafc_{\dtotal}"] \arrow[rr, "\iota"', hook] & {} \arrow[uu, "\sheafc_{\dtotal}" description, dotted] & \opensetc_2 \arrow[uu, "\sheafc_\dtotal"']                    

### 2021-12-01 paper/main.tex
more E|U

@@
 Sheaves can be thought of as an "algebraic data structure" \cite{ghristElementaryAppliedTopology2014} for data that lives over topological spaces.
-A presheaf $\sheafc$ is a map from the open set to the set of sections $\sheafc: \opensetc \rightarrow \cgamma{\openset}{ \dtotal}$ \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}.
+A presheaf $\sheafc$ is a map from the open set to the set of sections $\sheafc: \opensetc \rightarrow \cgamma{\opensetc}{ \dtotalc|_{\opensetc}}$ \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}.
 The presheaf preserves inclusion maps $\iota$ between open sets and the inclusion map $\iota^*$ between the sets of sections.
@@
 
-This means that a smaller space $\openset_1$ is included in a larger space $\openset_2$ and a function that is continuous over a larger space $\openset_2$ is continuous over a subspace $\openset_1 \subset\openset_2$.
-Sheaves are presheafs where sets of sections are defined over unions of open sets over a topology $\bigcup_{j \in I} \openset)_j \in \dtotal$ \cite{spanier1989algebraic, nlab:sheaf}.
+This means that a smaller space $\openset_1$ is included in a larger space $\openset_2$ and a function $\dsection \in \Gamma(\openset, \dtotal|_{\openset})$ that is continuous over a larger space $\openset_2$ is continuous over a subspace $\openset_1 \subset\openset_2$.
+Sheaves are presheafs where sets of sections are defined over unions of open sets over a topology $\bigcup_{j \in I} \openset)_j \in \dbase$ \cite{spanier1989algebraic, nlab:sheaf}.
 Sheaves provide a mathematical abstraction of data, the space it lives over, and relationships between subsets.

### 2021-12-01 paper/main.tex
switch side of E label

@@
     {} & & {} \\
-    & \opensetc_i \in Ob(\mathcal{\dbasec}^{\textcolor{base}{op}}) \arrow[uu, "\sheafc(\dtotalc)", shift left = 10.5, color=sheaf] &   
+    & \opensetc_i \in Ob(\mathcal{\dbasec}^{\textcolor{base}{op}}) \arrow[uu, "\sheafc(\dtotalc)"', shift left = 10.5, color=sheaf] &   
 \end{tikzcd}

### 2021-12-01 paper/main.tex
sheaf equation tweaking

--- 2021-12-01 paper/main.tex
@@
   \begin{tikzcd}
-    & \cgamma{\opensetc_i}{\dtotalc} \in Ob(\setc) & \\
+    & \cgamma{\opensetc}{\dtotalc|_{\opensetc}} \in Ob(\setc) & \\
     {} & & {} \\
-    & \opensetc_i \in Ob(\mathcal{\dbasec}^{\textcolor{base}{op}}) \arrow[uu, "\sheafc(\dtotalc)", shift left = 10.5, color=sheaf] &   
+    & \opensetc \in Ob(\mathcal{\dbasec}^{\textcolor{base}{op}}) \arrow[uu, "\sheafc(\dtotalc)"', shift left = 10, color=sheaf] &   
 \end{tikzcd}

### 2021-12-01 paper/main.tex
stripped out indexing

@@
   \label{eq:artist:graphic_sections}
-  \cgamma{\opensetgc}{\gtotalc} \coloneqq \textcolor{set}{\{}\gsectionc^{i}:\opensetgc\rightarrow \gtotalc\textcolor{set}{\}}|_{i\in I}
+    \begin{split}
+    \cgamma{\opensetgc}{\gtotalc|_{\opensetgc}} \coloneqq \{&\gsectionc^{0}, \cdots, \gsectionc^{i}, \cdots, \gsectionc^{n}\} \\
+     &\gsectionc: \opensetgc\rightarrow \gtotalc|_{\opensetgc}
+    \end{split}
+  \end{equation}
 \end{equation}
@@
   \begin{tikzcd}[row sep=huge]
-    {\cgamma{\opensetgc_i}{\gtotalc}}&  & \textcolor{set}{\{}\gsectionc^{j}:\opensetgc_i\rightarrow \gtotalc\textcolor{set}{\}}|_{j\in J} \arrow[ll, "\iota"', hook', color=set] \\ &  & \\
-    \opensetgc_i \arrow[uu, "\sheafc(\gtotalc)", color=sheaf] \arrow[rruu, "\vartistc(\sheafc(\dtotalc))\coloneqq \sheaf_{\vartist}(\gtotalc)" description, dashed, color=sheaf] &  & 
+    {\cgamma{\opensetgc}{\gtotalc|_{\opensetgc}}}&  & 
+    \textcolor{set}{\{}\gsectionc:\opensetgc\rightarrow \gtotalc\textcolor{set}{\}}
+    \arrow[ll, "\iota"', hook', color=set] \\ &  & \\
+    \opensetgc \arrow[uu, "\sheafc(\gtotalc)", color=sheaf] \arrow[rruu, "\vartistc(\sheafc(\dtotalc))\coloneqq \sheafc_{\vartistc}(\gtotalc)" description, dashed, color=sheaf] &  & 
     \end{tikzcd}
 \end{equation}
-The set of graphics reachable through an artist function is a subset $\{\gsection^{j}\}|_{j\in J}$ of the full set of graphics.
+The set of graphics reachable through an artist function is a subset of sections of the full set of graphics $\{\gsection\} \subset \Gamma(\opensetg, \gtotal|_{\opensetg})$.
 These sets of reachable graphics over open sets \opensetg\ are also objects of \textbf{Set}.

### 2021-12-01 paper/main.tex
pulled back artist

@@
   \label{eq:artist:data_pullback_sheaf_gamma}
-\sheafc: \mathcal{\gbasec^{\textcolor{base}{op}}} \textcolor{sheaf}{\rightarrow} \cgamma{\gtotalc}{\vindexc^*{\dtotalc}}
+\sheafc_{\vindexc^*}(\dtotalc): \mathcal{\gbasec^{\textcolor{base}{op}}} \textcolor{sheaf}{\rightarrow} \cgamma{\gbasec}{\vindexc^*{\dtotalc}}
 \end{equation}
@@
 \begin{equation}
+  \label{eq:artist:natural-transform-diagram}
   \begin{tikzcd}
-    & {} \arrow[dd, "\vartistc", Rightarrow, shorten=1.25em, color=artist] & \\
+    & {} \arrow[dd, "\vartistc", Rightarrow, shorten <=1.35em,  shorten >=1.25em, color=artist] & \\
 \mathcal{\gbasec}^{\textcolor{base}{op}} \arrow[rr, "\sheafc_{\vindexc_*}(\dtotalc)", bend left, color=sheaf] \arrow[rr, "\sheafc_{\vartistc}(\gtotalc)"', bend right, color=sheaf] &  & {\setc}\\
@@
 
-which means that the artist, pulled back through $\vindex^*$ is a map of sections 
+which means that the artist can be pulled back through $\vindex^*$\footnote{Defintion 3.68 in Seven Sketches in Compositionality by Fong and Spivak}
+\begin{equation}
+  \begin{tikzcd}
+    &                                                                                                                                                         & {} \arrow[dd, "\vartistc" description, Rightarrow, shorten <=1.35em,  shorten >=1.25em, color=artist] &       \\
+\mathcal{\dbasec}^{\textcolor{base}{op}} \arrow[r, "\vindexc^*", color=artist] & \mathcal{\gbasec}^{\textcolor{base}{op}} \arrow[rr, "\sheafc_{\vindex^*}(\dtotalc)", bend left, color=sheaf] \arrow[rr, "\sheaf_{\vartistc}(\gtotalc)"', bend right, color=sheaf] &                                                    & \setc \\
+    &                                                                                                                                                         & {}                                                 &
+\end{tikzcd}
+\end{equation}
+such that the artist can be constructed over data space 
+\begin{equation}
+  \begin{tikzcd}
+    & {} \arrow[dd, "\vartistc_{\vindexc^*}" description, Rightarrow, shorten <=1.35em,  shorten >=1.25em,, color=artist] &       \\
+\mathcal{\dbasec}^{\textcolor{base}{op}} \arrow[rr, "\sheafc_{\vindexc^*}(\dtotalc)\circ\vindexc^* = \sheafc(\dtotalc)", bend left, color=sheaf] \arrow[rr, "\sheafc_{\vartistc}(\gtotalc)"', bend right, color=sheaf] &                                                                 & \setc \\
+    & {}                                                              &      
+\end{tikzcd}
+\end{equation}
+
+which in turn provides a way to construct the artist as a map from data section in data space to corresponding graphic section
 \begin{equation}
@@
 such that for every dataset, the artists constructs a corresponding graphic generating function. 
-Besides preserving inclusions, as described in \autoref{eq:artist:inclusion_morphisms}, the artist function is constructed such that transformations on $\dsection$ and $\gsection$ are equivariant with respect to the morphism between data sheaf functors $\dfunc$ which was introduced in \autoref{eq:artist:morphism-of-sheaf-data}.
+\note{maybe mention here Munzner, Zeminsky, and the other dude on glyphs map back to data}
+\note{possibly more here as this is the type signature of the thing we implement}
 
 \subsection{Equivariance}
+Besides preserving inclusions, as described in \autoref{eq:artist:inclusion_morphisms}, the artist function is constructed such that transformations on $\dsection$ and $\gsection$ are equivariant with respect to the morphism between data sheaf functors $\dfunc$ which was introduced in \autoref{eq:artist:morphism-of-sheaf-data}.
 

### 2021-12-02 paper/main.tex
shuffle up sections explaination

@@
 
-As proposed by Butler et al \cite{butlerVisualizationModelBased1989,butlerVectorBundleClassesForm1992}, data can be represented as a map from the base space \dbase\ to the fiber space \dfiber.
-This map is called a section $\dsection: \dbase \rightarrow \dtotal$ of the fiber bundle
+ There is a map from a point on the base space \dbase\ which returns values in the fiber space \dfiber. This map is called a section $\dsection: \dbase \rightarrow \dtotal$ of the fiber bundle
 
@@
 
-and the set of all sections over \dbase\ is denoted $\Gamma(\dbase, \dtotal)$.
-Sections can also be defined locally, meaning over an open set $\openset \subseteq \dtotal$, and the set of all sections over an arbitrary open set is
-
+and the set of all sections over \dbase\ is denoted $\Gamma(\dbase, \dtotal)$ .
+Sections can also be defined locally, meaning over an open set $\openset \subseteq \dtotal$ and the set of all sections over an arbitrary open set is $\Gamma(\openset, \dtotal)$.
 \begin{equation}
@@
 \end{equation}
- is the set of all data sets with a shared fiber \dfiber\ and continuity \dbase. The set of all sections defined over the total base space is denoted $\Gamma(\dbase, \dtotal)$.
+Butler et al proposed that \cite{butlerVisualizationModelBased1989,butlerVectorBundleClassesForm1992} sections could be used to model data sets since they are maps from a point on the continuity to a point in the fiber, and therefore they encapsulate both.
+The set of sections is the space of all data sets with the same continuity and fields.
 
@@
   \begin{itemize}
-    \item \textit{objects} open sets $\openset_{i} \in \{openset\}$ in the topological space $(\dbase, \mathscr{T})$), including the empty set $\varnothing$ and the maximum set \dbase. 
+    \item \textit{objects} open sets $\openset_{i} \in \{\openset\}$ in the topological space $(\dbase, \mathscr{T})$), including the empty set $\varnothing$ and the maximum set \dbase. 
     \item \textit{morphisms} $\iota: \openset_{i} \hookrightarrow \openset_{j}$ for all  $\openset_i, \openset_j \subset \dtotal$
@@
 
-We model data as the sheaf  $\sheafc({\dtotalc}): \mathcal{\dbasec}^{op} \textcolor{sheaf}{\rightarrow} \setc$.
+We model a space of data sets with the same continuity and fields as the sheaf  $\sheafc({\dtotalc}): \mathcal{\dbasec}^{op} \textcolor{sheaf}{\rightarrow} \setc$.
 \begin{equation}
@@
 The objects of \textbf{Set} in this sheaf are the sets of sections $\Gamma(\dbase, \dtotal)$ introduced in \autoref{eq:related-work:continuity:sections} and the morphisms are the inclusion maps $(\iota, \iota^*)$ introduced in \autoref{eq:related-work:continuity:presheaf}.
+As mentioned in \autoref{re} sections $\dsection$ are
 
-The continuity of the graphic is encapsulated as a category $\mathcal{\mathcal{\gbase}}$ containing open set objects $\opensetg$, the complete set of opensets $\gbase$ that is the full space of the graphic, and the empty set.
-As with $\mathcal{\dbase}$, the morphisms between the open sets  $\opensetg$ are inclusion morphisms $\iota$.
-Over the graphic space $\opensetg \subseteq \gbase$ are a set of sections
-\begin{equation}
-  \label{eq:artist:graphic_sections}
-    \begin{split}
-    \cgamma{\opensetgc}{\gtotalc|_{\opensetgc}} \coloneqq \{&\gsectionc^{0}, \cdots, \gsectionc^{i}, \cdots, \gsectionc^{n}\} \\
-     &\gsectionc: \opensetgc\rightarrow \gtotalc|_{\opensetgc}
-    \end{split}
-  \end{equation}
-\end{equation}
-which are the set of all graphics with a shared base space and fiber and are a subcategory of \textbf{Set}. This allows us to represent the full space of graphics as a sheaf $\sheaf(\gtotal)$
-\begin{equation}
-  \label{eq:artist:artist_sheaf}
-  \begin{tikzcd}[row sep=huge]
-    {\cgamma{\opensetgc}{\gtotalc|_{\opensetgc}}}&  & 
-    \textcolor{set}{\{}\gsectionc:\opensetgc\rightarrow \gtotalc\textcolor{set}{\}}
-    \arrow[ll, "\iota"', hook', color=set] \\ &  & \\
-    \opensetgc \arrow[uu, "\sheafc(\gtotalc)", color=sheaf] \arrow[rruu, "\vartistc(\sheafc(\dtotalc))\coloneqq \sheafc_{\vartistc}(\gtotalc)" description, dashed, color=sheaf] &  & 
-    \end{tikzcd}
-\end{equation}
-The set of graphics reachable through an artist function is a subset of sections of the full set of graphics $\{\gsection\} \subset \Gamma(\opensetg, \gtotal|_{\opensetg})$.
-These sets of reachable graphics over open sets \opensetg\ are also objects of \textbf{Set}.
-This allows us to formulate the graphic output by the artist as a sheaf $\sheaf_{\vartist}$ from the graphic base space to the set of reachable graphics.
+\note{add transition sentence about artist }
+
+The continuity of the graphic is encapsulated as a category $\mathcal{\gbase}}$
+
+\begin{definition} The category of open sets $\mathcal{\gbase}$ in the topological space $(\dbase, \mathscr{T})$).
+This category consists of open set \textit{objects} $\opensetg \in \{\opensetg\}$, including the empty set \varnothing\ and the total set $\gtotal$, and inclusion \textmorphisms $\iota: \opensetg_i \rightarrow \opensetg_j$.
+\end{definition}
 
@@
   \label{eq:artist:data_pullback_sheaf}
-  \sheafc_{\vindexc^*}: \mathcal{\gbasec^{\textcolor{base}{op}}} \textcolor{sheaf}{\rightarrow} \setc
+  \sheafc_{\vindexc^*}(\dtotalc): \mathcal{\gbasec^{\textcolor{base}{op}}} \textcolor{sheaf}{\rightarrow} \setc
 \end{equation}
 
-Following from \autoref{eq:artist:data_pullback_sheaf} and the construction of the artist sheaf in \autoref{eq:artist:artist_sheaf}, we propose that the artist function $\vartist$ is a natural transformation
+As with $\mathcal{\dbase}$, the morphisms between the open sets  $\opensetg$ are inclusion morphisms $\iota$. Over the graphic space $\opensetg \subseteq \gbase$ are a set of sections 
+\begin{equation}
+  \label{eq:artist:graphic_sections}
+    \begin{split}
+    \cgamma{\opensetgc}{\gtotalc|_{\opensetgc}} \coloneqq \{&\gsectionc^{0}, \cdots, \gsectionc^{i}, \cdots, \gsectionc^{n}\} \\
+     &\gsectionc: \opensetgc\rightarrow \gtotalc|_{\opensetgc}
+    \end{split}
+  \end{equation}
+\end{equation}
+which are the set of all graphics with a shared base space and fiber and are a subcategory of \textbf{Set}. This allows us to represent the full space of graphics as a sheaf $\sheaf(\gtotal)$
+\begin{equation}
+  \label{eq:artist:artist_sheaf}
+  \begin{tikzcd}[row sep=huge]
+    {\cgamma{\opensetgc}{\gtotalc|_{\opensetgc}}}&  & 
+    \textcolor{set}{\{}\gsectionc:\opensetgc\rightarrow \gtotalc\textcolor{set}{\}}
+    \arrow[ll, "\iota"', hook', color=set] \\ &  & \\
+    \opensetgc \arrow[uu, "\sheafc(\gtotalc)", color=sheaf] \arrow[rruu, "\vartistc(\sheafc(\dtotalc))\coloneqq \sheafc_{\vartistc}(\gtotalc)" description, dashed, color=sheaf] &  & 
+    \end{tikzcd}
+\end{equation}
+The set of graphics reachable through an artist function is a subset of sections of the full set of graphics $\{\gsection\} \subset \Gamma(\opensetg, \gtotal|_{\opensetg})$.
+These sets of reachable graphics over open sets \opensetg\ are also objects of \textbf{Set}.
+This allows us to formulate the graphic output by the artist as a sheaf
+\begin{equation}
+  \label{eq:artist:artist_sheaf}
+  \sheafc_{\vartist}(\gtotalc): \gbasec^{\textcolor{base}{op}} \textcolor{sheaf}{\rightarrow} \setc
+\end{equation}
+from the graphic base space to the set of reachable graphics.
 
+Following from \autoref{eq:artist:data_pullback_sheaf} and \autoref{eq:artist:artist_sheaf},  we propose that the artist function $\vartist$ is a natural transformation
 \begin{equation}

### 2021-12-02 paper/main.tex
S,K, I

@@
 \label{sec:related-work:continuity}
-\begin{figure}[!h]
+\begin{figure}[h!]
   \includegraphics[width=\columnwidth]{k_different_types.png}
@@
 
-\begin{figure}[!h]
+\begin{figure}[h!]
   \includegraphics[width=\columnwidth]{whycontinuity.png}
@@
 
-\begin{figure}[!h]
+\begin{figure}[h!]
   \label{fig:related-work:equivariance}
@@
 The objects of \textbf{Set} in this sheaf are the sets of sections $\Gamma(\dbase, \dtotal)$ introduced in \autoref{eq:related-work:continuity:sections} and the morphisms are the inclusion maps $(\iota, \iota^*)$ introduced in \autoref{eq:related-work:continuity:presheaf}.
-As mentioned in \autoref{re} sections $\dsection$ are
 
-\note{add transition sentence about artist }
-
-The continuity of the graphic is encapsulated as a category $\mathcal{\gbase}}$
+An element of the sheaf, which is a section \dsection, is converted into a graphic.
+The topology of the graphic \gbase\ is determined both by the data continuity \dbase\ and the dimensionality of the display space
+\begin{equation}
+  \gbase = \dbase \times I^{\text{dim}(display)-\text{dim}(\dbase)} 
+\end{equation}
+where $I$ is an arbitrary 1 dimensional interval. 
+\begin{figure}[h!]
+  \label{fig:artist:gbase}
+  \includegraphics{scatter_s.png}
+  \caption{The topology of the data, which is a 0 dimensional point, is expanded to a region in the 2D screen display space.
+It is expanded via the intervals $I_1, I_2$ such that all coordinates on the disk map back to the same K \note{maybe also include line example,or just use line example}}
+\end{figure}
+As illustrated in \autoref{fig:artist:gbase}, the interval $I$ is used to increase the dimensionality of the data topology \dbase, by padding it with intervals, such that it is compatible with the display space.
 
@@
 \subsubsection{Shared Index}
-\begin{figure}[!h]
+\begin{figure}[h!]
   \includegraphics[width=\columnwidth]{k_union.png}
@@
 
-\begin{figure}[!h]
+\begin{figure}[h!]
   \centering
@@
 \paragraph{Multiview Constraints}
-\begin{figure}[!h]
+\begin{figure}[h!]
   \centering
@@
 
-\begin{figure}[!h]
+\begin{figure}[h!]
   \includegraphics[width=\columnwidth]{diff_type_q.png}

### 2021-12-02 paper/main.tex
deform retract + figuring out artist

--- 2021-12-02 paper/main.tex
@@
 \usepackage{mathrsfs} % math script fonts
+\usepackage{relsize} %bigger
 \usepackage{bm}
@@
 \subsection{Sheaves}
+\label{sec:related-work:continuity:sheaves}
 Sheaves can be thought of as an "algebraic data structure" \cite{ghristElementaryAppliedTopology2014} for data that lives over topological spaces.
@@
 
-We introduce the category $\mathcal{\dbase}$ to encapsulate the subsets of the continuity and the inclusion maps that glue them together.
-
+We introduce the category $\mathcal{\dbase}$ to encapsulate the continuity of the data, for example if the data are discrete points, continuous lines, images, or networks. 
 \begin{definition} \label{def:category:K} The category of open sets $\mathcal{\dbasec}$ consists of
@@
 
-We model a space of data sets with the same continuity and fields as the sheaf  $\sheafc({\dtotalc}): \mathcal{\dbasec}^{op} \textcolor{sheaf}{\rightarrow} \setc$.
+We also introduce an analogous category $\mathcal{\gbase}$ that encapsulates the topology of the space in which the graphic is displayed, for example whether it is a 2D screen or a 3D printer.
+
+\begin{definition} The category of open sets $\mathcal{\gbasec}$ in the topological space $(\dbase, \mathscr{T})$ consists of open set \textit{objects} $\opensetg \in \{\opensetg\}$, including the empty set $\varnothing$ and the maximum set $\gtotal$ and \textit{morphisms} $\iota: \opensetg_i \rightarrow \opensetg_j$.
+\end{definition}
+
+The data space \dbase\ is a deformation retract of the graphic topology \gbase\
+\begin{equation}
+  \label{eq:artist:gbase}
+  \gbase = \dbase \times I^{\text{dim}(display)-\text{dim}(\dbase)} 
+\end{equation}
+meaning that regions in \gbase\ collapse to points in \dbase\ such that the relative positions of the points are preserved \cite{nlab:deformation_retract}.
+The dimensionality of the data \dbase\ is expanded to match the dimensionality of the graphic display space by multiplying the data base space \dbase\ by intervals I.
+
+\begin{figure}[h!]
+  \label{fig:artist:gbase}
+  \includegraphics[width=\columnwidth]{deform_retract.png}
+  \caption{Given a 2D screen display, the graphic space \gbase\ must be 2D.
+It is constructed by multiplying the line \dbase\ by the interval [0,1] such that every vertical band in \gbase\ corresponds to a unique point $\dbasepoint \in \dbase$.}
+\end{figure}
+
+For example, as illustrated in \autoref{fig:artist:gbase}, the display space is a 2D screen and the data space is a 1D line \dbase.
+For that line to have any thickness, it must be rendered as a band.
+This means the region $(\dbasepoint, \alpha)|_{\alpha \in I} \subset \gbase$ maps to the corresponding point $\dbasepoint \in \dbase$.
+
+\subsubsection{Data}
+\note{clean up why sheaf}
+As mentioned in \autoref{sec:related-work:continuity:sheaves}, sheaves are a powerful tool for keeping track of data that lies over topological spaces.
+Therefore, we model the type of data an artist expects as a sheaf $\sheafc({\dtotalc}): \mathcal{\dbasec}^{op} \textcolor{sheaf}{\rightarrow} \setc$.
 \begin{equation}
@@
 The objects of \textbf{Set} in this sheaf are the sets of sections $\Gamma(\dbase, \dtotal)$ introduced in \autoref{eq:related-work:continuity:sections} and the morphisms are the inclusion maps $(\iota, \iota^*)$ introduced in \autoref{eq:related-work:continuity:presheaf}.
-As mentioned in \autoref{re} sections $\dsection$ are
-
-\note{add transition sentence about artist }
-
-The continuity of the graphic is encapsulated as a category $\mathcal{\gbase}}$
-
-\begin{definition} The category of open sets $\mathcal{\gbase}$ in the topological space $(\dbase, \mathscr{T})$).
-This category consists of open set \textit{objects} $\opensetg \in \{\opensetg\}$, including the empty set \varnothing\ and the total set $\gtotal$, and inclusion \textmorphisms $\iota: \opensetg_i \rightarrow \opensetg_j$.
-\end{definition}
 
@@
 \end{equation}
+such that the data continuity $\dbase$ is a deformation retract of $\gbase$, as expressed in \autoref{eq:artist:gbase}.
+
 As shown in \autoref{eq:artist:vindex}, the map between spaces $\vindex$ can be composed with section map $\dsection$ such that there is a map from the graphic base space $\gbase$ to the fiber bundle in which the data lives $\dtotal$.
-
 \begin{figure}[h!]
@@
 \end{equation}
-to generate a set of equivalent sections over the graphic $\Gamma(\gtotal, \vindex^*{\dtotal})$.
+to generate a set of equivalent sections over the graphic space $\Gamma(\gbase, \vindex^*{\dtotal})$.
 As a consequence of \autoref{eq:artist:data_gamma_pullback}, there exists a data sheaf over graphic space
@@
 
-As with $\mathcal{\dbase}$, the morphisms between the open sets  $\opensetg$ are inclusion morphisms $\iota$. Over the graphic space $\opensetg \subseteq \gbase$ are a set of sections 
+\subsubsection{Graphic}
+We model the space of all graphic generating functions over a specific graphic topology $\opensetg \subseteq \gbase$ and targeting a specific display as a set of sections 
 \begin{equation}
@@
 \end{equation}
-which are the set of all graphics with a shared base space and fiber and are a subcategory of \textbf{Set}. This allows us to represent the full space of graphics as a sheaf $\sheaf(\gtotal)$
+The set of sets of sections $\{\Gamma(\opensetg, \gtotal)\}|_{\opensetg \in gbase}$ is subcategory of \textbf{Set}; therefore the full space of graphics is a sheaf $\sheaf(\gtotal)$
 \begin{equation}
@@
     {\cgamma{\opensetgc}{\gtotalc|_{\opensetgc}}}&  & 
-    \textcolor{set}{\{}\gsectionc:\opensetgc\rightarrow \gtotalc\textcolor{set}{\}}
+    \textcolor{set}{\{}\gsectionc:\opensetgc\rightarrow \gtotalc|_{\opensetgc}\textcolor{set}{\}}
     \arrow[ll, "\iota"', hook', color=set] \\ &  & \\
@@
 \end{equation}
+
+\begin{equation}
+  \label{eq:artist:artist_sheaf}
+  \begin{tikzcd}[column sep=tiny, row sep=huge]
+    {\cgamma{\opensetgc}{\gtotalc|_{\opensetgc}}}&  & 
+    \textcolor{set}{\{}\gsectionc:\opensetgc\rightarrow \gtotalc|_{\opensetgc}\textcolor{set}{\}}
+    \arrow[ll, "\textcolor{set}{\mathlarger{\mathlarger{\mathlarger{\supset}}}}" description, hook', color=white] \\ &  & \\
+    \opensetgc \arrow[uu, "\sheafc(\gtotalc)", color=sheaf] \arrow[rruu, "\vartistc(\sheafc(\dtotalc))\coloneqq \sheafc_{\vartistc}(\gtotalc)" description, dashed, color=sheaf] &  & 
+    \end{tikzcd}
+\end{equation}
+
 The set of graphics reachable through an artist function is a subset of sections of the full set of graphics $\{\gsection\} \subset \Gamma(\opensetg, \gtotal|_{\opensetg})$.
@@
 
+\subsubsection{Data to Graphic}
 Following from \autoref{eq:artist:data_pullback_sheaf} and \autoref{eq:artist:artist_sheaf},  we propose that the artist function $\vartist$ is a natural transformation
@@
 
-which means that the artist can be pulled back through $\vindex^*$\footnote{Defintion 3.68 in Seven Sketches in Compositionality by Fong and Spivak}
+which means that the artist can be pulled back through $\vindex^*$\footnote{Definition 3.68 in Seven Sketches in Compositionality by Fong and Spivak}
 \begin{equation}

### 2021-12-03 paper/main.tex
changed restriction symbols + set builder notation

@@
 \usepackage{amsmath, amsthm, amsfonts,amssymb,eulervm,xspace, mathtools}
+\renewcommand{\restriction}{\mathord{\upharpoonright}} %restriction w/p space
 \usepackage{mathrsfs} % math script fonts
@@
   \label{eq:related-work:continuity:sections}
-  \begin{split}
-  \cgamma{\opensetc}{\dtotalc|_{\opensetc}} \coloneqq \{&\dsectionc^{0}, \cdots, \dsectionc^{i}, \cdots, \dsectionc^{n}\} \\
-   &\dsectionc: \opensetc\rightarrow \dtotalc|_{\openset}
-  \end{split}
+  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\openset} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;\forall \dbasepointc \in \opensetc \big\}
 \end{equation}
@@
 Sheaves can be thought of as an "algebraic data structure" \cite{ghristElementaryAppliedTopology2014} for data that lives over topological spaces.
-A presheaf $\sheafc$ is a map from the open set to the set of sections $\sheafc: \opensetc \rightarrow \cgamma{\opensetc}{ \dtotalc|_{\opensetc}}$ \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}.
+A presheaf $\sheafc$ is a map from the open set to the set of sections $\sheafc: \opensetc \rightarrow \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$ \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}.
 The presheaf preserves inclusion maps $\iota$ between open sets and the inclusion map $\iota^*$ between the sets of sections.
@@
   \begin{tikzcd}[column sep=small]
-    {\cgamma{\opensetc_1}{\dtotalc|_{\opensetc_1}}} & {} & {\cgamma{\opensetc_2}{\dtotalc|_{\opensetc_2}}} \arrow[ll, "\iota^*"', hook'] \\& &\\
+    {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} & {} & {\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}} \arrow[ll, "\iota^*"', hook'] \\& &\\
     \opensetc_1 \arrow[uu, "\sheafc_{\dtotal}"] \arrow[rr, "\iota"', hook] & {} \arrow[uu, "\sheafc_{\dtotal}" description, dotted] & \opensetc_2 \arrow[uu, "\sheafc_\dtotal"']                    
@@
 
-This means that a smaller space $\openset_1$ is included in a larger space $\openset_2$ and a function $\dsection \in \Gamma(\openset, \dtotal|_{\openset})$ that is continuous over a larger space $\openset_2$ is continuous over a subspace $\openset_1 \subset\openset_2$.
+This means that a smaller space $\openset_1$ is included in a larger space $\openset_2$ and a function $\dsection \in \Gamma(\openset, \dtotal\restriction_{\openset})$ that is continuous over a larger space $\openset_2$ is continuous over a subspace $\openset_1 \subset\openset_2$.
 Sheaves are presheafs where sets of sections are defined over unions of open sets over a topology $\bigcup_{j \in I} \openset)_j \in \dbase$ \cite{spanier1989algebraic, nlab:sheaf}.
@@
 For that line to have any thickness, it must be rendered as a band.
-This means the region $(\dbasepoint, \alpha)|_{\alpha \in I} \subset \gbase$ maps to the corresponding point $\dbasepoint \in \dbase$.
+This means the region $(\dbasepoint, \alpha)\vert_{\alpha \in I} \subset \gbase$ maps to the corresponding point $\dbasepoint \in \dbase$.
 
@@
   \begin{tikzcd}
-    & \cgamma{\opensetc}{\dtotalc|_{\opensetc}} \in Ob(\setc) & \\
+    & \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \in Ob(\setc) & \\
     {} & & {} \\
@@
 \begin{equation}
-  \label{eq:artist:graphic_sections}
-    \begin{split}
-    \cgamma{\opensetgc}{\gtotalc|_{\opensetgc}} \coloneqq \{&\gsectionc^{0}, \cdots, \gsectionc^{i}, \cdots, \gsectionc^{n}\} \\
-     &\gsectionc: \opensetgc\rightarrow \gtotalc|_{\opensetgc}
-    \end{split}
-  \end{equation}
-\end{equation}
-The set of sets of sections $\{\Gamma(\opensetg, \gtotal)\}|_{\opensetg \in gbase}$ is subcategory of \textbf{Set}; therefore the full space of graphics is a sheaf $\sheaf(\gtotal)$
-\begin{equation}
-  \label{eq:artist:artist_sheaf}
-  \begin{tikzcd}[row sep=huge]
-    {\cgamma{\opensetgc}{\gtotalc|_{\opensetgc}}}&  & 
-    \textcolor{set}{\{}\gsectionc:\opensetgc\rightarrow \gtotalc|_{\opensetgc}\textcolor{set}{\}}
-    \arrow[ll, "\iota"', hook', color=set] \\ &  & \\
-    \opensetgc \arrow[uu, "\sheafc(\gtotalc)", color=sheaf] \arrow[rruu, "\vartistc(\sheafc(\dtotalc))\coloneqq \sheafc_{\vartistc}(\gtotalc)" description, dashed, color=sheaf] &  & 
-    \end{tikzcd}
-\end{equation}
+  \label{eq:artist:graphic_sections}}
+    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \coloneqq \big\{\gsectionc: \opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} \bigm{\vert} \pi(\gsectionc(\gbasepointc)) = \gbasepointc \; \forall \gbasepointc \in \opensetgc \big\}
+\end{equation}
+
+The set of sets of sections $\{\Gamma(\opensetg, \gtotal)\}\vert_{{\opensetg \in gbase}}$ is subcategory of \textbf{Set}; therefore the full space of graphics is a sheaf $\sheaf(\gtotal)$
 
@@
   \begin{tikzcd}[column sep=tiny, row sep=huge]
-    {\cgamma{\opensetgc}{\gtotalc|_{\opensetgc}}}&  & 
-    \textcolor{set}{\{}\gsectionc:\opensetgc\rightarrow \gtotalc|_{\opensetgc}\textcolor{set}{\}}
-    \arrow[ll, "\textcolor{set}{\mathlarger{\mathlarger{\mathlarger{\supset}}}}" description, hook', color=white] \\ &  & \\
+    {\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}}&  & 
+    \textcolor{set}{\big\{}\gsectionc:\opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} 
+    \bigm{\vert} \vartist(\dsectionc) = \gsectionc \textcolor{set} \;, \dsectionc \in \sheafc(\dtotalc){\big\}} 
+    \arrow[ll, "\textcolor{set}{\mathlarger{\mathlarger{\supset}}}" description, hook', color=white] \\ &  & \\
     \opensetgc \arrow[uu, "\sheafc(\gtotalc)", color=sheaf] \arrow[rruu, "\vartistc(\sheafc(\dtotalc))\coloneqq \sheafc_{\vartistc}(\gtotalc)" description, dashed, color=sheaf] &  & 
@@
 
-The set of graphics reachable through an artist function is a subset of sections of the full set of graphics $\{\gsection\} \subset \Gamma(\opensetg, \gtotal|_{\opensetg})$.
+The set of graphics reachable through an artist function is a subset of sections of the full set of graphics $\{\gsection\} \subset \Gamma(\opensetg, \gtotal\restriction_{\opensetg})$.
 These sets of reachable graphics over open sets \opensetg\ are also objects of \textbf{Set}.
@@
 The bifunctor $\otimes$ combines categories in a way that  preserves identity and composition \cite{fongInvitationAppliedCategory2019}.
-For example, given a pair of categories $\mathcal{\dfiber_a},\mathcal{\dfiber_b}$, the bifunctor yields a set of pairs $\{(\dfiber_a, \dfiber_b) | \dfiber_a \in \mathcal{\dfiber_a}, \dfiber_b \in \mathcal{\dfiber_b}\}$.
+For example, given a pair of categories $\mathcal{\dfiber_a},\mathcal{\dfiber_b}$, the bifunctor yields a set of pairs $\{(\dfiber_a, \dfiber_b) \big{\mvert} \dfiber_a \in \mathcal{\dfiber_a}, \dfiber_b \in \mathcal{\dfiber_b}\}$.
 For the associated morphisms $g: \mathcal{\dfiber_a}\rightarrow \mathcal{\dfiber_a}$ and $h: \mathcal{\dfiber_b}\rightarrow \mathcal{\dfiber_b}$, the function $(g\times h): (\mathcal{\dfiber_a}\times \mathcal{\dfiber_b}) \rightarrow (\mathcal{\dfiber_a}\times \mathcal{\dfiber_b})$ is pointwise $(g\times h)(\dfiber_a, \dfiber_b)\coloneqq (g(\dfiber_a), h(\dfiber_b))$.
@@
 When a fiber bundle is trivial, we can dispense with localization and directly consider the fiber maps $Hom(\dfiber_{\dbasepoint}, \dfiber_{\dbasepoint})$ over a point \dbasepoint\ as the basis of our equivariance.
-This is because a sheaf $\sheaf(\dtotal)$ over a limit of open sets that contain a point $\dbasepoint \in \dbase$ is approximately the same as a sheaf over a point \dbasepoint\ and is called the stalk $\sheaf(\dtotal)|_{\dbasepoint}\coloneqq \lim\limits_{\openset\ni \dbasepoint} \Gamma(\openset, \dtotal)$ \cite{StalkSheaf2019}.
+This is because a sheaf $\sheaf(\dtotal)$ over a limit of open sets that contain a point $\dbasepoint \in \dbase$ is approximately the same as a sheaf over a point \dbasepoint\ and is called the stalk $\sheaf(\dtotal)\restriction_{\dbasepoint}\coloneqq \lim\limits_{\openset\ni \dbasepoint} \Gamma(\openset, \dtotal)$ \cite{StalkSheaf2019}.
 The fiber space over a point is embedded in the stalk over that point $\dfiber_{\dbasepoint}\hookrightarrow \mathscr{\dfiber}_{\dbasepoint}$.
@@
 
-where the projection functions $i_{\dbase_a},i_{\dbase_b}$ behave such that $i_{\dbase_a}(\dbasepoint)|_{\dbasepoint \in \dbase_b} = \left[\dbasepoint\right]$, $i_{\dbase_b}(\dbasepoint)|_{\dbasepoint \in \dbase_b} = \left[\dbasepoint\right]$, meaning the projection functions yield equivalent points in the disjoint union $\dbase_a \bigsqcup_{\dbase_c} \dbase_b$.
+where the projection functions $i_{\dbase_a},i_{\dbase_b}$ behave such that $i_{\dbase_a}(\dbasepoint)\vert_{\dbasepoint \in \dbase_b} = \left[\dbasepoint\right]$, $i_{\dbase_b}(\dbasepoint)\vert_{\dbasepoint \in \dbase_b} = \left[\dbasepoint\right]$, meaning the projection functions yield equivalent points in the disjoint union $\dbase_a \bigsqcup_{\dbase_c} \dbase_b$.
 Data that is completely disjoint in the index, as shown in \autoref{fig:artist:union:disjoint_k}, is a special case where $\dbase_c=\varnothing$ and therefore there is no constraint on the artist with regards to how this data is rendered.

### 2021-12-06 paper/main.tex
K to S clean up

@@
 
-We also introduce an analogous category $\mathcal{\gbase}$ that encapsulates the topology of the space in which the graphic is displayed, for example whether it is a 2D screen or a 3D printer.
+We also introduce an analogous category $\mathcal{\gbase}$ that encapsulates the topology of the graphic, which is the topology of the rendering in the display space; for example, this rendering can be a 2D or volumetric image or a 3D print.
 
@@
 
-The data space \dbase\ is a deformation retract of the graphic topology \gbase\
-\begin{equation}
-  \label{eq:artist:gbase}
-  \gbase = \dbase \times I^{\text{dim}(display)-\text{dim}(\dbase)} 
-\end{equation}
-meaning that regions in \gbase\ collapse to points in \dbase\ such that the relative positions of the points are preserved \cite{nlab:deformation_retract}.
-The dimensionality of the data \dbase\ is expanded to match the dimensionality of the graphic display space by multiplying the data base space \dbase\ by intervals I.
+The data space \dbase\ is a deformation retract of the graphic topology \gbase\meaning that regions in \gbase\ collapse to points in \dbase\ such that the relative positions of the points are preserved \cite{nlab:deformation_retract}.
+One method of constructing the graphic space is by expanding the dimensionality of the data \dbase\ to match the spatial dimensionality of the rendering by multiplying the data space by  intervals $I$.
 
@@
 For example, as illustrated in \autoref{fig:artist:gbase}, the display space is a 2D screen and the data space is a 1D line \dbase.
-For that line to have any thickness, it must be rendered as a band.
-This means the region $(\dbasepoint, \alpha)\vert_{\alpha \in I} \subset \gbase$ maps to the corresponding point $\dbasepoint \in \dbase$.
+For that line to have any thickness, it must be rendered as a ban.
+One method of obtaining this graphic continuity while preserving the constraint that it is a deformation retract is by multiplying the base space \dbase\ by an interval $I=\[0,1\]$ such that $\gbase = \dbase \times I^{1}$.
+This construction ensures that every region $(\dbasepoint, \alpha)\vert_{\alpha \in I} \subset \gbase$ maps to a  corresponding point $\dbasepoint \in \dbase$.
 
@@
     \textcolor{set}{\big\{}\gsectionc:\opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} 
-    \bigm{\vert} \vartist(\dsectionc) = \gsectionc \textcolor{set} \;, \dsectionc \in \sheafc(\dtotalc){\big\}} 
+    \bigm{\vert} \vartist(\dsectionc) = \gsectionc \;, \dsectionc \in \sheafc(\dtotalc)\textcolor{set}{\big\}} 
     \arrow[ll, "\textcolor{set}{\mathlarger{\mathlarger{\supset}}}" description, hook', color=white] \\ &  & \\

### 2021-12-15 paper/main.tex
fixing op
saving to clear out scraps

@@
 
- \begin{definition} A presheaf \sheafc\, as introduced in \autoref{eq:related-work:continuity:presheaf}, is a functor $\sheaf: \mathcal{C}^{op} \rightarrow \textbf{Set}$.\cite{nlab:presheaf} \end{definition}
+ \begin{definition} A presheaf \sheafc\, as introduced in \autoref{eq:related-work:continuity:presheaf}, is a functor $\sheaf: \mathcal{C}^{op} \rightarrow \textbf{Set}$.\cite{nlab:presheaf } \end{definition} \note{possibly add in from Hader lecture notes}
 
@@
   \label{eq:artist:data_pullback_sheaf_gamma}
-\sheafc_{\vindexc^*}(\dtotalc): \mathcal{\gbasec^{\textcolor{base}{op}}} \textcolor{sheaf}{\rightarrow} \cgamma{\gbasec}{\vindexc^*{\dtotalc}}
+\sheafc_{\vindexc^*}(\dtotalc): \mathcal{\gbasec}^{\textcolor{base}{op}} \textcolor{sheaf}{\rightarrow} \cgamma{\gbasec}{\vindexc^*{\dtotalc}}
 \end{equation}
@@
   \label{eq:artist:data_pullback_sheaf}
-  \sheafc_{\vindexc^*}(\dtotalc): \mathcal{\gbasec^{\textcolor{base}{op}}} \textcolor{sheaf}{\rightarrow} \setc
+  \sheafc_{\vindexc^*}(\dtotalc): \mathcal{\gbasec}^{\textcolor{base}{op}} \textcolor{sheaf}{\rightarrow} \setc
 \end{equation}

### 2021-12-16 paper/main.tex
defining rho in terms of phi

@@
 As mentioned in \autoref{sec:related-work:continuity:sheaves}, sheaves are a powerful tool for keeping track of data that lies over topological spaces.
-Therefore, we model the type of data an artist expects as a sheaf $\sheafc({\dtotalc}): \mathcal{\dbasec}^{op} \textcolor{sheaf}{\rightarrow} \setc$.
+Therefore, we model the type of data an artist expects as a sheaf $\sheafc_{\dtotalc}: \mathcal{\dbasec}^{op} \textcolor{sheaf}{\rightarrow} \setc$.
 \begin{equation}
@@
     {} & & {} \\
-    & \opensetc \in Ob(\mathcal{\dbasec}^{\textcolor{base}{op}}) \arrow[uu, "\sheafc(\dtotalc)"', shift left = 10, color=sheaf] &   
+    & \opensetc \in Ob(\mathcal{\dbasec}^{\textcolor{base}{op}}) \arrow[uu, "\sheafc_{\dtotalc}"', shift left = 10, color=sheaf] &   
 \end{tikzcd}
@@
   \begin{tikzcd}
-    \dtotalc                        & \vindexc^*\dtotalc \arrow[l, "\vindexc^*"', dashed, color=artist]                                                                           \\
-    \dbasec \arrow[u, "\dsectionc", color=section] & \gbasec \arrow[l, "\vindexc", color=artist] \arrow[lu, "\dsectionc \circ \vindexc" description, color=section] \arrow[u, "\vindexc^*\dsectionc"', dashed, color=section]
+    \dtotalc& \dtotalpullc \arrow[l, "\vindexpullc"', dashed, color=artist]                                                                           \\
+    \dbasec \arrow[u, "\dsectionc", color=section] & \gbasec \arrow[l, "\vindexc", color=artist] \arrow[lu, "\dsectionc \circ \vindexc" description, color=section] \arrow[u, "\dsectionpullc"', dashed, color=section]
     \end{tikzcd}
@@
   \label{fig:artist:pushforward_xi}
-  \caption{The map $\vindex^*$ is a map from a repetition of the 1D fiber over every point $\dbasepoint \in \dbase$ in data space, turning the fiber into a 2D plane of repeating values over the corresponding region $\dbasepoint = \vindex(\gbasepoint)$ in the graphic space \gbase.
+  \caption{The map $\vindexpull$ is a map from a repetition of the 1D fiber over every point $\dbasepoint \in \dbase$ in data space, turning the fiber into a 2D plane of repeating values over the corresponding region $\dbasepoint = \vindex(\gbasepoint)$ in the graphic space \gbase.
 }
 \end{figure}
-Since there is a mapping from the graphic space to the total bundle $\dsection \circ \vindex: \gbase \rightarrow \dtotal$, the mapping $\vindex$ can be pulled through the composition such that there is a map $\vindex^*$ between a total space over graphics $\vindex^*\dtotal$ and its equivalent space over data $\dtotal$.
+Since there is a mapping from the graphic space to the total bundle $\dsection \circ \vindex: \gbase \rightarrow \dtotal$, the mapping $\vindex$ can be pulled through the composition such that there is a map $\vindexpull$ between a total space over graphics $\dtotalpull$ and its equivalent space over data $\dtotal$.
 As illustrated in the example given in \autoref{fig:artist:pushforward_xi}, a bundle $\dtotal$ over a point $\dbasepoint \in \dbase$ is pulled back through \vindex\ \cite{nlab:pullback_bundle} such that it is copied over every point in the corresponding region in graphic space $\vindex(\gbasepoint) = \dbasepoint, \gbasepoint \subset \gbase$.
-The pullback $\vindex^*$ can then be composed with the elements of the set of sections
+The pullback $\vindexpull$ can then be composed with the elements of the set of sections
 \begin{equation}
   \label{eq:artist:data_gamma_pullback}
- \cgamma{\dbasec}{\dtotalc} \textcolor{artist}{\xrightarrow{\vindex^*}} \cgamma{\gbasec}{\vindexc^*\dtotalc}
-\end{equation}
-to generate a set of equivalent sections over the graphic space $\Gamma(\gbase, \vindex^*{\dtotal})$.
+  \vindexpullc: \cgamma{\dbasec}{\dtotalc} \textcolor{artist}{\rightarrow} \cgamma{\gbasec}{\dtotalpullc}
+\end{equation}
+to generate a set of equivalent sections over the graphic space $\Gamma(\gbase, \dtotalpull)$.
 As a consequence of \autoref{eq:artist:data_gamma_pullback}, there exists a data sheaf over graphic space
@@
   \label{eq:artist:data_pullback_sheaf_gamma}
-\sheafc_{\vindexc^*}(\dtotalc): \mathcal{\gbasec}^{\textcolor{base}{op}} \textcolor{sheaf}{\rightarrow} \cgamma{\gbasec}{\vindexc^*{\dtotalc}}
+\sheafc_{\dtotalpullc}: \mathcal{\gbasec}^{\textcolor{base}{op}} \textcolor{sheaf}{\rightarrow} \cgamma{\gbasec}{\dtotalpullc}}
 \end{equation}
@@
   \label{eq:artist:data_pullback_sheaf}
-  \sheafc_{\vindexc^*}(\dtotalc): \mathcal{\gbasec}^{\textcolor{base}{op}} \textcolor{sheaf}{\rightarrow} \setc
+  \sheafc_{\dtotalpullc}: \mathcal{\gbasec}^{\textcolor{base}{op}} \textcolor{sheaf}{\rightarrow} \setc
 \end{equation}
@@
 
-The set of sets of sections $\{\Gamma(\opensetg, \gtotal)\}\vert_{{\opensetg \in gbase}}$ is subcategory of \textbf{Set}; therefore the full space of graphics is a sheaf $\sheaf(\gtotal)$
-
+The set of sets of sections $\{\Gamma(\opensetg, \gtotal)\}\vert_{{\opensetg \in gbase}}$ is subcategory of \textbf{Set}; therefore the full space of graphics is a sheaf $\sheaf_{\gtotal}$
+
+\note{valid $\rho$ is determined by equivariance, which we haven't introduced yet, but we can introduce equivariance w/ just $\sheaf_{\gtotal}$}
 \begin{equation}
@@
     \textcolor{set}{\big\{}\gsectionc:\opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} 
-    \bigm{\vert} \vartist(\dsectionc) = \gsectionc \;, \dsectionc \in \sheafc(\dtotalc)\textcolor{set}{\big\}} 
+    \bigm{\vert} \vartistc\circ\dfuncc \circ \vartistc^{-1}[\gsectionc] = \dfuncc^{\prime} \circ \gsectionc, \dsectionc \in \sheafc_{\dtotal}
+    \textcolor{set}{\big\}} 
     \arrow[ll, "\textcolor{set}{\mathlarger{\mathlarger{\supset}}}" description, hook', color=white] \\ &  & \\
-    \opensetgc \arrow[uu, "\sheafc(\gtotalc)", color=sheaf] \arrow[rruu, "\vartistc(\sheafc(\dtotalc))\coloneqq \sheafc_{\vartistc}(\gtotalc)" description, dashed, color=sheaf] &  & 
+    \opensetgc \arrow[uu, "\sheafc_{\gtotalc}", color=sheaf] \arrow[rruu, "\vartistc(\sheafc_{\dtotalc})\coloneqq \sheafc_{\vartistc}" description, dashed, color=sheaf] &  & 
     \end{tikzcd}
@@
   \label{eq:artist:artist_sheaf}
-  \sheafc_{\vartist}(\gtotalc): \gbasec^{\textcolor{base}{op}} \textcolor{sheaf}{\rightarrow} \setc
+  \sheafc_{\vartist}: \gbasec^{\textcolor{base}{op}} \textcolor{sheaf}{\rightarrow} \setc
 \end{equation}
@@
     & {} \arrow[dd, "\vartistc", Rightarrow, shorten <=1.35em,  shorten >=1.25em, color=artist] & \\
-\mathcal{\gbasec}^{\textcolor{base}{op}} \arrow[rr, "\sheafc_{\vindexc_*}(\dtotalc)", bend left, color=sheaf] \arrow[rr, "\sheafc_{\vartistc}(\gtotalc)"', bend right, color=sheaf] &  & {\setc}\\
+\mathcal{\gbasec}^{\textcolor{base}{op}} \arrow[rr, "\sheafc_{\vindexc_*}(\dtotalpullc)", bend left, color=sheaf] \arrow[rr, "\sheafc_{\vartistc}(\gtotalc)"', bend right, color=sheaf] &  & {\setc}\\
     & {} &                            
@@
   \label{eq:artist:sheaf_map}
-  \vartistc: \sheafc_{\vindexc^*}(\dtotalc) \textcolor{artist}{\rightarrow} \sheafc_{\vartistc}(\gtotalc)
+  \vartistc: \sheafc_{\vindexc^*}(\dtotalpullc) \textcolor{artist}{\rightarrow} \sheafc_{\vartistc}(\gtotalc)
 \end{equation}
@@
   \begin{tikzcd}
-    \cgamma{\opensetgc_1}{\dtotalc} \arrow[dd, "\iota^*"', hook, color=set] \arrow[rrrr, "\vartistc_{\opensetgc_1}", bend left, color=artist] &  & \opensetgc_1 \arrow[ll, "\sheafc_{\vindexc^*}", color=sheaf] \arrow[rr, "\sheafc_{\vartistc}"', color=sheaf] &  & \cgamma{\opensetgc_1}{\gtotalc} \arrow[dd, "\iota^*", hook', color=set] \\
+    \cgamma{\opensetgc_1}{\dtotalpullc} \arrow[dd, "\iota^*"', hook, color=set] \arrow[rrrr, "\vartistc_{\opensetgc_1}", bend left, color=artist] &  & \opensetgc_1 \arrow[ll, "\sheafc_{\dtotalpullc}", color=sheaf] \arrow[rr, "\sheafc_{\vartistc}"', color=sheaf] &  & \cgamma{\opensetgc_1}{\gtotalc} \arrow[dd, "\iota^*", hook', color=set] \\
     {} \arrow[rrrr, dotted, color=artist] &&&& {} \\
-    \cgamma{\opensetgc_2}{\dtotalc} \arrow[rrrr, "\vartistc_{\opensetgc_2}"', bend right, color=artist]                            &  & \opensetgc_2 \arrow[ll, "\sheafc_{\vindexc^*}"', color=sheaf] \arrow[rr, "\sheafc_{\vartistc}", color=sheaf] \arrow[uu, "\iota" description, hook, color=base] &  & \cgamma{\opensetgc_2}{\gtotalc}
+    \cgamma{\opensetgc_2}{\dtotalpullc} \arrow[rrrr, "\vartistc_{\opensetgc_2}"', bend right, color=artist]                            &  & \opensetgc_2 \arrow[ll, "\sheafc_{\dtotalpullc}"', color=sheaf] \arrow[rr, "\sheafc_{\vartistc}", color=sheaf] \arrow[uu, "\iota" description, hook, color=base] &  & \cgamma{\opensetgc_2}{\gtotalc}
     \end{tikzcd}
@@
   \label{eq:artist:gamma_map}
-  \cgamma{\dbasec}{\dtotalc} \textcolor{artist}{\xrightarrow{\vindexc^*}} \cgamma{\gbasec}{\dtotalc} \textcolor{artist}{\xrightarrow{\vartist}} \cgamma{\gbasec}{\gtotalc}
+  \cgamma{\dbasec}{\dtotalc} \textcolor{artist}{\xrightarrow{\vindexc^*}} \cgamma{\gbasec}{\dtotalpullc} \textcolor{artist}{\xrightarrow{\vartist}} \cgamma{\gbasec}{\gtotalc}
 \end{equation}
@@
     & & {} \arrow[dd, "\vartistc" description, Rightarrow, shorten <=1.35em,  shorten >=1.25em, color=artist] &       \\
-\mathcal{\dbasec}^{\textcolor{base}{op}} \arrow[r, "\vindexc^*", color=artist] & \mathcal{\gbasec}^{\textcolor{base}{op}} \arrow[rr, "\sheafc_{\vindex^*}(\dtotalc)", bend left, color=sheaf] \arrow[rr, "\sheaf_{\vartistc}(\gtotalc)"', bend right, color=sheaf] &                                                    & \setc \\
+\mathcal{\dbasec}^{\textcolor{base}{op}} \arrow[r, "\vindexc^*", color=artist] & \mathcal{\gbasec}^{\textcolor{base}{op}} \arrow[rr, "\sheafc_{\dtotalpullc}", bend left, color=sheaf] \arrow[rr, "\sheaf_{\vartistc}"', bend right, color=sheaf] &                                                    & \setc \\
     &                                                                                                                                                         & {}                                                 &
@@
     & {} \arrow[dd, "\vartistc_{\vindexc^*}" description, Rightarrow, shorten <=1.35em,  shorten >=1.25em,, color=artist] &       \\
-\mathcal{\dbasec}^{\textcolor{base}{op}} \arrow[rr, "\sheafc_{\vindexc^*}(\dtotalc)\circ\vindexc^* = \sheafc(\dtotalc)", bend left, color=sheaf] \arrow[rr, "\sheafc_{\vartistc}(\gtotalc)"', bend right, color=sheaf] &                                                                 & \setc \\
+\mathcal{\dbasec}^{\textcolor{base}{op}} \arrow[rr, "\sheafc_{\vindexpullc} \circ\vindexc^* = \sheafc_{\dtotalc}", bend left, color=sheaf] \arrow[rr, "\sheafc_{\vartistc}"', bend right, color=sheaf] &                                                                 & \setc \\
     & {}                                                              &      
@@
 which in turn provides a way to construct the artist as a map from data section in data space to corresponding graphic section
-\begin{equation}
+
+\note{have I said anything about stalks and fibers yet? this may need to be moved down...}
+
+\begin{align}
   \label{eq:artist:section_map}
-  \vartistc_{\vindexc^*}: \dsectionc \textcolor{artist}{\mapsto}\gsectionc
-\end{equation}
-such that for every dataset, the artists constructs a corresponding graphic generating function. 
+  \vartistc_{\vindexpullc}:\{\dsectionc: \dbasepointc \rightarrow \dfiberc_{\dbasepointc} \mid \dbasepointc \in \opensetc \subseteq \dbasec\} &\rightarrow \{\gsectionc: \gbasepointc \rightarrow \gfiberc\restriction_{\gbasepointc} \mid \gbasepointc = \opensetgc \subseteq \gbasec \}
+\end{align}
+where, as defined in \autoref{eq:artist:construction:xi}, $\vindex(\gbasepoint) = \dbasepoint$ and \gsection\ is in the set of reachable sections $\gsection \in \sheaf_{\vartist}(\gbase)$
+
+the artists constructs a corresponding graphic generating function. 
 \note{maybe mention here Munzner, Zeminsky, and the other dude on glyphs map back to data}
@@
 \end{definition}
+
 The bifunctor $\otimes$ combines categories in a way that  preserves identity and composition \cite{fongInvitationAppliedCategory2019}.

### 2021-12-16 paper/main.tex
O_A to O_H

@@
   \label{eq:artist:data_pullback_sheaf_gamma}
-\sheafc_{\dtotalpullc}: \mathcal{\gbasec}^{\textcolor{base}{op}} \textcolor{sheaf}{\rightarrow} \cgamma{\gbasec}{\dtotalpullc}}
+\sheafc_{\dtotalpullc}: \opensetgc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetgc}{\dtotalpullc}}
 \end{equation}
@@
 
-The set of sets of sections $\{\Gamma(\opensetg, \gtotal)\}\vert_{{\opensetg \in gbase}}$ is subcategory of \textbf{Set}; therefore the full space of graphics is a sheaf $\sheaf_{\gtotal}$
+The set of sets of sections $\{\Gamma(\opensetg, \gtotal)\}\vert_{{\opensetg \in gbase}}$ is subcategory of \textbf{Set}; therefore the full space of graphics is a sheaf 
+\begin{equation}
+  \label{eq:artist:graphic_sheaf_gamma}
+  \sheafc_{\gtotalc}: \opensetgc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetgc}{\gtotalc}
+\end{equation}
+which means it is 
+\begin{equation}
+  \label{eq:artist:graphic_sheaf_set}
+  \sheafc_{\gtotalc}: \gbasec^{\textcolor{base}{op}} \textcolor{sheaf}{\rightarrow} \setc
+\end{equation}
+
+\subsubsection{Data to Graphic}
+Following from \autoref{eq:artist:data_pullback_sheaf} and \autoref{eq:artist:artist_sheaf},  we propose that the artist function $\vartist$ is a natural transformation of sheaf functors 
+\begin{equation}
+  \label{eq:artist:natural-transform-diagram}
+  \begin{tikzcd}
+    & {} \arrow[dd, "\vartistc", Rightarrow, shorten <=1.35em,  shorten >=1.25em, color=artist] & \\
+\mathcal{\gbasec}^{\textcolor{base}{op}} \arrow[rr, "\sheafc_{\dtotalpullc}", bend left, color=sheaf] \arrow[rr, "\sheafc_{\gtotalc}"', bend right, color=sheaf] &  & {\setc}\\
+    & {} &                            
+\end{tikzcd}
+\end{equation}
+
+where each sheaf is a map from an open set in the graphic base space $\opensetg \subset \gbase$ into a set of sections of a fiber bundle over that open set.
+As defined in \autoref{eq:related-work:equivariance:natural-transform}, the artist is a natural transformation.
+Following from \autoref{eq:related-work:equivariance:natural_transform_commute}, the artist specifies that there are compatible inclusion maps in the data and graphic spaces
+
+\begin{equation}
+  \begin{tikzcd}
+    \cgamma{\opensetgc_1}{\dtotalpullc} \arrow[dd, "\iota^*"', hook, color=set] \arrow[rrrr, "\vartistc_{\opensetgc_1}", bend left, color=artist] &  & \opensetgc_1 \arrow[ll, "\sheafc_{\dtotalpullc}", color=sheaf] \arrow[rr, "\sheafc_{\gtotalc}"', color=sheaf] &  & \cgamma{\opensetgc_1}{\gtotalc} \arrow[dd, "\iota^*", hook', color=set] \\
+    {} \arrow[rrrr, dotted, color=artist] &&&& {} \\
+    \cgamma{\opensetgc_2}{\dtotalpullc} \arrow[rrrr, "\vartistc_{\opensetgc_2}"', bend right, color=artist]                            &  & \opensetgc_2 \arrow[ll, "\sheafc_{\dtotalpullc}"', color=sheaf] \arrow[rr, "\sheafc_{\gtotalc}", color=sheaf] \arrow[uu, "\iota" description, hook, color=base] &  & \cgamma{\opensetgc_2}{\gtotalc}
+    \end{tikzcd}
+\end{equation}
+Given that $\opensetg \subseteqq \gbasec$, \autoref{eq:artist:inclusion_morphisms} can be generalized to sections over $\gbase$.
+This fact can be combined with \autoref{eq:artist:data_gamma_pullback} to describe the path from data sections over data space to graphic sections over graphic space
+\begin{equation}
+  \label{eq:artist:gamma_map}
+  \cgamma{\dbasec}{\dtotalc} \textcolor{artist}{\xrightarrow{\vindexc^*}} \cgamma{\gbasec}{\dtotalpullc} \textcolor{artist}{\xrightarrow{\vartist}} \cgamma{\gbasec}{\gtotalc}
+\end{equation}
+
+which means that the artist can be pulled back through $\vindex^*$\footnote{Definition 3.68 in Seven Sketches in Compositionality by Fong and Spivak}
+\begin{equation}
+  \begin{tikzcd}
+    & & {} \arrow[dd, "\vartistc" description, Rightarrow, shorten <=1.35em,  shorten >=1.25em, color=artist] &       \\
+\mathcal{\dbasec}^{\textcolor{base}{op}} \arrow[r, "\vindexc^*", color=artist] & \mathcal{\gbasec}^{\textcolor{base}{op}} \arrow[rr, "\sheafc_{\dtotalpullc}", bend left, color=sheaf] \arrow[rr, "\sheaf_{\gtotalc}"', bend right, color=sheaf] &                                                    & \setc \\
+    &                                                                                                                                                         & {}                                                 &
+\end{tikzcd}
+\end{equation}
+such that the artist can be constructed over data space 
+\begin{equation}
+  \begin{tikzcd}
+    & {} \arrow[dd, "\vartistc_{\vindexc^*}" description, Rightarrow, shorten <=1.35em,  shorten >=1.25em,, color=artist] &       \\
+\mathcal{\dbasec}^{\textcolor{base}{op}} \arrow[rr, "\sheafc_{\dtotalpullc} \circ\vindexpullc = \sheafc_{\dtotalc}", bend left, color=sheaf] \arrow[rr, "\sheafc_{\gtotalc}"', bend right, color=sheaf] &                                                                 & \setc \\
+    & {}                                                              &      
+\end{tikzcd}
+\end{equation}
+
+which in turn provides a way to construct the artist as a map from data section in data space to corresponding graphic section
+
+\subsection{Equivariance}
+Besides preserving inclusions, as described in \autoref{eq:artist:inclusion_morphisms}, the artist function is constructed such that transformations on $\dsection$ and $\gsection$ are equivariant with respect to the morphism between data sheaf functors $\dfunc$ which was introduced in \autoref{eq:artist:morphism-of-sheaf-data}.
+
+\begin{equation}
+  \label{eq:artist:morphism-of-sheaf-data}
+  \begin{tikzcd}
+    & \cgamma{\opensetc_i}{\dtotalc} \in Ob(\setc) & \\
+{} \arrow[rr, "\dfuncc", Rightarrow, pos=.3, shorten <=3.5em, shorten >=8.85em, color=action] & & {} \\
+    & \opensetc_i \in Ob(\mathcal{\dbasec}^{\textcolor{base}{op}}) \arrow[uu, "\sheafc(\dtotalc)", bend left, shift left = 10.5, color=sheaf] \arrow[uu, "\sheafc(\dtotalc)"', bend right, shift left=10.5, color=sheaf] &
+\end{tikzcd}
+\end{equation}
+
+The map $\dfunc: \sheaf(\dtotal) \rightarrow \sheaf(\dtotal)$ is any arbitrary morphism between sheaf functors $\sheaf(\dtotal)$.
+Following from \autoref{eq:related-work:equivariance:natural-transform}, \dfunc\ is a natural transform, which means all inclusions $\iota, \iota^*$ are preserved as part of any $\dfunc$ transformation.
+
+\begin{definition} 
+  \label{def:category:F} The fiber category $\mathcal{\dfiberc}$ is a monoidal category\cite{milewskiCategoryTheoryProgrammers}, which means it is a category equipped with a bifunctur $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$
+  \begin{itemize}
+    \item \textit{object} a monoid object \dfiber. A monoid is a set with a binary operator that is associative, closed, and for which the set contains an identity element $i$ \cite{nlan:monoid}.
+    \item \textit{morphisms} $\dfiber \otimes \dfiber \rightarrow \dfiber$ and unit $i \rightarrow \dfiber$
+  \end{itemize}
+\end{definition}
+
+The bifunctor $\otimes$ combines categories in a way that  preserves identity and composition \cite{fongInvitationAppliedCategory2019}.
+For example, given a pair of categories $\mathcal{\dfiber_a},\mathcal{\dfiber_b}$, the bifunctor yields a set of pairs $\{(\dfiber_a, \dfiber_b) \big{\mvert} \dfiber_a \in \mathcal{\dfiber_a}, \dfiber_b \in \mathcal{\dfiber_b}\}$.
+For the associated morphisms $g: \mathcal{\dfiber_a}\rightarrow \mathcal{\dfiber_a}$ and $h: \mathcal{\dfiber_b}\rightarrow \mathcal{\dfiber_b}$, the function $(g\times h): (\mathcal{\dfiber_a}\times \mathcal{\dfiber_b}) \rightarrow (\mathcal{\dfiber_a}\times \mathcal{\dfiber_b})$ is pointwise $(g\times h)(\dfiber_a, \dfiber_b)\coloneqq (g(\dfiber_a), h(\dfiber_b))$.
+Because of the bifunctor, objects of the category $\mathcal{\dfiber}$ can be the product of any arbitrary number and types of categories, including \textbf{Set}, \textbf{Graph}, and topological spaces \textbf{Top}.
+
+The set of all morphism of the fiber category $Hom_{\mathcal{\dfiber}}(\dfiber, \dfiber)$ are all possible morphisms in a given category.
+This includes any of the functions that serve as the basis of equivariance, such as the binary operations, group actions, and measurement scale discussed in  \autoref{sec:related-work:equivariance}.
+The generalization to $Hom_\mathcal{\dfiber}$ also allows for the inclusion of structures such as monoid actions\cite{barrCategoryTheoryComputing}, which provide a way of applying partial order relations to data\cite{fongInvitationAppliedCategory2019}, such as to build multi-ranked indicators\cite{bruggemannRankingPrioritizationMultiindicator2011}.
+
+When a fiber bundle is trivial, we can dispense with localization and directly consider the fiber maps $Hom(\dfiber_{\dbasepoint}, \dfiber_{\dbasepoint})$ over a point \dbasepoint\ as the basis of our equivariance.
+This is because a sheaf $\sheaf(\dtotal)$ over a limit of open sets that contain a point $\dbasepoint \in \dbase$ is approximately the same as a sheaf over a point \dbasepoint\ and is called the stalk $\sheaf(\dtotal)\restriction_{\dbasepoint}\coloneqq \lim\limits_{\openset\ni \dbasepoint} \Gamma(\openset, \dtotal)$ \cite{StalkSheaf2019}.
+The fiber space over a point is embedded in the stalk over that point $\dfiber_{\dbasepoint}\hookrightarrow \mathscr{\dfiber}_{\dbasepoint}$.
+The sheaf maps \dfunc\ restricted to the stalk are members of $hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$.
+The maps $\dfunc \in Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ include the types of data transformations described in \autoref{sec:related-work:equivariance}In \autoref{sec:artist:construction}, we assume that the fiber bundles are trivial.
+
+\note{}
+
+$\vartist: \sheaf^{\vindex^*} \rightarrow \sheaf_{\vartist}$, it follows from \autoref{eq:related-work:equivariance:functor_diagram} that an arbitrary morphism $\dfunc$ on the data sheaf has a compatible morphism $\dfunc^{\prime}$ on the graphic sheaf.
+\begin{equation}
+  \label{eq:artist:morphism_of_functors}
+  \begin{tikzcd}
+    \sheafc_{\vindexc^{*}} \arrow[d, "\dfuncc"', color=action] \arrow[r, "\vartistc", color=artist] & \sheafc_{\vartistc} \arrow[d, "\dfuncpc \coloneqq\vartistc(\dfuncc)", color=action] \\
+    \sheafc_{\vindexc^{*}} \arrow[r, "\vartistc"', color=artist]                      & \sheafc_{\vartistc}     
+    \end{tikzcd}
+\end{equation}
+As a consequence of \autoref{eq:artist:morphism_of_functors},
+
+\subsubsection{Equivariant Artist}
 
@@
 
-\subsubsection{Data to Graphic}
-Following from \autoref{eq:artist:data_pullback_sheaf} and \autoref{eq:artist:artist_sheaf},  we propose that the artist function $\vartist$ is a natural transformation
-\begin{equation}
-  \label{eq:artist:natural-transform-diagram}
-  \begin{tikzcd}
-    & {} \arrow[dd, "\vartistc", Rightarrow, shorten <=1.35em,  shorten >=1.25em, color=artist] & \\
-\mathcal{\gbasec}^{\textcolor{base}{op}} \arrow[rr, "\sheafc_{\vindexc_*}(\dtotalpullc)", bend left, color=sheaf] \arrow[rr, "\sheafc_{\vartistc}(\gtotalc)"', bend right, color=sheaf] &  & {\setc}\\
-    & {} &                            
-\end{tikzcd}
-\end{equation}
-
-of sheaf functors 
-\begin{equation}
-  \label{eq:artist:sheaf_map}
-  \vartistc: \sheafc_{\vindexc^*}(\dtotalpullc) \textcolor{artist}{\rightarrow} \sheafc_{\vartistc}(\gtotalc)
-\end{equation}
-where each sheaf is a map from an open set in the graphic base space $\opensetg \subset \gbase$ into a set of sections of a fiber bundle over that open set.
-As defined in \autoref{eq:related-work:equivariance:natural-transform}, the artist is a natural transformation.
-Following from \autoref{eq:related-work:equivariance:natural_transform_commute}, the artist specifies that there are compatible inclusion maps in the data and graphic spaces
-
-\begin{equation}
-  \begin{tikzcd}
-    \cgamma{\opensetgc_1}{\dtotalpullc} \arrow[dd, "\iota^*"', hook, color=set] \arrow[rrrr, "\vartistc_{\opensetgc_1}", bend left, color=artist] &  & \opensetgc_1 \arrow[ll, "\sheafc_{\dtotalpullc}", color=sheaf] \arrow[rr, "\sheafc_{\vartistc}"', color=sheaf] &  & \cgamma{\opensetgc_1}{\gtotalc} \arrow[dd, "\iota^*", hook', color=set] \\
-    {} \arrow[rrrr, dotted, color=artist] &&&& {} \\
-    \cgamma{\opensetgc_2}{\dtotalpullc} \arrow[rrrr, "\vartistc_{\opensetgc_2}"', bend right, color=artist]                            &  & \opensetgc_2 \arrow[ll, "\sheafc_{\dtotalpullc}"', color=sheaf] \arrow[rr, "\sheafc_{\vartistc}", color=sheaf] \arrow[uu, "\iota" description, hook, color=base] &  & \cgamma{\opensetgc_2}{\gtotalc}
-    \end{tikzcd}
-\end{equation}
-Given that $\opensetg \subseteqq \gbasec$, \autoref{eq:artist:inclusion_morphisms} can be generalized to sections over $\gbase$.
-This fact can be combined with \autoref{eq:artist:data_gamma_pullback} to describe the path from data sections over data space to graphic sections over graphic space
-\begin{equation}
-  \label{eq:artist:gamma_map}
-  \cgamma{\dbasec}{\dtotalc} \textcolor{artist}{\xrightarrow{\vindexc^*}} \cgamma{\gbasec}{\dtotalpullc} \textcolor{artist}{\xrightarrow{\vartist}} \cgamma{\gbasec}{\gtotalc}
-\end{equation}
-
-which means that the artist can be pulled back through $\vindex^*$\footnote{Definition 3.68 in Seven Sketches in Compositionality by Fong and Spivak}
-\begin{equation}
-  \begin{tikzcd}
-    & & {} \arrow[dd, "\vartistc" description, Rightarrow, shorten <=1.35em,  shorten >=1.25em, color=artist] &       \\
-\mathcal{\dbasec}^{\textcolor{base}{op}} \arrow[r, "\vindexc^*", color=artist] & \mathcal{\gbasec}^{\textcolor{base}{op}} \arrow[rr, "\sheafc_{\dtotalpullc}", bend left, color=sheaf] \arrow[rr, "\sheaf_{\vartistc}"', bend right, color=sheaf] &                                                    & \setc \\
-    &                                                                                                                                                         & {}                                                 &
-\end{tikzcd}
-\end{equation}
-such that the artist can be constructed over data space 
-\begin{equation}
-  \begin{tikzcd}
-    & {} \arrow[dd, "\vartistc_{\vindexc^*}" description, Rightarrow, shorten <=1.35em,  shorten >=1.25em,, color=artist] &       \\
-\mathcal{\dbasec}^{\textcolor{base}{op}} \arrow[rr, "\sheafc_{\vindexpullc} \circ\vindexc^* = \sheafc_{\dtotalc}", bend left, color=sheaf] \arrow[rr, "\sheafc_{\vartistc}"', bend right, color=sheaf] &                                                                 & \setc \\
-    & {}                                                              &      
-\end{tikzcd}
-\end{equation}
-
-which in turn provides a way to construct the artist as a map from data section in data space to corresponding graphic section
-
 \note{have I said anything about stalks and fibers yet? this may need to be moved down...}
@@
 \note{possibly more here as this is the type signature of the thing we implement}
-
-\subsection{Equivariance}
-Besides preserving inclusions, as described in \autoref{eq:artist:inclusion_morphisms}, the artist function is constructed such that transformations on $\dsection$ and $\gsection$ are equivariant with respect to the morphism between data sheaf functors $\dfunc$ which was introduced in \autoref{eq:artist:morphism-of-sheaf-data}.
-
-\begin{equation}
-  \label{eq:artist:morphism-of-sheaf-data}
-  \begin{tikzcd}
-    & \cgamma{\opensetc_i}{\dtotalc} \in Ob(\setc) & \\
-{} \arrow[rr, "\dfuncc", Rightarrow, pos=.3, shorten <=3.5em, shorten >=8.85em, color=action] & & {} \\
-    & \opensetc_i \in Ob(\mathcal{\dbasec}^{\textcolor{base}{op}}) \arrow[uu, "\sheafc(\dtotalc)", bend left, shift left = 10.5, color=sheaf] \arrow[uu, "\sheafc(\dtotalc)"', bend right, shift left=10.5, color=sheaf] &
-\end{tikzcd}
-\end{equation}
-
-The map $\dfunc: \sheaf(\dtotal) \rightarrow \sheaf(\dtotal)$ is any arbitrary morphism between sheaf functors $\sheaf(\dtotal)$.
-Following from \autoref{eq:related-work:equivariance:natural-transform}, \dfunc\ is a natural transform, which means all inclusions $\iota, \iota^*$ are preserved as part of any $\dfunc$ transformation.
-
-\begin{definition} 
-  \label{def:category:F} The fiber category $\mathcal{\dfiberc}$ is a monoidal category\cite{milewskiCategoryTheoryProgrammers}, which means it is a category equipped with a bifunctur $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$
-  \begin{itemize}
-    \item \textit{object} a monoid object \dfiber. A monoid is a set with a binary operator that is associative, closed, and for which the set contains an identity element $i$ \cite{nlan:monoid}.
-    \item \textit{morphisms} $\dfiber \otimes \dfiber \rightarrow \dfiber$ and unit $i \rightarrow \dfiber$
-  \end{itemize}
-\end{definition}
-
-The bifunctor $\otimes$ combines categories in a way that  preserves identity and composition \cite{fongInvitationAppliedCategory2019}.
-For example, given a pair of categories $\mathcal{\dfiber_a},\mathcal{\dfiber_b}$, the bifunctor yields a set of pairs $\{(\dfiber_a, \dfiber_b) \big{\mvert} \dfiber_a \in \mathcal{\dfiber_a}, \dfiber_b \in \mathcal{\dfiber_b}\}$.
-For the associated morphisms $g: \mathcal{\dfiber_a}\rightarrow \mathcal{\dfiber_a}$ and $h: \mathcal{\dfiber_b}\rightarrow \mathcal{\dfiber_b}$, the function $(g\times h): (\mathcal{\dfiber_a}\times \mathcal{\dfiber_b}) \rightarrow (\mathcal{\dfiber_a}\times \mathcal{\dfiber_b})$ is pointwise $(g\times h)(\dfiber_a, \dfiber_b)\coloneqq (g(\dfiber_a), h(\dfiber_b))$.
-Because of the bifunctor, objects of the category $\mathcal{\dfiber}$ can be the product of any arbitrary number and types of categories, including \textbf{Set}, \textbf{Graph}, and topological spaces \textbf{Top}.
-
-The set of all morphism of the fiber category $Hom_{\mathcal{\dfiber}}(\dfiber, \dfiber)$ are all possible morphisms in a given category.
-This includes any of the functions that serve as the basis of equivariance, such as the binary operations, group actions, and measurement scale discussed in  \autoref{sec:related-work:equivariance}.
-The generalization to $Hom_\mathcal{\dfiber}$ also allows for the inclusion of structures such as monoid actions\cite{barrCategoryTheoryComputing}, which provide a way of applying partial order relations to data\cite{fongInvitationAppliedCategory2019}, such as to build multi-ranked indicators\cite{bruggemannRankingPrioritizationMultiindicator2011}.
-
-When a fiber bundle is trivial, we can dispense with localization and directly consider the fiber maps $Hom(\dfiber_{\dbasepoint}, \dfiber_{\dbasepoint})$ over a point \dbasepoint\ as the basis of our equivariance.
-This is because a sheaf $\sheaf(\dtotal)$ over a limit of open sets that contain a point $\dbasepoint \in \dbase$ is approximately the same as a sheaf over a point \dbasepoint\ and is called the stalk $\sheaf(\dtotal)\restriction_{\dbasepoint}\coloneqq \lim\limits_{\openset\ni \dbasepoint} \Gamma(\openset, \dtotal)$ \cite{StalkSheaf2019}.
-The fiber space over a point is embedded in the stalk over that point $\dfiber_{\dbasepoint}\hookrightarrow \mathscr{\dfiber}_{\dbasepoint}$.
-The sheaf maps \dfunc\ restricted to the stalk are members of $hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$.
-The maps $\dfunc \in Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ include the types of data transformations described in \autoref{sec:related-work:equivariance}In \autoref{sec:artist:construction}, we assume that the fiber bundles are trivial.
-
-\note{}
-
-$\vartist: \sheaf^{\vindex^*} \rightarrow \sheaf_{\vartist}$, it follows from \autoref{eq:related-work:equivariance:functor_diagram} that an arbitrary morphism $\dfunc$ on the data sheaf has a compatible morphism $\dfunc^{\prime}$ on the graphic sheaf.
-\begin{equation}
-  \label{eq:artist:morphism_of_functors}
-  \begin{tikzcd}
-    \sheafc_{\vindexc^{*}} \arrow[d, "\dfuncc"', color=action] \arrow[r, "\vartistc", color=artist] & \sheafc_{\vartistc} \arrow[d, "\dfuncpc \coloneqq\vartistc(\dfuncc)", color=action] \\
-    \sheafc_{\vindexc^{*}} \arrow[r, "\vartistc"', color=artist]                      & \sheafc_{\vartistc}     
-    \end{tikzcd}
-\end{equation}
-As a consequence of \autoref{eq:artist:morphism_of_functors},
 

### 2021-12-16 paper/main.tex
more rearranging/strippng out O_A

@@
 
-$\vartist: \sheaf^{\vindex^*} \rightarrow \sheaf_{\vartist}$, it follows from \autoref{eq:related-work:equivariance:functor_diagram} that an arbitrary morphism $\dfunc$ on the data sheaf has a compatible morphism $\dfunc^{\prime}$ on the graphic sheaf.
+It follows from \autoref{eq:related-work:equivariance:functor_diagram} that an arbitrary morphism $\dfunc$ on the data sheaf has a compatible morphism $\dfunc^{\prime}$ on the graphic sheaf. 
 \begin{equation}
@@
   \begin{tikzcd}
-    \sheafc_{\vindexc^{*}} \arrow[d, "\dfuncc"', color=action] \arrow[r, "\vartistc", color=artist] & \sheafc_{\vartistc} \arrow[d, "\dfuncpc \coloneqq\vartistc(\dfuncc)", color=action] \\
-    \sheafc_{\vindexc^{*}} \arrow[r, "\vartistc"', color=artist]                      & \sheafc_{\vartistc}     
+    \sheafc_{\dtotalc} \arrow[d, "\dfuncc"', color=action] \arrow[r, "\vartistc_{\vindexpullc}", color=artist] & \sheafc_{\gtotalc} \arrow[d, "\dfuncpc \coloneqq\vartistc_{\vindexpullc}(\dfuncc)", color=action] \\
+    \sheafc_{\dtotalc} \arrow[r, "\vartistc_{\vindexpullc}"', color=artist]                      & \sheafc_{\gtotalc}     
     \end{tikzcd}
@@
     \textcolor{set}{\big\{}\gsectionc:\opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} 
-    \bigm{\vert} \vartistc\circ\dfuncc \circ \vartistc^{-1}[\gsectionc] = \dfuncc^{\prime} \circ \gsectionc, \dsectionc \in \sheafc_{\dtotal}
-    \textcolor{set}{\big\}} 
+    \bigm{\vert} \vartistc\circ (\dfuncc \circ \vartistc^{-1}[\gsectionc]) = \dfuncc^{\prime} \circ \gsectionc \textcolor{set}{\big\}} 
     \arrow[ll, "\textcolor{set}{\mathlarger{\mathlarger{\supset}}}" description, hook', color=white] \\ &  & \\

### 2021-12-20 paper/main.tex
moved library to libraries section

@@
 This means that often there is no way to express data continuity independent of visualization type, meaning most visualization libraries will allow, for example, visualizing discrete data as a line plot or an image.
-General purpose visualization libraries-such as Matplotlib\cite{hunterMatplotlib2DGraphics2007}, Vtk\cite{hanwellVisualizationToolkitVTK2015,geveciVTK2012}, and D3 \cite{bostockDataDrivenDocuments2011}-carry distinct data models as part of the implementation of each visual algorithm.
-The lack of unified data model means that each plot in a linked\cite{beckerBrushingScatterplots1987,bujaInteractiveData1991} visualization is treated as independent, as are the transforms converting each field in the data to a visual equivalent.
-
-Domain specific libraries can often guarantee consistency because they have a single model of the data in their software design, as discussed in Heer and Agarwal \cite{HeerSoftware2006}'s survey of visualization software design patterns.
-For example, the relational database is core to tools influenced by APT, such as Tableau\cite{StoltePolaris2002,hanrahanVizQL2006,MackinlayShowme2007} and the Grammar of Graphics\cite{wilkinsonGrammarGraphics2005} inspired ggplot\cite{wickhamGgplot2ElegantGraphics2016a}, Vega\cite{satyanarayanDeclarativeInteractionDesign2014} and Altair\cite{vanderplasAltairInteractiveStatistical2018}.
-Images underpin scientific visualization tools such as Napari\cite{nicholas_sofroniew_2021_4533308} and ImageJ\cite{schneiderNIHImageImageJ2012} and the digital humanities oriented ImagePlot\cite{studiesCulturevisImageplot2021} macro; the need to visualize and manipulate graphs has spawned tools like Gephi\cite{bastianGephiOpenSource2009}, Graphviz\cite{ellsonGraphvizOpenSource2002}, and Networkx\cite{HagbergExploringNetwork2008}.
 
@@
 \end{figure}
+
+\subsection{Visualization Library Architecture}
+General purpose visualization libraries-such as Matplotlib\cite{hunterMatplotlib2DGraphics2007}, Vtk\cite{hanwellVisualizationToolkitVTK2015,geveciVTK2012}, and D3 \cite{bostockDataDrivenDocuments2011}-carry distinct data models as part of the implementation of each visual algorithm.
+The lack of unified data model means that each plot in a linked\cite{beckerBrushingScatterplots1987,bujaInteractiveData1991} visualization is treated as independent, as are the transforms converting each field in the data to a visual equivalent.
+
+Domain specific libraries can often guarantee consistency because they have a single model of the data in their software design, as discussed in Heer and Agarwal \cite{HeerSoftware2006}'s survey of visualization software design patterns.
+For example, the relational database is core to tools influenced by APT, such as Tableau\cite{StoltePolaris2002,hanrahanVizQL2006,MackinlayShowme2007} and the Grammar of Graphics\cite{wilkinsonGrammarGraphics2005} inspired ggplot\cite{wickhamGgplot2ElegantGraphics2016a}, Vega\cite{satyanarayanDeclarativeInteractionDesign2014} and Altair\cite{vanderplasAltairInteractiveStatistical2018}.
+Images underpin scientific visualization tools such as Napari\cite{nicholas_sofroniew_2021_4533308} and ImageJ\cite{schneiderNIHImageImageJ2012} and the digital humanities oriented ImagePlot\cite{studiesCulturevisImageplot2021} macro; the need to visualize and manipulate graphs has spawned tools like Gephi\cite{bastianGephiOpenSource2009}, Graphviz\cite{ellsonGraphvizOpenSource2002}, and Networkx\cite{HagbergExploringNetwork2008}.
 

### 2021-12-20 paper/main.tex
cleaning up latex compile errors + intro of pullback/pushforward

@@
 \section{Algebraic Topology \& Category Theory}
-\note{potentially move this to appendix, if so push Butler & Vickers up to RW}
+\note{potentially move this to appendix, push Butler \& Vickers up to RW}
 
@@
 The model described in this work provides a model for expressing data sets with different topological properties.
+
+\subsection{Fiber Bundles}
 We obtain this generality by using a mathematical structure called a fiber bundles as the basis of our abstraction, as proposed by Butler, Bryson, and Pendley\cite{butlerVisualizationModelBased1989,
@@
 
-\subsection{Sheaves}
-\label{sec:related-work:continuity:sheaves}
-Sheaves can be thought of as an "algebraic data structure" \cite{ghristElementaryAppliedTopology2014} for data that lives over topological spaces.
-A presheaf $\sheafc$ is a map from the open set to the set of sections $\sheafc: \opensetc \rightarrow \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$ \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}.
-The presheaf preserves inclusion maps $\iota$ between open sets and the inclusion map $\iota^*$ between the sets of sections.
-
-\begin{equation}
-  \label{eq:related-work:continuity:presheaf}
-  \begin{tikzcd}[column sep=small]
-    {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} & {} & {\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}} \arrow[ll, "\iota^*"', hook'] \\& &\\
-    \opensetc_1 \arrow[uu, "\sheafc_{\dtotal}"] \arrow[rr, "\iota"', hook] & {} \arrow[uu, "\sheafc_{\dtotal}" description, dotted] & \opensetc_2 \arrow[uu, "\sheafc_\dtotal"']                    
-    \end{tikzcd}
-\end{equation}
-
-This means that a smaller space $\openset_1$ is included in a larger space $\openset_2$ and a function $\dsection \in \Gamma(\openset, \dtotal\restriction_{\openset})$ that is continuous over a larger space $\openset_2$ is continuous over a subspace $\openset_1 \subset\openset_2$.
-Sheaves are presheafs where sets of sections are defined over unions of open sets over a topology $\bigcup_{j \in I} \openset)_j \in \dbase$ \cite{spanier1989algebraic, nlab:sheaf}.
-Sheaves provide a mathematical abstraction of data, the space it lives over, and relationships between subsets.
-
 \subsection{Category Theory}
@@
 
- \begin{definition} A presheaf \sheafc\, as introduced in \autoref{eq:related-work:continuity:presheaf}, is a functor $\sheaf: \mathcal{C}^{op} \rightarrow \textbf{Set}$.\cite{nlab:presheaf } \end{definition} \note{possibly add in from Hader lecture notes}
-
  \begin{definition} A \textcolor{nattran}{natural transformation} is a morphism of functors $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\rightarrow} \textcolor{functor}{G}$ where $F, G$ are functors from $\mathcal{C}$ to $\mathcal{D}$.
@@
 \note{unpack this a drop more}
+
+\subsection{Sheaves}
+\label{sec:related-work:continuity:sheaves}
+Sheaves provide a mathematical abstraction of data, the space it lives over, and relationships between subsets.
+Sheaves can be thought of as an "algebraic data structure" \cite{ghristElementaryAppliedTopology2014} for data that lives over topological spaces.
+A presheaf  $\sheaf$ applied to fiber bundles is a map from an open set in a topological space, such $\openset \subset \dbase$, to the set of sections
+\begin{equation}
+  \label{eq:atct:presheaf}
+  \sheafc: \opensetc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
+\end{equation}
+over that openset \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}.
+The presheaf preserves the inclusion maps $\iota$ between open sets and the restriction maps $\iota^*$ between the sets of sections over that open sets.
+\begin{equation}
+  \label{eq:related-work:continuity:presheaf}
+  \begin{tikzcd}[column sep=small]
+    {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} & {} & {\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}} \arrow[ll, "\iota^*"', hook', color=set] \\& &\\
+    \opensetc_1 \arrow[uu, "\sheafc_{\dtotalc}", color=sheaf] \arrow[rr, "\iota"', hook, color=base] & {} \arrow[uu, "\sheafc_{\dtotalc}" description, dotted, color=sheaf] & \opensetc_2 \arrow[uu, "\sheafc_{\dtotalc}"', color=sheaf]
+    \end{tikzcd}
+\end{equation}
+
+This means that a smaller space $\openset_1$ is included in a larger space $\openset_2$ and a function $\dsection \in \Gamma(\openset, \dtotal\restriction_{\openset})$ that is continuous over a larger space $\openset_2$ is continuous over a subspace $\openset_1 \subset\openset_2$.
+
+A presheaf \sheafc\, as introduced in \autoref{eq:related-work:continuity:presheaf}, is a contravariant functor\cite{nlab:presheaf} from a category to a set. 
+\begin{equation}
+  \label{eq:actc:sheaf}
+ \sheafc: \textcolor{source}{\mathcal{C}^{op}} \rightarrow \textcolor{target}{\textbf{Set}}
+\end{equation}
+
+A presheaf $\sheaf(\dtotal)$ over a limit of open sets that contain a point $\dbasepoint \in \dbase$ is approximately the same as a sheaf over a point \dbasepoint\ and is called the stalk\cite{StalkSheaf2019}
+\begin{equation}
+  \label{eq:atct:stalk}
+  \sheafc(\dbasec)\restriction_{\dbasepointc}\coloneqq \lim\limits_{\opensetc\ni \dbasepointc} \Gamma(\opensetc, \dtotalc\restriction_{\opensetc}) 
+\end{equation}
+
+Sheaves are presheafs where sections are continuous over all unions of open sets in $\dbase$\cite{harder2008lectures, spanier1989algebraic, nlab:sheaf}.
+This means that a section which is defined over a stalk $\dsection \in \sheaf(\dbase)\vert_{\dbasepoint_i}$ and evaluated at a point $\dbasepoint_j$ is equal to a section in the stalk over that point $\dsection_{dbasepoint_i}\vert_{dbasepoint_j} = \dsection_{\dbasepoint_j}$
+
+We can define natural transforms, introduced in \autoref{eq:related-work:equivariance:natural-transform}, between sheaves over different spaces, for example spaces representing the topology of the data and the graphic, because sheaves can be moved over different topological spaces if there is a map between those spaces.
+Given a map $\textcolor{functor}{f}:\textcolor{source}{X} \textcolor{functor}{\rightarrow} \textcolor{target}{Y}$ between topological spaces $\textcolor{source}{X}, \textcolor{target}{Y}$ and sheaves on each space $\sheafc_{\textcolor{source}{X}}, \sheafc_{\textcolor{target}{Y}}$, we can construct the pullback functor $f^*$ and pushforward functor $f_*$ \cite{harder2008lectures} such that the following diagram commutes
+\begin{equation}
+  \label{eq:atct:pullback-pushforward}
+  abc
+\end{equation}
+
+\note{possibly add in from Hader lecture notes}
 
@@
 For that line to have any thickness, it must be rendered as a ban.
-One method of obtaining this graphic continuity while preserving the constraint that it is a deformation retract is by multiplying the base space \dbase\ by an interval $I=\[0,1\]$ such that $\gbase = \dbase \times I^{1}$.
+One method of obtaining this graphic continuity while preserving the constraint that it is a deformation retract is by multiplying the base space \dbase\ by an interval $I=\[0,1\]$ such that $\gbase = \dbase \times I$.
 This construction ensures that every region $(\dbasepoint, \alpha)\vert_{\alpha \in I} \subset \gbase$ maps to a  corresponding point $\dbasepoint \in \dbase$.
@@
 When a fiber bundle is trivial, we can dispense with localization and directly consider the fiber maps $Hom(\dfiber_{\dbasepoint}, \dfiber_{\dbasepoint})$ over a point \dbasepoint\ as the basis of our equivariance.
-This is because a sheaf $\sheaf(\dtotal)$ over a limit of open sets that contain a point $\dbasepoint \in \dbase$ is approximately the same as a sheaf over a point \dbasepoint\ and is called the stalk $\sheaf(\dtotal)\restriction_{\dbasepoint}\coloneqq \lim\limits_{\openset\ni \dbasepoint} \Gamma(\openset, \dtotal)$ \cite{StalkSheaf2019}.
-The fiber space over a point is embedded in the stalk over that point $\dfiber_{\dbasepoint}\hookrightarrow \mathscr{\dfiber}_{\dbasepoint}$.
+The fiber space over a point is embedded in the stalk, introduced in
+
+over that point $\dfiber_{\dbasepoint}\hookrightarrow \mathscr{\dfiber}_{\dbasepoint}$.
 The sheaf maps \dfunc\ restricted to the stalk are members of $hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$.

### 2021-12-20 paper/main.tex
top sheave then cat sheaf

@@
 Sheaves can be thought of as an "algebraic data structure" \cite{ghristElementaryAppliedTopology2014} for data that lives over topological spaces.
-A presheaf  $\sheaf$ applied to fiber bundles is a map from an open set in a topological space, such $\openset \subset \dbase$, to the set of sections
+A presheaf  $\sheaf$ applied to fiber bundles is a map from an open set in a topological space, such as $\openset \subset \dbase$ where $(\mathcal{T}, \dbase)$ is a topological space, to the set of sections of a bundle $\dtotal$ over that openset \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}.
 \begin{equation}
@@
 \end{equation}
-over that openset \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}.
 The presheaf preserves the inclusion maps $\iota$ between open sets and the restriction maps $\iota^*$ between the sets of sections over that open sets. 
@@
     {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} & {} & {\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}} \arrow[ll, "\iota^*"', hook', color=set] \\& &\\
-    \opensetc_1 \arrow[uu, "\sheafc_{\dtotalc}", color=sheaf] \arrow[rr, "\iota"', hook, color=base] & {} \arrow[uu, "\sheafc_{\dtotalc}" description, dotted, color=sheaf] & \opensetc_2 \arrow[uu, "\sheafc_{\dtotalc}"', color=sheaf]
-    \end{tikzcd}
-\end{equation}
-
-This means that a smaller space $\openset_1$ is included in a larger space $\openset_2$ and a function $\dsection \in \Gamma(\openset, \dtotal\restriction_{\openset})$ that is continuous over a larger space $\openset_2$ is continuous over a subspace $\openset_1 \subset\openset_2$.
-
-A presheaf \sheafc\, as introduced in \autoref{eq:related-work:continuity:presheaf}, is a contravariant functor\cite{nlab:presheaf} from a category to a set. 
-\begin{equation}
-  \label{eq:actc:sheaf}
- \sheafc: \textcolor{source}{\mathcal{C}^{op}} \rightarrow \textcolor{target}{\textbf{Set}}
-\end{equation}
-
-A presheaf $\sheaf(\dtotal)$ over a limit of open sets that contain a point $\dbasepoint \in \dbase$ is approximately the same as a sheaf over a point \dbasepoint\ and is called the stalk\cite{StalkSheaf2019}
+    \opensetc_1 \arrow[uu, "\sheafc_{\dbasec}", color=sheaf] \arrow[rr, "\iota"', hook, color=base] & {} \arrow[uu, "\sheafc_{\dbasec}" description, dotted, color=sheaf] & \opensetc_2 \arrow[uu, "\sheafc_{\dbasec}"', color=sheaf]
+    \end{tikzcd}
+\end{equation}
+
+This means that a smaller space $\openset_1$ is included in a larger space $\openset_2$ and a function in a set of global sections over a bundle $\dsection \in \Gamma(\openset_2, \dtotal\restriction_{\openset2})$ that is continuous over a larger space $\openset_2$ is continuous when restricted to a subspace $\dsection \in \Gamma(\openset_1,\dtotal\restriction_{\openset2}),\openset_1 \subset\openset_2$.
+
+A presheaf $\sheaf$ over a limit of open sets that contain a point $\dbasepoint \in \dbase$ is approximately the same as a sheaf over a point \dbasepoint\ and is called the stalk\cite{StalkSheaf2019} 
 \begin{equation}
@@
 Sheaves are presheafs where sections are continuous over all unions of open sets in $\dbase$\cite{harder2008lectures, spanier1989algebraic, nlab:sheaf}.
-This means that a section which is defined over a stalk $\dsection \in \sheaf(\dbase)\vert_{\dbasepoint_i}$ and evaluated at a point $\dbasepoint_j$ is equal to a section in the stalk over that point $\dsection_{dbasepoint_i}\vert_{dbasepoint_j} = \dsection_{\dbasepoint_j}$
-
+This means that a section which is defined over a stalk $\dsection \in \sheaf(\dbase)\vert_{\dbasepoint_i}$ and evaluated at a point $\dbasepoint_j$ is equal to a section in the stalk over that point $\dsection_{\dbasepoint_i}\vert_{\dbasepoint_j} = \dsection_{\dbasepoint_j}$
+
+In category theory, a presheaf \sheafc\, is generalized to be a contravariant functor\cite{nlab:presheaf} from a category to a set. 
+\begin{equation}
+  \label{eq:actc:sheaf}
+ \sheafc: \textcolor{source}{\mathcal{C}^{op}} \rightarrow \setc
+\end{equation}
 We can define natural transforms, introduced in \autoref{eq:related-work:equivariance:natural-transform}, between sheaves over different spaces, for example spaces representing the topology of the data and the graphic, because sheaves can be moved over different topological spaces if there is a map between those spaces.
-Given a map $\textcolor{functor}{f}:\textcolor{source}{X} \textcolor{functor}{\rightarrow} \textcolor{target}{Y}$ between topological spaces $\textcolor{source}{X}, \textcolor{target}{Y}$ and sheaves on each space $\sheafc_{\textcolor{source}{X}}, \sheafc_{\textcolor{target}{Y}}$, we can construct the pullback functor $f^*$ and pushforward functor $f_*$ \cite{harder2008lectures} such that the following diagram commutes
+Given a map $\textcolor{functor}{f}:\textcolor{source}{X} \textcolor{functor}{\rightarrow} \textcolor{target}{Y}$ between topological spaces $\textcolor{source}{X}, \textcolor{target}{Y}$ and sheaves on each space $\sheafc_{\textcolor{source}{X}}, \sheafc_{\textcolor{target}{Y}}$, we can construct a pullback functor $f^*$ and pushforward functor $f_*$ \cite{harder2008lectures} that can be composed with the sheaf functors.
 \begin{equation}
   \label{eq:atct:pullback-pushforward}
-  abc
-\end{equation}
+  \begin{tikzcd}[row sep=huge]
+    \sheafc_{\textcolor{source}{X}}(\textcolor{source}{X}) \arrow[rr, "\textcolor{functor}{f_*}", color=functor]                            &  & (\textcolor{functor}{f_*}\sheafc_{\textcolor{source}{X}})(\textcolor{target}{Y})    \\
+    \textcolor{source}{X} \arrow[rr, "\textcolor{functor}{f}", color=functor] \arrow[u, "\sheaf_{\textcolor{source}{X}}", color=sheaf] \arrow[d, dashed, color=sheaf] &  & \textcolor{target}{Y} \arrow[d, "\sheaf_{\textcolor{target}{Y}}", color=sheaf] \arrow[u, dashed, color=sheaf] \\
+    (\textcolor{functor}{f^*}\sheafc_{\textcolor{target}{Y}})(\textcolor{source}{X}) \arrow[rr, "\textcolor{functor}{f^*}", color=functor]  &  & \sheafc_{\textcolor{target}{Y}}(\textcolor{target}{Y})
+  \end{tikzcd}
+\end{equation}
+As shown in \autoref{eq:atct:pullback-pushforward}, the composition $f_*\sheaf_{X}$ modifies the functor $\sheaf_{X}$ such that sections $\dsection \in \sheaf_{X}
 

### 2021-12-20 paper/main.tex
section function mods

@@
 \end{equation}
-As shown in \autoref{eq:atct:pullback-pushforward}, the composition $f_*\sheaf_{X}$ modifies the functor $\sheaf_{X}$ such that sections $\dsection \in \sheaf_{X}
+As shown in \autoref{eq:atct:pullback-pushforward}, the pushforward $f_*\sheaf_{X}$ modifies the sheaf $\sheaf_{X}$ such that sections $s \in \sheaf_{X}(X)$ can be evaluated over points in $y$
+\begin{equation}
+  \label{eq:atct:pushforward-section}
+  (s: x \rightarrow \setb) \xrightarrow{f_*} (s_*: y \rightarrow \setb)
+\end{equation}
+while the pullback functor $f^*$ modifies the sheaf  $\sheaf_{Y}$ such that sections $\t \in \sheaf_{Y}(Y)$ can be evaluated over points in $x$
+\begin{equation}
+  \label{eq:atct:pullback-section}
+  (t^*: x \rightarrow \setb) \xrightarrow{f^*} (t: y \rightarrow \setb)
+\end{equation}
+As shown in \autoref{eq:atct:pushforward-section}, the pushforward modifies the signature of the functions returned by the functor , while in \autoref{eq:atct:pullback-section} the pullback modifies the signature of the argument to the functor.
 

### 2021-12-22 paper/main.tex
sheaves section

--- 2021-12-20 paper/main.tex
@@
   \label{eq:actc:sheaf}
- \sheafc: \textcolor{source}{\mathcal{C}^{op}} \rightarrow \setc
-\end{equation}
-We can define natural transforms, introduced in \autoref{eq:related-work:equivariance:natural-transform}, between sheaves over different spaces, for example spaces representing the topology of the data and the graphic, because sheaves can be moved over different topological spaces if there is a map between those spaces.
+ \sheafc: \textcolor{source}{\mathcal{C}^{op}} \textcolor{sheaf}{\rightarrow} \setc
+\end{equation}
+
+Sheaves can be moved over different topological spaces if there is a map between those spaces, which we can use to describe how sections of a sheaf over one space behave when moved over to a different space and the maps between these spaces.
 Given a map $\textcolor{functor}{f}:\textcolor{source}{X} \textcolor{functor}{\rightarrow} \textcolor{target}{Y}$ between topological spaces $\textcolor{source}{X}, \textcolor{target}{Y}$ and sheaves on each space $\sheafc_{\textcolor{source}{X}}, \sheafc_{\textcolor{target}{Y}}$, we can construct a pullback functor $f^*$ and pushforward functor $f_*$ \cite{harder2008lectures} that can be composed with the sheaf functors.
@@
 \end{equation}
-As shown in \autoref{eq:atct:pullback-pushforward}, the composition $f_*\sheaf_{X}$ modifies the functor $\sheaf_{X}$ such that sections $\dsection \in \sheaf_{X}
-
-\note{possibly add in from Hader lecture notes}
+As shown in \autoref{eq:atct:pullback-pushforward}, the pushforward $f_*\sheaf_{X}$ modifies the sheaf $\sheaf_{X}$ such that sections $s \in \sheaf_{X}$ can be evaluated over points in $y$
+\begin{equation}
+  \label{eq:atct:pushforward-section}
+  (\textcolor{section}{s}: \textcolor{source}{x} \textcolor{section}{\rightarrow} \setc) \textcolor{functor}{\xrightarrow{\mathmakebox[3em]{f_*}}} (\textcolor{functor}{f_*}\textcolor{section}{s}: \textcolor{target}{y} \textcolor{section}{\rightarrow} \setc)
+\end{equation}
+which means it decorates the section function $s$ so that it can take as input values in $Y$ and return the corresponding section in $X$ 
+\begin{equation}
+  \label{eq:atct:pushforward-f}
+  \textcolor{section}{s}(\textcolor{source}{x}) 
+  = \textcolor{section}{s}\textcolor{functor}{_*}(\textcolor{target}{y}) =\textcolor{section}{s}(\textcolor{functor}{f^{-1}}(\textcolor{target}{y}))
+\end{equation}
+The pullback functor $f^*$ modifies the sheaf  $\sheaf_{Y}$ such that sections $t \in \sheaf_{Y}$ can be evaluated over points in $x$
+\begin{equation}
+  \label{eq:atct:pullback-section}
+  (\textcolor{functor}{f^*}\textcolor{section}{t}: \textcolor{source}{x} \textcolor{section}{\rightarrow} \setc) \textcolor{functor}{\xrightarrow{\mathmakebox[3em]{f^*}}} (\textcolor{section}{t}: \textcolor{target}{y} \textcolor{section}{\rightarrow} \setc)
+\end{equation}
+which means it decorates the section function $t$ so that it can take as input values in $X$ and return the corresponding section in $Y$ 
+\begin{equation}
+  \label{eq:atct:pullback-f}
+  \textcolor{section}{t}\textcolor{functor}{^*}(\textcolor{source}{x})
+  =\textcolor{section}{t}(\textcolor{functor}{f}(\textcolor{source}{x}))
+  =\textcolor{section}{t}(\textcolor{target}{y}) 
+\end{equation}
+The pushforward and pullback functors are isomorphic \cite{harder2008lectures}
+\begin{equation}
+  \label{eq:atct:pullback-pushforward-isomorphic}
+  \textcolor{homset}{hom}_{\sheafc_{\textcolor{source}{X}}}(\textcolor{functor}{f*}\sheafc_{\textcolor{target}{Y}}, \sheafc_{\textcolor{source}{X}}) = 
+  \textcolor{homset}{hom}_{\sheafc_{\textcolor{target}{Y}}}(\sheafc_{\textcolor{target}{Y}}, \textcolor{functor}{f_*}\sheafc_{\textcolor{source}{X}})
+\end{equation}
+such that the diagram is commutative
+\begin{equation}
+  \label{eq:pullback-pushforward-homset}
+  \begin{tikzcd}
+    \sheafc_{\textcolor{source}{X}}(\textcolor{source}{X}) \arrow[rr, "\textcolor{functor}{f_*}", color=functor]  &  & (\textcolor{functor}{f_*}\sheafc_{\textcolor{source}{X}})(\textcolor{target}{Y}) \\
+    && \\
+    (\textcolor{functor}{f^*}\sheafc_{\textcolor{target}{Y}})(\textcolor{source}{X}) \arrow[rr, "\textcolor{functor}{f^*}", color=functor] \arrow[uu, "\textcolor{homset}{Hom_{\sheaf_X}}", color=homset] &  & \sheafc_{\textcolor{target}{Y}}(\textcolor{target}{Y}) \arrow[uu, "\textcolor{homset}{Hom_{\sheaf_Y}}"', color=homset] \arrow[lluu, "\textcolor{homset}{Hom_{XY}}"' description, dashed, color=homset]
+    \end{tikzcd}
+\end{equation}
+meaning that transformations of sections of one sheaf to sections of another are equivalent regardless of the space the sheaf is over.
 

### 2021-12-22 paper/main.tex
homset

@@
 
+\note{Munzer describes continuiity as,  We obtain this generality by using a mathematical structure called a fiber bundles as the basis of our abstraction, as proposed by Butler, Bryson, and Pendley\cite{butlerVisualizationModelBased1989,
+butlerVectorBundleClassesForm1992}}
+
 \subsection{Equivariance}
@@
 
+\note{fig and K&S model, vickers}
+
 \subsection{Visualization Library Architecture}
@@
 The model described in this work provides a model for expressing data sets with different topological properties.
-
-\subsection{Fiber Bundles}
 We obtain this generality by using a mathematical structure called a fiber bundles as the basis of our abstraction, as proposed by Butler, Bryson, and Pendley\cite{butlerVisualizationModelBased1989,
@@
     && \\
-    (\textcolor{functor}{f^*}\sheafc_{\textcolor{target}{Y}})(\textcolor{source}{X}) \arrow[rr, "\textcolor{functor}{f^*}", color=functor] \arrow[uu, "\textcolor{homset}{Hom_{\sheaf_X}}", color=homset] &  & \sheafc_{\textcolor{target}{Y}}(\textcolor{target}{Y}) \arrow[uu, "\textcolor{homset}{Hom_{\sheaf_Y}}"', color=homset] \arrow[lluu, "\textcolor{homset}{Hom_{XY}}"' description, dashed, color=homset]
-    \end{tikzcd}
-\end{equation}
-meaning that transformations of sections of one sheaf to sections of another are equivalent regardless of the space the sheaf is over.
+    (\textcolor{functor}{f^*}\sheafc_{\textcolor{target}{Y}})(\textcolor{source}{X}) \arrow[rr, "\textcolor{functor}{f^*}", color=functor] \arrow[uu, "\textcolor{homset}{hom_{\sheaf_X}}", color=homset] &  & \sheafc_{\textcolor{target}{Y}}(\textcolor{target}{Y}) \arrow[uu, "\textcolor{homset}{hom_{\sheaf_Y}}"', color=homset] \arrow[lluu, "\textcolor{homset}{hom_{XY}}"' description, dashed, color=homset]
+    \end{tikzcd}
+\end{equation}
+meaning that morphisms between sections over different topological spaces are equivalent when composed with the pullback and pushforward functor.
 

### 2021-12-22 paper/main.tex
artist sheaf map

@@
 \section{Algebraic Topology \& Category Theory}
+\label{sec:atct}
 \note{potentially move this to appendix, push Butler \& Vickers up to RW}
@@
 \subsection{Fiber Bundles}
-\label{sec:related-work:fiber-bundles}
+\label{sec:atct:fiber-bundles}
 \begin{figure}[h!]
@@
 \subsection{Category Theory}
-\label{sec:related-work:equivariance:category}
+\label{sec:atct:equivariance:category}
 In this work, we propose that equivariance constraints can be expressed using category theory.
@@
 \subsection{Sheaves}
-\label{sec:related-work:continuity:sheaves}
+\label{sec:atct:continuity:sheaves}
 Sheaves provide a mathematical abstraction of data, the space it lives over, and relationships between subsets.
@@
   \label{eq:artist:artist}
-  \vartistc: \mathcal{\dtotalc} \rightarrow \mathcal{\gtotalc}
+  \mathcal{\dtotalc} \textcolor{artist}{\xrightarrow{\vartist}} \mathcal{\gtotalc}
 \end{equation}
@@
   \begin{itemize}
-    \item \textit{objects} open sets $\openset_{i} \in \{\openset\}$ in the topological space $(\dbase, \mathscr{T})$), including the empty set $\varnothing$ and the maximum set \dbase. 
-    \item \textit{morphisms} $\iota: \openset_{i} \hookrightarrow \openset_{j}$ for all  $\openset_i, \openset_j \subset \dtotal$
+    \item \textit{objects} open sets $\opensetc_{i} \in \{\opensetc\}$ in the topological space $(\dbasec, \mathscr{T})$), including the empty set $\varnothing$ and the maximum set \dbase. 
+    \item \textit{morphisms} $\iota: \opensetc_{i} \hookrightarrow \openset_{j}$ for all  $\opensetc_i, \opensetc_j \subset \dtotalc$
   \end{itemize} 
@@
 
-\begin{definition} The category of open sets $\mathcal{\gbasec}$ in the topological space $(\dbase, \mathscr{T})$ consists of open set \textit{objects} $\opensetg \in \{\opensetg\}$, including the empty set $\varnothing$ and the maximum set $\gtotal$ and \textit{morphisms} $\iota: \opensetg_i \rightarrow \opensetg_j$.
+\begin{definition} The category of open sets $\mathcal{\gbasec}$ in the topological space $(\gbasec, \mathscr{T})$ consists of open set \textit{objects} $\opensetgc \in \{\opensetgc\}$, including the empty set $\varnothing$ and the maximum set $\gtotalc$ and \textit{morphisms} $\iota: \opensetgc_i \rightarrow \opensetgc_j$.
 \end{definition}
 
-The data space \dbase\ is a deformation retract of the graphic topology \gbase\meaning that regions in \gbase\ collapse to points in \dbase\ such that the relative positions of the points are preserved \cite{nlab:deformation_retract}.
+The data space \dbase\ is defined via the surjective function $\vindex$ such that regions in \gbase\ collapse to points in \dbase\
+\begin{equation}
+  \label{eq:artist:vindex}
+  \vindexc: \gtotalc \textcolor{artist}{\rightarrow} \dtotalc
+\end{equation}
+Since \dbase\ is defined as a deformation retract of the graphic topology \gbase, the continuity of the points in data space and graphic space are equivalent \cite{nlab:deformation_retract}.
 One method of constructing the graphic space is by expanding the dimensionality of the data \dbase\ to match the spatial dimensionality of the rendering by multiplying the data space by  intervals $I$.
@@
 \subsubsection{Data}
-\note{clean up why sheaf}
-As mentioned in \autoref{sec:related-work:continuity:sheaves}, sheaves are a powerful tool for keeping track of data that lies over topological spaces.
-Therefore, we model the type of data an artist expects as a sheaf $\sheafc_{\dtotalc}: \mathcal{\dbasec}^{op} \textcolor{sheaf}{\rightarrow} \setc$.
+As mentioned in \autoref{sec:atct:sheaves} sheaves are a powerful tool for keeping track of data that lies over topological spaces; therefore, we model the type of data an artist expects as a sheaf on a fiber bundle with total space \dtotalc and basespace \dbase.
+
 \begin{equation}
@@
 
-To establish a mapping between graphic and data base spaces, we introduce the surjective map $\vindex: \gbase \rightarrow \dbase$. 
-\begin{equation}
-  \label{eq:artist:vindex}
-  \begin{tikzcd}
-    \dtotalc& \dtotalpullc \arrow[l, "\vindexpullc"', dashed, color=artist]                                                                           \\
-    \dbasec \arrow[u, "\dsectionc", color=section] & \gbasec \arrow[l, "\vindexc", color=artist] \arrow[lu, "\dsectionc \circ \vindexc" description, color=section] \arrow[u, "\dsectionpullc"', dashed, color=section]
-    \end{tikzcd}
-\end{equation}
-such that the data continuity $\dbase$ is a deformation retract of $\gbase$, as expressed in \autoref{eq:artist:gbase}.
-
-As shown in \autoref{eq:artist:vindex}, the map between spaces $\vindex$ can be composed with section map $\dsection$ such that there is a map from the graphic base space $\gbase$ to the fiber bundle in which the data lives $\dtotal$.
-\begin{figure}[h!]
-  \includegraphics[width=\columnwidth]{pullback.png}
-  \label{fig:artist:pushforward_xi}
-  \caption{The map $\vindexpull$ is a map from a repetition of the 1D fiber over every point $\dbasepoint \in \dbase$ in data space, turning the fiber into a 2D plane of repeating values over the corresponding region $\dbasepoint = \vindex(\gbasepoint)$ in the graphic space \gbase.
-}
-\end{figure}
-Since there is a mapping from the graphic space to the total bundle $\dsection \circ \vindex: \gbase \rightarrow \dtotal$, the mapping $\vindex$ can be pulled through the composition such that there is a map $\vindexpull$ between a total space over graphics $\dtotalpull$ and its equivalent space over data $\dtotal$.
-As illustrated in the example given in \autoref{fig:artist:pushforward_xi}, a bundle $\dtotal$ over a point $\dbasepoint \in \dbase$ is pulled back through \vindex\ \cite{nlab:pullback_bundle} such that it is copied over every point in the corresponding region in graphic space $\vindex(\gbasepoint) = \dbasepoint, \gbasepoint \subset \gbase$.
-The pullback $\vindexpull$ can then be composed with the elements of the set of sections
-\begin{equation}
-  \label{eq:artist:data_gamma_pullback}
-  \vindexpullc: \cgamma{\dbasec}{\dtotalc} \textcolor{artist}{\rightarrow} \cgamma{\gbasec}{\dtotalpullc}
-\end{equation}
-to generate a set of equivalent sections over the graphic space $\Gamma(\gbase, \dtotalpull)$.
-As a consequence of \autoref{eq:artist:data_gamma_pullback}, there exists a data sheaf over graphic space
-\begin{equation}
-  \label{eq:artist:data_pullback_sheaf_gamma}
-\sheafc_{\dtotalpullc}: \opensetgc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetgc}{\dtotalpullc}}
-\end{equation}
-which is equivalent to 
-\begin{equation}
-  \label{eq:artist:data_pullback_sheaf}
-  \sheafc_{\dtotalpullc}: \mathcal{\gbasec}^{\textcolor{base}{op}} \textcolor{sheaf}{\rightarrow} \setc
-\end{equation}
-
 \subsubsection{Graphic}
-We model the space of all graphic generating functions over a specific graphic topology $\opensetg \subseteq \gbase$ and targeting a specific display as a set of sections 
+We model graphics as a sheaf from an open set in the graphic base space to sets of sections of graphic generating functions
+\begin{equation}
+  \label{eq:artist:graphic_sheaf_gamma}
+  \sheafc_{\gbasec}: \opensetgc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetgc}{\gtotalc}
+\end{equation}
+and we define the set of sections such that 
 \begin{equation}
@@
 
-The set of sets of sections $\{\Gamma(\opensetg, \gtotal)\}\vert_{{\opensetg \in gbase}}$ is subcategory of \textbf{Set}; therefore the full space of graphics is a sheaf 
-\begin{equation}
-  \label{eq:artist:graphic_sheaf_gamma}
-  \sheafc_{\gtotalc}: \opensetgc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetgc}{\gtotalc}
-\end{equation}
-which means it is 
-\begin{equation}
-  \label{eq:artist:graphic_sheaf_set}
-  \sheafc_{\gtotalc}: \gbasec^{\textcolor{base}{op}} \textcolor{sheaf}{\rightarrow} \setc
-\end{equation}
-
 \subsubsection{Data to Graphic}
+Given the data sheaf $\sheaf_{\dbase}$, graphic sheaf $sheaf_{gbase}$, functor $\vindex$ between $\gbase$ and $\dbase$, we can define the artist as a morphism between sections of sheaves
+\begin{equation}
+  \begin{tikzcd}[row sep=huge]
+    \{\gsectionc: \opensetgc \textcolor{section}{\rightarrow} \gtotalc\restriction_{\opensetgc}\} \arrow[rr, "\vindexpushc", color=functor]                                                     &  & \{\gsectionc: \vindexprec(\opensetc) \textcolor{section}{\rightarrow} \gtotalc\restriction_{\vindexprec(\opensetc)}\}\\&  &\\
+    \{\dsectionc: \vindexc(\opensetgc) \textcolor{section}{\rightarrow} \dtotalc\restriction_{\vindexc(\opensetgc)}\} \arrow[uu, "\vartistc_{\vindexpullc}: \vindexpullc \dsectionc \textcolor{artist}{\rightarrow} \gsectionc", color=homset] \arrow[rr, "\vindexpullc", color=functor] &  & \{\dsectionc: \opensetc \textcolor{section}{\rightarrow} \dtotalc\restriction_{\opensetc}\} \arrow[uu, "\vartistc_{\vindexpushc}: \dsectionc \textcolor{artist}{\rightarrow} \vindexpushc \gsectionc"', color=homset] \arrow[lluu, "\vartistc" description, dashed, color=artist]
+    \end{tikzcd}
+\end{equation}
 Following from \autoref{eq:artist:data_pullback_sheaf} and \autoref{eq:artist:artist_sheaf},  we propose that the artist function $\vartist$ is a natural transformation of sheaf functors 
@@
 
-\note{}
-
 It follows from \autoref{eq:related-work:equivariance:functor_diagram} that an arbitrary morphism $\dfunc$ on the data sheaf has a compatible morphism $\dfunc^{\prime}$ on the graphic sheaf. 
@@
 We propose that we can express expected consistency between these elements as a preservation of morphisms between objects of different data categories $\mathcal{\dtotal}$.
+\note{We can define a monoid on the set of artists w/ + such that artists are composoble. doing so lets us}
 
@@
 
+\subsubsection{Pulled Back Data}
+
+We model the type of data we have as fiberbundles, building on the work of Butler et al.
+\cite{butlerVisualizationModelBased1989,butlerVectorBundleClassesForm1992}.
+Specifically we denote our data as the bundle \dtotal\ over the base space \dbase.
+We pull this bundle over the graphic space \gbase\, establishing an association between regions in the graphic and corresponding data.
+
+Given the map \vindex\ (\autoref{eq:artist:vindex}) between data and graphic basespaces, we can introduce a pullback function that copies sections of the data bundle over the data base space to the graphic base space.
+. 
+\begin{equation}
+  \label{eq:artist:vindex}
+  \begin{tikzcd}
+    \dtotalpullc \arrow[rr, "\vindexpullc", color=artist]                                                   &  & \dtotalc                          \\
+                                                                                                                                    &  &                                   \\
+    \gbasec \arrow[rr, "\vindexc", color=artist] \arrow[uu, "\dsectionpullc", dashed, color=section] \arrow[rruu, "\vindexc\circ\dsectionc" description, dashed, color=section] &  & \dbasec \arrow[uu, "\dsectionc"', color=section]
+    \end{tikzcd}
+\end{equation}
+
+As shown in \autoref{eq:artist:vindex}, the map between spaces $\vindex$ can be composed with section map $\dsection$ such that there is a map from the graphic base space $\gbase$ to the fiber bundle in which the data lives $\dtotal$.
+\begin{figure}[h!]
+  \includegraphics[width=\columnwidth]{pullback.png}
+  \label{fig:artist:pushforward_xi}
+  \caption{The map $\vindexpull$ is a map from a repetition of the 1D fiber over every point $\dbasepoint \in \dbase$ in data space, turning the fiber into a 2D plane of repeating values over the corresponding region $\dbasepoint = \vindex(\gbasepoint)$ in the graphic space \gbase.
+}
+\end{figure}
+Since there is a mapping from the graphic space to the total bundle $\dsection \circ \vindex: \gbase \rightarrow \dtotal$, the mapping $\vindex$ can be pulled through the composition such that there is a map $\vindexpull$ between a total space over graphics $\dtotalpull$ and its equivalent space over data $\dtotal$.
+As illustrated in the example given in \autoref{fig:artist:pushforward_xi}, a bundle $\dtotal$ over a point $\dbasepoint \in \dbase$ is pulled back through \vindex\ \cite{nlab:pullback_bundle} such that it is copied over every point in the corresponding region in graphic space $\vindex(\gbasepoint) = \dbasepoint, \gbasepoint \subset \gbase$.
+
 \subsection{Graphic Codomain} 

### 2021-12-23 slides/main.tex
not compiling on equation

\documentclass[xcolor={dvipsnames}, handout]{beamer}
%\documentclass[xcolor={dvipsnames}]{beamer}
%\usepackage{amsmath,amsfonts,amssymb,pxfonts,eulervm,xspace}
\usepackage{amsmath, amsfonts, amssymb, mathtools, eulervm, xspace}
\usepackage{bm}
\usepackage{mathrsfs} % math script fonts
\usepackage{tikz}
\usetikzlibrary{cd} % commutative diagrams
\newtheorem{prop}{Proposition} %math?

\usepackage{subcaption} % subfigure float captioning

\usepackage{tabulary}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{array}

\usepackage{graphicx}

\usepackage{appendixnumberbeamer} 
\usepackage{comment}
\usepackage{minted}
\setminted[python]{fontsize=\scriptsize, 
                   linenos,
                   numbersep=8pt,
                   autogobble, 
                   frame=lines,
                   bgcolor=bg,
                   framesep=3mm} 
\usepackage{notation} % move this later

\graphicspath{.figures/}

\usepackage[backend=bibtex, sorting=none, doi=false,isbn=false,url=false, giveninits=true]{biblatex}
\bibliographystyle{plain}
\bibliography{../paper/bibliography.bib}

\usetheme{ccnycrest}

\newenvironment{changemargin}[2]{%
\begin{list}{}{%
\setlength{\topsep}{0pt}%
\setlength{\leftmargin}{#1}%
\setlength{\rightmargin}{#2}%
\setlength{\listparindent}{\parindent}%
\setlength{\itemindent}{\parindent}%
\setleng{}th{\parsep}{\parskip}%
}%
\item[]}{\end{list}}

\begin{document}

\title{Topological Equivariant Artist Model}

\begin{frame}
	\titlepage
    Hannah Aizenman, Tom Caswell, Michael Grossberg\\
    %Committee: Dr. Robert Haralick, Dr. Lev Manovich, Dr. Huy Vo\\
    %External Member: Dr. Marcus Hanwell
\end{frame}

\section{Introduction}
\begin{frame}{What are we doing?}
    \begin{itemize}
        \item develop a model for describing data to graphic transformations
        \item specify a visualization library architecture based on this model
        \item implement functional(ish) components based on this model using ideas from functional programming
    \end{itemize}
\end{frame}

\begin{frame}{What do visualization libraries do?}
    \begin{figure}
        \begin{overprint}
            \onslide<1|handout:1>\includegraphics[width=\linewidth]{figures/flow/s2.png}
            \onslide<2|handout:2>\includegraphics[width=\linewidth]{figures/flow/s_vc.png}
            \onslide<3|handout:3>\includegraphics[width=\linewidth]{figures/flow/s_mark.png}
            \onslide<4|handout:4>\includegraphics[width=\linewidth]{figures/flow/s3.png}
        \end{overprint}
    \end{figure}
\end{frame}

\begin{frame}{Case Study: Matplotlib}
    \begin{figure}
       \includegraphics[width=\linewidth]{figures/flow/artists.png}
    \end{figure}
\end{frame}

\begin{frame}{Visualizations are structure preserving maps}
    \begin{description}
        \item [continuity] how elements in a dataset are organized, e.g. discrete rows in a table, networked nodes, pixels in an image, points on a line
        \item [equivariance] data transformations have an equivalent effect on the graphical representation, e.g. rotating a matrix has a matching rotation of the image, translating the points on a line has a matching visual shift in the line plot
      \end{description}
\end{frame}

\begin{frame}{Continuity}
    \begin{figure}
        \includegraphics[width=\linewidth]{../paper/figures/k_different_types.png}
    \end{figure}
\end{frame}

\begin{frame}{Expressed in Indexing}
    \begin{figure}
        \includegraphics[width=1\textwidth]{figures/intro/munzner_datatypes.png}
        \caption{Figure 2.8 in Munzner's Visualization Analysis and Design\cite{munznerVisualizationAnalysisDesign2014}}
    \end{figure}
\end{frame}

\begin{frame}{Why is indexing not enough?}
    \includegraphics[width=1\linewidth]{../paper/figures/whycontinuity.png}
\end{frame}

\begin{frame}{Fiber bundle as abstract data structure\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}}
\begin{columns}
    \column{0.49\columnwidth}
        \begin{equation}
            \label{eq:related-work:continuity:fiber-bundle}
            \begin{tikzcd}[ampersand replacement=\&, row sep=large]
                \dfiberc \arrow[r, hook] & \dtotalc \arrow[d, "\pi"'] \\
                            & \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
            \end{tikzcd}
        \end{equation}
        %\begin{equation}
        %    \label{eq:related-work:continuity:sections}
        %    \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\openset} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;\forall \dbasepointc \in \opensetc \big\}
        %\end{equation}
 \column{0.49\columnwidth}
\end{columns}  
\end{frame}

\begin{comment}
\begin{frame}{Equivariance}
    \begin{description}
        \item[Retinal Variables \& Marks] visual encodings match properties of the data \cite{bertinSemiologyGraphicsDiagrams2011a}
        \pause
        \item[Expressiveness] structure preserving mappings\cite{mackinlayAutomatingDesignGraphical1986}
        \pause
        \item[Graphical Integrity] graphs show \textbf{only} the data\cite{tufteVisualDisplayQuantitative2001}
        \pause 
        \item[Naturalness] easier to understand when properties match\cite{norman_things_smart}
    \end{description}
\end{frame}

\begin{frame}{Equivariance}

\end{frame}

\begin{frame}{Domain specific libraries assume data structure\cite{HeerSoftware2006}}
    \begin{table}
        %\renewcommand{\arraystretch}{2}
        %
        \begin{tabular}{>{\onslide<1->}l>{\onslide<2->}l>{\onslide<3->}l}
            \includegraphics[width=.24\textwidth]{figures/intro/table.png} & \includegraphics[width=.3\textwidth]{figures/intro/landsat.png} & \includegraphics[width=.33\textwidth]{figures/math/graph.png} \\
            ggplot\cite{wickhamGgplot2ElegantGraphics2016a}  & ImageJ\cite{schneiderNIHImageImageJ2012}& Gephi\cite{bastianGephiOpenSource2009}\\
            Vega\cite{satyanarayanDeclarativeInteractionDesign2014} & ImagePlot\cite{studiesCulturevisImageplot2021} & Graphviz\cite{ellsonGraphvizOpenSource2002}\\
            Altair\cite{vanderplasAltairInteractiveStatistical2018}& Napari\cite{nicholas_sofroniew_2021_4533308} & Networkx\cite{HagbergExploringNetwork2008}\\
             Tableau \cite{StoltePolaris2002}& &\\
            \cite{hanrahanVizQL2006,MackinlayShowme2007}&&\\

        \end{tabular}
    \end{table}
\end{frame}

\begin{frame}{General purpose libraries can't\cite{toryRethinkingVisualizationHighlevel2004}}
    \begin{figure}
        \includegraphics[height=.5\textheight]{figures/intro/dataset_diagram.png}
        \caption{Data Representation, MayaVi 4.7.2 docs\cite{DataRepresentationMayavi}}
    \end{figure}
    \begin{enumerate}
        \item Matplotlib\cite{hunterMatplotlib2DGraphics2007} $\rightarrow$Seaborn\cite{waskom2020seaborn}, xarray \cite{hoyer2017xarray} 
        \item D3 \cite{bostockDataDrivenDocuments2011}
        \item VTK \cite{hanwellVisualizationToolkitVTK2015,geveciVTK2012}(MayaVi\cite{RamachandranMayaVI2011})$\rightarrow$ Titan\cite{brianwylieUnifiedToolkitInformation2009}, ParaView\cite{ahrens2005paraview}
    \end{enumerate}
\end{frame}

\subsection{Contributions}
\begin{frame}{Contributions}
    \begin{description}
        \item[Topological] continuity  
        \item[Equivariant] monoid action 
        \item[Artist] Matplotlib $\textit{artist}: \textit{data} \rightarrow \textit{graphic}$ 
        \item [Model] 
    \end{description}
\end{frame}

\section{Math}

\begin{frame}{Topological Equivariant Artist Model}
    \centering
    An artist $\mathscr{\vartist}$ is an equivariant map 
    \begingroup 
    \Huge
    \begin{equation*}
        \mathscr{\vartist}: \mathscr{\dtotal} \rightarrow \mathscr{\gtotal}
    \end{equation*}
    \endgroup
    from data $\mathscr{\dtotal}$ space to graphic $\mathscr{\gtotal}$ space. 
\end{frame}

\subsection{Data}
\begin{frame}{Model data as a fiber bundle \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}}
    A fiber bundle is a tuple $(\dtotal,\,\dbase,\,\pi ,\,\dfiber)$ defined by the map $\pi$
    \begin{equation*}
        \label{eq:fiber_bundle}
        \begin{tikzcd}[ampersand replacement=\&]
            \dfiber \arrow[r, hook] \& \dtotal \arrow[r, "\pi"] \& \dbase
        \end{tikzcd}
    \end{equation*}
    \begin{description}
        \pause
        \item[total space \dtotal] topology
        \pause
        \item[fiber space \dfiber] schema
        \pause
        \item[base space \dbase] continuity
    \end{description}
\end{frame}

\begin{frame}{Encode variable types in a schema like fiber \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES}}
    \begin{figure}[H]
        \centering
        \includegraphics[width=\textwidth]{figures/math/fiber.png}
    \label{fig:data_fiber_example}
    \end{figure}
\end{frame}

\begin{frame}{Monoids are the structure of the components of \dfiber}
    \begin{figure}[H]
        \centering
        \includegraphics[height=.3\textheight]{figures/math/fiber.png}
    \label{fig:data_fiber_example}
    \end{figure}
        \begin{equation*}
            \dfiber = \dfiber_0 \times \ldots \times \dfiber_i \times \ldots \times \dfiber_n
        \end{equation*}
        \pause
        Monoid actions $\monoid_{i}$ (e.g. rotation, partial ordering) define the structure on $\dfiber_i$
        \begin{equation*}
            \bullet: \monoid_i\times \dfiber_i \rightarrow \dfiber_i
        \end{equation*}
        where $\bullet$ is associative and has an identity action. 
\end{frame}

\begin{frame}{\dbase\ is an indexing space}
    \begin{figure}[H]
        \includegraphics[width=1\textwidth]{figures/math/base.png}
    \label{fig:base_example}
    \end{figure}
\end{frame}

\begin{frame}{Data are sections \dsection\ on \dtotal}
    For any fiber bundle, there exists a map
    \begin{equation*}
        \begin{tikzcd}[ampersand replacement=\&]
            \dfiber \arrow[r, hook] \& \dtotal \arrow[d, "\pi"'] \\
                              \& \dbase \arrow[u, "\dsection"', bend right]
        \end{tikzcd}
    \end{equation*}
     s.t. $\pi(\dsection(\dbasepoint)) = \dbasepoint$. $\Gamma(\dtotal)$ is the set of all global sections.
     \begin{figure}[H]
        \includegraphics[width=1\linewidth]{figures/math/fiberbundle.png}
        \label{fig:data_sections}
    \end{figure}
\end{frame}

\subsection{Graphic}
\begin{frame}{Graphic Bundle $(\gtotal,\,\gbase,\,\pi ,\,\gfiber)$} % watch dr explaination
    \only<2->{Continuity is preserved via the many \gbasepoint\ to one \dbasepoint\ map $\vindex: \gbase\rightarrow\dbase$}

    \only<1>{
        \begin{equation*} 
            \begin{tikzcd}[ampersand replacement=\&]
                F \arrow[r, hook] \& E \arrow[d, "\pi"']              \& D \arrow[r, hook] \& H \arrow[d, "\pi"']              \\
                                  \& K \arrow[u, "\tau"', bend right] \&                   \& S \arrow[u, "\rho"', bend right]
                \end{tikzcd}
            \end{equation*}}
    \only<2->{
    \begin{equation*} 
        \begin{tikzcd}[ampersand replacement=\&]
            F \arrow[r, hook] \& E \arrow[d, "\pi"']              \& D \arrow[r, hook] \& H \arrow[d, "\pi"']                                 \\
                              \& K \arrow[u, "\tau"', bend right] \&                   \& S \arrow[ll, "\xi"'] \arrow[u, "\rho"', bend right]
            \end{tikzcd}
        \end{equation*}}
    \only<3->{   
    \gfiber\ is a proxy for the target display, for example $(x,\, y,\, r,\, g,\, b) \in \gfiber$}
    \pause
    \begin{figure}[H] %%%bring up one at a time
        \begin{overprint}
            \onslide<4|handout:0>\includegraphics[width=1\linewidth]{figures/math/scatter_s.png}
            \onslide<5|handout:0>\includegraphics[width=1\linewidth]{figures/math/line_s.png}
            \onslide<6>\includegraphics[width=1\linewidth]{figures/math/image_s.png}
        \end{overprint}
    \end{figure}
\end{frame}

\subsection{Artist}
\begin{frame}{Visual bundle $(\vtotal,\,\dbase,\,\pi ,\,\vfiber)$}
    \begin{equation*}
        \mathcal{A}:\mathcal{E}\rightarrow\mathcal{H}
    \end{equation*}
\pause
\begin{equation*}
    \label{eq:artist_diagram}
    \begin{tikzcd}[ampersand replacement=\&]
        \dtotal^{\prime} \arrow[r, "\vchannel"] \arrow[rd, "\pi"'] \& \vtotal \arrow[d, "\pi"] \& \vindex^*\vtotal \arrow[r, "\vmark"] \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"'] \& \gtotal \arrow[ld, "\pi"] \\
                                              \& \dbase                  \& \gbase \arrow[l, "\vindex"']                                              \&                    
        \end{tikzcd}
\end{equation*}
\pause
\begin{equation*}
    \vartist: \mathcal{O}(\dtotal) \rightarrow \mathcal{O}(\gtotal)
\end{equation*}
\end{frame}

\begin{frame}{Visualization Assembly Function}
    \begin{figure}
        \includegraphics[width=\textwidth]{figures/math/path_of_q.png}
    \end{figure}
    \pause
    \begin{equation*}
        \label{eq:nu_expanded}
        \{\vchannel_{0}, \ldots, \vchannel_{n}\}: \{\dsection_{0}, \ldots, \dsection_{n}\} \mapsto \{\vsection_{0}, \ldots, \vsection_{n}\}
    \end{equation*}
    \pause
    \begin{equation*}
        Q = \nu \circ \tau
    \end{equation*}
\end{frame}

\begin{frame}{Group Equivariance: Stevens' Scales \cite{stevensTheoryScalesMeasurement1946}}
\begin{table}[H]
    \begin{tabularx}{\textwidth}{|l|l|X|}\toprule
        \textbf{scale} & \textbf{group} & \textbf{constraint} \\\midrule
        nominal & permutation &  $\text{if } \delement_1 \neq \delement_2 \text{ then } \vchannel (\delement_1) \neq\vchannel(\delement_2)$\\
        ordinal &  monotonic & $\text{if } \delement_1 \leq \delement_2 \text{ then } \vchannel (\delement_1) \leq \vchannel(\delement_2)$\\
        interval &  translation &  $\vchannel (x + c) = \vchannel(x) + c$ \\
        ratio &  scaling &  $\vchannel(xc) = \vchannel(x)*c $\\ \bottomrule
    \end{tabularx}
\end{table}
\end{frame}

\begin{frame}{Monoid Equivariance: Partial Orders}
    \begin{figure} 
        \begin{overprint}
            \onslide<1|handout:0>\includegraphics[width=1\linewidth]{figures/math/partial.png}
            \onslide<2|handout:0>\includegraphics[width=1\linewidth]{figures/math/partial_invalid.png}
            \onslide<3>\includegraphics[width=1\linewidth]{figures/math/partial_fixed.png}
        \end{overprint}
    \end{figure}
\end{frame}

\begin{frame}{Visualization Equivariance}
    \begin{figure}[H]
        \includegraphics[width=\textwidth]{figures/math/diff_type_q.png}
    \end{figure}
    \begin{equation*}
        \vmark(\vsection) = \vmark(\vsection^{\prime})\implies \vmark(m\circ\vsection) = \vmark(m\circ\vsection^{\prime})
    \end{equation*}

\end{frame}

\begin{frame}{Scatter: $\vmark(xpos, ypos)(\alpha, \beta)$}
    \begin{figure}[H]
        \begin{overprint}
            \onslide<1|handout:0>\includegraphics[width=1\linewidth]{figures/math/scatter_q.png}
            \onslide<2|handout:0>\includegraphics[width=1\linewidth]{figures/math/scatter_with_s.png}
        \end{overprint}
    \end{figure}   
\end{frame}

\begin{frame}{Line: $\vmark(xpos, \hat{n_{1}}, ypos, \hat{n_{2}})(\alpha, \beta)$ }
    \begin{figure}[H]
        \begin{overprint}
            \onslide<1|handout:0>\includegraphics[width=1\linewidth]{figures/math/line_q.png}
            \onslide<2|handout:0>\includegraphics[width=1\linewidth]{figures/math/line_with_s.png}
        \end{overprint}
    \end{figure}   
\end{frame}

\begin{frame}{Image $\vmark(xpos, ypos, color)$}
    \begin{figure}[H]
        \includegraphics[width=1\textwidth]{figures/math/image.png}
    \end{figure}
\end{frame}

\begin{frame}{Build \vmark\ over \dbase: \vmarkd}
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/q_hat.png}
\end{figure}
\begin{equation*}\label{eq:qhat_q_s}
    \vmarkd(\vsection(\dbasepoint))(\gbasepoint) \coloneqq \vmark((\vsectionpull)(\gbasepoint))
\end{equation*} 
\end{frame}

\begin{frame}{Composition of artists $+ \coloneqq \sqcup \dtotal_{i}$} 
    \begin{figure}
        \includegraphics[width=1\textwidth]{figures/math/exploding_artist.png}
    \end{figure}
\end{frame}

\section{Code}
\begin{frame}{TEAM driven rearchitecture of Matplotlib}
    \begin{itemize}
        \item complex visualizations
        \pause 
        \item structure preserving maps from data to visual
        \begin{itemize}
            \item data and graphics have equivalent continuity
            \item properties are equivariant under monoid actions
        \end{itemize}  
        \pause
        \item fiber bundles are an abstraction
        \begin{itemize}
            \item topologically complex heterogenous data 
            \item target display spaces
        \end{itemize}
    \end{itemize}
\end{frame}

\begin{frame}[fragile]{How do we make things?}
    \begin{figure}[H]
        \begin{subfigure}{0.49\textwidth}
            \includegraphics[width=\textwidth]{figures/code/scatter_0.png}
        \end{subfigure}
        \begin{subfigure}{0.49\textwidth}
            \includegraphics[width=\textwidth]{figures/code/line_1.png}
        \end{subfigure}
    \end{figure}
    \begin{columns}
    \column{0.49\textwidth}
    \begin{minted}{python}
    fig, ax = plt.subplots()
    artist = Point(data, transforms)
    ax.add_artist(artist)
    \end{minted}
    \column{.49\textwidth}
    \begin{minted}{python}
    fig, ax = plt.subplots()
    artist = Line(data, transforms)
    ax.add_artist(artist)
    \end{minted}
    \end{columns}
\end{frame}

\subsection{Artist}
\begin{frame}[fragile]{ \vchannel}
    \begin{minted}{python}
    cmap =  color.Categorical({'true':'deeppink', 'false':'deepskyblue'})
    transforms = {'x': {'name': 'v4', 'encoder': lambda x: x},
                    'y': {'name': 'v2', 'encoder': lambda x: x},
                    'facecolors': {'name':'v3', 'encoder': cmap}, 
                    's':{'name': None , 
                        'encoder': lambda _: itertools.repeat(.02)}}
    \end{minted}
    \begin{itemize}
        \item \mintinline{python}{lambda x: x} is identity \vchannel\
        \item \mintinline{python}|{'name':None}| map into \vfiber\ without corresponding \dsection\
        \item \mintinline{python}{color.Categorical} is custom \vchannel
    \end{itemize}
\end{frame}

\begin{frame}[fragile]{\vartist} %% rewrite w/ letters in talk
    \begin{minted}{python}
        class ArtistClass(matplotlib.artist.Artist):
            def __init__(self, E, V, *args, **kwargs):
                # set properties that are specific to the artist
                # stash the input E and V
                super().__init__(*args, **kwargs)

            def qhat(self, **args):
                # set the properties of the graphic

            def draw(self, renderer):
                # returns tau, indexed on fiber then key 
                tau = self.E.view(self.axes) 
                # visual channel encoding applied fiberwise 
                visual = {p_i: nu_i(tau_i)
                          for p_i, nu_i, tau)i 
                          in zip(self.V, tau)}
                self.qhat(**visual)
                # pass configurations off to the renderer
                super().draw(renderer)
        \end{minted}
\end{frame}

\begin{frame}[fragile]{\vmarkd}
\begin{minted}{python}
    class Point(mcollections.Collection):
        def assemble(self, x, y, s, facecolors='C0' ):
            # construct geometries of the circle glyphs in visual coordinates
            # set attributes of glyphs

    class Line(mcollections.LineCollection):
        def assemble(self, x, y, color='C0'):
            # assemble line marks as set of segments 
    \end{minted}
\end{frame}

\subsection{Data}
\begin{frame}[fragile]{Continuity}
    \begin{minted}{python}
class PointData: 
    # Fiberbundle is consistent across all sections
    FB = FiberBundle({'tables': ['vertex']},  
        {'v1': float, 'v2': str, 'v3': float})
    def tau(self, k):
        return # tau evaluated at one point k

class LineData: 
    FB = FiberBundle({'tables': ['edge']},
                {'x' : float, 'y':  float, 'color':mtypes.Color()})
    def tau(self, k): 
        return # tau evaluated on interval k
    \end{minted}
\end{frame}

\begin{frame}[fragile]{Same Artist, Different \dtotal}
    \begin{figure}[H]
        \begin{subfigure}{0.49\textwidth}
            \includegraphics[width=\textwidth]{figures/code/linec_1.png}
        \end{subfigure}
        \begin{subfigure}{0.49\textwidth}
            \includegraphics[width=\textwidth]{figures/code/lined_1.png}
        \end{subfigure}
    \end{figure}
\begin{minted}{python}
LineData(FB, edge_table, vertex_table, connect=True)
LineData(FB, edge_table, vertex_table, num_samples=2, connect=False)
\end{minted}
\end{frame}

\begin{frame}{Proposed Work}
\begin{figure}
    \includegraphics[width=1\textwidth]{figures/intro/gantt.png}
\end{figure}
\end{frame}

\begin{frame}{Acknowledgments}
    \begin{itemize}
        \item Professor Michael Grossberg and Dr. Thomas Caswell
        \item Professor Haralick, Professor Vo, Professor Manovich, Dr. Hanwell
        \item Matplotlib development team 
        \item CZI EOSS (grant number 2019-207333) from the Chan Zuckerberg Initiative DAF, an advised fund of Silicon Valley Community Foundation
    \end{itemize}

\end{frame}

\begin{frame}[allowframebreaks]{References}
\printbibliography
\end{frame}
\appendix

\begin{frame}{Mathematical Models of Vizualization}
    \begin{description}
        \item[algebraic process] data and viz transforms are symmetric \cite{kindlmannAlgebraicProcessVisualization2014}
        \begin{equation*}
            \begin{tikzcd}[ampersand replacement=\&]
                \dtotal \arrow[d, "m"'] \arrow[r, "\nu"] \& \vtotal \arrow[r, "Q"]  \& \gtotal \arrow[d, "\varphi(m)"] \\
                \dtotal \arrow[r, "\nu"']                     \& \vtotal \arrow[r, "Q"'] \& \gtotal                  
            \end{tikzcd}
        \end{equation*}

        \item[language] APT and GoG: syntax, semantics, and grammar\\ 
        \cite{mackinlayAutomatingDesignGraphical1986,mackinlayAUTOMATICDESIGNGRAPHICAL1987,wilkinsonGrammarGraphics2005}
        \item[functional dependencies] relationship between components \cite{sugibuchiFramwork2009}
        \item[category theory]  \textit{understanding} = \textit{read} $\circ$ \textit{render} \cite{vickersUnderstandingVisualizationFormal2013}    
    \end{description}
\end{frame}

\begin{frame}{Fiber is all possible values a variable can be \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES}}
    Given a space of all possible values \ftotal\
    \begin{equation*}
        \label{eq:data_types}
        \begin{tikzcd}[ampersand replacement=\&]
            \fttype \arrow[r] \arrow[d, "\pi_{\fsection}"'] \& \ftotal \arrow[d, "\pi"] \\
            \fnames \arrow[r, "\fsection"']                  \& \ftypes       
        \end{tikzcd}
    \end{equation*}
    a fiber component is the restricted space $\ftotal_{\fsection(\fname)}$. 
    \begin{equation*}
        \dfiber = \ftotal_{\fsection(\fname)} = \ftotal_{\ftype} 
    \end{equation*}
    \begin{description}
        \item[\ftypes] data types of the variables in the dataset 
        \item[\ftotal] disjoint union of all values of type $\ftype \in \ftypes$ 
        \item[\fnames] variable names, $\fname \in \fnames$
        \item[\fttype] \ftotal\ restricted to the data type of a named variable   
    \end{description}
\end{frame}

\begin{frame}{Monoid actions}
    A monoid \monoid\ is a set with
    \begin{description}
        \item[associative binary operator] $\ast:\monoid \times \monoid\rightarrow \monoid$
        \item[identity element] $e\in \monoid$ such that $e\ast a= a \ast e = a$ for all $a \in \monoid$. 
    \end{description}
    \begin{block}{left monoid action}
    A set \dfiber\ with an action\cite{nlab:action} $\bullet: \monoid\times \dfiber \rightarrow \dfiber$ with the properties:
        \begin{align*}
            \textbf{associativity}\;& \text{for all } f,g \in \monoid \text{ and } x\in \dfiber,\, f\bullet(g\bullet x) = (f\ast g) \bullet x\\
            \textbf{identity}\;& \text{for all } x\in \dfiber, e\in \monoid,\,  e\bullet x = x 
        \end{align*}
    \end{block}
\end{frame}

\begin{frame}{Keeping track of sections with sheafs}
    Restriction maps of a sheaf describe how local $\iota^*\dsection$ can be glued into larger sections \cite{ghristElementaryAppliedTopology2014,ghristHomologicalAlgebraData2018}
    \begin{equation*}
        \label{eq:sheaf}
        \begin{tikzcd}[ampersand replacement=\&]
            \iota^*\dtotal \arrow[d, "\pi"'] \arrow[r, "\iota^*", hook]             \& \dtotal \arrow[d, "\pi"']                  \\
            U \arrow[r, "\iota", hook] \arrow[u, "\iota^{*}\dsection"', bend right] \& \dbase \arrow[u, "\dsection"', bend right]
        \end{tikzcd}
    \end{equation*}
    The inclusion map $\iota: U \rightarrow \dbase$ pulls \dtotal\ over $U$ such that the pulled back $\iota^*\dsection$ only contains records over $U \subset \dbase$.
\end{frame}

\begin{frame}{Rendering: Define a Pixel}
    \begin{columns}
        \column{0.5\textwidth}
        Given a pixel
        \begin{equation*}
        p=\left[y_{top}, y_{bottom}, x_{right}, x_{left}\right]
        \end{equation*}
        the inverse map of the bounding box 
        \begin{equation*}
        \gbase_{p} ={\gsection_{xy}}^{-1}(p)
        \end{equation*}
        is a region $\gbase_p \subset \gbase$ such that 
        \begin{align}
            \scriptstyle r_p &= \scriptstyle \iint\limits_{S_p} \rho_r(s)ds^{2}\\
            \scriptstyle g_p &= \scriptstyle \iint\limits_{S_p} \rho_g(s)ds^{2}\\
            \scriptstyle  b_p &= \scriptstyle \iint\limits_{S_p} \rho_b(s)ds^{2}
        \end{align}
        yields the color of the pixel. 
        \column{0.5\textwidth}
        \begin{figure}[H]
            \includegraphics[width=\textwidth]{figures/math/render.png}
        \end{figure}
    \end{columns}
\end{frame}

\begin{frame}{$\mathcal{A}:\mathcal{E}\rightarrow\mathcal{E}$}
    The topological artist is a sheaf map
    \begin{equation*}
    \vartist: \mathcal{O}(\dtotal) \rightarrow \mathcal{O}(\gtotal)
    \end{equation*}
    that carries homomorphism of monoid actions $\varphi: \monoid \rightarrow \monoid^{\prime}$ 
    \cite{cegarraCohomologyMonoidsOperators2019} 
    \begin{equation*}
    \vartist(m\cdot \delement) = \varphi(m)\cdot \vartist(\delement) 
    \end{equation*}
\end{frame}

\begin{frame}{Visual Channel Encoders}
    We define the visual transformers \vchannel\ on components of the data bundle $\dsection_{i}$
    \begin{equation*}
        \label{eq:nu_expanded}
        \{\vchannel_{0}, \ldots, \vchannel_{n}\}: \{\dsection_{0}, \ldots, \dsection_{n}\} \mapsto \{\vsection_{0}, \ldots, \vsection_{n}\}
    \end{equation*}
    as the set of equivariant maps with the constraint 
    \begin{equation*}
        \vchannel_i(m_{\delement}(\dtotal_i)) = \varphi(m_{\delement})(\vchannel_i(\dtotal_i))
    \end{equation*} 
    where $\varphi:\monoid\rightarrow \monoid^{\prime}$ carries a homomorphism of monoid actions. 
\end{frame}

\begin{frame}{\vfiber\ Components}    
\begin{table}[H]
    \renewcommand{\arraystretch}{2}
    \begin{tabulary}{\textwidth}{|l|L|l|}\hline
     $\bm{\vchannel_{i}}$                      & $\bm{\vsection_{i}}$                                                            & $\bm{codomain(\vchannel_{i}) \subset \vfiber_{i}}$  \\ \hline                                              
    position                    & x, y, z, theta, r                                                          & $\mathbb{R}$   \\ \hline
    size                        & linewidth, markersize                                            & $\mathbb{R}^{+}$   \\ \hline
    shape                       & markerstyle                                                      & $\{f_{0}, \ldots, f_{n}\}$ \\ \hline
    color                       & color, facecolor, markerfacecolor, edgecolor  & $\mathbb{R}^{4}$ \\ \hline
    \multirow{2}{*}{texture}    & hatch                                                            & $\mathbb{N}^{10}$\\\cline{2-3}
                                & linestyle                                                        & $(\mathbb{R}, \mathbb{R^+}^{n, n\%2=0})$ \\ \hline              
    \end{tabulary}
    \label{tab:mpl_visual_variable_fiber}
\end{table}
\end{frame}
\begin{frame}{Monoid Equivariance: Partial Orders}
    \begin{figure} %break out steps/reveal
        \includegraphics[width=1\textwidth]{figures/math/monoid_equivariant.png}
    \end{figure}
\end{frame}

\begin{frame}{Glyph}
    The glyph is the graphic generated by $\vmark(\gbase_{\dbasepathpoint})$ where the path connected components $\dbasepath \subset \dbase$ are defined 
    \begin{equation*}
    \dbasepath = \{\dbasepathpoint \in \dbase \text{ s. t. } \exists \gamma \text{ s.t. } \gamma(0)=\dbasepoint \text{ and }\gamma(1)=\dbasepathpoint\}
    \end{equation*}
    such that the  path $\gamma$ from \dbasepoint\ to \dbasepathpoint\ is a continuous function from the interval [0,1] and $\gbase_{\dbasepathpoint}$ is the region
    \begin{equation*}
        \begin{tikzcd}[ampersand replacement=\&]
            \gtotal \arrow[r, shift left] \& \gbase_\dbasepathpoint \arrow[rr, "\vindex(\gbasepoint)", shift left] \arrow[l, "\gsection(\gbase_\dbasepathpoint)"] \& \& \dbasepath_{\dbasepoint} \arrow[ll, "\vindex^{-1}(\dbasepath)"]
            \end{tikzcd}
        \label{eq:mark}
    \end{equation*}
    such that the glyph is differentiable, in keeping with Ziemkiewicz and Kosara's description of a glyph\cite{ziemkiewiczEmbeddingInformationVisualization2009}.
\end{frame}

\begin{frame}{Artist Equivalance class}
    When artists share a base space 
    \begin{equation*}
        \dbase_2 \hookrightarrow \dbase_1
    \end{equation*}
    a composition operator can be defined such that the the artists can be considered to be acting on different components of the same section. 
\end{frame}

\begin{frame}[fragile]{Complex \vchannel}
    \begin{minted}{python}
    class Categorical:
        def __init__(self, mapping):
            # check that the conversion is to valid colors
            assert(mcolors.is_color_like(color) for color in mapping.values())
            self._mapping = mapping

        def __call__(self, value):
            # convert value to a color
            return [mcolors.to_rgba(self._mapping[v]) for v in values]
    \end{minted}
    That we can test for action equivariance
    \begin{minted}{python}
    def test_nominal(values, encoder):
        m1 = list(zip(values, encoder(values)))
        random.shuffle(values)
        m2 = list(zip(values, encoder(values)))
        assert sorted(m1) == sorted(m2)
    \end{minted}
\end{frame}

\begin{frame}[fragile]{Artist}
    \begin{minted}{python}
        class ArtistClass(matplotlib.artist.Artist):
            def __init__(self, data, transforms, *args, **kwargs):
                # properties that are specific to the graphic
                self.data = data 
                self.transforms = transforms
                super().__init__(*args, **kwargs)

            def assemble(self, **args):
                # set the properties of the graphic

            def draw(self, renderer):
                # returns K, indexed on fiber then key 
                view = self.data.view(self.axes) 
                # visual channel encoding applied fiberwise 
                visual = {p: t['encoder'](view[t['name']])
                          for p, t in self.transforms.items()}
                self.assemble(**visual)
                # pass configurations off to the renderer
                super().draw(renderer)
        \end{minted}
\end{frame}

\begin{frame}[fragile]{Artists: Scatter \& Line}
\begin{minted}{python}
class Point(mcollections.Collection):
    def assemble(self, x, y, s, facecolors='C0' ):
        # construct geometries of the circle glyphs in visual coordinates
        self._paths = [mpath.Path.circle(center=(xi,yi), radius=si) 
                    for (xi, yi, si) in zip(x, y, s)] 
        # set attributes of glyphs, these are vectorized 
        # circles and facecolors are lists of the same size
        self.set_facecolors(facecolors)

class Line(mcollections.LineCollection):
    def assemble(self, x, y, color='C0'):
        #assemble line marks as set of segments 
        segments = [np.vstack((vx, vy)).T for vx, vy in zip(x, y)]
        self.set_segments(segments)
        self.set_color(color)
\end{minted}
\end{frame}

\begin{frame}[fragile]{View}
\begin{minted}{python}
def view(self, axes):
    table = defaultdict(list)
    for k in self.keys:
        table['index'].append(k)
        for (name, value) in zip(self.FB.fiber.keys(), self.tau(k)[1]):
            table[name].append(value)
    return table
\end{minted}
\begin{description}
    \item [\mintinline{python}{VertexSimplex}] (name, value), value is scaler
    \item [\mintinline{python}{EdgeSimplex}]  (name, value), value is [x0, ..., xn]
\end{description}
\end{frame}

\begin{frame}[fragile]{Fiber Bundle}
    \begin{minted}{python}
    @dataclass
    class FiberBundle:
    """
    Attributes
    ----------
    K: {'tables': []}
    F: {variable name: type}
    """
        K: dict 
        F: dict
    \end{minted}
\end{frame}

\begin{frame}[fragile]{GraphLine Data Model}
\begin{minted}{python}
class GraphLine:
    def __init__(self, FB, edge_table, vertex_table, num_samples=1000,
                        connect=False):
        #set args as attributes and generate distance
        if connect: # test connectivity if edges are continuous
            assert edge_table.keys() == self.FB.F.keys()
            assert is_continuous(vertex_table)

    def tau(self, k):
        # evaluates functions defined in edge table
        return(k, (self.edges[c][k](self.distances) 
                        for c in self.FB.F.keys()))

    def view(self, axes):
        # walk the edge_vertex table to return the edge function
        table = defaultdict(list)
        for (i, (start, end)) in sorted(zip(self.ids, self.vertices), 
                                            key=lambda v:v[1][0]):
            table['index'].append(i)
            # same as view for line, returns nested list
            for (name, value) in zip(self.FB.F.keys(), self.tau(i)[1]):
                table[name].append(value)
        return table
\end{minted}
\end{frame}
\end{comment}
\end{document}

### 2021-12-27 slides/main.tex
clean up intro, new slide summing math

@@
 \usepackage{amsmath, amsfonts, amssymb, mathtools, eulervm, xspace}
+\renewcommand{\restriction}{\mathord{\upharpoonright}} %restriction w/p space
 \usepackage{bm}
@@
 
-\begin{frame}{Visualizations are structure preserving maps}
+\begin{frame}{Visualizations Preserve Structure}
     \begin{description}
-        \item [continuity] how elements in a dataset are organized, e.g. discrete rows in a table, networked nodes, pixels in an image, points on a line
-        \item [equivariance] data transformations have an equivalent effect on the graphical representation, e.g.
-rotating a matrix has a matching rotation of the image, translating the points on a line has a matching visual shift in the line plot
+        \item [continuity] topological properties \cite{wilkinsonGrammarGraphics2005}, i.e.
+how elements in a dataset are organized, e.g.
+discrete rows in a table, networked nodes, pixels in an image, points on a line
+        \item [equivariance] data and visual encodings are matched such that transformations have an equivalent effect on data and graphical representations, e.g.
+rotating a matrix and image, shifting  points on a line and a line graphic
       \end{description}
@@
 
-\begin{frame}{Expressed in Indexing}
+\begin{frame}{Expressed in container}    
     \begin{figure}
@@
 
-\begin{frame}{Why is indexing not enough?}
+\begin{frame}{Why is container type not enough?}
     \includegraphics[width=1\linewidth]{../paper/figures/whycontinuity.png}
@@
 
-\begin{frame}{Fiber bundle as abstract data structure\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}}
-\begin{columns}
-    \column{0.49\columnwidth}
-        \begin{equation}
-            \label{eq:related-work:continuity:fiber-bundle}
-            \begin{tikzcd}[ampersand replacement=\&, row sep=large]
-                \dfiberc \arrow[r, hook] & \dtotalc \arrow[d, "\pi"'] \\
-                            & \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
-            \end{tikzcd}
-        \end{equation}
-        %\begin{equation}
-        %    \label{eq:related-work:continuity:sections}
-        %    \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\openset} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;\forall \dbasepointc \in \opensetc \big\}
-        %\end{equation}
- \column{0.49\columnwidth}
-\end{columns}  
-\end{frame}
-
-\begin{comment}
 \begin{frame}{Equivariance}
@@
 
+\subsection{Algebraic Topology & Category Theory}
+
+\begin{frame}{Mathematical Data Abstraction}
+    \begin{description}
+        \item[Fiber Bundles] "unified, dimension-independent framework" that expresses data as the mapping between continuity and fields \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}
+        \item[Simplicial Databases] systemic way to apply a schema to a fiber, i.e. provides a way to name fields and bind them to types system (e.g. \texttt{int}, \texttt{float}) 
+         \cite{spivakSIMPLICIALDATABASES, spivakDatabasesAreCategories2010} 
+        \item[Sheaves on Bundles] "algebraic data structure" for representing data over topological spaces \cite{ghristElementaryAppliedTopology2014}
+    \end{description}
+\end{frame}
+
+\begin{frame}{Fiber Bundles}
+    \begin{columns}
+        \column{0.49\columnwidth}
+            \begin{equation*}
+                %\label{eq:related-work:continuity:fiber-bundle}
+                \begin{tikzcd}[ampersand replacement=\&]
+                  \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
+                   \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
+                \end{tikzcd}
+            \end{equation*}
+     \column{0.49\columnwidth}
+        \begin{description}
+            \item[\textcolor{base}{Base Space}] $( \dbasec, \mathcal{T})$
+            \item[\textcolor{fiber}{Fiber Space}] $\dfiberc_{\dbasepointc} = \pi^{-1}(\dbasepointc)$
+            \item[\textcolor{total}{Total Space}]  $\dtotalc = \dbasec \times \dfiberc$
+        \end{description}
+    \end{columns}
+    \pause
+    \begin{block}{Sections of the bundle}
+        \begin{equation*}
+            \label{eq:related-work:continuity:sections}
+            \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\openset} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;\forall \dbasepointc \in \opensetc \subset \dbasec \big\}
+        \end{equation*}
+    \end{block}
+\end{frame}
 \subsection{Contributions}
@@
 \end{frame}
-\end{comment}
+%\end{comment}
 \end{document}

### 2021-12-27 slides/main.tex
most of math section:

@@
 \usepackage{graphicx}
+\usepackage{adjustbox}%scale tikz-cd
 
@@
     \begin{description}
-        \item[Retinal Variables \& Marks] visual encodings match properties of the data \cite{bertinSemiologyGraphicsDiagrams2011a}
-        \pause
-        \item[Expressiveness] structure preserving mappings\cite{mackinlayAutomatingDesignGraphical1986}
+        \item[Retinal Variables \& Marks] visual encodings should match properties of the data \cite{bertinSemiologyGraphicsDiagrams2011a}
         \pause
@@
         \item[Naturalness] easier to understand when properties match\cite{norman_things_smart}
+        \pause
+        \item[Expressiveness] which structure preserving mappings can a tool implement\cite{mackinlayAutomatingDesignGraphical1986}] 
     \end{description}
@@
 
-\begin{frame}{Equivariance}
+\begin{frame}{Frameworks for Expressing Visual Equivariance}
+    \begin{description}
+        \item[linguistically] visualization has syntax, semantics, grammar expresses how to design structure preserving visualizations
+        \cite{mackinlayAutomatingDesignGraphical1986,mackinlayAUTOMATICDESIGNGRAPHICAL1987,wilkinsonGrammarGraphics2005}
 
+        \item[algebraically] transformations on data and graphics are equivalent symmetric \cite{kindlmannAlgebraicProcessVisualization2014}
+        \begin{columns}
+            \column{.5\textwidth}
+            \begin{description}
+                \item[D] data 
+                \item[R] data representation 
+                \item[V] visualization
+            \end{description}
+            \column{.5\textwidth}
+            \begin{equation*}
+                \begin{tikzcd}[ampersand replacement=\&]
+                    D \arrow[d, "\alpha"'] \arrow[r, "r_1"] \& R \arrow[r, "\nu"]  \& V \arrow[d, "\omega"] \\
+                    D \arrow[r, "r_2"']                     \& R \arrow[r, "\nu"'] \& V                    
+                \end{tikzcd}
+                \end{equation*}
+        \end{columns} 
+    \item[categorically] \textit{understanding} = \textit{read} $\circ$ \textit{render} \cite{vickersUnderstandingVisualizationFormal2013}    
+    \end{description}
 \end{frame}
 
-\begin{frame}{Domain specific libraries assume data structure\cite{HeerSoftware2006}}
+\begin{frame}{Domain specific libraries know their structure\cite{HeerSoftware2006}}
     \begin{table}
@@
 
-\begin{frame}{General purpose libraries can't\cite{toryRethinkingVisualizationHighlevel2004}}
+\begin{frame}{General purpose libraries generally can't\cite{toryRethinkingVisualizationHighlevel2004}}
     \begin{figure}
-        \includegraphics[height=.5\textheight]{figures/intro/dataset_diagram.png}
-        \caption{Data Representation, MayaVi 4.7.2 docs\cite{DataRepresentationMayavi}}
+        \includegraphics[height=.3\textheight]{../paper/figures/k_different_types.png}
     \end{figure}
@@
     \begin{columns}
-        \column{0.49\columnwidth}
+        \column{0.39\textwidth}
             \begin{equation*}
                 %\label{eq:related-work:continuity:fiber-bundle}
-                \begin{tikzcd}[ampersand replacement=\&]
+                \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
                   \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
@@
             \end{equation*}
-     \column{0.49\columnwidth}
+     \column{0.59\textwidth}
         \begin{description}
-            \item[\textcolor{base}{Base Space}] $( \dbasec, \mathcal{T})$
+            \item[\textcolor{base}{Base Space}] $( \dbasec, \mathcal{T})$ with
+                \begin{itemize}
+                   \item open sets $\opensetc \subset \dbasec$
+                   \item points $\dbasepointc \in \opensetc \in \dbasec$
+                 \end{itemize}  
             \item[\textcolor{fiber}{Fiber Space}] $\dfiberc_{\dbasepointc} = \pi^{-1}(\dbasepointc)$
-            \item[\textcolor{total}{Total Space}]  $\dtotalc = \dbasec \times \dfiberc$
+            \item[\textcolor{total}{Total Space}]  $\dtotalc\restriction_{\opensetc} = \dbasec\restriction_{\opensetc} \times \dfiberc\restriction_{\opensetc}$ \\ for all $\opensetc \subset K$ 
         \end{description}
@@
     \pause
-    \begin{block}{Sections of the bundle}
+    \begin{alertblock}{Sections}
+    $\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\openset} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}$
+    \end{alertblock}
+\end{frame}
+
+\begin{frame}{Category $\textcolor{source}{\mathcal{C}}$}
+\begin{columns}
+    \column{0.39\textwidth}
 \begin{equation*}
-            \label{eq:related-work:continuity:sections}
-            \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\openset} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;\forall \dbasepointc \in \opensetc \subset \dbasec \big\}
+    \begin{tikzcd}[ampersand replacement=\&]
+        \textcolor{source}{C_1} \arrow[r, "f", color=source] \arrow[rd, "g\circ f"', color=source] \arrow["id_{C_1}"', loop, distance=2em, in=215, out=145, color=source] \& \textcolor{source}{C_2} \arrow[d, "g", color=source] \arrow["id_{C_2}"', loop, distance=2em, in=125, out=55, color=source] \\
+    \& \textcolor{source}{C_3} \arrow["id_{C_3}"', loop, distance=2em, in=305, out=235, color=source]              
+    \end{tikzcd}
     \end{equation*}
-    \end{block}
+    \column{0.59\textwidth}
+    \begin{alertblock}{associativity} 
+        if $f: C_1 \rightarrow C_2$, $g: C_2 \rightarrow C_3$ and $h: C_3 \rightarrow C_4$ then $h\circ (g \circ f) = (h \circ g) \circ f$
+    \end{alertblock}
+    \begin{alertblock}{identity} 
+        for every $f: C_1 \rightarrow C_2$ there exists identity morphisms $f \circ id_{C_1} = f = id_{C_2} \circ f$
+    \end{alertblock}
+    \end{columns}
 \end{frame}
+
+\begin{frame}{Opposite Category $\textcolor{source}{\mathcal{C}^{op}}$}
+    \begin{columns}
+        \column{0.49\textwidth}
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement=\&]
+            \textcolor{fade}{C_1} \arrow[r, "f", color=fade] \arrow[rd, "g\circ f"', color=fade] \arrow["id_{C_1}"', loop, distance=2em, in=215, out=145, color=fade] \& \textcolor{fade}{C_2} \arrow[d, "g", color=fade] \arrow["id_{C_2}"', loop, distance=2em, in=125, out=55, color=fade] \\
+        \& \textcolor{fade}{C_3} \arrow["id_{C_3}"', loop, distance=2em, in=305, out=235, color=fade]              
+        \end{tikzcd}
+        \end{equation*}
+        \column{0.49\textwidth}
+        \begin{equation*}
+        \begin{tikzcd}[ampersand replacement = \&]
+            \textcolor{source}{C_1} \arrow["id_{C_1}"', loop, distance=2em, in=215, out=145, color=source] \& \textcolor{source}{C_2} \arrow[l, "f"',color=source] \arrow["id_{C_2}"', loop, distance=2em, in=125, out=55, color=source]\\
+            \& \textcolor{source}{C_3} \arrow[u, "g"', color=source] \arrow[lu, "f \circ g", color=source] \arrow["id_{C_3}"', loop, distance=2em, in=305, out=235, color=source]
+            \end{tikzcd}
+        \end{equation*}
+        \end{columns}
+\end{frame}
+
+\begin{frame}{Functor $\textcolor{sheaf}{\bm{F}}: \textcolor{source}{\mathcal{C}} \rightarrow \textcolor{target}{\mathcal{D}}$}
+\begin{columns}
+    \column{.39\textwidth}
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement = \&, ]
+            \textcolor{source}{c} \arrow[d, "f"', shift right=0, color=source] \arrow[r, "F", color=functor] \& \textcolor{target}{F(c)} \arrow[d, "F(f)", shift right=0, color=target] \\
+            \textcolor{source}{c^{\prime}} \arrow[r, "F"', color=functor]                       \& \textcolor{target}{F(c^{\prime})}                      
+        \end{tikzcd}
+    \end{equation*}
+    \column{.59\textwidth}
+    \begin{alertblock}{composition}
+    $\textcolor{functor}{F}(\textcolor{source}{g}) \circ  \textcolor{functor}{F}(\textcolor{source}{f}) = \textcolor{functor}{F} (\textcolor{source}{g}\circ \textcolor{source}{f})$
+    \end{alertblock}
+    \begin{alertblock}{identity}
+    $\textcolor{functor}{F}(\textcolor{source}{id_c}) = \textcolor{target}{id}_{\textcolor{functor}{F}(\textcolor{source}{c})}$
+    \end{alertblock}
+\end{columns}
+\end{frame}
+
+\begin{frame}{Natural Transformation $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\rightarrow} \textcolor{functor}{G}$}
+
+\adjustbox{scale=3, center}{
+    \begin{tikzcd}[ampersand replacement=\&]
+        \& {} \arrow[dd, "\alpha", Rightarrow, shorten=1.5em, color=nattran] \&             \\
+      \textcolor{source}{\mathcal{C}} \arrow[rr, "F", bend left, color=functor] \arrow[rr, "G"', bend right, color=functor] \& \& \textcolor{target}{\mathcal{D}} \\
+        \& {} \&            
+      \end{tikzcd}}
+\end{frame}
+
+\begin{frame}{Natural Transformation $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\rightarrow} \textcolor{functor}{G}$}
+\begin{equation*}
+    \begin{tikzcd}[column sep=huge, ampersand replacement = \&]
+        \textcolor{target}{F(c)} \arrow[dd, "F(f)"', color=target] \arrow[rr, "\alpha_{c}", bend left, color=nattran] \& \textcolor{source}{c} \arrow[dd, "f" description, color=source] \arrow[r, "G"', color=functor] \arrow[l, "F", color=functor] \& \textcolor{target}{G(c)} \arrow[dd, "G(f)", color=target] \\
+        {} \arrow[rr, dotted, color=nattran] \& \& {}                      \\
+        \textcolor{target}{F(c^{\prime})} \arrow[rr, "\alpha_{c^{\prime}}"', bend right, color=nattran] \& \textcolor{source}{c^{\prime}} \arrow[l, "F"', color=functor] \arrow[r, "G", color=functor] \& \textcolor{target}{G(c^{\prime})}
+    \end{tikzcd}
+\end{equation*}    
+\end{frame}
+
+\begin{frame}{Presheaf: $\sheafc: \opensetc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$}
+\begin{equation*}
+    \begin{tikzcd}[ampersand replacement = \&, column sep=small]
+        {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} \& {} \& {\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}} \arrow[ll, "\iota^*"', hook', color=set] \\\& \&\\
+        \opensetc_1 \arrow[uu, "\sheafc_{\dbasec}", color=sheaf] \arrow[rr, "\iota"', hook, color=base] \& {} \arrow[uu, "\sheafc_{\dbasec}" description, dotted, color=sheaf] \& \opensetc_2 \arrow[uu, "\sheafc_{\dbasec}"', color=sheaf]
+        \end{tikzcd}
+\end{equation*}
+\end{frame}
+
+\begin{frame}{Sheaf}
+
+\end{frame}
+\begin{frame}{Stalk \& Germ}
+
+\adjustbox{scale=1}{
+    \begin{description}
+    \item[Stalk] $\sheafc(\dbasec)\restriction_{\dbasepointc}\coloneqq \lim\limits_{\opensetc\ni \dbasepointc} \Gamma(\opensetc, \dtotalc\restriction_{\opensetc})$
+    \item[Germ] $\dsectionc(\dbasepointc) \in \sheafc(\dbasec)$
+\end{description}
+}
+\end{frame}
+
+\begin{comment}
 \subsection{Contributions}
@@
 
-\begin{frame}{Mathematical Models of Vizualization}
-    \begin{description}
-        \item[algebraic process] data and viz transforms are symmetric \cite{kindlmannAlgebraicProcessVisualization2014}
-        \begin{equation*}
-            \begin{tikzcd}[ampersand replacement=\&]
-                \dtotal \arrow[d, "m"'] \arrow[r, "\nu"] \& \vtotal \arrow[r, "Q"]  \& \gtotal \arrow[d, "\varphi(m)"] \\
-                \dtotal \arrow[r, "\nu"']                     \& \vtotal \arrow[r, "Q"'] \& \gtotal                  
-            \end{tikzcd}
-        \end{equation*}
-
-        \item[language] APT and GoG: syntax, semantics, and grammar\\ 
-        \cite{mackinlayAutomatingDesignGraphical1986,mackinlayAUTOMATICDESIGNGRAPHICAL1987,wilkinsonGrammarGraphics2005}
-        \item[functional dependencies] relationship between components \cite{sugibuchiFramwork2009}
-        \item[category theory]  \textit{understanding} = \textit{read} $\circ$ \textit{render} \cite{vickersUnderstandingVisualizationFormal2013}    
-    \end{description}
+\begin{frame}{Bertin Retinal Variables}
+    \begin{figure}
+        \includegraphics[width=.7\textwidth]{figures/intro/retinal_variables.png}
+        \caption{This tabular form of Bertin's retinal variables is from Understanding Graphics \cite{malamedInformationDisplayTips2010} who reproduced it from Krygier and Wood's \textit{Making Maps: A Visual Guide to Map Design for GIS}\cite{krygierMakingMapsVisual2005}}
+    \end{figure}
 \end{frame}
@@
 %\end{comment}
+
+\end{comment}
 \end{document}

### 2021-12-29 slides/main.tex
sheaves breakdown

@@
 
-\usetheme{ccnycrest}
+%\usetheme{ccnycrest}
 
@@
 \title{Topological Equivariant Artist Model}
-
+\begin{comment}
 \begin{frame}
@@
 
+%% maybe pull out the middle for now? move the 3 stages down to construction? 
 \begin{frame}{What do visualization libraries do?}
@@
 
+%flip columns
 \begin{frame}{Frameworks for Expressing Visual Equivariance}
@@
 \end{frame}
-
+\end{comment}
 \subsection{Algebraic Topology & Category Theory}
@@
 
-\begin{frame}{Presheaf: $\sheafc: \opensetc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$}
-\begin{equation*}
+\begin{frame}{Subsets?}
+    \begin{columns}
+    \column{0.29\textwidth}
+    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
+          \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
+           \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
+        \end{tikzcd}
+    \column{0.69\textwidth}
+        \begin{tikzcd}[ampersand replacement = \&, column sep=small]
+            {} \& {} \& \\\& \&\\
+            \opensetc_1 \subset \dbasec \arrow[uu, "", color=white] \& {}  \&                    
+            \end{tikzcd}
+    \end{columns}
+\end{frame}
+
+\begin{frame}{Subsets?}
+    \begin{columns}
+    \column{0.29\textwidth}
+    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
+          \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
+           \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
+        \end{tikzcd}
+    \column{0.69\textwidth}
+        \begin{tikzcd}[ampersand replacement = \&, column sep=small]
+            {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} \& {} \& \\\& \&\\
+            \opensetc_1 \subset \dbasec \arrow[uu, "", color=white] \& {}  \&                    
+            \end{tikzcd}
+    \end{columns}
+\end{frame}
+
+\begin{frame}{Presheaf on Bundle: $\sheafc: \opensetc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$}
+    \begin{columns}
+    \column{0.29\textwidth}
+    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
+          \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
+           \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
+        \end{tikzcd}
+    \column{0.69\textwidth}
+        \begin{tikzcd}[ampersand replacement = \&, column sep=small]
+            {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} \& {} \& \\\& \&\\
+            \opensetc_1 \subset \dbasec \arrow[uu, "\sheafc_{\dbasec}", color=sheaf] \& {}  \&                    
+            \end{tikzcd}
+    \end{columns}
+\end{frame}
+
+\begin{frame}{Presheaf on Bundle: $\sheafc: \opensetc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$}
+    \begin{columns}
+    \column{0.29\textwidth}
+    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
+          \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
+           \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
+        \end{tikzcd}
+    \column{0.69\textwidth}
+        \begin{tikzcd}[ampersand replacement = \&, column sep=small]
+            {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} \& {} \& {\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}}  \\\& \&\\
+            \opensetc_1 \arrow[uu, "\sheafc_{\dbasec}", color=sheaf]  \& {}  \& \opensetc_2 \arrow[uu, "\sheafc_{\dbasec}"', color=sheaf]                    
+            \end{tikzcd}
+    \end{columns}
+\end{frame}
+
+\begin{frame}{Presheaf on Bundle: $\sheafc: \opensetc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$}
+    \begin{columns}
+    \column{0.29\textwidth}
+    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
+          \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
+           \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
+        \end{tikzcd}
+    \column{0.69\textwidth}
+        \begin{tikzcd}[ampersand replacement = \&, column sep=small]
+            {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} \& {} \& {\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}} \arrow[ll, "\iota^*"', hook', color=set] \\\& \&\\
+            \opensetc_1 \arrow[uu, "\sheafc_{\dbasec}", color=sheaf] \arrow[rr, "\iota"', hook, color=base] \& {}  \& \opensetc_2 \arrow[uu, "\sheafc_{\dbasec}"', color=sheaf]                    
+            \end{tikzcd}
+    \end{columns}
+\end{frame}
+
+\begin{frame}{Presheaf on Bundle: $\sheafc: \opensetc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$}
+    \begin{columns}
+    \column{0.29\textwidth}
+    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
+          \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
+           \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
+        \end{tikzcd}
+    \column{0.69\textwidth}
         \begin{tikzcd}[ampersand replacement = \&, column sep=small]
@@
             \end{tikzcd}
-\end{equation*}
+    \end{columns}
+
+    \begin{description}
+        \item<2-|handout:2->[stalk]$\sheafc(\dbasec)\restriction_{\dbasepointc}\coloneqq \lim\limits_{\opensetc\ni \dbasepointc} \Gamma(\opensetc, \dtotalc\restriction_{\opensetc}) \approx \dfiberc_{\dbasepointc}$
+        \item<3-|handout:3>[germ]   $\dsectionc(\dbasepointc) \in \sheafc_{\dbasec}$
+    \end{description}
 \end{frame}
@@
 
-\end{frame}
-\begin{frame}{Stalk \& Germ}
-
-\adjustbox{scale=1}{
-    \begin{description}
-    \item[Stalk] $\sheafc(\dbasec)\restriction_{\dbasepointc}\coloneqq \lim\limits_{\opensetc\ni \dbasepointc} \Gamma(\opensetc, \dtotalc\restriction_{\opensetc})$
-    \item[Germ] $\dsectionc(\dbasepointc) \in \sheafc(\dbasec)$
-\end{description}
-}
 \end{frame}

### 2021-12-29 slides/main.tex
sheaf equation
breakdown of bundle
reworked functors

@@
                 \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-                  \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
-                   \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
+                  \onslide*<2-|handout:2->{\dfiberc} \onslide*<4-|handout:4->{\arrow[r, hook]} \& \onslide*<3|handout:3->{\dtotalc} \onslide*<4-|handout:4->{\arrow[d, "\pi"']} \\
+                   \& \onslide*<1-|handout:1->{\dbasec} \onslide*<5-|handout:5->{\arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]}
                 \end{tikzcd}
@@
         \begin{description}
-            \item[\textcolor{base}{Base Space}] $( \dbasec, \mathcal{T})$ with
-                \begin{itemize}
-                   \item open sets $\opensetc \subset \dbasec$
-                   \item points $\dbasepointc \in \opensetc \in \dbasec$
-                 \end{itemize}  
-            \item[\textcolor{fiber}{Fiber Space}] $\dfiberc_{\dbasepointc} = \pi^{-1}(\dbasepointc)$
-            \item[\textcolor{total}{Total Space}]  $\dtotalc\restriction_{\opensetc} = \dbasec\restriction_{\opensetc} \times \dfiberc\restriction_{\opensetc}$ \\ for all $\opensetc \subset K$ 
-        \end{description}
-    \end{columns}
-    \pause
-    \begin{alertblock}{Sections}
+            \item<1-|handout:1->[\textcolor{base}{Base Space}] $( \dbasec, \mathcal{T})$, $\dbasepointc \in \opensetc \subset \dbasec$
+            \item<2-|handout:2->[\textcolor{fiber}{Fiber Space}] $\dfiberc_{\dbasepointc} = \pi^{-1}(\dbasepointc)$
+            \item<3-|handout:3->[\textcolor{total}{Total Space}]  $\dtotalc\restriction_{\opensetc} = \dbasec\restriction_{\opensetc} \times \dfiberc\restriction_{\opensetc}$ \\ for all $\opensetc \subset K$
+        \end{description}
+    \end{columns}
+    \pause
+    \begin{alertblock}<5-|handout:5->{Sections}
     $\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\openset} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}$
@@
 
+\subsection{Category Theory}
 \begin{frame}{Category $\textcolor{source}{\mathcal{C}}$}
@@
 \end{frame}
+\begin{frame}{Category $\mathcal{\dbasec}$}
+\begin{columns}
+    \column{0.49\textwidth}
+\begin{equation*}
+    \begin{tikzcd}[ampersand replacement=\&]
+        \textcolor{source}{C_1} \arrow[r, "f", color=source] \arrow[rd, "g\circ f"', color=source] \arrow["id_{C_1}"', loop, distance=2em, in=215, out=145, color=source] \& \textcolor{source}{C_2} \arrow[d, "g", color=source] \arrow["id_{C_2}"', loop, distance=2em, in=125, out=55, color=source] \\
+    \& \textcolor{source}{C_3} \arrow["id_{C_3}"', loop, distance=2em, in=305, out=235, color=source]              
+    \end{tikzcd}
+    \end{equation*}
+    \column{0.49\textwidth}
+    \onslide<2-|handout:2->{
+    \begin{tikzcd}[ampersand replacement=\&]
+            \opensetc_1 \arrow[r, "\iota_{1,2}", color=base] \arrow[rd, "\iota_{1,2}\circ \iota_{2,3}"', color=base] \arrow["id_{\opensetc_1}"', loop, distance=2em, in=215, out=145, color=base] \& \opensetc_2 \arrow[d, "\iota_{2,3}", color=base] \arrow["id_{\opensetc_2}"', loop, distance=2em, in=125, out=55, color=base] \\
+        \& \opensetc_3 \arrow["id_{\opensetc_3}"', loop, distance=2em, in=305, out=235, color=base]              
+    \end{tikzcd}
+    }
+    \end{columns}
+\end{frame}
 
@@
 
-\begin{frame}{Functor $\textcolor{sheaf}{\bm{F}}: \textcolor{source}{\mathcal{C}} \rightarrow \textcolor{target}{\mathcal{D}}$}
+\begin{frame}{$\mathcal{C}^{op}$ with Objects of \setc}
+    \begin{columns}
+        \column{0.29\textwidth}
+        \begin{tikzcd}[ampersand replacement=\&]
+            \opensetc_1 \arrow[r, "\iota_{1,2}", color=base] \arrow[rd, "\iota_{1,2}\circ \iota_{2,3}"', color=base] \arrow["id_{\opensetc_1}"', loop, distance=2em, in=215, out=145, color=base] \& \opensetc_2 \arrow[d, "\iota_{2,3}", color=base] \arrow["id_{\opensetc_2}"', loop, distance=2em, in=125, out=55, color=base] \\
+        \& \opensetc_3 \arrow["id_{\opensetc_3}"', loop, distance=2em, in=305, out=235, color=base]              
+    \end{tikzcd}
+        \column{0.69\textwidth}
+        \onslide<2-|handout:2->{
+            \begin{tikzcd}[ampersand replacement=\&]
+            \cgamma{\opensetc_1}{\dtotalc\restriction_{\opensetc_1}} \&  \& \cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}} \arrow[ll, "\iota^*\restriction_{\opensetc_1}"', color=set]\\
+            \&\&\\\&  \& \cgamma{\opensetc_3}{\dtotalc\restriction_{\opensetc_3}} \arrow[uu, "\iota^*\restriction_{\opensetc_2}"', color=set] \arrow[lluu, "\iota^*\restriction_{\opensetc_2} \circ \iota^*\restriction_{\opensetc_1}", color=set]
+            \end{tikzcd}
+
+        }
+
+        \end{columns}
+\end{frame}
+
+\begin{frame}{Functor $\textcolor{functor}{\bm{F}}: \textcolor{source}{\mathcal{C}} \rightarrow \textcolor{target}{\mathcal{D}}$}
 \begin{columns}
@@
         \begin{tikzcd}[ampersand replacement = \&, ]
-            \textcolor{source}{c} \arrow[d, "f"', shift right=0, color=source] \arrow[r, "F", color=functor] \& \textcolor{target}{F(c)} \arrow[d, "F(f)", shift right=0, color=target] \\
-            \textcolor{source}{c^{\prime}} \arrow[r, "F"', color=functor]                       \& \textcolor{target}{F(c^{\prime})}                      
+            \onslide*<1-|handout:1->{\textcolor{source}{c}} 
+            \onslide*<1-|handout:1->{\arrow[d, "f"', shift right=0, color=source]} \onslide*<3-|handout:3->{\arrow[r, "F", color=functor]} \& 
+            \onslide*<2|handout:2>{\textcolor{target}{d}}
+            \onslide*<3-|handout:3->{\textcolor{target}{F(c)}} 
+            \onslide*<2|handout:2>{\arrow[d, shift right=0, color=target]}
+            \onslide*<3-|handout:3->{\arrow[d, "F(f)", shift right=0, color=target]} \\
+            \onslide*<1-|handout:1->{\textcolor{source}{c^{\prime}}} 
+            \onslide*<3-|handout:3->{\arrow[r, "F"', color=functor]}\& 
+            \onslide*<2|handout:2>{\textcolor{target}{d^{\prime}}}
+            \onslide*<3-|handout:3->{\textcolor{target}{F(c^{\prime})}}                      
         \end{tikzcd}
@@
     \column{.59\textwidth}
-    \begin{alertblock}{composition}
+    \begin{alertblock}<4-|handout:4->{composition}
     $\textcolor{functor}{F}(\textcolor{source}{g}) \circ  \textcolor{functor}{F}(\textcolor{source}{f}) = \textcolor{functor}{F} (\textcolor{source}{g}\circ \textcolor{source}{f})$
     \end{alertblock}
-    \begin{alertblock}{identity}
+    \begin{alertblock}<5-|handout:5->{identity}
     $\textcolor{functor}{F}(\textcolor{source}{id_c}) = \textcolor{target}{id}_{\textcolor{functor}{F}(\textcolor{source}{c})}$
@@
 
+\begin{frame}{Presheaf on Bundle: $\sheafc: \opensetc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$}
+    \begin{columns}
+    \column{0.29\textwidth}
+    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
+          \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
+           \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
+        \end{tikzcd}
+    \column{0.69\textwidth}
+        \begin{tikzcd}[ampersand replacement = \&, column sep=small]
+            {} \& {} \& \\\& \&\\
+            \opensetc_1 \subset \dbasec \arrow[uu, "", color=white] \& {}  \&                    
+            \end{tikzcd}
+    \end{columns}
+\end{frame}
+
+\begin{frame}{Presheaf on Bundle: $\sheafc: \opensetc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$}
+    \begin{columns}
+    \column{0.29\textwidth}
+    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
+          \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
+           \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
+        \end{tikzcd}
+    \column{0.69\textwidth}
+        \begin{tikzcd}[ampersand replacement = \&, column sep=small]
+            {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} \& {} \& \\\& \&\\
+            \opensetc_1 \subset \dbasec \arrow[uu, "", color=white] \& {}  \&                    
+            \end{tikzcd}
+    \end{columns}
+\end{frame}
+
+\begin{frame}{Presheaf on Bundle: $\sheafc: \opensetc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$}
+    \begin{columns}
+    \column{0.29\textwidth}
+    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
+          \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
+           \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
+        \end{tikzcd}
+    \column{0.69\textwidth}
+        \begin{tikzcd}[ampersand replacement = \&, column sep=small]
+            {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} \& {} \& \\\& \&\\
+            \opensetc_1 \subset \dbasec \arrow[uu, "\sheafc_{\dbasec}", color=sheaf] \& {}  \&                    
+            \end{tikzcd}
+    \end{columns}
+\end{frame}
+
+\begin{frame}{Presheaf on Bundle: $\sheafc: \opensetc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$}
+    \begin{columns}
+    \column{0.29\textwidth}
+    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
+          \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
+           \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
+        \end{tikzcd}
+    \column{0.69\textwidth}
+        \begin{tikzcd}[ampersand replacement = \&, column sep=small]
+            {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} \& {} \& {\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}}  \\\& \&\\
+            \opensetc_1 \arrow[uu, "\sheafc_{\dbasec}", color=sheaf]  \& {}  \& \opensetc_2 \arrow[uu, "\sheafc_{\dbasec}"', color=sheaf]                    
+            \end{tikzcd}
+    \end{columns}
+\end{frame}
+
+\begin{frame}{Presheaf on Bundle: $\sheafc: \opensetc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$}
+    \begin{columns}
+    \column{0.29\textwidth}
+    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
+          \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
+           \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
+        \end{tikzcd}
+    \column{0.69\textwidth}
+        \begin{tikzcd}[ampersand replacement = \&, column sep=small]
+            {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} \& {} \& {\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}} \arrow[ll, "\iota^*"', hook', color=set] \\\& \&\\
+            \opensetc_1 \arrow[uu, "\sheafc_{\dbasec}", color=sheaf] \arrow[rr, "\iota"', hook, color=base] \& {}  \& \opensetc_2 \arrow[uu, "\sheafc_{\dbasec}"', color=sheaf]                    
+            \end{tikzcd}
+    \end{columns}
+\end{frame}
+
+\begin{frame}{Presheaf on Bundle: $\sheafc: \opensetc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$}
+    \begin{columns}
+    \column{0.29\textwidth}
+    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
+          \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
+           \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
+        \end{tikzcd}
+    \column{0.69\textwidth}
+        \begin{tikzcd}[ampersand replacement = \&, column sep=small]
+            {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} \& {} \& {\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}} \arrow[ll, "\iota^*"', hook', color=set] \\\& \&\\
+            \opensetc_1 \arrow[uu, "\sheafc_{\dbasec}", color=sheaf] \arrow[rr, "\iota"', hook, color=base] \& {}  \arrow[uu, "\sheafc_{\dbasec}" description, dotted, color=sheaf]  \& \opensetc_2 \arrow[uu, "\sheafc_{\dbasec}"', color=sheaf]
+            \end{tikzcd}
+    \end{columns}
+
+    \begin{description}
+        \item<2-|handout:2->[stalk]$\sheafc(\dbasec)\restriction_{\dbasepointc}\coloneqq \lim\limits_{\opensetc\ni \dbasepointc} \Gamma(\opensetc, \dtotalc\restriction_{\opensetc}) \approx \dfiberc_{\dbasepointc}$
+        \item<3-|handout:3>[germ]   $\dsectionc(\dbasepointc) \in \sheafc_{\dbasec}$
+    \end{description}
+\end{frame}
+
+\begin{frame}{Sheaves}
+    A sheaf is a presheaf that satisfies the following two axioms
+
+    \onslide<2-|handout:2->{Given open cover $\mathcal{\opensetc} = \{\opensetc_i\}_{i\in I}$ of $\opensetc$:}
+    \begin{block}<3-|handout:3->{locality}
+    $\dsectionc^{1}, \dsectionc^{2} \in \sheafc(\opensetc_i)\\ \dsectionc^{1}\restriction_{\opensetc_i} = \dsectionc^{2}\restriction_{\opensetc_i} \implies \dsectionc^{1} = \dsectionc^{2}$
+    \end{block} 
+    \begin{block}<4-|handout:4>{gluing}
+    $\dsectionc_{i} \in \sheafc(\opensetc_i), \dsectionc_{j} \in \sheafc(\opensetc_j), \opensetc_i, \opensetc_j \in \mathcal{U}\\ \dsectionc_{i}\restriction_{\opensetc_i\cap \opensetc_j} = \dsectionc_{j}\restriction_{\opensetc_i \cap \opensetc_j} \implies \dsectionc\restriction_{\opensetc_i} = \dsectionc_i$
+    \end{block}    
+\end{description}    
+\end{frame}
+
+\subsubsection{Sheaves}
 \begin{frame}{Natural Transformation $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\rightarrow} \textcolor{functor}{G}$}
@@
 
-\begin{frame}{Subsets?}
-    \begin{columns}
-    \column{0.29\textwidth}
-    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-          \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
-           \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
-        \end{tikzcd}
-    \column{0.69\textwidth}
-        \begin{tikzcd}[ampersand replacement = \&, column sep=small]
-            {} \& {} \& \\\& \&\\
-            \opensetc_1 \subset \dbasec \arrow[uu, "", color=white] \& {}  \&                    
-            \end{tikzcd}
-    \end{columns}
-\end{frame}
-
-\begin{frame}{Subsets?}
-    \begin{columns}
-    \column{0.29\textwidth}
-    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-          \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
-           \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
-        \end{tikzcd}
-    \column{0.69\textwidth}
-        \begin{tikzcd}[ampersand replacement = \&, column sep=small]
-            {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} \& {} \& \\\& \&\\
-            \opensetc_1 \subset \dbasec \arrow[uu, "", color=white] \& {}  \&                    
-            \end{tikzcd}
-    \end{columns}
-\end{frame}
-
-\begin{frame}{Presheaf on Bundle: $\sheafc: \opensetc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$}
-    \begin{columns}
-    \column{0.29\textwidth}
-    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-          \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
-           \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
-        \end{tikzcd}
-    \column{0.69\textwidth}
-        \begin{tikzcd}[ampersand replacement = \&, column sep=small]
-            {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} \& {} \& \\\& \&\\
-            \opensetc_1 \subset \dbasec \arrow[uu, "\sheafc_{\dbasec}", color=sheaf] \& {}  \&                    
-            \end{tikzcd}
-    \end{columns}
-\end{frame}
-
-\begin{frame}{Presheaf on Bundle: $\sheafc: \opensetc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$}
-    \begin{columns}
-    \column{0.29\textwidth}
-    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-          \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
-           \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
-        \end{tikzcd}
-    \column{0.69\textwidth}
-        \begin{tikzcd}[ampersand replacement = \&, column sep=small]
-            {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} \& {} \& {\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}}  \\\& \&\\
-            \opensetc_1 \arrow[uu, "\sheafc_{\dbasec}", color=sheaf]  \& {}  \& \opensetc_2 \arrow[uu, "\sheafc_{\dbasec}"', color=sheaf]                    
-            \end{tikzcd}
-    \end{columns}
-\end{frame}
-
-\begin{frame}{Presheaf on Bundle: $\sheafc: \opensetc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$}
-    \begin{columns}
-    \column{0.29\textwidth}
-    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-          \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
-           \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
-        \end{tikzcd}
-    \column{0.69\textwidth}
-        \begin{tikzcd}[ampersand replacement = \&, column sep=small]
-            {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} \& {} \& {\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}} \arrow[ll, "\iota^*"', hook', color=set] \\\& \&\\
-            \opensetc_1 \arrow[uu, "\sheafc_{\dbasec}", color=sheaf] \arrow[rr, "\iota"', hook, color=base] \& {}  \& \opensetc_2 \arrow[uu, "\sheafc_{\dbasec}"', color=sheaf]                    
-            \end{tikzcd}
-    \end{columns}
-\end{frame}
-
-\begin{frame}{Presheaf on Bundle: $\sheafc: \opensetc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$}
-    \begin{columns}
-    \column{0.29\textwidth}
-    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-          \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
-           \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
-        \end{tikzcd}
-    \column{0.69\textwidth}
-        \begin{tikzcd}[ampersand replacement = \&, column sep=small]
-            {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} \& {} \& {\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}} \arrow[ll, "\iota^*"', hook', color=set] \\\& \&\\
-            \opensetc_1 \arrow[uu, "\sheafc_{\dbasec}", color=sheaf] \arrow[rr, "\iota"', hook, color=base] \& {}  \arrow[uu, "\sheafc_{\dbasec}" description, dotted, color=sheaf]  \& \opensetc_2 \arrow[uu, "\sheafc_{\dbasec}"', color=sheaf]
-            \end{tikzcd}
-    \end{columns}
-
-    \begin{description}
-        \item<2-|handout:2->[stalk]$\sheafc(\dbasec)\restriction_{\dbasepointc}\coloneqq \lim\limits_{\opensetc\ni \dbasepointc} \Gamma(\opensetc, \dtotalc\restriction_{\opensetc}) \approx \dfiberc_{\dbasepointc}$
-        \item<3-|handout:3>[germ]   $\dsectionc(\dbasepointc) \in \sheafc_{\dbasec}$
-    \end{description}
-\end{frame}
-
-\begin{frame}{Sheaf}
-
-\end{frame}
-
 \begin{comment}
+
 \subsection{Contributions}

### 2021-12-29 slides/main.tex
morphism on sheaves

@@
     \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-          \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
-           \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
+        \onslide*<8-|handout:8->{\dfiberc}
+        \onslide*<8-|handout:8->{\arrow[r, hook]} \& 
+        \onslide*<1-|handout:1->{\dtotalc}
+        \onslide*<1-|handout:1->{\arrow[d, "\pi"']} \\
+           \& 
+        \onslide*<1-|handout:1->{\dbasec}
+        \onslide*<3-|handout:3->{\arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]}
         \end{tikzcd}
@@
         \begin{tikzcd}[ampersand replacement = \&, column sep=small]
-            {} \& {} \& \\\& \&\\
-            \opensetc_1 \subset \dbasec \arrow[uu, "", color=white] \& {}  \&                    
+            \onslide*<3-|handout:3->{{\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}} \& 
+            {} \& 
+            \onslide*<5-|handout:5->{{\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}}} 
+            \onslide*<6-|handout:6->{\arrow[ll, "\iota^*"', hook', color=set]} \\\& \&\\
+            \onslide*<2|handout:2>{\opensetc_1 \subset \dbasec}
+            \onslide*<3-|handout:3->{\opensetc_1} 
+            \onslide*<4-|handout:4->{\arrow[uu, "\sheafc_{\dbasec}", color=sheaf]} 
+            \onslide*<6-|handout:6->{\arrow[rr, "\iota"', hook, color=base]} \&   
+            \onslide*<7-|handout:7->{\arrow[uu, "\sheafc_{\dbasec}" description, dotted, color=sheaf]} \& 
+            \onslide*<5-|handout:5->{\opensetc_2} 
+            \onslide*<5-|handout:5->{\arrow[uu, "\sheafc_{\dbasec}"', color=sheaf]}                   
             \end{tikzcd}
     \end{columns}
-\end{frame}
-
-\begin{frame}{Presheaf on Bundle: $\sheafc: \opensetc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$}
-    \begin{columns}
-    \column{0.29\textwidth}
-    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-          \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
-           \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
-        \end{tikzcd}
-    \column{0.69\textwidth}
-        \begin{tikzcd}[ampersand replacement = \&, column sep=small]
-            {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} \& {} \& \\\& \&\\
-            \opensetc_1 \subset \dbasec \arrow[uu, "", color=white] \& {}  \&                    
-            \end{tikzcd}
-    \end{columns}
-\end{frame}
-
-\begin{frame}{Presheaf on Bundle: $\sheafc: \opensetc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$}
-    \begin{columns}
-    \column{0.29\textwidth}
-    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-          \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
-           \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
-        \end{tikzcd}
-    \column{0.69\textwidth}
-        \begin{tikzcd}[ampersand replacement = \&, column sep=small]
-            {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} \& {} \& \\\& \&\\
-            \opensetc_1 \subset \dbasec \arrow[uu, "\sheafc_{\dbasec}", color=sheaf] \& {}  \&                    
-            \end{tikzcd}
-    \end{columns}
-\end{frame}
-
-\begin{frame}{Presheaf on Bundle: $\sheafc: \opensetc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$}
-    \begin{columns}
-    \column{0.29\textwidth}
-    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-          \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
-           \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
-        \end{tikzcd}
-    \column{0.69\textwidth}
-        \begin{tikzcd}[ampersand replacement = \&, column sep=small]
-            {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} \& {} \& {\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}}  \\\& \&\\
-            \opensetc_1 \arrow[uu, "\sheafc_{\dbasec}", color=sheaf]  \& {}  \& \opensetc_2 \arrow[uu, "\sheafc_{\dbasec}"', color=sheaf]                    
-            \end{tikzcd}
-    \end{columns}
-\end{frame}
-
-\begin{frame}{Presheaf on Bundle: $\sheafc: \opensetc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$}
-    \begin{columns}
-    \column{0.29\textwidth}
-    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-          \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
-           \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
-        \end{tikzcd}
-    \column{0.69\textwidth}
-        \begin{tikzcd}[ampersand replacement = \&, column sep=small]
-            {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} \& {} \& {\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}} \arrow[ll, "\iota^*"', hook', color=set] \\\& \&\\
-            \opensetc_1 \arrow[uu, "\sheafc_{\dbasec}", color=sheaf] \arrow[rr, "\iota"', hook, color=base] \& {}  \& \opensetc_2 \arrow[uu, "\sheafc_{\dbasec}"', color=sheaf]                    
-            \end{tikzcd}
-    \end{columns}
-\end{frame}
-
-\begin{frame}{Presheaf on Bundle: $\sheafc: \opensetc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$}
-    \begin{columns}
-    \column{0.29\textwidth}
-    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-          \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
-           \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
-        \end{tikzcd}
-    \column{0.69\textwidth}
-        \begin{tikzcd}[ampersand replacement = \&, column sep=small]
-            {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} \& {} \& {\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}} \arrow[ll, "\iota^*"', hook', color=set] \\\& \&\\
-            \opensetc_1 \arrow[uu, "\sheafc_{\dbasec}", color=sheaf] \arrow[rr, "\iota"', hook, color=base] \& {}  \arrow[uu, "\sheafc_{\dbasec}" description, dotted, color=sheaf]  \& \opensetc_2 \arrow[uu, "\sheafc_{\dbasec}"', color=sheaf]
-            \end{tikzcd}
-    \end{columns}
-
-    \begin{description}
-        \item<2-|handout:2->[stalk]$\sheafc(\dbasec)\restriction_{\dbasepointc}\coloneqq \lim\limits_{\opensetc\ni \dbasepointc} \Gamma(\opensetc, \dtotalc\restriction_{\opensetc}) \approx \dfiberc_{\dbasepointc}$
-        \item<3-|handout:3>[germ]   $\dsectionc(\dbasepointc) \in \sheafc_{\dbasec}$
+    \begin{description}
+        \item<8-|handout:8->[stalk]$\sheafc(\dbasec)\restriction_{\dbasepointc}\coloneqq \lim\limits_{\opensetc\ni \dbasepointc} \Gamma(\opensetc, \dtotalc\restriction_{\opensetc}) \approx \dfiberc_{\dbasepointc}$
+        \item<9-|handout:9>[germ]   $\dsectionc(\dbasepointc) \in \sheafc_{\dbasec}$
     \end{description}
@@
     \end{block}      
-\end{description}    
-\end{frame}
-
-\subsubsection{Sheaves}
+\end{frame}
+
+\begin{frame}{Morphism on Sheaves:\onslide<5-6,8-|handout:5-6,8->{ pullback}\onslide<8-|handout:8->{ and }\onslide<7-|handout:7->{pushforward}}
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
+            \onslide*<7-|handout:7->{\sheafc_{\gbasec}(\gbasec)} 
+            \onslide*<7-|handout:7->{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]} \& \& 
+            \onslide*<7-|handout:7->{(\textcolor{functor}{\vindexpush}\sheafc_{\gbasec})(\dbasec)}  \\
+            \onslide*<2-|handout:2->{\gbasec}
+            \onslide*<3-|handout:3->{\arrow[rr, "\vindex", color=functor]}
+            \onslide*<7-|handout:7->{\arrow[u, "\sheaf_{\gbasec}", color=sheaf]}
+            \onslide*<6,8-|handout:6,8->{\arrow[d, dashed, color=sheaf]} \&  \& 
+            \onslide*<1-|handout:1->{\dbasec}
+            \onslide*<4-6,8-|handout:4-6,8->{\arrow[d, "\sheaf_{\dbasec}", color=sheaf]}
+            \onslide*<7-|handout:7->{\arrow[u, dashed, color=sheaf]} \\
+            \onslide*<5-6,8-|handout:5-6,8->{(\textcolor{functor}{\vindexpull}\sheafc_{\dbasec})(\gbasec)} 
+            \onslide*<5-6,8-|handout:5-6,8->{\arrow[rr, "\textcolor{functor}{\vindexpull}", color=functor]}  \& \& 
+            \onslide*<4-6,8-|handout:4-6,8->{\sheafc_{\dbasec}(\dbasec)}                             
+        \end{tikzcd}
+    \end{equation*}
+\end{frame}
+
 \begin{frame}{Natural Transformation $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\rightarrow} \textcolor{functor}{G}$}

### 2021-12-29 slides/main.tex
sheaves morphism

@@
 
-\begin{frame}{Morphism on Sheaves:\onslide<5-6,8-|handout:5-6,8->{ pullback}\onslide<8-|handout:8->{ and }\onslide<7-|handout:7->{pushforward}}
+\begin{frame}{
+\only<-8|handout:-8>{Morphism on Sheaves:}
+\only<5-6|handout:5-6>{pullback}\only<7|handout:7>{pushforward} 
+\only<8|handout:8>{pullback \& pushforward}\only}
     \begin{equation*}
@@
 
+\begin{frame}{\only<1-3:handout:1-3>{Pullback \& Pushforward are Isomorphic}\only<4-|handout:4>{Data to Graphic is a Sheaf Morphism}}
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement=\&]
+          \onslide*<1-|handout:1->{\sheafc_{\gbasec}(\gbasec)} 
+          \onslide*<1-3|handout:1-3>{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]}  
+          \onslide*<4-|handout:4->{\arrow[rr, "\vindexpushc", color=artist]} 
+          \& \& 
+          \onslide*<1-|handout:1->{(\textcolor{functor}{\vindexpush}\sheafc_{\gbasec})(\dbase)} \\
+          \& \& \\
+          \onslide*<1-|handout:1->{(\textcolor{functor}{\vindexpull}\sheafc_{\dbase})(\gbasec)}
+          \onslide*<1-3|handout:1-3>{\arrow[rr, "\textcolor{functor}{\vindexpull}", color=functor]}
+          \onslide*<4-|handout:4->{\arrow[rr, "\vindexpullc", color=artist]}
+          \onslide*<2-3|handout:2-3>{\arrow[uu, "\textcolor{homset}{hom_{\gbasec}}", color=homset]} 
+          \onslide*<4-|handout:4->{\arrow[uu, "\vartistc_{\vindexpushc}", color=artist]}
+          \&  \& 
+          \onslide*<1-|handout:1->{\sheafc_{\dbase}(\dbase)} 
+          \onslide*<2-3|handout:2-3>{\arrow[uu, "\textcolor{homset}{hom_{\dbasec}}"', color=homset]}
+          \onslide*<4-|handout:4->{\arrow[uu, "\vartistc_{\vindexpullc}"', color=artist]}
+          \onslide*<3|handout:3>{\arrow[lluu, "\textcolor{homset}{hom_{\dbasec,\gbasec}}"' description, dashed, color=homset]}
+          \onslide*<4-|handout:4->{\arrow[lluu, "\vartistc"' description, dashed, color=artist]}
+        \end{tikzcd}
+      \end{equation*}
+
+    \onslide<3|handout:3>{
+        \begin{equation*}
+        \textcolor{homset}{hom}_{\gbasec}(\textcolor{functor}{\vindexpull}\sheafc_{\dbase}, \sheafc_{\gbasec}) = 
+    \textcolor{homset}{hom}_{\dbase}(\sheafc_{\dbase}, \textcolor{functor}{\vindexpush}\sheafc_{\gbasec})
+        \end{equation*}
+    }
+    \onslide<4|handout:4>{
+        \begin{equation*}
+            \vartistc_{\vindexpull} = \vartistc_{\vindexpush}
+        \end{equation*}
+    }
+\end{frame}
+
+\begin{frame}{}
+
+\end{frame}
+
 \begin{frame}{Natural Transformation $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\rightarrow} \textcolor{functor}{G}$}

### 2021-12-29 slides/main.tex
pushed sections
section morphisms
slides up to artist is homset
nt

@@
 
-\begin{frame}{$\mathcal{C}^{op}$ with Objects of \setc}
+\begin{frame}{$\mathcal{C}^{op}$}
     \begin{columns}
@@
 
-\begin{frame}{Presheaf on Bundle: $\sheafc: \opensetc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$}
+\begin{frame}{Presheaf: $\sheafc: \mathcal{C}^{op} \textcolor{sheaf}{\rightarrow} \setc $}
     \begin{columns}
@@
 \begin{frame}{
-\only<-8|handout:-8>{Morphism on Sheaves:}
-\only<5-6|handout:5-6>{pullback}\only<7|handout:7>{pushforward} 
-\only<8|handout:8>{pullback \& pushforward}\only}
+\only<-8|handout:-8>{Morphism of sheaves:}
+\only<5-6|handout:5-6>{pullback}
+\only<7|handout:7>{pushforward} 
+\only<8,11-|handout:8,11->{pullback \& pushforward}
+\only<9|handout:9>{pullback sections}
+\only<10|handout:10>{pushforward sections}
+}
     \begin{equation*}
         \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-            \onslide*<7-|handout:7->{\sheafc_{\gbasec}(\gbasec)} 
-            \onslide*<7-|handout:7->{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]} \& \& 
-            \onslide*<7-|handout:7->{(\textcolor{functor}{\vindexpush}\sheafc_{\gbasec})(\dbasec)}  \\
-            \onslide*<2-|handout:2->{\gbasec}
+            \onslide*<7-8|handout:7-8>{\sheafc_{\gbasec}(\gbasec)} 
+            \onslide*<10-|handout:10->{\{\gsectionc: \gbasepointc \scarr \setc \}}
+            \onslide*<7-8,10-|handout:7-8,10->{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]} \& \& 
+            \onslide*<7-8|handout:7-8>{(\textcolor{functor}{\vindexpush}\sheafc_{\gbasec})(\dbasec)}  
+            \onslide*<10-|handout:10->{(\textcolor{functor}{\vindexpush}\gsectionc:\dbasepointc \scarr \setc)}
+            \\
+            \onslide*<2-8|handout:2-8>{\gbasec}
+            \onslide*<9-|handout:9->{\gbasepointc}
             \onslide*<3-|handout:3->{\arrow[rr, "\vindex", color=functor]}
-            \onslide*<7-|handout:7->{\arrow[u, "\sheaf_{\gbasec}", color=sheaf]}
-            \onslide*<6,8-|handout:6,8->{\arrow[d, dashed, color=sheaf]} \&  \& 
-            \onslide*<1-|handout:1->{\dbasec}
-            \onslide*<4-6,8-|handout:4-6,8->{\arrow[d, "\sheaf_{\dbasec}", color=sheaf]}
-            \onslide*<7-|handout:7->{\arrow[u, dashed, color=sheaf]} \\
-            \onslide*<5-6,8-|handout:5-6,8->{(\textcolor{functor}{\vindexpull}\sheafc_{\dbasec})(\gbasec)} 
-            \onslide*<5-6,8-|handout:5-6,8->{\arrow[rr, "\textcolor{functor}{\vindexpull}", color=functor]}  \& \& 
-            \onslide*<4-6,8-|handout:4-6,8->{\sheafc_{\dbasec}(\dbasec)}                             
-        \end{tikzcd}
-    \end{equation*}
-\end{frame}
-
-\begin{frame}{\only<1-3:handout:1-3>{Pullback \& Pushforward are Isomorphic}\only<4-|handout:4>{Data to Graphic is a Sheaf Morphism}}
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement=\&]
-          \onslide*<1-|handout:1->{\sheafc_{\gbasec}(\gbasec)} 
+            \onslide*<7-8,10-|handout:7-8,10->{\arrow[u, "\sheaf_{\gbasec}", color=sheaf]}
+            \onslide*<6,8-9,11-|handout:6,8-9,11->{\arrow[d, dashed, color=sheaf]} \&  \& 
+            \onslide*<1-8|handout:1-8>{\dbasec}
+            \onslide*<9-|handout:9->{\dbasepointc}
+            \onslide*<4-6,8-9,11-|handout:4-6,8-9,11->{\arrow[d, "\sheaf_{\dbasec}", color=sheaf]}
+            \onslide*<7-8,10-|handout:7-8,10->{\arrow[u, dashed, color=sheaf]} \\
+            \onslide*<5-6,8|handout:5-6,8>{(\textcolor{functor}{\vindexpull}\sheafc_{\dbasec})(\gbasec)} 
+            \onslide*<9,11-|handout:9,11->{\{\textcolor{functor}{\vindexpull}\dsectionc:\gbasepointc\scarr\setc\}}
+            \onslide*<5-6,8-9,11-|handout:5-6,8-9,11->{\arrow[rr, "\textcolor{functor}{\vindexpull}", color=functor]}  \& \& 
+            \onslide*<4-6,8|handout:4-6,8>{\sheafc_{\dbasec}(\dbasec)}                      
+            \onslide*<9,11-|handout:9,11->{\{\dsectionc: \dbasepointc \scarr \setc \}}
+        \end{tikzcd}
+    \end{equation*}
+    \onslide<9,11-|handout:9,11->{
+        \begin{equation*}
+            \dsectionc^{\textcolor{functor}{*}}(\gbasepointc) = \dsectionc(\textcolor{functor}{\vindex}(\gbasepointc)) = \dsectionc(\dbasepointc)
+        \end{equation*}
+    }
+    \onslide<10-|handout:10,11->{
+        \begin{equation*}
+            \gsectionc(\gbasepointc) = \gsectionc_{\textcolor{functor}{*}}(\dbasepointc) = \gsectionc(\textcolor{functor}{\vindexpre}(\dbasepointc))
+        \end{equation*}
+    }
+\end{frame}
+
+\begin{frame}{
+    \only<1-3|handout:1-3>{Pullback \& Pushforward are Isomorphic}
+    \only<4|handout:4>{$\vartistc:\sheafc_{\dbasec}\rightarrow \sheafc_{\gbasec}$}
+    \only<5-|handout:5->{$\vartistc: \dsectionc \mapsto \gsectionc$}
+    }
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement=\&]
+          \onslide*<1-4|handout:1-4>{\sheafc_{\gbasec}(\gbasec)} 
+          \onslide*<5-|handout:5->{\{\gsectionc:\gbasepointc \scarr \setc\}}
           \onslide*<1-3|handout:1-3>{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]}  
@@
           \& \& 
-          \onslide*<1-|handout:1->{(\textcolor{functor}{\vindexpush}\sheafc_{\gbasec})(\dbase)} \\
+          \onslide*<1-4|handout:1-4>{(\textcolor{functor}{\vindexpush}\sheafc_{\gbasec})(\dbase)}
+          \onslide*<5-|handout:5>{\{\vindexpushc \gsectionc: \dbasepointc \scarr\setc\}}
+          \\
           \& \& \\
-          \onslide*<1-|handout:1->{(\textcolor{functor}{\vindexpull}\sheafc_{\dbase})(\gbasec)}
+          \onslide*<1-4|handout:1-4>{(\textcolor{functor}{\vindexpull}\sheafc_{\dbase})(\gbasec)}
+          \onslide*<5-|handout:5>{\{ \vindexpullc\dsectionc:\gbasepointc\scarr \setc\}}
           \onslide*<1-3|handout:1-3>{\arrow[rr, "\textcolor{functor}{\vindexpull}", color=functor]}
@@
           \&  \& 
-          \onslide*<1-|handout:1->{\sheafc_{\dbase}(\dbase)} 
+          \onslide*<1-4|handout:1-4>{\sheafc_{\dbase}(\dbase)} 
+          \onslide*<5-|handout:5->{\{\dsectionc:\dbasepointc \scarr \setc\}}
           \onslide*<2-3|handout:2-3>{\arrow[uu, "\textcolor{homset}{hom_{\dbasec}}"', color=homset]}
@@
         \begin{tikzcd}[column sep=huge, ampersand replacement = \&]
-            \textcolor{target}{F(c)} \arrow[dd, "F(f)"', color=target] \arrow[rr, "\alpha_{c}", bend left, color=nattran] \& \textcolor{source}{c} \arrow[dd, "f" description, color=source] \arrow[r, "G"', color=functor] \arrow[l, "F", color=functor] \& \textcolor{target}{G(c)} \arrow[dd, "G(f)", color=target] \\
-            {} \arrow[rr, dotted, color=nattran] \& \& {}                      \\
-            \textcolor{target}{F(c^{\prime})} \arrow[rr, "\alpha_{c^{\prime}}"', bend right, color=nattran] \& \textcolor{source}{c^{\prime}} \arrow[l, "F"', color=functor] \arrow[r, "G", color=functor] \& \textcolor{target}{G(c^{\prime})}
-        \end{tikzcd}
-    \end{equation*}    
-    \end{frame}
-
+            \onslide*<2-|handout:2->{\textcolor{target}{F(c)}}
+            \onslide*<2-|handout:2->{\arrow[dd, "F(f)"', color=target]}
+            \onslide*<4-5|handout:4-5>{\arrow[rr, "\alpha_{c}", bend left, color=nattran]}  
+            \onslide*<6|handout:6>{\arrow[rr, "\alpha_{c}", color=nattran]} \& 
+            \onslide*<1-4|handout:1-4>{\textcolor{source}{c}} 
+            \onslide*<1-4|handout:1-4>{\arrow[dd, "f" description, color=source]}
+            \onslide*<3-4|handout:3-4>{\arrow[r, "G"', color=functor]}
+            \onslide*<2-4|handout:2-4>{\arrow[l, "F", color=functor]} \& 
+            \onslide*<3-|handout:3->{\textcolor{target}{G(c)}} 
+            \onslide*<3-|handout:3->{\arrow[dd, "G(f)", color=target]} \\
+            {} 
+            \onslide*<4|handout:4>{\arrow[rr, dotted, color=nattran]}
+            \onslide*<5|handout:5>{\arrow[rr, color=nattran]} \& \& {}                      \\
+            \onslide*<2-|handout:2->{\textcolor{target}{F(c^{\prime})}} 
+            \onslide*<4-5|handout:4-5>{\arrow[rr, "\alpha_{c^{\prime}}"', bend right, color=nattran]} 
+            \onslide*<4-5|handout:6>{\arrow[rr, "\alpha_{c^{\prime}}"', color=nattran]}\& 
+            \onslide*<1-4|handout:1-4>{\textcolor{source}{c^{\prime}}}
+            \onslide*<2-4|handout:2-4>{\arrow[l, "F"', color=functor]} 
+            \onslide*<3-4|handout:3-4>{\arrow[r, "G", color=functor]} \& 
+            \onslide*<3-|handout:3->{\textcolor{target}{G(c^{\prime})}}          
+        \end{tikzcd}
+    \end{equation*}    
+\end{frame}
+
+\begin{frame}{$\vartist_{\vindexpullc}: \vindexpullc\sheafc_{\dbasec} \rightarrow \sheafc_{\gbasec}$}
+
+\end{frame}
 \begin{comment}

### 2021-12-31 slides/main.tex
artist nt + category sheaves
page #s

@@
 %\documentclass[xcolor={dvipsnames}]{beamer}
+\setbeamertemplate{footline}[frame number]
 %\usepackage{amsmath,amsfonts,amssymb,pxfonts,eulervm,xspace}
@@
                 \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-                  \onslide*<2-|handout:2->{\dfiberc} \onslide*<4-|handout:4->{\arrow[r, hook]} \& \onslide*<3|handout:3->{\dtotalc} \onslide*<4-|handout:4->{\arrow[d, "\pi"']} \\
-                   \& \onslide*<1-|handout:1->{\dbasec} \onslide*<5-|handout:5->{\arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]}
+                  \onslide*<2-|handout:1>{\dfiberc} 
+                  \onslide*<4-|handout:1>{\arrow[r, hook]} \& 
+                  \onslide*<3|handout:1>{\dtotalc} 
+                  \onslide*<4-|handout:1>{\arrow[d, "\pi"']} \\
+                   \& \onslide*<1-|handout:1>{\dbasec} 
+                   \onslide*<5-|handout:1>{\arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]}
                 \end{tikzcd}
@@
         \begin{description}
-            \item<1-|handout:1->[\textcolor{base}{Base Space}] $( \dbasec, \mathcal{T})$, $\dbasepointc \in \opensetc \subset \dbasec$
-            \item<2-|handout:2->[\textcolor{fiber}{Fiber Space}] $\dfiberc_{\dbasepointc} = \pi^{-1}(\dbasepointc)$
-            \item<3-|handout:3->[\textcolor{total}{Total Space}]  $\dtotalc\restriction_{\opensetc} = \dbasec\restriction_{\opensetc} \times \dfiberc\restriction_{\opensetc}$ \\ for all $\opensetc \subset K$
+            \item<1-|handout:1>[\textcolor{base}{Base Space}] $( \dbasec, \mathcal{T})$, $\dbasepointc \in \opensetc \subset \dbasec$
+            \item<2-|handout:1>[\textcolor{fiber}{Fiber Space}] $\dfiberc_{\dbasepointc} = \pi^{-1}(\dbasepointc)$
+            \item<3-|handout:1>[\textcolor{total}{Total Space}]  $\dtotalc\restriction_{\opensetc} = \dbasec\restriction_{\opensetc} \times \dfiberc\restriction_{\opensetc}$ \\ for all $\opensetc \subset K$
         \end{description}
@@
     \pause
-    \begin{alertblock}<5-|handout:5->{Sections}
+    \begin{alertblock}<5-|handout:1>{Sections}
     $\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\openset} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}$
@@
     \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-        \onslide*<8-|handout:8->{\dfiberc}
-        \onslide*<8-|handout:8->{\arrow[r, hook]} \& 
+        \onslide*<9-|handout:9->{\dfiberc}
+        \onslide*<9-|handout:9->{\arrow[r, hook]} \& 
         \onslide*<1-|handout:1->{\dtotalc}
@@
         \begin{tikzcd}[ampersand replacement = \&, column sep=small]
-            \onslide*<3-|handout:3->{{\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}} \& 
+            \onslide*<3,6-|handout:3,6->{{\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}} 
+            \onslide*<4-5|handout:4-5>{{\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}\in \setc} 
+            \&  
             {} \& 
-            \onslide*<5-|handout:5->{{\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}}} 
-            \onslide*<6-|handout:6->{\arrow[ll, "\iota^*"', hook', color=set]} \\\& \&\\
-            \onslide*<2|handout:2>{\opensetc_1 \subset \dbasec}
-            \onslide*<3-|handout:3->{\opensetc_1} 
-            \onslide*<4-|handout:4->{\arrow[uu, "\sheafc_{\dbasec}", color=sheaf]} 
-            \onslide*<6-|handout:6->{\arrow[rr, "\iota"', hook, color=base]} \&   
-            \onslide*<7-|handout:7->{\arrow[uu, "\sheafc_{\dbasec}" description, dotted, color=sheaf]} \& 
-            \onslide*<5-|handout:5->{\opensetc_2} 
-            \onslide*<5-|handout:5->{\arrow[uu, "\sheafc_{\dbasec}"', color=sheaf]}                   
+            \onslide*<6-|handout:6->{{\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}}} 
+            \onslide*<7-|handout:7->{\arrow[ll, "\iota^*"', hook', color=set]} \\\& \&\\
+            \onslide*<3|handout:3>{\opensetc_1 \subset \dbasec}
+            \onslide*<4-5|handout:4-5>{\opensetc_1 \in Ob(\mathcal{\dbasec}^{\textcolor{base}{op}})}
+            \onslide*<6-|handout:6->{\opensetc_1} 
+            \onslide*<5-|handout:5->{\arrow[uu, "\sheafc_{\dbasec}", color=sheaf]} 
+            \onslide*<7-|handout:7->{\arrow[rr, "\iota"', hook, color=base]} \&   
+            \onslide*<8-|handout:8->{\arrow[uu, "\sheafc_{\dbasec}" description, color=sheaf]} \& 
+            \onslide*<6-|handout:6->{\opensetc_2} 
+            \onslide*<6-|handout:6->{\arrow[uu, "\sheafc_{\dbasec}"', color=sheaf]}                   
             \end{tikzcd}
@@
     \begin{description}
-        \item<8-|handout:8->[stalk]$\sheafc(\dbasec)\restriction_{\dbasepointc}\coloneqq \lim\limits_{\opensetc\ni \dbasepointc} \Gamma(\opensetc, \dtotalc\restriction_{\opensetc}) \approx \dfiberc_{\dbasepointc}$
-        \item<9-|handout:9>[germ]   $\dsectionc(\dbasepointc) \in \sheafc_{\dbasec}$
+        \item<8-|handout:9->[stalk]$\sheafc(\dbasec)\restriction_{\dbasepointc}\coloneqq \lim\limits_{\opensetc\ni \dbasepointc} \Gamma(\opensetc, \dtotalc\restriction_{\opensetc}) \approx \dfiberc_{\dbasepointc}$
+        \item<9-|handout:10>[germ]   $\dsectionc(\dbasepointc) \in \sheafc_{\dbasec}$
     \end{description}
@@
 
+\begin{frame}{$\vartist_{\vindexpushc}: \sheafc_{\dbasec} \rightarrow \vindexpushc\sheafc_{\gbasec}$}
+    \begin{equation}
+      \begin{tikzcd}[ampersand replacement =\&]
+        \cgamma{\opensetc_1}{\dtotalc} 
+        \arrow[dd, "\iota^*"', hook, color=set] 
+        \arrow[rrrr, "\vartistc_{\opensetc_1}", bend left, color=artist] \&  \& 
+        \opensetc_1 
+        \arrow[ll, "\sheafc_{\dtotalc}", color=sheaf] 
+        \arrow[rr, "\sheafc_{\gtotalpushc}"', color=sheaf] \&  \& 
+        \cgamma{\opensetc_1}{\gtotalpushc} 
+        \arrow[dd, "\iota^*", hook', color=set] \\
+        {} 
+        \arrow[rrrr, dotted, color=artist] &&&& {} \\
+        \cgamma{\opensetc_2}{\dtotalc} 
+        \arrow[rrrr, "\vartistc_{\opensetc_2}"', bend right, color=artist] \&  \& 
+        \opensetc_2 
+        \arrow[ll, "\sheafc_{\dtotalc}"', color=sheaf] 
+        \arrow[rr, "\sheafc_{\gtotalpushc}", color=sheaf] 
+        \arrow[uu, "\iota" description, hook, color=base] \& \& 
+        \cgamma{\opensetc_2}{\gtotalpushc}               
+    \end{tikzcd}
+    \end{equation}
+\end{frame}
+
 \begin{frame}{$\vartist_{\vindexpullc}: \vindexpullc\sheafc_{\dbasec} \rightarrow \sheafc_{\gbasec}$}
-
-\end{frame}
+    \begin{equation}
+  \begin{tikzcd}[ampersand replacement =\&]
+    \cgamma{\opensetgc_1}{\dtotalpullc} 
+    \arrow[dd, "\iota^*"', hook, color=set] 
+    \arrow[rrrr, "\vartistc_{\opensetgc_1}", bend left, color=artist] \&  \& 
+    \opensetgc_1 
+    \arrow[ll, "\sheafc_{\dtotalpullc}", color=sheaf] 
+    \arrow[rr, "\sheafc_{\gtotalc}"', color=sheaf] \&  \& 
+    \cgamma{\opensetgc_1}{\gtotalc} 
+    \arrow[dd, "\iota^*", hook', color=set] \\
+    {} 
+    \arrow[rrrr, dotted, color=artist] &&&& {} \\
+    \cgamma{\opensetgc_2}{\dtotalpullc} 
+    \arrow[rrrr, "\vartistc_{\opensetgc_2}"', bend right, color=artist] \&  \& 
+    \opensetgc_2 
+    \arrow[ll, "\sheafc_{\dtotalpullc}"', color=sheaf] 
+    \arrow[rr, "\sheafc_{\gtotalc}", color=sheaf] 
+    \arrow[uu, "\iota" description, hook, color=base] \& \& 
+    \cgamma{\opensetgc_2}{\gtotalc}                                                        
+\end{tikzcd}
+\end{equation}
+\end{frame}
+
 \begin{comment}

### 2021-12-31 slides/main.tex
functor to mapsto

@@
 \subsection{Category Theory}
+
 \begin{frame}{Category $\textcolor{source}{\mathcal{C}}$}
@@
     \begin{tikzcd}[ampersand replacement=\&]
-        \textcolor{source}{C_1} \arrow[r, "f", color=source] \arrow[rd, "g\circ f"', color=source] \arrow["id_{C_1}"', loop, distance=2em, in=215, out=145, color=source] \& \textcolor{source}{C_2} \arrow[d, "g", color=source] \arrow["id_{C_2}"', loop, distance=2em, in=125, out=55, color=source] \\
-    \& \textcolor{source}{C_3} \arrow["id_{C_3}"', loop, distance=2em, in=305, out=235, color=source]              
+        \onslide*<1-|handout:1>{\textcolor{source}{C_1}} 
+        \onslide*<3-|handout:1>{\arrow[r, "f", color=source]}
+        \onslide*<4-|handout:1>{\arrow[rd, "g\circ f"', color=source]}
+        \onslide*<2-|handout:1>{\arrow["id_{C_1}"', loop, distance=2em, in=215, out=145, color=source]} \& 
+        \onslide*<1-|handout:1>{\textcolor{source}{C_2}}
+        \onslide*<3-|handout:1>{\arrow[d, "g", color=source]} 
+        \onslide*<2-|handout:1>{\arrow["id_{C_2}"', loop, distance=2em, in=125, out=55, color=source]} \\\& 
+        \onslide*<1-|handout:1>{\textcolor{source}{C_3}}
+        \onslide*<2-|handout:1>{\arrow["id_{C_3}"', loop, distance=2em, in=305, out=235, color=source]}              
     \end{tikzcd}
@@
     \column{0.59\textwidth}
-    \begin{alertblock}{associativity} 
+    \begin{alertblock}<5-|handout:1>{associativity} 
         if $f: C_1 \rightarrow C_2$, $g: C_2 \rightarrow C_3$ and $h: C_3 \rightarrow C_4$ then $h\circ (g \circ f) = (h \circ g) \circ f$
     \end{alertblock}
-    \begin{alertblock}{identity} 
+    \begin{alertblock}<6-|handout:1>{identity} 
         for every $f: C_1 \rightarrow C_2$ there exists identity morphisms $f \circ id_{C_1} = f = id_{C_2} \circ f$
@@
 \end{frame}
+
 \begin{frame}{Category $\mathcal{\dbasec}$}
@@
     \column{0.49\textwidth}
-    \onslide<2-|handout:2->{
+    \onslide<2-|handout:1>{
     \begin{tikzcd}[ampersand replacement=\&]
@@
         \column{0.69\textwidth}
-        \onslide<2-|handout:2->{
+        \onslide<2-|handout:1>{
             \begin{tikzcd}[ampersand replacement=\&]
@@
         \begin{tikzcd}[ampersand replacement = \&, ]
-            \onslide*<1-|handout:1->{\textcolor{source}{c}} 
-            \onslide*<1-|handout:1->{\arrow[d, "f"', shift right=0, color=source]} \onslide*<3-|handout:3->{\arrow[r, "F", color=functor]} \& 
-            \onslide*<2|handout:2>{\textcolor{target}{d}}
-            \onslide*<3-|handout:3->{\textcolor{target}{F(c)}} 
-            \onslide*<2|handout:2>{\arrow[d, shift right=0, color=target]}
-            \onslide*<3-|handout:3->{\arrow[d, "F(f)", shift right=0, color=target]} \\
-            \onslide*<1-|handout:1->{\textcolor{source}{c^{\prime}}} 
-            \onslide*<3-|handout:3->{\arrow[r, "F"', color=functor]}\& 
-            \onslide*<2|handout:2>{\textcolor{target}{d^{\prime}}}
-            \onslide*<3-|handout:3->{\textcolor{target}{F(c^{\prime})}}                      
+            \onslide*<1-|handout:1>{\textcolor{source}{c}} 
+            \onslide*<1-|handout:1>{\arrow[d, "f"', shift right=0, color=source]} 
+            \onslide*<3-|handout:1>{\arrow[r, "F", color=functor, maps to]} \& 
+            \onslide*<2-3|handout:0>{\textcolor{target}{d}}
+            \onslide*<4|handout:0>{\textcolor{target}{F(c)}} 
+            \onslide*<0|handout:1>{\textcolor{target}{F(c)=d}} 
+            \onslide*<2|handout:1>{\arrow[d, shift right=0, color=target]}
+            \onslide*<3-|handout:1>{\arrow[d, "F(f)", shift right=0, color=target]} \\
+            \onslide*<1-|handout:1>{\textcolor{source}{c^{\prime}}} 
+            \onslide*<3-|handout:1>{\arrow[r, "F"', color=functor, maps to]}\& 
+            \onslide*<2-3|handout:0>{\textcolor{target}{d^{\prime}}}
+            \onslide*<4-|handout:0>{\textcolor{target}{F(c^{\prime})}}
+            \onslide*<0|handout:1>{\textcolor{target}{F(c^{\prime})= d^{\prime}}}                    
         \end{tikzcd}
@@
     \column{.59\textwidth}
-    \begin{alertblock}<4-|handout:4->{composition}
+    \begin{alertblock}<5-|handout:1>{composition}
     $\textcolor{functor}{F}(\textcolor{source}{g}) \circ  \textcolor{functor}{F}(\textcolor{source}{f}) = \textcolor{functor}{F} (\textcolor{source}{g}\circ \textcolor{source}{f})$
     \end{alertblock}
-    \begin{alertblock}<5-|handout:5->{identity}
+    \begin{alertblock}<6-|handout:1>{identity}
     $\textcolor{functor}{F}(\textcolor{source}{id_c}) = \textcolor{target}{id}_{\textcolor{functor}{F}(\textcolor{source}{c})}$

### 2021-12-31 slides/main.tex
presheaf + functor

@@
             \end{tikzcd}
-
         }
@@
             \onslide*<0|handout:1>{\textcolor{target}{F(c)=d}} 
-            \onslide*<2|handout:1>{\arrow[d, shift right=0, color=target]}
-            \onslide*<3-|handout:1>{\arrow[d, "F(f)", shift right=0, color=target]} \\
+            \onslide*<2|handout:0>{\arrow[d, shift right=0, color=target]}
+            \onslide*<3-|handout:0>{\arrow[d, "F(f)", shift right=0, color=target]}
+            \onslide*<0|handout:1>{\arrow[d, "F(f)", shift right=3, color=target]} \\
             \onslide*<1-|handout:1>{\textcolor{source}{c^{\prime}}} 
@@
     \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-        \onslide*<9-|handout:9->{\dfiberc}
-        \onslide*<9-|handout:9->{\arrow[r, hook]} \& 
+        \onslide*<9-|handout:1->{\dfiberc}
+        \onslide*<9-|handout:1->{\arrow[r, hook]} \& 
         \onslide*<1-|handout:1->{\dtotalc}
@@
         \onslide*<1-|handout:1->{\dbasec}
-        \onslide*<3-|handout:3->{\arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]}
+        \onslide*<3-|handout:1->{\arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]}
         \end{tikzcd}
@@
         \begin{tikzcd}[ampersand replacement = \&, column sep=small]
-            \onslide*<3,6-|handout:3,6->{{\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}} 
-            \onslide*<4-5|handout:4-5>{{\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}\in \setc} 
+            \onslide*<3,6-|handout:0>{{\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}} 
+            \onslide*<4-5|handout:0>{{\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}\in \setc}
+            \onslide*<0|handout:1>{\setc \ni {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}}
             \&  
             {} \& 
-            \onslide*<6-|handout:6->{{\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}}} 
-            \onslide*<7-|handout:7->{\arrow[ll, "\iota^*"', hook', color=set]} \\\& \&\\
-            \onslide*<3|handout:3>{\opensetc_1 \subset \dbasec}
-            \onslide*<4-5|handout:4-5>{\opensetc_1 \in Ob(\mathcal{\dbasec}^{\textcolor{base}{op}})}
-            \onslide*<6-|handout:6->{\opensetc_1} 
-            \onslide*<5-|handout:5->{\arrow[uu, "\sheafc_{\dbasec}", color=sheaf]} 
-            \onslide*<7-|handout:7->{\arrow[rr, "\iota"', hook, color=base]} \&   
-            \onslide*<8-|handout:8->{\arrow[uu, "\sheafc_{\dbasec}" description, color=sheaf]} \& 
-            \onslide*<6-|handout:6->{\opensetc_2} 
-            \onslide*<6-|handout:6->{\arrow[uu, "\sheafc_{\dbasec}"', color=sheaf]}                   
+            \onslide*<6-|handout:1>{{\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}}} 
+            \onslide*<7-|handout:1>{\arrow[ll, "\iota^*"', hook', color=set]} \\\& \&\\
+            \onslide*<3|handout:0>{\opensetc_1 \subset \dbasec}
+            \onslide*<4-5|handout:0>{\opensetc_1 \in Ob(\mathcal{\dbasec}^{\textcolor{base}{op}})}
+            \onslide*<6-|handout:0>{\opensetc_1}
+            \onslide*<0|handout:1>{Ob(\mathcal{\dbasec}^{\textcolor{base}{op}}) \ni \opensetc_1} 
+            \onslide*<5-|handout:0>{\arrow[uu, "\sheafc_{\dbasec}", maps to, color=sheaf]}
+            \onslide*<0|handout:1>{\arrow[uu, "\sheafc_{\dbasec}", maps to, color=sheaf, shift right = 10]}
+            \onslide*<7-|handout:1>{\arrow[rr, "\iota"', hook, color=base]} \&   
+            \onslide*<8-|handout:1>{\arrow[uu, "\sheafc_{\dbasec}" description, maps to, color=sheaf]} \& 
+            \onslide*<6-|handout:1>{\opensetc_2} 
+            \onslide*<6-|handout:1>{\arrow[uu, "\sheafc_{\dbasec}"', maps to, color=sheaf]}                   
             \end{tikzcd}
@@
     \begin{description}
-        \item<8-|handout:9->[stalk]$\sheafc(\dbasec)\restriction_{\dbasepointc}\coloneqq \lim\limits_{\opensetc\ni \dbasepointc} \Gamma(\opensetc, \dtotalc\restriction_{\opensetc}) \approx \dfiberc_{\dbasepointc}$
-        \item<9-|handout:10>[germ]   $\dsectionc(\dbasepointc) \in \sheafc_{\dbasec}$
+        \item<8-|handout:1>[stalk]$\sheafc(\dbasec)\restriction_{\dbasepointc}\coloneqq \lim\limits_{\opensetc\ni \dbasepointc} \Gamma(\opensetc, \dtotalc\restriction_{\opensetc}) \approx \dfiberc_{\dbasepointc}$
+        \item<9-|handout:1>[germ]   $\dsectionc(\dbasepointc) \in \sheafc_{\dbasec}$
     \end{description}

### 2022-01-06 slides/main.tex
reworking functor
sheaf reworking

@@
                    framesep=3mm}
+
 \usepackage{notation} % move this later
@@
 
-\begin{frame}{Fiber Bundles}
+\begin{frame}
+    \frametitle<1|handout:1>{Fiber Bundle}
+    \frametitle<6|handout:2>{Data Bundle}
+    \frametitle<7|handout:3>{Graphic Bundle}
     \begin{columns}
@@
                 \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-                  \onslide*<2-|handout:1>{\dfiberc} 
-                  \onslide*<4-|handout:1>{\arrow[r, hook]} \& 
-                  \onslide*<3|handout:1>{\dtotalc} 
-                  \onslide*<4-|handout:1>{\arrow[d, "\pi"']} \\
-                   \& \onslide*<1-|handout:1>{\dbasec} 
-                   \onslide*<5-|handout:1>{\arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]}
+                  \onslide*<2-6|handout:1-2>{\dfiberc}
+                  \onslide*<7|handout:3>{\gfiberc} 
+                  \onslide*<4-|handout:1->{\arrow[r, hook]} \& 
+                  \onslide*<3-6|handout:1-2>{\dtotalc} 
+                  \onslide*<7|handout:3>{\gtotalc}
+                  \onslide*<4-|handout:1->{\arrow[d, "\pi"']} \\
+                   \& 
+                   \onslide*<1-6|handout:1-2>{\dbasec}
+                   \onslide*<7|handout:3>{\gbasec} 
+                   \onslide*<5-6|handout:1-2>{\arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]}
+                   \onslide*<7|handout:3>{\arrow[u, "\gsectionc \in \cgamma{\gbasec}{\gtotalc}"', bend right, pos=.5]}
                 \end{tikzcd}
@@
      \column{0.59\textwidth}
+     \only<1-5|handout:1>{
         \begin{description}
-            \item<1-|handout:1>[\textcolor{base}{Base Space}] $( \dbasec, \mathcal{T})$, $\dbasepointc \in \opensetc \subset \dbasec$
-            \item<2-|handout:1>[\textcolor{fiber}{Fiber Space}] $\dfiberc_{\dbasepointc} = \pi^{-1}(\dbasepointc)$
-            \item<3-|handout:1>[\textcolor{total}{Total Space}]  $\dtotalc\restriction_{\opensetc} = \dbasec\restriction_{\opensetc} \times \dfiberc\restriction_{\opensetc}$ \\ for all $\opensetc \subset K$
+            \item<1-5|handout:1>[\textcolor{base}{Base Space}]{$( \dbasec, \mathcal{T})$, $\dbasepointc \in \opensetc \subset \dbasec$}
+            \item<2-5|handout:1>[\textcolor{fiber}{Fiber Space}]{$\dfiberc\restriction_{\dbasepointc} = \pi^{-1}(\dbasepointc)$}
+            \item<3-5|handout:1>[\textcolor{total}{Total Space}]{ $\dtotalc\restriction_{\opensetc} = \dbasec\restriction_{\opensetc} \times \dfiberc\restriction_{\opensetc}$ \\ for all $\opensetc \subset K$}
+        \end{description}}
+    \only<6|handout:2>{
+        \begin{description}
+            \item[\textcolor{base}{Continuity Type \dbasec}]{how data elements are organized (topological properties) \cite{wilkinsonGrammarGraphics2005})}
+            \item[\textcolor{fiber}{Fields Type \dfiberc}] {generalization of a schema - named and typed date fields \cite{spivakSIMPLICIALDATABASES, spivakDatabasesAreCategories2010}}
+            \item[\textcolor{total}{Data Type \dtotalc}]{continuity type + field type}
         \end{description}
+        }
+        \only<7|handout:3>{
+            \begin{description}
+                \item[\textcolor{base}{Continuity Type \gbasec}]{how pixels are organized in the display space}
+                \item[\textcolor{fiber}{Display Type \gfiberc}] {display specifications}
+                \item[\textcolor{total}{Graphic Type \gtotalc}]{continuity type + display type}
+            \end{description}
+            }
     \end{columns}
     \pause
-    \begin{alertblock}<5-|handout:1>{Sections}
-    $\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\openset} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}$
+    \only<5|handout:1>{
+        \begin{alertblock}{Sections}
+            \begin{aligned*}         
+        \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}
+        \end{aligned*}
         \end{alertblock}
+    }
+    \only<6|handout:2>{
+    \begin{alertblock}{Data \cite{butlerVisualizationModelBased1989,butlerVectorBundleClassesForm1992}}
+        \begin{aligned*}
+        \dsectionc(\dbasepointc) = &\delementc,\; \dbasepointc\in \opensetc\subset \dbasec,\; \delement \in \dfiberc\restriction_{\dbasepointc}\\
+        &\delementc=\{field_0:value, \cdots field_i:value, \cdots field_n:value\}
+        \end{aligned*}
+    \end{alertblock}
+    }
+    \only<7|handout:3>{
+        \begin{alertblock}{Graphic}
+            \begin{aligned*}         
+            \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \coloneqq \big\{\gsectionc: \opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} \; \bigm{\vert} \pi(\gsectionc(\gbasepointc)) = \gbasepointc\;for\, all\; \gbasepointc \in \opensetgc \big\}\\
+            \vspace{5pt}\\
+            \gsectionc(\gbasepointc) = d,\; \gbasepointc\in \opensetgc\subset \gbasec,\; d \in \dfiberc\restriction_{\gbasepointc}\\
+            for\;example:\; d=\{x,\,y,\,r,\,g,\,b\}
+        \end{aligned*}
+    \end{alertblock}
+    }
 \end{frame}
@@
 \subsection{Category Theory}
+\begin{frame}[fragile]{}
+    \begin{minted}{python}
+        Artist(data:Data) -> Graphic
+    \end{minted}
+\end{frame}
 
@@
             \cgamma{\opensetc_1}{\dtotalc\restriction_{\opensetc_1}} \&  \& \cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}} \arrow[ll, "\iota^*\restriction_{\opensetc_1}"', color=set]\\
-            \&\&\\\&  \& \cgamma{\opensetc_3}{\dtotalc\restriction_{\opensetc_3}} \arrow[uu, "\iota^*\restriction_{\opensetc_2}"', color=set] \arrow[lluu, "\iota^*\restriction_{\opensetc_2} \circ \iota^*\restriction_{\opensetc_1}", color=set]
+            \&\&\\\&  \& \cgamma{\opensetc_3}{\dtotalc\restriction_{\opensetc_3}} \arrow[uu, "\iota^*\restriction_{\opensetc_2}"', color=set] \arrow[lluu, "\iota^*\restriction_{\opensetc_2} \circ \textcolor{set}{\iota^*\restriction}_{\opensetc_1}", color=set]
             \end{tikzcd}
@@
             \onslide*<2-3|handout:0>{\textcolor{target}{d}}
-            \onslide*<4|handout:0>{\textcolor{target}{F(c)}} 
+            \onslide*<4-|handout:0>{\textcolor{target}{F(c)}} 
             \onslide*<0|handout:1>{\textcolor{target}{F(c)=d}} 
@@
     \begin{alertblock}<5-|handout:1>{composition}
-    $\textcolor{functor}{F}(\textcolor{source}{g}) \circ  \textcolor{functor}{F}(\textcolor{source}{f}) = \textcolor{functor}{F} (\textcolor{source}{g}\circ \textcolor{source}{f})$
+        \begin{aligned*}
+            \textcolor{functor}{F}(\textcolor{source}{g}) \circ  \textcolor{functor}{F}(\textcolor{source}{f}) = \textcolor{functor}{F} (\textcolor{source}{g}\circ \textcolor{source}{f})
+        \end{aligned*}
     \end{alertblock}
     \begin{alertblock}<6-|handout:1>{identity}
-    $\textcolor{functor}{F}(\textcolor{source}{id_c}) = \textcolor{target}{id}_{\textcolor{functor}{F}(\textcolor{source}{c})}$
+        \begin{aligned*}
+            \textcolor{functor}{F}(\textcolor{source}{id_c}) = \textcolor{target}{id}_{\textcolor{functor}{F}(\textcolor{source}{c})}
+        \end{aligned*}
     \end{alertblock}
@@
             \onslide*<0|handout:1>{Ob(\mathcal{\dbasec}^{\textcolor{base}{op}}) \ni \opensetc_1} 
-            \onslide*<5-|handout:0>{\arrow[uu, "\sheafc_{\dbasec}", maps to, color=sheaf]}
-            \onslide*<0|handout:1>{\arrow[uu, "\sheafc_{\dbasec}", maps to, color=sheaf, shift right = 10]}
+            \onslide*<5-|handout:0>{\arrow[uu, "\sheafc_{\dbasec, \dtotalc}", maps to, color=sheaf]}
+            \onslide*<0|handout:1>{\arrow[uu, "\sheafc_{\dbasec, \dtotalc}", maps to, color=sheaf, shift right = 10]}
             \onslide*<7-|handout:1>{\arrow[rr, "\iota"', hook, color=base]} \&   
-            \onslide*<8-|handout:1>{\arrow[uu, "\sheafc_{\dbasec}" description, maps to, color=sheaf]} \& 
+            \onslide*<8-|handout:1>{\arrow[uu, "\sheafc_{\dbasec, \dtotalc}" description, maps to, color=sheaf]} \& 
             \onslide*<6-|handout:1>{\opensetc_2} 
-            \onslide*<6-|handout:1>{\arrow[uu, "\sheafc_{\dbasec}"', maps to, color=sheaf]}                   
+            \onslide*<6-|handout:1>{\arrow[uu, "\sheafc_{\dbasec, \dtotalc}"', maps to, color=sheaf]}                   
             \end{tikzcd}
     \end{columns}
-    \begin{description}
-        \item<8-|handout:1>[stalk]$\sheafc(\dbasec)\restriction_{\dbasepointc}\coloneqq \lim\limits_{\opensetc\ni \dbasepointc} \Gamma(\opensetc, \dtotalc\restriction_{\opensetc}) \approx \dfiberc_{\dbasepointc}$
-        \item<9-|handout:1>[germ]   $\dsectionc(\dbasepointc) \in \sheafc_{\dbasec}$
-    \end{description}
+    \begin{center}
+        \begin{alertblock}<8-|handout:1>{stalk}
+            \begin{aligned*}
+                \sheafc_{\dbasec, \dtotalc}\restriction_{\dbasepointc}\coloneqq \lim\limits_{\opensetc\ni \dbasepointc} \Gamma(\opensetc, \dtotalc\restriction_{\opensetc}) \approx \dfiberc_{\dbasepointc}
+            \end{aligned*}
+        \end{alertblock}
+
+        \begin{alertblock}<9-|handout:1>{germ}
+            \begin{aligned*}
+                \dsectionc(\dbasepointc) \in \sheafc_{\dbasec, \dtotalc}\restriction_{\dbasepointc}
+            \end{aligned*}
+        \end{alertblock}
+    \end{center}
 \end{frame}
@@
 \begin{frame}{Sheaves}
-    A sheaf is a presheaf that satisfies the following two axioms
-
-    \onslide<2-|handout:2->{Given open cover $\mathcal{\opensetc} = \{\opensetc_i\}_{i\in I}$ of $\opensetc$:}
-    \begin{block}<3-|handout:3->{locality}
-    $\dsectionc^{1}, \dsectionc^{2} \in \sheafc(\opensetc_i)\\ \dsectionc^{1}\restriction_{\opensetc_i} = \dsectionc^{2}\restriction_{\opensetc_i} \implies \dsectionc^{1} = \dsectionc^{2}$
+    A sheaf is a presheaf that satisfies the following two axioms\cite{bakerMathsSheaf}\\
+    \begin{block}<2-|handout:1>{locality}
+        if $\mathcal{\opensetc} = \{\opensetc_i\}_{i\in I}$ is an open covering of \opensetc
+        \onslide<3-|handout:1>{
+        and if $\dsectionc^{a}, \dsectionc^{b} \in \sheafc(\opensetc_i)$ have the property $\dsectionc^{a}\restriction_{\opensetc_i} = \dsectionc^{b}\restriction_{\opensetc_i}$ for each $\opensetc_i \in \mathcal{\opensetc}$,}
+        \onslide<4|handout:1>{ then $\dsectionc^{a} = \dsectionc^{b}$}
     \end{block} 
-    \begin{block}<4-|handout:4>{gluing}
-    $\dsectionc_{i} \in \sheafc(\opensetc_i), \dsectionc_{j} \in \sheafc(\opensetc_j), \opensetc_i, \opensetc_j \in \mathcal{U}\\ \dsectionc_{i}\restriction_{\opensetc_i\cap \opensetc_j} = \dsectionc_{j}\restriction_{\opensetc_i \cap \opensetc_j} \implies \dsectionc\restriction_{\opensetc_i} = \dsectionc_i$
+    \begin{block}<5-|handout:1>{gluing}
+        if $\mathcal{\opensetc} = \{\opensetc_i\}_{i\in I}$ is an open covering of \opensetc\ and if $\dsectionc^{i} \in \sheafc(\opensetc_i)$ is given for each $i \in I$
+        \onslide<6-|handout:1>{such that ${\dsectionc^{i}}\restriction_{\opensetc_i\cap \opensetc_j} = {\dsectionc^{j}}\restriction_{\opensetc_i \cap \opensetc_j}$ for each pair $\opensetc_i, \opensetc_j \in \mathcal{\opensetc}$,}
+        \onslide<7-|handout:1>{ then there exists $\dsectionc \in \sheafc(\opensetc)$ such that $\dsectionc\restriction_{\opensetc_i} = {\dsectionc^{i}}$ for each $i \in I$} 
     \end{block}      
@@
 
-\begin{frame}{
-\only<-8|handout:-8>{Morphism of sheaves:}
-\only<5-6|handout:5-6>{pullback}
-\only<7|handout:7>{pushforward} 
-\only<8,11-|handout:8,11->{pullback \& pushforward}
-\only<9|handout:9>{pullback sections}
-\only<10|handout:10>{pushforward sections}
+\begin{frame}
+    \frametitle<0|handout:1>{Functor: \vindexc: \mathcal{\gbasec} \rightarrow \mathcal{\dbasec}}
+    \frametitle<0|handout:2>{Data}
+    \frametitle<0|handout:3>{Pullback: data to region of the visualization} 
+    \frametitle<0|handout:4>{Graphic}
+    \frametitle<0|handout:5>{Pushforward: visualization to index of data}
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
+            \onslide*<7-8|handout:4>{\sheafc_{\gbasec, \gtotalc}} 
+            \onslide*<7-8,10-|handout:5>{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]} \& \& 
+            \onslide*<7-8|handout:5>{\textcolor{functor}{\vindexpush}\sheafc_{\gbasec, \gtotalc}}  
+            \\
+            \onslide*<2-|handout:1->{\gbasec}
+            \onslide*<3-|handout:1->{\arrow[rr, "\vindex", color=functor]}
+            \onslide*<7-8,10-|handout:4>{\arrow[u, "\sheaf_{\gbasec, \gtotalc}", maps to, color=sheaf]}
+            \onslide*<6,8-9,11-|handout:3>{\arrow[d, dashed, maps to, color=sheaf]} \&  \& 
+            \onslide*<1-|handout:1->{\dbasec}
+            \onslide*<4-6,8-9,11-|handout:2-3>{\arrow[d, "\sheaf_{\dbasec, \dtotalc}", maps to, color=sheaf]}
+            \onslide*<7-8,10-|handout:5>{\arrow[u, dashed, maps to, color=sheaf]} \\
+            \onslide*<5-6,8|handout:3>{\textcolor{functor}{\vindexpull}\sheafc_{\dbasec, \dtotalc}} 
+            \onslide*<5-6,8-9,11-|handout:3>{\arrow[rr, "\textcolor{functor}{\vindexpull}", color=functor]}  \& \& 
+            \onslide*<4-6,8|handout:2-3>{\sheafc_{\dbasec, \dtotalc}}
+        \end{tikzcd}
+    \end{equation*}
+    \only<4|handout:2>{
+        \begin{itemize}
+            \item $\dfiberc \hookrightarrow \dtotalc \xrightarrow{\pi} \dbasec$
+            \item $\sheafc_{\dbasec, \dtotalc}:\opensetc \mapsto \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}, \opensetc \subset \dbasec$
+            \item $\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \ni \dsectionc: \opensetc \rightarrow \dfiberc\restriction_{\opensetc}$
+            \item $\dsectionc(\dbasepointc) = \{f_0:v_0, \cdots,\}, \dbasepointc \in \opensetc$
+        \end{itemize}
     }
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-            \onslide*<7-8|handout:7-8>{\sheafc_{\gbasec}(\gbasec)} 
-            \onslide*<10-|handout:10->{\{\gsectionc: \gbasepointc \scarr \setc \}}
-            \onslide*<7-8,10-|handout:7-8,10->{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]} \& \& 
-            \onslide*<7-8|handout:7-8>{(\textcolor{functor}{\vindexpush}\sheafc_{\gbasec})(\dbasec)}  
-            \onslide*<10-|handout:10->{(\textcolor{functor}{\vindexpush}\gsectionc:\dbasepointc \scarr \setc)}
-            \\
-            \onslide*<2-8|handout:2-8>{\gbasec}
-            \onslide*<9-|handout:9->{\gbasepointc}
-            \onslide*<3-|handout:3->{\arrow[rr, "\vindex", color=functor]}
-            \onslide*<7-8,10-|handout:7-8,10->{\arrow[u, "\sheaf_{\gbasec}", color=sheaf]}
-            \onslide*<6,8-9,11-|handout:6,8-9,11->{\arrow[d, dashed, color=sheaf]} \&  \& 
-            \onslide*<1-8|handout:1-8>{\dbasec}
-            \onslide*<9-|handout:9->{\dbasepointc}
-            \onslide*<4-6,8-9,11-|handout:4-6,8-9,11->{\arrow[d, "\sheaf_{\dbasec}", color=sheaf]}
-            \onslide*<7-8,10-|handout:7-8,10->{\arrow[u, dashed, color=sheaf]} \\
-            \onslide*<5-6,8|handout:5-6,8>{(\textcolor{functor}{\vindexpull}\sheafc_{\dbasec})(\gbasec)} 
-            \onslide*<9,11-|handout:9,11->{\{\textcolor{functor}{\vindexpull}\dsectionc:\gbasepointc\scarr\setc\}}
-            \onslide*<5-6,8-9,11-|handout:5-6,8-9,11->{\arrow[rr, "\textcolor{functor}{\vindexpull}", color=functor]}  \& \& 
-            \onslide*<4-6,8|handout:4-6,8>{\sheafc_{\dbasec}(\dbasec)}                      
-            \onslide*<9,11-|handout:9,11->{\{\dsectionc: \dbasepointc \scarr \setc \}}
-        \end{tikzcd}
-    \end{equation*}
-    \onslide<9,11-|handout:9,11->{
-        \begin{equation*}
-            \dsectionc^{\textcolor{functor}{*}}(\gbasepointc) = \dsectionc(\textcolor{functor}{\vindex}(\gbasepointc)) = \dsectionc(\dbasepointc)
-        \end{equation*}
+    \only<6|handout:3>{
+        \begin{itemize}
+            \item $\vindexpullc\dfiberc \hookrightarrow \vindexpullc\dtotalc \xrightarrow{\pi} \gbasec$
+            \item $\vindexpullc\sheafc_{\dbasec,\dtotalc}:\opensetgc \mapsto \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}, \opensetgc \subset \gbasec$
+            \item $\cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} \ni \vindexpullc\dsectionc: \opensetgc \rightarrow \vindexpullc\dfiberc\restriction_{\opensetgc}$
+            \item $\vindexpullc\dsectionc(\gbasepointc) = \dsectionc(\vindexc(\gbasepointc)) = \dsectionc(\dbasepointc)$
+        \end{itemize}
     }
-    \onslide<10-|handout:10,11->{
-        \begin{equation*}
-            \gsectionc(\gbasepointc) = \gsectionc_{\textcolor{functor}{*}}(\dbasepointc) = \gsectionc(\textcolor{functor}{\vindexpre}(\dbasepointc))
-        \end{equation*}
+    \only<7|handout:4>{
+        \begin{itemize}
+            \item $\gfiberc \hookrightarrow \gtotalc \xrightarrow{\pi} \gbasec$
+            \item $\sheafc_{\gbasec, \gtotalc}:\opensetgc \mapsto \cgamma{\opensetgc}{\dtotalc\restriction_{\opensetgc}}, \opensetgc \subset \gbasec$
+            \item $\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \ni \gsectionc: \opensetgc \rightarrow \gfiberc\restriction_{\opensetgc}$
+            \item $\gsectionc(\gbasepointc) = \{d_0, \cdots,\}, \dbasepointc \in \opensetfc$
+        \end{itemize}
     }
+
 \end{frame}
@@
     \only<1-3|handout:1-3>{Pullback \& Pushforward are Isomorphic}
-    \only<4|handout:4>{$\vartistc:\sheafc_{\dbasec}\rightarrow \sheafc_{\gbasec}$}
+    \only<4|handout:4>{$\vartistc:\sheafc_{(\dbasec, \dtotalc)}\rightarrow \sheafc_{(\gbasec, \gtotalc)}$}
     \only<5-|handout:5->{$\vartistc: \dsectionc \mapsto \gsectionc$}
@@
         \begin{tikzcd}[ampersand replacement=\&]
-          \onslide*<1-4|handout:1-4>{\sheafc_{\gbasec}(\gbasec)} 
+          \onslide*<1-4|handout:1-4>{\sheafc_{(\gbasec, \gtotalc)}(\gbasec)} 
           \onslide*<5-|handout:5->{\{\gsectionc:\gbasepointc \scarr \setc\}}
@@
           \& \& 
-          \onslide*<1-4|handout:1-4>{(\textcolor{functor}{\vindexpush}\sheafc_{\gbasec})(\dbase)}
+          \onslide*<1-4|handout:1-4>{(\textcolor{functor}{\vindexpush}\sheafc_{(\gbasec, \gtotalc)})(\dbase)}
           \onslide*<5-|handout:5>{\{\vindexpushc \gsectionc: \dbasepointc \scarr\setc\}}
@@
           \& \& \\
-          \onslide*<1-4|handout:1-4>{(\textcolor{functor}{\vindexpull}\sheafc_{\dbase})(\gbasec)}
+          \onslide*<1-4|handout:1-4>{(\textcolor{functor}{\vindexpull}\sheafc_{(\dbasec, \dtotalc)})(\gbasec)}
           \onslide*<5-|handout:5>{\{ \vindexpullc\dsectionc:\gbasepointc\scarr \setc\}}
@@
           \&  \& 
-          \onslide*<1-4|handout:1-4>{\sheafc_{\dbase}(\dbase)} 
+          \onslide*<1-4|handout:1-4>{\sheafc_{(\dbasec, \dtotalc)}(\dbase)} 
           \onslide*<5-|handout:5->{\{\dsectionc:\dbasepointc \scarr \setc\}}
@@
         \begin{equation*}
-        \textcolor{homset}{hom}_{\gbasec}(\textcolor{functor}{\vindexpull}\sheafc_{\dbase}, \sheafc_{\gbasec}) = 
-    \textcolor{homset}{hom}_{\dbase}(\sheafc_{\dbase}, \textcolor{functor}{\vindexpush}\sheafc_{\gbasec})
+        \textcolor{homset}{hom}_{\gbasec}(\textcolor{functor}{\vindexpull}\sheafc_{(\dbasec, \dtotalc)}, \sheafc_{(\gbasec, \gtotalc)}) = 
+    \textcolor{homset}{hom}_{\dbase}(\sheafc_{(\dbasec, \dtotalc)}, \textcolor{functor}{\vindexpush}\sheafc_{(\gbasec, \gtotalc)})
         \end{equation*}
@@
 
-\begin{frame}{$\vartist_{\vindexpushc}: \sheafc_{\dbasec} \rightarrow \vindexpushc\sheafc_{\gbasec}$}
+\begin{frame}{$\vartist_{\vindexpushc}: \sheafc_{(\dbasec, \dtotalc)} \rightarrow \vindexpushc\sheafc_{(\gbasec, \gtotalc)}$}
     \begin{equation}
@@
 
-\begin{frame}{$\vartist_{\vindexpullc}: \vindexpullc\sheafc_{\dbasec} \rightarrow \sheafc_{\gbasec}$}
+\begin{frame}{$\vartist_{\vindexpullc}: \vindexpullc\sheafc_{(\dbasec, \dtotalc)} \rightarrow \sheafc_{(\gbasec, \gtotalc)}$}
     \begin{equation}

### 2022-01-06 slides/scraps.tex
sheaf reworking

\subsection{Data}

\begin{frame}{Visualizations are (mostly) evaluated on equivariance}
    \begin{description}
        \item[Expressiveness] structure preserving mappings from data to graphic (Mackinlay \cite{mackinlayAutomatingDesignGraphical1986})
        \item[Effectiveness] design choices made in deference to perceptual saliency (Mackinlay \cite{clevelandResearchStatisticalGraphics1987,clevelandGraphicalPerceptionTheory1984,chambersGraphicalMethodsData1983a, munznerVisualizationAnalysisDesign2014})
        \item[Naturalness] easier to understand when properties match (Norman \cite{norman_things_smart})
        \item[Graphical Integrity] graphs show \textbf{only} the data (Tufte \cite{tufteVisualDisplayQuantitative2001})
    \end{description}
\end{frame}

\begin{frame}{Visualization is commutative maps}

    \pause
    %% deconstruct D\R\V
    \begin{center}
        \textbf{Tam: Add topology and make it functional}
    \end{center}
\end{frame}

\begin{frame}{Models describe composition}
    \begin{description}
        \item[language model] APT, GoG: syntax, semantics, and grammar of graphics (Mackinlay, Wilkenson  \cite{mackinlayAutomatingDesignGraphical1986, mackinlayAUTOMATICDESIGNGRAPHICAL1987,wilkinsonGrammarGraphics2005})
        \item[functional dependencies] constrained maps between data and visual representation(Sugibuchi \cite{sugibuchiFramwork2009}) 
        \item[category theory] the semiotics of visualization are commutative (Vickers \cite{vickersUnderstandingViz2013})
        \item[algebraic process] data ($\alpha$) and viz ($\omega$) transforms are symmetric (Kindlmann and Scheidegger \cite{kindlmannAlgebraicProcessVisualization2014})
        \begin{columns}
            \column{.5\textwidth}

            \begin{description}
                \item[D] data 
                \item[R] representations
                \item[V] visualizations
            \end{description}
            \column{.5\textwidth}
            \begin{equation*}
                \begin{tikzcd}[ampersand replacement=\&]
                    D \arrow[d, "\alpha"'] \arrow[r, "r_1"] \& R \arrow[r, "\nu"]  \& V \arrow[d, "\omega"] \\
                    D \arrow[r, "r_2"']                     \& R \arrow[r, "\nu"'] \& V                    
                \end{tikzcd}
                \end{equation*}

        \end{columns}

    \end{description}
\end{frame}

\begin{frame}{Mathematical Frameworks for evaluating visualization}
    \begin{enumerate}
        \item APT: visualization has syntax and semantics like a language  (Mackinlay  \cite{mackinlayAutomatingDesignGraphical1986, mackinlayAUTOMATICDESIGNGRAPHICAL1987})
        \item visualization has functional dependencies that can be represented as graphs (Sugibuchi \cite{sugibuchiFramwork2009}) 
        \item the semiotics of visualization are commutative in a category theory framework (Vickers \cite{vickersUnderstandingViz2013})
        \item data ($\alpha$) and viz ($\omega$) symmetries (Kindlmann and Scheidegger \cite{kindlmann2014algebraic})
        \begin{block}{$v\circ r_2 \circ \alpha =\omega\circ v\circ r_1$}
            \begin{equation*}
            \begin{tikzcd}[ampersand replacement=\&]
                D \arrow[d, "\alpha"'] \arrow[r, "r_1"] \& R \arrow[r, "\nu"]  \& V \arrow[d, "\omega"] \\
                D \arrow[r, "r_2"']                     \& R \arrow[r, "\nu"'] \& V                    
            \end{tikzcd}
            \end{equation*}
        \end{block}
\end{frame}

\begin{frame}{Structure is encoded in variables and continuity}
    \begin{figure}
        \includegraphics[width=1\textwidth]{figures/intro/munzner_datatypes.png}
        \caption{Image is figure 2.8 in Munzner's Visualization Analysis and Design\cite{munznerVisualizationAnalysisDesign2014}}
    \end{figure}
    \begin{description}
        \item[binding] metadata are structural \textit{keys} with associated \textit{values} (Munzner \cite{munznerVisualizationAnalysisDesign2014})] 
        \item[continuity])
        \item[variables] Fibers can hold schema like encodings of variables (Spivak \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES})
    \end{description}
\end{frame}

begin{frame}{Monoid Actions: Permutation}
    \begin{figure}
        \includegraphics[width=1\linewidth]{figures/math/monoid_emoji.png}
    \end{figure}
\end{frame}

\begin{frame}{Why monoids? partial orders}
    \begin{figure}
        \begin{overprint}
            \onslide<1|handout:0>\includegraphics[width=1\linewidth]{figures/math/monoid_hasse.png}
            \onslide<2|handout:0>\includegraphics[width=1\linewidth]{figures/math/monoid_monotone.png}
            \onslide<3>\includegraphics[width=1\linewidth]{figures/math/monoid_maps.png}
        \end{overprint}
    \caption{Inspired by definition 1.59 diagram in Spivak and Fong's An Invitation to Applied Category Theory \cite{fongInvitationAppliedCategory2019}}
    \end{figure}
\end{frame}

%appendix 
\begin{frame}{\dbase\ is an indexing space independent of component semantics}
    \begin{equation}
        \pi:\dtotal_1\oplus\ldots\oplus \dtotal_i \oplus\ldots \oplus \dtotal_n \rightarrow \dbase
    \end{equation}
\end{frame}

%rename so everything is consistent

\begin{frame}{MayaVI}
    \begin{figure}
        \includegraphics[height=.5\textheight]{figures/intro/dataset_diagram.png}
        \caption{Data Representation, MayaVi 4.7.2 docs\cite{DataRepresentationMayavi}}
    \end{figure}
\end{frame}

### 2022-01-07 slides/main.tex
sheaves

@@
         \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-            \onslide*<7-8|handout:4>{\sheafc_{\gbasec, \gtotalc}} 
+            \onslide*<7-8|handout:4-5>{\sheafc_{\gbasec, \gtotalc}} 
             \onslide*<7-8,10-|handout:5>{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]} \& \& 
@@
             \\
-            \onslide*<2-|handout:1->{\gbasec}
-            \onslide*<3-|handout:1->{\arrow[rr, "\vindex", color=functor]}
-            \onslide*<7-8,10-|handout:4>{\arrow[u, "\sheaf_{\gbasec, \gtotalc}", maps to, color=sheaf]}
+            \onslide*<2-|handout:1,3->{\gbasec}
+            \onslide*<3-|handout:1,3,5>{\arrow[rr, "\vindex", color=functor]}
+            \onslide*<7-8,10-|handout:4-5>{\arrow[u, "\sheaf_{\gbasec, \gtotalc}", maps to, color=sheaf]}
             \onslide*<6,8-9,11-|handout:3>{\arrow[d, dashed, maps to, color=sheaf]} \&  \& 
-            \onslide*<1-|handout:1->{\dbasec}
+            \onslide*<1-|handout:1-3,5>{\dbasec}
             \onslide*<4-6,8-9,11-|handout:2-3>{\arrow[d, "\sheaf_{\dbasec, \dtotalc}", maps to, color=sheaf]}
@@
             \item $\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \ni \gsectionc: \opensetgc \rightarrow \gfiberc\restriction_{\opensetgc}$
-            \item $\gsectionc(\gbasepointc) = \{d_0, \cdots,\}, \dbasepointc \in \opensetfc$
+            \item $\gsectionc(\gbasepointc) = \{d_0, \cdots\}, \gbasepointc \in \opensetgc$
         \end{itemize}
     }
-
-\end{frame}
-
-\begin{frame}{
-    \only<1-3|handout:1-3>{Pullback \& Pushforward are Isomorphic}
-    \only<4|handout:4>{$\vartistc:\sheafc_{(\dbasec, \dtotalc)}\rightarrow \sheafc_{(\gbasec, \gtotalc)}$}
-    \only<5-|handout:5->{$\vartistc: \dsectionc \mapsto \gsectionc$}
-    }
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement=\&]
-          \onslide*<1-4|handout:1-4>{\sheafc_{(\gbasec, \gtotalc)}(\gbasec)} 
-          \onslide*<5-|handout:5->{\{\gsectionc:\gbasepointc \scarr \setc\}}
-          \onslide*<1-3|handout:1-3>{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]}  
-          \onslide*<4-|handout:4->{\arrow[rr, "\vindexpushc", color=artist]} 
-          \& \& 
-          \onslide*<1-4|handout:1-4>{(\textcolor{functor}{\vindexpush}\sheafc_{(\gbasec, \gtotalc)})(\dbase)}
-          \onslide*<5-|handout:5>{\{\vindexpushc \gsectionc: \dbasepointc \scarr\setc\}}
-          \\
-          \& \& \\
-          \onslide*<1-4|handout:1-4>{(\textcolor{functor}{\vindexpull}\sheafc_{(\dbasec, \dtotalc)})(\gbasec)}
-          \onslide*<5-|handout:5>{\{ \vindexpullc\dsectionc:\gbasepointc\scarr \setc\}}
-          \onslide*<1-3|handout:1-3>{\arrow[rr, "\textcolor{functor}{\vindexpull}", color=functor]}
-          \onslide*<4-|handout:4->{\arrow[rr, "\vindexpullc", color=artist]}
-          \onslide*<2-3|handout:2-3>{\arrow[uu, "\textcolor{homset}{hom_{\gbasec}}", color=homset]} 
-          \onslide*<4-|handout:4->{\arrow[uu, "\vartistc_{\vindexpushc}", color=artist]}
-          \&  \& 
-          \onslide*<1-4|handout:1-4>{\sheafc_{(\dbasec, \dtotalc)}(\dbase)} 
-          \onslide*<5-|handout:5->{\{\dsectionc:\dbasepointc \scarr \setc\}}
-          \onslide*<2-3|handout:2-3>{\arrow[uu, "\textcolor{homset}{hom_{\dbasec}}"', color=homset]}
-          \onslide*<4-|handout:4->{\arrow[uu, "\vartistc_{\vindexpullc}"', color=artist]}
-          \onslide*<3|handout:3>{\arrow[lluu, "\textcolor{homset}{hom_{\dbasec,\gbasec}}"' description, dashed, color=homset]}
-          \onslide*<4-|handout:4->{\arrow[lluu, "\vartistc"' description, dashed, color=artist]}
-        \end{tikzcd}
-      \end{equation*}
-
-    \onslide<3|handout:3>{
-        \begin{equation*}
-        \textcolor{homset}{hom}_{\gbasec}(\textcolor{functor}{\vindexpull}\sheafc_{(\dbasec, \dtotalc)}, \sheafc_{(\gbasec, \gtotalc)}) = 
-    \textcolor{homset}{hom}_{\dbase}(\sheafc_{(\dbasec, \dtotalc)}, \textcolor{functor}{\vindexpush}\sheafc_{(\gbasec, \gtotalc)})
-        \end{equation*}
-    }
-    \onslide<4|handout:4>{
-        \begin{equation*}
-            \vartistc_{\vindexpull} = \vartistc_{\vindexpush}
-        \end{equation*}
-    }
-\end{frame}
-
-\begin{frame}{}
-
-\end{frame}
+    \only<7|handout:5>{
+        \begin{itemize}
+            \item $\vindexpushc\gfiberc \hookrightarrow \vindexpushc\gtotalc \xrightarrow{\pi} \dbasec$
+            \item $\vindexpushc\sheafc_{\gbasec, \gtotalc}:\opensetc \mapsto \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}, \opensetc \subset \dbasec$
+            \item $\cgamma{\opensetc}{\gtotalc\restriction_{\opensetc}} \ni \gsectionc: \opensetc \rightarrow \gfiberc\restriction_{\opensetc}$
+            \item $\vindexpushc\gsectionc(\gbasepointc) = \gsectionc\restriction_{\vindexpre(\dbasepointc)} = \gsectionc(\gbasepointc) \forall \gbasepointc \in \vindexpre(\dbasepointc)$ 
+        \end{itemize}
+    }
+
+\end{frame}
+.
 
@@
 
+\begin{frame}{
+    \only<1-3|handout:1-3>{Natural Transforms Between Sheaves}
+    \only<4|handout:4>{$\vartistc:\sheafc_{(\dbasec, \dtotalc)}\rightarrow \sheafc_{(\gbasec, \gtotalc)}$}
+    \only<5-|handout:5->{$\vartistc: \dsectionc \mapsto \gsectionc$}
+    }
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement=\&]
+          \onslide*<1-4|handout:1>{\sheafc_{(\gbasec, \gtotalc)}(\gbasec)} 
+          \onslide*<1-3|handout:1>{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]}  
+          \onslide*<4-|handout:2->{\arrow[rr, "\vindexpushc", color=artist]} 
+          \& \& 
+          \onslide*<1-4|handout:1>{(\textcolor{functor}{\vindexpush}\sheafc_{(\gbasec, \gtotalc)})(\dbase)}
+          \\
+          \& \& \\
+          \onslide*<1-4|handout:1>{(\textcolor{functor}{\vindexpull}\sheafc_{(\dbasec, \dtotalc)})(\gbasec)}
+          \onslide*<1-3|handout:1>{\arrow[rr, "\textcolor{functor}{\vindexpull}", color=functor]}
+          \onslide*<4-|handout:2->{\arrow[rr, "\vindexpullc", color=artist]}
+          \onslide*<2-3|handout:1>{\arrow[uu, "\textcolor{nattran}{Nat_{\gbasec}}", color=nattran]} 
+          \onslide*<4-|handout:2->{\arrow[uu, "\vartistc_{\vindexpushc}", color=artist]}
+          \&  \& 
+          \onslide*<1-4|handout:1->{\sheafc_{\dbasec, \dtotalc}} 
+          \onslide*<2-3|handout:1>{\arrow[uu, "\textcolor{nattran}{Nat_{\dbasec}}"', color=nattran]}
+          \onslide*<4-|handout:2->{\arrow[uu, "\vartistc_{\vindexpullc}"', color=artist]}
+          \onslide*<3|handout:1>{\arrow[lluu, " "' description, dashed, color=nattran]}
+          \onslide*<4-|handout:2->{\arrow[lluu, "\vartistc"' description, dashed, color=artist]}
+        \end{tikzcd}
+      \end{equation*}
+
+    \onslide<3|handout:3>{
+        \begin{equation*}
+        \textcolor{nattran}{Nat}_{\gbasec}(\textcolor{functor}{\vindexpull}\sheafc_{(\dbasec, \dtotalc)}, \sheafc_{(\gbasec, \gtotalc)}) = 
+    \textcolor{nattran}{Nat}_{\dbase}(\sheafc_{(\dbasec, \dtotalc)}, \textcolor{functor}{\vindexpush}\sheafc_{(\gbasec, \gtotalc)})
+        \end{equation*}
+    }
+    \onslide<4|handout:4>{
+        \begin{equation*}
+            \vartistc_{\vindexpull} = \vartistc_{\vindexpush}
+        \end{equation*}
+    }
+\end{frame}
+
 \begin{frame}{$\vartist_{\vindexpushc}: \sheafc_{(\dbasec, \dtotalc)} \rightarrow \vindexpushc\sheafc_{(\gbasec, \gtotalc)}$}

### 2022-01-10 slides/main.tex
equivariance

@@
 \end{frame}
-\end{comment}
+
 \subsection{Algebraic Topology & Category Theory}
@@
         \item[Fiber Bundles] "unified, dimension-independent framework" that expresses data as the mapping between continuity and fields \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}
+        %% change SD to Category Theory Langauge - Express structure and constraints [software papers]
         \item[Simplicial Databases] systemic way to apply a schema to a fiber, i.e. provides a way to name fields and bind them to types system (e.g. \texttt{int}, \texttt{float}) 
@@
 
+%drop category K and \gamma op slides
 \begin{frame}{Category $\mathcal{\dbasec}$}
@@
 
+%pull back in keys, values 
 \begin{frame}{Presheaf: $\sheafc: \mathcal{C}^{op} \textcolor{sheaf}{\rightarrow} \setc $}
@@
         \begin{alertblock}<8-|handout:1>{stalk}
+            % F_k is subset of O_{K,E} is 
+            % O_{K,E} is the Jet
             \begin{aligned*}
-                \sheafc_{\dbasec, \dtotalc}\restriction_{\dbasepointc}\coloneqq \lim\limits_{\opensetc\ni \dbasepointc} \Gamma(\opensetc, \dtotalc\restriction_{\opensetc}) \approx \dfiberc_{\dbasepointc}
+                \sheafc_{\dbasec, \dtotalc}\restriction_{\dbasepointc}\coloneqq \lim\limits_{\opensetc\ni \dbasepointc} \Gamma(\opensetc, \dtotalc\restriction_{\opensetc}) \ \dfiberc_{\dbasepointc}
             \end{aligned*}
@@
         \begin{alertblock}<9-|handout:1>{germ}
+            %germ  is tau at limit, \tau(k) \in germ
             \begin{aligned*}
@@
 
-\begin{frame}{Sheaves}
+\begin{frame}{Sheaves} %why we care, either drop out or define what an open covering is
+    %%% footnote on open covering
     A sheaf is a presheaf that satisfies the following two axioms\cite{bakerMathsSheaf}\\
@@
     \end{block} 
-    \begin{block}<5-|handout:1>{gluing}
+    \begin{block}<5-|handout:1>{gluing} %gluing w/o open covering rule
         if $\mathcal{\opensetc} = \{\opensetc_i\}_{i\in I}$ is an open covering of \opensetc\ and if $\dsectionc^{i} \in \sheafc(\opensetc_i)$ is given for each $i \in I$
@@
 
+%flip the ordering: K, then G, then functor 
 \begin{frame}
@@
 
+%add transition slide
+
+%move other side to when I introduce Q 
 \begin{frame}{Natural Transformation $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\rightarrow} \textcolor{functor}{G}$}
@@
 
+% pull up artist^*, artist_*
 \begin{frame}{
@@
         \begin{tikzcd}[ampersand replacement=\&]
-          \onslide*<1-4|handout:1>{\sheafc_{(\gbasec, \gtotalc)}(\gbasec)} 
+          \onslide*<1-4|handout:1>{\sheafc_{\gbasec, \gtotalc}\gbasec} 
           \onslide*<1-3|handout:1>{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]}  
@@
 \end{frame}
+\end{comment}
+
+\section{Equivariant Artist}
+\begin{frame}{"Valid" viz?}
+    \begin{equation*}
+        \label{eq:artist:artist_sheaf}
+        \begin{tikzcd}[ampersand replacement = \&, column sep=tiny, row sep=huge]
+            \onslide*<1-|handout:1>{\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}} 
+            \&  \& 
+            \onslide*<2-|handout:1>{\textcolor{set}{\big\{}\gsectionc:\opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} 
+              \bigm{\vert} \vartistc\circ (\dfuncc \circ \vartistc^{-1}[\gsectionc]) = \dfuncc^{\prime} \circ \gsectionc \textcolor{set}{\big\}}} 
+            \onslide*<2-|handout:1>{\arrow[ll, "\textcolor{set}{\mathlarger{\mathlarger{\supset}}}" description, hook', color=white]} \\ 
+            \&  \& \\
+            \onslide*<1-|handout:1>{\opensetgc} 
+            \onslide*<1-|handout:1>{\arrow[uu, "\sheafc_{\gtotalc}", color=sheaf]} 
+            \onslide*<2-|handout:1>{\arrow[rruu, "\vartistc(\sheafc_{\dtotalc})\coloneqq \sheafc_{\vartistc}" description, dashed, color=sheaf]} 
+            \&  \& 
+        \end{tikzcd}
+    \end{equation*}
+\end{frame}
+
+\begin{frame}{$\dfuncc = (\hat\dfuncc, \hat\dfuncc_*, \tilde\dfuncc)$?}
+\begin{equation*}
+    \begin{tikzcd}[ampersand replacement = \&, row sep=large]
+        {\Gamma(U, E)} 
+        \arrow[r, "\hat\varphi_*"]                             
+        \& 
+        {\Gamma(U^{\prime}, \hat{\varphi_*}E)} 
+        \arrow[r, "\tilde{\varphi}"]                                                                                                             
+        \& {\Gamma(U^{\prime}, \tilde{\varphi}\hat{\varphi_*}E)} \\
+        {U \subset K} 
+        \arrow[r, "\hat\varphi"] 
+        \arrow[u, "{\mathcal{O}_{K,E}}"] 
+        \& 
+        {U^{\prime}\subset K^{\prime}}
+        \arrow[u, "{\mathcal{O}_{K^{\prime}, \hat\varphi_*E}}", dashed] 
+        \arrow[ru, "{\mathcal{O}_{\hat\varphi K, \tilde\varphi\hat\varphi_*E}}"', dotted] 
+        \&                                                                                                          
+    \end{tikzcd}                                   
+\end{equation*}
+\end{frame}
+
+\begin{frame}{Continuty Equivariance $\vindexpullc\hat\dfuncc = \dfuncc|_{\vindexprec}$}
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement =\&]
+            S 
+            \arrow[r, "\xi"] 
+            \arrow[d, "\xi^*\hat\varphi"'] 
+            \& 
+            K 
+            \arrow[d, "\hat\varphi"] \\
+            S^{\prime} 
+            \arrow[r, "\xi"]
+            \& 
+            K^{\prime}
+            \end{tikzcd}
+    \end{equation*}
+\end{frame}
+\begin{frame}{Field Equivariance $\tilde\dfuncc \cong \tilde\dfuncc_\gtotalc$}
+    \begin{equation*}
+    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
+        {\Gamma(U, E)} 
+        \arrow[r, "A"] 
+        \arrow[d, "\tilde\varphi\circ\hat\varphi_*"'] 
+        \& 
+        {\Gamma(W, H)} 
+        \arrow[d, "\tilde\varphi_H\circ\xi_*\hat{\varphi}_*"] \\
+        {\Gamma(\hat\varphi U, \tilde\varphi\hat\varphi_*E)} 
+        \arrow[r, "A"] 
+        \& 
+        {\Gamma(\xi^*\hat\varphi W, \tilde\varphi_H\xi^*\hat\varphi_*H)}    
+    \end{tikzcd}
+    \end{equation*}
+\end{frame}
+
+\begin{frame}{Equivariant Artist}
+    \begin{equation*}
+    \begin{tikzcd}[ampersand replacement=\&]
+        {\mathcal{O}_{K, E}} 
+        \arrow[r, "A"] 
+        \arrow[d, "\varphi"'] \& 
+        {\mathcal{O}_{S,H}} 
+        \arrow[d, "\varphi_H"] \\
+        {\mathcal{O}_{K^{\prime}, E^{\prime}}} 
+        \arrow[r, "A"]    \& 
+        {\mathcal{O}_{S^{\prime}, H^{\prime}}}    
+        \end{tikzcd} 
+    \end{equation*}
+\begin{block}
+    \vartistc\ is equivariant when there exists a $\dfuncc_H$ s.t the diagram commutes. 
+    \begin{equation}
+
+    \end{equation}
+\end{block}
+    \end{frame}
+
+\begin{frame}{More than one field?}
+    %cartestian categories
+\end{frame}
 
 \begin{comment}
-
 \subsection{Contributions}

### 2022-01-10 slides/main.tex
monoidal category

@@
             \vartistc\ is equivariant when there exists a $\dfuncc_H$ s.t the diagram commutes. 
-    \begin{equation}
-
-    \end{equation}
+            \begin{equation*}
+            \end{equation*}
         \end{block}
@@
 \begin{frame}{More than one field?}
-    %cartestian categories
+    \begin{block}{} 
+        \label{def:category:F} The fiber category $\mathcal{\dfiberc}$ is a monoidal category\cite{milewskiCategoryTheoryProgrammers}, which means it is a category equipped with a bifunctur $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$
+        \begin{itemize}
+          \item \textit{object} a monoid object \dfiber.
+A monoid is a set with a binary operator that is associative, closed, and for which the set contains an identity element $i$ \cite{nlab:monoid}.
+          \item \textit{morphisms} $\dfiber \otimes \dfiber \rightarrow \dfiber$ and unit $i \rightarrow \dfiber$
+        \end{itemize}
+    \end{block}
 \end{frame}

### 2022-01-12 slides/main.tex
flipped arrows, corrected sheaves and presheaves formulas

@@
 \end{frame}
-
-\subsection{Algebraic Topology & Category Theory}
+\end{comment}
+
+\section{Algebraic Topology & Category Theory}
 
@@
         \item[Fiber Bundles] "unified, dimension-independent framework" that expresses data as the mapping between continuity and fields \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}
-        %% change SD to Category Theory Langauge - Express structure and constraints [software papers]
-        \item[Simplicial Databases] systemic way to apply a schema to a fiber, i.e. provides a way to name fields and bind them to types system (e.g. \texttt{int}, \texttt{float}) 
-         \cite{spivakSIMPLICIALDATABASES, spivakDatabasesAreCategories2010} 
+        \item[Category Theory Language] express constraints in specifications \cite{wielsManagementEvolvingSpecifications1998}   
         \item[Sheaves on Bundles] "algebraic data structure" for representing data over topological spaces \cite{ghristElementaryAppliedTopology2014}
@@
         \begin{description}
-            \item[\textcolor{base}{Continuity Type \dbasec}]{how data elements are organized (topological properties) \cite{wilkinsonGrammarGraphics2005})}
+            \item[\textcolor{base}{Continuity Type \dbasec}]{how data elements are organized (topological properties) \cite{wilkinsonGrammarGraphics2005,munznerWhatDataAbstraction2014})}
             \item[\textcolor{fiber}{Fields Type \dfiberc}] {generalization of a schema - named and typed date fields \cite{spivakSIMPLICIALDATABASES, spivakDatabasesAreCategories2010}}
@@
 
-%drop category K and \gamma op slides
-\begin{frame}{Category $\mathcal{\dbasec}$}
-\begin{columns}
-    \column{0.49\textwidth}
-\begin{equation*}
-    \begin{tikzcd}[ampersand replacement=\&]
-        \textcolor{source}{C_1} \arrow[r, "f", color=source] \arrow[rd, "g\circ f"', color=source] \arrow["id_{C_1}"', loop, distance=2em, in=215, out=145, color=source] \& \textcolor{source}{C_2} \arrow[d, "g", color=source] \arrow["id_{C_2}"', loop, distance=2em, in=125, out=55, color=source] \\
-    \& \textcolor{source}{C_3} \arrow["id_{C_3}"', loop, distance=2em, in=305, out=235, color=source]              
-    \end{tikzcd}
-    \end{equation*}
-    \column{0.49\textwidth}
-    \onslide<2-|handout:1>{
-    \begin{tikzcd}[ampersand replacement=\&]
-            \opensetc_1 \arrow[r, "\iota_{1,2}", color=base] \arrow[rd, "\iota_{1,2}\circ \iota_{2,3}"', color=base] \arrow["id_{\opensetc_1}"', loop, distance=2em, in=215, out=145, color=base] \& \opensetc_2 \arrow[d, "\iota_{2,3}", color=base] \arrow["id_{\opensetc_2}"', loop, distance=2em, in=125, out=55, color=base] \\
-        \& \opensetc_3 \arrow["id_{\opensetc_3}"', loop, distance=2em, in=305, out=235, color=base]              
-    \end{tikzcd}
-    }
-    \end{columns}
-\end{frame}
-
 \begin{frame}{Opposite Category $\textcolor{source}{\mathcal{C}^{op}}$}
@@
         \end{equation*}
-        \end{columns}
-\end{frame}
-
-\begin{frame}{$\mathcal{C}^{op}$}
-    \begin{columns}
-        \column{0.29\textwidth}
-        \begin{tikzcd}[ampersand replacement=\&]
-            \opensetc_1 \arrow[r, "\iota_{1,2}", color=base] \arrow[rd, "\iota_{1,2}\circ \iota_{2,3}"', color=base] \arrow["id_{\opensetc_1}"', loop, distance=2em, in=215, out=145, color=base] \& \opensetc_2 \arrow[d, "\iota_{2,3}", color=base] \arrow["id_{\opensetc_2}"', loop, distance=2em, in=125, out=55, color=base] \\
-        \& \opensetc_3 \arrow["id_{\opensetc_3}"', loop, distance=2em, in=305, out=235, color=base]              
-    \end{tikzcd}
-        \column{0.69\textwidth}
-        \onslide<2-|handout:1>{
-            \begin{tikzcd}[ampersand replacement=\&]
-            \cgamma{\opensetc_1}{\dtotalc\restriction_{\opensetc_1}} \&  \& \cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}} \arrow[ll, "\iota^*\restriction_{\opensetc_1}"', color=set]\\
-            \&\&\\\&  \& \cgamma{\opensetc_3}{\dtotalc\restriction_{\opensetc_3}} \arrow[uu, "\iota^*\restriction_{\opensetc_2}"', color=set] \arrow[lluu, "\iota^*\restriction_{\opensetc_2} \circ \textcolor{set}{\iota^*\restriction}_{\opensetc_1}", color=set]
-            \end{tikzcd}
-        }
-
         \end{columns}
@@
 
-%pull back in keys, values 
 \begin{frame}{Presheaf: $\sheafc: \mathcal{C}^{op} \textcolor{sheaf}{\rightarrow} \setc $}
@@
         \begin{alertblock}<8-|handout:1>{stalk}
-            % F_k is subset of O_{K,E} is 
-            % O_{K,E} is the Jet
             \begin{aligned*}
-                \sheafc_{\dbasec, \dtotalc}\restriction_{\dbasepointc}\coloneqq \lim\limits_{\opensetc\ni \dbasepointc} \Gamma(\opensetc, \dtotalc\restriction_{\opensetc}) \ \dfiberc_{\dbasepointc}
+                \sheafc_{\dbasec, \dtotalc}\restriction_{\dbasepointc}\coloneqq \lim\limits_{\opensetc\ni \dbasepointc} \Gamma(\opensetc, \dtotalc\restriction_{\opensetc}) \\
+                 \dfiberc_{\dbasepointc} \subset  \sheafc_{\dbasec, \dtotalc}\restriction_{\dbasepointc}
             \end{aligned*}
@@
 
-\begin{frame}{Sheaves} %why we care, either drop out or define what an open covering is
-    %%% footnote on open covering
+\begin{frame}{Sheaves} 
     A sheaf is a presheaf that satisfies the following two axioms\cite{bakerMathsSheaf}\\
     \begin{block}<2-|handout:1>{locality}
-        if $\mathcal{\opensetc} = \{\opensetc_i\}_{i\in I}$ is an open covering of \opensetc
+        if there exists the open covering $\mathcal{\opensetc} = \{\opensetc_i\}_{i\in I}$ of \opensetc
         \onslide<3-|handout:1>{
@@
     \end{block} 
-    \begin{block}<5-|handout:1>{gluing} %gluing w/o open covering rule
-        if $\mathcal{\opensetc} = \{\opensetc_i\}_{i\in I}$ is an open covering of \opensetc\ and if $\dsectionc^{i} \in \sheafc(\opensetc_i)$ is given for each $i \in I$
+    \begin{block}<5-|handout:1>{gluing} 
+        if there exists the open covering $\mathcal{\opensetc} = \{\opensetc_i\}_{i\in I}$ of \opensetc\ and if $\dsectionc^{i} \in \sheafc(\opensetc_i)$ is given for each $i \in I$
         \onslide<6-|handout:1>{such that ${\dsectionc^{i}}\restriction_{\opensetc_i\cap \opensetc_j} = {\dsectionc^{j}}\restriction_{\opensetc_i \cap \opensetc_j}$ for each pair $\opensetc_i, \opensetc_j \in \mathcal{\opensetc}$,}
@@
 
-%flip the ordering: K, then G, then functor 
 \begin{frame}
-    \frametitle<0|handout:1>{Functor: \vindexc: \mathcal{\gbasec} \rightarrow \mathcal{\dbasec}}
-    \frametitle<0|handout:2>{Data}
-    \frametitle<0|handout:3>{Pullback: data to region of the visualization} 
-    \frametitle<0|handout:4>{Graphic}
-    \frametitle<0|handout:5>{Pushforward: visualization to index of data}
+    \frametitle<3|handout:3>{Functor: \vindexc: \mathcal{\gbasec} \rightarrow \mathcal{\dbasec}}
+    \frametitle<1|handout:1>{Data}
+    \frametitle<4-6|handout:4>{Pullback: data to region of the visualization} 
+    \frametitle<2|handout:2>{Graphic}
+    \frametitle<7-|handout:5>{Pushforward: visualization to index of data}
     \begin{equation*}
         \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-            \onslide*<7-8|handout:4-5>{\sheafc_{\gbasec, \gtotalc}} 
-            \onslide*<7-8,10-|handout:5>{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]} \& \& 
-            \onslide*<7-8|handout:5>{\textcolor{functor}{\vindexpush}\sheafc_{\gbasec, \gtotalc}}  
+            \onslide*<0|handout:2,3,5>{\sheafc_{\gbasec, \gtotalc}} 
+            \onslide*<0|handout:5>{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]} \& \& 
+            \onslide*<0|handout:5>{\textcolor{functor}{\vindexpush}\sheafc_{\gbasec, \gtotalc}}  
             \\
-            \onslide*<2-|handout:1,3->{\gbasec}
-            \onslide*<3-|handout:1,3,5>{\arrow[rr, "\vindex", color=functor]}
-            \onslide*<7-8,10-|handout:4-5>{\arrow[u, "\sheaf_{\gbasec, \gtotalc}", maps to, color=sheaf]}
-            \onslide*<6,8-9,11-|handout:3>{\arrow[d, dashed, maps to, color=sheaf]} \&  \& 
-            \onslide*<1-|handout:1-3,5>{\dbasec}
-            \onslide*<4-6,8-9,11-|handout:2-3>{\arrow[d, "\sheaf_{\dbasec, \dtotalc}", maps to, color=sheaf]}
-            \onslide*<7-8,10-|handout:5>{\arrow[u, dashed, maps to, color=sheaf]} \\
-            \onslide*<5-6,8|handout:3>{\textcolor{functor}{\vindexpull}\sheafc_{\dbasec, \dtotalc}} 
-            \onslide*<5-6,8-9,11-|handout:3>{\arrow[rr, "\textcolor{functor}{\vindexpull}", color=functor]}  \& \& 
-            \onslide*<4-6,8|handout:2-3>{\sheafc_{\dbasec, \dtotalc}}
-        \end{tikzcd}
-    \end{equation*}
-    \only<4|handout:2>{
+            \onslide*<0|handout:2->{\gbasec}
+            \onslide*<0|handout:3->{\arrow[rr, "\vindex", color=functor]}
+            \onslide*<0|handout:2,3,5>{\arrow[u, "\sheaf_{\gbasec, \gtotalc}", maps to, color=sheaf]}
+            \onslide*<0|handout:4>{\arrow[d, dashed, maps to, color=sheaf]} \&  \& 
+            \onslide*<0|handout:1,3->{\dbasec}
+            \onslide*<0|handout:1,3-4>{\arrow[d, "\sheaf_{\dbasec, \dtotalc}", maps to, color=sheaf]}
+            \onslide*<0|handout:5>{\arrow[u, dashed, maps to, color=sheaf]} \\
+            \onslide*<0|handout:4>{\textcolor{functor}{\vindexpull}\sheafc_{\dbasec, \dtotalc}} 
+             \& \& 
+            \onslide*<0|handout:1,3-4>{\sheafc_{\dbasec, \dtotalc}}
+            \onslide*<0|handout:4>{\arrow[ll, "\textcolor{functor}{\vindexpull}", color=functor]} 
+        \end{tikzcd}
+    \end{equation*}
+    \only<0|handout:1>{
         \begin{itemize}
@@
     }
-    \only<6|handout:3>{
+    \only<0|handout:4>{
         \begin{itemize}
@@
     }
-    \only<7|handout:4>{
+    \only<0|handout:2>{
         \begin{itemize}
@@
     }
-    \only<7|handout:5>{
+    \only<0|handout:5>{
         \begin{itemize}
@@
     }
-
-\end{frame}
-.
+\end{frame}
 
@@
 
-% pull up artist^*, artist_*
-\begin{frame}{
-    \only<1-3|handout:1-3>{Natural Transforms Between Sheaves}
-    \only<4|handout:4>{$\vartistc:\sheafc_{(\dbasec, \dtotalc)}\rightarrow \sheafc_{(\gbasec, \gtotalc)}$}
-    \only<5-|handout:5->{$\vartistc: \dsectionc \mapsto \gsectionc$}
-    }
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement=\&]
-          \onslide*<1-4|handout:1>{\sheafc_{\gbasec, \gtotalc}\gbasec} 
-          \onslide*<1-3|handout:1>{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]}  
-          \onslide*<4-|handout:2->{\arrow[rr, "\vindexpushc", color=artist]} 
-          \& \& 
-          \onslide*<1-4|handout:1>{(\textcolor{functor}{\vindexpush}\sheafc_{(\gbasec, \gtotalc)})(\dbase)}
-          \\
-          \& \& \\
-          \onslide*<1-4|handout:1>{(\textcolor{functor}{\vindexpull}\sheafc_{(\dbasec, \dtotalc)})(\gbasec)}
-          \onslide*<1-3|handout:1>{\arrow[rr, "\textcolor{functor}{\vindexpull}", color=functor]}
-          \onslide*<4-|handout:2->{\arrow[rr, "\vindexpullc", color=artist]}
-          \onslide*<2-3|handout:1>{\arrow[uu, "\textcolor{nattran}{Nat_{\gbasec}}", color=nattran]} 
-          \onslide*<4-|handout:2->{\arrow[uu, "\vartistc_{\vindexpushc}", color=artist]}
-          \&  \& 
-          \onslide*<1-4|handout:1->{\sheafc_{\dbasec, \dtotalc}} 
-          \onslide*<2-3|handout:1>{\arrow[uu, "\textcolor{nattran}{Nat_{\dbasec}}"', color=nattran]}
-          \onslide*<4-|handout:2->{\arrow[uu, "\vartistc_{\vindexpullc}"', color=artist]}
-          \onslide*<3|handout:1>{\arrow[lluu, " "' description, dashed, color=nattran]}
-          \onslide*<4-|handout:2->{\arrow[lluu, "\vartistc"' description, dashed, color=artist]}
-        \end{tikzcd}
-      \end{equation*}
-
-    \onslide<3|handout:3>{
-        \begin{equation*}
-        \textcolor{nattran}{Nat}_{\gbasec}(\textcolor{functor}{\vindexpull}\sheafc_{(\dbasec, \dtotalc)}, \sheafc_{(\gbasec, \gtotalc)}) = 
-    \textcolor{nattran}{Nat}_{\dbase}(\sheafc_{(\dbasec, \dtotalc)}, \textcolor{functor}{\vindexpush}\sheafc_{(\gbasec, \gtotalc)})
-        \end{equation*}
-    }
-    \onslide<4|handout:4>{
-        \begin{equation*}
-            \vartistc_{\vindexpull} = \vartistc_{\vindexpush}
-        \end{equation*}
-    }
-\end{frame}
-
 \begin{frame}{$\vartist_{\vindexpushc}: \sheafc_{(\dbasec, \dtotalc)} \rightarrow \vindexpushc\sheafc_{(\gbasec, \gtotalc)}$}
@@
 \end{frame}
-\end{comment}
+
+\begin{frame}{
+    \only<1-3|handout:1-3>{Natural Transforms Between Sheaves}
+    \only<4|handout:4>{$\vartistc:\sheafc_{(\dbasec, \dtotalc)}\rightarrow \sheafc_{(\gbasec, \gtotalc)}$}
+    \only<5-|handout:5->{$\vartistc: \dsectionc \mapsto \gsectionc$}
+    }
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement=\&]
+          \onslide*<1-4|handout:1>{\sheafc_{\gbasec, \gtotalc}\gbasec} 
+          \onslide*<1-3|handout:1>{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]}  
+          \onslide*<4-|handout:2->{\arrow[rr, "\vindexpushc", color=artist]} 
+          \& \& 
+          \onslide*<1-4|handout:1>{(\textcolor{functor}{\vindexpush}\sheafc_{(\gbasec, \gtotalc)})(\dbase)}
+          \\
+          \& \& \\
+          \onslide*<1-4|handout:1>{(\textcolor{functor}{\vindexpull}\sheafc_{(\dbasec, \dtotalc)})(\gbasec)}
+          \onslide*<1-3|handout:1>{\arrow[rr, "\textcolor{functor}{\vindexpull}", color=functor]}
+          \onslide*<4-|handout:2->{\arrow[rr, "\vindexpullc", color=artist]}
+          \onslide*<2-3|handout:1>{\arrow[uu, "\textcolor{nattran}{Nat_{\gbasec}}", color=nattran]} 
+          \onslide*<4-|handout:2->{\arrow[uu, "\vartistc_{\vindexpushc}", color=artist]}
+          \&  \& 
+          \onslide*<1-4|handout:1->{\sheafc_{\dbasec, \dtotalc}} 
+          \onslide*<2-3|handout:1>{\arrow[uu, "\textcolor{nattran}{Nat_{\dbasec}}"', color=nattran]}
+          \onslide*<4-|handout:2->{\arrow[uu, "\vartistc_{\vindexpullc}"', color=artist]}
+          \onslide*<3|handout:1>{\arrow[lluu, " "' description, dashed, color=nattran]}
+          \onslide*<4-|handout:2->{\arrow[lluu, "\vartistc"' description, dashed, color=artist]}
+        \end{tikzcd}
+      \end{equation*}
+
+    \onslide<3|handout:3>{
+        \begin{equation*}
+        \textcolor{nattran}{Nat}_{\gbasec}(\textcolor{functor}{\vindexpull}\sheafc_{(\dbasec, \dtotalc)}, \sheafc_{(\gbasec, \gtotalc)}) = 
+    \textcolor{nattran}{Nat}_{\dbase}(\sheafc_{(\dbasec, \dtotalc)}, \textcolor{functor}{\vindexpush}\sheafc_{(\gbasec, \gtotalc)})
+        \end{equation*}
+    }
+    \onslide<4|handout:4>{
+        \begin{equation*}
+            \vartistc_{\vindexpull} = \vartistc_{\vindexpush}
+        \end{equation*}
+    }
+\end{frame}
 
@@
 
-\begin{frame}{Continuty Equivariance $\vindexpullc\hat\dfuncc = \dfuncc|_{\vindexprec}$}
+\begin{frame}{Continuty Equivariance $\vindexpullc\hat\dfuncc = \hat\dfuncc|_{\vindexprec}$}
     \begin{equation*}
@@
         \begin{block}
-            \vartistc\ is equivariant when there exists a $\dfuncc_H$ s.t the diagram commutes. 
+            \vartistc\ is equivariant if and only if there exists a $\dfuncc_H$ s.t the diagram commutes. 
             \begin{equation*}

### 2022-01-12 slides/main.tex
flip pullback arrow

@@
 
-%add transition slide
-
-%move other side to when I introduce Q 
-\begin{frame}{Natural Transformation $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\rightarrow} \textcolor{functor}{G}$}
+%add transition slide->possibly call back 
+\begin{frame}
+    {\huge
+        \begin{equation*}
+            \sheafc_{\dbasec,\dtotalc} \textcolor{artist}{\xrightarrow{\vartistc}} \sheafc_{\gbasec, \gtotalc}
+        \end{equation*}
+    }
+\end{frame}
+
+\begin{frame}{Natural Transformation $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\mapsto} \textcolor{functor}{G}$}
 
@@
         \begin{tikzcd}[ampersand replacement=\&]
-            \& {} \arrow[dd, "\alpha", Rightarrow, shorten=1.5em, color=nattran] \&             \\
-          \textcolor{source}{\mathcal{C}} \arrow[rr, "F", bend left, color=functor] \arrow[rr, "G"', bend right, color=functor] \& \& \textcolor{target}{\mathcal{D}} \\
+            \& {} \arrow[dd, "\alpha" , maps to, shorten=1.5em, color=nattran] \&             \\
+          \textcolor{source}{\mathcal{C}} \arrow[rr, "F", bend left, color=functor, maps to] \arrow[rr, "G"', bend right, color=functor, maps to] \& \& \textcolor{target}{\mathcal{D}} \\
             \& {} \&            
@@
         \begin{tikzcd}[column sep=huge, ampersand replacement = \&]
-            \onslide*<2-|handout:2->{\textcolor{target}{F(c)}}
-            \onslide*<2-|handout:2->{\arrow[dd, "F(f)"', color=target]}
-            \onslide*<4-5|handout:4-5>{\arrow[rr, "\alpha_{c}", bend left, color=nattran]}  
-            \onslide*<6|handout:6>{\arrow[rr, "\alpha_{c}", color=nattran]} \& 
-            \onslide*<1-4|handout:1-4>{\textcolor{source}{c}} 
-            \onslide*<1-4|handout:1-4>{\arrow[dd, "f" description, color=source]}
-            \onslide*<3-4|handout:3-4>{\arrow[r, "G"', color=functor]}
-            \onslide*<2-4|handout:2-4>{\arrow[l, "F", color=functor]} \& 
-            \onslide*<3-|handout:3->{\textcolor{target}{G(c)}} 
-            \onslide*<3-|handout:3->{\arrow[dd, "G(f)", color=target]} \\
+            \onslide*<2-|handout:1>{\textcolor{target}{F(c)}}
+            \onslide*<2-|handout:1>{\arrow[dd, "F(f)"', color=target]}
+            \onslide*<4-5|handout:1>{\arrow[rr, "\alpha_{c}", bend left, color=nattran, maps to]}  
+            \onslide*<6|handout:0>{\arrow[rr, "\alpha_{c}", color=nattran, maps to]} \& 
+            \onslide*<1-4|handout:1>{\textcolor{source}{c}} 
+            \onslide*<1-4|handout:1>{\arrow[dd, "f" description, color=source]}
+            \onslide*<3-4|handout:1>{\arrow[r, "G"', color=functor,maps to ]}
+            \onslide*<2-4|handout:1>{\arrow[l, "F", color=functor, maps to]} \& 
+            \onslide*<3-|handout:1>{\textcolor{target}{G(c)}} 
+            \onslide*<3-|handout:1>{\arrow[dd, "G(f)", color=target, maps to]} \\
             {} 
-            \onslide*<4|handout:4>{\arrow[rr, dotted, color=nattran]}
-            \onslide*<5|handout:5>{\arrow[rr, color=nattran]} \& \& {}                      \\
-            \onslide*<2-|handout:2->{\textcolor{target}{F(c^{\prime})}} 
-            \onslide*<4-5|handout:4-5>{\arrow[rr, "\alpha_{c^{\prime}}"', bend right, color=nattran]} 
-            \onslide*<4-5|handout:6>{\arrow[rr, "\alpha_{c^{\prime}}"', color=nattran]}\& 
-            \onslide*<1-4|handout:1-4>{\textcolor{source}{c^{\prime}}}
-            \onslide*<2-4|handout:2-4>{\arrow[l, "F"', color=functor]} 
-            \onslide*<3-4|handout:3-4>{\arrow[r, "G", color=functor]} \& 
-            \onslide*<3-|handout:3->{\textcolor{target}{G(c^{\prime})}}          
-        \end{tikzcd}
-    \end{equation*}    
-\end{frame}
-
-\begin{frame}{$\vartist_{\vindexpushc}: \sheafc_{(\dbasec, \dtotalc)} \rightarrow \vindexpushc\sheafc_{(\gbasec, \gtotalc)}$}
-    \begin{equation}
+            \onslide*<4|handout:1>{\arrow[rr, dotted, color=nattran, maps to]}
+            \onslide*<5|handout:0>{\arrow[rr, color=nattran, maps to]} \& \& {}                      \\
+            \onslide*<2-|handout:1>{\textcolor{target}{F(c^{\prime})}} 
+            \onslide*<4-5|handout:1>{\arrow[rr, "\alpha_{c^{\prime}}"', bend right, color=nattran, maps to]} 
+            \onslide*<4-5|handout:0>{\arrow[rr, "\alpha_{c^{\prime}}"', color=nattran, maps to]}\& 
+            \onslide*<1-4|handout:1>{\textcolor{source}{c^{\prime}}}
+            \onslide*<2-4|handout:1>{\arrow[l, "F"', color=functor, maps to]} 
+            \onslide*<3-4|handout:1>{\arrow[r, "G", color=functor, maps to]} \& 
+            \onslide*<3-|handout:1>{\textcolor{target}{G(c^{\prime})}}          
+        \end{tikzcd}
+    \end{equation*}    
+\end{frame}
+
+\begin{frame}
+    \frametitle<1-2|handout:1>{Data Space: $\vartistc_{\vindexpushc}: \sheafc_{\dbasec, \dtotalc} \textcolor{artist}{\mapsto} \vindexpushc\sheafc_{\gbasec, \gtotalc}$}
+    \frametitle<3-4|handout:2>{Display Space: $\vartist_{\vindexpullc}: \vindexpullc\sheafc_{\dbasec, \dtotalc} \rightarrow \sheafc_{\gbasec, \gtotalc}$}
+    \frametitle<5|handout:3->{\textcolor{artist}{Artist} \vartistc}
+
+    \onslide*<1|handout:1>{
+        \begin{equation*}
       \begin{tikzcd}[ampersand replacement =\&]
@@
         \arrow[dd, "\iota^*"', hook, color=set] 
-        \arrow[rrrr, "\vartistc_{\opensetc_1}", bend left, color=artist] \&  \& 
+        \arrow[rrrr, "\vartistc_{\opensetc_1}", bend left, color=artist, maps to] \&  \& 
         \opensetc_1 
-        \arrow[ll, "\sheafc_{\dtotalc}", color=sheaf] 
-        \arrow[rr, "\sheafc_{\gtotalpushc}"', color=sheaf] \&  \& 
+        \arrow[ll, "\sheafc_{\dtotalc}", color=sheaf, maps to] 
+        \arrow[rr, "\sheafc_{\gtotalpushc}"', color=sheaf, maps to] \&  \& 
         \cgamma{\opensetc_1}{\gtotalpushc} 
@@
         {} 
-        \arrow[rrrr, dotted, color=artist] &&&& {} \\
+        \arrow[rrrr, dotted, color=artist, maps to] \&\&\&\& {} \\
         \cgamma{\opensetc_2}{\dtotalc} 
-        \arrow[rrrr, "\vartistc_{\opensetc_2}"', bend right, color=artist] \&  \& 
+        \arrow[rrrr, "\vartistc_{\opensetc_2}"', bend right, color=artist, maps to] \&  \& 
         \opensetc_2 
-        \arrow[ll, "\sheafc_{\dtotalc}"', color=sheaf] 
-        \arrow[rr, "\sheafc_{\gtotalpushc}", color=sheaf] 
+        \arrow[ll, "\sheafc_{\dtotalc}"', color=sheaf, maps to] 
+        \arrow[rr, "\sheafc_{\gtotalpushc}", color=sheaf, maps to] 
         \arrow[uu, "\iota" description, hook, color=base] \& \& 
@@
     \end{tikzcd}
-    \end{equation}
-\end{frame}
-
-\begin{frame}{$\vartist_{\vindexpullc}: \vindexpullc\sheafc_{(\dbasec, \dtotalc)} \rightarrow \sheafc_{(\gbasec, \gtotalc)}$}
-    \begin{equation}
+        \end{equation*}}
+
+    \onslide*<0|handout:2>{
+        \begin{equation*}
   \begin{tikzcd}[ampersand replacement =\&]
@@
     \arrow[dd, "\iota^*"', hook, color=set] 
-    \arrow[rrrr, "\vartistc_{\opensetgc_1}", bend left, color=artist] \&  \& 
+    \arrow[rrrr, "\vartistc_{\opensetgc_1}", bend left, color=artist, maps to] \&  \& 
     \opensetgc_1 
-    \arrow[ll, "\sheafc_{\dtotalpullc}", color=sheaf] 
-    \arrow[rr, "\sheafc_{\gtotalc}"', color=sheaf] \&  \& 
+    \arrow[ll, "\sheafc_{\dtotalpullc}", color=sheaf, maps to] 
+    \arrow[rr, "\sheafc_{\gtotalc}"', color=sheaf, maps to] \&  \& 
     \cgamma{\opensetgc_1}{\gtotalc} 
@@
     {} 
-    \arrow[rrrr, dotted, color=artist] &&&& {} \\
+    \arrow[rrrr, dotted, color=artist, maps to] \&\&\&\& {} \\
     \cgamma{\opensetgc_2}{\dtotalpullc} 
-    \arrow[rrrr, "\vartistc_{\opensetgc_2}"', bend right, color=artist] \&  \& 
+    \arrow[rrrr, "\vartistc_{\opensetgc_2}"', bend right, color=artist, maps to] \&  \& 
     \opensetgc_2 
-    \arrow[ll, "\sheafc_{\dtotalpullc}"', color=sheaf] 
-    \arrow[rr, "\sheafc_{\gtotalc}", color=sheaf] 
+    \arrow[ll, "\sheafc_{\dtotalpullc}"', color=sheaf, maps to] 
+    \arrow[rr, "\sheafc_{\gtotalc}", color=sheaf, maps to] 
     \arrow[uu, "\iota" description, hook, color=base] \& \& 
@@
 \end{tikzcd}
-\end{equation}
-\end{frame}
-
-\begin{frame}{
-    \only<1-3|handout:1-3>{Natural Transforms Between Sheaves}
-    \only<4|handout:4>{$\vartistc:\sheafc_{(\dbasec, \dtotalc)}\rightarrow \sheafc_{(\gbasec, \gtotalc)}$}
-    \only<5-|handout:5->{$\vartistc: \dsectionc \mapsto \gsectionc$}
-    }
+    \end{equation*}}
+
     \begin{equation*}
         \begin{tikzcd}[ampersand replacement=\&]
-          \onslide*<1-4|handout:1>{\sheafc_{\gbasec, \gtotalc}\gbasec} 
-          \onslide*<1-3|handout:1>{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]}  
-          \onslide*<4-|handout:2->{\arrow[rr, "\vindexpushc", color=artist]} 
+          \onslide*<0|handout:3>{\sheafc_{\gbasec, \gtotalc}} 
+          \onslide*<0|handout:3>{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]}  
           \& \& 
-          \onslide*<1-4|handout:1>{(\textcolor{functor}{\vindexpush}\sheafc_{(\gbasec, \gtotalc)})(\dbase)}
+          \onslide*<0|handout:3>{\textcolor{functor}{\vindexpush}\sheafc_{\gbasec, \gtotalc}}
           \\
           \& \& \\
-          \onslide*<1-4|handout:1>{(\textcolor{functor}{\vindexpull}\sheafc_{(\dbasec, \dtotalc)})(\gbasec)}
-          \onslide*<1-3|handout:1>{\arrow[rr, "\textcolor{functor}{\vindexpull}", color=functor]}
-          \onslide*<4-|handout:2->{\arrow[rr, "\vindexpullc", color=artist]}
-          \onslide*<2-3|handout:1>{\arrow[uu, "\textcolor{nattran}{Nat_{\gbasec}}", color=nattran]} 
-          \onslide*<4-|handout:2->{\arrow[uu, "\vartistc_{\vindexpushc}", color=artist]}
+          \onslide*<0|handout:3>{\textcolor{functor}{\vindexpull}\sheafc_{\dbasec, \dtotalc}}
+          \onslide*<0|handout:3>{\arrow[uu, "\vartistc_{\vindexpushc}", color=artist, maps to]}
           \&  \& 
-          \onslide*<1-4|handout:1->{\sheafc_{\dbasec, \dtotalc}} 
-          \onslide*<2-3|handout:1>{\arrow[uu, "\textcolor{nattran}{Nat_{\dbasec}}"', color=nattran]}
-          \onslide*<4-|handout:2->{\arrow[uu, "\vartistc_{\vindexpullc}"', color=artist]}
-          \onslide*<3|handout:1>{\arrow[lluu, " "' description, dashed, color=nattran]}
-          \onslide*<4-|handout:2->{\arrow[lluu, "\vartistc"' description, dashed, color=artist]}
-        \end{tikzcd}
-      \end{equation*}
-
-    \onslide<3|handout:3>{
+          \onslide*<0|handout:3>{\sheafc_{\dbasec, \dtotalc}}
+          \onslide*<0|handout:3>{\arrow[ll, "\textcolor{functor}{\vindexpull}", color=functor]} 
+          \onslide*<0|handout:3>{\arrow[uu, "\vartistc_{\vindexpullc}"', color=artist, maps to]}
+          \onslide*<4-|handout:3>{\arrow[lluu, "\vartistc"' description, dashed, color=artist, maps to]}
+        \end{tikzcd}
+      \end{equation*}
+
+    \onslide<0|handout:3>{
         \begin{equation*}
@@
     \textcolor{nattran}{Nat}_{\dbase}(\sheafc_{(\dbasec, \dtotalc)}, \textcolor{functor}{\vindexpush}\sheafc_{(\gbasec, \gtotalc)})
-        \end{equation*}
-    }
-    \onslide<4|handout:4>{
-        \begin{equation*}
-            \vartistc_{\vindexpull} = \vartistc_{\vindexpush}
         \end{equation*}

### 2022-01-18 slides/main.tex
fiber and equiv

@@
 \renewcommand{\restriction}{\mathord{\upharpoonright}} %restriction w/p space
+\usepackage{relsize}%mathlarger
 \usepackage{bm}
+\usepackage{centernot}
 \usepackage{mathrsfs} % math script fonts
@@
         \begin{alertblock}{Sections}
-            \begin{aligned*}         
+            \begin{align*}         
         \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}
-        \end{aligned*}
+        \end{align*}
         \end{alertblock}
@@
     \begin{alertblock}{Data \cite{butlerVisualizationModelBased1989,butlerVectorBundleClassesForm1992}}
-        \begin{aligned*}
+        \begin{align*}
         \dsectionc(\dbasepointc) = &\delementc,\; \dbasepointc\in \opensetc\subset \dbasec,\; \delement \in \dfiberc\restriction_{\dbasepointc}\\
         &\delementc=\{field_0:value, \cdots field_i:value, \cdots field_n:value\}
-        \end{aligned*}
+        \end{align*}
     \end{alertblock}
@@
         \begin{alertblock}{Graphic}
-            \begin{aligned*}         
+            \begin{align*}         
             \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \coloneqq \big\{\gsectionc: \opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} \; \bigm{\vert} \pi(\gsectionc(\gbasepointc)) = \gbasepointc\;for\, all\; \gbasepointc \in \opensetgc \big\}\\
@@
             for\;example:\; d=\{x,\,y,\,r,\,g,\,b\}
-        \end{aligned*}
+            \end{align*}
     \end{alertblock}
@@
     \begin{alertblock}<5-|handout:1>{composition}
-        \begin{aligned*}
+        \begin{align*}
             \textcolor{functor}{F}(\textcolor{source}{g}) \circ  \textcolor{functor}{F}(\textcolor{source}{f}) = \textcolor{functor}{F} (\textcolor{source}{g}\circ \textcolor{source}{f})
-        \end{aligned*}
+        \end{align*}
     \end{alertblock}
     \begin{alertblock}<6-|handout:1>{identity}
-        \begin{aligned*}
+        \begin{align*}
             \textcolor{functor}{F}(\textcolor{source}{id_c}) = \textcolor{target}{id}_{\textcolor{functor}{F}(\textcolor{source}{c})}
-        \end{aligned*}
+        \end{align*}
     \end{alertblock}
@@
         \begin{alertblock}<8-|handout:1>{stalk}
-            \begin{aligned*}
+            \begin{align*}
                 \sheafc_{\dbasec, \dtotalc}\restriction_{\dbasepointc}\coloneqq \lim\limits_{\opensetc\ni \dbasepointc} \Gamma(\opensetc, \dtotalc\restriction_{\opensetc}) \\
                  \dfiberc_{\dbasepointc} \subset  \sheafc_{\dbasec, \dtotalc}\restriction_{\dbasepointc}
-            \end{aligned*}
+            \end{align*}
         \end{alertblock}
@@
             %germ  is tau at limit, \tau(k) \in germ
-            \begin{aligned*}
+            \begin{align*}
                 \dsectionc(\dbasepointc) \in \sheafc_{\dbasec, \dtotalc}\restriction_{\dbasepointc}
-            \end{aligned*}
+            \end{align*}
         \end{alertblock}
@@
 \begin{frame}
-    \frametitle<3|handout:3>{Functor: \vindexc: \mathcal{\gbasec} \rightarrow \mathcal{\dbasec}}
+    \frametitle<3|handout:3>{Functor: $\vindexc: \mathcal{\gbasec} \rightarrow \mathcal{\dbasec}$}
     \frametitle<1|handout:1>{Data}
@@
 
-\begin{frame}{Natural Transformation $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\mapsto} \textcolor{functor}{G}$}
+\begin{frame}{Natural Transformation $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\Rightarrow} \textcolor{functor}{G}$}
 
@@
         \begin{tikzcd}[ampersand replacement=\&]
-            \& {} \arrow[dd, "\alpha" , maps to, shorten=1.5em, color=nattran] \&             \\
+            \& {} \arrow[dd, "\alpha" , Rightarrow, shorten=1.5em, color=nattran] \&             \\
           \textcolor{source}{\mathcal{C}} \arrow[rr, "F", bend left, color=functor, maps to] \arrow[rr, "G"', bend right, color=functor, maps to] \& \& \textcolor{target}{\mathcal{D}} \\
@@
 
-\begin{frame}{Natural Transformation $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\rightarrow} \textcolor{functor}{G}$}
+\begin{frame}{Natural Transformation $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\Rightarrow} \textcolor{functor}{G}$}
     \begin{equation*}
@@
             \onslide*<2-|handout:1>{\arrow[dd, "F(f)"', color=target]}
-            \onslide*<4-5|handout:1>{\arrow[rr, "\alpha_{c}", bend left, color=nattran, maps to]}  
-            \onslide*<6|handout:0>{\arrow[rr, "\alpha_{c}", color=nattran, maps to]} \& 
+            \onslide*<4-5|handout:1>{\arrow[rr, "\alpha_{c}", bend left, color=nattran, Rightarrow]}  
+            \onslide*<6|handout:0>{\arrow[rr, "\alpha_{c}", color=nattran, Rightarrow]} \& 
             \onslide*<1-4|handout:1>{\textcolor{source}{c}} 
@@
             {} 
-            \onslide*<4|handout:1>{\arrow[rr, dotted, color=nattran, maps to]}
-            \onslide*<5|handout:0>{\arrow[rr, color=nattran, maps to]} \& \& {}                      \\
+            \onslide*<4|handout:1>{\arrow[rr, dotted, color=nattran, Rightarrow]}
+            \onslide*<5|handout:0>{\arrow[rr, color=nattran, Rightarrow]} \& \& {}                      \\
             \onslide*<2-|handout:1>{\textcolor{target}{F(c^{\prime})}} 
-            \onslide*<4-5|handout:1>{\arrow[rr, "\alpha_{c^{\prime}}"', bend right, color=nattran, maps to]} 
+            \onslide*<4-5|handout:1>{\arrow[rr, "\alpha_{c^{\prime}}"', bend right, color=nattran, Rightarrow]} 
             \onslide*<4-5|handout:0>{\arrow[rr, "\alpha_{c^{\prime}}"', color=nattran, maps to]}\& 
@@
 \begin{frame}
-    \frametitle<1-2|handout:1>{Data Space: $\vartistc_{\vindexpushc}: \sheafc_{\dbasec, \dtotalc} \textcolor{artist}{\mapsto} \vindexpushc\sheafc_{\gbasec, \gtotalc}$}
-    \frametitle<3-4|handout:2>{Display Space: $\vartist_{\vindexpullc}: \vindexpullc\sheafc_{\dbasec, \dtotalc} \rightarrow \sheafc_{\gbasec, \gtotalc}$}
+    \frametitle<1-2|handout:1>{Data Space: $\vartistc_{\opensetc}: \sheafc_{\dbasec, \dtotalc} \textcolor{artist}{\Rightarrow} \vindexpushc\sheafc_{\gbasec, \gtotalc}$}
+    \frametitle<3-4|handout:2>{Display Space: $\vartist_{\opensetgc}: \vindexpullc\sheafc_{\dbasec, \dtotalc} \textcolor{artist}{\Rightarrow} \sheafc_{\gbasec, \gtotalc}$}
     \frametitle<5|handout:3->{\textcolor{artist}{Artist} \vartistc}
@@
         \arrow[dd, "\iota^*"', hook, color=set] 
-        \arrow[rrrr, "\vartistc_{\opensetc_1}", bend left, color=artist, maps to] \&  \& 
+        \arrow[rrrr, "\vartistc_{\opensetc_1}", bend left, color=artist, Rightarrow] \&  \& 
         \opensetc_1 
@@
         {} 
-        \arrow[rrrr, dotted, color=artist, maps to] \&\&\&\& {} \\
+        \arrow[rrrr, dotted, color=artist,Rightarrow] \&\&\&\& {} \\
         \cgamma{\opensetc_2}{\dtotalc} 
-        \arrow[rrrr, "\vartistc_{\opensetc_2}"', bend right, color=artist, maps to] \&  \& 
+        \arrow[rrrr, "\vartistc_{\opensetc_2}"', bend right, color=artist, Rightarrow] \&  \& 
         \opensetc_2 
@@
     \arrow[dd, "\iota^*"', hook, color=set] 
-    \arrow[rrrr, "\vartistc_{\opensetgc_1}", bend left, color=artist, maps to] \&  \& 
+    \arrow[rrrr, "\vartistc_{\opensetgc_1}", bend left, color=artist, Rightarrow] \&  \& 
     \opensetgc_1 
@@
     {} 
-    \arrow[rrrr, dotted, color=artist, maps to] \&\&\&\& {} \\
+    \arrow[rrrr, dotted, color=artist, Rightarrow] \&\&\&\& {} \\
     \cgamma{\opensetgc_2}{\dtotalpullc} 
-    \arrow[rrrr, "\vartistc_{\opensetgc_2}"', bend right, color=artist, maps to] \&  \& 
+    \arrow[rrrr, "\vartistc_{\opensetgc_2}"', bend right, color=artist, Rightarrow] \&  \& 
     \opensetgc_2 
@@
         \begin{tikzcd}[ampersand replacement=\&]
-          \onslide*<0|handout:3>{\sheafc_{\gbasec, \gtotalc}} 
-          \onslide*<0|handout:3>{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]}  
+          \onslide*<0|handout:3->{\sheafc_{\gbasec, \gtotalc}} 
+          \onslide*<0|handout:3->{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]}  
           \& \& 
-          \onslide*<0|handout:3>{\textcolor{functor}{\vindexpush}\sheafc_{\gbasec, \gtotalc}}
+          \onslide*<0|handout:3->{\textcolor{functor}{\vindexpush}\sheafc_{\gbasec, \gtotalc}}
           \\
           \& \& \\
-          \onslide*<0|handout:3>{\textcolor{functor}{\vindexpull}\sheafc_{\dbasec, \dtotalc}}
-          \onslide*<0|handout:3>{\arrow[uu, "\vartistc_{\vindexpushc}", color=artist, maps to]}
+          \onslide*<0|handout:3->{\textcolor{functor}{\vindexpull}\sheafc_{\dbasec, \dtotalc}}
+          \onslide*<0|handout:3->{\arrow[uu, "\vartistc_{\opensetgc}", color=artist, Rightarrow]}
           \&  \& 
-          \onslide*<0|handout:3>{\sheafc_{\dbasec, \dtotalc}}
-          \onslide*<0|handout:3>{\arrow[ll, "\textcolor{functor}{\vindexpull}", color=functor]} 
-          \onslide*<0|handout:3>{\arrow[uu, "\vartistc_{\vindexpullc}"', color=artist, maps to]}
-          \onslide*<4-|handout:3>{\arrow[lluu, "\vartistc"' description, dashed, color=artist, maps to]}
+          \onslide*<0|handout:3->{\sheafc_{\dbasec, \dtotalc}}
+          \onslide*<0|handout:3->{\arrow[ll, "\textcolor{functor}{\vindexpull}", color=functor]} 
+          \onslide*<0|handout:3->{\arrow[uu, "\vartistc_{\opensetc}"', color=artist, Rightarrow]}
+          \onslide*<4-|handout:3->{\arrow[lluu, "\vartistc"' description, dashed, color=artist, Rightarrow]}
         \end{tikzcd}
@@
         \begin{equation*}
-        \textcolor{nattran}{Nat}_{\gbasec}(\textcolor{functor}{\vindexpull}\sheafc_{(\dbasec, \dtotalc)}, \sheafc_{(\gbasec, \gtotalc)}) = 
-    \textcolor{nattran}{Nat}_{\dbase}(\sheafc_{(\dbasec, \dtotalc)}, \textcolor{functor}{\vindexpush}\sheafc_{(\gbasec, \gtotalc)})
+        \textcolor{nattran}{Nat}_{\opensetgc}(\textcolor{functor}{\vindexpull}\sheafc_{\dbasec, \dtotalc}, \sheafc_{\gbasec, \gtotalc}) = 
+    \textcolor{nattran}{Nat}_{\opensetc}(\sheafc_{\dbasec, \dtotalc}, \textcolor{functor}{\vindexpush}\sheafc_{\gbasec, \gtotalc})
+        \end{equation*}
+    }
+    \onslide<0|handout:4>{
+        \begin{equation*}
+            \sheafc_{\dbasec, \dtotalc} \textcolor{artist}{\xrightarrow{\vartistc}} \sheafc_{\gbasec, \gtotalc}
+        \end{equation*}
+        \begin{equation*}
+             (\sheafc_{\dbasec,\dtotalc} \textcolor{functor}{\xmapsto{\vindexpullc}} \vindexpullc \sheafc_{\dbasec,\dtotalc}) \textcolor{artist}{\xRightarrow{\vartistc_{\opensetgc}}} \sheafc_{\gbasec, \gtotalc} \tag{$\vartistc_{\opensetgc} \circ \vindexpullc = \vartistc$}
+        \end{equation*}
+        \begin{equation*}
+            (\sheafc_{\dbasec,\dtotalc} \textcolor{artist}{\xRightarrow{\vartistc}} \sheafc_{\gbasec, \gtotalc}) \textcolor{functor}{\xmapsto{\vindexpushc}} \vindexpushc \sheafc_{\gbasec, \gtotalc} \tag{$\vindexpushc \circ \vartistc = \vartistc_{\opensetc}$}
+        \end{equation*}
+        \begin{equation*}
+            ((\sheafc_{\dbasec,\dtotalc} \textcolor{functor}{\xmapsto{\vindexpullc}} \vindexpullc \sheafc_{\dbasec,\dtotalc}) \textcolor{artist}{\xRightarrow{\vartistc_{\opensetgc}}} \sheafc_{\gbasec, \gtotalc})\textcolor{functor}{\xmapsto{\vindexpushc}} \vindexpushc \sheafc_{\gbasec, \gtotalc} \tag{$\vindexpushc \circ \vartistc_{\opensetgc} \circ \vindexpullc = \vartistc_{\opensetc}$}
         \end{equation*}
@@
             \onslide*<2-|handout:1>{\textcolor{set}{\big\{}\gsectionc:\opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} 
-              \bigm{\vert} \vartistc\circ (\dfuncc \circ \vartistc^{-1}[\gsectionc]) = \dfuncc^{\prime} \circ \gsectionc \textcolor{set}{\big\}}} 
+              \bigm{\vert} \vartistc \circ \dfuncc_{\dsectionc} \circ \dsectionc = \dfuncc_{\gsectionc} \circ \gsectionc \textcolor{set}{\big\}}} 
             \onslide*<2-|handout:1>{\arrow[ll, "\textcolor{set}{\mathlarger{\mathlarger{\supset}}}" description, hook', color=white]} \\ 
@@
             \onslide*<1-|handout:1>{\opensetgc} 
-            \onslide*<1-|handout:1>{\arrow[uu, "\sheafc_{\gtotalc}", color=sheaf]} 
-            \onslide*<2-|handout:1>{\arrow[rruu, "\vartistc(\sheafc_{\dtotalc})\coloneqq \sheafc_{\vartistc}" description, dashed, color=sheaf]} 
+            \onslide*<1-|handout:1>{\arrow[uu, "\sheafc_{\gtotalc}", color=sheaf, maps to]} 
+            \onslide*<2-|handout:1>{\arrow[rruu, "\vartistc(\sheafc_{\dtotalc})\coloneqq \sheafc_{\vartistc}" description, dashed, color=sheaf, maps to]} 
             \&  \& 
@@
 
-\begin{frame}{$\dfuncc = (\hat\dfuncc, \hat\dfuncc_*, \tilde\dfuncc)$?}
-\begin{equation*}
-    \begin{tikzcd}[ampersand replacement = \&, row sep=large]
-        {\Gamma(U, E)} 
-        \arrow[r, "\hat\varphi_*"]                             
-        \& 
-        {\Gamma(U^{\prime}, \hat{\varphi_*}E)} 
-        \arrow[r, "\tilde{\varphi}"]                                                                                                             
-        \& {\Gamma(U^{\prime}, \tilde{\varphi}\hat{\varphi_*}E)} \\
-        {U \subset K} 
-        \arrow[r, "\hat\varphi"] 
-        \arrow[u, "{\mathcal{O}_{K,E}}"] 
-        \& 
-        {U^{\prime}\subset K^{\prime}}
-        \arrow[u, "{\mathcal{O}_{K^{\prime}, \hat\varphi_*E}}", dashed] 
-        \arrow[ru, "{\mathcal{O}_{\hat\varphi K, \tilde\varphi\hat\varphi_*E}}"', dotted] 
-        \&                                                                                                          
-    \end{tikzcd}                                   
-\end{equation*}
-\end{frame}
-
-\begin{frame}{Continuty Equivariance $\vindexpullc\hat\dfuncc = \hat\dfuncc|_{\vindexprec}$}
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement =\&]
-            S 
-            \arrow[r, "\xi"] 
-            \arrow[d, "\xi^*\hat\varphi"'] 
-            \& 
-            K 
-            \arrow[d, "\hat\varphi"] \\
-            S^{\prime} 
-            \arrow[r, "\xi"]
-            \& 
-            K^{\prime}
-            \end{tikzcd}
-    \end{equation*}
-\end{frame}
-\begin{frame}{Field Equivariance $\tilde\dfuncc \cong \tilde\dfuncc_\gtotalc$}
-    \begin{equation*}
-    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-        {\Gamma(U, E)} 
-        \arrow[r, "A"] 
-        \arrow[d, "\tilde\varphi\circ\hat\varphi_*"'] 
-        \& 
-        {\Gamma(W, H)} 
-        \arrow[d, "\tilde\varphi_H\circ\xi_*\hat{\varphi}_*"] \\
-        {\Gamma(\hat\varphi U, \tilde\varphi\hat\varphi_*E)} 
-        \arrow[r, "A"] 
-        \& 
-        {\Gamma(\xi^*\hat\varphi W, \tilde\varphi_H\xi^*\hat\varphi_*H)}    
-    \end{tikzcd}
-    \end{equation*}
-\end{frame}
-
-\begin{frame}{Equivariant Artist}
-    \begin{equation*}
-    \begin{tikzcd}[ampersand replacement=\&]
-        {\mathcal{O}_{K, E}} 
-        \arrow[r, "A"] 
-        \arrow[d, "\varphi"'] \& 
-        {\mathcal{O}_{S,H}} 
-        \arrow[d, "\varphi_H"] \\
-        {\mathcal{O}_{K^{\prime}, E^{\prime}}} 
-        \arrow[r, "A"]    \& 
-        {\mathcal{O}_{S^{\prime}, H^{\prime}}}    
-        \end{tikzcd} 
-    \end{equation*}
-        \begin{block}
-            \vartistc\ is equivariant if and only if there exists a $\dfuncc_H$ s.t the diagram commutes. 
-            \begin{equation*}
-            \end{equation*}
-        \end{block}
-    \end{frame}
-
-\begin{frame}{More than one field?}
-    \begin{block}{} 
+\begin{frame}{$\dfuncc = (\dfunchc, \dfuncpullc, \dfunctc)$}
+    \begin{tikzcd}[ampersand replacement = \&]
+        \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
+        \arrow[rr, "\dfuncpullc", color=action] 
+        \&  \& 
+        \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}} 
+        \arrow[rr, "\dfunctc", color=action] 
+        \&  \& 
+        \cgamma{\opensetc^{\prime}}{\dfunctc\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}} \\
+        \&  \&  
+        \&  \& \\
+        \opensetc 
+        \arrow[uu, "{\sheafc_{\dbasec, \dtotalc}}", maps to, color=sheaf]                 
+        \&  \& 
+        \opensetc^{\prime} 
+        \arrow[ll, "\dfunchc", color=action] 
+        \arrow[uu, "{\dfuncpullc\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf] 
+        \arrow[rruu, "{\dfunctc\dfuncpullc\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf] 
+        \&  \& 
+    \end{tikzcd}
+    \begin{itemize}
+        \item Base $\dbasec$ Type stays same: $\dfunchc: \dbasec \rightarrow \dbasec$
+        \item Fiber $\dfiberc$ Type stays the same: $\dfunctc: \dfiberc \rightarrow \dfiberc$
+        \item Section Transform: $\dfuncc: \dsectionc^{a}\restriction_{\opensetc} \mapsto \dsectionc^{b}\restriction_{\opensetc^{\prime}}$
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{\dfunctc}
+    $\dfunctc$ applies a monoid action $\monoidc$ on a fiber set $\dfiberc$
+    \begin{equation*}
+        \dfunctc: \monoidc \times \dfiberc \rightarrow \dfiberc
+    \end{equation*}
+    meaning that, by definition of a monoid, the following properties are satisfied
+    \begin{description}
+        \item[identity]: $\dfunctc(\textcolor{monoid}{e}, \delementc) = \delementc, \delementc \in \dfiberc, \textcolor{monoid}{e} \in \monoidc$
+        \item[associativity]  $\dfunctc(\textcolor{monoid}{g}, \dfunctc(\textcolor{monoid}{h}, \delementc)) = \dfunctc(\textcolor{monoid}{g}\circ \textcolor{monoid}{h}, \delementc), \textcolor{monoid}{g}, \textcolor{monoid}{h} \in \monoidc, \delementc \in \dfiberc$
+    \end{description}
+\begin{block}{Multivariate?} 
     \label{def:category:F} The fiber category $\mathcal{\dfiberc}$ is a monoidal category\cite{milewskiCategoryTheoryProgrammers}, which means it is a category equipped with a bifunctur $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$
     \begin{itemize}
-          \item \textit{object} a monoid object \dfiber.
-A monoid is a set with a binary operator that is associative, closed, and for which the set contains an identity element $i$ \cite{nlab:monoid}.
-          \item \textit{morphisms} $\dfiber \otimes \dfiber \rightarrow \dfiber$ and unit $i \rightarrow \dfiber$
+      \item \textit{object} set \dfiberc with actions \monoidc \cite{nlab:monoid}.
+      \item \textit{morphisms} $\dfiberc \otimes \dfiberc \rightarrow \dfiberc$, $\monoidc \otimes \monoidc \rightarrow \monoidc$ and unit
     \end{itemize}
 \end{block}
+\end{frame}
+
+\begin{frame}{Continuity Equivariance: $\dfunchc_{\gbasec}\circ \vindexc = \vindexc \circ \dfunchc_{\dbasec}$}
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement=\&]
+        \opensetgc 
+        \arrow[rr, "\vindexc", color=functor] 
+        \arrow[dd, "\dfunchc_{\gbasec}"', dashed, color=action] 
+        \&  \& 
+        \opensetc 
+        \arrow[dd, "\dfunchc_{\dbasec}", color=action] \\
+        \&  \& \\
+        \opensetgc^{\prime} 
+        \arrow[rr, "\vindexc"', color=functor]
+        \&  \& 
+        \opensetc^{\prime}\\
+        \underbrace{\dbasec \times {[0,1]}^{m}}_{\gbasec} \arrow[rr,"\vindex", maps to, color=functor] \& \& \dbasec
+        \end{tikzcd}
+    \end{equation*}
+
+\end{frame}
+
+\begin{frame}{Field Equivariance: $\vartist \circ (\dfunctc_{\dtotalc}\circ \dfuncpullc_{\dbasec}) =  (\dfunctc_{\gtotalc} \circ \dfuncpullc_{\gbasec}) \circ \vartist$}
+    \begin{equation*}
+    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
+            \cgamma{\opensetc^{\prime}}{\dfuncpullc_{\dbasec}\dtotalc\restriction_{\opensetc^{\prime}}} 
+            \arrow[rr, "\vartistc", color=artist, Rightarrow] 
+            \arrow[dd, "\dfunctc_{\dtotalc}\circ \dfuncpullc_{\dbasec}"', color=action] 
+            \&  \& 
+            \vartistc[\dfuncpullc\sheafc_{\dbasec,\dtotalc}] \subset \cgamma{\opensetgc^{\prime}}{\dfuncpullc_{\gbasec}\gtotalc\restriction_{\opensetgc^{\prime}}} 
+            \arrow[dd, "\dfunctc_{\gtotalc}\circ \dfuncpullc_{\gbasec}", dashed, color=action, shift right=5em] \\
+            \&  \&
+            \\
+            \cgamma{\opensetc^{\prime}}{\dfunctc_{\dtotalc}\dfuncpullc_{\dbasec}\dtotalc\restriction_{\opensetc^{\prime}}} 
+            \arrow[rr, "\vartistc", color=artist, Rightarrow]
+            \&  \& 
+            \vartistc[\dfunctc\dfuncpullc\sheafc_{\dbasec,\dtotalc}] \subset \cgamma{\opensetgc^{\prime}}{\dfunctc_{\gtotalc}\dfuncpullc_{\gbasec}\gtotalc\restriction_{\opensetgc^{\prime}}}
+    \end{tikzcd}
+    \end{equation*}
+\end{frame}
+
+\begin{frame}{Equivariant Artist}
+    \begin{equation*}
+    \begin{tikzcd}[ampersand replacement=\&]
+        \sheafc_{\dbasec, \dtotalc} 
+        \arrow[r, "\vartistc", Rightarrow, color=artist] 
+        \arrow[d, "\dfuncc_{\dsectionc}"', color=action] \& 
+        \sheafc_{\vartistc} 
+        \arrow[d, "\dfuncc_{\gsectionc}", color=action, dashed] \\
+        \dfuncc_{\dsectionc}\sheafc_{\dbasec, \dtotalc} 
+        \arrow[r, "\vartistc", Rightarrow, color=artist]    \& 
+        \dfuncc_{\gsectionc}\sheafc_{\vartistc}    
+        \end{tikzcd} 
+    \end{equation*}
+
+    \begin{proposition}{}
+        \begin{align*}
+            \vartistc(\dsectionc^{a}) = \vartistc(\dsectionc^{b}) &\centernot\implies \dsectionc^a = \dsectionc^b \\
+            \vartistc(\dsectionc^{a})  = \vartistc(\dsectionc^{b}) &\implies \vartistc(\dfuncc_{\dsectionc}(\dsectionc^{a})) = \vartistc(\dfuncc_{\dsectionc}(\dsectionc^{b})) \
+        \end{align*}
+    \end{proposition}
+
+\end{frame}
+
+\begin{frame}
+    dummy sheet
 \end{frame}
@@
 \end{frame}
-%\end{comment}
-
 \end{comment}

### 2022-01-26 slides/main.tex
fiber bundle def fix

@@
 \usepackage{tabularx}
+\usepackage{enumitem}
 \usepackage{booktabs}
@@
     \begin{columns}
-        \column{0.39\textwidth}
+        \column{0.31\textwidth}
         \begin{equation*}
@@
             \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-                  \onslide*<2-6|handout:1-2>{\dfiberc}
+              \onslide*<3-6|handout:1-2>{\dfiberc}
               \onslide*<7|handout:3>{\gfiberc} 
               \onslide*<4-|handout:1->{\arrow[r, hook]} \& 
-                  \onslide*<3-6|handout:1-2>{\dtotalc} 
+              \onslide*<1-6|handout:1-2>{\dtotalc} 
               \onslide*<7|handout:3>{\gtotalc}
@@
                \& 
-                   \onslide*<1-6|handout:1-2>{\dbasec}
+               \onslide*<2-6|handout:1-2>{\dbasec}
                \onslide*<7|handout:3>{\gbasec} 
@@
         \end{equation*}
-     \column{0.59\textwidth}
+     \column{0.69\textwidth}
      \only<1-5|handout:1>{
-        \begin{description}
-            \item<1-5|handout:1>[\textcolor{base}{Base Space}]{$( \dbasec, \mathcal{T})$, $\dbasepointc \in \opensetc \subset \dbasec$}
-            \item<2-5|handout:1>[\textcolor{fiber}{Fiber Space}]{$\dfiberc\restriction_{\dbasepointc} = \pi^{-1}(\dbasepointc)$}
-            \item<3-5|handout:1>[\textcolor{total}{Total Space}]{ $\dtotalc\restriction_{\opensetc} = \dbasec\restriction_{\opensetc} \times \dfiberc\restriction_{\opensetc}$ \\ for all $\opensetc \subset K$}
-        \end{description}}
+        \begin{description}[style=nextline]
+            \item<1-5|handout:1>[\textcolor{total}{Total Space}]{$(\dtotalc, \mathcal{T}_{\dtotalc})$}
+            \item<2-5|handout:1>[\textcolor{base}{Base Space}]{$(\dbasec, \mathcal{T}_{\dbasec}),\; \dbasepointc \in \opensetc \subseteq \dbasec$}
+            \item<3-5|handout:1>[\textcolor{fiber}{Fiber Space}]{$\dfiberc\restriction_{\dbasepointc} = \pi^{-1}(\dbasepointc), \dfiberc = \dfiberc\restriction_{\dbasepointc}\; \forall \dbasepointc \in \dbasec$}
+        \end{description}
+        }
     \only<6|handout:2>{
-        \begin{description}
-            \item[\textcolor{base}{Continuity Type \dbasec}]{how data elements are organized (topological properties) \cite{wilkinsonGrammarGraphics2005,munznerWhatDataAbstraction2014})}
-            \item[\textcolor{fiber}{Fields Type \dfiberc}] {generalization of a schema - named and typed date fields \cite{spivakSIMPLICIALDATABASES, spivakDatabasesAreCategories2010}}
-            \item[\textcolor{total}{Data Type \dtotalc}]{continuity type + field type}
+        \begin{description}[style=nextline]
+            \item[\textcolor{total}{Data \dtotalc}]{continuity + fields}
+            \item[\textcolor{base}{Continuity \dbasec}]{how data elements are organized (topological properties) \cite{wilkinsonGrammarGraphics2005}, index (key) space \cite{munznerWhatDataAbstraction2014})}
+            \item[\textcolor{fiber}{Fields \dfiberc}] {generalization of a schema - named and typed date fields \cite{spivakSIMPLICIALDATABASES, spivakDatabasesAreCategories2010}}
         \end{description}
@@
         \only<7|handout:3>{
-            \begin{description}
-                \item[\textcolor{base}{Continuity Type \gbasec}]{how pixels are organized in the display space}
-                \item[\textcolor{fiber}{Display Type \gfiberc}] {display specifications}
-                \item[\textcolor{total}{Graphic Type \gtotalc}]{continuity type + display type}
+            \begin{description}[style=nextline]
+                \item[\textcolor{total}{Graphic \gtotalc}]{virtual rendered output}
+                \item[\textcolor{base}{Continuity \gbasec}]{topology of graphic in rendered space}
+                \item[\textcolor{fiber}{Display \gfiberc}] {generalization of renderer specifications, e.g. \{xy,rgba\}, \{xy, cymk\}, \{xyz, rgba\}}
             \end{description}
@@
     \pause 
-    \only<5|handout:1>{
-        \begin{alertblock}{Sections}
-            \begin{align*}         
-        \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}
-        \end{align*}
+    \only<4-5|handout:1>{
+        \begin{alertblock}{}
+            \begin{description}[style=nextline]
+                \item[\textcolor{section}{Sections}]{$\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}$}
+                \item[Locally Trivial] {for every point $\dbasepointc \in \dbasec$, there exists an open neighborhood $\dbasepointc \in \opensetc \subseteq \dbasec$ s.t.
+there is a homeomorphism $\pi^{-1}(\opensetc)\xrightarrow{\varphi} \opensetc \times \dfiberc$}
+                \item[(Globally) Trivial]{$\dtotalc = \dbasec \times \dfiberc$}
+            \end{description}
         \end{alertblock}
@@
     \only<6|handout:2>{
-    \begin{alertblock}{Data \cite{butlerVisualizationModelBased1989,butlerVectorBundleClassesForm1992}}
-        \begin{align*}
-        \dsectionc(\dbasepointc) = &\delementc,\; \dbasepointc\in \opensetc\subset \dbasec,\; \delement \in \dfiberc\restriction_{\dbasepointc}\\
-        &\delementc=\{field_0:value, \cdots field_i:value, \cdots field_n:value\}
-        \end{align*}
+    \begin{alertblock}{}
+        \begin{description}[style=nextline]
+            \item[\textcolor{section}{Data} \cite{butlerVisualizationModelBased1989,butlerVectorBundleClassesForm1992}]{$
+        \dsectionc(\dbasepointc) = \;\delementc,\; \dbasepointc\in \opensetc \subseteq \dbasec,\; \delementc \in \dfiberc\restriction_{\dbasepointc}$} 
+        \item[]{$\delementc=\{field_0:value, \cdots field_i:value, \cdots field_n:value\}$}
+        \end{description}
     \end{alertblock}
@@
     \only<7|handout:3>{
-        \begin{alertblock}{Graphic}
-            \begin{align*}         
-            \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \coloneqq \big\{\gsectionc: \opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} \; \bigm{\vert} \pi(\gsectionc(\gbasepointc)) = \gbasepointc\;for\, all\; \gbasepointc \in \opensetgc \big\}\\
-            \vspace{5pt}\\
-            \gsectionc(\gbasepointc) = d,\; \gbasepointc\in \opensetgc\subset \gbasec,\; d \in \dfiberc\restriction_{\gbasepointc}\\
-            for\;example:\; d=\{x,\,y,\,r,\,g,\,b\}
-            \end{align*}
+        \begin{alertblock}{}
+            \begin{description}[style=nextline]
+                \item[\textcolor{section}{Graphic}]{$\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \coloneqq \big\{\gsectionc: \opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} \; \bigm{\vert} \pi(\gsectionc(\gbasepointc)) = \gbasepointc\;for\, all\; \gbasepointc \in \opensetgc \big\}$}
+                \item[] {$\gsectionc(\gbasepointc) = \gelementc,\; \gbasepointc\in \opensetgc\subseteq \gbasec,\; \gelementc \in \dfiberc\restriction_{\gbasepointc}$}
+                \item[] $\gelementc=\{x,\,y,\,r,\,g,\,b\}$
+
+        \end{description}    
     \end{alertblock}
@@
 
+\begin{frame}{What is \dfuncc?}
+    \begin{block}{Fiber Bundle Category}
+        \begin{description}[style=newline]
+            \item[object] $\dfiberc \hookrightarrow \dtotalc \xrightarrow{\pi} \dbasec$
+            \item[morphisms] $\dfuncc: \dtotalc \rightarrow \dtotalc$ 
+        \end{description}
+    \end{block}
+
+    \begin{block}{fiber transform equivariance}
+        The fiber $\dfiberc$ is the sole object of an arbitrary category $\mathcal{C}$. The bundle morphisms 
+        \begin{equation*}
+            \dfuncc \in Hom(\dfiberc,\dfiberc)
+        \end{equation*} 
+        are the set of morphisms that also preserve fiber structure.
+    \end{block}
+\end{frame}
+
 \begin{frame}{$\dfuncc = (\dfunchc, \dfuncpullc, \dfunctc)$}
@@
         \&  \& 
-        \cgamma{\opensetc^{\prime}}{\dfunctc\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}} \\
+        \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}} \\
         \&  \&  
@@
         \arrow[uu, "{\dfuncpullc\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf] 
-        \arrow[rruu, "{\dfunctc\dfuncpullc\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf] 
+        \arrow[rruu, "{\dfuncpullc\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf] 
         \&  \& 
@@
         \item Base $\dbasec$ Type stays same: $\dfunchc: \dbasec \rightarrow \dbasec$
-        \item Fiber $\dfiberc$ Type stays the same: $\dfunctc: \dfiberc \rightarrow \dfiberc$
+        \item Fiber $\dfiberc$ Type stays the same: $\dfunctc: \dfuncpullc \dtotalc \rightarrow \dfuncpullc\dtotalc$ s.t. $\pi(\dfuncpullc\dtotalc) = \pi(\dfunctc(\dfuncpullc\dtotalc))$ 
         \item Section Transform: $\dfuncc: \dsectionc^{a}\restriction_{\opensetc} \mapsto \dsectionc^{b}\restriction_{\opensetc^{\prime}}$
@@
 
-\begin{frame}{\dfunctc}
-    $\dfunctc$ applies a monoid action $\monoidc$ on a fiber set $\dfiberc$
-    \begin{equation*}
-        \dfunctc: \monoidc \times \dfiberc \rightarrow \dfiberc
-    \end{equation*}
-    meaning that, by definition of a monoid, the following properties are satisfied
-    \begin{description}
-        \item[identity]: $\dfunctc(\textcolor{monoid}{e}, \delementc) = \delementc, \delementc \in \dfiberc, \textcolor{monoid}{e} \in \monoidc$
-        \item[associativity]  $\dfunctc(\textcolor{monoid}{g}, \dfunctc(\textcolor{monoid}{h}, \delementc)) = \dfunctc(\textcolor{monoid}{g}\circ \textcolor{monoid}{h}, \delementc), \textcolor{monoid}{g}, \textcolor{monoid}{h} \in \monoidc, \delementc \in \dfiberc$
-    \end{description}
-\begin{block}{Multivariate?} 
-    \label{def:category:F} The fiber category $\mathcal{\dfiberc}$ is a monoidal category\cite{milewskiCategoryTheoryProgrammers}, which means it is a category equipped with a bifunctur $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$
-    \begin{itemize}
-      \item \textit{object} set \dfiberc with actions \monoidc \cite{nlab:monoid}.
-      \item \textit{morphisms} $\dfiberc \otimes \dfiberc \rightarrow \dfiberc$, $\monoidc \otimes \monoidc \rightarrow \monoidc$ and unit
-    \end{itemize}
-\end{block}
-\end{frame}
-
-\begin{frame}{Continuity Equivariance: $\dfunchc_{\gbasec}\circ \vindexc = \vindexc \circ \dfunchc_{\dbasec}$}
+\begin{frame}{Continuity Equivariance: $\dfunchc_{\dbasec}\circ \vindexc = \vindexc \circ \dfunchc_{\gbasec}$}
     \begin{equation*}
@@
     \end{proposition}
-
-\end{frame}
-
-\begin{frame}
-    dummy sheet
-\end{frame}
-
-\begin{comment}
-\subsection{Contributions}
-\begin{frame}{Contributions}
-    \begin{description}
-        \item[Topological] continuity  
-        \item[Equivariant] monoid action 
-        \item[Artist] Matplotlib $\textit{artist}: \textit{data} \rightarrow \textit{graphic}$ 
-        \item [Model] 
-    \end{description}
-\end{frame}
-
-\section{Math}
-
-\begin{frame}{Topological Equivariant Artist Model}
-    \centering
-    An artist $\mathscr{\vartist}$ is an equivariant map 
-    \begingroup 
-    \Huge
-    \begin{equation*}
-        \mathscr{\vartist}: \mathscr{\dtotal} \rightarrow \mathscr{\gtotal}
-    \end{equation*}
-    \endgroup
-    from data $\mathscr{\dtotal}$ space to graphic $\mathscr{\gtotal}$ space. 
-\end{frame}
-
-\subsection{Data}
-\begin{frame}{Model data as a fiber bundle \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}}
-    A fiber bundle is a tuple $(\dtotal,\,\dbase,\,\pi ,\,\dfiber)$ defined by the map $\pi$
-    \begin{equation*}
-        \label{eq:fiber_bundle}
-        \begin{tikzcd}[ampersand replacement=\&]
-            \dfiber \arrow[r, hook] \& \dtotal \arrow[r, "\pi"] \& \dbase
-        \end{tikzcd}
-    \end{equation*}
-    \begin{description}
-        \pause
-        \item[total space \dtotal] topology
-        \pause
-        \item[fiber space \dfiber] schema
-        \pause
-        \item[base space \dbase] continuity
-    \end{description}
-\end{frame}
-
-\begin{frame}{Encode variable types in a schema like fiber \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES}}
-    \begin{figure}[H]
-        \centering
-        \includegraphics[width=\textwidth]{figures/math/fiber.png}
-    \label{fig:data_fiber_example}
-    \end{figure}
-\end{frame}
-
-\begin{frame}{Monoids are the structure of the components of \dfiber}
-    \begin{figure}[H]
-        \centering
-        \includegraphics[height=.3\textheight]{figures/math/fiber.png}
-    \label{fig:data_fiber_example}
-    \end{figure}
-        \begin{equation*}
-            \dfiber = \dfiber_0 \times \ldots \times \dfiber_i \times \ldots \times \dfiber_n
-        \end{equation*}
-        \pause
-        Monoid actions $\monoid_{i}$ (e.g. rotation, partial ordering) define the structure on $\dfiber_i$
-        \begin{equation*}
-            \bullet: \monoid_i\times \dfiber_i \rightarrow \dfiber_i
-        \end{equation*}
-        where $\bullet$ is associative and has an identity action. 
-\end{frame}
-
-\begin{frame}{\dbase\ is an indexing space}
-    \begin{figure}[H]
-        \includegraphics[width=1\textwidth]{figures/math/base.png}
-    \label{fig:base_example}
-    \end{figure}
-\end{frame}
-
-\begin{frame}{Data are sections \dsection\ on \dtotal}
-    For any fiber bundle, there exists a map
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement=\&]
-            \dfiber \arrow[r, hook] \& \dtotal \arrow[d, "\pi"'] \\
-                              \& \dbase \arrow[u, "\dsection"', bend right]
-        \end{tikzcd}
-    \end{equation*}
-     s.t. $\pi(\dsection(\dbasepoint)) = \dbasepoint$. $\Gamma(\dtotal)$ is the set of all global sections.
-     \begin{figure}[H]
-        \includegraphics[width=1\linewidth]{figures/math/fiberbundle.png}
-        \label{fig:data_sections}
-    \end{figure}
-\end{frame}
-
-\subsection{Graphic}
-\begin{frame}{Graphic Bundle $(\gtotal,\,\gbase,\,\pi ,\,\gfiber)$} % watch dr explaination
-    \only<2->{Continuity is preserved via the many \gbasepoint\ to one \dbasepoint\ map $\vindex: \gbase\rightarrow\dbase$}
-
-    \only<1>{
-        \begin{equation*} 
-            \begin{tikzcd}[ampersand replacement=\&]
-                F \arrow[r, hook] \& E \arrow[d, "\pi"']              \& D \arrow[r, hook] \& H \arrow[d, "\pi"']              \\
-                                  \& K \arrow[u, "\tau"', bend right] \&                   \& S \arrow[u, "\rho"', bend right]
-                \end{tikzcd}
-            \end{equation*}}
-    \only<2->{
-    \begin{equation*} 
-        \begin{tikzcd}[ampersand replacement=\&]
-            F \arrow[r, hook] \& E \arrow[d, "\pi"']              \& D \arrow[r, hook] \& H \arrow[d, "\pi"']                                 \\
-                              \& K \arrow[u, "\tau"', bend right] \&                   \& S \arrow[ll, "\xi"'] \arrow[u, "\rho"', bend right]
-            \end{tikzcd}
-        \end{equation*}}
-    \only<3->{   
-    \gfiber\ is a proxy for the target display, for example $(x,\, y,\, r,\, g,\, b) \in \gfiber$}
-    \pause
-    \begin{figure}[H] %%%bring up one at a time
-        \begin{overprint}
-            \onslide<4|handout:0>\includegraphics[width=1\linewidth]{figures/math/scatter_s.png}
-            \onslide<5|handout:0>\includegraphics[width=1\linewidth]{figures/math/line_s.png}
-            \onslide<6>\includegraphics[width=1\linewidth]{figures/math/image_s.png}
-        \end{overprint}
-    \end{figure}
-\end{frame}
-
-\subsection{Artist}
-\begin{frame}{Visual bundle $(\vtotal,\,\dbase,\,\pi ,\,\vfiber)$}
-    \begin{equation*}
-        \mathcal{A}:\mathcal{E}\rightarrow\mathcal{H}
-    \end{equation*}
-\pause
-\begin{equation*}
-    \label{eq:artist_diagram}
-    \begin{tikzcd}[ampersand replacement=\&]
-        \dtotal^{\prime} \arrow[r, "\vchannel"] \arrow[rd, "\pi"'] \& \vtotal \arrow[d, "\pi"] \& \vindex^*\vtotal \arrow[r, "\vmark"] \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"'] \& \gtotal \arrow[ld, "\pi"] \\
-                                              \& \dbase                  \& \gbase \arrow[l, "\vindex"']                                              \&                    
-        \end{tikzcd}
-\end{equation*}
-\pause
-\begin{equation*}
-    \vartist: \mathcal{O}(\dtotal) \rightarrow \mathcal{O}(\gtotal)
-\end{equation*}
-\end{frame}
-
-\begin{frame}{Visualization Assembly Function}
+\end{frame}
+
+\begin{frame}{Multiple Fields: $\dfuncc = (\dfunchc, \dfuncpullc, \prod\limits_{i=0}^{n}\dfunctc_i)$} 
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement = \&, column sep=2ex]
+                \dfunctc_0 \circ \dfuncpullc \dtotalc_0 
+                \arrow[rdd, "\pi" description] 
+                \arrow[r, "\cdots \otimes \cdots", phantom] \& 
+                \dfunctc_i \circ \dfuncpullc \dtotalc_i 
+                \arrow[dd, "\pi" description] 
+                \arrow[r, "\cdots \otimes \cdots", phantom] \& 
+                \dfunctc_n \circ \dfuncpullc \dtotalc_n 
+                \arrow[ldd, "\pi" description] \& = \& 
+                \prod\limits_{i=0}^{n} 
+                \dfunctc_i \circ \dfuncpullc \dtotalc 
+                \arrow[dd, "\pi" description] \\    
+                \&\&\&\&\\    
+                \& 
+                {(\dfunchc=\dfunchc_{i \in [0,n]}) \circ \dbasec }                                             \& \& = \& 
+                \dfunchc \circ \dbasec                                                        
+                \end{tikzcd}                                              
+    \end{equation*}
+
+    \begin{block}{Add New Field to E}
+        \begin{description}[style=newline]
+            \item[construct] {$\odot:\dfiberc_{field} \times \dbasec \rightarrow \dtotalc_{field}$}
+            \item[concatenate]{$\otimes: \dtotalc_{field} \times \dtotalc \mapsto \dtotalc$ s.t. $\pi(\dtotalc_{field}) = \pi(\dtotalc)$}
+        \end{description}
+    \end{block}
+
+\end{frame}
+
+\begin{frame}{Field Equivariance}
+
+    \begin{equation*}
+        \oslash: \vartistc \times \dfiberc \rightarrow visual\; encoding
+    \end{equation*}
+
+\begin{equation*}
+    \begin{tikzcd}[ampersand replacement = \&]
+            \& 
+            {\sheafc_{\dbasec, \dtotalc}}
+            \arrow[rr, "\vartistc"]                   
+            \&  \& 
+            \sheafc_{\vartist} 
+            \arrow[rd, "\oslash_{\dfiber_{field}}"]           \&                        \\
+            \dfiberc_{field} 
+            \arrow[ru, "\otimes_{\dtotalc} \circ \odot_{\dbasec}"] 
+            \arrow[rd, "\otimes_{\dtotalc^{\prime}} \circ \odot_{\dbasec^{\prime}}"'] 
+            \& \& \& \& 
+            visual\;encoding \\
+            \& 
+            {\sheafc_{\dbasec^{\prime}, \dtotalc^{\prime}}} 
+            \arrow[rr, "\vartistc"'] 
+            \& \& 
+            \sheafc_{\vartist}^{\prime} 
+            \arrow[ru, "\oslash_{\dfiber_{field}}"'] 
+            \&                       
+    \end{tikzcd}
+\end{equation*}
+
+\end{frame}
+
+\section{Construction}
+\begin{frame}{What's in the $\vartist$ black box?}
+    \note{Add box around figure & lable it artist}
     \begin{figure}
-        \includegraphics[width=\textwidth]{figures/math/path_of_q.png}
-    \end{figure}
-    \pause
-    \begin{equation*}
-        \label{eq:nu_expanded}
-        \{\vchannel_{0}, \ldots, \vchannel_{n}\}: \{\dsection_{0}, \ldots, \dsection_{n}\} \mapsto \{\vsection_{0}, \ldots, \vsection_{n}\}
-    \end{equation*}
-    \pause
-    \begin{equation*}
-        Q = \nu \circ \tau
-    \end{equation*}
-\end{frame}
-
-\begin{frame}{Group Equivariance: Stevens' Scales \cite{stevensTheoryScalesMeasurement1946}}
+        \includegraphics[width=1\textwidth]{../paper/figures/path_of_q.png}
+    \end{figure}
+\end{frame}
+
+\begin{frame}{Encoding + Assembly Stages}
+    %E-V and Q diagram    
+\end{frame}
+\begin{frame}{Encoding + Assembly}
+    \begin{figure}
+        \includegraphics{../paper/figures/q_hat.p}
+    \end{figure}
+\end{frame}
+\begin{frame}{Domain and Range}
+    \begin{figure}
+        \begin{subfigure}{.5\textwidth}
+            \includegraphics[.5\textwidth]{../paper/figures/fiberbundle.png}
+        \end{subfigure}
+        \begin{subfigure}{.5\textwidth}
+            \includegraphics[.5\textwidth]{../paper/figures/render.png}
+        \end{subfigure}
+    \end{figure}
+\end{frame}
+
+\begin{frame}{Example $\dfunctc$: Stevens' Scales \cite{stevensTheoryScalesMeasurement1946}}
 \begin{table}[H]
@@
 
-\begin{frame}{Monoid Equivariance: Partial Orders}
-    \begin{figure} 
-        \begin{overprint}
-            \onslide<1|handout:0>\includegraphics[width=1\linewidth]{figures/math/partial.png}
-            \onslide<2|handout:0>\includegraphics[width=1\linewidth]{figures/math/partial_invalid.png}
-            \onslide<3>\includegraphics[width=1\linewidth]{figures/math/partial_fixed.png}
-        \end{overprint}
-    \end{figure}
-\end{frame}
-
-\begin{frame}{Visualization Equivariance}
-    \begin{figure}[H]
-        \includegraphics[width=\textwidth]{figures/math/diff_type_q.png}
-    \end{figure}
-    \begin{equation*}
-        \vmark(\vsection) = \vmark(\vsection^{\prime})\implies \vmark(m\circ\vsection) = \vmark(m\circ\vsection^{\prime})
-    \end{equation*}
-
-\end{frame}
-
+\begin{comment}
 \begin{frame}{Scatter: $\vmark(xpos, ypos)(\alpha, \beta)$}

### 2022-01-26 slides/main.tex
graphic bundle
sheaves

@@
             \begin{description}[style=nextline]
-                \item[\textcolor{total}{Graphic \gtotalc}]{virtual rendered output}
-                \item[\textcolor{base}{Continuity \gbasec}]{topology of graphic in rendered space}
-                \item[\textcolor{fiber}{Display \gfiberc}] {generalization of renderer specifications, e.g. \{xy,rgba\}, \{xy, cymk\}, \{xyz, rgba\}}
+                \item[\textcolor{total}{Graphic \gtotalc}]{continuity + renderer fields}
+                \item[\textcolor{base}{Continuity \gbasec}]{parameterization of the graphic's geometry}
+                \item[\textcolor{fiber}{Display \gfiberc}] {renderer fields, e.g. \{xy,rgba\}, \{xy, cymk\}, \{xyz, rgba\}}
             \end{description}
@@
     \begin{block}<2-|handout:1>{locality}
-        if there exists the open covering $\mathcal{\opensetc} = \{\opensetc_i\}_{i\in I}$ of \opensetc
         \onslide<3-|handout:1>{
-        and if $\dsectionc^{a}, \dsectionc^{b} \in \sheafc(\opensetc_i)$ have the property $\dsectionc^{a}\restriction_{\opensetc_i} = \dsectionc^{b}\restriction_{\opensetc_i}$ for each $\opensetc_i \in \mathcal{\opensetc}$,}
-        \onslide<4|handout:1>{ then $\dsectionc^{a} = \dsectionc^{b}$}
+        given $\opensetc = \bigcup\limits_{i\in I} \opensetc_i$ and $\dsectionc^{a}, \dsectionc^{b} \in \sheafc(\opensetc)$,}
+        \onslide<4|handout:1>{\\
+        if $\dsectionc^{a}\restriction_{\opensetc_i} = \dsectionc^{b}\restriction_{\opensetc_i}$ for each $\opensetc_i \in \opensetc$}
+        \onslide<4|handout:1>{then $\dsectionc^{a} = \dsectionc^{b}$}
     \end{block} 
     \begin{block}<5-|handout:1>{gluing} 
-        if there exists the open covering $\mathcal{\opensetc} = \{\opensetc_i\}_{i\in I}$ of \opensetc\ and if $\dsectionc^{i} \in \sheafc(\opensetc_i)$ is given for each $i \in I$
-        \onslide<6-|handout:1>{such that ${\dsectionc^{i}}\restriction_{\opensetc_i\cap \opensetc_j} = {\dsectionc^{j}}\restriction_{\opensetc_i \cap \opensetc_j}$ for each pair $\opensetc_i, \opensetc_j \in \mathcal{\opensetc}$,}
-        \onslide<7-|handout:1>{ then there exists $\dsectionc \in \sheafc(\opensetc)$ such that $\dsectionc\restriction_{\opensetc_i} = {\dsectionc^{i}}$ for each $i \in I$} 
+         given $\dsectionc^{i} \in \sheafc(\opensetc_i)$
+        \onslide<6-|handout:1>{s.t.
+${\dsectionc^{i}}\restriction_{\opensetc_i\cap \opensetc_j} = {\dsectionc^{j}}\restriction_{\opensetc_i \cap \opensetc_j}$ for $\opensetc_i, \opensetc_j \in \opensetc$,}
+        \onslide<7-|handout:1>{\\there exists $\dsectionc \in \sheafc(\opensetc)$ such that $\dsectionc\restriction_{\opensetc_i} = {\dsectionc^{i}}$} 
     \end{block}      

### 2022-01-26 slides/main.tex
openset gamma notation

@@
 \begin{frame}
-    \frametitle<3|handout:3>{Functor: $\vindexc: \mathcal{\gbasec} \rightarrow \mathcal{\dbasec}$}
+    \frametitle<3|handout:3>{Functor: $\vindexc: \opensetgc \rightarrow \opensetc$}
     \frametitle<1|handout:1>{Data}
@@
         \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-            \onslide*<0|handout:2,3,5>{\sheafc_{\gbasec, \gtotalc}} 
+            \onslide*<0|handout:2,3,5>{\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}} 
             \onslide*<0|handout:5>{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]} \& \& 
-            \onslide*<0|handout:5>{\textcolor{functor}{\vindexpush}\sheafc_{\gbasec, \gtotalc}}  
+            \onslide*<0|handout:5>{\cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}}  
             \\
-            \onslide*<0|handout:2->{\gbasec}
+            \onslide*<0|handout:2->{\opensetgc}
             \onslide*<0|handout:3->{\arrow[rr, "\vindex", color=functor]}
             \onslide*<0|handout:2,3,5>{\arrow[u, "\sheaf_{\gbasec, \gtotalc}", maps to, color=sheaf]}
-            \onslide*<0|handout:4>{\arrow[d, dashed, maps to, color=sheaf]} \&  \& 
-            \onslide*<0|handout:1,3->{\dbasec}
+            \onslide*<0|handout:4>{\arrow[d, dashed, maps to, "\sheaf_{\gbasec, \vindexpullc\dtotalc}"',color=sheaf]} \&  \& 
+            \onslide*<0|handout:1,3->{\opensetc}
             \onslide*<0|handout:1,3-4>{\arrow[d, "\sheaf_{\dbasec, \dtotalc}", maps to, color=sheaf]}
-            \onslide*<0|handout:5>{\arrow[u, dashed, maps to, color=sheaf]} \\
-            \onslide*<0|handout:4>{\textcolor{functor}{\vindexpull}\sheafc_{\dbasec, \dtotalc}} 
-             \& \& 
-            \onslide*<0|handout:1,3-4>{\sheafc_{\dbasec, \dtotalc}}
+            \onslide*<0|handout:5>{\arrow[u, dashed, maps to, "\sheaf_{\dbasec, \vindexpushc\gtotalc}"', color=sheaf]} \\
+            \onslide*<0|handout:4>{\cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}} 
+             \& \& 
+            \onslide*<0|handout:1,3-4>{\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}}
             \onslide*<0|handout:4>{\arrow[ll, "\textcolor{functor}{\vindexpull}", color=functor]} 

### 2022-01-26 slides/main.tex
phi

@@
 \section{Equivariant Artist}
-\begin{frame}{"Valid" viz?}
-    \begin{equation*}
-        \label{eq:artist:artist_sheaf}
-        \begin{tikzcd}[ampersand replacement = \&, column sep=tiny, row sep=huge]
-            \onslide*<1-|handout:1>{\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}} 
-            \&  \& 
-            \onslide*<2-|handout:1>{\textcolor{set}{\big\{}\gsectionc:\opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} 
-              \bigm{\vert} \vartistc \circ \dfuncc_{\dsectionc} \circ \dsectionc = \dfuncc_{\gsectionc} \circ \gsectionc \textcolor{set}{\big\}}} 
-            \onslide*<2-|handout:1>{\arrow[ll, "\textcolor{set}{\mathlarger{\mathlarger{\supset}}}" description, hook', color=white]} \\ 
-            \&  \& \\
-            \onslide*<1-|handout:1>{\opensetgc} 
-            \onslide*<1-|handout:1>{\arrow[uu, "\sheafc_{\gtotalc}", color=sheaf, maps to]} 
-            \onslide*<2-|handout:1>{\arrow[rruu, "\vartistc(\sheafc_{\dtotalc})\coloneqq \sheafc_{\vartistc}" description, dashed, color=sheaf, maps to]} 
-            \&  \& 
-        \end{tikzcd}
-    \end{equation*}
-\end{frame}
-
-\begin{frame}{What is \dfuncc?}
+
+\begin{frame}{Bundle Map \dfuncc?}
     \begin{block}{Fiber Bundle Category}
@@
 
-    \begin{block}{fiber transform equivariance}
-        The fiber $\dfiberc$ is the sole object of an arbitrary category $\mathcal{C}$. The bundle morphisms 
-        \begin{equation*}
-            \dfuncc \in Hom(\dfiberc,\dfiberc)
-        \end{equation*} 
-        are the set of morphisms that also preserve fiber structure.
+    \begin{block}{Fiber Category}
+        The fiber $\dfiberc$ is the sole object of an arbitrary category $\mathcal{C}$.
+        The morphisms on the fiber are $Hom(\dfiberc, \dfiberc)$
+    \end{block}
+
+    \begin{block}{Data Transformations}
+    $\dfuncc:\dtotalc\restriction_{\dbasepointc} \rightarrow \dtotalc\restriction_{\dbasepointc} \in Hom(\dfiberc\restriction_{\dbasepointc},\dfiberc\restriction_{\dbasepointc})$ 
     \end{block} 
@@
 
+\end{frame}
+
+\begin{frame}{"Valid" viz?}
+    \begin{equation*}
+        \label{eq:artist:artist_sheaf}
+        \begin{tikzcd}[ampersand replacement = \&, column sep=tiny, row sep=huge]
+            \onslide*<1-|handout:1>{\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}} 
+            \&  \& 
+            \onslide*<2-|handout:1>{\textcolor{set}{\big\{}\gsectionc:\opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} 
+              \bigm{\vert} \vartistc \circ \dfuncc_{\dsectionc} \circ \dsectionc = \dfuncc_{\gsectionc} \circ \gsectionc \textcolor{set}{\big\}}} 
+            \onslide*<2-|handout:1>{\arrow[ll, "\textcolor{set}{\mathlarger{\mathlarger{\supset}}}" description, hook', color=white]} \\ 
+            \&  \& \\
+            \onslide*<1-|handout:1>{\opensetgc} 
+            \onslide*<1-|handout:1>{\arrow[uu, "\sheafc_{\gtotalc}", color=sheaf, maps to]} 
+            \onslide*<2-|handout:1>{\arrow[rruu, "\vartistc(\sheafc_{\dtotalc})\coloneqq \sheafc_{\vartistc}" description, dashed, color=sheaf, maps to]} 
+            \&  \& 
+        \end{tikzcd}
+    \end{equation*}
 \end{frame}

### 2022-01-27 slides/main.tex
equivariance
defining output/delta
complex data

@@
                 \item[\textcolor{total}{Graphic \gtotalc}]{continuity + renderer fields}
-                \item[\textcolor{base}{Continuity \gbasec}]{parameterization of the graphic's geometry}
+                \item[\textcolor{base}{Continuity \gbasec}]{parameterization of graphic area (e.g. "inked" bounding box\cite{CairographicsOrg})}
                 \item[\textcolor{fiber}{Display \gfiberc}] {renderer fields, e.g. \{xy,rgba\}, \{xy, cymk\}, \{xyz, rgba\}}
@@
 \begin{frame}
-    \frametitle<3|handout:3>{Functor: $\vindexc: \opensetgc \rightarrow \opensetc$}
+    \frametitle<3|handout:3>{Function: $\vindexc: \gbasec \rightarrow \dbasec$}
     \frametitle<1|handout:1>{Data}
@@
 
+\begin{frame}{Output}
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
+            \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
+            \arrow[rr, "\vartistc", Rightarrow, color=artist] \&  \& 
+            \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+            \arrow[rrd, "\delta"'] 
+            \arrow[rr, "render"] \&  \& visualization 
+            \arrow[d, "measure" description] \\ 
+            \&  \&  \&  \& M
+            \end{tikzcd}
+    \end{equation*}
+\end{frame}
+
 \section{Equivariant Artist}
-
-\begin{frame}{Bundle Map \dfuncc?}
+\begin{frame}{Transform Data}
     \begin{block}{Fiber Bundle Category}
@@
     \end{block}
-
     \begin{block}{Fiber Category}
@@
 
-\begin{frame}{$\dfuncc = (\dfunchc, \dfuncpullc, \dfunctc)$}
+\begin{frame}{$\dfuncc = (\dfunchc, \dfunctc)$}
     \begin{tikzcd}[ampersand replacement = \&]
@@
 
-\begin{frame}{Continuity Equivariance: $\dfunchc_{\dbasec}\circ \vindexc = \vindexc \circ \dfunchc_{\gbasec}$}
+\begin{frame}{Equivariant Artist}
+    \begin{equation*}
+    \begin{tikzcd}[ampersand replacement=\&]
+        \sheafc_{\dbasec, \dtotalc} 
+        \arrow[r, "\vartistc", Rightarrow, color=artist] 
+        \arrow[d, "\dfuncc_{\dtotalc}"', color=action] \& 
+        \sheafc_{\gbasec, \gtotalc} 
+        \arrow[d, "\dfuncc_{\gtotalc}", color=action, dashed] \\
+        \sheafc_{\dbasec^{\prime}, \dfuncpullc\dtotalc} 
+        \arrow[r, "\vartistc", Rightarrow, color=artist]    \& 
+        \sheafc_{\gbasec^{\prime},\dfuncpullc{\gbasec}\gtotalc}    
+        \end{tikzcd} 
+    \end{equation*}
+
+    \begin{proposition}{}
+        \begin{align*}
+            \vartistc(\dsectionc^{a}) = \vartistc(\dsectionc^{b}) &\centernot\implies \dsectionc^a = \dsectionc^b \\
+            \vartistc(\dsectionc^{a})  = \vartistc(\dsectionc^{b}) &\implies \vartistc(\dfuncc_{\dtotalc}(\dsectionc^{a})) = \vartistc(\dfuncc_{\dtotalc}(\dsectionc^{b})) \
+        \end{align*}
+    \end{proposition}
+\end{frame}
+
+\begin{frame}{Equivariant $\dfunchc$}
+    \note{Maybe move to implementation section}
         \begin{equation*}
@@
         \end{equation*}
-
-\end{frame}
-
-\begin{frame}{Field Equivariance: $\vartist \circ (\dfunctc_{\dtotalc}\circ \dfuncpullc_{\dbasec}) =  (\dfunctc_{\gtotalc} \circ \dfuncpullc_{\gbasec}) \circ \vartist$}
-    \begin{equation*}
-    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-            \cgamma{\opensetc^{\prime}}{\dfuncpullc_{\dbasec}\dtotalc\restriction_{\opensetc^{\prime}}} 
-            \arrow[rr, "\vartistc", color=artist, Rightarrow] 
-            \arrow[dd, "\dfunctc_{\dtotalc}\circ \dfuncpullc_{\dbasec}"', color=action] 
-            \&  \& 
-            \vartistc[\dfuncpullc\sheafc_{\dbasec,\dtotalc}] \subset \cgamma{\opensetgc^{\prime}}{\dfuncpullc_{\gbasec}\gtotalc\restriction_{\opensetgc^{\prime}}} 
-            \arrow[dd, "\dfunctc_{\gtotalc}\circ \dfuncpullc_{\gbasec}", dashed, color=action, shift right=5em] \\
-            \&  \&
-            \\
-            \cgamma{\opensetc^{\prime}}{\dfunctc_{\dtotalc}\dfuncpullc_{\dbasec}\dtotalc\restriction_{\opensetc^{\prime}}} 
-            \arrow[rr, "\vartistc", color=artist, Rightarrow]
-            \&  \& 
-            \vartistc[\dfunctc\dfuncpullc\sheafc_{\dbasec,\dtotalc}] \subset \cgamma{\opensetgc^{\prime}}{\dfunctc_{\gtotalc}\dfuncpullc_{\gbasec}\gtotalc\restriction_{\opensetgc^{\prime}}}
-    \end{tikzcd}
-    \end{equation*}
-\end{frame}
-
-\begin{frame}{Equivariant Artist}
+\end{frame}
+
+\begin{frame}{Checking if \vartistc\ is equivariant}
     \begin{equation*}
         \begin{tikzcd}[ampersand replacement=\&]
-        \sheafc_{\dbasec, \dtotalc} 
-        \arrow[r, "\vartistc", Rightarrow, color=artist] 
-        \arrow[d, "\dfuncc_{\dsectionc}"', color=action] \& 
-        \sheafc_{\vartistc} 
-        \arrow[d, "\dfuncc_{\gsectionc}", color=action, dashed] \\
-        \dfuncc_{\dsectionc}\sheafc_{\dbasec, \dtotalc} 
-        \arrow[r, "\vartistc", Rightarrow, color=artist]    \& 
-        \dfuncc_{\gsectionc}\sheafc_{\vartistc}    
-        \end{tikzcd} 
-    \end{equation*}
-
-    \begin{proposition}{}
-        \begin{align*}
-            \vartistc(\dsectionc^{a}) = \vartistc(\dsectionc^{b}) &\centernot\implies \dsectionc^a = \dsectionc^b \\
-            \vartistc(\dsectionc^{a})  = \vartistc(\dsectionc^{b}) &\implies \vartistc(\dfuncc_{\dsectionc}(\dsectionc^{a})) = \vartistc(\dfuncc_{\dsectionc}(\dsectionc^{b})) \
-        \end{align*}
-    \end{proposition}
-\end{frame}
-
-\begin{frame}{Multiple Fields: $\dfuncc = (\dfunchc, \dfuncpullc, \prod\limits_{i=0}^{n}\dfunctc_i)$} 
+            \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
+            \arrow[rr, "\vartistc"', Rightarrow, color=artist] 
+            \arrow[d, "\dfuncc_{\dtotalc}"', color=action] 
+            \arrow[rrr, "\equivc", bend left, color=monoid, dashed] \&  \& 
+            \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+            \arrow[r, "\delta"']\& M 
+            \arrow[d, "\dfuncc_{M}", color=action] \\
+            \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}} 
+            \arrow[rr, "\vartistc", Rightarrow, color=artist] 
+            \arrow[rrr, "\equivc"', bend right, color=monoid, dashed] \&  \& 
+            \cgamma{\opensetgc^{\prime}}{\dfuncpullc\gtotalc\restriction_{\opensetgc^{\prime}}} 
+            \arrow[r, "\delta"] \& M^{\prime}                        
+        \end{tikzcd}
+    \end{equation*}
+        there exists a monoid homomorphism $\equivc: \dfuncc_{\dtotalc} \rightarrow \dfuncc_{M}$ such that 
+        \begin{description}
+            \item[equivariance] $\equivc(\dfuncc_{\dtotalc}(\dsectionc(\dbasepointc))) = \dfuncc_{M}(\equivc(\dsectionc(\dbasepointc)))$, for all $\dbasepointc \in \dbasec$ 
+            \item[continuity] $\lim\limits_{x\to \dbasepointc}(\equivc(\dsectionc(x))) = \equivc(\dsectionc(\dbasepointc))$ for all $\dbasepointc \in \dbasec$ 
+        \end{description} 
+\end{frame}
+
+\section{Complex Structure}
+\begin{frame}{Multiple Bundles: $\dfuncc = (\dfunchc, \prod\limits_{i=0}^{n}\dfunctc_i)$} 
     \begin{equation*}
@@
 
-    \begin{block}{Add New Field to E}
+\end{frame}
+
+\begin{frame}{complex data}
     \begin{description}[style=newline]
-            \item[construct] {$\odot:\dfiberc_{field} \times \dbasec \rightarrow \dtotalc_{field}$}
-            \item[concatenate]{$\otimes: \dtotalc_{field} \times \dtotalc \mapsto \dtotalc$ s.t. $\pi(\dtotalc_{field}) = \pi(\dtotalc)$}
-        \end{description}
-    \end{block}
-
-\end{frame}
-
-\begin{frame}{Field Equivariance}
-
-    \begin{equation*}
-        \oslash: \vartistc \times \dfiberc \rightarrow visual\; encoding
-    \end{equation*}
-
+        \item[combining continuities]{given $\dbasepointc_{a} \in \dbasec_{c} \subset \dbasec_{a}$ and $\dbasepointc_{b} \in \dbasec_{c} \subset \dbasec_{b}$, if $\dbasepointc_{a} = \dbasepointc_{b}$ then $\equivc(\dsectionc(\dbasepointc_{a})) = \equivc(\dsectionc(\dbasepointc_{b}))$ for all $\dbasepointc_{a}, \dbasepointc_{b} \in \dbasec_{a}\bigsqcup\limits_{\dbasec_{c}} \dbasec_{b}$}
+        \item[shared fibers]{
+            if $p = p^{\prime}$ then $\equivc \circ q$ = $\equivc^{\prime} \circ q^{\prime}$
         \begin{equation*}
         \begin{tikzcd}[ampersand replacement = \&]
-            \& 
-            {\sheafc_{\dbasec, \dtotalc}}
-            \arrow[rr, "\vartistc"]                   
-            \&  \& 
-            \sheafc_{\vartist} 
-            \arrow[rd, "\oslash_{\dfiber_{field}}"]           \&                        \\
-            \dfiberc_{field} 
-            \arrow[ru, "\otimes_{\dtotalc} \circ \odot_{\dbasec}"] 
-            \arrow[rd, "\otimes_{\dtotalc^{\prime}} \circ \odot_{\dbasec^{\prime}}"'] 
-            \& \& \& \& 
-            visual\;encoding \\
-            \& 
-            {\sheafc_{\dbasec^{\prime}, \dtotalc^{\prime}}} 
-            \arrow[rr, "\vartistc"'] 
-            \& \& 
-            \sheafc_{\vartist}^{\prime} 
-            \arrow[ru, "\oslash_{\dfiber_{field}}"'] 
-            \&                       
-    \end{tikzcd}
-\end{equation*}
-
+        \& \dfiberc^a \times \dfiberc^b 
+        \arrow[ld, "p"', color=fiber] 
+        \arrow[rd, "q \circ \equivc"] \&     \\
+        \dfiberc^a \& \& M^a \\
+        \& \dfiberc^a \times \dfiberc^c 
+        \arrow[lu, "p^{\prime}", color=fiber] 
+        \arrow[ru, " q^{\prime} \circ\equivc^{\prime}"'] \&    
+      \end{tikzcd}
+    \end{equation*}}         
+        \item[both]{given $\dsectionc^{a} \in \sheafc_{\dbasec_a, \dtotalc_a}, \dsectionc^{b} \in \sheafc_{\dbasec_b, \dtotalc_b}$ and $\dbasepointc \in \dbasec_{a}\bigsqcup\limits_{\dbasec_{c}} \dbasec_{b}$ then $q(\equivc(\dsection^a(\dbasepointc))) = q^{\prime}(\equivc(\dsectionc^b(\dbasepointc)))$ when $p(\dfiberc^{a}\restriction_{\dbasepointc}) = p^{\prime}(\dfiberc^{b}\restriction_{\dbasepointc})$}
+    \end{description}
 \end{frame}

### 2022-01-27 slides/main.tex
valid artist
nu slides

@@
                \onslide*<7|handout:3>{\gbasec} 
-               \onslide*<5-6|handout:1-2>{\arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]}
-               \onslide*<7|handout:3>{\arrow[u, "\gsectionc \in \cgamma{\gbasec}{\gtotalc}"', bend right, pos=.5]}
+               \onslide*<5-6|handout:1-2>{\arrow[u, "\dsectionc \in \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}"', bend right, pos=.5]}
+               \onslide*<7|handout:3>{\arrow[u, "\gsectionc \in \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}"', bend right, pos=.5]}
             \end{tikzcd}
@@
 \end{frame}
+
+\begin{comment}
+    \note{move this to top where I introduce data/graphics}
+\begin{frame}{Domain and Range}
+    \begin{figure}
+        \begin{subfigure}{.5\textwidth}
+            \includegraphics[.5\textwidth]{../paper/figures/fiberbundle.png}
+        \end{subfigure}
+        \begin{subfigure}{.5\textwidth}
+            \includegraphics[.5\textwidth]{../paper/figures/render.png}
+        \end{subfigure}
+    \end{figure}
+\end{frame}
+\end{comment}
 
@@
             \onslide*<2-|handout:1>{\textcolor{set}{\big\{}\gsectionc:\opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} 
-              \bigm{\vert} \vartistc \circ \dfuncc_{\dsectionc} \circ \dsectionc = \dfuncc_{\gsectionc} \circ \gsectionc \textcolor{set}{\big\}}} 
+              \bigm{\vert} \delta(\vartistc(\dsectionc)) = \equivc(\dsectionc) \textcolor{set}{\big\}}} 
             \onslide*<2-|handout:1>{\arrow[ll, "\textcolor{set}{\mathlarger{\mathlarger{\supset}}}" description, hook', color=white]} \\ 
@@
 
-\begin{frame}{Encoding + Assembly Stages}
-    %E-V and Q diagram    
-\end{frame}
-\begin{frame}{Encoding + Assembly}
-    \begin{figure}
-        \includegraphics{../paper/figures/q_hat.p}
-    \end{figure}
-\end{frame}
-\begin{frame}{Domain and Range}
-    \begin{figure}
-        \begin{subfigure}{.5\textwidth}
-            \includegraphics[.5\textwidth]{../paper/figures/fiberbundle.png}
-        \end{subfigure}
-        \begin{subfigure}{.5\textwidth}
-            \includegraphics[.5\textwidth]{../paper/figures/render.png}
-        \end{subfigure}
-    \end{figure}
-\end{frame}
-
-\begin{frame}{Example $\dfunctc$: Stevens' Scales \cite{stevensTheoryScalesMeasurement1946}}
+\begin{frame}{Measurable Visual Components}
+    \begin{columns}
+        \column{0.31\textwidth}
+        \begin{equation*}
+            \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
+                \onslide*<1|handout:1>{\vfiberc}
+                \onslide*<1|handout:1>{\arrow[r, hook]} \& 
+                \onslide*<1|handout:1>{\vtotalc} 
+                \onslide*<1|handout:1>{\arrow[d, "\pi"']} \\
+                \& 
+                \onslide*<1|handout:1>{\dbasec} 
+                \onslide*<1|handout:1>{\arrow[u, "\vsectionc \in \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}}"', bend right, pos=.5]}
+            \end{tikzcd}
+        \end{equation*}
+        \column{0.69\textwidth}
+    \only<1|handout:1>{
+        \begin{description}[style=nextline]
+            \item[\textcolor{total}{Data \vtotalc}]{continuity + visual fields}
+            \item[\textcolor{base}{Continuity \dbasec}]{data continuity}
+            \item[\textcolor{fiber}{components \vfiberc}] {visual components \cite{bertinIIPropertiesGraphic2011} of a graphic, e.g. x and y position, area, color, line thickness}
+        \end{description}
+        }
+    \end{columns}
+    \pause 
+    \only<1|handout:1>{
+        \begin{alertblock}{}
+            \begin{description}[style=nextline]
+                \item[\textcolor{section}{visual Components}]{$\cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} \coloneqq \big\{\vsectionc: \opensetc\rightarrow \vtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\vsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}$}
+            \end{description}
+        \end{alertblock}
+    }
+\end{frame}
+
+\begin{frame}{Equivariant Data to Components: $\vchannel: \sheaf_{\dbasec,\dtotalc} \rightarrow \sheaf_{\dbasec, \vtotalc}$}
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
+            \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \arrow[rr, "\vchannelc"] \& \& 
+            \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} \\
+             \& \opensetc 
+             \arrow[lu, "{\sheafc_{\dbasec,\dtotalc}}" description, maps to] 
+             \arrow[ru, "{\sheafc_{\dbasec, \vtotalc}}" description, maps to] \& 
+            \end{tikzcd}
+    \end{equation*}
+
+    \begin{alertblock}{Build in Equivariance}
+    implement $\vchannelc: \dsectionc \mapsto \vsectionc$ as a monoid homomorphism $\vchannelc: \dfuncc_{\dtotalc} \rightarrow \dfuncc_{\vtotalc}$ such that 
+        \begin{description}
+            \item[equivariance] $\vchannelc(\dfuncc_{\dtotalc}(\dsectionc(\dbasepointc))) = \dfuncc_{\vtotalc}(\vchannelc(\dsectionc(\dbasepointc)))$, for all $\dbasepointc \in \dbasec$ 
+            \item[continuity] $\lim\limits_{x\to \dbasepointc}(\vchannelc(\dsectionc(x))) = \vchannelc(\dsectionc(\dbasepointc))$ for all $\dbasepointc \in \dbasec$ 
+    \end{description} 
+\end{alertblock}
+\end{frame}
+
+\begin{frame}{Sample $\dfuncc$: Stevens' Scales \cite{stevensTheoryScalesMeasurement1946}}
     \begin{table}[H]
@@
             ordinal &  monotonic & $\text{if } \delement_1 \leq \delement_2 \text{ then } \vchannel (\delement_1) \leq \vchannel(\delement_2)$\\
-        interval &  translation &  $\vchannel (x + c) = \vchannel(x) + c$ \\
-        ratio &  scaling &  $\vchannel(xc) = \vchannel(x)*c $\\ \bottomrule
+            interval &  translation &  $\vchannel (\delement + c) = \vchannel(\delement) + c$ \\
+            ratio &  scaling &  $\vchannel(\delement* c) = \vchannel(\delement)*c $\\ \bottomrule
         \end{tabularx}
     \end{table}
+\end{frame}
+
+\begin{frame}{Shared Components: $\vchannelc = \prod\limits_{i=0}^{n}\vchannelc_i$}
+    \begin{equation}
+        \begin{tikzcd}[ampersand replacement=\&]
+            \dfiberc_a\times \dfiberc_b 
+            \arrow[rr, "\vchannelc"] 
+            \arrow[rd, "p_{\dfiberc}"] \& \& \vfiberc_a\times \vfiberc_b 
+            \arrow[rd, "p_{\vfiberc}"] \& \\    \& \dfiberc_{a} \arrow[rr, "\vchannelc_{a}", dashed] \& \& \vfiberc_a \\
+            \dfiberc_a\times \dfiberc_c 
+            \arrow[rr, "\vchannelc^{\prime}"] \arrow[ru, "p_{\dfiberc^{\prime}}"'] \& \& {\vfiberc_a}^{\prime}\times {\vfiberc_c}^{prime} \arrow[ru, "p_{\vfiberc^{\prime}}"'] \&           
+            \end{tikzcd}
+    \end{equation}
+    \begin{alertblock}{Consistent Transformations \cite{hullmanKeeping2018}}
+        if $p_{\dfiberc} = p_{\dfiberc^{\prime}}$ then $p_{\vfiberc}(\vchannelc(\dsectionc)) = p_{\vfiberc^{\prime}}(\vchannelc^{\prime}(\dsectionc^{\prime}))$ s.t.
+there exists a transformation $\vchannelc_a: \dfiberc_{a}\rightarrow \vfiberc_{a}$
+    \end{alertblock}
+\end{frame}
+
+\begin{frame}{Components to Graphic}
+    \begin{figure}
+        \includegraphics{../paper/figures/q_hat.p}
+    \end{figure}
 \end{frame}

### 2022-02-18 slides/main.tex
implementation construction

@@
     \begin{block}{Data Transformations}
-    $\dfuncc:\dtotalc\restriction_{\dbasepointc} \rightarrow \dtotalc\restriction_{\dbasepointc} \in Hom(\dfiberc\restriction_{\dbasepointc},\dfiberc\restriction_{\dbasepointc})$ 
+    $\dfuncc:\dtotalc\restriction_{\dbasepointc} \rightarrow \dtotalc\restriction_{\dbasepointc^{\prime}} \in Hom(\dfiberc\restriction_{\dbasepointc},\dfiberc\restriction_{\dbasepointc^{\prime}})$ 
     \end{block} 
@@
 
-\begin{frame}{complex data}
+\begin{frame}{Complex data}
     \begin{description}[style=newline]
@@
     \begin{alertblock}{Build in Equivariance}
-    implement $\vchannelc: \dsectionc \mapsto \vsectionc$ as a monoid homomorphism $\vchannelc: \dfuncc_{\dtotalc} \rightarrow \dfuncc_{\vtotalc}$ such that 
+    implement $\vchannelc$ as a functor such that 
         \begin{description}
             \item[equivariance] $\vchannelc(\dfuncc_{\dtotalc}(\dsectionc(\dbasepointc))) = \dfuncc_{\vtotalc}(\vchannelc(\dsectionc(\dbasepointc)))$, for all $\dbasepointc \in \dbasec$ 
-            \item[continuity] $\lim\limits_{x\to \dbasepointc}(\vchannelc(\dsectionc(x))) = \vchannelc(\dsectionc(\dbasepointc))$ for all $\dbasepointc \in \dbasec$ 
+            \item[continuity] $\lim\limits_{x\to \dbasepointc}(\vchannelc(\dsectionc(x))) = \vchannelc(\dsectionc(\dbasepointc))$ for all $\dbasepointc \in \dbasec$ \note{for all sections} 
     \end{description} 
@@
             ordinal &  monotonic & $\text{if } \delement_1 \leq \delement_2 \text{ then } \vchannel (\delement_1) \leq \vchannel(\delement_2)$\\
-            interval &  translation &  $\vchannel (\delement + c) = \vchannel(\delement) + c$ \\
-            ratio &  scaling &  $\vchannel(\delement* c) = \vchannel(\delement)*c $\\ \bottomrule
+            interval &  translation &  $\vchannel (\delement + c) = \vchannel(\delement) + \vchannel(c)$ \\
+            ratio &  scaling &  $\vchannel(\delement* c) = \vchannel(\delement)*\vchannel(c) $\\ \bottomrule
         \end{tabularx}
@@
             \dfiberc_a\times \dfiberc_c 
-            \arrow[rr, "\vchannelc^{\prime}"] \arrow[ru, "p_{\dfiberc^{\prime}}"'] \& \& {\vfiberc_a}^{\prime}\times {\vfiberc_c}^{prime} \arrow[ru, "p_{\vfiberc^{\prime}}"'] \&           
+            \arrow[rr, "\vchannelc^{\prime}"] \arrow[ru, "p_{\dfiberc^{\prime}}"'] \& \& {\vfiberc_a}^{\prime}\times {\vfiberc_c}^{\prime} \arrow[ru, "p_{\vfiberc^{\prime}}"'] \&           
             \end{tikzcd}
@@
 
+\begin{frame}{Remapping Components}
+    \begin{block}<>{Functor $\vchannel$}
+        $\vchannelc$ is a functor from on fiber categories $\vchannel: \mathcal{\dfiberc} \rightarrow \mathcal{\vfiberc}$
+
+    \end{block}
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement=\&]
+            \dfiberc 
+            \arrow[rr, "\vchannelc_{\dfiberc}"] 
+            \arrow[rrdd, "{\vchannelc_{\dfiberc}}^{\prime}"', dotted] 
+            \&  \& 
+            \vfiberc \coloneqq \dfiberc^{\prime}  
+            \arrow[dd, "\vchannelc_{\vfiberc}"] \\
+            \&  \& \\
+            \&  \& \vfiberc^{\prime}                           
+        \end{tikzcd}
+    \end{equation*}
+\end{frame}
+
 \begin{frame}{Components to Graphic}
@@
     \end{figure}
+\end{frame}
+
+\begin{frame}{implement}
+    \begin{tikzcd}[ampersand replacement = \&]
+    \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
+    \arrow[r, "\vchannelc"] \& 
+    \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} 
+    \arrow[r, "\vmarkc_*"] 
+    \arrow[rr, "\vmarkc", bend left] \& 
+    \cgamma{\opensetc}{\gtotalc\restriction_{\opensetc}} \& 
+    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+    \arrow[l, "\vindexcpush"]\\ 
+    \& \\ \& 
+    \opensetc 
+    \arrow[luu, "{\sheafc_{\dbasec, \dtotalc}}" description, maps to] 
+    \arrow[uu, "{\sheafc_{\dbasec, \vtotalc}}" description, maps to] 
+    \arrow[ruu, "{\sheafc_{\dbasec, \gtotalc}}" description, maps to] 
+    \& \& 
+    \opensetgc \arrow[uu, "{\sheafc_{\gbasec, \gtotalc}}" description, maps to] 
+    \arrow[ll, "\vindexc"', dashed]
+    \end{tikzcd}
 \end{frame}

### 2022-02-21 slides/main.tex
reworking notation and pushpull story and finishing out Q

@@
 
+\subsection{Fiber Bundles}
 \begin{frame}
     \frametitle<1|handout:1>{Fiber Bundle}
-    \frametitle<6|handout:2>{Data Bundle}
-    \frametitle<7|handout:3>{Graphic Bundle}
+    \frametitle<6|handout:2-3>{Data Bundle}
+    \frametitle<7|handout:4-5>{Graphic Bundle}
     \begin{columns}
@@
               \onslide*<3-6|handout:1-2>{\dfiberc}
-              \onslide*<7|handout:3>{\gfiberc} 
-              \onslide*<4-|handout:1->{\arrow[r, hook]} \& 
+              \onslide*<7|handout:4>{\gfiberc} 
+              \onslide*<4-|handout:1-2,4>{\arrow[r, hook]} \& 
               \onslide*<1-6|handout:1-2>{\dtotalc} 
-              \onslide*<7|handout:3>{\gtotalc}
-              \onslide*<4-|handout:1->{\arrow[d, "\pi"']} \\
+              \onslide*<7|handout:4>{\gtotalc}
+              \onslide*<4-|handout:1-2,4>{\arrow[d, "\pi"']} \\
                \& 
                \onslide*<2-6|handout:1-2>{\dbasec}
-               \onslide*<7|handout:3>{\gbasec} 
+               \onslide*<7|handout:4>{\gbasec} 
                \onslide*<5-6|handout:1-2>{\arrow[u, "\dsectionc \in \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}"', bend right, pos=.5]}
-               \onslide*<7|handout:3>{\arrow[u, "\gsectionc \in \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}"', bend right, pos=.5]}
+               \onslide*<7|handout:4>{\arrow[u, "\gsectionc \in \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}"', bend right, pos=.5]}
             \end{tikzcd}
@@
         }
-        \only<7|handout:3>{
+        \only<7|handout:4>{
             \begin{description}[style=nextline]
@@
     }
-    \only<7|handout:3>{
+    \only<7|handout:4>{
         \begin{alertblock}{}
@@
     }
-\end{frame}
-
-\begin{comment}
-    \note{move this to top where I introduce data/graphics}
-\begin{frame}{Domain and Range}
+    \only<|handout:3>{
         \begin{figure}
-        \begin{subfigure}{.5\textwidth}
-            \includegraphics[.5\textwidth]{../paper/figures/fiberbundle.png}
-        \end{subfigure}
-        \begin{subfigure}{.5\textwidth}
-            \includegraphics[.5\textwidth]{../paper/figures/render.png}
-        \end{subfigure}
-    \end{figure}
-\end{frame}
-\end{comment}
-
-\subsection{Category Theory}
+            \includegraphics[scale=.65]{figures/math/spherebundle.png}
+        \end{figure}
+    }
+    \only<|handout:5>{
+        \begin{figure}
+            \includegraphics[scale=.65]{figures/math/render.png}
+        \end{figure}
+    }
+\end{frame}
+
+\subsection{Category Theory Language}
 \begin{frame}[fragile]{}
@@
 
-\begin{frame}{Presheaf: $\sheafc: \mathcal{C}^{op} \textcolor{sheaf}{\rightarrow} \setc $}
-    \begin{columns}
-    \column{0.29\textwidth}
-    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-        \onslide*<9-|handout:1->{\dfiberc}
-        \onslide*<9-|handout:1->{\arrow[r, hook]} \& 
-        \onslide*<1-|handout:1->{\dtotalc}
-        \onslide*<1-|handout:1->{\arrow[d, "\pi"']} \\
-           \& 
-        \onslide*<1-|handout:1->{\dbasec}
-        \onslide*<3-|handout:1->{\arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]}
-        \end{tikzcd}
-    \column{0.69\textwidth}
-        \begin{tikzcd}[ampersand replacement = \&, column sep=small]
-            \onslide*<3,6-|handout:0>{{\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}} 
-            \onslide*<4-5|handout:0>{{\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}\in \setc}
-            \onslide*<0|handout:1>{\setc \ni {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}}
-            \&  
-            {} \& 
-            \onslide*<6-|handout:1>{{\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}}} 
-            \onslide*<7-|handout:1>{\arrow[ll, "\iota^*"', hook', color=set]} \\\& \&\\
-            \onslide*<3|handout:0>{\opensetc_1 \subset \dbasec}
-            \onslide*<4-5|handout:0>{\opensetc_1 \in Ob(\mathcal{\dbasec}^{\textcolor{base}{op}})}
-            \onslide*<6-|handout:0>{\opensetc_1}
-            \onslide*<0|handout:1>{Ob(\mathcal{\dbasec}^{\textcolor{base}{op}}) \ni \opensetc_1} 
-            \onslide*<5-|handout:0>{\arrow[uu, "\sheafc_{\dbasec, \dtotalc}", maps to, color=sheaf]}
-            \onslide*<0|handout:1>{\arrow[uu, "\sheafc_{\dbasec, \dtotalc}", maps to, color=sheaf, shift right = 10]}
-            \onslide*<7-|handout:1>{\arrow[rr, "\iota"', hook, color=base]} \&   
-            \onslide*<8-|handout:1>{\arrow[uu, "\sheafc_{\dbasec, \dtotalc}" description, maps to, color=sheaf]} \& 
-            \onslide*<6-|handout:1>{\opensetc_2} 
-            \onslide*<6-|handout:1>{\arrow[uu, "\sheafc_{\dbasec, \dtotalc}"', maps to, color=sheaf]}                   
-            \end{tikzcd}
-    \end{columns}
-    \begin{center}
-        \begin{alertblock}<8-|handout:1>{stalk}
-            \begin{align*}
-                \sheafc_{\dbasec, \dtotalc}\restriction_{\dbasepointc}\coloneqq \lim\limits_{\opensetc\ni \dbasepointc} \Gamma(\opensetc, \dtotalc\restriction_{\opensetc}) \\
-                 \dfiberc_{\dbasepointc} \subset  \sheafc_{\dbasec, \dtotalc}\restriction_{\dbasepointc}
-            \end{align*}
-        \end{alertblock}
-
-        \begin{alertblock}<9-|handout:1>{germ}
-            %germ  is tau at limit, \tau(k) \in germ
-            \begin{align*}
-                \dsectionc(\dbasepointc) \in \sheafc_{\dbasec, \dtotalc}\restriction_{\dbasepointc}
-            \end{align*}
-        \end{alertblock}
-    \end{center}
-\end{frame}
-
-\begin{frame}{Sheaves} 
-    A sheaf is a presheaf that satisfies the following two axioms\cite{bakerMathsSheaf}\\
-    \begin{block}<2-|handout:1>{locality}
-        \onslide<3-|handout:1>{
-        given $\opensetc = \bigcup\limits_{i\in I} \opensetc_i$ and $\dsectionc^{a}, \dsectionc^{b} \in \sheafc(\opensetc)$,}
-        \onslide<4|handout:1>{\\
-        if $\dsectionc^{a}\restriction_{\opensetc_i} = \dsectionc^{b}\restriction_{\opensetc_i}$ for each $\opensetc_i \in \opensetc$}
-        \onslide<4|handout:1>{then $\dsectionc^{a} = \dsectionc^{b}$}
-    \end{block} 
-    \begin{block}<5-|handout:1>{gluing} 
-         given $\dsectionc^{i} \in \sheafc(\opensetc_i)$
-        \onslide<6-|handout:1>{s.t.
-${\dsectionc^{i}}\restriction_{\opensetc_i\cap \opensetc_j} = {\dsectionc^{j}}\restriction_{\opensetc_i \cap \opensetc_j}$ for $\opensetc_i, \opensetc_j \in \opensetc$,}
-        \onslide<7-|handout:1>{\\there exists $\dsectionc \in \sheafc(\opensetc)$ such that $\dsectionc\restriction_{\opensetc_i} = {\dsectionc^{i}}$} 
-    \end{block}      
-\end{frame}
-
-\begin{frame}
-    \frametitle<3|handout:3>{Function: $\vindexc: \gbasec \rightarrow \dbasec$}
-    \frametitle<1|handout:1>{Data}
-    \frametitle<4-6|handout:4>{Pullback: data to region of the visualization} 
-    \frametitle<2|handout:2>{Graphic}
-    \frametitle<7-|handout:5>{Pushforward: visualization to index of data}
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-            \onslide*<0|handout:2,3,5>{\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}} 
-            \onslide*<0|handout:5>{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]} \& \& 
-            \onslide*<0|handout:5>{\cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}}  
-            \\
-            \onslide*<0|handout:2->{\opensetgc}
-            \onslide*<0|handout:3->{\arrow[rr, "\vindex", color=functor]}
-            \onslide*<0|handout:2,3,5>{\arrow[u, "\sheaf_{\gbasec, \gtotalc}", maps to, color=sheaf]}
-            \onslide*<0|handout:4>{\arrow[d, dashed, maps to, "\sheaf_{\gbasec, \vindexpullc\dtotalc}"',color=sheaf]} \&  \& 
-            \onslide*<0|handout:1,3->{\opensetc}
-            \onslide*<0|handout:1,3-4>{\arrow[d, "\sheaf_{\dbasec, \dtotalc}", maps to, color=sheaf]}
-            \onslide*<0|handout:5>{\arrow[u, dashed, maps to, "\sheaf_{\dbasec, \vindexpushc\gtotalc}"', color=sheaf]} \\
-            \onslide*<0|handout:4>{\cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}} 
-             \& \& 
-            \onslide*<0|handout:1,3-4>{\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}}
-            \onslide*<0|handout:4>{\arrow[ll, "\textcolor{functor}{\vindexpull}", color=functor]} 
-        \end{tikzcd}
-    \end{equation*}
-    \only<0|handout:1>{
-        \begin{itemize}
-            \item $\dfiberc \hookrightarrow \dtotalc \xrightarrow{\pi} \dbasec$
-            \item $\sheafc_{\dbasec, \dtotalc}:\opensetc \mapsto \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}, \opensetc \subset \dbasec$
-            \item $\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \ni \dsectionc: \opensetc \rightarrow \dfiberc\restriction_{\opensetc}$
-            \item $\dsectionc(\dbasepointc) = \{f_0:v_0, \cdots,\}, \dbasepointc \in \opensetc$
-        \end{itemize}
-    }
-    \only<0|handout:4>{
-        \begin{itemize}
-            \item $\vindexpullc\dfiberc \hookrightarrow \vindexpullc\dtotalc \xrightarrow{\pi} \gbasec$
-            \item $\vindexpullc\sheafc_{\dbasec,\dtotalc}:\opensetgc \mapsto \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}, \opensetgc \subset \gbasec$
-            \item $\cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} \ni \vindexpullc\dsectionc: \opensetgc \rightarrow \vindexpullc\dfiberc\restriction_{\opensetgc}$
-            \item $\vindexpullc\dsectionc(\gbasepointc) = \dsectionc(\vindexc(\gbasepointc)) = \dsectionc(\dbasepointc)$
-        \end{itemize}
-    }
-    \only<0|handout:2>{
-        \begin{itemize}
-            \item $\gfiberc \hookrightarrow \gtotalc \xrightarrow{\pi} \gbasec$
-            \item $\sheafc_{\gbasec, \gtotalc}:\opensetgc \mapsto \cgamma{\opensetgc}{\dtotalc\restriction_{\opensetgc}}, \opensetgc \subset \gbasec$
-            \item $\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \ni \gsectionc: \opensetgc \rightarrow \gfiberc\restriction_{\opensetgc}$
-            \item $\gsectionc(\gbasepointc) = \{d_0, \cdots\}, \gbasepointc \in \opensetgc$
-        \end{itemize}
-    }
-    \only<0|handout:5>{
-        \begin{itemize}
-            \item $\vindexpushc\gfiberc \hookrightarrow \vindexpushc\gtotalc \xrightarrow{\pi} \dbasec$
-            \item $\vindexpushc\sheafc_{\gbasec, \gtotalc}:\opensetc \mapsto \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}, \opensetc \subset \dbasec$
-            \item $\cgamma{\opensetc}{\gtotalc\restriction_{\opensetc}} \ni \gsectionc: \opensetc \rightarrow \gfiberc\restriction_{\opensetc}$
-            \item $\vindexpushc\gsectionc(\gbasepointc) = \gsectionc\restriction_{\vindexpre(\dbasepointc)} = \gsectionc(\gbasepointc) \forall \gbasepointc \in \vindexpre(\dbasepointc)$ 
-        \end{itemize}
-    }
-\end{frame}
-
-%add transition slide->possibly call back 
-\begin{frame}
-    {\huge
-        \begin{equation*}
-            \sheafc_{\dbasec,\dtotalc} \textcolor{artist}{\xrightarrow{\vartistc}} \sheafc_{\gbasec, \gtotalc}
-        \end{equation*}
-    }
-\end{frame}
-
 \begin{frame}{Natural Transformation $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\Rightarrow} \textcolor{functor}{G}$}
-
+    \onslide<|handout:1>{
     \adjustbox{scale=3, center}{
@@
             \& {} \&            
-          \end{tikzcd}}
+          \end{tikzcd}}}
 \end{frame}
@@
 
+\subsection{Sheaves}
+
+\begin{frame}{Presheaf: $\sheafc: \mathcal{C}^{op} \textcolor{sheaf}{\rightarrow} \setc $}
+    \begin{columns}
+    \column{0.29\textwidth}
+    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
+        \onslide*<9-|handout:1->{\dfiberc}
+        \onslide*<9-|handout:1->{\arrow[r, hook]} \& 
+        \onslide*<1-|handout:1->{\dtotalc}
+        \onslide*<1-|handout:1->{\arrow[d, "\pi"']} \\
+           \& 
+        \onslide*<1-|handout:1->{\dbasec}
+        \onslide*<3-|handout:1->{\arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]}
+        \end{tikzcd}
+    \column{0.69\textwidth}
+        \begin{tikzcd}[ampersand replacement = \&, column sep=small]
+            \onslide*<3,6-|handout:0>{{\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}} 
+            \onslide*<4-5|handout:0>{{\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}\in \setc}
+            \onslide*<0|handout:1>{\setc \ni {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}}
+            \&  
+            {} \& 
+            \onslide*<6-|handout:1>{{\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}}} 
+            \onslide*<7-|handout:1>{\arrow[ll, "\iota^*"', hook', color=set]} \\\& \&\\
+            \onslide*<3|handout:0>{\opensetc_1 \subset \dbasec}
+            \onslide*<4-5|handout:0>{\opensetc_1 \in Ob(\mathcal{\dbasec}^{\textcolor{base}{op}})}
+            \onslide*<6-|handout:0>{\opensetc_1}
+            \onslide*<0|handout:1>{Ob(\mathcal{\dbasec}^{\textcolor{base}{op}}) \ni \opensetc_1} 
+            \onslide*<5-|handout:0>{\arrow[uu, "\sheafc_{\dbasec, \dtotalc}", maps to, color=sheaf]}
+            \onslide*<0|handout:1>{\arrow[uu, "\sheafc_{\dbasec, \dtotalc}", maps to, color=sheaf, shift right = 10]}
+            \onslide*<7-|handout:1>{\arrow[rr, "\iota"', hook, color=base]} \&   
+            \onslide*<8-|handout:1>{\arrow[uu, "\sheafc_{\dbasec, \dtotalc}" description, maps to, color=sheaf]} \& 
+            \onslide*<6-|handout:1>{\opensetc_2} 
+            \onslide*<6-|handout:1>{\arrow[uu, "\sheafc_{\dbasec, \dtotalc}"', maps to, color=sheaf]}                   
+            \end{tikzcd}
+    \end{columns}
+    \begin{center}
+        \begin{alertblock}<8-|handout:1>{stalk}
+            \begin{align*}
+                \sheafc_{\dbasec, \dtotalc}\restriction_{\dbasepointc}\coloneqq \lim\limits_{\opensetc\ni \dbasepointc} \Gamma(\opensetc, \dtotalc\restriction_{\opensetc}) \\
+                 \dfiberc_{\dbasepointc} \subset  \sheafc_{\dbasec, \dtotalc}\restriction_{\dbasepointc}
+            \end{align*}
+        \end{alertblock}
+
+        \begin{alertblock}<9-|handout:1>{germ}
+            %germ  is tau at limit, \tau(k) \in germ
+            \begin{align*}
+                \dsectionc(\dbasepointc) \in \sheafc_{\dbasec, \dtotalc}\restriction_{\dbasepointc}
+            \end{align*}
+        \end{alertblock}
+    \end{center}
+\end{frame}
+
+\begin{frame}{Sheaves on Bundles} 
+    A sheaf is a presheaf that satisfies the following two axioms\cite{bakerMathsSheaf}\\
+    \begin{block}<2-|handout:1>{locality}
+        \onslide<3-|handout:1>{
+        given $\opensetc = \bigcup\limits_{i\in I} \opensetc_i$ and $\dsectionc^{a}, \dsectionc^{b} \in \sheafc(\opensetc)$,}
+        \onslide<4|handout:1>{\\
+        if $\dsectionc^{a}\restriction_{\opensetc_i} = \dsectionc^{b}\restriction_{\opensetc_i}$ for each $\opensetc_i \in \opensetc$}
+        \onslide<4|handout:1>{then $\dsectionc^{a} = \dsectionc^{b}$}
+    \end{block} 
+    \begin{block}<5-|handout:1>{gluing} 
+         given $\dsectionc^{i} \in \sheafc(\opensetc_i)$
+        \onslide<6-|handout:1>{s.t.
+${\dsectionc^{i}}\restriction_{\opensetc_i\cap \opensetc_j} = {\dsectionc^{j}}\restriction_{\opensetc_i \cap \opensetc_j}$ for $\opensetc_i, \opensetc_j \in \opensetc$,}
+        \onslide<7-|handout:1>{\\there exists $\dsectionc \in \sheafc(\opensetc)$ such that $\dsectionc\restriction_{\opensetc_i} = {\dsectionc^{i}}$} 
+    \end{block}      
+\end{frame}
+
 \begin{frame}
-    \frametitle<1-2|handout:1>{Data Space: $\vartistc_{\opensetc}: \sheafc_{\dbasec, \dtotalc} \textcolor{artist}{\Rightarrow} \vindexpushc\sheafc_{\gbasec, \gtotalc}$}
-    \frametitle<3-4|handout:2>{Display Space: $\vartist_{\opensetgc}: \vindexpullc\sheafc_{\dbasec, \dtotalc} \textcolor{artist}{\Rightarrow} \sheafc_{\gbasec, \gtotalc}$}
-    \frametitle<5|handout:3->{\textcolor{artist}{Artist} \vartistc}
-
-    \onslide*<1|handout:1>{
+    \frametitle<3|handout:3>{Function: $\vindexc: \gbasec \rightarrow \dbasec$}
+    \frametitle<1|handout:1>{Data}
+    \frametitle<4-6|handout:4>{Pullback: data to region of the visualization} 
+    \frametitle<2|handout:2>{Graphic}
+    \frametitle<7-|handout:5>{Pushforward: visualization to index of data}
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
+            \onslide*<0|handout:2,3,5>{\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}} 
+            \onslide*<0|handout:5>{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]} \& \& 
+            \onslide*<0|handout:5>{\cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}}  
+            \\
+            \onslide*<0|handout:2->{\opensetgc}
+            \onslide*<0|handout:3->{\arrow[rr, "\vindex", color=functor]}
+            \onslide*<0|handout:2,3,5>{\arrow[u, "\sheaf_{\gbasec, \gtotalc}", maps to, color=sheaf]}
+            \onslide*<0|handout:4>{\arrow[d, dashed, maps to, "\sheaf_{\gbasec, \vindexpullc\dtotalc}"',color=sheaf]} \&  \& 
+            \onslide*<0|handout:1,3->{\opensetc}
+            \onslide*<0|handout:1,3-4>{\arrow[d, "\sheaf_{\dbasec, \dtotalc}", maps to, color=sheaf]}
+            \onslide*<0|handout:5>{\arrow[u, dashed, maps to, "\sheaf_{\dbasec, \vindexpushc\gtotalc}"', color=sheaf]} \\
+            \onslide*<0|handout:4>{\cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}} 
+             \& \& 
+            \onslide*<0|handout:1,3-4>{\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}}
+            \onslide*<0|handout:4>{\arrow[ll, "\textcolor{functor}{\vindexpull}", color=functor]} 
+        \end{tikzcd}
+    \end{equation*}
+    \only<0|handout:1>{
+        \begin{itemize}
+            \item $\dfiberc \hookrightarrow \dtotalc \xrightarrow{\pi} \dbasec$
+            \item $\sheafc_{\dbasec, \dtotalc}:\opensetc \mapsto \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}, \opensetc \subset \dbasec$
+            \item $\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \ni \dsectionc: \opensetc \rightarrow \dfiberc\restriction_{\opensetc}$
+            \item $\dsectionc(\dbasepointc) = \{f_0:v_0, \cdots,\}, \dbasepointc \in \opensetc$
+        \end{itemize}
+    }
+    \only<0|handout:4>{
+        \begin{itemize}
+            \item $\vindexpullc\dfiberc \hookrightarrow \vindexpullc\dtotalc \xrightarrow{\pi} \gbasec$
+            \item $\vindexpullc\sheafc_{\dbasec,\dtotalc}:\opensetgc \mapsto \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}, \opensetgc \subset \gbasec$
+            \item $\cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} \ni \vindexpullc\dsectionc: \opensetgc \rightarrow \vindexpullc\dfiberc\restriction_{\opensetgc}$
+            \item $\vindexpullc\dsectionc(\gbasepointc) = \dsectionc(\vindexc(\gbasepointc)) = \dsectionc(\dbasepointc)$
+        \end{itemize}
+    }
+    \only<0|handout:2>{
+        \begin{itemize}
+            \item $\gfiberc \hookrightarrow \gtotalc \xrightarrow{\pi} \gbasec$
+            \item $\sheafc_{\gbasec, \gtotalc}:\opensetgc \mapsto \cgamma{\opensetgc}{\dtotalc\restriction_{\opensetgc}}, \opensetgc \subset \gbasec$
+            \item $\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \ni \gsectionc: \opensetgc \rightarrow \gfiberc\restriction_{\opensetgc}$
+            \item $\gsectionc(\gbasepointc) = \{d_0, \cdots\}, \gbasepointc \in \opensetgc$
+        \end{itemize}
+    }
+    \only<0|handout:5>{
+        \begin{itemize}
+            \item $\vindexpushc\gfiberc \hookrightarrow \vindexpushc\gtotalc \xrightarrow{\pi} \dbasec$
+            \item $\vindexpushc\sheafc_{\gbasec, \gtotalc}:\opensetc \mapsto \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}, \opensetc \subset \dbasec$
+            \item $\cgamma{\opensetc}{\gtotalc\restriction_{\opensetc}} \ni \gsectionc: \opensetc \rightarrow \gfiberc\restriction_{\opensetc}$
+            \item $\vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexpre(\dbasepointc)} = \gsectionc(\gbasepointc) \forall \gbasepointc \in \vindexpre(\dbasepointc)$ 
+        \end{itemize}
+    }
+\end{frame}
+
+%add transition slide->possibly call back 
+\subsection{Artist}
+\begin{frame}
+    \frametitle<0|handout:1>{Artist: \vartistc: \sheafc_{\dbasec, \dtotalc} \rightarrow \sheafc_{\gbasec, \gotalc}}
+    \frametitle<0|handout:2>{Data Space: $\vartistc_{\opensetc}: \sheafc_{\dbasec, \dtotalc} \textcolor{artist}{\Rightarrow} \vindexpushc\sheafc_{\gbasec, \gtotalc}$}
+    \frametitle<0|handout:3>{Display Space: $\vartist_{\opensetgc}: \vindexpullc\sheafc_{\dbasec, \dtotalc} \textcolor{artist}{\Rightarrow} \sheafc_{\gbasec, \gtotalc}$}
+    \frametitle<0|handout:4>{Artist: \vartistc: \cgamma{\opensetc}{\dtotalc} \rightarrow \cgamma{\opensetgc}{\gtotalc}}
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement=\&]
+          \onslide*<0|handout:1>{\sheafc_{\gbasec, \gtotalc}} 
+          \onslide*<0|handout:4>{\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}}
+          \onslide*<0|handout:1,4>{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]}  
+          \& \& 
+          \onslide*<0|handout:1>{\vindexpushc\sheafc_{\gbasec, \gtotalc}}
+          \onslide*<0|handout:4>{\cgamma{\opensetc}{\vindexpushc \gtotalc\restriction_{\opensetc}}}
+          \\
+          \& \& \\
+          \onslide*<0|handout:1>{\textcolor{functor}{\vindexpull}\sheafc_{\dbasec, \dtotalc}}
+          \onslide*<0|handout:4>{\cgamma{\opensetgc}{\vindexpullc \dtotalc\restriction_{\opensetgc}}}
+          \onslide*<0|handout:1,4>{\arrow[uu, "\vartistc_{\opensetgc}", color=artist, Rightarrow]}
+          \&  \& 
+          \onslide*<0|handout:1>{\sheafc_{\dbasec, \dtotalc}}
+          \onslide*<0|handout:4>{\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}}
+          \onslide*<0|handout:1,4>{\arrow[ll, "\textcolor{functor}{\vindexpull}", color=functor]} 
+          \onslide*<0|handout:1,4>{\arrow[uu, "\vartistc_{\opensetc}"', color=artist, Rightarrow]}
+          \onslide*<4-|handout:1,4>{\arrow[lluu, "\vartistc"' description, dashed, color=artist, Rightarrow]}
+        \end{tikzcd}
+      \end{equation*}
+
+      \onslide*<1|handout:2>{
         \begin{equation*}
@@
         \opensetc_1 
-        \arrow[ll, "\sheafc_{\dtotalc}", color=sheaf, maps to] 
-        \arrow[rr, "\sheafc_{\gtotalpushc}"', color=sheaf, maps to] \&  \& 
+        \arrow[ll, "\sheafc_{\dbasec,\dtotalc}", color=sheaf, maps to] 
+        \arrow[rr, "\sheafc_{\dbasec,\gtotalpushc}"', color=sheaf, maps to] \&  \& 
         \cgamma{\opensetc_1}{\gtotalpushc} 
@@
         \opensetc_2 
-        \arrow[ll, "\sheafc_{\dtotalc}"', color=sheaf, maps to] 
-        \arrow[rr, "\sheafc_{\gtotalpushc}", color=sheaf, maps to] 
+        \arrow[ll, "\sheafc_{\dbasec,\dtotalc}"', color=sheaf, maps to] 
+        \arrow[rr, "\sheafc_{\dbasec,\gtotalpushc}", color=sheaf, maps to] 
         \arrow[uu, "\iota" description, hook, color=base] \& \& 
@@
 
-    \onslide*<0|handout:2>{
+    \onslide*<0|handout:3>{
         \begin{equation*}
@@
     \opensetgc_1 
-    \arrow[ll, "\sheafc_{\dtotalpullc}", color=sheaf, maps to] 
-    \arrow[rr, "\sheafc_{\gtotalc}"', color=sheaf, maps to] \&  \& 
+    \arrow[ll, "\sheafc_{\gbasec, \dtotalpullc}", color=sheaf, maps to] 
+    \arrow[rr, "\sheafc_{\gbasec,\gtotalc}"', color=sheaf, maps to] \&  \& 
     \cgamma{\opensetgc_1}{\gtotalc} 
@@
     \opensetgc_2 
-    \arrow[ll, "\sheafc_{\dtotalpullc}"', color=sheaf, maps to] 
-    \arrow[rr, "\sheafc_{\gtotalc}", color=sheaf, maps to] 
+    \arrow[ll, "\sheafc_{\gbasec,\dtotalpullc}"', color=sheaf, maps to] 
+    \arrow[rr, "\sheafc_{\gbasec, \gtotalc}", color=sheaf, maps to] 
     \arrow[uu, "\iota" description, hook, color=base] \& \& 
@@
     \end{equation*}}    
-
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement=\&]
-          \onslide*<0|handout:3->{\sheafc_{\gbasec, \gtotalc}} 
-          \onslide*<0|handout:3->{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]}  
-          \& \& 
-          \onslide*<0|handout:3->{\textcolor{functor}{\vindexpush}\sheafc_{\gbasec, \gtotalc}}
-          \\
-          \& \& \\
-          \onslide*<0|handout:3->{\textcolor{functor}{\vindexpull}\sheafc_{\dbasec, \dtotalc}}
-          \onslide*<0|handout:3->{\arrow[uu, "\vartistc_{\opensetgc}", color=artist, Rightarrow]}
-          \&  \& 
-          \onslide*<0|handout:3->{\sheafc_{\dbasec, \dtotalc}}
-          \onslide*<0|handout:3->{\arrow[ll, "\textcolor{functor}{\vindexpull}", color=functor]} 
-          \onslide*<0|handout:3->{\arrow[uu, "\vartistc_{\opensetc}"', color=artist, Rightarrow]}
-          \onslide*<4-|handout:3->{\arrow[lluu, "\vartistc"' description, dashed, color=artist, Rightarrow]}
-        \end{tikzcd}
-      \end{equation*}
-
-    \onslide<0|handout:3>{
+    \onslide<0|handout:1>{
         \begin{equation*}
@@
     \onslide<0|handout:4>{
-        \begin{equation*}
-            \sheafc_{\dbasec, \dtotalc} \textcolor{artist}{\xrightarrow{\vartistc}} \sheafc_{\gbasec, \gtotalc}
-        \end{equation*}
-        \begin{equation*}
-             (\sheafc_{\dbasec,\dtotalc} \textcolor{functor}{\xmapsto{\vindexpullc}} \vindexpullc \sheafc_{\dbasec,\dtotalc}) \textcolor{artist}{\xRightarrow{\vartistc_{\opensetgc}}} \sheafc_{\gbasec, \gtotalc} \tag{$\vartistc_{\opensetgc} \circ \vindexpullc = \vartistc$}
-        \end{equation*}
-        \begin{equation*}
-            (\sheafc_{\dbasec,\dtotalc} \textcolor{artist}{\xRightarrow{\vartistc}} \sheafc_{\gbasec, \gtotalc}) \textcolor{functor}{\xmapsto{\vindexpushc}} \vindexpushc \sheafc_{\gbasec, \gtotalc} \tag{$\vindexpushc \circ \vartistc = \vartistc_{\opensetc}$}
-        \end{equation*}
-        \begin{equation*}
-            ((\sheafc_{\dbasec,\dtotalc} \textcolor{functor}{\xmapsto{\vindexpullc}} \vindexpullc \sheafc_{\dbasec,\dtotalc}) \textcolor{artist}{\xRightarrow{\vartistc_{\opensetgc}}} \sheafc_{\gbasec, \gtotalc})\textcolor{functor}{\xmapsto{\vindexpushc}} \vindexpushc \sheafc_{\gbasec, \gtotalc} \tag{$\vindexpushc \circ \vartistc_{\opensetgc} \circ \vindexpullc = \vartistc_{\opensetc}$}
-        \end{equation*}
-        }
-\end{frame}
-
-\begin{frame}{Output}
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
+        \begin{alertblock}{pull data section \dsectionc\ over graphic space $\gbasepointc \in \gbasec$}
+            $\vindexpullc\dsectionc(\gbasepointc) = \dsectionc(\vindexc(\gbasepointc)) = \dsectionc(\dbasepointc)$
+        \end{alertblock}
+        \begin{alertblock}{push graphic \gsectionc\ section over data space $\dbasepointc \in \dbasec$}
+            $\vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexpre(\dbasepointc)} = \gsectionc(\gbasepointc) \forall \gbasepointc \in \vindexpre(\dbasepointc)$ 
+        \end{alertblock}
+        }
+\end{frame}
+
+\begin{frame}{Artist as Profunctor}
+    \begin{tikzcd}[ampersand replacement=\&]
         \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
-            \arrow[rr, "\vartistc", Rightarrow, color=artist] \&  \& 
-            \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
-            \arrow[rrd, "\delta"'] 
-            \arrow[rr, "render"] \&  \& visualization 
-            \arrow[d, "measure" description] \\ 
-            \&  \&  \&  \& M
-            \end{tikzcd}
-    \end{equation*}
+        \arrow[rr, "\vartistc", Rightarrow, color=artist] 
+        \&  \& \
+        \cgamma{\opensetgc}{\dtotalc\restriction_{\opensetgc}}\\
+        \& \&                                                                                 \\
+        \opensetc 
+        \arrow[uu, "{\sheafc_{\dbasec, \dtotalc}}", maps to, color=sheaf] 
+        \&  \& 
+        \opensetgc 
+        \arrow[ll, "\vindexc"', color=functor] 
+        \arrow[uu, "{\sheafc_{\gbasec, \gtotalc}}"', maps to, color=sheaf]
+        \end{tikzcd}
 \end{frame}
@@
     \begin{equation*}
+        \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
+            \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
+            \arrow[rr, "\vartistc", Rightarrow, color=artist] \&  \& 
+            \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+            \arrow[rrd, "\delta"'] 
+            \arrow[rr, "render"] \&  \& visualization 
+            \arrow[d, "measure" description] \\ 
+            \&  \&  \&  \& M
+            \end{tikzcd}
+    \end{equation*}
+\end{frame}
+
+\begin{frame}{Checking if \vartistc\ is equivariant}
+    \begin{equation*}
         \begin{tikzcd}[ampersand replacement=\&]
@@
     \end{equation*}
-        there exists a monoid homomorphism $\equivc: \dfuncc_{\dtotalc} \rightarrow \dfuncc_{M}$ such that 
+        there exists a functor $\equivc: \dfuncc_{\dtotalc} \rightarrow \dfuncc_{M}$ such that 
         \begin{description}
@@
         \begin{tikzcd}[ampersand replacement = \&, column sep=2ex]
-                \dfunctc_0 \circ \dfuncpullc \dtotalc_0 
+                \dfunctc_0 \dfuncpullc \dtotalc_0 
                 \arrow[rdd, "\pi" description] 
                 \arrow[r, "\cdots \otimes \cdots", phantom] \& 
-                \dfunctc_i \circ \dfuncpullc \dtotalc_i 
+                \dfunctc_i  \dfuncpullc \dtotalc_i 
                 \arrow[dd, "\pi" description] 
                 \arrow[r, "\cdots \otimes \cdots", phantom] \& 
-                \dfunctc_n \circ \dfuncpullc \dtotalc_n 
+                \dfunctc_n \dfuncpullc \dtotalc_n 
                 \arrow[ldd, "\pi" description] \& = \& 
                 \prod\limits_{i=0}^{n} 
-                \dfunctc_i \circ \dfuncpullc \dtotalc 
+                \dfunctc_i \dfuncpullc \dtotalc 
                 \arrow[dd, "\pi" description] \\    
@@
                 \& 
-                {(\dfunchc=\dfunchc_{i \in [0,n]}) \circ \dbasec }                                             \& \& = \& 
-                \dfunchc \circ \dbasec                                                        
+                {(\dfunchc=\dfunchc_{i \in [0,n]})\dbasec }                                             \& \& = \& 
+                \dfunchc \dbasec                                                        
                 \end{tikzcd}                                              
@@
             \&  \& 
-            \onslide*<2-|handout:1>{\textcolor{set}{\big\{}\gsectionc:\opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} 
-              \bigm{\vert} \delta(\vartistc(\dsectionc)) = \equivc(\dsectionc) \textcolor{set}{\big\}}} 
+            \onslide*<2-|handout:1>{\textcolor{set}{\big\{}\gsectionc 
+              \bigm{\vert} \delta(\vartistc(\dfuncc_{\dtotalc}(\dsectionc))) = \dfuncc_{M}(\delta(\gsectionc)) \textcolor{set}{\big\}}} 
             \onslide*<2-|handout:1>{\arrow[ll, "\textcolor{set}{\mathlarger{\mathlarger{\supset}}}" description, hook', color=white]} \\ 
@@
             \onslide*<1-|handout:1>{\opensetgc} 
-            \onslide*<1-|handout:1>{\arrow[uu, "\sheafc_{\gtotalc}", color=sheaf, maps to]} 
+            \onslide*<1-|handout:1>{\arrow[uu, "\sheafc_{\gbasec,\gtotalc}", color=sheaf, maps to]} 
             \onslide*<2-|handout:1>{\arrow[rruu, "\vartistc(\sheafc_{\dtotalc})\coloneqq \sheafc_{\vartistc}" description, dashed, color=sheaf, maps to]} 
@@
 \end{alertblock}
+\end{frame}
+\begin{frame}{Remapping Components}
+    \begin{block}<>{Functor $\vchannel$}
+        $\vchannelc$ is a functor from on fiber categories $\vchannelc: \mathcal{\dfiberc} \rightarrow \mathcal{\vfiberc}$
+
+    \end{block}
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement=\&]
+            \dfiberc 
+            \arrow[rr, "\vchannelc_{\dfiberc}"] 
+            \arrow[rrdd, "{\vchannelc_{\dfiberc}}^{\prime}"', dotted] 
+            \&  \& 
+            \vfiberc \coloneqq \dfiberc^{\prime}  
+            \arrow[dd, "\vchannelc_{\vfiberc}"] \\
+            \&  \& \\
+            \&  \& \vfiberc^{\prime}                           
+        \end{tikzcd}
+    \end{equation*}
 \end{frame}
@@
 
-\begin{frame}{Remapping Components}
-    \begin{block}<>{Functor $\vchannel$}
-        $\vchannelc$ is a functor from on fiber categories $\vchannel: \mathcal{\dfiberc} \rightarrow \mathcal{\vfiberc}$
-
-    \end{block}
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement=\&]
-            \dfiberc 
-            \arrow[rr, "\vchannelc_{\dfiberc}"] 
-            \arrow[rrdd, "{\vchannelc_{\dfiberc}}^{\prime}"', dotted] 
+\begin{frame}{Combining Components}
+    \begin{tikzcd}[ampersand replacement = \&]
+    \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
+    \arrow[r, "\vchannelc", color=artist] \& 
+    \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} 
+    \arrow[r, "\vmarkc_*", color=artist] 
+    \arrow[rr, "\vmarkc", bend left, color=artist] \& 
+    \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} \& 
+    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+    \arrow[l, "\vindexpushc"', color=functor, dashed]\\ 
+    \& \\ \& 
+    \opensetc 
+    \arrow[luu, "{\sheafc_{\dbasec, \dtotalc}}" description, maps to, color=sheaf] 
+    \arrow[uu, "{\sheafc_{\dbasec, \vtotalc}}" description, maps to, color=sheaf] 
+    \arrow[ruu, "{\sheafc_{\dbasec, \gtotalc}}" description, maps to, color=sheaf] 
     \& \& 
-            \vfiberc \coloneqq \dfiberc^{\prime}  
-            \arrow[dd, "\vchannelc_{\vfiberc}"] \\
-            \&  \& \\
-            \&  \& \vfiberc^{\prime}                           
-        \end{tikzcd}
-    \end{equation*}
+    \opensetgc 
+    \arrow[uu, "{\sheafc_{\gbasec, \gtotalc}}" description, maps to, color=sheaf] 
+    \arrow[ll, "\vindexc"', dashed, color=functor]
+    \end{tikzcd}
+
+    \begin{alertblock}{Equivariance}
+        \item[equivariance] $\delta(\vmarkc(\vchannelc\dfuncc_{\dtotalc}(\dsectionc))) = \dfuncc_{M}(\delta(\dsectionc))$ 
+        \item[continuity] the assembly function $\vmarkc$ is defined as  
+        \begin{equation}
+        \vmarkc_*(\vsection(\dbasepointc))(\gbasepointc) = \vmarkc(\vsectionc(\vindex(\gbasepointc)))   
+        \end{equation}
+        such that there must exist a deformation retract $\vindexc:\gbasec \rightarrow \dbasec$ for all points $\gbasepointc \in \gbase$. 
+    \end{alertblock}
 \end{frame}
@@
     \begin{figure}
-        \includegraphics{../paper/figures/q_hat.p}
-    \end{figure}
-\end{frame}
-
-\begin{frame}{implement}
-    \begin{tikzcd}[ampersand replacement = \&]
-    \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
-    \arrow[r, "\vchannelc"] \& 
-    \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} 
-    \arrow[r, "\vmarkc_*"] 
-    \arrow[rr, "\vmarkc", bend left] \& 
-    \cgamma{\opensetc}{\gtotalc\restriction_{\opensetc}} \& 
-    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
-    \arrow[l, "\vindexcpush"]\\ 
-    \& \\ \& 
-    \opensetc 
-    \arrow[luu, "{\sheafc_{\dbasec, \dtotalc}}" description, maps to] 
-    \arrow[uu, "{\sheafc_{\dbasec, \vtotalc}}" description, maps to] 
-    \arrow[ruu, "{\sheafc_{\dbasec, \gtotalc}}" description, maps to] 
-    \& \& 
-    \opensetgc \arrow[uu, "{\sheafc_{\gbasec, \gtotalc}}" description, maps to] 
-    \arrow[ll, "\vindexc"', dashed]
-    \end{tikzcd}
+        \includegraphics[scale=.15]{../paper/figures/qhat.png}
+    \end{figure}
 \end{frame}

### 2022-02-23 slides/main.tex
fixed broken slides

@@
 %add transition slide->possibly call back 
-\subsection{Artist}
+\section{Artist}
+
 \begin{frame}
-    \frametitle<0|handout:1>{Artist: \vartistc: \sheafc_{\dbasec, \dtotalc} \rightarrow \sheafc_{\gbasec, \gotalc}}
+    \frametitle<0|handout:1>{Artist: $\vartistc: \sheafc_{\dbasec, \dtotalc} \rightarrow \sheafc_{\gbasec, \gtotalc}$}
     \frametitle<0|handout:2>{Data Space: $\vartistc_{\opensetc}: \sheafc_{\dbasec, \dtotalc} \textcolor{artist}{\Rightarrow} \vindexpushc\sheafc_{\gbasec, \gtotalc}$}
     \frametitle<0|handout:3>{Display Space: $\vartist_{\opensetgc}: \vindexpullc\sheafc_{\dbasec, \dtotalc} \textcolor{artist}{\Rightarrow} \sheafc_{\gbasec, \gtotalc}$}
-    \frametitle<0|handout:4>{Artist: \vartistc: \cgamma{\opensetc}{\dtotalc} \rightarrow \cgamma{\opensetgc}{\gtotalc}}
+    \frametitle<0|handout:4>{Artist: $\vartistc: \cgamma{\opensetc}{\dtotalc} \rightarrow \cgamma{\opensetgc}{\gtotalc}$}
+
     \begin{equation*}
@@
     \begin{tikzcd}[ampersand replacement =\&]
-    \cgamma{\opensetgc_1}{\dtotalpullc} 
+    \cgamma{\opensetgc_1}{\dtotalpullc\restriction_{\opensetgc_1}} 
     \arrow[dd, "\iota^*"', hook, color=set] 
@@
     \arrow[rr, "\sheafc_{\gbasec,\gtotalc}"', color=sheaf, maps to] \&  \& 
-    \cgamma{\opensetgc_1}{\gtotalc} 
+    \cgamma{\opensetgc_1}{\gtotalc\restriction_{\opensetgc_1}} 
     \arrow[dd, "\iota^*", hook', color=set] \\
@@
     \arrow[rrrr, dotted, color=artist, Rightarrow] \&\&\&\& {} \\
-    \cgamma{\opensetgc_2}{\dtotalpullc} 
+    \cgamma{\opensetgc_2}{\dtotalpullc\restriction_{\opensetgc_2}} 
     \arrow[rrrr, "\vartistc_{\opensetgc_2}"', bend right, color=artist, Rightarrow] \&  \& 
@@
     \arrow[uu, "\iota" description, hook, color=base] \& \& 
-    \cgamma{\opensetgc_2}{\gtotalc}   
+    \cgamma{\opensetgc_2}{\gtotalc\restriction_{\opensetgc_2}}   
     \end{tikzcd}
@@
         \begin{alertblock}{pull data section \dsectionc\ over graphic space $\gbasepointc \in \gbasec$}
-            $\vindexpullc\dsectionc(\gbasepointc) = \dsectionc(\vindexc(\gbasepointc)) = \dsectionc(\dbasepointc)$
+            \begin{equation*}
+                \vindexpullc\dsectionc(\gbasepointc) = \dsectionc(\vindexc(\gbasepointc)) = \dsectionc(\dbasepointc)
+            \end{equation*}
         \end{alertblock}
         \begin{alertblock}{push graphic \gsectionc\ section over data space $\dbasepointc \in \dbasec$}
-            $\vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexpre(\dbasepointc)} = \gsectionc(\gbasepointc) \forall \gbasepointc \in \vindexpre(\dbasepointc)$ 
+            \begin{equation*}
+                \vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexpre(\dbasepointc)} = \gsectionc(\gbasepointc) \forall \gbasepointc \in \vindexpre(\dbasepointc)
+            \end{equation*}
         \end{alertblock}
@@
 
-\section{Equivariant Artist}
+\subsection{Equivariance}
+
 \begin{frame}{Transform Data}
@@
             \sheafc_{\dbasec^{\prime}, \dfuncpullc\dtotalc} 
-        \arrow[r, "\vartistc", Rightarrow, color=artist]    \& 
-        \sheafc_{\gbasec^{\prime},\dfuncpullc{\gbasec}\gtotalc}    
-        \end{tikzcd} 
-    \end{equation*}
-
-    \begin{proposition}{}
+            \arrow[r, "\vartistc", Rightarrow, color=artist]\& 
+            \sheafc_{\gbasec^{\prime},\dfuncpullc\gtotalc}
+        \end{tikzcd} 
+        \end{equation*}
+
+    \begin{alertblock}{}
         \begin{align*}
             \vartistc(\dsectionc^{a}) = \vartistc(\dsectionc^{b}) &\centernot\implies \dsectionc^a = \dsectionc^b \\
-            \vartistc(\dsectionc^{a})  = \vartistc(\dsectionc^{b}) &\implies \vartistc(\dfuncc_{\dtotalc}(\dsectionc^{a})) = \vartistc(\dfuncc_{\dtotalc}(\dsectionc^{b})) \
+            \vartistc(\dsectionc^{a})  = \vartistc(\dsectionc^{b}) &\implies \vartistc(\dfuncc_{\dtotalc}(\dsectionc^{a})) = \vartistc(\dfuncc_{\dtotalc}(\dsectionc^{b}))
         \end{align*}
-    \end{proposition}
-\end{frame}
-
+    \end{alertblock}
+\end{frame}
+
+\begin{comment}
 \begin{frame}{Equivariant $\dfunchc$}
@@
 
-\section{Complex Structure}
 \begin{frame}{Multiple Bundles: $\dfuncc = (\dfunchc, \prod\limits_{i=0}^{n}\dfunctc_i)$} 
@@
 
+\begin{frame}{Functor $\vchannelc: \dfiberc \rightarrow \vfiberc$}
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement=\&]
+            \dfiberc 
+            \arrow[rr, "\vchannelc_{\dfiberc,\vfiber}"] 
+            \arrow[rrdd, "{\vchannelc_{\dfiberc,{\vfiberc}^{\prime}}"', dotted] 
+            \&  \& 
+            \vfiberc \coloneqq \dfiberc^{\prime}  
+            \arrow[dd, "\vchannelc_{{\dfiberc}^{\prime}, {\vfiberc}^{\prime}}"] \\
+            \&  \& \\
+            \&  \& \vfiberc^{\prime}                           
+        \end{tikzcd}
+    \end{equation*}
+    \begin{alertblock}
+        $\vchannelc$ is a functor s.t. if there exists a map ${\vchannelc}_{\dfiber, \vfiber}: \dfiber \mapsto \vfiber$ and 
+        ${\vchannelc}_{{\dfiber}^{\prime}, {\vfiber}^{\prime}}: {\dfiber}^{\prime} \mapsto {\vfiber}^{\prime}$ then by composition there exists a map ${\vchannelc}_{\dfiberc, {\vfiberc}^{\prime}}: \dfiberc \rightarrow {\vfiberc}^{\prime}$
+    \end{alertblock}
+\end{frame}
+
 \begin{frame}{Equivariant Data to Components: $\vchannel: \sheaf_{\dbasec,\dtotalc} \rightarrow \sheaf_{\dbasec, \vtotalc}$}
@@
 \end{alertblock}
-\end{frame}
-\begin{frame}{Remapping Components}
-    \begin{block}<>{Functor $\vchannel$}
-        $\vchannelc$ is a functor from on fiber categories $\vchannelc: \mathcal{\dfiberc} \rightarrow \mathcal{\vfiberc}$
-
-    \end{block}
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement=\&]
-            \dfiberc 
-            \arrow[rr, "\vchannelc_{\dfiberc}"] 
-            \arrow[rrdd, "{\vchannelc_{\dfiberc}}^{\prime}"', dotted] 
-            \&  \& 
-            \vfiberc \coloneqq \dfiberc^{\prime}  
-            \arrow[dd, "\vchannelc_{\vfiberc}"] \\
-            \&  \& \\
-            \&  \& \vfiberc^{\prime}                           
-        \end{tikzcd}
-    \end{equation*}
 \end{frame}

### 2022-02-23 slides/main.tex
compiles up to construction

@@
 
+\subsubsection{Natural Transforms}
 \begin{frame}{Natural Transformation $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\Rightarrow} \textcolor{functor}{G}$}
@@
 
-\begin{frame}
+\subsubsection{Pullback and Pushforward}
+\begin{frame}{}
     \frametitle<3|handout:3>{Function: $\vindexc: \gbasec \rightarrow \dbasec$}
@@
     \end{equation*}
-    \only<0|handout:1>{
+    \only<0|handout:1>{% data sheaf
         \begin{itemize}
@@
     }
-    \only<0|handout:4>{
+    \only<0|handout:4>{%pulled back data sheaf
         \begin{itemize}
@@
     }
-    \only<0|handout:2>{
+    \only<0|handout:2>{% graphic sheaf
         \begin{itemize}
@@
     }
-    \only<0|handout:5>{
+    \only<0|handout:5>{%pushed forward sheaf
         \begin{itemize}
@@
             \item $\vindexpushc\sheafc_{\gbasec, \gtotalc}:\opensetc \mapsto \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}, \opensetc \subset \dbasec$
-            \item $\cgamma{\opensetc}{\gtotalc\restriction_{\opensetc}} \ni \gsectionc: \opensetc \rightarrow \gfiberc\restriction_{\opensetc}$
-            \item $\vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexpre(\dbasepointc)} = \gsectionc(\gbasepointc) \forall \gbasepointc \in \vindexpre(\dbasepointc)$ 
+            \item $\cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} \ni \vindexpushc\gsectionc: \opensetc \rightarrow \vindexpushc \gfiberc\restriction_{\opensetc}$
+            \item $\vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexpre(\dbasepointc)} = \gsectionc(\gbasepointc)\;\forall \gbasepointc \in \vindexpre(\dbasepointc)$ 
         \end{itemize}
     }
+\end{frame}
+
+\begin{frame}{Example: Scatter Plot}
+    \begin{figure}
+        \includegraphics[scale=.1]{figures/math/push_pull_scatter.png}
+    \end{figure}
+\end{frame}
+
+\begin{frame}{Example: Line Plot}
+    \begin{figure}
+        \includegraphics[scale=.1]{figures/math/push_pull_line.png}
+    \end{figure}
+\end{frame}
+
+\begin{frame}{Example: Image}
+    \begin{figure}
+        \includegraphics[scale=.1]{figures/math/push_pull_image.png}
+    \end{figure}
+\end{frame}
+
+\begin{frame}{Example: Heatmap}
+    \begin{figure}
+        \includegraphics[scale=.1]{figures/math/push_pull_heat.png}
+    \end{figure}
+\end{frame}
+
+\begin{frame}{Example: Sphere}
+    \begin{figure}
+        \includegraphics[scale=.1]{figures/math/push_pull_sphere.png}
+    \end{figure}
 \end{frame}
@@
     \frametitle<0|handout:4>{Artist: $\vartistc: \cgamma{\opensetc}{\dtotalc} \rightarrow \cgamma{\opensetgc}{\gtotalc}$}
-
+    \onslide*<0|handout:1,4>{
     \begin{equation*}
@@
         \end{tikzcd}
-    \end{equation*}
+    \end{equation*}}
 
@@
 \subsection{Equivariance}
-
 \begin{frame}{Transform Data}
@@
 
-    \begin{alertblock}{}
+    \begin{alertblock}{Artist Constraints}
         \begin{align*}
@@
 
-\begin{comment}
 \begin{frame}{Equivariant $\dfunchc$}
@@
     \end{equation*}}         
-        \item[both]{given $\dsectionc^{a} \in \sheafc_{\dbasec_a, \dtotalc_a}, \dsectionc^{b} \in \sheafc_{\dbasec_b, \dtotalc_b}$ and $\dbasepointc \in \dbasec_{a}\bigsqcup\limits_{\dbasec_{c}} \dbasec_{b}$ then $q(\equivc(\dsection^a(\dbasepointc))) = q^{\prime}(\equivc(\dsectionc^b(\dbasepointc)))$ when $p(\dfiberc^{a}\restriction_{\dbasepointc}) = p^{\prime}(\dfiberc^{b}\restriction_{\dbasepointc})$}
+        \item[both]{given $\dsectionc^{d} \in \sheafc_{\dbasec_{d}, \dtotalc_{d}}, \dsectionc^{e} \in \sheafc_{\dbasec_{e}, \dtotalc_{e}}$ and $\dbasepointc \in \dbasec_{d}\bigsqcup\limits_{\dbasec_{f}} \dbasec_{e}$ then $q(\equivc(\dsection^{d}(\dbasepointc))) = q^{\prime}(\equivc(\dsectionc^{e}(\dbasepointc)))$ when $p(\dfiberc^{d}\restriction_{\dbasepointc}) = p^{\prime}(\dfiberc^{e} \restriction_{\dbasepointc})$}
     \end{description}
@@
     \begin{equation*}
-        \label{eq:artist:artist_sheaf}
         \begin{tikzcd}[ampersand replacement = \&, column sep=tiny, row sep=huge]
@@
             \&  \& 
-            \onslide*<2-|handout:1>{\textcolor{set}{\big\{}\gsectionc 
-              \bigm{\vert} \delta(\vartistc(\dfuncc_{\dtotalc}(\dsectionc))) = \dfuncc_{M}(\delta(\gsectionc)) \textcolor{set}{\big\}}} 
+            \onslide*<2-|handout:1>{\textcolor{set}{\big\{}\gsectionc \bigm{\vert} 
+              \delta(\vartistc(\dfuncc_{\dtotalc}(\dsectionc))) = \dfuncc_{M}(\delta(\gsectionc)) \textcolor{set}{\big\}}} 
             \onslide*<2-|handout:1>{\arrow[ll, "\textcolor{set}{\mathlarger{\mathlarger{\supset}}}" description, hook', color=white]} \\ 
@@
 \end{frame}
+
+\begin{comment}
 
@@
 \end{frame}
-
-\begin{comment}
-\begin{frame}{Scatter: $\vmark(xpos, ypos)(\alpha, \beta)$}
-    \begin{figure}[H]
-        \begin{overprint}
-            \onslide<1|handout:0>\includegraphics[width=1\linewidth]{figures/math/scatter_q.png}
-            \onslide<2|handout:0>\includegraphics[width=1\linewidth]{figures/math/scatter_with_s.png}
-        \end{overprint}
-    \end{figure}   
-\end{frame}
-
-\begin{frame}{Line: $\vmark(xpos, \hat{n_{1}}, ypos, \hat{n_{2}})(\alpha, \beta)$ }
-    \begin{figure}[H]
-        \begin{overprint}
-            \onslide<1|handout:0>\includegraphics[width=1\linewidth]{figures/math/line_q.png}
-            \onslide<2|handout:0>\includegraphics[width=1\linewidth]{figures/math/line_with_s.png}
-        \end{overprint}
-    \end{figure}   
-\end{frame}
-
-\begin{frame}{Image $\vmark(xpos, ypos, color)$}
-    \begin{figure}[H]
-        \includegraphics[width=1\textwidth]{figures/math/image.png}
-    \end{figure}
-\end{frame}
-
-\begin{frame}{Build \vmark\ over \dbase: \vmarkd}
-\begin{figure}[H]
-    \includegraphics[width=1\textwidth]{figures/math/q_hat.png}
-\end{figure}
-\begin{equation*}\label{eq:qhat_q_s}
-    \vmarkd(\vsection(\dbasepoint))(\gbasepoint) \coloneqq \vmark((\vsectionpull)(\gbasepoint))
-\end{equation*} 
-\end{frame}
-
-\begin{frame}{Composition of artists $+ \coloneqq \sqcup \dtotal_{i}$} 
-    \begin{figure}
-        \includegraphics[width=1\textwidth]{figures/math/exploding_artist.png}
-    \end{figure}
-\end{frame}
-
-\section{Code}
-\begin{frame}{TEAM driven rearchitecture of Matplotlib}
-    \begin{itemize}
-        \item complex visualizations
-        \pause 
-        \item structure preserving maps from data to visual
-        \begin{itemize}
-            \item data and graphics have equivalent continuity
-            \item properties are equivariant under monoid actions
-        \end{itemize}  
-        \pause
-        \item fiber bundles are an abstraction
-        \begin{itemize}
-            \item topologically complex heterogenous data 
-            \item target display spaces
-        \end{itemize}
-    \end{itemize}
-\end{frame}
-
-\begin{frame}[fragile]{How do we make things?}
-    \begin{figure}[H]
-        \begin{subfigure}{0.49\textwidth}
-            \includegraphics[width=\textwidth]{figures/code/scatter_0.png}
-        \end{subfigure}
-        \begin{subfigure}{0.49\textwidth}
-            \includegraphics[width=\textwidth]{figures/code/line_1.png}
-        \end{subfigure}
-    \end{figure}
-    \begin{columns}
-    \column{0.49\textwidth}
-    \begin{minted}{python}
-    fig, ax = plt.subplots()
-    artist = Point(data, transforms)
-    ax.add_artist(artist)
-    \end{minted}
-    \column{.49\textwidth}
-    \begin{minted}{python}
-    fig, ax = plt.subplots()
-    artist = Line(data, transforms)
-    ax.add_artist(artist)
-    \end{minted}
-    \end{columns}
-\end{frame}
-
-\subsection{Artist}
-\begin{frame}[fragile]{ \vchannel}
-    \begin{minted}{python}
-    cmap =  color.Categorical({'true':'deeppink', 'false':'deepskyblue'})
-    transforms = {'x': {'name': 'v4', 'encoder': lambda x: x},
-                    'y': {'name': 'v2', 'encoder': lambda x: x},
-                    'facecolors': {'name':'v3', 'encoder': cmap}, 
-                    's':{'name': None , 
-                        'encoder': lambda _: itertools.repeat(.02)}}
-    \end{minted}
-    \begin{itemize}
-        \item \mintinline{python}{lambda x: x} is identity \vchannel\
-        \item \mintinline{python}|{'name':None}| map into \vfiber\ without corresponding \dsection\
-        \item \mintinline{python}{color.Categorical} is custom \vchannel
-    \end{itemize}
-\end{frame}
-
-\begin{frame}[fragile]{\vartist} %% rewrite w/ letters in talk
-    \begin{minted}{python}
-        class ArtistClass(matplotlib.artist.Artist):
-            def __init__(self, E, V, *args, **kwargs):
-                # set properties that are specific to the artist
-                # stash the input E and V
-                super().__init__(*args, **kwargs)
-
-            def qhat(self, **args):
-                # set the properties of the graphic
-
-            def draw(self, renderer):
-                # returns tau, indexed on fiber then key 
-                tau = self.E.view(self.axes) 
-                # visual channel encoding applied fiberwise 
-                visual = {p_i: nu_i(tau_i)
-                          for p_i, nu_i, tau)i 
-                          in zip(self.V, tau)}
-                self.qhat(**visual)
-                # pass configurations off to the renderer
-                super().draw(renderer)
-        \end{minted}
-\end{frame}
-
-\begin{frame}[fragile]{\vmarkd}
-\begin{minted}{python}
-    class Point(mcollections.Collection):
-        def assemble(self, x, y, s, facecolors='C0' ):
-            # construct geometries of the circle glyphs in visual coordinates
-            # set attributes of glyphs
-
-    class Line(mcollections.LineCollection):
-        def assemble(self, x, y, color='C0'):
-            # assemble line marks as set of segments 
-    \end{minted}
-\end{frame}
-
-\subsection{Data}
-\begin{frame}[fragile]{Continuity}
-    \begin{minted}{python}
-class PointData: 
-    # Fiberbundle is consistent across all sections
-    FB = FiberBundle({'tables': ['vertex']},  
-        {'v1': float, 'v2': str, 'v3': float})
-    def tau(self, k):
-        return # tau evaluated at one point k
-
-class LineData: 
-    FB = FiberBundle({'tables': ['edge']},
-                {'x' : float, 'y':  float, 'color':mtypes.Color()})
-    def tau(self, k): 
-        return # tau evaluated on interval k
-    \end{minted}
-\end{frame}
-
-\begin{frame}[fragile]{Same Artist, Different \dtotal}
-    \begin{figure}[H]
-        \begin{subfigure}{0.49\textwidth}
-            \includegraphics[width=\textwidth]{figures/code/linec_1.png}
-        \end{subfigure}
-        \begin{subfigure}{0.49\textwidth}
-            \includegraphics[width=\textwidth]{figures/code/lined_1.png}
-        \end{subfigure}
-    \end{figure}
-\begin{minted}{python}
-LineData(FB, edge_table, vertex_table, connect=True)
-LineData(FB, edge_table, vertex_table, num_samples=2, connect=False)
-\end{minted}
-\end{frame}
-
-\begin{frame}{Proposed Work}
-\begin{figure}
-    \includegraphics[width=1\textwidth]{figures/intro/gantt.png}
-\end{figure}
-\end{frame}
-
-\begin{frame}{Acknowledgments}
-    \begin{itemize}
-        \item Professor Michael Grossberg and Dr. Thomas Caswell
-        \item Professor Haralick, Professor Vo, Professor Manovich, Dr. Hanwell
-        \item Matplotlib development team 
-        \item CZI EOSS (grant number 2019-207333) from the Chan Zuckerberg Initiative DAF, an advised fund of Silicon Valley Community Foundation
-    \end{itemize}
-
-\end{frame}
-
-\begin{frame}[allowframebreaks]{References}
-\printbibliography
-\end{frame}
-\appendix
-
-\begin{frame}{Bertin Retinal Variables}
-    \begin{figure}
-        \includegraphics[width=.7\textwidth]{figures/intro/retinal_variables.png}
-        \caption{This tabular form of Bertin's retinal variables is from Understanding Graphics \cite{malamedInformationDisplayTips2010} who reproduced it from Krygier and Wood's \textit{Making Maps: A Visual Guide to Map Design for GIS}\cite{krygierMakingMapsVisual2005}}
-    \end{figure}
-\end{frame}
-
-\begin{frame}{Fiber is all possible values a variable can be \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES}}
-    Given a space of all possible values \ftotal\
-    \begin{equation*}
-        \label{eq:data_types}
-        \begin{tikzcd}[ampersand replacement=\&]
-            \fttype \arrow[r] \arrow[d, "\pi_{\fsection}"'] \& \ftotal \arrow[d, "\pi"] \\
-            \fnames \arrow[r, "\fsection"']                  \& \ftypes       
-        \end{tikzcd}
-    \end{equation*}
-    a fiber component is the restricted space $\ftotal_{\fsection(\fname)}$. 
-    \begin{equation*}
-        \dfiber = \ftotal_{\fsection(\fname)} = \ftotal_{\ftype} 
-    \end{equation*}
-    \begin{description}
-        \item[\ftypes] data types of the variables in the dataset 
-        \item[\ftotal] disjoint union of all values of type $\ftype \in \ftypes$ 
-        \item[\fnames] variable names, $\fname \in \fnames$
-        \item[\fttype] \ftotal\ restricted to the data type of a named variable   
-    \end{description}
-\end{frame}
-
-\begin{frame}{Monoid actions}
-    A monoid \monoid\ is a set with
-    \begin{description}
-        \item[associative binary operator] $\ast:\monoid \times \monoid\rightarrow \monoid$
-        \item[identity element] $e\in \monoid$ such that $e\ast a= a \ast e = a$ for all $a \in \monoid$. 
-    \end{description}
-    \begin{block}{left monoid action}
-    A set \dfiber\ with an action\cite{nlab:action} $\bullet: \monoid\times \dfiber \rightarrow \dfiber$ with the properties:
-        \begin{align*}
-            \textbf{associativity}\;& \text{for all } f,g \in \monoid \text{ and } x\in \dfiber,\, f\bullet(g\bullet x) = (f\ast g) \bullet x\\
-            \textbf{identity}\;& \text{for all } x\in \dfiber, e\in \monoid,\,  e\bullet x = x 
-        \end{align*}
-    \end{block}
-\end{frame}
-
-\begin{frame}{Keeping track of sections with sheafs}
-    Restriction maps of a sheaf describe how local $\iota^*\dsection$ can be glued into larger sections \cite{ghristElementaryAppliedTopology2014,ghristHomologicalAlgebraData2018}
-    \begin{equation*}
-        \label{eq:sheaf}
-        \begin{tikzcd}[ampersand replacement=\&]
-            \iota^*\dtotal \arrow[d, "\pi"'] \arrow[r, "\iota^*", hook]             \& \dtotal \arrow[d, "\pi"']                  \\
-            U \arrow[r, "\iota", hook] \arrow[u, "\iota^{*}\dsection"', bend right] \& \dbase \arrow[u, "\dsection"', bend right]
-        \end{tikzcd}
-    \end{equation*}
-    The inclusion map $\iota: U \rightarrow \dbase$ pulls \dtotal\ over $U$ such that the pulled back $\iota^*\dsection$ only contains records over $U \subset \dbase$.
-\end{frame}
-
-\begin{frame}{Rendering: Define a Pixel}
-    \begin{columns}
-        \column{0.5\textwidth}
-        Given a pixel
-        \begin{equation*}
-        p=\left[y_{top}, y_{bottom}, x_{right}, x_{left}\right]
-        \end{equation*}
-        the inverse map of the bounding box 
-        \begin{equation*}
-        \gbase_{p} ={\gsection_{xy}}^{-1}(p)
-        \end{equation*}
-        is a region $\gbase_p \subset \gbase$ such that 
-        \begin{align}
-            \scriptstyle r_p &= \scriptstyle \iint\limits_{S_p} \rho_r(s)ds^{2}\\
-            \scriptstyle g_p &= \scriptstyle \iint\limits_{S_p} \rho_g(s)ds^{2}\\
-            \scriptstyle  b_p &= \scriptstyle \iint\limits_{S_p} \rho_b(s)ds^{2}
-        \end{align}
-        yields the color of the pixel. 
-        \column{0.5\textwidth}
-        \begin{figure}[H]
-            \includegraphics[width=\textwidth]{figures/math/render.png}
-        \end{figure}
-    \end{columns}
-\end{frame}
-
-\begin{frame}{$\mathcal{A}:\mathcal{E}\rightarrow\mathcal{E}$}
-    The topological artist is a sheaf map
-    \begin{equation*}
-    \vartist: \mathcal{O}(\dtotal) \rightarrow \mathcal{O}(\gtotal)
-    \end{equation*}
-    that carries homomorphism of monoid actions $\varphi: \monoid \rightarrow \monoid^{\prime}$ 
-    \cite{cegarraCohomologyMonoidsOperators2019} 
-    \begin{equation*}
-    \vartist(m\cdot \delement) = \varphi(m)\cdot \vartist(\delement) 
-    \end{equation*}
-\end{frame}
-
-\begin{frame}{Visual Channel Encoders}
-    We define the visual transformers \vchannel\ on components of the data bundle $\dsection_{i}$
-    \begin{equation*}
-        \label{eq:nu_expanded}
-        \{\vchannel_{0}, \ldots, \vchannel_{n}\}: \{\dsection_{0}, \ldots, \dsection_{n}\} \mapsto \{\vsection_{0}, \ldots, \vsection_{n}\}
-    \end{equation*}
-    as the set of equivariant maps with the constraint 
-    \begin{equation*}
-        \vchannel_i(m_{\delement}(\dtotal_i)) = \varphi(m_{\delement})(\vchannel_i(\dtotal_i))
-    \end{equation*} 
-    where $\varphi:\monoid\rightarrow \monoid^{\prime}$ carries a homomorphism of monoid actions. 
-\end{frame}
-
-\begin{frame}{\vfiber\ Components}    
-\begin{table}[H]
-    \renewcommand{\arraystretch}{2}
-    \begin{tabulary}{\textwidth}{|l|L|l|}\hline
-     $\bm{\vchannel_{i}}$                      & $\bm{\vsection_{i}}$                                                            & $\bm{codomain(\vchannel_{i}) \subset \vfiber_{i}}$  \\ \hline
-    position                    & x, y, z, theta, r                                                          & $\mathbb{R}$   \\ \hline
-    size                        & linewidth, markersize                                            & $\mathbb{R}^{+}$   \\ \hline
-    shape                       & markerstyle                                                      & $\{f_{0}, \ldots, f_{n}\}$ \\ \hline
-    color                       & color, facecolor, markerfacecolor, edgecolor  & $\mathbb{R}^{4}$ \\ \hline
-    \multirow{2}{*}{texture}    & hatch                                                            & $\mathbb{N}^{10}$\\\cline{2-3}
-                                & linestyle                                                        & $(\mathbb{R}, \mathbb{R^+}^{n, n\%2=0})$ \\ \hline              
-    \end{tabulary}
-    \label{tab:mpl_visual_variable_fiber}
-\end{table}
-\end{frame}
-\begin{frame}{Monoid Equivariance: Partial Orders}
-    \begin{figure} %break out steps/reveal
-        \includegraphics[width=1\textwidth]{figures/math/monoid_equivariant.png}
-    \end{figure}
-\end{frame}
-
-\begin{frame}{Glyph}
-    The glyph is the graphic generated by $\vmark(\gbase_{\dbasepathpoint})$ where the path connected components $\dbasepath \subset \dbase$ are defined 
-    \begin{equation*}
-    \dbasepath = \{\dbasepathpoint \in \dbase \text{ s. t. } \exists \gamma \text{ s.t. } \gamma(0)=\dbasepoint \text{ and }\gamma(1)=\dbasepathpoint\}
-    \end{equation*}
-    such that the  path $\gamma$ from \dbasepoint\ to \dbasepathpoint\ is a continuous function from the interval [0,1] and $\gbase_{\dbasepathpoint}$ is the region
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement=\&]
-            \gtotal \arrow[r, shift left] \& \gbase_\dbasepathpoint \arrow[rr, "\vindex(\gbasepoint)", shift left] \arrow[l, "\gsection(\gbase_\dbasepathpoint)"] \& \& \dbasepath_{\dbasepoint} \arrow[ll, "\vindex^{-1}(\dbasepath)"]
-            \end{tikzcd}
-        \label{eq:mark}
-    \end{equation*}
-    such that the glyph is differentiable, in keeping with Ziemkiewicz and Kosara's description of a glyph\cite{ziemkiewiczEmbeddingInformationVisualization2009}.
-\end{frame}
-
-\begin{frame}{Artist Equivalance class}
-    When artists share a base space 
-    \begin{equation*}
-        \dbase_2 \hookrightarrow \dbase_1
-    \end{equation*}
-    a composition operator can be defined such that the the artists can be considered to be acting on different components of the same section. 
-\end{frame}
-
-\begin{frame}[fragile]{Complex \vchannel}
-    \begin{minted}{python}
-    class Categorical:
-        def __init__(self, mapping):
-            # check that the conversion is to valid colors
-            assert(mcolors.is_color_like(color) for color in mapping.values())
-            self._mapping = mapping
-
-        def __call__(self, value):
-            # convert value to a color
-            return [mcolors.to_rgba(self._mapping[v]) for v in values]
-    \end{minted}
-    That we can test for action equivariance
-    \begin{minted}{python}
-    def test_nominal(values, encoder):
-        m1 = list(zip(values, encoder(values)))
-        random.shuffle(values)
-        m2 = list(zip(values, encoder(values)))
-        assert sorted(m1) == sorted(m2)
-    \end{minted}
-\end{frame}
-
-\begin{frame}[fragile]{Artist}
-    \begin{minted}{python}
-        class ArtistClass(matplotlib.artist.Artist):
-            def __init__(self, data, transforms, *args, **kwargs):
-                # properties that are specific to the graphic
-                self.data = data 
-                self.transforms = transforms
-                super().__init__(*args, **kwargs)
-
-            def assemble(self, **args):
-                # set the properties of the graphic
-
-            def draw(self, renderer):
-                # returns K, indexed on fiber then key 
-                view = self.data.view(self.axes) 
-                # visual channel encoding applied fiberwise 
-                visual = {p: t['encoder'](view[t['name']])
-                          for p, t in self.transforms.items()}
-                self.assemble(**visual)
-                # pass configurations off to the renderer
-                super().draw(renderer)
-        \end{minted}
-\end{frame}
-
-\begin{frame}[fragile]{Artists: Scatter \& Line}
-\begin{minted}{python}
-class Point(mcollections.Collection):
-    def assemble(self, x, y, s, facecolors='C0' ):
-        # construct geometries of the circle glyphs in visual coordinates
-        self._paths = [mpath.Path.circle(center=(xi,yi), radius=si) 
-                    for (xi, yi, si) in zip(x, y, s)] 
-        # set attributes of glyphs, these are vectorized 
-        # circles and facecolors are lists of the same size
-        self.set_facecolors(facecolors)
-
-class Line(mcollections.LineCollection):
-    def assemble(self, x, y, color='C0'):
-        #assemble line marks as set of segments 
-        segments = [np.vstack((vx, vy)).T for vx, vy in zip(x, y)]
-        self.set_segments(segments)
-        self.set_color(color)
-\end{minted}
-\end{frame}
-
-\begin{frame}[fragile]{View}
-\begin{minted}{python}
-def view(self, axes):
-    table = defaultdict(list)
-    for k in self.keys:
-        table['index'].append(k)
-        for (name, value) in zip(self.FB.fiber.keys(), self.tau(k)[1]):
-            table[name].append(value)
-    return table
-\end{minted}
-\begin{description}
-    \item [\mintinline{python}{VertexSimplex}] (name, value), value is scaler
-    \item [\mintinline{python}{EdgeSimplex}]  (name, value), value is [x0, ..., xn]
-\end{description}
-\end{frame}
-
-\begin{frame}[fragile]{Fiber Bundle}
-    \begin{minted}{python}
-    @dataclass
-    class FiberBundle:
-    """
-    Attributes
-    ----------
-    K: {'tables': []}
-    F: {variable name: type}
-    """
-        K: dict 
-        F: dict
-    \end{minted}
-\end{frame}
-
-\begin{frame}[fragile]{GraphLine Data Model}
-\begin{minted}{python}
-class GraphLine:
-    def __init__(self, FB, edge_table, vertex_table, num_samples=1000,
-                        connect=False):
-        #set args as attributes and generate distance
-        if connect: # test connectivity if edges are continuous
-            assert edge_table.keys() == self.FB.F.keys()
-            assert is_continuous(vertex_table)
-
-    def tau(self, k):
-        # evaluates functions defined in edge table
-        return(k, (self.edges[c][k](self.distances) 
-                        for c in self.FB.F.keys()))
-
-    def view(self, axes):
-        # walk the edge_vertex table to return the edge function
-        table = defaultdict(list)
-        for (i, (start, end)) in sorted(zip(self.ids, self.vertices), 
-                                            key=lambda v:v[1][0]):
-            table['index'].append(i)
-            # same as view for line, returns nested list
-            for (name, value) in zip(self.FB.F.keys(), self.tau(i)[1]):
-                table[name].append(value)
-        return table
-\end{minted}
-\end{frame}
 \end{comment}

### 2022-02-23 slides/scraps.tex
reworking notation and pushpull story and finishing out Q
compiles up to construction

@@
 \end{frame}
+
+\begin{equation}
+    \begin{equation*}
+        \sheafc_{\dbasec, \dtotalc} \textcolor{artist}{\xrightarrow{\vartistc}} \sheafc_{\gbasec, \gtotalc}
+    \end{equation*}
+    \begin{equation*}
+         (\sheafc_{\dbasec,\dtotalc} \textcolor{functor}{\xmapsto{\vindexpullc}} \vindexpullc \sheafc_{\dbasec,\dtotalc}) \textcolor{artist}{\xRightarrow{\vartistc_{\opensetgc}}} \sheafc_{\gbasec, \gtotalc} \tag{$\vartistc_{\opensetgc} \circ \vindexpullc = \vartistc$}
+    \end{equation*}
+    \begin{equation*}
+        (\sheafc_{\dbasec,\dtotalc} \textcolor{artist}{\xRightarrow{\vartistc}} \sheafc_{\gbasec, \gtotalc}) \textcolor{functor}{\xmapsto{\vindexpushc}} \vindexpushc \sheafc_{\gbasec, \gtotalc} \tag{$\vindexpushc \circ \vartistc = \vartistc_{\opensetc}$}
+    \end{equation*}
+    \begin{equation*}
+        ((\sheafc_{\dbasec,\dtotalc} \textcolor{functor}{\xmapsto{\vindexpullc}} \vindexpullc \sheafc_{\dbasec,\dtotalc}) \textcolor{artist}{\xRightarrow{\vartistc_{\opensetgc}}} \sheafc_{\gbasec, \gtotalc})\textcolor{functor}{\xmapsto{\vindexpushc}} \vindexpushc \sheafc_{\gbasec, \gtotalc} \tag{$\vindexpushc \circ \vartistc_{\opensetgc} \circ \vindexpullc = \vartistc_{\opensetc}$}
+    \end{equation*}
+\end{equation}
+
+    \begin{alertblock}
+        \begin{align*}
+            \vartistc(\dsectionc^{a}) = \vartistc(\dsectionc^{b}) &\centernot\implies \dsectionc^a = \dsectionc^b \\
+            \vartistc(\dsectionc^{a})  = \vartistc(\dsectionc^{b}) &\implies \vartistc(\dfuncc_{\dtotalc}(\dsectionc^{a})) = \vartistc(\dfuncc_{\dtotalc}(\dsectionc^{b}))
+        \end{align*}
+    \end{alertblock}
+
+\begin{frame}{Scatter: $\vmark(xpos, ypos)(\alpha, \beta)$}
+    \begin{figure}[H]
+        \begin{overprint}
+            \onslide<1|handout:0>\includegraphics[width=1\linewidth]{figures/math/scatter_q.png}
+            \onslide<2|handout:0>\includegraphics[width=1\linewidth]{figures/math/scatter_with_s.png}
+        \end{overprint}
+    \end{figure}   
+\end{frame}
+
+\begin{frame}{Line: $\vmark(xpos, \hat{n_{1}}, ypos, \hat{n_{2}})(\alpha, \beta)$ }
+    \begin{figure}[H]
+        \begin{overprint}
+            \onslide<1|handout:0>\includegraphics[width=1\linewidth]{figures/math/line_q.png}
+            \onslide<2|handout:0>\includegraphics[width=1\linewidth]{figures/math/line_with_s.png}
+        \end{overprint}
+    \end{figure}   
+\end{frame}
+
+\begin{frame}{Image $\vmark(xpos, ypos, color)$}
+    \begin{figure}[H]
+        \includegraphics[width=1\textwidth]{figures/math/image.png}
+    \end{figure}
+\end{frame}
+
+\begin{frame}{Build \vmark\ over \dbase: \vmarkd}
+\begin{figure}[H]
+    \includegraphics[width=1\textwidth]{figures/math/q_hat.png}
+\end{figure}
+\begin{equation*}\label{eq:qhat_q_s}
+    \vmarkd(\vsection(\dbasepoint))(\gbasepoint) \coloneqq \vmark((\vsectionpull)(\gbasepoint))
+\end{equation*} 
+\end{frame}
+
+\begin{frame}{Composition of artists $+ \coloneqq \sqcup \dtotal_{i}$} 
+    \begin{figure}
+        \includegraphics[width=1\textwidth]{figures/math/exploding_artist.png}
+    \end{figure}
+\end{frame}
+
+\section{Code}
+\begin{frame}{TEAM driven rearchitecture of Matplotlib}
+    \begin{itemize}
+        \item complex visualizations
+        \pause 
+        \item structure preserving maps from data to visual
+        \begin{itemize}
+            \item data and graphics have equivalent continuity
+            \item properties are equivariant under monoid actions
+        \end{itemize}  
+        \pause
+        \item fiber bundles are an abstraction
+        \begin{itemize}
+            \item topologically complex heterogenous data 
+            \item target display spaces
+        \end{itemize}
+    \end{itemize}
+\end{frame}
+
+\begin{frame}[fragile]{How do we make things?}
+    \begin{figure}[H]
+        \begin{subfigure}{0.49\textwidth}
+            \includegraphics[width=\textwidth]{figures/code/scatter_0.png}
+        \end{subfigure}
+        \begin{subfigure}{0.49\textwidth}
+            \includegraphics[width=\textwidth]{figures/code/line_1.png}
+        \end{subfigure}
+    \end{figure}
+    \begin{columns}
+    \column{0.49\textwidth}
+    \begin{minted}{python}
+    fig, ax = plt.subplots()
+    artist = Point(data, transforms)
+    ax.add_artist(artist)
+    \end{minted}
+    \column{.49\textwidth}
+    \begin{minted}{python}
+    fig, ax = plt.subplots()
+    artist = Line(data, transforms)
+    ax.add_artist(artist)
+    \end{minted}
+    \end{columns}
+\end{frame}
+
+\subsection{Artist}
+\begin{frame}[fragile]{ \vchannel}
+    \begin{minted}{python}
+    cmap =  color.Categorical({'true':'deeppink', 'false':'deepskyblue'})
+    transforms = {'x': {'name': 'v4', 'encoder': lambda x: x},
+                    'y': {'name': 'v2', 'encoder': lambda x: x},
+                    'facecolors': {'name':'v3', 'encoder': cmap}, 
+                    's':{'name': None , 
+                        'encoder': lambda _: itertools.repeat(.02)}}
+    \end{minted}
+    \begin{itemize}
+        \item \mintinline{python}{lambda x: x} is identity \vchannel\
+        \item \mintinline{python}|{'name':None}| map into \vfiber\ without corresponding \dsection\
+        \item \mintinline{python}{color.Categorical} is custom \vchannel
+    \end{itemize}
+\end{frame}
+
+\begin{frame}[fragile]{\vartist} %% rewrite w/ letters in talk
+    \begin{minted}{python}
+        class ArtistClass(matplotlib.artist.Artist):
+            def __init__(self, E, V, *args, **kwargs):
+                # set properties that are specific to the artist
+                # stash the input E and V
+                super().__init__(*args, **kwargs)
+
+            def qhat(self, **args):
+                # set the properties of the graphic
+
+            def draw(self, renderer):
+                # returns tau, indexed on fiber then key 
+                tau = self.E.view(self.axes) 
+                # visual channel encoding applied fiberwise 
+                visual = {p_i: nu_i(tau_i)
+                          for p_i, nu_i, tau)i 
+                          in zip(self.V, tau)}
+                self.qhat(**visual)
+                # pass configurations off to the renderer
+                super().draw(renderer)
+        \end{minted}
+\end{frame}
+
+\begin{frame}[fragile]{\vmarkd}
+\begin{minted}{python}
+    class Point(mcollections.Collection):
+        def assemble(self, x, y, s, facecolors='C0' ):
+            # construct geometries of the circle glyphs in visual coordinates
+            # set attributes of glyphs
+
+    class Line(mcollections.LineCollection):
+        def assemble(self, x, y, color='C0'):
+            # assemble line marks as set of segments 
+    \end{minted}
+\end{frame}
+
+\subsection{Data}
+\begin{frame}[fragile]{Continuity}
+    \begin{minted}{python}
+class PointData: 
+    # Fiberbundle is consistent across all sections
+    FB = FiberBundle({'tables': ['vertex']},  
+        {'v1': float, 'v2': str, 'v3': float})
+    def tau(self, k):
+        return # tau evaluated at one point k
+
+class LineData: 
+    FB = FiberBundle({'tables': ['edge']},
+                {'x' : float, 'y':  float, 'color':mtypes.Color()})
+    def tau(self, k): 
+        return # tau evaluated on interval k
+    \end{minted}
+\end{frame}
+
+\begin{frame}[fragile]{Same Artist, Different \dtotal}
+    \begin{figure}[H]
+        \begin{subfigure}{0.49\textwidth}
+            \includegraphics[width=\textwidth]{figures/code/linec_1.png}
+        \end{subfigure}
+        \begin{subfigure}{0.49\textwidth}
+            \includegraphics[width=\textwidth]{figures/code/lined_1.png}
+        \end{subfigure}
+    \end{figure}
+\begin{minted}{python}
+LineData(FB, edge_table, vertex_table, connect=True)
+LineData(FB, edge_table, vertex_table, num_samples=2, connect=False)
+\end{minted}
+\end{frame}
+
+\begin{frame}{Proposed Work}
+\begin{figure}
+    \includegraphics[width=1\textwidth]{figures/intro/gantt.png}
+\end{figure}
+\end{frame}
+
+\begin{frame}{Acknowledgments}
+    \begin{itemize}
+        \item Professor Michael Grossberg and Dr. Thomas Caswell
+        \item Professor Haralick, Professor Vo, Professor Manovich, Dr. Hanwell
+        \item Matplotlib development team 
+        \item CZI EOSS (grant number 2019-207333) from the Chan Zuckerberg Initiative DAF, an advised fund of Silicon Valley Community Foundation
+    \end{itemize}
+
+\end{frame}
+
+\begin{frame}[allowframebreaks]{References}
+\printbibliography
+\end{frame}
+\appendix
+
+\begin{frame}{Bertin Retinal Variables}
+    \begin{figure}
+        \includegraphics[width=.7\textwidth]{figures/intro/retinal_variables.png}
+        \caption{This tabular form of Bertin's retinal variables is from Understanding Graphics \cite{malamedInformationDisplayTips2010} who reproduced it from Krygier and Wood's \textit{Making Maps: A Visual Guide to Map Design for GIS}\cite{krygierMakingMapsVisual2005}}
+    \end{figure}
+\end{frame}
+
+\begin{frame}{Fiber is all possible values a variable can be \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES}}
+    Given a space of all possible values \ftotal\
+    \begin{equation*}
+        \label{eq:data_types}
+        \begin{tikzcd}[ampersand replacement=\&]
+            \fttype \arrow[r] \arrow[d, "\pi_{\fsection}"'] \& \ftotal \arrow[d, "\pi"] \\
+            \fnames \arrow[r, "\fsection"']                  \& \ftypes       
+        \end{tikzcd}
+    \end{equation*}
+    a fiber component is the restricted space $\ftotal_{\fsection(\fname)}$. 
+    \begin{equation*}
+        \dfiber = \ftotal_{\fsection(\fname)} = \ftotal_{\ftype} 
+    \end{equation*}
+    \begin{description}
+        \item[\ftypes] data types of the variables in the dataset 
+        \item[\ftotal] disjoint union of all values of type $\ftype \in \ftypes$ 
+        \item[\fnames] variable names, $\fname \in \fnames$
+        \item[\fttype] \ftotal\ restricted to the data type of a named variable   
+    \end{description}
+\end{frame}
+
+\begin{frame}{Monoid actions}
+    A monoid \monoid\ is a set with
+    \begin{description}
+        \item[associative binary operator] $\ast:\monoid \times \monoid\rightarrow \monoid$
+        \item[identity element] $e\in \monoid$ such that $e\ast a= a \ast e = a$ for all $a \in \monoid$. 
+    \end{description}
+    \begin{block}{left monoid action}
+    A set \dfiber\ with an action\cite{nlab:action} $\bullet: \monoid\times \dfiber \rightarrow \dfiber$ with the properties:
+        \begin{align*}
+            \textbf{associativity}\;& \text{for all } f,g \in \monoid \text{ and } x\in \dfiber,\, f\bullet(g\bullet x) = (f\ast g) \bullet x\\
+            \textbf{identity}\;& \text{for all } x\in \dfiber, e\in \monoid,\,  e\bullet x = x 
+        \end{align*}
+    \end{block}
+\end{frame}
+
+\begin{frame}{Keeping track of sections with sheafs}
+    Restriction maps of a sheaf describe how local $\iota^*\dsection$ can be glued into larger sections \cite{ghristElementaryAppliedTopology2014,ghristHomologicalAlgebraData2018}
+    \begin{equation*}
+        \label{eq:sheaf}
+        \begin{tikzcd}[ampersand replacement=\&]
+            \iota^*\dtotal \arrow[d, "\pi"'] \arrow[r, "\iota^*", hook]             \& \dtotal \arrow[d, "\pi"']                  \\
+            U \arrow[r, "\iota", hook] \arrow[u, "\iota^{*}\dsection"', bend right] \& \dbase \arrow[u, "\dsection"', bend right]
+        \end{tikzcd}
+    \end{equation*}
+    The inclusion map $\iota: U \rightarrow \dbase$ pulls \dtotal\ over $U$ such that the pulled back $\iota^*\dsection$ only contains records over $U \subset \dbase$.
+\end{frame}
+
+\begin{frame}{Rendering: Define a Pixel}
+    \begin{columns}
+        \column{0.5\textwidth}
+        Given a pixel
+        \begin{equation*}
+        p=\left[y_{top}, y_{bottom}, x_{right}, x_{left}\right]
+        \end{equation*}
+        the inverse map of the bounding box 
+        \begin{equation*}
+        \gbase_{p} ={\gsection_{xy}}^{-1}(p)
+        \end{equation*}
+        is a region $\gbase_p \subset \gbase$ such that 
+        \begin{align}
+            \scriptstyle r_p &= \scriptstyle \iint\limits_{S_p} \rho_r(s)ds^{2}\\
+            \scriptstyle g_p &= \scriptstyle \iint\limits_{S_p} \rho_g(s)ds^{2}\\
+            \scriptstyle  b_p &= \scriptstyle \iint\limits_{S_p} \rho_b(s)ds^{2}
+        \end{align}
+        yields the color of the pixel. 
+        \column{0.5\textwidth}
+        \begin{figure}[H]
+            \includegraphics[width=\textwidth]{figures/math/render.png}
+        \end{figure}
+    \end{columns}
+\end{frame}
+
+\begin{frame}{$\mathcal{A}:\mathcal{E}\rightarrow\mathcal{E}$}
+    The topological artist is a sheaf map
+    \begin{equation*}
+    \vartist: \mathcal{O}(\dtotal) \rightarrow \mathcal{O}(\gtotal)
+    \end{equation*}
+    that carries homomorphism of monoid actions $\varphi: \monoid \rightarrow \monoid^{\prime}$ 
+    \cite{cegarraCohomologyMonoidsOperators2019} 
+    \begin{equation*}
+    \vartist(m\cdot \delement) = \varphi(m)\cdot \vartist(\delement) 
+    \end{equation*}
+\end{frame}
+
+\begin{frame}{Visual Channel Encoders}
+    We define the visual transformers \vchannel\ on components of the data bundle $\dsection_{i}$
+    \begin{equation*}
+        \label{eq:nu_expanded}
+        \{\vchannel_{0}, \ldots, \vchannel_{n}\}: \{\dsection_{0}, \ldots, \dsection_{n}\} \mapsto \{\vsection_{0}, \ldots, \vsection_{n}\}
+    \end{equation*}
+    as the set of equivariant maps with the constraint 
+    \begin{equation*}
+        \vchannel_i(m_{\delement}(\dtotal_i)) = \varphi(m_{\delement})(\vchannel_i(\dtotal_i))
+    \end{equation*} 
+    where $\varphi:\monoid\rightarrow \monoid^{\prime}$ carries a homomorphism of monoid actions. 
+\end{frame}
+
+\begin{frame}{\vfiber\ Components}    
+\begin{table}[H]
+    \renewcommand{\arraystretch}{2}
+    \begin{tabulary}{\textwidth}{|l|L|l|}\hline
+     $\bm{\vchannel_{i}}$                      & $\bm{\vsection_{i}}$                                                            & $\bm{codomain(\vchannel_{i}) \subset \vfiber_{i}}$  \\ \hline
+    position                    & x, y, z, theta, r                                                          & $\mathbb{R}$   \\ \hline
+    size                        & linewidth, markersize                                            & $\mathbb{R}^{+}$   \\ \hline
+    shape                       & markerstyle                                                      & $\{f_{0}, \ldots, f_{n}\}$ \\ \hline
+    color                       & color, facecolor, markerfacecolor, edgecolor  & $\mathbb{R}^{4}$ \\ \hline
+    \multirow{2}{*}{texture}    & hatch                                                            & $\mathbb{N}^{10}$\\\cline{2-3}
+                                & linestyle                                                        & $(\mathbb{R}, \mathbb{R^+}^{n, n\%2=0})$ \\ \hline              
+    \end{tabulary}
+    \label{tab:mpl_visual_variable_fiber}
+\end{table}
+\end{frame}
+\begin{frame}{Monoid Equivariance: Partial Orders}
+    \begin{figure} %break out steps/reveal
+        \includegraphics[width=1\textwidth]{figures/math/monoid_equivariant.png}
+    \end{figure}
+\end{frame}
+
+\begin{frame}{Glyph}
+    The glyph is the graphic generated by $\vmark(\gbase_{\dbasepathpoint})$ where the path connected components $\dbasepath \subset \dbase$ are defined 
+    \begin{equation*}
+    \dbasepath = \{\dbasepathpoint \in \dbase \text{ s. t. } \exists \gamma \text{ s.t. } \gamma(0)=\dbasepoint \text{ and }\gamma(1)=\dbasepathpoint\}
+    \end{equation*}
+    such that the  path $\gamma$ from \dbasepoint\ to \dbasepathpoint\ is a continuous function from the interval [0,1] and $\gbase_{\dbasepathpoint}$ is the region
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement=\&]
+            \gtotal \arrow[r, shift left] \& \gbase_\dbasepathpoint \arrow[rr, "\vindex(\gbasepoint)", shift left] \arrow[l, "\gsection(\gbase_\dbasepathpoint)"] \& \& \dbasepath_{\dbasepoint} \arrow[ll, "\vindex^{-1}(\dbasepath)"]
+            \end{tikzcd}
+        \label{eq:mark}
+    \end{equation*}
+    such that the glyph is differentiable, in keeping with Ziemkiewicz and Kosara's description of a glyph\cite{ziemkiewiczEmbeddingInformationVisualization2009}.
+\end{frame}
+
+\begin{frame}{Artist Equivalance class}
+    When artists share a base space 
+    \begin{equation*}
+        \dbase_2 \hookrightarrow \dbase_1
+    \end{equation*}
+    a composition operator can be defined such that the the artists can be considered to be acting on different components of the same section. 
+\end{frame}
+
+\begin{frame}[fragile]{Complex \vchannel}
+    \begin{minted}{python}
+    class Categorical:
+        def __init__(self, mapping):
+            # check that the conversion is to valid colors
+            assert(mcolors.is_color_like(color) for color in mapping.values())
+            self._mapping = mapping
+
+        def __call__(self, value):
+            # convert value to a color
+            return [mcolors.to_rgba(self._mapping[v]) for v in values]
+    \end{minted}
+    That we can test for action equivariance
+    \begin{minted}{python}
+    def test_nominal(values, encoder):
+        m1 = list(zip(values, encoder(values)))
+        random.shuffle(values)
+        m2 = list(zip(values, encoder(values)))
+        assert sorted(m1) == sorted(m2)
+    \end{minted}
+\end{frame}
+
+\begin{frame}[fragile]{Artist}
+    \begin{minted}{python}
+        class ArtistClass(matplotlib.artist.Artist):
+            def __init__(self, data, transforms, *args, **kwargs):
+                # properties that are specific to the graphic
+                self.data = data 
+                self.transforms = transforms
+                super().__init__(*args, **kwargs)
+
+            def assemble(self, **args):
+                # set the properties of the graphic
+
+            def draw(self, renderer):
+                # returns K, indexed on fiber then key 
+                view = self.data.view(self.axes) 
+                # visual channel encoding applied fiberwise 
+                visual = {p: t['encoder'](view[t['name']])
+                          for p, t in self.transforms.items()}
+                self.assemble(**visual)
+                # pass configurations off to the renderer
+                super().draw(renderer)
+        \end{minted}
+\end{frame}
+
+\begin{frame}[fragile]{Artists: Scatter \& Line}
+\begin{minted}{python}
+class Point(mcollections.Collection):
+    def assemble(self, x, y, s, facecolors='C0' ):
+        # construct geometries of the circle glyphs in visual coordinates
+        self._paths = [mpath.Path.circle(center=(xi,yi), radius=si) 
+                    for (xi, yi, si) in zip(x, y, s)] 
+        # set attributes of glyphs, these are vectorized 
+        # circles and facecolors are lists of the same size
+        self.set_facecolors(facecolors)
+
+class Line(mcollections.LineCollection):
+    def assemble(self, x, y, color='C0'):
+        #assemble line marks as set of segments 
+        segments = [np.vstack((vx, vy)).T for vx, vy in zip(x, y)]
+        self.set_segments(segments)
+        self.set_color(color)
+\end{minted}
+\end{frame}
+
+\begin{frame}[fragile]{View}
+\begin{minted}{python}
+def view(self, axes):
+    table = defaultdict(list)
+    for k in self.keys:
+        table['index'].append(k)
+        for (name, value) in zip(self.FB.fiber.keys(), self.tau(k)[1]):
+            table[name].append(value)
+    return table
+\end{minted}
+\begin{description}
+    \item [\mintinline{python}{VertexSimplex}] (name, value), value is scaler
+    \item [\mintinline{python}{EdgeSimplex}]  (name, value), value is [x0, ..., xn]
+\end{description}
+\end{frame}
+
+\begin{frame}[fragile]{Fiber Bundle}
+    \begin{minted}{python}
+    @dataclass
+    class FiberBundle:
+    """
+    Attributes
+    ----------
+    K: {'tables': []}
+    F: {variable name: type}
+    """
+        K: dict 
+        F: dict
+    \end{minted}
+\end{frame}
+
+\begin{frame}[fragile]{GraphLine Data Model}
+\begin{minted}{python}
+class GraphLine:
+    def __init__(self, FB, edge_table, vertex_table, num_samples=1000,
+                        connect=False):
+        #set args as attributes and generate distance
+        if connect: # test connectivity if edges are continuous
+            assert edge_table.keys() == self.FB.F.keys()
+            assert is_continuous(vertex_table)
+
+    def tau(self, k):
+        # evaluates functions defined in edge table
+        return(k, (self.edges[c][k](self.distances) 
+                        for c in self.FB.F.keys()))
+
+    def view(self, axes):
+        # walk the edge_vertex table to return the edge function
+        table = defaultdict(list)
+        for (i, (start, end)) in sorted(zip(self.ids, self.vertices), 
+                                            key=lambda v:v[1][0]):
+            table['index'].append(i)
+            # same as view for line, returns nested list
+            for (name, value) in zip(self.FB.F.keys(), self.tau(i)[1]):
+                table[name].append(value)
+        return table
+\end{minted}
+\end{frame}
+\end{comment}

### 2022-02-23 slides/main.tex
construction/maps to

@@
 
-\begin{comment}
-
 \section{Construction}
@@
 \end{frame}
+
+\subsection{Measurable Visual Components $M$}
 
@@
             \begin{description}[style=nextline]
-                \item[\textcolor{section}{visual Components}]{$\cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} \coloneqq \big\{\vsectionc: \opensetc\rightarrow \vtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\vsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}$}
-            \end{description}
-        \end{alertblock}
-    }
-\end{frame}
-
-\begin{frame}{Functor $\vchannelc: \dfiberc \rightarrow \vfiberc$}
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement=\&]
-            \dfiberc 
-            \arrow[rr, "\vchannelc_{\dfiberc,\vfiber}"] 
-            \arrow[rrdd, "{\vchannelc_{\dfiberc,{\vfiberc}^{\prime}}"', dotted] 
-            \&  \& 
-            \vfiberc \coloneqq \dfiberc^{\prime}  
-            \arrow[dd, "\vchannelc_{{\dfiberc}^{\prime}, {\vfiberc}^{\prime}}"] \\
-            \&  \& \\
-            \&  \& \vfiberc^{\prime}                           
-        \end{tikzcd}
-    \end{equation*}
-    \begin{alertblock}
-        $\vchannelc$ is a functor s.t. if there exists a map ${\vchannelc}_{\dfiber, \vfiber}: \dfiber \mapsto \vfiber$ and 
-        ${\vchannelc}_{{\dfiber}^{\prime}, {\vfiber}^{\prime}}: {\dfiber}^{\prime} \mapsto {\vfiber}^{\prime}$ then by composition there exists a map ${\vchannelc}_{\dfiberc, {\vfiberc}^{\prime}}: \dfiberc \rightarrow {\vfiberc}^{\prime}$
-    \end{alertblock}
-\end{frame}
-
-\begin{frame}{Equivariant Data to Components: $\vchannel: \sheaf_{\dbasec,\dtotalc} \rightarrow \sheaf_{\dbasec, \vtotalc}$}
+                \item[\textcolor{section}{visual components}]{$\cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} \coloneqq \big\{\vsectionc: \opensetc\rightarrow \vtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\vsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}$}
+            \end{description}
+        \end{alertblock}
+    }
+\end{frame}
+
+\begin{frame}{Data to Visual Transformation: $\vchannelc: \dsectionc \mapsto \vsectionc$}
     \begin{equation*}
         \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-            \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \arrow[rr, "\vchannelc"] \& \& 
+            \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
+            \arrow[rr, "\vchannelc", color=artist, maps to] \& \& 
             \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} \\
              \& \opensetc 
-             \arrow[lu, "{\sheafc_{\dbasec,\dtotalc}}" description, maps to] 
-             \arrow[ru, "{\sheafc_{\dbasec, \vtotalc}}" description, maps to] \& 
-            \end{tikzcd}
-    \end{equation*}
-
-    \begin{alertblock}{Build in Equivariance}
-    implement $\vchannelc$ as a functor such that 
+             \arrow[lu, "{\sheafc_{\dbasec,\dtotalc}}" description, maps to, color=sheaf] 
+             \arrow[ru, "{\sheafc_{\dbasec, \vtotalc}}" description, maps to, color=sheaf] \& 
+            \end{tikzcd}
+    \end{equation*}
+
+    \begin{alertblock}{constraints} 
         \begin{description}
             \item[equivariance] $\vchannelc(\dfuncc_{\dtotalc}(\dsectionc(\dbasepointc))) = \dfuncc_{\vtotalc}(\vchannelc(\dsectionc(\dbasepointc)))$, for all $\dbasepointc \in \dbasec$ 
-            \item[continuity] $\lim\limits_{x\to \dbasepointc}(\vchannelc(\dsectionc(x))) = \vchannelc(\dsectionc(\dbasepointc))$ for all $\dbasepointc \in \dbasec$ \note{for all sections} 
-    \end{description} 
-\end{alertblock}
-\end{frame}
-
-\begin{frame}{Sample $\dfuncc$: Stevens' Scales \cite{stevensTheoryScalesMeasurement1946}}
+            \item[continuity] $\lim\limits_{x\to \dbasepointc}(\vchannelc(\dsectionc(x))) = \vchannelc(\dsectionc(\dbasepointc))$ for all $\dbasepointc \in \dbasec$
+            \item[pointwise] $\vchannelc: \dsectionc(\dbasepointc) \mapsto \vsectionc(\dbasepointc) = \vchannelc: \vindexpullc\dsectionc(\gbasepointc) \mapsto \vindexpullc\vsectionc(\gbasepointc)$ when $\vindexc(\gbasepointc) = \dbasepointc$
+    \end{description} 
+\end{alertblock}
+\end{frame}
+
+\begin{frame}{Functor $\vchannelc: \dsectionc \mapsto \vsectionc$}
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
+            \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
+            \arrow[rr, "\vchannelc_{\dfiberc,\vfiberc}", maps to, color=artist] 
+            \arrow[rrdd, "{\vchannelc_{\dfiberc,{\vfiberc}^{\prime}}}"', dotted, maps to, color=artist] 
+            \&  \& 
+            \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} \coloneqq \cgamma{\opensetc}{\dtotalc^{\prime}\restriction_{\opensetc}} 
+            \arrow[dd, "\vchannelc_{{\dfiberc}^{\prime}, {\vfiberc}^{\prime}}", maps to, color=artist] \\
+            \&  \& \\
+            \&  \& \cgamma{\opensetgc}{\vtotalc^{\prime}}                           
+        \end{tikzcd}
+    \end{equation*}
+\end{frame}
+
+\begin{frame}{$\vchannelc:\dfuncc_{\dtotalc} \mapsto \dfuncc_{\vtotalc}$: Stevens' Scales \cite{stevensTheoryScalesMeasurement1946}}
     \begin{table}[H]
@@
             \dfiberc_a\times \dfiberc_b 
-            \arrow[rr, "\vchannelc"] 
-            \arrow[rd, "p_{\dfiberc}"] \& \& \vfiberc_a\times \vfiberc_b 
-            \arrow[rd, "p_{\vfiberc}"] \& \\    \& \dfiberc_{a} \arrow[rr, "\vchannelc_{a}", dashed] \& \& \vfiberc_a \\
+            \arrow[rr, "\vchannelc", color=artist, maps to] 
+            \arrow[rd, "p_{\dfiberc}", color=fiber] \& \& \vfiberc_a\times \vfiberc_b 
+            \arrow[rd, "p_{\vfiberc}", color=fiber] \& \\    \& 
+            \dfiberc_{a} 
+            \arrow[rr, "\vchannelc_{a}", dashed, maps to, color=artist] \& \& \vfiberc_a \\
             \dfiberc_a\times \dfiberc_c 
-            \arrow[rr, "\vchannelc^{\prime}"] \arrow[ru, "p_{\dfiberc^{\prime}}"'] \& \& {\vfiberc_a}^{\prime}\times {\vfiberc_c}^{\prime} \arrow[ru, "p_{\vfiberc^{\prime}}"'] \&           
+            \arrow[rr, "\vchannelc^{\prime}", color=artist, maps to] 
+            \arrow[ru, "p_{\dfiberc^{\prime}}"', color=fiber] \& \& 
+            {\vfiberc_a}^{\prime}\times {\vfiberc_c}^{\prime} 
+            \arrow[ru, "p_{\vfiberc^{\prime}}"', color=fiber] \&           
             \end{tikzcd}
@@
 
-\begin{frame}{Combining Components}
-    \begin{tikzcd}[ampersand replacement = \&]
+\subsection{Visual Representations of Data}
+
+\begin{frame}{Building a Visual Representation}
+    \begin{equation*}
+    \begin{tikzcd}[ampersand replacement = \&, column sep=small]
         \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
-    \arrow[r, "\vchannelc", color=artist] \& 
+        \arrow[rr, "\vchannelc", color=artist, mapsto] 
+        \& \& 
+        \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} \& 
         \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} 
-    \arrow[r, "\vmarkc_*", color=artist] 
-    \arrow[rr, "\vmarkc", bend left, color=artist] \& 
-    \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} \& 
-    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
-    \arrow[l, "\vindexpushc"', color=functor, dashed]\\ 
-    \& \\ \& 
+        \arrow[rr, "\vmarkc", color=artist] 
+        \& \& 
+        \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}\\
+        \& \& \& \&  \& \\
+        \& \opensetc 
+        \arrow[luu, "{\sheafc_{\dbasec, \dtotalc}}" description, maps to, color=sheaf] 
+        \arrow[ruu, "{\sheafc_{\dbasec, \vtotalc}}" description, maps to, color=sheaf] 
+        \& \& 
         \opensetc 
-    \arrow[luu, "{\sheafc_{\dbasec, \dtotalc}}" description, maps to, color=sheaf] 
         \arrow[uu, "{\sheafc_{\dbasec, \vtotalc}}" description, maps to, color=sheaf]  
-    \arrow[ruu, "{\sheafc_{\dbasec, \gtotalc}}" description, maps to, color=sheaf] 
         \& \& 
@@
         \arrow[uu, "{\sheafc_{\gbasec, \gtotalc}}" description, maps to, color=sheaf] 
-    \arrow[ll, "\vindexc"', dashed, color=functor]
-    \end{tikzcd}
-
-    \begin{alertblock}{Equivariance}
+        \arrow[ll, "\vindexc"', color=functor]
+        \end{tikzcd}
+    \end{equation*}
+    \begin{alertblock}{}
+        \begin{description}[style=newline]
         \item[equivariance] $\delta(\vmarkc(\vchannelc\dfuncc_{\dtotalc}(\dsectionc))) = \dfuncc_{M}(\delta(\dsectionc))$ 
@@
         such that there must exist a deformation retract $\vindexc:\gbasec \rightarrow \dbasec$ for all points $\gbasepointc \in \gbase$. 
-    \end{alertblock}
-\end{frame}
-
+        \end{description}
+    \end{alertblock}
+\end{frame}
+
+\begin{frame}{Implementation Choices: $\vartistc_{\opensetc} = \vartistc_{\opensetgc}$}
+    \begin{equation*}
+    \begin{tikzcd}[ampersand replacement = \&, column sep=small]
+        \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}  
+        \arrow[r, "\vchannelc", maps to, color=artist] 
+        \arrow[rrrr, "\vartistc_{\opensetgc}", Rightarrow, bend left, color=artist] \& 
+        \cgamma{\opensetgc}{\vindexpullc\vtotalc\restriction_{\opensetgc}} \& 
+        \cgamma{\opensetgc}{\vindexpullc\vtotalc\restriction_{\opensetgc}}  
+        \arrow[rr, "\vmarkc^*", no head, Rightarrow, color=artist] \&  \& 
+        \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}  
+        \arrow[dd, "\vindexpushc", color=functor] \\
+        \& \& \& \& \\
+        \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}  
+        \arrow[uu, "\vindexpullc", color=functor] 
+        \arrow[r, "\vchannelc", maps to, color=artist] 
+        \arrow[rrrr, "\vartistc_{\opensetc}", Rightarrow, bend right, color=artist] 
+        \& 
+        \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}}  
+        \arrow[uu, "\vindexpullc"', color=functor] \& 
+        \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}}
+        \arrow[rr, "\vmarkc_*", Rightarrow, color=artist] 
+        \arrow[uu, "\vindexpullc", color=functor] 
+        \arrow[rruu, "\vmarkc", Rightarrow, color=artist] 
+        \&  \& 
+        \cgamma{\opensetgc}{\vindexpushc\gtotalc\restriction_{\opensetc}}
+    \end{tikzcd}
+    \end{equation*}
+\end{frame}
+
+\begin{comment}
 \begin{frame}{Components to Graphic}

### 2022-02-28 slides/main.tex
decomposing nu

@@
                 \item[Locally Trivial] {for every point $\dbasepointc \in \dbasec$, there exists an open neighborhood $\dbasepointc \in \opensetc \subseteq \dbasec$ s.t.
-there is a homeomorphism $\pi^{-1}(\opensetc)\xrightarrow{\varphi} \opensetc \times \dfiberc$}
+there is a homeomorphism $\pi^{-1}(\opensetc)\xrightarrow{\equivc} \opensetc \times \dfiberc$}
                 \item[(Globally) Trivial]{$\dtotalc = \dbasec \times \dfiberc$}
@@
             \begin{equation*}
-                \vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexpre(\dbasepointc)} = \gsectionc(\gbasepointc) \forall \gbasepointc \in \vindexpre(\dbasepointc)
+                \vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexprec(\dbasepointc)} = \gsectionc(\gbasepointc) \forall \gbasepointc \in \vindexprec(\dbasepointc)
             \end{equation*}
@@
 \subsection{Equivariance}
+
 \begin{frame}{Transform Data}
@@
     \end{equation*}
+    \begin{alertblock}{Visual Measurement $\delta$}
+        \begin{equation}
+            \delta: \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \rightarrow M
+        \end{equation}
+        $M$ is a measurable component of the rendered visual element, e.g. color, position, shape, texture
+    \end{alertblock}
 \end{frame}
@@
 
-\begin{frame}{Data to Visual Transformation: $\vchannelc: \dsectionc \mapsto \vsectionc$}
-    \begin{equation*}
+\begin{frame}{Data to Visual Transformation: $\vchannelc: \dfiberc_{\dbasepointc} \mapsto \vfiberc_{\dbasepointc}$}
+    \begin{columns}
+    \column{.5\textwidth}
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement=\&]
+        \dtotalc 
+        \arrow[rr, "\vchannelc_{\dbasec}", color=artist] 
+        \arrow[rd, "\pi" description, dotted] 
+        \arrow[ddd, "\dfuncc_{\dtotalc}"', color=action] 
+        \& \& 
+        \vtotalc 
+        \arrow[ld, "\pi" description, dotted] 
+        \arrow[ddd, "\dfuncc_{\vtotalc}", color=action] \\
+        \& \dbasec \&  \\
+        \& \dbasec^{\prime} 
+        \arrow[u, "\dfunchc" description, color=action] \& \\
+        \dfuncpullc\dtotalc 
+        \arrow[ru, "\pi" description, dotted] 
+        \arrow[rr, "\vchannel_{\dbasec}"', color=artist] 
+        \& \& 
+        \dfuncpullc\dtotalc 
+        \arrow[lu, "\pi" description, dotted]
+        \end{tikzcd}
+    \end{equation*}
+    \column{.5\textwidth} 
+    \begin{equation}
     \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-            \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
-            \arrow[rr, "\vchannelc", color=artist, maps to] \& \& 
-            \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} \\
-             \& \opensetc 
-             \arrow[lu, "{\sheafc_{\dbasec,\dtotalc}}" description, maps to, color=sheaf] 
-             \arrow[ru, "{\sheafc_{\dbasec, \vtotalc}}" description, maps to, color=sheaf] \& 
-            \end{tikzcd}
-    \end{equation*}
-
-    \begin{alertblock}{constraints} 
+        \dfiberc_{\dbasepointc} 
+        \arrow[rr, "\vchannelc_{\dtotalc} ", maps to, color=artist] 
+        \arrow[rrdd, "{\vchannelc_{\dtotalc,\vtotalc}}" description, dashed, maps to, color=artist] 
+        \&  \& 
+        \vfiberc_{\dbasepointc}\coloneqq {\dfiberc_{\dbasepointc}}^{\vtotalc} \arrow[dd, "\vchannelc_{\vtotalc}", maps to, color=artist] \\
+        \& \& \\
+        \&  \& {\vfiberc_{\dbasepointc}}^{\vtotalc}
+        \end{tikzcd}
+    \end{equation}
+    \end{columns}
+    \begin{alertblock}{equivariance} 
+       $\vchannelc: \dfuncc_{\dtotalc} \rightarrow \dfuncc_{\vtotalc}$ is a functor such that 
        \begin{description}
-            \item[equivariance] $\vchannelc(\dfuncc_{\dtotalc}(\dsectionc(\dbasepointc))) = \dfuncc_{\vtotalc}(\vchannelc(\dsectionc(\dbasepointc)))$, for all $\dbasepointc \in \dbasec$ 
+           \item[equivariance] {$\vchannelc(\dfuncc(\dsectionc(\dfunchc_{\dtotalc}^{-1}(\dbasepointc)))) = \dfuncc_{\vtotalc}(\vchannelc(\dsectionc(\dfunchc_{\vtotalc}^{-1}(\dbasepointc))))$\\$\dfunchc_{\dtotalc}(\dbasepointc^{\prime}) = \dfunchc_{\vtotalc}(\dbasepointc^{\prime})$}
            \item[continuity] $\lim\limits_{x\to \dbasepointc}(\vchannelc(\dsectionc(x))) = \vchannelc(\dsectionc(\dbasepointc))$ for all $\dbasepointc \in \dbasec$ 
-            \item[pointwise] $\vchannelc: \dsectionc(\dbasepointc) \mapsto \vsectionc(\dbasepointc) = \vchannelc: \vindexpullc\dsectionc(\gbasepointc) \mapsto \vindexpullc\vsectionc(\gbasepointc)$ when $\vindexc(\gbasepointc) = \dbasepointc$
-    \end{description} 
-\end{alertblock}
-\end{frame}
-
-\begin{frame}{Functor $\vchannelc: \dsectionc \mapsto \vsectionc$}
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-            \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
-            \arrow[rr, "\vchannelc_{\dfiberc,\vfiberc}", maps to, color=artist] 
-            \arrow[rrdd, "{\vchannelc_{\dfiberc,{\vfiberc}^{\prime}}}"', dotted, maps to, color=artist] 
-            \&  \& 
-            \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} \coloneqq \cgamma{\opensetc}{\dtotalc^{\prime}\restriction_{\opensetc}} 
-            \arrow[dd, "\vchannelc_{{\dfiberc}^{\prime}, {\vfiberc}^{\prime}}", maps to, color=artist] \\
-            \&  \& \\
-            \&  \& \cgamma{\opensetgc}{\vtotalc^{\prime}}                           
-        \end{tikzcd}
-    \end{equation*}
-\end{frame}
-
-\begin{frame}{$\vchannelc:\dfuncc_{\dtotalc} \mapsto \dfuncc_{\vtotalc}$: Stevens' Scales \cite{stevensTheoryScalesMeasurement1946}}
+       \end{description}
+\end{alertblock}
+\end{frame}
+
+\begin{frame}{$\vchannelc:\dfuncc_{\dtotalc} \rightarrow \dfuncc_{\vtotalc}$: Stevens' Scales \cite{stevensTheoryScalesMeasurement1946}}
     \begin{table}[H]

### 2022-03-02 slides/main.tex
code based slide hiding + reworked dfunc unpacking

@@
 \setbeamertemplate{footline}[frame number]
+\setbeamersize{description width=-\labelsep}
+
 %\usepackage{amsmath,amsfonts,amssymb,pxfonts,eulervm,xspace}
@@
 \renewcommand{\restriction}{\mathord{\upharpoonright}} %restriction w/p space
-\usepackage{relsize}%mathlarger
+\usepackage{relsize} %mathlarger
 \usepackage{bm}
@@
 \usepackage{tabularx}
-\usepackage{enumitem}
+%\usepackage{enumitem} $ won't work with beamer
 \usepackage{booktabs}
@@
 \title{Topological Equivariant Artist Model}
-\begin{comment}
-\begin{frame}
+
+\begin{frame}<presentation:1|handout:1>
 	\titlepage
@@
 \section{Introduction}
-\begin{frame}{What are we doing?}
+\begin{frame}<presentation:1|handout:0>{What are we doing?}
     \begin{itemize}
@@
 %% maybe pull out the middle for now? move the 3 stages down to construction? 
-\begin{frame}{What do visualization libraries do?}
+\begin{frame}<presentation:1|handout:0>{What do visualization libraries do?}
     \begin{figure}
@@
 
-\begin{frame}{Case Study: Matplotlib}
+\begin{frame}<presentation:1|handout:0>{Case Study: Matplotlib}
     \begin{figure}
@@
 
-\begin{frame}{Visualizations Preserve Structure}
+\begin{frame}<presentation:1|handout:0>{Visualizations Preserve Structure}
     \begin{description}
@@
 
-\begin{frame}{Continuity}
+\begin{frame}<presentation:1|handout:0>{Continuity}
     \begin{figure}
@@
 
-\begin{frame}{Expressed in container}    
+\begin{frame}<presentation:1|handout:0>{Expressed in container} 
     \begin{figure}
@@
 
-\begin{frame}{Why is container type not enough?}
+\begin{frame}<presentation:1|handout:0>{Why is container type not enough?}
     \includegraphics[width=1\linewidth]{../paper/figures/whycontinuity.png}
@@
 
-\begin{frame}{Equivariance}
+\begin{frame}<presentation:1|handout:0>{Equivariance}
     \begin{description}
@@
 %flip columns
-\begin{frame}{Frameworks for Expressing Visual Equivariance}
+\begin{frame}<presentation:1|handout:0>{Frameworks for Expressing Visual Equivariance}
     \begin{description}
@@
 
-\begin{frame}{Domain specific libraries know their structure\cite{HeerSoftware2006}}
+\begin{frame}<presentation:1|handout:0>
+    \frametitle{Domain specific libraries know their structure\cite{HeerSoftware2006}}
     \begin{table}
@@
             \cite{hanrahanVizQL2006,MackinlayShowme2007}&&\\        
-
         \end{tabular}
@@
 
-\begin{frame}{General purpose libraries generally can't\cite{toryRethinkingVisualizationHighlevel2004}}
+\begin{frame}<presentation:1|handout:0>
+    \frametitle{General purpose libraries generally can't\cite{toryRethinkingVisualizationHighlevel2004}}
     \begin{figure}
@@
     \end{figure}
+
     \begin{enumerate}
-        \item Matplotlib\cite{hunterMatplotlib2DGraphics2007} $\rightarrow$Seaborn\cite{waskom2020seaborn}, xarray \cite{hoyer2017xarray} 
+        \item Matplotlib\cite{hunterMatplotlib2DGraphics2007} $\rightarrow$ Seaborn\cite{waskom2020seaborn}, xarray \cite{hoyer2017xarray} 
         \item D3 \cite{bostockDataDrivenDocuments2011}
-        \item VTK \cite{hanwellVisualizationToolkitVTK2015,geveciVTK2012}(MayaVi\cite{RamachandranMayaVI2011})$\rightarrow$ Titan\cite{brianwylieUnifiedToolkitInformation2009}, ParaView\cite{ahrens2005paraview}
+        \item VTK \cite{hanwellVisualizationToolkitVTK2015,geveciVTK2012},MayaVi\cite{RamachandranMayaVI2011}$\rightarrow$ Titan\cite{brianwylieUnifiedToolkitInformation2009}, ParaView\cite{ahrens2005paraview}
     \end{enumerate}
 \end{frame}
-\end{comment}
 
 \section{Algebraic Topology & Category Theory}
-
-\begin{frame}{Mathematical Data Abstraction}
+\begin{frame}<presentation:1|handout:0>{Mathematical Data Abstraction}
     \begin{description}
@@
 \subsection{Fiber Bundles}
-\begin{frame}
+\begin{frame}<presentation:1|handout:0>
     \frametitle<1|handout:1>{Fiber Bundle}
@@
      \only<1-5|handout:1>{
-        \begin{description}[style=nextline]
+        \begin{description}
             \item<1-5|handout:1>[\textcolor{total}{Total Space}]{$(\dtotalc, \mathcal{T}_{\dtotalc})$}
@@
     \only<6|handout:2>{
-        \begin{description}[style=nextline]
+        \begin{description}
             \item[\textcolor{total}{Data \dtotalc}]{continuity + fields}
@@
         \only<7|handout:4>{
-            \begin{description}[style=nextline]
+            \begin{description}
                 \item[\textcolor{total}{Graphic \gtotalc}]{continuity + renderer fields}
@@
         \begin{alertblock}{}
-            \begin{description}[style=nextline]
+            \begin{description}
                 \item[\textcolor{section}{Sections}]{$\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}$}
@@
     \begin{alertblock}{}
-        \begin{description}[style=nextline]
+        \begin{description}
             \item[\textcolor{section}{Data} \cite{butlerVisualizationModelBased1989,butlerVectorBundleClassesForm1992}]{$
@@
         \begin{alertblock}{}
-            \begin{description}[style=nextline]
+            \begin{description}
                 \item[\textcolor{section}{Graphic}]{$\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \coloneqq \big\{\gsectionc: \opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} \; \bigm{\vert} \pi(\gsectionc(\gbasepointc)) = \gbasepointc\;for\, all\; \gbasepointc \in \opensetgc \big\}$}
@@
 \subsection{Category Theory Language}
-\begin{frame}[fragile]{}
+\begin{frame}<presentation:1|handout:0>[fragile]{}
     \begin{minted}{python}
@@
 
-\begin{frame}{Category $\textcolor{source}{\mathcal{C}}$}
+\begin{frame}<presentation:1|handout:0>{Category $\textcolor{source}{\mathcal{C}}$}
 \begin{columns}
@@
 
-\begin{frame}{Opposite Category $\textcolor{source}{\mathcal{C}^{op}}$}
+\begin{frame}<presentation:1|handout:0>{Opposite Category $\textcolor{source}{\mathcal{C}^{op}}$}
     \begin{columns}
@@
 
-\begin{frame}{Functor $\textcolor{functor}{\bm{F}}: \textcolor{source}{\mathcal{C}} \rightarrow \textcolor{target}{\mathcal{D}}$}
+\begin{frame}<presentation:1|handout:0>
+    \frametitle{Functor $\textcolor{functor}{\bm{F}}: \textcolor{source}{\mathcal{C}} \rightarrow \textcolor{target}{\mathcal{D}}$}
 \begin{columns}
@@
 \subsubsection{Natural Transforms}
-\begin{frame}{Natural Transformation $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\Rightarrow} \textcolor{functor}{G}$}
+\begin{frame}<presentation:1|handout:0>
+    \frametitle{Natural Transformation $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\Rightarrow} \textcolor{functor}{G}$}
     \onslide<|handout:1>{
@@
 
-\begin{frame}{Natural Transformation $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\Rightarrow} \textcolor{functor}{G}$}
+\begin{frame}<presentation:1|handout:0>
+    \frametitle{Natural Transformation $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\Rightarrow} \textcolor{functor}{G}$}
     \begin{equation*}
@@
 
-\begin{frame}{Presheaf: $\sheafc: \mathcal{C}^{op} \textcolor{sheaf}{\rightarrow} \setc $}
+\begin{frame}<presentation:1|handout:0>
+    \frametitle{Presheaf: $\sheafc: \mathcal{C}^{op} \textcolor{sheaf}{\rightarrow} \setc $}
     \begin{columns}
@@
 
-\begin{frame}{Sheaves on Bundles}
+\begin{frame}<presentation:1|handout:0>{Sheaves on Bundles}
     A sheaf is a presheaf that satisfies the following two axioms\cite{bakerMathsSheaf}\\
@@
 \subsubsection{Pullback and Pushforward}
-\begin{frame}{}
+\begin{frame}<presentation:1|handout:1->{}
     \frametitle<3|handout:3>{Function: $\vindexc: \gbasec \rightarrow \dbasec$}
@@
             \\
-            \onslide*<0|handout:2->{\opensetgc}
+            \onslide*<0|handout:2,4->{\opensetgc}
+            \onslide*<0|handout:3>{\underbrace{\opensetc\times[0,1]^{m}}_{\opensetgc}}
             \onslide*<0|handout:3->{\arrow[rr, "\vindex", color=functor]}
@@
 
-\begin{frame}{Example: Scatter Plot}
+\begin{frame}<presentation:1|handout:0>{Example: Scatter Plot}
     \begin{figure}
@@
 
-\begin{frame}{Example: Line Plot}
+\begin{frame}<presentation:1|handout:0>{Example: Line Plot}
     \begin{figure}
@@
 
-\begin{frame}{Example: Image}
+\begin{frame}<presentation:1|handout:0>{Example: Image}
     \begin{figure}
@@
 
-\begin{frame}{Example: Heatmap}
+\begin{frame}<presentation:1|handout:0>{Example: Heatmap}
     \begin{figure}
@@
 
-\begin{frame}{Example: Sphere}
+\begin{frame}<presentation:1|handout:0>{Example: Sphere}
     \begin{figure}
@@
 %add transition slide->possibly call back
+
 \section{Artist}
 
-\begin{frame}
+\begin{frame}<presentation:1|handout:0>
     \frametitle<0|handout:1>{Artist: $\vartistc: \sheafc_{\dbasec, \dtotalc} \rightarrow \sheafc_{\gbasec, \gtotalc}$}
@@
 
-\begin{frame}{Artist as Profunctor}
-    \begin{tikzcd}[ampersand replacement=\&]
+\subsection{Equivariance}
+
+\begin{frame}<presentation:1|handout:1>{Transform Data}
+    \begin{block}{Fiber Bundle Category}
+        \begin{description}[style=newline]
+            \item[object] $\dfiberc \hookrightarrow \dtotalc \xrightarrow{\pi} \dbasec$
+            \item[morphisms] $\dfuncc: \dtotalc \rightarrow \dtotalc^{\prime}$
+        \end{description}
+    \end{block}
+    \begin{block}{Fiber Category}
+        The fiber $\dfiberc$ is a  monoidal category (single object w/ bicartesian product operator on category) of an arbitrary type $\mathcal{C}$.
+The morphisms on the fiber are $Hom(\dfiberc, \dfiberc)$
+    \end{block}
+
+    \begin{block}{Data Transformations}
+    $\dfuncc:\dtotalc\restriction_{\dbasepointc} \rightarrow \dtotalc^{\prime}\restriction_{\dbasepointc^{\prime}} \in Hom(\dfiberc\restriction_{\dbasepointc},\dfiberc^{\prime}\restriction_{\dbasepointc^{\prime}})$
+    \end{block} 
+\end{frame}
+
+\begin{frame}<presentation:1|handout:1>{$\dfuncc = (\dfunchc, \dfunctc)$}
+    \begin{tikzcd}[ampersand replacement = \&]
         \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
-        \arrow[rr, "\vartistc", Rightarrow, color=artist] 
-        \&  \& \
-        \cgamma{\opensetgc}{\dtotalc\restriction_{\opensetgc}}\\
+        \arrow[rr, "\dfuncpullc", color=action] 
+        \&  \& 
+        \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}} 
+        \arrow[rr, "\dfunctc", color=action] 
+        \&  \& 
+        \cgamma{\opensetc^{\prime}}{\dtotalc^{\prime}\restriction_{\opensetc^{\prime}}} \\
+        \&  \&  
         \&  \& \\
@@
         \&  \& 
-        \opensetgc 
-        \arrow[ll, "\vindexc"', color=functor] 
-        \arrow[uu, "{\sheafc_{\gbasec, \gtotalc}}"', maps to, color=sheaf]
-        \end{tikzcd}
-\end{frame}
-
-\subsection{Equivariance}
-
-\begin{frame}{Transform Data}
-    \begin{block}{Fiber Bundle Category}
-        \begin{description}[style=newline]
-            \item[object] $\dfiberc \hookrightarrow \dtotalc \xrightarrow{\pi} \dbasec$
-            \item[morphisms] $\dfuncc: \dtotalc \rightarrow \dtotalc$ 
-        \end{description}
-    \end{block}
-    \begin{block}{Fiber Category}
-        The fiber $\dfiberc$ is the sole object of an arbitrary category $\mathcal{C}$.
-        The morphisms on the fiber are $Hom(\dfiberc, \dfiberc)$
-    \end{block}
-
-    \begin{block}{Data Transformations}
-    $\dfuncc:\dtotalc\restriction_{\dbasepointc} \rightarrow \dtotalc\restriction_{\dbasepointc^{\prime}} \in Hom(\dfiberc\restriction_{\dbasepointc},\dfiberc\restriction_{\dbasepointc^{\prime}})$ 
-    \end{block} 
-\end{frame}
-
-\begin{frame}{$\dfuncc = (\dfunchc, \dfunctc)$}
-    \begin{tikzcd}[ampersand replacement = \&]
-        \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
-        \arrow[rr, "\dfuncpullc", color=action] 
-        \&  \& 
-        \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}} 
-        \arrow[rr, "\dfunctc", color=action] 
-        \&  \& 
-        \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}} \\
-        \&  \&  
-        \&  \& \\
-        \opensetc 
-        \arrow[uu, "{\sheafc_{\dbasec, \dtotalc}}", maps to, color=sheaf]                 
-        \&  \& 
         \opensetc^{\prime} 
         \arrow[ll, "\dfunchc", color=action] 
-        \arrow[uu, "{\dfuncpullc\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf] 
-        \arrow[rruu, "{\dfuncpullc\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf] 
+        \arrow[uu, "{\sheafc_{\dbasec^{\prime}, \dfuncpullc\dtotalc}}"', maps to, color=sheaf] 
+        \arrow[rruu, "{\sheafc_{\dbasec^{\prime}, \dtotalc^{\prime}}}"', maps to, color=sheaf] 
         \&  \& 
@@
     \begin{itemize}
-        \item Base $\dbasec$ Type stays same: $\dfunchc: \dbasec \rightarrow \dbasec$
-        \item Fiber $\dfiberc$ Type stays the same: $\dfunctc: \dfuncpullc \dtotalc \rightarrow \dfuncpullc\dtotalc$ s.t. $\pi(\dfuncpullc\dtotalc) = \pi(\dfunctc(\dfuncpullc\dtotalc))$ 
-        \item Section Transform: $\dfuncc: \dsectionc^{a}\restriction_{\opensetc} \mapsto \dsectionc^{b}\restriction_{\opensetc^{\prime}}$
+        \item Base Transformation: $\dfunchc: \opensetc^{\prime} \rightarrow \opensetc$
+        \item Fiber Transformation: $\dfunctc: \dfuncpullc \dtotalc \rightarrow \dtotalc^{\prime}$ s.t. $\pi(\dfuncpullc\dtotalc) = \pi(\dtotalc^{\prime})$ 
+        \item Section Transform: $\dfuncc: \dsectionc\restriction_{\opensetc} \mapsto \dsectionc^{\prime}\restriction_{\opensetc^{\prime}}$
     \end{itemize}
@@
 
-\begin{frame}{Equivariant Artist}
+\begin{frame}<presentation:1|handout:1>{Equivariant Artist}
     \begin{equation*}
@@
 
-\begin{frame}{Equivariant $\dfunchc$}
+\begin{frame}<presentation:1|handout:1>{Equivariant $\dfunchc$}
     \note{Maybe move to implementation section}
@@
 
-\begin{frame}{Checking if \vartistc\ is equivariant}
+\begin{frame}<presentation:1|handout:1>{Checking if \vartistc\ is equivariant}
     \begin{equation*}
@@
 
-\begin{frame}{Multiple Bundles: $\dfuncc = (\dfunchc, \prod\limits_{i=0}^{n}\dfunctc_i)$} 
+\begin{frame}<presentation:1|handout:0>{Multiple Bundles: $\dfuncc = (\dfunchc, \prod\limits_{i=0}^{n}\dfunctc_i)$}
     \begin{equation*}
@@
 
-\begin{frame}{Complex data}
+\begin{frame}<presentation:1|handout:0>{Complex data}
     \begin{description}[style=newline]
@@
 
-\begin{frame}{"Valid" viz?}
+\begin{frame}<presentation:1|handout:1>{"Valid" viz?}
     \begin{equation*}
@@
 \section{Construction}
-\begin{frame}{What's in the $\vartist$ black box?}
+\begin{frame}<presentation:1|handout:0>{What's in the $\vartist$ black box?}
     \note{Add box around figure & lable it artist}
@@
 
-\begin{frame}{Measurable Visual Components}
+\begin{frame}<presentation:1|handout:0>{Measurable Visual Components}
     \begin{columns}
@@
 
-\begin{frame}{Data to Visual Transformation: $\vchannelc: \dfiberc_{\dbasepointc} \mapsto \vfiberc_{\dbasepointc}$}
+\begin{frame}<presentation:1|handout:0>
+    \frametitle{Data to Visual Transformation: $\vchannelc: \dfiberc_{\dbasepointc} \mapsto \vfiberc_{\dbasepointc}$}
     \begin{columns}
@@
     \end{columns}
+    % fix corner to be V 
+    % fix \phi to be \phi 
     \begin{alertblock}{equivariance} 
@@
 
-\begin{frame}{$\vchannelc:\dfuncc_{\dtotalc} \rightarrow \dfuncc_{\vtotalc}$: Stevens' Scales \cite{stevensTheoryScalesMeasurement1946}}
+\begin{frame}<presentation:1|handout:0>\frametitle{$\vchannelc:\dfuncc_{\dtotalc} \rightarrow \dfuncc_{\vtotalc}$: Stevens' Scales \cite{stevensTheoryScalesMeasurement1946}}
     \begin{table}[H]
@@
 
-\begin{frame}{Shared Components: $\vchannelc = \prod\limits_{i=0}^{n}\vchannelc_i$}
+\begin{frame}<presentation:1|handout:0>
+    \frametitle{Shared Components: $\vchannelc = \prod\limits_{i=0}^{n}\vchannelc_i$}
     \begin{equation}
@@
 
-\subsection{Visual Representations of Data}
-
-\begin{frame}{Building a Visual Representation}
-    \begin{equation*}
-    \begin{tikzcd}[ampersand replacement = \&, column sep=small]
-        \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
-        \arrow[rr, "\vchannelc", color=artist, mapsto] 
-        \& \& 
-        \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} \& 
+\subsection{Combining Components into Visualizations}
+
+\begin{frame}<presentation:1|handout:0>{Assembly Function $\vmarkc$}
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement = \&]
             \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} 
-        \arrow[rr, "\vmarkc", color=artist] 
-        \& \& 
-        \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}\\
-        \& \& \& \&  \& \\
-        \& \opensetc 
-        \arrow[luu, "{\sheafc_{\dbasec, \dtotalc}}" description, maps to, color=sheaf] 
-        \arrow[ruu, "{\sheafc_{\dbasec, \vtotalc}}" description, maps to, color=sheaf] 
-        \& \& 
+            \arrow[rrrr, "\vmarkc", Rightarrow, color=artist] 
+            \arrow[ddd, "\dfuncc_{\vtotalc}"', color=action] 
+            \& \& \& \& 
+            \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+            \arrow[ddd, "\dfuncc_{\gtotalc}", color=action]  \\
+            \& 
             \opensetc 
-        \arrow[uu, "{\sheafc_{\dbasec, \vtotalc}}" description, maps to, color=sheaf]  
+            \arrow[lu, maps to, color=sheaf] 
              \&  \& 
             \opensetgc 
-        \arrow[uu, "{\sheafc_{\gbasec, \gtotalc}}" description, maps to, color=sheaf] 
             \arrow[ll, "\vindexc"', color=functor] 
-        \end{tikzcd}
-    \end{equation*}
-    \begin{alertblock}{}
-        \begin{description}[style=newline]
-        \item[equivariance] $\delta(\vmarkc(\vchannelc\dfuncc_{\dtotalc}(\dsectionc))) = \dfuncc_{M}(\delta(\dsectionc))$ 
-        \item[continuity] the assembly function $\vmarkc$ is defined as  
-        \begin{equation}
-        \vmarkc_*(\vsection(\dbasepointc))(\gbasepointc) = \vmarkc(\vsectionc(\vindex(\gbasepointc)))   
-        \end{equation}
-        such that there must exist a deformation retract $\vindexc:\gbasec \rightarrow \dbasec$ for all points $\gbasepointc \in \gbase$. 
-        \end{description}
-    \end{alertblock}
-\end{frame}
-
-\begin{frame}{Implementation Choices: $\vartistc_{\opensetc} = \vartistc_{\opensetgc}$}
+            \arrow[ru, maps to, color=sheaf]  \& \\
+            \& 
+            \opensetc^{\prime} 
+            \arrow[u, "\dfunchc_{\dtotalc}", color=action] 
+            \arrow[ld, maps to, color=sheaf] 
+            \&  \& 
+            \opensetgc^{\prime} 
+            \arrow[ll, "\vindexc"', color=functor] 
+            \arrow[u, "\dfunchc_{\gtotalc}"', color=action] 
+            \arrow[rd, maps to, color=functor] \& \\
+            \cgamma{\opensetc^{\prime}}{\dfuncpullc_{\vtotalc}\vtotalc\restriction_{\opensetc^{\prime}}}
+            \arrow[rrrr, "\vmarkc", Rightarrow, color=artist] 
+            \& \&  \& \& 
+            \cgamma{\opensetgc^{\prime}}{\dfuncpullc_{\gtotalc}\gtotalc\restriction_{\opensetgc^{\prime}}}
+            \end{tikzcd}
+        \end{equation*}
+    \begin{alertblock}{equivariance}
+       \begin{equation*}
+        \vmarkc(\dfuncc_{\vtotalc}(\vsectionc(\dfunchc_{\vtotalc}^{-1}(\dbasepointc^{\prime})))) = \dfuncc_{\gtotalc}(\vmarkc((\vsectionc(\dfunchc_{\gtotalc}^{-1}(\vindexprec(\dbasepointc^{\prime}))))))
+       \end{equation*}
+    \end{alertblock}
+\end{frame}
+
+\begin{frame}<presentation:1|handout:0>{Implementation Choices: $\vartistc_{\opensetc} = \vartistc_{\opensetgc}$}
     \begin{equation*}
@@
 
-\begin{comment}
-\begin{frame}{Components to Graphic}
+\begin{frame}<presentation:1|handout:0>{Components to Graphic}
     \begin{figure}
@@
 \end{frame}
-\end{comment}
 \end{document}

### 2022-03-02 slides/main.tex
artist equiv

@@
             \onslide*<0|handout:2,4->{\opensetgc}
-            \onslide*<0|handout:3>{\underbrace{\opensetc\times[0,1]^{m}}_{\opensetgc}}
+            \onslide*<0|handout:3>{\overbrace{\opensetc\times[0,1]^{m}}^{\opensetgc}}
             \onslide*<0|handout:3->{\arrow[rr, "\vindex", color=functor]}
@@
 \subsection{Equivariance}
-
 \begin{frame}<presentation:1|handout:1>{Transform Data}
@@
         \begin{tikzcd}[ampersand replacement=\&]
-            \sheafc_{\dbasec, \dtotalc} 
-            \arrow[r, "\vartistc", Rightarrow, color=artist] 
-            \arrow[d, "\dfuncc_{\dtotalc}"', color=action] \& 
-            \sheafc_{\gbasec, \gtotalc} 
-            \arrow[d, "\dfuncc_{\gtotalc}", color=action, dashed] \\
-            \sheafc_{\dbasec^{\prime}, \dfuncpullc\dtotalc} 
-            \arrow[r, "\vartistc", Rightarrow, color=artist]\& 
-            \sheafc_{\gbasec^{\prime},\dfuncpullc\gtotalc}
-        \end{tikzcd} 
-        \end{equation*}
-
-    \begin{alertblock}{Artist Constraints}
+        \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
+        \arrow[rrr, "\vartistc", Rightarrow, color=artist] 
+        \arrow[d, "\dfuncpullc_{\dtotalc}"', color=action] 
+        \& \& \& 
+        \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+        \arrow[d, "\dfuncpullc_{\gtotalc}", dotted] \\
+        \cgamma{\opensetc^{\prime}}{\dfuncpullc_{\dtotalc}\dtotalc\restriction_{\opensetc^{\prime}}} 
+        \arrow[dd, "\dfunctc_{\dtotalc}"', color=action] \& 
+        \opensetc 
+        \arrow[lu, dashed, maps to, color=sheaf] \& 
+        \opensetgc 
+        \arrow[l, "\vindexc"', color=functor] 
+        \arrow[ru, dashed, maps to, color=sheaf] \& 
+        \cgamma{\opensetgc^{\prime}}{\dfuncpullc_{\gtotalc}\gtotalc\restriction_{\opensetgc^{\prime}}} 
+        \arrow[dd, "\dfunctc_{\gtotalc}", dotted, color=action] \\
+        \& 
+        \opensetc^{\prime} 
+        \arrow[u, "\dfunchc_{\dtotalc}", color=action] 
+        \arrow[ld, dashed, maps to, color=sheaf] 
+        \arrow[lu, dashed, maps to, color=sheaf] \& 
+        \opensetgc^{\prime} 
+        \arrow[l, "\vindexc"', color=functor] 
+        \arrow[u, "\dfunchc_{\gtotalc}"', dotted, color=action] 
+        \arrow[rd, dashed, maps to, color=sheaf] 
+        \arrow[ru, dashed, maps to, color=sheaf] \& \\
+        \cgamma{\opensetc^{\prime}}{\dtotalc^{\prime}\restriction_{\opensetc^{\prime}}} 
+        \arrow[rrr, "\vartistc^{\prime}", Rightarrow, color=artist]  
+        \& \& \& 
+        \cgamma{\opensetgc^{\prime}}{\gtotalc^{\prime}\restriction_{\opensetgc^{\prime}}}
+        \end{tikzcd}
+    \end{equation*}
+        $(\vartistc, \vartistc^{\prime})$ are equivariant with respect to $\dfuncc_{\dtotalc}$ if a compatible transform  $\dfuncc_{\dtotalc}$ can be defined such that 
+        \begin{equation*}
+            \vartistc^{\prime}(\dfunctc_{\dtotalc}(\dsectionc(\dfunchc_{\dtotalc}(\vindexc(\gbasepointc^{\prime}))))) = \dfunctc_{\gtotalc}(\vartistc(\dsectionc(\vindexc(\dfunchc_{\gtotalc}(\gbasepointc^{\prime})))))
+        \end{equation*}
+        for all points$
+        \gbasepointc^{\prime}\in \gbasec^{\prime$
+\end{frame}
+
+\begin{frame}<presentation:0|handout:0>{$\vartistc= \vartistc^{\prime}$}
     \begin{align*}
@@
     \end{align*}
-    \end{alertblock}
-\end{frame}
-
-\begin{frame}<presentation:1|handout:1>{Equivariant $\dfunchc$}
-    \note{Maybe move to implementation section}
-        \begin{equation*}
-            \begin{tikzcd}[ampersand replacement=\&]
-            \opensetgc 
-            \arrow[rr, "\vindexc", color=functor] 
-            \arrow[dd, "\dfunchc_{\gbasec}"', dashed, color=action] 
-            \&  \& 
-            \opensetc 
-            \arrow[dd, "\dfunchc_{\dbasec}", color=action] \\
-            \&  \& \\
-            \opensetgc^{\prime} 
-            \arrow[rr, "\vindexc"', color=functor]
-            \&  \& 
-            \opensetc^{\prime}\\
-            \underbrace{\dbasec \times {[0,1]}^{m}}_{\gbasec} \arrow[rr,"\vindex", maps to, color=functor] \& \& \dbasec
-            \end{tikzcd}
-        \end{equation*}
 \end{frame}

### 2022-03-07 slides/main.tex
testing equivariance

@@
         \end{equation*}
-        for all points$
-        \gbasepointc^{\prime}\in \gbasec^{\prime$
+        for all points $\gbasepointc^{\prime}\in \gbasec^{\prime}$
 \end{frame}
@@
 
-\begin{frame}<presentation:1|handout:1>{Checking if \vartistc\ is equivariant}
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-            \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
-            \arrow[rr, "\vartistc", Rightarrow, color=artist] \&  \& 
-            \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
-            \arrow[rrd, "\delta"'] 
-            \arrow[rr, "render"] \&  \& visualization 
-            \arrow[d, "measure" description] \\ 
-            \&  \&  \&  \& M
-            \end{tikzcd}
-    \end{equation*}
-    \begin{alertblock}{Visual Measurement $\delta$}
-        \begin{equation}
-            \delta: \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \rightarrow M
-        \end{equation}
-        $M$ is a measurable component of the rendered visual element, e.g. color, position, shape, texture
-    \end{alertblock}
-\end{frame}
-
-\begin{frame}<presentation:1|handout:1>{Checking if \vartistc\ is equivariant}
+\begin{frame}<presentation:1|handout:1>{Testing if \vartistc\ is equivariant}
+    $M$ is a (scaler, vector) measurable component (e.g. color, position, shape, texture, rotation, ) of the rendered visual element.
     \begin{equation*}
@@
     \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
-            \arrow[rr, "\vartistc"', Rightarrow, color=artist] 
+    \arrow[rr, "\vartistc", Rightarrow, color=artist] 
+    \arrow[dd, "\equivc"', color=monoid] \&  \& 
+    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+    \arrow[rr, "render"] 
+    \arrow[dd, "\extractm", color=monoid] 
+    \&  \& 
+    visualization 
+    \arrow[lldd, "\measure"] \\
+    \& \& \& \&  \\
+    {Hom(\opensetc, \measure )} 
+    \&  \& 
+    {Hom(\opensetgc,\measure )} 
+    \arrow[ll, "\vindexpushc"', color=functor] 
+    \& \& 
+    \end{tikzcd}
+    \end{equation*}
+    \begin{description}
+        \item[input]{$\equivc: \dsectionc \mapsto (\opensetc \xrightarrow{\equivc_{\dsectionc}} \measurec)$}
+        \item[output]{$\extractmc:\gsectionc \mapsto (\opensetgc \xrightarrow{\extractmc_{\gsectionc}} \measurec)$}
+        \item[] $\equivc_{\dsectionc}(\dbasepointc) = \extractmc_{\gsectionc}(\gbasepointc)$ for all $\vindexc(\gbasepointc) = \dbasepointc, \dbasepointc \in \dbasec, \gbasepointc \in \gbasec$ 
+    \end{description}
+\end{frame}
+
+\begin{frame}<presentation:0|handout:1->{Visual Measurement}
+    \begin{columns}
+    \column{.5\textwidth}
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement=\&]
+            \measurec 
+            \arrow[rr, "\dfuncc_{\measurec}", color=action] 
+            \& \& 
+            \measurec^{\prime}\\
+            \& \& \\
+            \opensetgc 
+            \arrow[uu, "\extractmc_{\gsectionc}"] 
+            \& \& 
+            \opensetgc^{\prime} 
+            \arrow[ll, "\dfunchc_{\gtotalc}", color=action] 
+            \arrow[uu, "\extractmc_{\dfunctc_{\gtotalc}\gsectionc}"'] 
+            \arrow[lluu, "\extractmc_{\dfuncpullc\gsectionc}" description]
+        \end{tikzcd}
+    \end{equation*}
+    \column{.5\textwidth}
+    \onslide*<presentation:0|handout:2>{
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement=\&]
+            \measurec 
+            \arrow[rr, "\dfuncc_{\measurec}", color=action] 
+            \& \& 
+            \measurec^{\prime}\\
+            \& \& \\
+            \opensetc 
+            \arrow[uu, "\equivc_{\dsectionc}"] 
+            \& \& 
+            \opensetc^{\prime} 
+            \arrow[ll, "\dfunchc_{\dtotalc}", color=action] 
+            \arrow[uu, "\equivc_{\dfunctc_{\dtotalc}\dsectionc}"'] 
+            \arrow[lluu, "\equivc_{\dfuncpullc\dsectionc}" description]
+        \end{tikzcd}
+    \end{equation*}
+    }
+    \end{columns}
+    \onslide*<presentation:0|handout:1>{
+    $\extractmc_{\gsectionc} \in Hom(\opensetgc, \measurec)$ is a mapping from graphic to visual measurement.
+$\extractmc_{\gsectionc}$ is a map from an openset $\opensetgc \subseteq \gbasec$ to a measurement $\measurec_{\opensetgc}$ that corresponds to the graphic representation at that region $\gsectionc\restriction_{\opensetgc}$ and the corresponding data $\dsectionc\restriction_{\vindexc\restriction_{\opensetgc}}$.}
+    \onslide*<presentation:0|handout:2>{
+    $\equivc_{\dsectionc} \in Hom(\opensetc, \measurec)$ is a mapping from data to visual measurement.
+$\equivc_{\dsectionc}:\dsectionc \mapsto$ is a map from an openset $\opensetc \subseteq \dbasec$ to a measurement $\measurec_{\opensetc}$ that corresponds to the data record at that region $\dsectionc\restriction_{\opensetc}$ and the corresponding graphic $\gsectionc\restriction_{\vindexprec\restriction_{\opensetc}}$
+    }
+\end{frame}
+
+\begin{frame}<presentation:1|handout:1->
+    \frametitle<presentation:0|handout:1>{Using output $\gsectionc$ to check if $\vartistc$ is equivariant}
+    \frametitle<presentation:0|handout:2>{Using input $\dsectionc$ to check if $\vartistc$ is equivariant}
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement=\&]
+            \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
+            \arrow[rr, "\vartistc", Rightarrow, color=artist] 
             \arrow[d, "\dfuncc_{\dtotalc}"', color=action] 
-            \arrow[rrr, "\equivc", bend left, color=monoid, dashed] \&  \& 
+            \onslide*<presentation:0|handout:2>{\arrow[rrr, "\equivc", bend left, color=monoid, dashed]} \&  \& 
             \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
-            \arrow[r, "\delta"']\& M 
-            \arrow[d, "\dfuncc_{M}", color=action] \\
-            \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}} 
+            \arrow[d, "\dfuncc_{\gtotalc}"', color=action]
+            \onslide*<presentation:0|handout:1>{\arrow[r, "\delta"]}
+            \onslide*<presentation:0|handout:2>{\arrow[r,"\delta \circ \vindexprec"]}
+            \& 
+            \onslide*<presentation:0|handout:1>{Hom(\opensetgc, \measurec)}
+            \onslide*<presentation:0|handout:2>{Hom(\opensetc, \measurec)} 
+            \arrow[d, "\dfuncc_{\measurec}", color=action] \\
+            \cgamma{\opensetc^{\prime}}{\dtotalc^{\prime}\restriction_{\opensetc^{\prime}}} 
             \arrow[rr, "\vartistc", Rightarrow, color=artist] 
-            \arrow[rrr, "\equivc"', bend right, color=monoid, dashed] \&  \& 
-            \cgamma{\opensetgc^{\prime}}{\dfuncpullc\gtotalc\restriction_{\opensetgc^{\prime}}} 
-            \arrow[r, "\delta"] \& M^{\prime}                        
-        \end{tikzcd}
-    \end{equation*}
-        there exists a functor $\equivc: \dfuncc_{\dtotalc} \rightarrow \dfuncc_{M}$ such that 
-        \begin{description}
-            \item[equivariance] $\equivc(\dfuncc_{\dtotalc}(\dsectionc(\dbasepointc))) = \dfuncc_{M}(\equivc(\dsectionc(\dbasepointc)))$, for all $\dbasepointc \in \dbasec$ 
-            \item[continuity] $\lim\limits_{x\to \dbasepointc}(\equivc(\dsectionc(x))) = \equivc(\dsectionc(\dbasepointc))$ for all $\dbasepointc \in \dbasec$ 
-        \end{description} 
+            \onslide*<presentation:0|handout:2>{\arrow[rrr, "\equivc"', bend right, color=monoid, dashed]} \&  \& 
+            \cgamma{\opensetgc^{\prime}}{\gtotalc^{\prime}\restriction_{\opensetgc^{\prime}}} 
+            \onslide*<presentation:0|handout:1>{\arrow[r, "\delta"]}
+            \onslide*<presentation:0|handout:2>{\arrow[r,"\delta \circ \vindexprec"]} \& 
+            \onslide*<presentation:0|handout:1>{Hom(\opensetgc^{\prime}, \measurec^{\prime})}
+            \onslide*<presentation:1|handout:2>{Hom(\opensetc^{\prime}, \measurec^{\prime})}                        
+        \end{tikzcd}
+    \end{equation*}
+    \onslide*<presentation:0|handout:1>{
+        \begin{align*}
+            \vartistc^{\prime}(\dfunctc_{\dtotalc}(\dsectionc(\dfunchc_{\dtotalc}(\vindexc(\gbasepointc^{\prime}))))) &= \dfunctc_{\gtotalc}(\vartistc(\dsectionc(\vindexc(\dfunchc_{\gtotalc}(\gbasepointc^{\prime})))))\\
+            \extractmc(\dfunctc_{\gtotalc}(\gsectionc(\dfunchc_{\gtotalc})))(\gbasepointc^{\prime}) &= \dfuncc_{\measurec}(\extractmc(\gsectionc))(\gbasepointc^{\prime}) = \extractmc_{\dfunctc_{\gtotalc}\gsectionc}(\gbasepointc^{\prime})
+        \end{align*}
+    }
+    \onslide*<presentation:0|handout:2>{
+        \begin{description}
+            \item[equivariance] $\equivc(\dfunctc_{\dtotalc}(\dsectionc(\dfunchc_{\dtotalc})))(\dbasepointc^{\prime}) =\dfuncc_{\measurec}(\equivc(\dsectionc))(\dbasepointc^{\prime})= \equivc_{\dfunctc_{\dtotalc}\dsectionc}(\dbasepointc^{\prime})$ for all $\dbasepointc^{\prime} \in \dbasec^{\prime}$
+            \item[continuity] $\lim\limits_{x\to \dbasepointc}\equivc_{\dsectionc}(x) = \equivc_{\dsectionc}(\dbasepointc)$ for all $\dbasepointc \in \dbasec$ 
+        \end{description} 
+    }
 \end{frame}
@@
 \subsection{Combining Components into Visualizations}
-
 \begin{frame}<presentation:1|handout:0>{Assembly Function $\vmarkc$}

### 2022-03-21 slides/main.tex
changes to nu implementation, #17

@@
 \section{Algebraic Topology & Category Theory}
-\begin{frame}<presentation:1|handout:0>{Mathematical Data Abstraction}
+\begin{frame}<presentation:1|handout:1>{Mathematical Data Abstraction}
     \begin{description}
@@
 \subsection{Fiber Bundles}
-\begin{frame}<presentation:1|handout:0>
+\begin{frame}<presentation:1|handout:1>
     \frametitle<1|handout:1>{Fiber Bundle}
@@
 
-\begin{frame}<presentation:1|handout:0>
+\begin{frame}<presentation:1|handout:1>
     \frametitle{Presheaf: $\sheafc: \mathcal{C}^{op} \textcolor{sheaf}{\rightarrow} \setc $}
@@
 
-\begin{frame}<presentation:1|handout:0>{Sheaves on Bundles}
+\begin{frame}<presentation:1|handout:1>{Sheaves on Bundles}
     A sheaf is a presheaf that satisfies the following two axioms\cite{bakerMathsSheaf}\\
@@
             \item $\dfiberc \hookrightarrow \dtotalc \xrightarrow{\pi} \dbasec$
-            \item $\sheafc_{\dbasec, \dtotalc}:\opensetc \mapsto \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}, \opensetc \subset \dbasec$
-            \item $\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \ni \dsectionc: \opensetc \rightarrow \dfiberc\restriction_{\opensetc}$
+            \item $\sheafc_{\dbasec, \dtotalc}:\opensetc \mapsto \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}, \opensetc \subseteq \dbasec$
+            \item $\dsectionc: \opensetc \rightarrow \dfiberc\restriction_{\opensetc} \in \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$
             \item $\dsectionc(\dbasepointc) = \{f_0:v_0, \cdots,\}, \dbasepointc \in \opensetc$
@@
             \item $\vindexpullc\dfiberc \hookrightarrow \vindexpullc\dtotalc \xrightarrow{\pi} \gbasec$
-            \item $\vindexpullc\sheafc_{\dbasec,\dtotalc}:\opensetgc \mapsto \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}, \opensetgc \subset \gbasec$
-            \item $\cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} \ni \vindexpullc\dsectionc: \opensetgc \rightarrow \vindexpullc\dfiberc\restriction_{\opensetgc}$
+            \item $\vindexpullc\sheafc_{\dbasec,\dtotalc}:\opensetgc \mapsto \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}, \opensetgc \subseteq \gbasec$
+            \item $\vindexpullc\dsectionc: \opensetgc \rightarrow \vindexpullc\dfiberc\restriction_{\opensetgc} \in \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}$
             \item $\vindexpullc\dsectionc(\gbasepointc) = \dsectionc(\vindexc(\gbasepointc)) = \dsectionc(\dbasepointc)$
@@
             \item $\gfiberc \hookrightarrow \gtotalc \xrightarrow{\pi} \gbasec$
-            \item $\sheafc_{\gbasec, \gtotalc}:\opensetgc \mapsto \cgamma{\opensetgc}{\dtotalc\restriction_{\opensetgc}}, \opensetgc \subset \gbasec$
-            \item $\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \ni \gsectionc: \opensetgc \rightarrow \gfiberc\restriction_{\opensetgc}$
+            \item $\sheafc_{\gbasec, \gtotalc}:\opensetgc \mapsto \cgamma{\opensetgc}{\dtotalc\restriction_{\opensetgc}}, \opensetgc \subseteq \gbasec$
+            \item $\gsectionc: \opensetgc \rightarrow \gfiberc\restriction_{\opensetgc} \in \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}$
             \item $\gsectionc(\gbasepointc) = \{d_0, \cdots\}, \gbasepointc \in \opensetgc$
@@
             \item $\vindexpushc\gfiberc \hookrightarrow \vindexpushc\gtotalc \xrightarrow{\pi} \dbasec$
-            \item $\vindexpushc\sheafc_{\gbasec, \gtotalc}:\opensetc \mapsto \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}, \opensetc \subset \dbasec$
-            \item $\cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} \ni \vindexpushc\gsectionc: \opensetc \rightarrow \vindexpushc \gfiberc\restriction_{\opensetc}$
-            \item $\vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexpre(\dbasepointc)} = \gsectionc(\gbasepointc)\;\forall \gbasepointc \in \vindexpre(\dbasepointc)$ 
+            \item $\vindexpushc\sheafc_{\gbasec, \gtotalc}:\opensetc \mapsto \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}, \opensetc \subseteq \dbasec$
+            \item $\vindexpushc\gsectionc: \opensetc \rightarrow \vindexpushc \gfiberc\restriction_{\opensetc} \in \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}$
+            \item $\vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexpre(\dbasepointc)} = \gsectionc(\gbasepointc)\;\forall\;                                                                                                                                           \gbasepointc \in \vindexpre(\dbasepointc)$
         \end{itemize}
@@
 
-\begin{frame}<presentation:1|handout:0>{Example: Scatter Plot}
+\begin{frame}<presentation:1|handout:1>{Example: Scatter Plot}
     \begin{figure}
@@
 
-\begin{frame}<presentation:1|handout:0>{Example: Line Plot}
+\begin{frame}<presentation:1|handout:1>{Example: Line Plot}
     \begin{figure}
@@
 
-\begin{frame}<presentation:1|handout:0>{Example: Image}
+\begin{frame}<presentation:1|handout:1>{Example: Image}
     \begin{figure}
@@
 
-\begin{frame}<presentation:1|handout:0>{Example: Heatmap}
+\begin{frame}<presentation:1|handout:1>{Example: Heatmap}
     \begin{figure}
@@
 
-\begin{frame}<presentation:1|handout:0>{Example: Sphere}
+\begin{frame}<presentation:1|handout:1>{Example: Sphere}
     \begin{figure}
@@
 
-\begin{frame}<presentation:1|handout:0>
+\begin{frame}<presentation:1|handout:1>
     \frametitle<0|handout:1>{Artist: $\vartistc: \sheafc_{\dbasec, \dtotalc} \rightarrow \sheafc_{\gbasec, \gtotalc}$}
@@
 
+\begin{frame}<presentation:1|handout:1>{Reachable \gsectionc?}
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement = \&, column sep=tiny, row sep=huge]
+            \onslide*<1-|handout:1>{\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}} 
+            \&  \& 
+            \onslide*<2-|handout:1>{\imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}}} 
+            \onslide*<2-|handout:1>{\arrow[ll, "\textcolor{set}{\mathlarger{\mathlarger{\supset}}}" description, hook', color=white]} \\ 
+            \&  \& \\
+            \onslide*<1-|handout:1>{\opensetgc} 
+            \onslide*<1-|handout:1>{\arrow[uu, "\sheafc_{\gbasec,\gtotalc}", color=sheaf, maps to]} 
+            \onslide*<2-|handout:1>{\arrow[rruu, "\vartistc(\sheafc_{\dbasec,\dtotalc})" description, dashed, color=sheaf, maps to]} 
+            \&  \& 
+        \end{tikzcd}
+    \end{equation*}
+    \begin{alertblock}{Output Subtype}
+        \begin{equation*}
+            \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} = \{\gsectionc\vert\;\exists\;\dsectionc \in \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}\;s.t.\; 
+            \vartistc(\dsectionc) = \gsectionc,\; \vindexc(\opensetgc) = \opensetc \}
+        \end{equation*}
+    \end{alertblock}
+\end{frame}
+
 \subsection{Equivariance}
 \begin{frame}<presentation:1|handout:1>{Transform Data}
+    %swap fiber & bundle, note that need section morphisms not bundle morphisms
+    \begin{block}{Fiber Category}
+        The fiber $\dfiberc$ is a  monoidal category (single object w/ bicartesian product operator on category) of an arbitrary type $\mathcal{C}$.
+The morphisms on the fiber are $\dfunctc \in Hom(\dfiberc, \dfiberc)$
+    \end{block}
+
     \begin{block}{Fiber Bundle Category}
@@
             \item[object] $\dfiberc \hookrightarrow \dtotalc \xrightarrow{\pi} \dbasec$
-            \item[morphisms] $\dfuncc: \dtotalc \rightarrow \dtotalc^{\prime}$
-        \end{description}
-    \end{block}
-    \begin{block}{Fiber Category}
-        The fiber $\dfiberc$ is a  monoidal category (single object w/ bicartesian product operator on category) of an arbitrary type $\mathcal{C}$.
-The morphisms on the fiber are $Hom(\dfiberc, \dfiberc)$
-    \end{block}
-
-    \begin{block}{Data Transformations}
-    $\dfuncc:\dtotalc\restriction_{\dbasepointc} \rightarrow \dtotalc^{\prime}\restriction_{\dbasepointc^{\prime}} \in Hom(\dfiberc\restriction_{\dbasepointc},\dfiberc^{\prime}\restriction_{\dbasepointc^{\prime}})$
+            \item[morphisms] $\dfuncc:(\dfunchc, \dfunctc)$
+        \end{description}
     \end{block}
@@
         \&  \& 
-        \cgamma{\opensetc^{\prime}}{\dtotalc^{\prime}\restriction_{\opensetc^{\prime}}} \\
+        \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}} \\
         \&  \&  
@@
         \arrow[ll, "\dfunchc", color=action] 
-        \arrow[uu, "{\sheafc_{\dbasec^{\prime}, \dfuncpullc\dtotalc}}"', maps to, color=sheaf] 
-        \arrow[rruu, "{\sheafc_{\dbasec^{\prime}, \dtotalc^{\prime}}}"', maps to, color=sheaf] 
-        \&  \& 
-    \end{tikzcd}
-    \begin{itemize}
-        \item Base Transformation: $\dfunchc: \opensetc^{\prime} \rightarrow \opensetc$
-        \item Fiber Transformation: $\dfunctc: \dfuncpullc \dtotalc \rightarrow \dtotalc^{\prime}$ s.t. $\pi(\dfuncpullc\dtotalc) = \pi(\dtotalc^{\prime})$ 
-        \item Section Transform: $\dfuncc: \dsectionc\restriction_{\opensetc} \mapsto \dsectionc^{\prime}\restriction_{\opensetc^{\prime}}$
-    \end{itemize}
+        \arrow[uu, "{\sheafc_{\dbasec, \dfuncpullc\dtotalc}}"', maps to, color=sheaf] 
+        \arrow[rruu, "{\sheafc_{\dbasec, \dfuncpullc\dtotalc}}"', maps to, color=sheaf] 
+        \&  \& 
+    \end{tikzcd}
+    \begin{description}
+        \item[Base Transformation:] {$\dfunchc: \opensetc^{\prime} \rightarrow \opensetc$ where $\opensetc, \opensetc^{\prime} \subseteq \dbasec$, \\
+        $\dfuncpullc\dsectionc\restriction_{\opensetc}: \dsectionc \mapsto \dsectionc\restriction_{\opensetc} \circ \dfunchc$}
+
+        \item [Fiber Transformation:]{$\dfunctc: \dfuncpullc \dtotalc_{\dbasepointc^{\prime}} \rightarrow {\dfuncpullc\dtotalc_{\dbasepointc^{\prime}}} \in Hom(\dfuncpullc\dfiberc\restriction_{\dbasepointc}, \dfuncpullclllllll\dfiberc\restriction_{\dbasepointc}), \dbasepointc^{\prime} \in \opensetc^{\prime}$\\
+        $\dfunctc: \dfuncpullc \dsectionc\restriction_{\opensetc} \mapsto \dfuncpullc \dsectionc^{\prime}\restriction_{\opensetc}$, $\dsectionc, \dsectionc^{\prime} \in \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}}$}
+        \item [Section Transform:] $\dfuncc: \dsectionc\restriction_{\opensetc} \mapsto \dfuncpullc\dsectionc^{\prime}\restriction_{\opensetc}$
+    \end{description}
 \end{frame}
@@
 \begin{frame}<presentation:1|handout:1>{Equivariant Artist}
+    $(\vartistc, \vartistc^{\prime})$ are equivariant with respect to $\dfuncc_{\dtotalc}$ if a compatible transform  $\dfuncc_{\gtotalc}$ can be defined such that 
     \begin{equation*}
@@
         \& \& \& 
-        \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+        \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
         \arrow[d, "\dfuncpullc_{\gtotalc}", dotted] \\
@@
         \arrow[ru, dashed, maps to, color=sheaf] \& 
-        \cgamma{\opensetgc^{\prime}}{\dfuncpullc_{\gtotalc}\gtotalc\restriction_{\opensetgc^{\prime}}} 
+        \imartist{\opensetgc^{\prime}}{\dfuncpullc_{\gtotalc}\gtotalc\restriction_{\opensetgc^{\prime}}} 
         \arrow[dd, "\dfunctc_{\gtotalc}", dotted, color=action] \\
@@
         \& \& \& 
-        \cgamma{\opensetgc^{\prime}}{\gtotalc^{\prime}\restriction_{\opensetgc^{\prime}}}
-        \end{tikzcd}
-    \end{equation*}
-        $(\vartistc, \vartistc^{\prime})$ are equivariant with respect to $\dfuncc_{\dtotalc}$ if a compatible transform  $\dfuncc_{\dtotalc}$ can be defined such that 
-        \begin{equation*}
-            \vartistc^{\prime}(\dfunctc_{\dtotalc}(\dsectionc(\dfunchc_{\dtotalc}(\vindexc(\gbasepointc^{\prime}))))) = \dfunctc_{\gtotalc}(\vartistc(\dsectionc(\vindexc(\dfunchc_{\gtotalc}(\gbasepointc^{\prime})))))
-        \end{equation*}
-        for all points $\gbasepointc^{\prime}\in \gbasec^{\prime}$
+        \imartist{\opensetgc^{\prime}}{\gtotalc^{\prime}\restriction_{\opensetgc^{\prime}}}
+        \end{tikzcd}
+    \end{equation*}
+    %$\vartistc^{\prime}(\dfunctc_{\dtotalc}(\dsectionc(\dfunchc_{\dtotalc}(\vindexc(\gbasepointc^{\prime}))))) = \dfunctc_{\gtotalc}(\vartistc(\dsectionc(\vindexc(\dfunchc_{\gtotalc}(\gbasepointc^{\prime})))))$
+    %for all points $\gbasepointc^{\prime}\in \gbasec^{\prime}$
 \end{frame}
@@
     \arrow[dd, "\equivc"', color=monoid] \&  \& 
-    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+    \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
     \arrow[rr, "render"] 
@@
     visualization 
-    \arrow[lldd, "\measure"] \\
+    \arrow[lldd, "measure"] \\
     \& \& \& \&  \\
-    {Hom(\opensetc, \measure )} 
-    \&  \& 
-    {Hom(\opensetgc,\measure )} 
-    \arrow[ll, "\vindexpushc"', color=functor] 
+    {Hom(\opensetc, \measurec)} 
+    \arrow[rr, "\vindexpullc"', color=functor] 
+    \&  \& 
+    {Hom(\opensetgc,\measurec )} 
     \& \& 
@@
 
-\begin{frame}<presentation:0|handout:1->{Visual Measurement}
+\begin{frame}<presentation:0|handout:->{Visual Measurement}
     \begin{columns}
@@
             \measurec 
-            \arrow[rr, "\dfuncc_{\measurec}", color=action] 
+            \arrow[rr, "\dfunctc_{\measurec}", color=action] 
             \& \& 
@@
             \arrow[d, "\dfuncc_{\dtotalc}"', color=action] 
-            \onslide*<presentation:0|handout:2>{\arrow[rrr, "\equivc", bend left, color=monoid, dashed]} \&  \& 
-            \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
-            \arrow[d, "\dfuncc_{\gtotalc}"', color=action]
-            \onslide*<presentation:0|handout:1>{\arrow[r, "\delta"]}
-            \onslide*<presentation:0|handout:2>{\arrow[r,"\delta \circ \vindexprec"]}
+            \onslide*<presentation:0|handout:2>{\arrow[rrr, "\equivc", bend left, color=monoid]} \&  \& 
+            \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+            \arrow[d, "\dfuncc_{\gtotalc}"', dotted, color=action]
+            \onslide*<presentation:0|handout:1>{\arrow[r, "\extractmc", color=monoid]}
+            \onslide*<presentation:0|handout:2>{\arrow[r,"\vindexprec \circ \extractmc", color=monoid]}
             \& 
@@
             \arrow[rr, "\vartistc", Rightarrow, color=artist] 
-            \onslide*<presentation:0|handout:2>{\arrow[rrr, "\equivc"', bend right, color=monoid, dashed]} \&  \& 
-            \cgamma{\opensetgc^{\prime}}{\gtotalc^{\prime}\restriction_{\opensetgc^{\prime}}} 
-            \onslide*<presentation:0|handout:1>{\arrow[r, "\delta"]}
-            \onslide*<presentation:0|handout:2>{\arrow[r,"\delta \circ \vindexprec"]} \& 
+            \onslide*<presentation:0|handout:2>{\arrow[rrr, "\equivc"', bend right, color=monoid]} \&  \& 
+            \imartist{\opensetgc^{\prime}}{\gtotalc^{\prime}\restriction_{\opensetgc^{\prime}}} 
+            \onslide*<presentation:0|handout:1>{\arrow[r, "\extractmc", color=monoid]}
+            \onslide*<presentation:0|handout:2>{\arrow[r,"\vindexprec \circ \extractmc", color=monoid]} \& 
             \onslide*<presentation:0|handout:1>{Hom(\opensetgc^{\prime}, \measurec^{\prime})}
@@
 
-\begin{frame}<presentation:1|handout:1>{"Valid" viz?}
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement = \&, column sep=tiny, row sep=huge]
-            \onslide*<1-|handout:1>{\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}} 
-            \&  \& 
-            \onslide*<2-|handout:1>{\textcolor{set}{\big\{}\gsectionc \bigm{\vert} 
-              \delta(\vartistc(\dfuncc_{\dtotalc}(\dsectionc))) = \dfuncc_{M}(\delta(\gsectionc)) \textcolor{set}{\big\}}} 
-            \onslide*<2-|handout:1>{\arrow[ll, "\textcolor{set}{\mathlarger{\mathlarger{\supset}}}" description, hook', color=white]} \\ 
-            \&  \& \\
-            \onslide*<1-|handout:1>{\opensetgc} 
-            \onslide*<1-|handout:1>{\arrow[uu, "\sheafc_{\gbasec,\gtotalc}", color=sheaf, maps to]} 
-            \onslide*<2-|handout:1>{\arrow[rruu, "\vartistc(\sheafc_{\dtotalc})\coloneqq \sheafc_{\vartistc}" description, dashed, color=sheaf, maps to]} 
-            \&  \& 
-        \end{tikzcd}
-    \end{equation*}
-\end{frame}
-
 \section{Construction}
-\begin{frame}<presentation:1|handout:0>{What's in the $\vartist$ black box?}
+\begin{frame}<presentation:1|handout:1>{Building an equivariant $\vartist$?}
     \note{Add box around figure & lable it artist}
@@
 
-\subsection{Measurable Visual Components $M$}
-
-\begin{frame}<presentation:1|handout:0>{Measurable Visual Components}
+\subsection{Measurement Specification $\vtotalc$}
+
+\begin{frame}<presentation:1|handout:1>{Typed Measurable Visual Components: $\vtotalc$}
     \begin{columns}
@@
 
-\begin{frame}<presentation:1|handout:0>
-    \frametitle{Data to Visual Transformation: $\vchannelc: \dfiberc_{\dbasepointc} \mapsto \vfiberc_{\dbasepointc}$}
+\begin{frame}<presentation:1|handout:1>
+    \frametitle{Data to Visual Transformation: $\vchannelc: \dtotalc \rightarrow \vtotalc$}
+    \begin{equation*}
+    \begin{tikzcd}[ampersand replacement = \&]
+        \dtotalc 
+        \arrow[dd, "\vchannelc_{\opensetc}"', color=artist] 
+        \arrow[rr, "\dfuncpullc"] 
+        \& \& 
+        \dfuncpullc\dtotalc 
+        \arrow[rr, "\dfunctc_{\dtotalc}"] 
+        \&  \& 
+        \dtotalc^{\prime} 
+        \arrow[dd, "{\vchannelc^{\prime}}_{\opensetc^{\prime}}", color=artist] \\
+        \& 
+        \opensetc 
+        \arrow[lu, dotted, maps to] 
+        \arrow[ld, dotted, maps to] \& 
+        \opensetc^{\prime} 
+        \arrow[l, "\dfunchc"] 
+        \arrow[u, dotted, maps to] 
+        \arrow[d, dotted, maps to] 
+        \arrow[rrd, dotted, maps to] 
+        \arrow[rru, dotted, maps to] 
+        \& \& \\
+        \vtotalc 
+        \arrow[rr, "\dfuncpullc"'] 
+        \& \& 
+        \dfuncpullc\vtotalc 
+        \arrow[rr, "\dfunctc_{\vtotalc}"'] 
+        \& \& 
+        \vtotalc^{\prime}
+    \end{tikzcd}
+    \end{equation*}
+    Given $\dsectionc(\dbasepointc) = \dsectionc(\dfunchc(\dbasepointc^{\prime}))$ and  $\vsectionc(\dbasepointc) = \vsectionc(\dfunchc(\dbasepointc^{\prime}))$
+    \begin{description}
+        \item[equivariance] $\dfunctc_{\vtotalc}(\vchannelc_{\opensetc}(\dsectionc))
+        = \vchannelc_{\opensetc}(\dfunctc_{\dtotalc}(\dsectionc))$, $\vchannelc_{\opensetc} = {\vchannelc^{\prime}}_{\opensetc^{\prime}}$
+        \item[continuity] $\lim\limits_{x\to \dbasepointc}(\vchannelc(\dsectionc(x))) = \vchannelc(\dsectionc(\dbasepointc))$ for all $\dbasepointc \in \dbasec$ 
+    \end{description}
+\end{frame}
+
+\begin{frame}{Data to Visual Transformation $\vchannelc: \dfiberc_{\dbasepointc} \mapsto \vfiberc_{\dbasepointc}$}
+$\pi(\dtotalc) = \pi(\vchannelc(\dtotalc))$ and $\vchannelc$ is composable s.t 
     \begin{columns}
@@
         \begin{tikzcd}[ampersand replacement=\&]
-        \dtotalc 
-        \arrow[rr, "\vchannelc_{\dbasec}", color=artist] 
-        \arrow[rd, "\pi" description, dotted] 
-        \arrow[ddd, "\dfuncc_{\dtotalc}"', color=action] 
-        \& \& 
-        \vtotalc 
-        \arrow[ld, "\pi" description, dotted] 
-        \arrow[ddd, "\dfuncc_{\vtotalc}", color=action] \\
-        \& \dbasec \&  \\
-        \& \dbasec^{\prime} 
-        \arrow[u, "\dfunchc" description, color=action] \& \\
-        \dfuncpullc\dtotalc 
-        \arrow[ru, "\pi" description, dotted] 
-        \arrow[rr, "\vchannel_{\dbasec}"', color=artist] 
-        \& \& 
-        \dfuncpullc\dtotalc 
-        \arrow[lu, "\pi" description, dotted]
+            \dfiberc_{\dbasepointc} 
+            \arrow[rr, "\vchannelc", maps to] 
+            \arrow[dd, "\dfunctc_{\dtotalc}"'] 
+            \& \& 
+            \vfiberc_{\dbasepointc} 
+            \arrow[dd, "\dfunctc_{\vtotalc}"] \\
+            {} 
+            \arrow[rr, dashed, maps to]
+            \& \& {} \\
+            {\dfiberc^{\prime}}_{\dbasepointc} 
+            \arrow[rr, "\vchannelc^{\prime}"', maps to]
+            \& \& 
+            {\vfiberc^{\prime}}_{\dbasepointc}
             \end{tikzcd}
@@
         \column{.5\textwidth} 
-    \begin{equation}
+        \begin{equation*}
         \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
@@
             \end{tikzcd}
-    \end{equation}
+        \end{equation*}
     \end{columns}
-    % fix corner to be V 
-    % fix \phi to be \phi 
-    \begin{alertblock}{equivariance} 
-       $\vchannelc: \dfuncc_{\dtotalc} \rightarrow \dfuncc_{\vtotalc}$ is a functor such that 
-       \begin{description}
-           \item[equivariance] {$\vchannelc(\dfuncc(\dsectionc(\dfunchc_{\dtotalc}^{-1}(\dbasepointc)))) = \dfuncc_{\vtotalc}(\vchannelc(\dsectionc(\dfunchc_{\vtotalc}^{-1}(\dbasepointc))))$\\$\dfunchc_{\dtotalc}(\dbasepointc^{\prime}) = \dfunchc_{\vtotalc}(\dbasepointc^{\prime})$}
-           \item[continuity] $\lim\limits_{x\to \dbasepointc}(\vchannelc(\dsectionc(x))) = \vchannelc(\dsectionc(\dbasepointc))$ for all $\dbasepointc \in \dbasec$ 
-       \end{description}
-\end{alertblock}
-\end{frame}
-
-\begin{frame}<presentation:1|handout:0>\frametitle{$\vchannelc:\dfuncc_{\dtotalc} \rightarrow \dfuncc_{\vtotalc}$: Stevens' Scales \cite{stevensTheoryScalesMeasurement1946}}
+
+\end{frame}
+
+\begin{frame}<presentation:1|handout:1>\frametitle{$\vchannelc:\dfuncc_{\dtotalc} \rightarrow \dfuncc_{\vtotalc}$: Stevens' Scales \cite{stevensTheoryScalesMeasurement1946}}
     \begin{table}[H]
@@
 
-\begin{frame}<presentation:1|handout:0>
+\begin{frame}<presentation:1|handout:1>
     \frametitle{Shared Components: $\vchannelc = \prod\limits_{i=0}^{n}\vchannelc_i$}
@@
 \subsection{Combining Components into Visualizations}
-\begin{frame}<presentation:1|handout:0>{Assembly Function $\vmarkc$}
+\begin{frame}<presentation:1|handout:1>{Assembly $\vmarkc$}
     \begin{equation*}
@@
             \arrow[rrrr, "\vmarkc", Rightarrow, color=artist] 
-            \arrow[ddd, "\dfuncc_{\vtotalc}"', color=action] 
+            \arrow[ddd, "\dfunctc_{\vtotalc} \circ \dfuncpullc_{\vtotalc}"', color=action] 
             \& \& \& \& 
             \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
-            \arrow[ddd, "\dfuncc_{\gtotalc}", color=action]  \\
+            \arrow[ddd, "\dfunctc_{\gtotalc} \circ \dfunchc_{\gtotalc}", color=action]  \\
             \& 
@@
             \arrow[rd, maps to, color=functor] \& \\
-            \cgamma{\opensetc^{\prime}}{\dfuncpullc_{\vtotalc}\vtotalc\restriction_{\opensetc^{\prime}}}
-            \arrow[rrrr, "\vmarkc", Rightarrow, color=artist] 
+            \cgamma{\opensetc^{\prime}}{\vtotalc^{\prime}\restriction_{\opensetc^{\prime}}}
+            \arrow[rrrr, "\vmarkc^{\prime}", Rightarrow, color=artist] 
             \& \&  \& \& 
-            \cgamma{\opensetgc^{\prime}}{\dfuncpullc_{\gtotalc}\gtotalc\restriction_{\opensetgc^{\prime}}}
+            \cgamma{\opensetgc^{\prime}}{\gtotalc^{\prime}\restriction_{\opensetgc^{\prime}}}
             \end{tikzcd}
@@
        \begin{equation*}
-        \vmarkc(\dfuncc_{\vtotalc}(\vsectionc(\dfunchc_{\vtotalc}^{-1}(\dbasepointc^{\prime})))) = \dfuncc_{\gtotalc}(\vmarkc((\vsectionc(\dfunchc_{\gtotalc}^{-1}(\vindexprec(\dbasepointc^{\prime}))))))
-       \end{equation*}
-    \end{alertblock}
-\end{frame}
-
-\begin{frame}<presentation:1|handout:0>{Implementation Choices: $\vartistc_{\opensetc} = \vartistc_{\opensetgc}$}
-    \begin{equation*}
-    \begin{tikzcd}[ampersand replacement = \&, column sep=small]
+        \vmarkc^{\prime}(\dfunctc_{\vtotalc}(\vsectionc(\dfunchc_{\dtotalc}(\vindexc(\gbasepointc^{\prime}))))) = \dfunctc_{\gtotalc}(\vmarkc(\vsectionc(\vindexc(\dfunchc_{\gtotalc}(\gbasepointc^{\prime})))))
+       \end{equation*}
+    \end{alertblock}
+\end{frame}
+
+\begin{frame}<presentation:1|handout:1>{Implementation Choices: $\vartistc_{\opensetc} = \vartistc_{\opensetgc}$}
+    \begin{equation*}
+    \begin{tikzcd}[ampersand replacement = \&, column sep=tiny]
         \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}  
-        \arrow[r, "\vchannelc", maps to, color=artist] 
+        \arrow[r, "\vchannelc_{\opensetgc}", maps to, color=artist] 
         \arrow[rrrr, "\vartistc_{\opensetgc}", Rightarrow, bend left, color=artist] \& 
@@
         \cgamma{\opensetgc}{\vindexpullc\vtotalc\restriction_{\opensetgc}}  
-        \arrow[rr, "\vmarkc^*", no head, Rightarrow, color=artist] \&  \& 
+        \arrow[rr, "\vmarkc_{\opensetgc}", no head, Rightarrow, color=artist] \&  \& 
         \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}  
@@
         \arrow[uu, "\vindexpullc", color=functor] 
-        \arrow[r, "\vchannelc", maps to, color=artist] 
+        \arrow[r, "\vchannelc_{\opensetc}", maps to, color=artist] 
         \arrow[rrrr, "\vartistc_{\opensetc}", Rightarrow, bend right, color=artist] 
@@
         \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}}
-        \arrow[rr, "\vmarkc_*", Rightarrow, color=artist] 
+        \arrow[rr, "\vmarkc_{\opensetc}", Rightarrow, color=artist] 
         \arrow[uu, "\vindexpullc", color=functor] 
@@
 
-\begin{frame}<presentation:1|handout:0>{Components to Graphic}
-    \begin{figure}
-        \includegraphics[scale=.15]{../paper/figures/qhat.png}
+\begin{frame}<presentation:1|handout:1>{Components to Graphic}
+    \begin{figure}
+        \includegraphics[scale=.35]{../paper/figures/qhat.png}
     \end{figure}

### 2022-03-21 matplottoy/lib/bar.py

@@ -70,12 +70,16 @@ class TopologicalArtist:
         return self.data.query(data_bounds, sampling_rate)
         # section.view, #xi seperates nicely 
 
+def rename_P(V, new_key):
+    return {(new_key[k] if k in new_key else k):v for (k, v) in V.items()}
+
 def barH(V):
     new_key = {'position': 'x', 'height':'y'}
     return Bar({(new_key[k] if k in new_key else k):v for (k, v) in V.items()})
 
 def barV(V):
     new_key = {'position': 'y', 'height':'x'}
+    # this can return Bar with nuBar & kew renaming should be a nu too maybe 
     return BarV({(new_key[k] if k in new_key else k):v for (k, v) in V.items()})
 
 class Bar(Rho): 
@@ -130,12 +134,19 @@ class Bar(Rho):
 
 class BarH(Bar):
     bounds = {'x': 'length', 'y': 'position'}
+    
+    @staticmethod
+    def bar_nu(length, width, position, floor, facecolor, edgecolor, linewidth, linestyle):
+        return {"position":floor, "width":length, 
+                "length":width, "floor":position, 
+                "facecolor":facecolor, "edgecolor":edgecolor, 
+                "linewidth":linewidth, "linestyle":linestyle}
 
     @staticmethod
     def qhat(length, width, position, floor, facecolor, edgecolor, linewidth, linestyle):
         # horizontal bar, x0 is floor, x1 is height, y is y0, width is y1
         # alternative is to do this renamin in nu calling parent nu first
-        return Bar.qhat(floor, length, width, position, facecolor, edgecolor, linewidth, linestyle)
+        return Bar.qhat(**BarH.bar_nu(length, width, position, floor, facecolor, edgecolor, linewidth, linestyle))
 
 class GenericArtist(martist.Artist):
     # start w/ working w/ an artist object then stripped down artist

### 2022-03-28 slides/main.tex
compiles, presentation mode

@@
-\documentclass[xcolor={dvipsnames}, handout]{beamer}
-%\documentclass[xcolor={dvipsnames}]{beamer}
+%\documentclass[xcolor={dvipsnames}, handout]{beamer}
+\documentclass[xcolor={dvipsnames}]{beamer}
 \setbeamertemplate{footline}[frame number]
@@
 %% maybe pull out the middle for now? move the 3 stages down to construction? 
-\begin{frame}<presentation:1|handout:0>{What do visualization libraries do?}
+\begin{frame}<presentation:1-|handout:0>{What do visualization libraries do?}
     \begin{figure}
@@
 
-\begin{frame}<presentation:1|handout:0>{Equivariance}
+\begin{frame}<presentation:1-|handout:0>{Equivariance}
     \begin{description}
@@
 %flip columns
-\begin{frame}<presentation:1|handout:0>{Frameworks for Expressing Visual Equivariance}
+\begin{frame}<presentation:1-|handout:0>{Frameworks for Expressing Visual Equivariance}
     \begin{description}
@@
 
-        \item[algebraically] transformations on data and graphics are equivalent symmetric \cite{kindlmannAlgebraicProcessVisualization2014}
+        \item[algebraically] transformations on data and graphics are equivalent and symmetric \cite{kindlmannAlgebraicProcessVisualization2014}
         \begin{columns}
@@
             \begin{description}
-                \item[D] data 
-                \item[R] data representation 
+                \item[D] raw data 
+                \item[R] representation (data container) 
                 \item[V] visualization
@@
 \subsection{Fiber Bundles}
-\begin{frame}<presentation:1|handout:1>
+\begin{frame}<presentation:1-|handout:1>
     \frametitle<1|handout:1>{Fiber Bundle}
@@
     \pause 
-    \only<4-5|handout:1>{
+    \only<5|handout:1>{
         \begin{alertblock}{}
@@
 \subsection{Category Theory Language}
-\begin{frame}<presentation:1|handout:0>[fragile]{}
+\begin{frame}<presentation:1|handout:1>[fragile]{}
     \begin{minted}{python}
-        Artist(data:Data) -> Graphic
+        Artist(Data) -> Graphic
     \end{minted}
@@
 
-\begin{frame}<presentation:1|handout:0>{Category $\textcolor{source}{\mathcal{C}}$}
+\begin{frame}<presentation:1-|handout:1>{Category $\textcolor{source}{\mathcal{C}}$}
 \begin{columns}
@@
 
-\begin{frame}<presentation:1|handout:0>{Opposite Category $\textcolor{source}{\mathcal{C}^{op}}$}
+\begin{frame}<presentation:1-|handout:1>{Opposite Category $\textcolor{source}{\mathcal{C}^{op}}$}
     \begin{columns}
@@
         \column{0.49\textwidth}
+        \onslide*<presentation:2|handout:1>{
         \begin{equation*}
@@
         \end{equation*}
+        }
         \end{columns}
@@
 
-\begin{frame}<presentation:1|handout:0>
+\begin{frame}<presentation:1-|handout:1>
     \frametitle{Functor $\textcolor{functor}{\bm{F}}: \textcolor{source}{\mathcal{C}} \rightarrow \textcolor{target}{\mathcal{D}}$}
@@
 \subsubsection{Natural Transforms}
-\begin{frame}<presentation:1|handout:0>
+\begin{frame}<presentation:1-|handout:1>
     \frametitle{Natural Transformation $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\Rightarrow} \textcolor{functor}{G}$}
-    \onslide<|handout:1>{
     \adjustbox{scale=3, center}{
@@
             \& {} \&            
-          \end{tikzcd}}}
-\end{frame}
-
-\begin{frame}<presentation:1|handout:0>
+          \end{tikzcd}}
+\end{frame}
+
+\begin{frame}<presentation:1-|handout:1>
     \frametitle{Natural Transformation $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\Rightarrow} \textcolor{functor}{G}$}
     \begin{equation*}
-        \begin{tikzcd}[column sep=huge, ampersand replacement = \&]
-            \onslide*<2-|handout:1>{\textcolor{target}{F(c)}}
-            \onslide*<2-|handout:1>{\arrow[dd, "F(f)"', color=target]}
-            \onslide*<4-5|handout:1>{\arrow[rr, "\alpha_{c}", bend left, color=nattran, Rightarrow]}  
-            \onslide*<6|handout:0>{\arrow[rr, "\alpha_{c}", color=nattran, Rightarrow]} \& 
-            \onslide*<1-4|handout:1>{\textcolor{source}{c}} 
-            \onslide*<1-4|handout:1>{\arrow[dd, "f" description, color=source]}
-            \onslide*<3-4|handout:1>{\arrow[r, "G"', color=functor,maps to ]}
-            \onslide*<2-4|handout:1>{\arrow[l, "F", color=functor, maps to]} \& 
-            \onslide*<3-|handout:1>{\textcolor{target}{G(c)}} 
-            \onslide*<3-|handout:1>{\arrow[dd, "G(f)", color=target, maps to]} \\
-            {} 
-            \onslide*<4|handout:1>{\arrow[rr, dotted, color=nattran, Rightarrow]}
-            \onslide*<5|handout:0>{\arrow[rr, color=nattran, Rightarrow]} \& \& {}                      \\
-            \onslide*<2-|handout:1>{\textcolor{target}{F(c^{\prime})}} 
-            \onslide*<4-5|handout:1>{\arrow[rr, "\alpha_{c^{\prime}}"', bend right, color=nattran, Rightarrow]} 
-            \onslide*<4-5|handout:0>{\arrow[rr, "\alpha_{c^{\prime}}"', color=nattran, maps to]}\& 
-            \onslide*<1-4|handout:1>{\textcolor{source}{c^{\prime}}}
-            \onslide*<2-4|handout:1>{\arrow[l, "F"', color=functor, maps to]} 
-            \onslide*<3-4|handout:1>{\arrow[r, "G", color=functor, maps to]} \& 
-            \onslide*<3-|handout:1>{\textcolor{target}{G(c^{\prime})}}          
+        \begin{tikzcd}[ampersand replacement=\&]                      
         \end{tikzcd}
@@
 
-\begin{frame}<presentation:1|handout:1>
+\begin{frame}<presentation:1-|handout:1>
     \frametitle{Presheaf: $\sheafc: \mathcal{C}^{op} \textcolor{sheaf}{\rightarrow} \setc $}
@@
             \onslide*<3,6-|handout:0>{{\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}} 
-            \onslide*<4-5|handout:0>{{\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}\in \setc}
-            \onslide*<0|handout:1>{\setc \ni {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}}
+            \onslide*<4-5|handout:0>{{\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}\in Ob(\setc)}
+            \onslide*<0|handout:1>{\Ob(\setc) \ni {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}}
             \&  
@@
 \subsubsection{Pullback and Pushforward}
-\begin{frame}<presentation:1|handout:1->{}
+\begin{frame}<presentation:1-|handout:1->{}
     \frametitle<3|handout:3>{Function: $\vindexc: \gbasec \rightarrow \dbasec$}
     \frametitle<1|handout:1>{Data}
-    \frametitle<4-6|handout:4>{Pullback: data to region of the visualization} 
+    \frametitle<4|handout:4>{Pullback: data to region of the visualization} 
     \frametitle<2|handout:2>{Graphic}
-    \frametitle<7-|handout:5>{Pushforward: visualization to index of data}
+    \frametitle<5|handout:5>{Pushforward: visualization to index of data}
     \begin{equation*}
         \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-            \onslide*<0|handout:2,3,5>{\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}} 
-            \onslide*<0|handout:5>{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]} \& \& 
-            \onslide*<0|handout:5>{\cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}}  
+            \onslide*<2,3,5|handout:2,3,5>{\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}} 
+            \onslide*<5|handout:5>{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]} \& \& 
+            \onslide*<5|handout:5>{\cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}}  
             \\
-            \onslide*<0|handout:2,4->{\opensetgc}
-            \onslide*<0|handout:3>{\overbrace{\opensetc\times[0,1]^{m}}^{\opensetgc}}
-            \onslide*<0|handout:3->{\arrow[rr, "\vindex", color=functor]}
-            \onslide*<0|handout:2,3,5>{\arrow[u, "\sheaf_{\gbasec, \gtotalc}", maps to, color=sheaf]}
-            \onslide*<0|handout:4>{\arrow[d, dashed, maps to, "\sheaf_{\gbasec, \vindexpullc\dtotalc}"',color=sheaf]} \&  \& 
-            \onslide*<0|handout:1,3->{\opensetc}
-            \onslide*<0|handout:1,3-4>{\arrow[d, "\sheaf_{\dbasec, \dtotalc}", maps to, color=sheaf]}
-            \onslide*<0|handout:5>{\arrow[u, dashed, maps to, "\sheaf_{\dbasec, \vindexpushc\gtotalc}"', color=sheaf]} \\
-            \onslide*<0|handout:4>{\cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}} 
-             \& \& 
-            \onslide*<0|handout:1,3-4>{\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}}
-            \onslide*<0|handout:4>{\arrow[ll, "\textcolor{functor}{\vindexpull}", color=functor]} 
-        \end{tikzcd}
-    \end{equation*}
-    \only<0|handout:1>{% data sheaf
+            \onslide*<2,4|handout:2,4->{\opensetgc}
+            \onslide*<3|handout:3>{\overbrace{\opensetc\times[0,1]^{m}}^{\opensetgc}}
+            \onslide*<3-|handout:3->{\arrow[rr, "\vindex", color=functor]}
+            \onslide*<2,3,5|handout:2,3,5>{\arrow[u, "\sheaf_{\gbasec, \gtotalc}", maps to, color=sheaf]}
+            \onslide*<4|handout:4>{\arrow[d, dashed, maps to, "\sheaf_{\gbasec, \vindexpullc\dtotalc}"',color=sheaf]} \&  \& 
+            \onslide*<1,3-|handout:1,3->{\opensetc}
+            \onslide*<1,3-4|handout:1,3-4>{\arrow[d, "\sheaf_{\dbasec, \dtotalc}", maps to, color=sheaf]}
+            \onslide*<5|handout:5>{\arrow[u, dashed, maps to, "\sheaf_{\dbasec, \vindexpushc\gtotalc}"', color=sheaf]} \\
+            \onslide*<4|handout:4>{\cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}} 
+             \& \& 
+            \onslide*<1,3-4|handout:1,3-4>{\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}}
+            \onslide*<4|handout:4>{\arrow[ll, "\textcolor{functor}{\vindexpull}", color=functor]} 
+        \end{tikzcd}
+    \end{equation*}
+    \only<1|handout:1>{% data sheaf
         \begin{itemize}
@@
     }
-    \only<0|handout:4>{%pulled back data sheaf
+    \only<4|handout:4>{%pulled back data sheaf
         \begin{itemize}
@@
     }
-    \only<0|handout:2>{% graphic sheaf
+    \only<2|handout:2>{% graphic sheaf
         \begin{itemize}
@@
     }
-    \only<0|handout:5>{%pushed forward sheaf
+    \only<5|handout:5>{%pushed forward sheaf
         \begin{itemize}
@@
 
-\begin{frame}<presentation:1|handout:1>
+\begin{frame}<presentation:1-|handout:1->
     \frametitle<0|handout:1>{Artist: $\vartistc: \sheafc_{\dbasec, \dtotalc} \rightarrow \sheafc_{\gbasec, \gtotalc}$}
@@
         \begin{tikzcd}[ampersand replacement=\&]
-          \onslide*<0|handout:1>{\sheafc_{\gbasec, \gtotalc}} 
+          \onslide*<0|handout:1>{\sheafc_{\gbasec, \g
+          totalc}} 
           \onslide*<0|handout:4>{\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}}
@@
 
-\begin{frame}<presentation:1|handout:1>{Reachable \gsectionc?}
+\begin{frame}<presentation:1-|handout:1>{Reachable \gsectionc?}
     \begin{equation*}
         \begin{tikzcd}[ampersand replacement = \&, column sep=tiny, row sep=huge]
-            \onslide*<1-|handout:1>{\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}} 
-            \&  \& 
-            \onslide*<2-|handout:1>{\imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}}} 
-            \onslide*<2-|handout:1>{\arrow[ll, "\textcolor{set}{\mathlarger{\mathlarger{\supset}}}" description, hook', color=white]} \\ 
+            \onslide*<presentation:1-|handout:1>{\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}} 
+            \&  \& 
+            \onslide*<presentation:2-|handout:1>{\imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}}} 
+            \onslide*<presentation:2-|handout:1>{\arrow[ll, "\textcolor{set}{\mathlarger{\mathlarger{\supset}}}" description, hook', color=white]} \\ 
             \&  \& \\
-            \onslide*<1-|handout:1>{\opensetgc} 
-            \onslide*<1-|handout:1>{\arrow[uu, "\sheafc_{\gbasec,\gtotalc}", color=sheaf, maps to]} 
-            \onslide*<2-|handout:1>{\arrow[rruu, "\vartistc(\sheafc_{\dbasec,\dtotalc})" description, dashed, color=sheaf, maps to]} 
-            \&  \& 
-        \end{tikzcd}
-    \end{equation*}
-    \begin{alertblock}{Output Subtype}
+            \onslide*<presentation:1-|handout:1>{\opensetgc} 
+            \onslide*<presentation:1-|handout:1>{\arrow[uu, "\sheafc_{\gbasec,\gtotalc}", color=sheaf, maps to]} 
+            \onslide*<presentation:2-|handout:1>{\arrow[rruu, "\vartistc(\sheafc_{\dbasec,\dtotalc})" description, dashed, color=sheaf, maps to]} 
+            \&  \& 
+        \end{tikzcd}
+    \end{equation*}
+    \begin{alertblock}<presentation:2-|handout:1>{Output Subtype}
         \begin{equation*}
@@
     \end{description}
+\end{frame}
+
+\begin{frame}<presentation:1|handout:1->
+    \frametitle<presentation:1-|handout:1->{Combining Bundles}
+    \frametitle<presentation:2|handout:2>{Transforming Combined Bundles: $\dfuncc = (\dfunchc, \prod\limits_{i=0}^{n}\dfunctc_i)$}
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement = \&, column sep=2ex]
+                \onslide*<presentation:2|handout:2>{\dfunctc_0 \dfuncpullc} \dtotalc_0 
+                \arrow[rdd, "\pi" description] 
+                \onslide*<presentation:1|handout:1>{\arrow[r, "\cdot \otimes \cdot;", phantom]}
+                \onslide*<presentation:2|handout:2>{\arrow[r, "\cdots \otimes \cdots", phantom]} 
+                \& 
+                \onslide*<presentation:2|handout:2>{\dfunctc_i \dfuncpullc} \dtotalc_i 
+                \arrow[dd, "\pi" description] 
+                \arrow[r, "\cdots \otimes \cdots", phantom] \& 
+                \onslide*<presentation:2|handout:2>{\dfunctc_n \dfuncpullc} \dtotalc_n 
+                \arrow[ldd, "\pi" description] \& = \& 
+                \prod\limits_{i=0}^{n} 
+                \onslide*<presentation:2|handout:2>{\dfunctc_i \dfuncpullc} \dtotalc_i 
+                \arrow[dd, "\pi" description] \\    
+                \&\&\&\&\\    
+                \& 
+                \onslide*<presentation:2|handout:2>{(\dfunchc=\dfunchc_{i \in [0,n]})}\dbasec 
+                \& \& = \&        
+                \onslide*<presentation:2|handout:2>{\dfunchc} \dbasec                                               
+                \end{tikzcd}                                              
+    \end{equation*}
+    \onslide*<presentation:1|handout:1>{
+        \begin{equation*}
+            \prod\limits_{i=0}^{n} \underbrace{Hom(\dbasec,\dtotalc_i)}_{\dsectionc_i} = 
+            \underbrace{Hom(\dbasec,  \prod\limits_{i=0}^{n}\dtotalc_i)}_{\dsectionc}
+        \end{equation*}}
+    \onslide*<presentation:2|handout:2>{
+    \begin{equation*}
+        \prod\limits_{i=0}^{n} \underbrace{Hom(\dfunchc\dbasec, \dfunctc_i\dfuncpullc \dtotalc_i)}_{\dfuncc_i\dsectionc_i} = 
+        \underbrace{Hom(\dfunchc\dbasec,  \prod\limits_{i=0}^{n} \dfunctc_i\dfuncpullc\dtotalc_i)}_{\dfuncc\dsectionc}
+    \end{equation*}}
+
+\end{frame}
+
+\begin{frame}<presentation:1|handout:1>{Composable $\dfuncc = (\dfunchc, \prod\limits_{i=0}^{n}\dfunctc_i)$}
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement = \&]
+            \dtotalc_a 
+            \arrow[dd, "\dfunc c_a"', color=action] \& \& 
+            \dtotalc_a \times \dtotalc_b 
+            \arrow[dd, "{\dfuncc_{a,b}}" description, color=action] 
+            \arrow[ll, "\pi_a"', color=total] 
+            \arrow[rr, "\pi_b", color=total] \& \& 
+            \dtotalc_b 
+            \arrow[dd, "\dfuncc_b", color=action] \\
+            \& \& \& \& \\
+            \dtotalc_a  \& \& 
+            \dtotalc_a \times \dtotalc_b 
+            \arrow[rr, "\pi_b", color=total] 
+            \arrow[ll, "\pi_a"', color=total] \& \& 
+            \dtotalc_b
+        \end{tikzcd}
+    \end{equation*}    
 \end{frame}
@@
 
-\begin{frame}<presentation:1|handout:0>{Multiple Bundles: $\dfuncc = (\dfunchc, \prod\limits_{i=0}^{n}\dfunctc_i)$}
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement = \&, column sep=2ex]
-                \dfunctc_0 \dfuncpullc \dtotalc_0 
-                \arrow[rdd, "\pi" description] 
-                \arrow[r, "\cdots \otimes \cdots", phantom] \& 
-                \dfunctc_i  \dfuncpullc \dtotalc_i 
-                \arrow[dd, "\pi" description] 
-                \arrow[r, "\cdots \otimes \cdots", phantom] \& 
-                \dfunctc_n \dfuncpullc \dtotalc_n 
-                \arrow[ldd, "\pi" description] \& = \& 
-                \prod\limits_{i=0}^{n} 
-                \dfunctc_i \dfuncpullc \dtotalc 
-                \arrow[dd, "\pi" description] \\    
-                \&\&\&\&\\    
-                \& 
-                {(\dfunchc=\dfunchc_{i \in [0,n]})\dbasec }                                             \& \& = \& 
-                \dfunchc \dbasec                                                        
-                \end{tikzcd}                                              
-    \end{equation*}
-
-\end{frame}
-
-\begin{frame}<presentation:1|handout:0>{Complex data}
+\begin{frame}<presentation:1|handout:1>{Complex data}
     \begin{description}[style=newline]
@@
 \begin{frame}<presentation:1|handout:1>
-    \frametitle{Data to Visual Transformation: $\vchannelc: \dtotalc \rightarrow \vtotalc$}
+    \frametitle{Transformations in Data $\vchannelc: \dsectionc \mapsto \vsectionc$}
+    %rewrite in gamma notation
     \begin{equation*}
@@
 \begin{frame}{Data to Visual Transformation $\vchannelc: \dfiberc_{\dbasepointc} \mapsto \vfiberc_{\dbasepointc}$}
+    %remove maps to 
 $\pi(\dtotalc) = \pi(\vchannelc(\dtotalc))$ and $\vchannelc$ is composable s.t 
@@
             \& \& \& \& 
-            \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
-            \arrow[ddd, "\dfunctc_{\gtotalc} \circ \dfunchc_{\gtotalc}", color=action]  \\
+            \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+            \arrow[ddd, "\dfunctc_{\gtotalc} \circ \dfunchc_{\gtotalc}", dotted, color=action]  \\
             \& 
@@
             \arrow[ll, "\vindexc"', color=functor] 
-            \arrow[u, "\dfunchc_{\gtotalc}"', color=action] 
+            \arrow[u, "\dfunchc_{\gtotalc}"', dotted, color=action] 
             \arrow[rd, maps to, color=functor] \& \\
@@
             \& \&  \& \& 
-            \cgamma{\opensetgc^{\prime}}{\gtotalc^{\prime}\restriction_{\opensetgc^{\prime}}}
+            \imartist{\opensetgc^{\prime}}{\gtotalc^{\prime}\restriction_{\opensetgc^{\prime}}}
             \end{tikzcd}
@@
         \arrow[rr, "\vmarkc_{\opensetgc}", no head, Rightarrow, color=artist] \&  \& 
-        \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}  
+        \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}}  
         \arrow[dd, "\vindexpushc", color=functor] \\
@@
         \&  \& 
-        \cgamma{\opensetgc}{\vindexpushc\gtotalc\restriction_{\opensetc}}
+        \imartist{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}
     \end{tikzcd}

### 2022-03-28 slides/main.tex
artist
compiles through nu

@@
 
-\begin{frame}<presentation:1|handout:0>
+\begin{frame}<presentation:1-|handout:1>
     \frametitle{Domain specific libraries know their structure\cite{HeerSoftware2006}}
@@
 
-\begin{frame}<presentation:1|handout:0>
+\begin{frame}<presentation:1-|handout:1>
     \frametitle{General purpose libraries generally can't\cite{toryRethinkingVisualizationHighlevel2004}}
@@
             \& {} \arrow[dd, "\alpha" , Rightarrow, shorten=1.5em, color=nattran] \&             \\
-          \textcolor{source}{\mathcal{C}} \arrow[rr, "F", bend left, color=functor, maps to] \arrow[rr, "G"', bend right, color=functor, maps to] \& \& \textcolor{target}{\mathcal{D}} \\
+          \textcolor{source}{\mathcal{C}} \arrow[rr, "F", bend left, color=functor] \arrow[rr, "G"', bend right, color=functor] \& \& \textcolor{target}{\mathcal{D}} \\
             \& {} \&            
@@
         \begin{tikzcd}[ampersand replacement=\&]  
+                \&  \& 
+                \textcolor{target}{d_s} \& 
+                \textcolor{source}{c} 
+                \arrow[r, "G", maps to, color=functor] 
+                \arrow[l, "F"', maps to, color=functor] \& 
+                \textcolor{target}{d_t} \\
+                \textcolor{source}{c_1} 
+                \arrow[dd, "f" description, color=source] \& \& 
+                \textcolor{target}{F(c_1)} 
+                \arrow[dd, "F(f)"', color=target] 
+                \arrow[rr, "\alpha_{c_1}", color=nattran] \& \& 
+                \textcolor{target}{G(c_1)} 
+                \arrow[dd, "G(f)", color=target] \\
+                \& \& {} 
+                \arrow[rr, "\alpha_{f}", dotted, color=nattran] \& \& {} \\
+                \textcolor{source}{c_2} \& \& 
+                \textcolor{target}{F(c_2)} 
+                \arrow[rr, "\alpha{c_2}", color=nattran] \& \& 
+                \textcolor{target}{G(c_2)} \\
+                \& \& 
+                \textcolor{target}{F} 
+                \arrow[rr, "\alpha", Rightarrow, color=nattran] \& \& 
+                \textcolor{target}{G}                        
         \end{tikzcd}                    
@@
             \\
-            \onslide*<2,4|handout:2,4->{\opensetgc}
+            \onslide*<2,4-|handout:2,4->{\opensetgc}
             \onslide*<3|handout:3>{\overbrace{\opensetc\times[0,1]^{m}}^{\opensetgc}}
@@
 \begin{frame}<presentation:1-|handout:1->
-    \frametitle<0|handout:1>{Artist: $\vartistc: \sheafc_{\dbasec, \dtotalc} \rightarrow \sheafc_{\gbasec, \gtotalc}$}
-    \frametitle<0|handout:2>{Data Space: $\vartistc_{\opensetc}: \sheafc_{\dbasec, \dtotalc} \textcolor{artist}{\Rightarrow} \vindexpushc\sheafc_{\gbasec, \gtotalc}$}
-    \frametitle<0|handout:3>{Display Space: $\vartist_{\opensetgc}: \vindexpullc\sheafc_{\dbasec, \dtotalc} \textcolor{artist}{\Rightarrow} \sheafc_{\gbasec, \gtotalc}$}
-    \frametitle<0|handout:4>{Artist: $\vartistc: \cgamma{\opensetc}{\dtotalc} \rightarrow \cgamma{\opensetgc}{\gtotalc}$}
-    \onslide*<0|handout:1,4>{
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement=\&]
-          \onslide*<0|handout:1>{\sheafc_{\gbasec, \g
-          totalc}} 
-          \onslide*<0|handout:4>{\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}}
-          \onslide*<0|handout:1,4>{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]}  
-          \& \& 
-          \onslide*<0|handout:1>{\vindexpushc\sheafc_{\gbasec, \gtotalc}}
-          \onslide*<0|handout:4>{\cgamma{\opensetc}{\vindexpushc \gtotalc\restriction_{\opensetc}}}
+    \frametitle<1|handout:1>{$\textcolor{homset}{Hom}_{\gbasec}(\sheaf_{\gbasec, \vindexpullc\dtotalc}, \sheaf_{\gbasec,\gtotalc}) = \textcolor{homset}{Hom}_{\dbasec}(\sheafc_{\dbasec, \dtotalc}, \sheafc_{\dbasec, \vindexpushc\gtotalc})$}
+    \frametitle<2|handout:2>{Data Space: $\vartistc_{\dbasec}: \sheafc_{\dbasec, \dtotalc} \textcolor{artist}{\Rightarrow} \sheafc_{\dbasec, \vindexpushc\gtotalc}$}
+    \frametitle<3|handout:3>{Display Space: $\vartist_{\gbasec}: \sheafc_{\gbasec, \vindexpullc\dtotalc} \textcolor{artist}{\Rightarrow} \sheafc_{\gbasec, \gtotalc}$}
+    \frametitle<4|handout:4>{Artist: $\vartistc: \dsectionc \mapsto \gsectionc$}
+
+    \onslide*<1,4|handout:1,4>{
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement=\&]
+          \onslide*<1|handout:1>{\sheafc_{\gbasec, \gtotalc}} 
+          \onslide*<4|handout:4>{\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}}
+          \onslide*<1,4|handout:1,4>{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]}  
+          \& \& 
+          \onslide*<1|handout:1>{\sheafc_{\dbasec, \vindexpushc\gtotalc}}
+          \onslide*<4|handout:4>{\cgamma{\opensetc}{\vindexpushc \gtotalc\restriction_{\opensetc}}}
           \\
           \& \& \\
-          \onslide*<0|handout:1>{\textcolor{functor}{\vindexpull}\sheafc_{\dbasec, \dtotalc}}
-          \onslide*<0|handout:4>{\cgamma{\opensetgc}{\vindexpullc \dtotalc\restriction_{\opensetgc}}}
-          \onslide*<0|handout:1,4>{\arrow[uu, "\vartistc_{\opensetgc}", color=artist, Rightarrow]}
-          \&  \& 
-          \onslide*<0|handout:1>{\sheafc_{\dbasec, \dtotalc}}
-          \onslide*<0|handout:4>{\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}}
-          \onslide*<0|handout:1,4>{\arrow[ll, "\textcolor{functor}{\vindexpull}", color=functor]} 
-          \onslide*<0|handout:1,4>{\arrow[uu, "\vartistc_{\opensetc}"', color=artist, Rightarrow]}
-          \onslide*<4-|handout:1,4>{\arrow[lluu, "\vartistc"' description, dashed, color=artist, Rightarrow]}
+          \onslide*<1|handout:1>{\sheafc_{\gbasec, \vindexpullc\dtotalc}}
+          \onslide*<4|handout:4>{\cgamma{\opensetgc}{\vindexpullc \dtotalc\restriction_{\opensetgc}}}
+          \onslide*<1|handout:1>{\arrow[uu, "\vartistc_{\gbasec} \in \textcolor{homset}{Hom}_{\gbasec}", color=homset]}
+          \onslide*<4|handout:4>{\arrow[uu, "\vartistc_{\gbasec}", color=artist]}
+          \&  \& 
+          \onslide*<1|handout:1>{\sheafc_{\dbasec, \dtotalc}}
+          \onslide*<4|handout:4>{\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}}
+          \onslide*<1,4|handout:1,4>{\arrow[ll, "\textcolor{functor}{\vindexpull}", color=functor]} 
+          \onslide*<1|handout:1>{\arrow[uu, "\vartistc_{\dbasec} \in \textcolor{homset}{Hom}_{\dbasec}"', color=homset]}
+          \onslide*<4|handout:4>{\arrow[uu, "\vartistc_{\dbasec}"', color=artist]}
+          \onslide*<1|handout:1>{\arrow[lluu, "\vartistc \in \textcolor{homset}{Hom}_{\dbasec,\gbasec}"' description, dashed, color=homset]}
+          \onslide*<4|handout:4>{\arrow[lluu, "\vartistc"' description, color=artist]}
         \end{tikzcd}
@@
 
-    \onslide*<1|handout:2>{
-        \begin{equation*}
-      \begin{tikzcd}[ampersand replacement =\&]
-        \cgamma{\opensetc_1}{\dtotalc} 
-        \arrow[dd, "\iota^*"', hook, color=set] 
-        \arrow[rrrr, "\vartistc_{\opensetc_1}", bend left, color=artist, Rightarrow] \&  \& 
-        \opensetc_1 
-        \arrow[ll, "\sheafc_{\dbasec,\dtotalc}", color=sheaf, maps to] 
-        \arrow[rr, "\sheafc_{\dbasec,\gtotalpushc}"', color=sheaf, maps to] \&  \& 
-        \cgamma{\opensetc_1}{\gtotalpushc} 
-        \arrow[dd, "\iota^*", hook', color=set] \\
-        {} 
-        \arrow[rrrr, dotted, color=artist,Rightarrow] \&\&\&\& {} \\
-        \cgamma{\opensetc_2}{\dtotalc} 
-        \arrow[rrrr, "\vartistc_{\opensetc_2}"', bend right, color=artist, Rightarrow] \&  \& 
-        \opensetc_2 
-        \arrow[ll, "\sheafc_{\dbasec,\dtotalc}"', color=sheaf, maps to] 
-        \arrow[rr, "\sheafc_{\dbasec,\gtotalpushc}", color=sheaf, maps to] 
-        \arrow[uu, "\iota" description, hook, color=base] \& \& 
-        \cgamma{\opensetc_2}{\gtotalpushc}               
+    \onslide*<2|handout:2>{
+        \begin{equation*}
+            \begin{tikzcd}[ampersand replacement=\&]  
+                    \&  \& 
+                    \sheafc_{\dbasec, \dtotalc} 
+                    \arrow[rr, "\vartistc_{\dbasec}", Rightarrow, color=artist] \& \& 
+                    \sheafc_{\dbasec, \vindexpushc\gtotalc}  \\
+                    \opensetc_1\& \& 
+                    \cgamma{\opensetc_1}{\dtotalc\restriction_{\opensetc_1}}
+                    \arrow[dd, "\iota^*"', color=set] 
+                    \arrow[rr, "\vartistc_{\opensetc_1}", color=artist] \& \& 
+                    \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc_1}} 
+                    \arrow[dd, "\iota^*", color=set] \\
+                    \& \& {} 
+                    \arrow[rr, "\vartistc_{\sheafc}", dotted, color=artist] \& \& {} \\
+                    \arrow[uu, "\iota" description, hook, color=base] 
+                    \opensetc_2 \& \& 
+                    \cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}
+                    \arrow[rr, "\vartistc_{\opensetc_2}", color=artist] \& \& 
+                    \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc_2}} \\
+                    \& \& 
+                    \cgamma{\opensetc}{\dtotalc\restriction\opensetc} \& 
+                    \opensetc 
+                    \arrow[r, "\sheafc_{\dbasec,\vindexpushc\gtotalc}", maps to, color=sheaf] 
+                    \arrow[l, "\sheafc_{\dbasec, \dtotalc}"', maps to, color=sheaf] \& 
+                    \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction\opensetc}
             \end{tikzcd}                    
@@
 
-    \onslide*<0|handout:3>{
-        \begin{equation*}
-    \begin{tikzcd}[ampersand replacement =\&]
-    \cgamma{\opensetgc_1}{\dtotalpullc\restriction_{\opensetgc_1}} 
-    \arrow[dd, "\iota^*"', hook, color=set] 
-    \arrow[rrrr, "\vartistc_{\opensetgc_1}", bend left, color=artist, Rightarrow] \&  \& 
-    \opensetgc_1 
-    \arrow[ll, "\sheafc_{\gbasec, \dtotalpullc}", color=sheaf, maps to] 
-    \arrow[rr, "\sheafc_{\gbasec,\gtotalc}"', color=sheaf, maps to] \&  \& 
-    \cgamma{\opensetgc_1}{\gtotalc\restriction_{\opensetgc_1}} 
-    \arrow[dd, "\iota^*", hook', color=set] \\
-    {} 
-    \arrow[rrrr, dotted, color=artist, Rightarrow] \&\&\&\& {} \\
-    \cgamma{\opensetgc_2}{\dtotalpullc\restriction_{\opensetgc_2}} 
-    \arrow[rrrr, "\vartistc_{\opensetgc_2}"', bend right, color=artist, Rightarrow] \&  \& 
-    \opensetgc_2 
-    \arrow[ll, "\sheafc_{\gbasec,\dtotalpullc}"', color=sheaf, maps to] 
-    \arrow[rr, "\sheafc_{\gbasec, \gtotalc}", color=sheaf, maps to] 
-    \arrow[uu, "\iota" description, hook, color=base] \& \& 
-    \cgamma{\opensetgc_2}{\gtotalc\restriction_{\opensetgc_2}}   
-    \end{tikzcd}
-        \end{equation*}}    
-    \onslide<0|handout:1>{
-        \begin{equation*}
-        \textcolor{nattran}{Nat}_{\opensetgc}(\textcolor{functor}{\vindexpull}\sheafc_{\dbasec, \dtotalc}, \sheafc_{\gbasec, \gtotalc}) = 
-    \textcolor{nattran}{Nat}_{\opensetc}(\sheafc_{\dbasec, \dtotalc}, \textcolor{functor}{\vindexpush}\sheafc_{\gbasec, \gtotalc})
-        \end{equation*}
-    }
-    \onslide<0|handout:4>{
+    \onslide*<3|handout:3>{
+        \begin{equation*}
+            \begin{tikzcd}[ampersand replacement=\&]  
+                    \&  \& 
+                    \sheafc_{\gbasec, \vindexpullc\dtotalc} 
+                    \arrow[rr, "\vartistc_{\gbasec}", Rightarrow, color=artist] \& \& 
+                    \sheafc_{\gbasec, \gtotalc}  \\
+                    \opensetgc_1\& \& 
+                    \cgamma{\opensetgc_1}{\vindexpullc\dtotalc\restriction_{\opensetgc_1}}
+                    \arrow[dd, "\iota^*"', color=set] 
+                    \arrow[rr, "\vartistc_{\opensetgc_1}", color=artist] \& \& 
+                    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc_1}} 
+                    \arrow[dd, "\iota^*", color=set] \\
+                    \& \& {} 
+                    \arrow[rr, "\vartistc_{\sheafc}", dotted, color=artist] \& \& {} \\
+                    \arrow[uu, "\iota" description, hook, color=base] 
+                    \opensetc_2 \& \& 
+                    \cgamma{\opensetgc_2}{\vindexpullc\dtotalc\restriction_{\opensetgc_2}}
+                    \arrow[rr, "\vartistc_{\opensetgc_2}", color=artist] \& \& 
+                    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc_2}} \\
+                    \& \& 
+                    \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction\opensetgc} \& 
+                    \opensetgc 
+                    \arrow[r, "\sheafc_{\gbasec,\gtotalc}", maps to, color=sheaf] 
+                    \arrow[l, "\sheafc_{\gbasec, \vindexpullc\dtotalc}"', maps to, color=sheaf] \& 
+                    \cgamma{\opensetc}{\gtotalc\restriction\opensetgc}
+            \end{tikzcd}                    
+        \end{equation*}
+    }
+    \onslide<4|handout:4>{
         \begin{alertblock}{pull data section \dsectionc\ over graphic space $\gbasepointc \in \gbasec$}
@@
 \begin{frame}<presentation:1|handout:1>{$\dfuncc = (\dfunchc, \dfunctc)$}
-    \begin{tikzcd}[ampersand replacement = \&]
-        \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
-        \arrow[rr, "\dfuncpullc", color=action] 
-        \&  \& 
-        \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}} 
-        \arrow[rr, "\dfunctc", color=action] 
-        \&  \& 
-        \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}} \\
-        \&  \&  
-        \&  \& \\
-        \opensetc 
-        \arrow[uu, "{\sheafc_{\dbasec, \dtotalc}}", maps to, color=sheaf]                 
-        \&  \& 
-        \opensetc^{\prime} 
-        \arrow[ll, "\dfunchc", color=action] 
-        \arrow[uu, "{\sheafc_{\dbasec, \dfuncpullc\dtotalc}}"', maps to, color=sheaf] 
-        \arrow[rruu, "{\sheafc_{\dbasec, \dfuncpullc\dtotalc}}"', maps to, color=sheaf] 
-        \&  \& 
-    \end{tikzcd}
+    \begin{figure}
+        \includegraphics[width=0.5\textwidth]{figures/math/dfunc.png}
+    \end{figure}
     \begin{description}
@@
 
-        \item [Fiber Transformation:]{$\dfunctc: \dfuncpullc \dtotalc_{\dbasepointc^{\prime}} \rightarrow {\dfuncpullc\dtotalc_{\dbasepointc^{\prime}}} \in Hom(\dfuncpullc\dfiberc\restriction_{\dbasepointc}, \dfuncpullclllllll\dfiberc\restriction_{\dbasepointc}), \dbasepointc^{\prime} \in \opensetc^{\prime}$\\
+        \item [Fiber Transformation:]{$\dfunctc: \dfuncpullc \dtotalc_{\dbasepointc^{\prime}} \rightarrow {\dfuncpullc\dtotalc_{\dbasepointc^{\prime}}} \in Hom(\dfuncpullc\dfiberc\restriction_{\dbasepointc}, \dfuncpullc\dfiberc\restriction_{\dbasepointc}), \dbasepointc^{\prime} \in \opensetc^{\prime}$\\
         $\dfunctc: \dfuncpullc \dsectionc\restriction_{\opensetc} \mapsto \dfuncpullc \dsectionc^{\prime}\restriction_{\opensetc}$, $\dsectionc, \dsectionc^{\prime} \in \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}}$}
@@
     \end{description}
-\end{frame}
-
-\begin{frame}<presentation:1|handout:1->
-    \frametitle<presentation:1-|handout:1->{Combining Bundles}
-    \frametitle<presentation:2|handout:2>{Transforming Combined Bundles: $\dfuncc = (\dfunchc, \prod\limits_{i=0}^{n}\dfunctc_i)$}
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement = \&, column sep=2ex]
-                \onslide*<presentation:2|handout:2>{\dfunctc_0 \dfuncpullc} \dtotalc_0 
-                \arrow[rdd, "\pi" description] 
-                \onslide*<presentation:1|handout:1>{\arrow[r, "\cdot \otimes \cdot;", phantom]}
-                \onslide*<presentation:2|handout:2>{\arrow[r, "\cdots \otimes \cdots", phantom]} 
-                \& 
-                \onslide*<presentation:2|handout:2>{\dfunctc_i \dfuncpullc} \dtotalc_i 
-                \arrow[dd, "\pi" description] 
-                \arrow[r, "\cdots \otimes \cdots", phantom] \& 
-                \onslide*<presentation:2|handout:2>{\dfunctc_n \dfuncpullc} \dtotalc_n 
-                \arrow[ldd, "\pi" description] \& = \& 
-                \prod\limits_{i=0}^{n} 
-                \onslide*<presentation:2|handout:2>{\dfunctc_i \dfuncpullc} \dtotalc_i 
-                \arrow[dd, "\pi" description] \\    
-                \&\&\&\&\\    
-                \& 
-                \onslide*<presentation:2|handout:2>{(\dfunchc=\dfunchc_{i \in [0,n]})}\dbasec 
-                \& \& = \&        
-                \onslide*<presentation:2|handout:2>{\dfunchc} \dbasec                                               
-                \end{tikzcd}                                              
-    \end{equation*}
-    \onslide*<presentation:1|handout:1>{
-        \begin{equation*}
-            \prod\limits_{i=0}^{n} \underbrace{Hom(\dbasec,\dtotalc_i)}_{\dsectionc_i} = 
-            \underbrace{Hom(\dbasec,  \prod\limits_{i=0}^{n}\dtotalc_i)}_{\dsectionc}
-        \end{equation*}}
-    \onslide*<presentation:2|handout:2>{
-    \begin{equation*}
-        \prod\limits_{i=0}^{n} \underbrace{Hom(\dfunchc\dbasec, \dfunctc_i\dfuncpullc \dtotalc_i)}_{\dfuncc_i\dsectionc_i} = 
-        \underbrace{Hom(\dfunchc\dbasec,  \prod\limits_{i=0}^{n} \dfunctc_i\dfuncpullc\dtotalc_i)}_{\dfuncc\dsectionc}
-    \end{equation*}}
-
 \end{frame}
@@
         \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
-        \arrow[rrr, "\vartistc", Rightarrow, color=artist] 
+        \arrow[rrr, "\vartistc", color=artist] 
         \arrow[d, "\dfuncpullc_{\dtotalc}"', color=action] 
@@
         \opensetc 
-        \arrow[lu, dashed, maps to, color=sheaf] \& 
+         \& 
         \opensetgc 
         \arrow[l, "\vindexc"', color=functor] 
-        \arrow[ru, dashed, maps to, color=sheaf] \& 
+        \& 
         \imartist{\opensetgc^{\prime}}{\dfuncpullc_{\gtotalc}\gtotalc\restriction_{\opensetgc^{\prime}}} 
@@
         \arrow[u, "\dfunchc_{\dtotalc}", color=action] 
-        \arrow[ld, dashed, maps to, color=sheaf] 
-        \arrow[lu, dashed, maps to, color=sheaf] \& 
+        \& 
         \opensetgc^{\prime} 
@@
         \arrow[u, "\dfunchc_{\gtotalc}"', dotted, color=action] 
-        \arrow[rd, dashed, maps to, color=sheaf] 
-        \arrow[ru, dashed, maps to, color=sheaf] \& \\
+        \& \\
         \cgamma{\opensetc^{\prime}}{\dtotalc^{\prime}\restriction_{\opensetc^{\prime}}} 
-        \arrow[rrr, "\vartistc^{\prime}", Rightarrow, color=artist]  
+        \arrow[rrr, "\vartistc^{\prime}", color=artist]  
         \& \& \& 
@@
     \end{equation*}
-    %$\vartistc^{\prime}(\dfunctc_{\dtotalc}(\dsectionc(\dfunchc_{\dtotalc}(\vindexc(\gbasepointc^{\prime}))))) = \dfunctc_{\gtotalc}(\vartistc(\dsectionc(\vindexc(\dfunchc_{\gtotalc}(\gbasepointc^{\prime})))))$
-    %for all points $\gbasepointc^{\prime}\in \gbasec^{\prime}$
+    For all points $\gbasepointc^{\prime}\in \gbasec^{\prime}$:
+    \begin{equation*}
+        \vartistc^{\prime}(\dfunctc_{\dtotalc}(\dsectionc(\dfunchc_{\dtotalc}(\vindexc(\gbasepointc^{\prime}))))) = \dfunctc_{\gtotalc}(\vartistc(\dsectionc(\vindexc(\dfunchc_{\gtotalc}(\gbasepointc^{\prime})))))
+    \end{equation*}
+
 \end{frame}
@@
     \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
-    \arrow[rr, "\vartistc", Rightarrow, color=artist] 
+    \arrow[rr, "\vartistc", color=artist] 
     \arrow[dd, "\equivc"', color=monoid] \&  \& 
@@
 
-\begin{frame}<presentation:0|handout:->{Visual Measurement}
+\begin{frame}<presentation:1-|handout:->{Visual Measurement}
     \begin{columns}
@@
     \column{.5\textwidth}
-    \onslide*<presentation:0|handout:2>{
+    \onslide*<presentation:2|handout:2>{
     \begin{equation*}
@@
     \end{columns}
-    \onslide*<presentation:0|handout:1>{
+    \onslide*<presentation:1|handout:1>{
     $\extractmc_{\gsectionc} \in Hom(\opensetgc, \measurec)$ is a mapping from graphic to visual measurement.
 $\extractmc_{\gsectionc}$ is a map from an openset $\opensetgc \subseteq \gbasec$ to a measurement $\measurec_{\opensetgc}$ that corresponds to the graphic representation at that region $\gsectionc\restriction_{\opensetgc}$ and the corresponding data $\dsectionc\restriction_{\vindexc\restriction_{\opensetgc}}$.}
-    \onslide*<presentation:0|handout:2>{
+    \onslide*<presentation:2|handout:2>{
     $\equivc_{\dsectionc} \in Hom(\opensetc, \measurec)$ is a mapping from data to visual measurement.
@@
 
-\begin{frame}<presentation:1|handout:1->
-    \frametitle<presentation:0|handout:1>{Using output $\gsectionc$ to check if $\vartistc$ is equivariant}
-    \frametitle<presentation:0|handout:2>{Using input $\dsectionc$ to check if $\vartistc$ is equivariant}
+\begin{frame}<presentation:1-|handout:1->
+    \frametitle<presentation:1|handout:1>{Using output $\gsectionc$ to check if $\vartistc$ is equivariant}
+    \frametitle<presentation:2|handout:2>{Using input $\dsectionc$ to check if $\vartistc$ is equivariant}
     \begin{equation*}
@@
             \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
-            \arrow[rr, "\vartistc", Rightarrow, color=artist] 
+            \arrow[rr, "\vartistc", color=artist] 
             \arrow[d, "\dfuncc_{\dtotalc}"', color=action] 
-            \onslide*<presentation:0|handout:2>{\arrow[rrr, "\equivc", bend left, color=monoid]} \&  \& 
+            \onslide*<presentation:2|handout:2>{\arrow[rrr, "\equivc", bend left, color=monoid]} \&  \& 
             \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
             \arrow[d, "\dfuncc_{\gtotalc}"', dotted, color=action]
-            \onslide*<presentation:0|handout:1>{\arrow[r, "\extractmc", color=monoid]}
-            \onslide*<presentation:0|handout:2>{\arrow[r,"\vindexprec \circ \extractmc", color=monoid]}
-            \& 
-            \onslide*<presentation:0|handout:1>{Hom(\opensetgc, \measurec)}
-            \onslide*<presentation:0|handout:2>{Hom(\opensetc, \measurec)} 
+            \onslide*<presentation:1|handout:1>{\arrow[r, "\extractmc", color=monoid]}
+            \onslide*<presentation:2|handout:2>{\arrow[r,"\vindexprec \circ \extractmc", color=monoid]}
+            \& 
+            \onslide*<presentation:1|handout:1>{Hom(\opensetgc, \measurec)}
+            \onslide*<presentation:2|handout:2>{Hom(\opensetc, \measurec)} 
             \arrow[d, "\dfuncc_{\measurec}", color=action] \\
             \cgamma{\opensetc^{\prime}}{\dtotalc^{\prime}\restriction_{\opensetc^{\prime}}} 
-            \arrow[rr, "\vartistc", Rightarrow, color=artist] 
-            \onslide*<presentation:0|handout:2>{\arrow[rrr, "\equivc"', bend right, color=monoid]} \&  \& 
+            \arrow[rr, "\vartistc", color=artist] 
+            \onslide*<presentation:2|handout:2>{\arrow[rrr, "\equivc"', bend right, color=monoid]} \&  \& 
             \imartist{\opensetgc^{\prime}}{\gtotalc^{\prime}\restriction_{\opensetgc^{\prime}}} 
-            \onslide*<presentation:0|handout:1>{\arrow[r, "\extractmc", color=monoid]}
-            \onslide*<presentation:0|handout:2>{\arrow[r,"\vindexprec \circ \extractmc", color=monoid]} \& 
-            \onslide*<presentation:0|handout:1>{Hom(\opensetgc^{\prime}, \measurec^{\prime})}
-            \onslide*<presentation:1|handout:2>{Hom(\opensetc^{\prime}, \measurec^{\prime})}                        
-        \end{tikzcd}
-    \end{equation*}
-    \onslide*<presentation:0|handout:1>{
+            \onslide*<presentation:1|handout:1>{\arrow[r, "\extractmc", color=monoid]}
+            \onslide*<presentation:2|handout:2>{\arrow[r,"\vindexprec \circ \extractmc", color=monoid]} \& 
+            \onslide*<presentation:1|handout:1>{Hom(\opensetgc^{\prime}, \measurec^{\prime})}
+            \onslide*<presentation:2|handout:2>{Hom(\opensetc^{\prime}, \measurec^{\prime})}                        
+        \end{tikzcd}
+    \end{equation*}
+    \onslide*<presentation:1|handout:1>{
         \begin{align*}
@@
     }
-    \onslide*<presentation:0|handout:2>{
+    \onslide*<presentation:2|handout:2>{
         \begin{description}
@@
 \end{frame}
+\end{comment}
 \end{document}

### 2022-03-29 slides/main.tex
flipped arrows

@@
 \begin{document}
-
-\title{Topological Equivariant Artist Model}
+\begin{frame}<presentation:1-|handout:0>{About}
+    Slides: https://bit.ly/team329
+    \begin{block}{Me (twitter/github: story645)}
+        \begin{itemize}
+            \item nth year grad student (on my 3rd EO)
+            \item former adjunct at CCNY, former Digital Fellow
+            \item Matplotlib Community Manager \& Core Developer
+        \end{itemize}
+    \end{block}
+
+    \begin{block}<presentation:2|handout:0>{Project}
+        \begin{itemize}
+            \item Funded by Chan Zuckerberg Initiative EOSS 1 \& 3
+            \item paper rejected by vizweek last spring
+            \item work has since gone all in on category theory 
+        \end{itemize}
+    \end{block}
+\end{frame}
 
@@
 \section{Introduction}
-\begin{frame}<presentation:1|handout:0>{What are we doing?}
+\begin{frame}<presentation:0|handout:0>{What are we doing?}
     \begin{itemize}
@@
 
-\begin{frame}<presentation:1|handout:0>{Case Study: Matplotlib}
+\begin{frame}<presentation:1|handout:0>{Matplotlib}
     \begin{figure}
@@
 
-\begin{frame}<presentation:1|handout:0>{Visualizations Preserve Structure}
-    \begin{description}
-        \item [continuity] topological properties \cite{wilkinsonGrammarGraphics2005}, i.e.
-how elements in a dataset are organized, e.g.
-discrete rows in a table, networked nodes, pixels in an image, points on a line
-        \item [equivariance] data and visual encodings are matched such that transformations have an equivalent effect on data and graphical representations, e.g.
-rotating a matrix and image, shifting  points on a line and a line graphic
-      \end{description}
-\end{frame}
-
-\begin{frame}<presentation:1|handout:0>{Continuity}
-    \begin{figure}
-        \includegraphics[width=\linewidth]{../paper/figures/k_different_types.png}
-    \end{figure}
-\end{frame}
-
-\begin{frame}<presentation:1|handout:0>{Expressed in container} 
+\begin{frame}<presentation:1-|handout:0>
+    \frametitle<presentation:1-2|handout:0>{Everything is an Artist}
+    \frametitle<presentation:3|handout:0>{Transformations Change Artists into Different Coordinates}
+    \begin{overprint}
+        \onslide<1|handout:0>\includegraphics[scale=.25]{figures/math/sphx_glr_anatomy_001_2_0x.png}
+        \onslide<2|handout:0>\includegraphics[width=\textwidth]{figures/math/artist_tree.png}
+        \onslide<3|handout:0>\includegraphics[width=\textwidth]{figures/math/transforms.png}
+    \end{overprint}
+\end{frame}
+
+\begin{frame}[fragile]{A Simple Artist}
+\begin{minted}{python}
+    class SomeArtist(Artist):
+        'An example Artist that implements the draw method'
+
+        def draw(self, renderer):
+            # create some objects and use renderer to draw self here
+            renderer.draw_path(graphics_context, path, transform)
+\end{minted}
+\end{frame}
+
+\begin{frame}<presentation:1-|handout:0>{Goals}
+    \begin{figure}
+        \includegraphics[width=\linewidth]{figures/flow/artists.png}
+    \end{figure}
+\end{frame}
+
+\begin{frame}<presentation:1|handout:0>{How do we express structure?} 
     \begin{figure}
@@
 
-\begin{frame}<presentation:1|handout:0>{Why is container type not enough?}
-    \includegraphics[width=1\linewidth]{../paper/figures/whycontinuity.png}
+\begin{frame}<presentation:1|handout:0>{Continuity}
+    \begin{figure}
+        \includegraphics[width=\linewidth]{../paper/figures/k_different_types.png}
+    \end{figure}
+    \begin{block}{Topological Properties \cite{wilkinsonGrammarGraphics2005}}
+        how elements in a dataset are organized, e.g. discrete rows in a table, networked nodes, pixels in an image, points on a line
+    \end{block}
+\end{frame}
+
+\begin{frame}<presentation:1|handout:0>{Visual Algorithms \& Continuity}
+    \begin{figure}
+        \includegraphics[width=1\linewidth]{figures/math/plottypes.png}
+    \end{figure}
 \end{frame}
@@
     \begin{description}
-        \item[Retinal Variables \& Marks] visual encodings should match properties of the data \cite{bertinSemiologyGraphicsDiagrams2011a}
+        \item[What] Retinal Variables \& Marks: visual encodings should match properties of the data \cite{bertinSemiologyGraphicsDiagrams2011a}
         \pause
-        \item[Graphical Integrity] graphs show \textbf{only} the data\cite{tufteVisualDisplayQuantitative2001}
+        \item[Why] Graphical Integrity: graphs show \textbf{only} the data\cite{tufteVisualDisplayQuantitative2001}
         \pause 
-        \item[Naturalness] easier to understand when properties match\cite{norman_things_smart}
+        \item[Why] Naturalness: easier to understand when properties match\cite{norman_things_smart}
         \pause
-        \item[Expressiveness] which structure preserving mappings can a tool implement\cite{mackinlayAutomatingDesignGraphical1986}] 
+        \item[How] Expressiveness: which structure preserving mappings can a tool implement\cite{mackinlayAutomatingDesignGraphical1986}] 
     \end{description}
@@
 %flip columns
-\begin{frame}<presentation:1-|handout:0>{Frameworks for Expressing Visual Equivariance}
+\begin{frame}<presentation:0-|handout:0>{Theoretical Frameworks}
     \begin{description}
@@
         \end{columns} 
-    \item[categorically] \textit{understanding} = \textit{read} $\circ$ \textit{render} \cite{vickersUnderstandingVisualizationFormal2013}    
+    %\item[categorically] \textit{understanding} = \textit{read} $\circ$ \textit{render} \cite{vickersUnderstandingVisualizationFormal2013}    
     \end{description}
@@
 \begin{frame}<presentation:1-|handout:1>
-    \frametitle{Domain specific libraries know their structure\cite{HeerSoftware2006}}
+    \frametitle{Domain Specific Library: library assumes structure \cite{HeerSoftware2006}}
     \begin{table}
@@
 \begin{frame}<presentation:1-|handout:1>
-    \frametitle{General purpose libraries generally can't\cite{toryRethinkingVisualizationHighlevel2004}}
+    \frametitle{Building Block Library\cite{wongsuphasawatNavigatingWideWorld2021}: visual algorithms assume structure \cite{toryRethinkingVisualizationHighlevel2004}}
     \begin{figure}
@@
 \section{Algebraic Topology & Category Theory}
-\begin{frame}<presentation:1|handout:1>{Mathematical Data Abstraction}
+\begin{frame}<presentation:1|handout:1>{Design Composable Structure Preserving API}
     \begin{description}
@@
                \onslide*<7|handout:4>{\gbasec} 
-               \onslide*<5-6|handout:1-2>{\arrow[u, "\dsectionc \in \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}"', bend right, pos=.5]}
-               \onslide*<7|handout:4>{\arrow[u, "\gsectionc \in \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}"', bend right, pos=.5]}
+               \onslide*<5-6|handout:1-2>{\arrow[u, "\dsectionc"', bend right, pos=.5]}
+               \onslide*<7|handout:4>{\arrow[u, "\gsectionc"', bend right, pos=.5]}
             \end{tikzcd}
@@
     \end{columns}
-    \pause 
-    \only<5|handout:1>{
+    \only<4-5|handout:1>{
         \begin{alertblock}{}
             \begin{description}
-                \item[\textcolor{section}{Sections}]{$\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}$}
+                \only<5:handout:1>{\item[\textcolor{section}{Sections}]{$\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}$}}
+            \end{description}
+        \end{alertblock}
+    }
+    \only<4:handout:1>{
+        \begin{alertblock}{}
+    \begin{description}
     \item[Locally Trivial] {for every point $\dbasepointc \in \dbasec$, there exists an open neighborhood $\dbasepointc \in \opensetc \subseteq \dbasec$ s.t.
@@
 
-\begin{frame}<presentation:1-|handout:1>{Category $\textcolor{source}{\mathcal{C}}$}
+\begin{frame}<presentation:1-|handout:1->
+    \frametitle<presentation:1-6|handout:1>{Category \setc}
+    \frametitle<presentation:7|handout:2>{Category \setc^{op}}
 \begin{columns}
@@
     \begin{tikzcd}[ampersand replacement=\&]
-        \onslide*<1-|handout:1>{\textcolor{source}{C_1}} 
-        \onslide*<3-|handout:1>{\arrow[r, "f", color=source]}
-        \onslide*<4-|handout:1>{\arrow[rd, "g\circ f"', color=source]}
-        \onslide*<2-|handout:1>{\arrow["id_{C_1}"', loop, distance=2em, in=215, out=145, color=source]} \& 
-        \onslide*<1-|handout:1>{\textcolor{source}{C_2}}
-        \onslide*<3-|handout:1>{\arrow[d, "g", color=source]} 
-        \onslide*<2-|handout:1>{\arrow["id_{C_2}"', loop, distance=2em, in=125, out=55, color=source]} \\\& 
-        \onslide*<1-|handout:1>{\textcolor{source}{C_3}}
-        \onslide*<2-|handout:1>{\arrow["id_{C_3}"', loop, distance=2em, in=305, out=235, color=source]}              
+        \onslide*<1-|handout:1>{\setc_1} 
+        \onslide*<3-6|handout:1>{\arrow[r, "r_{1,2}", color=set]}
+        \onslide*<4-6|handout:1>{\arrow[rd, "r_{1,2}\circ r_{2,3}"', color=set]}
+        \onslide*<2-6|handout:1>{\arrow["id_{\setc_1}"', loop, distance=2em, in=215, out=145, color=set]} \& 
+        \onslide*<1-|handout:1>{\setc_2}
+        \onslide*<7|handout:2>{\arrow[l, color=set]}
+        \onslide*<3-6|handout:1>{\arrow[d, "r_{2,3}", color=set]} 
+        \onslide*<2-6|handout:1>{\arrow["id_{\setc_2}"', loop, distance=2em, in=125, out=55, color=set]} \\\& 
+        \onslide*<1-|handout:1>{\setc_3}
+        \onslide*<7|handout:2>{\arrow[u, color=set]}
+        \onslide*<7|handout:2>{\arrow[lu, color=set]}
+        \onslide*<2-6|handout:1>{\arrow["id_{\setc_3}"', loop, distance=2em, in=305, out=235, color=set]}              
     \end{tikzcd}
@@
     \column{0.59\textwidth}
-    \begin{alertblock}<5-|handout:1>{associativity} 
-        if $f: C_1 \rightarrow C_2$, $g: C_2 \rightarrow C_3$ and $h: C_3 \rightarrow C_4$ then $h\circ (g \circ f) = (h \circ g) \circ f$
-    \end{alertblock}
-    \begin{alertblock}<6-|handout:1>{identity} 
-        for every $f: C_1 \rightarrow C_2$ there exists identity morphisms $f \circ id_{C_1} = f = id_{C_2} \circ f$
+    \begin{alertblock}<5-6|handout:1>{associativity} 
+        if $r_{1,2}: \setc_1 \rightarrow \setc_2$, $r_{2,3}: \setc_2 \rightarrow \setc_3$ and $r_{3,4}: \setc_3 \rightarrow \setc_4$ then $r_{3,4}\circ (r_{2,3} \circ r_{1,2}) = (r_{3,4} \circ r_{2,3}) \circ r_{1,2}$
+    \end{alertblock}
+    \begin{alertblock}<6|handout:1>{identity} 
+        for every $r_{1,2}: \setc_1 \rightarrow \setc_2$ there exists identity morphisms $r_{1,2} \circ id_{\setc_1} = r_{1,2} = id_{\setc_2} \circ r_{1,2}$
     \end{alertblock}
@@
 
-\begin{frame}<presentation:1-|handout:1>{Opposite Category $\textcolor{source}{\mathcal{C}^{op}}$}
+\begin{frame}<presentation:0|handout:1>{Opposite Category $\textcolor{source}{\mathcal{C}^{op}}$}
     \begin{columns}
@@
     \end{description}
-\end{frame}
-
-\begin{frame}<presentation:1|handout:1>{Composable $\dfuncc = (\dfunchc, \prod\limits_{i=0}^{n}\dfunctc_i)$}
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement = \&]
-            \dtotalc_a 
-            \arrow[dd, "\dfunc c_a"', color=action] \& \& 
-            \dtotalc_a \times \dtotalc_b 
-            \arrow[dd, "{\dfuncc_{a,b}}" description, color=action] 
-            \arrow[ll, "\pi_a"', color=total] 
-            \arrow[rr, "\pi_b", color=total] \& \& 
-            \dtotalc_b 
-            \arrow[dd, "\dfuncc_b", color=action] \\
-            \& \& \& \& \\
-            \dtotalc_a  \& \& 
-            \dtotalc_a \times \dtotalc_b 
-            \arrow[rr, "\pi_b", color=total] 
-            \arrow[ll, "\pi_a"', color=total] \& \& 
-            \dtotalc_b
-        \end{tikzcd}
-    \end{equation*}    
 \end{frame}
@@
 
+\begin{frame}<presentation:1|handout:1>{Composable $\dfuncc = (\dfunchc, \prod\limits_{i=0}^{n}\dfunctc_i)$}
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement = \&]
+            \dtotalc_a 
+            \arrow[dd, "\dfunc a"', color=action] \& \& 
+            \dtotalc_a \times \dtotalc_b 
+            \arrow[dd, "{\dfuncc_{a,b}}" description, color=action] 
+            \arrow[ll, "\pi_a"', color=total] 
+            \arrow[rr, "\pi_b", color=total] \& \& 
+            \dtotalc_b 
+            \arrow[dd, "\dfuncc_b", color=action] \\
+            \& \& \& \& \\
+            \dtotalc_a  \& \& 
+            \dtotalc_a \times \dtotalc_b 
+            \arrow[rr, "\pi_b", color=total] 
+            \arrow[ll, "\pi_a"', color=total] \& \& 
+            \dtotalc_b
+        \end{tikzcd}
+    \end{equation*}
+    if there exists functions $\dfuncc_{a,b}: \dtotal_a \times \dtotal_b \rightarrow \dtotal_a \times \dtotal_b$, $\dfunc_{a}: \dtotal_a \rightarrow \dtotal_a$ and $\dfuncc_{b}: \dtotal_b \rightarrow \dtotal_b$ s.t.
+$\pi_a \circ \dfuncc_a = \dfuncc_{a,b} \circ \pi_a$ and $\pi_b \circ \dfuncc_b = \dfuncc_{a,b} \circ \pi_b$ then $\dfuncc_{a,b} = (\dfuncc_a, \dfuncc_b)$
+
+\end{frame}
+
 \section{Construction}
@@
     %rewrite in gamma notation
-    \begin{equation*}
-    \begin{tikzcd}[ampersand replacement = \&]
-        \dtotalc 
-        \arrow[dd, "\vchannelc_{\opensetc}"', color=artist] 
-        \arrow[rr, "\dfuncpullc"] 
-        \& \& 
-        \dfuncpullc\dtotalc 
-        \arrow[rr, "\dfunctc_{\dtotalc}"] 
-        \&  \& 
-        \dtotalc^{\prime} 
-        \arrow[dd, "{\vchannelc^{\prime}}_{\opensetc^{\prime}}", color=artist] \\
-        \& 
-        \opensetc 
-        \arrow[lu, dotted, maps to] 
-        \arrow[ld, dotted, maps to] \& 
-        \opensetc^{\prime} 
-        \arrow[l, "\dfunchc"] 
-        \arrow[u, dotted, maps to] 
-        \arrow[d, dotted, maps to] 
-        \arrow[rrd, dotted, maps to] 
-        \arrow[rru, dotted, maps to] 
-        \& \& \\
-        \vtotalc 
-        \arrow[rr, "\dfuncpullc"'] 
-        \& \& 
-        \dfuncpullc\vtotalc 
-        \arrow[rr, "\dfunctc_{\vtotalc}"'] 
-        \& \& 
-        \vtotalc^{\prime}
-    \end{tikzcd}
-    \end{equation*}
+    \begin{figure}
+        \includegraphics[.35\textwidth]{figures/math/nu_table.png}
+    \end{figure}
     Given $\dsectionc(\dbasepointc) = \dsectionc(\dfunchc(\dbasepointc^{\prime}))$ and  $\vsectionc(\dbasepointc) = \vsectionc(\dfunchc(\dbasepointc^{\prime}))$
@@
     \begin{columns}
-        \column{.5\textwidth}
+        \column{.15\textwidth}
         \begin{equation*}
@@
             \dfiberc_{\dbasepointc} 
-            \arrow[rr, "\vchannelc", maps to] 
+            \arrow[rr, "\vchannelc"] 
             \arrow[dd, "\dfunctc_{\dtotalc}"'] 
@@
             {} 
-            \arrow[rr, dashed, maps to]
+            \arrow[rr, dashed]
             \& \& {} \\
             {\dfiberc^{\prime}}_{\dbasepointc} 
-            \arrow[rr, "\vchannelc^{\prime}"', maps to]
+            \arrow[rr, "\vchannelc^{\prime}"']
             \& \& 
@@
             \dfiberc_{\dbasepointc} 
-            \arrow[rr, "\vchannelc_{\dtotalc} ", maps to, color=artist] 
-            \arrow[rrdd, "{\vchannelc_{\dtotalc,\vtotalc}}" description, dashed, maps to, color=artist] 
-            \&  \& 
-            \vfiberc_{\dbasepointc}\coloneqq {\dfiberc_{\dbasepointc}}^{\vtotalc} \arrow[dd, "\vchannelc_{\vtotalc}", maps to, color=artist] \\
+            \arrow[rr, "\vchannelc_{\dtotalc} ", color=artist] 
+            \arrow[rrdd, "{\vchannelc_{\dtotalc,\vtotalc}}" description, dashed, color=artist] 
+            \&  \& 
+            \vfiberc_{\dbasepointc}\coloneqq {\dfiberc_{\dbasepointc}}^{\vtotalc} 
+            \arrow[dd, "\vchannelc_{\vtotalc}", color=artist] \\
             \& \& \\
@@
             \dfiberc_a\times \dfiberc_b 
-            \arrow[rr, "\vchannelc", color=artist, maps to] 
+            \arrow[rr, "\vchannelc", color=artist] 
             \arrow[rd, "p_{\dfiberc}", color=fiber] \& \& \vfiberc_a\times \vfiberc_b 
@@
             \dfiberc_{a} 
-            \arrow[rr, "\vchannelc_{a}", dashed, maps to, color=artist] \& \& \vfiberc_a \\
+            \arrow[rr, "\vchannelc_{a}", dashed, color=artist] \& \& \vfiberc_a \\
             \dfiberc_a\times \dfiberc_c 
-            \arrow[rr, "\vchannelc^{\prime}", color=artist, maps to] 
+            \arrow[rr, "\vchannelc^{\prime}", color=artist] 
             \arrow[ru, "p_{\dfiberc^{\prime}}"', color=fiber] \& \& 
@@
             \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} 
-            \arrow[rrrr, "\vmarkc", Rightarrow, color=artist] 
+            \arrow[rrrr, "\vmarkc", color=artist] 
             \arrow[ddd, "\dfunctc_{\vtotalc} \circ \dfuncpullc_{\vtotalc}"', color=action] 
@@
             \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
-            \arrow[ddd, "\dfunctc_{\gtotalc} \circ \dfunchc_{\gtotalc}", dotted, color=action]  \\
+            \arrow[ddd, "\dfunctc_{\gtotalc} \circ \dfuncpullc_{\gtotalc}", dotted, color=action]  \\
             \& 
             \opensetc 
-            \arrow[lu, maps to, color=sheaf] 
              \&  \& 
             \opensetgc 
-            \arrow[ll, "\vindexc"', color=functor] 
-            \arrow[ru, maps to, color=sheaf]  \& \\
+            \arrow[ll, "\vindexc"', color=functor]   \& \\
             \& 
@@
             \arrow[u, "\dfunchc_{\dtotalc}", color=action] 
-            \arrow[ld, maps to, color=sheaf] 
             \&  \& 
@@
             \arrow[u, "\dfunchc_{\gtotalc}"', dotted, color=action] 
-            \arrow[rd, maps to, color=functor] \& \\
+            \& \\
             \cgamma{\opensetc^{\prime}}{\vtotalc^{\prime}\restriction_{\opensetc^{\prime}}}
-            \arrow[rrrr, "\vmarkc^{\prime}", Rightarrow, color=artist] 
+            \arrow[rrrr, "\vmarkc^{\prime}", color=artist] 
             \& \&  \& \& 
@@
 
-\begin{frame}<presentation:1|handout:1>{Implementation Choices: $\vartistc_{\opensetc} = \vartistc_{\opensetgc}$}
-    \begin{equation*}
-    \begin{tikzcd}[ampersand replacement = \&, column sep=tiny]
-        \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}  
-        \arrow[r, "\vchannelc_{\opensetgc}", maps to, color=artist] 
-        \arrow[rrrr, "\vartistc_{\opensetgc}", Rightarrow, bend left, color=artist] \& 
-        \cgamma{\opensetgc}{\vindexpullc\vtotalc\restriction_{\opensetgc}} \& 
-        \cgamma{\opensetgc}{\vindexpullc\vtotalc\restriction_{\opensetgc}}  
-        \arrow[rr, "\vmarkc_{\opensetgc}", no head, Rightarrow, color=artist] \&  \& 
-        \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}}  
-        \arrow[dd, "\vindexpushc", color=functor] \\
-        \& \& \& \& \\
-        \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}  
-        \arrow[uu, "\vindexpullc", color=functor] 
-        \arrow[r, "\vchannelc_{\opensetc}", maps to, color=artist] 
-        \arrow[rrrr, "\vartistc_{\opensetc}", Rightarrow, bend right, color=artist] 
-        \& 
-        \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}}  
-        \arrow[uu, "\vindexpullc"', color=functor] \& 
-        \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}}
-        \arrow[rr, "\vmarkc_{\opensetc}", Rightarrow, color=artist] 
-        \arrow[uu, "\vindexpullc", color=functor] 
-        \arrow[rruu, "\vmarkc", Rightarrow, color=artist] 
-        \&  \& 
-        \imartist{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}
-    \end{tikzcd}
-    \end{equation*}
-\end{frame}
-
 \begin{frame}<presentation:1|handout:1>{Components to Graphic}
     \begin{figure}
-        \includegraphics[scale=.35]{../paper/figures/qhat.png}
-    \end{figure}
-\end{frame}
-\end{comment}
+        \includegraphics[scale=.15]{../paper/figures/qhat.png}
+    \end{figure}
+\end{frame}
+
+\begin{frame}<presentation:1|handout:1>{Implementation Choices: $\vartistc_{\dbasec} = \vartistc_{\gbasec}$}
+    \begin{figure}
+        \includegraphics[scale=.5]{figures/math/implement_paths).png}
+    \end{figure}
+\end{frame}
+
+\begin{frame}{Combining Qs}
+    The codomain of all $\vmarkc$ targeting the same output space is the bundle $\gtotalc$ and $Hom(\gbasec_1, \gtotalc) + Hom(\gbasec_2, \gtotalc) = Hom(\gbasec_1 + \gbasec_2, \gtotalc)$; therefore 
+    $\sqcup \vmarkc_i(\cgamma{\opensetc_i}{\dtotalc_i\restriction_{\opensetc_i}}) = \cgamma{\sqcup \opensetgc_i}{\gtotalc\restriction_{\sqcup\opensetgc_i}}$
+    \begin{figure}
+        \includegraphics[scale=.25]{figures/math/qcom.png}
+    \end{figure}
+\end{frame}
+
+\begin{frame}{Compatible Qs}
+    \includegraphics[scale=.5]{figures/math/compatiblw_q.png}
+\end{frame}
 \end{document}

### 2022-03-29 slides/main.tex
artist nts

@@
 
-\begin{frame}<presentation:1-|handout:1>
+\begin{frame}<presentation:0|handout:0>
     \frametitle{Functor $\textcolor{functor}{\bm{F}}: \textcolor{source}{\mathcal{C}} \rightarrow \textcolor{target}{\mathcal{D}}$}
@@
 \subsubsection{Natural Transforms}
-\begin{frame}<presentation:1-|handout:1>
+\begin{frame}<presentation:0|handout:0>
     \frametitle{Natural Transformation $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\Rightarrow} \textcolor{functor}{G}$}
@@
 
-\begin{frame}<presentation:1-|handout:1>
+\begin{frame}<presentation:0|handout:0>
     \frametitle{Natural Transformation $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\Rightarrow} \textcolor{functor}{G}$}
@@
 \begin{frame}<presentation:1-|handout:1>
-    \frametitle{Presheaf: $\sheafc: \mathcal{C}^{op} \textcolor{sheaf}{\rightarrow} \setc $}
+    \frametitle{Presheaf Functor: $\sheafc: \mathcal{C}^{op} \textcolor{sheaf}{\rightarrow} \setc $}
     \begin{columns}
@@
     \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-        \onslide*<9-|handout:1->{\dfiberc}
-        \onslide*<9-|handout:1->{\arrow[r, hook]} \& 
+        \onslide*<8-|handout:1->{\dfiberc}
+        \onslide*<8-|handout:1->{\arrow[r, hook]} \& 
         \onslide*<1-|handout:1->{\dtotalc}
@@
         \onslide*<1-|handout:1->{\dbasec}
-        \onslide*<3-|handout:1->{\arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]}
+        \onslide*<2-|handout:1->{\arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]}
         \end{tikzcd}
@@
         \begin{tikzcd}[ampersand replacement = \&, column sep=small]
-            \onslide*<3,6-|handout:0>{{\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}} 
-            \onslide*<4-5|handout:0>{{\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}\in Ob(\setc)}
+            \onslide*<2,5-|handout:0>{{\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}} 
+            \onslide*<3-4|handout:0>{{\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}\in Ob(\setc)}
             \onslide*<0|handout:1>{\Ob(\setc) \ni {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}}
@@
             {} \& 
-            \onslide*<6-|handout:1>{{\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}}} 
-            \onslide*<7-|handout:1>{\arrow[ll, "\iota^*"', hook', color=set]} \\\& \&\\
-            \onslide*<3|handout:0>{\opensetc_1 \subset \dbasec}
-            \onslide*<4-5|handout:0>{\opensetc_1 \in Ob(\mathcal{\dbasec}^{\textcolor{base}{op}})}
-            \onslide*<6-|handout:0>{\opensetc_1}
+            \onslide*<5-|handout:1>{{\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}}} 
+            \onslide*<6-|handout:1>{\arrow[ll, "\iota^*"', hook', color=set]} \\\& \&\\
+            \onslide*<2|handout:0>{\opensetc_1 \subset \dbasec}
+            \onslide*<3-4|handout:0>{\opensetc_1 \in Ob(\mathcal{\dbasec}^{\textcolor{base}{op}})}
+            \onslide*<5-|handout:0>{\opensetc_1}
             \onslide*<0|handout:1>{Ob(\mathcal{\dbasec}^{\textcolor{base}{op}}) \ni \opensetc_1} 
-            \onslide*<5-|handout:0>{\arrow[uu, "\sheafc_{\dbasec, \dtotalc}", maps to, color=sheaf]}
+            \onslide*<4-|handout:0>{\arrow[uu, "\sheafc_{\dbasec, \dtotalc}", maps to, color=sheaf]}
             \onslide*<0|handout:1>{\arrow[uu, "\sheafc_{\dbasec, \dtotalc}", maps to, color=sheaf, shift right = 10]}
-            \onslide*<7-|handout:1>{\arrow[rr, "\iota"', hook, color=base]} \&   
-            \onslide*<8-|handout:1>{\arrow[uu, "\sheafc_{\dbasec, \dtotalc}" description, maps to, color=sheaf]} \& 
-            \onslide*<6-|handout:1>{\opensetc_2} 
-            \onslide*<6-|handout:1>{\arrow[uu, "\sheafc_{\dbasec, \dtotalc}"', maps to, color=sheaf]}                   
+            \onslide*<6-|handout:1>{\arrow[rr, "\iota"', hook, color=base]} \&   
+            \onslide*<7-|handout:1>{\arrow[uu, "\sheafc_{\dbasec, \dtotalc}" description, maps to, color=sheaf]} \& 
+            \onslide*<5-|handout:1>{\opensetc_2} 
+            \onslide*<5-|handout:1>{\arrow[uu, "\sheafc_{\dbasec, \dtotalc}"', maps to, color=sheaf]}                   
             \end{tikzcd}
@@
     \begin{center}
-        \begin{alertblock}<8-|handout:1>{stalk}
+        \begin{alertblock}<7-|handout:1>{stalk}
             \begin{align*}
@@
 
-        \begin{alertblock}<9-|handout:1>{germ}
+        \begin{alertblock}<8-|handout:1>{germ}
             %germ  is tau at limit, \tau(k) \in germ
@@
 
-\begin{frame}<presentation:1|handout:1>{Sheaves on Bundles}
+\begin{frame}<presentation:1-|handout:1>{Sheaves on Bundles}
     A sheaf is a presheaf that satisfies the following two axioms\cite{bakerMathsSheaf}\\
@@
 
-\begin{frame}<presentation:1|handout:1>{Example: Scatter Plot}
+\begin{frame}<presentation:0|handout:0>{Example: Scatter Plot}
     \begin{figure}
@@
 
-\begin{frame}<presentation:1|handout:1>{Example: Line Plot}
+\begin{frame}<presentation:0|handout:0>{Example: Line Plot}
     \begin{figure}
@@
 
-\begin{frame}<presentation:1|handout:1>{Example: Image}
+\begin{frame}<presentation:0|handout:0>{Example: Image}
     \begin{figure}
@@
 
-\begin{frame}<presentation:1|handout:1>{Example: Heatmap}
+\begin{frame}<presentation:0|handout:0>{Example: Heatmap}
     \begin{figure}
@@
 
-\begin{frame}<presentation:1|handout:1>{Example: Sphere}
+\begin{frame}<presentation:0|handout:0>{Example: Sphere}
     \begin{figure}
@@
     \frametitle<3|handout:3>{Display Space: $\vartist_{\gbasec}: \sheafc_{\gbasec, \vindexpullc\dtotalc} \textcolor{artist}{\Rightarrow} \sheafc_{\gbasec, \gtotalc}$}
-    \frametitle<4|handout:4>{Artist: $\vartistc: \dsectionc \mapsto \gsectionc$}
-
-    \onslide*<1,4|handout:1,4>{
+    \frametitle<6|handout:6>{Artist: $\vartistc: \dsectionc \mapsto \gsectionc$}
+
+    \onslide*<1,8|handout:1,8>{
     \begin{equation*}
@@
           \onslide*<1|handout:1>{\sheafc_{\gbasec, \gtotalc}} 
-          \onslide*<4|handout:4>{\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}}
-          \onslide*<1,4|handout:1,4>{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]}  
+          \onslide*<8|handout:4>{\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}}
+          \onslide*<1,8|handout:1,4>{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]}  
           \& \& 
           \onslide*<1|handout:1>{\sheafc_{\dbasec, \vindexpushc\gtotalc}}
-          \onslide*<4|handout:4>{\cgamma{\opensetc}{\vindexpushc \gtotalc\restriction_{\opensetc}}}
+          \onslide*<8|handout:4>{\cgamma{\opensetc}{\vindexpushc \gtotalc\restriction_{\opensetc}}}
           \\
@@
           \onslide*<1|handout:1>{\sheafc_{\gbasec, \vindexpullc\dtotalc}}
-          \onslide*<4|handout:4>{\cgamma{\opensetgc}{\vindexpullc \dtotalc\restriction_{\opensetgc}}}
+          \onslide*<8|handout:4>{\cgamma{\opensetgc}{\vindexpullc \dtotalc\restriction_{\opensetgc}}}
           \onslide*<1|handout:1>{\arrow[uu, "\vartistc_{\gbasec} \in \textcolor{homset}{Hom}_{\gbasec}", color=homset]}
-          \onslide*<4|handout:4>{\arrow[uu, "\vartistc_{\gbasec}", color=artist]}
+          \onslide*<8|handout:4>{\arrow[uu, "\vartistc_{\gbasec}", color=artist]}
           \&  \& 
           \onslide*<1|handout:1>{\sheafc_{\dbasec, \dtotalc}}
-          \onslide*<4|handout:4>{\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}}
-          \onslide*<1,4|handout:1,4>{\arrow[ll, "\textcolor{functor}{\vindexpull}", color=functor]} 
+          \onslide*<8|handout:4>{\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}}
+          \onslide*<1,8|handout:1,4>{\arrow[ll, "\textcolor{functor}{\vindexpull}", color=functor]} 
           \onslide*<1|handout:1>{\arrow[uu, "\vartistc_{\dbasec} \in \textcolor{homset}{Hom}_{\dbasec}"', color=homset]}
-          \onslide*<4|handout:4>{\arrow[uu, "\vartistc_{\dbasec}"', color=artist]}
+          \onslide*<8|handout:4>{\arrow[uu, "\vartistc_{\dbasec}"', color=artist]}
           \onslide*<1|handout:1>{\arrow[lluu, "\vartistc \in \textcolor{homset}{Hom}_{\dbasec,\gbasec}"' description, dashed, color=homset]}
-          \onslide*<4|handout:4>{\arrow[lluu, "\vartistc"' description, color=artist]}
+          \onslide*<8|handout:4>{\arrow[lluu, "\vartistc"' description, color=artist]}
         \end{tikzcd}
@@
 
-    \onslide*<2|handout:2>{
-        \begin{equation*}
-            \begin{tikzcd}[ampersand replacement=\&]  
-                    \&  \& 
-                    \sheafc_{\dbasec, \dtotalc} 
-                    \arrow[rr, "\vartistc_{\dbasec}", Rightarrow, color=artist] \& \& 
-                    \sheafc_{\dbasec, \vindexpushc\gtotalc}  \\
-                    \opensetc_1\& \& 
-                    \cgamma{\opensetc_1}{\dtotalc\restriction_{\opensetc_1}}
-                    \arrow[dd, "\iota^*"', color=set] 
-                    \arrow[rr, "\vartistc_{\opensetc_1}", color=artist] \& \& 
-                    \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc_1}} 
-                    \arrow[dd, "\iota^*", color=set] \\
+    \onslide*<2-6|handout:2>{
+        \begin{equation*}
+            \begin{tikzcd}[ampersand replacement=\&]  
+                    \&  \& 
+                    \onslide*<6|handout:2>{\sheafc_{\dbasec, \dtotalc}} 
+                    \onslide*<6|handout:2>{\arrow[rr, "\vartistc_{\dbasec}", Rightarrow, color=artist]} \& \& 
+                    \onslide*<6|handout:2>{\sheafc_{\dbasec, \vindexpushc\gtotalc}}  \\
+                    \onslide*<4-6|handout:2>{\opensetc_1}\& \& 
+                    \onslide*<3-6|handout:2>{\cgamma{\opensetc_1}{\dtotalc\restriction_{\opensetc_1}}}
+                    \onslide*<5-6|handout:2>{\arrow[dd, "\iota^*"', color=set]} 
+                    \onslide*<3-6|handout:2>{\arrow[rr, "\vartistc_{\opensetc_1}", color=artist]} \& \& 
+                    \onslide*<3-6|handout:2>{\cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc_1}}} 
+                    \onslide*<5-6|handout:2>{\arrow[dd, "\iota^*", color=set]} \\
                     \& \& {} 
-                    \arrow[rr, "\vartistc_{\sheafc}", dotted, color=artist] \& \& {} \\
-                    \arrow[uu, "\iota" description, hook, color=base] 
-                    \opensetc_2 \& \& 
-                    \cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}
-                    \arrow[rr, "\vartistc_{\opensetc_2}", color=artist] \& \& 
-                    \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc_2}} \\
-                    \& \& 
+                    \onslide*<5-6|handout:2>{\arrow[rr, "\vartistc_{\sheafc}", dotted, color=artist]} \& \& {} \\
+                    \onslide*<4-6|handout:2>{\arrow[uu, "\iota" description, hook, color=base]} 
+                    \onslide*<4-6|handout:2>{\opensetc_2} \& \& 
+                    \onslide*<3-6|handout:2>{\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}}
+                    \onslide*<3-6|handout:2>{\arrow[rr, "\vartistc_{\opensetc_2}", color=artist]} \& \& 
+                    \onslide*<3-6|handout:2>{\cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc_2}}} \\
+                    \& \& 
+                    \onslide*<2-6|handout:2>{
                     \cgamma{\opensetc}{\dtotalc\restriction\opensetc} \& 
@@
                     \arrow[l, "\sheafc_{\dbasec, \dtotalc}"', maps to, color=sheaf] \& 
-                    \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction\opensetc}
+                    \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction\opensetc}}
             \end{tikzcd}                    
@@
 
-    \onslide*<3|handout:3>{
+    \onslide*<7|handout:3>{
         \begin{equation*}
@@
                     \arrow[uu, "\iota" description, hook, color=base] 
-                    \opensetc_2 \& \& 
+                    \opensetgc_2 \& \& 
                     \cgamma{\opensetgc_2}{\vindexpullc\dtotalc\restriction_{\opensetgc_2}}
@@
     }
-    \onslide<4|handout:4>{
+    \onslide<8|handout:4>{
         \begin{alertblock}{pull data section \dsectionc\ over graphic space $\gbasepointc \in \gbasec$}
@@
 \subsection{Equivariance}
-\begin{frame}<presentation:1|handout:1>{Transform Data}
+\begin{frame}<presentation:1|handout:1>{Expressing Equivariance via Morphisms $\dfuncc$}
     %swap fiber & bundle, note that need section morphisms not bundle morphisms
@@
 The morphisms on the fiber are $\dfunctc \in Hom(\dfiberc, \dfiberc)$
+        The category is equipped with the bifunctor $\otimes: \dfiberc \times \dfiberc \rightarrow \dfiberc$
     \end{block}

### 2022-03-29 slides/main.tex
compiled arrows w/ q

@@
 \begin{frame}<presentation:1|handout:1>{$\dfuncc = (\dfunchc, \dfunctc)$}
-    \begin{figure}
-        \includegraphics[width=0.5\textwidth]{figures/math/dfunc.png}
-    \end{figure}
+
+    \begin{equation*}
+            \begin{tikzcd}[ampersand replacement=\&, column sep=.3em]
+                \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
+                \arrow[rr, "\dfuncpullc", color=dfuncc] \&  \& 
+                \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}} \&  \& 
+                \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}}
+                \arrow[rr, "\dfunctc", color=dfuncc] 
+                \&  \& 
+                \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}} \\
+                \& \& \& \&  \&  \&  \\
+                \opensetc 
+                \arrow[uu, maps to, color=sheaf] \&  \& 
+                \opensetc^{\prime} 
+                \arrow[uu, maps to, color=sheaf] 
+                \arrow[ll, "\dfunchc"', color=dfuncc] 
+                \& \& \& \&                                                                       
+            \end{tikzcd}
+    \end{equation*}
     \begin{description}
@@
 \section{Construction}
+\begin{frame}{How Do We Get From Data to Graphic?}
+    \begin{figure}
+        \includegraphics[width=1\textwidth]{figures/math/path_of_math.png}
+    \end{figure}
+\end{frame}
+
 \begin{frame}<presentation:1|handout:1>{Building an equivariant $\vartist$?}
-    \note{Add box around figure & lable it artist}
     \begin{figure}
@@
 \subsection{Measurement Specification $\vtotalc$}
+\begin{frame}<presentation:1|handout:1>{Data to Measurable Components}
+    \begin{figure}
+        \includegraphics[width=1\textwidth]{figures/math/justnu.png}
+    \end{figure}
+\end{frame}
 
@@
 
-\begin{frame}<presentation:1|handout:1>
+\begin{frame}<presentation:0|handout:1>
     \frametitle{Transformations in Data $\vchannelc: \dsectionc \mapsto \vsectionc$}
@@
 \subsection{Combining Components into Visualizations}
+\begin{frame}<presentation:1|handout:1>{Components to Graphic}
+    \begin{figure}
+        \includegraphics[width=1\textwidth]{figures/math/justq.png}
+    \end{figure}
+\end{frame}
+
 \begin{frame}<presentation:1|handout:1>{Assembly $\vmarkc$}
@@
 
-\begin{frame}<presentation:1|handout:1>{Components to Graphic}
-    \begin{figure}
-        \includegraphics[scale=.15]{../paper/figures/qhat.png}
-    \end{figure}
+\begin{frame}{Combining Qs}
+    \begin{columns}
+    \column{.55\textwidth}       
+    The codomain of all $\vmarkc$ targeting the same output space is the bundle $\gtotalc$ and $Hom(\gbasec_1, \gtotalc) + Hom(\gbasec_2, \gtotalc) = Hom(\gbasec_1 + \gbasec_2, \gtotalc)$; therefore 
+    \begin{equation*}
+        \bigsqcup_i \vmarkc_i(\cgamma{\opensetc_i}{\dtotalc_i\restriction_{\opensetc_i}}) = \cgamma{\mathop{\sqcup}_i\opensetgc_i}{\gtotalc\restriction_{\mathop{\sqcup}_i\opensetgc_i}})
+    \end{equation*}
+    when $\vindexc(\opensetgc_i) = \opensetc_i$
+
+    \column{.45\textwidth}
+    \begin{figure}
+        \includegraphics[scale=.25]{figures/math/qcom.png}
+    \end{figure}
+    \end{columns}
+\end{frame}
+
+\begin{frame}{Compatible Qs}
+    \begin{equation*}
+    \begin{tikzcd}[ampersand replacement=\&, column sep=small]
+        \& 
+        \cgamma{\opensetc}{\vtotalc^{a}\restriction_{\opensetc}} 
+        \arrow[dd, "\vchannelc^{a\rightarrow b}", color=artist] 
+        \arrow[rrd, "{\vmarkc_{\opensetc}}^{a}", color=artist] 
+        \arrow[rrrrd, "\vmarkc^{a}", bend left, color=artist] 
+        \&  \&  \& \&   \\
+        \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
+        \arrow[ru, "\vchannelc^a", color=artist] 
+        \arrow[rd, "\vchannelc^b"', color=artist] 
+        \& \&  \& 
+        \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} 
+        \&  \& 
+        \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+        \arrow[ll, "\vindexpushc"', color=functor] \\
+        \& \cgamma{\opensetc}{\vtotalc^{b}\restriction_{\opensetc}} 
+        \arrow[rru, "{\vmarkc_{\opensetc}}^{b}"', color=artist] 
+        \arrow[rrrru, "{\vmarkc}^{b}"', bend right, color=artist] 
+         \&  \&  \&  \&  
+    \end{tikzcd}
+    \end{equation*}
 \end{frame}
@@
 \begin{frame}<presentation:1|handout:1>{Implementation Choices: $\vartistc_{\dbasec} = \vartistc_{\gbasec}$}
-    \begin{figure}
-        \includegraphics[scale=.5]{figures/math/implement_paths).png}
-    \end{figure}
-\end{frame}
-
-\begin{frame}{Combining Qs}
-    The codomain of all $\vmarkc$ targeting the same output space is the bundle $\gtotalc$ and $Hom(\gbasec_1, \gtotalc) + Hom(\gbasec_2, \gtotalc) = Hom(\gbasec_1 + \gbasec_2, \gtotalc)$; therefore 
-    $\sqcup \vmarkc_i(\cgamma{\opensetc_i}{\dtotalc_i\restriction_{\opensetc_i}}) = \cgamma{\sqcup \opensetgc_i}{\gtotalc\restriction_{\sqcup\opensetgc_i}}$
-    \begin{figure}
-        \includegraphics[scale=.25]{figures/math/qcom.png}
-    \end{figure}
-\end{frame}
-
-\begin{frame}{Compatible Qs}
-    \includegraphics[scale=.5]{figures/math/compatiblw_q.png}
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement=\&, column sep=small]
+            \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} 
+            \arrow[rr, "\vchannelc_{\gbasec}", color=artist] 
+            \arrow[rrrr, "\vartistc_{\gbasec}", bend left, color=artist]  
+            \&  \& 
+            \cgamma{\opensetgc}{\vindexpullc\vtotalc\restriction_{\opensetgc}} 
+            \arrow[rr, "\vmarkc_{\gbasec}", color=artist] \&  \& 
+            \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+            \arrow[dd, "\vindexpushc", color=functor] \\
+             \&  \&   \&  \& \\
+            \cgamma{\openset}{\dtotalc\restriction_{\opensetc}} 
+            \arrow[rr, "\vchannelc_{\dbasec}", color=artist] 
+            \arrow[rrrr, "\vartistc_{\dbasec}", bend right, color=artist] 
+            \arrow[uu, "\vindexpullc", color=functor] \&  \& 
+            \cgamma{\openset}{\vtotalc\restriction_{\opensetc}} 
+            \arrow[rr, "\vmarkc_{\dbasec}", color=artist ] 
+            \arrow[uu, "\vindexpullc", , color=functor] \&  \& 
+            \imartist{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}                 
+            \end{tikzcd}
+    \end{equation*}
 \end{frame}

### 2022-04-04 paper/main.tex
equivariance
cleared out, rewrote intro to math section

--- 2021-12-22 paper/main.tex
@@
 
-\IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
+\IEEEraisesectionheading{\section{Introduction}\label{sec:intro}}
 
-\IEEEPARstart{T}his paper uses methods from algebraic topology and category theory to develop a model of the transformation from data to graphical representation.
-This model provides a language to specify how data is structured and how this structure is carried through in the visualization, and serves as the basis for a functional approach to implementing visualization library components.
-Topology allows us to describe the structure of the data and graphics in a generalizable, scalable, and trackable way.
-Category theory provides a framework for separating the transformations implemented by visualization libraries from the various stages of visualization and therefore can be used to describe the constraints imposed on the library components \cite{wielsManagementEvolvingSpecifications1998,goguenCategoricalManifesto1991}.
-Well constrained modular components are inherently functional\cite{hughesWhyFunctionalProgramming1989}, and a functional framework yields a library implementation that is likely to be shorter, clearer, and more suited to distributed, concurrent, and on demand tasks\cite{huHowFunctionalProgramming2015}.
-Using this functional approach, this paper contributes a practical framework for decoupling data processing from visualization generation in a way that allows for modular visualization components that are applicable to a variety of data sets in different formats.
-\note{is it OK that this is something reviewer 4 wrote}
+\IEEEPARstart{T}
 
 \section{Related Work}
-This work aims to develop a model for describing visualization transformations that can serve as guidance for how to architecture a general purpose visualization library.
-We define a general purpose visualization library as one that provides non domain specific building block components\cite{wongsuphasawatNavigatingWideWorld2021} for building visualizations, for example functions for converting data to color or encoding data as dots.
-We propose that visualization library components should preserve continuity and equivariance, which we define as
 
-\begin{LaTeXdescription}
-  \item [continuity] how elements in a dataset are organized, e.g. discrete rows in a table, networked nodes, pixels in an image, points on a line
-  \item [equivariance] data transformations that have an equivalent effect on the graphical representation, e.g.
-rotating a matrix has a matching rotation of the image, translating the points on a line has a matching visual shift in the line plot
-\end{LaTeXdescription}
-
-In \autoref{sec:related-work:continuity} we discuss how data continuity influences visualization library architecture and in \autoref{sec:related-work:equivariance} we summarize how visualization researchers have approached equivariance.
-
-\subsection{Continuity}
-\label{sec:related-work:continuity}
-\begin{figure}[h!]
-  \includegraphics[width=\columnwidth]{k_different_types.png}
-  \caption{Continuity is how elements in a data set are connected to each other, which is distinct from how the data is structured.
-The rows in (a) are discrete, therefore they have discrete continuity as illustrated by the discrete dots.
-The gaussian in (b) is a 1D continuous function, therefore the continuity of the elements of the gaussian can be represented as a line on an interval (0,1).
-In (c), every element of the globe is connected to its nearest neighbors, which yields a 2D continuous continuity as illustrated by the square.}
-  \label{fig:related-work:continuity}
-\end{figure}
-Continuity describes elements in a data set are organized; this concept is termed topological properties by Wilkinson\cite{wilkinsonGrammarGraphics2005}.
-Wilkinson provides the examples of values that are isolated from each other, and therefore discrete, and values lying on a continuum or in a compact region.
-For example, in \autoref{fig:related-work:continuity}, each station record in the table is independent of the others; therefore, the continuity of the table is discrete.
-The data provided by the gaussian are points sampled along the curve, therefore the continuity of the points on the line is 1D continuous.
-Every point on the globe is connected to its 6 nearest cardinal neighboring points (NW, N, NE, E, SE, S, SW, W).
-We propose that a robust model of continuity provides a way to develop library components that can work on the data in small pieces in a manner where the overall continuity of the data is preserved in the visual transformation.
-
-\begin{figure}[h!]
-  \includegraphics[width=\columnwidth]{whycontinuity.png}
-  \caption{Continuity is implicit in choice of visualization rather than explicitly in choice of data container.
-The line plots in (b) are generated by a 2D table (a).
-Structurally this table can be identical to the 2D matrix (a) that generates the image in (c).}
-  \label{fig:related-work:visual-algorithm}
-\end{figure}
-
- The preservation of continuity can be made explicit, as in the transformation of table to parallel coordinates in Ruchikachorn and Mueller \cite{ruchikachornLearningVisualizationsAnalogy2015}, but is often expressed implicitly in the choice of visual algorithm (visualization type), as explored in taxonomies by Tory and M\"{o}ller \cite{toryRethinkingVisualizationHighlevel2004} and Chi\cite{chiTaxonomyVisualizationTechniques2000}.
-
-For example, in \autoref{fig:related-work:visual-algorithm} the same table can be interpreted as a set of 1D continuous curves when visualized as a collection of line plots or as a 2D surface when visualized as an image.
-This means that often there is no way to express data continuity independent of visualization type, meaning most visualization libraries will allow, for example, visualizing discrete data as a line plot or an image.
-
-\subsection{Equivariance}
-\label{sec:related-work:equivariance}
-When introducing the retinal variables, Bertin informally specifies that continuity is preserved in the mark and defines equivariance constraints in terms of data and visual variables being selective, associative, ordered, or quantitative\cite{bertinSemiologyGraphicsDiagrams2011a}.
-In the \textit{A Presentation Tool}(APT) model, Mackinlay embeds the continuity constraint in the choice of visualization type and generalizes the equivariance constraint to preserving a binary operator from one domain to another.
-The algebraic model of visualization\cite{kindlmannAlgebraicProcessVisualization2014}, proposed by Kindlmann and Scheidegger, restricts equivariance to invertible transformations.
-
-\begin{figure}[h!]
-  \label{fig:related-work:equivariance}
-  \includegraphics[width=\columnwidth]{equiv.png}
-  \caption{Equivariance is that a transformation on the data has a corresponding transformation in the graphical representation.
-For example, in this figure the data is scaled by a factor 10.
-Equivalently the line plot is scaled by factor of 10, resulting in a shrunken line plot.
-Either a transformation on the data side can induce a transformation on the visual side, or a transformation on the visual side indicates that there is also a transformation on the data side.
-}
-\end{figure}
-
-\subsection{Visualization Library Architecture}
-General purpose visualization libraries-such as Matplotlib\cite{hunterMatplotlib2DGraphics2007}, Vtk\cite{hanwellVisualizationToolkitVTK2015,geveciVTK2012}, and D3 \cite{bostockDataDrivenDocuments2011}-carry distinct data models as part of the implementation of each visual algorithm.
-The lack of unified data model means that each plot in a linked\cite{beckerBrushingScatterplots1987,bujaInteractiveData1991} visualization is treated as independent, as are the transforms converting each field in the data to a visual equivalent.
-
-Domain specific libraries can often guarantee consistency because they have a single model of the data in their software design, as discussed in Heer and Agarwal \cite{HeerSoftware2006}'s survey of visualization software design patterns.
-For example, the relational database is core to tools influenced by APT, such as Tableau\cite{StoltePolaris2002,hanrahanVizQL2006,MackinlayShowme2007} and the Grammar of Graphics\cite{wilkinsonGrammarGraphics2005} inspired ggplot\cite{wickhamGgplot2ElegantGraphics2016a}, Vega\cite{satyanarayanDeclarativeInteractionDesign2014} and Altair\cite{vanderplasAltairInteractiveStatistical2018}.
-Images underpin scientific visualization tools such as Napari\cite{nicholas_sofroniew_2021_4533308} and ImageJ\cite{schneiderNIHImageImageJ2012} and the digital humanities oriented ImagePlot\cite{studiesCulturevisImageplot2021} macro; the need to visualize and manipulate graphs has spawned tools like Gephi\cite{bastianGephiOpenSource2009}, Graphviz\cite{ellsonGraphvizOpenSource2002}, and Networkx\cite{HagbergExploringNetwork2008}.
-
-\section{Algebraic Topology \& Category Theory}
-\note{potentially move this to appendix, push Butler \& Vickers up to RW}
-
-Category theory can be used to express software specifications because it provides a method of expressing structure and how that structure can be composed \cite{wielsManagementEvolvingSpecifications1998} algebraic topology provides a rich language for expressing data continuity.
-In this section, we introduce concepts from algebraic topology and category theory that we will use in \autoref{sec:artist:category} to develop a formal model of specifying visualization component.
+\section{Expressing Structure}
+\label{sec:atct}
+\note{why why why?}
+We use concepts from algberaic topolology and category theory to formally express the structure visualization components are expected to preserve.
+Category theory is a mathematical langauge for describing how objects are specified, constrained, and composed \cite{wielsManagementEvolvingSpecifications1998}\; therefore in this work we propose implementable in code categorical constructions of visualization components.
+Fiber bundles, are a "unified, dimension-independent framework", as described by Butler\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}, that allow us to seperatly describe the topololigical properties and fields of a data source and also the mapping between topolology and data values.
+Sheaves on bundles are an "algebraic data structure", as described by Ghrist \cite{ghristElementaryAppliedTopology2014}, for for expressing the bookeeping involved in subsetting data and composing data that is distributed or streaming.
 
 \subsection{Fiber Bundles}
-\label{sec:related-work:fiber-bundles}
-\begin{figure}[h!]
-  \includegraphics[width=\columnwidth]{fiberbundle.png}
-  \caption{A fiber bundle is mathematical construct that allows us to express the relationship between data and continuity.
-The \textcolor{total}{total} space \dtotal is the topological space in which the data is embedded.
-The \textcolor{fiber}{fiber} space \dfiber\ is embedded in \dtotal\ and is the set of all possible values that any
-  \note{add big rectangle E}}
-  \label{fig:related-work:fiber-bundle}
-\end{figure}
+\label{sec:atct:fiber-bundles}
 
-The model described in this work provides a model for expressing data sets with different topological properties.
-
-\subsection{Fiber Bundles}
-We obtain this generality by using a mathematical structure called a fiber bundles as the basis of our abstraction, as proposed by Butler, Bryson, and Pendley\cite{butlerVisualizationModelBased1989,
-butlerVectorBundleClassesForm1992}.
-As described by them, a fiber bundle is a formal model of the mapping between data points and the underlying topological space they lie in.
-For example, nodes on a network and the graph of the network, an image and the underlying grid at which the image is sampled, or table columns and the table index.
-
-Formally, a fiber bundle is a mathematical structure $(\dtotalc, \dbasec, \pi, \dfiberc)$
-\begin{equation}
-  \label{eq:fiber_bundle}
-  \begin{tikzcd}
-      \dfiberc \arrow[r, hook] & \dtotalc \arrow[r, "\pi"] & \dbasec
-  \end{tikzcd}
-\end{equation}
-
-\begin{definition}
-The \textcolor{base}{base space} \dbasec\ is a topological space, which means it is a set $\dbase$ with a collection of open sets \cite{TopologyVsTopology} surrounding each element in the set.
-Open sets are a collection of subsets $\{\opensetc\}$ in a set \dbase, including the empty set, the whole set \dbase, and every union and intersection of its member subsets.
-For each point $\dbasepoint \in \dbase$, there is a collection of subsets $\{\openset\} \subset \dbase$ such that $\dbasepoint \in \openset \subseteq \dbase$ \cite{spanier1989algebraic, bradleyTopologyCategoricalApproach2020}
-\end{definition}
-\note{intuition about base space}
-
-\begin{definition} The \textcolor{fiber}{fiber space} \dfiberc\ is a topological space such that for every $\dbasepoint \in \dbase$, the fiber over that point is isomorphic to the preimage of that point $\dfiber_{\dbasepoint} \cong \pi^{-1}(\{\dbasepoint\})$.
-All fibers $\dfiber_{\dbasepoint}$ are homeomorphic to each other \cite{spanier1989algebraic, hatcherAlgebraicTopology2002}
-\end{definition}
-
-The fibers can be thought of as encapsulating data types according to Spivak, who proposes that databases can be represented as fiber bundles \cite{spivakSIMPLICIALDATABASES,spivakDatabasesAreCategories2010}.
-Spivak formulates the fiber as a (field name, data domain) simple schema where field names and field types can be formally mapped to the set of values associated with the field type, for example $\mathbb{R}$ for a \texttt{float} column named \textit{temperature}.
-
-\begin{definition} The \textcolor{total}{total space} \dtotalc\ is the space reachable via the projection map $\pi$.
-The total space is always locally trivial, meaning $\dtotal = \dbase \times \dfiber$ over any open neighborhood $\openset_{\dbasepoint}$.
-\end{definition}
-
-The fiber bundle is trivial when $\dtotal = \dbase \times \dfiber$ is true globally; a non-trivial bundle can be thought of as the twisted fiber product $\dtotal=\dbase times_{T} \dfiber$\cite{hatcherAlgebraicTopology2002,nlab:twisting_function}.
-The twisting refers to the fiber \dfiber\ not aligning in the same direction over all $\openset \subseteq \dbase$.
-
- There is a map from a point on the base space \dbase\ which returns values in the fiber space \dfiber. This map is called a section $\dsection: \dbase \rightarrow \dtotal$ of the fiber bundle
-
-\begin{equation}
-  \label{eq:related-work:continuity:fiber-bundle}
-  \begin{tikzcd}[row sep=large]
-      \dfiberc \arrow[r, hook] & \dtotalc \arrow[d, "\pi"'] \\
-                        & \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
-  \end{tikzcd}
-\end{equation}
-
-and the set of all sections over \dbase\ is denoted $\Gamma(\dbase, \dtotal)$ .
-Sections can also be defined locally, meaning over an open set $\openset \subseteq \dtotal$ and the set of all sections over an arbitrary open set is $\Gamma(\openset, \dtotal)$.
-\begin{equation}
-  \label{eq:related-work:continuity:sections}
-  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\openset} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;\forall \dbasepointc \in \opensetc \big\}
-\end{equation}
-Butler et al proposed that \cite{butlerVisualizationModelBased1989,butlerVectorBundleClassesForm1992} sections could be used to model data sets since they are maps from a point on the continuity to a point in the fiber, and therefore they encapsulate both.
-The set of sections is the space of all data sets with the same continuity and fields.
-
-\subsection{Category Theory}
-\label{sec:related-work:equivariance:category}
-In this work, we propose that equivariance constraints can be expressed using category theory.
-Vickers et.
-al \cite{vickersUnderstandingVisualizationFormal2013} provide a brief introduction to category theory for visualization practitioners, but their work focuses on reasoning about visualization design, while this paper aims to provide guidance on designing visualization library components.
-Briefly we introduce concepts in category theory that we use to describe our model in \autoref{sec:artist:category} and \autoref{sec:artist:construction}.
-
-\begin{definition} A \textcolor{source}{category} $\textcolor{source}{\mathcal{C}}$ consists of a collection of objects $c$ with identity $id_c: C\rightarrow C$ and the set of morphisms between every two objects $f:C_1 \rightarrow C_2$, \cite{fongInvitationAppliedCategory2019,maclaneCategoriesWorkingMathematician2013}.
-\end{definition}
-The set of morphisms between objects is termed the hom set $hom_{\mathcal{C}}(C_1, C_2)$. The morphisms on the category  compose
-\begin{equation*}
-  \begin{tikzcd}
-    \textcolor{source}{C_1} \arrow[r, "f", color=source] \arrow[rd, "g\circ f"', color=source] \arrow["id_{C_1}"', loop, distance=2em, in=215, out=145, color=source] & \textcolor{source}{C_2} \arrow[d, "g", color=source] \arrow["id_{C_2}"', loop, distance=2em, in=125, out=55, color=source] \\
-  & \textcolor{source}{C_3} \arrow["id_{C_3}"', loop, distance=2em, in=305, out=235, color=source]              
-  \end{tikzcd}
-\end{equation*}
- and are constructed such that the following axioms hold \cite{riehlCategoryTheoryContext}:
- \begin{LaTeXdescription}
-   \item[associativity] if $f: C_1 \rightarrow C_2$, $g: C_2 \rightarrow C_3$ and $h: C_3 \rightarrow C_4$ then $h\circ (g \circ f) = (h \circ g) \circ f$
-   \item[identity] for every $f: C_1 \rightarrow C_2$ there exists identity morphisms $f \circ id_{C_1} = f = id_{C_2} \circ f$
- \end{LaTeXdescription}
-
- \begin{definition} An opposite category $\mathcal{C}^{op}$ is a category with all the same objects of category $\mathcal{C}$ and morphisms that are reversed.
-For example $f:C_1 \rightarrow C_2$ in $\mathcal{C}$ is $f:C_2 \rightarrow C_1$ in $\mathcal{C}^{op}$.
- \end{definition}
-
- \begin{definition} A functor is a morphism between objects of any two  categories $\textcolor{sheaf}{\bm{F}}: \textcolor{source}{\mathcal{C}} \rightarrow \textcolor{target}{\mathcal{D}}$.
-A functor has the properties of identity and composition \cite{riehlCategoryTheoryContext}\end{definition}
-
-A functor preserves the structure of the category, namely morphisms and the source and target objects of those morphisms, composition of morphisms, and identity morphisms \cite{riehlCategoryTheoryContext}.
-For example, the functor $\textcolor{functor}{F}:\mathcal{C} \textcolor{functor}{\rightarrow} \mathcal{D}$ maps an object in $\textcolor{source}{\mathcal{C}}$ into an object in $\textcolor{target}{\mathcal{D}}$
-\begin{equation}
-\label{eq:related-work:equivariance:functor_diagram}
-\begin{tikzcd}
-  \textcolor{source}{c} \arrow[d, "f"', shift right=0, color=source] \arrow[r, "F", color=functor] & \textcolor{target}{F(c)} \arrow[d, "F(f)", shift right=0, color=target] \\
-  \textcolor{source}{c^{\prime}} \arrow[r, "F"', color=functor]                       & \textcolor{target}{F(c^{\prime})}                      
-  \end{tikzcd}
-\end{equation}
- such that for every morphism $f: c \rightarrow c^{\prime} \in Hom(\mathcal{C})$ there is a compatible morphism $F(f): F(c) \rightarrow F(c^{\prime}) \in Hom(\mathcal{D})$.
-Functors $\bm{F}: \mathcal{C}^{op} \rightarrow \mathcal{D}$ are called contravariant functors.
-
- \begin{definition} A \textcolor{nattran}{natural transformation} is a morphism of functors $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\rightarrow} \textcolor{functor}{G}$ where $F, G$ are functors from $\mathcal{C}$ to $\mathcal{D}$.
-\cite{riehlCategoryTheoryContext, bradleyWhatNaturalTransformation}
- \end{definition}
-
-\begin{equation}
-  \label{eq:related-work:equivariance:natural-transform}
-\begin{tikzcd}
-  & {} \arrow[dd, "\alpha", Rightarrow, shorten=1.5em, color=nattran] &             \\
-\textcolor{source}{\mathcal{C}} \arrow[rr, "F", bend left, color=functor] \arrow[rr, "G"', bend right, color=functor] &                                     & \textcolor{target}{\mathcal{D}} \\
-  & {}                                  &            
-\end{tikzcd}
-\end{equation}
-The natural transformation composes with a functor such that the output is a compatible functor between the same categories.
-This means that for every morphism $f: c \rightarrow c^{\prime}$ between objects in $\mathcal{C}$, there is a commutating naturality square for the corresponding objects in $\mathcal{D}$ \cite{milewskiCategoryTheoryProgrammers}.
-\begin{equation}
-  \label{eq:related-work:equivariance:natural_transform_commute}
-  \begin{tikzcd}[column sep=huge]
-    \textcolor{target}{F(c)} \arrow[dd, "F(f)"', color=target] \arrow[rr, "\alpha_{c}", bend left, color=nattran] & \textcolor{source}{c} \arrow[dd, "f" description, color=source] \arrow[r, "G"', color=functor] \arrow[l, "F", color=functor] & \textcolor{target}{G(c)} \arrow[dd, "G(f)", color=target] \\
-    {} \arrow[rr, dotted, color=nattran] & & {}                      \\
-    \textcolor{target}{F(c^{\prime})} \arrow[rr, "\alpha_{c^{\prime}}"', bend right, color=nattran] & \textcolor{source}{c^{\prime}} \arrow[l, "F"', color=functor] \arrow[r, "G", color=functor] & \textcolor{target}{G(c^{\prime})}
-    \end{tikzcd}
-\end{equation}
-\note{unpack this a drop more}
-
-\subsection{Sheaves}
-\label{sec:related-work:continuity:sheaves}
-Sheaves provide a mathematical abstraction of data, the space it lives over, and relationships between subsets.
-Sheaves can be thought of as an "algebraic data structure" \cite{ghristElementaryAppliedTopology2014} for data that lives over topological spaces.
-A presheaf  $\sheaf$ applied to fiber bundles is a map from an open set in a topological space, such as $\openset \subset \dbase$ where $(\mathcal{T}, \dbase)$ is a topological space, to the set of sections of a bundle $\dtotal$ over that openset \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}.
-\begin{equation}
-  \label{eq:atct:presheaf}
-  \sheafc: \opensetc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
-\end{equation}
-The presheaf preserves the inclusion maps $\iota$ between open sets and the restriction maps $\iota^*$ between the sets of sections over that open sets. 
-\begin{equation}
-  \label{eq:related-work:continuity:presheaf}
-  \begin{tikzcd}[column sep=small]
-    {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} & {} & {\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}} \arrow[ll, "\iota^*"', hook', color=set] \\& &\\
-    \opensetc_1 \arrow[uu, "\sheafc_{\dbasec}", color=sheaf] \arrow[rr, "\iota"', hook, color=base] & {} \arrow[uu, "\sheafc_{\dbasec}" description, dotted, color=sheaf] & \opensetc_2 \arrow[uu, "\sheafc_{\dbasec}"', color=sheaf]
-    \end{tikzcd}
-\end{equation}
-
-This means that a smaller space $\openset_1$ is included in a larger space $\openset_2$ and a function in a set of global sections over a bundle $\dsection \in \Gamma(\openset_2, \dtotal\restriction_{\openset2})$ that is continuous over a larger space $\openset_2$ is continuous when restricted to a subspace $\dsection \in \Gamma(\openset_1,\dtotal\restriction_{\openset2}),\openset_1 \subset\openset_2$.
-
-A presheaf $\sheaf$ over a limit of open sets that contain a point $\dbasepoint \in \dbase$ is approximately the same as a sheaf over a point \dbasepoint\ and is called the stalk\cite{StalkSheaf2019} 
-\begin{equation}
-  \label{eq:atct:stalk}
-  \sheafc(\dbasec)\restriction_{\dbasepointc}\coloneqq \lim\limits_{\opensetc\ni \dbasepointc} \Gamma(\opensetc, \dtotalc\restriction_{\opensetc}) 
-\end{equation}
-
-Sheaves are presheafs where sections are continuous over all unions of open sets in $\dbase$\cite{harder2008lectures, spanier1989algebraic, nlab:sheaf}.
-This means that a section which is defined over a stalk $\dsection \in \sheaf(\dbase)\vert_{\dbasepoint_i}$ and evaluated at a point $\dbasepoint_j$ is equal to a section in the stalk over that point $\dsection_{\dbasepoint_i}\vert_{\dbasepoint_j} = \dsection_{\dbasepoint_j}$
-
-In category theory, a presheaf \sheafc\, is generalized to be a contravariant functor\cite{nlab:presheaf} from a category to a set. 
-\begin{equation}
-  \label{eq:actc:sheaf}
- \sheafc: \textcolor{source}{\mathcal{C}^{op}} \textcolor{sheaf}{\rightarrow} \setc
-\end{equation}
-
-Sheaves can be moved over different topological spaces if there is a map between those spaces, which we can use to describe how sections of a sheaf over one space behave when moved over to a different space and the maps between these spaces.
-Given a map $\textcolor{functor}{f}:\textcolor{source}{X} \textcolor{functor}{\rightarrow} \textcolor{target}{Y}$ between topological spaces $\textcolor{source}{X}, \textcolor{target}{Y}$ and sheaves on each space $\sheafc_{\textcolor{source}{X}}, \sheafc_{\textcolor{target}{Y}}$, we can construct a pullback functor $f^*$ and pushforward functor $f_*$ \cite{harder2008lectures} that can be composed with the sheaf functors.
-\begin{equation}
-  \label{eq:atct:pullback-pushforward}
-  \begin{tikzcd}[row sep=huge]
-    \sheafc_{\textcolor{source}{X}}(\textcolor{source}{X}) \arrow[rr, "\textcolor{functor}{f_*}", color=functor]                            &  & (\textcolor{functor}{f_*}\sheafc_{\textcolor{source}{X}})(\textcolor{target}{Y})    \\
-    \textcolor{source}{X} \arrow[rr, "\textcolor{functor}{f}", color=functor] \arrow[u, "\sheaf_{\textcolor{source}{X}}", color=sheaf] \arrow[d, dashed, color=sheaf] &  & \textcolor{target}{Y} \arrow[d, "\sheaf_{\textcolor{target}{Y}}", color=sheaf] \arrow[u, dashed, color=sheaf] \\
-    (\textcolor{functor}{f^*}\sheafc_{\textcolor{target}{Y}})(\textcolor{source}{X}) \arrow[rr, "\textcolor{functor}{f^*}", color=functor]  &  & \sheafc_{\textcolor{target}{Y}}(\textcolor{target}{Y})
-  \end{tikzcd}
-\end{equation}
-As shown in \autoref{eq:atct:pullback-pushforward}, the pushforward $f_*\sheaf_{X}$ modifies the sheaf $\sheaf_{X}$ such that sections $s \in \sheaf_{X}$ can be evaluated over points in $y$
-\begin{equation}
-  \label{eq:atct:pushforward-section}
-  (\textcolor{section}{s}: \textcolor{source}{x} \textcolor{section}{\rightarrow} \setc) \textcolor{functor}{\xrightarrow{\mathmakebox[3em]{f_*}}} (\textcolor{functor}{f_*}\textcolor{section}{s}: \textcolor{target}{y} \textcolor{section}{\rightarrow} \setc)
-\end{equation}
-which means it decorates the section function $s$ so that it can take as input values in $Y$ and return the corresponding section in $X$ 
-\begin{equation}
-  \label{eq:atct:pushforward-f}
-  \textcolor{section}{s}(\textcolor{source}{x}) 
-  = \textcolor{section}{s}\textcolor{functor}{_*}(\textcolor{target}{y}) =\textcolor{section}{s}(\textcolor{functor}{f^{-1}}(\textcolor{target}{y}))
-\end{equation}
-The pullback functor $f^*$ modifies the sheaf  $\sheaf_{Y}$ such that sections $t \in \sheaf_{Y}$ can be evaluated over points in $x$
-\begin{equation}
-  \label{eq:atct:pullback-section}
-  (\textcolor{functor}{f^*}\textcolor{section}{t}: \textcolor{source}{x} \textcolor{section}{\rightarrow} \setc) \textcolor{functor}{\xrightarrow{\mathmakebox[3em]{f^*}}} (\textcolor{section}{t}: \textcolor{target}{y} \textcolor{section}{\rightarrow} \setc)
-\end{equation}
-which means it decorates the section function $t$ so that it can take as input values in $X$ and return the corresponding section in $Y$ 
-\begin{equation}
-  \label{eq:atct:pullback-f}
-  \textcolor{section}{t}\textcolor{functor}{^*}(\textcolor{source}{x})
-  =\textcolor{section}{t}(\textcolor{functor}{f}(\textcolor{source}{x}))
-  =\textcolor{section}{t}(\textcolor{target}{y}) 
-\end{equation}
-The pushforward and pullback functors are isomorphic \cite{harder2008lectures}
-\begin{equation}
-  \label{eq:atct:pullback-pushforward-isomorphic}
-  \textcolor{homset}{hom}_{\sheafc_{\textcolor{source}{X}}}(\textcolor{functor}{f*}\sheafc_{\textcolor{target}{Y}}, \sheafc_{\textcolor{source}{X}}) = 
-  \textcolor{homset}{hom}_{\sheafc_{\textcolor{target}{Y}}}(\sheafc_{\textcolor{target}{Y}}, \textcolor{functor}{f_*}\sheafc_{\textcolor{source}{X}})
-\end{equation}
-such that the diagram is commutative
-\begin{equation}
-  \label{eq:pullback-pushforward-homset}
-  \begin{tikzcd}
-    \sheafc_{\textcolor{source}{X}}(\textcolor{source}{X}) \arrow[rr, "\textcolor{functor}{f_*}", color=functor]  &  & (\textcolor{functor}{f_*}\sheafc_{\textcolor{source}{X}})(\textcolor{target}{Y}) \\
-    && \\
-    (\textcolor{functor}{f^*}\sheafc_{\textcolor{target}{Y}})(\textcolor{source}{X}) \arrow[rr, "\textcolor{functor}{f^*}", color=functor] \arrow[uu, "\textcolor{homset}{Hom_{\sheaf_X}}", color=homset] &  & \sheafc_{\textcolor{target}{Y}}(\textcolor{target}{Y}) \arrow[uu, "\textcolor{homset}{Hom_{\sheaf_Y}}"', color=homset] \arrow[lluu, "\textcolor{homset}{Hom_{XY}}"' description, dashed, color=homset]
-    \end{tikzcd}
-\end{equation}
-meaning that transformations of sections of one sheaf to sections of another are equivalent regardless of the space the sheaf is over.
-
-\section{Artist} 
-In this section we describe the properties a data to visualization transform must satisfy to be considered equivariant and continuity preserving.
-We propose that by explicitly defining these properties, we can better incorporate them into visualization library design.
-
-We formulate the visualization transformation as a functor, which we call the \textcolor{artist}{Artist} \vartist \footnote{We call this transformation the artist because the \mintinline{python}{Artist} object in Matplotlib\cite{hunterMatplotlib2DGraphics2007}}.
-The artist \vartist\ converts \textcolor{total}{data}, which we denote $\mathcal{\dtotalc}$ to \textcolor{total}{graphics} $\mathcal{\gtotalc}$.
-We formulate this association as
-
-\begin{equation}
-  \label{eq:artist:artist}
-  \vartistc: \mathcal{\dtotalc} \rightarrow \mathcal{\gtotalc}
-\end{equation}
-
-In this section we encapsulate the structure of the data and graphic spaces as categories.
-We then use these definitions to express the structure that an artist preserves.
-Finally, we introduce rules for composing artists in a way where the structure between the inputs to the individual artists is preserved.
-
-\subsection{Continuity}
-\label{sec:artist:category}
-
-Fiber bundles, as introduced in \autoref{sec:related-work:fiber-bundles}, serve as the model of data and graphic spaces in out model.
-In this section, we introduce categorical formulations of the topological spaces that make up a fiber bundle.
-
-We introduce the category $\mathcal{\dbase}$ to encapsulate the continuity of the data, for example if the data are discrete points, continuous lines, images, or networks. 
-\begin{definition} \label{def:category:K} The category of open sets $\mathcal{\dbasec}$ consists of
-  \begin{itemize}
-    \item \textit{objects} open sets $\openset_{i} \in \{\openset\}$ in the topological space $(\dbase, \mathscr{T})$), including the empty set $\varnothing$ and the maximum set \dbase. 
-    \item \textit{morphisms} $\iota: \openset_{i} \hookrightarrow \openset_{j}$ for all  $\openset_i, \openset_j \subset \dtotal$
-  \end{itemize} 
-\end{definition}
-
-We also introduce an analogous category $\mathcal{\gbase}$ that encapsulates the topology of the graphic, which is the topology of the rendering in the display space; for example, this rendering can be a 2D or volumetric image or a 3D print.
-
-\begin{definition} The category of open sets $\mathcal{\gbasec}$ in the topological space $(\dbase, \mathscr{T})$ consists of open set \textit{objects} $\opensetg \in \{\opensetg\}$, including the empty set $\varnothing$ and the maximum set $\gtotal$ and \textit{morphisms} $\iota: \opensetg_i \rightarrow \opensetg_j$.
-\end{definition}
-
-The data space \dbase\ is a deformation retract of the graphic topology \gbase\meaning that regions in \gbase\ collapse to points in \dbase\ such that the relative positions of the points are preserved \cite{nlab:deformation_retract}.
-One method of constructing the graphic space is by expanding the dimensionality of the data \dbase\ to match the spatial dimensionality of the rendering by multiplying the data space by  intervals $I$.
-
-\begin{figure}[h!]
-  \label{fig:artist:gbase}
-  \includegraphics[width=\columnwidth]{deform_retract.png}
-  \caption{Given a 2D screen display, the graphic space \gbase\ must be 2D.
-It is constructed by multiplying the line \dbase\ by the interval [0,1] such that every vertical band in \gbase\ corresponds to a unique point $\dbasepoint \in \dbase$.}
-\end{figure}
-
-For example, as illustrated in \autoref{fig:artist:gbase}, the display space is a 2D screen and the data space is a 1D line \dbase.
-For that line to have any thickness, it must be rendered as a ban.
-One method of obtaining this graphic continuity while preserving the constraint that it is a deformation retract is by multiplying the base space \dbase\ by an interval $I=\[0,1\]$ such that $\gbase = \dbase \times I$.
-This construction ensures that every region $(\dbasepoint, \alpha)\vert_{\alpha \in I} \subset \gbase$ maps to a  corresponding point $\dbasepoint \in \dbase$.
-
-\subsubsection{Data}
-\note{clean up why sheaf}
-As mentioned in \autoref{sec:related-work:continuity:sheaves}, sheaves are a powerful tool for keeping track of data that lies over topological spaces.
-Therefore, we model the type of data an artist expects as a sheaf $\sheafc_{\dtotalc}: \mathcal{\dbasec}^{op} \textcolor{sheaf}{\rightarrow} \setc$.
-\begin{equation}
-  \label{eq:artist:morphism-of-sheaf-data}
-  \begin{tikzcd}
-    & \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \in Ob(\setc) & \\
-    {} & & {} \\
-    & \opensetc \in Ob(\mathcal{\dbasec}^{\textcolor{base}{op}}) \arrow[uu, "\sheafc_{\dtotalc}"', shift left = 10, color=sheaf] &   
-\end{tikzcd}
-\end{equation}
-The objects of \textbf{Set} in this sheaf are the sets of sections $\Gamma(\dbase, \dtotal)$ introduced in \autoref{eq:related-work:continuity:sections} and the morphisms are the inclusion maps $(\iota, \iota^*)$ introduced in \autoref{eq:related-work:continuity:presheaf}.
-
-To establish a mapping between graphic and data base spaces, we introduce the surjective map $\vindex: \gbase \rightarrow \dbase$. 
-\begin{equation}
-  \label{eq:artist:vindex}
-  \begin{tikzcd}
-    \dtotalc& \dtotalpullc \arrow[l, "\vindexpullc"', dashed, color=artist]                                                                           \\
-    \dbasec \arrow[u, "\dsectionc", color=section] & \gbasec \arrow[l, "\vindexc", color=artist] \arrow[lu, "\dsectionc \circ \vindexc" description, color=section] \arrow[u, "\dsectionpullc"', dashed, color=section]
-    \end{tikzcd}
-\end{equation}
-such that the data continuity $\dbase$ is a deformation retract of $\gbase$, as expressed in \autoref{eq:artist:gbase}.
-
-As shown in \autoref{eq:artist:vindex}, the map between spaces $\vindex$ can be composed with section map $\dsection$ such that there is a map from the graphic base space $\gbase$ to the fiber bundle in which the data lives $\dtotal$.
-\begin{figure}[h!]
-  \includegraphics[width=\columnwidth]{pullback.png}
-  \label{fig:artist:pushforward_xi}
-  \caption{The map $\vindexpull$ is a map from a repetition of the 1D fiber over every point $\dbasepoint \in \dbase$ in data space, turning the fiber into a 2D plane of repeating values over the corresponding region $\dbasepoint = \vindex(\gbasepoint)$ in the graphic space \gbase.
-}
-\end{figure}
-Since there is a mapping from the graphic space to the total bundle $\dsection \circ \vindex: \gbase \rightarrow \dtotal$, the mapping $\vindex$ can be pulled through the composition such that there is a map $\vindexpull$ between a total space over graphics $\dtotalpull$ and its equivalent space over data $\dtotal$.
-As illustrated in the example given in \autoref{fig:artist:pushforward_xi}, a bundle $\dtotal$ over a point $\dbasepoint \in \dbase$ is pulled back through \vindex\ \cite{nlab:pullback_bundle} such that it is copied over every point in the corresponding region in graphic space $\vindex(\gbasepoint) = \dbasepoint, \gbasepoint \subset \gbase$.
-The pullback $\vindexpull$ can then be composed with the elements of the set of sections
-\begin{equation}
-  \label{eq:artist:data_gamma_pullback}
-  \vindexpullc: \cgamma{\dbasec}{\dtotalc} \textcolor{artist}{\rightarrow} \cgamma{\gbasec}{\dtotalpullc}
-\end{equation}
-to generate a set of equivalent sections over the graphic space $\Gamma(\gbase, \dtotalpull)$.
-As a consequence of \autoref{eq:artist:data_gamma_pullback}, there exists a data sheaf over graphic space
-\begin{equation}
-  \label{eq:artist:data_pullback_sheaf_gamma}
-\sheafc_{\dtotalpullc}: \opensetgc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetgc}{\dtotalpullc}}
-\end{equation}
-which is equivalent to 
-\begin{equation}
-  \label{eq:artist:data_pullback_sheaf}
-  \sheafc_{\dtotalpullc}: \mathcal{\gbasec}^{\textcolor{base}{op}} \textcolor{sheaf}{\rightarrow} \setc
-\end{equation}
-
-\subsubsection{Graphic}
-We model the space of all graphic generating functions over a specific graphic topology $\opensetg \subseteq \gbase$ and targeting a specific display as a set of sections 
-\begin{equation}
-  \label{eq:artist:graphic_sections}}
-    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \coloneqq \big\{\gsectionc: \opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} \bigm{\vert} \pi(\gsectionc(\gbasepointc)) = \gbasepointc \; \forall \gbasepointc \in \opensetgc \big\}
-\end{equation}
-
-The set of sets of sections $\{\Gamma(\opensetg, \gtotal)\}\vert_{{\opensetg \in gbase}}$ is subcategory of \textbf{Set}; therefore the full space of graphics is a sheaf 
-\begin{equation}
-  \label{eq:artist:graphic_sheaf_gamma}
-  \sheafc_{\gtotalc}: \opensetgc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetgc}{\gtotalc}
-\end{equation}
-which means it is 
-\begin{equation}
-  \label{eq:artist:graphic_sheaf_set}
-  \sheafc_{\gtotalc}: \gbasec^{\textcolor{base}{op}} \textcolor{sheaf}{\rightarrow} \setc
-\end{equation}
-
-\subsubsection{Data to Graphic}
-Following from \autoref{eq:artist:data_pullback_sheaf} and \autoref{eq:artist:artist_sheaf},  we propose that the artist function $\vartist$ is a natural transformation of sheaf functors 
-\begin{equation}
-  \label{eq:artist:natural-transform-diagram}
-  \begin{tikzcd}
-    & {} \arrow[dd, "\vartistc", Rightarrow, shorten <=1.35em,  shorten >=1.25em, color=artist] & \\
-\mathcal{\gbasec}^{\textcolor{base}{op}} \arrow[rr, "\sheafc_{\dtotalpullc}", bend left, color=sheaf] \arrow[rr, "\sheafc_{\gtotalc}"', bend right, color=sheaf] &  & {\setc}\\
-    & {} &                            
-\end{tikzcd}
-\end{equation}
-
-where each sheaf is a map from an open set in the graphic base space $\opensetg \subset \gbase$ into a set of sections of a fiber bundle over that open set.
-As defined in \autoref{eq:related-work:equivariance:natural-transform}, the artist is a natural transformation.
-Following from \autoref{eq:related-work:equivariance:natural_transform_commute}, the artist specifies that there are compatible inclusion maps in the data and graphic spaces
-
-\begin{equation}
-  \begin{tikzcd}
-    \cgamma{\opensetgc_1}{\dtotalpullc} \arrow[dd, "\iota^*"', hook, color=set] \arrow[rrrr, "\vartistc_{\opensetgc_1}", bend left, color=artist] &  & \opensetgc_1 \arrow[ll, "\sheafc_{\dtotalpullc}", color=sheaf] \arrow[rr, "\sheafc_{\gtotalc}"', color=sheaf] &  & \cgamma{\opensetgc_1}{\gtotalc} \arrow[dd, "\iota^*", hook', color=set] \\
-    {} \arrow[rrrr, dotted, color=artist] &&&& {} \\
-    \cgamma{\opensetgc_2}{\dtotalpullc} \arrow[rrrr, "\vartistc_{\opensetgc_2}"', bend right, color=artist]                            &  & \opensetgc_2 \arrow[ll, "\sheafc_{\dtotalpullc}"', color=sheaf] \arrow[rr, "\sheafc_{\gtotalc}", color=sheaf] \arrow[uu, "\iota" description, hook, color=base] &  & \cgamma{\opensetgc_2}{\gtotalc}
-    \end{tikzcd}
-\end{equation}
-Given that $\opensetg \subseteqq \gbasec$, \autoref{eq:artist:inclusion_morphisms} can be generalized to sections over $\gbase$.
-This fact can be combined with \autoref{eq:artist:data_gamma_pullback} to describe the path from data sections over data space to graphic sections over graphic space
-\begin{equation}
-  \label{eq:artist:gamma_map}
-  \cgamma{\dbasec}{\dtotalc} \textcolor{artist}{\xrightarrow{\vindexc^*}} \cgamma{\gbasec}{\dtotalpullc} \textcolor{artist}{\xrightarrow{\vartist}} \cgamma{\gbasec}{\gtotalc}
-\end{equation}
-
-which means that the artist can be pulled back through $\vindex^*$\footnote{Definition 3.68 in Seven Sketches in Compositionality by Fong and Spivak}
-\begin{equation}
-  \begin{tikzcd}
-    & & {} \arrow[dd, "\vartistc" description, Rightarrow, shorten <=1.35em,  shorten >=1.25em, color=artist] &       \\
-\mathcal{\dbasec}^{\textcolor{base}{op}} \arrow[r, "\vindexc^*", color=artist] & \mathcal{\gbasec}^{\textcolor{base}{op}} \arrow[rr, "\sheafc_{\dtotalpullc}", bend left, color=sheaf] \arrow[rr, "\sheaf_{\gtotalc}"', bend right, color=sheaf] &                                                    & \setc \\
-    &                                                                                                                                                         & {}                                                 &
-\end{tikzcd}
-\end{equation}
-such that the artist can be constructed over data space 
-\begin{equation}
-  \begin{tikzcd}
-    & {} \arrow[dd, "\vartistc_{\vindexc^*}" description, Rightarrow, shorten <=1.35em,  shorten >=1.25em,, color=artist] &       \\
-\mathcal{\dbasec}^{\textcolor{base}{op}} \arrow[rr, "\sheafc_{\dtotalpullc} \circ\vindexpullc = \sheafc_{\dtotalc}", bend left, color=sheaf] \arrow[rr, "\sheafc_{\gtotalc}"', bend right, color=sheaf] &                                                                 & \setc \\
-    & {}                                                              &      
-\end{tikzcd}
-\end{equation}
-
-which in turn provides a way to construct the artist as a map from data section in data space to corresponding graphic section
-
-\subsection{Equivariance}
-Besides preserving inclusions, as described in \autoref{eq:artist:inclusion_morphisms}, the artist function is constructed such that transformations on $\dsection$ and $\gsection$ are equivariant with respect to the morphism between data sheaf functors $\dfunc$ which was introduced in \autoref{eq:artist:morphism-of-sheaf-data}.
-
-\begin{equation}
-  \label{eq:artist:morphism-of-sheaf-data}
-  \begin{tikzcd}
-    & \cgamma{\opensetc_i}{\dtotalc} \in Ob(\setc) & \\
-{} \arrow[rr, "\dfuncc", Rightarrow, pos=.3, shorten <=3.5em, shorten >=8.85em, color=action] & & {} \\
-    & \opensetc_i \in Ob(\mathcal{\dbasec}^{\textcolor{base}{op}}) \arrow[uu, "\sheafc(\dtotalc)", bend left, shift left = 10.5, color=sheaf] \arrow[uu, "\sheafc(\dtotalc)"', bend right, shift left=10.5, color=sheaf] &
-\end{tikzcd}
-\end{equation}
-
-The map $\dfunc: \sheaf(\dtotal) \rightarrow \sheaf(\dtotal)$ is any arbitrary morphism between sheaf functors $\sheaf(\dtotal)$.
-Following from \autoref{eq:related-work:equivariance:natural-transform}, \dfunc\ is a natural transform, which means all inclusions $\iota, \iota^*$ are preserved as part of any $\dfunc$ transformation.
-
-\begin{definition} 
-  \label{def:category:F} The fiber category $\mathcal{\dfiberc}$ is a monoidal category\cite{milewskiCategoryTheoryProgrammers}, which means it is a category equipped with a bifunctur $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$
-  \begin{itemize}
-    \item \textit{object} a monoid object \dfiber. A monoid is a set with a binary operator that is associative, closed, and for which the set contains an identity element $i$ \cite{nlan:monoid}.
-    \item \textit{morphisms} $\dfiber \otimes \dfiber \rightarrow \dfiber$ and unit $i \rightarrow \dfiber$
-  \end{itemize}
-\end{definition}
-
-The bifunctor $\otimes$ combines categories in a way that  preserves identity and composition \cite{fongInvitationAppliedCategory2019}.
-For example, given a pair of categories $\mathcal{\dfiber_a},\mathcal{\dfiber_b}$, the bifunctor yields a set of pairs $\{(\dfiber_a, \dfiber_b) \big{\mvert} \dfiber_a \in \mathcal{\dfiber_a}, \dfiber_b \in \mathcal{\dfiber_b}\}$.
-For the associated morphisms $g: \mathcal{\dfiber_a}\rightarrow \mathcal{\dfiber_a}$ and $h: \mathcal{\dfiber_b}\rightarrow \mathcal{\dfiber_b}$, the function $(g\times h): (\mathcal{\dfiber_a}\times \mathcal{\dfiber_b}) \rightarrow (\mathcal{\dfiber_a}\times \mathcal{\dfiber_b})$ is pointwise $(g\times h)(\dfiber_a, \dfiber_b)\coloneqq (g(\dfiber_a), h(\dfiber_b))$.
-Because of the bifunctor, objects of the category $\mathcal{\dfiber}$ can be the product of any arbitrary number and types of categories, including \textbf{Set}, \textbf{Graph}, and topological spaces \textbf{Top}.
-
-The set of all morphism of the fiber category $Hom_{\mathcal{\dfiber}}(\dfiber, \dfiber)$ are all possible morphisms in a given category.
-This includes any of the functions that serve as the basis of equivariance, such as the binary operations, group actions, and measurement scale discussed in  \autoref{sec:related-work:equivariance}.
-The generalization to $Hom_\mathcal{\dfiber}$ also allows for the inclusion of structures such as monoid actions\cite{barrCategoryTheoryComputing}, which provide a way of applying partial order relations to data\cite{fongInvitationAppliedCategory2019}, such as to build multi-ranked indicators\cite{bruggemannRankingPrioritizationMultiindicator2011}.
-
-When a fiber bundle is trivial, we can dispense with localization and directly consider the fiber maps $Hom(\dfiber_{\dbasepoint}, \dfiber_{\dbasepoint})$ over a point \dbasepoint\ as the basis of our equivariance.
-The fiber space over a point is embedded in the stalk, introduced in
-
-over that point $\dfiber_{\dbasepoint}\hookrightarrow \mathscr{\dfiber}_{\dbasepoint}$.
-The sheaf maps \dfunc\ restricted to the stalk are members of $hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$.
-The maps $\dfunc \in Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ include the types of data transformations described in \autoref{sec:related-work:equivariance}In \autoref{sec:artist:construction}, we assume that the fiber bundles are trivial.
-
-\note{}
-
-It follows from \autoref{eq:related-work:equivariance:functor_diagram} that an arbitrary morphism $\dfunc$ on the data sheaf has a compatible morphism $\dfunc^{\prime}$ on the graphic sheaf. 
-\begin{equation}
-  \label{eq:artist:morphism_of_functors}
-  \begin{tikzcd}
-    \sheafc_{\dtotalc} \arrow[d, "\dfuncc"', color=action] \arrow[r, "\vartistc_{\vindexpullc}", color=artist] & \sheafc_{\gtotalc} \arrow[d, "\dfuncpc \coloneqq\vartistc_{\vindexpullc}(\dfuncc)", color=action] \\
-    \sheafc_{\dtotalc} \arrow[r, "\vartistc_{\vindexpullc}"', color=artist]                      & \sheafc_{\gtotalc}     
-    \end{tikzcd}
-\end{equation}
-As a consequence of \autoref{eq:artist:morphism_of_functors},
-
-\subsubsection{Equivariant Artist}
-
-\note{valid $\rho$ is determined by equivariance, which we haven't introduced yet, but we can introduce equivariance w/ just $\sheaf_{\gtotal}$}
-\begin{equation}
-  \label{eq:artist:artist_sheaf}
-  \begin{tikzcd}[column sep=tiny, row sep=huge]
-    {\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}}&  & 
-    \textcolor{set}{\big\{}\gsectionc:\opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} 
-    \bigm{\vert} \vartistc\circ (\dfuncc \circ \vartistc^{-1}[\gsectionc]) = \dfuncc^{\prime} \circ \gsectionc \textcolor{set}{\big\}} 
-    \arrow[ll, "\textcolor{set}{\mathlarger{\mathlarger{\supset}}}" description, hook', color=white] \\ &  & \\
-    \opensetgc \arrow[uu, "\sheafc_{\gtotalc}", color=sheaf] \arrow[rruu, "\vartistc(\sheafc_{\dtotalc})\coloneqq \sheafc_{\vartistc}" description, dashed, color=sheaf] &  & 
-    \end{tikzcd}
-\end{equation}
-
-The set of graphics reachable through an artist function is a subset of sections of the full set of graphics $\{\gsection\} \subset \Gamma(\opensetg, \gtotal\restriction_{\opensetg})$.
-These sets of reachable graphics over open sets \opensetg\ are also objects of \textbf{Set}.
-This allows us to formulate the graphic output by the artist as a sheaf
-\begin{equation}
-  \label{eq:artist:artist_sheaf}
-  \sheafc_{\vartist}: \gbasec^{\textcolor{base}{op}} \textcolor{sheaf}{\rightarrow} \setc
-\end{equation}
-from the graphic base space to the set of reachable graphics.
-
-\note{have I said anything about stalks and fibers yet? this may need to be moved down...}
-
-\begin{align}
-  \label{eq:artist:section_map}
-  \vartistc_{\vindexpullc}:\{\dsectionc: \dbasepointc \rightarrow \dfiberc_{\dbasepointc} \mid \dbasepointc \in \opensetc \subseteq \dbasec\} &\rightarrow \{\gsectionc: \gbasepointc \rightarrow \gfiberc\restriction_{\gbasepointc} \mid \gbasepointc = \opensetgc \subseteq \gbasec \}
-\end{align}
-where, as defined in \autoref{eq:artist:construction:xi}, $\vindex(\gbasepoint) = \dbasepoint$ and \gsection\ is in the set of reachable sections $\gsection \in \sheaf_{\vartist}(\gbase)$
-
-the artists constructs a corresponding graphic generating function. 
-\note{maybe mention here Munzner, Zeminsky, and the other dude on glyphs map back to data}
-\note{possibly more here as this is the type signature of the thing we implement}
-
-\subsection{Composition of Artists}
-\label{sec:artist:union}
-Visualizations generally consist of more than one visual element, for example line plots, a legend, axis ticks, spines, and labels.
-We propose that we can express expected consistency between these elements as a preservation of morphisms between objects of different data categories $\mathcal{\dtotal}$.
-
-\subsubsection{Shared Index}
-\begin{figure}[h!]
-  \includegraphics[width=\columnwidth]{k_union.png}
-  \label{fig:artist:union:disjoint_k}
-  \caption{In \autoref{fig:artist:union:k}, the input object $\sheaf_{\dtotal}$ encodes a continuous function over spaces $\dtotal_a$, $\dtotal_b$ 
-  and overlapping space $\dtotal_c$.
-In this figure, the visual transformation is to a point on the line on screen.
-An artist that preserves the continuity of the function must generate graphics such that functions evaluated  $\dbasepoint \in \dbase_a \cap \dbase_b$}
-\end{figure}
-
-An example of overlapping indexing is a parallelized version of a sliding window algorithm where window overlaps must resolve to the same value at the same position \cite{chuTimeSeriesSegmentation1995};
-
-\begin{equation}
-\label{eq:artist:intersection:basespace}
-\begin{tikzcd}
-  \dbasec_{c} \arrow[d, "\iota", hook] \arrow[r, "\iota", hook] & \dbasec_{b} \arrow[d, "i_{\dbasec_b}"]      \\
-  \dbasec_{a} \arrow[r, "i_{\dbasec_a}"]                         & \dbasec_{a} \bigsqcup_{\dbasec_{c}}\dbasec_{b}
-  \end{tikzcd}
-\end{equation}
-
-where the projection functions $i_{\dbase_a},i_{\dbase_b}$ behave such that $i_{\dbase_a}(\dbasepoint)\vert_{\dbasepoint \in \dbase_b} = \left[\dbasepoint\right]$, $i_{\dbase_b}(\dbasepoint)\vert_{\dbasepoint \in \dbase_b} = \left[\dbasepoint\right]$, meaning the projection functions yield equivalent points in the disjoint union $\dbase_a \bigsqcup_{\dbase_c} \dbase_b$.
-Data that is completely disjoint in the index, as shown in \autoref{fig:artist:union:disjoint_k}, is a special case where $\dbase_c=\varnothing$ and therefore there is no constraint on the artist with regards to how this data is rendered.
-
-\subsubsection{Shared Fields}
-
-\begin{figure}[h!]
-  \includegraphics[width=\columnwidth]{intersection_f.png}
-  \caption{The fiber spaces $F_a, F_b$ have a shared fiber space $F_c$.
-If they are input into the same artist $A_1$, then an equivariant transformation is one where $F_c$ is transformed to a visual element in a consistent matter.
-In this figure, $F_c$ is mapped to the x-axis for both $F_a$ and $F_b$.}
-  \label{fig:artist:compose:union_fiber}
-\end{figure}
-
-\begin{equation}
-  \label{eq:artist:intersection:fiber}
-  \begin{tikzcd}
-    \dfiberc_{a}\times \dfiberc_{b} \arrow[d, "proj_b"] \arrow[rr, "proj_a"] &  & \dfiberc_{a} \arrow[d, "proj_c"] \\
-    \dfiberc_{b} \arrow[rr, "proj_b"]                                                  &  & \dfiberc_{c}                    
-    \end{tikzcd}
-\end{equation}
-%%% the fiber objects need to be composable into product categories, coproduct? -> OR on return values 
-\autoref{eq:artist:intersection:basespace} and \autoref{eq:artist:intersection:fiber} can be composed to specify how different aspects of the structure of a multivariate nested dimensional dataset, such as a spatio-temporal weather dataset, should be preserved in a visualization.
-\note{this is probably the power of this model and should maybe get a figure?}
-
-\section{Construction and formal properties of Artists}
-\label{sec:artist:construction}
-\begin{figure*}[h!]
-  \includegraphics[width=\linewidth]{q.png}
-  \caption{}
-  \label{fig:constraints:q-overall}
-\end{figure*}
-
-Following from \autoref{eq:artist:equivariance}, the artist \vartistc\ is a natural transformation
-  \begin{equation}
-    \vartist: \sheaf_{\mathcal{\dtotal}}
-  \end{equation}
-that we construct to preserve continuity and equivariance.
-
-\begin{definition} We define the natural transformation \textcolor{artist}{Artist}  \vartist\ as the tuple $(\vindexc, \vchannelc, \vmarkc, \mathcal{\dtotalc},\mathcal{\vtotalc}, \mathcal{\gtotalc})$ where
-  \begin{enumerate}
-    \item $\vchannelc: \dtotalc \rightarrow \vtotalc$ is a bundle map from data values to the visual variables they are mapped to
-    \item $\vmarkc: \sheafc_{\mathcal{\vtotalc} \rightarrow \sheafc_{\mathcal{\vartistc}}$ is a sheaf map that builds a graphic generating function parameterized by the visual variables 
-    \item $\vindexc: \gbasec \rightarrow \dbasec$ is a surjective map from the graphic topological base to the data topological base
-  \end{enumerate}
-\end{definition}
-and $\mathcal{\dtotalc}$, $\mathcal{\vtotalc}$, and $\mathcal{\gtotalc}$ are the categorical representations of fiber bundles that model the data, visual variable, and graphic space of the data to visual transformation.
-We construct the artist to have two stages
-
-\begin{equation}
-  \label{eq:artist:construction:all}
-  \begin{tikzcd}
-    \dtotalc \arrow[r, "\vindexc"] \arrow[dd, "\pi"'] & \vtotalc \arrow[ldd, "\pi"] & \mathcal{\vtotalc}                                                        & \gtotalc                                           \\
-    &                             & {} \arrow[r, "\vmarkc"]                                                   & {}                                                 \\
-    \dbasec                                           &                             & \mathcal{\dbasec} \arrow[uu, "\sheaf_{\vtotalc}"] \arrow[r, "\vindexc^*"] & \mathcal{\gbasec} \arrow[uu, "\sheaf_{\vartist}"']
-    \end{tikzcd}
-\end{equation}
-which are the data to visual variable map $\vchannel$ at the bundle level and the visual variable to graphic map $\vmark$ at the sheaf level.
-Usage of the bundle directly in an equation or diagram - $\dotal, \vtotal, \gtotal$ - denotes that the function can be evaluated pointwise over $\dbasepoint \in \dbase$.
-Use of the categorical form -- $\mathcal{\dtotalc}, \mathcal{\vtotalc}, \mathcal{\gtotalc}$ -- denotes that the function is evaluated over an openset object, meaning that the function needs knowledge of both a value over a point and some information about neighboring values.
-
-\subsection{Data Domain}
-We model data as sections of a fiber bundle $(\dtotal, \pi, \dbase, \dfiber)$.
-We encode the continuity of the data as the \textcolor{base}{base space} $\dbase$.
-We model the data as the section \dsection\ because, as described in \autoref{sec:related-work:fiber-bundles}, the section is the map from the indexing space \dbase\ to the space of possible data values \dfiber.
-
-\begin{figure}[h!]
-  \includegraphics[width=\columnwidth]{fiberbundle.png}
-  \caption{\note{replace with more concrete}}
-  \label{fig:artist:data}
-\end{figure}
-
-One example of encoding data as a section of a fiber bundle is illustrated in \autoref{fig:artist:data}.
-In this example, the data is a \note{not totally decided yet} table of weather station data.
-Here we are interested in the time series of temperature values in the data, so we encode the continuity as the 1D interval \dbase.
-In this multivariate data set, the fields we want to visualize are \texttt{time}, \texttt{temperature}, and \texttt{station}.
-The fiber space \dfiber\ is the cartesian cross product of the fibers of each field
-\begin{equation*}
-  \dfiber = \dfiber_{time} \times \dfiber_{temperature} \times \dfiber_{station}
-\end{equation*}
-where each field fiber is the set of values that are valid for the field: 
-\begin{align*}
-  \dfiber_{time} &= \mathbb{R} \\
-  \dfiber_{temperature} &= \mathbb{R}\\
-  \dfiber_{station} &= \{s_0, s_1, \cdots, s_i, \cdots, s_n\} 
-\end{align*}
-
-The section \dsection\ is the abstraction of the data being visualized. The section at a point $k \in K$ in the base space returns a value from each field in the fiber.
-\begin{equation*}
-  \tau(k) = ((time, t_k); (precipitation, p_k); (station\;name, n_k))
-\end{equation*}
-
-\subsection{Graphic Codomain} 
-The object of the graphic category $\mathcal{\gtotal}$ is the fiber bundle \gtotal. The bundle \gtotal\ has the same structure as the data bundle \dtotal
-
-\begin{equation}
-  \begin{tikzcd}[ampersand replacement=\&]
-      \gfiber \arrow[r, hook] \& \gtotal \arrow[d, "\pi"'] \\
-                        \& \gbase \arrow[u, "\gsection"', bend right]
-  \end{tikzcd}
-\end{equation}
-with a fiber space \gfiber\ embedded in the total space \gtotal\ and a section map $\gsection:\gbase\rightarrow\gtotal$.
-The attributes of the graphic bundle $(\gtotal, \pi, \gfiber, \gbase)$ encode attributes of the graphic and display space
-
-\begin{LaTeXdescription}
-\item [\textcolor{base}{base space} \gbase] continuity of display space (e.g. screen, 3D print)
-\item [\textcolor{fiber}{fiber space} \gfiber] attributes of the display space (e.g a pixel = (x,y,r,g,b,a))
-\item [\textcolor{section}{section} \gsection] graphic generating function
-\end{LaTeXdescription}.
-
-We represent the graphic output as a fiber bundle because it is an abstraction that is generalizable to various output mediums (screens, 3D prints) and types of graphics.
-
-\begin{figure}[h!]
-  \includegraphics[width=\columnwidth]{render.png}
-  \caption{}
-  \label{fig:artist:graphic}
-\end{figure}
-
-As illustrated in \autoref{fig:artist:graphic}, in this work, \gtotal\ assumes the the display is an idealized 2D screen.
-\note{include some more about the figure} The graphic section \gsection\ is an abstraction of rendering.
-For example, \gsection\ can be a specification such as PDF\cite{bienz1993portable}, SVG\cite{quintScalable2003} or an OpenGL scene graph\cite{CarsonOpenGL1997}, or a rendering engine such as Cairo\cite{CairographicsOrg} or AGG\cite{shemanarevAntiGrainGeometry}.
-
-\subsection{Visualization Library Components}
-
-\subsubsection{Visual Bundle \vtotal}
-\begin{equation}
-  \begin{tikzcd}[ampersand replacement=\&]
-      \vfiberc \arrow[r, hook] \& \vtotalc \arrow[d, "\pi"'] \\
-                        \& \dbasec \arrow[u, "\vsectionc"', bend right]
-  \end{tikzcd}
-\end{equation}
-
-\subsubsection{Data to Visual Encodings: \vchannel} 
-\note{maybe figure out how to put these side by side nicely}
-
-\begin{equation}
-  \label{eq:artist:construction:nu}
-  \begin{tikzcd}
-    \dtotalc \arrow[d, "\pi"'] \arrow[r, "\vchannelc"] & \vtotalc \arrow[ld, "\pi"] &  & \dfiberc \arrow[r, "\vchannelc"]                          & \vfiberc \\
-    \dbasec                                            &                            &  & \dbasec \arrow[u, "\dsectionc"] \arrow[ru, "\vsectionc"'] &         
-    \end{tikzcd}
-\end{equation}
-Since the functor \vindex\ is bundle wise, it can be evaluated at the bundle at a point, which is the fiber space; therefore \vindex\ is a functors from the product category $\mathcal{\dfiber}$ to the product categrory $\mathcal{\vfiber}$.
-This constraint is expressed in our construction of \vchannel
-
-which means equivariance of 
-\begin{equation}
-\label{eq:artist:construction:nu:equivariance}
-\begin{tikzcd}
-  \dsectionc \arrow[d, "\dfuncc"'] \arrow[r, "\vchannelc"] & \vsectionc \arrow[d, "\vchannelc_{\dfuncc}"] \\
-  \dsectionc^{\prime} \arrow[r, "\vchannelc"]                       & \vsectionc^{\prime}                                 
-  \end{tikzcd}
-\end{equation}
-
-Since the functor $\nu$ acts on product categories, it can be decomposed into $nu_i$ component functors that act on corresponding sections $\dsection_i$ of the fiber $\dfiber_i$.
-
-\begin{equation}
-  \label{eq:math:artist:nu}
-  \{\vchannelc_{0}, \ldots, \vchannelc_{n}\}: \{\dsectionc_{0}, \ldots, \dsectionc_{n}\} \mapsto \{\vsectionc_{0}, \ldots, \vsectionc_{n}\}
-\end{equation}
-
-One example of this is
-
-\begin{figure}[h!]
-  \centering
-  \subfloat[\note{insert title}]{\includegraphics[width=1.2in]{partial_fixed.png}%
-  \label{fig_first_case}}
-  %\hfill
-  \subfloat[\note{insert title}]{\includegraphics[width=1.2in]{partial_invalid.png}%
-  \label{fig_second_case}}
-  \caption{is a weird nu, colors are a but much}
-  \label{fig_sim}
-\end{figure}
-
-\paragraph{Multiview Constraints}
-\begin{figure}[h!]
-  \centering
-  \subfloat[No shared fields]]{\includegraphics[width=.4\columnwidth]{exploding_artist.png}%
-  \label{fig:artist:union:k}}
-  %\hfill
-  \subfloat[Shared fields]{\includegraphics[width=.4\columnwidth]{combined_artist.png}%
-  \label{fig:artist:union:disjoint_k}}
-  \caption{In \autoref{fig:artist:union:k}, the input object $\sheaf_{\dtotal}$ encodes a continuous function over spaces $\dtotal_a$, $\dtotal_b$ 
-  and overlapping space $\dtotal_c$.
-In this figure, the visual transformation is to a point on the line on screen.
-An artist that preserves the continuity of the function must generate graphics such that functions evaluted  $\dbasepoint \in \dbase_a \cap \dbase_b$
-  \note{replace a) w/ fiber cross diagram? - use first half introduced in \autoref{fig:artist:compose:union_fiber}}}
-\end{figure}
-
-The concept that shared data fields should be encoded visually in a consistent is formally expressed by Qu and Hullman \cite{hullmanKeeping2018} as the notion that the same field should have the same scale.
-
-We enforce the equivariance constraint specified in \autoref{eq:artist:construction:multi_nu} by constructing artists that apply \vchannel\ encoders such that 
-\begin{equation}
-  \label{eq:artist:construction:multi_nu}
-  \begin{tikzcd}[row sep=.5em]
-    \dfiberc_a \arrow[rrd, "proj_a" description] \arrow[rrrr, "\vchannelc_c"]  &  &  &  & \vfiberc_c \\
-    &  & \dfiberc_{c} \arrow[rru, "\vchannelc_c" description] \arrow[rrd, "\vchannelc^{\prime}_c" description] &  & \\
-    \dfiberc_b \arrow[rru, "proj_b" description] \arrow[rrrr, "{\vchannelc_c}^{\prime}"'] &  &                                                                                                       &  & {\vfiberc_c}^{\prime}
-    \end{tikzcd}
-\end{equation}
-
-\subsubsection{Visual to Graphic: \vmark}
-
-\begin{equation}
-  \label{eq:artist:construction:q}
-  \begin{tikzcd}
-    \mathcal{\vtotalc}                                & \gtotalc                                           \\
-    {} \arrow[r, "\vmarkc"]                           & {}                                                 \\
-    \mathcal{\dbasec} \arrow[uu, "\sheaf_{\vtotalc}"] & \mathcal{\gbasec} \arrow[uu, "\sheaf_{\vartist}"']
-    \end{tikzcd}
-\end{equation}
-
-\begin{figure}[h!]
-  \includegraphics[width=\columnwidth]{diff_type_q.png}
-  \caption{rework this as a commutative box w/ the r in E row associated w/ this qhat(k)}
-\end{figure}
-
-\begin{equation} 
-  \label{eq:artist:construction:latent_vars}
-\begin{tikzcd}
-  \mathcal{\vfiberc_a} \arrow[rrr, "\vmarkc_a"] \arrow[rd, "p_c"]          &                      &  & \sheafc_{\vartist} \arrow[lld, "l_c", dashed]                     \\
-                                                                           & \mathcal{\vfiberc_c} &  &                                                                   \\
-  \mathcal{\vfiberc_b} \arrow[rrr, "\vmarkc_b"] \arrow[ru, "p^{\prime}_c"] &                      &  & \sheafc_{\vartistc^{\prime}} \arrow[llu, "l^{\prime}_c"', dashed]
-\end{tikzcd}
-\end{equation}
-
-\subsubsection{Graphic to Data: \vindex}
-\begin{equation}
-  \label{eq:artist:construction:xi}
-  \begin{tikzcd}
-    \textcolor{fade}{\dtotal} \arrow[rd, "\pi"', color=fade] &         & \textcolor{fade}{\vtotal} \arrow[ld, "\pi", color=fade] &  & \textcolor{fade}{\gtotal} \arrow[d, "\pi", color=fade] \\
-  & \dbasec &                                             &  & \gbasec \arrow[lll, "\vindexc"']          
-    \end{tikzcd}
-\end{equation}
-
-The functor $\vindex$ is a deformation retract, which means....\cite{hatcherAlgebraicTopology2002,spanier1989algebraic}
-
-By \note{homotopy lifting?}
-\begin{equation}
-  \label{eq:artist:construction:xi:lifted}
-  \begin{tikzcd}
-    & \opensetc_j \arrow[d, "\iota", hook] \\
-\dbasec \arrow[ru, "\xi^*", dashed] & \gbasec \arrow[l, "\xi"]            
-\end{tikzcd}
-\end{equation}
-
-\begin{figure}[h!]
-  \centering
-  \subfloat[\dbasec_a \sqcup_{\dbasec_c} \dbase_b]{\includegraphics[width=1.2in]{union_k.png}%
-  \label{fig_first_case}}
-  %\hfil
-  \subfloat[\dbasec_c=\varnothing]{\includegraphics[width=1.2in]{disjoint_k.png}%
-  \label{fig_second_case}}
-  \caption{\note{probably doesn't need the k=0}}
-  \label{fig_sim}
-\end{figure}
-
-\begin{equation}
-  \label{eq:artist:construction:xi:multi_base}
-  \begin{tikzcd}
-    \dbasec_{c} \arrow[d, hook] \arrow[r, hook] & \dbasec_{b}                 & \gbasec_b \arrow[l, "\vindexc"]                                                \\
-    \dbasec_{a}                                 & \gbasec_a \arrow[l, "\vindexc"'] & \gbasec_c \arrow[l, hook'] \arrow[u, hook] \arrow[llu, "\vindexc" description]
-  \end{tikzcd}
-\end{equation}
-
-\note{talk about xi even though is not explicitely implemented}
-The mapping between graphic and data space expresses how...
-This is what allows the artist to generate graphics where the subset of data on view is dynamically updated, such as pan and zoom navigation\cite{NekrasovskiEvaluationPanZoom2006} and sliding windows on streaming data\cite{crouchDynamicGraphsSlidingwindow2013,chuTimeSeriesSegmentation1995}.
-
-\section{Case Study}
-\label{sec:case-study}
-\begin{figure}[h!]
-  \includegraphics[width=\columnwidth]{path_of_q.png}
-    \caption{\vchannel \& \mark are what we're implementing}
-  \label{fig:api}
-\end{figure}
-We implement the \textbf{arrows} in \autoref{fig:api}.
-\mintinline{python}{axesArtist} is a parent artist that acts as a screen.
-This allows for the composition described in                                           \autoref{sec:artist:union}
-
-\subsection{\vartist}
-\begin{minted}{python}
-for local_tau in axesArtist.artist.data.query(screen_bounds, dpi):
-    mu = axesArtist.artist.graphic.mu(local_tau)
-    rho = axesArtist.artist.graphic.qhat(**mu)
-    H = rho(renderer)
-\end{minted}
-
-where the artist is already parameterized with the \vindex\ functions and which fibers they are associated to:
-
-\begin{minted}{python}
-\end{minted}
-
-\subsubsection{\vindex}
-\subsubsection{\vchannel}
-\subsubsection{\vmarkd}
+\section{Preserving Structure}
 
@@
 \appendices
+\section{Category Theory}
+\subsection*{Category}
+\subsection*{}
 

### 2022-04-04 paper/main.tex
fiber bundles

@@
 We use concepts from algberaic topolology and category theory to formally express the structure visualization components are expected to preserve.
-Category theory is a mathematical langauge for describing how objects are specified, constrained, and composed \cite{wielsManagementEvolvingSpecifications1998}\; therefore in this work we propose implementable in code categorical constructions of visualization components.
-Fiber bundles, are a "unified, dimension-independent framework", as described by Butler\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}, that allow us to seperatly describe the topololigical properties and fields of a data source and also the mapping between topolology and data values.
-Sheaves on bundles are an "algebraic data structure", as described by Ghrist \cite{ghristElementaryAppliedTopology2014}, for for expressing the bookeeping involved in subsetting data and composing data that is distributed or streaming.
+Fiber bundles and sheaves, which come from algebraic topolology, provide a uniform generalizable way to describe data and visualizations.
+Category theory is a method to describe how objects are specified, constrained, and composed \cite{wielsManagementEvolvingSpecifications1998}\; therefore in this work we propose categorical constructions of visualization components as specifications for implementable components.
 
@@
 \label{sec:atct:fiber-bundles}
+Fiber bundles, are a "unified, dimension-independent framework", as described by Butler\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}, that allow us to seperatly describe the topololigical properties and fields of a data source and also the mapping between topolology and data values.
+A fiber bundle $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and continuous surjective map $\pi: \dtotalc \rightarrow \dbasec$\cite{FiberBundle2020}.
+
+\begin{equation}
+  \label{eq:atct:fb_intro}
+  \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
+   \dfiberc
+    \arrow[r, hook] \& 
+    \dtotalc
+    \arrow[d, "\pi"'] \\
+     \& 
+  \dbasec
+     \arrow[u, "\dsectionc"', bend right, pos=.5]
+  \end{tikzcd}
+\end{equation}
+
+The \textcolor{base}{base space} is a topological space \dbasec\ with points $\dbasepointc \in \dbasec$ and topology $\mathcal{T}_{\dbasec}$; $\mathcal_{T}_k$ is a set of opensets $ \dbasepointc \in \opensetc \subseteqq \dbasec$ \cite{munkresElementsAlgebraicTopology1984} that cover the base space \dbasec.
+The \textcolor{fiber space} is a topological space \dfiberc\ that is the preimage of the projection function $\pi$ at a point in the base space $\dfiberc|_{\dbasepointc} = \pi^{-1}(\dbasepointc)$ and fibers of a bundle are isomorphic $\dfiberc \simeq \dfiberc_{\dbasepointc}\;\forall \dbasepointc \in \dbase$.
+The \textcolor{section}{sections} of a bundle $\dsection$ are maps from the base space to points in the bundles over that base space and $\textcolor{set}{\Gamma}$ denotes the set of all sections in a bundle over an openset.
+\begin{equation}
+  \label{eq:atct:fb_sections}
+  \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \coloneqq \big\{\gsectionc: \opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} \; \bigm{\vert} \pi(\gsectionc(\gbasepointc)) = \gbasepointc\;for\, all\; \gbasepointc \in \opensetgc \big\}
+\end{equation}
+
+A fiber bundle is locally trivial, which means that for every point \dbasepointc\ there exists an open neighborhood $\dbasepointc \in \opensetc \subseteq \dbasec$ such that there is a homeomorphism $\pi^{-1}(\opensetc)\xrightarrow{\equivc} \opensetc \times \dfiberc$.
+A bundle may be globally trivial, meaning that $\dtotalc = \dbasec \times \dfiberc$, and generally can be thought of as a twisted product of total and base spaces \cite{munkresElementsAlgebraicTopology1984}.
+
+\begin{figure}
+  \label{fig:atct:fb}
+  \includegraphics[width=\columnwidth]{fiberbundle.png}
+  \caption{change to concrete example}
+\end{figure}
+
+Fiber bundles can be used to model data for visualization, as proposed by Butler\cite{butlerVectorBundleClassesForm1992, butlerVisualizationModelBased1989}.
+We encode the topological properties of the data \cite{wilkinsonGrammarGraphics2005} as the base space \dbasec.
+The topological properties, as described by Wilkenson, are how the data is organized - for example \dbase\ could be discrete points, a line, a 2D surface, a 3D volume, or a graph.
+The names and data types of the data fields can be expressed as the fiber space \dfiberc, as described by Spivak \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES}.
+For example, a temperature field could be expressed as a fiber $\dfiber_{temperature}=\mathbb{R}$, and a color field could be $\dfiber_{color}=\mathbb{R}^{3}$.
+Even though fibers on non-trivial spaces, such as vector fields, are different, we can simplify to one type because fibers are isomorphic.\note{unsure of this claim but I think this is why, and also I need to tell you why you care that they're isomorphic} The total bundle space \dtotalc\ acts as a way of expressing all data with the same topolology \dbasec\ and fields \dfiberc, for example a database schema.
+A specific dataset is represented as a section \dsectionc, and the return values of a section evaluated at a point \dbasepointc\ is the record $\{field_{0}: value_{0}, ..., field_n:value_n\}$ at that point.
+
+We also use fiber bundles to represent the output of a visualization algorithm, which we term the graphic but generalizes to output on any display space, such as a screen or 3D print. 
+\begin{equation}
+  \label{eq:atct:fb_graphic}
+  \gfiberc \hookrightarrow \gtotalc \xrightarrow{\pi} \gbasec
+\end{equation}
+The base space \gbasec\ is a parameterization of the display area, for example the inked bounding box in cairo \cite{CairographicsOrg}.
+The fiber space \gfiberc\ is an abstraction of the renderer fields, for example $\{x,\,y,\,r,\,g,\,b,\,a\}$.
+The sections of the graphic
+\begin{equation}
+  \label{eq:atct:fb_graphic_section}
+  $\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \coloneqq \big\{\gsectionc: \opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} \; \bigm{\vert} \pi(\gsectionc(\gbasepointc)) = \gbasepointc\;for\, all\; \gbasepointc \in \opensetgc \big\}$
+\end{equation}
+are functions that generate different graphics, for example markers for a scatter plot, bars for a bar chart, cells of a heatmap, or pixels of an image.
+
+\subsubsection{Input and Output Types}
+
+\subsection{Sheaves}
+Sheaves on bundles are an "algebraic data structure", as described by Ghrist \cite{ghristElementaryAppliedTopology2014}, for expressing the bookkeeping involved in subsetting data and composing data that is distributed or streaming.
+
+\subsection{Transforms on Sheaves}
+
+\subsection{Transforms Between Sheaves}
 

### 2022-04-05 paper/main.tex
classes
sheave transforms"

@@
 \usetikzlibrary{cd} % commutative diagrams
+\newtheorem{axiom}{Axiom}
 \newtheorem{prop}{Proposition} %math?
@@
 
-\begin{equation}
+\begin{equation}[h]
   \label{eq:atct:fb_intro}
@@
    \dfiberc
-    \arrow[r, hook] \& 
+    \arrow[r, hook, color=total] \& 
     \dtotalc
-    \arrow[d, "\pi"'] \\
+    \arrow[d, "\pi"',color=total] \\
      \& 
   \dbasec
-     \arrow[u, "\dsectionc"', bend right, pos=.5]
+     \arrow[u, "\dsectionc"', bend right, pos=.5, color=section]
   \end{tikzcd}
@@
 
-\begin{figure}
+\begin{figure}[h]
   \label{fig:atct:fb}
@@
 Fiber bundles can be used to model data for visualization, as proposed by Butler\cite{butlerVectorBundleClassesForm1992, butlerVisualizationModelBased1989}.
-We encode the topological properties of the data \cite{wilkinsonGrammarGraphics2005} as the base space \dbasec.
+We describe the topological properties of the data \cite{wilkinsonGrammarGraphics2005} as the base space \dbasec.
 The topological properties, as described by Wilkenson, are how the data is organized - for example \dbase\ could be discrete points, a line, a 2D surface, a 3D volume, or a graph.
@@
 A specific dataset is represented as a section \dsectionc, and the return values of a section evaluated at a point \dbasepointc\ is the record $\{field_{0}: value_{0}, ..., field_n:value_n\}$ at that point.
+
+\begin{figure}
+  \label{fig:atct:fb_graphic_bundle}
+  \includegraphics[width=1\columnwidth]{render.png}
+  \caption{}
+\end{figure}
 
@@
 
-\subsubsection{Input and Output Types}
+\subsection{Data: Base $\rightarrow$ Fiber}
+As described in \autoref{sec:atct:fiber-bundles}, we model data as a map from base space to fiber space.To encapsulate the structure carried in these maps, and provide a way to define these maps in reusable and composable manner, we define categories of base and fiber spaces.
+
+\begin{definition}
+  The base space category $\mathcal{\opensetc}$ has open set objects $\opensetc$ and inclusion morphisms $\iota: \opensetc_1 \rightarrow \opensetc_2$ such that $\opensetc_1 \subseteq \opensetc_2$. 
+\end{definition}
+
+As with all categories, each object has an identity morphism.
+Morphisms are also associative and commutative \cite{barrCategoryTheoryComputing}, which is what allows for currying of functions and composition of objects.
+
+\begin{definition}
+  The fiber category $\mathcal{\dfiberc}$ is a monoidal category, meaning the category has a single object $\dfiberc$ of an arbitrary type.
+The morphisms on the fiber object are $\dfuncc \in Hom(\dfiberc, \dfiberc)$.
+\end{definition}
+
+The fiber category is equipped with a bifunctor $\otimes: \mathcal{\dfiber} \times {\dfiber}$.
+The bifunctor provides a method for composing fibers into more complex fibers, thereby allowing for complex fields.
 
@@
 Sheaves on bundles are an "algebraic data structure", as described by Ghrist \cite{ghristElementaryAppliedTopology2014}, for expressing the bookkeeping involved in subsetting data and composing data that is distributed or streaming.
+Formally, pre-sheaves are functors, meaning they are functions from objects of one category to objects of another category \cite{WhatFunctorDefinitions}.
+In this work we consider the presheaf
+\begin{equation}
+  \label{eq:atct:presheaf}
+  \begin{tikzcd}
+    \cgamma{\opensetc_1}{\dtotalc\restriction_{\opensetc_1}}  &  & \cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}} 
+    \arrow[ll, "\iota^*"', color=set] \\
+    & & \\
+    \opensetc_1 
+    \arrow[rr, "\iota", hook, color=base] 
+    \arrow[uu, "{\sheafc_{\dbasec, \dtotalc}}", maps to, color=sheaf] &  & \opensetc_2 
+    \arrow[uu, "{\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf]              
+    \end{tikzcd}
+\end{equation}
 
-\subsection{Transforms on Sheaves}
+which associates a set of sections $\cgamma{\openset}{\dtotal\restriction_{\openset}}$ with an openset \openset.
+The sets of sections are objects of the category \setc\ and the morphisms of the category are restrictions $\iota^*$.
+The sheaf over a an open set $\openset$ surrounding a point $\dbasepoint$ is called a stalk\cite{StalkSheaf2019}
+\begin{equation}
+  \label{eq:atct:sheaf:stalk}
+    \sheaf_{\dbase, \dtotalc}\restriction_{\dbasepoint}\coloneqq \lim\limits_{\openset\ni \dbasepoint} \Gamma(\openset, \dtotal\restriction_{\openset}) 
+\end{equation}
+where the fiber is contained inside the stalk  $\dfiber_{\dbasepoint} \subset  \sheaf_{\dbase, \dtotal}\restriction_{\dbasepoint}$.
+The germ is the section evaluated at a point in the stalk  $\dsection(\dbasepoint) \in \sheaf_{\dbase, \dtotal}\restriction_{\dbasepoint}$ and is the data.
+\note{I never use stalk and germ terminology at any point in this paper}
 
-\subsection{Transforms Between Sheaves}
+A sheaf is a presheaf that satisfies the locality and gluing axioms \cite{bakerMathsSheaf}.
+The locality axiom is that given a union of opensets $\mathscr{\opensetc} = \bigcup\limits_{i\in I} \opensetc_i$ and $\dsectionc^{a}, \dsectionc^{b} \in \sheafc(\mathscr{\opensetc})$,  if $\dsectionc^{a}\restriction_{\opensetc_i} = \dsectionc^{b}\restriction_{\opensetc_i}$ for each $\opensetc_i \in \opensetc$ then $\dsectionc^{a} = \dsectionc^{b}$.
+This means that two sections in a sheaf are equal when they evaluate to the same values over all open sets in a collection.
+The gluing axiom is that given $\dsectionc^{i} \in \sheafc(\opensetc_i)$ such that ${\dsectionc^{i}}\restriction_{\opensetc_i\cap \opensetc_j} = {\dsectionc^{j}}\restriction_{\opensetc_i \cap \opensetc_j}$ for $\opensetc_i, \opensetc_j \in \mathscr{\opensetc}$, there exists $\dsectionc \in \sheafc(\mathscr{\opensetc})$ such that $\dsectionc\restriction_{\opensetc_i} = {\dsectionc^{i}}$.
+This means we can construct a section $\dsection$ on the collection of open sets that is the union of sections defined on specific open sets.
+This expresses an equivalency between a concatenated represetation of data and the distributed represetation of the same data.
 
-\section{Preserving Structure}
+In this model, we represent data and graphics as a sheaf; doing so allows us to express that data sources are expected to have the bookkeeping that a sheaf requires, meaning inclusions, restrictions, locality, and gluing must be preserved.
+
+\subsubsection{Transforms on Sheaves: $\dfuncc$}
+We express transformations of data, such as scaling and rotation, as function on sheaves $\dfuncc=(\dfunchc, \dfunctc)$
+\begin{equation}
+  \label{eq:atct:sheaves:morphism}
+  \begin{tikzcd}
+    \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
+    \arrow[rr, "\dfuncpullc", color=action] &  & \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}} & \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}}
+    \arrow[dd, "\dfunctc", color=action] \\
+     &  & &       \\
+    \opensetc 
+    \arrow[uu, maps to,color=sheaf]  &  & \opensetc^{\prime} 
+    \arrow[ll, "\dfunchc"', color=action] 
+    \arrow[uu, maps to, color=sheaf]                    & \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}}                       
+    \end{tikzcd}
+\end{equation}
+
+The base transformation $\dfunch: \openset^{\prime}\rightarrow \openset$ where both opensets are in the same basespace $\openset, \openset^{\prime} \subseteq \dbase$ induces a pullback of sections 
+\begin{equation}
+  \dfuncpull \dsection \restriction_{\openset}: \dsection\restriction_{\openset} \mapsto \dsection \restriction_{\openset} \circ \dfunch 
+\end{equation}
+An example of a base transformation is a remapping of an underlying indexing, such as a sorting of primary keys for a database of discrete values.
+
+The fiber transformation $\dfunctc: \dfuncpull \dtotal_{\dbasepoint^{\prime}} \rightarrow \dfuncpull \dtotal_{\dbasepoint^{prime}}$ is a morphism on the fiber $\dfunct \in Hom(\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}},\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}})$ restricted to a point $\dbasepoint^{\prime} \in \openset^{\prime}$.
+The fiber transformation is a change in section
+\begin{equation}
+  \dfunct: \dfuncpull \dsection \restriction_{\openset} \mapsto \dfuncpull \dsection^{\prime} \restriction_{\openset}
+\end{equation}
+where $\dsection, \dsection^{\prime} \in \Gamma(\openset^{\prime}, \dfuncpull\dtotal\restriction_{\openset^{\prime}})$.
+Examples of a fiber transformation are actions on fibers, such as the permutation, ordering, translation, and scaling codified as Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946}.
+
+The transformation 
+\begin{equation}
+  \dfunc: \dsection\restriction_{\openset} \mapsto \dsection^{\prime}\restriction_{\openset} \circ \dfunch
+\end{equation}
+is used to express transformation such as rotation that induce both a remapping of the index space and a change n the values.
+
+\subsection{ Data Index \dbasec\ \& Visualization Region \gbasec}
+
+\begin{equation}
+  \begin{tikzcd}[row sep=huge]
+    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \arrow[rr, "\vindexpushc"]                                                                              &  & \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}                                                                       \\
+    \opensetgc \arrow[rr, "\vindexc"] \arrow[u, "{\sheafc_{\gbasec, \gtotalc}}", maps to] \arrow[d, "{\sheafc_{\gbasec, \vindexpullc\dtotalc}}"', dashed, maps to] &  & \opensetc \arrow[d, "{\sheafc_{\dbasec,\dtotalc}}", maps to] \arrow[u, "{\sheafc_{\dbasec, \vindexpushc \gtotalc}}"', dashed, maps to] \\
+    \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}                                                                                             &  & \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \arrow[ll, "\vindexpullc"']
+    \end{tikzcd}
+\end{equation}
+
+\section{Artist: Data $\rightarrow$ Graphic}
 

### 2022-04-06 paper/main.tex
diagram out to end of artist section

@@
 
-\section{Expressing Structure}
+\section{Data And Visualizations}
 \label{sec:atct}
@@
 
-\subsubsection{Transforms on Sheaves: $\dfuncc$}
-We express transformations of data, such as scaling and rotation, as function on sheaves $\dfuncc=(\dfunchc, \dfunctc)$
+\subsubsection{Data Transformations $\dfuncc=(\dfunchc, \dfunctc)$}
+We express transformations of data, such as scaling and rotation, as function on sheaves: 
+\note{may not need the diagram here}
 \begin{equation}
@@
 
-The base transformation $\dfunch: \openset^{\prime}\rightarrow \openset$ where both opensets are in the same basespace $\openset, \openset^{\prime} \subseteq \dbase$ induces a pullback of sections 
+In this work, we assume that the map
+\begin{equation}
+\dfunch: \openset^{\prime}\rightarrow \openset
+\end{equation}
+is from one open set to another open set in the same base space $\openset, \openset^{\prime} \subseteq \dbase$.
+For example a remapping of an underlying indexing, such as a sorting of primary keys for a database of discrete values.
+This base space transfromation induces a pullback section
 \begin{equation}
@@
 \end{equation}
-An example of a base transformation is a remapping of an underlying indexing, such as a sorting of primary keys for a database of discrete values.
+such that $\dsection|_{\openset} = \dfuncpull\dsection|_{\opensetc^{\prime}}$ because $\dsection|_{\openset} = \dsection|_{\dfunch(\opensetc^{prime})}$.
+This means that the base space transformation \dfunch\ alone does not change the data values.
 
-The fiber transformation $\dfunctc: \dfuncpull \dtotal_{\dbasepoint^{\prime}} \rightarrow \dfuncpull \dtotal_{\dbasepoint^{prime}}$ is a morphism on the fiber $\dfunct \in Hom(\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}},\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}})$ restricted to a point $\dbasepoint^{\prime} \in \openset^{\prime}$.
+The fiber transformation $\dfunct: \dfuncpull \dtotal_{\dbasepoint^{\prime}} \rightarrow \dfuncpull \dtotal_{\dbasepoint^{\prime}}$ is a morphism on the fiber $\dfunct \in Hom(\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}},\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}})$ restricted to a point $\dbasepoint^{\prime} \in \openset^{\prime}$.
 The fiber transformation is a change in section
@@
 
-The transformation 
+The full transformation 
 \begin{equation}
@@
 \end{equation}
-is used to express transformation such as rotation that induce both a remapping of the index space and a change n the values.
+is used to express actions such as rotation that induce both a remapping of the index space and a change in the data values.
+The data transform \dfunc\ is composable if there exists functions $\dfunc_{a,b}: \dtotal_a \times \dtotal_b \rightarrow \dtotal_a \times \dtotal_b$, $\dfunc_{a}: \dtotal_a \rightarrow \dtotal_a$ and $\dfunc_{b}: \dtotal_b \rightarrow \dtotal_b$ s.t.
+$\pi_a \circ \dfunc_a = \dfunc_{a,b} \circ \pi_a$ and $\pi_b \circ \dfunc_b = \dfunc_{a,b} \circ \pi_b$ then $\dfunc_{a,b} = (\dfunc_a, \dfunc_b)$.
+This allows us to define a data transform $\dfunc = (\dfunch, \prod\limits_{i=0}^{n}\dfunct_i)$ where each fiber transform $\dfunct_{i}$ is applied to a different fiber field $\dfiber_i$.
 
@@
   \begin{tikzcd}[row sep=huge]
-    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \arrow[rr, "\vindexpushc"]                                                                              &  & \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}                                                                       \\
-    \opensetgc \arrow[rr, "\vindexc"] \arrow[u, "{\sheafc_{\gbasec, \gtotalc}}", maps to] \arrow[d, "{\sheafc_{\gbasec, \vindexpullc\dtotalc}}"', dashed, maps to] &  & \opensetc \arrow[d, "{\sheafc_{\dbasec,\dtotalc}}", maps to] \arrow[u, "{\sheafc_{\dbasec, \vindexpushc \gtotalc}}"', dashed, maps to] \\
-    \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}                                                                                             &  & \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \arrow[ll, "\vindexpullc"']
+    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+    \arrow[rr, "\vindexpushc", color=functor]   &  & \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}\\
+    \opensetgc \arrow[rr, "\vindexc", color=functor] 
+    \arrow[u, "{\sheafc_{\gbasec, \gtotalc}}", maps to, color=sheaf] 
+    \arrow[d, "{\sheafc_{\gbasec, \vindexpullc\dtotalc}}"', dashed, maps to, color=sheaf] &  & \opensetc 
+    \arrow[d, "{\sheafc_{\dbasec,\dtotalc}}", maps to, color=sheaf] 
+    \arrow[u, "{\sheafc_{\dbasec, \vindexpushc \gtotalc}}"', dashed, maps to, color=sheaf] \\
+    \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}  &  & \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
+    \arrow[ll, "\vindexpullc"', color=functor]      
     \end{tikzcd}
@@
 \section{Artist: Data $\rightarrow$ Graphic}
+Because $Hom_{\gbase}(\sheaf_{\gbase, \vindexpull\dtotal}, \sheaf_{\gbase,\gtotal}) = Hom_{\dbase}(\sheaf_{\dbase, \dtotal}, \sheaf_{\dbase, \vindexpush\gtotal})$
 
+\begin{equation}
+\begin{tikzcd}
+  {\sheafc_{\gbasec,\gtotalc}} 
+  \arrow[rr, "\vindexpushc", color=functor]  
+  &  & {\sheafc_{\dbasec,\vindexpushc\gtotalc}}   \\
+   &  &             \\
+  {\sheafc_{\gbasec,\vindexpullc\dtotalc}} 
+  \arrow[uu, "\vartistc^{\gbasec}", Rightarrow, color=artist] &  & {\sheafc_{\dbasec,\dtotalc}} 
+  \arrow[ll, "\vindexpullc"', color=functor]
+   \arrow[uu, "\vartistc^{\dbasec}"', Rightarrow, color=artist] 
+   \arrow[lluu, "\vartistc" description, color=artist]
+  \end{tikzcd}
+\end{equation}
+
+\begin{equation*}
+  \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \coloneqq \{\gsectionc\vert\;\exists\;\dsectionc \in \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}\;s.t.\; 
+  \vartistc(\dsectionc) = \gsectionc,\; \vindexc(\opensetgc) = \opensetc \} 
+\end{equation*}
+$\imartist{\opensetg}{\gtotal\restriction_{\opensetg}} \subset \cgamma{\opensetg}{\gtotal\restriction_{\opensetg}}$
+
+\subsection{Equivariant Artist}
+\begin{equation*}
+  \begin{tikzcd}[ampersand replacement=\&, column sep=small]
+  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
+  \arrow[rrr, "\vartistc", color=artist] 
+  \arrow[d, "\dfuncpullc_{\dtotalc}"', color=action] 
+  \& \& \& 
+  \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+  \arrow[d, "\dfuncpullc_{\gtotalc}", dotted] \\
+  \cgamma{\opensetc^{\prime}}{\dfuncpullc_{\dtotalc}\dtotalc\restriction_{\opensetc^{\prime}}} 
+  \arrow[dd, "\dfunctc_{\dtotalc}"', color=action] \& 
+  \opensetc 
+   \& 
+  \opensetgc 
+  \arrow[l, "\vindexc"', color=functor] 
+  \& 
+  \imartist{\opensetgc^{\prime}}{\dfuncpullc_{\gtotalc}\gtotalc\restriction_{\opensetgc^{\prime}}} 
+  \arrow[dd, "\dfunctc_{\gtotalc}", dotted, color=action] \\
+  \& 
+  \opensetc^{\prime} 
+  \arrow[u, "\dfunchc_{\dtotalc}", color=action] 
+  \& 
+  \opensetgc^{\prime} 
+  \arrow[l, "\vindexc"', color=functor] 
+  \arrow[u, "\dfunchc_{\gtotalc}"', dotted, color=action] 
+  \& \\
+  \cgamma{\opensetc^{\prime}}{\dtotalc^{\prime}\restriction_{\opensetc^{\prime}}} 
+  \arrow[rrr, "\vartistc^{\prime}", color=artist]  
+  \& \& \& 
+  \imartist{\opensetgc^{\prime}}{\gtotalc^{\prime}\restriction_{\opensetgc^{\prime}}}
+  \end{tikzcd}
+\end{equation*}
+For all points $\gbasepointc^{\prime}\in \gbasec^{\prime}$:
+\begin{equation*}
+  \vartistc^{\prime}(\dfunctc_{\dtotalc}(\dsectionc(\dfunchc_{\dtotalc}(\vindexc(\gbasepointc^{\prime}))))) = \dfunctc_{\gtotalc}(\vartistc(\dsectionc(\vindexc(\dfunchc_{\gtotalc}(\gbasepointc^{\prime})))))
+\end{equation*}
+\subsection*{Testing Equivariance}
+\begin{equation*}
+  \begin{tikzcd}[row sep=huge]
+    \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
+    \arrow[d, "\equivc", color=monoid]
+     \arrow[rr, "\vartistc", color=artist] &  & \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+     \arrow[d, "\extractmc", color=monoid] \arrow[r, "render"] & VIZ \arrow[ld, "measure"] \\
+    {Hom(\opensetc, \measurec)} \arrow[rr, "\vindexpullc", color=functor]  &  & {Hom(\opensetgc, \measurec)}   & 
+    \end{tikzcd}
+  \end{equation*}
+  \begin{description}
+      \item[input]{$\equivc: \dsectionc \mapsto (\opensetc \xrightarrow{\equivc_{\dsectionc}} \measurec)$}
+      \item[output]{$\extractmc:\gsectionc \mapsto (\opensetgc \xrightarrow{\extractmc_{\gsectionc}} \measurec)$}
+      \item[] $\equivc_{\dsectionc}(\dbasepointc) = \extractmc_{\gsectionc}(\gbasepointc)$ for all $\vindexc(\gbasepointc) = \dbasepointc, \dbasepointc \in \dbasec, \gbasepointc \in \gbasec$ 
+  \end{description}
+
+  \note{introduce \phi_M as basis of equivariance}
 \section{Discussion}

### 2022-04-08 paper/main.tex
flipping arrows

@@
 
-\subsection{Data: Base $\rightarrow$ Fiber}
+\subsection{Input and Output Types}
+\label{sec:atct:io}
 As described in \autoref{sec:atct:fiber-bundles}, we model data as a map from base space to fiber space.To encapsulate the structure carried in these maps, and provide a way to define these maps in reusable and composable manner, we define categories of base and fiber spaces.
@@
 \begin{definition}
+  \label{sec:atct:io:base}
   The base space category $\mathcal{\opensetc}$ has open set objects $\opensetc$ and inclusion morphisms $\iota: \opensetc_1 \rightarrow \opensetc_2$ such that $\opensetc_1 \subseteq \opensetc_2$. 
@@
 \begin{definition}
+  \label{def:atct:io:fiber}
   The fiber category $\mathcal{\dfiberc}$ is a monoidal category, meaning the category has a single object $\dfiberc$ of an arbitrary type.
-The morphisms on the fiber object are $\dfuncc \in Hom(\dfiberc, \dfiberc)$.
+The morphisms on the fiber object are $\dfunctc \in Hom(\dfiberc, \dfiberc)$.
 \end{definition}
@@
 Formally, pre-sheaves are functors, meaning they are functions from objects of one category to objects of another category \cite{WhatFunctorDefinitions}.
-In this work we consider the presheaf
+Here the sheaf
+\begin{equation}
+  \sheafc_{\dbasec, \dtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
+\end{equation}
+denotes a sheaf from opensets on the base space $\dbase$ to sets of sections on the bundle $\dtotal$. As a contravariant functor it preserves morphism s
+
 \begin{equation}
@@
 
-\subsubsection{Data Transformations $\dfuncc=(\dfunchc, \dfunctc)$}
-We express transformations of data, such as scaling and rotation, as function on sheaves: 
-\note{may not need the diagram here}
+\subsubsection{Measurement Types}
+
+We specify that structure on a sheaf includes the actions $\dfuncc$ on sections of that sheaf, which is a generalization of Steven's \cite{stevensTheoryScalesMeasurement1946} classification of measurement scales by their mathematical group structure.
+These sheaf actions describe which changes in the data we expect to be equivariant to changes in the visualization.
+
 \begin{equation}
@@
     \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
-    \arrow[rr, "\dfuncpullc", color=action] &  & \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}} & \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}}
+    \arrow[rr, "\dfuncpullc", color=action, maps to] &  & 
+    \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}} & 
+    \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}} 
     \arrow[dd, "\dfunctc", color=action] \\
@@
     \arrow[uu, maps to,color=sheaf]  &  & \opensetc^{\prime} 
-    \arrow[ll, "\dfunchc"', color=action] 
+    \arrow[ll, "\dfunchc"', color=action, maps to] 
     \arrow[uu, maps to, color=sheaf]                    & \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}}                       
@@
 
-The fiber transformation $\dfunct: \dfuncpull \dtotal_{\dbasepoint^{\prime}} \rightarrow \dfuncpull \dtotal_{\dbasepoint^{\prime}}$ is a morphism on the fiber $\dfunct \in Hom(\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}},\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}})$ restricted to a point $\dbasepoint^{\prime} \in \openset^{\prime}$.
+As introduced in \autoref{def:atct:io:fiber}, the fiber transformation $\dfunct: \dfuncpull \dtotal_{\dbasepoint^{\prime}} \rightarrow \dfuncpull \dtotal_{\dbasepoint^{\prime}}$ is a morphism on the fiber $\dfunct \in Hom(\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}},\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}})$ restricted to a point $\dbasepoint^{\prime} \in \openset^{\prime}$.
 The fiber transformation is a change in section
@@
 
-\subsection{ Data Index \dbasec\ \& Visualization Region \gbasec}
-
+\subsection{Map Between Base Spaces}
+\begin{figure}[ht]
+  \includegraphics*[width=1\columnwidth]{../../slides/figures/math/push_pull_scatter.png}
+\end{figure}
+We define a functor
 \begin{equation}
+  \vindexc: \opensetgc \rightarrow \openset 
+\end{equation}
+that expresses the constraint that every region in a visualization $\opensetg \subseteq \gbase$ must map to a corresponding set $\openset \subset \dbase$.
+We define the functor $\vindexc$ such that the data base space \dbasec\ is a deformation retraction\cite{nlab:deformation_retract} of the graphic space $\gbasec$.
+To simplify matters, in this paper, we construct the graphic space as a constant multiple of the base space such that
+\begin{equation}
+  \vindexc: \underbrace{\opensetc\times[0,1]^{n}}_{\opensetgc} \rightarrow \opensetc 
+\end{equation}
+where $ n = dim(\dbase) - dim(\gbase)$ if $dim(\dbase)<dim(\gbase)$ and otherwise $ n = dim(\dbase)$. 
+\begin{equation}
+  \label{eq:atct:morphisms:xi}
 \begin{tikzcd}[row sep=huge]
+  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
+  \arrow[rr, "\vindexpullc", maps to, color=functor] &  & 
+  \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}                                                                                                       \\
+  \opensetc 
+  \arrow[d, "{\sheafc_{\dbasec, \vindexpushc \gtotalc}}", dashed, maps to, color=sheaf] 
+  \arrow[u, "{\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf] &  & 
+  \opensetgc 
+  \arrow[d, "{\sheafc_{\gbasec, \gtotalc}}", maps to, color=sheaf] 
+  \arrow[u, "{\sheafc_{\gbasec, \vindexpullc\dtotalc}}"', dashed, maps to, color=sheaf] 
+  \arrow[ll, "\vindexc"', maps to, color=functor] \\
+  \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} &  & 
   \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
-    \arrow[rr, "\vindexpushc", color=functor]   &  & \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}\\
-    \opensetgc \arrow[rr, "\vindexc", color=functor] 
-    \arrow[u, "{\sheafc_{\gbasec, \gtotalc}}", maps to, color=sheaf] 
-    \arrow[d, "{\sheafc_{\gbasec, \vindexpullc\dtotalc}}"', dashed, maps to, color=sheaf] &  & \opensetc 
-    \arrow[d, "{\sheafc_{\dbasec,\dtotalc}}", maps to, color=sheaf] 
-    \arrow[u, "{\sheafc_{\dbasec, \vindexpushc \gtotalc}}"', dashed, maps to, color=sheaf] \\
-    \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}  &  & \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
-    \arrow[ll, "\vindexpullc"', color=functor]      
+  \arrow[ll, "\vindexpushc"', maps to, color=functor]                                                         
   \end{tikzcd}
@@
 
-\section{Artist: Data $\rightarrow$ Graphic}
+\section{Artist: Data to Graphic}
 Because $Hom_{\gbase}(\sheaf_{\gbase, \vindexpull\dtotal}, \sheaf_{\gbase,\gtotal}) = Hom_{\dbase}(\sheaf_{\dbase, \dtotal}, \sheaf_{\dbase, \vindexpush\gtotal})$
@@
   \begin{tikzcd}
-  {\sheafc_{\gbasec,\gtotalc}} 
-  \arrow[rr, "\vindexpushc", color=functor]  
-  &  & {\sheafc_{\dbasec,\vindexpushc\gtotalc}}   \\
+    \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
+    \arrow[dd, "\vartistc^{\dbasec}"', color=artist] 
+    \arrow[rrdd, "\vartistc", color=artist] 
+    \arrow[rr, "\vindexpullc", color=functor] &  &
+    \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} 
+    \arrow[dd, "\vartistc^{\gbasec}", color=artist] \\
      & & \\
-  {\sheafc_{\gbasec,\vindexpullc\dtotalc}} 
-  \arrow[uu, "\vartistc^{\gbasec}", Rightarrow, color=artist] &  & {\sheafc_{\dbasec,\dtotalc}} 
-  \arrow[ll, "\vindexpullc"', color=functor]
-   \arrow[uu, "\vartistc^{\dbasec}"', Rightarrow, color=artist] 
-   \arrow[lluu, "\vartistc" description, color=artist]
+    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} &  & 
+    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+    \arrow[ll, "\vindexpushc"', color=functor]                  
     \end{tikzcd}
@@
 
-  \note{introduce \phi_M as basis of equivariance}
+  \note{introduce $\phi_M$ as basis of equivariance}
 \section{Discussion}

### 2022-04-11 paper/main.tex
sheaves, fabricating Q fibers

@@
   \label{eq:atct:fb_sections}
-  \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \coloneqq \big\{\gsectionc: \opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} \; \bigm{\vert} \pi(\gsectionc(\gbasepointc)) = \gbasepointc\;for\, all\; \gbasepointc \in \opensetgc \big\}
+  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}
 \end{equation}
@@
 
-As with all categories, each object has an identity morphism.
-Morphisms are also associative and commutative \cite{barrCategoryTheoryComputing}, which is what allows for currying of functions and composition of objects.
+By definition, each object in a category has an identity morphism and morphisms are associative and commutative \cite{barrCategoryTheoryComputing}.
 
@@
 The fiber category is equipped with a bifunctor $\otimes: \mathcal{\dfiber} \times {\dfiber}$.
-The bifunctor provides a method for composing fibers into more complex fibers, thereby allowing for complex fields.
+The bifunctor provides a method for combining fibers, thereby allowing for fields that contain multityped values.
+For example, RGB color can either be represented as three fiber fields $\dfiber_{red}, \dfiber_{green}, \dfiber_{blue}$ or a composite fiber field $\dfiber_{red} \otimes \dfiber_{green} \otimes \dfiber_{blue} = \dfiber_{RGB}$.
 
 \subsection{Sheaves}
-Sheaves on bundles are an "algebraic data structure", as described by Ghrist \cite{ghristElementaryAppliedTopology2014}, for expressing the bookkeeping involved in subsetting data and composing data that is distributed or streaming.
+Sheaves on bundles are an "algebraic data structure", as described by Ghrist \cite{ghristElementaryAppliedTopology2014}, for expressing the bookkeeping involved in keeping track of sections of data over opensets.
+Sheaves are a way of expressing that subsets, distributed data, and streaming data are local sections that can be glued together.
 Formally, pre-sheaves are functors, meaning they are functions from objects of one category to objects of another category \cite{WhatFunctorDefinitions}.
@@
     \cgamma{\opensetc_1}{\dtotalc\restriction_{\opensetc_1}}  &  & \cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}} 
-    \arrow[ll, "\iota^*"', color=set] \\
+    \arrow[ll, "\iota^*"', maps to, color=set] \\
     & & \\
     \opensetc_1 
-    \arrow[rr, "\iota", hook, color=base] 
+    \arrow[rr, "\iota", maps to, color=base] 
     \arrow[uu, "{\sheafc_{\dbasec, \dtotalc}}", maps to, color=sheaf] &  & \opensetc_2 
@@
 which associates a set of sections $\cgamma{\openset}{\dtotal\restriction_{\openset}}$ with an openset \openset.
-The sets of sections are objects of the category \setc\ and the morphisms of the category are restrictions $\iota^*$.
+The sets of sections are objects of the category \setc\ and the morphisms of the category are restrictions\footnote{restriction functions } $\iota^*$.
 The sheaf over a an open set $\openset$ surrounding a point $\dbasepoint$ is called a stalk\cite{StalkSheaf2019}
@@
 \end{equation}
-is used to express actions such as rotation that induce both a remapping of the index space and a change in the data values.
-The data transform \dfunc\ is composable if there exists functions $\dfunc_{a,b}: \dtotal_a \times \dtotal_b \rightarrow \dtotal_a \times \dtotal_b$, $\dfunc_{a}: \dtotal_a \rightarrow \dtotal_a$ and $\dfunc_{b}: \dtotal_b \rightarrow \dtotal_b$ s.t.
+is used to express actions such as rotation that induce both a remapping of the index space and a change in the data values. The data transform \dfunc\ is composable
+\begin{equation}
+  \dfuncc = (\dfunchc, \prod\limits_{i=0}^{n}\dfunctc_i)
+\end{equation}
+if there exists functions $\dfunc_{a,b}: \dtotal_a \times \dtotal_b \rightarrow \dtotal_a \times \dtotal_b$, $\dfunc_{a}: \dtotal_a \rightarrow \dtotal_a$ and $\dfunc_{b}: \dtotal_b \rightarrow \dtotal_b$ s.t.
 $\pi_a \circ \dfunc_a = \dfunc_{a,b} \circ \pi_a$ and $\pi_b \circ \dfunc_b = \dfunc_{a,b} \circ \pi_b$ then $\dfunc_{a,b} = (\dfunc_a, \dfunc_b)$.
-This allows us to define a data transform $\dfunc = (\dfunch, \prod\limits_{i=0}^{n}\dfunct_i)$ where each fiber transform $\dfunct_{i}$ is applied to a different fiber field $\dfiber_i$.
+This allows us to define a data transform where each fiber transform $\dfunct_{i}$ can be applied to a different fiber field $\dfiber_i$.
 
 \subsection{Map Between Base Spaces}
-\begin{figure}[ht]
-  \includegraphics*[width=1\columnwidth]{../../slides/figures/math/push_pull_scatter.png}
-\end{figure}
-We define a functor
+\note{add couple of sentences introducing image functors}
+We define a \textcolor{functor}{functor} between the graphic and data base spaces 
 \begin{equation}
-  \vindexc: \opensetgc \rightarrow \openset 
+  \vindexc: \opensetgc \rightarrow \opensetc 
 \end{equation}
-that expresses the constraint that every region in a visualization $\opensetg \subseteq \gbase$ must map to a corresponding set $\openset \subset \dbase$.
-We define the functor $\vindexc$ such that the data base space \dbasec\ is a deformation retraction\cite{nlab:deformation_retract} of the graphic space $\gbasec$.
-To simplify matters, in this paper, we construct the graphic space as a constant multiple of the base space such that
-\begin{equation}
-  \vindexc: \underbrace{\opensetc\times[0,1]^{n}}_{\opensetgc} \rightarrow \opensetc 
-\end{equation}
-where $ n = dim(\dbase) - dim(\gbase)$ if $dim(\dbase)<dim(\gbase)$ and otherwise $ n = dim(\dbase)$. 
+such that every $\opensetg \subseteq \gbase$ must map to a corresponding set $\openset \subset \dbase$. In doing so, we can define the pullback \vindexpullc\  and pushforward \vindexpushc\
+
 \begin{equation}
@@
 
+\begin{figure}[ht]
+  \label{fig:atct:morphisms:sheaf}
+  \includegraphics*[width=1\columnwidth]{../../slides/figures/math/push_pull_scatter.png}
+  \caption{}
+\end{figure}
+
+As illustrated in \autoref{fig:atct:morphisms:sheaf}, the \note{image functors?} between sheaves are a way of expressising the bookkeeping of data in screen and graphic space and graphic in screen and graphic space.
+
 \section{Artist: Data to Graphic}
@@
      & & \\
-    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} &  & 
+    \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} &  & 
     \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
@@
 
-\begin{equation*}
+\begin{equation}
   \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \coloneqq \{\gsectionc\vert\;\exists\;\dsectionc \in \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}\;s.t.\; 
   \vartistc(\dsectionc) = \gsectionc,\; \vindexc(\opensetgc) = \opensetc \} 
-\end{equation*}
+\end{equation}
 $\imartist{\opensetg}{\gtotal\restriction_{\opensetg}} \subset \cgamma{\opensetg}{\gtotal\restriction_{\opensetg}}$
@@
   \note{introduce $\phi_M$ as basis of equivariance}
+\section{Construction}
+\subsection{Measurable Visual Components}
+
+\begin{equation}
+  \vfiberc \hookrightarrow \vtotalc \xrightarrow{\pi} \dbasec
+\end{equation}
+where
+\cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \vtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}
+\begin{equation}
+
+\end{equation}
+\subsection{$\vindexc: \gbasec \rightarrow \dbasec$}
+We define the functor $\vindex$ such that the data base space \dbase\ is a deformation retraction\cite{nlab:deformation_retract} of the graphic space $\gbase$, meaning that there is a continuous surjective mapping from every point $\dbasepoint\in\dbase$ to a point $\dbasepoint \in dbase$ .
+To simplify matters, in this paper, we construct the graphic space as a constant multiple of the base space such that
+\begin{equation}
+  \underbrace{\opensetc\times[0,1]^{n}}_{\opensetgc} \textcolor{functor}{\xmapsto{\hspace{1em}\vindexc\hspace{1em}}} \opensetc} 
+\end{equation}
+\begin{equation*}
+  n = \begin{cases}
+    dim(\gbase) - dim(\dbase) & dim(\dbase)<dim(\gbase)\\
+  0 & otherwise
+  \end{cases}
+\end{equation*}
+
+\subsection{$\nu: \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \rightarrow \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}}$}
+
+\subsection{$\vmarkc: \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} \rightarrow \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}$}
+\subsubsection{Fabricating \vmarkc's}
+
+\begin{equation}
+  \begin{tikzcd}
+    \cgamma{\opensetc}{\vtotalc^{a}\restriction_{\opensetc}} 
+    \arrow[rr, "{{\vmarkc^{a}_1}}", color=artist] 
+    \arrow[dd, "\vchannelc"', color=artist] 
+    \arrow[rrdd, "{{\vmarkc^{a}_2}}", color=artist] &  & 
+    \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+    \arrow[dd, "\dfunctc_{\gtotalc}", dotted, color=action] \\
+   &  &  \\
+    \cgamma{\opensetc}{\vtotalc^{b}\restriction_{\opensetc}} 
+    \arrow[rr, "{{\vmarkc^{b}_2}}", color=artist]   &  & 
+    \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+  \end{tikzcd}
+\end{equation}
+
 \section{Discussion}

### 2022-04-14 paper/main.tex
testing nu and q

@@
 \label{sec:atct:fiber-bundles}
-Fiber bundles, are a "unified, dimension-independent framework", as described by Butler\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}, that allow us to seperatly describe the topololigical properties and fields of a data source and also the mapping between topolology and data values.
+Fiber bundles, are a "unified, dimension-independent framework", as described by Butler\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}, that allow us to seperatly describe the topological properties and fields of a data source and also the mapping between topolology and data values.
 A fiber bundle $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and continuous surjective map $\pi: \dtotalc \rightarrow \dbasec$\cite{FiberBundle2020}.
@@
   \label{eq:atct:fb_graphic_section}
-  $\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \coloneqq \big\{\gsectionc: \opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} \; \bigm{\vert} \pi(\gsectionc(\gbasepointc)) = \gbasepointc\;for\, all\; \gbasepointc \in \opensetgc \big\}$
+  \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \coloneqq \big\{\gsectionc: \opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} \; \bigm{\vert} \pi(\gsectionc(\gbasepointc)) = \gbasepointc\;for\, all\; \gbasepointc \in \opensetgc \big\}
 \end{equation}
@@
 This means we can construct a section $\dsection$ on the collection of open sets that is the union of sections defined on specific open sets.
-This expresses an equivalency between a concatenated represetation of data and the distributed represetation of the same data.
+This expresses an equivalency between a concatenated represetation of data and the distributed representation of the same data.
 
@@
 
-\subsubsection{Measurement Types}
-
+\subsubsection{Measurement Scales}
+\label{sec:atct:sheaves:measurments}
 We specify that structure on a sheaf includes the actions $\dfuncc$ on sections of that sheaf, which is a generalization of Steven's \cite{stevensTheoryScalesMeasurement1946} classification of measurement scales by their mathematical group structure.
@@
 For example a remapping of an underlying indexing, such as a sorting of primary keys for a database of discrete values.
-This base space transfromation induces a pullback section
+This base space transformation induces a pullback section
 \begin{equation}
@@
 \end{equation}
-such that $\dsection|_{\openset} = \dfuncpull\dsection|_{\opensetc^{\prime}}$ because $\dsection|_{\openset} = \dsection|_{\dfunch(\opensetc^{prime})}$.
+such that $\dsection|_{\openset} = \dfuncpull\dsection|_{\opensetc^{\prime}}$ because $\dsection|_{\openset} = \dsection|_{\dfunch(\opensetc^{\prime})}$.
 This means that the base space transformation \dfunch\ alone does not change the data values.
@@
 \begin{equation}
-  \dfunc: \dsection\restriction_{\openset} \mapsto \dsection^{\prime}\restriction_{\openset} \circ \dfunch
+  \dfuncc: \dsectionc\restriction_{\opensetc} \mapsto \dsectionc^{\prime}\restriction_{\opensetc} \circ \dfunchc
 \end{equation}
@@
 \end{equation}
-if there exists functions $\dfunc_{a,b}: \dtotal_a \times \dtotal_b \rightarrow \dtotal_a \times \dtotal_b$, $\dfunc_{a}: \dtotal_a \rightarrow \dtotal_a$ and $\dfunc_{b}: \dtotal_b \rightarrow \dtotal_b$ s.t.
-$\pi_a \circ \dfunc_a = \dfunc_{a,b} \circ \pi_a$ and $\pi_b \circ \dfunc_b = \dfunc_{a,b} \circ \pi_b$ then $\dfunc_{a,b} = (\dfunc_a, \dfunc_b)$.
+if there exists functions $\dfunc_{a,b}: \dtotal_a \times \dtotal_b \rightarrow \dtotal_a \times \dtotal_b$, $\dfunc_{a}: \dtotal_a \rightarrow \dtotal_a$ and $\dfunc_{b}: \dtotal_b \rightarrow \dtotal_b$ such that $\pi_a \circ \dfunc_a = \dfunc_{a,b} \circ \pi_a$ and $\pi_b \circ \dfunc_b = \dfunc_{a,b} \circ \pi_b$ then $\dfunc_{a,b} = (\dfunc_a, \dfunc_b)$.
 This allows us to define a data transform where each fiber transform $\dfunct_{i}$ can be applied to a different fiber field $\dfiber_i$.
 
-\subsection{Map Between Base Spaces}
+\subsection{Maps Between Spaces}
 \note{add couple of sentences introducing image functors}
-We define a \textcolor{functor}{functor} between the graphic and data base spaces 
+We propose that there must exist a continuous \textcolor{functor}{map between topological spaces} \vindexc 
 \begin{equation}
@@
 \end{equation}
-such that every $\opensetg \subseteq \gbase$ must map to a corresponding set $\openset \subset \dbase$. In doing so, we can define the pullback \vindexpullc\  and pushforward \vindexpushc\
+such that every $\opensetg \subseteq \gbase$ must map to a corresponding set $\openset \subset \dbase$.
+Given \vindex\ we can construct two functors \vindexpullc\ and \vindexpushc\ which transport sheaves
 
@@
 \end{equation}
+such that $\vindexpullc$ pulls sheaves of sections on $\openset \subseteq \dbase$ over $\opensetg \subseteq \gbase$
+
+and $\vindexpushc$ pushes sheaves of sections on $\opensetgc$ over $\openset$
+
+The continuous map $\vindex$ and transport functors $\vindexpullc, \vindexpushc$ allow us to express the expectation that every graphic section has a corresponding data section.
 
@@
 
-As illustrated in \autoref{fig:atct:morphisms:sheaf}, the \note{image functors?} between sheaves are a way of expressising the bookkeeping of data in screen and graphic space and graphic in screen and graphic space.
+Functors between sheaves are a way of expressing the bookkeeping involved in keeping track of which graphic section corresponds to which data section.
+This allows us to construct graphic specifications for each data index $\vindexpush\gsection$ and retrieve the data $\vindexpull\dsection$ for any graphic section generating any piece of a graphic.
+As illustrated in \autoref{fig:atct:morphisms:sheaf}, the \note{image functors?}
 
-\section{Artist: Data to Graphic}
-Because $Hom_{\gbase}(\sheaf_{\gbase, \vindexpull\dtotal}, \sheaf_{\gbase,\gtotal}) = Hom_{\dbase}(\sheaf_{\dbase, \dtotal}, \sheaf_{\dbase, \vindexpush\gtotal})$
-
+We propose that visualization libraries implement morphisms from a data sheaf $\sheafc_{\dtotal, \dbase}$ to a graphic sheaf $\sheafc_{\gtotal, \gbase}$ 
 \begin{equation}
+  \label{eq:atct:sheaves:homset}
   \begin{tikzcd}
     \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
-    \arrow[dd, "\vartistc^{\dbasec}"', color=artist] 
-    \arrow[rrdd, "\vartistc", color=artist] 
+    \arrow[dd, "\textcolor{set}{Hom}_{\sheafc_{\dbasec}}"', color=homset] 
+    \arrow[rrdd, "\textcolor{set}{Hom}_{\sheafc_{\dbasec},\sheafc_{\gbasec}}", color=homset] 
     \arrow[rr, "\vindexpullc", color=functor] &  &
     \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} 
-    \arrow[dd, "\vartistc^{\gbasec}", color=artist] \\
+    \arrow[dd, "\textcolor{set}{Hom}_{\sheafc_{\gbasec}}", color=homset] \\
      & & \\
@@
 \end{equation}
+and the hom set is the set of all morphisms from one sheaf to the other.
+The functors $\vindexpullc, \vindexpushc$ are adjoint, meaning that there is a functorial isomorphism \cite{harder2008lectures} such that
+\begin{equation}
+\begin{split}
+  & \textcolor{set}{Hom}_{\sheafc_{\gbasec}}(\vindexpullc\sheafc_{\dtotalc, \dbasec},\sheafc_{\gtotalc,\gbasec})\\
+  \simeq & \textcolor{set}{Hom}_{\sheafc_{\dbasec}}(\sheafc_{\dtotalc, \dbasec},\vindexpushc\sheafc_{\gtotalc,\gbasec}) \\
+  \simeq & \textcolor{set}{Hom}_{\sheafc_{\dbasec},\sheafc_{\gbasec}}(\sheafc_{\dtotalc, \dbasec},\sheafc_{\gtotalc,\gbasec}) \\
+\end{split} 
+\end{equation}
+which means that the diagram of sheaf morphisms \autoref{eq:atct:sheaves:homset} is commutative; therefore the functors $\vindexpullc$ and $\vindexpushc$ can be used to adapt morphisms written over one space to another space.
+This provides us flexibility over which space to construct data to graphic transforms.
+For example, specifications such as svg\cite{quintScalable2003} and Vega\cite{satyanarayanDeclarativeInteractionDesign2014}, typically define their transformations in data space and return specifications for the render to evaluate.
+It may also be
+An artist may also want to specify the transformations in a proxy of graphic space, for example for operations like dynamic resampling \note{citation:root maybe}.
 
+\section{Artist: Data to Graphic}
+\label{sec:artist}
+In this work we propose that visualization libraries are implementing a subset of the functions in the hom sets in \autoref{eq:atct:sheaves:homset} and we call these subset of functions the artist:
+\begin{align}
+  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} &\textcolor{artist}{\xrightarrow{\vartistc}} \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} &\in \textcolor{set}{Hom}_{\sheafc_{\dbase}, \sheafc_{\gbase}}\\
+  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} &\textcolor{artist}{\xrightarrow{\vartistc^{\dbasec}}} \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} &\in \textcolor{set}{Hom}{\sheafc_{\dbase}}\\
+  \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} &\textcolor{artist}{\xrightarrow{\vartistc^{\gbasec}}} \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} &\in \textcolor{set}{Hom}_{\sheafc_{\gbase}}\\
+\end{align}
+
+The space of all sections $\cgamma{\opensetg}{\gtotal\restriction_{\openset}}$ includes sections that may not be structure preserving.
+For example, a section may go from every point in the graphic space to the same single point in the graphic fiber $\gsection(\gbasepoint_i) = d\; \forall \gbasepoint \in \gbase$ such that the visual output is a single inked pixel on a screen.
+
+The output of an artist \vartist\ is a restricted subset of graphic sections
 \begin{equation}
@@
   \vartistc(\dsectionc) = \gsectionc,\; \vindexc(\opensetgc) = \opensetc \} 
-\end{equation}
-$\imartist{\opensetg}{\gtotal\restriction_{\opensetg}} \subset \cgamma{\opensetg}{\gtotal\restriction_{\opensetg}}$
+\end{equation} that are by definition only reachable through a structure preserving artist, which we describe in \autoref{sec:artist:equivariance}.
 
-\subsection{Equivariant Artist}
+\subsection*{Equivariance}
+\label{sec:artist:equivariance}
 \begin{equation*}
@@
 \end{equation*}
+
 \subsection*{Testing Equivariance}
@@
   \begin{description}
-      \item[input]{$\equivc: \dsectionc \mapsto (\opensetc \xrightarrow{\equivc_{\dsectionc}} \measurec)$}
-      \item[output]{$\extractmc:\gsectionc \mapsto (\opensetgc \xrightarrow{\extractmc_{\gsectionc}} \measurec)$}
+      \item[expected]{$\equivc: \dsectionc \mapsto (\opensetc \xrightarrow{\equivc_{\dsectionc}} \measurec)$}
+      \item[actual]{$\extractmc:\gsectionc \mapsto (\opensetgc \xrightarrow{\extractmc_{\gsectionc}} \measurec)$}
       \item[] $\equivc_{\dsectionc}(\dbasepointc) = \extractmc_{\gsectionc}(\gbasepointc)$ for all $\vindexc(\gbasepointc) = \dbasepointc, \dbasepointc \in \dbasec, \gbasepointc \in \gbasec$ 
@@
   \note{introduce $\phi_M$ as basis of equivariance}
+
 \section{Construction}
+\label{sec:construction}
 \subsection{Measurable Visual Components}
@@
 where
-\cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \vtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}
 \begin{equation}
+\cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} \coloneqq \big\{\vsectionc: \opensetc\rightarrow \vtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\vsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}
+\end{equation}
 
-\end{equation}
+The bundle $\vtotal$ is structurally equivalent to the bundle $\dtotal$, but allows for separating the data that is input into the artist $\dsection$ from the data internal to the artist $\vsection$.
+This separation also allows for defining many of the target measurements \measure\ and the equivariant transformations to the measurements as internal components of the artists.
+
 \subsection{$\vindexc: \gbasec \rightarrow \dbasec$}
@@
 \subsection{$\nu: \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \rightarrow \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}}$}
+The transformation $\vchannel$ is a pointwise transformation $\pi(\dtotal) = \pi(\vchannel(\dtotal))$ such that 
+\begin{equation}
+  \label{eq:constrution:nu}
+  \vchannelc: \dfiberc_{\dfiberc} \rightarrow \vfiberc_{\dbasepoint}
+\end{equation}
+which means that every point in a data fiber $\delement \in \dfiber_{\dbasepointc}$ can be mapped into a correspoindg point in the visual fiber $\velement \in \vfiber_{\dbasepointc}$.
+A pair of transformations $(\vchannel, \vchannel^{\prime})$ are equivariant if
+\begin{equation}
+  \dfunct_{\vtotal} \vchannel(\dsection(\dbasepoint)) = \vchannel^{\prime}(\dfunct_{\dtotal}(\dsection(\dbasepoint)))\; \forall \dbasepoint \in \dbase
+\end{equation}
+and a transformation $\vchannel$ is equivariant when $\vchannel = \vchannel^{\prime}$
+\subsection*{Compositing \vchannelc's}
+
+\begin{equation}
+  \begin{tikzcd}[col sep=Huge]
+    \dfiberc_{\dbasepointc} 
+    \arrow[rr, "\vchannelc", color=artist] 
+    \arrow[rrrr, "\vchannelc^{\prime\prime}", dashed, bend right, color=artist] &  & 
+    \vfiberc_{\dbasepointc}\coloneqq{\dfiberc_{\dbasepointc}^{\prime}} 
+    \arrow[rr, "\vchannelc^{\prime}", color=artist] &  & 
+    \vfiberc^{\prime}_{\dbasepointc}
+  \end{tikzcd}
+\end{equation}
+
+\subsection*{Sharing \vchannelc's}
+\begin{equation}
+  \begin{tikzcd}[column sep=4em]
+    {\dfiberc_{\dbasepointc}^{a}} \times {\dfiberc_{\dbasepointc}^{b}} 
+    \arrow[d, "\pi_a"', color=fiber] 
+    \arrow[r, "\vchannelc_{ab}", color=artist] 
+    \arrow[rr, "\equivc_{ab}", bend left, color=monoid]  & 
+    {\vfiberc_{\dbasepointc}^{a}} \times {\vfiberc_{\dbasepointc}^{b}} 
+    \arrow[d, "\pi_a", color=fiber] & 
+    \measurec_{\dbasepointc}^{ab} 
+    \arrow[d, "\measurec^{ab}\restriction_a", color=set] \\
+    \dfiberc_{\dbasepointc}^a 
+    \arrow[r, "\vchannelc_{a}", dashed, color=artist] & 
+    \vfiberc_{\dbasepointc}^a 
+    \arrow[r, "\simeq", dotted]  & 
+    \measurec_{\dbasepointc}^a   \\
+    {\dfiberc_{\dbasepointc}^{a}} \times {\dfiberc_{\dbasepointc}^{c}} 
+    \arrow[u, "\pi_a", color=fiber] 
+    \arrow[r, "\vchannelc_{ac}", color=artist] 
+    \arrow[rr, "\equivc_{ac}", bend right, color=monoid] & 
+    {\vfiberc_{\dbasepointc}^{a}} \times {\vfiberc_{\dbasepointc}^{c}} 
+    \arrow[u, "\pi_a"', color=fiber] & 
+    \measurec_{\dbasepointc}^{ac} 
+    \arrow[u, "\measurec^{ac}\restriction_a"', color=set]            
+    \end{tikzcd}
+\end{equation}
 
@@
   \begin{tikzcd}
-    \cgamma{\opensetc}{\vtotalc^{a}\restriction_{\opensetc}} 
-    \arrow[rr, "{{\vmarkc^{a}_1}}", color=artist] 
-    \arrow[dd, "\vchannelc"', color=artist] 
-    \arrow[rrdd, "{{\vmarkc^{a}_2}}", color=artist] &  & 
-    \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
-    \arrow[dd, "\dfunctc_{\gtotalc}", dotted, color=action] \\
-   &  &  \\
-    \cgamma{\opensetc}{\vtotalc^{b}\restriction_{\opensetc}} 
-    \arrow[rr, "{{\vmarkc^{b}_2}}", color=artist]   &  & 
-    \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+      \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} 
+      \arrow[rr, "\vchannel", color=artist] 
+      \arrow[rrrr, "\vmarkc^{\prime}"', bend right, color=artist, dashed] &  & \cgamma{\opensetc}{\vtotalc^{\prime}\restriction_{\opensetc}} 
+      \arrow[rr, "\vmarkc", color=artist] &  & \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}
       \end{tikzcd} 
 \end{equation}
+
+\subsubsection{Combining \vmarkc's}
+
+\subsubsection{Shared Visual Encodings}
+\begin{equation}
+  \begin{tikzcd}[row sep=2.5em, column sep=.75sem]
+    &  & \imartistsub{ab}{\opensetgc}{\gtotalc\restriction\opensetgc}
+    \arrow[rrd, "\extractmc_{bc}", color=monoid] &  &  \\
+\cgamma{\opensetc}{\vtotalc^{a}\restriction_{\opensetc}\times \vtotalc^{b}\restriction_{\opensetc}} 
+\arrow[rru, "\vmarkc_{ab}", color=artist] 
+\arrow[d, "\pi_a"', color=total] &  &  &  & 
+{{\textcolor{set}{Hom}(\opensetgc, \measurec^{ab})}} 
+\arrow[d, "\measurec^{ab}\restriction_a", color=set]  \\
+\cgamma{\opensetc}{\vtotalc^{a}\restriction_{\opensetc}} 
+\arrow[rr, "\simeq", dotted]  &  & 
+{{\textcolor{set}{Hom}(\opensetc, \measurec^{a})}} 
+\arrow[rr, "\vindexpullc", color=functor] &  & 
+{{\textcolor{set}{Hom}(\opensetgc, \measurec^{a})}} \\
+\cgamma{\opensetc}{\vtotalc^{a}\restriction_{\opensetc}\times \vtotalc^{c}\restriction_{\opensetc}} 
+\arrow[rrd, "\vmarkc_{bc}", color=artist] 
+\arrow[u, "\pi_a", color=total]  &  &  &  & 
+{{\textcolor{set}{Hom}(\opensetgc, \measurec^{ac})}} 
+\arrow[u, "\measurec^{ab}\restriction_a"', color=set] \\
+ &  & \imartistsub{ac}{\opensetgc}{\gtotalc\restriction\opensetgc}
+ \arrow[rru, "\extractmc_{ab}", color=monoid] &  & 
+\end{tikzcd}
+\end{equation}
+\note{when Q_{ab} = Q_{bc}, this says only things that change are the parameter controlled pieces}
 

### 2022-04-21 paper/main.tex
math implementation bridge

@@
 
-The \textcolor{base}{base space} is a topological space \dbasec\ with points $\dbasepointc \in \dbasec$ and topology $\mathcal{T}_{\dbasec}$; $\mathcal_{T}_k$ is a set of opensets $ \dbasepointc \in \opensetc \subseteqq \dbasec$ \cite{munkresElementsAlgebraicTopology1984} that cover the base space \dbasec.
-The \textcolor{fiber space} is a topological space \dfiberc\ that is the preimage of the projection function $\pi$ at a point in the base space $\dfiberc|_{\dbasepointc} = \pi^{-1}(\dbasepointc)$ and fibers of a bundle are isomorphic $\dfiberc \simeq \dfiberc_{\dbasepointc}\;\forall \dbasepointc \in \dbase$.
+The \textcolor{base}{base space} is a topological space \dbasec\ with points $\dbasepointc \in \dbasec$ and topology $\mathcal{T}_{\dbasec}$; $\mathcal{T}_k$ is a set of opensets $ \dbasepointc \in \opensetc \subseteqq \dbasec$ \cite{munkresElementsAlgebraicTopology1984} that cover the base space \dbasec.
+The \textcolor{fiber}{fiber space} is a topological space \dfiberc\ that is the preimage of the projection function $\pi$ at a point in the base space $\dfiber_{\dbasepointc} = \pi^{-1}(\dbasepointc)$ and fibers of a bundle are isomorphic $\dfiberc \simeq \dfiberc_{\dbasepointc}\;\forall \dbasepointc \in \dbase$.
 The \textcolor{section}{sections} of a bundle $\dsection$ are maps from the base space to points in the bundles over that base space and $\textcolor{set}{\Gamma}$ denotes the set of all sections in a bundle over an openset.
@@
 For example, a temperature field could be expressed as a fiber $\dfiber_{temperature}=\mathbb{R}$, and a color field could be $\dfiber_{color}=\mathbb{R}^{3}$.
-Even though fibers on non-trivial spaces, such as vector fields, are different, we can simplify to one type because fibers are isomorphic.\note{unsure of this claim but I think this is why, and also I need to tell you why you care that they're isomorphic} The total bundle space \dtotalc\ acts as a way of expressing all data with the same topolology \dbasec\ and fields \dfiberc, for example a database schema.
+Even though fibers on non-trivial spaces, such as vector fields, are different, we can simplify to one type because fibers are isomorphic.\note{unsure of this claim but I think this is why, and also I need to tell you why you care that they're isomorphic} The total bundle space \dtotalc\ acts as a way of expressing all data with the same topology \dbasec\ and fields \dfiberc, for example a database schema.
 A specific dataset is represented as a section \dsectionc, and the return values of a section evaluated at a point \dbasepointc\ is the record $\{field_{0}: value_{0}, ..., field_n:value_n\}$ at that point.
@@
 \subsection*{Testing Equivariance}
-\begin{equation*}
+\begin{equation}
   \begin{tikzcd}[row sep=huge]
     \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
-    \arrow[d, "\equivc", color=monoid]
-     \arrow[rr, "\vartistc", color=artist] &  & \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
-     \arrow[d, "\extractmc", color=monoid] \arrow[r, "render"] & VIZ \arrow[ld, "measure"] \\
-    {Hom(\opensetc, \measurec)} \arrow[rr, "\vindexpullc", color=functor]  &  & {Hom(\opensetgc, \measurec)}   & 
+    \arrow[rr, "\vartistc", color=artist] 
+    \arrow[d, "\equivc"', color=monoid] &  & 
+    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+    \arrow[d, "render"] 
+    \arrow[lld, "\extractmc"', color=monoid, dashed] \\
+    {\textcolor{set}{Hom}(\opensetc, \measurec)}  &  & visualization 
+    \arrow[ll, "measure",]
     \end{tikzcd}
-  \end{equation*}
+  \end{equation}
   \begin{description}
-      \item[expected]{$\equivc: \dsectionc \mapsto (\opensetc \xrightarrow{\equivc_{\dsectionc}} \measurec)$}
-      \item[actual]{$\extractmc:\gsectionc \mapsto (\opensetgc \xrightarrow{\extractmc_{\gsectionc}} \measurec)$}
-      \item[] $\equivc_{\dsectionc}(\dbasepointc) = \extractmc_{\gsectionc}(\gbasepointc)$ for all $\vindexc(\gbasepointc) = \dbasepointc, \dbasepointc \in \dbasec, \gbasepointc \in \gbasec$ 
+      \item[expected]{\;$\equivc: \dsectionc \mapsto (\opensetc \xrightarrow{\equivc_{\dsectionc}} \measurec)$}
+      \item[actual]{$\extractmc:\gsectionc \mapsto (\opensetc \xrightarrow{\extractmc_{\gsectionc}} \measurec)$}
     \end{description}
+    such that
+    \begin{equation}
+    \equivc(\dsectionc)(\dbasepointc) = \extractmc(\vartistc(\dsectionc))(\dbasepointc) = \extractmc(\gsectionc\circ\vindexprec)(\dbasepointc)
+    \end{equation}
 
@@
 \begin{equation}
-  \underbrace{\opensetc\times[0,1]^{n}}_{\opensetgc} \textcolor{functor}{\xmapsto{\hspace{1em}\vindexc\hspace{1em}}} \opensetc} 
+  \underbrace{\opensetc\times[0,1]^{n}}_{\opensetgc} \textcolor{functor}{\xmapsto{\hspace{1em}\vindexc\hspace{1em}}} \opensetc 
 \end{equation}
@@
 and a transformation $\vchannel$ is equivariant when $\vchannel = \vchannel^{\prime}$
-\subsection*{Compositing \vchannelc's}
+\subsection*{Fabricating \vchannelc's}
 
@@
 \end{equation}
+\subsubsection{Combining \vchannelc's} 
+We can add visual parameters as
 
-\subsection*{Sharing \vchannelc's}
+\subsection*{Validating \vchannelc's}
 \begin{equation}
@@
 
-\subsection{$\vmarkc: \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} \rightarrow \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}$}
+\subsection{Graphic Compositing Function}
+The compositing function $\vmarkc$ transforms the sections of visual elements $\vsectionc$ into sections of graphics $\gsectionc$.
+\begin{equation}
+  \vmarkc: \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} \rightarrow \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}
+\end{equation}
+
+\note{some transition about sheaf morphisms}
+Therefore an equivariant $\vmarkc$ satisfies the conditions....
+
 \subsubsection{Fabricating \vmarkc's}
@@
 \subsubsection{Combining \vmarkc's}
+A set of $\vmarkc$ functions that output graphic sections, as introduced in \autoref{eq:atct:fb_graphic_section}, that go to the same space, for example to the same figure in a screen, are returning a set of functions into the same total bundle $\gtotal$;
+\begin{equation}
+  \label{eq:construction:qc1}
+  \bigsqcup_i \vmark_i(\Gamma(\openset_i,\dtotal_i\restriction_{\openset_i})) = \bigsqcup_i \Gamma(\opensetg_i, \gtotal\restriction_{\opensetg_i})
+\end{equation}
+A property of hom-sets is that $Hom(W_i,H) + Hom(W_j,H) = Hom(W_i+W_j,H)$, which means that the union of hom-sets with the same target is equal to the hom-set of the union of sources to the target; therefore
+\begin{equation}
+  \label{eq:construction:qc2}
+  \bigsqcup_i\Gamma(\opensetg_i, \gtotal\restriction_{\opensetg_i}) = \Gamma(\mathop{\sqcup}_{i}\opensetg_i, \gtotal\restriction_{{\mathop{\sqcup}_i}\opensetg_i})
+\end{equation}
+the union of sets of functions that generate parts of an image is equal to a set of functions that can generate the whole image.
+This implies that we can define a $\vmarkc$ that generates the set of sections over the whole image
+\begin{equation}
+  \label{eq:construction:qc3}
+\Gamma(\mathop{\sqcup}_i\opensetg_i, \gtotal\restriction_{\mathop{\sqcup}_{i}\opensetg_i}) \implies \vmark(\mathop{\sqcup}_{i}\Gamma(\openset_i,\vtotal_i\restriction_{\openset_i}) 
+\end{equation}
+which means that, following from \autoref{eq:constrution:qc1} and \autoref{eq:construction:qc3},
+\begin{equation}
+  \label{eq:construction:qc4}
+  \vmarkc(\mathop{\sqcup}_{i}\cgamma{\opensetc_i}{\vtotalc_i\restriction_{\opensetc_i}}) =   
+  \bigsqcup_i \vmarkc_i(\cgamma{\opensetc_i}{\vtotalc_i\restriction_{\opensetc_i}})
+\end{equation}
+we can construct a $\vmarkc = \sqcup_i\vmarkc_i$.v This allows us to .
 
-\subsubsection{Shared Visual Encodings}
-\begin{equation}
-  \begin{tikzcd}[row sep=2.5em, column sep=.75sem]
-    &  & \imartistsub{ab}{\opensetgc}{\gtotalc\restriction\opensetgc}
-    \arrow[rrd, "\extractmc_{bc}", color=monoid] &  &  \\
-\cgamma{\opensetc}{\vtotalc^{a}\restriction_{\opensetc}\times \vtotalc^{b}\restriction_{\opensetc}} 
-\arrow[rru, "\vmarkc_{ab}", color=artist] 
-\arrow[d, "\pi_a"', color=total] &  &  &  & 
-{{\textcolor{set}{Hom}(\opensetgc, \measurec^{ab})}} 
-\arrow[d, "\measurec^{ab}\restriction_a", color=set]  \\
-\cgamma{\opensetc}{\vtotalc^{a}\restriction_{\opensetc}} 
-\arrow[rr, "\simeq", dotted]  &  & 
-{{\textcolor{set}{Hom}(\opensetc, \measurec^{a})}} 
-\arrow[rr, "\vindexpullc", color=functor] &  & 
-{{\textcolor{set}{Hom}(\opensetgc, \measurec^{a})}} \\
-\cgamma{\opensetc}{\vtotalc^{a}\restriction_{\opensetc}\times \vtotalc^{c}\restriction_{\opensetc}} 
-\arrow[rrd, "\vmarkc_{bc}", color=artist] 
-\arrow[u, "\pi_a", color=total]  &  &  &  & 
-{{\textcolor{set}{Hom}(\opensetgc, \measurec^{ac})}} 
-\arrow[u, "\measurec^{ab}\restriction_a"', color=set] \\
- &  & \imartistsub{ac}{\opensetgc}{\gtotalc\restriction\opensetgc}
- \arrow[rru, "\extractmc_{ab}", color=monoid] &  & 
+\begin{figure}
+  \label{fig:construction:combined_q}
+  \includegraphics[width=1\colunwidth]{qsketchh.png}
+  \caption{}
+\end{figure}
+
+As illustrated in \autoref{fig:construction:combined_q}, this composition rule expresses the construction of a multipart circle-square graphic out of circles and squares produced by different $\vmarkc$.
+\note{rewrite:Also figures are the unions of their parts}.
+
+\subsubsection{Validating \vmarkc's}
+
+\begin{tikzcd}[column sep=.5em, row sep=1em]
+  \cgamma{\opensetc}{(\vtotalc^{a}\times \vtotalc^{b})\restriction_{\opensetc}} 
+    \arrow[rr, "\vmarkc_{ab}", color=artist] 
+    \arrow[dd, "\pi_a"', color=total]                &  & \imartistsub{ab}{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \arrow[rd, "\extractmc_{ab}",color=monoid]   &                                                                                                   \\
+                                                                                                                               &  &                                                                                                            & {{\textcolor{set}{Hom}(\opensetc, \measurec^{ab})}} \arrow[ld, "\measurec\restriction_{a}", color=set]  \\
+  \cgamma{\opensetc}{\vtotalc^{a}\restriction_{\opensetc}} \arrow[rr, "\simeq", dotted]                                        &  & {{\textcolor{set}{Hom}(\opensetc, \measurec^{a})}}                                                         &                                                                                                   \\
+                                                                                                                               &  &                                                                                                            & {{\textcolor{set}{Hom}(\opensetc, \measurec^{ac})}} \arrow[lu, "\measurec\restriction_{a}"', color=set] \\
+  \cgamma{\opensetc}{(\vtotalc^{a}\times \vtotalc^{c})\restriction_{\opensetc}} \arrow[rr, "\vmarkc_{ac}", color=artist] \arrow[uu, "\pi_a", color=total] &  & \imartistsub{ac}{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \arrow[ru, "\extractmc_{ac}"', color=monoid] &
   \end{tikzcd}
-\end{equation}
-\note{when Q_{ab} = Q_{bc}, this says only things that change are the parameter controlled pieces}
+\note{when $Q_{ab} = Q_{bc}$, this says only things that change are the parameter controlled pieces}
+
+\subsection{Implementation Options: $\vartistc=\vmarkc\circ\vchannelc$}
+As mentioned in \autoref{sec:atct:sheaves}, sheaves equipped with transport functors have isomorphic maps between the sheaves.
+This property allows us to specify the transformation from data to graphic space in data or graphic space or a mix of both such that the following diagram commutes:
+
+\begin{tikzcd}[row sep=2.5em, column sep=1.5em]
+  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
+  \arrow[rr, "\vchannelc", color=artist] 
+  \arrow[rrrr, "\vartistc^{\dbasec}", bend left, color=artist] 
+  \arrow[dd, "\vindexpullc"', color=functor] 
+  \arrow[rrrrdd, "\vartistc", color=artist, pos=.2] &  & 
+  \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} 
+  \arrow[rrdd, "\vmarkc", color=artist] 
+  \arrow[rr, "\vmarkc^{\dbasec}", color=artist] 
+  \arrow[dd, "\vindexpullc", color=functor, pos=.2] &  & \imartist{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}  \\
+   & & & & \\
+  \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} \arrow[rr, "\vchannelc", color=artist] 
+  \arrow[rrrr, "\vartistc^{\gbasec}", bend right, color=artist] & & 
+  \cgamma{\opensetgc}{\vindexpullc\vtotalc\restriction_{\opensetgc}} \arrow[rr, "\vmarkc^{\gbasec}", color=artist] &  & 
+  \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \arrow[uu, "\vindexpushc"', color=functor]
+\end{tikzcd}
 

### 2022-04-25 paper/main.tex
fleshed out push pull natural transforms implementation q testing

@@
 \usepackage{amsmath, amsthm, amsfonts,amssymb,eulervm,xspace, mathtools}
+\usepackage{stmaryrd}%mapsfrom 
 \renewcommand{\restriction}{\mathord{\upharpoonright}} %restriction w/p space
@@
 \begin{equation}
-  \label{eq:atct:sheaves:morphism}
+  \label{eq:atct:sheaves:monoid_morphism}
   \begin{tikzcd}
@@
 such that every $\opensetg \subseteq \gbase$ must map to a corresponding set $\openset \subset \dbase$.
-Given \vindex\ we can construct two functors \vindexpullc\ and \vindexpushc\ which transport sheaves
+Given \vindex\ we can construct two functors \vindexpull\ and \vindexpush\ which transport sheaves
 
@@
   \opensetc 
-  \arrow[d, "{\sheafc_{\dbasec, \vindexpushc \gtotalc}}", dashed, maps to, color=sheaf] 
+  \arrow[d, "{\vindexpushc\sheafc_{\gbasec,  \gtotalc}}", dashed, maps to, color=sheaf] 
   \arrow[u, "{\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf] &  & 
@@
   \arrow[d, "{\sheafc_{\gbasec, \gtotalc}}", maps to, color=sheaf] 
-  \arrow[u, "{\sheafc_{\gbasec, \vindexpullc\dtotalc}}"', dashed, maps to, color=sheaf] 
+  \arrow[u, "{\vindexpullc\sheafc_{\dbasec, \dtotalc}}"', dashed, maps to, color=sheaf] 
   \arrow[ll, "\vindexc"', maps to, color=functor] \\
@@
 \end{equation}
-such that $\vindexpullc$ pulls sheaves of sections on $\openset \subseteq \dbase$ over $\opensetg \subseteq \gbase$
+such that there is an association between graphic sections \gsection\ that take as input \gbasepoint\ and data sections that take input \dbasepointc\ when $\vindex(\gbasepoint) = \dbasepoint$. 
+The \textcolor{functor}{pullback functor} $\vindexpullc$ transports sheaves of sections on $\openset \subseteq \dbase$ over $\opensetg \subseteq \gbase$
+\begin{equation}
+  \vindexpullc\dsectionc: \opensetgc \rightarrow \vindexpullc \dtotalc\restriction_{\opensetgc} \in \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} 
+\end{equation}
+such that there is a way to then look up what data values correspond with a graphic index
+\begin{equation}
+  \vindexpullc\dsectionc(\gbasepointc) = \dsectionc(\vindex(\gbasepointc)) = \dsectionc(\dbasepointc)
+\end{equation}
 
-and $\vindexpushc$ pushes sheaves of sections on $\opensetgc$ over $\openset$
+ The \textcolor{functor}{pushforward functor} $\vindexpushc$ transports sheaves of sections on $\opensetgc$ over $\openset$
+ \begin{equation}  
+  \vindexpushc\gsectionc: \opensetc \rightarrow \vindexpushc \gtotalc\restriction_{\opensetc} \in \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} 
+\end{equation}
+such that it provides a way to look up which graphic  corresponds with a data index
+\begin{equation}
+  \vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexprec(\dbasepointc)} = \gsectionc(\gbasepointc)\;\forall \gbasepointc\in \vindexprec(\dbasepointc)
+\end{equation}
 
-The continuous map $\vindex$ and transport functors $\vindexpullc, \vindexpushc$ allow us to express the expectation that every graphic section has a corresponding data section.
+Therefore, the continuous map $\vindex$ and transport functors $\vindexpull, \vindexpush$ allow us to express the correspondence between graphic section and data section.
 
@@
 \begin{equation}
+  \label{eq:atct:sheaves:homset:functors}
 \begin{split}
@@
 \begin{align}
-  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} &\textcolor{artist}{\xrightarrow{\vartistc}} \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} &\in \textcolor{set}{Hom}_{\sheafc_{\dbase}, \sheafc_{\gbase}}\\
-  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} &\textcolor{artist}{\xrightarrow{\vartistc^{\dbasec}}} \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} &\in \textcolor{set}{Hom}{\sheafc_{\dbase}}\\
-  \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} &\textcolor{artist}{\xrightarrow{\vartistc^{\gbasec}}} \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} &\in \textcolor{set}{Hom}_{\sheafc_{\gbase}}\\
+  \label{eq:artist:hom_transport}
+  \vartistc:& \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}\textcolor{artist}{\rightarrow} \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \in \textcolor{set}{Hom}_{\sheafc_{\dbase}, \sheafc_{\gbase}}\\
+  \vartistc^{\dbasec}:&\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}  \textcolor{artist}{\rightarrow} \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} \in \textcolor{set}{Hom}{\sheafc_{\dbase}}\\
+  \vartistc^{\gbasec}:& \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} \textcolor{artist}{\rightarrow} \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \in \textcolor{set}{Hom}_{\sheafc_{\gbase}}
 \end{align}
+ Because the artists can be constructed as morphisms of sheave over the same base spaces, as listed in \autoref{eq:atct:sheaves:homset:functors}, they are natural transformations
+\begin{align}
+  \label{eq:artist:natural_transform}
+  \vartistc^{\dbasec}:& \sheafc_{\dbasec, \dtotalc} \textcolor{artist}{\Rightarrow} \vindexpushc \sheafc_{\gbasec, \gtotalc}\\
+  \vartistc^{\gbasec}:&\vindexpullc\sheafc_{\dbasec, \dtotalc} \textcolor{artist}{\Rightarrow}\sheafc_{\gbasec, \gtotalc}
+\end{align}
+which means that they are maps of functors that take the same input object and return objects in the same category\cite{milewskiCategoryTheoryProgrammers}.
+As illustrated in \autoref{eq:atct:sheaves:homset}, the sheaf functors
+\begin{equation}
+    \begin{tikzcd}
+      \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} &  & \opensetc \arrow[ll, "{\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf] \arrow[rr, "{\vindexpushc\sheafc_{\gbasec, \gtotalc}}", maps to, color=sheaf] &  & \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}
+      \end{tikzcd}
+\end{equation}
+take as input the same object \opensetc\ and return sets of data and graphic sections that are objects in \setb.
+As a map between these sheaf functors, the artist has to preserve the $\iota, \iota^*$ morphisms of the presheaf functor, described in \autoref{eq:atct:presheaf}, such that the following diagram commutes
 
-The space of all sections $\cgamma{\opensetg}{\gtotal\restriction_{\openset}}$ includes sections that may not be structure preserving.
-For example, a section may go from every point in the graphic space to the same single point in the graphic fiber $\gsection(\gbasepoint_i) = d\; \forall \gbasepoint \in \gbase$ such that the visual output is a single inked pixel on a screen.
+\begin{equation}
+  \label{eq:artist:natural_transform:inclusions}
+  \begin{tikzcd}
+    \opensetc_1 & \cgamma{\opensetc_{1}}{\dtotalc\restruction_{\opensetc_1}} \arrow[dd, "\iota^*"', color=set] \arrow[rr, "\vartistc", color=artist] &  & \cgamma{\opensetc_{1}}{\vindexpushc\gtotalc\restruction_{\opensetc_1}} \arrow[dd, "\iota^*", color=set] \\
+      &  &  &  \\
+    \opensetc_{2} \arrow[uu, "\iota", hook, color=base] & \cgamma{\opensetc_{2}}{\dtotalc\restruction_{\opensetc_2}} \arrow[rr, "\vartistc", color=artist]                        &  & \cgamma{\opensetc_{2}}{\vindexpushc\gtotalc\restruction_{\opensetc_2}}
+    \end{tikzcd}
+\end{equation}
+ The diagram in \autoref{eq:artist:natural_transform} shows that restricting a set of outputs of an artist to a set of graphic sections over a subspace is equivalent to restricting the inputs to data sections over the same subspace.
+Since the input and output of the artist is always sections over the same subspace, this is why the by construction the artist preserves the bookkeeping provided through the use of sheaves.
 
@@
 \end{equation} that are by definition only reachable through a structure preserving artist, which we describe in \autoref{sec:artist:equivariance}.
+We define this subset because the space of all sections $\cgamma{\opensetg}{\gtotal\restriction_{\openset}}$ includes sections that may not be structure preserving.
+For example, a section may go from every point in the graphic space to the same single point in the graphic fiber $\gsection(\gbasepoint_i) = d\; \forall \gbasepoint \in \gbase$ such that the visual output is a single inked pixel on a screen.
 
@@
 
-\subsection{$\nu: \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \rightarrow \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}}$}
-The transformation $\vchannel$ is a pointwise transformation $\pi(\dtotal) = \pi(\vchannel(\dtotal))$ such that 
+\subsection{}
+The transformation $\vchannel$ is a map from data sections to graphic sections
 \begin{equation}
-  \label{eq:constrution:nu}
-  \vchannelc: \dfiberc_{\dfiberc} \rightarrow \vfiberc_{\dbasepoint}
+\nu: \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \rightarrow \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}}
 \end{equation}
-which means that every point in a data fiber $\delement \in \dfiber_{\dbasepointc}$ can be mapped into a correspoindg point in the visual fiber $\velement \in \vfiber_{\dbasepointc}$.
-A pair of transformations $(\vchannel, \vchannel^{\prime})$ are equivariant if
+such that the sections project to the same point on the base space $\pi(\dtotal) = \pi(\vchannel(\dtotal))$. A pair of transformations $(\vchannel, \vchannel^{\prime})$ are equivariant if 
 \begin{equation}
@@
 and a transformation $\vchannel$ is equivariant when $\vchannel = \vchannel^{\prime}$
+
 \subsection*{Fabricating \vchannelc's}
+A consequence of $\pi(\dtotal) = \pi(\vchannel(\dtotal))$ is that $\vchannelc$ can be constructed as a pointwise transformation such that  
+\begin{equation}
+  \label{eq:constrution:nu}
+  \vchannelc: \dfiberc_{\dbasepointc} \rightarrow \vfiberc_{\dbasepoint}
+\end{equation}
 
+This property means that a point in a single data fiber $\delement \in \dfiber_{\dbasepointc}$ can be mapped into a corresponding point in a visual fiber $\velement \in \vfiber_{\dbasepointc}$.
+We use this property to construct compositions of $\vchannelc$ functions
 \begin{equation}
+  \label{eq:construction:nu:fabrication}
   \begin{tikzcd}[col sep=Huge]
@@
 \end{equation}
+such that any new $\vchannel$ can be constructed as a composition of two or more $\vchannel$s.
+
 \subsubsection{Combining \vchannelc's} 
@@
     \measurec_{\dbasepointc}^{ab} 
-    \arrow[d, "\measurec^{ab}\restriction_a", color=set] \\
+    \arrow[d, "\measurec\restriction_a", color=set] \\
     \dfiberc_{\dbasepointc}^a 
@@
     \measurec_{\dbasepointc}^{ac} 
-    \arrow[u, "\measurec^{ac}\restriction_a"', color=set]            
+    \arrow[u, "\measurec\restriction_a"', color=set]            
     \end{tikzcd}
@@
 \begin{equation}
+  \label{eq:construction:q:fabrication}
   \begin{tikzcd}
@@
 \subsubsection{Validating \vmarkc's}
+\begin{equation}
+  \begin{tikzcd}[row sep=huge]
+    \cgamma{\opensetc}{(\vtotalc^{a}\times \vtotalc^{b})\restriction_{\opensetc}}  
+    \arrow[rr, "\vmarkc_{ab}", color=artist] \arrow[d, "\pi_a"', color=total] &  &  \imartistsub{ab}{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+    \arrow[d, "\measurec\restriction_a \circ \extractmc_{ab}", color=monoid]  \\
+   \cgamma{\opensetc}{\vtotalc^{a}\restriction_{\opensetc}} 
+   \arrow[rr, "\simeq", dotted] &  & {\textcolor{set}{Hom}(\opensetc, \measurec^{a})}                                                                             \\
+    \cgamma{\opensetc}{(\vtotalc^{a}\times \vtotalc^{c})\restriction_{\opensetc}}  \arrow[rr, "\vmarkc_{ab}", color=artist] \arrow[u, "\pi_a", color=total]  &  &  \imartistsub{ac}{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \arrow[u, "\measurec\restriction_a \circ \extractmc_{ac}"', color=monoid]
+   \end{tikzcd}
+\end{equation}
 
-\begin{tikzcd}[column sep=.5em, row sep=1em]
-  \cgamma{\opensetc}{(\vtotalc^{a}\times \vtotalc^{b})\restriction_{\opensetc}} 
-    \arrow[rr, "\vmarkc_{ab}", color=artist] 
-    \arrow[dd, "\pi_a"', color=total]                &  & \imartistsub{ab}{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \arrow[rd, "\extractmc_{ab}",color=monoid]   &                                                                                                   \\
-                                                                                                                               &  &                                                                                                            & {{\textcolor{set}{Hom}(\opensetc, \measurec^{ab})}} \arrow[ld, "\measurec\restriction_{a}", color=set]  \\
-  \cgamma{\opensetc}{\vtotalc^{a}\restriction_{\opensetc}} \arrow[rr, "\simeq", dotted]                                        &  & {{\textcolor{set}{Hom}(\opensetc, \measurec^{a})}}                                                         &                                                                                                   \\
-                                                                                                                               &  &                                                                                                            & {{\textcolor{set}{Hom}(\opensetc, \measurec^{ac})}} \arrow[lu, "\measurec\restriction_{a}"', color=set] \\
-  \cgamma{\opensetc}{(\vtotalc^{a}\times \vtotalc^{c})\restriction_{\opensetc}} \arrow[rr, "\vmarkc_{ac}", color=artist] \arrow[uu, "\pi_a", color=total] &  & \imartistsub{ac}{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \arrow[ru, "\extractmc_{ac}"', color=monoid] &
-  \end{tikzcd}
 \note{when $Q_{ab} = Q_{bc}$, this says only things that change are the parameter controlled pieces}
@@
 \subsection{Implementation Options: $\vartistc=\vmarkc\circ\vchannelc$}
-As mentioned in \autoref{sec:atct:sheaves}, sheaves equipped with transport functors have isomorphic maps between the sheaves.
-This property allows us to specify the transformation from data to graphic space in data or graphic space or a mix of both such that the following diagram commutes:
 
+As shown in \autoref{eq:atct:sheaves:homset}, sheaves equipped with transport functors have isomorphic maps between the sheaves.
+This property, allows us to specify the transformation from data to graphic space in data or graphic space or a combination, as listed in \autoref{eq:artist:hom_transport}, such that the following diagram commutes:
+\begin{equation}
+  \label{eq:construction:artist:path}
 \begin{tikzcd}[row sep=2.5em, column sep=1.5em]
   \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
-  \arrow[rr, "\vchannelc", color=artist] 
+  \arrow[rr, "\vchannelc^{\dbasec}", color=artist] 
   \arrow[rrrr, "\vartistc^{\dbasec}", bend left, color=artist] 
@@
    & & & & \\
-  \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} \arrow[rr, "\vchannelc", color=artist] 
+  \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} \arrow[rr, "\vchannelc^{\gbasec}", color=artist] 
   \arrow[rrrr, "\vartistc^{\gbasec}", bend right, color=artist] & & 
@@
 \end{tikzcd}  
+\end{equation}
+The diagram in \autoref{eq:construction:artist:path} illustrates how a developer can connect transformations over data space, denoted with a subset $\dbase$, with transformations over graphic space $\gbase$, using $\vindexpushc$ and $\vindexpullc$ adaptors.
+This allows developers to for example connect transformers that transform data on a line to a color in dataspace, but build a line compositing function that dynamically resamples what is on screen in graphic space.
+x
 
@@
 \appendices
-\section{Category Theory}
-\subsection*{Category}
-\subsection*{}
 

### 2022-04-27 paper/main.tex
equiv rewrite + fleshing bundles out

@@
 \usepackage[]{footmisc}
+%\usepackage[notref, notcite]{showkeys}
 \usepackage{hyperref} % autoref
@@
 % *** SPECIALIZED LIST PACKAGES ***
+
 \usepackage{xcolor}
@@
 
+\usepackage[inline]{showlabels}
+
 \usepackage{notation} %notation conventions
 % correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
+\hyphenation{}
 
@@
 \begin{equation}[h]
-  \label{eq:atct:fb_intro}
+  \label{eq:atct:fb:intro}
   \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
@@
 \begin{equation}
-  \label{eq:atct:fb_sections}
+  \label{eq:atct:fb:sections}
   \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}
@@
 A bundle may be globally trivial, meaning that $\dtotalc = \dbasec \times \dfiberc$, and generally can be thought of as a twisted product of total and base spaces \cite{munkresElementsAlgebraicTopology1984}.
+
+Fiber bundles can be used to model data for visualization, as proposed by Butler\cite{butlerVectorBundleClassesForm1992, butlerVisualizationModelBased1989}.
+We describe the topological properties of the data \cite{wilkinsonGrammarGraphics2005} as the base space \dbasec.
+The topological properties, as described by Wilkenson, are how the data is organized - for example \dbase\ could be discrete points, a line, a 2D surface, a 3D volume, or a graph.
+The names and data types of the data fields can be expressed as the fiber space \dfiberc, as described by Spivak \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES}.
+
+Fiber bundles can be used to model data for visualization, as proposed by Butler\cite{butlerVectorBundleClassesForm1992, butlerVisualizationModelBased1989}.
+Even though fibers on non-trivial spaces, such as vector fields, are different, we simplify and represent all fibers with the same schema because fibers are isomorphic.
 
@@
 
-Fiber bundles can be used to model data for visualization, as proposed by Butler\cite{butlerVectorBundleClassesForm1992, butlerVisualizationModelBased1989}.
-We describe the topological properties of the data \cite{wilkinsonGrammarGraphics2005} as the base space \dbasec.
-The topological properties, as described by Wilkenson, are how the data is organized - for example \dbase\ could be discrete points, a line, a 2D surface, a 3D volume, or a graph.
-The names and data types of the data fields can be expressed as the fiber space \dfiberc, as described by Spivak \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES}.
-For example, a temperature field could be expressed as a fiber $\dfiber_{temperature}=\mathbb{R}$, and a color field could be $\dfiber_{color}=\mathbb{R}^{3}$.
-Even though fibers on non-trivial spaces, such as vector fields, are different, we can simplify to one type because fibers are isomorphic.\note{unsure of this claim but I think this is why, and also I need to tell you why you care that they're isomorphic} The total bundle space \dtotalc\ acts as a way of expressing all data with the same topology \dbasec\ and fields \dfiberc, for example a database schema.
-A specific dataset is represented as a section \dsectionc, and the return values of a section evaluated at a point \dbasepointc\ is the record $\{field_{0}: value_{0}, ..., field_n:value_n\}$ at that point.
-
-\begin{figure}
-  \label{fig:atct:fb_graphic_bundle}
-  \includegraphics[width=1\columnwidth]{render.png}
-  \caption{}
-\end{figure}
+As illustrated in \autoref{fig:atct:fb}, \note{figure out what the actual data example should be}
 
@@
 \begin{equation}
-  \label{eq:atct:fb_graphic}
+  \label{eq:atct:fb:graphic}
   \gfiberc \hookrightarrow \gtotalc \xrightarrow{\pi} \gbasec
@@
 are functions that generate different graphics, for example markers for a scatter plot, bars for a bar chart, cells of a heatmap, or pixels of an image.
+
+\begin{figure}
+  \label{fig:atct:fb:graphic}
+  \includegraphics[width=1\columnwidth]{render.png}
+  \caption{}
+\end{figure}
+In \autoref{fig:atct:fb:graphic}, the section function $\gsection$ encodes the drawing of the \note{pin down what's beeing drawn}.   
+The renderer subsamples the position fibers  \textit{x} and \textit{y} of the graphic fiber space \gfiber.
+The preimage of the section at these coordinates returns a set of coordinates ${\gsection^{-1}_{x,y}} = \{\gbasepoint | \gsection(\gbasepoint)\restriction_{x,y} = \(x,y\)\}$ at the points $\gbasepoint \in \gbase$ in the graphic base topological space.
+The section is then evaluated at these points $\gsection\restriction_{\gsection^{-1}_{x,y}}$ to return a set of points in the fiber.
+Each point has an RGB tuple and these tuples are integrated to return the color of the pixel at that $(x,y)$ point.
 
@@
 \begin{equation}
+ \label{eq:atct:morphism:base}
 \dfunch: \openset^{\prime}\rightarrow \openset
@@
 \begin{equation}
-  \dfuncpull \dsection \restriction_{\openset}: \dsection\restriction_{\openset} \mapsto \dsection \restriction_{\openset} \circ \dfunch 
+  \label{eq:atct:morphism:basepull}
+  \dfuncpull \dsection \restriction_{\openset^{\prime}}: \dsection\restriction_{\openset^{\prime}} \mapsto \dsection \restriction_{\openset^{\prime} \circ \dfunch} 
 \end{equation}
@@
 \begin{equation}
-  \dfunct: \dfuncpull \dsection \restriction_{\openset} \mapsto \dfuncpull \dsection^{\prime} \restriction_{\openset}
+  \label{eq:atct:morphism:fiber}
+  \dfunct: \dfuncpull \dsection \restriction_{\openset^{\prime}} \mapsto \dfuncpull \dsection^{\prime} \restriction_{\openset}
 \end{equation}
@@
 \begin{equation}
+  \label{eq:atct:morphism:all}
   \dfuncc: \dsectionc\restriction_{\opensetc} \mapsto \dsectionc^{\prime}\restriction_{\opensetc} \circ \dfunchc
@@
 For example, specifications such as svg\cite{quintScalable2003} and Vega\cite{satyanarayanDeclarativeInteractionDesign2014}, typically define their transformations in data space and return specifications for the render to evaluate.
-It may also be
 An artist may also want to specify the transformations in a proxy of graphic space, for example for operations like dynamic resampling \note{citation:root maybe}.
@@
 
-\subsection*{Equivariance}
+\subsection{Operators}
+\subsubsection{More Base Spaces: +}
+\subsubsection{More Fibers: \times}
+
+\subsection{Equivariance}
+The notion of structure preservation discussed in \autoref{sec:related-work:continuity} is that data and components of the graphic vary in equivalent ways.
+We describe the changes on the graphic side as changes in measurments \measure\, which are scaler or vector components of the rendered graphic that can be quantified, such as the color, position, shape, texture, or rotation angle of the graphic.
+The visual variables \cite{bertinIIPropertiesGraphic2011} are a subset of measurable components.
+For example, a measurement of a scatter marker could be its color (e.g.
+red) or its x position (e.g.
+5).
+
+We formalize this structure preservation as equivariance, which is that for every morphism on the data $(\dfunch_{\dtotal}, \dfunct_{\dtotal})$ there is an equivalent morphism on the graphic  $(\dfunch_{\gtotal}, \dfunct_{\gtotal})$ The artist is an equivariant map if the diagram commutes for all points $\gbasepointc^{\prime}\in \gbasec^{\prime}$
 \label{sec:artist:equivariance}
@@
 \end{equation*}
-For all points $\gbasepointc^{\prime}\in \gbasec^{\prime}$:
+such that starting at an arbitrary data point $\dsectionc(\dbasepointc)$ and transforming it into a different data point and then into a graphic 
 \begin{equation*}
@@
 \end{equation*}
+is equivalent to transforming the original data point into a graphic and then transforming the graphic into another graphic.
+The function $\dfunch_{\gtotal}$ induces a change in graphic generating function that matches the change in data.
 
-\subsection*{Testing Equivariance}
+It is difficult to formally define or implement $\dfunch_{\gtotal}$ because the equivalent changes in the graphic, such as a scaling or rotation, would have to be evaluated solely on the basis of the fiber elements $\gsection(\dbasepoint)$ which are tuples of output space specifications.
+For example, in a 2D screen, evaluating that a scatter marker respects a change in scaling would mean translating a table of $\{x,y,r,g,b\}$ values into measurements \measure\ and then checking how those measurements change.
+This is equivalent to measuring components of the rendered image, as shown in the diagram
+
 \begin{equation}
+  \label{eq:artist:inout:diagram}
   \begin{tikzcd}[row sep=huge]
@@
 \end{equation}
-  \begin{description}
-      \item[expected]{\;$\equivc: \dsectionc \mapsto (\opensetc \xrightarrow{\equivc_{\dsectionc}} \measurec)$}
-      \item[actual]{$\extractmc:\gsectionc \mapsto (\opensetc \xrightarrow{\extractmc_{\gsectionc}} \measurec)$}
-    \end{description}
-    such that
+so instead we introduce a measurement extraction function $\extractm = measure\circ render$ that decomposes the rendered output into measurable components and returns a map from the data index into the measurement space
 \begin{equation}
-    \equivc(\dsectionc)(\dbasepointc) = \extractmc(\vartistc(\dsectionc))(\dbasepointc) = \extractmc(\gsectionc\circ\vindexprec)(\dbasepointc)
+  \label{eq:artist:actual}
+  \extractmc:\gsectionc \mapsto (\opensetc \xrightarrow{\extractmc_{\gsectionc}} \measurec)
+\end{equation}
+such that $\extractmc_{\gsectionc}(\dbasepoint)$ returns the measurements $\measurec_{\dbasepointc}$ of the visual element generated for the data at that point $\gsectionc|_{\vindexprec(\dbasepointc)}$.
+
+We also introduce a function \equivc\ that maps data to the measurement space directly 
+\begin{equation}
+\equivc: \dsectionc \mapsto (\opensetc \xrightarrow{\equivc_{\dsectionc}} \measurec)
+\end{equation}
+such that $\equivc_{\dsection}(\dbasepointc)$ is the expected set of measurements $\measure_{\dbasepoint}$.
+The pair of \textcolor{monoid}{testing functions} (\equivc, \extractmc) can be used to test that the expected encoding of the data matches the actual encoding
+\begin{equation}
+    \equivc(\dsectionc)(\dbasepointc) = \extractmc(\vartistc(\dsectionc))(\dbasepointc) = \extractmc(\gsectionc\circ\vindexprec)(\dbasepointc)=\measurec_{\dbasepointc}
 \end{equation}
 
-  \note{introduce $\phi_M$ as basis of equivariance}
+An artist is equivariant when changes to expected and actual are equivariant.
+As introduced in \autoref{eq:atct:morphism:base}, the base space transformation \dfunch\ is invariant because $\dsection\restriction_{\openset} = \dsection\restriction_{\dfunch(\openset^{prime})}$.
+This means that, for all points in the data $\dbasepoint \in \dbase$, the measurement should not change if only the base space is transformed
+\begin{equation}
+  \equivc(\vartistc(\dsectionc))(\dfunch(\dbasepointc^{\prime})) = \extractmc(\vartistc(\dsectionc))(\dbasepointc)
+\end{equation}
+On the other hand, a change in sections \autoref{eq:atct:morphism:fiber} induces an equivalent change in measurements
+\begin{equation}
+  \equivc(\dfunctc(\dsectionc))(\dbasepoint) = \dfunctc_{\measure}(\extractmc(\vartistc(\dsectionc))(\dbasepoint))
+\end{equation}
+The change in measurements $\dfunctc_{\measure}$ is defined by the developer as the symmetry between data and graphic that the artist is expected to preserve.
+For example a data to color transformation could be defined with the constraint that it must respect permutation since the data is categorical or with the constraint that it respect scaling if the data is quantative.
 
@@
 \end{equation}
-we can construct a $\vmarkc = \sqcup_i\vmarkc_i$.v This allows us to .
+we can construct a $\vmarkc = \sqcup_i\vmarkc_i$. This allows us to .
 

### 2022-05-09 paper/main.tex
figure code + related work + color sync

@@
 
+\subsection{Topology}
+The topological properties of visualization data are ..., as described by Wilkenson\cite{wilkinsonGrammarGraphics2005}.
+
+\begin{figure}[h!]
+  \includegraphics[width=1\columnwidth]{continuity.png}
+\end{figure}
+Vsisualization libraries designed around continuity,  
+Visualization library algorithms assume
+
+\begin{figure}[h!]
+  \includegraphics[width=1\columnwidth]{k_different_types.png}
+\end{figure}
+Fiber bundles are a mathematical structure that can be use to encode the topological properties of visualization data, as proposed by Butler\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}.
+Broadly, fiber bundles
+
+\subsection{Equivariance}
+\begin{figure}[h!]
+  \includegraphics[width=1\columnwidth]{equiviz.png}
+\end{figure}
+
 \section{Data And Visualizations}
@@
 \subsection{Sheaves}
+\label{sec:atct:sheaves}
 Sheaves on bundles are an "algebraic data structure", as described by Ghrist \cite{ghristElementaryAppliedTopology2014}, for expressing the bookkeeping involved in keeping track of sections of data over opensets.
@@
 \subsection{Maps Between Spaces}
-\note{add couple of sentences introducing image functors}
-We propose that there must exist a continuous \textcolor{functor}{map between topological spaces} \vindexc 
+The data sections defined in \autoref{eq:atct:fb} are sections of sheaves, introduced in \autoref{sec:atct:sheaves}, over open sets $\openset \subseteq \dbase$ in a base space \dbase\ that encodes the topology of the data topology.
+The graphic sections, defined in \autoref{eq:atct:fb_graphic_section}, are sections of sheaves over open sets $\opensetg \subseteq \gbase$ in a base space \gbase\ that encodes the topology of the graphic.
+We propose that there exists a continuous \textcolor{functor}{map between topological spaces} \vindexc
 \begin{equation}
@@
 such that every $\opensetg \subseteq \gbase$ must map to a corresponding set $\openset \subset \dbase$.
-Given \vindex\ we can construct two functors \vindexpull\ and \vindexpush\ which transport sheaves
 
+Given the map between spaces \vindex\, a property of sheaves is that there exist image functors\cite{ImageFunctorsSheaves2021} functors \vindexpull\ and \vindexpush.
+The \textcolor{functor}{pullback functor} $\vindexpullc$ sends sheaves over $\dbase$ to $\gbase$ and the \textcolor{functor}{pushforward functor} $\vindexpushc$  sends sheaves over $\gbase$ to $\dbase$
 \begin{equation}
@@
 such that there is an association between graphic sections \gsection\ that take as input \gbasepoint\ and data sections that take input \dbasepointc\ when $\vindex(\gbasepoint) = \dbasepoint$.
-The \textcolor{functor}{pullback functor} $\vindexpullc$ transports sheaves of sections on $\openset \subseteq \dbase$ over $\opensetg \subseteq \gbase$
+
+The pull back $\vindexpullc$ transports sheaves of sections on $\openset \subseteq \dbase$ over $\opensetg \subseteq \gbase$
 \begin{equation}
@@
 
- The \textcolor{functor}{pushforward functor} $\vindexpushc$ transports sheaves of sections on $\opensetgc$ over $\openset$
+ The pushforward $\vindexpushc$ transports sheaves of sections on $\opensetgc$ over $\openset$
  \begin{equation}  
@@
   \begin{tikzcd}
-    \opensetc_1 & \cgamma{\opensetc_{1}}{\dtotalc\restruction_{\opensetc_1}} \arrow[dd, "\iota^*"', color=set] \arrow[rr, "\vartistc", color=artist] &  & \cgamma{\opensetc_{1}}{\vindexpushc\gtotalc\restruction_{\opensetc_1}} \arrow[dd, "\iota^*", color=set] \\
+    \opensetc_1 & \cgamma{\opensetc_{1}}{\dtotalc\restruction_{\opensetc_1}} \arrow[dd, "\iota^*"', color=set] \arrow[rr, "\vartistc_{\opensetc_1}", color=artist] &  & \cgamma{\opensetc_{1}}{\vindexpushc\gtotalc\restruction_{\opensetc_1}} \arrow[dd, "\iota^*", color=set] \\
       &  &  &  \\
-    \opensetc_{2} \arrow[uu, "\iota", hook, color=base] & \cgamma{\opensetc_{2}}{\dtotalc\restruction_{\opensetc_2}} \arrow[rr, "\vartistc", color=artist]                        &  & \cgamma{\opensetc_{2}}{\vindexpushc\gtotalc\restruction_{\opensetc_2}}
+    \opensetc_{2} \arrow[uu, "\iota", hook, color=base] & \cgamma{\opensetc_{2}}{\dtotalc\restruction_{\opensetc_2}} \arrow[rr, "\vartistc_{\opensetc_2}", color=artist]                        &  & \cgamma{\opensetc_{2}}{\vindexpushc\gtotalc\restruction_{\opensetc_2}}
     \end{tikzcd}
@@
 \begin{equation}
-  \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \coloneqq \{\gsectionc\vert\;\exists\;\dsectionc \in \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}\;s.t.\; 
+  \begin{split}
+  \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} &\coloneqq\\ 
+  &\{\gsectionc\vert\;\exists\;\dsectionc \in \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}\;s.t.\; 
   \vartistc(\dsectionc) = \gsectionc,\; \vindexc(\opensetgc) = \opensetc \} 
-\end{equation} that are by definition only reachable through a structure preserving artist, which we describe in \autoref{sec:artist:equivariance}.
+  \end{split}
+\end{equation} 
+that are, by definition, only reachable through a structure preserving artist, which we describe in \autoref{sec:artist:equivariance}.
 We define this subset because the space of all sections $\cgamma{\opensetg}{\gtotal\restriction_{\openset}}$ includes sections that may not be structure preserving.
@@
 
-\subsection{Operators}
-\subsubsection{More Base Spaces: +}
-\subsubsection{More Fibers: \times}
+\subsection{Artist Operators}
+\subsubsection{Addition: +}
+\label{sec:artist:addition}
+ $\vartist \coloneqq \vartist^{a}(\dsectionc^{a}) + \vartist^{b}(\dsectionc^{b})$ &\\
+
+ with input 
+ \begin{equation}
+    \begin{tikzcd}[column sep=small]
+      \dtotalc^{a} \arrow[dd, "\pi"', color=total]  &  & \dtotalc^{b} \arrow[dd, "\pi"', color=total]                              \\
+      {}    &  & {}                                                           \\
+      \dbasec^{a} \arrow[uu, "\dsectionc^{a}"', dashed, bend right, color=section] &  & \dbasec^{b} \arrow[uu, "\dsectionc^{b}"', dashed, bend right, color=section]
+      \end{tikzcd}
+  \end{equation}
+such that output 
+\begin{equation}
+  \label{eq:artist:plus:output}
+  \gsectionc(\gbasepointc) = \begin{cases} \gsectionc^{a}(\dbasepointc) & \gbasepointc \in \vindexprec(\dbasec^{a}) \\
+    \gsectionc^{b}(\dbasepointc) & \gbasepointc \in \vindexprec(\dbasec^{b})
+  \end{cases}
+\end{equation}
+
+\subsubsection{Multiplication: $\times$}
+$\vartist \coloneqq \vartist^{a}(\dsectionc^{a}) \times \vartist^{b}(\dsectionc^{b})$
+with input 
+\begin{equation}
+  \begin{tikzcd}
+    \dtotalc^{a} \arrow[rdd, "\pi"', color=total] & & \dtotalc^{b} \arrow[ldd, "\pi"', color=total] \\
+       & & \\
+      & \dbasec \arrow[luu, "\dsectionc^{a}"', dashed, bend right, color=section] 
+      \arrow[ruu, "\dsectionc^{b}"', dashed, bend right, color=section] &                                 
+    \end{tikzcd}
+  \end{equation}
+  \\
+ The output 
+\begin{equation}
+  \{\gsectionc^{a}(\gbasepointc), \gsectionc^{b}(\gbasepointc)\}, \gbasepointc \in \vindexprec(\dbasec)
+\end{equation} 
+which would mean s
+
+which would require returning two points in the fiber $\gbase \rightarrow \dfiber \times \dfiber$. This function is adjoint to the addition function in \autoref{sec:artist:addition} 
+\begin{equation}
+  hom(\dbase, \dfiber\times \dfiber) \simeq hom(\dbase \plus \dbase, \dfiber)
+\end{equation}
+such that the $\times$ operator can be implemented as a special case of \autoref{eq:artist:plus:output} where $\gbase^a$ is a copy of $\gbase$ associated with $\gsection^a$ and $\gbase^b$ is a copy of $\gbase$ associated with $\gsection^{b}$.
 
@@
 \label{sec:artist:equivariance}
-\begin{equation*}
+\begin{equation}
+  \autoref{eq:artist:equivariance}
   \begin{tikzcd}[ampersand replacement=\&, column sep=small]
@@
   \end{tikzcd}
-\end{equation*}
+\end{equation}
 such that starting at an arbitrary data point $\dsectionc(\dbasepointc)$ and transforming it into a different data point and then into a graphic 
@@
     \arrow[d, "\equivc"', color=monoid] &  & 
-    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+    \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
     \arrow[d, "render"] 
@@
 \end{equation}
-such that $\extractmc_{\gsectionc}(\dbasepoint)$ returns the measurements $\measurec_{\dbasepointc}$ of the visual element generated for the data at that point $\gsectionc|_{\vindexprec(\dbasepointc)}$.
+such that $\extractmc_{\gsectionc}(\dbasepoint)$ returns a quantification of the measurable components $\measurec_{\dbasepointc}$ of the visual element generated for the data at that point $\gsectionc|_{\vindexprec(\dbasepointc)}$.
 
@@
 such that $\equivc_{\dsection}(\dbasepointc)$ is the expected set of measurements $\measure_{\dbasepoint}$.
-The pair of \textcolor{monoid}{testing functions} (\equivc, \extractmc) can be used to test that the expected encoding of the data matches the actual encoding
+The pair of \textcolor{monoid}{verification functions} (\equivc, \extractmc) can be used to test that the expected encoding $\equivb_{\dsection}$ of the data matches the actual encoding $\extractm_{\gsection}$
 \begin{equation}
+  \label{eq:artist:verification}
     \equivc(\dsectionc)(\dbasepointc) = \extractmc(\vartistc(\dsectionc))(\dbasepointc) = \extractmc(\gsectionc\circ\vindexprec)(\dbasepointc)=\measurec_{\dbasepointc}
@@
 An artist is equivariant when changes to expected and actual are equivariant.
-As introduced in \autoref{eq:atct:morphism:base}, the base space transformation \dfunch\ is invariant because $\dsection\restriction_{\openset} = \dsection\restriction_{\dfunch(\openset^{prime})}$.
+As introduced in \autoref{eq:atct:morphism:base}, the base space transformation \dfunch\ is invariant because $\dsection\restriction_{\openset} = \dsection\restriction_{\dfunch(\openset^{\prime})}$.
 This means that, for all points in the data $\dbasepoint \in \dbase$, the measurement should not change if only the base space is transformed
@@
 
-\subsection{}
+\subsection{Transforming Data into Measurable Components}
 The transformation $\vchannel$ is a map from data sections to graphic sections
@@
 \end{equation}
-such that the sections project to the same point on the base space $\pi(\dtotal) = \pi(\vchannel(\dtotal))$. A pair of transformations $(\vchannel, \vchannel^{\prime})$ are equivariant if 
-\begin{equation}
-  \dfunct_{\vtotal} \vchannel(\dsection(\dbasepoint)) = \vchannel^{\prime}(\dfunct_{\dtotal}(\dsection(\dbasepoint)))\; \forall \dbasepoint \in \dbase
-\end{equation}
-and a transformation $\vchannel$ is equivariant when $\vchannel = \vchannel^{\prime}$
+such that the sections project to the same point on the base space $\pi(\dtotal) = \pi(\vchannel(\dtotal))$.
+As with artists, $\vchannel$ are maps of sections such that the composition operators defined in \autoref{sec:artist:composition} can also act on transformers $\vchannel$.
 
-\subsection*{Fabricating \vchannelc's}
 A consequence of $\pi(\dtotal) = \pi(\vchannel(\dtotal))$ is that $\vchannelc$ can be constructed as a pointwise transformation such that  
@@
 \end{equation}
-
-This property means that a point in a single data fiber $\delement \in \dfiber_{\dbasepointc}$ can be mapped into a corresponding point in a visual fiber $\velement \in \vfiber_{\dbasepointc}$.
-We use this property to construct compositions of $\vchannelc$ functions
+which means that means that a point in a single data fiber $\delement \in \dfiber_{\dbasepointc}$ can be mapped into a corresponding point in a visual fiber $\velement \in \vfiber_{\dbasepointc}$.
+Since $\vchannel$ can be constructed as a pointwise function property, we can construct $\vchannelc$ functions
 \begin{equation}
@@
 
-\subsubsection{Combining \vchannelc's} 
-We can add visual parameters as
+\subsubsection{Transformer Equivariance}
 
-\subsection*{Validating \vchannelc's}
+A pair of transformations $(\vchannel, \vchannel^{\prime})$ are equivariant if 
 \begin{equation}
+  \dfunct_{\vtotal} \vchannel(\dsection(\dbasepoint)) = \vchannel(\dfunct_{\dtotal}(\dsection(\dbasepoint)))
+\end{equation}
+for all $\dbasepoint \in \dbase$.  Since the visual sections and data sections are over the same space $\dfunc$
+
+\begin{equation}
+  \label{def:construction:nu:validate}
   \begin{tikzcd}[column sep=4em]
@@
 
-\subsection{Graphic Compositing Function}
+\subsection{Compositing Measurable Components into Output}
 The compositing function $\vmarkc$ transforms the sections of visual elements $\vsectionc$ into sections of graphics $\gsectionc$.
@@
 \end{equation}
-
-\note{some transition about sheaf morphisms}
-Therefore an equivariant $\vmarkc$ satisfies the conditions....
-
-\subsubsection{Fabricating \vmarkc's}
 
@@
 
-\subsubsection{Combining \vmarkc's}
-A set of $\vmarkc$ functions that output graphic sections, as introduced in \autoref{eq:atct:fb_graphic_section}, that go to the same space, for example to the same figure in a screen, are returning a set of functions into the same total bundle $\gtotal$;
+\subsubsection{Compositor Equivariance}
+
+As mentioned in \autoref{sec:construction:nu}, the distinction between data sections $\dsection$ and visual sections $\vsection$ is semantic rather than structural.
+Therefore, the equivariance condition on the sheaf map with a change of base space $\vmarkc$ is identical to the one for artists shown in \autoref{eq:artist:equivariance} such that an equivariant \vmark\ must satisify the condition
 \begin{equation}
-  \label{eq:construction:qc1}
-  \bigsqcup_i \vmark_i(\Gamma(\openset_i,\dtotal_i\restriction_{\openset_i})) = \bigsqcup_i \Gamma(\opensetg_i, \gtotal\restriction_{\opensetg_i})
+  \vmarkc(\dfunctc_{\vtotalc}(\vsectionc(\dfunchc_{\vtotalc}(\vindexc(\gbasepointc^{\prime}))))) = \dfunctc_{\gtotalc}(\vmarkc(\vsectionc(\vindexc(\dfunchc_{\gtotalc}(\gbasepointc^{\prime})))))  
 \end{equation}
-A property of hom-sets is that $Hom(W_i,H) + Hom(W_j,H) = Hom(W_i+W_j,H)$, which means that the union of hom-sets with the same target is equal to the hom-set of the union of sources to the target; therefore
+which means that any change to a measurable component input must have a measurably equivalent change in the output.
+
+As with the artist, it is difficult to work with the output section \gsection\ directly and instead we test the compositer \vmark\ using measurements derived from the output.
+A compositer is expected to pass visual transformers \vsection\ through to their expected visual measurements, as described in \autoref{def:construction:nu:validate}, such that any shared visual components at a point $\vsection^{a}(\dbasepoint)$ must render to the same measurement $\measure^{a}_{\dbasepoint}$
 \begin{equation}
-  \label{eq:construction:qc2}
-  \bigsqcup_i\Gamma(\opensetg_i, \gtotal\restriction_{\opensetg_i}) = \Gamma(\mathop{\sqcup}_{i}\opensetg_i, \gtotal\restriction_{{\mathop{\sqcup}_i}\opensetg_i})
-\end{equation}
-the union of sets of functions that generate parts of an image is equal to a set of functions that can generate the whole image.
-This implies that we can define a $\vmarkc$ that generates the set of sections over the whole image
-\begin{equation}
-  \label{eq:construction:qc3}
-\Gamma(\mathop{\sqcup}_i\opensetg_i, \gtotal\restriction_{\mathop{\sqcup}_{i}\opensetg_i}) \implies \vmark(\mathop{\sqcup}_{i}\Gamma(\openset_i,\vtotal_i\restriction_{\openset_i}) 
-\end{equation}
-which means that, following from \autoref{eq:constrution:qc1} and \autoref{eq:construction:qc3},
-\begin{equation}
-  \label{eq:construction:qc4}
-  \vmarkc(\mathop{\sqcup}_{i}\cgamma{\opensetc_i}{\vtotalc_i\restriction_{\opensetc_i}}) =   
-  \bigsqcup_i \vmarkc_i(\cgamma{\opensetc_i}{\vtotalc_i\restriction_{\opensetc_i}})
-\end{equation}
-we can construct a $\vmarkc = \sqcup_i\vmarkc_i$. This allows us to .
-
-\begin{figure}
-  \label{fig:construction:combined_q}
-  \includegraphics[width=1\colunwidth]{qsketchh.png}
-  \caption{}
-\end{figure}
-
-As illustrated in \autoref{fig:construction:combined_q}, this composition rule expresses the construction of a multipart circle-square graphic out of circles and squares produced by different $\vmarkc$.
-\note{rewrite:Also figures are the unions of their parts}.
-
-\subsubsection{Validating \vmarkc's}
-\begin{equation}
+  \label{eq:construction:q:validate}
   \begin{tikzcd}[row sep=huge]
@@
 
-\subsection{Implementation Options: $\vartistc=\vmarkc\circ\vchannelc$}
+\subsection{Constructing the Artist: $\vartistc=\vmarkc\circ\vchannelc$}
 

### 2022-05-09 paper/figcode/papercolors.py

@@ -1,0 +1,13 @@
+import re
+
+sub = re.compile(r"\\definecolor\{(?P<name>.*)\}\{HTML\}\{(?P<color>.*)\}")
+
+def build_colors(filepath = '../notation.sty'):
+    colordict = {}
+    with open(filepath) as f:
+        for line in f.readlines():
+            if groups := sub.match(line):
+                colordict[groups.group('name')] = "#"+ groups.group('color')
+    return colordict
+
+colordict = build_colors()

### 2022-05-11 paper/main.tex
related work, measurment scale, bundles

@@
 \begin{figure}[h!]
+  \label{fig:related-work:equivariance}
   \includegraphics[width=1\columnwidth]{equiviz.png}
@@
 
-\begin{figure}[h]
+\begin{figure}[h!]
   \label{fig:atct:fb}
-  \includegraphics[width=\columnwidth]{fiberbundle.png}
+  \includegraphics[width=\columnwidth]{fb_tau.png}
   \caption{change to concrete example}
@@
 
-\begin{figure}
+\begin{figure}[h!]
   \label{fig:atct:fb:graphic}
-  \includegraphics[width=1\columnwidth]{render.png}
+  \includegraphics[width=1\columnwidth]{fb_rho.png}
   \caption{}
@@
 such that $\dsection|_{\openset} = \dfuncpull\dsection|_{\opensetc^{\prime}}$ because $\dsection|_{\openset} = \dsection|_{\dfunch(\opensetc^{\prime})}$.
-This means that the base space transformation \dfunch\ alone does not change the data values.
-
+This means that the base space transformation \dfunch\ does not change the data values at a given point
+\begin{equation}
+  \label{eq:atct:morphism:verify_base}
+  \dsection(\dbase) = \dfuncpull\dsectionc(\dbasepoint^{\prime}) = \dsection(\dfunch(\dbase^{prime}))
+\end{equation} 
+where $\dfunch(\dbasepoint^{\prime}) = \dfunch(\dbasepoint)$.
 As introduced in \autoref{def:atct:io:fiber}, the fiber transformation $\dfunct: \dfuncpull \dtotal_{\dbasepoint^{\prime}} \rightarrow \dfuncpull \dtotal_{\dbasepoint^{\prime}}$ is a morphism on the fiber $\dfunct \in Hom(\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}},\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}})$ restricted to a point $\dbasepoint^{\prime} \in \openset^{\prime}$.
@@
 where $\dsection, \dsection^{\prime} \in \Gamma(\openset^{\prime}, \dfuncpull\dtotal\restriction_{\openset^{\prime}})$.
-Examples of a fiber transformation are actions on fibers, such as the permutation, ordering, translation, and scaling codified as Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946}.
+Examples of a fiber transformation are actions such as partial ordering for ranking purpose \cite{bruggemannRankingPrioritizationMultiindicator2011} and the permutation, ordering, translation, and scaling codified as Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946}.
 
-The full transformation 
+We define a full data transformation as one that induces both a remapping of the index space and a change in the data values
 \begin{equation}
@@
 \end{equation}
-is used to express actions such as rotation that induce both a remapping of the index space and a change in the data values. The data transform \dfunc\ is composable
+which gives us an equation that can express complex transformations.
+For example, wind vectors on a sphere are a non-trivial bundle, meaning their fibers aren't identical at every point in the base space; therefore a rotation (change of base space) of the sphere must include an adjustment to the values that have moved closer to the pole (a change in fiber).
+The data transform \dfunc\ is composable 
 \begin{equation}
@@
 \end{equation}
-if there exists functions $\dfunc_{a,b}: \dtotal_a \times \dtotal_b \rightarrow \dtotal_a \times \dtotal_b$, $\dfunc_{a}: \dtotal_a \rightarrow \dtotal_a$ and $\dfunc_{b}: \dtotal_b \rightarrow \dtotal_b$ such that $\pi_a \circ \dfunc_a = \dfunc_{a,b} \circ \pi_a$ and $\pi_b \circ \dfunc_b = \dfunc_{a,b} \circ \pi_b$ then $\dfunc_{a,b} = (\dfunc_a, \dfunc_b)$.
+if each (identical) component base space is transformed in the same way $\dfunch$ and there exists functions $\dfunc_{a,b}: \dtotal_a \times \dtotal_b \rightarrow \dtotal_a \times \dtotal_b$, $\dfunc_{a}: \dtotal_a \rightarrow \dtotal_a$ and $\dfunc_{b}: \dtotal_b \rightarrow \dtotal_b$ such that $\pi_a \circ \dfunc_a = \dfunc_{a,b} \circ \pi_a$ and $\pi_b \circ \dfunc_b = \dfunc_{a,b} \circ \pi_b$ then $\dfunc_{a,b} = (\dfunc_a, \dfunc_b)$.
 This allows us to define a data transform where each fiber transform $\dfunct_{i}$ can be applied to a different fiber field $\dfiber_i$.
+
+\begin{figure}[h!]
+  \label{fig:atct:phi}
+  \includegraphics[width=\columnwidth]{phi.png}
+\end{figure}
+Transposition, as illustrated in \autoref{fig:atct:phi}, is one example of a transformation on the base space $(\dfunch, \dfuncpull)$; transposition is a continuity preserving data transformation that moves every data value to a new location such that the values are unchanged and connected in the same way.
+Scaling, as shown in \autoref{fig:atct:phi} is a transform $\dfunct$ that changes each data value, in this example by multiplying each value by two.
+As defined in \autoref{eq:atct:morphism:all}, these transformations can combine to yield a matrix that is both transposed and scaled.
+The matrix in \autoref{fig:atct:phi} is visualized as a heatmap \autoref{fig:related-work:equivariance} and, as discussed in \autoref{sec:related-work:equivariance}, the heatmap is illustrating equivariance.
+In the $(\dfunch, \dfuncpull)$ case, the data to color map has not changed, but the column bands of color have become row bands.
+In the $\dfunct$ case, the data has been scaled by two and the color encoding has been scaled by two, but the bands remain column ordered.
+When the data is transposed and scaled, then the heatmap is also transposed and the colors are scaled.
 
@@
 
-\begin{figure}[ht]
+\begin{figure}[h!]
   \label{fig:atct:morphisms:sheaf}
-  \includegraphics*[width=1\columnwidth]{../../slides/figures/math/push_pull_scatter.png}
+  \includegraphics*[width=1\columnwidth]{xi_scatter.png}
   \caption{}

### 2022-05-13 paper/main.tex
composition overhaul

@@
 \subsection{Topology}
-The topological properties of visualization data are ..., as described by Wilkenson\cite{wilkinsonGrammarGraphics2005}.
+The topological properties of visualization data are , as described by Wilkenson\cite{wilkinsonGrammarGraphics2005}.
 
@@
 \end{figure}
-Vsisualization libraries designed around continuity,  
+Visualization libraries designed around continuity,  
 Visualization library algorithms assume
@@
 \subsection{Equivariance}
-\begin{figure}[h!]
-  \label{fig:related-work:equivariance}
-  \includegraphics[width=1\columnwidth]{equiviz.png}
-\end{figure}
 
@@
 A sheaf is a presheaf that satisfies the locality and gluing axioms \cite{bakerMathsSheaf}. 
-The locality axiom is that given a union of opensets $\mathscr{\opensetc} = \bigcup\limits_{i\in I} \opensetc_i$ and $\dsectionc^{a}, \dsectionc^{b} \in \sheafc(\mathscr{\opensetc})$,  if $\dsectionc^{a}\restriction_{\opensetc_i} = \dsectionc^{b}\restriction_{\opensetc_i}$ for each $\opensetc_i \in \opensetc$ then $\dsectionc^{a} = \dsectionc^{b}$.
+\begin{definition}
+\label{def:atct:sheaf:locality}
+The locality axiom is that given a union of opensets $\mathscr{\opensetc} = \bigcup\limits_{i\in I} \opensetc_i$ and $\dsectionc^{a}, \dsectionc^{b} \in \sheafc(\mathscr{\opensetc})$,  if $\dsectionc^{a}\restriction_{\opensetc_i} = \dsectionc^{b}\restriction_{\opensetc_i}$ for each $\opensetc_i \in \opensetc$ then $\dsectionc^{a} = \dsectionc^{b}$
+\end{definition}
 This means that two sections in a sheaf are equal when they evaluate to the same values over all open sets in a collection. 
+\begin{definition}
+\label{def:atct:sheaf:gluing}
 The gluing axiom is that given $\dsectionc^{i} \in \sheafc(\opensetc_i)$ such that ${\dsectionc^{i}}\restriction_{\opensetc_i\cap \opensetc_j} = {\dsectionc^{j}}\restriction_{\opensetc_i \cap \opensetc_j}$ for $\opensetc_i, \opensetc_j \in \mathscr{\opensetc}$, there exists $\dsectionc \in \sheafc(\mathscr{\opensetc})$ such that $\dsectionc\restriction_{\opensetc_i} = {\dsectionc^{i}}$.
+\end{definition}
 This means we can construct a section $\dsection$ on the collection of open sets that is the union of sections defined on specific open sets.
@@
 
-\subsubsection{Measurement Scales}
+\subsubsection{Structure on Data}
 \label{sec:atct:sheaves:measurments}
-We specify that structure on a sheaf includes the actions $\dfuncc$ on sections of that sheaf, which is a generalization of Steven's \cite{stevensTheoryScalesMeasurement1946} classification of measurement scales by their mathematical group structure.
-These sheaf actions describe which changes in the data we expect to be equivariant to changes in the visualization.
+We define structure on data as the allowable \textcolor{action}{transformations on the bundle}, which we formalize as actions $\dfuncc$ on sections of that sheaf.
+Our construction of $\dfunc$ is a generalization of the classification of measurement scales by their mathematical structure \cite{stevensTheoryScalesMeasurement1946,leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}, which often serves as the basis of the equivariance described in \autoref{sec:related-work:equivariance}.
 
+We separate data transformations into two types, transformations on the base space $(\dfunchc, \dfuncpullc)$ and transformations on the fiber space $\dfunctc$. 
 \begin{equation}
@@
 \end{equation}
-
-In this work, we assume that the map
+As shown in \autoref{eq:atct:sheaves:monoid_morphism}, the base space transform 
 \begin{equation}
@@
 \end{figure}
-Transposition, as illustrated in \autoref{fig:atct:phi}, is one example of a transformation on the base space $(\dfunch, \dfuncpull)$; transposition is a continuity preserving data transformation that moves every data value to a new location such that the values are unchanged and connected in the same way.
-Scaling, as shown in \autoref{fig:atct:phi} is a transform $\dfunct$ that changes each data value, in this example by multiplying each value by two.
-As defined in \autoref{eq:atct:morphism:all}, these transformations can combine to yield a matrix that is both transposed and scaled.
-The matrix in \autoref{fig:atct:phi} is visualized as a heatmap \autoref{fig:related-work:equivariance} and, as discussed in \autoref{sec:related-work:equivariance}, the heatmap is illustrating equivariance.
-In the $(\dfunch, \dfuncpull)$ case, the data to color map has not changed, but the column bands of color have become row bands.
-In the $\dfunct$ case, the data has been scaled by two and the color encoding has been scaled by two, but the bands remain column ordered.
-When the data is transposed and scaled, then the heatmap is also transposed and the colors are scaled.
+\autoref{fig:atct:phi} provides an example of a base space change \dfunch, a fiber space change \dfunct, and a composition of the two \dfunc\ applied to each data point $x_{\dbasepoint} \in \texttt{data}$.
+As defined in \autoref{eq:atct:morphism:verify_base},
+since transposition is a $(\dfunch, \dfuncpull)$, every data value $x_{\dbasepoint}$ must be equal to the value at the corresponding location in the transposed matrix $x_{\dbasepoint} = x_{\dbasepoint^{\prime}}, \dfunch(\dbasepoint^{\prime}) = \dbasepoint$ for all $x_{\dbasepoint^{\prime}} \in texttt{data.T}$ .
+In \autoref{fig:atct:phi}, this is demonstrated in the columns becoming rows such that there are no new values and each value retains the same neighboring values.
+As defined in \autoref{eq:atct:morphism:fiber} scaling is a transform $\dfunct$ that changes the value of each point but not its location.
+In \autoref{fig:atct:phi}, this means that $x_{\dbasepoint^{\prime}}= 2*x_{\dbasepoint}, \dbasepoint = \dbasepoint_{\prime}$ for all indices in \textt{data}, \textt{data.T}.
+As defined in \autoref{eq:atct:morphism:all} and illustrated in \autoref{fig:atct:phi}, these transformations can combine to yield a matrix that is both transposed and scaled such that $x_{\dbasepoint^{\prime}} = 2*x_{\dbasepoint}$ for all
+$\dfunch(\dbasepoint^{\prime}) = \dbasepoint$
 
@@
 
-\subsection{Artist Operators}
-\subsubsection{Addition: +}
-\label{sec:artist:addition}
- $\vartist \coloneqq \vartist^{a}(\dsectionc^{a}) + \vartist^{b}(\dsectionc^{b})$ &\\
-
- with input 
- \begin{equation}
-    \begin{tikzcd}[column sep=small]
-      \dtotalc^{a} \arrow[dd, "\pi"', color=total]  &  & \dtotalc^{b} \arrow[dd, "\pi"', color=total]                              \\
-      {}    &  & {}                                                           \\
-      \dbasec^{a} \arrow[uu, "\dsectionc^{a}"', dashed, bend right, color=section] &  & \dbasec^{b} \arrow[uu, "\dsectionc^{b}"', dashed, bend right, color=section]
-      \end{tikzcd}
-  \end{equation}
-such that output 
-\begin{equation}
-  \label{eq:artist:plus:output}
-  \gsectionc(\gbasepointc) = \begin{cases} \gsectionc^{a}(\dbasepointc) & \gbasepointc \in \vindexprec(\dbasec^{a}) \\
-    \gsectionc^{b}(\dbasepointc) & \gbasepointc \in \vindexprec(\dbasec^{b})
-  \end{cases}
-\end{equation}
-
-\subsubsection{Multiplication: $\times$}
-$\vartist \coloneqq \vartist^{a}(\dsectionc^{a}) \times \vartist^{b}(\dsectionc^{b})$
-with input 
-\begin{equation}
-  \begin{tikzcd}
-    \dtotalc^{a} \arrow[rdd, "\pi"', color=total] & & \dtotalc^{b} \arrow[ldd, "\pi"', color=total] \\
-       & & \\
-      & \dbasec \arrow[luu, "\dsectionc^{a}"', dashed, bend right, color=section] 
-      \arrow[ruu, "\dsectionc^{b}"', dashed, bend right, color=section] &                                 
-    \end{tikzcd}
-  \end{equation}
-  \\
- The output 
-\begin{equation}
-  \{\gsectionc^{a}(\gbasepointc), \gsectionc^{b}(\gbasepointc)\}, \gbasepointc \in \vindexprec(\dbasec)
-\end{equation} 
-which would mean s
-
-which would require returning two points in the fiber $\gbase \rightarrow \dfiber \times \dfiber$. This function is adjoint to the addition function in \autoref{sec:artist:addition} 
-\begin{equation}
-  hom(\dbase, \dfiber\times \dfiber) \simeq hom(\dbase \plus \dbase, \dfiber)
-\end{equation}
-such that the $\times$ operator can be implemented as a special case of \autoref{eq:artist:plus:output} where $\gbase^a$ is a copy of $\gbase$ associated with $\gsection^a$ and $\gbase^b$ is a copy of $\gbase$ associated with $\gsection^{b}$.
-
 \subsection{Equivariance}
@@
 \begin{equation}
+  \label{eq:atct:equivarance:verify:base}
   \equivc(\vartistc(\dsectionc))(\dfunch(\dbasepointc^{\prime})) = \extractmc(\vartistc(\dsectionc))(\dbasepointc)
@@
 \begin{equation}
+  \label{eq:atct:equivarance:verify:fiber}
   \equivc(\dfunctc(\dsectionc))(\dbasepoint) = \dfunctc_{\measure}(\extractmc(\vartistc(\dsectionc))(\dbasepoint))
@@
 The change in measurements $\dfunctc_{\measure}$ is defined by the developer as the symmetry between data and graphic that the artist is expected to preserve.
-For example a data to color transformation could be defined with the constraint that it must respect permutation since the data is categorical or with the constraint that it respect scaling if the data is quantative.
+
+\begin{figure}[h!]
+  \label{fig:artist:equivariance}
+  \includegraphics[width=1\columnwidth]{equivariance.png}
+  \caption{}
+\end{figure}
+For example, in \autoref{fig:artist:equivarance}, the measurable variable is color.
+This is a visual representation of the data shown in \autoref{fig:atct:phi}, and as such the equivariant transformations are an equivalent transposition and scaling of the colors.
+This visualization is equivariant with respect to base space transformations, as defined in \autoref{eq:atct:equivarance:verify:base}, because the color values at the new position at the old position $measure_\dbasepoint^{\prime} = \measure_{\dbasepoint}$.
+This visualization is also equivariant with respect to fiber wise transformations, as defined in \autoref{eq:atct:equivarance:verify:fiber}, because the colors are consistently scaled in the same was the data.
+For example, the values that have become 2 and 4 in the $\dfunctc$ and $\dfunc$ panels are colored the same as the original 2 and 4 values in the first panel.
+The equivariance in this visualization is composable, as shown in the colors being both transposed and scaled correctly in the $\dfunc$ panel.
+
+\subsection{Artist Operators}\label{sec:artist:operators}
+A common use of category theory in software engineering is the specification of modular components \cite{wielsManagementEvolvingSpecifications1998} such that we can build systems where the structure preserved by components is preserved in the composition of the components.
+We propose two types of artist composition, addition and multiplication.
+Artist addition is the composition of artists acting on distinct bundles, while artist multiplication is the composition of artists acting on different bundles over the same base space.
+
+\begin{equation}
+  \label{eq:artist:operator}
+  \begin{tikzcd}[]
+      & \dfiberc^{a} \times_{\dfiberc^{c}} \dfiber^{b} 
+    \arrow[ld, "{\pi_{a\times b, a}}"', color=fiber] 
+    \arrow[rd, "{\pi_{a\times b, b}}", color=fiber] & & 
+    \dtotalc^{a} \oplus_{\dfiberc^{c}} \dtotalc^{b} 
+    \arrow[ddd, "\pi"', color=total] \\
+    \dfiberc^{a} 
+    %\arrow[d, "\pi"', color=fiber] 
+    \arrow[r, "{\pi_{a,c}}", color=fiber] 
+    & \dfiberc^{c} & \dfiberc^{b} 
+    %\arrow[d, "\pi", color=fiber] 
+    \arrow[l, "{\pi_{b,c}}"', color=fiber] & \\
+    \dbasec^{a} 
+    \arrow[rd, "{\iota_{a, a+b}}"', color=base] 
+    & \dbasec^{c} 
+    \arrow[r, "{\iota_{c,b}}", color=base] 
+    \arrow[l, "{\iota_{c,a}}"', color=base] 
+    & \dbasec^{b} 
+    \arrow[ld, "{\iota_{b, a+b}}", color=base] 
+    &  \\
+    & \dbasec^{a}\sqcup_{\dbasec^c}\dbasec^{b} & & 
+    \dbasec^{a}\sqcup_{\dbasec^c}\dbasec^{b} 
+    \arrow[uuu, "{\dsectionc^{a,b}}"', dashed, bend right, color=section]
+  \end{tikzcd}
+\end{equation}
+
+As illustrated in \autoref{eq:artist:operator}, data bundles can be combined by taking the disjoint  union of base spaces $\dbase^{a} \sqcup_{\dbase^c} \dbase^b$.
+In the trivial case where $\dtotal^{a} = \dtotal_b$, this operation is a concatenation that adds values without adding fields.
+In this case, given $\dbasepointc_{a} \in \dbasec_{c} \subset \dbasec_{a}$ and $\dbasepointc_{b} \in \dbasec_{c} \subset \dbasec_{b}$, if $\dbasepointc_{a} = \dbasepointc_{b}$ \note{map to same thing on overlap for same measurments}
+
+\begin{equation}
+  \equivc(\dsectionc(\dbasepointc_{a})) = \equivc(\dsectionc(\dbasepointc_{b}))
+\end{equation}
+for all $\dbasepointc_{a}, \dbasepointc_{b} \in \dbasec_{a}\bigsqcup\limits_{\dbasec_{c}} \dbasec_{b}$
+
+The fibers can be combined by taking the cartesian product of the fibers $\dfiber^{a} \times_{\dfiber^c} \dfiber^{b}$. An artist acting on a cartesian product of fibers is a
+
+The expectation that the same fiber be mapped consistently is a generalization of the expectation that the same field in different views should be encoded in the same manner\cite{hullmanKeeping2018}.
+
+\subsubsection{Addition: +}
+\label{sec:artist:addition}
+ $\vartist \coloneqq \vartist^{a}(\dsectionc^{a}) + \vartist^{b}(\dsectionc^{b})$ &\\
+
+is defined to have output 
+\begin{equation}
+  \label{eq:artist:plus:output}
+  \gsectionc(\gbasepointc) = \begin{cases} \gsectionc^{a}(\dbasepointc) & \gbasepointc \in \vindexprec(\dbasec^{a}) \\
+    \gsectionc^{b}(\dbasepointc) & \gbasepointc \in \vindexprec(\dbasec^{b})
+  \end{cases}
+\end{equation}
+
+\subsubsection{Multiplication: $\times$}
+$\vartist \coloneqq \vartist^{a}(\dsectionc^{a}) \times \vartist^{b}(\dsectionc^{b})$
+where  where $\dbase_a = \dbase_b$
+
+Given that $\vindex(\gbase) = \dbase$
+\begin{equation}
+  \{\gsectionc^{a}(\gbasepointc), \gsectionc^{b}(\gbasepointc)\}, \gbasepointc \in \vindexprec(\dbasec)
+\end{equation}
+
+which would require returning two points in the fiber $\gbase \rightarrow \dfiber \times \dfiber$. This function is adjoint to the addition function in \autoref{sec:artist:addition} 
+\begin{equation}
+  hom(\gbase, \gfiber\times \gfiber) \simeq hom(\gbase + \gbase, \gfiber)
+\end{equation}
+such that the $\times$ operator can be implemented as a special case of \autoref{eq:artist:plus:output} where $\gbase^a$ is a copy of $\gbase$ associated with $\gsection^a$ and $\gbase^b$ is a copy of $\gbase$ associated with $\gsection^{b}$.
+
+For example, a box plot \cite{wickham40YearsBoxplots2011} typically consists of a rectangle, multiple lines, and scatter points.
 

### 2022-05-16 paper/main.tex
fb figure captions and explainations, operators

@@
 The \textcolor{fiber}{fiber space} is a topological space \dfiberc\ that is the preimage of the projection function $\pi$ at a point in the base space $\dfiber_{\dbasepointc} = \pi^{-1}(\dbasepointc)$ and fibers of a bundle are isomorphic $\dfiberc \simeq \dfiberc_{\dbasepointc}\;\forall \dbasepointc \in \dbase$.
-The \textcolor{section}{sections} of a bundle $\dsection$ are maps from the base space to points in the bundles over that base space and $\textcolor{set}{\Gamma}$ denotes the set of all sections in a bundle over an openset.
+The \textcolor{section}{sections} of a bundle $\dsection$ are maps from the base space to points in the bundles over that base space $\textcolor{set}{\Gamma}$ denotes the set of all sections in a bundle over an openset.
 \begin{equation}
@@
 A bundle may be globally trivial, meaning that $\dtotalc = \dbasec \times \dfiberc$, and generally can be thought of as a twisted product of total and base spaces \cite{munkresElementsAlgebraicTopology1984}.
+Throughout this work we will assume that fiber bundles are trivial because that allows us to model the section functions as from $\texttt{base} \rightarrow \texttt{fiber}$ since the fiber is the same at all points in the base space.
 
@@
 The names and data types of the data fields can be expressed as the fiber space \dfiberc, as described by Spivak \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES}.
-
-Fiber bundles can be used to model data for visualization, as proposed by Butler\cite{butlerVectorBundleClassesForm1992, butlerVisualizationModelBased1989}.
-Even though fibers on non-trivial spaces, such as vector fields, are different, we simplify and represent all fibers with the same schema because fibers are isomorphic.
+Extending Spivak's fiber bundle description of a table, we define the values returned by a section as named and typed records of the dataset.
 
@@
   \includegraphics[width=\columnwidth]{fb_tau.png}
-  \caption{change to concrete example}
+  \caption{This trivial dataset consists of a single point $\dbasec$ and the fields \texttt{speed, direction, signal, and strength}.
+Each of these fields has an associated datatype, which is the set of possible valid values for that field.
+The fiber space is the cartesian product of these sets $\dfiberc$.
+Each section $\dsectionc$ function returns a single record in the fiber, here the data record outlined in green.
+  \note{maybe change to temperature}}
 \end{figure} 
-
-As illustrated in \autoref{fig:atct:fb}, \note{figure out what the actual data example should be}
+A fiber bundle is a mathematical structure for encoding continuity, type, and datasets with that continuity and type.
+In \autoref{fig:atct:fb}, the continuity is a discrete point \dbase.
+The fiber space that encodes the types is the cartesian product of the field types, and the section returns a single record with values of each field type.
+While this is a very trivial example, a fiber bundle base space can have almost any topology.
+The fiber space also does not really have restrictions on which types it can encode or the structure of those types - for example, \texttt{speed} and \texttt{direction} are separate fibers in \autoref{fig:atct:fb} but can be combined into a single two dimensional fiber $\dfiber_{speed, direction} = {\mathbb{R}^{+}}^{s}$ The generality in base space and fiber space makes fiber bundles a unified dimension and type independent data abstraction that can encode most of the common visualization data set structures, such as continuous functions, tables, data cubes, images and networks.
 
@@
   \includegraphics[width=1\columnwidth]{fb_rho.png}
-  \caption{}
+  \caption{A bundle representation of a section that generates a scatter marker on a 2D screen.
+The fiber space encodes the potential color and position of each pixel in a prerender space.
+The marker is a 2d graphic; therefore the base space that parameterizes the graphic area \dbasec\ is a 2d disc.
+This section \gsection\ describes how to build a square red marker; a section evaluated on a single point $\gbasepoint \in \gbase$ is an infinitesimally small piece of the square, here illustrated as the small piece of the square outlined in green.
+}
 \end{figure}
-In \autoref{fig:atct:fb:graphic}, the section function $\gsection$ encodes the drawing of the \note{pin down what's beeing drawn}.   
+
+In \autoref{fig:atct:fb:graphic}, the section function $\gsection$ encodes the drawing of a red square marker.
 The renderer subsamples the position fibers  \textit{x} and \textit{y} of the graphic fiber space \gfiber.
@@
 The section is then evaluated at these points $\gsection\restriction_{\gsection^{-1}_{x,y}}$ to return a set of points in the fiber.
-Each point has an RGB tuple and these tuples are integrated to return the color of the pixel at that $(x,y)$ point.
+One of these points, enlarged for visibility, is illustrated as the small square swatch inside the larger marker in \autoref{fig:atct:fb}.
+The renderer is handed some version of $\gsection$ to produce the final graphic.
 
@@
 This means we can construct a section $\dsection$ on the collection of open sets that is the union of sections defined on specific open sets.
-This expresses an equivalency between a concatenated represetation of data and the distributed representation of the same data.
-
+This expresses an equivalency between a concatenated representation of data and the distributed representation of the same data.
 In this model, we represent data and graphics as a sheaf; doing so allows us to express that data sources are expected to have the bookkeeping that a sheaf requires, meaning inclusions, restrictions, locality, and gluing must be preserved.
@@
   \includegraphics[width=\columnwidth]{phi.png}
+  \caption{Values in a data set can be transformed in three ways: $\dfunch$-values can change position, .e.g transposed;  $\dfunct$-values can change, e.g.
+doubled; $\dfunc$ - values can change position and value}
 \end{figure}
@@
 \begin{equation}
+  \label{eq:atct:xi}
   \vindexc: \opensetgc \rightarrow \opensetc 
@@
   \includegraphics*[width=1\columnwidth]{xi_scatter.png}
-  \caption{}
+  \caption{The base space functor $\vindexc$ is the mapping from each disc $\gbasec_i$ to it's corresponding point $\dbasepointc_i$.
+Over each point is a data record $\dsectionc$.
+The pullback function $\vindexpullc$ copies these records over $\gbase$ so that each piece of each marker knows the data associated with it.
+Each graphic disc has a corresponding graphic function $\gsectionc$.
+The pushforward function $\vindexpushc$ converts $\gsectionc$ into the visual parameters associated with each data point $\dbasepoint$.}
 \end{figure}
@@
 This allows us to construct graphic specifications for each data index $\vindexpush\gsection$ and retrieve the data $\vindexpull\dsection$ for any graphic section generating any piece of a graphic.
-As illustrated in \autoref{fig:atct:morphisms:sheaf}, the \note{image functors?}
+In \autoref{fig:atct:morphisms:sheaf}, the graphic specifications are the set of visual parameters associated with each point $\dbasepoint_i$.
+Each of these parameters describes the corresponding graphic $\gsectionc|_{\gbase_i}$ generated by evaluating the graphic section on the graphic base space associated with that data point $\vindexpre^{\dbasepoint_i} = \gbase_i$.
+The pullback functor $\vindexpull$ copies each data section at each point $\dsectionc(\dbasepointc_i)$ over the corresponding disk; in \autoref{fig:atct:morphisms:sheaf} this means that every pixel in a given marker, for example the yellow triangle, maps back into a single data record.
+Visualization specification languages such as vega \cite{heerDeclarative2010} and svg \cite{quintScalable2003} assume that there is a graphic  specification for every point.
+Interactive tooltips assume that there is a map between the graphics rendered on screen and the data such that a glyph maps back to a record.
+The functors $(\vindex, \vindexpush, \vindexpull)$ codify the expectation of these mappings between data and graphic space.
 
@@
 A common use of category theory in software engineering is the specification of modular components \cite{wielsManagementEvolvingSpecifications1998} such that we can build systems where the structure preserved by components is preserved in the composition of the components.
-We propose two types of artist composition, addition and multiplication.
-Artist addition is the composition of artists acting on distinct bundles, while artist multiplication is the composition of artists acting on different bundles over the same base space.
-
+This allows us to express that an artist that works on a dataset can be composed of artists that work on sub parts of that dataset.
 \begin{equation}
   \label{eq:artist:operator}
-  \begin{tikzcd}[]
-      & \dfiberc^{a} \times_{\dfiberc^{c}} \dfiber^{b} 
+  \begin{tikzcd}[column sep=small]
+      & \dfiberc^{a} \times_{\dfiberc^{c}} \dfiberc^{b} 
     \arrow[ld, "{\pi_{a\times b, a}}"', color=fiber] 
@@
 
-As illustrated in \autoref{eq:artist:operator}, data bundles can be combined by taking the disjoint  union of base spaces $\dbase^{a} \sqcup_{\dbase^c} \dbase^b$.
-In the trivial case where $\dtotal^{a} = \dtotal_b$, this operation is a concatenation that adds values without adding fields.
-In this case, given $\dbasepointc_{a} \in \dbasec_{c} \subset \dbasec_{a}$ and $\dbasepointc_{b} \in \dbasec_{c} \subset \dbasec_{b}$, if $\dbasepointc_{a} = \dbasepointc_{b}$ \note{map to same thing on overlap for same measurments}
+\subsubsection{Addition: +}
+\label{sec:artist:addition}
+As illustrated in \autoref{eq:artist:operator}, data bundles can be combined by taking the disjoint  union of base spaces $\dbase^{a} \sqcup_{\dbase^c} \dbase^b$, where $\dbase^c$ is an overlap.
+When the fibers of each bundle are isomorphic $\dfiber^{a} \simeq \dfiber^{b}$, which we denote as $\dtotal$, this is analogous to adding more observations to a dataset.
+We propose an addition operator that states that an artist that takes in a dataset can be constructed using artists that take as inputs subsets of the dataset
+
+\begin{equation*}
+  \label{eq:artist:addition}
+  \vartistc_{a+b}(\cgamma{\dbasec^{a} \sqcup_{\dbase^c} \dbasec^b}{\dtotalc}) \coloneqq \vartistc_{a}(\cgamma{\dbasec^{a}}{\dtotalc}) + \vartistc_{b}(\cgamma{\dbasec^{b}}{\dtotalc}) 
+\end{equation*}
+As introduce in $\autoref{eq:artist:hom_transport}$, the artist returns a function $\gsection$.
+We assume that the output space is a trivial bundle, which means that $\gsectionc \in Hom(\gbase, \gfiber)$ because the output specification is the same at each point $\gbase$.
+This allows us to make use of the hom set adjoint property\note{find citation}\
+\begin{equation*}
+  Hom(\gbase^{a} + \gbasec^b, \gfiber) = Hom(\gbase^{a}, \gfiber) + Hom(\gbase^b, \gfiber)
+\end{equation*} 
+to define an artist constructed via addition as consisting of two distinct graphic sections
+\begin{equation}
+  \label{eq:artist:plus:output}
+  \gsectionc(\gbasepointc) \coloneqq \begin{cases} \gsectionc^{a}(\gbasepointc) & \gbasepointc \in \vindexprec(\dbasec^{a}) \\
+    \gsectionc^{b}(\gbasepointc) & \gbasepointc \in \vindexprec(\dbasec^{b})
+  \end{cases}
+\end{equation}
+that are evaluated only if the input graphic point is an the graphic area that graphic section acts on.
+
+One way to verify that these artists are composable is to check that the return the same graphic on points in the intersection $\dbase^{c}$.
+Given $\dbasepointc_{a} \in \dbasec_{c} \subset \dbasec_{a}$ and $\dbasepointc_{b} \in \dbasec_{c} \subset \dbasec_{b}$, if $\dbasepointc_{a} = \dbasepointc_{b}$ then
 
 \begin{equation}
-  \equivc(\dsectionc(\dbasepointc_{a})) = \equivc(\dsectionc(\dbasepointc_{b}))
+  \label{eq:artist:plus:verify}
+  \begin{split}
+  \vartistc_{a+b}(\dsectionc^{a+b}(\dbasepointc_{a})) \\ 
+  = \vartistc_{a}(\dsectionc^{a}(\dbasepointc_{a})) = \vartistc_{b}(\dsectionc^{b}(\dbasepointc_{b}))
+  \end{split}
 \end{equation}
@@
 
-The fibers can be combined by taking the cartesian product of the fibers $\dfiber^{a} \times_{\dfiber^c} \dfiber^{b}$. An artist acting on a cartesian product of fibers is a
+ One example of an artist that is a sum of artists is a sphere drawer that draws different quadrants of a sphere $\vartist(\dsection) = \vartist_{1}(\dsection_{1}) + \vartist_{2}(\dsection_{2}) + \vartist_{3}(\dsection_{3}) \vartist_{4}(\dsection_{4})$.
+Given an input $\dbasepoint \in \dbase_4$ in the 4th quadrant, then the graphic section that would be executed is $\gsection_{4}$.
+If that point is also in the 3rd quadrant  $\dbasepoint \in \dbase_3$, then both artist outputs must return the same values $\gsection_{4}(\vindexprec(\dbasepoint)) = \gsection_{3}(\vindexprec(\dbasepoint))$.
 
-The expectation that the same fiber be mapped consistently is a generalization of the expectation that the same field in different views should be encoded in the same manner\cite{hullmanKeeping2018}.
+\subsubsection{Multiplication: $\times$}
+As illustrated in \autoref{eq:artist:operator},  fibers that are a cartesian product of fiberspaces $\dfiber^{a} \times_{\dfiber^c} \dfiber^{b}$, where $\dfiber^c$ is any fiber that is present in both fibers, can be projected down into component fibers.
+In the trivial case where the base spaces are the same $\dbase^{a} = \dbase^{b} = \dbase$, this is equivalent to adding more fields to a dataset.
 
-\subsubsection{Addition: +}
-\label{sec:artist:addition}
- $\vartist \coloneqq \vartist^{a}(\dsectionc^{a}) + \vartist^{b}(\dsectionc^{b})$ &\\
+\begin{equation*}
+  \label{eq:artist:multiplication}
+  \vartistc_{a \times b}(\cgamma{\dbase}{\dtotalc^{a\times b}}) \coloneqq \vartistc_{a}(\cgamma{\dbasec}{\dtotalc^{a}}) \times \vartistc_{b}(\cgamma{\dbasec}{\dtotalc^{b}}) 
+\end{equation*}
+which following from an adjoint property of homsets \note{find citation}
+\begin{equation*}
+  Hom(\gbase, \gfiber) \times Hom(\gbase, \gfiber) = Hom(\gbase, \gfiber\times \gfiber)
+\end{equation*}
+which means that the artists on the subsets of fibers can be defined 
+\begin{equation}
+  \gsectionc^{a \times b} = \{\gsectionc^{a}(\gbasepointc), \gsectionc^{b}(\gbasepointc)\}, \gbasepointc \in \vindexprec(\dbasec)
+\end{equation} 
+but that the signature of $\gsectionc^{a \times b}$ would be $\gbase \rightarrow \gfiber \times \gfiber$.
+Instead of having to special case the return type of artists that are compositions of multiple case, the hom adjoint \note{find cite} property
+\begin{equation*}
+  Hom(\gbase, \gfiber \times \gfiber) = Hom(\gbase+\gbase, \gfiber)
+\end{equation*}
+ means that multiplication can be considered as a special case of addition where $\dbase^{a} = \dbase^{b}$.
+While we discussed the trivial case in \autoref{sec:artist:addition}, there is no strict  requirement that $\dfiber^{a} = \dfiber^{b}$.
 
-is defined to have output 
+One way to verify that these artists are composable is to check that they encode any shared fiber $\dfiber^{c}$ in the same way.
+
 \begin{equation}
-  \label{eq:artist:plus:output}
-  \gsectionc(\gbasepointc) = \begin{cases} \gsectionc^{a}(\dbasepointc) & \gbasepointc \in \vindexprec(\dbasec^{a}) \\
-    \gsectionc^{b}(\dbasepointc) & \gbasepointc \in \vindexprec(\dbasec^{b})
-  \end{cases}
+  \begin{split}
+    \extractmc(\vartistc_{a\times b}(\dsectionc^{a\times b}(\dbasepointc)))\restriction_{dfiberc^{c}}\\ = 
+    \extractmc(\vartistc_{a}(\dsectionc^{a}(\dbasepointc_{a})))\restriction_{\dfiber^{c}} = \extractmc(\vartistc_{b}(\dsectionc^{b}(\dbasepointc_{b})))\restriction_{\dfiber^{c}}
+  \end{split}
 \end{equation}
 
-\subsubsection{Multiplication: $\times$}
-$\vartist \coloneqq \vartist^{a}(\dsectionc^{a}) \times \vartist^{b}(\dsectionc^{b})$
-where  where $\dbase_a = \dbase_b$
-
-Given that $\vindex(\gbase) = \dbase$
-\begin{equation}
-  \{\gsectionc^{a}(\gbasepointc), \gsectionc^{b}(\gbasepointc)\}, \gbasepointc \in \vindexprec(\dbasec)
-\end{equation}
-
-which would require returning two points in the fiber $\gbase \rightarrow \dfiber \times \dfiber$. This function is adjoint to the addition function in \autoref{sec:artist:addition} 
-\begin{equation}
-  hom(\gbase, \gfiber\times \gfiber) \simeq hom(\gbase + \gbase, \gfiber)
-\end{equation}
-such that the $\times$ operator can be implemented as a special case of \autoref{eq:artist:plus:output} where $\gbase^a$ is a copy of $\gbase$ associated with $\gsection^a$ and $\gbase^b$ is a copy of $\gbase$ associated with $\gsection^{b}$.
-
-For example, a box plot \cite{wickham40YearsBoxplots2011} typically consists of a rectangle, multiple lines, and scatter points.
+This expectation of using the same encoding for the same variable is a generalization of the concept of consistency checking of multiple view encodings discussed by Zening and Hullman \cite{hullmanKeeping2018}.
+This expectation can also be used to check that a multipart glyph is assembled correctly.
+For example, a box plot \cite{wickham40YearsBoxplots2011} typically consists of a rectangle, multiple lines, and scatter points; therefore a boxplot artist $\vartist_{boxplot} = \vartist_{rect} \times \vartist_{errors} \times \vartist_{line} \times \vartist_{points}$ must be constructed such that all the sub artists draw a graphic at or around the same x value.
+\note{maybe stripped down box figure}
 

### 2022-05-16 paper/main.tex
xi

@@
 \subsection{Maps Between Spaces}
+\label{sec:atct:xi}
 The data sections defined in \autoref{eq:atct:fb} are sections of sheaves, introduced in \autoref{sec:atct:sheaves}, over open sets $\openset \subseteq \dbase$ in a base space \dbase\ that encodes the topology of the data topology.
@@
 
-\subsection{Artist Operators}\label{sec:artist:operators}
+\subsection{Composing Artists}\label{sec:artist:operators}
 A common use of category theory in software engineering is the specification of modular components \cite{wielsManagementEvolvingSpecifications1998} such that we can build systems where the structure preserved by components is preserved in the composition of the components.
@@
 \subsubsection{Multiplication: $\times$}
+\label{sec:artist:operator:multiplication}
 As illustrated in \autoref{eq:artist:operator},  fibers that are a cartesian product of fiberspaces $\dfiber^{a} \times_{\dfiber^c} \dfiber^{b}$, where $\dfiber^c$ is any fiber that is present in both fibers, can be projected down into component fibers.
@@
 For example, a box plot \cite{wickham40YearsBoxplots2011} typically consists of a rectangle, multiple lines, and scatter points; therefore a boxplot artist $\vartist_{boxplot} = \vartist_{rect} \times \vartist_{errors} \times \vartist_{line} \times \vartist_{points}$ must be constructed such that all the sub artists draw a graphic at or around the same x value.
-\note{maybe stripped down box figure}
+
+\begin{figure}[h!]
+  \label{fig:artist:operator}
+  \includegraphics*[width=1\columnwidth]{qcom.png}
+  \caption{The circle-line glyph can either be considered as composed of two separate shapes, a circle and line, which each have their own graphic base spaces or as a single glyph which has an equivalent graphic base  $\gbase_{circle} \sqcup \gbase_{line} = \gbase_{circle-line}$ }
+\end{figure}
+
+The circle-line glyph in \autoref{fig:artist:operator} can be generated via a line plot and highlighted points in that plot, which is the addition case of two datasets with the same fields and different continuities.
+The same circle-line glyph can also be generated via the multiplication case as different fields of the same bundle.
+As mentioned in \autoref{sec:artist:operator:multiplication}, multiplication is implemented as addition of two copies of a base space.
+Although we have been discussing the trivial cases of adding observations or adding fields, this merging of artists in datasets can be generalized:
+\begin{equation}
+  \vartist(\cgamma{\bigsqcup_{i} \dbase^{i}}{\bigoplus_{i}\dtotalc^{i}}) = \sum_{i}
+  \vartist_{i}(\cgamma{\dbase^{i}}{\dtotal^{i}}) 
+\end{equation}
+
+As shown in \autoref{eq:artist:operator}, bundles over a union of base spaces can be joined as a product of the fibers.
+This allows us to consider all the data inputs in a complex visualization as a combined input, where some sections evaluate to null in fields for which there are no values for that point in the combined base space $\dbasepoint \in \bigsqcup_{i} \dbase^{i}$ The combined construction of the data is a method for expressing what each data input has in common with another data input-for example the data for labeling tick marks or legends-
+and therefore which commonalities need to be preserved in the artists that act on these inputs.
 
@@
 \label{sec:construction}
+\begin{figure}[h!]
+  \label{fig:construction:flow-chart}
+  \includegraphics*[width=1\columnwidth]{path_of_q.png}
+  \caption{One method of constructing the artist is by formulating the artist as having two stages: (1) encoding functions $\vchannelc$ that convert different parts of the input into measurable visual components, such as color or position; (2) compositing functions $\vmarkc$ that assemble the measurable visual components into a visual element generating function.
+In this construction, we propose an intermediate visual bundle $\vtotalc$ that encodes the space of possible visual encodings.}
+\end{figure}
+We propose that one way of constructing artist functions is to separate generating a visualization, as shown in \autoref{fig:construction:flow-chart}, is into an equivariant continuity preserving encoding stage $\vchannelc$ and a equivariant continuity preserving compositing stage $\vmarkc$.
+In the \textcolor{artist}{encoding} stage $\vchannelc$, a data bundle is treated as separable fields and each field is mapped to a measurable visual variable.
+In the encoding stage, the expected visual mappings $\equivb$ can be implemented inside the library.
+Factoring out the encoding stage leaves the \textcolor{artist}{compositing} stage $\vmarkc$ responsible for faithfully translating those measurable visual components into a visual element.
+
 \subsection{Measurable Visual Components}
-
+\label{sec:construction:vtotal}
+As shown in \autoref{fig:construction:flow-chart}, we propose an intermediate visual fiber bundle 
 \begin{equation}
@@
 \end{equation}
-where
+where the space of possible visual encodings is the fiber space $\vfiber$. The space of visual sections which return visual encoding specifications is 
 \begin{equation}
@@
 \end{equation}
-
-The bundle $\vtotal$ is structurally equivalent to the bundle $\dtotal$, but allows for separating the data that is input into the artist $\dsection$ from the data internal to the artist $\vsection$.
-This separation also allows for defining many of the target measurements \measure\ and the equivariant transformations to the measurements as internal components of the artists.
-
-\subsection{$\vindexc: \gbasec \rightarrow \dbasec$}
-We define the functor $\vindex$ such that the data base space \dbase\ is a deformation retraction\cite{nlab:deformation_retract} of the graphic space $\gbase$, meaning that there is a continuous surjective mapping from every point $\dbasepoint\in\dbase$ to a point $\dbasepoint \in dbase$ .
+As shown in \autoref{eq:artist:inout:diagram},  measurable visual components are defined as having the same continuity as the data $\dbase$.
+This means that every data record $\dsection(\dbasepoint)$ has a corresponding visual section such that $\pi(\dsection(\dbasepoint)) = \pi(\vsection(\dbasepoint))$.
+
+Although the bundle $\vtotal$ is structurally equivalent to the bundle $\dtotal$, its existence allows for separating the data that is input into the artist $\dsection$ from the data internal to the artist $\vsection$.
+This allows a visualization library to define a visual bundle space $\vtotalc$ the holds the internal representation standard for measurable visual components.
+For example, a visualization library could define a visual color fiber as an RGBA tuple $\vfiber_{color} = \mathbb{R}^{3} \times [0,1]$.
+This would then set the expectation that arbitrary color encoding functions would need to return an RGBA tuple for the library to recognize the encoding as a color.
+
+\subsection{Defining Maps Between Spaces $\vindex$}
+In \autoref{sec:atct:sheaves}, the functors $\vindex, \vindexpull, \vindexpushc$ are introduced to describe the expected relationship between screen and data base spaces.
+In the construction of the artist, the functor $\vindex$ is defined such that the data base space \dbase\ is a deformation retraction\cite{nlab:deformation_retract} of the graphic space $\gbase$.
+This means that there is a continuous surjective mapping from every point $\dbasepoint\in\dbase$ to a point $\dbasepoint \in dbase$.
 To simplify matters, in this paper, we construct the graphic space as a constant multiple of the base space such that
@@
 \end{equation}
+where n is a thickening of the graphic base space $\gbase$ to account for the dimensionality of the output space
 \begin{equation*}
@@
 \end{equation*}
+because the data dimensionality $\dbase$ may be too small for a graphic representation.
+
+\begin{figure}[h!]
+  \label{fig:construction:xi}
+  \includegraphics[width=1\columwidth]{deform_retract.png}
+  \caption{The graphic base space $\gbase$ is collapsable to the line $\dbase$ such that every band $(\dbasepoint_i, [0,1])$ on $\gbase$ maps to corresponding point $\dbasepoint_i \in \dbase$ and the band $[0,1]$ determines which pixels are filled in to generate the shape of the line.\note{maybe put in a phantom line plot in s}}
+\end{figure}
+For example, as shown in \autoref{fig:construction:xi}, a line is 1D but is a 2D glyph on a screen; therefore the graphic space $\gbase$ is constructed by multiplying the base space $\dbase$ with an interval $[0,1]$.
+Because $\gbase$ is collapsible into $\dbase$, every band $(\dbasepoint_i, [0,1])$ corresponds to a point in the base space $\dbasepoint_i \in \dbase$.
+The first coordinate $\alpha=\dbasepoint_i$ provides a lookup to retrieve the associated visual variables.
+The second coordinate, which is a point in the interval $\beta=[0,1]$.
+Together they are a point $\gbasepoint=(\alpha,\beta) \in gbase$ in the graphic base space.
+This point $\gbasepoint$ is the input into the graphic section $\gsection(\gbasepoint)$ that is used to determine which pixels are colored, which in turn determines the thickness, texture, and color of the line.
 

### 2022-05-17 paper/main.tex
q

@@
 
-The fiber category is equipped with a bifunctor $\otimes: \mathcal{\dfiber} \times {\dfiber}$.
+The fiber category is equipped with a bifunctor $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$.
 The bifunctor provides a method for combining fibers, thereby allowing for fields that contain multityped values.
@@
 This expresses an equivalency between a concatenated representation of data and the distributed representation of the same data.
-In this model, we represent data and graphics as a sheaf; doing so allows us to express that data sources are expected to have the bookkeeping that a sheaf requires, meaning inclusions, restrictions, locality, and gluing must be preserved.
-
-\subsubsection{Structure on Data}
+In this model, we represent data and graphics as a sheaf; doing so allows us to express that data sources are expected to implement the bookkeeping that a sheaf requires, meaning inclusions, restrictions, locality, and gluing must be preserved.
+
+\subsubsection{Structure}
 \label{sec:atct:sheaves:measurments}
@@
  The diagram in \autoref{eq:artist:natural_transform} shows that restricting a set of outputs of an artist to a set of graphic sections over a subspace is equivalent to restricting the inputs to data sections over the same subspace.
-Since the input and output of the artist is always sections over the same subspace, this is why the by construction the artist preserves the bookkeeping provided through the use of sheaves.
+By construction, the artist preserves the bookkeeping provided through the use of sheaves and transport functors.
+Specifically, the artist always takes as input a section over a data subspace and returns a section over the corresponding graphic subspace.
+Since these are sections of sheaves, they must be implemented such that they satisfy the sheaf axioms (\autoref{def:atct:sheaf:gluing}, \autoref{def:atct:sheaf:locality}) and therefore can be glued together and seperated as needed.
 
@@
   \begin{tikzcd}[row sep=huge]
-    \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
+    \cgamma{\dbasec}{\dtotalc} 
     \arrow[rr, "\vartistc", color=artist] 
     \arrow[d, "\equivc"', color=monoid] &  & 
-    \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+    \imartist{\gbasec}{\gtotalc} 
     \arrow[d, "render"] 
     \arrow[lld, "\extractmc"', color=monoid, dashed] \\
-    {\textcolor{set}{Hom}(\opensetc, \measurec)}  &  & visualization 
+    {\textcolor{set}{Hom}(\dbasec, \measurec)}  &  & visualization 
     \arrow[ll, "measure",]
@@
   \label{eq:artist:actual}
-  \extractmc:\gsectionc \mapsto (\opensetc \xrightarrow{\extractmc_{\gsectionc}} \measurec)
+  \extractmc: (\gsectionc \circ \vindexpre) \mapsto (\dbasec \xrightarrow{\extractmc_{\gsectionc}} \measurec)
 \end{equation}
 such that $\extractmc_{\gsectionc}(\dbasepoint)$ returns a quantification of the measurable components $\measurec_{\dbasepointc}$ of the visual element generated for the data at that point $\gsectionc|_{\vindexprec(\dbasepointc)}$.
+As shown in \autoref{eq:artist:actual}, the function $\extractmc$ takes in $\gsectionc$ and must evaluate the graphic section on all the graphic space associated with a point $\vindexpre\restriction_{\dbasepoint}$.
+This is because the measurable aspects, such as thickness or marker shape, refer to the entire visual element for that data point and as such can only be computed on the whole element.
 
@@
 \begin{equation}
-\equivc: \dsectionc \mapsto (\opensetc \xrightarrow{\equivc_{\dsectionc}} \measurec)
+\equivc: \dsectionc \mapsto (\dbase \xrightarrow{\equivc_{\dsectionc}} \measurec)
 \end{equation}
@@
   \label{eq:atct:equivarance:verify:base}
-  \equivc(\vartistc(\dsectionc))(\dfunch(\dbasepointc^{\prime})) = \extractmc(\vartistc(\dsectionc))(\dbasepointc)
+  \equivc(\dsectionc)(\dfunch(\dbasepointc^{\prime})) = \extractmc(\vartistc(\dsectionc))(\dbasepointc)
 \end{equation}
@@
   \begin{split}
-  \vartistc_{a+b}(\dsectionc^{a+b}(\dbasepointc_{a})) \\ 
-  = \vartistc_{a}(\dsectionc^{a}(\dbasepointc_{a})) = \vartistc_{b}(\dsectionc^{b}(\dbasepointc_{b}))
+  &\vartistc_{a+b}(\dsectionc^{a+b}(\dbasepointc_{a})) \\ 
+  & = \vartistc_{a}(\dsectionc^{a}(\dbasepointc_{a})) = \vartistc_{b}(\dsectionc^{b}(\dbasepointc_{b}))
   \end{split}
@@
   \begin{split}
-    \extractmc(\vartistc_{a\times b}(\dsectionc^{a\times b}(\dbasepointc)))\restriction_{dfiberc^{c}}\\ = 
-    \extractmc(\vartistc_{a}(\dsectionc^{a}(\dbasepointc_{a})))\restriction_{\dfiber^{c}} = \extractmc(\vartistc_{b}(\dsectionc^{b}(\dbasepointc_{b})))\restriction_{\dfiber^{c}}
+    &\extractmc(\vartistc_{a\times b}(\dsectionc^{a\times b}(\dbasepointc)))\restriction_{\dfiberc^{c}}\\ 
+    &= 
+    \extractmc(\vartistc_{a}(\dsectionc^{a}(\dbasepointc_{a})))\restriction_{\dfiberc^{c}} = \extractmc(\vartistc_{b}(\dsectionc^{b}(\dbasepointc_{b})))\restriction_{\dfiberc^{c}}
   \end{split}
@@
 \begin{equation}
-  \vartist(\cgamma{\bigsqcup_{i} \dbase^{i}}{\bigoplus_{i}\dtotalc^{i}}) = \sum_{i}
-  \vartist_{i}(\cgamma{\dbase^{i}}{\dtotal^{i}}) 
+  \vartistc(\cgamma{\bigsqcup_{i} \dbasec^{i}}{\bigoplus_{i}\dtotalc^{i}}) \coloneqq \sum_{i}
+  \vartisct_{i}(\cgamma{\dbasec^{i}}{\dtotalc^{i}}) 
 \end{equation}
@@
   \label{fig:construction:xi}
-  \includegraphics[width=1\columwidth]{deform_retract.png}
+  \includegraphics[width=1\columnwidth]{deform_retract.png}
   \caption{The graphic base space $\gbase$ is collapsable to the line $\dbase$ such that every band $(\dbasepoint_i, [0,1])$ on $\gbase$ maps to corresponding point $\dbasepoint_i \in \dbase$ and the band $[0,1]$ determines which pixels are filled in to generate the shape of the line.\note{maybe put in a phantom line plot in s}}
@@
 
-\subsection{Transforming Data into Measurable Components}
-The transformation $\vchannel$ is a map from data sections to graphic sections
+\subsection{Encode Data as Measurable Components}
+\label{sec:construction:nu}
+As introduced in \autoref{fig:construction:flow-chart}, the encoding function $\vchannel$ is a map from data sections to graphic sections
 \begin{equation}
@@
 such that the sections project to the same point on the base space $\pi(\dtotal) = \pi(\vchannel(\dtotal))$.
-As with artists, $\vchannel$ are maps of sections such that the composition operators defined in \autoref{sec:artist:composition} can also act on transformers $\vchannel$.
-
-A consequence of $\pi(\dtotal) = \pi(\vchannel(\dtotal))$ is that $\vchannelc$ can be constructed as a pointwise transformation such that  
+A consequence of this property is that $\vchannelc$ can be constructed as a pointwise transformation such that
 \begin{equation}
@@
 which means that means that a point in a single data fiber $\delement \in \dfiber_{\dbasepointc}$ can be mapped into a corresponding point in a visual fiber $\velement \in \vfiber_{\dbasepointc}$.
-Since $\vchannel$ can be constructed as a pointwise function property, we can construct $\vchannelc$ functions
+This means that an encoding function $\vchannel$ can convert a single record and may not need the whole dataset.
+
+The difference between $\dtotal$ and $\vtotal$ are semantic rather than structural; they are both maps from the topological space \dbase\ to sets of functions that return records of values.
+This means any $\vtotal$ can be redefined as $\dtotal$,
 \begin{equation}
@@
 \end{equation}
-such that any new $\vchannel$ can be constructed as a composition of two or more $\vchannel$s.
-
-\subsubsection{Transformer Equivariance}
-
-A pair of transformations $(\vchannel, \vchannel^{\prime})$ are equivariant if 
-\begin{equation}
-  \dfunct_{\vtotal} \vchannel(\dsection(\dbasepoint)) = \vchannel(\dfunct_{\dtotal}(\dsection(\dbasepoint)))
-\end{equation}
-for all $\dbasepoint \in \dbase$.  Since the visual sections and data sections are over the same space $\dfunc$
-
-\begin{equation}
-  \label{def:construction:nu:validate}
+which means that, as shown in \autoref{eq:construction:nu:fabrication}, any collection of $\vchannel$ functions can be composed such that they are equivalent to a $\vchannel$ that directly converts the input to the output.
+This means that a developer has the flexibility to implement either the direct convertor or the building block version (or both) depending on other library concerns.
+As with artists, $\vchannel$ are maps of sections such that the composition operators defined in \autoref{sec:artist:composition} can also act on transformers $\vchannel$, meaning that encoders can be added and multiplied.
+
+\subsubsection{Encoder Verification}
+\label{sec:construction:nu:verification}
+A  motivation for constructing an artist with an encoder stage $\vchannel$ is so that the conversion from data to measurable component can be tested separately from the assembly of components into a glyph.
+\begin{equation}
+  \label{eq:construction:nu:validate}
   \begin{tikzcd}[column sep=4em]
@@
 \end{equation}
-
-\subsection{Compositing Measurable Components into Output}
+As shown in \autoref{eq:construction:nu:validate}, an encoder is considered valid if there is an ismorphism between the actual outputted visual component and the expected measurable component encoding.
+An encoder is consistent if it encodes the same field in the same way even if coming from different data sources.
+
+An encoding function $\vchannelc$ is equivariant if the change in data, as defined in \autoref{sec:atct:sheaves:measurments}, and change in visual components are equivariant.
+Since $\dtotal$ and $\vtotal$ are over the same base space and are pointwise, the base space change $\dfunch_{\dtotal}$ applies to both sides of the equation
+\begin{equation}
+  \vchannelc(\dsectionc_{\dtotalc}(\dfunchc_{\dbasec}(\dbasepointc^{\prime}))) = \vsectionc(\dfunchc_{\dbasec}(\dbasepointc^{\prime}))
+\end{equation}
+and therefore there should not be a change in encoding. On the other hand, a change in the data values $\dfunct_{\dtotal}$ must have an equivalent change in visual components
+\begin{equation}
+  \dfunctc_{\vtotal} \vchannelc(\dsectionc(\dbasepoint)) = \vchannelc(\dfunctc_{\dtotalc}(\dsectionc(\dbasepointc)))
+\end{equation}
+The change in visual components $\dfunct_{\vtotal}$ is dependent both on $\dfunct_{\dtotal}$ and the choice of visual encoding.
+As mentioned in \autoref{sec:related-work:equivariance}, this is why Bertin and many others since have advocated choosing an encoding that has a structure that matches the data structure\cite{bertinSemiologyGraphicsDiagrams2011a}.
+For example choosing a quantative colormap to encode quantative data if the $\dfunctc$ operation is scaling, as in \autoref{fig:artist:equivariance}.
+
+\subsection{Composite Measurable Components Into Visual Elements}
+As shown in \autoref{fig:construction:flow-chart}, the compositor function $\vmarkc$ transforms the measurable components into properties of a visual element.
 The compositing function $\vmarkc$ transforms the sections of visual elements $\vsectionc$ into sections of graphics $\gsectionc$.
@@
 
-\subsubsection{Compositor Equivariance}
-
-As mentioned in \autoref{sec:construction:nu}, the distinction between data sections $\dsection$ and visual sections $\vsection$ is semantic rather than structural.
-Therefore, the equivariance condition on the sheaf map with a change of base space $\vmarkc$ is identical to the one for artists shown in \autoref{eq:artist:equivariance} such that an equivariant \vmark\ must satisify the condition
-\begin{equation}
-  \vmarkc(\dfunctc_{\vtotalc}(\vsectionc(\dfunchc_{\vtotalc}(\vindexc(\gbasepointc^{\prime}))))) = \dfunctc_{\gtotalc}(\vmarkc(\vsectionc(\vindexc(\dfunchc_{\gtotalc}(\gbasepointc^{\prime})))))  
-\end{equation}
-which means that any change to a measurable component input must have a measurably equivalent change in the output.
-
-As with the artist, it is difficult to work with the output section \gsection\ directly and instead we test the compositer \vmark\ using measurements derived from the output.
-A compositer is expected to pass visual transformers \vsection\ through to their expected visual measurements, as described in \autoref{def:construction:nu:validate}, such that any shared visual components at a point $\vsection^{a}(\dbasepoint)$ must render to the same measurement $\measure^{a}_{\dbasepoint}$
+\subsubsection{Compositor Verification}
+\label{sec:construction:q:verification}
+An advantage of factoring out encoding and verification, as discussed in \autoref{sec:construction:nu:verification}, is that the responsibility of the compositor can be scoped to translating measurable components into visual elements.
 \begin{equation}
@@
    \arrow[rr, "\simeq", dotted] &  & {\textcolor{set}{Hom}(\opensetc, \measurec^{a})}                                                                             \\
-    \cgamma{\opensetc}{(\vtotalc^{a}\times \vtotalc^{c})\restriction_{\opensetc}}  \arrow[rr, "\vmarkc_{ab}", color=artist] \arrow[u, "\pi_a", color=total]  &  &  \imartistsub{ac}{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \arrow[u, "\measurec\restriction_a \circ \extractmc_{ac}"', color=monoid]
-   \end{tikzcd}
-\end{equation}
-
-\note{when $Q_{ab} = Q_{bc}$, this says only things that change are the parameter controlled pieces}
+    \cgamma{\opensetc}{(\vtotalc^{a}\times \vtotalc^{c})\restriction_{\opensetc}}  \arrow[rr, "\vmarkc_{ac}", color=artist] \arrow[u, "\pi_a", color=total]  &  &  \imartistsub{ac}{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \arrow[u, "\measurec\restriction_a \circ \extractmc_{ac}"', color=monoid]
+   \end{tikzcd}
+\end{equation}
+As illustrated in \autoref{eq:construction:q:validate}, a compositor is valid if there is an isomorphism between the actual outputted measured visual component and the expected measurable component that is the input.
+One way of verifying that a compositor is consistent is by verifying that it passes through one encoding even while changing others.
+For example, when $\vmark_{ab}=\vmark_{ac}$ then the output should differ in the same measurable components as $\vsectionc_{ab}$ and $\vsection_{ac}$.
+
+A compositor function \vmark\ is equivariant if the renderer output changes in a way equivariant to the data transformation defined in \autoref{sec:atct:sheaves:measurments}.
+This means that a change in base space $\dfunch_{\dtotalc}$ should have an equivalent change in visual element base space.
+This means that there should be no change in visual measurement
+\begin{equation}
+  \label{eq:construction:q:verification:base}
+  \vsectionc(\dfunchc_{\dbasec}(\dbasepointc^{\prime})) = \extractmc(\vmarkc(\vsectionc)(\dfunch_{\dbasec}(\vindexpre^{\dbasepointc}))) = \measurec_{\dbasepointc}
+\end{equation}
+As discussed in \autoref{fig:artist:equivariance}, the change in base space may induce a change in locations of measurements relative to each other in the output; this can be verified via checking that all the measurements have not changed relative to the original positions $\measurec_{\dbasepoint} = \measurec_{\dbasepointc^{\prime}}$ and through separate measurable variables that encode holistic data properties, such as orientation or origin.
+
+The compositor function is also expected to be equivariant with respect to changes in data and measurable components
+\begin{equation}
+  \dfunctc_{\vtotalc}(\vsectionc(\dbasepointc)) = \dfunctc_{\gtotalc}(\vmarkc(\vsectionc(\dbasepointc)))  
+\end{equation}
+which means that any change to a measurable component input must have a measurably equivalent change in the output.
+As illustrated in \autoref{fig:artist:equivariance}, the compositor $\vmarkc$ is expected to assemble the measurable components such that base space changes, for example transposition, are reflected in the output; faithfully pass through equivariant measurable components, such as scaled colors; and ensure that both types of transformations, here scaling and transposition, are present in the final glyph.
 
@@
 
-As shown in \autoref{eq:atct:sheaves:homset}, sheaves equipped with transport functors have isomorphic maps between the sheaves.
+As shown in \autoref{eq:atct:sheaves:homset}, if a sheaf is equipped with transport functors, then the functions between sheaves over one space are isomorphic to functions between sheave over the other space.
 This property, allows us to specify the transformation from data to graphic space in data or graphic space or a combination, as listed in \autoref{eq:artist:hom_transport}, such that the following diagram commutes:
@@
 This allows developers to for example connect transformers that transform data on a line to a color in dataspace, but build a line compositing function that dynamically resamples what is on screen in graphic space.
-x
-
-\section{Discussion}
+
+\section{Discussion: Implementing The Math}
 \subsection{Limitations}

### 2022-05-17 paper/main.tex
atct

@@
 \IEEEraisesectionheading{\section{Introduction}\label{sec:intro}}
+\note{please prioritize flagging content that needs major revisions: (1) Wrong (2) Missing (3) Horrific  (4) Incoherant}
 
@@
 \label{sec:atct}
-\note{why why why?}
-We use concepts from algberaic topolology and category theory to formally express the structure visualization components are expected to preserve.
-Fiber bundles and sheaves, which come from algebraic topolology, provide a uniform generalizable way to describe data and visualizations.
+A goal of this work is to develop guidelines for building composable structure preserving functions for translating data into visual representations.
+Concepts in algebraic topology and category theory provide a language for formally expressing the structure that these visualizations are expected to preserve.
+Fiber bundles and sheaves, which come from algebraic topolology, provide a uniform generalizable way to describe data and visualization.
 Category theory is a method to describe how objects are specified, constrained, and composed \cite{wielsManagementEvolvingSpecifications1998}\; therefore in this work we propose categorical constructions of visualization components as specifications for implementable components.
@@
   \includegraphics[width=1\columnwidth]{equivariance.png}
-  \caption{}
+  \caption{This artist is equivariant because the input data and output graphic representation are transposed $\dfunch$, scaled $\dfunct$ and both transposed and scaled $\dfunc$ in the same manner.}
 \end{figure}
@@
   \vartistc(\cgamma{\bigsqcup_{i} \dbasec^{i}}{\bigoplus_{i}\dtotalc^{i}}) \coloneqq \sum_{i}
-  \vartisct_{i}(\cgamma{\dbasec^{i}}{\dtotalc^{i}}) 
+  \vartistc_{i}(\cgamma{\dbasec^{i}}{\dtotalc^{i}}) 
 \end{equation}
@@
 \begin{equation}
-  \dfunctc_{\vtotal} \vchannelc(\dsectionc(\dbasepoint)) = \vchannelc(\dfunctc_{\dtotalc}(\dsectionc(\dbasepointc)))
+  \dfunctc_{\vtotalc} \vchannelc(\dsectionc(\dbasepoint)) = \vchannelc(\dfunctc_{\dtotalc}(\dsectionc(\dbasepointc)))
 \end{equation}
@@
 \end{equation}
+The compositing function is map from sheaves over $\dbase$ to sheaves over $\gbase$.
+This is because, as described in \autoref{fig:construction:xi}, the graphic section must be evaluated on all points in the graphic space to generate the visual element corresponding to a data record at a single point $\vartist(\dsection(\dbasepoint)) = \gsection(\vindexpre(\dbasepoint))$.
+
+Since encoder functions are infinitely composable, as described in \autoref{eq:construction:nu:fabrication}, a new compositor function $\vmarkc$ can be constructed by precomposing $\vchannelc$ functions with the existing $\vmarkc$.
 
@@
 \end{equation}
+The composition in \autoref{eq:construction:q:fabrication} means that different measurable components can yield the same visual elements.
+
+\begin{figure}[h!]
+  \label{fig:construction:q}
+  \includegraphics[width=1\columnwidth]{q.png}
+  \caption{This simple $\vmarkc$ assembles a circular visual element that is the color specified in $\vsection(\dbasepoint)$ and is at the intersection specified in $\vsection(\dbasepoint)$}
+\end{figure}
+As shown in \autoref{fig:construction:q}, a set of  $\vchannel$ functions individually convert the values in the data record to visual components.
+Then the $\vmarkc$ function combines these visual encodings to produce a graphic section \gsection.
+When this section is evaluated on the graphic space associated with the data $\gsection(\vindexpre(\dbasepoint))$, it produces a blue circular marker at the intersection of the x and y positions listed in \vsection.
+The composition rule in \autoref{eq:construction:q:fabrication} means that developers can implement $\vmarkc$ as drawing circles or can implement a $\vmarkc$ that draws arbitrary shapes, and then provide different $\vchannelc$ adapters, such as one that specifies that the shape is a circle.
 

### 2022-05-17 paper/main.tex
why sheaf

@@
 
-\section{Data And Visualizations}
+\section{Formally Describing Structure}
 \label{sec:atct}
@@
  The diagram in \autoref{eq:artist:natural_transform} shows that restricting a set of outputs of an artist to a set of graphic sections over a subspace is equivalent to restricting the inputs to data sections over the same subspace.
-By construction, the artist preserves the bookkeeping provided through the use of sheaves and transport functors.
-Specifically, the artist always takes as input a section over a data subspace and returns a section over the corresponding graphic subspace.
-Since these are sections of sheaves, they must be implemented such that they satisfy the sheaf axioms (\autoref{def:atct:sheaf:gluing}, \autoref{def:atct:sheaf:locality}) and therefore can be glued together and seperated as needed.
+By construction, the artist is expected to preserve the sheaf axioms of locality and gluing (\autoref{def:atct:sheaf:gluing}, \autoref{def:atct:sheaf:locality}).
+This bookkeeping is necessary for any visualization technique that selectively acts on different pieces of a data set; for example panning and zooming \cite{NekrasovskiEvaluationPanZoom2006} and streaming and animated visualizations.
 

### 2022-05-17 paper/main.tex
discussion

@@
                    numbersep=8pt, 
+                   frame=lines,
                    autogobble,
-                   frame=lines,
-                   framesep=3mm} 
+                   framesep=3mm, 
+                   breaklines=True} 
 % *** ALIGNMENT PACKAGES ***
@@
 
-\section{Formally Describing Structure}
+\section{IO: Data, Graphic}
 \label{sec:atct}
@@
 
-\begin{equation}[h]
+\begin{equation}
   \label{eq:atct:fb:intro}
@@
 By construction, the artist is expected to preserve the sheaf axioms of locality and gluing (\autoref{def:atct:sheaf:gluing}, \autoref{def:atct:sheaf:locality}).
-This bookkeeping is necessary for any visualization technique that selectively acts on different pieces of a data set; for example panning and zooming \cite{NekrasovskiEvaluationPanZoom2006} and streaming and animated visualizations.
+This bookkeeping is necessary for any visualization technique that selectively acts on different pieces of a data set; for example streaming visualizations \cite{krstajicVisualizationStreamingData2013} and panning and zooming \cite{NekrasovskiEvaluationPanZoom2006}
 
@@
 \begin{figure}[h!]
-  \label{fig:construction:flow-chart}
+  \label{fig:construction:flow}
   \includegraphics*[width=1\columnwidth]{path_of_q.png}
@@
 \end{figure}
-We propose that one way of constructing artist functions is to separate generating a visualization, as shown in \autoref{fig:construction:flow-chart}, is into an equivariant continuity preserving encoding stage $\vchannelc$ and a equivariant continuity preserving compositing stage $\vmarkc$.
+We propose that one way of constructing artist functions is to separate generating a visualization, as shown in \autoref{fig:construction:flow}, is into an equivariant continuity preserving encoding stage $\vchannelc$ and a equivariant continuity preserving compositing stage $\vmarkc$.
 In the \textcolor{artist}{encoding} stage $\vchannelc$, a data bundle is treated as separable fields and each field is mapped to a measurable visual variable.
@@
 \label{sec:construction:vtotal}
-As shown in \autoref{fig:construction:flow-chart}, we propose an intermediate visual fiber bundle 
+As shown in \autoref{fig:construction:flow}, we propose an intermediate visual fiber bundle 
 \begin{equation}
@@
 \label{sec:construction:nu}
-As introduced in \autoref{fig:construction:flow-chart}, the encoding function $\vchannel$ is a map from data sections to graphic sections
+As introduced in \autoref{fig:construction:flow}, the encoding function $\vchannel$ is a map from data sections to graphic sections
 \begin{equation}
@@
 \subsection{Composite Measurable Components Into Visual Elements}
-As shown in \autoref{fig:construction:flow-chart}, the compositor function $\vmarkc$ transforms the measurable components into properties of a visual element.
+As shown in \autoref{fig:construction:flow}, the compositor function $\vmarkc$ transforms the measurable components into properties of a visual element.
 The compositing function $\vmarkc$ transforms the sections of visual elements $\vsectionc$ into sections of graphics $\gsectionc$.
@@
 \section{Discussion: Implementing The Math}
-\subsection{Limitations}
-\subsection{future work}
+he framework specified in \autoref{sec:artist} and \autoref{sec:construction} describes how to build structure preserving visualization components, but it is left to the library developer to follow these guidelines when building and reusing components.
+In this section, we introduce a toy example of building an artist out of the components introduced in \autoref{sec:construction} to illustrate how components that adhere to these specifications are maintainable, extendible, scalable, and support concurrency.
+Specially, we introduce artists for building different types of bar graphs because it is a visualization type that allows us to demonstrate  composability and multivariate data encoding.
+We build our visualization components by extending the Python visualization library Matplotlib's artist\footnote{Matplotlib artists are our artist's namesake}\cite{hunterMatplotlib2DGraphics2007,hunterArchitectureOpenSource} to show that components using this model can be incorporated into existing visualization libraries iteratively.
+While the architecture specified in \autoref{sec:construction} can be implemented fully functionally, we make use of objects to keep track of parameters passed into artists.
+
+We construct a toy dataset with a discrete $\dbase$ of 4 points and a fiber space of $\dfiber=\{apple,\, orange,\, lemon\}\times \mathbb{Z}^{+} \times \{\texttt{True}, \texttt{False}\}$. 
+\begin{table}[h!]
+\begin{tabular}{|llrl|}
+  \hline \\
+  &   fruit &  calories &  juice \\
+  \hline
+  &   apple &        95 &   True \\ &  orange &        67 &   True \\ &   lemon &        17 &  False \\ &    lime &        20 &  False \\
+  \hline
+\end{tabular}
+\end{table}
+that we thinly wrap so that the common data interface function is that $\tau = \ttexttt{DataContainerObject.query}$. 
+\begin{minted}{python}
+  class FruitFrameWrapper:
+    def query(self, data_bounds, sampling_rate):
+      # local sections are a list of
+      # {field: local_batch_of_values}
+      return local_sections
+\end{minted}
+
+This interface provides a uniform way of  accessing subsets of the data, which are local sections.
+Because it is more efficient to work with the data in column order, we often treat the data as a collection of single fiber bundles.
+This is equivalent to the total bundle, as shown in \autoref{eq:artist:operator}.
+
+To encode the values in the dataset, we enforce equivariance by writing $\vchannel$ encoders that match the structure of our dataset.
+For example, the fruit column is a nominal measurement scale.
+Therefore we implement a position encoder that respects permutation $\dfunch$ transformations.
+The most simple form of this $\vchannel$ is a python dictionary that returns an integer position, because Matplotlib's internal parameter space expects a numerical position type.
+\begin{minted}{python}
+  def position_encoder(val):
+    return {'apple': 0, 'orange': 2, 'lemon': 4, 'lime': 6}[val]
+\end{minted}
+As mentioned in \autoref{eq:construction:nu:fabrication}, the encoders can be composed up.
+For example, the compositor $\vchannel$ may need the position to be converted to screen coordinates.
+Here the screen coordinate $\vchannel$ is a method of a Matplotlib axes object; a Matplotlib axes is akin to a container artist that holds all information about the sub artists plotted within it.
+\begin{minted}{python}
+def composite_x_transform(ax, nu):
+    return lambda x: ax.transData.transform(
+            (position_encoder(x), 0))[0]
+\end{minted}
+This encoder returns a function that is \texttt{transData.transform} $\vchannel_{transData}$ composed with the position encoder $\vchannel_{position}$ and takes as input a record to be encoded.
+As with the position encoder, the transData encoder respects permutation transforms because it returns reals; therefore the composite encoder respects permutation transforms.
+In this model, developers implement $\vchannel$ encoders that are explicit about which $\dfunc_{\vtotal}$ they support.
+Writing semantically correct encoders is also the responsiblity of the developer and is not addressed in the model.
+For example \mintinline{python}{fruit_encoder = lamda x: {'apple': green, 'orange':'yellow', 'lemon':'red', 'lime':'orange'}} is a valid color encoding with respect to permutation, but none of those colors are intuitive to the data.
+It is therefore left to the user, or domain specific library developer, to choose $\vchannel$ encoders that are appropriate for their data.
+
+After converting each record into an intermediate visual component $\vsection$, the set of visual records is passed into $\vmarkc$.
+Here the $\vmarkc$ includes one last encoder, as illustrated in \autoref{eq:construction:q:fabrication}, that assembles the independent visual components into a rectangle.
+This $\vchannel$ is inside the $\vmarkc$ to hide that library preferred format from the user.
+It is called \texttt{qhat} to indicate that this is the $\vartist^{\dbase}$ path in \autoref{eq:construction:artist:path}.
+This means that the parameters are constructed in data space $\dbase$ and this function returns a pushed forward $\vindexpush\gsection$.
+
+\begin{minted}{python}
+   def qhat(position, width, length, floor, facecolor, edgecolor, linewidth, linestyle): 
+        box = box_nu(position, width, length, floor)
+        def fake_draw(render, transform=mtransforms.IdentityTransform()):
+            for (bx, fc, ec, lw, ls) in zip(box, facecolor, edgecolor, linewidth, linestyle):
+                gc = render.new_gc()
+                gc.set_foreground((ec.r, ec.g, ec.b, ec.a))
+                gc.set_dashes(*ls)
+                gc.set_linewidth(lw)
+                render.draw_path(gc=gc, path=bx, transform=transform, rgbFace=(fc.r, fc.g, fc.b, fc.a))
+        return fake_draw
+\end{minted}
+The function \texttt{fake\_draw} is the analog of $\vindexpush\gsection$.
+This function builds the rendering spec through the renderer API, and this curried function is returned.
+The transform here is required for the code to run, but is set to identity meaning that this function directly uses the output of the position encoders.
+The curried $\texttt{fake\_draw} \approx \vindexpush \gsection$ is evaluated using a renderer object.
+In our model, as shown in \autoref{eq:artist:inout:diagram}, the renderer is supposed to take $\gsection$ as input such that $renderer(\gsection) = visualization$, but here that would require an out of scope patching of the Matplotlib render objects.
+
+The $\vchannel$ and $\vmark$ are wrapped in a container object that stores the $\vartist = \vmark \circ \vchannel$ composition and a method for computing the $\vsectionc$.
+\begin{minted}{python}
+  class Bar(Rho): 
+    def compose_with_nu(self, pfield, ffield, 
+          nu, nu_inv:):
+        # returns a new copy of the Bar artist 
+        # with the additional nu that converts 
+        # from a data (F) field value to a 
+        # visual (P) field value
+        return new
+
+    def mu(self, tau_local): #draw
+       # uses the stored nus to convert data
+       # stored nus have F->P field info
+      return mus
+
+    @staticmethod
+    def qhat(position, width, length, floor, facecolor, edgecolor, linewidth, linestyle):
+
+        return fake_draw
+\end{minted}
+
+This artist is then passed along to a shim artist that makes it compatible with existing Matplotlib objects
+\begin{minted}{python}
+class GenericArtist(martist.Artist):
+    def __init__(self, artist:TopologicalArtist):
+        super().__init__()
+        self.artist = artist
+
+    def compose_with_tau(self, section):
+        self.section = section
+
+    def draw(self, renderer, bounds, rate):
+        for tau_local in self.section.query(bounds, rate): 
+            mu = self.artist.nu(tau_local)
+            rho = self.artist.qhat(**mu)
+            output = rho(renderer)
+\end{minted}
+As shown in the \texttt{draw} method, generating a graphic section $\gsection$ is implemented as the composition of $\texttt{qhat} \approx \vmark$ and $\texttt{nu} \approx \vchannel$ applied to a local section of the sheaf $\textttt{self.section.query} \approx \desection^{i}$ such  $\texttt{draw} \approx \vmarkc\circ\vchannelc\circ\dsectionc  = \vartistc\circ \dsectionc$.
+The $\vchannel$ and $\vmark$ functions shown here are written such that they can generate a visual element given a local section $\dsection\restriction_{\dbase^{i}}$ which can be as little or large as needed.
+This flexibility is a prerequisite for building scalable and streaming visualizations that may not have access to all the data.
+
+\begin{figure}[h!]
+  \centering
+  \subfloat[]{
+    \includegraphics[width=.45\columnwidth]{bar.png}
+  \label{fig:discussion:vbar}}
+  \subfloat[]{
+    \includegraphics[width=.45\columnwidth]{barh.png}
+  \label{fig:discussion:hbar}
+  }
+  \label{fig:discussion:bar}
+\end{figure}
+
+The \texttt{GenericArtist} is a a standard Matplotlib object; therefore it can be hooked into the Matplotlib draw tree to produce the vertical bar chart in \autoref{fig:discussion:vbar}.
+Using the Matplotlib artist framework means this new artist can be composed with existing artists, such as the ones that draw the axes and ticks.
+
+One of the advantages of this model is that it allows for succinctly expressing the difference between two very similar visualizations, such as \autoref{fig:discussion:vbar} and \autoref{fig:discussion:hbar}.
+In this model, the horizontal bar is implemented as a composition of a $\vchannel$ that renames fields in $\vsection_{barh}$ and the $\vmark$ implementation for the horizontal bar.
+\begin{minted}{python}
+def qhat(length, width, position, floor, facecolor, edgecolor, linewidth, linestyle):
+  return Bar.qhat(**BarH.bar_nu(length, width, position, floor, facecolor, edgecolor, linewidth, linestyle))
+\end{minted}
+This composition is equivalent to $\vmark_{barh} = \vmark_{bar} \circ \vchannel_{vtoh}$, which is an example of \autoref{eq:construction:q:fabrication}.
+These functions can be further added together, as described in \autoref{sec:artist:operators} to build more complex visualizations.
+
+The example in this section is intentionally trivial to illustrate that the math to code translation is fairly straightforward and results in fairly self contained composable functions.
+Further research could investigate building new systems using this model, specifically libraries for visualizing domain specific structured data and domain specific artists.
+More research could also explore applying this model to visualizing high dimensional data, particularly building artists that take as input distributed data and artists that are concurrent.
+Developing complex systems could also be an avenue to codify how interactive techniques are expressed in this framework.
 

### 2022-05-17 paper/main.tex
equiv

@@
 
+\note{Functionl, Types, What is viz library architecture}
 \section{Related Work}
+Many visualization researchers have discussed structure preservation in a visualization context.
+In this work, we build on and generalize the notion that the structure of a dataset can be described as a combination of the topological properties of the dataset and the mathematical structure of the dataset fields.
 
 \subsection{Topology}
-The topological properties of visualization data are , as described by Wilkenson\cite{wilkinsonGrammarGraphics2005}.
-
+\label{sec:related-work:continuity}
 \begin{figure}[h!]
+  \label{fig:related-work:continuity:ktypes}
+  \includegraphics[width=1\columnwidth]{k_different_types.png}
+  \caption{A topological representation of the data (in purple) is a simplified representation of how the records in a dataset are connected.
+Discrete rows of a table are discrete points, a 1D continuous function is evaluated on a 1D continuous interval, and a map is 2D continuous surface.
+}
+\end{figure}
+The topological properties of a dataset are the ways in which elements are organized, as described by Wilkenson\cite{wilkinsonGrammarGraphics2005}.
+Some examples, as shown in \autoref{fig:related-work:continuity:ktypes} are discrete rows of a table, a continuous 1D line, or a surface.
+The topological space (in purple) and the data form a key-value pair, as described by Munzner \cite{munznerVisualizationAnalysisDesign2014}.
+She imbues semantic meaning to the topology, wherein the key is an independent attribute used as a look up key to retrieve the value in the dataset.
+She defines the value as a dependent attribute.
+In this work we use a more general notion of the topological space as an indexing space for the data records.
+For example the data record in the yellow row in \autoref{fig:related-work:continuity:ktypes} can be expressed as a constant function of the yellow dot; A point on the green line is mapped to by evaluating the gaussian on a point on the interval; and each pixel in the map can be found via a lookup on the 2D surface.
+This general formulation of topology gives us a language for expressing the continuity of a dataset.
+
+\begin{figure}[h!]
+  \label{fig:related-work:continuity:borked}
   \includegraphics[width=1\columnwidth]{continuity.png}
 \end{figure}
-Visualization libraries designed around continuity,  
-Visualization library algorithms assume
-
-\begin{figure}[h!]
-  \includegraphics[width=1\columnwidth]{k_different_types.png}
-\end{figure}
-Fiber bundles are a mathematical structure that can be use to encode the topological properties of visualization data, as proposed by Butler\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}.
-Broadly, fiber bundles
+The continuity matters to visualization libraries because visual algorithms, such as a the line plot and image plot in \autoref{fig:related-work:continuity:borked}, assume that the continuity of their data, as described in taxonomies of visualization algorithms by Troy, M\"{o}ller, and Chi \cite{toryRethinkingVisualizationHighlevel2004,chiTaxonomyVisualizationTechniques2000}.
+In \autoref{fig:related-work:continuity:borked}, both algorithms are fed the same input matrix.
+The line plot assumes this matrix is a list of lines and plots it as such.
+The image map assumes that the data is an image and plots it as such.
+Neither function has a way of asking the matrix what it is, nor does the matrix encode information about the continuity of the data it contains.
+While this work is not about choosing the right visualization type for the given continuity, the ability to encode the continuity expectations in the visual algorithm and the data container would provide a way of ensuring that the assumptions match.
+
+Current visualization libraries handle continuity in two main ways.
+Libraries that provide "building block" \cite{wongsuphasawatNavigatingWideWorld2021} components for building new visual algorithms embed their assumptions in their algorithms.
+This flexibility allows libraries such as matplotlib\cite{hunterMatplotlib2DGraphics2007}, VTK\cite{hanwellVisualizationToolkitVTK2015,geveciVTK2012}, and D3\cite{bostockDataDrivenDocuments2011} to support a wide variety of data types, but at the cost of the ambiguity of \autoref{fig:related-work:continuity:borked}.
+These libraries can also have visual algorithms with  parameters that are tuned to the expected data in a way that leaves the over all library feeling uncohesive.
+On the other hand, domain specific libraries are designed with the assumption of continuities that are common in the domain \cite{HeerSoftware2006}.
+Tabular data is a very common data topology and is assumed by
+A Presentation Tool\cite{mackinlayAutomatingDesignGraphical1986, mackinlayAutomatingDesignGraphical1986} and grammar of graphics\cite{wilkinsonGrammarGraphics2005} and the ggplot\cite{wickhamGgplot2ElegantGraphics2016a}, vega\cite{satyanarayanDeclarativeInteractionDesign2014}, and altair\cite{vanderplasAltairInteractiveStatistical2018} libraries built on these frameworks.
+Image libraries such as Napari\cite{nicholas_sofroniew_2021_4533308} and ImageJ\cite{schneiderNIHImageImageJ2012} and its humanaties data ImagePlot\cite{studiesCulturevisImageplot2021} plugin assumed that the input is 2D continuous.
+Networking libraries such as gephi\cite{bastianGephiOpenSource2009} and networkx\cite{HagbergExploringNetwork2008} assume a graph like structure.
+These libraries tend to benefit from more cohesive apis, but at the cost of a much more limited set of visualization algorithms than the building block libraries.
+
+To bridge the need for flexibility and cohesiveness, we build on the proposal of fiber bundles as a general abstraction model for visualization data, as proposed by Butler\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}.
+Fiber bundles are a mathematical structure that can be used to encode the topological properties of visualization data.
+There is also a methodology for encoding field types in fiber bundles, as described in Spivak's description of tables as fiber bundles\cite{spivakSIMPLICIALDATABASES}.
+Fiber bundles are discussed in more detail in \autoref{sec:atct:fiber-bundles}, but broadly they provide a mathematical representation of data that can encode topological properties, field types, and data values in a way that is both uniform and continuity and dimension independent way.
 
 \subsection{Equivariance}
+Equivariant maps are functions where variations in the input match variations in the output.
+The expectation that visualizations are equivariant maps from data to visual representation was first formalized by Bertin \cite{bertinSemiologyGraphicsDiagrams2011a}.
+He classified datasets into point, line, and surface topologies and proposed that these datasets should be represented by visual elements with the same sort of topology - points should be represented by discrete entities, lines by lines, and surfaces through area marks.
+Bertin also codified the idea that visual encoding of a data field should match the structure of the field; for example ordered data should be encoded as an ordered visual element or quantitative data as a quantitative visual element.
+This structure mapping is necessary to ensure that the visual representations in graphs are not distorting the data\cite{tufteVisualDisplayQuantitative2001} and graphs are easier to understand when properties match \cite{norman_things_smart}.
+
+The expectation of matching structure is generalized as a mapping of a binary operator that acts on the data domain to a binary operator that acts on the graphic domain by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987} An invertible binary operator is the basis of Kindlemannn and Scheideggers's algebraic model of visualization\cite{kindlmannAlgebraicProcessVisualization2014}.
+They propose three specific types of equivariance: that the visualization should not change if the data representation (i.e.
+the data container) changes, that changing the data should change the visualization in the same way in a perceptually significant manner.
+While the data representation introduced in \autoref{sec:atct:io} dispenses with a distinction between containers and the data, the definition of equivariance introduced in  \autoref{sec:artist:equivariance} includes invariant transformations and visually measurable transformations.
+
+Visualization libraries are in part measured by how expressive the components of the library are; expressiveness is a measure of which structure preserving mappings a tool can implement \cite{mackinlayAutomaticDesignGraphical1987}.
+While some visualization tools aim to automate the pairing of data with structure preserving automated tool, such as Tableau\cite{StoltePolaris2002,hanrahanVizQL2006,MackinlayShowme2007}, many visualization libraries leave that choice to the user.
+Some libraries, such as Matplotlib, set default encoders for their visual algorithms and allow the user to change the default.
+In this work, we propose guidelines for building expressive structure informed data to visual transformers.
+As with visual algorithms, we aim to provide a way to check that the structure of the data matches the structure expected by the component.
+Doing so allows users to more easily develop equivariant visual algorithms.
 
@@
 \section{Discussion: Implementing The Math}
-he framework specified in \autoref{sec:artist} and \autoref{sec:construction} describes how to build structure preserving visualization components, but it is left to the library developer to follow these guidelines when building and reusing components.
+The framework specified in \autoref{sec:artist} and \autoref{sec:construction} describes how to build structure preserving visualization components, but it is left to the library developer to follow these guidelines when building and reusing components.
 In this section, we introduce a toy example of building an artist out of the components introduced in \autoref{sec:construction} to illustrate how components that adhere to these specifications are maintainable, extendible, scalable, and support concurrency.
@@
 While the architecture specified in \autoref{sec:construction} can be implemented fully functionally, we make use of objects to keep track of parameters passed into artists.
-
-We construct a toy dataset with a discrete $\dbase$ of 4 points and a fiber space of $\dfiber=\{apple,\, orange,\, lemon\}\times \mathbb{Z}^{+} \times \{\texttt{True}, \texttt{False}\}$. 
+In this toy example, the small composable components allow for more easily verifying that each component does its transformation correctly before assembling them into larger systems.
+
 \begin{table}[h!]
+  \label{tab:discussion:data}
+  \centering
 \begin{tabular}{|llrl|}
@@
   &   fruit &  calories &  juice \\
-  \hline
+
   &   apple &        95 &   True \\ &  orange &        67 &   True \\ &   lemon &        17 &  False \\ &    lime &        20 &  False \\
@@
 \end{table}
-that we thinly wrap so that the common data interface function is that $\tau = \ttexttt{DataContainerObject.query}$. 
+We construct a toy dataset with a discrete $\dbase$ of 4 points and a fiber space of $\dfiber=\{apple,\, orange,\, lemon\}\times \mathbb{Z}^{+} \times \{\texttt{True}, \texttt{False}\}$.
+We thinly wrap \autoref{tab:discussion:data} in an object so that the common data interface function is that $\tau = \ttexttt{DataContainerObject.query}$.
 \begin{minted}{python}
@@
 This interface provides a uniform way of accessing subsets of the data, which are local sections.
+The motivation for a common data interface is that it would allow the artist to talk to different common python data containers, such as numpy\cite{harris2020array}, pandas\cite{jeff_reback_2020_3715232}, xarray \cite{hoyer2017xarray}, and networkx\cite{HagbergExploringNetwork2008}.
+Currently, data stored in these containers must be unpacked and converted into arrays and matrices in ways that either destroy or recreate the structure encoded in the container.
+For example a pandas data frame must be unpacked into its columns before it is sent into most artists and continuity is implicit in the columns being the same length rather than a tracked base space $\dbase$.
+
 Because it is more efficient to work with the data in column order, we often treat the data as a collection of single fiber bundles.
 This is equivalent to the total bundle, as shown in \autoref{eq:artist:operator}.
-
-To encode the values in the dataset, we enforce equivariance by writing $\vchannel$ encoders that match the structure of our dataset.
+To encode the values in the dataset, we enforce equivariance by writing $\vchannel$ encoders that match the structure of the fields in the dataset.
 For example, the fruit column is a nominal measurement scale.
@@
 In this model, developers implement $\vchannel$ encoders that are explicit about which $\dfunc_{\vtotal}$ they support.
-Writing semantically correct encoders is also the responsiblity of the developer and is not addressed in the model.
+Writing semantically correct encoders is also the responsibility of the developer and is not addressed in the model.
 For example \mintinline{python}{fruit_encoder = lamda x: {'apple': green, 'orange':'yellow', 'lemon':'red', 'lime':'orange'}} is a valid color encoding with respect to permutation, but none of those colors are intuitive to the data.
@@
 \begin{minted}{python}
-  class Bar(Rho): 
+  class Bar: 
     def compose_with_nu(self, pfield, ffield, 
@@
 
-    def mu(self, tau_local): #draw
+    def nu(self, tau_local): #draw
        # uses the stored nus to convert data

### 2022-05-18 paper/main.tex
sent off draft

@@
 
-\usepackage[inline]{showlabels}
+%\usepackage[inline]{showlabels}
 
@@
 \IEEEraisesectionheading{\section{Introduction}\label{sec:intro}}
-\note{please prioritize flagging content that needs major revisions: (1) Wrong (2) Missing (3) Horrific  (4) Incoherant}
-
-\IEEEPARstart{T}
-
-\note{Functionl, Types, What is viz library architecture}
+\note{please prioritize flagging content that needs major revisions: (1) Wrong (2) Missing (3) Horrific  (4) Incoherant (5) Movable to appendix - TCVG page limit is 12 pages content + supplemental}
+
+\IEEEPARstart{V}isualization software libraries consist of a set of functions for translating data into visual representations.
+These functions are expected to preserve the structure of the input data, however that structure is specified.
+Over time, library developers may need to develop new components to adapt to changing data needs, especially the growth of complex structured and distributed data.
+We propose that the complex structure of data can be described using concepts from algebraic topology and category theory, and that category theory can be used to express the constraints these components must satisfy in order to produce structure preserving data to visual transformations.
+This abstraction of the data structure and specificaton of the components can then be used to guide the developement of functional\footnote{functional programming paradigm} components.
+Since these components are functional, they are expected to be more composable and modular than existing components, have fewer side effects and therefore be easier to incorporate alignside existing library elements, and be verifiable and therefore easier to test \cite{huHowFunctionalProgramming2015,hughesWhyFunctionalProgramming1989}.
+In this paper, we introduce a mathematical framework for describing, constructing, and testing visualization library components and provide a toy example illustrating these concepts.
+
 \section{Related Work}
-Many visualization researchers have discussed structure preservation in a visualization context.
-In this work, we build on and generalize the notion that the structure of a dataset can be described as a combination of the topological properties of the dataset and the mathematical structure of the dataset fields.
+Visualization library components are required to preserve structure; in a visualization context this has widely come to be defined as a combination of the topological properties of the dataset and the mathematical structure of the dataset fields.
 
@@
 
+\note{maybe move the following to \autoref{sec:atct:io} and consolidate?}
 To bridge the need for flexibility and cohesiveness, we build on the proposal of fiber bundles as a general abstraction model for visualization data, as proposed by Butler\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}.
@@
 
-\section{IO: Data, Graphic}
+\section{Bundles and Sheaves}
 \label{sec:atct}
@@
 Category theory is a method to describe how objects are specified, constrained, and composed \cite{wielsManagementEvolvingSpecifications1998}\; therefore in this work we propose categorical constructions of visualization components as specifications for implementable components.
+A brief introduction to category theory for visualization practitioners is in Vickers et al \cite{vickersUnderstandingVisualizationFormal2013}, but they are applying category theory to semantic concerns about visualization design rather than library architecture.
 
@@
 The marker is a 2d graphic; therefore the base space that parameterizes the graphic area \dbasec\ is a 2d disc.
-This section \gsection\ describes how to build a square red marker; a section evaluated on a single point $\gbasepoint \in \gbase$ is an infinitesimally small piece of the square, here illustrated as the small piece of the square outlined in green.
-}
+This section \gsection\ describes how to build a square red marker; a section evaluated on a single point $\gbasepoint \in \gbase$ is an infinitesimally small piece of the square, here illustrated as the small piece of the square outlined in green.}
 \end{figure}
@@
 The renderer subsamples the position fibers  \textit{x} and \textit{y} of the graphic fiber space \gfiber.
-The preimage of the section at these coordinates returns a set of coordinates ${\gsection^{-1}_{x,y}} = \{\gbasepoint | \gsection(\gbasepoint)\restriction_{x,y} = \(x,y\)\}$ at the points $\gbasepoint \in \gbase$ in the graphic base topological space.
-The section is then evaluated at these points $\gsection\restriction_{\gsection^{-1}_{x,y}}$ to return a set of points in the fiber.
+The preimage of the section at these coordinates returns a set of coordinates \({\gsection^{-1}_{x,y}} = \{\gbasepoint | \gsection(\gbasepoint)\restriction_{x,y} = \(x,y\)\}\) at the points \(\gbasepoint \in \gbase\) in the graphic base topological space.
+The section is then evaluated at these points \(\gsection\restriction_{\gsection^{-1}_{x,y}}\) to return a set of points in the fiber.
 One of these points, enlarged for visibility, is illustrated as the small square swatch inside the larger marker in \autoref{fig:atct:fb}.
@@
 \begin{equation}
+  \label{eq:atct:sheaf:functor}
   \sheafc_{\dbasec, \dtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
@@
 As defined in \autoref{eq:atct:morphism:fiber} scaling is a transform $\dfunct$ that changes the value of each point but not its location.
-In \autoref{fig:atct:phi}, this means that $x_{\dbasepoint^{\prime}}= 2*x_{\dbasepoint}, \dbasepoint = \dbasepoint_{\prime}$ for all indices in \textt{data}, \textt{data.T}.
+In \autoref{fig:atct:phi}, this means that $x_{\dbasepoint^{\prime}}= 2*x_{\dbasepoint}, \dbasepoint = \dbasepoint_{\prime}$ for all indices in \texttt{data}, \texttt{data.T}.
 As defined in \autoref{eq:atct:morphism:all} and illustrated in \autoref{fig:atct:phi}, these transformations can combine to yield a matrix that is both transposed and scaled such that $x_{\dbasepoint^{\prime}} = 2*x_{\dbasepoint}$ for all
@@
 
-\subsection{Maps Between Spaces}
+\subsection{Image Functors}
 \label{sec:atct:xi}
@@
 For example, specifications such as svg\cite{quintScalable2003} and Vega\cite{satyanarayanDeclarativeInteractionDesign2014}, typically define their transformations in data space and return specifications for the render to evaluate.
-An artist may also want to specify the transformations in a proxy of graphic space, for example for operations like dynamic resampling \note{citation:root maybe}.
+An artist may also want to specify the transformations in a proxy of graphic space, for example to generate on-demand dynamically resampled continuous lines.
 
@@
 \label{sec:artist}
-In this work we propose that visualization libraries are implementing a subset of the functions in the hom sets in \autoref{eq:atct:sheaves:homset} and we call these subset of functions the artist:
+In this work we propose that visualization libraries are implementing a subset of the functions in the hom sets in \autoref{eq:atct:sheaves:homset}. We call these subset of functions the artist:
 \begin{align}
   \label{eq:artist:hom_transport}
-  \vartistc:& \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}\textcolor{artist}{\rightarrow} \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \in \textcolor{set}{Hom}_{\sheafc_{\dbase}, \sheafc_{\gbase}}\\
-  \vartistc^{\dbasec}:&\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}  \textcolor{artist}{\rightarrow} \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} \in \textcolor{set}{Hom}{\sheafc_{\dbase}}\\
-  \vartistc^{\gbasec}:& \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} \textcolor{artist}{\rightarrow} \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \in \textcolor{set}{Hom}_{\sheafc_{\gbase}}
+  \vartistc:& \cgamma{\dbasec}{\dtotalc}\textcolor{artist}{\rightarrow} \cgamma{\gbasec}{\gtotalc} \in \textcolor{set}{Hom}_{\sheafc_{\dbase}, \sheafc_{\gbase}}\\
+  \vartistc^{\dbasec}:&\cgamma{\dbasec}{\dtotalc}  \textcolor{artist}{\rightarrow} \cgamma{\dbase}{\vindexpushc\gtotalc} \in \textcolor{set}{Hom}{\sheafc_{\dbase}}\\
+  \vartistc^{\gbasec}:& \cgamma{\gbasec}{\vindexpullc\dtotalc} \textcolor{artist}{\rightarrow} \cgamma{\gbasec}{\gtotalc} \in \textcolor{set}{Hom}_{\sheafc_{\gbase}}
 \end{align}
@@
 \begin{equation}
+  \label{eq:artist:sheaf:base}
     \begin{tikzcd}
-      \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} &  & \opensetc \arrow[ll, "{\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf] \arrow[rr, "{\vindexpushc\sheafc_{\gbasec, \gtotalc}}", maps to, color=sheaf] &  & \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}
+      \cgamma{\dbasec}{\dtotalc} &  & \dbasec \arrow[ll, "{\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf] \arrow[rr, "{\vindexpushc\sheafc_{\gbasec, \gtotalc}}", maps to, color=sheaf] &  & \cgamma{\dbase}{\vindexpushc\gtotalc}
       \end{tikzcd}
@@
   \begin{tikzcd}
-    \opensetc_1 & \cgamma{\opensetc_{1}}{\dtotalc\restruction_{\opensetc_1}} \arrow[dd, "\iota^*"', color=set] \arrow[rr, "\vartistc_{\opensetc_1}", color=artist] &  & \cgamma{\opensetc_{1}}{\vindexpushc\gtotalc\restruction_{\opensetc_1}} \arrow[dd, "\iota^*", color=set] \\
+    \dbasec_1 & \cgamma{\dbasec_1}{\dtotalc} 
+    \arrow[dd, "\iota^*"', color=set] 
+    \arrow[rr, "\vartistc_{\dbase_1}", color=artist] &  & 
+    \cgamma{\dbasec_{1}}{\vindexpushc\gtotalc} 
+    \arrow[dd, "\iota^*", color=set] \\
       &  &  &  \\
-    \opensetc_{2} \arrow[uu, "\iota", hook, color=base] & \cgamma{\opensetc_{2}}{\dtotalc\restruction_{\opensetc_2}} \arrow[rr, "\vartistc_{\opensetc_2}", color=artist]                        &  & \cgamma{\opensetc_{2}}{\vindexpushc\gtotalc\restruction_{\opensetc_2}}
+    \dbasec_{2} \arrow[uu, "\iota", hook, color=base] & 
+    \cgamma{\dbasec_{2}}{\dtotalc} 
+    \arrow[rr, "\vartistc_{\dbasec_2}", color=artist] &  & 
+    \cgamma{\dbasec_{2}}{\vindexpushc\gtotalc}                      
     \end{tikzcd}
@@
 \begin{equation}
-  \begin{split}
-  \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} &\coloneqq\\ 
-  &\{\gsectionc\vert\;\exists\;\dsectionc \in \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}\;s.t.\; 
-  \vartistc(\dsectionc) = \gsectionc,\; \vindexc(\opensetgc) = \opensetc \} 
-  \end{split}
+  \label{eq:artist:output}
+  \imartist{\gbasec}{\gtotalc} \coloneqq\\ 
+  \{\gsectionc \mid\;\exists\;\dsectionc \in \cgamma{\dbasec}{\dtotalc}\;s.t.\; 
+  \vartistc(\dsectionc) = \gsectionc,\; \vindexc(\gbasec) = \dbasec \}
 \end{equation} 
@@
   \begin{tikzcd}[ampersand replacement=\&, column sep=small]
-  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
+  \cgamma{\dbasec}{\dtotalc} 
   \arrow[rrr, "\vartistc", color=artist] 
@@
   \& \& \& 
-  \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+  \imartist{\gbasec}{\gtotalc} 
   \arrow[d, "\dfuncpullc_{\gtotalc}", dotted] \\
-  \cgamma{\opensetc^{\prime}}{\dfuncpullc_{\dtotalc}\dtotalc\restriction_{\opensetc^{\prime}}} 
+  \cgamma{\dbasec^{\prime}}{\dfuncpullc_{\dtotalc}\dtotalc} 
   \arrow[dd, "\dfunctc_{\dtotalc}"', color=action] \& 
@@
   \& 
-  \imartist{\opensetgc^{\prime}}{\dfuncpullc_{\gtotalc}\gtotalc\restriction_{\opensetgc^{\prime}}} 
+  \imartist{\gbase^{\prime}}{\dfuncpullc_{\gtotalc}\gtotalc} 
   \arrow[dd, "\dfunctc_{\gtotalc}", dotted, color=action] \\
@@
   \& \\
-  \cgamma{\opensetc^{\prime}}{\dtotalc^{\prime}\restriction_{\opensetc^{\prime}}} 
+  \cgamma{\dbasec^{\prime}}{\dtotalc^{\prime}} 
   \arrow[rrr, "\vartistc^{\prime}", color=artist]  
   \& \& \& 
-  \imartist{\opensetgc^{\prime}}{\gtotalc^{\prime}\restriction_{\opensetgc^{\prime}}}
+  \imartist{\gbasec^{\prime}}{\gtotalc^{\prime}}
   \end{tikzcd}
@@
 \begin{equation}
-  \vartistc(\cgamma{\bigsqcup_{i} \dbasec^{i}}{\bigoplus_{i}\dtotalc^{i}}) \coloneqq \sum_{i}
+  \vartistc(\cgamma{\mathop{\sqcup}_{i} \dbasec^{i}}{\mathop{\oplus}_{i}\dtotalc^{i}}) \coloneqq \sum_{i}
   \vartistc_{i}(\cgamma{\dbasec^{i}}{\dtotalc^{i}}) 
@@
 As shown in \autoref{eq:artist:operator}, bundles over a union of base spaces can be joined as a product of the fibers.
-This allows us to consider all the data inputs in a complex visualization as a combined input, where some sections evaluate to null in fields for which there are no values for that point in the combined base space $\dbasepoint \in \bigsqcup_{i} \dbase^{i}$ The combined construction of the data is a method for expressing what each data input has in common with another data input-for example the data for labeling tick marks or legends-
+This allows us to consider all the data inputs in a complex visualization as a combined input, where some sections evaluate to null in fields for which there are no values for that point in the combined base space $\dbasepoint \in \mathop{\sqcup}_{i} \dbase^{i}$ The combined construction of the data is a method for expressing what each data input has in common with another data input-for example the data for labeling tick marks or legends-
 and therefore which commonalities need to be preserved in the artists that act on these inputs.
@@
 \begin{equation}
-  \underbrace{\opensetc\times[0,1]^{n}}_{\opensetgc} \textcolor{functor}{\xmapsto{\hspace{1em}\vindexc\hspace{1em}}} \opensetc 
+  \underbrace{\dbasec\times[0,1]^{n}}_{\gbasec} \textcolor{functor}{\xmapsto{\hspace{1em}\vindexc\hspace{1em}}} \dbasec
 \end{equation}
@@
 \begin{equation}
-\nu: \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \rightarrow \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}}
+\nu: \cgamma{\dbasec}{\dtotalc} \rightarrow \cgamma{\dbasec}{\vtotalc}
 \end{equation}
@@
 \begin{equation}
-  \vmarkc: \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} \rightarrow \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}
+  \vmarkc: \cgamma{\dbasec}{\vtotalc} \rightarrow \cgamma{\gbasec}{\gtotalc}
 \end{equation}
@@
   \begin{tikzcd}
-      \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} 
+      \cgamma{\dbasec}{\vtotalc} 
       \arrow[rr, "\vchannel", color=artist] 
-      \arrow[rrrr, "\vmarkc^{\prime}"', bend right, color=artist, dashed] &  & \cgamma{\opensetc}{\vtotalc^{\prime}\restriction_{\opensetc}} 
-      \arrow[rr, "\vmarkc", color=artist] &  & \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}
+      \arrow[rrrr, "\vmarkc^{\prime}", bend right, color=artist, dashed] &  & \cgamma{\dbasec}{\vtotalc^{\prime}} 
+      \arrow[rr, "\vmarkc", color=artist] &  & \cgamma{\gbasec}{\gtotalc}
       \end{tikzcd} 
@@
   \begin{tikzcd}[row sep=huge]
-    \cgamma{\opensetc}{(\vtotalc^{a}\times \vtotalc^{b})\restriction_{\opensetc}}  
-    \arrow[rr, "\vmarkc_{ab}", color=artist] \arrow[d, "\pi_a"', color=total] &  &  \imartistsub{ab}{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+    \cgamma{\dbasec}{\vtotalc^{a}\times \vtotalc^{b}}  
+    \arrow[rr, "\vmarkc_{ab}", color=artist] 
+    \arrow[d, "\pi_a"', color=total] &  &  \imartistsub{ab}{\gbasec}{\gtotalc} 
     \arrow[d, "\measurec\restriction_a \circ \extractmc_{ab}", color=monoid]  \\
-   \cgamma{\opensetc}{\vtotalc^{a}\restriction_{\opensetc}} 
-   \arrow[rr, "\simeq", dotted] &  & {\textcolor{set}{Hom}(\opensetc, \measurec^{a})}                                                                             \\
-    \cgamma{\opensetc}{(\vtotalc^{a}\times \vtotalc^{c})\restriction_{\opensetc}}  \arrow[rr, "\vmarkc_{ac}", color=artist] \arrow[u, "\pi_a", color=total]  &  &  \imartistsub{ac}{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \arrow[u, "\measurec\restriction_a \circ \extractmc_{ac}"', color=monoid]
+   \cgamma{\dbasec}{\vtotalc^{a}} 
+   \arrow[rr, "\simeq", dotted] &  & 
+   {\textcolor{set}{Hom}(\dbasec, \measurec^{a})}  \\
+    \cgamma{\dbasec}{\vtotalc^{a}\times \vtotalc^{c}}  
+    \arrow[rr, "\vmarkc_{ac}", color=artist] 
+    \arrow[u, "\pi_a", color=total]  &  &  \imartistsub{ac}{\gbasec}{\gtotalc} 
+    \arrow[u, "\measurec\restriction_a \circ \extractmc_{ac}"', color=monoid]
    \end{tikzcd}
@@
 \begin{equation}
+  \label{eq:construnction:nu:verify:base}
   \dfunctc_{\vtotalc}(\vsectionc(\dbasepointc)) = \dfunctc_{\gtotalc}(\vmarkc(\vsectionc(\dbasepointc)))  
@@
 \begin{tikzcd}[row sep=2.5em, column sep=1.5em]
-  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
+  \cgamma{\dbasec}{\dtotalc} 
   \arrow[rr, "\vchannelc^{\dbasec}", color=artist] 
@@
   \arrow[rrrrdd, "\vartistc", color=artist, pos=.2] &  & 
-  \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} 
+  \cgamma{\dbasec}{\vtotalc} 
   \arrow[rrdd, "\vmarkc", color=artist] 
   \arrow[rr, "\vmarkc^{\dbasec}", color=artist] 
-  \arrow[dd, "\vindexpullc", color=functor, pos=.2] &  & \imartist{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}  \\
+  \arrow[dd, "\vindexpullc", color=functor, pos=.2] &  & \imartist{\dbasec}{\vindexpushc\gtotalc}  \\
    & & & & \\
-  \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} \arrow[rr, "\vchannelc^{\gbasec}", color=artist] 
+  \cgamma{\gbasec}{\vindexpullc\dtotalc} 
+  \arrow[rr, "\vchannelc^{\gbasec}", color=artist] 
   \arrow[rrrr, "\vartistc^{\gbasec}", bend right, color=artist] & & 
-  \cgamma{\opensetgc}{\vindexpullc\vtotalc\restriction_{\opensetgc}} \arrow[rr, "\vmarkc^{\gbasec}", color=artist] &  & 
-  \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \arrow[uu, "\vindexpushc"', color=functor]
+  \cgamma{\gbasec}{\vindexpullc\vtotalc} 
+  \arrow[rr, "\vmarkc^{\gbasec}", color=artist] &  & 
+  \imartist{\gbasec}{\gtotalc} 
+  \arrow[uu, "\vindexpushc"', color=functor]
 \end{tikzcd}  
@@
 \section{Discussion: Implementing The Math}
+\label{sec:discussion}
 The framework specified in \autoref{sec:artist} and \autoref{sec:construction} describes how to build structure preserving visualization components, but it is left to the library developer to follow these guidelines when building and reusing components.
@@
 \begin{table}[h!]
-  \label{tab:discussion:data}
   \centering
-\begin{tabular}{|llrl|}
+\begin{tabular}{|lcl|}
   \hline \\
-  &   fruit &  calories &  juice \\
-
-  &   apple &        95 &   True \\ &  orange &        67 &   True \\ &   lemon &        17 &  False \\ &    lime &        20 &  False \\
+   fruit &  calories &  juice \\
+  \hline\\
+    apple &        95 &   True \\ 
+   orange &        67 &   True \\ 
+  lemon &        17 &  False \\ 
+      lime &        20 &  False \\
   \hline
 \end{tabular}
+\label{tab:discussion:data}
 \end{table}
@@
 \section{Conclusion}
-The conclusion goes here.
+The toy example presented in \autoref{sec:discussion} demonstrates that it is relatively straightforward to build working visualization library components using the construction described in \autoref{sec:construction}.
+Since these components are defined with single record inputs, the can be implemented such that they are concurrent.
+The cost of building a new function using these components is sometimes as small as renaming fields, meaning the new feature is relatively easy to maintain.
+These new components are also a lower maintenance burden because, by definition, they are designed in conjunction with tests that verify that they are equivariant.
+These new components are also compatible with the existing library architecture, allowing for a slow iterative transition to components built using this framework. 
+The framework introduced in this paper is a marriage of the ways the graphic and data visualization communities approach visualization.
+The graphic community prioritizes \note{?} how input is translated to output, which is encapsulated in the artist $\vartist$.
+The data visualization community prioritizes the manner in which that input is encoded, which is encapsulated in the separation of stages $\vmark \circ \vchannel$.
+Formalizing that both views are equivalent $\vartist=\vmark \circ \vchannel$ gives library developers the flexibility to build visualization components in the manner that makes more sense for the domain without having to sacrifice the equivariance of the translation.
 

### 2022-05-26 paper/main.tex
section renaming + reworking of bundles

@@
 We propose that the complex structure of data can be described using concepts from algebraic topology and category theory, and that category theory can be used to express the constraints these components must satisfy in order to produce structure preserving data to visual transformations.
-This abstraction of the data structure and specificaton of the components can then be used to guide the developement of functional\footnote{functional programming paradigm} components.
+This abstraction of the data structure and specification of the components can then be used to guide the developement of functional\footnote{functional programming paradigm} components.
 Since these components are functional, they are expected to be more composable and modular than existing components, have fewer side effects and therefore be easier to incorporate alignside existing library elements, and be verifiable and therefore easier to test \cite{huHowFunctionalProgramming2015,hughesWhyFunctionalProgramming1989}.
@@
   \caption{A topological representation of the data (in purple) is a simplified representation of how the records in a dataset are connected.
-Discrete rows of a table are discrete points, a 1D continuous function is evaluated on a 1D continuous interval, and a map is 2D continuous surface.
-}
+For example, a 1D continuous function is evaluated on a 1D continuous interval, and a map is 2D continuous surface.
+A network is a union of discrete nodes and continuous lines.
+The nodes represent the continuity of a table of discrete weather station data and the graph edges are encoded with the distances between the nodes.}
 \end{figure}
@@
 In this work we use a more general notion of the topological space as an indexing space for the data records.
-For example the data record in the yellow row in \autoref{fig:related-work:continuity:ktypes} can be expressed as a constant function of the yellow dot; A point on the green line is mapped to by evaluating the gaussian on a point on the interval; and each pixel in the map can be found via a lookup on the 2D surface.
+For example  in \autoref{fig:related-work:continuity:ktypes}, a point on the green line is mapped to by evaluating the gaussian on a point on the interval and each pixel in the map can be found via a lookup on the 2D surface.
 This general formulation of topology gives us a language for expressing the continuity of a dataset.
+The network in \autoref{fig:related-work:continuity:ktypes} is a union of a 0D discrete points and 1D lines connecting those points.
+The data record in the yellow row can be expressed as a constant function of the yellow network node, while the data on the graph edge between any two points is a function of the distance between those points.
 
@@
   \includegraphics[width=1\columnwidth]{continuity.png}
+  \caption{The \texttt{line} and \texttt{image} visual algorithms take in the same array of data.
+The \texttt{line} assumes that the array is a list of lines, while the \texttt{image} assumes that the array is an image.}
 \end{figure}
-The continuity matters to visualization libraries because visual algorithms, such as a the line plot and image plot in \autoref{fig:related-work:continuity:borked}, assume that the continuity of their data, as described in taxonomies of visualization algorithms by Troy, M\"{o}ller, and Chi \cite{toryRethinkingVisualizationHighlevel2004,chiTaxonomyVisualizationTechniques2000}.
-In \autoref{fig:related-work:continuity:borked}, both algorithms are fed the same input matrix.
+The continuity matters to visualization libraries because visual algorithms, such as a the line plot and image plot in \autoref{fig:related-work:continuity:borked}, assume the continuity of their data, as described in taxonomies of visualization algorithms by Troy, M\"{o}ller, and Chi \cite{toryRethinkingVisualizationHighlevel2004,chiTaxonomyVisualizationTechniques2000}.
+In \autoref{fig:related-work:continuity:borked}, both \texttt{line} and \textt{image} algorithms are fed the same input matrix.
 The line plot assumes this matrix is a list of lines and plots it as such.
@@
 
-\section{Bundles and Sheaves}
+\section{Formal Properties of Data \& Graphics}
 \label{sec:atct}
@@
 
-\subsection{Fiber Bundles}
+\subsection{Uniform Abstract Data Representation}
 \label{sec:atct:fiber-bundles}
@@
 
+\begin{figure}[h!]
+  \includegraphics[width=\columnwidth]{fbtau.png}
+  \caption{A non-trivial bundle provides a way of expressing data where the fiber changes along the indexing base space.
+Here $\dfiber_{\dbasepoint_1}$ and $\dfiber_{\dbasepoint_2}$ go in reverse directions \note{do fibers have to be isomorphic?}.
+The global section $sin$ returns $sin(\dbasepoint_{i})$ which is a value between $[0,1]$
+  \label{fig:atct:stripbundle}}
+\end{figure}
+
+The bundle in \autoref{fig:atct:stripbundle} is non-trivial, meaning that the fiber is not identical over every point in the base space $\dbase$.
+One section over this 1D continuous space is a sin function that takes as input a point in the base space $\dbase$ and returns the sin of the point.
+
 A fiber bundle is locally trivial, which means that for every point \dbasepointc\ there exists an open neighborhood $\dbasepointc \in \opensetc \subseteq \dbasec$ such that there is a homeomorphism $\pi^{-1}(\opensetc)\xrightarrow{\equivc} \opensetc \times \dfiberc$.
@@
 \begin{figure}[h!]
-  \label{fig:atct:fb}
-  \includegraphics[width=\columnwidth]{fb_tau.png}
-  \caption{This trivial dataset consists of a single point $\dbasec$ and the fields \texttt{speed, direction, signal, and strength}.
-Each of these fields has an associated datatype, which is the set of possible valid values for that field.
-The fiber space is the cartesian product of these sets $\dfiberc$.
-Each section $\dsectionc$ function returns a single record in the fiber, here the data record outlined in green.
-  \note{maybe change to temperature}}
+  \includegraphics[width=\columnwidth]{fbnet.png}
+  \caption{The fibers over the discrete points $\dbase_{1}, \dbase_{2}$  encode the fields in the temperature table and the fiber over the continuous line $\dbase_{3}$ encodes the fields in the distance table \autoref{fig:related-work:continuity:ktypes}.
+Those tables can be modeled as sections of the bundle in this figure, where the section evaluated at a point in the base space returns the associated row, and on the edge is a constant function that returns the distance.
+  \label{fig:atct:networkbundle}}
 \end{figure}
+
+The bundle in \autoref{fig:atct:networkbundle} is constructed from a union of bundles over discrete points $\dbase_{1}, \dbase_{3}$ and a 1D continuous connecting line $\dbase_{2}$ that connects the two points.
+
 A fiber bundle is a mathematical structure for encoding continuity, type, and datasets with that continuity and type.
@@
 
-\subsection{Input and Output Types}
+\subsection{Encoding Topology and Field Types}
 \label{sec:atct:io}
@@
 
-\subsection{Sheaves}
+\subsection{Formal Properties of Data Containers}
 \label{sec:atct:sheaves}
@@
 
-\subsubsection{Structure}
+\subsubsection{Mathematical Structure of Data}
 \label{sec:atct:sheaves:measurments}
@@
 
-\subsection{Image Functors}
+\subsection{Associations Between Data and Graphics}
 \label{sec:atct:xi}
@@
 
-\subsection{Defining Maps Between Spaces $\vindex$}
+\subsection{Map Between Graphics and Data}
 In \autoref{sec:atct:sheaves}, the functors $\vindex, \vindexpull, \vindexpushc$ are introduced to describe the expected relationship between screen and data base spaces.
@@
 
-\subsection{Composite Measurable Components Into Visual Elements}
+\subsection{Assemble Components Into a Graphic}
 As shown in \autoref{fig:construction:flow}, the compositor function $\vmarkc$ transforms the measurable components into properties of a visual element.
@@
 
-\subsection{Constructing the Artist: $\vartistc=\vmarkc\circ\vchannelc$}
+\subsection{Implementing the Artist: $\vartistc=\vmarkc\circ\vchannelc$}
 
@@
 
-\section{Discussion: Implementing The Math}
+\section{Discussion: Translate Spec to Design}
 \label{sec:discussion}
@@
 In this section, we introduce a toy example of building an artist out of the components introduced in \autoref{sec:construction} to illustrate how components that adhere to these specifications are maintainable, extendible, scalable, and support concurrency.
-Specially, we introduce artists for building different types of bar graphs because it is a visualization type that allows us to demonstrate composability and multivariate data encoding.
+
+\begin{figure}[h!]
+  \centering
+  \subfloat[]{
+    \includegraphics[width=.45\columnwidth]{bar.png}
+  \label{fig:discussion:vbar}}
+  \subfloat[]{
+    \includegraphics[width=.45\columnwidth]{barh.png}
+  \label{fig:discussion:hbar}
+  }
+  \label{fig:discussion:bar}
+\end{figure}
+
+Specially, we introduce artists for building the graphical elements shown in \autoref{fig:discussion:bar} because it is a visualization type that allows us to demonstrate composability and multivariate data encoding.
 We build our visualization components by extending the Python visualization library Matplotlib's artist\footnote{Matplotlib artists are our artist's namesake}\cite{hunterMatplotlib2DGraphics2007,hunterArchitectureOpenSource} to show that components using this model can be incorporated into existing visualization libraries iteratively.
@@
 
+\subsection{Bundle Inspired Data Containers}
 \begin{table}[h!]
@@
 For example a pandas data frame must be unpacked into its columns before it is sent into most artists and continuity is implicit in the columns being the same length rather than a tracked base space $\dbase$.
-
 Because it is more efficient to work with the data in column order, we often treat the data as a collection of single fiber bundles.
 This is equivalent to the total bundle, as shown in \autoref{eq:artist:operator}.
+
+\subsection{Component Encoders}
 To encode the values in the dataset, we enforce equivariance by writing $\vchannel$ encoders that match the structure of the fields in the dataset.
@@
 
+\subsection{Graphic Compositors}
 After converting each record into an intermediate visual component $\vsection$, the set of visual records is passed into $\vmarkc$.
@@
 
+One of the advantages of this model is that it allows for succinctly expressing the difference between two very similar visualizations, such as \autoref{fig:discussion:vbar} and \autoref{fig:discussion:hbar}.
+In this model, the horizontal bar is implemented as a composition of a $\vchannel$ that renames fields in $\vsection_{barh}$ and the $\vmark$ implementation for the horizontal bar.
+\begin{minted}{python}
+def qhat(length, width, position, floor, facecolor, edgecolor, linewidth, linestyle):
+  return Bar.qhat(**BarH.bar_nu(length, width, position, floor, facecolor, edgecolor, linewidth, linestyle))
+\end{minted}
+This composition is equivalent to $\vmark_{barh} = \vmark_{bar} \circ \vchannel_{vtoh}$, which is an example of \autoref{eq:construction:q:fabrication}.
+These functions can be further added together, as described in \autoref{sec:artist:operators} to build more complex visualizations.
+
+\subsection{Integrating Components Into Existing Library}
 The $\vchannel$ and $\vmark$ are wrapped in a container object that stores the $\vartist = \vmark \circ \vchannel$ composition and a method for computing the $\vsectionc$.
@@
 
-\begin{figure}[h!]
-  \centering
-  \subfloat[]{
-    \includegraphics[width=.45\columnwidth]{bar.png}
-  \label{fig:discussion:vbar}}
-  \subfloat[]{
-    \includegraphics[width=.45\columnwidth]{barh.png}
-  \label{fig:discussion:hbar}
-  }
-  \label{fig:discussion:bar}
-\end{figure}
-
 The \texttt{GenericArtist} is a a standard Matplotlib object; therefore it can be hooked into the Matplotlib draw tree to produce the vertical bar chart in \autoref{fig:discussion:vbar}.
 Using the Matplotlib artist framework means this new artist can be composed with existing artists, such as the ones that draw the axes and ticks.
-
-One of the advantages of this model is that it allows for succinctly expressing the difference between two very similar visualizations, such as \autoref{fig:discussion:vbar} and \autoref{fig:discussion:hbar}.
-In this model, the horizontal bar is implemented as a composition of a $\vchannel$ that renames fields in $\vsection_{barh}$ and the $\vmark$ implementation for the horizontal bar.
-\begin{minted}{python}
-def qhat(length, width, position, floor, facecolor, edgecolor, linewidth, linestyle):
-  return Bar.qhat(**BarH.bar_nu(length, width, position, floor, facecolor, edgecolor, linewidth, linestyle))
-\end{minted}
-This composition is equivalent to $\vmark_{barh} = \vmark_{bar} \circ \vchannel_{vtoh}$, which is an example of \autoref{eq:construction:q:fabrication}.
-These functions can be further added together, as described in \autoref{sec:artist:operators} to build more complex visualizations.
 

### 2022-05-27 paper/main.tex
#19

@@
 
-%\usepackage[inline]{showlabels}
+\usepackage[inline]{showlabels}
 
@@
 \begin{figure}[h!]
-  \label{fig:related-work:continuity:ktypes}
   \includegraphics[width=1\columnwidth]{k_different_types.png}
@@
 A network is a union of discrete nodes and continuous lines.
-The nodes represent the continuity of a table of discrete weather station data and the graph edges are encoded with the distances between the nodes.}
+The nodes represent the continuity of a table of discrete weather station data and the graph edges are encoded with the distances between the nodes.
+\label{fig:related-work:continuity:ktypes}}
 \end{figure}
@@
 \begin{figure}[h!]
-  \label{fig:related-work:continuity:borked}
   \includegraphics[width=1\columnwidth]{continuity.png}
   \caption{The \texttt{line} and \texttt{image} visual algorithms take in the same array of data.
-The \texttt{line} assumes that the array is a list of lines, while the \texttt{image} assumes that the array is an image.}
+The \texttt{line} assumes that the array is a list of lines, while the \texttt{image} assumes that the array is an image.
+\label{fig:related-work:continuity:borked}}
 \end{figure}
@@
   \includegraphics[width=\columnwidth]{fbtau.png}
-  \caption{A non-trivial bundle provides a way of expressing data where the fiber changes along the indexing base space.
-Here $\dfiber_{\dbasepoint_1}$ and $\dfiber_{\dbasepoint_2}$ go in reverse directions \note{do fibers have to be isomorphic?}.
-The global section $sin$ returns $sin(\dbasepoint_{i})$ which is a value between $[0,1]$
+  \caption{A non-trivial fiber bundle is bundle where the (by definition isomorphic) fibers change along the indexing base space.
+Here $\dfiber_{\dbasepoint_1}$ and $\dfiber_{\dbasepoint_2}$ go in reverse directions and the global section $sin$ returns $sin(\dbasepoint_{i})$ which is a value between $[0,1]$
   \label{fig:atct:stripbundle}}
@@
 
-We also use fiber bundles to represent the output of a visualization algorithm, which we term the graphic but generalizes to output on any display space, such as a screen or 3D print. 
+\subsection{Uniform Abstract Graphic Representation}
+Fiber bundles are such a generalizable data abstraction that we can also use them to encode the output of a visual bundle; bundles allow for fully expressing the properties of the graphic that must be symmetric to the data in an equivariant (\autoref{sec:related-work:equivariance}) data to graphic transformation.
+We denote the output as a graphic, but the use of bundles allows us to generalize to output on any display space, such as a screen or 3D print.
 \begin{equation}
@@
 \end{equation}
+\begin{equation}
+  \label{eq:atct:fb_graphic_section}
+  \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \coloneqq \big\{\gsectionc: \opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} \; \bigm{\vert} \pi(\gsectionc(\gbasepointc)) = \gbasepointc\;for\, all\; \gbasepointc \in \opensetgc \big\}
+\end{equation}
 The base space \gbasec\ is a parameterization of the display area, for example the inked bounding box in cairo \cite{CairographicsOrg}.
 The fiber space \gfiberc\ is an abstraction of the renderer fields, for example $\{x,\,y,\,r,\,g,\,b,\,a\}$.
-The sections of the graphic
-\begin{equation}
-  \label{eq:atct:fb_graphic_section}
-  \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \coloneqq \big\{\gsectionc: \opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} \; \bigm{\vert} \pi(\gsectionc(\gbasepointc)) = \gbasepointc\;for\, all\; \gbasepointc \in \opensetgc \big\}
-\end{equation}
-are functions that generate different graphics, for example markers for a scatter plot, bars for a bar chart, cells of a heatmap, or pixels of an image.
+The sections $\gsection$ of the graphic are functions that go from a point in the graphic base to elements in a prerendering space.
 
 \begin{figure}[h!]
-  \label{fig:atct:fb:graphic}
   \includegraphics[width=1\columnwidth]{fb_rho.png}
-  \caption{A bundle representation of a section that generates a scatter marker on a 2D screen.
-The fiber space encodes the potential color and position of each pixel in a prerender space.
-The marker is a 2d graphic; therefore the base space that parameterizes the graphic area \dbasec\ is a 2d disc.
-This section \gsection\ describes how to build a square red marker; a section evaluated on a single point $\gbasepoint \in \gbase$ is an infinitesimally small piece of the square, here illustrated as the small piece of the square outlined in green.}
+  \caption{For a 2D display a section $\gsection$ maps into a fiber $\dfiber$ that encodes an RGB prerender space with infinite resolution $\mathcal{R}^{2}$.
+Each tiny colored box is an approximation of the return value of the same section function $\gsection$ evaluated on different points $\gbasepoint \in \gbase$ in the base space.
+\label{fig:atct:fb:graphic}}
 \end{figure}
 
-In \autoref{fig:atct:fb:graphic}, the section function $\gsection$ encodes the drawing of a red square marker.
-The renderer subsamples the position fibers  \textit{x} and \textit{y} of the graphic fiber space \gfiber.
-The preimage of the section at these coordinates returns a set of coordinates \({\gsection^{-1}_{x,y}} = \{\gbasepoint | \gsection(\gbasepoint)\restriction_{x,y} = \(x,y\)\}\) at the points \(\gbasepoint \in \gbase\) in the graphic base topological space.
-The section is then evaluated at these points \(\gsection\restriction_{\gsection^{-1}_{x,y}}\) to return a set of points in the fiber.
-One of these points, enlarged for visibility, is illustrated as the small square swatch inside the larger marker in \autoref{fig:atct:fb}.
-The renderer is handed some version of $\gsection$ to produce the final graphic.
-
-\subsection{Encoding Topology and Field Types}
+In \autoref{fig:atct:fb:graphic}, the section function $\gsection$ maps into the fiber for a simplified 2D RGB infinite resolution prerender space and returns the $\{x,y,r,g,b\}$ values of a pixel in an infinite resolution space.
+In \autoref{fig:atct:fb:graphic} these pixels are approximated as the small orange and black colored boxes.
+Each box is the output of the $\gsection(\gbasepoint)$ section that intersects the box.
+The set of all pixels returned by a section evaluated on a given visual base space $\gsection|_{\gbase}$ can yield a visual element, such as a marker, line, or piece of a glyph.
+
+\subsection{Topology and Field Categories}
 \label{sec:atct:io}
-As described in \autoref{sec:atct:fiber-bundles}, we model data as a map from base space to fiber space.To encapsulate the structure carried in these maps, and provide a way to define these maps in reusable and composable manner, we define categories of base and fiber spaces.
+We define categorical types for the topology and the fibers because categories are defined both by the objects of the category and the valid morphisms on the object.
+This allows for the formal construction of an equivariance aware type that holds the topology, field, type, and what changes to either must be preserved.
 
@@
   \label{sec:atct:io:base}
-  The base space category $\mathcal{\opensetc}$ has open set objects $\opensetc$ and inclusion morphisms $\iota: \opensetc_1 \rightarrow \opensetc_2$ such that $\opensetc_1 \subseteq \opensetc_2$. 
+  The base space category $\mathcal{\dbasec}$ has open set objects $\opensetc$ and inclusion morphisms $\iota: \opensetc_1 \rightarrow \opensetc_2$ such that $\opensetc_1 \subseteq \opensetc_2$. 
 \end{definition}
@@
 
+The conditions on sheaves express the presheaf properties of continuity preservation as applied to sections, which as mentioned in \autoref{sec:atct:fiber-bundles} are how data is modeled in this framework.
 A sheaf is a presheaf that satisfies the locality and gluing axioms \cite{bakerMathsSheaf}.
@@
 \begin{figure}[h!]
-  \label{fig:atct:phi}
   \includegraphics[width=\columnwidth]{phi.png}
   \caption{Values in a data set can be transformed in three ways: $\dfunch$-values can change position, .e.g transposed;  $\dfunct$-values can change, e.g.
-doubled; $\dfunc$ - values can change position and value}
+doubled; $\dfunc$ - values can change position and value  \label{fig:atct:phi}}
 \end{figure}
@@
 \begin{figure}[h!]
-  \label{fig:atct:morphisms:sheaf}
   \includegraphics*[width=1\columnwidth]{xi_scatter.png}
@@
 Each graphic disc has a corresponding graphic function $\gsectionc$.
-The pushforward function $\vindexpushc$ converts $\gsectionc$ into the visual parameters associated with each data point $\dbasepoint$.}
+The pushforward function $\vindexpushc$ converts $\gsectionc$ into the visual parameters associated with each data point $\dbasepoint$.\label{fig:atct:morphisms:sheaf}}
 \end{figure}
@@
   \label{eq:atct:sheaves:homset:functors}
-\begin{split}
-  & \textcolor{set}{Hom}_{\sheafc_{\gbasec}}(\vindexpullc\sheafc_{\dtotalc, \dbasec},\sheafc_{\gtotalc,\gbasec})\\
-  \simeq & \textcolor{set}{Hom}_{\sheafc_{\dbasec}}(\sheafc_{\dtotalc, \dbasec},\vindexpushc\sheafc_{\gtotalc,\gbasec}) \\
-  \simeq & \textcolor{set}{Hom}_{\sheafc_{\dbasec},\sheafc_{\gbasec}}(\sheafc_{\dtotalc, \dbasec},\sheafc_{\gtotalc,\gbasec}) \\
-\end{split} 
-\end{equation}
-which means that the diagram of sheaf morphisms \autoref{eq:atct:sheaves:homset} is commutative; therefore the functors $\vindexpullc$ and $\vindexpushc$ can be used to adapt morphisms written over one space to another space.
+   \textcolor{set}{Hom}_{\sheafc_{\gbasec}}(\vindexpullc\sheafc_{\dtotalc, \dbasec},\sheafc_{\gtotalc,\gbasec})
+  \simeq  \textcolor{set}{Hom}_{\sheafc_{\dbasec}}(\sheafc_{\dtotalc, \dbasec},\vindexpushc\sheafc_{\gtotalc,\gbasec}) 
+\end{equation}
+such that both are isomorphic to  $\simeq  \textcolor{set}{Hom}_{\sheafc_{\dbasec},\sheafc_{\gbasec}}(\sheafc_{\dtotalc, \dbasec},\sheafc_{\gtotalc,\gbasec})$.
+This set of ismorphic homsets means that the diagram of sheaf morphisms \autoref{eq:atct:sheaves:homset} is commutative; therefore the functors $\vindexpullc$ and $\vindexpushc$ can be used to adapt morphisms written over one space to another space.
+\note{maybe move to construction}
 This provides us flexibility over which space to construct data to graphic transforms.
@@
   \label{eq:artist:hom_transport}
-  \vartistc:& \cgamma{\dbasec}{\dtotalc}\textcolor{artist}{\rightarrow} \cgamma{\gbasec}{\gtotalc} \in \textcolor{set}{Hom}_{\sheafc_{\dbase}, \sheafc_{\gbase}}\\
-  \vartistc^{\dbasec}:&\cgamma{\dbasec}{\dtotalc}  \textcolor{artist}{\rightarrow} \cgamma{\dbase}{\vindexpushc\gtotalc} \in \textcolor{set}{Hom}{\sheafc_{\dbase}}\\
-  \vartistc^{\gbasec}:& \cgamma{\gbasec}{\vindexpullc\dtotalc} \textcolor{artist}{\rightarrow} \cgamma{\gbasec}{\gtotalc} \in \textcolor{set}{Hom}_{\sheafc_{\gbase}}
+  \vartistc:& \cgamma{\dbasec}{\dtotalc}\textcolor{artist}{\rightarrow} \cgamma{\gbasec}{\gtotalc} \in \textcolor{set}{Hom}_{\sheafc_{\dbase}, \sheafc_{\gbase}}
 \end{align}
- Because the artists can be constructed as morphisms of sheave over the same base spaces, as listed in \autoref{eq:atct:sheaves:homset:functors}, they are natural transformations
-\begin{align}
-  \label{eq:artist:natural_transform}
-  \vartistc^{\dbasec}:& \sheafc_{\dbasec, \dtotalc} \textcolor{artist}{\Rightarrow} \vindexpushc \sheafc_{\gbasec, \gtotalc}\\
-  \vartistc^{\gbasec}:&\vindexpullc\sheafc_{\dbasec, \dtotalc} \textcolor{artist}{\Rightarrow}\sheafc_{\gbasec, \gtotalc}
-\end{align}
-which means that they are maps of functors that take the same input object and return objects in the same category\cite{milewskiCategoryTheoryProgrammers}.
+ Because the artists can be constructed as morphisms of sheave over the same base spaces, as listed in \autoref{eq:atct:sheaves:homset:functors}, they are natural transformations, which means that they are maps of functors that take the same input object and return objects in the same category\cite{milewskiCategoryTheoryProgrammers}.
 As illustrated in \autoref{eq:atct:sheaves:homset}, the sheaf functors
@@
   \cgamma{\dbasec^{\prime}}{\dtotalc^{\prime}} 
-  \arrow[rrr, "\vartistc^{\prime}", color=artist]  
+  \arrow[rrr, "\vartistc", color=artist]  
   \& \& \& 
@@
 \begin{equation*}
-  \vartistc^{\prime}(\dfunctc_{\dtotalc}(\dsectionc(\dfunchc_{\dtotalc}(\vindexc(\gbasepointc^{\prime}))))) = \dfunctc_{\gtotalc}(\vartistc(\dsectionc(\vindexc(\dfunchc_{\gtotalc}(\gbasepointc^{\prime})))))
+  \vartistc(\dfunctc_{\dtotalc}(\dsectionc(\dfunchc_{\dtotalc}(\vindexc(\gbasepointc^{\prime}))))) = \dfunctc_{\gtotalc}(\vartistc(\dsectionc(\vindexc(\dfunchc_{\gtotalc}(\gbasepointc^{\prime})))))
 \end{equation*}
@@
 The function $\dfunch_{\gtotal}$ induces a change in graphic generating function that matches the change in data.
-
-It is difficult to formally define or implement $\dfunch_{\gtotal}$ because the equivalent changes in the graphic, such as a scaling or rotation, would have to be evaluated solely on the basis of the fiber elements $\gsection(\dbasepoint)$ which are tuples of output space specifications.
-For example, in a 2D screen, evaluating that a scatter marker respects a change in scaling would mean translating a table of $\{x,y,r,g,b\}$ values into measurements \measure\ and then checking how those measurements change.
-This is equivalent to measuring components of the rendered image, as shown in the diagram
-
+The graphic transformation $\dfunch_{\gtotal}$ is difficult to define because by definition it acts on a single record, for example a pixel in an idealized 2D screen.
+
+Instead, we define an output \textcolor{action}{verification} function $\extractmc$ that takes as input the section evaluated on all the graphic space associated with a point $\gsection_{\vindexpre\restriction_{\dbasepoint}}$ and returns the corresponding \textcolor{action}{measurable visual components} $\measurec_{k}$.
+\begin{equation}
+  \label{eq:artist:actual}
+  \extractmc: (\gsectionc \circ \vindexpre) \mapsto (\dbasec \xrightarrow{\extractmc_{\gsectionc}} \measurec)
+\end{equation}
+The measurable elements can only be computed over the entire preimage because these aspects, such as thickness or marker shape, refer to the entire visual element. 
 \begin{equation}
@@
 \end{equation}
-so instead we introduce a measurement extraction function $\extractm = measure\circ render$ that decomposes the rendered output into measurable components and returns a map from the data index into the measurement space
-\begin{equation}
-  \label{eq:artist:actual}
-  \extractmc: (\gsectionc \circ \vindexpre) \mapsto (\dbasec \xrightarrow{\extractmc_{\gsectionc}} \measurec)
-\end{equation}
-such that $\extractmc_{\gsectionc}(\dbasepoint)$ returns a quantification of the measurable components $\measurec_{\dbasepointc}$ of the visual element generated for the data at that point $\gsectionc|_{\vindexprec(\dbasepointc)}$.
-As shown in \autoref{eq:artist:actual}, the function $\extractmc$ takes in $\gsectionc$ and must evaluate the graphic section on all the graphic space associated with a point $\vindexpre\restriction_{\dbasepoint}$.
-This is because the measurable aspects, such as thickness or marker shape, refer to the entire visual element for that data point and as such can only be computed on the whole element.
+The extraction function is equivalent to measuring components of the rendered image $\extractm = measure\circ render$, which means an alternative way of implementing the function when $\gbase$ is not accessible is by decomposing the output into its measurable components.
 
@@
   \label{eq:atct:equivarance:verify:base}
-  \equivc(\dsectionc)(\dfunch(\dbasepointc^{\prime})) = \extractmc(\vartistc(\dsectionc))(\dbasepointc)
+  \equivc(\dsectionc)(\dfunchc(\dbasepointc^{\prime})) = \extractmc(\vartistc(\dsectionc))(\dbasepointc)
 \end{equation}
@@
   \label{eq:atct:equivarance:verify:fiber}
-  \equivc(\dfunctc(\dsectionc))(\dbasepoint) = \dfunctc_{\measure}(\extractmc(\vartistc(\dsectionc))(\dbasepoint))
+  \equivc(\dfunctc(\dsectionc))(\dbasepointc) = \dfunctc_{\measure}(\extractmc(\vartistc(\dsectionc))(\dbasepoint))
 \end{equation}
@@
 \begin{figure}[h!]
-  \label{fig:artist:equivariance}
   \includegraphics[width=1\columnwidth]{equivariance.png}
-  \caption{This artist is equivariant because the input data and output graphic representation are transposed $\dfunch$, scaled $\dfunct$ and both transposed and scaled $\dfunc$ in the same manner.}
+  \caption{This artist is equivariant because the input data and output graphic representation are transposed $\dfunch$, scaled $\dfunct$ and both transposed and scaled $\dfunc$ in the same manner.
+\label{fig:artist:equivariance}}
 \end{figure}
@@
 \begin{figure}[h!]
-  \label{fig:artist:operator}
   \includegraphics*[width=1\columnwidth]{qcom.png}
-  \caption{The circle-line glyph can either be considered as composed of two separate shapes, a circle and line, which each have their own graphic base spaces or as a single glyph which has an equivalent graphic base  $\gbase_{circle} \sqcup \gbase_{line} = \gbase_{circle-line}$ }
+  \caption{The circle-line glyph can either be considered as composed of two separate shapes, a circle and line, which each have their own graphic base spaces or as a single glyph which has an equivalent graphic base  $\gbase_{circle} \sqcup \gbase_{line} = \gbase_{circle-line}$  \label{fig:artist:operator}}
 \end{figure}
@@
 \begin{figure}[h!]
-  \label{fig:construction:xi}
+
   \includegraphics[width=1\columnwidth]{deform_retract.png}
-  \caption{The graphic base space $\gbase$ is collapsable to the line $\dbase$ such that every band $(\dbasepoint_i, [0,1])$ on $\gbase$ maps to corresponding point $\dbasepoint_i \in \dbase$ and the band $[0,1]$ determines which pixels are filled in to generate the shape of the line.\note{maybe put in a phantom line plot in s}}
+  \caption{The graphic base space $\gbase$ is collapsable to the line $\dbase$ such that every band $(\dbasepoint_i, [0,1])$ on $\gbase$ maps to corresponding point $\dbasepoint_i \in \dbase$ and the band $[0,1]$ determines which pixels are filled in to generate the shape of the line.\note{maybe put in a phantom line plot in s}\label{fig:construction:xi}}
 \end{figure}
@@
 
-\subsection{Encode Data as Measurable Components}
+\subsection{Component Encoders}
 \label{sec:construction:nu}
@@
 
-\subsection{Assemble Components Into a Graphic}
+\subsection{Graphic Compositor}
 As shown in \autoref{fig:construction:flow}, the compositor function $\vmarkc$ transforms the measurable components into properties of a visual element.
@@
 \begin{figure}[h!]
-  \label{fig:construction:q}
   \includegraphics[width=1\columnwidth]{q.png}
-  \caption{This simple $\vmarkc$ assembles a circular visual element that is the color specified in $\vsection(\dbasepoint)$ and is at the intersection specified in $\vsection(\dbasepoint)$}
+  \caption{This simple $\vmarkc$ assembles a circular visual element that is the color specified in $\vsection(\dbasepoint)$ and is at the intersection specified in $\vsection(\dbasepoint)$ \label{fig:construction:q}}
 \end{figure}
@@
   \label{eq:construction:q:verification:base}
-  \vsectionc(\dfunchc_{\dbasec}(\dbasepointc^{\prime})) = \extractmc(\vmarkc(\vsectionc)(\dfunch_{\dbasec}(\vindexpre^{\dbasepointc}))) = \measurec_{\dbasepointc}
+  \vsectionc(\dfunchc_{\dbasec}(\dbasepointc^{\prime})) = \extractmc(\vmarkc(\vsectionc)(\dfunch_{\dbasec}(\vindexprec^{\dbasepointc}))) = \measurec_{\dbasepointc}
 \end{equation}
@@
 \begin{equation}
-  \label{eq:construnction:nu:verify:base}
-  \dfunctc_{\vtotalc}(\vsectionc(\dbasepointc)) = \dfunctc_{\gtotalc}(\vmarkc(\vsectionc(\dbasepointc)))  
+  \label{eq:construnction:q:verify:base}
+  \dfunctc_{\vtotalc}(\vsectionc(\dbasepointc)) = \dfunctc_{\measurec}(\vmarkc(\vsectionc(\dbasepointc)))  
 \end{equation}
@@
 
-\section{Discussion: Translate Spec to Design}
+\section{Discussion: Feasibility of Math as Design Spec}
 \label{sec:discussion}
@@
 
-\subsection{Integrating Components Into Existing Library}
+\subsection{Integrating Components into an Existing Library}
 The $\vchannel$ and $\vmark$ are wrapped in a container object that stores the $\vartist = \vmark \circ \vchannel$ composition and a method for computing the $\vsectionc$.

### 2022-05-27 paper/main.tex
changes through xi

@@
 
-\IEEEPARstart{V}isualization software libraries consist of a set of functions for translating data into visual representations.
-These functions are expected to preserve the structure of the input data, however that structure is specified.
-Over time, library developers may need to develop new components to adapt to changing data needs, especially the growth of complex structured and distributed data.
-We propose that the complex structure of data can be described using concepts from algebraic topology and category theory, and that category theory can be used to express the constraints these components must satisfy in order to produce structure preserving data to visual transformations.
+\IEEEPARstart{V}isualizations are expected to be a faithful translation of the inherent structure of the data being visualized; therefore the components of visualization software libraries that translate data to graphics are also expected to preserve this structure.
+Over time, library developers may need to develop new components to adapt to changing data needs, especially high dimensional multivariate and distributed data.
+This paper provides a method for formally describing structure that should be preserved and this work provides guidance on how to verify this preservation.
 This abstraction of the data structure and specification of the components can then be used to guide the developement of functional\footnote{functional programming paradigm} components.
@@
 \section{Related Work}
-Visualization library components are required to preserve structure; in a visualization context this has widely come to be defined as a combination of the topological properties of the dataset and the mathematical structure of the dataset fields.
+We present an abstraction of data that is agnostic to the types of fields or number of dimensions of the data by building on work that has codified structure and how it is preserved.
+Structure is mainly described as a combination of the topological properties of the data and the mathematical structure of the dataset fields.
+Structure preservation is expressed as equivalent changes on the data input and graphics output of a visualization algorithm.
+Our model encapsulates all these notions of structure by providing a language for expressing topology, field structure, and changes to the data model.
 
@@
 \end{figure}
-The topological properties of a dataset are the ways in which elements are organized, as described by Wilkenson\cite{wilkinsonGrammarGraphics2005}.
+The topological properties of a dataset are the ways in which elements are connected, as described by Wilkenson\cite{wilkinsonGrammarGraphics2005}.
 Some examples, as shown in \autoref{fig:related-work:continuity:ktypes} are discrete rows of a table, a continuous 1D line, or a surface.
-The topological space (in purple) and the data form a key-value pair, as described by Munzner \cite{munznerVisualizationAnalysisDesign2014}.
-She imbues semantic meaning to the topology, wherein the key is an independent attribute used as a look up key to retrieve the value in the dataset.
-She defines the value as a dependent attribute.
-In this work we use a more general notion of the topological space as an indexing space for the data records.
-For example  in \autoref{fig:related-work:continuity:ktypes}, a point on the green line is mapped to by evaluating the gaussian on a point on the interval and each pixel in the map can be found via a lookup on the 2D surface.
+The topological space, in purple, acts as a generalized indexing space for the data.
+For example, a point on the green line is mapped to by evaluating the gaussian on a point on the interval and each pixel in the map can be found via a lookup on the 2D surface.
 This general formulation of topology gives us a language for expressing the continuity of a dataset.
-The network in \autoref{fig:related-work:continuity:ktypes} is a union of a 0D discrete points and 1D lines connecting those points.
+The network is a union of a 0D discrete points and 1D lines connecting those points.
 The data record in the yellow row can be expressed as a constant function of the yellow network node, while the data on the graph edge between any two points is a function of the distance between those points.
 
-\begin{figure}[h!]
-  \includegraphics[width=1\columnwidth]{continuity.png}
-  \caption{The \texttt{line} and \texttt{image} visual algorithms take in the same array of data.
-The \texttt{line} assumes that the array is a list of lines, while the \texttt{image} assumes that the array is an image.
-\label{fig:related-work:continuity:borked}}
-\end{figure}
-The continuity matters to visualization libraries because visual algorithms, such as a the line plot and image plot in \autoref{fig:related-work:continuity:borked}, assume the continuity of their data, as described in taxonomies of visualization algorithms by Troy, M\"{o}ller, and Chi \cite{toryRethinkingVisualizationHighlevel2004,chiTaxonomyVisualizationTechniques2000}.
-In \autoref{fig:related-work:continuity:borked}, both \texttt{line} and \textt{image} algorithms are fed the same input matrix.
-The line plot assumes this matrix is a list of lines and plots it as such.
-The image map assumes that the data is an image and plots it as such.
-Neither function has a way of asking the matrix what it is, nor does the matrix encode information about the continuity of the data it contains.
-While this work is not about choosing the right visualization type for the given continuity, the ability to encode the continuity expectations in the visual algorithm and the data container would provide a way of ensuring that the assumptions match.
-
-Current visualization libraries handle continuity in two main ways.
-Libraries that provide "building block" \cite{wongsuphasawatNavigatingWideWorld2021} components for building new visual algorithms embed their assumptions in their algorithms.
-This flexibility allows libraries such as matplotlib\cite{hunterMatplotlib2DGraphics2007}, VTK\cite{hanwellVisualizationToolkitVTK2015,geveciVTK2012}, and D3\cite{bostockDataDrivenDocuments2011} to support a wide variety of data types, but at the cost of the ambiguity of \autoref{fig:related-work:continuity:borked}.
-These libraries can also have visual algorithms with  parameters that are tuned to the expected data in a way that leaves the over all library feeling uncohesive.
-On the other hand, domain specific libraries are designed with the assumption of continuities that are common in the domain \cite{HeerSoftware2006}.
-Tabular data is a very common data topology and is assumed by
+Bertin stated that data should be represented by visual elements with the same type of topology - points should be represented by discrete entities, lines by lines, and surfaces through area marks \cite{bertinSemiologyGraphicsDiagrams2011a}.
+Visual algorithms assume the continuity of their data, as described in taxonomies of visualization algorithms by Troy, M\"{o}ller, and Chi \cite{toryRethinkingVisualizationHighlevel2004,chiTaxonomyVisualizationTechniques2000} and try to preserve it, but many data structures do not provide a method for encoding the topology of the data they store.
+For example, a \texttt{line} algorithm often take as input a list of (x,y) coordinates.
+The \texttt{line} algorithm assumes that its input is continuous and therefore displays the data as if it was continuous, regardless of the actual data topology.
+The ability to encode the continuity expectations in the visual algorithm and the data container would provide a way of verifying that the assumptions match and therefore that the components are preserving the topological structure.
+
+\subsection{Equivariance}
+Visualizations that preserve the mathematical structure of the data are called equivariant maps.
+Equivariant maps are maps where every change to an input has an equivalent change in the output.
+Bertin codified the idea that visual encoding of a data field should match the structure of the field; for example ordered data should be encoded as an ordered visual element or quantitative data as a quantitative visual element \cite{bertinSemiologyGraphicsDiagrams2011a}.
+This structure mapping is necessary to ensure that the visual representations in graphs are not distorting the data\cite{tufteVisualDisplayQuantitative2001} and graphs are easier to understand when properties match \cite{norman_things_smart}.
+The expectation of matching structure is generalized as a mapping of a binary operator that acts on the data domain to a binary operator that acts on the graphic domain by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}.
+An invertible binary operator is the basis of Kindlemannn and Scheideggers's algebraic model of visualization\cite{kindlmannAlgebraicProcessVisualization2014}.
+They propose three specific types of equivariance: that the visualization should not change if the data representation (i.e.
+the data container) changes, that data maps unambiguously to visual elements \cite{ziemkiewiczEmbeddingInformationVisualization2009}, that changing the data should coresspond to changes in the visualization in a perceptually significant manner.
+While the data representation introduced in \autoref{sec:atct:io} dispenses with a distinction between containers and the data, the expectation of unambiguous mappings is expressed in \autoref{sec:atct:xi} and the definition of equivariance introduced in \autoref{sec:artist:equivariance} includes the correspondence between data and visual changes.
+
+\subsection{Structure Preservation In Software}
+\label{sec:related-work:software}
+Visualization libraries are in part measured by how expressive the components of the library are; expressiveness is a measure of which structure preserving mappings a tool can implement \cite{mackinlayAutomaticDesignGraphical1987}.
+While some visualization tools aim to automate the pairing of data with structure preserving visual representations, such as Tableau\cite{StoltePolaris2002,hanrahanVizQL2006,MackinlayShowme2007}, many visualization libraries leave that choice to the user.
+Libraries that provide small buildable components \cite{wongsuphasawatNavigatingWideWorld2021}, such as components for making boxes or translating data values to colors, mostly embed their connectivity assumptions in their visual algorithms and allow the user choose almost any component, regardless of structure preservation.
+Domain specific libraries are designed with the assumption of continuities that are common in the domain \cite{HeerSoftware2006}, and therefore can somewhat restrict the space of choices to algorithms that are appropriate for the data.
+Buildable component libraries such as matplotlib\cite{hunterMatplotlib2DGraphics2007} and D3\cite{bostockDataDrivenDocuments2011} support a wide variety of data types, but at the cost of visual algorithms with  parameters that are tuned to the expected data in a way that leaves the over all library feeling uncohesive.
+VTK\cite{hanwellVisualizationToolkitVTK2015,geveciVTK2012} provides a language for expressing the topological properties of the data, but does so in a non-uniform manner.
+On the other hand, the topology of a table as a set of discrete rows, as illustrated in \autoref{fig:related-work:continuity:ktypes}, is assumed by
 A Presentation Tool\cite{mackinlayAutomatingDesignGraphical1986, mackinlayAutomatingDesignGraphical1986} and grammar of graphics\cite{wilkinsonGrammarGraphics2005} and the ggplot\cite{wickhamGgplot2ElegantGraphics2016a}, vega\cite{satyanarayanDeclarativeInteractionDesign2014}, and altair\cite{vanderplasAltairInteractiveStatistical2018} libraries built on these frameworks.
@@
 Networking libraries such as gephi\cite{bastianGephiOpenSource2009} and networkx\cite{HagbergExploringNetwork2008} assume a graph like structure.
-These libraries tend to benefit from more cohesive apis, but at the cost of a much more limited set of visualization algorithms than the building block libraries.
-
-\note{maybe move the following to \autoref{sec:atct:io} and consolidate?}
-To bridge the need for flexibility and cohesiveness, we build on the proposal of fiber bundles as a general abstraction model for visualization data, as proposed by Butler\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}.
-Fiber bundles are a mathematical structure that can be used to encode the topological properties of visualization data.
-There is also a methodology for encoding field types in fiber bundles, as described in Spivak's description of tables as fiber bundles\cite{spivakSIMPLICIALDATABASES}.
-Fiber bundles are discussed in more detail in \autoref{sec:atct:fiber-bundles}, but broadly they provide a mathematical representation of data that can encode topological properties, field types, and data values in a way that is both uniform and continuity and dimension independent way.
-
-\subsection{Equivariance}
-Equivariant maps are functions where variations in the input match variations in the output.
-The expectation that visualizations are equivariant maps from data to visual representation was first formalized by Bertin \cite{bertinSemiologyGraphicsDiagrams2011a}.
-He classified datasets into point, line, and surface topologies and proposed that these datasets should be represented by visual elements with the same sort of topology - points should be represented by discrete entities, lines by lines, and surfaces through area marks.
-Bertin also codified the idea that visual encoding of a data field should match the structure of the field; for example ordered data should be encoded as an ordered visual element or quantitative data as a quantitative visual element.
-This structure mapping is necessary to ensure that the visual representations in graphs are not distorting the data\cite{tufteVisualDisplayQuantitative2001} and graphs are easier to understand when properties match \cite{norman_things_smart}.
-
-The expectation of matching structure is generalized as a mapping of a binary operator that acts on the data domain to a binary operator that acts on the graphic domain by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987} An invertible binary operator is the basis of Kindlemannn and Scheideggers's algebraic model of visualization\cite{kindlmannAlgebraicProcessVisualization2014}.
-They propose three specific types of equivariance: that the visualization should not change if the data representation (i.e.
-the data container) changes, that changing the data should change the visualization in the same way in a perceptually significant manner.
-While the data representation introduced in \autoref{sec:atct:io} dispenses with a distinction between containers and the data, the definition of equivariance introduced in  \autoref{sec:artist:equivariance} includes invariant transformations and visually measurable transformations.
-
-Visualization libraries are in part measured by how expressive the components of the library are; expressiveness is a measure of which structure preserving mappings a tool can implement \cite{mackinlayAutomaticDesignGraphical1987}.
-While some visualization tools aim to automate the pairing of data with structure preserving automated tool, such as Tableau\cite{StoltePolaris2002,hanrahanVizQL2006,MackinlayShowme2007}, many visualization libraries leave that choice to the user.
-Some libraries, such as Matplotlib, set default encoders for their visual algorithms and allow the user to change the default.
-In this work, we propose guidelines for building expressive structure informed data to visual transformers.
-As with visual algorithms, we aim to provide a way to check that the structure of the data matches the structure expected by the component.
-Doing so allows users to more easily develop equivariant visual algorithms.
+These domain specific libraries tend to benefit from more cohesive apis, but at the cost of a much more limited set of visualization algorithms than the building block libraries.
 
@@
 \label{sec:atct}
-A goal of this work is to develop guidelines for building composable structure preserving functions for translating data into visual representations.
-Concepts in algebraic topology and category theory provide a language for formally expressing the structure that these visualizations are expected to preserve.
-Fiber bundles and sheaves, which come from algebraic topolology, provide a uniform generalizable way to describe data and visualization.
-Category theory is a method to describe how objects are specified, constrained, and composed \cite{wielsManagementEvolvingSpecifications1998}\; therefore in this work we propose categorical constructions of visualization components as specifications for implementable components.
-A brief introduction to category theory for visualization practitioners is in Vickers et al \cite{vickersUnderstandingVisualizationFormal2013}, but they are applying category theory to semantic concerns about visualization design rather than library architecture.
+In this section, we propose a mathematical abstraction of a data container and graphic prerendered output.
+This mathematical abstraction has a robust language for expressing topology and fields; expresses how to verify that data continuity is preserved on subset, distributed, and streaming data; and formalizes the expectation of a correspondence between data and visual elements.
 
@@
 \label{sec:atct:fiber-bundles}
-Fiber bundles, are a "unified, dimension-independent framework", as described by Butler\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}, that allow us to seperatly describe the topological properties and fields of a data source and also the mapping between topolology and data values.
+To bridge the need for flexibility and cohesiveness, we model data using a mathematical representation of data that can encode topological properties, field types, and data values in a way that is both uniform and continuity and dimension independent way.
+This mathematical representation is called a fiber bundle and its use as an abstract data representation in visualization was first proposed by Butler\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}.
 A fiber bundle $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and continuous surjective map $\pi: \dtotalc \rightarrow \dbasec$\cite{FiberBundle2020}.
@@
 
-The \textcolor{base}{base space} is a topological space \dbasec\ with points $\dbasepointc \in \dbasec$ and topology $\mathcal{T}_{\dbasec}$; $\mathcal{T}_k$ is a set of opensets $ \dbasepointc \in \opensetc \subseteqq \dbasec$ \cite{munkresElementsAlgebraicTopology1984} that cover the base space \dbasec.
-The \textcolor{fiber}{fiber space} is a topological space \dfiberc\ that is the preimage of the projection function $\pi$ at a point in the base space $\dfiber_{\dbasepointc} = \pi^{-1}(\dbasepointc)$ and fibers of a bundle are isomorphic $\dfiberc \simeq \dfiberc_{\dbasepointc}\;\forall \dbasepointc \in \dbase$.
-The \textcolor{section}{sections} of a bundle $\dsection$ are maps from the base space to points in the bundles over that base space $\textcolor{set}{\Gamma}$ denotes the set of all sections in a bundle over an openset.
+The \textcolor{base}{base space} is a topological space \dbasec\ with points $\dbasepoint \in \dbase$ and topology $\mathcal{T}_{\dbase}$; $\mathcal{T}_\dbasepoint$ is a set of opensets $ \dbasepoint \in \openset \subseteqq \dbase$ \cite{munkresElementsAlgebraicTopology1984} that cover the base space \dbasec.
+
+The \textcolor{fiber}{fiber space} is a topological space \dfiberc\ that is the preimage of the projection function $\pi$ at a point in the base space $\dfiber_{\dbasepoint} = \pi^{-1}(\dbasepoint)$, meaning that all points in a fiber over a space $\dfiber_{\dbasepoint}$ map to the same point $\dbasepoint$.
+Fibers of a bundle are isomorphic $\dfiberc \simeq \dfiberc_{\dbasepointc}\;\forall \dbasepointc \in \dbase$, which means that the space of possible values is the same set of values over all points in the base space.
+The fiber space can be used to encode the space of all possible values of a data field, which is analogous to a CS data type, and the fibers can be named, as described by Spivak \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES}.
+
+The \textcolor{section}{sections} $\dsection$ of a bundle are continuous functions that bind the base space with a set of values in the total space $\dtotal$ and $\textcolor{set}{\Gamma}$ denotes the set of all sections in a bundle over an openset.
+
 \begin{equation}
@@
 \end{equation}
+Each section function takes as input a point in the basespace $\dbasepoint$ and returns a single point in the fiber.
+Extending Spivak's fiber bundle description of a table, we define the values returned by a section as named and typed records.
+Since the fiber space is a topological space, it does not really have restrictions on the dimension of the types it can encode or the structure of those types.
+For example color can be represented as the cartesian product of three distinct sets, one for each color, or a three dimensional topological space $\mathbb{R} \times \mathbb{R} \times \mathbb{R} = \times \mathbb{R}^{3}$.
+
+The base space and the section evaluated on the base space form a sort of key-value pair, as described by Munzner \cite{munznerVisualizationAnalysisDesign2014}.
+She imbues semantic meaning to the topology, wherein the key is an independent attribute used as a look up key to retrieve the value in the dataset.
+She defines the value as a dependent attribute.
+In this work we use a more general notion of the topological space as an indexing space for the data records.
+Therefore, we do not impose any semantic constraints on what the fiber and base space encode.
+For example, a data cube of temperature can be encoded such that every fiber is a set of 2D temperature surfaces $\dfiber_{tempsurf}$, the base space is a 1D continuous interval, and each section returns a surface $\dsection(\dbasepoint) = f(x,y)$ where x and y are already parameterized.
+Alternatively, the fiber can encode each cell of the data cube $\dfiber_{lat} \times \dfiber_{lon} \times \dfiber_{temp}$, the base space could be a 3D volume, and each section returns a $\dsection(\dbasepoint) = \{lat, lon, temp\}$ record.
+
+Fiber bundles where the fibers are not identical are called non-trivial bundles.
+Non-trivial bundles allow us to generalize bundles to many different types of total spaces.
+We can also decompose non-trivial bundles into trivial bundles, which are bundles where all the fibers are identical.
+This is because all fiber bundles are locally trivial, which means that for a small enough subset of the data $\opensetc \subseteq \dbasec$, all the fibers are equivalent $\pi^{-1}(\opensetc)\xrightarrow{\equivc} \opensetc \times \dfiberc$.
+A bundle may be globally trivial, meaning that $\dtotalc = \dbasec \times \dfiberc$, and generally can be thought of as a twisted product of total and base spaces \cite{munkresElementsAlgebraicTopology1984}.
+Using trivial and locally trivial bundles allows us to encode the possible data records in a bundle as a single fiber $\dfiber$ type rather than having to maintain a fiber type for each point \dbasepoint.
 
@@
 
-The bundle in \autoref{fig:atct:stripbundle} is non-trivial, meaning that the fiber is not identical over every point in the base space $\dbase$.
-One section over this 1D continuous space is a sin function that takes as input a point in the base space $\dbase$ and returns the sin of the point.
-
-A fiber bundle is locally trivial, which means that for every point \dbasepointc\ there exists an open neighborhood $\dbasepointc \in \opensetc \subseteq \dbasec$ such that there is a homeomorphism $\pi^{-1}(\opensetc)\xrightarrow{\equivc} \opensetc \times \dfiberc$.
-A bundle may be globally trivial, meaning that $\dtotalc = \dbasec \times \dfiberc$, and generally can be thought of as a twisted product of total and base spaces \cite{munkresElementsAlgebraicTopology1984}.
-Throughout this work we will assume that fiber bundles are trivial because that allows us to model the section functions as from $\texttt{base} \rightarrow \texttt{fiber}$ since the fiber is the same at all points in the base space.
-
-Fiber bundles can be used to model data for visualization, as proposed by Butler\cite{butlerVectorBundleClassesForm1992, butlerVisualizationModelBased1989}.
-We describe the topological properties of the data \cite{wilkinsonGrammarGraphics2005} as the base space \dbasec.
-The topological properties, as described by Wilkenson, are how the data is organized - for example \dbase\ could be discrete points, a line, a 2D surface, a 3D volume, or a graph.
-The names and data types of the data fields can be expressed as the fiber space \dfiberc, as described by Spivak \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES}.
-Extending Spivak's fiber bundle description of a table, we define the values returned by a section as named and typed records of the dataset.
+The bundle in \autoref{fig:atct:stripbundle} is a fiber bundle representation of a mobius band, which is the total space \dtotal\ of the bundle.
+Each fiber \dfiber\ collapses to a point in the base space via the projection function $\pi$.
+The bundle is non-trivial because the fibers are not identical $[-1, 1] \not \equiv [1, 1]$.
+One section over this 1D continuous space is a \texttt{sin} function that takes as input a base point $\dbasepoint_{i}$ and returns a point in the fiber $sin(\dbasepoint_{i})$.
 
@@
 
-The bundle in \autoref{fig:atct:networkbundle} is constructed from a union of bundles over discrete points $\dbase_{1}, \dbase_{3}$ and a 1D continuous connecting line $\dbase_{2}$ that connects the two points.
-
-A fiber bundle is a mathematical structure for encoding continuity, type, and datasets with that continuity and type.
-In \autoref{fig:atct:fb}, the continuity is a discrete point \dbase.
-The fiber space that encodes the types is the cartesian product of the field types, and the section returns a single record with values of each field type.
-While this is a very trivial example, a fiber bundle base space can have almost any topology.
-The fiber space also does not really have restrictions on which types it can encode or the structure of those types - for example, \texttt{speed} and \texttt{direction} are separate fibers in \autoref{fig:atct:fb} but can be combined into a single two dimensional fiber $\dfiber_{speed, direction} = {\mathbb{R}^{+}}^{s}$ The generality in base space and fiber space makes fiber bundles a unified dimension and type independent data abstraction that can encode most of the common visualization data set structures, such as continuous functions, tables, data cubes, images and networks.
+In \autoref{fig:atct:networkbundle}, the network from \autoref{fig:related-work:continuity:ktypes} is encoded as a section of a bundle.
+This bundle is constructed from a union of bundles over discrete points $\dbase_{0}, \dbase_{2}$ and a 1D continuous connecting line $\dbase_{1}$ that connects these two points.
+The fiber over the data points $\dfiber_{node}$ expresses that the name column type is a subset of \textttt{string}, the temperature column type is \texttt{float}, and the precipitation column type is positive \texttt{float}.
+The fiber over the network edge encodes that the distance from the edge table has type \texttt{float}.
+The data section evaluated at a discrete point $\dbase_{0}$ returns the values for the first row in the table.
+Evaluating the function on $\dbase_{2}$ would return a different row in the table.
+The section $\dsection$ knows which nodes $\dbase_{1}$ is between and so returns the corresponding data value.
+The network edge function is constant and returns the same value any where along $\dbase_{1}$.
+
+The only constraint on sections is that they are (piecewise) continuous and that they return values in the fiber.
+They can be constant or non-constant and they can be defined over a single type of topology or a union of topologies.
+The generality in base space, fiber space, and section makes fiber bundles a unified dimension and type independent data abstraction.
+Therefore fiber bundles provide guidance on constructing a structure aware data interface layer that can support most of the common visualization data set structures, such as continuous functions, tables, data cubes, images and networks.
 
@@
 \label{sec:atct:io}
-We define categorical types for the topology and the fibers because categories are defined both by the objects of the category and the valid morphisms on the object.
-This allows for the formal construction of an equivariance aware type that holds the topology, field, type, and what changes to either must be preserved.
+We define categorical types for the topology and the fiber because this allows us to describe  how objects that encode topology and fiber should be specified, constrained, and composed \cite{wielsManagementEvolvingSpecifications1998}, which we can then translate to the implementation of these objects.
+In particular, the categorical construction allows us to bundle the topology and fiber as objects together with the changes to these objects, which are termed morphisms, that the visualization is expected to preserve.
 
@@
 
+A brief introduction to category theory for visualization practitioners is in Vickers et al \cite{vickersUnderstandingVisualizationFormal2013}, but they are applying category theory to semantic concerns about visualization design rather than library architecture.
+
 \subsection{Formal Properties of Data Containers}
 \label{sec:atct:sheaves}
-Sheaves on bundles are an "algebraic data structure", as described by Ghrist \cite{ghristElementaryAppliedTopology2014}, for expressing the bookkeeping involved in keeping track of sections of data over opensets.
-Sheaves are a way of expressing that subsets, distributed data, and streaming data are local sections that can be glued together.
-Formally, pre-sheaves are functors, meaning they are functions from objects of one category to objects of another category \cite{WhatFunctorDefinitions}.
-Here the sheaf
+Sheaves are a mathematical way to describe the expectation that data structures implement the bookkeeping that keeps track of the continuity of the data and graphic \cite{ghristElementaryAppliedTopology2014}.
+These constraints generalize to data that is subsetted, distributed, streaming, and on-demand.
+
+Presheaves are functors, meaning they are functions from objects of one category to objects of another category\cite{WhatFunctorDefinitions} that preserve the morphisms in each category.
+By definition, presheaves are maps from objects in an arbitrary category to an object in the category \setb\cite{nlab:presheaf}.
+Here the arbitrary category is the base space category and the object in \setb is the set of sections
+
 \begin{equation}
@@
 \end{equation}
-denotes a sheaf from opensets on the base space $\dbase$ to sets of sections on the bundle $\dtotal$. As a contravariant functor it preserves morphism s
-
+
+Presheaves are a contraviant functor, which means that the inclusion morphisms $\iota$ between base space objects go in the opposite direction from the restriction morphisms $\iota^*$ between objects in set.
 \begin{equation}
@@
 \end{equation}
-
-which associates a set of sections $\cgamma{\openset}{\dtotal\restriction_{\openset}}$ with an openset \openset.
-The sets of sections are objects of the category \setc\ and the morphisms of the category are restrictions\footnote{restriction functions } $\iota^*$.
-The sheaf over a an open set $\openset$ surrounding a point $\dbasepoint$ is called a stalk\cite{StalkSheaf2019}
+A presheaf describes the expectation that any set of sections defined over a larger space $\dsection \in Gamma(\opensetc_2, \dtotal\restriction_{\openset_{2}})$ is also defined over subspaces of that space $\dsection\restriction_{\opensetc_{1}}$.
+
+The locality and gluing axioms \cite{bakerMathsSheaf} of a sheaf describe the conditions for verifying that the sections over subspaces are part of a cohesive section over the whole space.
+The locality axiom asserts that two sections in a sheaf are equal when they evaluate to the same values over the same set of open sets.
+The gluing axiom asserts that the union of sections defined on specific open sets is equivalent to one big section over the union of spaces.
+The gluing axiom therefore expresses an equivalency between a concatenated representation of data and the distributed representation of the same data.
+In \autoref{fig:atct:networkbundle} this axiom means that the sections over each node and the edge can be combined into one section over the piece of the network in the figure.
+
+Each section of a sheaf over a point returns a single record in the fiber. The sheaf over an open set $\openset$ surrounding a point $\dbasepoint$ is called a stalk\cite{StalkSheaf2019}
 \begin{equation}
@@
 The germ is the section evaluated at a point in the stalk  $\dsection(\dbasepoint) \in \sheaf_{\dbase, \dtotal}\restriction_{\dbasepoint}$ and is the data.
-\note{I never use stalk and germ terminology at any point in this paper}
-
-The conditions on sheaves express the presheaf properties of continuity preservation as applied to sections, which as mentioned in \autoref{sec:atct:fiber-bundles} are how data is modeled in this framework.
-A sheaf is a presheaf that satisfies the locality and gluing axioms \cite{bakerMathsSheaf}.
-\begin{definition}
-\label{def:atct:sheaf:locality}
-The locality axiom is that given a union of opensets $\mathscr{\opensetc} = \bigcup\limits_{i\in I} \opensetc_i$ and $\dsectionc^{a}, \dsectionc^{b} \in \sheafc(\mathscr{\opensetc})$,  if $\dsectionc^{a}\restriction_{\opensetc_i} = \dsectionc^{b}\restriction_{\opensetc_i}$ for each $\opensetc_i \in \opensetc$ then $\dsectionc^{a} = \dsectionc^{b}$
-\end{definition}
-This means that two sections in a sheaf are equal when they evaluate to the same values over all open sets in a collection. 
-\begin{definition}
-\label{def:atct:sheaf:gluing}
-The gluing axiom is that given $\dsectionc^{i} \in \sheafc(\opensetc_i)$ such that ${\dsectionc^{i}}\restriction_{\opensetc_i\cap \opensetc_j} = {\dsectionc^{j}}\restriction_{\opensetc_i \cap \opensetc_j}$ for $\opensetc_i, \opensetc_j \in \mathscr{\opensetc}$, there exists $\dsectionc \in \sheafc(\mathscr{\opensetc})$ such that $\dsectionc\restriction_{\opensetc_i} = {\dsectionc^{i}}$.
-\end{definition}
-This means we can construct a section $\dsection$ on the collection of open sets that is the union of sections defined on specific open sets.
-This expresses an equivalency between a concatenated representation of data and the distributed representation of the same data.
-In this model, we represent data and graphics as a sheaf; doing so allows us to express that data sources are expected to implement the bookkeeping that a sheaf requires, meaning inclusions, restrictions, locality, and gluing must be preserved.
-
-\subsubsection{Mathematical Structure of Data}
+The stalk and germ in the stalk provide the derivative of the record, which is needed for some computational and visualization tasks.
+
+\subsection{Mathematical Structure of Data}
 \label{sec:atct:sheaves:measurments}
 We define structure on data as the allowable \textcolor{action}{transformations on the bundle}, which we formalize as actions $\dfuncc$ on sections of that sheaf.
-Our construction of $\dfunc$ is a generalization of the classification of measurement scales by their mathematical structure \cite{stevensTheoryScalesMeasurement1946,leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}, which often serves as the basis of the equivariance described in \autoref{sec:related-work:equivariance}.
+Actions  $act: X \times Y \tightarrow Y$ are elements of one set $X$ applied to elements of another set $Y$ such that the action in X applied to Y returns an element in Y\cite{nlab:action}.
+Actions are a generalization of applying a function to an element and returning an element of the same type.
+Our sheaf action  $\dfunc$ is a generalization of the classification of measurement scales by the actions allowed on the data values\cite{stevensTheoryScalesMeasurement1946,leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}, which often serves as the basis of the equivariance described in \autoref{sec:related-work:equivariance}.
 
 We separate data transformations into two types, transformations on the base space $(\dfunchc, \dfuncpullc)$ and transformations on the fiber space $\dfunctc$.
+These diagrams are seperated because the base space transformation and the pullback are functors on the openset object and associated set of sections, while the fiber transformation takes a single fiber to a different fiber in the set of sections.
+
 \begin{equation}
@@
 \end{equation}
-is from one open set to another open set in the same base space $\openset, \openset^{\prime} \subseteq \dbase$.
-For example a remapping of an underlying indexing, such as a sorting of primary keys for a database of discrete values.
-This base space transformation induces a pullback section
+The base space transformation is from one open set to another open set in the same base space $\openset, \openset^{\prime} \subseteq \dbase$.
+To correctly align the sections with the remapped base space, there is a a corresponding section pullback function
 \begin{equation}
@@
 Examples of a fiber transformation are actions such as partial ordering for ranking purpose \cite{bruggemannRankingPrioritizationMultiindicator2011} and the permutation, ordering, translation, and scaling codified as Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946}.
+Since there are many different fiber transformations,  there is not a uniform way of expressing how this morphism acts on sections unlike in \label{eq:atct:morphism:verify_base}.
 
@@
 \end{equation}
-which gives us an equation that can express complex transformations.
-For example, wind vectors on a sphere are a non-trivial bundle, meaning their fibers aren't identical at every point in the base space; therefore a rotation (change of base space) of the sphere must include an adjustment to the values that have moved closer to the pole (a change in fiber).
+which gives us an equation that can express transformations that have both a base space change and a fiber change.
+
 The data transform \dfunc\ is composable 
@@
 \end{figure}
-\autoref{fig:atct:phi} provides an example of a base space change \dfunch, a fiber space change \dfunct, and a composition of the two \dfunc\ applied to each data point $x_{\dbasepoint} \in \texttt{data}$.
-As defined in \autoref{eq:atct:morphism:verify_base},
-since transposition is a $(\dfunch, \dfuncpull)$, every data value $x_{\dbasepoint}$ must be equal to the value at the corresponding location in the transposed matrix $x_{\dbasepoint} = x_{\dbasepoint^{\prime}}, \dfunch(\dbasepoint^{\prime}) = \dbasepoint$ for all $x_{\dbasepoint^{\prime}} \in texttt{data.T}$ .
-In \autoref{fig:atct:phi}, this is demonstrated in the columns becoming rows such that there are no new values and each value retains the same neighboring values.
-As defined in \autoref{eq:atct:morphism:fiber} scaling is a transform $\dfunct$ that changes the value of each point but not its location.
-In \autoref{fig:atct:phi}, this means that $x_{\dbasepoint^{\prime}}= 2*x_{\dbasepoint}, \dbasepoint = \dbasepoint_{\prime}$ for all indices in \texttt{data}, \texttt{data.T}.
-As defined in \autoref{eq:atct:morphism:all} and illustrated in \autoref{fig:atct:phi}, these transformations can combine to yield a matrix that is both transposed and scaled such that $x_{\dbasepoint^{\prime}} = 2*x_{\dbasepoint}$ for all
-$\dfunch(\dbasepoint^{\prime}) = \dbasepoint$
-
-\subsection{Associations Between Data and Graphics}
+\autoref{fig:atct:phi} provides an example of a transposition base space change \dfunch, a scaling fiber space change \dfunct, and a composition of the two \dfunc\ applied to each data point $x_{\dbasepoint} \in \texttt{data}$.
+In the transposition only case, the values in $\dfuncpull\dsection$ retain their neighbors from $\dsection$ because \dfunc\ does not change the continuity.
+Each value in $\dfuncpull\dsection$ is also the same as in $\dsection$, just moved to the new position.
+In $\dfunct\dsection$, each value is scaled by two but remains in the same location as in $\dsection$.
+And in $\dfunc\dsection$ each function is transposed such that it retains its neighbors and all values are scaled consistently.
+
+\subsection{Data and Graphic Correspondence}
 \label{sec:atct:xi}
-The data sections defined in \autoref{eq:atct:fb} are sections of sheaves, introduced in \autoref{sec:atct:sheaves}, over open sets $\openset \subseteq \dbase$ in a base space \dbase\ that encodes the topology of the data topology.
+A property of sheaves is that, given two sheaves, if there exists a map between the base spaces of each sheave, then we can construct transport functors that can move each sheaves over the other base space.
+This allows us to move data over the corresponding graphic base space and graphics over the corresponding data base space.
+The data sections defined in \autoref{eq:atct:fb} are sections of sheaves, introduced in \autoref{sec:atct:sheaves}, over open sets $\openset \subseteq \dbase$ in a base space \dbase\ that encode the topology of the data topology.
 The graphic sections, defined in \autoref{eq:atct:fb_graphic_section}, are sections of sheaves over open sets $\opensetg \subseteq \gbase$ in a base space \gbase\ that encodes the topology of the graphic.
@@
   \includegraphics*[width=1\columnwidth]{xi_scatter.png}
-  \caption{The base space functor $\vindexc$ is the mapping from each disc $\gbasec_i$ to it's corresponding point $\dbasepointc_i$.
-Over each point is a data record $\dsectionc$.
-The pullback function $\vindexpullc$ copies these records over $\gbase$ so that each piece of each marker knows the data associated with it.
-Each graphic disc has a corresponding graphic function $\gsectionc$.
-The pushforward function $\vindexpushc$ converts $\gsectionc$ into the visual parameters associated with each data point $\dbasepoint$.\label{fig:atct:morphisms:sheaf}}
+  \caption{The graphic base space $\gbase_{0}$ constracts to the data base space $\dbase_{0}$, and $\gbase_{1}$ contracts to $\dbase_{1}$.
+Therefore the data section $\dsection(\dbase_{0})$ can be copied over such that every pulled back section evaluates to the same record $\vindexpull\dsection|_{\gbase_{0}} = \dsection_{\dbasepoint_{0}}$.
+Each of these pulled back sections $\vindexpull(\gbasepoint_i)$ has a corresponding graphic section $\gsection(\gbasepoint_i)$ that returns a record in an idealized prerendered infinite resolution space.
+Here we approximate that space and show $\gsection|_{\gbase_0}, \gsection|_{\gbase_1}$ to show how a section produces a visual element that has corresponding data.
+The way each element is parameterized is encapsulated in the pushforward $\vindexpush\gsection$.
+The pushforward is the curried section $\gsection|_{\vindexpre(\dbasepoint)}$, so we've chosen to represent x and y in data units and color in hex as an example of the type of parameterization encapsulated in the function.
+\label{fig:atct:morphisms:sheaf}}
 \end{figure}
@@
 Functors between sheaves are a way of expressing the bookkeeping involved in keeping track of which graphic section corresponds to which data section.
+The $(\dbasepoint_i, \gbase_i)$ pairing expressed in \autoref{eq:atct:xi} establishes that there is a correspondence between sections evaluated over $\dbasepoint_i$ and $\gbase_i$.
 This allows us to construct graphic specifications for each data index $\vindexpush\gsection$ and retrieve the data $\vindexpull\dsection$ for any graphic section generating any piece of a graphic.
-In \autoref{fig:atct:morphisms:sheaf}, the graphic specifications are the set of visual parameters associated with each point $\dbasepoint_i$.
-Each of these parameters describes the corresponding graphic $\gsectionc|_{\gbase_i}$ generated by evaluating the graphic section on the graphic base space associated with that data point $\vindexpre^{\dbasepoint_i} = \gbase_i$.
-The pullback functor $\vindexpull$ copies each data section at each point $\dsectionc(\dbasepointc_i)$ over the corresponding disk; in \autoref{fig:atct:morphisms:sheaf} this means that every pixel in a given marker, for example the yellow triangle, maps back into a single data record.
+In \autoref{fig:atct:morphisms:sheaf}, the sections return rows in the table from \autoref{fig:related-work:continuity:ktypes}.
+Each section $\dsection(\dbasepoint_{i})$ has a corresponding visual element produced by evaluating the graphic section on the graphic base space $\gsection|_{\gbase_{i}}$.
+In \autoref{fig:atct:morphisms:sheaf} the output of $\gsection$ as pixels in an idealized prerendered space is approximated as the circles.
+Each graphic section $\gsection(\gbasepoint), \gbasepoint \in \gbase_{i}$ that is a piece of the graphic has a corresponding pulled back data value $\vindexpull\dsectionc(\gbasepoint)$.
+The pullback copies the data points such that $\vindexpull\dsectionc|_{\gbase_i} = \dsection(\dbasepoint_i)$.
+Each visual element $\gsection|_{\gbase_i}$ is generated from a single record $\dsection(\dbasepoint_i)$; therefore the pushforward is the curried graphic section $\gsection|_{\vindexpre(\dbasepoint_i)}$.
+The graphic parameterized in \autoref{fig:atct:morphisms:sheaf} is intended as an approximation of $\vindexpush\gsection$.
+
 Visualization specification languages such as vega \cite{heerDeclarative2010} and svg \cite{quintScalable2003} assume that there is a graphic  specification for every point.
@@
 which means that, as shown in \autoref{eq:construction:nu:fabrication}, any collection of $\vchannel$ functions can be composed such that they are equivalent to a $\vchannel$ that directly converts the input to the output.
-This means that a developer has the flexibility to implement either the direct convertor or the building block version (or both) depending on other library concerns.
-As with artists, $\vchannel$ are maps of sections such that the composition operators defined in \autoref{sec:artist:composition} can also act on transformers $\vchannel$, meaning that encoders can be added and multiplied.
+As with artists, $\vchannel$ are maps of sections such that the composition operators defined in \autoref{sec:artist:composition} can also act on transformers $\vchannel$, meaning that encoders can be added $\vchannel_{a+b} = \vchannel_{a} + \vchannel_{b}$ and multiplied d $\vchannel_{a\times b} = \vchannel_{a}  \vchannel_{b}$.
+Encoders designed to satisfy these composability constraints provide for a rich set of building blocks for implementing complex encoders.
 
@@
 The composition in \autoref{eq:construction:q:fabrication} means that different measurable components can yield the same visual elements.
+The composition operators defined in \autoref{sec:artist:composition} can also act on compositors $\vmark$ such that $\vmarkc_{a+b} = \vmark_{a} + \vmark_{b}$ and multiplied d $\vmark_{a\times b} = \vmark_{a}  \vmark_{b}$.
 

### 2022-05-28 paper/main.tex
intro

@@
 \IEEEraisesectionheading{\section{Introduction}\label{sec:intro}}
-\note{please prioritize flagging content that needs major revisions: (1) Wrong (2) Missing (3) Horrific  (4) Incoherant (5) Movable to appendix - TCVG page limit is 12 pages content + supplemental}
+\note{need a math check + rough cut on sections 4+}
 
 \IEEEPARstart{V}isualizations are expected to be a faithful translation of the inherent structure of the data being visualized; therefore the components of visualization software libraries that translate data to graphics are also expected to preserve this structure.
-Over time, library developers may need to develop new components to adapt to changing data needs, especially high dimensional multivariate and distributed data.
-This paper provides a method for formally describing structure that should be preserved and this work provides guidance on how to verify this preservation.
-This abstraction of the data structure and specification of the components can then be used to guide the developement of functional\footnote{functional programming paradigm} components.
-Since these components are functional, they are expected to be more composable and modular than existing components, have fewer side effects and therefore be easier to incorporate alignside existing library elements, and be verifiable and therefore easier to test \cite{huHowFunctionalProgramming2015,hughesWhyFunctionalProgramming1989}.
+Motivated by a need to develop new library components that adapt to modern data needs, especially high dimensional multivariate and distributed data, we propose a framework that provides a language to express that topological and field structure of the input data that must be mapped to the visualization and which changes to data must be expressed in visualization, and a functional framework that describes composability in terms of functional algebraic operations.
+This allows complex visualizations to be built from simple verifiable parts\cite{huHowFunctionalProgramming2015,hughesWhyFunctionalProgramming1989}, which makes these components easier to drop into the existing library and be used by domain specific library developers to build new libraries that work with existing ones.
+While many of these elements have been partially described elsewhere, our framework naturally combines them in a unified framework based on ideas in category theory.
+Further we see the application of these ideas in low level visualization library design as our motivation, and we hope these ideas will help shape the reorganization of critical data visualization libraries, such as Matplotlib.
 In this paper, we introduce a mathematical framework for describing, constructing, and testing visualization library components and provide a toy example illustrating these concepts.
@@
 \section{Related Work}
-We present an abstraction of data that is agnostic to the types of fields or number of dimensions of the data by building on work that has codified structure and how it is preserved.
-Structure is mainly described as a combination of the topological properties of the data and the mathematical structure of the dataset fields.
-Structure preservation is expressed as equivalent changes on the data input and graphics output of a visualization algorithm.
-Our model encapsulates all these notions of structure by providing a language for expressing topology, field structure, and changes to the data model.
+In this work, we propose a uniform method for expressing the structure that visualizations are expected to preserved by combining and generalizing previous work that describes structure as a combination of the topological properties of the data, the mathematical structure of the data fields, and equivalent changes to data and graphics.
+We propose that a unified abstraction model can guide developers in building structure aware composable, reusable software library components with more consistent interfaces.
 
@@
 \label{sec:related-work:software}
-Visualization libraries are in part measured by how expressive the components of the library are; expressiveness is a measure of which structure preserving mappings a tool can implement \cite{mackinlayAutomaticDesignGraphical1987}.
+Visualization libraries are in part measured by how expressive the components of the library are; where expressiveness is a measure of which structure preserving mappings a tool can implement \cite{mackinlayAutomaticDesignGraphical1987}.
 While some visualization tools aim to automate the pairing of data with structure preserving visual representations, such as Tableau\cite{StoltePolaris2002,hanrahanVizQL2006,MackinlayShowme2007}, many visualization libraries leave that choice to the user.
@@
 \begin{figure}[h!]
-  \includegraphics[width=\columnwidth]{fbtau.png}
+  \centering
+  %needs cropping
+  \includegraphics[scale=.5]{fbtau.png}
   \caption{A non-trivial fiber bundle is bundle where the (by definition isomorphic) fibers change along the indexing base space.
@@
 
-An artist is equivariant when changes to expected and actual are equivariant.
+An artist is equivariant when changes to the input and output are equivariant.
+
 As introduced in \autoref{eq:atct:morphism:base}, the base space transformation \dfunch\ is invariant because $\dsection\restriction_{\openset} = \dsection\restriction_{\dfunch(\openset^{\prime})}$.
@@
 As illustrated in \autoref{eq:artist:operator}, data bundles can be combined by taking the disjoint  union of base spaces $\dbase^{a} \sqcup_{\dbase^c} \dbase^b$, where $\dbase^c$ is an overlap.
-When the fibers of each bundle are isomorphic $\dfiber^{a} \simeq \dfiber^{b}$, which we denote as $\dtotal$, this is analogous to adding more observations to a dataset.
+When the fibers of each bundle are isomorphic $\dfiber^{a} \simeq \dfiber^{b}$, which we denote as $\dtotal$, this is analogous to adding more records to a dataset.
 We propose an addition operator that states that an artist that takes in a dataset can be constructed using artists that take as inputs subsets of the dataset

### 2022-05-29 paper/main.tex
folded back in a bit of the gluing stuff to make things concrete

@@
 The locality and gluing axioms \cite{bakerMathsSheaf} of a sheaf describe the conditions for verifying that the sections over subspaces are part of a cohesive section over the whole space.
-The locality axiom asserts that two sections in a sheaf are equal when they evaluate to the same values over the same set of open sets.
-The gluing axiom asserts that the union of sections defined on specific open sets is equivalent to one big section over the union of spaces.
-The gluing axiom therefore expresses an equivalency between a concatenated representation of data and the distributed representation of the same data.
+The locality axiom asserts that two sections in a sheaf are equal $\dsection^{a} = \dsection^{b}$ when they evaluate to the same values over the same set of open sets $\desection^{a}|{\openset} =  \dsection^{b}|_{openset}$.
+The gluing axiom asserts that the union of sections defined on specific open sets is equivalent to one big section over the union of spaces $\dsection|_{\openset_{i} \sqcup \openset_{j}} = \dsection^{i}|_{\openset_i} \sqcup \dsection^{j}|_{\openset_j}$ if these sections agree on overlaps $\dsection^{i}|_{\openset_i\cap\openset_j} =  \dsection^{j}|_{\openset_i\cap\openset_j}$ The gluing axiom generalizes to expresses an equivalency between a concatenated representation of data and distributed representation of the same data.
 In \autoref{fig:atct:networkbundle} this axiom means that the sections over each node and the edge can be combined into one section over the piece of the network in the figure.

### 2022-05-29 paper/main.tex
xi figure caption

@@
   \includegraphics*[width=1\columnwidth]{xi_scatter.png}
-  \caption{The graphic base space $\gbase_{0}$ constracts to the data base space $\dbase_{0}$, and $\gbase_{1}$ contracts to $\dbase_{1}$.
-Therefore the data section $\dsection(\dbase_{0})$ can be copied over such that every pulled back section evaluates to the same record $\vindexpull\dsection|_{\gbase_{0}} = \dsection_{\dbasepoint_{0}}$.
-Each of these pulled back sections $\vindexpull(\gbasepoint_i)$ has a corresponding graphic section $\gsection(\gbasepoint_i)$ that returns a record in an idealized prerendered infinite resolution space.
-Here we approximate that space and show $\gsection|_{\gbase_0}, \gsection|_{\gbase_1}$ to show how a section produces a visual element that has corresponding data.
-The way each element is parameterized is encapsulated in the pushforward $\vindexpush\gsection$.
-The pushforward is the curried section $\gsection|_{\vindexpre(\dbasepoint)}$, so we've chosen to represent x and y in data units and color in hex as an example of the type of parameterization encapsulated in the function.
+  \caption{The functors $\vindex, \vindexpull, \vindexpush$ establish the correspondence between a data section \dsection\ and a graphic generation function \gsection.
+For every data record $\dsection(\dbasepoint)$, there is a graphic generating function $\vindexpush\gsection = \gsection|_{\vindexpre\dbasepoint}$.
+This curried function is here represented as the visual parameters associated with that function to demonstrate that it generates the visual element for that data record.
+Every data section can also be copied over to the graphic space such that each piece of the (idealized in infinite resolution) graphic $\gsection(\dbasepoint)$ has a corresponding data value $\vindexpull\dsection(\dbasepoint)$.
   \label{fig:atct:morphisms:sheaf}}
@@
 
-Functors between sheaves are a way of expressing the bookkeeping involved in keeping track of which graphic section corresponds to which data section.
+Functors between sheaves are a way of expressing the bookkeeping involved in keeping track of which graphic section \gsection\ corresponds to which data section \dsection\.
 The $(\dbasepoint_i, \gbase_i)$ pairing expressed in \autoref{eq:atct:xi} establishes that there is a correspondence between sections evaluated over $\dbasepoint_i$ and $\gbase_i$.

### 2022-06-01 paper/main.tex
xi and operator figs

@@
 \IEEEPARstart{V}isualizations are expected to be a faithful translation of the inherent structure of the data being visualized; therefore the components of visualization software libraries that translate data to graphics are also expected to preserve this structure.
-Motivated by a need to develop new library components that adapt to modern data needs, especially high dimensional multivariate and distributed data, we propose a framework that provides a language to express that topological and field structure of the input data that must be mapped to the visualization and which changes to data must be expressed in visualization, and a functional framework that describes composability in terms of functional algebraic operations.
-This allows complex visualizations to be built from simple verifiable parts\cite{huHowFunctionalProgramming2015,hughesWhyFunctionalProgramming1989}, which makes these components easier to drop into the existing library and be used by domain specific library developers to build new libraries that work with existing ones.
+Motivated by a need to develop new library components that adapt to modern data needs, especially high dimensional multivariate and distributed data, we propose a framework that topological and field structure of the input data that must be mapped to the visualization and which changes to data must be expressed in visualization, and a a functional design framework that describes composability in terms of functional algebraic operations.
+\note{this may be a good place to incorporate types} This allows complex visualizations to be built from simple verifiable parts\cite{huHowFunctionalProgramming2015,hughesWhyFunctionalProgramming1989}, which makes these components easier to drop into the existing library and be used by domain specific library developers to build new libraries that work with existing ones.
 While many of these elements have been partially described elsewhere, our framework naturally combines them in a unified framework based on ideas in category theory.
 Further we see the application of these ideas in low level visualization library design as our motivation, and we hope these ideas will help shape the reorganization of critical data visualization libraries, such as Matplotlib.
-In this paper, we introduce a mathematical framework for describing, constructing, and testing visualization library components and provide a toy example illustrating these concepts.
+The contribution of this paper is a uniform method for describing the structure in data and a mathematical framework for describing, constructing, and testing structure preserving visualization library components.
 
 \section{Related Work}
-In this work, we propose a uniform method for expressing the structure that visualizations are expected to preserved by combining and generalizing previous work that describes structure as a combination of the topological properties of the data, the mathematical structure of the data fields, and equivalent changes to data and graphics.
+We build on the exentsive work codifying the structure of visualization data to develop a more generalizable method for expressing the structure that visualizations are expected to preserved.
+Broadly, previous work describes structure as a combination of the topological properties of the data, the mathematical structure of the data fields, and equivalent changes to data and graphics.
 We propose that a unified abstraction model can guide developers in building structure aware composable, reusable software library components with more consistent interfaces.
@@
 The locality and gluing axioms \cite{bakerMathsSheaf} of a sheaf describe the conditions for verifying that the sections over subspaces are part of a cohesive section over the whole space.
-The locality axiom asserts that two sections in a sheaf are equal $\dsection^{a} = \dsection^{b}$ when they evaluate to the same values over the same set of open sets $\desection^{a}|{\openset} =  \dsection^{b}|_{openset}$.
-The gluing axiom asserts that the union of sections defined on specific open sets is equivalent to one big section over the union of spaces $\dsection|_{\openset_{i} \sqcup \openset_{j}} = \dsection^{i}|_{\openset_i} \sqcup \dsection^{j}|_{\openset_j}$ if these sections agree on overlaps $\dsection^{i}|_{\openset_i\cap\openset_j} =  \dsection^{j}|_{\openset_i\cap\openset_j}$ The gluing axiom generalizes to expresses an equivalency between a concatenated representation of data and distributed representation of the same data.
+The locality axiom asserts that two sections in a sheaf are equal $\dsection^{a} = \dsection^{b}$ when they evaluate to the same values over the same set of open sets $\dsection^{a}|{\openset} =  \dsection^{b}|_{\openset}$.
+The gluing axiom asserts that the union of sections defined on specific open sets is equivalent to one big section over the union of spaces $\dsection|_{\openset_{i} \cup \openset_{j}} = \dsection^{i}|_{\openset_i} \cup \dsection^{j}|_{\openset_j}$ if these sections agree on overlaps $\dsection^{i}|_{\openset_i\cap\openset_j} =  \dsection^{j}|_{\openset_i\cap\openset_j}$ The gluing axiom generalizes to expresses an equivalency between a concatenated representation of data and distributed representation of the same data.
 In \autoref{fig:atct:networkbundle} this axiom means that the sections over each node and the edge can be combined into one section over the piece of the network in the figure.
@@
 such that every $\opensetg \subseteq \gbase$ must map to a corresponding set $\openset \subset \dbase$.
+The functor $\vindex$ is surjective such that for every point $\dbasepoint \in \openset$ there is a corresponding set of points $\{\gbasepoint | \gbasepoint \in  \vindexpre(k)\}$ for all $\gbasepoint in \opensetg$.
 
@@
 \begin{equation}
-  \vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexprec(\dbasepointc)} = \gsectionc(\gbasepointc)\;\forall \gbasepointc\in \vindexprec(\dbasepointc)
+  \vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexprec(\dbasepointc)} = \{\gsectionc(\gbasepointc)\;for\;all\; \gbasepointc\in \vindexprec(\dbasepointc)\}
 \end{equation}

### 2022-06-07 paper/main.tex
related work, start of topology revisions

@@
 \section{Related Work}
-We build on the exentsive work codifying the structure of visualization data to develop a more generalizable method for expressing the structure that visualizations are expected to preserved.
+We build on the exentsive work codifying the structure of visualization data to develop a more generalizable method for expressing the structure that visualizations are expected to preserve.
 Broadly, previous work describes structure as a combination of the topological properties of the data, the mathematical structure of the data fields, and equivalent changes to data and graphics.
@@
   \includegraphics[width=1\columnwidth]{k_different_types.png}
-  \caption{A topological representation of the data (in purple) is a simplified representation of how the records in a dataset are connected.
-For example, a 1D continuous function is evaluated on a 1D continuous interval, and a map is 2D continuous surface.
-A network is a union of discrete nodes and continuous lines.
-The nodes represent the continuity of a table of discrete weather station data and the graph edges are encoded with the distances between the nodes.
+  \caption{
+    A topological representation of the data (in purple) is a simplified representation of how the records in a dataset are connected.
+The table is a set of independent records; therefore it has a 0D continuity represented by the disconnected dots.
+These records can also be samples from a timeseries, which has a 1D continuity.
+\note{started making a map of the temperature data but was taking a while so I didn't  }
     \label{fig:related-work:continuity:ktypes}}
 \end{figure}
-The topological properties of a dataset are the ways in which elements are connected, as described by Wilkenson\cite{wilkinsonGrammarGraphics2005}.
-Some examples, as shown in \autoref{fig:related-work:continuity:ktypes} are discrete rows of a table, a continuous 1D line, or a surface.
-The topological space, in purple, acts as a generalized indexing space for the data.
-For example, a point on the green line is mapped to by evaluating the gaussian on a point on the interval and each pixel in the map can be found via a lookup on the 2D surface.
-This general formulation of topology gives us a language for expressing the continuity of a dataset.
-The network is a union of a 0D discrete points and 1D lines connecting those points.
-The data record in the yellow row can be expressed as a constant function of the yellow network node, while the data on the graph edge between any two points is a function of the distance between those points.
-
-Bertin stated that data should be represented by visual elements with the same type of topology - points should be represented by discrete entities, lines by lines, and surfaces through area marks \cite{bertinSemiologyGraphicsDiagrams2011a}.
-Visual algorithms assume the continuity of their data, as described in taxonomies of visualization algorithms by Troy, M\"{o}ller, and Chi \cite{toryRethinkingVisualizationHighlevel2004,chiTaxonomyVisualizationTechniques2000} and try to preserve it, but many data structures do not provide a method for encoding the topology of the data they store.
+
+Roughly, topology is a formal way of encoding the connectivity of the elements of the dataset \cite{wilkinsonGrammarGraphics2005} and there is an expectation that the visual elements should match this connectivity.
+Formally, a topological space is a set $X$ with a structure called a topology.
+The set $X$ is a collection of points, where these points are \note{getting tripped up on how much of the definition to include here, but this is the natural place to give a definition of topology and opensets}
+Some common topologies, as illustrated in\autoref{fig:related-work:continuity:ktypes}, are point, lines, and surfaces.
+The table is an example of disconnected discrete elements and therefore has the topology of discrete points.
+The timeseries is  sampled from a 1d continuous temperature measurement, and therefore has a topology of a 1D space.
+The map in \autoref{fig:related-work:continuity:ktypes} shows a 2D continuous surface.
+As illustrated in the figure, it is not uncommon for the same dataset to have multiple topologies.
+For the common topologies of discrete points, lines, and surfaces, Bertin recommend that they be represented visually by point, line, and area type visual elements respectively \cite{bertinSemiologyGraphicsDiagrams2011a}, and this expectation has been generalized in the development of visual algoriithms.
+
+Visual algorithms assume the continuity of their data, as described in taxonomies of visualization algorithms Chi and by Troy and M\"{o}ller \cite{chiTaxonomyVisualizationTechniques2000,toryRethinkingVisualizationHighlevel2004} and try to preserve it, but many data structures do not provide a method for encoding the topology of the data they store.
 For example, a \texttt{line} algorithm often take as input a list of (x,y) coordinates.
@@
   %needs cropping
-  \includegraphics[scale=.5]{fbtau.png}
+  \includegraphics[trim={2cm 2cm 2cm 2cm}, clip]{fbtau.png}
   \caption{A non-trivial fiber bundle is bundle where the (by definition isomorphic) fibers change along the indexing base space.
@@
 Presheaves are a contraviant functor, which means that the inclusion morphisms $\iota$ between base space objects go in the opposite direction from the restriction morphisms $\iota^*$ between objects in set.
+\note{maybe unpack each piece of this equation into a seperate equation/walk through, from U to fiber when trivial}  
 \begin{equation}
@@
 \end{equation}
-A presheaf describes the expectation that any set of sections defined over a larger space $\dsection \in Gamma(\opensetc_2, \dtotal\restriction_{\openset_{2}})$ is also defined over subspaces of that space $\dsection\restriction_{\opensetc_{1}}$.
+A presheaf describes the expectation that any set of sections defined over a larger space $\dsection \in \Gamma(\opensetc_2, \dtotal\restriction_{\openset_{2}})$ is also defined over subspaces of that space $\dsection\restriction_{\opensetc_{1}}$.
 
@@
 \end{equation}
-
+\note{$\xi_*rho(k)(s)=rho(s) for s in \xi^{-1}(k)$
+ and move {} into narrative}
 Therefore, the continuous map $\vindex$ and transport functors $\vindexpull, \vindexpush$ allow us to express the correspondence between graphic section and data section.
@@
   \includegraphics[width=1\columnwidth]{equivariance.png}
-  \caption{This artist is equivariant because the input data and output graphic representation are transposed $\dfunch$, scaled $\dfunct$ and both transposed and scaled $\dfunc$ in the same manner.
+  \caption{This artist is equivariant because when the input data $\dsection$ is transposed, $\dfunch$, scaled $\dfunct$, and transposed and scaled $\dfunc$, the corresponding colored cells are transposed, scaled such that the color is moved two steps, and both transposed and scaled.
   \label{fig:artist:equivariance}}
@@
 
+ \note{replace w/ a line plot w/markers}
  One example of an artist that is a sum of artists is a sphere drawer that draws different quadrants of a sphere $\vartist(\dsection) = \vartist_{1}(\dsection_{1}) + \vartist_{2}(\dsection_{2}) + \vartist_{3}(\dsection_{3}) \vartist_{4}(\dsection_{4})$.
@@
 \begin{figure}[h!]
-  \includegraphics*[width=1\columnwidth]{qcom.png}
-  \caption{The circle-line glyph can either be considered as composed of two separate shapes, a circle and line, which each have their own graphic base spaces or as a single glyph which has an equivalent graphic base  $\gbase_{circle} \sqcup \gbase_{line} = \gbase_{circle-line}$  \label{fig:artist:operator}}
+  \centering
+  \includegraphics*[scale=.75]{qcom.png}
+  \caption{The circle-line visual element can be constructed via $\gsection_{circle}$ + $\gsection_{line}$ functions that generate the circle and line elements respectively.
+This is equivalent to a $\gsection_{circle+line}$ function that takes as input the combined base space $\gbase_{circle} \sqcup \gbase_{line} = \gbase_{circle-line}$ and returns pixels in the circle-line element.
+\label{fig:artist:operator}}
 \end{figure}
-
-The circle-line glyph in \autoref{fig:artist:operator} can be generated via a line plot and highlighted points in that plot, which is the addition case of two datasets with the same fields and different continuities.
-The same circle-line glyph can also be generated via the multiplication case as different fields of the same bundle.
-As mentioned in \autoref{sec:artist:operator:multiplication}, multiplication is implemented as addition of two copies of a base space.
+There is no way to visually determine whether a visual element is the output of a single artist or a multiplied or added collection of artists.
+The circle-line visual element in \autoref{fig:artist:operator} can be a visual representation of a highlighted point intersecting with a line plot with the same fields.
+The same element can also be encoding some fields of a section in the circle and other fields of that section in the lines.
+\note{+*equive}
 Although we have been discussing the trivial cases of adding observations or adding fields, this merging of artists in datasets can be generalized:
@@
 \label{sec:construction}
-\begin{figure}[h!]
-  \label{fig:construction:flow}
-  \includegraphics*[width=1\columnwidth]{path_of_q.png}
-  \caption{One method of constructing the artist is by formulating the artist as having two stages: (1) encoding functions $\vchannelc$ that convert different parts of the input into measurable visual components, such as color or position; (2) compositing functions $\vmarkc$ that assemble the measurable visual components into a visual element generating function.
-In this construction, we propose an intermediate visual bundle $\vtotalc$ that encodes the space of possible visual encodings.}
-\end{figure}
+
 We propose that one way of constructing artist functions is to separate generating a visualization, as shown in \autoref{fig:construction:flow}, is into an equivariant continuity preserving encoding stage $\vchannelc$ and a equivariant continuity preserving compositing stage $\vmarkc$.
@@
   \includegraphics[width=1\columnwidth]{deform_retract.png}
-  \caption{The graphic base space $\gbase$ is collapsable to the line $\dbase$ such that every band $(\dbasepoint_i, [0,1])$ on $\gbase$ maps to corresponding point $\dbasepoint_i \in \dbase$ and the band $[0,1]$ determines which pixels are filled in to generate the shape of the line.\note{maybe put in a phantom line plot in s}\label{fig:construction:xi}}
+  \caption{The graphic base space $\gbase$ is collapsable to the line $\dbase$ such that every band $(\dbasepoint_i, [0,1])$ on $\gbase$ maps to corresponding point $\dbasepoint_i \in \dbase$.
+The band $[0,1]$ determines the thickness of a rendered line for a given point $\dbasepoint_{i}$ by specifying how pixels corresponding to that point are colored.
+\label{fig:construction:xi}}
 \end{figure}
+
 For example, as shown in \autoref{fig:construction:xi}, a line is 1D but is a 2D glyph on a screen; therefore the graphic space $\gbase$ is constructed by multiplying the base space $\dbase$ with an interval $[0,1]$.
@@
 The composition in \autoref{eq:construction:q:fabrication} means that different measurable components can yield the same visual elements.
-The composition operators defined in \autoref{sec:artist:composition} can also act on compositors $\vmark$ such that $\vmarkc_{a+b} = \vmark_{a} + \vmark_{b}$ and multiplied d $\vmark_{a\times b} = \vmark_{a}  \vmark_{b}$.
+The composition operators defined in \autoref{sec:artist:composition} can also act on compositors $\vmark$ such that $\vmark_{a+b} = \vmark_{a} + \vmark_{b}$ and multiplied d $\vmark_{a\times b} = \vmark_{a}  \vmark_{b}$.
 
 \begin{figure}[h!]
-  \includegraphics[width=1\columnwidth]{q.png}
+  \includegraphics[width=1\columnwidth]{full_scatter.png}
   \caption{This simple $\vmarkc$ assembles a circular visual element that is the color specified in $\vsection(\dbasepoint)$ and is at the intersection specified in $\vsection(\dbasepoint)$ \label{fig:construction:q}}
@@
 As shown in \autoref{fig:construction:q}, a set of  $\vchannel$ functions individually convert the values in the data record to visual components.
-Then the $\vmarkc$ function combines these visual encodings to produce a graphic section \gsection.
+Then the $\vmark$ function combines these visual encodings to produce a graphic section \gsection.
 When this section is evaluated on the graphic space associated with the data $\gsection(\vindexpre(\dbasepoint))$, it produces a blue circular marker at the intersection of the x and y positions listed in \vsection.
-The composition rule in \autoref{eq:construction:q:fabrication} means that developers can implement $\vmarkc$ as drawing circles or can implement a $\vmarkc$ that draws arbitrary shapes, and then provide different $\vchannelc$ adapters, such as one that specifies that the shape is a circle.
+The composition rule in \autoref{eq:construction:q:fabrication} means that developers can implement $\vmark$ as drawing circles or can implement a $\vmark$ that draws arbitrary shapes, and then provide different $\vchannel$ adapters, such as one that specifies that the shape is a circle.
 
@@
 This allows developers to for example connect transformers that transform data on a line to a color in dataspace, but build a line compositing function that dynamically resamples what is on screen in graphic space.
+\begin{figure}
+  \includegraphics[width=1\columnwidth]{full_line.png}
+  \caption{\label{fig:construction:artist:combo}}
+\end{figure}
 

### 2022-06-10 paper/main.tex
top rework

@@
 \section{Related Work}
-We build on the exentsive work codifying the structure of visualization data to develop a more generalizable method for expressing the structure that visualizations are expected to preserve.
+We build on the extensive work codifying the structure of visualization data to develop a more generalizable method for expressing the structure that visualizations are expected to preserve.
 Broadly, previous work describes structure as a combination of the topological properties of the data, the mathematical structure of the data fields, and equivalent changes to data and graphics.
@@
   \caption{
-    A topological representation of the data (in purple) is a simplified representation of how the records in a dataset are connected.
-The table is a set of independent records; therefore it has a 0D continuity represented by the disconnected dots.
-These records can also be samples from a timeseries, which has a 1D continuity.
-\note{started making a map of the temperature data but was taking a while so I didn't  }
+  A a topological representation of the data (in purple) is a simplified representation of a view of how the records in a dataset are connected.
+The table can be viewed as a set of independent records; therefore it has a 0D continuity represented by the disconnected dots.
+These weather records are also samples of continuous measurements at each location, which can be modeled as the time series measurements being samples on a 1d continuous interval and the geospatial station measurements as sparse samples on a 2D continuous space.
   \label{fig:related-work:continuity:ktypes}}
@@
 
-Roughly, topology is a formal way of encoding the connectivity of the elements of the dataset \cite{wilkinsonGrammarGraphics2005} and there is an expectation that the visual elements should match this connectivity.
-Formally, a topological space is a set $X$ with a structure called a topology.
-The set $X$ is a collection of points, where these points are \note{getting tripped up on how much of the definition to include here, but this is the natural place to give a definition of topology and opensets}
-Some common topologies, as illustrated in\autoref{fig:related-work:continuity:ktypes}, are point, lines, and surfaces.
-The table is an example of disconnected discrete elements and therefore has the topology of discrete points.
-The timeseries is  sampled from a 1d continuous temperature measurement, and therefore has a topology of a 1D space.
-The map in \autoref{fig:related-work:continuity:ktypes} shows a 2D continuous surface.
-As illustrated in the figure, it is not uncommon for the same dataset to have multiple topologies.
-For the common topologies of discrete points, lines, and surfaces, Bertin recommend that they be represented visually by point, line, and area type visual elements respectively \cite{bertinSemiologyGraphicsDiagrams2011a}, and this expectation has been generalized in the development of visual algoriithms.
-
-Visual algorithms assume the continuity of their data, as described in taxonomies of visualization algorithms Chi and by Troy and M\"{o}ller \cite{chiTaxonomyVisualizationTechniques2000,toryRethinkingVisualizationHighlevel2004} and try to preserve it, but many data structures do not provide a method for encoding the topology of the data they store.
-For example, a \texttt{line} algorithm often take as input a list of (x,y) coordinates.
-The \texttt{line} algorithm assumes that its input is continuous and therefore displays the data as if it was continuous, regardless of the actual data topology.
-The ability to encode the continuity expectations in the visual algorithm and the data container would provide a way of verifying that the assumptions match and therefore that the components are preserving the topological structure.
+There is an assumption in visualization that the connectivity of the elements in a dataset, which is roughly what topology encodes\cite{wilkinsonGrammarGraphics2005}, are preserved in the visual representation.
+Bertin gives the examples of discrete data values should be represented by discrete entities such as points, quantitive continuous data by lines, and surfaces through area marks \cite{bertinSemiologyGraphicsDiagrams2011a}.
+In \autoref{fig:related-work:continuity:ktypes} the records of the GHCN\cite{lawrimoreGlobalHistoricalClimatology2011} weather station table can be considered disconnected from each other and therefore have a 0D continuity.
+This means each distinct row can be reduced to a discrete point and it is expected  that each rows will be encoded as a discrete visual element, such as a standalone scatter point.
+Each station can also be considered a time series of weather observation since temperature and pressure are sampled from a 1D continuous measurement space.
+The connectivity of these observations can be modeled as an interval and the data is encoded as a line since that visual element is also 1D continuous.
+The measurements at set of stations on any given day is a sparse sampling of a 2D geospatial grid and the connectivity can be modeled as a 2D plane.
+Instead of interpolation, which is how continuity was preserved in the timeseries plot, in this instance the connectivity is preserved by plotting the stations in the map such that their relative distances are preserved.
+Topological continuity allows us to encode the connectivity of the data in a uniform manner such that we can then generalize the preservation of that continuity in the visual representation beyond points, lines, areas, and other case by case continuities.
+
+Visual algorithms assume the topology of their input data, as described in taxonomies of visualization algorithms Chi and by Troy and M\"{o}ller \cite{chiTaxonomyVisualizationTechniques2000,toryRethinkingVisualizationHighlevel2004}.
+For example, a \texttt{line} algorithm often does not have a way to query whether a list of (x,y) coordinates is the distinct rows, the time series, or the list of stations in \autoref{fig:related-work:continuity:ktypes}.
+While plotting the time series as a continuous line would be correct, it would be incorrect for a visualization to indicate that the distinct rows or stations are connected in a 1D continuous manner.
+Since topological continuity is generalizable, we propose the construction of topology types such that data can explicitly state its topology type and visual algorithms can check that this type matches their assumptions.
 
@@
 \label{sec:atct}
-In this section, we propose a mathematical abstraction of a data container and graphic prerendered output.
-This mathematical abstraction has a robust language for expressing topology and fields; expresses how to verify that data continuity is preserved on subset, distributed, and streaming data; and formalizes the expectation of a correspondence between data and visual elements.
+In this section, we propose a mathematical abstraction of the data input and graphic prerendered output.
+This mathematical abstraction has a robust language for expressing topology and fields; expresses how to verify that data continuity is preserved on subset, distributed, and streaming data representations; and formalizes the expectation of a correspondence between data and visual elements.
 
@@
 \label{sec:atct:fiber-bundles}
-To bridge the need for flexibility and cohesiveness, we model data using a mathematical representation of data that can encode topological properties, field types, and data values in a way that is both uniform and continuity and dimension independent way.
+We model data using a mathematical representation of data that can encode topological properties, field types, and data values in a uniform.
 This mathematical representation is called a fiber bundle and its use as an abstract data representation in visualization was first proposed by Butler\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}.
@@
 
-The \textcolor{base}{base space} is a topological space \dbasec\ with points $\dbasepoint \in \dbase$ and topology $\mathcal{T}_{\dbase}$; $\mathcal{T}_\dbasepoint$ is a set of opensets $ \dbasepoint \in \openset \subseteqq \dbase$ \cite{munkresElementsAlgebraicTopology1984} that cover the base space \dbasec.
+The \textcolor{base}{base space} is a topological space \dbasec\ with points $\dbasepoint \in \dbase$ and topology $\mathcal{T}_{\dbase}$.
+For our purposes, $\dbase$ is a set of points in a mathematical space-such as a metric space, manifold, or euclidean space-that encodes the underlying connectivity of the dataset.
+For example, $\dbase$ could be the discrete points, interval, or plane in \autoref{fig:related-work:continuity:ktypes}.
+On this set \dbase, we can define subsets of connected elements, called opensets \openset.
+An openset \openset\ is a generalization of an open interval, meaning an openset contains the points inside a bounded region on X and excludes the points on the boundary.
+A topology $\mathcal{T}_{\dbase}$ defined on \dbase\ is a collection of opensets such that the empty set and \dbase are in $\mathcal{T}_{\dbase}$,and every union and intersection of opensets in $\mathcal{T}_{\dbase}$ are also in $\mathcal{T}_{\dbase}$ \cite{TopologyVsTopology,bradleyTopologyCategoricalApproach2020}.
+The topology is encoding the notion that the connectivity is expected to be preserved regardless of how small an openset or large a union of opensets.
 
@@
 \end{equation}
-\note{$\xi_*rho(k)(s)=rho(s) for s in \xi^{-1}(k)$
- and move {} into narrative}
+\note{\ci_*rho(k)(s)=rho(s) for s in \xi^{-1}(k)
+} and move {} into narrative
 Therefore, the continuous map $\vindex$ and transport functors $\vindexpull, \vindexpush$ allow us to express the correspondence between graphic section and data section.

### 2022-06-10 paper/main.tex
equivariance + brought up related work + comments

@@
 
-There is an assumption in visualization that the connectivity of the elements in a dataset, which is roughly what topology encodes\cite{wilkinsonGrammarGraphics2005}, are preserved in the visual representation.
+There is an assumption in visualization that the connectivity of the elements in a dataset are preserved in the visual representation; roughly topology provides a method for encoding this connectivity \cite{wilkinsonGrammarGraphics2005}.
 Bertin gives the examples of discrete data values should be represented by discrete entities such as points, quantitive continuous data by lines, and surfaces through area marks \cite{bertinSemiologyGraphicsDiagrams2011a}.
@@
 Since topological continuity is generalizable, we propose the construction of topology types such that data can explicitly state its topology type and visual algorithms can check that this type matches their assumptions.
+The typing system proposed in this paper builds on Butler's proposal of using a mathematical structure called fiber bundles as an abstract data representation in visualization \cite{butlerVectorBundleClassesForm1992, butlerVisualizationModelBased1989}.
+We sketch out fiber bundles in \autoref{sec:atct:fiber-bundles}, but his work provides a thorough introduction to bundles for visualization practitioners.
 
 \subsection{Equivariance}
-Visualizations that preserve the mathematical structure of the data are called equivariant maps.
-Equivariant maps are maps where every change to an input has an equivalent change in the output.
+Alongside preserving topology, visualization algorithms are expected to preserve the structure in data values when they translate data to visual representations.
+The mathematical term for this is equivariance and an equivariant map is a map that transforms input with structure to output with compatible structure.
 Bertin codified the idea that visual encoding of a data field should match the structure of the field; for example ordered data should be encoded as an ordered visual element or quantitative data as a quantitative visual element \cite{bertinSemiologyGraphicsDiagrams2011a}.
@@
 They propose three specific types of equivariance: that the visualization should not change if the data representation (i.e.
-the data container) changes, that data maps unambiguously to visual elements \cite{ziemkiewiczEmbeddingInformationVisualization2009}, that changing the data should coresspond to changes in the visualization in a perceptually significant manner.
-While the data representation introduced in \autoref{sec:atct:io} dispenses with a distinction between containers and the data, the expectation of unambiguous mappings is expressed in \autoref{sec:atct:xi} and the definition of equivariance introduced in \autoref{sec:artist:equivariance} includes the correspondence between data and visual changes.
+the data container) changes, that data maps unambiguously to visual elements \cite{ziemkiewiczEmbeddingInformationVisualization2009}, that changing the data should correspond to changes in the visualization in a perceptually significant manner.
+In \autoref{sec:atct:sheaves} we introduce a uniform abstract data representation layer, the expectation of unambiguous mappings is expressed in \autoref{sec:atct:xi}, and the definition of equivariance introduced in \autoref{sec:artist:equivariance} includes the correspondence between data and visual changes.
+In this work, we formally specify the conditions necessary to build structure preserving components using category theory.
+A brief visualization oriented introduction to category theory is in Vickers et al \cite{vickersUnderstandingVisualizationFormal2013}, but they are applying category theory to semantic concerns about visualization design rather than library architecture.
 
@@
 \label{sec:related-work:software}
-Visualization libraries are in part measured by how expressive the components of the library are; where expressiveness is a measure of which structure preserving mappings a tool can implement \cite{mackinlayAutomaticDesignGraphical1987}.
+Visualization libraries are in part measured by how expressive the components of the library are, where expressiveness is a measure of which structure preserving mappings a tool can implement \cite{mackinlayAutomaticDesignGraphical1987}.
 While some visualization tools aim to automate the pairing of data with structure preserving visual representations, such as Tableau\cite{StoltePolaris2002,hanrahanVizQL2006,MackinlayShowme2007}, many visualization libraries leave that choice to the user.
@@
 The \textcolor{base}{base space} is a topological space \dbasec\ with points $\dbasepoint \in \dbase$ and topology $\mathcal{T}_{\dbase}$.
-For our purposes, $\dbase$ is a set of points in a mathematical space-such as a metric space, manifold, or euclidean space-that encodes the underlying connectivity of the dataset.
+For our purposes, $\dbase$ is a set of points that encodes the underlying connectivity of the dataset.
 For example, $\dbase$ could be the discrete points, interval, or plane in \autoref{fig:related-work:continuity:ktypes}.
@@
 Since the fiber space is a topological space, it does not really have restrictions on the dimension of the types it can encode or the structure of those types.
-For example color can be represented as the cartesian product of three distinct sets, one for each color, or a three dimensional topological space $\mathbb{R} \times \mathbb{R} \times \mathbb{R} = \times \mathbb{R}^{3}$.
+For example, wind has a speed component and a direction component and we can consider those separately $\dfiber_{speed} = \mathbb{R}, \dfiber_{direction} = \mathbb{R}$ or as one fiber$\dfiber_{wind} = \mathbb{R}^{2}$.
 
@@
 For example, RGB color can either be represented as three fiber fields $\dfiber_{red}, \dfiber_{green}, \dfiber_{blue}$ or a composite fiber field $\dfiber_{red} \otimes \dfiber_{green} \otimes \dfiber_{blue} = \dfiber_{RGB}$.
-
-A brief introduction to category theory for visualization practitioners is in Vickers et al \cite{vickersUnderstandingVisualizationFormal2013}, but they are applying category theory to semantic concerns about visualization design rather than library architecture.
 

### 2022-06-15 paper/main.tex
cleaning up why trivial/non-trivial bundles matter

@@
 
+\note{Spelling out vis library component behavior w/ alg top + cat theory lets us be super general in terms of inputs/outputs (what data, what vis) while super specific about the constraints the components need to satisfy to transform that input to that output in a structure preseving way}
+
 \IEEEPARstart{V}isualizations are expected to be a faithful translation of the inherent structure of the data being visualized; therefore the components of visualization software libraries that translate data to graphics are also expected to preserve this structure.
@@
 Instead of interpolation, which is how continuity was preserved in the timeseries plot, in this instance the connectivity is preserved by plotting the stations in the map such that their relative distances are preserved.
-Topological continuity allows us to encode the connectivity of the data in a uniform manner such that we can then generalize the preservation of that continuity in the visual representation beyond points, lines, areas, and other case by case continuities.
+Encoding connectivity using the formalism of topological spaces allows us to encode the connectivity of the data in a uniform manner such that we can then generalize the preservation of that connectivity in the visual representation rather than constructing a condition specifically for each type of topology.
 
@@
 \label{sec:atct:fiber-bundles}
-We model data using a mathematical representation of data that can encode topological properties, field types, and data values in a uniform.
-This mathematical representation is called a fiber bundle and its use as an abstract data representation in visualization was first proposed by Butler\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}.
+We model data using a mathematical representation of data that can encode topological properties, field types, and data values in a uniform manner using a structure from algebraic topology called a fiber bundle.
+We extend Butler's proposal of bundles as abstract visualization data type \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989} by incorporating Spivak's methodology for encoding named data types from his fiber bundle representation of relational databases \cite{spivakDatabasesAreCategories2010, spivakSIMPLICIALDATABASES}.
+We build on this work to describe how to encode the connectivity of the data as a topological space, separately encode the fields as their own topological space with a typing system, and express the mappings between these two spaces.
+
 A fiber bundle $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and continuous surjective map $\pi: \dtotalc \rightarrow \dbasec$\cite{FiberBundle2020}.
@@
 
+\subsubsection{Connectivity: Base Space \dbase}
 The \textcolor{base}{base space} is a topological space \dbasec\ with points $\dbasepoint \in \dbase$ and topology $\mathcal{T}_{\dbase}$.
@@
 On this set \dbase, we can define subsets of connected elements, called opensets \openset.
-An openset \openset\ is a generalization of an open interval, meaning an openset contains the points inside a bounded region on X and excludes the points on the boundary.
+An openset \openset\ is a generalization of an open interval, meaning an openset contains points inside a bounded region on \dbase and excludes the points on the boundary.
 A topology $\mathcal{T}_{\dbase}$ defined on \dbase\ is a collection of opensets such that the empty set and \dbase are in $\mathcal{T}_{\dbase}$,and every union and intersection of opensets in $\mathcal{T}_{\dbase}$ are also in $\mathcal{T}_{\dbase}$ \cite{TopologyVsTopology,bradleyTopologyCategoricalApproach2020}.
@@
 
+\subsection{Field Types: Fiber Space \dfiber}
 The \textcolor{fiber}{fiber space} is a topological space \dfiberc\ that is the preimage of the projection function $\pi$ at a point in the base space $\dfiber_{\dbasepoint} = \pi^{-1}(\dbasepoint)$, meaning that all points in a fiber over a space $\dfiber_{\dbasepoint}$ map to the same point $\dbasepoint$.
-Fibers of a bundle are isomorphic $\dfiberc \simeq \dfiberc_{\dbasepointc}\;\forall \dbasepointc \in \dbase$, which means that the space of possible values is the same set of values over all points in the base space.
-The fiber space can be used to encode the space of all possible values of a data field, which is analogous to a CS data type, and the fibers can be named, as described by Spivak \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES}.
-
+The fiber space can be used to encode the space of all possible values of a data field and the fibers can be named, as described by Spivak \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES}.
+The space of all possible values is analogous to a CS data type; therefore the fiber space can be used to encode the data types of the data fields.
+This set theoretic formulation of field types also supports encoding valid operations on the type as part of the type, akin to methods on an object, as part of the field structure.
+For example, the Steven's measurement types are defined by the mathematical group structure on the measurment fields \cite{stevensTheoryScalesMeasurement1946,leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}, and often serve as the basis of the equivariance described in \autoref{sec:related-work:equivariance}.
+
+\subsection{Data: Section \dsection}
 The \textcolor{section}{sections} $\dsection$ of a bundle are continuous functions that bind the base space with a set of values in the total space $\dtotal$ and $\textcolor{set}{\Gamma}$ denotes the set of all sections in a bundle over an openset.
@@
 \end{equation}
-Each section function takes as input a point in the basespace $\dbasepoint$ and returns a single point in the fiber.
-Extending Spivak's fiber bundle description of a table, we define the values returned by a section as named and typed records.
-Since the fiber space is a topological space, it does not really have restrictions on the dimension of the types it can encode or the structure of those types.
-For example, wind has a speed component and a direction component and we can consider those separately $\dfiber_{speed} = \mathbb{R}, \dfiber_{direction} = \mathbb{R}$ or as one fiber$\dfiber_{wind} = \mathbb{R}^{2}$.
-
-The base space and the section evaluated on the base space form a sort of key-value pair, as described by Munzner \cite{munznerVisualizationAnalysisDesign2014}.
-She imbues semantic meaning to the topology, wherein the key is an independent attribute used as a look up key to retrieve the value in the dataset.
-She defines the value as a dependent attribute.
-In this work we use a more general notion of the topological space as an indexing space for the data records.
-Therefore, we do not impose any semantic constraints on what the fiber and base space encode.
-For example, a data cube of temperature can be encoded such that every fiber is a set of 2D temperature surfaces $\dfiber_{tempsurf}$, the base space is a 1D continuous interval, and each section returns a surface $\dsection(\dbasepoint) = f(x,y)$ where x and y are already parameterized.
-Alternatively, the fiber can encode each cell of the data cube $\dfiber_{lat} \times \dfiber_{lon} \times \dfiber_{temp}$, the base space could be a 3D volume, and each section returns a $\dsection(\dbasepoint) = \{lat, lon, temp\}$ record.
-
-Fiber bundles where the fibers are not identical are called non-trivial bundles.
-Non-trivial bundles allow us to generalize bundles to many different types of total spaces.
-We can also decompose non-trivial bundles into trivial bundles, which are bundles where all the fibers are identical.
-This is because all fiber bundles are locally trivial, which means that for a small enough subset of the data $\opensetc \subseteq \dbasec$, all the fibers are equivalent $\pi^{-1}(\opensetc)\xrightarrow{\equivc} \opensetc \times \dfiberc$.
-A bundle may be globally trivial, meaning that $\dtotalc = \dbasec \times \dfiberc$, and generally can be thought of as a twisted product of total and base spaces \cite{munkresElementsAlgebraicTopology1984}.
-Using trivial and locally trivial bundles allows us to encode the possible data records in a bundle as a single fiber $\dfiber$ type rather than having to maintain a fiber type for each point \dbasepoint.
+Each section function takes as input a point in the basespace  and returns a single point in the fiber  $\dsection: \dbasepoint \rightarrow \dfiber_{\dbasepoint}$.
+In a trivial bundle $\dtotalc = \dbasec \times \dfiberc$ such that all fibers are equal $\dfiberc \equiv \dfiberc_{\dbasepointc}\;\forall \dbasepointc \in \dbase$; therefore we can define a global section $\dsection \in \Gamma(\dbase, \dfiber)$.
+In non-trivial (general) bundles, fibers are isomorphic $\dfiberc \simeq \dfiberc_{\dbasepointc}\;\forall \dbasepointc \in \dbase$ but not necessarily equal, which means that we usually cannot define a global section.
+Instead, all fiber bundles are locally trivializable, $\dtotal\restriction_{\openset_{k}} = \openset_{k}\times \dfiber$ for all open neighborhoods $\dbasepoint \openset \dbase$ for all points $\dbasepoint \in \dbase$.
+This allows us to define sections locally, meaning $\dsection\restriction_{\openset} \in \Gamma(\openset_{\dbasepoint}, \dtotal\restriction_{\openset_{\dbasepoint}})$.
+All data can be expressed as a collection of local sections, which are maps from the indexing space \dbase\ to the field space \dbase, and data that is embedded in trivial bundles can be expressed as a single section.
 
@@
 One section over this 1D continuous space is a \texttt{sin} function that takes as input a base point $\dbasepoint_{i}$ and returns a point in the fiber $sin(\dbasepoint_{i})$.
+
+\subsection{intrepreation}
+Extending Spivak's fiber bundle description of a table, we define the values returned by a section as named and typed records.
+Since the fiber space is a topological space, it does not really have restrictions on the dimension of the types it can encode or the structure of those types.
+For example, wind has a speed component and a direction component and we can consider those separately $\dfiber_{speed} = \mathbb{R}, \dfiber_{direction} = \mathbb{R}$ or as one fiber$\dfiber_{wind} = \mathbb{R}^{2}$.
+The base space and the section evaluated on the base space form a sort of key-value pair, as described by Munzner \cite{munznerVisualizationAnalysisDesign2014}.
+She imbues semantic meaning to the topology, wherein the key is an independent attribute used as a look up key to retrieve the value in the dataset.
+She defines the value as a dependent attribute.
+In this work we use a more general notion of the topological space as an indexing space for the data records.
+Therefore, we do not impose any semantic constraints on what the fiber and base space encode.
+For example, a data cube of temperature can be encoded such that every fiber is a set of 2D temperature surfaces $\dfiber_{tempsurf}$, the base space is a 1D continuous interval, and each section returns a surface $\dsection(\dbasepoint) = f(x,y)$ where x and y are already parameterized.
+Alternatively, the fiber can encode each cell of the data cube $\dfiber_{lat} \times \dfiber_{lon} \times \dfiber_{temp}$, the base space could be a 3D volume, and each section returns a $\dsection(\dbasepoint) = \{lat, lon, temp\}$ record.
 
@@
 \label{sec:atct:sheaves:measurments}
-We define structure on data as the allowable \textcolor{action}{transformations on the bundle}, which we formalize as actions $\dfuncc$ on sections of that sheaf.
-Actions  $act: X \times Y \tightarrow Y$ are elements of one set $X$ applied to elements of another set $Y$ such that the action in X applied to Y returns an element in Y\cite{nlab:action}.
-Actions are a generalization of applying a function to an element and returning an element of the same type.
-Our sheaf action  $\dfunc$ is a generalization of the classification of measurement scales by the actions allowed on the data values\cite{stevensTheoryScalesMeasurement1946,leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}, which often serves as the basis of the equivariance described in \autoref{sec:related-work:equivariance}.
+Equivariance in a visual context was formally defined as the preservation of a binary operator from one domain to another by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}.
+We generalize binary operations to actions $\dfunc \in \dfuncset$ on elements of the bundle $\delement \in \dtotal$.
+This is expressed with a composition function  $act: \dfuncset \times E \tightarrow \dtotal$ that applies the action to the element such that $\dfunc(\delement_0) = \delement_1, \delement_0, \delement_1 \in \dfunc$.
+We generalize to a family of actions because that allows for expanding the set of allowable transformations on the data beyond a single operator.
 
 We separate data transformations into two types, transformations on the base space $(\dfunchc, \dfuncpullc)$ and transformations on the fiber space $\dfunctc$.
-These diagrams are seperated because the base space transformation and the pullback are functors on the openset object and associated set of sections, while the fiber transformation takes a single fiber to a different fiber in the set of sections.
+These diagrams are separated because the base space transformation and the pullback are functors on the openset object and associated set of sections, while the fiber transformation takes a single fiber to a different fiber in the set of sections.
 

### 2022-06-24 paper/main.tex
more responding to feedback

@@
 
-\subsection{Topology}
+\subsection{Topological Structure}
 \label{sec:related-work:continuity}
@@
 
-\subsection{Equivariance}
-Alongside preserving topology, visualization algorithms are expected to preserve the structure in data values when they translate data to visual representations.
+\subsection{Equivariant Transformations}
+Along with preserving topology, visualization algorithms are expected to preserve the structure of the data fields, i.e.
+the column types in a table, when they translate data to visual representations.
 The mathematical term for this is equivariance and an equivariant map is a map that transforms input with structure to output with compatible structure.
@@
 
-A fiber bundle $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and continuous surjective map $\pi: \dtotalc \rightarrow \dbasec$\cite{FiberBundle2020}.
-
+A fiber bundle $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and continuous surjective map $\pi: \dtotalc \rightarrow \dbasec$ \cite{FiberBundle2020}.
 \begin{equation}
@@
   \dbasec
-     \arrow[u, "\dsectionc"', bend right, pos=.5, color=section]
+     \arrow[u, "\dsectionc"', bend right, pos=.5, color=section, dashed]
   \end{tikzcd}
@@
 
-\subsubsection{Connectivity: Base Space \dbase}
-The \textcolor{base}{base space} is a topological space \dbasec\ with points $\dbasepoint \in \dbase$ and topology $\mathcal{T}_{\dbase}$.
-For our purposes, $\dbase$ is a set of points that encodes the underlying connectivity of the dataset.
-For example, $\dbase$ could be the discrete points, interval, or plane in \autoref{fig:related-work:continuity:ktypes}.
-On this set \dbase, we can define subsets of connected elements, called opensets \openset.
-An openset \openset\ is a generalization of an open interval, meaning an openset contains points inside a bounded region on \dbase and excludes the points on the boundary.
-A topology $\mathcal{T}_{\dbase}$ defined on \dbase\ is a collection of opensets such that the empty set and \dbase are in $\mathcal{T}_{\dbase}$,and every union and intersection of opensets in $\mathcal{T}_{\dbase}$ are also in $\mathcal{T}_{\dbase}$ \cite{TopologyVsTopology,bradleyTopologyCategoricalApproach2020}.
-The topology is encoding the notion that the connectivity is expected to be preserved regardless of how small an openset or large a union of opensets.
+ $\textcolor{set}{\Gamma}$ denotes the set of all sections. The set of sections $\dsection$ of a bundle are continuous functions 
+\begin{equation}
+  \label{eq:atct:fb:sections}
+  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}
+\end{equation}
+
+\subsubsection{Topology: Base Space \dbase}
+We encode the underlying connectivity of the data as the \textcolor{base}{base space} \dbase\ of the fiber bundle because it provides a uniform language for describing the topology regardless of how the data is subset.
+The base space \dbase\ is a topological space $(\dbase,\mathcal{T}_{\dbase})$, which means it consists of the set of points ${\dbasepoint \in dbasec}$ and topology $\mathcal{T}_{\dbase}$.
+A topology $\mathcal{T}_{\dbase}$ is a collection of open sets $\openset$ such that the empty set and \dbase are in $\mathcal{T}_{\dbase}$,and every union and intersection of opensets in $\mathcal{T}_{\dbase}$ are also in $\mathcal{T}_{\dbase}$ \cite{bradleyTopologyCategoricalApproach2020}.
+An openset \openset\ is a generalization of an open interval to a region, meaning an open set contains points inside a bounded region on \dbase and excludes the points on the boundary \cite{TopologyVsTopology}.
+A property of the base space \dbase\ is that it is a quotient topology, ,meaning it has the finest topology of $\dbase$, which  means that it contains the largest possible collection of distinct open sets on $\dbase$ \cite{QuotientSpaceTopology2020}.
+For our purposes, $\dbase$ is a set of points that encodes the underlying connectivity of the our dataset, for example, $\dbase$ could be the discrete points, interval, or plane in \autoref{fig:related-work:continuity:ktypes}.
+The topology on $\dbase$ encodes the space of smallest possible subsets of the data, and the union and intersection rules express that these subsets must be combinable into larger subsets of the data.
+
+We encapsulate the notion of a programmatic topology type as the base space category $\mathcal{\dbasec}$.
+The base space category $\mathcal{\dbasec}$ has open set objects $\opensetc$.
+The morphisms of the category are inclusions $\iota: \opensetc_1 \rightarrow \opensetc_2$ such that $\opensetc_1 \subseteq \opensetc_2$.
+By definition of a catagory\cite{barrCategoryTheoryComputing}, morphisms are commutative $(\iota_{1,2}: \openset_{1} \rightarrow \openset_{2}) \circ (\iota_{2,3}: \openset_{2} \rightarrow \openset_{3}) = \iota_{1,3}: \openset_{1} \rightarrow \openset_{3}$.
+Another property of morphisms is that they are associative such that $\iota_{3,4}\circ(\iota_{2,3}\circ\iota_{1,2}) = (\iota_{3,4} \circ \iota{2,3})\circ \iota_{1,2}$ where $\iota_{3,4}:\openset_{3}\rightarrow \openset_{4}$.
+Each object also has an identity morphism $1_{\openset_i}: \openset_i \rightarrow \openset_i$.
+We propose that these categorical types can be mapped to
+programmatic types that describe the data topology and that the inclusion morphisms can be used to assert that the continuity has been preserved when defining subsets of data.
 
 \subsection{Field Types: Fiber Space \dfiber}
-The \textcolor{fiber}{fiber space} is a topological space \dfiberc\ that is the preimage of the projection function $\pi$ at a point in the base space $\dfiber_{\dbasepoint} = \pi^{-1}(\dbasepoint)$, meaning that all points in a fiber over a space $\dfiber_{\dbasepoint}$ map to the same point $\dbasepoint$.
+We encode the field types of a dataset as the  \textcolor{fiber}{fiber space} \dfiberc of the fiber bundle because it allows us to encode fields of varying types and dimensionalities in a consistent manner.
+The fiber \dfiber\ is the preimage of the projection function $\pi$ at a point in the base space $\dfiber_{\dbasepoint} = \pi^{-1}(\dbasepoint)$, meaning that all points in a fiber space $\dfiber_{\dbasepoint}$ map to the same point $\dbasepoint$.
 The fiber space can be used to encode the space of all possible values of a data field and the fibers can be named, as described by Spivak \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES}.
 The space of all possible values is analogous to a CS data type; therefore the fiber space can be used to encode the data types of the data fields.
-This set theoretic formulation of field types also supports encoding valid operations on the type as part of the type, akin to methods on an object, as part of the field structure.
-For example, the Steven's measurement types are defined by the mathematical group structure on the measurment fields \cite{stevensTheoryScalesMeasurement1946,leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}, and often serve as the basis of the equivariance described in \autoref{sec:related-work:equivariance}.
+The fiber is also a topological space, meaning there is no restriction on the dimensionality of the field, thus allowing for the encoding of multi-dimensional fields in the same manner as single dimensional fields.
+
+As with the base space category $\mathcal{\dbase}$, we propose a fiber category $\mathcal{\dfiberc}$ to encapsulate the field types of the data, which is akin to a schema in a database \cite{ullmanFirstCourseDatabase2008}.
+The fiber category $\mathcal{\dfiberc}$ is a monoidal category, meaning the category has a single object $\dfiberc$ of an arbitrary type.
+The morphisms on the fiber object are $\dfunctc \in Hom(\dfiberc, \dfiberc)$.
+The fiber category is equipped with a bifunctor $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$.
+The bifunctor provides a method for combining fibers, thereby allowing for fields that contain multityped values.
+For example, wind can be represented as two fields $\dfiber_{speed} \times \dfiber_{direction}$ or a composite fiber field $\dfiber_{speed} \otimes \dfiber_{direction} = \dfiber_{wind}$.
+The $\otimes$ encapsulates both the sets associated with each fiber $\mathbb{R} \times \mathbb{R} = \mathbb{R}^{2}$ and the morphisms associated with each functor $(\dfunct_{speed}, \dfunct_{direction}) = \dfunct_{wind}$.
+Defining the field schema as a monoidal category allows us to formally express the structure of the field, as defined by its sets and functions, at different levels of composition as needed by different visualization tasks.
 
 \subsection{Data: Section \dsection}
-The \textcolor{section}{sections} $\dsection$ of a bundle are continuous functions that bind the base space with a set of values in the total space $\dtotal$ and $\textcolor{set}{\Gamma}$ denotes the set of all sections in a bundle over an openset.
-
-\begin{equation}
-  \label{eq:atct:fb:sections}
-  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}
-\end{equation}
-Each section function takes as input a point in the basespace  and returns a single point in the fiber  $\dsection: \dbasepoint \rightarrow \dfiber_{\dbasepoint}$.
-In a trivial bundle $\dtotalc = \dbasec \times \dfiberc$ such that all fibers are equal $\dfiberc \equiv \dfiberc_{\dbasepointc}\;\forall \dbasepointc \in \dbase$; therefore we can define a global section $\dsection \in \Gamma(\dbase, \dfiber)$.
-In non-trivial (general) bundles, fibers are isomorphic $\dfiberc \simeq \dfiberc_{\dbasepointc}\;\forall \dbasepointc \in \dbase$ but not necessarily equal, which means that we usually cannot define a global section.
-Instead, all fiber bundles are locally trivializable, $\dtotal\restriction_{\openset_{k}} = \openset_{k}\times \dfiber$ for all open neighborhoods $\dbasepoint \openset \dbase$ for all points $\dbasepoint \in \dbase$.
-This allows us to define sections locally, meaning $\dsection\restriction_{\openset} \in \Gamma(\openset_{\dbasepoint}, \dtotal\restriction_{\openset_{\dbasepoint}})$.
-All data can be expressed as a collection of local sections, which are maps from the indexing space \dbase\ to the field space \dbase, and data that is embedded in trivial bundles can be expressed as a single section.
-
+We encode data as a \textcolor{section}{section} \dsectionc\ of a bundle because this allows us to incorporate the topology and field types in the data definition.
+Sections  $\dsection: \dbasepoint \mapsto \delement$ map from a point in the basespace $\dbasepoint \in \dbase$ to a record in the fiber space $\delement \in \dfiber_{\dbasepoint}$.
+We can therefore model data as structures that map from an index like point $\dbasepoint$ to a data record $\delement$.
+Bundles can have multiple sections, as denoted by $\Gamma(\openset, \dtotal\restriction{\openset})$ which encapsulates the notion of multiple datasets with a shared fiber and base space.
+For example, all functions on the interval [0, 1] or a set of database tables with the same schema.
+
+In a trivial bundle all fibers are equal $\dfiberc \equiv \dfiberc_{\dbasepointc}\;\forall \dbasepointc \in \dbase$, which means that $\dtotalc = \dbasec \times \dfiberc$ .
+This allows us to define global sections $\dsection: \dbase \rightarrow \dfiber \in \Gamma(\dbase, \dfiber)$ which we translate into a data signature of the form
+\begin{equation}
+  \textcolor{section}{\texttt{dataset}}: \textcolor{base}{\texttt{topology}} \rightarrow \textcolor{fiber}{\texttt{field}}
+\end{equation}
+where $\dsection=\texttt{dataset}$, $\dbase=\texttt{topology}$ and $\dfiber=\texttt{fields}$. This type signature encapsulates the topology and field types of the data without much loss of generality.
+
+Non-trivial bundles allow us to generalize even further.
+In a non-trivial (general) bundles, fibers are isomorphic $\dfiber \simeq \dfiber_{\dbasepoint}\;\forall \dbasepoint \in \dbase$, which means there is no single fiber type $\dfiber$.
+Instead, we can define sections locally $\dsection\restriction_{\openset} \in \Gamma(\openset_{\dbasepoint}, \dtotal\restriction_{\openset_{\dbasepoint}})$, which means a section defined only on an open neighborhood $\dbasepoint \in \openset \in \dbase$.
+This is because all fiber bundles are locally trivializable, $\dtotal\restriction_{\openset_{k}} = \openset_{k}\times \dfiber$ such that there's always a subset where the fibers are equal.
+This means almost all data can be expressed as a collection of local sections $\{\dsection\restriction_{\openset_{\dbasepoint}}| \dbasepoint \in \dbase \}$ which translate to a set of signatures
+\begin{equation}
+  \begin{split}
+\{\textcolor{section}{\texttt{data-subset}}: \textcolor{base}{\texttt{topology}} \rightarrow \textcolor{fiber}{\texttt{fields}} \\
+ s.\;t.\;\textcolor{section}{\texttt{data-subset}} \in \texttt{dataset}\}  
+  \end{split}
+\end{equation}
+that can be glued together on overlaps to generate the dataset. We discuss this gluing in \autoref{sec:atct:sheaves}.
+
+\subsection{intrepretation/example}
 \begin{figure}[h!]
-  \centering
-  %needs cropping
-  \includegraphics[trim={2cm 2cm 2cm 2cm}, clip]{fbtau.png}
-  \caption{A non-trivial fiber bundle is bundle where the (by definition isomorphic) fibers change along the indexing base space.
-Here $\dfiber_{\dbasepoint_1}$ and $\dfiber_{\dbasepoint_2}$ go in reverse directions and the global section $sin$ returns $sin(\dbasepoint_{i})$ which is a value between $[0,1]$
-  \label{fig:atct:stripbundle}}
+  \includegraphics[width=\columnwidth]{fbnet.png}
+  \caption{\label{fig:atct:networkbundle}}
 \end{figure}
-
-The bundle in \autoref{fig:atct:stripbundle} is a fiber bundle representation of a mobius band, which is the total space \dtotal\ of the bundle.
-Each fiber \dfiber\ collapses to a point in the base space via the projection function $\pi$.
-The bundle is non-trivial because the fibers are not identical $[-1, 1] \not \equiv [1, 1]$.
-One section over this 1D continuous space is a \texttt{sin} function that takes as input a base point $\dbasepoint_{i}$ and returns a point in the fiber $sin(\dbasepoint_{i})$.
-
-\subsection{intrepreation}
+\note{figure out picture/example}
 Extending Spivak's fiber bundle description of a table, we define the values returned by a section as named and typed records.
@@
 In this work we use a more general notion of the topological space as an indexing space for the data records.
-Therefore, we do not impose any semantic constraints on what the fiber and base space encode.
 For example, a data cube of temperature can be encoded such that every fiber is a set of 2D temperature surfaces $\dfiber_{tempsurf}$, the base space is a 1D continuous interval, and each section returns a surface $\dsection(\dbasepoint) = f(x,y)$ where x and y are already parameterized.
 Alternatively, the fiber can encode each cell of the data cube $\dfiber_{lat} \times \dfiber_{lon} \times \dfiber_{temp}$, the base space could be a 3D volume, and each section returns a $\dsection(\dbasepoint) = \{lat, lon, temp\}$ record.
-
-\begin{figure}[h!]
-  \includegraphics[width=\columnwidth]{fbnet.png}
-  \caption{The fibers over the discrete points $\dbase_{1}, \dbase_{2}$  encode the fields in the temperature table and the fiber over the continuous line $\dbase_{3}$ encodes the fields in the distance table \autoref{fig:related-work:continuity:ktypes}.
-Those tables can be modeled as sections of the bundle in this figure, where the section evaluated at a point in the base space returns the associated row, and on the edge is a constant function that returns the distance.
-  \label{fig:atct:networkbundle}}
-\end{figure}
-
-In \autoref{fig:atct:networkbundle}, the network from \autoref{fig:related-work:continuity:ktypes} is encoded as a section of a bundle.
-This bundle is constructed from a union of bundles over discrete points $\dbase_{0}, \dbase_{2}$ and a 1D continuous connecting line $\dbase_{1}$ that connects these two points.
-The fiber over the data points $\dfiber_{node}$ expresses that the name column type is a subset of \textttt{string}, the temperature column type is \texttt{float}, and the precipitation column type is positive \texttt{float}.
-The fiber over the network edge encodes that the distance from the edge table has type \texttt{float}.
-The data section evaluated at a discrete point $\dbase_{0}$ returns the values for the first row in the table.
-Evaluating the function on $\dbase_{2}$ would return a different row in the table.
-The section $\dsection$ knows which nodes $\dbase_{1}$ is between and so returns the corresponding data value.
-The network edge function is constant and returns the same value any where along $\dbase_{1}$.
-
-The only constraint on sections is that they are (piecewise) continuous and that they return values in the fiber.
-They can be constant or non-constant and they can be defined over a single type of topology or a union of topologies.
-The generality in base space, fiber space, and section makes fiber bundles a unified dimension and type independent data abstraction.
-Therefore fiber bundles provide guidance on constructing a structure aware data interface layer that can support most of the common visualization data set structures, such as continuous functions, tables, data cubes, images and networks.
 
@@
 \end{equation}
+The total space \gtotalc\ is an abstraction of an ideal (infinite resolution) space into which the graphic can be rendered.
+The base space \gbasec\ is a parameterization of the display area, for example the inked bounding box in cairo \cite{CairographicsOrg}.
+The fiber space \gfiberc\ is an abstraction of the renderer fields, for example $\{x,\,y,\,r,\,g,\,b,\,a\}$.
+
+As with data, we model the graphic generating functions as sections $\gsection$ of the graphic bundle 
 \begin{equation}
@@
 \end{equation}
-The base space \gbasec\ is a parameterization of the display area, for example the inked bounding box in cairo \cite{CairographicsOrg}.
-The fiber space \gfiberc\ is an abstraction of the renderer fields, for example $\{x,\,y,\,r,\,g,\,b,\,a\}$.
-The sections $\gsection$ of the graphic are functions that go from a point in the graphic base to elements in a prerendering space.
+that map from a point in an openset in the graphic space $\gbasepoint \in \opensetg \in \gbase$ to a point in the graphic fiber $\gfiber$.
+The section evaluated on a single point $\gbasepoint$ returns a single graphic record, for example one pixel in an ideal resolution space.
+In our model, the unevaluated graphic section us passed to a renderer to generate graphics.
 
@@
   \includegraphics[width=1\columnwidth]{fb_rho.png}
-  \caption{For a 2D display a section $\gsection$ maps into a fiber $\dfiber$ that encodes an RGB prerender space with infinite resolution $\mathcal{R}^{2}$.
+  \caption{For a 2D display a section $\gsection$ maps from each point $\gbasepoint$ into a fiber $\gfiber_{\gbasepoint}$ that encodes an RGB prerender space with infinite resolution $\mathcal{R}^{2}$.
 Each tiny colored box is an approximation of the return value of the same section function $\gsection$ evaluated on different points $\gbasepoint \in \gbase$ in the base space.
+We represent all $\gfiber_{\gbasepoint}$ as the fiber $\gfiber$ because a 2D bundle is a trivial bundle so $\gfiber_{\gbasepoint} = \gfiber$ for all \gbasepoint.
   \label{fig:atct:fb:graphic}}
@@
 The set of all pixels returned by a section evaluated on a given visual base space $\gsection|_{\gbase}$ can yield a visual element, such as a marker, line, or piece of a glyph.
-
-\subsection{Topology and Field Categories}
-\label{sec:atct:io}
-We define categorical types for the topology and the fiber because this allows us to describe  how objects that encode topology and fiber should be specified, constrained, and composed \cite{wielsManagementEvolvingSpecifications1998}, which we can then translate to the implementation of these objects.
-In particular, the categorical construction allows us to bundle the topology and fiber as objects together with the changes to these objects, which are termed morphisms, that the visualization is expected to preserve.
-
-\begin{definition}
-  \label{sec:atct:io:base}
-  The base space category $\mathcal{\dbasec}$ has open set objects $\opensetc$ and inclusion morphisms $\iota: \opensetc_1 \rightarrow \opensetc_2$ such that $\opensetc_1 \subseteq \opensetc_2$. 
-\end{definition}
-
-By definition, each object in a category has an identity morphism and morphisms are associative and commutative \cite{barrCategoryTheoryComputing}.
-
-\begin{definition}
-  \label{def:atct:io:fiber}
-  The fiber category $\mathcal{\dfiberc}$ is a monoidal category, meaning the category has a single object $\dfiberc$ of an arbitrary type.
-The morphisms on the fiber object are $\dfunctc \in Hom(\dfiberc, \dfiberc)$.
-\end{definition}
-
-The fiber category is equipped with a bifunctor $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$.
-The bifunctor provides a method for combining fibers, thereby allowing for fields that contain multityped values.
-For example, RGB color can either be represented as three fiber fields $\dfiber_{red}, \dfiber_{green}, \dfiber_{blue}$ or a composite fiber field $\dfiber_{red} \otimes \dfiber_{green} \otimes \dfiber_{blue} = \dfiber_{RGB}$.
+\autoref{fig:atct:fb:graphic} illustrates a highly idealized space with no overlaps.
+To support overlaps, we can add another fiber element $\gfiber_{z}$ and leave it to the renderer to appropriately blend layers.
 
@@
 Presheaves are functors, meaning they are functions from objects of one category to objects of another category\cite{WhatFunctorDefinitions} that preserve the morphisms in each category. 
-By definition, presheaves are maps from objects in an arbitrary category to an object in the category \setb\cite{nlab:presheaf}.
-Here the arbitrary category is the base space category and the object in \setb is the set of sections
-
+A presheaf is a contravariant functor from an object in an arbitrary category to an object in the category \setb\cite{nlab:presheaf}. 
 \begin{equation}
@@
 \end{equation}
-
-Presheaves are a contraviant functor, which means that the inclusion morphisms $\iota$ between base space objects go in the opposite direction from the restriction morphisms $\iota^*$ between objects in set.
-\note{maybe unpack each piece of this equation into a seperate equation/walk through, from U to fiber when trivial}  
-\begin{equation}
-  \label{eq:atct:presheaf}
-  \begin{tikzcd}
-    \cgamma{\opensetc_1}{\dtotalc\restriction_{\opensetc_1}}  &  & \cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}} 
-    \arrow[ll, "\iota^*"', maps to, color=set] \\
-    & & \\
-    \opensetc_1 
-    \arrow[rr, "\iota", maps to, color=base] 
-    \arrow[uu, "{\sheafc_{\dbasec, \dtotalc}}", maps to, color=sheaf] &  & \opensetc_2 
-    \arrow[uu, "{\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf]              
-    \end{tikzcd}
-\end{equation}
-A presheaf describes the expectation that any set of sections defined over a larger space $\dsection \in \Gamma(\opensetc_2, \dtotal\restriction_{\openset_{2}})$ is also defined over subspaces of that space $\dsection\restriction_{\opensetc_{1}}$.
+The presheaf expresses that every subset of the data $\openset$ has an associated set of data sections $\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$.
+The presheaf is a contravariant functor because the morphisms between input objects
+\begin{equation*}
+  \iota: \openset_1 \rightarrow \openset_2
+\end{equation*}
+go in the opposite direction from the restriction morphisms $\iota^*$ between the sets
+\begin{equation*}
+  \iota^*: \cgamma{\openset_2}{\dtotal_{\openset_2}} \rightarrow \cgamma{\openset_1}{\dtotal_{\openset_1}}
+\end{equation*}
+which are the outputs of the sheaf $\sheaf(\openset) = \cgamma{\openset}{\dtotal\restriction_{\openset}}$.
+We use a  presheaf to describe the expectation that any set of sections defined over a larger space $\dsection \in \Gamma(\opensetc_2, \dtotal\restriction_{\openset_{2}})$ is also defined over subspaces of that space $\dsection\restriction_{\opensetc_{1}}$.
 
@@
 We generalize binary operations to actions $\dfunc \in \dfuncset$ on elements of the bundle $\delement \in \dtotal$.
-This is expressed with a composition function  $act: \dfuncset \times E \tightarrow \dtotal$ that applies the action to the element such that $\dfunc(\delement_0) = \delement_1, \delement_0, \delement_1 \in \dfunc$.
+This is expressed with a composition function  $act: \dfuncset \times E \rightarrow \dtotal$ that applies the action to an element of \dtotal\ such that $\dfunc(\delement_0) = \delement_1, \delement_0, \delement_1 \in \dfunc$.
 We generalize to a family of actions because that allows for expanding the set of allowable transformations on the data beyond a single operator.
 
-We separate data transformations into two types, transformations on the base space $(\dfunchc, \dfuncpullc)$ and transformations on the fiber space $\dfunctc$.
+We separate data transformations into two components, transformations on the base space $(\dfunchc, \dfuncpullc)$ and transformations on the fiber space $\dfunctc$.
 These diagrams are separated because the base space transformation and the pullback are functors on the openset object and associated set of sections, while the fiber transformation takes a single fiber to a different fiber in the set of sections.
@@
 where $\dsection, \dsection^{\prime} \in \Gamma(\openset^{\prime}, \dfuncpull\dtotal\restriction_{\openset^{\prime}})$.
-Examples of a fiber transformation are actions such as partial ordering for ranking purpose \cite{bruggemannRankingPrioritizationMultiindicator2011} and the permutation, ordering, translation, and scaling codified as Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946}.
 Since there are many different fiber transformations,  there is not a uniform way of expressing how this morphism acts on sections unlike in \label{eq:atct:morphism:verify_base}.
+Examples of a fiber transformation are actions such as partial ordering for ranking purpose \cite{bruggemannRankingPrioritizationMultiindicator2011} and the permutation, ordering, translation, and scaling codified as Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946m, leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}.
 
@@
 \label{sec:atct:xi}
-A property of sheaves is that, given two sheaves, if there exists a map between the base spaces of each sheave, then we can construct transport functors that can move each sheaves over the other base space.
-This allows us to move data over the corresponding graphic base space and graphics over the corresponding data base space.
-The data sections defined in \autoref{eq:atct:fb} are sections of sheaves, introduced in \autoref{sec:atct:sheaves}, over open sets $\openset \subseteq \dbase$ in a base space \dbase\ that encode the topology of the data topology.
-The graphic sections, defined in \autoref{eq:atct:fb_graphic_section}, are sections of sheaves over open sets $\opensetg \subseteq \gbase$ in a base space \gbase\ that encodes the topology of the graphic.
-We propose that there exists a continuous \textcolor{functor}{map between topological spaces} \vindexc
+There is an expectation that for a visualization to readable, the visual elements must correspond to distinct data elements\cite{ziemkiewiczEmbeddingInformationVisualization2009} and we can use the properties of sheaves to formally express this correspondence.
+First we establish that the graphic continuity corresponds to the data continuity by proposing that there exists a continuous \textcolor{functor}{map between topological spaces} \vindexc
 \begin{equation}
@@
 
-Given the map between spaces \vindex\, a property of sheaves is that there exist image functors\cite{ImageFunctorsSheaves2021} functors \vindexpull\ and \vindexpush.
+Given the map between spaces \vindex\, a property of sheaves is that there exist image functors\cite{ImageFunctorsSheaves2021} functors \vindexpull\ and \vindexpush that can transport sheaves over other base spaces.
 The \textcolor{functor}{pullback functor} $\vindexpullc$ sends sheaves over $\dbase$ to $\gbase$ and the \textcolor{functor}{pushforward functor} $\vindexpushc$  sends sheaves over $\gbase$ to $\dbase$
@@
 \end{equation}
-\note{\ci_*rho(k)(s)=rho(s) for s in \xi^{-1}(k)
-} and move {} into narrative
+\note{$\xi_*rho(k)(s)=rho(s) for s in \xi^{-1}(k)
+$} and move {} into narrative
 Therefore, the continuous map $\vindex$ and transport functors $\vindexpull, \vindexpush$ allow us to express the correspondence between graphic section and data section.

### 2022-06-24 matplottoy/lib/bar.py
old code changes

@@ -63,7 +63,7 @@ class TopologicalArtist:
     def compose_with_tau(self, section): #doesn't need to be there
         # do I care about bounds yet? or only at draw time?
         new = copy.copy(self)
-        new.data = section
+        new.section = section
         return new
 
     def query(self, data_bounds:Optional[dict]=None, sampling_rate:Optional[int]=None)->dict:
@@ -98,7 +98,7 @@ class Bar(Rho):
         new.V[p] = FtoV(nu=nu, f=f, nu_inv=nu_inv) 
         return new
 
-    def mu(self, tau_restricted: Records)->Records: #draw
+    def nu(self, tau_restricted: Records)->Records: #draw
         #required nus, also done first to get length
         mus = {p: self.V[p].nu(*(tau_restricted[c] 
                 for c in self.V[p].f)) for p in self.P_required}
@@ -115,20 +115,25 @@ class Bar(Rho):
                 mus[p] = self.V[p].nu(*(tau_restricted[c] 
                                         for c in self.V[p].f))
         return mus
-
+    
+    @staticmethod
+    def box_nu(position, width, length, floor):
+        return [mpath.Path([(xi, fl), (xi, fl + yi), 
+                           (xi + wd, fl + yi), 
+                           (xi + wd, fl), (xi, fl)], closed=True) for 
+                (xi, wd, yi, fl) in zip(position, width, length, floor)]
+    
     @staticmethod
     def qhat(position, width, length, floor, facecolor, edgecolor, linewidth, linestyle):
         # are matplotlib 
+        box = Bar.box_nu(position, width, length, floor)
         def fake_draw(render, transform=mtransforms.IdentityTransform()):
-            for (xi, wd, yi, fl, fc, ec, lw, ls) in zip(position, width, length, floor, facecolor, edgecolor, linewidth, linestyle):
+            for (bx, fc, ec, lw, ls) in zip(box, facecolor, edgecolor, linewidth, linestyle):
                 gc = render.new_gc()
                 gc.set_foreground((ec.r, ec.g, ec.b, ec.a))
                 gc.set_dashes(*ls)
                 gc.set_linewidth(lw)
-                path = mpath.Path([(xi, fl), (xi, fl + yi), 
-                           (xi + wd, fl + yi), 
-                           (xi + wd, fl), (xi, fl)], closed=True)
-                render.draw_path(gc=gc, path=path, transform=transform, 
+                render.draw_path(gc=gc, path=bx, transform=transform, 
[…]

### 2022-06-24 paper/main.tex
associativity

@@
 
-\subsection{Equivariant Transformations}
+\subsection{Data Field Structure}
 Along with preserving topology, visualization algorithms are expected to preserve the structure of the data fields, i.e.
@@
 In \autoref{sec:atct:sheaves} we introduce a uniform abstract data representation layer, the expectation of unambiguous mappings is expressed in \autoref{sec:atct:xi}, and the definition of equivariance introduced in \autoref{sec:artist:equivariance} includes the correspondence between data and visual changes.
-In this work, we formally specify the conditions necessary to build structure preserving components using category theory.
+In this work, we formally specify the conditions necessary to build structure preserving components using category theory, because it provides a rich language for expressing the structure of objects, functions between objects, and the constraints that must hold for these functions to compose \cite{wielsManagementEvolvingSpecifications1998,yorgeyMonoidsThemeVariations}.
 A brief visualization oriented introduction to category theory is in Vickers et al \cite{vickersUnderstandingVisualizationFormal2013}, but they are applying category theory to semantic concerns about visualization design rather than library architecture.
@@
 
-\subsubsection{Topology: Base Space \dbase}
+\subsubsection{Topological Structure: Base Space \dbase}
 We encode the underlying connectivity of the data as the \textcolor{base}{base space} \dbase\ of the fiber bundle because it provides a uniform language for describing the topology regardless of how the data is subset.
@@
 
-We encapsulate the notion of a programmatic topology type as the base space category $\mathcal{\dbasec}$.
+We propose that we can express topological structure programmatically by  constructing data types that encapsulates the class of topological structure-.i.e point, line, plane, network, etc.
+These data types can be formally specified as objects of a category $\mathcal{\dbasec}$.
 The base space category $\mathcal{\dbasec}$ has open set objects $\opensetc$.
-The morphisms of the category are inclusions $\iota: \opensetc_1 \rightarrow \opensetc_2$ such that $\opensetc_1 \subseteq \opensetc_2$.
-By definition of a catagory\cite{barrCategoryTheoryComputing}, morphisms are commutative $(\iota_{1,2}: \openset_{1} \rightarrow \openset_{2}) \circ (\iota_{2,3}: \openset_{2} \rightarrow \openset_{3}) = \iota_{1,3}: \openset_{1} \rightarrow \openset_{3}$.
-Another property of morphisms is that they are associative such that $\iota_{3,4}\circ(\iota_{2,3}\circ\iota_{1,2}) = (\iota_{3,4} \circ \iota{2,3})\circ \iota_{1,2}$ where $\iota_{3,4}:\openset_{3}\rightarrow \openset_{4}$.
-Each object also has an identity morphism $1_{\openset_i}: \openset_i \rightarrow \openset_i$.
-We propose that these categorical types can be mapped to
-programmatic types that describe the data topology and that the inclusion morphisms can be used to assert that the continuity has been preserved when defining subsets of data.
-
-\subsection{Field Types: Fiber Space \dfiber}
+The morphisms between objects of the category are inclusions $\opensetc_i \xrightarrow{\iota} \opensetc_j$ such that $\opensetc_i \subseteq \opensetc_j$.
+Given two morphisms $\opensetc_i \xrightarrow{\iota_{i,j}} \opensetc_j$ and $\opensetc_j \xrightarrow{\iota_{j,k}} \opensetc_j$, these morphisms must compose
+\begin{equation}
+  \label{eq:atct:base:cat:comp}
+\begin{tikzcd}
+  \openset_{i} \arrow[rr, "{\iota_{i,j}}"] \arrow[rrrr, "{\iota_{i,k}}", bend right] &  & \openset_{j} \arrow[rr, "{\iota_{j,k}}"] &  & \openset_{k}
+\end{tikzcd}
+\end{equation}
+such that there exists a function $\opensetc_i \xrightarrow{\iota_{i,k}} \opensetc_k$.
+The expected composition of functions is often expressed as commutative diagrams, such as \label{eq:atct:base:cat:comp}.
+Compositions of morphisms are associative
+\begin{align}
+
+ & ((\openset_{i}\xrightarrow{\iota_{i,j}} \openset_{j})) \xrightarrow{\iota_{j,k}} \openset_{k} \xrightarrow{\iota_{k,l}} \openset_{l}\label{eq:atct:cat:as1}
+  \\
+  = & ((\openset_{j}\xrightarrow{\iota_{j,k}} \openset_{k})\xrightarrow{\iota_{k,l}} \openset_{l})
+
+    \openset_{i}\xrightarrow{\iota_{i,j}} \openset_{j}\label{eq:atct:cat:as2}
+\end{align}
+\note{this needs to be sorted cause this format ends up w/ the same order}
+
+such that \autoref{eq:atct:cat:as1} and \autoref{eq:atct:cat:as2}  return the same value. In general associativity allows us flexibility in building call stacks of curried functio
+
+ While here we discuss composition, associativity, and identity in the context of $\mathcal{\dbase}$, these are general properties of categories \cite{riehlCategoryTheoryContext,barrCategoryTheoryComputing,fongInvitationAppliedCategory2019}.
+
+\subsection{Data Fields: Fiber Space \dfiber}
 We encode the field types of a dataset as the  \textcolor{fiber}{fiber space} \dfiberc of the fiber bundle because it allows us to encode fields of varying types and dimensionalities in a consistent manner.

### 2022-07-04 paper/main.tex
all the spivak fiber goodness

--- 2022-06-24 paper/main.tex
@@
 
-\subsection{Equivariant Transformations}
+\subsection{Data Field Structure}
+\label{sec:related-work:equivariance}
 Along with preserving topology, visualization algorithms are expected to preserve the structure of the data fields, i.e.
@@
 
-A fiber bundle $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and continuous surjective map $\pi: \dtotalc \rightarrow \dbasec$ \cite{FiberBundle2020}.
+\begin{definition} A \textbf{fiber bundle} $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and continuous surjective map $\pi: \dtotalc \rightarrow \dbasec$ \cite{FiberBundle2020,spanier1989algebraic}.
 \begin{equation}
@@
 \end{equation} 
-
- $\textcolor{set}{\Gamma}$ denotes the set of all sections. The set of sections $\dsection$ of a bundle are continuous functions 
-\begin{equation}
-  \label{eq:atct:fb:sections}
-  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}
-\end{equation}
-
-\subsubsection{Topology: Base Space \dbase}
+A map $\pi$ is a bundle projection map when 
+\begin{enumerate}
+  \item the \textcolor{fiber}{fiber space} \dfiberc\ is the preimage of the projection function $\pi$ at a point $\dbasepoint$ in the \textcolor{base}{base space} \dbasec\ such that $\dfiber_{\dbasepoint} = \pi^{-1}(\dbasepoint)$
+  \item there is an open neighborhood $\openset_{\dbasepoint}$ surrounding each point in the base space $\dbasepoint \in \openset_{\dbasepoint} \subset \dbase$ such that the \textcolor{total}{total space} \dtotalc\ over the neighborhood is locally trivial $\dtotal\restriction_{\openset} = \openset \times \dfiber$
+\end{enumerate}
+\end{definition}
+The first condition for a fiber bundle means that all points in a fiber space over a point in the base space map to the same base space point, which we rely on to establish that there is a formal binding between typed data value, discussed in \autoref{sec:atct:fiber:fiber} and the corresponding index, discussed in \autoref{sec:atct:fiber:base}.
+The condition of local triviality means that the fibers are identical over open neighborhoods and is what allows fiber bundles to be extremely generalizable, as discussed in \autoref{sec:atct:fiber:sections}
+
+\begin{definition} A \textcolor{section}{\textbf{section}} $\dsectionc: \dbasec \rightarrow \dtotalc$ over a fiber bundle is a smooth right inverse of $\pi$ such that $\pi(\dsection(\dbasepoint)) = \dbasepoint$ for all $\dbasepoint \in \dbase$
+\end{definition}
+In \autoref{sec:atct:fiber:sections}, we propose that modeling data as sections provides a way to encapsulate topological and field structure in a uniform dimension and type independent manner.
+
+\subsubsection{Topological Structure: Base Space \dbase}
+\autoref{sec:atct:fiber:base}
 We encode the underlying connectivity of the data as the \textcolor{base}{base space} \dbase\ of the fiber bundle because it provides a uniform language for describing the topology regardless of how the data is subset.
-The base space \dbase\ is a topological space $(\dbase,\mathcal{T}_{\dbase})$, which means it consists of the set of points ${\dbasepoint \in dbasec}$ and topology $\mathcal{T}_{\dbase}$.
-A topology $\mathcal{T}_{\dbase}$ is a collection of open sets $\openset$ such that the empty set and \dbase are in $\mathcal{T}_{\dbase}$,and every union and intersection of opensets in $\mathcal{T}_{\dbase}$ are also in $\mathcal{T}_{\dbase}$ \cite{bradleyTopologyCategoricalApproach2020}.
-An openset \openset\ is a generalization of an open interval to a region, meaning an open set contains points inside a bounded region on \dbase and excludes the points on the boundary \cite{TopologyVsTopology}.
-A property of the base space \dbase\ is that it is a quotient topology, ,meaning it has the finest topology of $\dbase$, which  means that it contains the largest possible collection of distinct open sets on $\dbase$ \cite{QuotientSpaceTopology2020}.
+The base space \dbase\ is a topological space $(\dbase,\mathcal{T}_{\dbase})$, which means it  consists of the set of points ${\dbasepoint \in \dbase}$ and topology $\mathcal{T}_{\dbase}$ \cite{munkresElementsAlgebraicTopology1984}.
+The set \dbase\ is equipped with open subsets \openset, which are a generalization of an open interval.
+This means opensets contain points inside a bounded region on the topological space and excludes the points on the boundary \cite{TopologyVsTopology}.
+An openset that surrounds a point \dbasepoint\ is called an open neighborhood $\openset_{\dbasepoint}$.
+Opensets are key to constructing topologies, and are used in this paper to model subsets of the data contextualized within the larger continuity those subsets belong to.
+
+\begin{definition}
+A \textbf{topology} $\mathcal{T}_{\dbase}$ on a space \dbase\ is a collection of open sets \openset\ of \dbase\ that satisfy the properties 
+  \begin{itemize}
+    \item the empty set $\varnothing$ and $\dbase$ are in $\mathcal{T}_{\dbase}$
+    \item for all $\openset_i, \openset_j \in \mathcal{T}_{\dbase}$, the intersection $\openset_i \cap \openset_j$  is in $ \mathcal{T}_{\dbase}$
+    \item for all $\openset_i, \openset_j \in \mathcal{T}_{\dbase}$, the union $\openset_i \cup \openset_j$ is in $ \mathcal{T}_{\dbase}$
+  \end{itemize}
+  such that the the topology provides an axiomatic way to construct the topological space $\dbase \coloneqq (\dbase,\mathcal{T}_{\dbase})$\cite{bradleyTopologyCategoricalApproach2020}.
+\end{definition}
+ A property of the base space \dbase\ of a fiber bundle is that it is a quotient topology, which  means that it contains the largest possible collection of distinct opensets, meaning each openset is as small as possible, of $\dbase$ that still preserve the underlying continuity of \dbase \cite{QuotientSpaceTopology2020}.
 For our purposes, $\dbase$ is a set of points that encodes the underlying connectivity of the our dataset, for example, $\dbase$ could be the discrete points, interval, or plane in \autoref{fig:related-work:continuity:ktypes}.
-The topology on $\dbase$ encodes the space of smallest possible subsets of the data, and the union and intersection rules express that these subsets must be combinable into larger subsets of the data.
-
-We encapsulate the notion of a programmatic topology type as the base space category $\mathcal{\dbasec}$.
-The base space category $\mathcal{\dbasec}$ has open set objects $\opensetc$.
-The morphisms of the category are inclusions $\iota: \opensetc_1 \rightarrow \opensetc_2$ such that $\opensetc_1 \subseteq \opensetc_2$.
-By definition of a catagory\cite{barrCategoryTheoryComputing}, morphisms are commutative $(\iota_{1,2}: \openset_{1} \rightarrow \openset_{2}) \circ (\iota_{2,3}: \openset_{2} \rightarrow \openset_{3}) = \iota_{1,3}: \openset_{1} \rightarrow \openset_{3}$.
-Another property of morphisms is that they are associative such that $\iota_{3,4}\circ(\iota_{2,3}\circ\iota_{1,2}) = (\iota_{3,4} \circ \iota{2,3})\circ \iota_{1,2}$ where $\iota_{3,4}:\openset_{3}\rightarrow \openset_{4}$.
-Each object also has an identity morphism $1_{\openset_i}: \openset_i \rightarrow \openset_i$.
-We propose that these categorical types can be mapped to
-programmatic types that describe the data topology and that the inclusion morphisms can be used to assert that the continuity has been preserved when defining subsets of data.
-
-\subsection{Field Types: Fiber Space \dfiber}
-We encode the field types of a dataset as the  \textcolor{fiber}{fiber space} \dfiberc of the fiber bundle because it allows us to encode fields of varying types and dimensionalities in a consistent manner.
-The fiber \dfiber\ is the preimage of the projection function $\pi$ at a point in the base space $\dfiber_{\dbasepoint} = \pi^{-1}(\dbasepoint)$, meaning that all points in a fiber space $\dfiber_{\dbasepoint}$ map to the same point $\dbasepoint$.
-The fiber space can be used to encode the space of all possible values of a data field and the fibers can be named, as described by Spivak \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES}.
-The space of all possible values is analogous to a CS data type; therefore the fiber space can be used to encode the data types of the data fields.
-The fiber is also a topological space, meaning there is no restriction on the dimensionality of the field, thus allowing for the encoding of multi-dimensional fields in the same manner as single dimensional fields.
-
-As with the base space category $\mathcal{\dbase}$, we propose a fiber category $\mathcal{\dfiberc}$ to encapsulate the field types of the data, which is akin to a schema in a database \cite{ullmanFirstCourseDatabase2008}.
+We express the allowable ways to subset and compose points such that the connectivity is preserved as the topology $\mathcal{T}_{\dbase}$ on the data continuity.
+
+The topological structure of the data can be expressed programmatically by constructing data types that encapsulate the class of topological structure-.i.e point, line, plane, network, etc.
+We propose that these data types can be formally specified as objects of a category.
+\begin{definition} 
+   An arbitrary \textbf{category} $\mathcal{C}$ consists of the following \textit{data}: 
+\begin{enumerate}
+  \item \textit{object}: X
+  \item \textit{identity morphism}: $X \xrightarrow {id_x} X$
+  \item \textit{morphisms}: $X \xrightarrow{f} Y$ between objects X, Y in $\mathcal{C}$
+  \item \textit{composite morphism}: given the morphisms $X \xrightarrow{f} Y$ and $Y \xrightarrow{g} Z$, there exists a morphism $X \xrightarrow{g\circ f} Z$
+\end{enumerate}
+The data must satisfy the following \textit{rules}
+\begin{enumerate}
+  \item \textit{unitality:} for every morphism $ X \xrightarrow{f} Y$, there exists identity morphisms $id_x, id_y$ such that $f \circ id_x = f = id_y \circ f$
+  \item \textit{associativity:} if any three morphisms $f, g, h$ are composable, 
+    \begin{equation*}
+      \begin{tikzcd}
+        X \arrow[r, "f"] \arrow[rrr, "h\circ(g\circ f) = (h\circ g)\circ f"', bend right, dashed] & Y  \arrow[r, "g"] & Z \arrow[r, "h"] & W
+        \end{tikzcd}}
+  \end{equation*}
+  then they are associative such that $h\circ(g\circ f) = (h \circ g) \circ f$  \cite{lawvere2009conceptual,riehlCategoryTheoryContext,maclaneCategoriesWorkingMathematician2013,fongInvitationAppliedCategory2019}.
+  \end{enumerate}
+\end{definition}
+
+We construct a base space category $\mathcal{\dbasec}$ to encapsulate the topological structure of the data.
+The category $\mathcal{\dbase}$ has open set objects $\opensetc$ and inclusion morphisms $\opensetc_i \xrightarrow{\iota} \opensetc_j$ such that $\opensetc_i \subseteq \opensetc_j$.
+The composability property expresses that inclusion is transitive, while associativity expresses that the inclusion functions can be curried in various equivalent groupings.
+By formally specifying the properties of the topological structure datatypes as $\mathcal{\dbase}$, we can express that these are the properties that are required as part of the implementation of the data type objects.
+
+\subsubsection{Data Fields: Fiber Space \dfiber}
+\autoref{sec:atct:fb:fiber}
+As mentioned in \autoref{sec:related-work:equivariance}, equivariance is traditionally evaluated as the preservation of field structure, which is based on the field type.
+To have a rich language for expressing field type, we generalize Spivak's fiber bundle formulation of table \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES}.
+
+\begin{definition} A \textbf{type specification} is the bundle map $\pi: \mathscr{U} \rightarrow \textbf{DT}$.
+The base space $\textbf{DT}$ is the set of data types, and the total space $\mathscr{U}$ is the set of domains of those data types.
+Given an arbitrary data type $T \in \textbf{DT}$, the preimage $\pi^{-1} \subset \mathscr{U}$ is the domain of $T$ and $x \in \pi^{-1}(T)$ is an object of type $T$.
+\end{definition}
+For example, if $T=\texttt{int}$, then the image $\pi^{-1}(\texttt{int}) = \mathbb{Z} \subset \mathscr{U}$ is the set of all integers.
+The domain $\mathscr{U}_{{\sigma}}$ is also a topological space, meaning there is no restriction on the dimensionality of the field, thus allowing for the encoding of multi-dimensional fields in the same manner as single dimensional fields.
+Since many fields can have the same datatype, Spivak formally defines a mapping from field name to field data type, akin to a  data base schema \cite{ullmanFirstCourseDatabase2008}.
+\begin{definition} A \textbf{schema} consists of a pair $(\fnames, \sigma)$ where $\fnames$ is the set of field names and $\sigma: \fnames \rightarrow \textbf{DT}$ is a function from field name to field data type.
+\end{definition}
+The function $\sigma$ is composed with $\pi$ such that the bundle  $\pi_{\sigma}:\mathscr{U}_{\sigma} \rightarrow \fnames$ associates a field name $c \in C$ with the corresponding data type domain.
+Spivak then defines data records as sections on the bundle defined by $\pi_{\sigma}$.
+\begin{definition} A \textbf{record} is a function $\delement: \fnames \rightarrow \mathscr{U}_{\sigma}$ and the set of records on $\pi_{\sigma}$ is denoted $\Gamma^{\pi}(\sigma)$.
+\end{definition}
+Records must return an object of type $\sigma(\fname) \in \textbf{DT}$ for each field $c \in C$.
+Spivak then constructs tables as sections $\dsection: \dbase \rightarrow \Gamma^{\pi}(\sigma)$ from a key space $\dspace$ to the set of records $\Gamma^{\pi}(\sigma)$ on the schema bundle.
+Building on Spivak's construction, we define the fiber space in this work as the set of records
+\begin{equation*}
+  \dfiberc \coloneqq \{\delement: \fnames \rightarrow \mathscr{U}_{\sigma} \bigm{\vert} \pi_{\sigma}(\delement(\fname)) = \fname\;for\;all\; \fname \in \fnames \}
+\end{equation*}
+such that the preimage of a point is the corresponding domain $\pi^{-1}(\dbasepoint) = \dfiber_{k} = \mathscr{U}_{{\sigma}_{\dbasepoint}}$.
+This construction of the fiber as a schema bundle allows for the unique representation of data fields that have identical domains because they have the same underlying data type while still retaining the richness of a formal typing system.
+
+As with the base space category $\mathcal{\dbase}$, we propose a fiber category $\mathcal{\dfiberc}$ to encapsulate the field types of the data.
 The fiber category $\mathcal{\dfiberc}$ is a monoidal category, meaning the category has a single object $\dfiberc$ of an arbitrary type.
@@
 
-\subsection{Data: Section \dsection}
+\subsubsection{Data: Section \dsection} 
+\autoref{sec:atct:fb:sections}
 We encode data as a \textcolor{section}{section} \dsectionc\ of a bundle because this allows us to incorporate the topology and field types in the data definition.
-Sections  $\dsection: \dbasepoint \mapsto \delement$ map from a point in the basespace $\dbasepoint \in \dbase$ to a record in the fiber space $\delement \in \dfiber_{\dbasepoint}$.
-We can therefore model data as structures that map from an index like point $\dbasepoint$ to a data record $\delement$.
-Bundles can have multiple sections, as denoted by $\Gamma(\openset, \dtotal\restriction{\openset})$ which encapsulates the notion of multiple datasets with a shared fiber and base space.
-For example, all functions on the interval [0, 1] or a set of database tables with the same schema.
+We can define these section functions locally, meaning over a specific open subset \openset\ of \dbase\
+\begin{equation}
+  \label{eq:atct:fb:sections}
+  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}
+\end{equation}
+such that each section function $\dsection: \dbasepoint \mapsto \delement$ maps from a point $\dbasepoint \in \openset$ to a record in the fiber space $\delement \in \dfiber_{\dbasepoint}$ over that point.
+Bundles can have multiple sections, as denoted by $\Gamma(\openset, \dtotal\restriction{\openset})$.
+We can therefore model data as structures that map from an index like point $\dbasepoint$ to a data record $\delement$, and encapsulate multiple datasets with the same fiber and base space as different sections of the same bundle.
 
@@
 
-\subsection{intrepretation/example}
+\subsubsection{intrepretation/example}
 \begin{figure}[h!]
@@
 
-\subsection{Uniform Abstract Graphic Representation}
+\subsubsection{Uniform Abstract Graphic Representation}
 Fiber bundles are such a generalizable data abstraction that we can also use them to encode the output of a visual bundle; bundles allow for fully expressing the properties of the graphic that must be symmetric to the data in an equivariant (\autoref{sec:related-work:equivariance}) data to graphic transformation.
@@
 
+\note{add definition of functor + presheaf}
 Presheaves are functors, meaning they are functions from objects of one category to objects of another category\cite{WhatFunctorDefinitions} that preserve the morphisms in each category. 
@@
 We use a  presheaf to describe the expectation that any set of sections defined over a larger space $\dsection \in \Gamma(\opensetc_2, \dtotal\restriction_{\openset_{2}})$ is also defined over subspaces of that space $\dsection\restriction_{\opensetc_{1}}$.
-
+\note{make these formal definitions again}
 The locality and gluing axioms \cite{bakerMathsSheaf} of a sheaf describe the conditions for verifying that the sections over subspaces are part of a cohesive section over the whole space.
@@
 \label{sec:atct:sheaves:measurments}
+\note{move to Artist }
 Equivariance in a visual context was formally defined as the preservation of a binary operator from one domain to another by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}.
@@
 Given the map between spaces \vindex\, a property of sheaves is that there exist image functors\cite{ImageFunctorsSheaves2021} functors \vindexpull\ and \vindexpush that can transport sheaves over other base spaces.
+\note{insert formal def of pull back and pushforward functors}
 The \textcolor{functor}{pullback functor} $\vindexpullc$ sends sheaves over $\dbase$ to $\gbase$ and the \textcolor{functor}{pushforward functor} $\vindexpushc$  sends sheaves over $\gbase$ to $\dbase$

### 2022-07-04 paper/main.tex
sections + examples

@@
 \begin{definition} A \textbf{record} is a function $\delement: \fnames \rightarrow \mathscr{U}_{\sigma}$ and the set of records on $\pi_{\sigma}$ is denoted $\Gamma^{\pi}(\sigma)$.
+Records must return an object of type $\sigma(\fname) \in \textbf{DT}$ for each field $c \in C$.
 \end{definition}
-Records must return an object of type $\sigma(\fname) \in \textbf{DT}$ for each field $c \in C$.
 Spivak then constructs tables as sections $\dsection: \dbase \rightarrow \Gamma^{\pi}(\sigma)$ from a key space $\dspace$ to the set of records $\Gamma^{\pi}(\sigma)$ on the schema bundle.
-Building on Spivak's construction, we define the fiber space in this work as the set of records
-\begin{equation*}
+Generalizing Spivak's formulation of tables, we define the fiber space in this work as the set of record functions
+\begin{equation}
   \dfiberc \coloneqq \{\delement: \fnames \rightarrow \mathscr{U}_{\sigma} \bigm{\vert} \pi_{\sigma}(\delement(\fname)) = \fname\;for\;all\; \fname \in \fnames \}
-\end{equation*}
-such that the preimage of a point is the corresponding domain $\pi^{-1}(\dbasepoint) = \dfiber_{k} = \mathscr{U}_{{\sigma}_{\dbasepoint}}$.
-This construction of the fiber as a schema bundle allows for the unique representation of data fields that have identical domains because they have the same underlying data type while still retaining the richness of a formal typing system.
+\end{equation}
+such that the preimage of a point is the corresponding data type domain $\pi^{-1}(\dbasepoint) = \dfiber_{k} = \mathscr{U}_{{\sigma}_{\dbasepoint}}$.
+The construction of the fiber as the space of records provides a way to distinguish data fields with the same datatype while allowing that datatype to be defined as almost any arbitrary topological space.
 
@@
 \end{equation}
-where $\dsection=\texttt{dataset}$, $\dbase=\texttt{topology}$ and $\dfiber=\texttt{fields}$. This type signature encapsulates the topology and field types of the data without much loss of generality.
-
-Non-trivial bundles allow us to generalize even further.
-In a non-trivial (general) bundles, fibers are isomorphic $\dfiber \simeq \dfiber_{\dbasepoint}\;\forall \dbasepoint \in \dbase$, which means there is no single fiber type $\dfiber$.
+where $\dsection=\texttt{dataset}$, $\dbase=\texttt{topology}$ and $\dfiber=\texttt{fields}$.
+This type signature provides a method of explicitly stating the topology and field type of the data and generalizes to almost any topology and fiber type, provided that the total space is trivial.
+
+When the total space is non-trivial, we can use the fiber bundle property of local-triviality to generate type signatures.
+In a non-trivial bundle, fibers are isomorphic $\dfiber \simeq \dfiber_{\dbasepoint}\;\forall \dbasepoint \in \dbase$, which means there is no single fiber type $\dfiber$.
 Instead, we can define sections locally $\dsection\restriction_{\openset} \in \Gamma(\openset_{\dbasepoint}, \dtotal\restriction_{\openset_{\dbasepoint}})$, which means a section defined only on an open neighborhood $\dbasepoint \in \openset \in \dbase$.
@@
 \end{equation}
-that can be glued together on overlaps to generate the dataset. We discuss this gluing in \autoref{sec:atct:sheaves}.
-
-\subsubsection{intrepretation/example}
+that can be glued together on overlaps to generate the dataset. This gluing can be implemented as part of the data container, which we discuss in \autoref{sec:atct:sheaves}.
+
+\subsubsection{Examples}
 \begin{figure}[h!]
-  \includegraphics[width=\columnwidth]{fbnet.png}
-  \caption{\label{fig:atct:networkbundle}}
+  \includegraphics[width=\columnwidth]{tbundle.png}
+  \caption{The mobius band is a non-trivial total space because the twist means that the fibers are sometimes [-1, 1] and other times [1,-1].
+The mobius band can be represented as two trivial bundles, one with fiber [-1,1] and one with fiber [1,-1] that can be glued on the edges into the band.
+The gluing rules ensure that the arrows on the edges go in the same direction.\label{fig:atct:nontrivialbundle}}
 \end{figure}
-\note{figure out picture/example}
-Extending Spivak's fiber bundle description of a table, we define the values returned by a section as named and typed records.
-Since the fiber space is a topological space, it does not really have restrictions on the dimension of the types it can encode or the structure of those types.
-For example, wind has a speed component and a direction component and we can consider those separately $\dfiber_{speed} = \mathbb{R}, \dfiber_{direction} = \mathbb{R}$ or as one fiber$\dfiber_{wind} = \mathbb{R}^{2}$.
-The base space and the section evaluated on the base space form a sort of key-value pair, as described by Munzner \cite{munznerVisualizationAnalysisDesign2014}.
-She imbues semantic meaning to the topology, wherein the key is an independent attribute used as a look up key to retrieve the value in the dataset.
-She defines the value as a dependent attribute.
-In this work we use a more general notion of the topological space as an indexing space for the data records.
-For example, a data cube of temperature can be encoded such that every fiber is a set of 2D temperature surfaces $\dfiber_{tempsurf}$, the base space is a 1D continuous interval, and each section returns a surface $\dsection(\dbasepoint) = f(x,y)$ where x and y are already parameterized.
-Alternatively, the fiber can encode each cell of the data cube $\dfiber_{lat} \times \dfiber_{lon} \times \dfiber_{temp}$, the base space could be a 3D volume, and each section returns a $\dsection(\dbasepoint) = \{lat, lon, temp\}$ record.
+
+The mobius band is an example of the non-trivial bundle because the twist in the middle means that the fibers are not the same, as shown in \autoref{fig:atct:nontrivialbundle}.
+But, as with all bundles, the mobius band can be decomposed into trivial bundle components, here the two sides of the mobius band.
+Each side has fibers going in the same direction, with the mismatch only apparent on the edges.
+These trivial bundles can be glued into the mobius band so long as the fibers (arrows on the edges) go in the same direction when glued.
+The data is encoded as the section of the bundle, here the sin curve.
+
+\begin{figure}[h!]
+  \includegraphics[width=\columnwidth]{dbundle.png}
+  \caption{The table from \autoref{fig:related-work:continuity:ktypes} has a 3 dimensional fiber (name, temperature, precipitation), a 0D base space, and the row is the section.
+The time series has a 2D fiber (time, precipitation) and a 1D base space since the data is sampled along a continuous space (physical time).
+Each section returns a data point along the line.
+The map is sparsely sampled points along a 2D plane encoding spatial continuity, and the fiber is a 3D space encoding (latitude, longitude, precipitation)
+    \label{fig:atct:trivialbundle}}
+\end{figure}
+
+In \autoref{fig:atct:trivialbundle}, the base space $\dbase$ acts as an indexing space into the fiber space $\dfiber$.
+In the bundle encoding of the table, the indexing space is random keys; in the time series bundle, the base space is the interval [0,1]; and the map is sparse samples from a continuous space $[0,1]^{2}$ In contrast to a notion of a semantic binding between indexing space and field values, as proposed by Munzner\cite{munznerVisualizationAnalysisDesign2014}, the fields describing the continuity are part of the fiber.
+For example, the time is a fiber in the timeseries and the latitude and longitude are part of the map's fiber.
+This separation between connectivity and what it is called means the time or location can change units without a change to structure.
+It also provides a way to express data that may seem continuous but isn't, for example independent measurements over time.
+The data in \autoref{fig:atct:trivialbundle} comes from the same dataset; therefore we know that the bundles share connectivity and fiber space.
+It is expected that shared components be translated to visual elements in a consistent manner \cite{hullmanKeeping2018}, and in \autoref{sec:artist:operators} we introduce operators for expressing which components are shared and how to verify that they have been mapped into visual elements in a consistent manner.
 
 \subsubsection{Uniform Abstract Graphic Representation}
-Fiber bundles are such a generalizable data abstraction that we can also use them to encode the output of a visual bundle; bundles allow for fully expressing the properties of the graphic that must be symmetric to the data in an equivariant (\autoref{sec:related-work:equivariance}) data to graphic transformation.
+One of the advantages of fiber bundles is that they are general enough that we can also encode the output of a visual algorithm as a bundle.
+This allows us to use the same structure to express the properties of data and the graphic that must be symmetric to the data in an equivariant (\autoref{sec:related-work:equivariance}) transformation.
 We denote the output as a graphic, but the use of bundles allows us to generalize to output on any display space, such as a screen or 3D print.

### 2022-07-04 paper/main.tex
sheaf fun

@@
 The section evaluated on a single point $\gbasepoint$ returns a single graphic record, for example one pixel in an ideal resolution space.
-In our model, the unevaluated graphic section us passed to a renderer to generate graphics.
+In our model, the unevaluated graphic section is passed to a renderer to generate graphics.
 
@@
 These constraints generalize to data that is subsetted, distributed, streaming, and on-demand.
-
-\note{add definition of functor + presheaf}
-Presheaves are functors, meaning they are functions from objects of one category to objects of another category\cite{WhatFunctorDefinitions} that preserve the morphisms in each category. 
-A presheaf is a contravariant functor from an object in an arbitrary category to an object in the category \setb\cite{nlab:presheaf}. 
+Sheaves are built out of presheaves, which are functors, meaning they are functions from objects of one category to objects of another category\cite{WhatFunctorDefinitions} that preserve the morphisms in each category.
+\begin{definition} A \textbf{functor} is map $F: \mathcal{C} \rightarrow \mathcal{D}$ such that every object $C \in \mathcal{OB(C)}$ has a corresponding object $F(C) \in \mathcal{OB(D)}$ in category $\mathcal{D}$.
+The functor also acts on morphisms such that for every morphism $f: C_1 \rightarrow C_2$ in category $\mathcal{C}$, there is a compatible morphism $F(f): F(C_{1}) \rightarrow F(C_2)$.
+A functor must satisfy the properties 
+\begin{itemize}
+  \item \textit{identity}: $F(id_{C}(C)) = id_{D}(F(C))$
+  \item \textit{composition}:$F(g)\circ F(f) = F(g\circ f)$ for any composable morphisms $C_{1}\xrightarrow{f} C_2$, $C_2 \xrightarrow{g} C_3$
+\end{itemize}
+\end{definition}
+Functors preserve all the structure of the categories, namely the domains, codomains, composition, and identities\cite{riehlCategoryTheoryContext}$.
+Specifically, functors allow us to express which structure must be preserved when an object is transformed from one category to another.
+
+The presheaf encodes the expectation that continuity is preserved across sections.
+Specifically, a presheaf is a contravariant functor from an object in an arbitrary category to an object in the category \setb\cite{nlab:presheaf}.
 \begin{equation}
@@
 We use a presheaf to describe the expectation that any set of sections defined over a larger space $\dsection \in \Gamma(\opensetc_2, \dtotal\restriction_{\openset_{2}})$ is also defined over subspaces of that space $\dsection\restriction_{\opensetc_{1}}$.
-\note{make these formal definitions again}
-The locality and gluing axioms \cite{bakerMathsSheaf} of a sheaf describe the conditions for verifying that the sections over subspaces are part of a cohesive section over the whole space.
-The locality axiom asserts that two sections in a sheaf are equal $\dsection^{a} = \dsection^{b}$ when they evaluate to the same values over the same set of open sets $\dsection^{a}|{\openset} =  \dsection^{b}|_{\openset}$.
-The gluing axiom asserts that the union of sections defined on specific open sets is equivalent to one big section over the union of spaces $\dsection|_{\openset_{i} \cup \openset_{j}} = \dsection^{i}|_{\openset_i} \cup \dsection^{j}|_{\openset_j}$ if these sections agree on overlaps $\dsection^{i}|_{\openset_i\cap\openset_j} =  \dsection^{j}|_{\openset_i\cap\openset_j}$ The gluing axiom generalizes to expresses an equivalency between a concatenated representation of data and distributed representation of the same data.
-In \autoref{fig:atct:networkbundle} this axiom means that the sections over each node and the edge can be combined into one section over the piece of the network in the figure.
+This means any data container must implement subsetting in a manner that  makes sense for the continuity, for example an appropriate interpolation for continuous data.
+
+While the presheaf is a generalization of the concept of a global space of sections $\Gamma(\dbase, \dtotal)$ \cite{spanier1989algebraic}, the locality and gluing axioms \cite{bakerMathsSheaf} of a sheaf describe the conditions for verifying that the sections over subspaces are part of a cohesive section over the whole space.
+\begin{definition} A \textbf{sheaf} is a presheaf that satisfies the following two axioms
+\begin{itemize}
+  \item \textit{locality} two sections in a sheaf are equal $\dsection^{a} = \dsection^{b}$ when they evaluate to the same values over the same set of open sets $\dsection^{a}|{\openset} =  \dsection^{b}|_{\openset}$.
+  \item \textit{gluing} the union of sections defined on specific open sets is equivalent to one big section over the union of spaces $\dsection|_{\openset_{i} \cup \openset_{j}} = \dsection^{i}|_{\openset_i} \cup \dsection^{j}|_{\openset_j}$ if these sections agree on overlaps $\dsection^{i}|_{\openset_i\cap\openset_j} =  \dsection^{j}|_{\openset_i\cap\openset_j}$
+  \end{itemize}
+\end{definition}
+The gluing axiom says that a distributed representation of a dataset, which is a set of local sections, is equivalent to a section over the union of the opensets of the local sections.
+The locality axiom asserts that the glued section function is equivalent to a function over the union if they evaluate t the same values.
+The gluing axiom can also be used to generate the gluing rules used to construct non-trivial bundles from the set of trivial local section.
+Generally, the sheaf asserts the expectation that the data container is implemented such that the connectivity between the opensets (underlying data continuity) is preserved.
 

### 2022-07-04 paper/main.tex
pushpull

@@
 
-\subsection{Uniform Abstract Data Representation}
+\subsection{Abstract Data Representation}
 \label{sec:atct:fiber-bundles}
@@
 
-\subsection{Formal Properties of Data Containers}
+\subsection{Abstract Data Containers}
 \label{sec:atct:sheaves}
-Sheaves are a mathematical way to describe the expectation that data structures implement the bookkeeping that keeps track of the continuity of the data and graphic \cite{ghristElementaryAppliedTopology2014}.
+While bundles provide a way to describe the structure of the data, sheaves are a mathematical way of describing the data container.
+Sheaves are an algebraic data structure that provides a way of abstractly discussing the bookkeeping that data containers must implement to keep track of the continuity of the data \cite{ghristElementaryAppliedTopology2014}.
+This abstraction facilitates representational invariance, as introduced by Kindlemannn and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}, since the container level is uniformally specified as satisfying sheaf constraints.
 These constraints generalize to data that is subsetted, distributed, streaming, and on-demand.
+
 Sheaves are built out of presheaves, which are functors, meaning they are functions from objects of one category to objects of another category\cite{WhatFunctorDefinitions} that preserve the morphisms in each category.
-\begin{definition} A \textbf{functor} is map $F: \mathcal{C} \rightarrow \mathcal{D}$ such that every object $C \in \mathcal{OB(C)}$ has a corresponding object $F(C) \in \mathcal{OB(D)}$ in category $\mathcal{D}$.
+\begin{definition} A \textbf{functor} is a map $F: \mathcal{C} \rightarrow \mathcal{D}$ such that every object $C \in \mathcal{OB(C)}$ has a corresponding object $F(C) \in \mathcal{OB(D)}$ in category $\mathcal{D}$.
 The functor also acts on morphisms such that for every morphism $f: C_1 \rightarrow C_2$ in category $\mathcal{C}$, there is a compatible morphism $F(f): F(C_{1}) \rightarrow F(C_2)$.
@@
 \end{definition}
-Functors preserve all the structure of the categories, namely the domains, codomains, composition, and identities\cite{riehlCategoryTheoryContext}$.
+Functors preserve all the structure of the categories, namely the domains, codomains, composition, and identities\cite{riehlCategoryTheoryContext}.
 Specifically, functors allow us to express which structure must be preserved when an object is transformed from one category to another.
 
-The presheaf encodes the expectation that continuity is preserved across sections.
-Specifically, a presheaf is a contravariant functor from an object in an arbitrary category to an object in the category \setb\cite{nlab:presheaf}.
+We use a presheaf to encode the expectation that continuity is preserved across sections. 
+\begin{definition}
+  A \textbf{presheaf} $F:\mathcal{C}^{op} \rightarrow \setb$ is a contravariant functor from an object in an arbitrary category to an object in the category \setb\cite{nlab:presheaf}. 
+  \end{definition}
+Specifically, we use a presheaf to describe the expectation that any set of sections defined over a larger space $\dsection \in \Gamma(\opensetc_2, \dtotal\restriction_{\openset_{2}})$ is also defined over subspaces of that space $\dsection\restriction_{\opensetc_{1}}$.
+We do this by defining the presheaf as the association between an openset and its corresponding set of sections
 \begin{equation}
@@
 \end{equation}
-The presheaf expresses that every subset of the data $\openset$ has an associated set of data sections $\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$.
-The presheaf is a contravariant functor because the morphisms between input objects
+
+The presheaf preserves continuity by stating that given a morphism $\iota$ between two objects of $\mathcal{\dbase}$ 
 \begin{equation*}
@@
 \end{equation*}
-go in the opposite direction from the restriction morphisms $\iota^*$ between the sets
+there must be a corresponding restriction morphiism $\iota^*$ between the sets of sections 
 \begin{equation*}
@@
 which are the outputs of the sheaf $\sheaf(\openset) = \cgamma{\openset}{\dtotal\restriction_{\openset}}$.
-We use a presheaf to describe the expectation that any set of sections defined over a larger space $\dsection \in \Gamma(\opensetc_2, \dtotal\restriction_{\openset_{2}})$ is also defined over subspaces of that space $\dsection\restriction_{\opensetc_{1}}$.
-This means any data container must implement subsetting in a manner that  makes sense for the continuity, for example an appropriate interpolation for continuous data.
+The presheaf is contravariant because the morphisms between the input openset objects go in the opposite direction from the morphisms between the output set objects.
+The expectation that data containers implement inclusion and restriction morphisms means that, for example, if a data container implements subsetting (inclusion) of a continuous base space than it must ensure that the sections can be interpolated such that they can be defined over the subspace.
 
@@
 
-\subsection{Mathematical Structure of Data}
-\label{sec:atct:sheaves:measurments}
-\note{move to Artist }
+\subsection{Data and Graphic Correspondence}
+\label{sec:atct:xi}
+There is an expectation that for a visualization to be readable, the visual elements must correspond to distinct data elements\cite{ziemkiewiczEmbeddingInformationVisualization2009} and we can use the properties of sheaves to formally express this correspondence.
+First we establish that the graphic continuity corresponds to the data continuity by proposing that there exists a continuous \textcolor{functor}{map between topological spaces} \vindexc
+\begin{equation}
+  \label{eq:atct:xi}
+  \vindexc: \opensetgc \textcolor{functor}{\rightarrow} \opensetc 
+\end{equation}
+such that every $\opensetg \subseteq \gbase$ must map to a corresponding set $\openset \subset \dbase$.
+The functor $\vindex$ is surjective such that for every point $\dbasepoint \in \openset$ there is a corresponding set of points $\{\gbasepoint | \gbasepoint \in  \vindexpre(k)\}$ for all $\gbasepoint in \opensetg$.
+
+\subsubsection{Graphic Corresponding to Data}
+The pushforward (direct image) sheaf establishes which graphic generating function $\gsection$ corresponds to a point $\dbasepoint \in dbase$ in the data base space.
+
+\begin{definition} Given a sheaf $\sheaf_{\gbase, \gtotal}$ on $\gbase$, the \textbf{pushforward} sheaf  $\vindexpush\sheaf_{\gbase, \gtotal}$ on $\dbase$ is defined as 
+  \begin{equation*}
+    \vindexpushc(\sheafc_{\gbasec, \gtotalc})(\opensetc)  = \sheafc_{\gbasec, \gtotalc}(\vindexc^{-1}(\opensetc))
+  \end{equation*}
+for all opensets $\opensetc \subset \dbasec$\cite{harder2008lectures}.
+\end{definition}
+The pushforward sheaf returns the set of graphic sections over the data base space that corresponds to the graphic space $\vindex^{-1}(\openset) = \opensetgc$.
+The pushforward functor $\vindexpushc$ transports sheaves of sections on $\opensetgc$ over $\openset$
+ \begin{equation}  
+  \vindexpushc\gsectionc: \opensetc \rightarrow \vindexpushc \gtotalc\restriction_{\opensetc} \in \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} 
+\end{equation}
+such that it provides a way to look up which graphic  corresponds with a data index
+\begin{equation}
+  \vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexprec(\dbasepointc)}
+\end{equation}
+such that $\vindexpush\gsection(\dbasepoint))(\gbasepoint) = \gsection(\gbasepoint)$ for all $\gbasepoint \in \vindexpre(\dbasepoint)$.
+Therefore, the continuous map $\vindex$ and transport functors $\vindexpull, \vindexpush$ allow us to express the correspondence between graphic section and data section.
+
+\subsubsection{ Data Corresponding to Graphic}
+The pullback (inverse image) sheaf establishes which data record  returned by $\dsection$ corresponds to a point $\gbasepoint \in \gbase$ in the graphic base space. 
+\begin{definition} Given a sheaf $\sheaf_{\dbase, \dtotal}$ on $\dbase$, the \textbf{pullback} sheaf $\vindexpullc\sheafc_{\dbasec, \dtotalc}$ on $\gbasec$ is defined as the sheaf associated to the presheaf
+  \begin{equation*}
+    \vindexpullc(\sheafc_{\dbasec, \dtotalc})(\opensetgc) = \sheafc_{\dbasec, \dtotalc}(\vindexc(\opensetgc))   
+  \end{equation*}
+for $\vindex(\opensetgc) \in \dbase$ \cite{harder2008lectures}.
+\end{definition}
+The pullback sheaf returns the set of data sections over the graphic base space that corresponds to the graphic space $\vindex(\opensetg) = \openset$.
+The pullback $\vindexpullc$ transports sheaves of sections on $\openset \subseteq \dbase$ over $\opensetg \subseteq \gbase$
+\begin{equation}
+  \vindexpullc\dsectionc: \opensetgc \rightarrow \vindexpullc \dtotalc\restriction_{\opensetgc} \in \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} 
+\end{equation}
+such that there is a way to then look up what data values correspond with a graphic index
+\begin{equation}
+  \vindexpullc\dsectionc(\gbasepointc) = \dsectionc(\vindex(\gbasepointc)) = \dsectionc(\dbasepointc)
+\end{equation}
+As \vindex\ is surjective, there are many points $\gbasepoint \in \opensetg\subseteq\gbase$ in the graphic space that correspond to a single point $\vindex(\gbasepoint) = \dbasepoint$.
+
+\subsubsection{Example: Correspondence Between Graphic and Data}
+\begin{figure}[h!]
+  \includegraphics*[width=1\columnwidth]{xi_scatter.png}
+  \caption{The functors $\vindex, \vindexpull, \vindexpush$ establish the correspondence between a data section \dsection\ and a graphic generation function \gsection.
+For every data record $\dsection(\dbasepoint)$, there is a graphic generating function $\vindexpush\gsection = \gsection|_{\vindexpre\dbasepoint}$.
+This curried function is here represented as the visual parameters associated with that function to demonstrate that it generates the visual element for that data record.
+Every data section can also be copied over to the graphic space such that each piece of the (idealized in infinite resolution) graphic $\gsection(\dbasepoint)$ has a corresponding data value $\vindexpull\dsection(\dbasepoint)$.
+  \label{fig:atct:morphisms:sheaf}}
+\end{figure}
+
+Functors between sheaves are a way of expressing the bookkeeping involved in keeping track of which graphic section \gsection\ corresponds to which data section \dsection\.
+The $(\dbasepoint_i, \gbase_i)$ pairing expressed in \autoref{eq:atct:xi} establishes that there is a correspondence between sections evaluated over $\dbasepoint_i$ and $\gbase_i$.
+This allows us to construct graphic specifications for each data index $\vindexpush\gsection$ and retrieve the data $\vindexpull\dsection$ for any graphic section generating any piece of a graphic.
+In \autoref{fig:atct:morphisms:sheaf}, the sections return rows in the table from \autoref{fig:related-work:continuity:ktypes}.
+Each section $\dsection(\dbasepoint_{i})$ has a corresponding visual element produced by evaluating the graphic section on the graphic base space $\gsection|_{\gbase_{i}}$.
+In \autoref{fig:atct:morphisms:sheaf} the output of $\gsection$ as pixels in an idealized prerendered space is approximated as the circles.
+Each graphic section $\gsection(\gbasepoint), \gbasepoint \in \gbase_{i}$ that is a piece of the graphic has a corresponding pulled back data value $\vindexpull\dsectionc(\gbasepoint)$.
+The pullback copies the data points such that $\vindexpull\dsectionc|_{\gbase_i} = \dsection(\dbasepoint_i)$.
+Each visual element $\gsection|_{\gbase_i}$ is generated from a single record $\dsection(\dbasepoint_i)$; therefore the pushforward is the curried graphic section $\gsection|_{\vindexpre(\dbasepoint_i)}$.
+The graphic parameterized in \autoref{fig:atct:morphisms:sheaf} is intended as an approximation of $\vindexpush\gsection$.
+
+Visualization specification languages such as vega \cite{heerDeclarative2010} and svg \cite{quintScalable2003} assume that there is a graphic  specification for every point.
+Interactive tooltips assume that there is a map between the graphics rendered on screen and the data such that a glyph maps back to a record.
+The functors $(\vindex, \vindexpush, \vindexpull)$ codify the expectation of these mappings between data and graphic space.
+
+\section{Artist: Data to Graphic}
+\label{sec:artist}
+In this work we propose that visualization libraries are implementing transformations from data sheaf to graphic sheaf. We call these subset of functions the artist:
+\begin{align}
+  \label{eq:artist:hom_transport}
+  \vartistc:& \cgamma{\dbasec}{\dtotalc}\textcolor{artist}{\rightarrow} \cgamma{\gbasec}{\gtotalc}
+\end{align}
+ Because the artists can be constructed as morphisms of sheave over the same base spaces, as listed in \autoref{eq:atct:sheaves:homset:functors}, they are natural transformations.
+ \begin{definition}
+    \note{natural transform definition}
+ \end{definition}
+ which means that they are maps of functors that take the same input object and return objects in the same category\cite{milewskiCategoryTheoryProgrammers}.
+As illustrated in \autoref{eq:atct:sheaves:homset}, the sheaf functors
+\begin{equation}
+  \label{eq:artist:sheaf:base}
+    \begin{tikzcd}
+      \cgamma{\dbasec}{\dtotalc} &  & \dbasec \arrow[ll, "{\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf] \arrow[rr, "{\vindexpushc\sheafc_{\gbasec, \gtotalc}}", maps to, color=sheaf] &  & \cgamma{\dbase}{\vindexpushc\gtotalc}
+      \end{tikzcd}
+\end{equation}
+take as input the same object \opensetc\ and return sets of data and graphic sections that are objects in \setb.
+As a map between these sheaf functors, the artist has to preserve the $\iota, \iota^*$ morphisms of the presheaf functor, described in \autoref{eq:atct:presheaf}, such that the following diagram commutes
+
+\begin{equation}
+  \label{eq:artist:natural_transform:inclusions}
+  \begin{tikzcd}
+    \dbasec_1 & \cgamma{\dbasec_1}{\dtotalc} 
+    \arrow[dd, "\iota^*"', color=set] 
+    \arrow[rr, "\vartistc_{\dbase_1}", color=artist] &  & 
+    \cgamma{\dbasec_{1}}{\vindexpushc\gtotalc} 
+    \arrow[dd, "\iota^*", color=set] \\
+      &  &  &  \\
+    \dbasec_{2} \arrow[uu, "\iota", hook, color=base] & 
+    \cgamma{\dbasec_{2}}{\dtotalc} 
+    \arrow[rr, "\vartistc_{\dbasec_2}", color=artist] &  & 
+    \cgamma{\dbasec_{2}}{\vindexpushc\gtotalc}                      
+    \end{tikzcd}
+\end{equation}
+ The diagram in \autoref{eq:artist:natural_transform} shows that restricting a set of outputs of an artist to a set of graphic sections over a subspace is equivalent to restricting the inputs to data sections over the same subspace.
+By construction, the artist is expected to preserve the sheaf axioms of locality and gluing (\autoref{def:atct:sheaf:gluing}, \autoref{def:atct:sheaf:locality}).
+This bookkeeping is necessary for any visualization technique that selectively acts on different pieces of a data set; for example streaming visualizations \cite{krstajicVisualizationStreamingData2013} and panning and zooming \cite{NekrasovskiEvaluationPanZoom2006}
+
+The output of an artist \vartist\ is a restricted subset of graphic sections
+\begin{equation}
+  \label{eq:artist:output}
+  \imartist{\gbasec}{\gtotalc} \coloneqq\\ 
+  \{\gsectionc \mid\;\exists\;\dsectionc \in \cgamma{\dbasec}{\dtotalc}\;s.t.\; 
+  \vartistc(\dsectionc) = \gsectionc,\; \vindexc(\gbasec) = \dbasec \}
+\end{equation} 
+that are, by definition, only reachable through a structure preserving artist, which we describe in \autoref{sec:artist:equivariance}.
+We define this subset because the space of all sections $\cgamma{\opensetg}{\gtotal\restriction_{\openset}}$ includes sections that may not be structure preserving.
+For example, a section may go from every point in the graphic space to the same single point in the graphic fiber $\gsection(\gbasepoint_i) = d\; \forall \gbasepoint \in \gbase$ such that the visual output is a single inked pixel on a screen.
+
+\subsection{Equivariance}
+\label{sec:artist:equiv}
 Equivariance in a visual context was formally defined as the preservation of a binary operator from one domain to another by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}.
@@
 We generalize to a family of actions because that allows for expanding the set of allowable transformations on the data beyond a single operator.
-
+We describe the changes on the graphic side as changes in measurements \measure\, which are scaler or vector components of the rendered graphic that can be quantified, such as the color, position, shape, texture, or rotation angle of the graphic.
+The visual variables \cite{bertinIIPropertiesGraphic2011} are a subset of measurable components.
+For example, a measurement of a scatter marker could be its color (e.g.
+red) or its x position (e.g.
+5).
+
+\subsubsection{Mathematical Structure of Data}
+\label{sec:artist:equivariant:data}
 We separate data transformations into two components, transformations on the base space $(\dfunchc, \dfuncpullc)$ and transformations on the fiber space $\dfunctc$.
@@
 \end{equation}
-As shown in \autoref{eq:atct:sheaves:monoid_morphism}, the base space transform 
+
+The base space transformation is from one open set to another open set in the same base space
 \begin{equation}
@@
  \end{equation}
-The base space transformation is from one open set to another open set in the same base space $\openset, \openset^{\prime} \subseteq \dbase$.
+such that $\openset, \openset^{\prime} \subseteq \dbase$.
+This means $\openset$ and $\openset^{\prime}$ are of the same topology type.
 To correctly align the sections with the remapped base space, there is a a corresponding section pullback function
@@
 
-\subsection{Data and Graphic Correspondence}
-\label{sec:atct:xi}
-There is an expectation that for a visualization to readable, the visual elements must correspond to distinct data elements\cite{ziemkiewiczEmbeddingInformationVisualization2009} and we can use the properties of sheaves to formally express this correspondence.
-First we establish that the graphic continuity corresponds to the data continuity by proposing that there exists a continuous \textcolor{functor}{map between topological spaces} \vindexc
-\begin{equation}
-  \label{eq:atct:xi}
-  \vindexc: \opensetgc \rightarrow \opensetc 
-\end{equation}
-such that every $\opensetg \subseteq \gbase$ must map to a corresponding set $\openset \subset \dbase$.
-The functor $\vindex$ is surjective such that for every point $\dbasepoint \in \openset$ there is a corresponding set of points $\{\gbasepoint | \gbasepoint \in  \vindexpre(k)\}$ for all $\gbasepoint in \opensetg$.
-
-Given the map between spaces \vindex\, a property of sheaves is that there exist image functors\cite{ImageFunctorsSheaves2021} functors \vindexpull\ and \vindexpush that can transport sheaves over other base spaces.
-\note{insert formal def of pull back and pushforward functors}
-The \textcolor{functor}{pullback functor} $\vindexpullc$ sends sheaves over $\dbase$ to $\gbase$ and the \textcolor{functor}{pushforward functor} $\vindexpushc$  sends sheaves over $\gbase$ to $\dbase$
-\begin{equation}
-  \label{eq:atct:morphisms:xi}
-\begin{tikzcd}[row sep=huge]
-  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
-  \arrow[rr, "\vindexpullc", maps to, color=functor] &  & 
-  \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}  \\
-  \opensetc 
-  \arrow[d, "{\vindexpushc\sheafc_{\gbasec,  \gtotalc}}", dashed, maps to, color=sheaf] 
-  \arrow[u, "{\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf] &  & 
-  \opensetgc 
-  \arrow[d, "{\sheafc_{\gbasec, \gtotalc}}", maps to, color=sheaf] 
-  \arrow[u, "{\vindexpullc\sheafc_{\dbasec, \dtotalc}}"', dashed, maps to, color=sheaf] 
-  \arrow[ll, "\vindexc"', maps to, color=functor] \\
-  \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} &  & 
-  \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
-  \arrow[ll, "\vindexpushc"', maps to, color=functor]                                                         
-  \end{tikzcd}
-\end{equation}
-such that there is an association between graphic sections \gsection\ that take as input \gbasepoint\ and data sections that take input \dbasepointc\ when $\vindex(\gbasepoint) = \dbasepoint$.
-
-The pull back $\vindexpullc$ transports sheaves of sections on $\openset \subseteq \dbase$ over $\opensetg \subseteq \gbase$
-\begin{equation}
-  \vindexpullc\dsectionc: \opensetgc \rightarrow \vindexpullc \dtotalc\restriction_{\opensetgc} \in \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} 
-\end{equation}
-such that there is a way to then look up what data values correspond with a graphic index
-\begin{equation}
-  \vindexpullc\dsectionc(\gbasepointc) = \dsectionc(\vindex(\gbasepointc)) = \dsectionc(\dbasepointc)
-\end{equation}
-
- The pushforward $\vindexpushc$ transports sheaves of sections on $\opensetgc$ over $\openset$
- \begin{equation}  
-  \vindexpushc\gsectionc: \opensetc \rightarrow \vindexpushc \gtotalc\restriction_{\opensetc} \in \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} 
-\end{equation}
-such that it provides a way to look up which graphic  corresponds with a data index
-\begin{equation}
-  \vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexprec(\dbasepointc)} = \{\gsectionc(\gbasepointc)\;for\;all\; \gbasepointc\in \vindexprec(\dbasepointc)\}
-\end{equation}
-\note{$\xi_*rho(k)(s)=rho(s) for s in \xi^{-1}(k)
-$} and move {} into narrative
-Therefore, the continuous map $\vindex$ and transport functors $\vindexpull, \vindexpush$ allow us to express the correspondence between graphic section and data section.
-
-\begin{figure}[h!]
-  \includegraphics*[width=1\columnwidth]{xi_scatter.png}
-  \caption{The functors $\vindex, \vindexpull, \vindexpush$ establish the correspondence between a data section \dsection\ and a graphic generation function \gsection.
-For every data record $\dsection(\dbasepoint)$, there is a graphic generating function $\vindexpush\gsection = \gsection|_{\vindexpre\dbasepoint}$.
-This curried function is here represented as the visual parameters associated with that function to demonstrate that it generates the visual element for that data record.
-Every data section can also be copied over to the graphic space such that each piece of the (idealized in infinite resolution) graphic $\gsection(\dbasepoint)$ has a corresponding data value $\vindexpull\dsection(\dbasepoint)$.
-  \label{fig:atct:morphisms:sheaf}}
-\end{figure}
-
-Functors between sheaves are a way of expressing the bookkeeping involved in keeping track of which graphic section \gsection\ corresponds to which data section \dsection\.
-The $(\dbasepoint_i, \gbase_i)$ pairing expressed in \autoref{eq:atct:xi} establishes that there is a correspondence between sections evaluated over $\dbasepoint_i$ and $\gbase_i$.
-This allows us to construct graphic specifications for each data index $\vindexpush\gsection$ and retrieve the data $\vindexpull\dsection$ for any graphic section generating any piece of a graphic.
-In \autoref{fig:atct:morphisms:sheaf}, the sections return rows in the table from \autoref{fig:related-work:continuity:ktypes}.
-Each section $\dsection(\dbasepoint_{i})$ has a corresponding visual element produced by evaluating the graphic section on the graphic base space $\gsection|_{\gbase_{i}}$.
-In \autoref{fig:atct:morphisms:sheaf} the output of $\gsection$ as pixels in an idealized prerendered space is approximated as the circles.
-Each graphic section $\gsection(\gbasepoint), \gbasepoint \in \gbase_{i}$ that is a piece of the graphic has a corresponding pulled back data value $\vindexpull\dsectionc(\gbasepoint)$.
-The pullback copies the data points such that $\vindexpull\dsectionc|_{\gbase_i} = \dsection(\dbasepoint_i)$.
-Each visual element $\gsection|_{\gbase_i}$ is generated from a single record $\dsection(\dbasepoint_i)$; therefore the pushforward is the curried graphic section $\gsection|_{\vindexpre(\dbasepoint_i)}$.
-The graphic parameterized in \autoref{fig:atct:morphisms:sheaf} is intended as an approximation of $\vindexpush\gsection$.
-
-Visualization specification languages such as vega \cite{heerDeclarative2010} and svg \cite{quintScalable2003} assume that there is a graphic  specification for every point.
-Interactive tooltips assume that there is a map between the graphics rendered on screen and the data such that a glyph maps back to a record.
-The functors $(\vindex, \vindexpush, \vindexpull)$ codify the expectation of these mappings between data and graphic space.
-
-We propose that visualization libraries implement morphisms from a data sheaf $\sheafc_{\dtotal, \dbase}$ to a graphic sheaf $\sheafc_{\gtotal, \gbase}$ 
-\begin{equation}
-  \label{eq:atct:sheaves:homset}
-  \begin{tikzcd}
-    \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
-    \arrow[dd, "\textcolor{set}{Hom}_{\sheafc_{\dbasec}}"', color=homset] 
-    \arrow[rrdd, "\textcolor{set}{Hom}_{\sheafc_{\dbasec},\sheafc_{\gbasec}}", color=homset] 
-    \arrow[rr, "\vindexpullc", color=functor] &  &
-    \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} 
-    \arrow[dd, "\textcolor{set}{Hom}_{\sheafc_{\gbasec}}", color=homset] \\
-     & & \\
-    \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} &  & 
-    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
-    \arrow[ll, "\vindexpushc"', color=functor]                  
-    \end{tikzcd}
-\end{equation}
-and the hom set is the set of all morphisms from one sheaf to the other.
-The functors $\vindexpullc, \vindexpushc$ are adjoint, meaning that there is a functorial isomorphism \cite{harder2008lectures} such that
-\begin{equation}
-  \label{eq:atct:sheaves:homset:functors}
-   \textcolor{set}{Hom}_{\sheafc_{\gbasec}}(\vindexpullc\sheafc_{\dtotalc, \dbasec},\sheafc_{\gtotalc,\gbasec})
-  \simeq  \textcolor{set}{Hom}_{\sheafc_{\dbasec}}(\sheafc_{\dtotalc, \dbasec},\vindexpushc\sheafc_{\gtotalc,\gbasec}) 
-\end{equation}
-such that both are isomorphic to  $\simeq  \textcolor{set}{Hom}_{\sheafc_{\dbasec},\sheafc_{\gbasec}}(\sheafc_{\dtotalc, \dbasec},\sheafc_{\gtotalc,\gbasec})$.
-This set of ismorphic homsets means that the diagram of sheaf morphisms \autoref{eq:atct:sheaves:homset} is commutative; therefore the functors $\vindexpullc$ and $\vindexpushc$ can be used to adapt morphisms written over one space to another space.
-\note{maybe move to construction}
-This provides us flexibility over which space to construct data to graphic transforms.
-For example, specifications such as svg\cite{quintScalable2003} and Vega\cite{satyanarayanDeclarativeInteractionDesign2014}, typically define their transformations in data space and return specifications for the render to evaluate.
-An artist may also want to specify the transformations in a proxy of graphic space, for example to generate on-demand dynamically resampled continuous lines.
-
-\section{Artist: Data to Graphic}
-\label{sec:artist}
-In this work we propose that visualization libraries are implementing a subset of the functions in the hom sets in \autoref{eq:atct:sheaves:homset}. We call these subset of functions the artist:
-\begin{align}
-  \label{eq:artist:hom_transport}
-  \vartistc:& \cgamma{\dbasec}{\dtotalc}\textcolor{artist}{\rightarrow} \cgamma{\gbasec}{\gtotalc} \in \textcolor{set}{Hom}_{\sheafc_{\dbase}, \sheafc_{\gbase}}
-\end{align}
- Because the artists can be constructed as morphisms of sheave over the same base spaces, as listed in \autoref{eq:atct:sheaves:homset:functors}, they are natural transformations, which means that they are maps of functors that take the same input object and return objects in the same category\cite{milewskiCategoryTheoryProgrammers}.
-As illustrated in \autoref{eq:atct:sheaves:homset}, the sheaf functors
-\begin{equation}
-  \label{eq:artist:sheaf:base}
-    \begin{tikzcd}
-      \cgamma{\dbasec}{\dtotalc} &  & \dbasec \arrow[ll, "{\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf] \arrow[rr, "{\vindexpushc\sheafc_{\gbasec, \gtotalc}}", maps to, color=sheaf] &  & \cgamma{\dbase}{\vindexpushc\gtotalc}
-      \end{tikzcd}
-\end{equation}
-take as input the same object \opensetc\ and return sets of data and graphic sections that are objects in \setb.
-As a map between these sheaf functors, the artist has to preserve the $\iota, \iota^*$ morphisms of the presheaf functor, described in \autoref{eq:atct:presheaf}, such that the following diagram commutes
-
-\begin{equation}
-  \label{eq:artist:natural_transform:inclusions}
-  \begin{tikzcd}
-    \dbasec_1 & \cgamma{\dbasec_1}{\dtotalc} 
-    \arrow[dd, "\iota^*"', color=set] 
-    \arrow[rr, "\vartistc_{\dbase_1}", color=artist] &  & 
-    \cgamma{\dbasec_{1}}{\vindexpushc\gtotalc} 
-    \arrow[dd, "\iota^*", color=set] \\
-      &  &  &  \\
-    \dbasec_{2} \arrow[uu, "\iota", hook, color=base] & 
-    \cgamma{\dbasec_{2}}{\dtotalc} 
-    \arrow[rr, "\vartistc_{\dbasec_2}", color=artist] &  & 
-    \cgamma{\dbasec_{2}}{\vindexpushc\gtotalc}                      
-    \end{tikzcd}
-\end{equation}
- The diagram in \autoref{eq:artist:natural_transform} shows that restricting a set of outputs of an artist to a set of graphic sections over a subspace is equivalent to restricting the inputs to data sections over the same subspace.
-By construction, the artist is expected to preserve the sheaf axioms of locality and gluing (\autoref{def:atct:sheaf:gluing}, \autoref{def:atct:sheaf:locality}).
-This bookkeeping is necessary for any visualization technique that selectively acts on different pieces of a data set; for example streaming visualizations \cite{krstajicVisualizationStreamingData2013} and panning and zooming \cite{NekrasovskiEvaluationPanZoom2006}
-
-The output of an artist \vartist\ is a restricted subset of graphic sections
-\begin{equation}
-  \label{eq:artist:output}
-  \imartist{\gbasec}{\gtotalc} \coloneqq\\ 
-  \{\gsectionc \mid\;\exists\;\dsectionc \in \cgamma{\dbasec}{\dtotalc}\;s.t.\; 
-  \vartistc(\dsectionc) = \gsectionc,\; \vindexc(\gbasec) = \dbasec \}
-\end{equation} 
-that are, by definition, only reachable through a structure preserving artist, which we describe in \autoref{sec:artist:equivariance}.
-We define this subset because the space of all sections $\cgamma{\opensetg}{\gtotal\restriction_{\openset}}$ includes sections that may not be structure preserving.
-For example, a section may go from every point in the graphic space to the same single point in the graphic fiber $\gsection(\gbasepoint_i) = d\; \forall \gbasepoint \in \gbase$ such that the visual output is a single inked pixel on a screen.
-
-\subsection{Equivariance}
-The notion of structure preservation discussed in \autoref{sec:related-work:continuity} is that data and components of the graphic vary in equivalent ways.
-We describe the changes on the graphic side as changes in measurments \measure\, which are scaler or vector components of the rendered graphic that can be quantified, such as the color, position, shape, texture, or rotation angle of the graphic.
-The visual variables \cite{bertinIIPropertiesGraphic2011} are a subset of measurable components.
-For example, a measurement of a scatter marker could be its color (e.g.
-red) or its x position (e.g.
-5).
-
+\subsubsection{Equivariant Artist}
+\label{sec:artist:equivariant:artist}
 We formalize this structure preservation as equivariance, which is that for every morphism on the data $(\dfunch_{\dtotal}, \dfunct_{\dtotal})$ there is an equivalent morphism on the graphic  $(\dfunch_{\gtotal}, \dfunct_{\gtotal})$ The artist is an equivariant map if the diagram commutes for all points $\gbasepointc^{\prime}\in \gbasec^{\prime}$
@@
   \arrow[dd, "\dfunctc_{\dtotalc}"', color=action] \& 
-  \opensetc 
+  \dbasec 
    \& 
-  \opensetgc 
+  \gbasec 
   \arrow[l, "\vindexc"', color=functor] 
@@
   \& 
-  \opensetc^{\prime} 
+  \dbasec^{\prime} 
   \arrow[u, "\dfunchc_{\dtotalc}", color=action] 
   \& 
-  \opensetgc^{\prime} 
+  \gbasec^{\prime} 
   \arrow[l, "\vindexc"', color=functor] 

### 2022-07-04 paper/main.tex
nt

@@
 
-\subsubsection{Example: Correspondence Between Graphic and Data}
+\subsubsection{Example: Graphic and Data}
 \begin{figure}[h!]
@@
 \end{align}
- Because the artists can be constructed as morphisms of sheave over the same base spaces, as listed in \autoref{eq:atct:sheaves:homset:functors}, they are natural transformations.
- \begin{definition}
-    \note{natural transform definition}
+ Because the artists can be constructed as morphisms of sheaves over the same base spaces, they are natural transformations.
+ \begin{definition} Given two functors $F,G$ with the same domain $\mathcal{C}$ and codomain $\mathcal{D}$, a \textbf{natural transformation} $\alpha: F \Rightarrow G$ is a map
+\begin{itemize}
+  \item \textit{data}: morphism $F(c) \xrightarrow{\alpha_{c}} G(c)$ for each object $c \in \mathcal{C}$
+  \item \textit{property} when $f: c_1 \rightarrow c_2$ is a morphism in $\mathcal{C}$, the components of the natural transform commute $G(f) \circ \alpha_{c_1} = \alpha_{c_2} \circ F(f)$  
+\end{itemize}
+such that $\alpha = (\alpha_{c})_{c\in\mathcal{C}}$ is the set of all natural transformation components $\alpha_{c}$.\cite{bradleyWhatNaturalTransformation}
  \end{definition}
- which means that they are maps of functors that take the same input object and return objects in the same category\cite{milewskiCategoryTheoryProgrammers}.
+ which means that natural transforms are maps of functors that take the same input object and return objects in the same category\cite{milewskiCategoryTheoryProgrammers}.
 As illustrated in \autoref{eq:atct:sheaves:homset}, the sheaf functors
@@
 
-As shown in \autoref{eq:atct:sheaves:homset}, if a sheaf is equipped with transport functors, then the functions between sheaves over one space are isomorphic to functions between sheave over the other space.
+ If a sheaf is equipped with transport functors, then the functions between sheaves over one space are isomorphic to functions between sheaves over the other space\cite{harder2008lectures}.
 This property, allows us to specify the transformation from data to graphic space in data or graphic space or a combination, as listed in \autoref{eq:artist:hom_transport}, such that the following diagram commutes:
@@
 \end{equation}
-The diagram in \autoref{eq:construction:artist:path} illustrates how a developer can connect transformations over data space, denoted with a subset $\dbase$, with transformations over graphic space $\gbase$, using $\vindexpushc$ and $\vindexpullc$ adaptors.
-This allows developers to for example connect transformers that transform data on a line to a color in dataspace, but build a line compositing function that dynamically resamples what is on screen in graphic space.
+The diagram in \autoref{eq:construction:artist:path} illustrates how a developer can connect transformations over data space, denoted with a subset $\dbase$, with transformations over graphic space $\gbase$, using $\vindexpushc$ and $\vindexpullc$ adaptors.This allows developers to for example connect transformers that transform data on a line to a color in dataspace, but build a line compositing function that dynamically resamples what is on screen in graphic space.
+
+\section{Discussion: Feasibility of Math as Design Spec}
+\label{sec:discussion}
 \begin{figure}
@@
 \end{figure}
-
-\section{Discussion: Feasibility of Math as Design Spec}
-\label{sec:discussion}
 The framework specified in \autoref{sec:artist} and \autoref{sec:construction} describes how to build structure preserving visualization components, but it is left to the library developer to follow these guidelines when building and reusing components.

### 2022-07-06 paper/main.tex
reworked action def + brought homset stuff into implementation + more minor moving around

@@
 \end{align}
- Because the artists can be constructed as morphisms of sheaves over the same base spaces, they are natural transformations.
+ The artists can be constructed as morphisms of sheaves over the same base spaces through the application of pushforward and pullback functors; therefore they are natural transformations.
  \begin{definition} Given two functors $F,G$ with the same domain $\mathcal{C}$ and codomain $\mathcal{D}$, a \textbf{natural transformation} $\alpha: F \Rightarrow G$ is a map
@@
  \end{definition}
- which means that natural transforms are maps of functors that take the same input object and return objects in the same category\cite{milewskiCategoryTheoryProgrammers}.
+ This means that natural transforms are maps of functors that take the same input object and return objects in the same category\cite{milewskiCategoryTheoryProgrammers}.
 As illustrated in \autoref{eq:atct:sheaves:homset}, the sheaf functors
@@
 \end{equation}
-take as input the same object \opensetc\ and return sets of data and graphic sections that are objects in \setb.
+take as input an openset object \opensetc\ or \opensetgc\ and return sets of data and graphic sections that are objects in \setb.
 As a map between these sheaf functors, the artist has to preserve the $\iota, \iota^*$ morphisms of the presheaf functor, described in \autoref{eq:atct:presheaf}, such that the following diagram commutes
@@
  The diagram in \autoref{eq:artist:natural_transform} shows that restricting a set of outputs of an artist to a set of graphic sections over a subspace is equivalent to restricting the inputs to data sections over the same subspace.
-By construction, the artist is expected to preserve the sheaf axioms of locality and gluing (\autoref{def:atct:sheaf:gluing}, \autoref{def:atct:sheaf:locality}).
+Because the artist is a functor of sheaves, the artist is expected to tra nslate the data continuity to graphic continuity such that the connectivity of subsets is preserved.
 This bookkeeping is necessary for any visualization technique that selectively acts on different pieces of a data set; for example streaming visualizations \cite{krstajicVisualizationStreamingData2013} and panning and zooming \cite{NekrasovskiEvaluationPanZoom2006}
@@
 Equivariance in a visual context was formally defined as the preservation of a binary operator from one domain to another by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}.
-We generalize binary operations to actions $\dfunc \in \dfuncset$ on elements of the bundle $\delement \in \dtotal$.
-This is expressed with a composition function  $act: \dfuncset \times E \rightarrow \dtotal$ that applies the action to an element of \dtotal\ such that $\dfunc(\delement_0) = \delement_1, \delement_0, \delement_1 \in \dfunc$.
-We generalize to a family of actions because that allows for expanding the set of allowable transformations on the data beyond a single operator.
+We generalize binary operations to actions $\dfunc \in \dfuncset$ on the sheaf $\sheaf_{\dbase, \dtotal}$.
+\begin{definition} An \textbf{action}  $\act: \dfuncsetc \times \dtotalc \rightarrow \sheaf_{\dbase, \dtotal}$ such that 
+  \begin{equation}
+    act(\dfuncc_{j}, act(\dfuncc_{i}, \dsectionc)) = act(\dfuncc_{j} \circ \dfuncc_{i}, \dsectionc)
+  \end{equation}
+where $\dfunc_{i}, \dfunc_{j}$ are elements of the set $\dfuncset$ equipped with operator $\circ$ and \dsection\ is an element of the sheaf  $\sheaf_{\dbase, \dtotal}$.\cite{nlab:action}
+\end{definition}.
+
+We generalize from binary operations to a family of actions because that allows for expanding the set of allowable transformations on the data beyond a single operator.
 We describe the changes on the graphic side as changes in measurements \measure\, which are scaler or vector components of the rendered graphic that can be quantified, such as the color, position, shape, texture, or rotation angle of the graphic.
@@
 We separate data transformations into two components, transformations on the base space $(\dfunchc, \dfuncpullc)$ and transformations on the fiber space $\dfunctc$. 
-These diagrams are separated because the base space transformation and the pullback are functors on the openset object and associated set of sections, while the fiber transformation takes a single fiber to a different fiber in the set of sections.
-
 \begin{equation}
@@
 \end{equation}
+The base space transformation transforms one openset object $\openset^{\prime}$ to another object $\openset$, and the pullback functor transports the entire set of sections $\Gamma(\openset, \dtotal\restriction_{\openset})$ over the new base space $\Gamma(\openset^{\prime}, \dfuncpull\dtotal\restriction_{\openset^{\prime}})$.
+The fiber transformation transforms a single section $\dfuncpull\dsection$ to a different section $\dfuncpull\dsection$.
+
+\paragraph{Transform Continuity}
 
@@
   \label{eq:atct:morphism:verify_base}
-  \dsection(\dbase) = \dfuncpull\dsectionc(\dbasepoint^{\prime}) = \dsection(\dfunch(\dbase^{prime}))
+  \dsection(\dbase) = \dfuncpull\dsection(\dbasepoint^{\prime}) = \dsection(\dfunch(\dbase^{prime}))
 \end{equation} 
 where $\dfunch(\dbasepoint^{\prime}) = \dfunch(\dbasepoint)$.
+
+\paragraph{Transform Records}
+
 As introduced in \autoref{def:atct:io:fiber}, the fiber transformation $\dfunct: \dfuncpull \dtotal_{\dbasepoint^{\prime}} \rightarrow \dfuncpull \dtotal_{\dbasepoint^{\prime}}$ is a morphism on the fiber $\dfunct \in Hom(\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}},\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}})$ restricted to a point $\dbasepoint^{\prime} \in \openset^{\prime}$.
@@
 Examples of a fiber transformation are actions such as partial ordering for ranking purpose \cite{bruggemannRankingPrioritizationMultiindicator2011} and the permutation, ordering, translation, and scaling codified as Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946m, leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}.
+
+\paragraph{Transform Continuity and Records}
 
@@
 
- If a sheaf is equipped with transport functors, then the functions between sheaves over one space are isomorphic to functions between sheaves over the other space\cite{harder2008lectures}.
-This property, allows us to specify the transformation from data to graphic space in data or graphic space or a combination, as listed in \autoref{eq:artist:hom_transport}, such that the following diagram commutes:
+ When a sheaf is equipped with transport functors, then the functions between sheaves over one space are isomorphic to functions between sheaves over the other space\cite{harder2008lectures} such that the following diagram commutes
+
+ \begin{equation}
+  \label{eq:atct:sheaves:homset}
+  \begin{tikzcd}
+    \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
+    \arrow[dd, "\textcolor{set}{Hom}_{\sheafc_{\dbasec}}"', color=homset] 
+    \arrow[rrdd, "\textcolor{set}{Hom}_{\sheafc_{\dbasec},\sheafc_{\gbasec}}", color=homset] 
+    \arrow[rr, "\vindexpullc", color=functor] &  &
+    \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} 
+    \arrow[dd, "\textcolor{set}{Hom}_{\sheafc_{\gbasec}}", color=homset] \\
+     & & \\
+    \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} &  & 
+    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
+    \arrow[ll, "\vindexpushc"', color=functor]                  
+    \end{tikzcd}
+\end{equation}
+Since the artist is a family of functions in the homeset between sheaves, the isomorphism allows for the specification of the transformation from data as combination of functions over different spaces such that the following diagram commutes:
 \begin{equation}
@@
 \end{equation}
-The diagram in \autoref{eq:construction:artist:path} illustrates how a developer can connect transformations over data space, denoted with a subset $\dbase$, with transformations over graphic space $\gbase$, using $\vindexpushc$ and $\vindexpullc$ adaptors.This allows developers to for example connect transformers that transform data on a line to a color in dataspace, but build a line compositing function that dynamically resamples what is on screen in graphic space.
+This means that an artist over data space $\vartist_{\dbase}: \dsection \mapsto \vindexpush \gsection$, an artist over graphic space $vartist_{\gbase}: \vindexpull \dsection \mapsto \gsection$, and an artist $\vartist: \dsection \mapsto \gsection$ are equivalent such that:
+\begin{equation*}
+  \begin{split}
+  & \dsection(\dbasepoint) = \vindexpull\dsection(\gbasepoint)  \\
+   & \implies
+  \vartist_{\dbase}(\dsection(\dbasepoint)) = \vartist_{\gbase}(\vindexpull\dsection(\gbasepoint)) = \vartist(\dsection(\dbasepoint))\\
+  & \implies \vindexpush\gsection(\gbasepoint) = \gsection(\gbasepoint) 
+  \end{split}
+\end{equation*}
+when $\vindex(\gbasepoint) = \dbasepoint$.
+This equivalence allows a  developer to connect transformations over data space, denoted with a subset $\dbase$, with transformations over graphic space $\gbase$, using $\vindexpushc$ and $\vindexpullc$ adaptors.This allows developers to for example connect transformers that transform data on a line to a color in dataspace, but build a line compositing function that dynamically resamples what is on screen in graphic space.
 

### 2022-07-12 paper/main.tex
working references

@@
 Motivated by a need to develop new library components that adapt to modern data needs, especially high dimensional multivariate and distributed data, we propose a framework that topological and field structure of the input data that must be mapped to the visualization and which changes to data must be expressed in visualization, and a a functional design framework that describes composability in terms of functional algebraic operations.
-\note{this may be a good place to incorporate types} This allows complex visualizations to be built from simple verifiable parts\cite{huHowFunctionalProgramming2015,hughesWhyFunctionalProgramming1989}, which makes these components easier to drop into the existing library and be used by domain specific library developers to build new libraries that work with existing ones.
+\note{this may be a good place to incorporate types} This allows complex visualizations to be built from simple verifiable parts\cite{huHowFunctionalProgramming2015, hughesWhyFunctionalProgramming1989}, which makes these components easier to drop into the existing library and be used by domain specific library developers to build new libraries that work with existing ones.
 While many of these elements have been partially described elsewhere, our framework naturally combines them in a unified framework based on ideas in category theory.
@@
 There is an assumption in visualization that the connectivity of the elements in a dataset are preserved in the visual representation; roughly topology provides a method for encoding this connectivity \cite{wilkinsonGrammarGraphics2005}.
-Bertin gives the examples of discrete data values should be represented by discrete entities such as points, quantitive continuous data by lines, and surfaces through area marks \cite{bertinSemiologyGraphicsDiagrams2011a}.
+Bertin gives the examples of discrete data values should be represented by discrete entities such as points, quantitative continuous data by lines, and surfaces through area marks \cite{bertinSemiologyGraphicsDiagrams2011}.
 In \autoref{fig:related-work:continuity:ktypes} the records of the GHCN\cite{lawrimoreGlobalHistoricalClimatology2011} weather station table can be considered disconnected from each other and therefore have a 0D continuity.
@@
         X \arrow[r, "f"] \arrow[rrr, "h\circ(g\circ f) = (h\circ g)\circ f"', bend right, dashed] & Y  \arrow[r, "g"] & Z \arrow[r, "h"] & W
-        \end{tikzcd}}
+        \end{tikzcd}
   \end{equation*}
@@
 \end{definition}
-Spivak then constructs tables as sections $\dsection: \dbase \rightarrow \Gamma^{\pi}(\sigma)$ from a key space $\dspace$ to the set of records $\Gamma^{\pi}(\sigma)$ on the schema bundle.
+Spivak then constructs tables as sections $\dsection: \dbase \rightarrow \Gamma^{\pi}(\sigma)$ from a key space $\dbase$ to the set of records $\Gamma^{\pi}(\sigma)$ on the schema bundle.
 Generalizing Spivak's formulation of tables, we define the fiber space in this work as the set of record functions
@@
 The presheaf is contravariant because the morphisms between the input openset objects go in the opposite direction from the morphisms between the output set objects.
-The expectation that data containers implement inclusion and restriction morphisms means that, for example, if a data container implements subsetting (inclusion) of a continuous base space than it must ensure that the sections can be interpolated such that they can be defined over the subspace.
+The expectation that data containers implement inclusion and restriction morphisms means that, for example, a data container mist ensure that sections can be interpolated over any subset that the container can provide.
 
@@
 We generalize binary operations to actions $\dfunc \in \dfuncset$ on the sheaf $\sheaf_{\dbase, \dtotal}$.
-\begin{definition} An \textbf{action}  $\act: \dfuncsetc \times \dtotalc \rightarrow \sheaf_{\dbase, \dtotal}$ such that 
+
+\begin{definition} 
+  An \textbf{action}  $act: \dfuncsetc \times \dtotalc \rightarrow \sheaf_{\dbase, \dtotal}$ is a function that expresses that elements of set $\dfuncset$ act on elements of the sheaf of sections.Actions are associative such that
   \begin{equation}
@@
   \end{equation}
-where $\dfunc_{i}, \dfunc_{j}$ are elements of the set $\dfuncset$ equipped with operator $\circ$ and \dsection\ is an element of the sheaf  $\sheaf_{\dbase, \dtotal}$.\cite{nlab:action}
+ for all elements  $\dfunc_{i}, \dfunc_{j}$ of the set $\dfuncset$ equipped with operator $\circ$ and all elements \dsection\ of the sheaf  $\sheaf_{\dbase, \dtotal}$. \cite{nlab:action}
 \end{definition}.
@@
 where $\dsection, \dsection^{\prime} \in \Gamma(\openset^{\prime}, \dfuncpull\dtotal\restriction_{\openset^{\prime}})$.
-Since there are many different fiber transformations,  there is not a uniform way of expressing how this morphism acts on sections unlike in \label{eq:atct:morphism:verify_base}.
+Since there are many different fiber transformations,  there is not a uniform way of expressing how this morphism acts on sections unlike in \autoref{eq:atct:morphism:verify_base}.
 Examples of a fiber transformation are actions such as partial ordering for ranking purpose \cite{bruggemannRankingPrioritizationMultiindicator2011} and the permutation, ordering, translation, and scaling codified as Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946m, leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}.
@@
   \label{eq:construction:nu:fabrication}
-  \begin{tikzcd}[col sep=Huge]
+  \begin{tikzcd}[col sep=huge]
     \dfiberc_{\dbasepointc} 
@@
 \end{minted}
-As shown in the \texttt{draw} method, generating a graphic section $\gsection$ is implemented as the composition of $\texttt{qhat} \approx \vmark$ and $\texttt{nu} \approx \vchannel$ applied to a local section of the sheaf $\textttt{self.section.query} \approx \desection^{i}$ such  $\texttt{draw} \approx \vmarkc\circ\vchannelc\circ\dsectionc  = \vartistc\circ \dsectionc$.
+As shown in the \texttt{draw} method, generating a graphic section $\gsection$ is implemented as the composition of $\texttt{qhat} \approx \vmark$ and $\texttt{nu} \approx \vchannel$ applied to a local section of the sheaf $\texttt{self.section.query} \approx \desection^{i}$ such  $\texttt{draw} \approx \vmarkc\circ\vchannelc\circ\dsectionc  = \vartistc\circ \dsectionc$.
 The $\vchannel$ and $\vmark$ functions shown here are written such that they can generate a visual element given a local section $\dsection\restriction_{\dbase^{i}}$ which can be as little or large as needed.
@@
 % argument is your BibTeX string definitions and bibliography database(s)
-\bibliography{bibliography}
+\bibliography{IEEEabrv, bibliography}
 

### 2022-07-13 paper/main.tex
fixed most reference issues

@@
 \usepackage[utf8]{inputenc}
-\usepackage{subfig} %ieee does not like subfigure
 \usepackage{multicol}
@@
 \usepackage{url}
-
+% \usepackage[breaklinks=true]{hyperref} breaks compilation?
 % *** Do not adjust lengths that control margins, column widths, etc. ***
@@
 The mathematical term for this is equivariance and an equivariant map is a map that transforms input with structure to output with compatible structure.
-Bertin codified the idea that visual encoding of a data field should match the structure of the field; for example ordered data should be encoded as an ordered visual element or quantitative data as a quantitative visual element \cite{bertinSemiologyGraphicsDiagrams2011a}.
+Bertin codified the idea that visual encoding of a data field should match the structure of the field; for example ordered data should be encoded as an ordered visual element or quantitative data as a quantitative visual element \cite{bertinSemiologyGraphicsDiagrams2011}.
 This structure mapping is necessary to ensure that the visual representations in graphs are not distorting the data\cite{tufteVisualDisplayQuantitative2001} and graphs are easier to understand when properties match \cite{norman_things_smart}.
@@
 the data container) changes, that data maps unambiguously to visual elements \cite{ziemkiewiczEmbeddingInformationVisualization2009}, that changing the data should correspond to changes in the visualization in a perceptually significant manner.
-In \autoref{sec:atct:sheaves} we introduce a uniform abstract data representation layer, the expectation of unambiguous mappings is expressed in \autoref{sec:atct:xi}, and the definition of equivariance introduced in \autoref{sec:artist:equivariance} includes the correspondence between data and visual changes.
+In \autoref{sec:atct:sheaves} we introduce a uniform abstract data representation layer, the expectation of unambiguous mappings is expressed in \autoref{sec:atct:xi}, and the definition of equivariance introduced in \autoref{sec:artist:equivariant:artist} includes the correspondence between data and visual changes.
 In this work, we formally specify the conditions necessary to build structure preserving components using category theory, because it provides a rich language for expressing the structure of objects, functions between objects, and the constraints that must hold for these functions to compose \cite{wielsManagementEvolvingSpecifications1998,yorgeyMonoidsThemeVariations}.
@@
 On the other hand, the topology of a table as a set of discrete rows, as illustrated in \autoref{fig:related-work:continuity:ktypes}, is assumed by
-A Presentation Tool\cite{mackinlayAutomatingDesignGraphical1986, mackinlayAutomatingDesignGraphical1986} and grammar of graphics\cite{wilkinsonGrammarGraphics2005} and the ggplot\cite{wickhamGgplot2ElegantGraphics2016a}, vega\cite{satyanarayanDeclarativeInteractionDesign2014}, and altair\cite{vanderplasAltairInteractiveStatistical2018} libraries built on these frameworks.
+A Presentation Tool\cite{mackinlayAutomatingDesignGraphical1986, mackinlayAutomatingDesignGraphical1986} and grammar of graphics\cite{wilkinsonGrammarGraphics2005} and the ggplot\cite{wickhamGgplot2ElegantGraphics2016}, vega\cite{satyanarayanDeclarativeInteractionDesign2014}, and altair\cite{vanderplasAltairInteractiveStatistical2018} libraries built on these frameworks.
 Image libraries such as Napari\cite{nicholas_sofroniew_2021_4533308} and ImageJ\cite{schneiderNIHImageImageJ2012} and its humanaties data ImagePlot\cite{studiesCulturevisImageplot2021} plugin assumed that the input is 2D continuous.
@@
 \end{definition}
-The first condition for a fiber bundle means that all points in a fiber space over a point in the base space map to the same base space point, which we rely on to establish that there is a formal binding between typed data value, discussed in \autoref{sec:atct:fiber:fiber} and the corresponding index, discussed in \autoref{sec:atct:fiber:base}.
-The condition of local triviality means that the fibers are identical over open neighborhoods and is what allows fiber bundles to be extremely generalizable, as discussed in \autoref{sec:atct:fiber:sections}
+The first condition for a fiber bundle means that all points in a fiber space over a point in the base space map to the same base space point, which we rely on to establish that there is a formal binding between typed data value, discussed in \autoref{sec:atct:fb:fiber} and the corresponding index, discussed in \autoref{sec:atct:fb:base}.
+The condition of local triviality means that the fibers are identical over open neighborhoods and is what allows fiber bundles to be extremely generalizable, as discussed in \autoref{sec:atct:fb:sections}
 
@@
 \end{definition}
-In \autoref{sec:atct:fiber:sections}, we propose that modeling data as sections provides a way to encapsulate topological and field structure in a uniform dimension and type independent manner.
+In \autoref{sec:atct:fb:sections}, we propose that modeling data as sections provides a way to encapsulate topological and field structure in a uniform dimension and type independent manner.
 
 \subsubsection{Topological Structure: Base Space \dbase}
-\autoref{sec:atct:fiber:base}
+\label{sec:atct:fb:base}
 We encode the underlying connectivity of the data as the \textcolor{base}{base space} \dbase\ of the fiber bundle because it provides a uniform language for describing the topology regardless of how the data is subset.
@@
 \subsubsection{Data Fields: Fiber Space \dfiber}
-\autoref{sec:atct:fb:fiber}
+\label{sec:atct:fb:fiber}
 As mentioned in \autoref{sec:related-work:equivariance}, equivariance is traditionally evaluated as the preservation of field structure, which is based on the field type.
@@
 
-\subsubsection{Data: Section \dsection} 
-\autoref{sec:atct:fb:sections}
+\subsubsection{Data: Section} 
+\label{sec:atct:fb:sections}
 We encode data as a \textcolor{section}{section} \dsectionc\ of a bundle because this allows us to incorporate the topology and field types in the data definition.
@@
 
-Sheaves are built out of presheaves, which are functors, meaning they are functions from objects of one category to objects of another category\cite{WhatFunctorDefinitions} that preserve the morphisms in each category.
+Sheaves are built out of presheaves, which are functors, meaning they are functions from objects of one category to objects of another category\cite{bradleyWhatFunctorDefinitions} that preserve the morphisms in each category.
 \begin{definition} A \textbf{functor} is a map $F: \mathcal{C} \rightarrow \mathcal{D}$ such that every object $C \in \mathcal{OB(C)}$ has a corresponding object $F(C) \in \mathcal{OB(D)}$ in category $\mathcal{D}$.
@@
 \begin{equation*}
+  \label{eq:atct:sheaf:inclusion}
   \iota: \openset_1 \rightarrow \openset_2
@@
 \begin{equation*}
+  \label{eq:atct:sheaf:restriction}
   \iota^*: \cgamma{\openset_2}{\dtotal_{\openset_2}} \rightarrow \cgamma{\openset_1}{\dtotal_{\openset_1}}
@@
 take as input an openset object \opensetc\ or \opensetgc\ and return sets of data and graphic sections that are objects in \setb.
-As a map between these sheaf functors, the artist has to preserve the $\iota, \iota^*$ morphisms of the presheaf functor, described in \autoref{eq:atct:presheaf}, such that the following diagram commutes
+As a map between these sheaf functors, the artist has to preserve the $\iota, \iota^*$ morphisms of the presheaf functor, described in \autoref{eq:atct:sheaf:inclusion} and \autoref{eq:atct:sheaf:restriction}, such that the following diagram commutes:
 
@@
 \end{equation}
- The diagram in \autoref{eq:artist:natural_transform} shows that restricting a set of outputs of an artist to a set of graphic sections over a subspace is equivalent to restricting the inputs to data sections over the same subspace.
+ The diagram in \autoref{eq:artist:natural_transform:inclusions} shows that restricting a set of outputs of an artist to a set of graphic sections over a subspace is equivalent to restricting the inputs to data sections over the same subspace.
 Because the artist is a functor of sheaves, the artist is expected to tra nslate the data continuity to graphic continuity such that the connectivity of subsets is preserved.
@@
 \end{equation} 
-that are, by definition, only reachable through a structure preserving artist, which we describe in \autoref{sec:artist:equivariance}.
+that are, by definition, only reachable through a structure preserving artist, which we describe in \autoref{sec:artist:equivariant:artist}.
 We define this subset because the space of all sections $\cgamma{\opensetg}{\gtotal\restriction_{\openset}}$ includes sections that may not be structure preserving.
@@
 We separate data transformations into two components, transformations on the base space $(\dfunchc, \dfuncpullc)$ and transformations on the fiber space $\dfunctc$.
-\begin{equation}
-  \label{eq:atct:sheaves:monoid_morphism}
+
+\begin{equation}
+  \label{eq:artist:sheaves:monoid_morphism}
   \begin{tikzcd}
@@
 
-As introduced in \autoref{def:atct:io:fiber}, the fiber transformation $\dfunct: \dfuncpull \dtotal_{\dbasepoint^{\prime}} \rightarrow \dfuncpull \dtotal_{\dbasepoint^{\prime}}$ is a morphism on the fiber $\dfunct \in Hom(\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}},\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}})$ restricted to a point $\dbasepoint^{\prime} \in \openset^{\prime}$.
+As introduced in \autoref{eq:artist:sheaves:monoid_morphism}, the fiber transformation $\dfunct: \dfuncpull \dtotal_{\dbasepoint^{\prime}} \rightarrow \dfuncpull \dtotal_{\dbasepoint^{\prime}}$ is a morphism on the fiber $\dfunct \in Hom(\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}},\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}})$ restricted to a point $\dbasepoint^{\prime} \in \openset^{\prime}$.
 The fiber transformation is a change in section
@@
 Since there are many different fiber transformations,  there is not a uniform way of expressing how this morphism acts on sections unlike in \autoref{eq:atct:morphism:verify_base}.
-Examples of a fiber transformation are actions such as partial ordering for ranking purpose \cite{bruggemannRankingPrioritizationMultiindicator2011} and the permutation, ordering, translation, and scaling codified as Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946m, leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}.
+Examples of a fiber transformation are actions such as partial ordering for ranking purpose \cite{bruggemannRankingPrioritizationMultiindicator2011} and the permutation, ordering, translation, and scaling codified as Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946, leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}.
 
@@
 We formalize this structure preservation as equivariance, which is that for every morphism on the data $(\dfunch_{\dtotal}, \dfunct_{\dtotal})$ there is an equivalent morphism on the graphic  $(\dfunch_{\gtotal}, \dfunct_{\gtotal})$ The artist is an equivariant map if the diagram commutes for all points $\gbasepointc^{\prime}\in \gbasec^{\prime}$
-\label{sec:artist:equivariance}
-\begin{equation}
-  \autoref{eq:artist:equivariance}
+
+\begin{equation}
+  \label{eq:artist:equivariance}
   \begin{tikzcd}[ampersand replacement=\&, column sep=small]
@@
 \end{figure}
-For example, in \autoref{fig:artist:equivarance}, the measurable variable is color.
+
+For example, in \autoref{fig:artist:equivariance}, the measurable variable is color.
 This is a visual representation of the data shown in \autoref{fig:atct:phi}, and as such the equivariant transformations are an equivalent transposition and scaling of the colors.
@@
 
-\subsubsection{Addition: +}
+\subsubsection{Addition}
 \label{sec:artist:addition}
@@
 
-\subsubsection{Multiplication: $\times$}
+\subsubsection{Multiplication}
 \label{sec:artist:operator:multiplication}
@@
 
-We propose that one way of constructing artist functions is to separate generating a visualization, as shown in \autoref{fig:construction:flow}, is into an equivariant continuity preserving encoding stage $\vchannelc$ and a equivariant continuity preserving compositing stage $\vmarkc$.
+We propose that one way of constructing artist functions is to separate generating a visualization into an encoding stage $\vchannelc$ and a compositing stage $\vmarkc$.
 In the \textcolor{artist}{encoding} stage $\vchannelc$, a data bundle is treated as separable fields and each field is mapped to a measurable visual variable.
@@
 \label{sec:construction:vtotal}
-As shown in \autoref{fig:construction:flow}, we propose an intermediate visual fiber bundle 
+We propose an intermediate visual fiber bundle 
 \begin{equation}
@@
 \label{sec:construction:nu}
-As introduced in \autoref{fig:construction:flow}, the encoding function $\vchannel$ is a map from data sections to graphic sections
+The encoding function $\vchannel$ is a map from data sections to graphic sections
 \begin{equation}
@@
   \label{eq:construction:nu:fabrication}
-  \begin{tikzcd}[col sep=huge]
+  \begin{tikzcd}[column sep=huge]
     \dfiberc_{\dbasepointc} 
@@
 which means that, as shown in \autoref{eq:construction:nu:fabrication}, any collection of $\vchannel$ functions can be composed such that they are equivalent to a $\vchannel$ that directly converts the input to the output.
-As with artists, $\vchannel$ are maps of sections such that the composition operators defined in \autoref{sec:artist:composition} can also act on transformers $\vchannel$, meaning that encoders can be added $\vchannel_{a+b} = \vchannel_{a} + \vchannel_{b}$ and multiplied d $\vchannel_{a\times b} = \vchannel_{a}  \vchannel_{b}$.
+As with artists, $\vchannel$ are maps of sections such that the operators defined in \autoref{sec:artist:operators} can also act on transformers $\vchannel$, meaning that encoders can be added $\vchannel_{a+b} = \vchannel_{a} + \vchannel_{b}$ and multiplied d $\vchannel_{a\times b} = \vchannel_{a}  \vchannel_{b}$.
 Encoders designed to satisfy these composability constraints provide for a rich set of building blocks for implementing complex encoders.
@@
 
-An encoding function $\vchannelc$ is equivariant if the change in data, as defined in \autoref{sec:atct:sheaves:measurments}, and change in visual components are equivariant.
+An encoding function $\vchannelc$ is equivariant if the change in data, as defined in \autoref{sec:artist:equivariant:data}, and change in visual components are equivariant.
 Since $\dtotal$ and $\vtotal$ are over the same base space and are pointwise, the base space change $\dfunch_{\dtotal}$ applies to both sides of the equation
@@
 The change in visual components $\dfunct_{\vtotal}$ is dependent both on $\dfunct_{\dtotal}$ and the choice of visual encoding.
-As mentioned in \autoref{sec:related-work:equivariance}, this is why Bertin and many others since have advocated choosing an encoding that has a structure that matches the data structure\cite{bertinSemiologyGraphicsDiagrams2011a}.
+As mentioned in \autoref{sec:related-work:equivariance}, this is why Bertin and many others since have advocated choosing an encoding that has a structure that matches the data structure\cite{bertinSemiologyGraphicsDiagrams2011}.
 For example choosing a quantative colormap to encode quantative data if the $\dfunctc$ operation is scaling, as in \autoref{fig:artist:equivariance}.
@@
 \subsection{Graphic Compositor}
-As shown in \autoref{fig:construction:flow}, the compositor function $\vmarkc$ transforms the measurable components into properties of a visual element.
+The compositor function $\vmarkc$ transforms the measurable components into properties of a visual element.
 The compositing function $\vmarkc$ transforms the sections of visual elements $\vsectionc$ into sections of graphics $\gsectionc$.
@@
 The composition in \autoref{eq:construction:q:fabrication} means that different measurable components can yield the same visual elements.
-The composition operators defined in \autoref{sec:artist:composition} can also act on compositors $\vmark$ such that $\vmark_{a+b} = \vmark_{a} + \vmark_{b}$ and multiplied d $\vmark_{a\times b} = \vmark_{a}  \vmark_{b}$.
+The operators defined in \autoref{sec:artist:operators} can also act on compositors $\vmark$ such that $\vmark_{a+b} = \vmark_{a} + \vmark_{b}$ and multiplied d $\vmark_{a\times b} = \vmark_{a}  \vmark_{b}$.
 
@@
 
-A compositor function \vmark\ is equivariant if the renderer output changes in a way equivariant to the data transformation defined in \autoref{sec:atct:sheaves:measurments}.
+A compositor function \vmark\ is equivariant if the renderer output changes in a way equivariant to the data transformation defined in \autoref{sec:artist:equivariant:data}.
 This means that a change in base space $\dfunch_{\dtotalc}$ should have an equivalent change in visual element base space.
@@
 
-\subsection{Implementing the Artist: $\vartistc=\vmarkc\circ\vchannelc$}
+\subsection{Implementing the Artist}
 
@@
 \end{equation}
+
 Since the artist is a family of functions in the homeset between sheaves, the isomorphism allows for the specification of the transformation from data as combination of functions over different spaces such that the following diagram commutes:
@@
 We construct a toy dataset with a discrete $\dbase$ of 4 points and a fiber space of $\dfiber=\{apple,\, orange,\, lemon\}\times \mathbb{Z}^{+} \times \{\texttt{True}, \texttt{False}\}$.
-We thinly wrap \autoref{tab:discussion:data} in an object so that the common data interface function is that $\tau = \ttexttt{DataContainerObject.query}$.
+We thinly wrap \autoref{tab:discussion:data} in an object so that the common data interface function is that $\tau = \texttt{DataContainerObject.query}$.
 \begin{minted}{python}
@@
 \end{minted}
-As shown in the \texttt{draw} method, generating a graphic section $\gsection$ is implemented as the composition of $\texttt{qhat} \approx \vmark$ and $\texttt{nu} \approx \vchannel$ applied to a local section of the sheaf $\texttt{self.section.query} \approx \desection^{i}$ such  $\texttt{draw} \approx \vmarkc\circ\vchannelc\circ\dsectionc  = \vartistc\circ \dsectionc$.
+As shown in the \texttt{draw} method, generating a graphic section $\gsection$ is implemented as the composition of $\texttt{qhat} \approx \vmark$ and $\texttt{nu} \approx \vchannel$ applied to a local section of the sheaf $\texttt{self.section.query} \approx \dsection^{i}$ such  $\texttt{draw} \approx \vmarkc\circ\vchannelc\circ\dsectionc  = \vartistc\circ \dsectionc$.
 The $\vchannel$ and $\vmark$ functions shown here are written such that they can generate a visual element given a local section $\dsection\restriction_{\dbase^{i}}$ which can be as little or large as needed.
@@
 % argument is your BibTeX string definitions and bibliography database(s)
-\bibliography{IEEEabrv, bibliography}
+\bibliography{bibliography}
 

### 2022-07-27 paper/main.tex
intro

@@
 % *** PDF, URL AND HYPERLINK PACKAGES ***
-\usepackage{url}
-% \usepackage[breaklinks=true]{hyperref} breaks compilation?
+\usepackage{xurl} 
+%\usepackage{hyperref} breaks compilation?
 % *** Do not adjust lengths that control margins, column widths, etc. ***
@@
 \label{sec:related-work:software}
-Visualization libraries are in part measured by how expressive the components of the library are, where expressiveness is a measure of which structure preserving mappings a tool can implement \cite{mackinlayAutomaticDesignGraphical1987}.
+Visualization libraries are in part measured by how expressive the components of the library are, where expressiveness is a measure of which structure preserving mappings a tool can implement \cite{mackinlayAutomatingDesignGraphical1986}.
 While some visualization tools aim to automate the pairing of data with structure preserving visual representations, such as Tableau\cite{StoltePolaris2002,hanrahanVizQL2006,MackinlayShowme2007}, many visualization libraries leave that choice to the user.
@@
 We model data using a mathematical representation of data that can encode topological properties, field types, and data values in a uniform manner using a structure from algebraic topology called a fiber bundle.
-We extend Butler's proposal of bundles as abstract visualization data type \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989} by incorporating Spivak's methodology for encoding named data types from his fiber bundle representation of relational databases \cite{spivakDatabasesAreCategories2010, spivakSIMPLICIALDATABASES}.
+We extend Butler's proposal of bundles as abstract visualization data type \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989} by incorporating Spivak's methodology for encoding named data types from his fiber bundle representation of relational databases \cite{spivakDatabasesAreCategories2010,spivakSimplicialDatabases2009}.
 We build on this work to describe how to encode the connectivity of the data as a topological space, separately encode the fields as their own topological space with a typing system, and express the mappings between these two spaces.
@@
 As mentioned in \autoref{sec:related-work:equivariance}, equivariance is traditionally evaluated as the preservation of field structure, which is based on the field type.
-To have a rich language for expressing field type, we generalize Spivak's fiber bundle formulation of table \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES}.
-
+To have a rich language for expressing field type, we generalize Spivak's fiber bundle formulation of table \cite{spivakDatabasesAreCategories2010,spivakSimplicialDatabases2009}.
+\note{def 3.5 and 3.6 should maybe be pushed back to prose as 3.7 is the only one we rely on - it defines what tau returns}
 \begin{definition} A \textbf{type specification} is the bundle map $\pi: \mathscr{U} \rightarrow \textbf{DT}$.
@@
 
-While the presheaf is a generalization of the concept of a global space of sections $\Gamma(\dbase, \dtotal)$ \cite{spanier1989algebraic}, the locality and gluing axioms \cite{bakerMathsSheaf} of a sheaf describe the conditions for verifying that the sections over subspaces are part of a cohesive section over the whole space.
+While the presheaf is a generalization of the concept of a global space of sections $\Gamma(\dbase, \dtotal)$ \cite{spanier1989algebraic}, the locality and gluing axioms \cite{bakerEuclideanSpaceMathsSheaf} of a sheaf describe the conditions for verifying that the sections over subspaces are part of a cohesive section over the whole space.
 \begin{definition} A \textbf{sheaf} is a presheaf that satisfies the following two axioms
@@
 % http://www.michaelshell.org/tex/ieeetran/bibtex/
+
 \bibliographystyle{IEEEtran}

### 2022-07-28 paper/main.tex
intro

@@
 \IEEEraisesectionheading{\section{Introduction}\label{sec:intro}}
-\note{need a math check + rough cut on sections 4+}
-
-\note{Spelling out vis library component behavior w/ alg top + cat theory lets us be super general in terms of inputs/outputs (what data, what vis) while super specific about the constraints the components need to satisfy to transform that input to that output in a structure preseving way}
-
-\IEEEPARstart{V}isualizations are expected to be a faithful translation of the inherent structure of the data being visualized; therefore the components of visualization software libraries that translate data to graphics are also expected to preserve this structure.
-Motivated by a need to develop new library components that adapt to modern data needs, especially high dimensional multivariate and distributed data, we propose a framework that topological and field structure of the input data that must be mapped to the visualization and which changes to data must be expressed in visualization, and a a functional design framework that describes composability in terms of functional algebraic operations.
-\note{this may be a good place to incorporate types} This allows complex visualizations to be built from simple verifiable parts\cite{huHowFunctionalProgramming2015, hughesWhyFunctionalProgramming1989}, which makes these components easier to drop into the existing library and be used by domain specific library developers to build new libraries that work with existing ones.
-While many of these elements have been partially described elsewhere, our framework naturally combines them in a unified framework based on ideas in category theory.
-Further we see the application of these ideas in low level visualization library design as our motivation, and we hope these ideas will help shape the reorganization of critical data visualization libraries, such as Matplotlib.
-The contribution of this paper is a uniform method for describing the structure in data and a mathematical framework for describing, constructing, and testing structure preserving visualization library components.
+
+\IEEEPARstart{V}isualizations are expected to be a faithful translation of data, meaning that inherent structure in the data should be present in the visualization; therefore the components of visualization software libraries that translate data to graphics are also expected to be structure preserving.
+This inherent structure can be quite complex, such as when the data is high dimensional, multivariate, deeply nested, or encoded in a distributed manner.
+We propose that we can rigorously specify what structure the library components are expected to preserve using category theory.
+We also propose that these specifications can generalize to most structure types by extending previous work on encoding  inherent structure in a consistent manner using mathematical structures from algebraic topology.
+These algebraic structures allow us to encode inherent structure as programmatic types, and the use of category theory facilitates developing a functional design framework that describes composability in terms of functional algebraic operations.
+The typing system and composability inherent in a functional design encourage library developers to build complex visualization components from simple verifiable parts that have few side effects by design\cite{huHowFunctionalProgramming2015, hughesWhyFunctionalProgramming1989}.
+These categorically specified components could be built as a standalone library, but can also be integrated into existing libraries because they are inherently self-contained.
+Furthermore we see the application of these ideas in low level visualization library design as our motivation, and we hope these ideas will help shape the reorganization of critical data visualization libraries, such as Matplotlib.
+The contribution of this paper is a mathematical framework for describing structure preserving visualization components in a manner that is generalizable to different types of inherent structure.
+This framework provides also provides guidence for the construction and testing of structure preserving visualization library components.
 

### 2022-07-28 paper/main.tex
related work tp and sw

@@
 We also propose that these specifications can generalize to most structure types by extending previous work on encoding  inherent structure in a consistent manner using mathematical structures from algebraic topology.
-These algebraic structures allow us to encode inherent structure as programmatic types, and the use of category theory facilitates developing a functional design framework that describes composability in terms of functional algebraic operations.
-The typing system and composability inherent in a functional design encourage library developers to build complex visualization components from simple verifiable parts that have few side effects by design\cite{huHowFunctionalProgramming2015, hughesWhyFunctionalProgramming1989}.
+These algebraic structures allow us to encode inherent structure as programmatic types, and the use of category theory facilitates developing a functional design framework that describes function composition in terms of functional algebraic operations.
+The typing system and composition inherent in a functional design encourage library developers to build complex visualization components from simple verifiable parts that have few side effects by design\cite{huHowFunctionalProgramming2015, hughesWhyFunctionalProgramming1989}.
 These categorically specified components could be built as a standalone library, but can also be integrated into existing libraries because they are inherently self-contained.
@@
 The contribution of this paper is a mathematical framework for describing structure preserving visualization components in a manner that is generalizable to different types of inherent structure.
-This framework provides also provides guidence for the construction and testing of structure preserving visualization library components.
+This framework provides also provides guidance for the construction and testing of structure preserving visualization library components.
 
@@
 In \autoref{fig:related-work:continuity:ktypes} the records of the GHCN\cite{lawrimoreGlobalHistoricalClimatology2011} weather station table can be considered disconnected from each other and therefore have a 0D continuity.
-This means each distinct row can be reduced to a discrete point and it is expected  that each rows will be encoded as a discrete visual element, such as a standalone scatter point.
+This means each distinct row can be reduced to a discrete point and each row will be encoded as a discrete visual element, such as a standalone scatter point.
 Each station can also be considered a time series of weather observation since temperature and pressure are sampled from a 1D continuous measurement space.
@@
 Instead of interpolation, which is how continuity was preserved in the timeseries plot, in this instance the connectivity is preserved by plotting the stations in the map such that their relative distances are preserved.
-Encoding connectivity using the formalism of topological spaces allows us to encode the connectivity of the data in a uniform manner such that we can then generalize the preservation of that connectivity in the visual representation rather than constructing a condition specifically for each type of topology.
-
-Visual algorithms assume the topology of their input data, as described in taxonomies of visualization algorithms Chi and by Troy and M\"{o}ller \cite{chiTaxonomyVisualizationTechniques2000,toryRethinkingVisualizationHighlevel2004}.
+Encoding connectivity using the formalism of topological spaces allows us to encode the connectivity of the data in a uniform manner such that we can then generalize the preservation of that connectivity in the visual representation rather than constructing a condition specifically for each type of topological structure.
+
+Visual algorithms assume the topology of their input data, as described in taxonomies of visualization algorithms Chi\cite{chiTaxonomyVisualizationTechniques2000} and by Troy and M\"{o}ller \cite{toryRethinkingVisualizationHighlevel2004}.
 For example, a \texttt{line} algorithm often does not have a way to query whether a list of (x,y) coordinates is the distinct rows, the time series, or the list of stations in \autoref{fig:related-work:continuity:ktypes}.
@@
 The typing system proposed in this paper builds on Butler's proposal of using a mathematical structure called fiber bundles as an abstract data representation in visualization \cite{butlerVectorBundleClassesForm1992, butlerVisualizationModelBased1989}.
+We extend Butler's work because he discusses how bundles can be used to encode data in a consistent manner for a number of different \note{arbitrary?} topological and field structures.
  We sketch out fiber bundles in \autoref{sec:atct:fiber-bundles}, but his work provides a thorough introduction to bundles for visualization practitioners.
@@
 While some visualization tools aim to automate the pairing of data with structure preserving visual representations, such as Tableau\cite{StoltePolaris2002,hanrahanVizQL2006,MackinlayShowme2007}, many visualization libraries leave that choice to the user.
-Libraries that provide small buildable components \cite{wongsuphasawatNavigatingWideWorld2021}, such as components for making boxes or translating data values to colors, mostly embed their connectivity assumptions in their visual algorithms and allow the user choose almost any component, regardless of structure preservation.
-Domain specific libraries are designed with the assumption of continuities that are common in the domain \cite{HeerSoftware2006}, and therefore can somewhat restrict the space of choices to algorithms that are appropriate for the data.
-Buildable component libraries such as matplotlib\cite{hunterMatplotlib2DGraphics2007} and D3\cite{bostockDataDrivenDocuments2011} support a wide variety of data types, but at the cost of visual algorithms with  parameters that are tuned to the expected data in a way that leaves the over all library feeling uncohesive.
-VTK\cite{hanwellVisualizationToolkitVTK2015,geveciVTK2012} provides a language for expressing the topological properties of the data, but does so in a non-uniform manner.
-On the other hand, the topology of a table as a set of discrete rows, as illustrated in \autoref{fig:related-work:continuity:ktypes}, is assumed by
+For example, connectivity assumptions tend to be embedded in each of the visual algorithms of `building block` libraries, a term coined by by Wongsuphasawat \cite{wongsuphasawatNavigatingWideWorld2021} to describe libraries that provide modular components for building elements of a visualization, such as functions for making boxes or translating data values to colors.
+In building block libraries such as Matplotlib\cite{hunterMatplotlib2DGraphics2007} and D3\cite{bostockDataDrivenDocuments2011} the connectivity assumptions lead to very different interfaces; for example in Matplotlib methods for updating the data in a plot differ across the various plotting objects and the parameter list is inconsistent as to which parameters can be applied per data point or for the whole element, and in D3 the API for network visualizers is mostly distinct from the rest of the modules.
+While VTK\cite{hanwellVisualizationToolkitVTK2015,geveciVTK2012} provides a language for expressing the topological properties of the data, and therefore can embed that information in its visual algorithms, it does so in a non-uniform manner.
+On the other hand, domain specific libraries are designed with the assumption of continuities that are common in the domain \cite{HeerSoftware2006}, and therefore can somewhat restrict the interface to choices that that are appropriate for the domain.
+For example, a tabluar topological structure of discret rows, as illustrated in \autoref{fig:related-work:continuity:ktypes}, is assumed by
 A Presentation Tool\cite{mackinlayAutomatingDesignGraphical1986, mackinlayAutomatingDesignGraphical1986} and grammar of graphics\cite{wilkinsonGrammarGraphics2005} and the ggplot\cite{wickhamGgplot2ElegantGraphics2016}, vega\cite{satyanarayanDeclarativeInteractionDesign2014}, and altair\cite{vanderplasAltairInteractiveStatistical2018} libraries built on these frameworks.
-Image libraries such as Napari\cite{nicholas_sofroniew_2021_4533308} and ImageJ\cite{schneiderNIHImageImageJ2012} and its humanaties data ImagePlot\cite{studiesCulturevisImageplot2021} plugin assumed that the input is 2D continuous.
-Networking libraries such as gephi\cite{bastianGephiOpenSource2009} and networkx\cite{HagbergExploringNetwork2008} assume a graph like structure.
-These domain specific libraries tend to benefit from more cohesive apis, but at the cost of a much more limited set of visualization algorithms than the building block libraries.
+Image libraries such as Napari\cite{nicholas_sofroniew_2021_4533308} and ImageJ\cite{schneiderNIHImageImageJ2012} and its humanities ImagePlot\cite{studiesCulturevisImageplot2021} plugin assume that the input is 2D continuous.
+Networking libraries such as gephi\cite{bastianGephiOpenSource2009} and networkx\cite{HagbergExploringNetwork2008} assume a graph-like structure.
+By assuming the structure of their data, these domain specific libraries can provide more cohesive interfaces for a much more limited set of visualization algorithms than the building block libraries offer.
 

### 2022-07-28 paper/main.tex
equivariance section of rw

@@
 Along with preserving topology, visualization algorithms are expected to preserve the structure of the data fields, i.e.
-the column types in a table, when they translate data to visual representations.
-The mathematical term for this is equivariance and an equivariant map is a map that transforms input with structure to output with compatible structure.
-Bertin codified the idea that visual encoding of a data field should match the structure of the field; for example ordered data should be encoded as an ordered visual element or quantitative data as a quantitative visual element \cite{bertinSemiologyGraphicsDiagrams2011}.
-This structure mapping is necessary to ensure that the visual representations in graphs are not distorting the data\cite{tufteVisualDisplayQuantitative2001} and graphs are easier to understand when properties match \cite{norman_things_smart}.
-The expectation of matching structure is generalized as a mapping of a binary operator that acts on the data domain to a binary operator that acts on the graphic domain by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}.
-An invertible binary operator is the basis of Kindlemannn and Scheideggers's algebraic model of visualization\cite{kindlmannAlgebraicProcessVisualization2014}.
-They propose three specific types of equivariance: that the visualization should not change if the data representation (i.e.
+the columns in a table, when they translate data to visual representations.
+The field structure of data has traditionally been described using the measurment scales-i.e.
+nominal, ordinal, interval, ratio - which Steven's classified by their mathematical group structure\cite{stevensTheoryScalesMeasurement1946a} and others have formally expanded to include more types \cite{leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}.
+The expectation that visual transformation is one where the measurement structure be matched by the visual encoding was codified by Bertin \cite{bertinSemiologyGraphicsDiagrams2011}.
+An \textit{equivariant} map is the formal term for a function where the input and output have symmetric group structure.
+Given a group $G$ that acts on input $X$ and output $Y$ and a function $f: X \rightarrow Y$, the function $f$ is \textit{equivariant} if $f(g(x)) = g(f(x))$ for all $g$ in $G$ and for all $x$ in $X$\cite{nlab:equivariant}.
+For example, nominal measurements are permutable, meaning that they are acted on by elements of a permutation group.
+For the visual transformation to be equivariant, that nominal measurement must be mapped to a visual encoding that is also permutable, such as distinct marker shapes or color hues.
+Mackinlay\cite{mackinlayAutomaticDesignGraphical1987} more generally definines visualization transformations as \textit{homomorphic} maps.
+A map $f: X \rightarrow Y$ is a \textit{homomorphism} when $X$ and $Y$ have the same algebraic structure type and $f(x_1 \oslash x_2) = f(x_1) \oslash f(x_2)$ for every pair $x_1, x_2$ in $X$\cite{grimaldiDiscreteCombinatorialMathematics2006}.
+Mackinlay's construction encompasses non-invertable structures such as partial orderings.
+In \autoref{sec:artist:equivariant:data}, we will also describe the structure of the data in terms of the functions acting on the data.
+
+Structure preservation is important because it means visualzations are easier to understand \cite{norman_things_smart} and ensures that the visual representations are not distortions of the data\cite{tufteVisualDisplayQuantitative2001}.
+Kindlemannn and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014} propose three specific types of structure preservation: that the visualization should not change if the data representation (i.e.
 the data container) changes, that data maps unambiguously to visual elements \cite{ziemkiewiczEmbeddingInformationVisualization2009}, that changing the data should correspond to changes in the visualization in a perceptually significant manner.
@@
 \subsection{Equivariance}
-\label{sec:artist:equiv}
-Equivariance in a visual context was formally defined as the preservation of a binary operator from one domain to another by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}.
-We generalize binary operations to actions $\dfunc \in \dfuncset$ on the sheaf $\sheaf_{\dbase, \dtotal}$.
+\label{sec:artist:equiv}\note{our basis of structure preservation - is this even equivariance}
+to actions $\dfunc \in \dfuncset$ on the sheaf $\sheaf_{\dbase, \dtotal}$.
 

### 2022-07-28 paper/main.tex
related work equiv

@@
 We build on the extensive work codifying the structure of visualization data to develop a more generalizable method for expressing the structure that visualizations are expected to preserve.
-Broadly, previous work describes structure as a combination of the topological properties of the data, the mathematical structure of the data fields, and equivalent changes to data and graphics.
-We propose that a unified abstraction model can guide developers in building structure aware composable, reusable software library components with more consistent interfaces.
+Structure preservation is important because it means visualzations are easier to understand \cite{norman_things_smart} and ensures that the visual representations are not distortions of the data\cite{tufteVisualDisplayQuantitative2001}.
+Broadly, previous work describes structure as a combination of the topological properties of the data, the mathematical structure of the data fields, and equivalent or compatible changes to data and graphics.
+We propose that a unified abstraction model can guide developers in building structure aware composable software library components with more consistent interfaces.
 
@@
 \label{sec:related-work:equivariance}
-Along with preserving topology, visualization algorithms are expected to preserve the structure of the data fields, i.e.
-the columns in a table, when they translate data to visual representations.
+Along with preserving topology, visualization algorithms are expected to preserve the structure of the data fields, i.e. the columns in a table, when they translate data to visual representations. 
 The field structure of data has traditionally been described using the measurment scales-i.e. 
@@
 The expectation that visual transformation is one where the measurement structure be matched by the visual encoding was codified by Bertin \cite{bertinSemiologyGraphicsDiagrams2011}.
-An \textit{equivariant} map is the formal term for a function where the input and output have symmetric group structure.
+Two mathematical concepts for expressing structure preservation are \textit{equivariance} and \textit{homomorphism}:
+\begin{itemize}
+\item an \textit{equivariant} function preserves symmetric group structure on its input and output.
 Given a group $G$ that acts on input $X$ and output $Y$ and a function $f: X \rightarrow Y$, the function $f$ is \textit{equivariant} if $f(g(x)) = g(f(x))$ for all $g$ in $G$ and for all $x$ in $X$\cite{nlab:equivariant}.
+\item a \textit{homomorphic} function preserves binary operations on input and output of the same algberaic type.
+A map $f: X \rightarrow Y$ is a \textit{homomorphism} when $X$ and $Y$ have the same algebraic structure type and $f(x_1 \oslash x_2) = f(x_1) \oslash f(x_2)$ for every pair $x_1, x_2$ in $X$\cite{grimaldiDiscreteCombinatorialMathematics2006}.
+\end{itemize}
 For example, nominal measurements are permutable, meaning that they are acted on by elements of a permutation group.
 For the visual transformation to be equivariant, that nominal measurement must be mapped to a visual encoding that is also permutable, such as distinct marker shapes or color hues.
-Mackinlay\cite{mackinlayAutomaticDesignGraphical1987} more generally definines visualization transformations as \textit{homomorphic} maps.
-A map $f: X \rightarrow Y$ is a \textit{homomorphism} when $X$ and $Y$ have the same algebraic structure type and $f(x_1 \oslash x_2) = f(x_1) \oslash f(x_2)$ for every pair $x_1, x_2$ in $X$\cite{grimaldiDiscreteCombinatorialMathematics2006}.
-Mackinlay's construction encompasses non-invertable structures such as partial orderings.
-In \autoref{sec:artist:equivariant:data}, we will also describe the structure of the data in terms of the functions acting on the data.
-
-Structure preservation is important because it means visualzations are easier to understand \cite{norman_things_smart} and ensures that the visual representations are not distortions of the data\cite{tufteVisualDisplayQuantitative2001}.
-Kindlemannn and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014} propose three specific types of structure preservation: that the visualization should not change if the data representation (i.e.
-the data container) changes, that data maps unambiguously to visual elements \cite{ziemkiewiczEmbeddingInformationVisualization2009}, that changing the data should correspond to changes in the visualization in a perceptually significant manner.
+Partial orders are non-invertable, so a visual encoding is homomorphic when data that is partially ordered is mapped to visual encodings that are partially ordered in the same way.Mackinlay\cite{mackinlayAutomaticDesignGraphical1987} proposes that data to visual encoders should be \textit{homomorphic} maps in his specification of \textit{A Presentation Tool}.
+The notion that visual transforms are equivariant underlies Kindlemannn and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}'s Algebraic Visualization Design.
+They propose three specific types of structure preservation: that the visualization should not change if the data representation (i.e.
+the data container) changes, that data maps unambiguously to visual elements \cite{ziemkiewiczEmbeddingInformationVisualization2009}, and that changing the data should correspond to changes in the visualization in a perceptually significant manner.
 In \autoref{sec:atct:sheaves} we introduce a uniform abstract data representation layer, the expectation of unambiguous mappings is expressed in \autoref{sec:atct:xi}, and the definition of equivariance introduced in \autoref{sec:artist:equivariant:artist} includes the correspondence between data and visual changes.

### 2022-08-17 paper/main.tex
why fb in human readable terms

@@
 \usepackage{stmaryrd}%mapsfrom 
-\renewcommand{\restriction}{\mathord{\upharpoonright}} %restriction w/p space
+%\renewcommand{\restriction}{\mathord{\upharpoonright}} %restriction w/p space
 \usepackage{mathrsfs} % math script fonts
@@
 We also propose that these specifications can generalize to most structure types by extending previous work on encoding  inherent structure in a consistent manner using mathematical structures from algebraic topology.
-These algebraic structures allow us to encode inherent structure as programmatic types, and the use of category theory facilitates developing a functional design framework that describes function composition in terms of functional algebraic operations.
-The typing system and composition inherent in a functional design encourage library developers to build complex visualization components from simple verifiable parts that have few side effects by design\cite{huHowFunctionalProgramming2015, hughesWhyFunctionalProgramming1989}.
-These categorically specified components could be built as a standalone library, but can also be integrated into existing libraries because they are inherently self-contained.
+These algebraic structures can be translated into analogous programmatic types for encoding structure, while the use of category theory facilitates developing a functional design framework that describes function composition in terms of functional algebraic operations.
+The typing system and composition inherent in a functional design encourage library developers to build complex visualization components from simple verifiable parts that have few side effects \cite{huHowFunctionalProgramming2015, hughesWhyFunctionalProgramming1989}.
+These categorically specified components could be built as a standalone library and integrated into existing libraries because they are inherently self-contained.
 Furthermore we see the application of these ideas in low level visualization library design as our motivation, and we hope these ideas will help shape the reorganization of critical data visualization libraries, such as Matplotlib.
 The contribution of this paper is a mathematical framework for describing structure preserving visualization components in a manner that is generalizable to different types of inherent structure.
-This framework provides also provides guidance for the construction and testing of structure preserving visualization library components.
+This framework also provides guidance for the construction and testing of structure preserving visualization library components.
 
@@
 The typing system proposed in this paper builds on Butler's proposal of using a mathematical structure called fiber bundles as an abstract data representation in visualization \cite{butlerVectorBundleClassesForm1992, butlerVisualizationModelBased1989}.
-We extend Butler's work because he discusses how bundles can be used to encode data in a consistent manner for a number of different \note{arbitrary?} topological and field structures.
+We extend Butler's work because he discusses how bundles can be used to encode data in a consistent manner for the topological and field structures that are common in visualization.
 We sketch out fiber bundles in \autoref{sec:atct:fiber-bundles}, but his work provides a thorough introduction to bundles for visualization practitioners.
 
-\subsection{Data Field Structure}
+\subsection{Field Structure} 
 \label{sec:related-work:equivariance}
-Along with preserving topology, visualization algorithms are expected to preserve the structure of the data fields, i.e. the columns in a table, when they translate data to visual representations. 
-The field structure of data has traditionally been described using the measurment scales-i.e. 
+As with topology, Bertin \cite{bertinSemiologyGraphicsDiagrams2011} codified the expectation that a data value and its visual encoding are expected to have compatible structure.
+The structure on values has traditionally been described  using the measurement scales-i.e.
 nominal, ordinal, interval, ratio - which Steven's classified by their mathematical group structure\cite{stevensTheoryScalesMeasurement1946a} and others have formally expanded to include more types \cite{leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}.
-The expectation that visual transformation is one where the measurement structure be matched by the visual encoding was codified by Bertin \cite{bertinSemiologyGraphicsDiagrams2011}.
-Two mathematical concepts for expressing structure preservation are \textit{equivariance} and \textit{homomorphism}:
-\begin{itemize}
-\item an \textit{equivariant} function preserves symmetric group structure on its input and output.
-Given a group $G$ that acts on input $X$ and output $Y$ and a function $f: X \rightarrow Y$, the function $f$ is \textit{equivariant} if $f(g(x)) = g(f(x))$ for all $g$ in $G$ and for all $x$ in $X$\cite{nlab:equivariant}.
-\item a \textit{homomorphic} function preserves binary operations on input and output of the same algberaic type.
-A map $f: X \rightarrow Y$ is a \textit{homomorphism} when $X$ and $Y$ have the same algebraic structure type and $f(x_1 \oslash x_2) = f(x_1) \oslash f(x_2)$ for every pair $x_1, x_2$ in $X$\cite{grimaldiDiscreteCombinatorialMathematics2006}.
-\end{itemize}
+Generally the structure on each field of the dataset is preserved separately, i.e.
+each column in a table is mapped to a distinct encoding.
+Two mathematical concepts for describing a function that preserves structure are \textit{equivariance} and \textit{homomorphism}.
+An \textit{equivariant} function preserves symmetric group structure on its input and output.
+Given a group $G$ that acts on input $X$ and output $Y$ and a function $f: X \rightarrow Y$, the function $f$ is \textit{equivariant} when $f(g(x)) = g(f(x))$ for all $g$ in $G$ and for all $x$ in $X$\cite{nlab:equivariant}.
+A \textit{homomorphic} function $f$ preserves binary operations on input and output of the same algebraic type.
+A map $f: X \rightarrow Y$ is a \textit{homomorphism} when  $f(x_1 \oslash x_2) = f(x_1) \oslash f(x_2)$ for every pair $x_1, x_2$ in $X$\cite{grimaldiDiscreteCombinatorialMathematics2006}.
 For example, nominal measurements are permutable, meaning that they are acted on by elements of a permutation group.
 For the visual transformation to be equivariant, that nominal measurement must be mapped to a visual encoding that is also permutable, such as distinct marker shapes or color hues.
-Partial orders are non-invertable, so a visual encoding is homomorphic when data that is partially ordered is mapped to visual encodings that are partially ordered in the same way.Mackinlay\cite{mackinlayAutomaticDesignGraphical1987} proposes that data to visual encoders should be \textit{homomorphic} maps in his specification of \textit{A Presentation Tool}.
-The notion that visual transforms are equivariant underlies Kindlemannn and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}'s Algebraic Visualization Design.
-They propose three specific types of structure preservation: that the visualization should not change if the data representation (i.e.
+Partial orders are non-invertable, so a visual encoding is homomorphic when data that is partially ordered is mapped to visual encodings that are partially ordered in the same way.
+
+The notion that visual encodings should be \textit{homomorphic} was proposed by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987} his specification of \textit{A Presentation Tool} and the idea that visual transforms are \textit{equivariant} underlies Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}'s Algebraic Visualization Design.
+Kindlemann and Scheidegger propose three specific types of structure preservation: that the visualization should not change if the data representation (i.e.
 the data container) changes, that data maps unambiguously to visual elements \cite{ziemkiewiczEmbeddingInformationVisualization2009}, and that changing the data should correspond to changes in the visualization in a perceptually significant manner.
@@
 While some visualization tools aim to automate the pairing of data with structure preserving visual representations, such as Tableau\cite{StoltePolaris2002,hanrahanVizQL2006,MackinlayShowme2007}, many visualization libraries leave that choice to the user.
-For example, connectivity assumptions tend to be embedded in each of the visual algorithms of `building block` libraries, a term coined by by Wongsuphasawat \cite{wongsuphasawatNavigatingWideWorld2021} to describe libraries that provide modular components for building elements of a visualization, such as functions for making boxes or translating data values to colors.
-In building block libraries such as Matplotlib\cite{hunterMatplotlib2DGraphics2007} and D3\cite{bostockDataDrivenDocuments2011} the connectivity assumptions lead to very different interfaces; for example in Matplotlib methods for updating the data in a plot differ across the various plotting objects and the parameter list is inconsistent as to which parameters can be applied per data point or for the whole element, and in D3 the API for network visualizers is mostly distinct from the rest of the modules.
+For example, connectivity assumptions tend to be embedded in each of the visual algorithms of `building block` libraries, a term used by Wongsuphasawat \cite{wongsuphasawatNavigatingWideWorld2021} to describe libraries that provide modular components for building elements of a visualization, such as functions for making boxes or translating data values to colors.
+In building block libraries such as Matplotlib\cite{hunterMatplotlib2DGraphics2007} and D3\cite{bostockDataDrivenDocuments2011} the connectivity assumptions lead to very different interfaces; for example in Matplotlib methods for updating data and parameters for controlling aesthetics differ between plots.
 While VTK\cite{hanwellVisualizationToolkitVTK2015,geveciVTK2012} provides a language for expressing the topological properties of the data, and therefore can embed that information in its visual algorithms, it does so in a non-uniform manner.
 On the other hand, domain specific libraries are designed with the assumption of continuities that are common in the domain \cite{HeerSoftware2006}, and therefore can somewhat restrict the interface to choices that that are appropriate for the domain.
-For example, a tabluar topological structure of discret rows, as illustrated in \autoref{fig:related-work:continuity:ktypes}, is assumed by
+For example, a tabular topological structure of discret rows, as illustrated in \autoref{fig:related-work:continuity:ktypes}, is assumed by
 A Presentation Tool\cite{mackinlayAutomatingDesignGraphical1986, mackinlayAutomatingDesignGraphical1986} and grammar of graphics\cite{wilkinsonGrammarGraphics2005} and the ggplot\cite{wickhamGgplot2ElegantGraphics2016}, vega\cite{satyanarayanDeclarativeInteractionDesign2014}, and altair\cite{vanderplasAltairInteractiveStatistical2018} libraries built on these frameworks.
@@
 
-\begin{definition} A \textbf{fiber bundle} $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and continuous surjective map $\pi: \dtotalc \rightarrow \dbasec$ \cite{FiberBundle2020,spanier1989algebraic}.
+\begin{definition} A \textbf{fiber bundle} $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and  bundle projection map $\pi: \dtotalc \rightarrow \dbasec$ \cite{FiberBundle2020,spanier1989algebraic}.
 \begin{equation}
@@
     \dtotalc
-    \arrow[d, "\pi"',color=total] \\
+    \arrow[d, "\pi"',color=total, two heads] \\
      \& 
@@
 \end{equation} 
-A map $\pi$ is a bundle projection map when 
+A continuous surjective map \textbf{$\pi$ is bundle projection} map when 
 \begin{enumerate}
-  \item the \textcolor{fiber}{fiber space} \dfiberc\ is the preimage of the projection function $\pi$ at a point $\dbasepoint$ in the \textcolor{base}{base space} \dbasec\ such that $\dfiber_{\dbasepoint} = \pi^{-1}(\dbasepoint)$
-  \item there is an open neighborhood $\openset_{\dbasepoint}$ surrounding each point in the base space $\dbasepoint \in \openset_{\dbasepoint} \subset \dbase$ such that the \textcolor{total}{total space} \dtotalc\ over the neighborhood is locally trivial $\dtotal\restriction_{\openset} = \openset \times \dfiber$
+  \item the \textcolor{fiber}{fiber space} \dfiberc\ is the preimage of the projection function $\pi$ at a point $\dbasepoint$ in the \textcolor{base}{base space} \dbasec\ such that $\dfiber_{\dbasepoint} = \pi^{-1}(\dbasepoint)$.
+All fibers in a bundle are isomorphic such that $\dfiber \cong \dfiber_{\dbasepoint}$ for all points $\dbasepoint \in \dbase$.
+  \item there is an open neighborhood $\openset_{\dbasepoint}$ surrounding each point in the base space $\dbasepoint \in \openset_{\dbasepoint} \subset \dbase$ such that the \textcolor{total}{total space} \dtotalc\ over the neighborhood, denoted $\dtotal\restriction\openset$\footnote{The symbol $\restriction$ is the restriction operator\cite{RestrictionMathematics2022}.
+For example $\pi^{-1}\restriction\openset$ is the inverse function $\pi^{-1}$ only defined on $\openset$}, is locally trivial.
+The condition for local triviality is that $\dtotal\restriction_{\openset} = \openset \times \dfiber = \pi^{-1}\restriction_{\openset}$
 \end{enumerate}
 \end{definition}
-The first condition for a fiber bundle means that all points in a fiber space over a point in the base space map to the same base space point, which we rely on to establish that there is a formal binding between typed data value, discussed in \autoref{sec:atct:fb:fiber} and the corresponding index, discussed in \autoref{sec:atct:fb:base}.
-The condition of local triviality means that the fibers are identical over open neighborhoods and is what allows fiber bundles to be extremely generalizable, as discussed in \autoref{sec:atct:fb:sections}
+We propose that the map $\pi$ can express the formal binding between having a typed data field, discussed in \autoref{sec:atct:fb:fiber} and a corresponding index, as discussed in \autoref{sec:atct:fb:base}.
+For our purposes, the advantage of fiber bundles is that the field type and topological structure are unconstrained in terms of dimensionality and the only conditions that must be satisfied are that every index point has a corresponding field type and the field types must be the same (isomorphic) for every index point.
 
@@
 \end{definition}
-In \autoref{sec:atct:fb:sections}, we propose that modeling data as sections provides a way to encapsulate topological and field structure in a uniform dimension and type independent manner.
+We propose that modeling data as sections provides a way to encapsulate topological and field structure in a uniform dimension and type independent manner, as discussed in \autoref{sec:atct:fb:sections}.
 
@@
 \begin{definition} 
-   An arbitrary \textbf{category} $\mathcal{C}$ consists of the following \textit{data}: 
+   An \textbf{category} $\mathcal{C}$ consists of the following \textit{data}: 
 \begin{enumerate}
@@
   \label{eq:atct:fb:graphic}
-  \gfiberc \hookrightarrow \gtotalc \xrightarrow{\pi} \gbasec
+  \begin{tikzcd}
+      \gfiberc \arrow[r, hook, color=total] & \gtotalc \arrow[r, "\pi", two heads, color=total] & \gbasec
+  \end{tikzcd}
 \end{equation}

### 2022-08-19 paper/main.tex
rewrote top def

@@
   \item there is an open neighborhood $\openset_{\dbasepoint}$ surrounding each point in the base space $\dbasepoint \in \openset_{\dbasepoint} \subset \dbase$ such that the \textcolor{total}{total space} \dtotalc\ over the neighborhood, denoted $\dtotal\restriction\openset$\footnote{The symbol $\restriction$ is the restriction operator\cite{RestrictionMathematics2022}.
-For example $\pi^{-1}\restriction\openset$ is the inverse function $\pi^{-1}$ only defined on $\openset$}, is locally trivial.
+For example $\pi^{-1}\restriction\openset$ is the inverse function $\pi^{-1}$ only defined on the points in $\openset$}, is locally trivial.
 The condition for local triviality is that $\dtotal\restriction_{\openset} = \openset \times \dfiber = \pi^{-1}\restriction_{\openset}$
@@
 \end{definition}
-We propose that the map $\pi$ can express the formal binding between having a typed data field, discussed in \autoref{sec:atct:fb:fiber} and a corresponding index, as discussed in \autoref{sec:atct:fb:base}.
+We propose that the map $\pi$ can express the formal binding between having a typed data field, discussed in \autoref{sec:atct:fb:fiber}, and a corresponding point in the topological structure, as discussed in \autoref{sec:atct:fb:base}.
 For our purposes, the advantage of fiber bundles is that the field type and topological structure are unconstrained in terms of dimensionality and the only conditions that must be satisfied are that every index point has a corresponding field type and the field types must be the same (isomorphic) for every index point.
@@
 \label{sec:atct:fb:base}
-We encode the underlying connectivity of the data as the \textcolor{base}{base space} \dbase\ of the fiber bundle because it provides a uniform language for describing the topology regardless of how the data is subset.
+We encode the topological structure of the data as the \textcolor{base}{base space} \dbase\ of the fiber bundle since the base space acts as indexing space where every point in the base space has a corresponding fiber space $\pi^{-1}(\dbasepoint) = \dfiber_{\dbasepoint}$.
 The base space \dbase\ is a topological space $(\dbase,\mathcal{T}_{\dbase})$, which means it  consists of the set of points ${\dbasepoint \in \dbase}$ and topology $\mathcal{T}_{\dbase}$ \cite{munkresElementsAlgebraicTopology1984}.
-The set \dbase\ is equipped with open subsets \openset, which are a generalization of an open interval.
-This means opensets contain points inside a bounded region on the topological space and excludes the points on the boundary \cite{TopologyVsTopology}.
-An openset that surrounds a point \dbasepoint\ is called an open neighborhood $\openset_{\dbasepoint}$.
-Opensets are key to constructing topologies, and are used in this paper to model subsets of the data contextualized within the larger continuity those subsets belong to.
+A topology $\dbase \coloneqq (\dbase,\mathcal{T}_{\dbase})$ is an axiomatic way of defining a topological structure\cite{bradleyTopologyCategoricalApproach2020}  as a collection of subsets, called open sets \footnote{Open sets (open subsets) are a generalization of open intervals to n dimensional spaces.
+For example, an open ball is the set of points inside the ball and excludes points on the surface of the ball.
+\cite{weissteinOpenSet}}, of that structure.
 
 \begin{definition}
-A \textbf{topology} $\mathcal{T}_{\dbase}$ on a space \dbase\ is a collection of open sets \openset\ of \dbase\ that satisfy the properties 
+A \textbf{topology} $\mathcal{T}$ on a space \dbase\ is a collection of open sets \openset\ of \dbase\ that satisfy the properties 
   \begin{itemize}
-    \item the empty set $\varnothing$ and $\dbase$ are in $\mathcal{T}_{\dbase}$
-    \item for all $\openset_i, \openset_j \in \mathcal{T}_{\dbase}$, the intersection $\openset_i \cap \openset_j$  is in $ \mathcal{T}_{\dbase}$
-    \item for all $\openset_i, \openset_j \in \mathcal{T}_{\dbase}$, the union $\openset_i \cup \openset_j$ is in $ \mathcal{T}_{\dbase}$
+    \item the empty set $\varnothing$ and $\dbase$ are in $\mathcal{T}$
+    \item the union of open sets in $\mathcal{T}$ is in also in $\mathcal{T}$
+    \item any finite intersection of open sets in $\mathcal{T}$ is also in $\mathscr{T}$
   \end{itemize}
-  such that the the topology provides an axiomatic way to construct the topological space $\dbase \coloneqq (\dbase,\mathcal{T}_{\dbase})$\cite{bradleyTopologyCategoricalApproach2020}.
 \end{definition}
- A property of the base space \dbase\ of a fiber bundle is that it is a quotient topology, which  means that it contains the largest possible collection of distinct opensets, meaning each openset is as small as possible, of $\dbase$ that still preserve the underlying continuity of \dbase \cite{QuotientSpaceTopology2020}.
-For our purposes, $\dbase$ is a set of points that encodes the underlying connectivity of the our dataset, for example, $\dbase$ could be the discrete points, interval, or plane in \autoref{fig:related-work:continuity:ktypes}.
-We express the allowable ways to subset and compose points such that the connectivity is preserved as the topology $\mathcal{T}_{\dbase}$ on the data continuity.
+The base space of a fiber bundle is a quotient topology\cite{QuotientSpaceTopology2020}, meaning that it divides the topological space into the smallest open sets such that $\pi$ remains a continuous function.
+For our purposes, $\dbase$ encodes the underlying connectivity of the our dataset, for example, $\dbase$ could be the discrete points, interval, or plane in \autoref{fig:related-work:continuity:ktypes}.
+We define $\dbase$ as a topology $\mathcal{T}_{\dbase}$ because a topology codifies that the underlying topological structure of the data can be subsetted and how these subsets can be composed in a structure preserving manner.
 
@@
 In a non-trivial bundle, fibers are isomorphic $\dfiber \simeq \dfiber_{\dbasepoint}\;\forall \dbasepoint \in \dbase$, which means there is no single fiber type $\dfiber$.
-Instead, we can define sections locally $\dsection\restriction_{\openset} \in \Gamma(\openset_{\dbasepoint}, \dtotal\restriction_{\openset_{\dbasepoint}})$, which means a section defined only on an open neighborhood $\dbasepoint \in \openset \in \dbase$.
+Instead, we can define sections locally $\dsection\restriction_{\openset} \in \Gamma(\openset_{\dbasepoint}, \dtotal\restriction_{\openset_{\dbasepoint}})$, which means a section defined only on an open neighborhood\footnote{An open set that surrounds a point \dbasepoint\ is called an open neighborhood $\openset_{\dbasepoint}$.} $\dbasepoint \in \openset \in \dbase$.
 This is because all fiber bundles are locally trivializable, $\dtotal\restriction_{\openset_{k}} = \openset_{k}\times \dfiber$ such that there's always a subset where the fibers are equal.

### 2022-08-25 paper/main.tex
fiber section

@@
 We encode the topological structure of the data as the \textcolor{base}{base space} \dbase\ of the fiber bundle since the base space acts as indexing space where every point in the base space has a corresponding fiber space $\pi^{-1}(\dbasepoint) = \dfiber_{\dbasepoint}$.
-The base space \dbase\ is a topological space $(\dbase,\mathcal{T}_{\dbase})$, which means it  consists of the set of points ${\dbasepoint \in \dbase}$ and topology $\mathcal{T}_{\dbase}$ \cite{munkresElementsAlgebraicTopology1984}.
+The \textcolor{base}{base space} \dbasec\ is a topological space $(\dbase,\mathcal{T}_{\dbase})$, which means it  consists of the set of points ${\dbasepoint \in \dbase}$ and topology $\mathcal{T}_{\dbase}$ \cite{munkresElementsAlgebraicTopology1984}.
 A topology $\dbase \coloneqq (\dbase,\mathcal{T}_{\dbase})$ is an axiomatic way of defining a topological structure\cite{bradleyTopologyCategoricalApproach2020}  as a collection of subsets, called open sets \footnote{Open sets (open subsets) are a generalization of open intervals to n dimensional spaces.
@@
 \begin{enumerate}
-  \item \textit{object}: X
-  \item \textit{identity morphism}: $X \xrightarrow {id_x} X$
-  \item \textit{morphisms}: $X \xrightarrow{f} Y$ between objects X, Y in $\mathcal{C}$
-  \item \textit{composite morphism}: given the morphisms $X \xrightarrow{f} Y$ and $Y \xrightarrow{g} Z$, there exists a morphism $X \xrightarrow{g\circ f} Z$
+  \item a collection of \textit{objects} $X \in \mathcal{Ob}(\mathcal{C})$
+  \item for every pair of objects $X, Y \in \mathcal{Ob}(\mathcal{C})$, a set of \textit{morphisms} $X \xrightarrow{f} Y \in Hom_{\mathcal{C}}(X, Y)$
+  \item for every object $X$, a distinct \textit{identity morphism} $X \xrightarrow {id_x} X$ in $Hom_{\mathcal{C}}(X, X)$
+  \item a \textit{composition function} $f \in Hom_{\mathcal{C}}(X, Y) \times  g \in Hom_{\mathcal{C}}(Y, Z) \rightarrow g \circ f \in Hom_{\mathcal{C}}(X, Z)$
 \end{enumerate}
-The data must satisfy the following \textit{rules}
+such that 
 \begin{enumerate}
-  \item \textit{unitality:} for every morphism $ X \xrightarrow{f} Y$, there exists identity morphisms $id_x, id_y$ such that $f \circ id_x = f = id_y \circ f$
+  \item \textit{unitality:} for every morphism $ X \xrightarrow{f} Y$, $f \circ id_x = f = id_y \circ f$
   \item \textit{associativity:} if any three morphisms $f, g, h$ are composable, 
@@
 
-We construct a base space category $\mathcal{\dbasec}$ to encapsulate the topological structure of the data.
-The category $\mathcal{\dbase}$ has open set objects $\opensetc$ and inclusion morphisms $\opensetc_i \xrightarrow{\iota} \opensetc_j$ such that $\opensetc_i \subseteq \opensetc_j$.
+We encapsulate the topological structure of the data as a category $\mathcal{\dbasec}$.
+The standard construction of a category from a topological space is that it has open set objects $\opensetc$ and inclusion morphisms $\opensetc_i \xrightarrow{\iota} \opensetc_j$ such that $\opensetc_i \subseteq \opensetc_j$\cite{riehlCategoryTheoryContext}.
 The composability property expresses that inclusion is transitive, while associativity expresses that the inclusion functions can be curried in various equivalent groupings.
@@
 \label{sec:atct:fb:fiber}
-As mentioned in \autoref{sec:related-work:equivariance}, equivariance is traditionally evaluated as the preservation of field structure, which is based on the field type.
-To have a rich language for expressing field type, we generalize Spivak's fiber bundle formulation of table \cite{spivakDatabasesAreCategories2010,spivakSimplicialDatabases2009}.
-\note{def 3.5 and 3.6 should maybe be pushed back to prose as 3.7 is the only one we rely on - it defines what tau returns}
-\begin{definition} A \textbf{type specification} is the bundle map $\pi: \mathscr{U} \rightarrow \textbf{DT}$.
-The base space $\textbf{DT}$ is the set of data types, and the total space $\mathscr{U}$ is the set of domains of those data types.
-Given an arbitrary data type $T \in \textbf{DT}$, the preimage $\pi^{-1} \subset \mathscr{U}$ is the domain of $T$ and $x \in \pi^{-1}(T)$ is an object of type $T$.
-\end{definition}
+As mentioned in \autoref{sec:related-work:equivariance}, visualization researchers traditionally describe equivariance as the preservation of field structure, which is based on the field type.
+Spivak shows that data typing can be expressed in a categorical framework in his fiber bundle formulation of tables in relational databases \cite{spivakDatabasesAreCategories2010,spivakSimplicialDatabases2009}.
+In this work, we adopt Spivak's definitions of \textit{type specification}, \textit{schema}, \textit{record} because that allows us to use a dimension agnostic named typing system for the fields of our dataset that is consistent with the abstraction we are using to express the continuity.
+Spivak introduces a \textit{type specification} as a bundle map $\pi: \mathscr{U} \rightarrow \textbf{DT}$.
+The base space $\textbf{DT}$ is the set of data types, and the total space $\mathscr{U}$ is the disjoint union of the domains of those data types.
+Given an arbitrary data type $T \in \textbf{DT}$, the preimage $ \pi^{-1}(T) \subset \mathscr{U}$ is the domain of $T$ and $x \in \pi^{-1}(T)$ is an object of type $T$\cite{spivakSimplicialDatabases2009}.
 For example, if $T=\texttt{int}$, then the image $\pi^{-1}(\texttt{int}) = \mathbb{Z} \subset \mathscr{U}$ is the set of all integers.
-The domain $\mathscr{U}_{{\sigma}}$ is also a topological space, meaning there is no restriction on the dimensionality of the field, thus allowing for the encoding of multi-dimensional fields in the same manner as single dimensional fields.
-Since many fields can have the same datatype, Spivak formally defines a mapping from field name to field data type, akin to a  data base schema \cite{ullmanFirstCourseDatabase2008}.
-\begin{definition} A \textbf{schema} consists of a pair $(\fnames, \sigma)$ where $\fnames$ is the set of field names and $\sigma: \fnames \rightarrow \textbf{DT}$ is a function from field name to field data type.
-\end{definition}
-The function $\sigma$ is composed with $\pi$ such that the bundle  $\pi_{\sigma}:\mathscr{U}_{\sigma} \rightarrow \fnames$ associates a field name $c \in C$ with the corresponding data type domain.
-Spivak then defines data records as sections on the bundle defined by $\pi_{\sigma}$.
+Since many fields can have the same datatype, Spivak formally defines a mapping from field name to field data type, akin to a database schema \cite{ullmanFirstCourseDatabase2008}.
+According to Spivak, a \textit{schema} consists of a pair $(\fnames, \sigma)$ where $\fnames$ is the set of field names and $\sigma: \fnames \rightarrow \textbf{DT}$ is a function from field name to field data type\cite{spivakSIMPLICIALDATABASESa}.
+The function $\sigma$ is composed with $\pi$ such that the bundle  $\pi_{\sigma}:\mathscr{U}_{\sigma} \rightarrow \fnames$ associates a field name $c \in C$ with the corresponding domain for that data type $\mathscr{\U}_{\sigma}$.
 \begin{definition} A \textbf{record} is a function $\delement: \fnames \rightarrow \mathscr{U}_{\sigma}$ and the set of records on $\pi_{\sigma}$ is denoted $\Gamma^{\pi}(\sigma)$.
@@
 \end{definition}
-Spivak then constructs tables as sections $\dsection: \dbase \rightarrow \Gamma^{\pi}(\sigma)$ from a key space $\dbase$ to the set of records $\Gamma^{\pi}(\sigma)$ on the schema bundle.
-Generalizing Spivak's formulation of tables, we define the fiber space in this work as the set of record functions
+Spivak then describes tables as sections $\dsection: \dbase \rightarrow \Gamma^{\pi}(\sigma)$ from a key space $\dbase$ to the set of all possible records $\Gamma^{\pi}(\sigma)$ on the schema bundle, and his notion of a table generalizes to our notion of a data container.
+
+To build on the rich typing system provided by Spivak, we define the \textcolor{fiber}{fiber space} \dfiberc\ to be the set of all data records
 \begin{equation}
@@
 such that the preimage of a point is the corresponding data type domain $\pi^{-1}(\dbasepoint) = \dfiber_{k} = \mathscr{U}_{{\sigma}_{\dbasepoint}}$. 
-The construction of the fiber as the space of records provides a way to distinguish data fields with the same datatype while allowing that datatype to be defined as almost any arbitrary topological space.
+Adopting Spivak's fiber bundle construction of types allows our model to reuse types so long as the field names are distinct and that field values can be accesed by field name,  since those are sections on $\mathcal{U}_{\sigma}$.
+Furthermore, since domains $\mathscr{U}_{{\sigma}}$ of types are a topological space, multi-dimensional fields can be encoded in the same manner as single dimensional fields and fields can have different names but the same type.
 
 As with the base space category $\mathcal{\dbase}$, we propose a fiber category $\mathcal{\dfiberc}$ to encapsulate the field types of the data.
-The fiber category $\mathcal{\dfiberc}$ is a monoidal category, meaning the category has a single object $\dfiberc$ of an arbitrary type.
-The morphisms on the fiber object are $\dfunctc \in Hom(\dfiberc, \dfiberc)$.
-The fiber category is equipped with a bifunctor $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$.
-The bifunctor provides a method for combining fibers, thereby allowing for fields that contain multityped values.
+The fiber category has a single object $\dfiberc$ of an arbitrary type and morphisms on the fiber object $\dfunctc \in Hom(\dfiberc, \dfiberc)$.
+The fiber category $\mathcal{\dfiberc}$ is a monoidal category, meaning that it is a category equipped with a bifunctor $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$.
+The bifunctor provides a method for combining fibers, thereby allowing us to express fields that contain multityped values.
 For example, wind can be represented as two fields $\dfiber_{speed} \times \dfiber_{direction}$ or a composite fiber field $\dfiber_{speed} \otimes \dfiber_{direction} = \dfiber_{wind}$.

### 2022-09-01 paper/main.tex
spivak reworking of typing bundles

@@
 These algebraic structures can be translated into analogous programmatic types for encoding structure, while the use of category theory facilitates developing a functional design framework that describes function composition in terms of functional algebraic operations.
-The typing system and composition inherent in a functional design encourage library developers to build complex visualization components from simple verifiable parts that have few side effects \cite{huHowFunctionalProgramming2015, hughesWhyFunctionalProgramming1989}.
+The typing system and composition inherent in a functional design encourage library developers to build complex visualization components from simple verifiable parts that have few side effects\cite{huHowFunctionalProgramming2015, hughesWhyFunctionalProgramming1989}.
 These categorically specified components could be built as a standalone library and integrated into existing libraries because they are inherently self-contained.
@@
 
-\begin{definition} A \textbf{fiber bundle} $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and  bundle projection map $\pi: \dtotalc \rightarrow \dbasec$ \cite{FiberBundle2020,spanier1989algebraic}.
+\begin{definition}
+   A \textbf{fiber bundle} $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and  bundle projection map $\pi: \dtotalc \rightarrow \dbasec$ \cite{FiberBundle2020,spanier1989algebraic}.
 \begin{equation}
@@
 \end{definition}
-
 We encapsulate the topological structure of the data as a category $\mathcal{\dbasec}$.
@@
 Spivak shows that data typing can be expressed in a categorical framework in his fiber bundle formulation of tables in relational databases \cite{spivakDatabasesAreCategories2010,spivakSimplicialDatabases2009}.
-In this work, we adopt Spivak's definitions of \textit{type specification}, \textit{schema}, \textit{record} because that allows us to use a dimension agnostic named typing system for the fields of our dataset that is consistent with the abstraction we are using to express the continuity.
+In this work, we adopt Spivak's definitions of \textit{type specification}, \textit{schema}, and \textit{record} because that allows us to use a dimension agnostic named typing system for the fields of our dataset that is consistent with the abstraction we are using to express the continuity.
 Spivak introduces a \textit{type specification} as a bundle map $\pi: \mathscr{U} \rightarrow \textbf{DT}$.
-The base space $\textbf{DT}$ is the set of data types, and the total space $\mathscr{U}$ is the disjoint union of the domains of those data types.
-Given an arbitrary data type $T \in \textbf{DT}$, the preimage $ \pi^{-1}(T) \subset \mathscr{U}$ is the domain of $T$ and $x \in \pi^{-1}(T)$ is an object of type $T$\cite{spivakSimplicialDatabases2009}.
-For example, if $T=\texttt{int}$, then the image $\pi^{-1}(\texttt{int}) = \mathbb{Z} \subset \mathscr{U}$ is the set of all integers.
+The base space $\textbf{DT}$ is a set of data types $T \in \textbf{DT}$ and the total space $\mathscr{U}$ is the disjoint union of the domains of each type \[\mathscr{U} = \bigsqcup_{T \in \textbf{DT}} \pi^{-1}(T)\] such that each element $x$ in the domain $\pi^{-1}(T)$ is one possible value of an object of type $T$ \cite{spivakSimplicialDatabases2009}.
+For example, if $T=\texttt{int}$, then the image $\pi^{-1}(\texttt{int}) = \mathbb{Z} \subset \mathscr{U}$ is the set of all integers and $x=3 \in \mathbb{Z}$ is the value of one $\texttt{int}$ object.
+
 Since many fields can have the same datatype, Spivak formally defines a mapping from field name to field data type, akin to a database schema \cite{ullmanFirstCourseDatabase2008}.
-According to Spivak, a \textit{schema} consists of a pair $(\fnames, \sigma)$ where $\fnames$ is the set of field names and $\sigma: \fnames \rightarrow \textbf{DT}$ is a function from field name to field data type\cite{spivakSIMPLICIALDATABASESa}.
-The function $\sigma$ is composed with $\pi$ such that the bundle  $\pi_{\sigma}:\mathscr{U}_{\sigma} \rightarrow \fnames$ associates a field name $c \in C$ with the corresponding domain for that data type $\mathscr{\U}_{\sigma}$.
+According to Spivak, a \textit{schema} consists of a pair $(\fnames, \sigma)$ where $\fnames$ is the set of field names and $\sigma: \fnames \rightarrow \textbf{DT}$ is a function from field name to field data type\cite{spivakSimplicialDatabases2009}.
+The function $\sigma$ is composed with $\pi$ such that $\pi^{-1}(\sigma(C)) \subseteq \mathscr{U}$; this composition induces a domain bundle $\pi_{\sigma}:\mathscr{U}_{\sigma} \rightarrow \fnames$ that associates a field name $c \in C$ with its corresponding domain $\pi^{-1}_{\sigma}(C) \subseteq \mathscr{U}_{\sigma}$.
 \begin{definition} A \textbf{record} is a function $\delement: \fnames \rightarrow \mathscr{U}_{\sigma}$ and the set of records on $\pi_{\sigma}$ is denoted $\Gamma^{\pi}(\sigma)$.
@@
 Defining the field schema as a monoidal category allows us to formally express the structure of the field, as defined by its sets and functions, at different levels of composition as needed by different visualization tasks.
-
+ \
 \subsubsection{Data: Section} 
@@
 We encode data as a \textcolor{section}{section} \dsectionc\ of a bundle because this allows us to incorporate the topology and field types in the data definition.
-We can define these section functions locally, meaning over a specific open subset \openset\ of \dbase\
+We can define these section functions locally, meaning that the section is (piece-wise) continuous over a specific open subset \openset\ of \dbase\
 \begin{equation}
@@
 \end{equation}
-such that each section function $\dsection: \dbasepoint \mapsto \delement$ maps from a point $\dbasepoint \in \openset$ to a record in the fiber space $\delement \in \dfiber_{\dbasepoint}$ over that point.
+such that each section function $\dsection: \dbasepoint \mapsto \delement$ maps from each point $\dbasepoint \in \openset$ to a corresponding record in the fiber space $\delement \in \dfiber_{\dbasepoint}$ over that point.
 Bundles can have multiple sections, as denoted by $\Gamma(\openset, \dtotal\restriction{\openset})$.
@@
 
-In a trivial bundle all fibers are equal $\dfiberc \equiv \dfiberc_{\dbasepointc}\;\forall \dbasepointc \in \dbase$, which means that $\dtotalc = \dbasec \times \dfiberc$ .
+In a trivial bundle, the fibers are equal $\dfiberc =  \dfiberc_{\dbasepointc}$ over all points $\dbasepointc \in \dbase$  in the base space such that the total space is the product of the fiber and base space $\dtotalc = \dbasec \times \dfiberc$.
 This allows us to define global sections $\dsection: \dbase \rightarrow \dfiber \in \Gamma(\dbase, \dfiber)$ which we translate into a data signature of the form
@@
 When the total space is non-trivial, we can use the fiber bundle property of local-triviality to generate type signatures.
-In a non-trivial bundle, fibers are isomorphic $\dfiber \simeq \dfiber_{\dbasepoint}\;\forall \dbasepoint \in \dbase$, which means there is no single fiber type $\dfiber$.
+In a non-trivial bundle fibers are isomorphic $\dfiber \simeq \dfiber_{\dbasepoint}$ for all points $\dbasepoint \in \dbase$, which means that there may be a \textit{twist} in the bundle that prevents defining global sections \cite{hatcherAlgebraicTopology2002}.
+As mentioned in \autoref{def}
+
+Fiber bundles are locally trivializable, $\dtotal\restriction_{\openset_{k}} = \openset_{k}\times \dfiber$ such that there's always a subset where the fibers are equal.
+
 Instead, we can define sections locally $\dsection\restriction_{\openset} \in \Gamma(\openset_{\dbasepoint}, \dtotal\restriction_{\openset_{\dbasepoint}})$, which means a section defined only on an open neighborhood\footnote{An open set that surrounds a point \dbasepoint\ is called an open neighborhood $\openset_{\dbasepoint}$.} $\dbasepoint \in \openset \in \dbase$.
-This is because all fiber bundles are locally trivializable, $\dtotal\restriction_{\openset_{k}} = \openset_{k}\times \dfiber$ such that there's always a subset where the fibers are equal.
  This means almost all data can be expressed as a collection of local sections $\{\dsection\restriction_{\openset_{\dbasepoint}}| \dbasepoint \in \dbase \}$ which translate to a set of signatures 
@@
   \caption{The mobius band is a non-trivial total space because the twist means that the fibers are sometimes [-1, 1] and other times [1,-1].
-The mobius band can be represented as two trivial bundles, one with fiber [-1,1] and one with fiber [1,-1] that can be glued on the edges into the band.
+The mobius band can be represented as two trivial bundles, one with fiber [-1, 1] and one with fiber [1,-1] that can be glued on the edges into the band.
 The gluing rules ensure that the arrows on the edges go in the same direction.\label{fig:atct:nontrivialbundle}}
@@
 \bibliography{bibliography}
-
 % biography section 

### 2022-09-01 paper/main.tex
moved intro of trivial around, incorporated transitions into it

@@
 \end{definition}
-We propose that the map $\pi$ can express the formal binding between having a typed data field, discussed in \autoref{sec:atct:fb:fiber}, and a corresponding point in the topological structure, as discussed in \autoref{sec:atct:fb:base}.
-For our purposes, the advantage of fiber bundles is that the field type and topological structure are unconstrained in terms of dimensionality and the only conditions that must be satisfied are that every index point has a corresponding field type and the field types must be the same (isomorphic) for every index point.
-
 \begin{definition} A \textcolor{section}{\textbf{section}} $\dsectionc: \dbasec \rightarrow \dtotalc$ over a fiber bundle is a smooth right inverse of $\pi$ such that $\pi(\dsection(\dbasepoint)) = \dbasepoint$ for all $\dbasepoint \in \dbase$
 \end{definition}
+
+Local triviality $\dtotal\restriction_{\openset_{k}} = \openset_{k}\times \dfiber$ means that there is always a subset of the base space over which the fibers are identical.
+A bundle can be subsetted into regions such that the bundled can be reconstructed by gluing the regions back together via transition maps on the border of each region.
+These borders are fibers, and in a trivial bundle transition maps are identity maps because fibers are identical $\dfiber = \dfiber_{k}$ for all points $\dbasepoint \in dbase$.
+In a non-trivial bundle fibers are isomorphic $\dfiber \simeq \dfiber_{\dbasepoint}$ for all points $\dbasepoint \in \dbase$, which means that there is at least one transition map that is not an identity \cite{hatcherAlgebraicTopology2002}.
+
+We propose that the map $\pi$ can express the formal binding between having a typed data field, discussed in \autoref{sec:atct:fb:fiber}, and a corresponding point in the topological structure, as discussed in \autoref{sec:atct:fb:base}.
+For our purposes, the advantage of fiber bundles is that the field type and topological structure are unconstrained in terms of dimensionality and the only conditions that must be satisfied are that every index point has a corresponding field type and the field types must be the same for every index point.
 We propose that modeling data as sections provides a way to encapsulate topological and field structure in a uniform dimension and type independent manner, as discussed in \autoref{sec:atct:fb:sections}.
@@
 
-In a trivial bundle, the fibers are equal $\dfiberc =  \dfiberc_{\dbasepointc}$ over all points $\dbasepointc \in \dbase$  in the base space such that the total space is the product of the fiber and base space $\dtotalc = \dbasec \times \dfiberc$.
+In a trivial bundle, the total space is the product of the fiber and base space $\dtotalc = \dbasec \times \dfiberc$.
 This allows us to define global sections $\dsection: \dbase \rightarrow \dfiber \in \Gamma(\dbase, \dfiber)$ which we translate into a data signature of the form
@@
 
-When the total space is non-trivial, we can use the fiber bundle property of local-triviality to generate type signatures.
-In a non-trivial bundle fibers are isomorphic $\dfiber \simeq \dfiber_{\dbasepoint}$ for all points $\dbasepoint \in \dbase$, which means that there may be a \textit{twist} in the bundle that prevents defining global sections \cite{hatcherAlgebraicTopology2002}.
-As mentioned in \autoref{def}
-
-Fiber bundles are locally trivializable, $\dtotal\restriction_{\openset_{k}} = \openset_{k}\times \dfiber$ such that there's always a subset where the fibers are equal.
-
-Instead, we can define sections locally $\dsection\restriction_{\openset} \in \Gamma(\openset_{\dbasepoint}, \dtotal\restriction_{\openset_{\dbasepoint}})$, which means a section defined only on an open neighborhood\footnote{An open set that surrounds a point \dbasepoint\ is called an open neighborhood $\openset_{\dbasepoint}$.} $\dbasepoint \in \openset \in \dbase$.
- This means almost all data can be expressed as a collection of local sections $\{\dsection\restriction_{\openset_{\dbasepoint}}| \dbasepoint \in \dbase \}$ which translate to a set of signatures 
+When the total space is non-trivial, we can use the fiber bundle property of local-triviality to define local sections $\dsection\restriction_{\openset} \in \Gamma(\openset_{\dbasepoint}, \dtotal\restriction_{\openset_{\dbasepoint}})$.
+A local section is defined over an open neighborhood  $\dbasepoint \in \openset \in \dbase$, which is an open set that surrounds a point \dbasepoint.
+Most data sets can be encoded as a collection of local sections $\{\dsection\restriction_{\openset_{\dbasepoint}}| \dbasepoint \in \dbase \}$ and this encoding can be translated into a set of signatures
 \begin{equation}
@@
 \{\textcolor{section}{\texttt{data-subset}}: \textcolor{base}{\texttt{topology}} \rightarrow \textcolor{fiber}{\texttt{fields}} \\
- s.\;t.\;\textcolor{section}{\texttt{data-subset}} \in \texttt{dataset}\}  
+ s.\;t.\;\textcolor{section}{\texttt{data-subset}} \subset \texttt{dataset}\}  
   \end{split}
 \end{equation}
-that can be glued together on overlaps to generate the dataset. This gluing can be implemented as part of the data container, which we discuss in \autoref{sec:atct:sheaves}.
+The subsets of the fiber bundle and the transition maps between these subsets are encoded in an atlas\cite{ghristElementaryAppliedTopology2014} and the notion of an atlas can be incorporated into the data container, as discussed in \autoref{sec:atct:sheaves}.
 
@@
 
-The mobius band is an example of the non-trivial bundle because the twist in the middle means that the fibers are not the same, as shown in \autoref{fig:atct:nontrivialbundle}.
+The mobius band is an example of a non-trivial bundle because the twist in the middle, as shown in \autoref{fig:atct:nontrivialbundle}, means that there is at least one local .
+
 But, as with all bundles, the mobius band can be decomposed into trivial bundle components, here the two sides of the mobius band.

### 2022-09-02 paper/main.tex
section 3.1

@@
 \begin{figure}[h!]
-  \includegraphics[width=\columnwidth]{tbundle.png}
-  \caption{The mobius band is a non-trivial total space because the twist means that the fibers are sometimes [-1, 1] and other times [1,-1].
-The mobius band can be represented as two trivial bundles, one with fiber [-1, 1] and one with fiber [1,-1] that can be glued on the edges into the band.
-The gluing rules ensure that the arrows on the edges go in the same direction.\label{fig:atct:nontrivialbundle}}
+  \includegraphics[width=\columnwidth]{trivnon.png}
+  \caption{One local trivialization of both a trivial cylinder bundle and a non-trivial mobius bundle is a plane over the interval $\left[0,2\pi\right]$.
+The transition map that transforms this trivialization into a cylinder is an identity map aligning two \textit{up} fibers, while the transition map that transforms this trivialization into a mobius bundle aligns an \textit{up} fiber with a \textit{down} fiber.
+  \note{not sure about wording here/ maybe should still have two transition maps so non-trivial has 1 id and 1 not-id}
+  \label{fig:atct:nontrivialbundle}}
 \end{figure}
 
-The mobius band is an example of a non-trivial bundle because the twist in the middle, as shown in \autoref{fig:atct:nontrivialbundle}, means that there is at least one local .
-
-But, as with all bundles, the mobius band can be decomposed into trivial bundle components, here the two sides of the mobius band.
-Each side has fibers going in the same direction, with the mismatch only apparent on the edges.
-These trivial bundles can be glued into the mobius band so long as the fibers (arrows on the edges) go in the same direction when glued.
-The data is encoded as the section of the bundle, here the sin curve.
+As shown in \autoref{fig:atct:nontrivialbundle}, a cylinder bundle and mobius band bundle over the same base space (here a unit circle) can be locally trivialized in the same manner.
+The local trivialization in \autoref{fig:atct:nontrivialbundle} is a plane over the interval $\left[0,2\pi\right]$.
+In the cylinder bundle, all fibers go in the same direction, \textit{up}, such that transition maps on local trivializations are all identity maps aliging two \textit{up} fibers.
+This property of transition maps is what charaterizes it as a trivial bundle, while local trivializations of  non-trivial bundles always have at least on non-identity transition map.
+In \autoref{fig:atct:nontrivialbundle} this is the sole transtion map and it aligns an \textit{up} fiber with a \textit{down} fiber, thereby creating the twist in the mobius band.
+The data is encoded as the section of the bundle, here exemplified as the sin curve.
+In the cylinder bundle, the sin curve is continuous on the whole space, while the sin curve is discontinuous at the twist in the mobius band and can be made continuous via the transition map.
 
@@
   \includegraphics[width=\columnwidth]{dbundle.png}
-  \caption{The table from \autoref{fig:related-work:continuity:ktypes} has a 3 dimensional fiber (name, temperature, precipitation), a 0D base space, and the row is the section.
-The time series has a 2D fiber (time, precipitation) and a 1D base space since the data is sampled along a continuous space (physical time).
-Each section returns a data point along the line.
-The map is sparsely sampled points along a 2D plane encoding spatial continuity, and the fiber is a 3D space encoding (latitude, longitude, precipitation)
+  \caption{The table from \autoref{fig:related-work:continuity:ktypes} has a 3 dimensional fiber (name, temperature, precipitation), a 0D base space, and each row is a section.
+Each time series is a section of a bundle with a 2D fiber (time, precipitation) and a 1D base spac encoding temporal continuity.
+Each location of the rain map is a section of a bundle with a 2D plane base encoding spatial continuity and a 3D fiber space encoding (latitude, longitude, precipitation)
     \label{fig:atct:trivialbundle}}
@@
 In \autoref{fig:atct:trivialbundle}, the base space $\dbase$ acts as an indexing space into the fiber space $\dfiber$.
-In the bundle encoding of the table, the indexing space is random keys; in the time series bundle, the base space is the interval [0,1]; and the map is sparse samples from a continuous space $[0,1]^{2}$ In contrast to a notion of a semantic binding between indexing space and field values, as proposed by Munzner\cite{munznerVisualizationAnalysisDesign2014}, the fields describing the continuity are part of the fiber.
+In the bundle encoding of the table, the indexing space is arbitrarily numbered keys; in the time series bundle, the base space is the interval [0,1]; and the map is sparse samples from a continuous space $[0,1]^{2}$ In contrast to a notion of a semantic binding between indexing space and field values, as proposed by Munzner\cite{munznerVisualizationAnalysisDesign2014}, the fields describing the continuity are part of the fiber.
 For example, the time is a fiber in the timeseries and the latitude and longitude are part of the map's fiber.
@@
 The base space \gbasec\ is a parameterization of the display area, for example the inked bounding box in cairo \cite{CairographicsOrg}.
-The fiber space \gfiberc\ is an abstraction of the renderer fields, for example $\{x,\,y,\,r,\,g,\,b,\,a\}$.
+The fiber space \gfiberc\ is an abstraction of the renderer fields; for example a 2 dimension screen has pixels that can be parameterized $ \gfiber=\{x,\,y\,z\,r,\,g,\,b,\,a\}$.
 
@@
 \end{equation}
-that map from a point in an openset in the graphic space $\gbasepoint \in \opensetg \in \gbase$ to a point in the graphic fiber $\gfiber$.
+that map from a point in an openset in the graphic space $\gbasepoint \in \opensetg \subseteq \gbase$ to a point in the graphic fiber $\gfiber$.
 The section evaluated on a single point $\gbasepoint$ returns a single graphic record, for example one pixel in an ideal resolution space.
@@
   \includegraphics[width=1\columnwidth]{fb_rho.png}
-  \caption{For a 2D display a section $\gsection$ maps from each point $\gbasepoint$ into a fiber $\gfiber_{\gbasepoint}$ that encodes an RGB prerender space with infinite resolution $\mathcal{R}^{2}$.
+  \caption{For a 2D display, a section $\gsection$ maps from each point $\gbasepoint$ into a fiber $\gfiber$ that encodes an RGB prerender space with infinite resolution $\mathcal{R}^{2}$.
 Each tiny colored box is an approximation of the return value of the same section function $\gsection$ evaluated on different points $\gbasepoint \in \gbase$ in the base space.
-We represent all $\gfiber_{\gbasepoint}$ as the fiber $\gfiber$ because a 2D bundle is a trivial bundle so $\gfiber_{\gbasepoint} = \gfiber$ for all \gbasepoint.
+\note{add alpha and z channels}
   \label{fig:atct:fb:graphic}}
@@
 In \autoref{fig:atct:fb:graphic} these pixels are approximated as the small orange and black colored boxes.
-Each box is the output of the $\gsection(\gbasepoint)$ section that intersects the box.
+Each pixel is the output of the $\gsection(\gbasepoint)$ section that intersects the box.
 The set of all pixels returned by a section evaluated on a given visual base space $\gsection|_{\gbase}$ can yield a visual element, such as a marker, line, or piece of a glyph.
 \autoref{fig:atct:fb:graphic} illustrates a highly idealized space with no overlaps.
-To support overlaps, we can add another fiber element $\gfiber_{z}$ and leave it to the renderer to appropriately blend layers.
+Overlaps can be indicated via fiber element $\gfiber_{z}$ and it can be left to the renderer to appropriately blend layers based on $\gfiber_{z}$ and $\gfiber_{a}$.
 

### 2022-09-06 paper/main.tex
reworked data example

@@
 \begin{enumerate}
-  \item a collection of \textit{objects} $X \in \mathcal{Ob}(\mathcal{C})$
-  \item for every pair of objects $X, Y \in \mathcal{Ob}(\mathcal{C})$, a set of \textit{morphisms} $X \xrightarrow{f} Y \in Hom_{\mathcal{C}}(X, Y)$
+  \item a collection of \textit{objects} $X \textbf{ob}(\mathcal{C})$
+  \item for every pair of objects $X, Y \in \textbf{ob}(\mathcal{C})$, a set of \textit{morphisms} $X \xrightarrow{f} Y \in Hom_{\mathcal{C}}(X, Y)$
   \item for every object $X$, a distinct \textit{identity morphism} $X \xrightarrow {id_x} X$ in $Hom_{\mathcal{C}}(X, X)$
@@
   \includegraphics[width=\columnwidth]{trivnon.png}
-  \caption{One local trivialization of both a trivial cylinder bundle and a non-trivial mobius bundle is a plane over the interval $\left[0,2\pi\right]$.
-The transition map that transforms this trivialization into a cylinder is an identity map aligning two \textit{up} fibers, while the transition map that transforms this trivialization into a mobius bundle aligns an \textit{up} fiber with a \textit{down} fiber.
-  \note{not sure about wording here/ maybe should still have two transition maps so non-trivial has 1 id and 1 not-id}
-  \label{fig:atct:nontrivialbundle}}
+  \caption{A fiber bundle with a cylinderical total space, which is a trivial bundle, and bundle with a mobius band total space, which is non-trivial, can be covered by the same set of local trivializations.
+For example, they can both be decomposed into two planes over two intervals $\left(-\epsilon,\pi+\epsilon\right), \left(\pi-\epsilon, 2\pi+\epsilon\right)$.
+The transition maps that transforms this set of local trivializations into a cylinder are two identity map aligning two \textit{up} fibers, while the transition maps that transform this set into a mobius bundle are an identity map and a map that aligns an \textit{up} fiber with a \textit{down} fiber.
+\note{in final may change this to something that's not half half to make the arbitraryness more clear/break it up into smaller pieces} \label{fig:atct:nontrivialbundle}}
 \end{figure}
 
-As shown in \autoref{fig:atct:nontrivialbundle}, a cylinder bundle and mobius band bundle over the same base space (here a unit circle) can be locally trivialized in the same manner.
-The local trivialization in \autoref{fig:atct:nontrivialbundle} is a plane over the interval $\left[0,2\pi\right]$.
-In the cylinder bundle, all fibers go in the same direction, \textit{up}, such that transition maps on local trivializations are all identity maps aliging two \textit{up} fibers.
-This property of transition maps is what charaterizes it as a trivial bundle, while local trivializations of  non-trivial bundles always have at least on non-identity transition map.
-In \autoref{fig:atct:nontrivialbundle} this is the sole transtion map and it aligns an \textit{up} fiber with a \textit{down} fiber, thereby creating the twist in the mobius band.
-The data is encoded as the section of the bundle, here exemplified as the sin curve.
-In the cylinder bundle, the sin curve is continuous on the whole space, while the sin curve is discontinuous at the twist in the mobius band and can be made continuous via the transition map.
+As shown in \autoref{fig:atct:nontrivialbundle}, a trivial bundle with a cylinderical total space and a non-trivial bundle with mobius band total spacee can be covered with the same local trivializations; the distinguishing factor between these bundles is that the transition maps on these trivializations differ.
+The fibers in the cylinder bundle align in the same direction, \textit{up}, such that the transition maps that assemble the local trivializations into the cylinderical total space are all identity maps alinging two \textit{up} fibers.
+In the mobius bundle, there is twist where the fibers align in different directions, shown in the inset in \autoref{fig:atct:nontrivialbundle}.
+This twist is not present in either local trivialization because, by definition, fibers need to be aligned in local trivializations; therefore the transition map that aligns the fiber at the overlap between $\left(-\epsilon, \epsilon\right)$ must account for this flip by aligning \textit{up} fibers \textit{down} fibers.
+In this example, the transition map on the $\left(\pi-\epsilon, \pi+\epsilon\right)$ is an identity map since the fibers are aligned in the same direction on that section of the mobius band.
+\note{add something in the figure highlighting this part of the cylinder and the mobius band}
 
@@
 The set of all pixels returned by a section evaluated on a given visual base space $\gsection|_{\gbase}$ can yield a visual element, such as a marker, line, or piece of a glyph.
-\autoref{fig:atct:fb:graphic} illustrates a highly idealized space with no overlaps.
-Overlaps can be indicated via fiber element $\gfiber_{z}$ and it can be left to the renderer to appropriately blend layers based on $\gfiber_{z}$ and $\gfiber_{a}$.
+While \autoref{fig:atct:fb:graphic} illustrates a highly idealized space with no overlaps, overlaps can be managed via a fiber element $\gfiber_{z}$ for ordering.
+It is left to the renderer to choose how to blend layers based on $\gfiber_{z}$ and $\gfiber_{a}$.
 
@@
 Sheaves are built out of presheaves, which are functors, meaning they are functions from objects of one category to objects of another category\cite{bradleyWhatFunctorDefinitions} that preserve the morphisms in each category.
-\begin{definition} A \textbf{functor} is a map $F: \mathcal{C} \rightarrow \mathcal{D}$ such that every object $C \in \mathcal{OB(C)}$ has a corresponding object $F(C) \in \mathcal{OB(D)}$ in category $\mathcal{D}$.
+\begin{definition} A \textbf{functor} is a map $F: \mathcal{C} \rightarrow \mathcal{D}$ such that every object $C \in \textbf{ob}(\mathcal{C})$ has a corresponding object $F(C) \in \textbf{ob}(\mathcal{D})$ in category $\mathcal{D}$.
 The functor also acts on morphisms such that for every morphism $f: C_1 \rightarrow C_2$ in category $\mathcal{C}$, there is a compatible morphism $F(f): F(C_{1}) \rightarrow F(C_2)$.
@@
   \item \textit{identity}: $F(id_{C}(C)) = id_{D}(F(C))$
-  \item \textit{composition}:$F(g)\circ F(f) = F(g\circ f)$ for any composable morphisms $C_{1}\xrightarrow{f} C_2$, $C_2 \xrightarrow{g} C_3$
+  \item \textit{composition}: $F(g)\circ F(f) = F(g\circ f)$ for any composable morphisms $C_{1}\xrightarrow{f} C_2$, $C_2 \xrightarrow{g} C_3$
 \end{itemize}
@@
 \subsection{Equivariance}
-\label{sec:artist:equiv}\note{our basis of structure preservation - is this even equivariance}
-to actions $\dfunc \in \dfuncset$ on the sheaf $\sheaf_{\dbase, \dtotal}$.
-
+\label{sec:artist:equiv}
+As introduced in \autoref{sec:related-work:equivariance}, data and the corresponding visual encoding are expected to have compatible structure.
+This structure can be formally expressed as actions $\dfunc \in \dfuncset$ on the sheaf $\sheaf_{\dbase, \dtotal}$.
+\note{rewrite for generalities probably}
 \begin{definition} 
-  An \textbf{action}  $act: \dfuncsetc \times \dtotalc \rightarrow \sheaf_{\dbase, \dtotal}$ is a function that expresses that elements of set $\dfuncset$ act on elements of the sheaf of sections.Actions are associative such that
+  An \textbf{action}  $act: \dfuncsetc \times \sheaf_{\dbase, \dtotal} \rightarrow \sheaf_{\dbase, \dtotal}$ is a function that expresses that elements of set $\dfuncset$ act on elements of the set of sections $\sheaf_{\dbase, \dtotal}$.
+Actions are associative such that
   \begin{equation}
@@
 
-\paragraph{Transform Continuity}
-
-The base space transformation is from one open set to another open set in the same base space
+\paragraph{Transform Continuity}\\
+
+The base space transformation is a pointwise continuous map from one open set to another open set in the same base space
 \begin{equation}
@@
   \label{eq:atct:morphism:verify_base}
-  \dsection(\dbase) = \dfuncpull\dsection(\dbasepoint^{\prime}) = \dsection(\dfunch(\dbase^{prime}))
+  \dsection(\dbase) = \dfuncpull\dsection(\dbasepoint^{\prime}) = \dsection(\dfunch(\dbasepoint^{\prime}))
 \end{equation} 
@@
 
-\paragraph{Transform Records}
-
-As introduced in \autoref{eq:artist:sheaves:monoid_morphism}, the fiber transformation $\dfunct: \dfuncpull \dtotal_{\dbasepoint^{\prime}} \rightarrow \dfuncpull \dtotal_{\dbasepoint^{\prime}}$ is a morphism on the fiber $\dfunct \in Hom(\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}},\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}})$ restricted to a point $\dbasepoint^{\prime} \in \openset^{\prime}$.
+\paragraph{Transform Records}\\
+
+As introduced in \autoref{eq:artist:sheaves:monoid_morphism}, the fiber transformation $\dfunct: \dfuncpull \dtotal_{\dbasepoint^{\prime}} \rightarrow \dfuncpull \dtotal_{\dbasepoint^{\prime}}$ is a monoid action on the fiber $\dfunct \in Hom(\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}},\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}})$ restricted to a point $\dbasepoint^{\prime} \in \openset^{\prime}$.
 The fiber transformation is a change in section

### 2022-09-08 paper/main.tex
sections for measurment transforms

@@
 
-We propose that the map $\pi$ can express the formal binding between having a typed data field, discussed in \autoref{sec:atct:fb:fiber}, and a corresponding point in the topological structure, as discussed in \autoref{sec:atct:fb:base}.
-For our purposes, the advantage of fiber bundles is that the field type and topological structure are unconstrained in terms of dimensionality and the only conditions that must be satisfied are that every index point has a corresponding field type and the field types must be the same for every index point.
+\begin{figure}[h!]
+  \includegraphics[width=\columnwidth]{trivnon.png}
+  \caption{A fiber bundle with a cylinderical total space, which is a trivial bundle, and bundle with a mobius band total space, which is non-trivial, can be covered by the same set of local trivializations.
+For example, they can both be decomposed into two planes over two intervals $\left(-\epsilon,\pi+\epsilon\right), \left(\pi-\epsilon, 2\pi+\epsilon\right)$.
+The transition maps that transforms this set of local trivializations into a cylinder are two identity map aligning two \textit{up} fibers, while the transition maps that transform this set into a mobius bundle are an identity map and a map that aligns an \textit{up} fiber with a \textit{down} fiber.
+\note{in final may change this to something that's not half half to make the arbitraryness more clear/break it up into smaller pieces} \label{fig:atct:nontrivialbundle}}
+\end{figure}
+
+As shown in \autoref{fig:atct:nontrivialbundle}, a trivial bundle with a cylinderical total space and a non-trivial bundle with mobius band total spacee can be covered with the same local trivializations; the distinguishing factor between these bundles is that the transition maps on these trivializations differ.
+The fibers in the cylinder bundle align in the same direction, \textit{up}, such that the transition maps that assemble the local trivializations into the cylinderical total space are all identity maps alinging two \textit{up} fibers.
+In the mobius bundle, there is twist where the fibers align in different directions, shown in the inset in \autoref{fig:atct:nontrivialbundle}.
+This twist is not present in either local trivialization because, by definition, fibers need to be aligned in local trivializations; therefore the transition map that aligns the fiber at the overlap between $\left(-\epsilon, \epsilon\right)$ must account for this flip by aligning \textit{up} fibers \textit{down} fibers.
+In this example, the transition map on the $\left(\pi-\epsilon, \pi+\epsilon\right)$ is an identity map since the fibers are aligned in the same direction on that section of the mobius band.
+\note{add something in the figure highlighting this part of the cylinder and the mobius band, possibly add in a section}
+
+We propose that the total space of a bundle can encode the mathematical space in which a  dataset is embedded, the base space can encode the topological properties of the dataset, and the fiber space can encode the data types of the record fields of the dataset.
+The map $\pi$ expresses the formal binding between having a typed data field, discussed in \autoref{sec:atct:fb:fiber}, and a corresponding point in the topological structure, which is discussed in \autoref{sec:atct:fb:base}.
+Fiber bundles are a good abstract data representation for visualization because the field type and topological structure are unconstrained in terms of dimensionality and the only conditions that must be satisfied are that every point in the base space has a corresponding field of values and the field types must be the same for every point in the base space.
 We propose that modeling data as sections provides a way to encapsulate topological and field structure in a uniform dimension and type independent manner, as discussed in \autoref{sec:atct:fb:sections}.
@@
 
-\subsubsection{Examples}
-\begin{figure}[h!]
-  \includegraphics[width=\columnwidth]{trivnon.png}
-  \caption{A fiber bundle with a cylinderical total space, which is a trivial bundle, and bundle with a mobius band total space, which is non-trivial, can be covered by the same set of local trivializations.
-For example, they can both be decomposed into two planes over two intervals $\left(-\epsilon,\pi+\epsilon\right), \left(\pi-\epsilon, 2\pi+\epsilon\right)$.
-The transition maps that transforms this set of local trivializations into a cylinder are two identity map aligning two \textit{up} fibers, while the transition maps that transform this set into a mobius bundle are an identity map and a map that aligns an \textit{up} fiber with a \textit{down} fiber.
-\note{in final may change this to something that's not half half to make the arbitraryness more clear/break it up into smaller pieces} \label{fig:atct:nontrivialbundle}}
-\end{figure}
-
-As shown in \autoref{fig:atct:nontrivialbundle}, a trivial bundle with a cylinderical total space and a non-trivial bundle with mobius band total spacee can be covered with the same local trivializations; the distinguishing factor between these bundles is that the transition maps on these trivializations differ.
-The fibers in the cylinder bundle align in the same direction, \textit{up}, such that the transition maps that assemble the local trivializations into the cylinderical total space are all identity maps alinging two \textit{up} fibers.
-In the mobius bundle, there is twist where the fibers align in different directions, shown in the inset in \autoref{fig:atct:nontrivialbundle}.
-This twist is not present in either local trivialization because, by definition, fibers need to be aligned in local trivializations; therefore the transition map that aligns the fiber at the overlap between $\left(-\epsilon, \epsilon\right)$ must account for this flip by aligning \textit{up} fibers \textit{down} fibers.
-In this example, the transition map on the $\left(\pi-\epsilon, \pi+\epsilon\right)$ is an identity map since the fibers are aligned in the same direction on that section of the mobius band.
-\note{add something in the figure highlighting this part of the cylinder and the mobius band}
+\subsubsection{Example}
 
@@
 
-\paragraph{Transform Continuity}\\
-
+\paragraph{\textbf{Topological structure}}
+\noindent
 The base space transformation is a pointwise continuous map from one open set to another open set in the same base space
@@
 
-\paragraph{Transform Records}\\
-
+\paragraph{\textbf{Records}}
+\noindent
 As introduced in \autoref{eq:artist:sheaves:monoid_morphism}, the fiber transformation $\dfunct: \dfuncpull \dtotal_{\dbasepoint^{\prime}} \rightarrow \dfuncpull \dtotal_{\dbasepoint^{\prime}}$ is a monoid action on the fiber $\dfunct \in Hom(\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}},\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}})$ restricted to a point $\dbasepoint^{\prime} \in \openset^{\prime}$.
@@
 
-\paragraph{Transform Continuity and Records}
-
-We define a full data transformation as one that induces both a remapping of the index space and a change in the data values
+\paragraph{\textbf{Topological structure and records}} 
+\noindent We define a full data transformation as one that induces both a remapping of the index space and a change in the data values
 \begin{equation}
@@
 
-\section{Discussion: Feasibility of Math as Design Spec}
+\section{Discussion: Feasibility as Design Spec}
 \label{sec:discussion}

### 2022-09-08 paper/main.tex
3.2

@@
 \end{equation} 
-A continuous surjective map \textbf{$\pi$ is bundle projection} map when 
+A continuous surjective map $\bm{\pi}$ is a \textbf{bundle projection} map when 
 \begin{enumerate}
@@
 Sheaves are built out of presheaves, which are functors, meaning they are functions from objects of one category to objects of another category\cite{bradleyWhatFunctorDefinitions} that preserve the morphisms in each category.
-\begin{definition} A \textbf{functor} is a map $F: \mathcal{C} \rightarrow \mathcal{D}$ such that every object $C \in \textbf{ob}(\mathcal{C})$ has a corresponding object $F(C) \in \textbf{ob}(\mathcal{D})$ in category $\mathcal{D}$.
-The functor also acts on morphisms such that for every morphism $f: C_1 \rightarrow C_2$ in category $\mathcal{C}$, there is a compatible morphism $F(f): F(C_{1}) \rightarrow F(C_2)$.
-A functor must satisfy the properties 
+\begin{definition} A \textbf{functor} is a map $F: \mathcal{C} \rightarrow \mathcal{D}$, which means it is a function between objects $F: \textbf{ob}(\mathcal{C}) \mapsto \textbf{ob}(\mathcal{D})$ and that for every morphism $f \in Hom(C_1, C_2)$  there is a corresponding function $F: Hom(C1, C2) \mapsto Hom(F(C_1), F( C_2))$.
+A \textbf{functor} must satisfy the properties 
 \begin{itemize}
@@
 \end{itemize}
+$F(C) \in \textbf{ob}(\mathcal{D})$ denotes the object to which an object $C$ is mapped, and $F(f) \in Hom(F_(C_1), F_(C_2))$ denotes the morphism that $f$ is mapped to. 
 \end{definition}
-Functors preserve all the structure of the categories, namely the domains, codomains, composition, and identities\cite{riehlCategoryTheoryContext}.
+
+Functors preserve the domains, codomains, composition, and identities\cite{riehlCategoryTheoryContext}.
 Specifically, functors allow us to express which structure must be preserved when an object is transformed from one category to another.
-
-We use a presheaf to encode the expectation that continuity is preserved across sections. 
+In this work, we use a presheaf functor to encode the constraints that ensure that the topological structure has been reserved.
+
 \begin{definition}
@@
 \begin{itemize}
-  \item \textit{locality} two sections in a sheaf are equal $\dsection^{a} = \dsection^{b}$ when they evaluate to the same values over the same set of open sets $\dsection^{a}|{\openset} =  \dsection^{b}|_{\openset}$.
+  \item \textit{locality} two sections in a sheaf are equal $\dsection^{a} = \dsection^{b}$ when they evaluate to the same values over the same set of open sets $\dsection^{a}|_{\openset} =  \dsection^{b}|_{\openset}$.
   \item \textit{gluing} the union of sections defined on specific open sets is equivalent to one big section over the union of spaces $\dsection|_{\openset_{i} \cup \openset_{j}} = \dsection^{i}|_{\openset_i} \cup \dsection^{j}|_{\openset_j}$ if these sections agree on overlaps $\dsection^{i}|_{\openset_i\cap\openset_j} =  \dsection^{j}|_{\openset_i\cap\openset_j}$
@@
 The gluing axiom says that a distributed representation of a dataset, which is a set of local sections, is equivalent to a section over the union of the opensets of the local sections.
-The locality axiom asserts that the glued section function is equivalent to a function over the union if they evaluate t the same values.
+The locality axiom asserts that the glued section function is equivalent to a function over the union if they evaluate to the same values.
 The gluing axiom can also be used to generate the gluing rules used to construct non-trivial bundles from the set of trivial local section.
@@
 
-Each section of a sheaf over a point returns a single record in the fiber. The sheaf over an open set $\openset$ surrounding a point $\dbasepoint$ is called a stalk\cite{StalkSheaf2019}
+Each section of a sheaf over a point returns a single record in the fiber. The sheaf over an open set $\openset$ surrounding a point $\dbasepoint$ is called a \textit{stalk}\cite{StalkSheaf2019}
 \begin{equation}
@@
 where the fiber is contained inside the stalk  $\dfiber_{\dbasepoint} \subset  \sheaf_{\dbase, \dtotal}\restriction_{\dbasepoint}$.
-The germ is the section evaluated at a point in the stalk  $\dsection(\dbasepoint) \in \sheaf_{\dbase, \dtotal}\restriction_{\dbasepoint}$ and is the data.
-The stalk and germ in the stalk provide the derivative of the record, which is needed for some computational and visualization tasks.
+The \textit{germ} is the section evaluated at a point in the stalk  $\dsection(\dbasepoint) \in \sheaf_{\dbase, \dtotal}\restriction_{\dbasepoint}$ and is the data.
+Since the stalk and the germ include the values near the limit of the point at \dbasepoint, the germ can be used to compute the mathematical derivative of the data for visualization tasks that require this information.
 

### 2022-09-09 paper/main.tex
summary table

@@
 
+\section{Objects & Functions}
+This table contains a summary of the topological structures introduced in \autoref{sec:atct:fiber-bundles} and \autoref{sec:construction:vtotal} and what they encode.
+These structures serve as the basis of abstractions for the data objects in \autoref{sec:discussion}.
+
+\begin{table}[h]
+  \centering
+  \begin{tabular}{|r|r|r|r|r|}
+    \hline\\
+   & Data  & Visual  & Graphic  & Encodes \\
+\hline \\
+\textcolor{total}{total space} & \dtotal & \vtotal & \gtotal &  \\
+\hline\\
+\textcolor{base}{base space} & \dbase & \dbase & \gbase & topological structure\\
+\hline \\ 
+\textcolor{base}{open set} & \openset & \openset & \opensetg & subspace \\
+\hline \\
+\textcolor{fiber}{fiber space} & \dfiber & \vfiber & \gfiber  & field structure\\
+\hline \\
+\textcolor{section}{Section} & \dsection & \vsection & \gsection & data set \\
+\hline
+  \end{tabular}
+\end{table}
+
+This table contains a summary of the functions introduced in \autoref{sec:atct:xi}, \autoref{sec:artist}, and \autoref{sec:construction}.
+These functions are an abstraction of the behavior implemented in visualization libraries.
+
 % use section* for acknowledgment

### 2022-09-09 paper/main.tex
appendix tables

@@
 Every data section can also be copied over to the graphic space such that each piece of the (idealized in infinite resolution) graphic $\gsection(\dbasepoint)$ has a corresponding data value $\vindexpull\dsection(\dbasepoint)$.
-  \label{fig:atct:morphisms:sheaf}}
+  \label{fig:atct:morphisms:sheaf}} \note{possibly reorient and expand this diagram to illustrate eq 55}
 \end{figure}
@@
 
+ \note{should either be oriented same as 55 and/or pushed back up to 3.3 as an intro to artist or squished a little. }
  \begin{equation}
@@
 
-\section{Objects & Functions}
+\section{Summary}
 This table contains a summary of the topological structures introduced in \autoref{sec:atct:fiber-bundles} and \autoref{sec:construction:vtotal} and what they encode.
@@
   \begin{tabular}{|r|r|r|r|r|}
-    \hline\\
+    \hline
    & Data  & Visual  & Graphic  & Encodes \\
-\hline \\
+\hline 
 \textcolor{total}{total space} & \dtotal & \vtotal & \gtotal &  \\
-\hline\\
+\hline
 \textcolor{base}{base space} & \dbase & \dbase & \gbase & topological structure\\
-\hline \\ 
+\hline 
 \textcolor{base}{open set} & \openset & \openset & \opensetg & subspace \\
-\hline \\
+\hline 
 \textcolor{fiber}{fiber space} & \dfiber & \vfiber & \gfiber  & field structure\\
-\hline \\
+\hline 
 \textcolor{section}{Section} & \dsection & \vsection & \gsection & data set \\
@@
 These functions are an abstraction of the behavior implemented in visualization libraries.
+
+\note{color and replace source/target with constraint}
+\begin{table}[h]
+  \centering
+\begin{tabular}{|r|l|r|r|}
+  \hline
+      & function & source & target \\
+  \hline
+  \textcolor{artist}{artist} & $\vartist: \Gamma(\dbase, \dtotal) \rightarrow \Gamma(\gbase, \gtotal)$ &  data & graphic \\
+  \hline 
+  \textcolor{functor}{lookup} & $\vindex: \gbase\rightarrow \dbase$ & graphic space & data space\\
+  \hline
+  \multicolumn{4}{c}{}\\
+  \hline
+  \textcolor{artist}{encoder} & $\vchannel: \Gamma(\dbase, \dtotal) \rightarrow \Gamma(\dbase, \vtotal) $ & data & visial \\
+  \hline
+  \textcolor{artist}{compositor} & $\vmark: \Gamma(\dbase,\vtotal) \rightarrow \Gamma(\gbase, \vtotal)$ & visual & graphic \\
+  \hline
+\end{tabular}
+\end{table}
+
+These functions can be verified using the following functions:
 

### 2022-09-14 paper/main.tex
tables

--- 2022-09-08 paper/main.tex
@@
 
-\subsection{Field Structure} 
+\subsection{Field Type Structure} 
 \label{sec:related-work:equivariance}
@@
 An \textit{equivariant} function preserves symmetric group structure on its input and output.
+A group $G$ is a set of elements $g \in G$ with an invertable binary operator, identity, and
+
 Given a group $G$ that acts on input $X$ and output $Y$ and a function $f: X \rightarrow Y$, the function $f$ is \textit{equivariant} when $f(g(x)) = g(f(x))$ for all $g$ in $G$ and for all $x$ in $X$\cite{nlab:equivariant}.
@@
 
-The notion that visual encodings should be \textit{homomorphic} was proposed by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987} his specification of \textit{A Presentation Tool} and the idea that visual transforms are \textit{equivariant} underlies Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}'s Algebraic Visualization Design.
+The notion that visual encodings should be \textit{homomorphic} was proposed by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987} in his specification of \textit{A Presentation Tool} and the idea that visual transforms are \textit{equivariant} underlies Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}'s Algebraic Visualization Design.
 Kindlemann and Scheidegger propose three specific types of structure preservation: that the visualization should not change if the data representation (i.e.
@@
 
-\subsubsection{Data Fields: Fiber Space \dfiber}
+\subsubsection{Data Field Types: Fiber Space \dfiber}
 \label{sec:atct:fb:fiber}
@@
 
+\section{Summary}
+The topological spaces and functions introduced throughout this paper are summarized here for reference.
+
+\begin{table}[h!]
+  \centering
+  {\renewcommand{\arraystretch}{1.2}
+  \begin{tabular}{|r c c c l|}
+    \hline
+ Topological Space  & Data  & Visual  & Graphic  & Encodes \\
+   \hline
+\textcolor{total}{total space} & \dtotal & \vtotal & \gtotal & fields + topology \\
+\textcolor{base}{base space} & \dbasec & \dbasec & \gbasec & topological structure\\
+\textcolor{base}{open set} & \opensetc & \opensetc & \opensetgc & subspace \\
+\textcolor{fiber}{fiber space} & \dfiberc & \vfiberc & \gfiberc  & field structure\\
+\hline
+  \end{tabular}
+  \caption{Topological spaces introduced in \autoref{sec:atct:fiber-bundles} and \autoref{sec:construction:vtotal}}
+  \label{tab:appendix:summary:objects}
+\end{table}
+
+\begin{table}[h!]
+  \centering
+  {\renewcommand{\arraystretch}{1.2}
+  \begin{tabular}{|r c c c l|}
+     & topological structure & subspace & field structure &  fields + topology\\
+     & \textcolor{base}{total space} & \textcolor{base}{openset} & \textcolor{fiber}{field types} && 
+  \end{tabular}
+  \caption{Topological spaces introduced in \autoref{sec:atct:fiber-bundles} and \autoref{sec:construction:vtotal}}
+  \label{tab:appendix:summary:objects}
+\end{table}
+
+\begin{table}[h!]
+  \centering
+  {\renewcommand{\arraystretch}{1.2}
+  \begin{tabular}{|r c c c l|}
+
+\textcolor{section}{section} & $\dsectionc: \dbase \textcolor{section}{\rightarrow} \dfiberc$ &  $\vsectionc: \dbase \textcolor{section}{\rightarrow} \vfiberc$ & $\gsectionc: \gbase \textcolor{section}{\rightarrow} \gfiberc$ & data set \\
+\textcolor{sheaf}{sheaf} & $\sheafc_{\dbasec, \dtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$ & $\sheafc_{\dbasec, \vtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}}$ & $\sheafc_{\gbasec, \gtotalc}: \opensetgc \rightarrow \cgamma{\opensetc}{\gtotalc\restriction_{\opensetgc}}$ & subspace to records map \\
+\hline
+  \end{tabular}
+  \caption{Functions that associate topological subspaces with records.
+Equivalent constructions can be applied to elements in visual and graphic sheaves, and these functions are distinguised through subscripts $\dfunc_{\dtotal}$, $\dfunc_{\vtotal}$ and $\dfunc_{\gtotal}$}
+  \label{tab:appendix:summary:datafunctions}
+\end{table}
+
+\begin{table}[h!]
+\centering
+{\renewcommand{\arraystretch}{1.2}
+\caption{Functions $\dfunc=(\dfunchc, \dfunctc)$ for modifying data records}
+\label{tab:appendix:summary:datamod}
+\end{table}
+
+\note{color and replace source/target with constraint}
+\begin{table}[h!]
+  \centering
+\begin{tabular}{|r|l|l|}
+  \hline
+      & function & constraint \\
+  \hline
+  \textcolor{artist}{artist} & $\vartist: \Gamma(\dbase, \dtotal) \rightarrow \Gamma(\gbase, \gtotal)$ &  \\
+  \hline 
+  \textcolor{functor}{lookup} & $\vindex: \gbase\rightarrow \dbase$  &\\
+  \multicolumn{4}{c}{}\\
+  \hline
+  \textcolor{artist}{encoder} & $\vchannel: \Gamma(\dbase, \dtotal) \rightarrow \Gamma(\dbase, \vtotal) $  & \\
+  \hline
+  \textcolor{artist}{compositor} & $\vmark: \Gamma(\dbase,\vtotal) \rightarrow \Gamma(\gbase, \vtotal)$ &  \\
+  \hline
+\end{tabular}
+\caption{artist, verification functions, and construction $\vartist = \vmarkc \circ \nu$ introduced in \autoref{sec:atct:xi}, \autoref{sec:artist}, and \autoref{sec:construction}}
+\label{tab:appedendix:summary:artist}
+\end{table}
+
+These functions can be verified using the following functions:
+
 % use section* for acknowledgment

### 2022-09-14 paper/main.tex
sec 2 intro

@@
 Two mathematical concepts for describing a function that preserves structure are \textit{equivariance} and \textit{homomorphism}.
-An \textit{equivariant} function preserves symmetric group structure on its input and output.
-A group $G$ is a set of elements $g \in G$ with an invertable binary operator, identity, and
-
-Given a group $G$ that acts on input $X$ and output $Y$ and a function $f: X \rightarrow Y$, the function $f$ is \textit{equivariant} when $f(g(x)) = g(f(x))$ for all $g$ in $G$ and for all $x$ in $X$\cite{nlab:equivariant}.
-A \textit{homomorphic} function $f$ preserves binary operations on input and output of the same algebraic type.
-A map $f: X \rightarrow Y$ is a \textit{homomorphism} when  $f(x_1 \oslash x_2) = f(x_1) \oslash f(x_2)$ for every pair $x_1, x_2$ in $X$\cite{grimaldiDiscreteCombinatorialMathematics2006}.
+An \textit{equivariant} function preserves symmetric group structure on its input and output, while a \textit{homomorphic} function $f$ preserves binary operations on input and output of the same algebraic type.
+Group structure and binary operations are both describing families of functions or operations that can be applied to data and visual encodings.
+A group $G$ is a set with an operator $\circ$.
+The set contains an identity element $e$ and the operator $\circ$ is closed, associative and invertible.
+Applying elements of a set, such as a group, to another set, such as data $X$, is called an \textit{action}.
+\begin{definition}
+  An \textcolor{action}{\textbf{action}} $act: G \times X \rightarrow X$ is a function that   satisfies the axioms of identity $act(e, x)$ for $e \in G$ and associativity $act(g, act(f, x)) = act(f \circ g, x)$ for $f,g \in G$ \cite{nlab:action}.
+\end{definition}
+Given a group $G$ that acts on input $X$ and output $Y$ and a function $f: X \rightarrow Y$, 
+ \begin{definition}
+$f$ is \textbf{equivariant} when $f(g(x)) = g(f(x))$ for all $g$ in $G$ and for all $x$ in $X$\cite{nlab:equivariant}.
+ \begin{definition}
+ $f$ is \textbf{homomorphic} when  $f(x_1 \oslash x_2) = f(x_1) \oslash f(x_2)$ for every pair $x_1, x_2$ in $X$\cite{grimaldiDiscreteCombinatorialMathematics2006}. 
+ \end{definition}
 For example, nominal measurements are permutable, meaning that they are acted on by elements of a permutation group.
 For the visual transformation to be equivariant, that nominal measurement must be mapped to a visual encoding that is also permutable, such as distinct marker shapes or color hues.
-Partial orders are non-invertable, so a visual encoding is homomorphic when data that is partially ordered is mapped to visual encodings that are partially ordered in the same way.
+We evaluate whether the encoding is homomorphic when the visual encoding does not have a group structure.
+For example, partial orders are not invertible; therefore a visual encoding of partially ordered data is structure preserving when the ordering is mapped to equivalently ordered visual encodings.
 
@@
 This structure can be formally expressed as actions $\dfunc \in \dfuncset$ on the sheaf $\sheaf_{\dbase, \dtotal}$.
-\note{rewrite for generalities probably}
-\begin{definition} 
-  An \textbf{action}  $act: \dfuncsetc \times \sheaf_{\dbase, \dtotal} \rightarrow \sheaf_{\dbase, \dtotal}$ is a function that expresses that elements of set $\dfuncset$ act on elements of the set of sections $\sheaf_{\dbase, \dtotal}$.
-Actions are associative such that
-  \begin{equation}
-    act(\dfuncc_{j}, act(\dfuncc_{i}, \dsectionc)) = act(\dfuncc_{j} \circ \dfuncc_{i}, \dsectionc)
-  \end{equation}
- for all elements  $\dfunc_{i}, \dfunc_{j}$ of the set $\dfuncset$ equipped with operator $\circ$ and all elements \dsection\ of the sheaf  $\sheaf_{\dbase, \dtotal}$. \cite{nlab:action}
-\end{definition}.
-
 We generalize from binary operations to a family of actions because that allows for expanding the set of allowable transformations on the data beyond a single operator.
@@
   \caption{Topological spaces introduced in \autoref{sec:atct:fiber-bundles} and \autoref{sec:construction:vtotal}}
-  \label{tab:appendix:summary:objects}
+  \label{}
 \end{table}
@@
   {\renewcommand{\arraystretch}{1.2}
-  \begin{tabular}{|r c c c l|}
+  \begin{tabular}{|r r r r r|}
      & topological structure & subspace & field structure &  fields + topology\\
-     & \textcolor{base}{total space} & \textcolor{base}{openset} & \textcolor{fiber}{field types} && 
+     & \textcolor{base}{base space} & \textcolor{base}{openset} & \textcolor{fiber}{field types} & \textcolor{total}{total space}
   \end{tabular}

### 2022-09-14 paper/main.tex
summary tables

@@
 $f$ is \textbf{equivariant} when $f(g(x)) = g(f(x))$ for all $g$ in $G$ and for all $x$ in $X$\cite{nlab:equivariant}.
+ \end{definition}
  \begin{definition}
@@
   \centering
-  {\renewcommand{\arraystretch}{1.2}
-  \begin{tabular}{|r c c c l|}
+  {\renewcommand{\arraystretch}{1.5}
+  \begin{tabular}{|r | c c c|}
     \hline
- Topological Space  & Data  & Visual  & Graphic  & Encodes \\
+    &\textcolor{base}{point}/\textcolor{base}{openset}/\textcolor{base}{base space} & \textcolor{fiber}{fiber space} & \textcolor{total}{total space}\\
+     &  location/subset/topology & fields &  dataset type\\
     \hline
-\textcolor{total}{total space} & \dtotal & \vtotal & \gtotal & fields + topology \\
-\textcolor{base}{base space} & \dbasec & \dbasec & \gbasec & topological structure\\
-\textcolor{base}{open set} & \opensetc & \opensetc & \opensetgc & subspace \\
-\textcolor{fiber}{fiber space} & \dfiberc & \vfiberc & \gfiberc  & field structure\\
+   Data & $\dbasepoint \in \opensetc \subseteq \dbasec$ & \dfiberc & \dtotalc\\
+   Visual & $\dbasepoint \in \opensetc \subseteq \dbasec$  & \vfiberc & \vtotalc\\
+   Graphic & $\gbasepoint \in \opensetgc \subseteq \gbasec$ & \gfiberc & \gtotalc\\
    \hline
   \end{tabular}
-  \caption{Topological spaces introduced in \autoref{sec:atct:fiber-bundles} and \autoref{sec:construction:vtotal}}
-  \label{}
+  \caption{Topological spaces introduced in \autoref{sec:atct:fiber-bundles}}
+  \label{tab:appendix:summary:objects}
+  }
+\end{table}
+
+\begin{table}[h!]
+  \centering
+  {\renewcommand{\arraystretch}{1.5}
+  \begin{tabular}{|r | l l | }
+    \hline
+     & \textcolor{section}{section} & \textcolor{sheaf}{sheaf} \\
+     & location to record &  subset to set of possible records  \\
+     \hline
+  Data & $ \cgamma{\dbasec}{\dtotalc} \ni \dsectionc: \dbase \textcolor{section}{\rightarrow} \dfiberc$ & $\sheafc_{\dbasec, \dtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$\\
+  Visual &  $\cgamma{\dbasec}{\vtotalc} \ni \vsectionc: \dbase \textcolor{section}{\rightarrow} \vfiberc$ & $\sheafc_{\dbasec, \vtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}}$\\
+  Graphic &    $\cgamma{\gbasec}{\gtotalc} \ni \gsectionc: \gbase \textcolor{section}{\rightarrow} \gfiberc$ &  $\sheafc_{\gbasec, \gtotalc}: \opensetgc \rightarrow \cgamma{\opensetc}{\gtotalc\restriction_{\opensetgc}}$ \\
+  \hline
+  \end{tabular}
+  \caption{Functions that associate topological subspaces with records, discussed in \autoref{sec:atct:fb:sections} and \autoref{sec:atct:sheaves}}
+  \label{tab:appendix:summary:datafunctions}
+  }
+\end{table}
+
+\begin{table}[h!]
+  \centering
+  {\renewcommand{\arraystretch}{1.5}
+  \begin{tabular}{|r | l  l |}
+\hline
+ \gbasepointc\ to \dbasepointc& $ \vindexc: \opensetgc \rightarrow \opensetc$ & for $\dbasepointc \in \opensetc$ exists $\gbasepointc \in \opensetgc$ \\
+ & & s.t. $\vindexc(\gbasepointc) = \dbasepointc$\\
+ graphic for \dbasepointc & $ \vindexpushc \gsectionc: \opensetc \rightarrow \vindexpushc\gtotalc\restriction_{\opensetc}$ & $\vindexpushc\gsectionc(\dbasepointc)(\gbasepointc) = \gsectionc(\gbasepointc)$ \\
+data for \gbasepointc & $\vindexpullc\dsectionc: \opensetgc \rightarrow  \vindexpullc\dtotalc\restriction_{\opensetgc}$ & $\vindexpullc\dsectionc(\gbasepointc)=\dsectionc(\vindexc(\gbasepointc)) = \dsectionc(\dbasepointc)  \\
+\hline
+  \end{tabular}
+  \caption{Functors between graphic and data base spaces  \autoref{sec:atct:xi}}
+  \label{tab:appendix:summary:transport}
+  }
 \end{table}
@@
 {\renewcommand{\arraystretch}{1.2}
-  \begin{tabular}{|r r r r r|}
-     & topological structure & subspace & field structure &  fields + topology\\
-     & \textcolor{base}{base space} & \textcolor{base}{openset} & \textcolor{fiber}{field types} & \textcolor{total}{total space}
+\begin{tabular}{}
+  \note{changing locations, moving records to new locations, changing records}
+
 \end{tabular}
-  \caption{Topological spaces introduced in \autoref{sec:atct:fiber-bundles} and \autoref{sec:construction:vtotal}}
-  \label{tab:appendix:summary:objects}
+
+\caption{Functions $\dfunc=(\dfunchc, \dfunctc)$ for modifying data records.
+Equivalent constructions can be applied to elements in visual and graphic sheaves, and these functions are distinguised through subscripts $\dfunc_{\dtotal}$, $\dfunc_{\vtotal}$ and $\dfunc_{\gtotal}$}
+\label{tab:appendix:summary:datamod}
+}
 \end{table}
 
+\note{color}
 \begin{table}[h!]
@@
   {\renewcommand{\arraystretch}{1.2}
-  \begin{tabular}{|r c c c l|}
-
-\textcolor{section}{section} & $\dsectionc: \dbase \textcolor{section}{\rightarrow} \dfiberc$ &  $\vsectionc: \dbase \textcolor{section}{\rightarrow} \vfiberc$ & $\gsectionc: \gbase \textcolor{section}{\rightarrow} \gfiberc$ & data set \\
-\textcolor{sheaf}{sheaf} & $\sheafc_{\dbasec, \dtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$ & $\sheafc_{\dbasec, \vtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}}$ & $\sheafc_{\gbasec, \gtotalc}: \opensetgc \rightarrow \cgamma{\opensetc}{\gtotalc\restriction_{\opensetgc}}$ & subspace to records map \\
-\hline
-  \end{tabular}
-  \caption{Functions that associate topological subspaces with records.
-Equivalent constructions can be applied to elements in visual and graphic sheaves, and these functions are distinguised through subscripts $\dfunc_{\dtotal}$, $\dfunc_{\vtotal}$ and $\dfunc_{\gtotal}$}
-  \label{tab:appendix:summary:datafunctions}
-\end{table}
-
-\begin{table}[h!]
-\centering
-{\renewcommand{\arraystretch}{1.2}
-\caption{Functions $\dfunc=(\dfunchc, \dfunctc)$ for modifying data records}
-\label{tab:appendix:summary:datamod}
-\end{table}
-
-\note{color and replace source/target with constraint}
-\begin{table}[h!]
-  \centering
 \begin{tabular}{|r|l|l|}
@@
   \textcolor{functor}{lookup} & $\vindex: \gbase\rightarrow \dbase$  &\\
-  \multicolumn{4}{c}{}\\
   \hline
@@
 \end{tabular}
-\caption{artist, verification functions, and construction $\vartist = \vmarkc \circ \nu$ introduced in \autoref{sec:atct:xi}, \autoref{sec:artist}, and \autoref{sec:construction}}
+\caption{artist, verification functions, and construction $\vartist = \vmarkc \circ \nu$ introduced in \autoref{sec:artist}, and \autoref{sec:construction}}
 \label{tab:appedendix:summary:artist}
+}
 \end{table}

### 2022-09-16 paper/main.tex
math aug paper

@@
 A group $G$ is a set with an operator $\circ$.
-The set contains an identity element $e$ and the operator $\circ$ is closed, associative and invertible.
+The set contains an identity element $e \in G$ and the operator $\circ$ is closed, associative and invertible.
 Applying elements of a set, such as a group, to another set, such as data $X$, is called an \textit{action}.
 \begin{definition}
-  An \textcolor{action}{\textbf{action}} $act: G \times X \rightarrow X$ is a function that   satisfies the axioms of identity $act(e, x)$ for $e \in G$ and associativity $act(g, act(f, x)) = act(f \circ g, x)$ for $f,g \in G$ \cite{nlab:action}.
+  An \textcolor{action}{\textbf{action}} \textit{act} of $G$ on $X$ is a function  $act: G \times X \rightarrow X$.
+An action has the properties of identity $act(e, x) = x$ for all $ x \in X$ and associativity $act(g, act(f, x)) = act(f \circ g, x)$ for $f,g \in G$ \cite{nlab:action, wki}.
 \end{definition}
-Given a group $G$ that acts on input $X$ and output $Y$ and a function $f: X \rightarrow Y$, 
+In \autoref{sec:artist:equivariant}, actions on data and compatible actions on graphics define structure in terms of testing whether a visualization is equivariant.
+Given a group $G$ that acts on input $X$ and output $Y$ and a function $f: X \rightarrow Y$
+
  \begin{definition}
@@
 \end{definition}
-
 Functors preserve the domains, codomains, composition, and identities\cite{riehlCategoryTheoryContext}.  
-Specifically, functors allow us to express which structure must be preserved when an object is transformed from one category to another.
-In this work, we use a presheaf functor to encode the constraints that ensure that the topological structure has been reserved.
-
 \begin{definition}
@@
 \end{definition}
-Specifically, we use a presheaf to describe the expectation that any set of sections defined over a larger space $\dsection \in \Gamma(\opensetc_2, \dtotal\restriction_{\openset_{2}})$ is also defined over subspaces of that space $\dsection\restriction_{\opensetc_{1}}$.
-We do this by defining the presheaf as the association between an openset and its corresponding set of sections
+Contravariance means that the morphisms between the input openset objects go in the opposite direction from the morphisms between the output set objects.
+We specify behavior using functors because it allows us to express which structure must be preserved when an object is transformed from one category to another.
+
+We use a presheaf functor to encode the binding between an openset and corresponding sets of sections 
 \begin{equation}
@@
 \end{equation}
-
-The presheaf preserves continuity by stating that given a morphism $\iota$ between two objects of $\mathcal{\dbase}$ 
+and the morphisms preserved by the functor describe how subsets of the data are expected to behave. The presheaf is contravariant because the inclusion morphisms between input objects 
 \begin{equation*}
   \label{eq:atct:sheaf:inclusion}
-  \iota: \openset_1 \rightarrow \openset_2
+  \iota: \opensetc_1 \rightarrow \opensetc_2
 \end{equation*}
-there must be a corresponding restriction morphiism $\iota^*$ between the sets of sections 
+are defined such that they correspond to the partial ordering $\openset_1 \subseteq \openset_2$, but the restriction morphisms $\iota^*$ between the sets of sections 
 \begin{equation*}
   \label{eq:atct:sheaf:restriction}
-  \iota^*: \cgamma{\openset_2}{\dtotal_{\openset_2}} \rightarrow \cgamma{\openset_1}{\dtotal_{\openset_1}}
+  \iota^*: \cgamma{\opensetc_2}{\dtotalc_{\opensetc_2}} \rightarrow \cgamma{\opensetc_1}{\dtotalc_{\openset_c1}}
 \end{equation*}
-which are the outputs of the sheaf $\sheaf(\openset) = \cgamma{\openset}{\dtotal\restriction_{\openset}}$.
-The presheaf is contravariant because the morphisms between the input openset objects go in the opposite direction from the morphisms between the output set objects.
-The expectation that data containers implement inclusion and restriction morphisms means that, for example, a data container mist ensure that sections can be interpolated over any subset that the container can provide.
+go in the opposite direction because any arbitrary function defined on an openset $\dsection\restriction_{\openset_2}$ must be definable on a subset of that space $\dsection\restriction_{\openset_1}$.
+The expectation that data containers implement inclusion and restriction morphisms means that, for example, a data container must ensure that continuous sections can be interpolated.
 
@@
     \hline
-   Data & $\dbasepoint \in \opensetc \subseteq \dbasec$ & \dfiberc & \dtotalc\\
-   Visual & $\dbasepoint \in \opensetc \subseteq \dbasec$  & \vfiberc & \vtotalc\\
+   Data & $\dbasepointc \in \opensetc \subseteq \dbasec$ & \dfiberc & \dtotalc\\
+   Visual & $\dbasepointc \in \opensetc \subseteq \dbasec$  & \vfiberc & \vtotalc\\
    Graphic & $\gbasepoint \in \opensetgc \subseteq \gbasec$ & \gfiberc & \gtotalc\\
@@
      & \textcolor{section}{section} & \textcolor{sheaf}{sheaf} \\
-     & location to record &  subset to set of possible records  \\
+     & record at location & set of possible records for subset \\
      \hline
@@
  graphic for \dbasepointc & $ \vindexpushc \gsectionc: \opensetc \rightarrow \vindexpushc\gtotalc\restriction_{\opensetc}$ & $\vindexpushc\gsectionc(\dbasepointc)(\gbasepointc) = \gsectionc(\gbasepointc)$ \\
-data for \gbasepointc & $\vindexpullc\dsectionc: \opensetgc \rightarrow  \vindexpullc\dtotalc\restriction_{\opensetgc}$ & $\vindexpullc\dsectionc(\gbasepointc)=\dsectionc(\vindexc(\gbasepointc)) = \dsectionc(\dbasepointc)  \\
+ record for \gbasepointc & $\vindexpullc\dsectionc: \opensetgc \rightarrow  \vindexpullc\dtotalc\restriction_{\opensetgc}$ & $\vindexpullc\dsectionc(\gbasepointc)=\dsectionc(\vindexc(\gbasepointc)) = \dsectionc(\dbasepointc)$  \\
 \hline
@@
 {\renewcommand{\arraystretch}{1.2}
-\begin{tabular}{}
+\begin{tabular}{|c|}
+
+\end{tabular}
 \note{changing locations, moving records to new locations, changing records}
-
-\end{tabular}
-
 \caption{Functions $\dfunc=(\dfunchc, \dfunctc)$ for modifying data records.

### 2022-09-16 paper/main.tex
actions + top rewrite

@@
 \begin{definition}
-  An \textcolor{action}{\textbf{action}} \textit{act} of $G$ on $X$ is a function  $act: G \times X \rightarrow X$.
-An action has the properties of identity $act(e, x) = x$ for all $ x \in X$ and associativity $act(g, act(f, x)) = act(f \circ g, x)$ for $f,g \in G$ \cite{nlab:action, wki}.
+  An \textcolor{action}{\textbf{action}}\footnote{Throughout this paper, we augment the math with color to group conceptually similar objects and functions, for example \textcolor{action}{actions} and \textcolor{action}{sets of actions}.
+These colors carry through to the diagrams.\cite{headMathAugmentationHow2022}} \textit{act} of \textcolor{action}{$G$} on $X$ is a function  $act: \textcolor{action}{G} \times X \rightarrow X$.
+An action has the properties of identity $act(\textcolor{action}{e}, x) = \x$ for all $ x \in X$ and associativity $act(\textcolor{action}{g}, act(\textcolor{action}{f}, x)) = act(\textcolor{action}{f} \circ \textcolor{action}{g}, x)$ for $\textcolor{action}{f},\textcolor{action}{g} \in \textcolor{action}{G}$ \cite{nlab:action, grimaldiDiscreteCombinatorialMathematics2006}.
 \end{definition}
@@
 For example, an open ball is the set of points inside the ball and excludes points on the surface of the ball.
-\cite{weissteinOpenSet}}, of that structure.
-
+\cite{weissteinOpenSet,bradleyTopologyVsTopology}}, of that structure.
 \begin{definition}
-A \textbf{topology} $\mathcal{T}$ on a space \dbase\ is a collection of open sets \openset\ of \dbase\ that satisfy the properties 
-  \begin{itemize}
-    \item the empty set $\varnothing$ and $\dbase$ are in $\mathcal{T}$
-    \item the union of open sets in $\mathcal{T}$ is in also in $\mathcal{T}$
-    \item any finite intersection of open sets in $\mathcal{T}$ is also in $\mathscr{T}$
-  \end{itemize}
+A \textbf{topology} $\mathcal{T}$ on a space $X$ is a collection of open sets \openset\ of $X$.
+This collection has the properties that the empty set $\varnothing$ and $\dbase$ are in $\mathcal{T}$, the union of open sets in $\mathcal{T}$ is also in $\mathcal{T}$, and any finite intersection of open sets in $\mathcal{T}$ is also in $\mathscr{T}$.
+\cite{bradleyTopologyCategoricalApproach2020}
 \end{definition}
-The base space of a fiber bundle is a quotient topology\cite{QuotientSpaceTopology2020}, meaning that it divides the topological space into the smallest open sets such that $\pi$ remains a continuous function.
-For our purposes, $\dbase$ encodes the underlying connectivity of the our dataset, for example, $\dbase$ could be the discrete points, interval, or plane in \autoref{fig:related-work:continuity:ktypes}.
-We define $\dbase$ as a topology $\mathcal{T}_{\dbase}$ because a topology codifies that the underlying topological structure of the data can be subsetted and how these subsets can be composed in a structure preserving manner.
+We propose that the topology on \dbase\ can act as a mathematical abstraction for the \textcolor{base}{indexing space} of a dataset because the properties of a topology are the same as would be expected of an index space, namely that the space has a mathematical structure, that the space can be subsetted, and that continuity is preserved when sets of indices are merged.
+For example, in \autoref{fig:related-work:continuity:ktypes}, a topology on the line would break the line up into subsets, and those subsets could only be joined in ways where they cover the line in the same order in which it was broken up.
+The base space of a fiber bundle is a quotient topology\cite{QuotientSpaceTopology2020,munkresElementsAlgebraicTopology1984}, meaning that it divides the topological space into the smallest possible (largest number of) open sets such that $\pi$ remains a continuous function.
+This means that the topology can be defined to have a resolution equal to the number of indicies in a dataset or at whatever resolution of continuois function is required for a given task.
 
@@
 \end{definition}
-Spivak then describes tables as sections $\dsection: \dbase \rightarrow \Gamma^{\pi}(\sigma)$ from a key space $\dbase$ to the set of all possible records $\Gamma^{\pi}(\sigma)$ on the schema bundle, and his notion of a table generalizes to our notion of a data container.
+Spivak then describes tables as sections $\dsection: \dbase \rightarrow \Gamma^{\pi}(\sigma)$ from an indexing space $\dbase$ to the set of all possible records $\Gamma^{\pi}(\sigma)$ on the schema bundle, and his notion of a table generalizes to our notion of a data container.
 
@@
 \subsection{Data and Graphic Correspondence}
+\note{index correspondence}
 \label{sec:atct:xi}
-There is an expectation that for a visualization to be readable, the visual elements must correspond to distinct data elements\cite{ziemkiewiczEmbeddingInformationVisualization2009} and we can use the properties of sheaves to formally express this correspondence.
-First we establish that the graphic continuity corresponds to the data continuity by proposing that there exists a continuous \textcolor{functor}{map between topological spaces} \vindexc
+There is an expectation that for a visualization to be readable, the visual elements must correspond to distinct data elements\cite{ziemkiewiczEmbeddingInformationVisualization2009} and we can use the properties of sheaves to formally express this correspondence.\note{making it clearer that xi is a map between indexing spaces, and that defining it as surjective bakes in these expectations - make it clearer that we are defining functions and the way in which we define it is a choice to express constraints}
+\note{two types of definitions: standard definitions \& constructed definitions : we define things this way to have this role to encode this thing }
+
+First we establish that the graphic continuity corresponds to the data continuity by proposing that there exists a continuous \textcolor{functor}{map} \vindexc between opensets in the data base space $\openset \subseteq \dbase$
 \begin{equation}
@@
 The functor $\vindex$ is surjective such that for every point $\dbasepoint \in \openset$ there is a corresponding set of points $\{\gbasepoint | \gbasepoint \in  \vindexpre(k)\}$ for all $\gbasepoint in \opensetg$.
-
+\note{add little figure for xi showing it mapping from graphic index space back into data index space 
+}
+\note{Harder: if there exists a function between spaces, can construct the transport functors}
 \subsubsection{Graphic Corresponding to Data}
@@
     &\textcolor{base}{point}/\textcolor{base}{openset}/\textcolor{base}{base space} & \textcolor{fiber}{fiber space} & \textcolor{total}{total space}\\
-     &  location/subset/topology & fields &  dataset type\\
+     &  location/subset/indicies & fields &  dataset type\\
     \hline
@@
    Visual & $\dbasepointc \in \opensetc \subseteq \dbasec$  & \vfiberc & \vtotalc\\
-   Graphic & $\gbasepoint \in \opensetgc \subseteq \gbasec$ & \gfiberc & \gtotalc\\
+   Graphic & $\gbasepointc \in \opensetgc \subseteq \gbasec$ & \gfiberc & \gtotalc\\
    \hline
@@
   \begin{tabular}{|r | l  l |}
+\hline
+& function & constraint\\
 \hline

### 2022-09-20 paper/main.tex
more human readable xi

@@
 Applying elements of a set, such as a group, to another set, such as data $X$, is called an \textit{action}.
-\begin{definition}
-  An \textcolor{action}{\textbf{action}}\footnote{Throughout this paper, we augment the math with color to group conceptually similar objects and functions, for example \textcolor{action}{actions} and \textcolor{action}{sets of actions}.
-These colors carry through to the diagrams.\cite{headMathAugmentationHow2022}} \textit{act} of \textcolor{action}{$G$} on $X$ is a function  $act: \textcolor{action}{G} \times X \rightarrow X$.
-An action has the properties of identity $act(\textcolor{action}{e}, x) = \x$ for all $ x \in X$ and associativity $act(\textcolor{action}{g}, act(\textcolor{action}{f}, x)) = act(\textcolor{action}{f} \circ \textcolor{action}{g}, x)$ for $\textcolor{action}{f},\textcolor{action}{g} \in \textcolor{action}{G}$ \cite{nlab:action, grimaldiDiscreteCombinatorialMathematics2006}.
+
+\begin{definition}\cite{nlab:action, grimaldiDiscreteCombinatorialMathematics2006}
+  An \textcolor{action}{\textbf{action}} \footnote{Throughout this paper, we augment the math with color to group conceptually similar objects and functions\cite{headMathAugmentationHow2022}, for example \textcolor{action}{actions} and \textcolor{action}{sets of actions}.
+This color coding carries through to the figures.} of \textcolor{action}{$G$} on $X$ is a function  $act: \textcolor{action}{G} \times X \rightarrow X$.
+An action has the properties of identity $act(\textcolor{action}{e}, x) = x$ for all  $x \in X$ and associativity $act(\textcolor{action}{g}, act(\textcolor{action}{f}, x)) = act(\textcolor{action}{f} \circ \textcolor{action}{g}, x)$ for $\textcolor{action}{f},\textcolor{action}{g} \in \textcolor{action}{G}$.
 \end{definition}
+
 In \autoref{sec:artist:equivariant}, actions on data and compatible actions on graphics define structure in terms of testing whether a visualization is equivariant.
@@
 
-Sheaves are built out of presheaves, which are functors, meaning they are functions from objects of one category to objects of another category\cite{bradleyWhatFunctorDefinitions} that preserve the morphisms in each category.
-\begin{definition} A \textbf{functor} is a map $F: \mathcal{C} \rightarrow \mathcal{D}$, which means it is a function between objects $F: \textbf{ob}(\mathcal{C}) \mapsto \textbf{ob}(\mathcal{D})$ and that for every morphism $f \in Hom(C_1, C_2)$  there is a corresponding function $F: Hom(C1, C2) \mapsto Hom(F(C_1), F( C_2))$.
+We can mathematically encode that we expect data containers to preserve the underlying continuity of the indexing space and the mappings beween indexing space and record space using a type of function called a functor.
+Functors are mappings between categories that preserve the domains, codomains, composition, and identities of the morphisms within the category\cite{riehlCategoryTheoryContext}.
+
+\begin{definition}\cite{bradleyWhatFunctorDefinitions,bradleyTopologyCategoricalApproach2020} A \textbf{functor} is a map $F: \mathcal{C} \rightarrow \mathcal{D}$, which means it is a function between objects $F: \textbf{ob}(\mathcal{C}) \mapsto \textbf{ob}(\mathcal{D})$ and that for every morphism $f \in Hom(C_1, C_2)$  there is a corresponding function $F: Hom(C1, C2) \mapsto Hom(F(C_1), F( C_2))$.
 A \textbf{functor} must satisfy the properties 
@@
 \end{definition}
-Functors preserve the domains, codomains, composition, and identities\cite{riehlCategoryTheoryContext}.  
+Modeling the data container as a functor allows us state that, just like a functor, the container is a map between index space objects and sets of data records that preserve morphisms between index space objects and data records.
+\begin{equation}
+  \label{eq:atct:sheaf:functor}
+  \sheafc_{\dbasec, \dtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
+\end{equation}
+A common way of encapsulating a map from a topological space to a category of sets is as a presheaf
 \begin{definition}
-  A \textbf{presheaf} $F:\mathcal{C}^{op} \rightarrow \setb$ is a contravariant functor from an object in an arbitrary category to an object in the category \setb\cite{nlab:presheaf}. 
+  A \textbf{presheaf} $F:\mathcal{C}^{op} \rightarrow \setb$ is a contravariant functor from an object in an arbitrary category to an object in the category \setb\cite{nlab:presheaf, spanier1989algebraic}.
 \end{definition}
 Contravariance means that the morphisms between the input openset objects go in the opposite direction from the morphisms between the output set objects.
-We specify behavior using functors because it allows us to express which structure must be preserved when an object is transformed from one category to another.
-
-We use a presheaf functor to encode the binding between an openset and corresponding sets of sections 
-\begin{equation}
-  \label{eq:atct:sheaf:functor}
-  \sheafc_{\dbasec, \dtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
-\end{equation}
-and the morphisms preserved by the functor describe how subsets of the data are expected to behave. The presheaf is contravariant because the inclusion morphisms between input objects 
+The presheaf is contravariant because the inclusion morphisms between input objects
 \begin{equation*}
@@
 \end{equation*}
-go in the opposite direction because any arbitrary function defined on an openset $\dsection\restriction_{\openset_2}$ must be definable on a subset of that space $\dsection\restriction_{\openset_1}$.
+go in the opposite direction.
+This is because any arbitrary function defined on an openset $\dsection\restriction_{\openset_2}$ must be definable on a subset of that space $\dsection\restriction_{\openset_1}$.
 The expectation that data containers implement inclusion and restriction morphisms means that, for example, a data container must ensure that continuous sections can be interpolated.
 
-While the presheaf is a generalization of the concept of a global space of sections $\Gamma(\dbase, \dtotal)$ \cite{spanier1989algebraic}, the locality and gluing axioms \cite{bakerEuclideanSpaceMathsSheaf} of a sheaf describe the conditions for verifying that the sections over subspaces are part of a cohesive section over the whole space.
-\begin{definition} A \textbf{sheaf} is a presheaf that satisfies the following two axioms
+While presheaves preserve the rules for ensuring continuity in the indexing space and restriction in the sets of data records, sheaves add on conditions for verifying that sections over subspaces are part of a cohesive section over the whole space.
+\begin{definition}\cite{bakerEuclideanSpaceMathsSheaf,spanier1989algebraic} A \textbf{sheaf} is a presheaf that satisfies the following two axioms
 \begin{itemize}
@@
 \label{sec:atct:xi}
-There is an expectation that for a visualization to be readable, the visual elements must correspond to distinct data elements\cite{ziemkiewiczEmbeddingInformationVisualization2009} and we can use the properties of sheaves to formally express this correspondence.\note{making it clearer that xi is a map between indexing spaces, and that defining it as surjective bakes in these expectations - make it clearer that we are defining functions and the way in which we define it is a choice to express constraints}
-\note{two types of definitions: standard definitions \& constructed definitions : we define things this way to have this role to encode this thing }
-
-First we establish that the graphic continuity corresponds to the data continuity by proposing that there exists a continuous \textcolor{functor}{map} \vindexc between opensets in the data base space $\openset \subseteq \dbase$
+There is an expectation that for a visualization to be readable, the visual elements must correspond to distinct data elements\cite{ziemkiewiczEmbeddingInformationVisualization2009} and we can use the properties of sheaves to formally express this correspondence.
+We first describe the relationship between the graphic indexing space \gbase\ and the data indexing space \dbase\, which we propose is one where multiple graphic indexes map to one data index, and every index in the graphic space can be mapped to an index in the data space.
+We encode these expectations as the \textcolor{functor}{map} \vindexc, which we define to be a surjective continuous map
 \begin{equation}
@@
 \end{equation}
-such that every $\opensetg \subseteq \gbase$ must map to a corresponding set $\openset \subset \dbase$.
-The functor $\vindex$ is surjective such that for every point $\dbasepoint \in \openset$ there is a corresponding set of points $\{\gbasepoint | \gbasepoint \in  \vindexpre(k)\}$ for all $\gbasepoint in \opensetg$.
-\note{add little figure for xi showing it mapping from graphic index space back into data index space 
-}
-\note{Harder: if there exists a function between spaces, can construct the transport functors}
+between a graphic subspace $\opensetg \subseteq \gbase$ and data subspace $\openset \subseteq \dbase$.
+The functor $\vindex$ is surjective such that for every point $\dbasepoint \in \openset$ there is a corresponding set of points $\{\gbasepoint | \gbasepoint \in  \vindexpre(k)\}$ for all $\gbasepoint \in \opensetg$.
+We construct the map as going from graphic to data because that encodes the notion that every visual element traces back to the data in some way.
+We define $\vindex$ as a surjective map because it allows us to express that a union of graphic spaces $S_i$ maps to single data point $\dbasepoint$, which allows us to express visual representations of a single record that are the union of many primatives, discussed in \autoref{eq:artist:operator}, such as multipart glyphs (e.g boxplots) and combinations of plot types (e.g line with point markers).
+
+\note{add little figure for xi showing it mapping from graphic index space back into data index space}
+
 \subsubsection{Graphic Corresponding to Data}
+Since we have defined a function \vindex\ between two spaces $\dbase, \gbase$, we can construct functors that transport sheaves over each space to the other\cite{harder2008lectures}.
+This allows us to describe what data we expect at each graphic index location and what graphic is expected at each data index location.
+Transport functors compose the indexing map \vindex\ with the sheave map to say that a record \dsection\ at \dbasepoint\ is at all corrresponding \gbasepoint\ and that a function \gsection\ over one point \gbasepoint\ is the same function at all points $\gbasepoint \in \gbase$ that correspond to the same record index \dbasepoint.
+
 The pushforward (direct image) sheaf establishes which graphic generating function $\gsection$ corresponds to a point $\dbasepoint \in dbase$ in the data base space.   
-
 \begin{definition} Given a sheaf $\sheaf_{\gbase, \gtotal}$ on $\gbase$, the \textbf{pushforward} sheaf  $\vindexpush\sheaf_{\gbase, \gtotal}$ on $\dbase$ is defined as 
@@
  \begin{equation}  
-  \vindexpushc\gsectionc: \opensetc \rightarrow \vindexpushc \gtotalc\restriction_{\opensetc} \in \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} 
+  \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}  \ni \vindexpushc\gsectionc: \opensetc \rightarrow \vindexpushc \gtotalc\restriction_{\opensetc} 
 \end{equation}
@@
 The pullback (inverse image) sheaf establishes which data record  returned by $\dsection$ corresponds to a point $\gbasepoint \in \gbase$ in the graphic base space. 
-\begin{definition} Given a sheaf $\sheaf_{\dbase, \dtotal}$ on $\dbase$, the \textbf{pullback} sheaf $\vindexpullc\sheafc_{\dbasec, \dtotalc}$ on $\gbasec$ is defined as the sheaf associated to the presheaf
+\begin{definition} \cite{harder2008lectures} Given a sheaf $\sheaf_{\dbase, \dtotal}$ on $\dbase$, the \textbf{pullback} sheaf $\vindexpullc\sheafc_{\dbasec, \dtotalc}$ on $\gbasec$ is defined as the sheaf associated to the presheaf
   \begin{equation*}
@@
   \end{equation*}
-for $\vindex(\opensetgc) \in \dbase$ \cite{harder2008lectures}.
+for $\vindex(\opensetgc) \in \dbase$.
 \end{definition}
@@
 \begin{equation}
-  \vindexpullc\dsectionc: \opensetgc \rightarrow \vindexpullc \dtotalc\restriction_{\opensetgc} \in \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} 
+  \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}} \ni \vindexpullc\dsectionc: \opensetgc \rightarrow \vindexpullc \dtotalc\restriction_{\opensetgc}
 \end{equation}
@@
 
-Functors between sheaves are a way of expressing the bookkeeping involved in keeping track of which graphic section \gsection\ corresponds to which data section \dsection\.
+Functors between sheaves are a way of expressing the bookkeeping involved in keeping track of which graphic section \gsection\ corresponds to which data section \dsection.
 The $(\dbasepoint_i, \gbase_i)$ pairing expressed in \autoref{eq:atct:xi} establishes that there is a correspondence between sections evaluated over $\dbasepoint_i$ and $\gbase_i$.

### 2022-09-21 paper/main.tex
reworked sheaves/presheaves - add figure

@@
 Each tiny colored box is an approximation of the return value of the same section function $\gsection$ evaluated on different points $\gbasepoint \in \gbase$ in the base space.
-\note{add alpha and z channels}
+\note{add alpha and z channels and very faded out marker point}
   \label{fig:atct:fb:graphic}}
@@
 \end{equation*}
-go in the opposite direction.
-This is because any arbitrary function defined on an openset $\dsection\restriction_{\openset_2}$ must be definable on a subset of that space $\dsection\restriction_{\openset_1}$.
-The expectation that data containers implement inclusion and restriction morphisms means that, for example, a data container must ensure that continuous sections can be interpolated.
-
-While presheaves preserve the rules for ensuring continuity in the indexing space and restriction in the sets of data records, sheaves add on conditions for verifying that sections over subspaces are part of a cohesive section over the whole space.
+restricts the larger set to the smaller one such that  $\cgamma{\openset_2}{\dtotal\restriction\openset_2} \subseteq \cgamma{\openset_1}{\dtotal\restriction\openset_1}$. 
+\begin{figure}[h]
+  \includegraphics[width=\columnwidth]{presheaf.png}
+  \caption{\note{amend this so if possible the restriction loses a function}
+  \label{fig:atct:presheaf}}
+\end{figure}
+For example, in \autoref{fig:atct:presheaf}, the set of sections over the interval $\openset_2 = \left[0, 2\pi\right]$ the set of sections includes $sin, cos, constant_0 \subset \cgamma{\openset_2}{\dtotal\restriction\openset_2}$ function.
+One subset of this interval is $\openset_1 = \left[\frac{\pi}{2}, \frac{3\pi}{2}\right]$, and over this interval is a set of corresponding sections $\cgamma{\openset_1}{\dtotal\restriction\openset_1}$.
+The constraint that the sheaf functor enforces is that given the sheaves $\sheafc_{\openset_1}, \sheafc_{\openset_2}$, and an inclusion between the opensets, there must be a corresponding restriction between the sections such the set of sections over $\openset_1$ can only contain section functions that were over $\openset_2$.
+This expectation means that when data containers implement subsetting, which is a combination of inclusion and restriction, no new data is introduced.
+
+While presheaves preserve the rules for sets of sections, sheaves add on conditions for verifying that individual sections over subspaces are part of a cohesive section over the whole space. 
 \begin{definition}\cite{bakerEuclideanSpaceMathsSheaf,spanier1989algebraic} A \textbf{sheaf} is a presheaf that satisfies the following two axioms
@@
 The gluing axiom can also be used to generate the gluing rules used to construct non-trivial bundles from the set of trivial local section.
-Generally, the sheaf asserts the expectation that the data container is implemented such that the connectivity between the opensets (underlying data continuity) is preserved.
+Generally, the sheaf asserts the expectation that the data container is implemented such that the connectivity between the opensets (indexing subspaces) is preserved.
 

### 2022-10-13 paper/main.tex
presheaf diagram and nlab citations

--- 2022-09-20 paper/main.tex
@@
  \begin{definition}
-$f$ is \textbf{equivariant} when $f(g(x)) = g(f(x))$ for all $g$ in $G$ and for all $x$ in $X$\cite{nlab:equivariant}.
+$f$ is \textbf{equivariant} when $f(g(x)) = g(f(x))$ for all $g$ in $G$ and for all $x$ in $X$\cite{nlab:equivariant, pittsNominalSetsNames2013}.
  \end{definition}
@@
   \label{eq:atct:sheaf:restriction}
-  \iota^*: \cgamma{\opensetc_2}{\dtotalc_{\opensetc_2}} \rightarrow \cgamma{\opensetc_1}{\dtotalc_{\openset_c1}}
+  \iota^*: \cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}} \rightarrow \cgamma{\opensetc_1}{\dtotalc\restriction_{\openset_c1}}
 \end{equation*}
-go in the opposite direction.
-This is because any arbitrary function defined on an openset $\dsection\restriction_{\openset_2}$ must be definable on a subset of that space $\dsection\restriction_{\openset_1}$.
-The expectation that data containers implement inclusion and restriction morphisms means that, for example, a data container must ensure that continuous sections can be interpolated.
-
-While presheaves preserve the rules for ensuring continuity in the indexing space and restriction in the sets of data records, sheaves add on conditions for verifying that sections over subspaces are part of a cohesive section over the whole space.
+restricts the larger set to the smaller one such that all functions that are continuous over a space must be continuous over a subspace $\Gamma_2 \subseteq \Gamma_1$, where $\Gamma_{i}\coloneqq\Gamma(\openset_{i}, \dtotal\restriction_{\openset_{i}})$.
+
+For example, lets define presheaves $\sheaf_1, \sheaf_2$. These are maps from intervals $\openset_1, \openset_2$ to a set of functions $\Gamma_1, \Gamma_2$ that are continuous over that interval: 
+\begin{equation*}
+\begin{tikzcd}
+  \textcolor{set}{\Gamma_2 = \textcolor{set}{\left\{\begin{array}{c} 
+      \textcolor{section}{constant}\\ 
+      \textcolor{section}{sin}\\
+      \textcolor{section}{cos}\\
+    \end{array}\right\}}}               
+  \arrow[rr, "\iota^*", color=set] &  & 
+  \textcolor{set}{\Gamma_1 = \left\{\begin{array}{c} 
+    \textcolor{section}{constant}\\ 
+    \textcolor{section}{sin}\\
+    \textcolor{section}{cos}\\
+    \textcolor{section}{tan}\\
+  \end{array}\right\}} \\
+  \textcolor{base}{\openset_2 = \left(0,1\right)} 
+  \arrow[u, "\sheafc_{1}", maps to, color=sheaf, shift right=4.5]
+  &  & 
+  \textcolor{base}{\openset_1=\left(\frac{\pi}{2}, \frac{3\pi}{2}\right)} 
+  \arrow[ll, "\iota"', color=base] 
+  \arrow[u, "\sheafc_{2}"', maps to, color=sheaf, shift right=4.5]
+  \end{tikzcd}
+\end{equation*}
+The constraints of a presheaf functor are that since the $constant, sin, cos$ functions are defined over the interval $\left[0,1\right]$, these functions must also be continuous over the sub-interval $\left(\frac{\pi}{2}, \frac{3\pi}{2}\right)$; therefore the sections in $\Gamma_{2}$ must also be included in the set of sections over the subspace $\Gamma_{1}$.
+The generalization of this constraint is that data structures that contain continuous functions must support interpolating them over arbitrarily small subspaces.
+
+While presheaves preserve the rules for sets of sections, sheaves add on conditions for gluing individual sections over subspaces into cohesive sections over the whole space. 
 \begin{definition}\cite{bakerEuclideanSpaceMathsSheaf,spanier1989algebraic} A \textbf{sheaf} is a presheaf that satisfies the following two axioms
@@
 The locality axiom asserts that the glued section function is equivalent to a function over the union if they evaluate to the same values.
-The gluing axiom can also be used to generate the gluing rules used to construct non-trivial bundles from the set of trivial local section.
-Generally, the sheaf asserts the expectation that the data container is implemented such that the connectivity between the opensets (underlying data continuity) is preserved.
+The gluing axiom can also be used to generate the gluing rules used to construct non-trivial bundles from the set of trivial local sections.
+Generally, the sheaf asserts the expectation that the data container is implemented such that the connectivity between the opensets (indexing subspaces) is preserved.
 
@@
 
-\subsection{Data and Graphic Correspondence}
-\note{index correspondence}
+\subsection{Data Index and Graphic Index Correspondence}
 \label{sec:atct:xi}
@@
 The functor $\vindex$ is surjective such that for every point $\dbasepoint \in \openset$ there is a corresponding set of points $\{\gbasepoint | \gbasepoint \in  \vindexpre(k)\}$ for all $\gbasepoint \in \opensetg$.
+
 We construct the map as going from graphic to data because that encodes the notion that every visual element traces back to the data in some way.
-We define $\vindex$ as a surjective map because it allows us to express that a union of graphic spaces $S_i$ maps to single data point $\dbasepoint$, which allows us to express visual representations of a single record that are the union of many primatives, discussed in \autoref{eq:artist:operator}, such as multipart glyphs (e.g boxplots) and combinations of plot types (e.g line with point markers).
-
-\note{add little figure for xi showing it mapping from graphic index space back into data index space}
-
-\subsubsection{Graphic Corresponding to Data}
-Since we have defined a function \vindex\ between two spaces $\dbase, \gbase$, we can construct functors that transport sheaves over each space to the other\cite{harder2008lectures}.
+As exemplified in \autoref{fig:atct:morphisms:sheaf}, we define $\vindex$ as a surjective map because it allows us to express that a union of graphic spaces $S_i$ maps to single data point $\dbasepoint$, which allows us to express visual representations of a single record that are the union of many primatives, discussed in \autoref{eq:artist:operator}, such as multipart glyphs (e.g boxplots) and combinations of plot types (e.g line with point markers).
+
+\subsubsection{Data and Graphic Correspondence}
+Since we have defined a function \vindex\ between two spaces $\dbase, \gbase$, we can then construct functors that transport sheaves over each space to the other\cite{harder2008lectures}.
 This allows us to describe what data we expect at each graphic index location and what graphic is expected at each data index location.
@@
 
+\paragraph{\textbf{Graphic Corresponding to Data}}
 The pushforward (direct image) sheaf establishes which graphic generating function $\gsection$ corresponds to a point $\dbasepoint \in dbase$ in the data base space.   
@@
 
-\subsubsection{ Data Corresponding to Graphic}
+\paragraph{\textbf{Data Corresponding to Graphic}}
 The pullback (inverse image) sheaf establishes which data record  returned by $\dsection$ corresponds to a point $\gbasepoint \in \gbase$ in the graphic base space. 
@@
 \begin{figure}[h!]
-  \includegraphics*[width=1\columnwidth]{xi_scatter.png}
-  \caption{The functors $\vindex, \vindexpull, \vindexpush$ establish the correspondence between a data section \dsection\ and a graphic generation function \gsection.
-For every data record $\dsection(\dbasepoint)$, there is a graphic generating function $\vindexpush\gsection = \gsection|_{\vindexpre\dbasepoint}$.
-This curried function is here represented as the visual parameters associated with that function to demonstrate that it generates the visual element for that data record.
-Every data section can also be copied over to the graphic space such that each piece of the (idealized in infinite resolution) graphic $\gsection(\dbasepoint)$ has a corresponding data value $\vindexpull\dsection(\dbasepoint)$.
-  \label{fig:atct:morphisms:sheaf}} \note{possibly reorient and expand this diagram to illustrate eq 55}
+  \includegraphics*[width=1\columnwidth]{xi_sin.png}
+  \caption{The data consists of the $sin$ and $cos$ functions over a unit circle base space.
+We choose to visualize this as a circle and two line plots.
+The indexing function \vindex\ bookkeeps which parts of the circle and each curve correspond to each point on the unit circle.
+The pushforward $\vindexpush$ matches each point in the data space to the specification of the graphic at that point, while the pullback $\vindexpull$ matches each point in the graphic space to the data over that point.
+\label{fig:atct:morphisms:sheaf}}
 \end{figure}
@@
 This allows us to construct graphic specifications for each data index $\vindexpush\gsection$ and retrieve the data $\vindexpull\dsection$ for any graphic section generating any piece of a graphic.
-In \autoref{fig:atct:morphisms:sheaf}, the sections return rows in the table from \autoref{fig:related-work:continuity:ktypes}.
-Each section $\dsection(\dbasepoint_{i})$ has a corresponding visual element produced by evaluating the graphic section on the graphic base space $\gsection|_{\gbase_{i}}$.
-In \autoref{fig:atct:morphisms:sheaf} the output of $\gsection$ as pixels in an idealized prerendered space is approximated as the circles.
-Each graphic section $\gsection(\gbasepoint), \gbasepoint \in \gbase_{i}$ that is a piece of the graphic has a corresponding pulled back data value $\vindexpull\dsectionc(\gbasepoint)$.
-The pullback copies the data points such that $\vindexpull\dsectionc|_{\gbase_i} = \dsection(\dbasepoint_i)$.
-Each visual element $\gsection|_{\gbase_i}$ is generated from a single record $\dsection(\dbasepoint_i)$; therefore the pushforward is the curried graphic section $\gsection|_{\vindexpre(\dbasepoint_i)}$.
+In \autoref{fig:atct:morphisms:sheaf}, the visualization is a graphic representation of a unit circle and the sin and cosine curves on that interval.
+The index lookup $\vindex$ describes which parts of the circle and curves are generated from which points on the unit circle.
+Given this correspondance, the pullback $\vindexpull$ looks up which values are being represented in a given part of the graphic.
+The pushforward $\vindexpush$ describes how a graphic is supposed to look for each point in the data space.
 The graphic parameterized in \autoref{fig:atct:morphisms:sheaf} is intended as an approximation of $\vindexpush\gsection$.
@@
 Interactive tooltips assume that there is a map between the graphics rendered on screen and the data such that a glyph maps back to a record.
-The functors $(\vindex, \vindexpush, \vindexpull)$ codify the expectation of these mappings between data and graphic space.
+The expectation that these mappings between data and graphic space exist are encapsulated in the functors $(\vindex, \vindexpush, \vindexpull)$.
 
@@
 In \autoref{sec:atct:sheaves}, the functors $\vindex, \vindexpull, \vindexpushc$ are introduced to describe the expected relationship between screen and data base spaces.
-In the construction of the artist, the functor $\vindex$ is defined such that the data base space \dbase\ is a deformation retraction\cite{nlab:deformation_retract} of the graphic space $\gbase$.
+In the construction of the artist, the functor $\vindex$ is defined such that the data base space \dbase\ is a deformation retraction\cite{nlab:deformation_retract,hatcherAlgebraicTopology2002} of the graphic space $\gbase$.
 This means that there is a continuous surjective mapping from every point $\dbasepoint\in\dbase$ to a point $\dbasepoint \in dbase$.

### 2022-10-13 paper/main.tex
wiki cites

@@
 All fibers in a bundle are isomorphic such that $\dfiber \cong \dfiber_{\dbasepoint}$ for all points $\dbasepoint \in \dbase$.
-  \item there is an open neighborhood $\openset_{\dbasepoint}$ surrounding each point in the base space $\dbasepoint \in \openset_{\dbasepoint} \subset \dbase$ such that the \textcolor{total}{total space} \dtotalc\ over the neighborhood, denoted $\dtotal\restriction\openset$\footnote{The symbol $\restriction$ is the restriction operator\cite{RestrictionMathematics2022}.
-For example $\pi^{-1}\restriction\openset$ is the inverse function $\pi^{-1}$ only defined on the points in $\openset$}, is locally trivial.
+  \item there is an open neighborhood $\openset_{\dbasepoint}$ surrounding each point in the base space $\dbasepoint \in \openset_{\dbasepoint} \subset \dbase$ such that the \textcolor{total}{total space} \dtotalc\ over the neighborhood, denoted $\dtotal\restriction\openset$\footnote{The symbol $\restriction$ is the restriction operator\cite{RestrictionMathematics2022} defined in \texttt{amssymb}.
+For example $\pi^{-1}\restriction\openset$ is the inverse function $\pi^{-1}$ defined only on the points in $\openset$.
+Since $\pi^{-1}(\dbase)=\dtotal$, we use the shorthand $\dtotal\restriction_{\openset} \coloneqq \pi^{-1}\restriction_{\openset}$}, is locally trivial.
 The condition for local triviality is that $\dtotal\restriction_{\openset} = \openset \times \dfiber = \pi^{-1}\restriction_{\openset}$
@@
 
-Each section of a sheaf over a point returns a single record in the fiber. The sheaf over an open set $\openset$ surrounding a point $\dbasepoint$ is called a \textit{stalk}\cite{StalkSheaf2019}
+Each section of a sheaf over a point returns a single record in the fiber.
+The sheaf over an open set $\openset$ surrounding a point $\dbasepoint$ is called a \textit{stalk}\cite{StalkSheaf2019,harder2008lectures}
 \begin{equation}

### 2022-10-26 paper/main.tex
record transformer constraints

@@
 
-\begin{definition}\cite{nlab:action, grimaldiDiscreteCombinatorialMathematics2006}
+\begin{definition}\label{def:related-work:action}\cite{nlab:action, grimaldiDiscreteCombinatorialMathematics2006}
   An \textcolor{action}{\textbf{action}} \footnote{Throughout this paper, we augment the math with color to group conceptually similar objects and functions\cite{headMathAugmentationHow2022}, for example \textcolor{action}{actions} and \textcolor{action}{sets of actions}.
@@
 
-To build on the rich typing system provided by Spivak, we define the \textcolor{fiber}{fiber space} \dfiberc\ to be the set of all data records
+To build on the rich typing system provided by Spivak, we define the \textcolor{fiber}{fiber space} \dfiberc\ to be the space of all possible data records
 \begin{equation}
@@
 Each section of a sheaf over a point returns a single record in the fiber.
-The sheaf over an open set $\openset$ surrounding a point $\dbasepoint$ is called a \textit{stalk}\cite{StalkSheaf2019,harder2008lectures}
+The sheaf over an open set $\openset$ surrounding a point $\dbasepoint$ is called a \textit{stalk}\cite{StalkSheaf2019, harder2008lectures}
 \begin{equation}
@@
 The index lookup $\vindex$ describes which parts of the circle and curves are generated from which points on the unit circle.
-Given this correspondance, the pullback $\vindexpull$ looks up which values are being represented in a given part of the graphic.
-The pushforward $\vindexpush$ describes how a graphic is supposed to look for each point in the data space.
-The graphic parameterized in \autoref{fig:atct:morphisms:sheaf} is intended as an approximation of $\vindexpush\gsection$.
-
-Visualization specification languages such as vega \cite{heerDeclarative2010} and svg \cite{quintScalable2003} assume that there is a graphic  specification for every point.
-Interactive tooltips assume that there is a map between the graphics rendered on screen and the data such that a glyph maps back to a record.
-The expectation that these mappings between data and graphic space exist are encapsulated in the functors $(\vindex, \vindexpush, \vindexpull)$.
+Given this correspondance, the pullback $\vindexpull\dsection$ looks up which values are being represented in a given part of the graphic.
+This type of lookup is critical for interactive techniques such as brushing, linking, and tooltips\cite{beckerBrushingScatterplots1987}.
+The pushforward $\vindexpush\gsection$ describes how a graphic is supposed to look for each point in the data space.
+The graphic parameterization in \autoref{fig:atct:morphisms:sheaf} is intended as an approximation of $\vindexpush\gsection$ and is akin to declarative visualization specs such as vega \cite{satyanarayanDeclarativeInteractionDesign2014} and svg \cite{quintScalable2003}.
+These specs and $\vindexpushc\gsection$ provide a renderer independent way of describing the graphic and are therefore useful for standardizing internal representation of the graphic and serializing the graphic for portability.
 
@@
   \label{eq:atct:morphism:base}
- \dfunch: \openset^{\prime}\rightarrow \openset
+ \dfunchc: \dbasepointc^{\prime}\mapsto \dbasepointc
  \end{equation}
@@
   \label{eq:atct:morphism:basepull}
-  \dfuncpull \dsection \restriction_{\openset^{\prime}}: \dsection\restriction_{\openset^{\prime}} \mapsto \dsection \restriction_{\openset^{\prime} \circ \dfunch} 
+  \dfuncpullc \dsectionc \restriction_{\opensetc^{\prime}}: \dsectionc\restriction_{\opensetc^{\prime}} \mapsto \dsectionc \restriction_{\opensetc^{\prime} \circ \dfunchc} 
 \end{equation}
@@
 where $\dsection, \dsection^{\prime} \in \Gamma(\openset^{\prime}, \dfuncpull\dtotal\restriction_{\openset^{\prime}})$.
-Since there are many different fiber transformations,  there is not a uniform way of expressing how this morphism acts on sections unlike in \autoref{eq:atct:morphism:verify_base}.
+Since there are many different fiber transformations,  there is not a uniform way of expressing how this morphism acts on sections unlike in \autoref{eq:atct:morphism:verify_base}.\note{This is where we reintroduce the equivariance and homomorphism definitions from the introduction}
+
 Examples of a fiber transformation are actions such as partial ordering for ranking purpose \cite{bruggemannRankingPrioritizationMultiindicator2011} and the permutation, ordering, translation, and scaling codified as Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946, leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}.
@@
     &\textcolor{base}{point}/\textcolor{base}{openset}/\textcolor{base}{base space} & \textcolor{fiber}{fiber space} & \textcolor{total}{total space}\\
-     &  location/subset/indicies & fields &  dataset type\\
-    \hline
-   Data & $\dbasepointc \in \opensetc \subseteq \dbasec$ & \dfiberc & \dtotalc\\
-   Visual & $\dbasepointc \in \opensetc \subseteq \dbasec$  & \vfiberc & \vtotalc\\
-   Graphic & $\gbasepointc \in \opensetgc \subseteq \gbasec$ & \gfiberc & \gtotalc\\
+     &  location/subset/indicies & record/fields &  dataset type\\
+    \hline
+   Data & $\dbasepointc \in \opensetc \subseteq \dbasec$ & $\delementc \in \dfiberc$ & \dtotalc\\
+   Visual & $\dbasepointc \in \opensetc \subseteq \dbasec$  & $\velementc \in \vfiberc$ & \vtotalc\\
+   Graphic & $\gbasepointc \in \opensetgc \subseteq \gbasec$ & $\gelementc \in \gfiberc$ & \gtotalc\\
    \hline
@@
   \end{tabular}
-  \caption{Functors between graphic and data base spaces  \autoref{sec:atct:xi}}
+  \caption{Functors between graphic and data indexing spaces  \autoref{sec:atct:xi}}
   \label{tab:appendix:summary:transport}
@@
 \centering
-{\renewcommand{\arraystretch}{1.2}
-\begin{tabular}{|c|}
-
+{\renewcommand{\arraystretch}{1.5}
+\begin{tabular}{|r|l|l|}
+  \hline
+  changes & function & constraints, for all $\dbasepointc \in \opensetc$ \\
+  \hline
+  index & $\dfunchc: \opensetc \rightarrow \opensetc^{\prime}$ &   $\dsectionc(\dbasepointc) = \dsectionc(\dfunchc(\dbasepointc^{\prime})) = \dfuncpullc \dsectionc(\dbasepointc^{\prime})$ \\
+  & & \\
+  record  & $\dfunctc: \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc}}$ &  $\lim\limits_{x\rightarrow \dbasepointc}\dfunctc(\dsectionc(x)) = \dfunctc(\dsectionc(\dbasepointc))$
+  \\
+  & $\textcolor{white}{\dfunct: } \rightarrow \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc}}$ &    
+  \\
+  & $\dfunctc: \dfiberc \rightarrow \dfiberc$ &  $\dfunctc(\dsectionc(\dbasepointc)) \in \dfiberc$  \\ 
+  & &  $\dfunctc(id_{\dfiberc}(\dsectionc(\dbasepointc))) = id_{\dfiberc}(\dfunctc(\dsectionc(\dbasepointc)))$\\
+  & & $\dfunctc(\dfunctc(\dsectionc(\dbasepointc))) = (\dfunctc\circ\dfunctc)(\dsectionc(\dbasepointc))$\\
+  \hline
 \end{tabular}
-\note{changing locations, moving records to new locations, changing records}
 \caption{Functions $\dfunc=(\dfunchc, \dfunctc)$ for modifying data records.
@@
 }
+\end{table}
+
+\begin{table}[h!]
+  \renewcommand{\arraystretch}{1.5}
+  \begin{tabular}{|lll|}\hline
+      scale & property & scale specific constraints \\ \hline
+      nominal & permutation &  $\dsectionc(\dbasepointc_1) \neq \dsectionc(\dbasepointc_2) \implies \dfunctc (\dsectionc(\dbasepointc_1)) \neq\dfunctc(\dsectionc(\dbasepointc_2)$)\\
+      ordinal &  monotonic &  $\dsectionc(\dbasepointc_1) \leq \dsectionc(\dbasepointc_2) \implies \dfunctc (\dsectionc(\dbasepointc_1)) \leq \dfunctc(\dsectionc(\dbasepointc_2)$) \\
+      interval &  translation &  $\dfunctc(\dsectionc(\dbasepointc) + C) = \dfunctc(\dsectionc(\dbasepointc)) + C$ \\
+      ratio &  scaling &  $\dfunctc(\dsectionc(\dbasepointc) * C) = \dfunctc(\dsectionc(\dbasepointc))*C $\\ \hline
+  \end{tabular}
+  \caption{The record transformer $\dfunctc$ is defined by the general constraints listed in \autoref{tab:appendix:summary:datamod} and it must respect any properties that are specific to the record field.
+For example, this table lists the properties of the Steven's measurement scale types\cite{stevensTheoryScalesMeasurement1946}}
+  \label{tab:appendix:summary:datatypes}
 \end{table}

### 2022-10-27 paper/main.tex
concretizing record transform

@@
 We propose that these data types can be formally specified as objects of a category.
-\begin{definition} 
+\begin{definition}\label{def:atct:category}
    An \textbf{category} $\mathcal{C}$ consists of the following \textit{data}: 
@@
 such that $\dsection|_{\openset} = \dfuncpull\dsection|_{\opensetc^{\prime}}$ because $\dsection|_{\openset} = \dsection|_{\dfunch(\opensetc^{\prime})}$.
-This means that the base space transformation \dfunch\ does not change the data values at a given point
+This means that the base space transformation  $\dfunch(\dbasepoint^{\prime}) = \dfunch(\dbasepoint)$ such that
 \begin{equation}
   \label{eq:atct:morphism:verify_base}
-  \dsection(\dbase) = \dfuncpull\dsection(\dbasepoint^{\prime}) = \dsection(\dfunch(\dbasepoint^{\prime}))
-\end{equation} 
-where $\dfunch(\dbasepoint^{\prime}) = \dfunch(\dbasepoint)$.
+  \dsectionc(\dbasec) = \dfuncpullc\dsectionc(\dbasepointc^{\prime}) = \dsectionc(\dfunchc(\dbasepointc^{\prime}))
+\end{equation} 
+which means that the index of the record changes from $\dbasepoint$ to $\dbasepoint^{\prime}$ but the values in the record are unmodified.
 
 \paragraph{\textbf{Records}}
-\noindent
-As introduced in \autoref{eq:artist:sheaves:monoid_morphism}, the fiber transformation $\dfunct: \dfuncpull \dtotal_{\dbasepoint^{\prime}} \rightarrow \dfuncpull \dtotal_{\dbasepoint^{\prime}}$ is a monoid action on the fiber $\dfunct \in Hom(\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}},\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}})$ restricted to a point $\dbasepoint^{\prime} \in \openset^{\prime}$.
-The fiber transformation is a change in section
+As introduced in \autoref{eq:artist:sheaves:monoid_morphism}, the fiber transformation $\dfunctc$ is a change in section 
 \begin{equation}
   \label{eq:atct:morphism:fiber}
-  \dfunct: \dfuncpull \dsection \restriction_{\openset^{\prime}} \mapsto \dfuncpull \dsection^{\prime} \restriction_{\openset}
+  \dfunctc: \dfuncpullc \dsectionc \restriction_{\opensetc^{\prime}} \mapsto \dfuncpullc \dsectionc^{\prime} \restriction_{\opensetc}
 \end{equation}
 where $\dsection, \dsection^{\prime} \in \Gamma(\openset^{\prime}, \dfuncpull\dtotal\restriction_{\openset^{\prime}})$.
-Since there are many different fiber transformations,  there is not a uniform way of expressing how this morphism acts on sections unlike in \autoref{eq:atct:morphism:verify_base}.\note{This is where we reintroduce the equivariance and homomorphism definitions from the introduction}
-
-Examples of a fiber transformation are actions such as partial ordering for ranking purpose \cite{bruggemannRankingPrioritizationMultiindicator2011} and the permutation, ordering, translation, and scaling codified as Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946, leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}.
+Since $\dfunct$ maps from one continuous function to another, it must itself be continuous such that
+\begin{equation}
+  \label{eq:atct:morphism:fiber:continuity}
+   \lim\limits_{x \rightarrow \dbasepointc^{\prime}}\dfunctc(\dfuncpullc\dsectionc(x)) =  \dfunctc(\dfuncpullc\dsectionc(\dbasepointc^{\prime}))
+\end{equation}
+As mentioned in \autoref{sec:atct:fb:fiber}, $\dfunct$ is also  
+a morphism on the fiber category $\dfunct \in Hom(\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}},\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}})$ restricted to a point $\dbasepoint^{\prime} \in \openset^{\prime}$.
+This means $\dfunct$ has to statisfy the properties of a morphism (\autoref{def:atct:category})
+\begin{itemize}
+  \item \textit{closed}: $\dfunctc(\dfuncpullc\dsectionc(\dbasepointc^{\prime})) \in \dfiberc$
+  \item \textit{unitality}: $\dfunctc(id_{\dfiberc}(\dfuncpullc\dsectionc(\dbasepointc^{\prime}))) = id_{\dfiberc}(\dfunctc(\dfuncpullc\dsectionc(\dbasepointc^{\prime})))$
+  \item \textit{composition and associativity}: \\
+  $\dfunctc(\dfunctc(\dfuncpullc\dsectionc(\dbasepointc^{\prime}))) = (\dfunctc\circ\dfunctc)(\dfuncpullc\dsectionc(\dbasepointc^{\prime}))$
+\end{itemize}
+
+Additionally, each fiber $\dfiber_{type}$ may have features in addition to the set of possible records. For example, monoids and groups are defined as sets with operators 
+\note{clarify if $\dfunctc$ are actions on F or encode the operators/extra structure on F}
+
+the monoids that underly partial ordering for ranking purpose \cite{bruggemannRankingPrioritizationMultiindicator2011} and the groups that define Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946, leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}.
 
@@
   \begin{tabular}{|lll|}\hline
-      scale & property & scale specific constraints \\ \hline
-      nominal & permutation &  $\dsectionc(\dbasepointc_1) \neq \dsectionc(\dbasepointc_2) \implies \dfunctc (\dsectionc(\dbasepointc_1)) \neq\dfunctc(\dsectionc(\dbasepointc_2)$)\\
-      ordinal &  monotonic &  $\dsectionc(\dbasepointc_1) \leq \dsectionc(\dbasepointc_2) \implies \dfunctc (\dsectionc(\dbasepointc_1)) \leq \dfunctc(\dsectionc(\dbasepointc_2)$) \\
-      interval &  translation &  $\dfunctc(\dsectionc(\dbasepointc) + C) = \dfunctc(\dsectionc(\dbasepointc)) + C$ \\
-      ratio &  scaling &  $\dfunctc(\dsectionc(\dbasepointc) * C) = \dfunctc(\dsectionc(\dbasepointc))*C $\\ \hline
+      scale & operators & sample constraint \\ \hline
+      nominal & $=,\neq$ &  $\dsectionc(\dbasepointc_1) \neq \dsectionc(\dbasepointc_2)\implies \dfunctc (\dsectionc(\dbasepointc_1)) \neq\dfunctc(\dsectionc(\dbasepointc_2))$\\ 
+      ordinal & $<, \leq, \geq, >$ &  $\dsectionc(\dbasepointc_1) \leq \dsectionc(\dbasepointc_2) \implies \dfunctc (\dsectionc(\dbasepointc_1)) \leq \dfunctc(\dsectionc(\dbasepointc_2)$) \\
+      interval & $+, -$ &  $\dfunctc(\dsectionc(\dbasepointc) + C) = \dfunctc(\dsectionc(\dbasepointc)) + C$ \\
+      ratio & $*,/$ &  $\dfunctc(\dsectionc(\dbasepointc) * C) = \dfunctc(\dsectionc(\dbasepointc))*C $\\ \hline
   \end{tabular}
-  \caption{The record transformer $\dfunctc$ is defined by the general constraints listed in \autoref{tab:appendix:summary:datamod} and it must respect any properties that are specific to the record field.
-For example, this table lists the properties of the Steven's measurement scale types\cite{stevensTheoryScalesMeasurement1946}}
+  \caption{The record transformer $\dfunctc$ must satisfy the constraints listed in \autoref{tab:appendix:summary:datamod} and $\dfunctc$ must also respect the mathamatical structure of $\dfiber$.
+This table lists examples of $\dfunct$ preserving one of the binary operators that are part of the definition of each of the Steven's measurement scale types\cite{stevensTheoryScalesMeasurement1946}}.
+A full implementation would ensure that all operators that are defined as part of of $\dfiber$ are preserved.
   \label{tab:appendix:summary:datatypes}

### 2022-10-28 paper/main.tex
we has operators

@@
 Adopting Spivak's fiber bundle construction of types allows our model to reuse types so long as the field names are distinct and that field values can be accesed by field name,  since those are sections on $\mathcal{U}_{\sigma}$.
-Furthermore, since domains $\mathscr{U}_{{\sigma}}$ of types are a topological space, multi-dimensional fields can be encoded in the same manner as single dimensional fields and fields can have different names but the same type.
+Furthermore, since domains $\mathscr{U}_{{\sigma}}$ of types are a mathematical space, multi-dimensional fields can be encoded in the same manner as single dimensional fields and fields can have different names but the same type.
 
@@
 The fiber category has a single object $\dfiberc$ of an arbitrary type and morphisms on the fiber object $\dfunctc \in Hom(\dfiberc, \dfiberc)$.
-The fiber category $\mathcal{\dfiberc}$ is a monoidal category, meaning that it is a category equipped with a bifunctor $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$.
-The bifunctor provides a method for combining fibers, thereby allowing us to express fields that contain multityped values.
+We can also equip the category with any operators or relations that are part of he mathematical structure of the field type.
+For example we can equip the category with a comparison operator, which is part of the definition of the monoidal structure of a partially ordered ranking variable \cite{bruggemannRankingPrioritizationMultiindicator2011} or the group structure of Steven's ordinal measurment scale \cite{stevensTheoryScalesMeasurement1946, leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}.
+Steven's other scales are summarized in \autoref{tab:appendix:summary:stevens}.
+
+The fiber category $\mathcal{\dfiberc}$ is also equipped with a bifunctor because it is a monoidal category and this functor provides a method for combining fiber types.
+The bifunctor $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$ allows us to express fields that contain multityped values.
 For example, wind can be represented as two fields $\dfiber_{speed} \times \dfiber_{direction}$ or a composite fiber field $\dfiber_{speed} \otimes \dfiber_{direction} = \dfiber_{wind}$.
@@
 \label{sec:artist:equivariant:data}
+\note{something something rotation etc}
 We separate data transformations into two components, transformations on the base space $(\dfunchc, \dfuncpullc)$ and transformations on the fiber space $\dfunctc$.
@@
 
-Additionally, each fiber $\dfiber_{type}$ may have features in addition to the set of possible records. For example, monoids and groups are defined as sets with operators 
-\note{clarify if $\dfunctc$ are actions on F or encode the operators/extra structure on F}
-
-the monoids that underly partial ordering for ranking purpose \cite{bruggemannRankingPrioritizationMultiindicator2011} and the groups that define Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946, leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}.
+Additionally, $\dfunctc$ must preserve any features of $\dfiber$, such as operators that are defined as part of the structure of $\dfiber$.
+Examples of testing that $\dfunctc$ preserves the operations, and therefore structure, of the Steven's measurement scales are shown in \autoref{tab:appendix:summary:stevens}.
+We do not provide a general rule here because these constraints are defined with respect to how specific properties of the mathematical structure of individual fields $\dfiber$ are expected to be preserved rather than as a general consequence of $\dfunctc$ being a section map and morphism of the category.
 
@@
 A full implementation would ensure that all operators that are defined as part of of $\dfiber$ are preserved.
-  \label{tab:appendix:summary:datatypes}
+  \label{tab:appendix:summary:stevens}
 \end{table}
@@
   \hline 
-  \textcolor{functor}{lookup} & $\vindex: \gbase\rightarrow \dbase$  &\\
+  \textcolor{functor}{lookup} & $\vindex: \gbase\rightarrow \dbase$  &\\ add [0,1] etc...
   \hline

### 2023-01-19 paper/main.tex
we has references
link to prototype
clean up

@@
 \usetikzlibrary{cd} % commutative diagrams
+\usepackage{pgfplots}
+\DeclareUnicodeCharacter{2212}{−}
+\usepgfplotslibrary{groupplots,dateplot}
+\usetikzlibrary{patterns,shapes.arrows}
+
+\pgfplotsset{compat=newest}
 \newtheorem{axiom}{Axiom}
@@
 On the other hand, domain specific libraries are designed with the assumption of continuities that are common in the domain \cite{HeerSoftware2006}, and therefore can somewhat restrict the interface to choices that that are appropriate for the domain.
-For example, a tabular topological structure of discret rows, as illustrated in \autoref{fig:related-work:continuity:ktypes}, is assumed by
+For example, a tabular topological structure of discrete rows, as illustrated in \autoref{fig:related-work:continuity:ktypes}, is assumed by
 A Presentation Tool\cite{mackinlayAutomatingDesignGraphical1986, mackinlayAutomatingDesignGraphical1986} and grammar of graphics\cite{wilkinsonGrammarGraphics2005} and the ggplot\cite{wickhamGgplot2ElegantGraphics2016}, vega\cite{satyanarayanDeclarativeInteractionDesign2014}, and altair\cite{vanderplasAltairInteractiveStatistical2018} libraries built on these frameworks.
@@
 
-\begin{definition}
+\begin{definition}\label{def:fiber_bundle}
    A \textbf{fiber bundle} $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and  bundle projection map $\pi: \dtotalc \rightarrow \dbasec$ \cite{FiberBundle2020,spanier1989algebraic}.
@@
 Local triviality $\dtotal\restriction_{\openset_{k}} = \openset_{k}\times \dfiber$ means that there is always a subset of the base space over which the fibers are identical. 
-A bundle can be subsetted into regions such that the bundled can be reconstructed by gluing the regions back together via transition maps on the border of each region.
-These borders are fibers, and in a trivial bundle transition maps are identity maps because fibers are identical $\dfiber = \dfiber_{k}$ for all points $\dbasepoint \in dbase$.
-In a non-trivial bundle fibers are isomorphic $\dfiber \simeq \dfiber_{\dbasepoint}$ for all points $\dbasepoint \in \dbase$, which means that there is at least one transition map that is not an identity \cite{hatcherAlgebraicTopology2002}.
-
 \begin{figure}[h!]
-  \includegraphics[width=\columnwidth]{trivnon.png}
-  \caption{A fiber bundle with a cylinderical total space, which is a trivial bundle, and bundle with a mobius band total space, which is non-trivial, can be covered by the same set of local trivializations.
-For example, they can both be decomposed into two planes over two intervals $\left(-\epsilon,\pi+\epsilon\right), \left(\pi-\epsilon, 2\pi+\epsilon\right)$.
-The transition maps that transforms this set of local trivializations into a cylinder are two identity map aligning two \textit{up} fibers, while the transition maps that transform this set into a mobius bundle are an identity map and a map that aligns an \textit{up} fiber with a \textit{down} fiber.
-\note{in final may change this to something that's not half half to make the arbitraryness more clear/break it up into smaller pieces} \label{fig:atct:nontrivialbundle}}
+  \includegraphics[width=\columnwidth]{figures/local_trivalizations.png}
+  \caption{Two local trivializations, $\dtotal_{0}$ and $\dtotal_{1}$, of an arbitrary bundle \dtotal.} 
 \end{figure}
 
-As shown in \autoref{fig:atct:nontrivialbundle}, a trivial bundle with a cylinderical total space and a non-trivial bundle with mobius band total spacee can be covered with the same local trivializations; the distinguishing factor between these bundles is that the transition maps on these trivializations differ.
-The fibers in the cylinder bundle align in the same direction, \textit{up}, such that the transition maps that assemble the local trivializations into the cylinderical total space are all identity maps alinging two \textit{up} fibers.
-In the mobius bundle, there is twist where the fibers align in different directions, shown in the inset in \autoref{fig:atct:nontrivialbundle}.
-This twist is not present in either local trivialization because, by definition, fibers need to be aligned in local trivializations; therefore the transition map that aligns the fiber at the overlap between $\left(-\epsilon, \epsilon\right)$ must account for this flip by aligning \textit{up} fibers \textit{down} fibers.
-In this example, the transition map on the $\left(\pi-\epsilon, \pi+\epsilon\right)$ is an identity map since the fibers are aligned in the same direction on that section of the mobius band.
-\note{add something in the figure highlighting this part of the cylinder and the mobius band, possibly add in a section}
+For example, in \autoref{fig:local_trivalizations}, the bundles $\dtotal_{0}=\dtotal\restriction_{[\epsilon^{-}, {\frac{2\pi}{5}}^{+}]}$ and $\dtotal_{1}=\dtotal\restriction_{[{\frac{2\pi}{5}}^{-}, \epsilon^{-}]}$ are local trivializations of an arbitrary bundle $\dtotal$.
+These local trivializations overlap on the intersections of the open subsets of each trivalization $[\epsilon^{-}, \epsilon^{+}]$ and $[{\frac{2\pi}{5}}^{-}, {\frac{2\pi}{5}}^{+}]$.
+A set of trivializations can be glued into a larger bundle by defining transition maps that describe how to align the fibers in the overlapping space such that the sections of the bundle remain continuous.
+\note{add transition map construction formula}
+
+\begin{figure}[h!]
+  \includegraphics[width=1\columwidth]{figures/cyl_mob.png}
+  \caption{\note{figure out short caption here}\label{fig:cyl_mob_bundles}}
+\end{figure}
+As demonstrated in \autoref{fig:cyl_mob_bundles} different types of total spaces can be decomposed into the same local trivializations; therefore the distinguishing factor between total spaces can be the transition maps for constructing the spaces from trivializations.
+
+In a \textit{trivial bundle}, the fibers are identical $\dfiber = \dfiber_{k}$ for all points $\dbasepoint \in dbase$ such that all the transtion maps are also identical.
+For example, in the cylinder case the fibers in the overlappping regions of the trivalizations point in the same direction $\uparrow$ to illustrate that they are equal $\dfiber_0\restriction_{\openset_1\cap\openset_2} = \dfiber_{1}\restriction_{\openset_1\cap\openset_2}$.
+Because the fibers point in the same direction, the  transition maps at both intersections map the fiber to itself $\uparrow\rightarrow\uparrow$.
+
+In a \textit{non-trivial} bundle fibers are isomorphic $\dfiber \simeq \dfiber_{\dbasepoint}$ for all points $\dbasepoint \in \dbase$ such that there is at least one transition map that is not an identity \cite{hatcherAlgebraicTopology2002}.
+For example, in the mobius band case, while $\dfiber_{0}\restriction_{[{\frac{2\pi}{5}}^{-}, {\frac{2\pi}{5}}^{+}]} = \dfiber_{0}\restriction_{[{\frac{2\pi}{5}}^{-}, {\frac{2\pi}{5}}^{+}]}$ such that the transition map is an identity map $\uparrow\rightarrow\uparrow$, the transition map must flip  $\uparrow \rightarrow \downarrow$ the fibers of $\dfiber_0\restriction_{[\epsilon^{-}, \epsilon^{+}]}$ in order for them to be equal to the fibers of $\dfiber_1\restriction_{[\epsilon^{-}, \epsilon^{+}]}$.
+\note{add summary sentence about power of triviality, probably pointing to it's power for abstraction}
 
@@
 \begin{equation}
-  \vindexpullc\dsectionc(\gbasepointc) = \dsectionc(\vindex(\gbasepointc)) = \dsectionc(\dbasepointc)
+  \vindexpullc\dsectionc(\gbasepointc) = \dsectionc(\vindexc(\gbasepointc)) = \dsectionc(\dbasepointc)
 \end{equation}
@@
 Instead, we define an output \textcolor{action}{verification} function $\extractmc$ that takes as input the section evaluated on all the graphic space associated with a point $\gsection_{\vindexpre\restriction_{\dbasepoint}}$ and returns the corresponding \textcolor{action}{measurable visual components} $\measurec_{k}$.
+\note{formall define M as a space of measurements }
 \begin{equation}
@@
 \subsection{Composing Artists}\label{sec:artist:operators}
+\note{addition: intersections mapped same, multiplication: fibers mapped same}
+\note{large big data glued together correctly}
 A common use of category theory in software engineering is the specification of modular components \cite{wielsManagementEvolvingSpecifications1998} such that we can build systems where the structure preserved by components is preserved in the composition of the components.
@@
 \label{sec:artist:addition}
-As illustrated in \autoref{eq:artist:operator}, data bundles can be combined by taking the disjoint  union of base spaces $\dbase^{a} \sqcup_{\dbase^c} \dbase^b$, where $\dbase^c$ is an overlap.
+As illustrated in \autoref{eq:artist:operator}, data bundles can be combined by taking the union of base spaces $\dbase^{a} \sqcup_{\dbase^c} \dbase^b$, where $\dbase^c$ is an overlap.
+\note{amalgamated union, or data bundles of the same type can be thought of as local trivializations of a larger space and glued togetehr as discussed in \autoref{sec:atct:fiber-bundles}}
+
 When the fibers of each bundle are isomorphic $\dfiber^{a} \simeq \dfiber^{b}$, which we denote as $\dtotal$, this is analogous to adding more records to a dataset.
@@
 \label{sec:artist:operator:multiplication}
+\note{fiber product vs cartesian product}
 As illustrated in \autoref{eq:artist:operator},  fibers that are a cartesian product of fiberspaces $\dfiber^{a} \times_{\dfiber^c} \dfiber^{b}$, where $\dfiber^c$ is any fiber that is present in both fibers, can be projected down into component fibers.
@@
 \label{sec:construction}
-
+\note{add a high level diagram Data->V->Screen}
+\note{add back in path of Q, use tikz backend to convert to pgf to then tweak}
 We propose that one way of constructing artist functions is to separate generating a visualization into an encoding stage $\vchannelc$ and a compositing stage $\vmarkc$.
@@
 \label{sec:construction:nu}
-The encoding function $\vchannel$ is a map from data sections to graphic sections
-\begin{equation}
-\nu: \cgamma{\dbasec}{\dtotalc} \rightarrow \cgamma{\dbasec}{\vtotalc}
-\end{equation}
+The encoding functions $\vchannel: \cgamma{\dbasec}{\dtotalc} \rightarrow \cgamma{\dbasec}{\vtotalc}$ are a subset of the data to visual encoding functions $\equivc$ described in \autoref{sec:artist:equiv}.
+\begin{figure*}[h!]
+  \includegraphics*[]{encoderset.png}
+\end{figure*}
+
 such that the sections project to the same point on the base space $\pi(\dtotal) = \pi(\vchannel(\dtotal))$.
@@
 The difference between $\dtotal$ and $\vtotal$ are semantic rather than structural; they are both maps from the topological space \dbase\ to sets of functions that return records of values.
-This means any $\vtotal$ can be redefined as $\dtotal$,
+This means any $\vtotal$ can be redefined as $\dtotal$, which means that, as shown in \autoref{eq:construction:nu:fabrication}, any collection of $\vchannel$ functions can be composed such that they are equivalent to a $\vchannel$ that directly converts the input to the output.
 \begin{equation}
   \label{eq:construction:nu:fabrication}
-  \begin{tikzcd}[column sep=huge]
+  \begin{tikzcd}
     \dfiberc_{\dbasepointc} 
@@
 \end{equation}
-which means that, as shown in \autoref{eq:construction:nu:fabrication}, any collection of $\vchannel$ functions can be composed such that they are equivalent to a $\vchannel$ that directly converts the input to the output.
  As with artists, $\vchannel$ are maps of sections such that the operators defined in \autoref{sec:artist:operators} can also act on transformers $\vchannel$, meaning that encoders can be added $\vchannel_{a+b} = \vchannel_{a} + \vchannel_{b}$ and multiplied d $\vchannel_{a\times b} = \vchannel_{a}  \vchannel_{b}$.
@@
   \includegraphics[width=1\columnwidth]{full_scatter.png}
-  \caption{This simple $\vmarkc$ assembles a circular visual element that is the color specified in $\vsection(\dbasepoint)$ and is at the intersection specified in $\vsection(\dbasepoint)$ \label{fig:construction:q}}
+  \caption{This simple $\vmarkc$ assembles a circular visual element that is the color specified in $\vsection(\dbasepoint)$ and is at the intersection specified in $\vsection(\dbasepoint)$ \label{fig:construction:q} \note{much better labeling, include semantic labeling, make everything bigger}}
 \end{figure}
@@
 \label{sec:discussion}
-\begin{figure}
-  \includegraphics[width=1\columnwidth]{full_line.png}
-  \caption{\label{fig:construction:artist:combo}}
-\end{figure}
+
 The framework specified in \autoref{sec:artist} and \autoref{sec:construction} describes how to build structure preserving visualization components, but it is left to the library developer to follow these guidelines when building and reusing components.
@@
 
-This artist is then passed along to a shim artist that makes it compatible with existing Matplotlib objects
-\begin{minted}{python}
-class GenericArtist(martist.Artist):
-    def __init__(self, artist:TopologicalArtist):
-        super().__init__()
-        self.artist = artist
-
-    def compose_with_tau(self, section):
-        self.section = section
-
-    def draw(self, renderer, bounds, rate):
-        for tau_local in self.section.query(bounds, rate): 
-            mu = self.artist.nu(tau_local)
-            rho = self.artist.qhat(**mu)
-            output = rho(renderer)
-\end{minted}
 As shown in the \texttt{draw} method, generating a graphic section $\gsection$ is implemented as the composition of $\texttt{qhat} \approx \vmark$ and $\texttt{nu} \approx \vchannel$ applied to a local section of the sheaf $\texttt{self.section.query} \approx \dsection^{i}$ such  $\texttt{draw} \approx \vmarkc\circ\vchannelc\circ\dsectionc  = \vartistc\circ \dsectionc$.
@@
 
-The \texttt{GenericArtist} is a a standard Matplotlib object; therefore it can be hooked into the Matplotlib draw tree to produce the vertical bar chart in \autoref{fig:discussion:vbar}.
+This artist is then passed along to a shim artist that makes it compatible with existing Matplotlib objects (\autoref{sec:appendix:artist_shim}).
+This shim object is hooked into the Matplotlib draw tree to produce the vertical bar chart in \autoref{fig:discussion:vbar}.
 Using the Matplotlib artist framework means this new artist can be composed with existing artists, such as the ones that draw the axes and ticks.
-
 The example in this section is intentionally trivial to illustrate that the math to code translation is fairly straightforward and results in fairly self contained composable functions.
+A library applying these ideas, created by Thomas Caswell, can be found at \url{https://github.com/matplotlib/data-prototype}.
 Further research could investigate building new systems using this model, specifically libraries for visualizing domain specific structured data and domain specific artists.
@@
 \section{Summary}
+\label{sec:appndix:summary}
 The topological spaces and functions introduced throughout this paper are summarized here for reference.
@@
 
-\note{color}
 \begin{table}[h!]
@@
   \hline
-      & function & constraint \\
-  \hline
-  \textcolor{artist}{artist} & $\vartist: \Gamma(\dbase, \dtotal) \rightarrow \Gamma(\gbase, \gtotal)$ &  \\
-  \hline 
-  \textcolor{functor}{lookup} & $\vindex: \gbase\rightarrow \dbase$  &\\ add [0,1] etc...
-  \hline
-  \textcolor{artist}{encoder} & $\vchannel: \Gamma(\dbase, \dtotal) \rightarrow \Gamma(\dbase, \vtotal) $  & \\
-  \hline
-  \textcolor{artist}{compositor} & $\vmark: \Gamma(\dbase,\vtotal) \rightarrow \Gamma(\gbase, \vtotal)$ &  \\
+      & function & constraints \\
+  \hline
+  \textcolor{artist}{artist} & $\vartistc: \cgamma{\dbasec}{\dtotalc} \rightarrow \imartist{\gbasec}{\gtotalc}$ &  \\
+  Data to Graphi  &  $\imartist{\gbasec}{\gtotalc} \subset \cgamma{\gbasec}{\gtotalc}$ & $\vindexc(\gbasec) = \dbasec$\\
+  & & \\
+  \hline 
+  Encoders  & & \\
+           & & \\
+  \hline
+  Decompositor & & \\
+              && \\
   \hline
@@
 
+\note{color}
+\begin{table}[h!]
+  \centering
+  {\renewcommand{\arraystretch}{1.2}
+\begin{   tabular}{|r|l|l|}
+  \hline
+      & function & constraint \\
+  \hline
+  \textcolor{artist}{artist} & $\vartist: \Gamma(\dbase, \dtotal) \rightarrow \Gamma(\gbase, \gtotal)$ &  \\
+  \hline 
+  \textcolor{functor}{lookup} & $\vindex: \gbase\rightarrow \dbase$  &add [0,1] etc...\\ 
+  \hline
+  \textcolor{artist}{encoders} & $\vchannel: \Gamma(\dbase, \dtotal) \rightarrow \Gamma(\dbase, \vtotal) $  & \\
+  \hline
+  \textcolor{artist}{compositor} & $\vmark: \Gamma(\dbase,\vtotal) \rightarrow \Gamma(\gbase, \vtotal)$ &  \\
+  \hline
+\end{tabular}
+\caption{artist, verification functions, and construction $\vartist = \vmarkc \circ \nu$ introduced in \autoref{sec:artist}, and \autoref{sec:construction}}
+\label{tab:appedendix:summary:artist}
+}
+\end{table}
+
 These functions can be verified using the following functions:
 
+\label{sec:appendix:artist_shim}
+
+As mentioned in \autoref{sec:discussion}, one advantage of using this type of  
+functional categorical approach to software design is that we can develop new 
+components that can be incorporated into the existing code base. For matplotlib, 
+we can use these functional artists by wrapping them in a very thin compatibility
+layer shim so that they behave like existing artists.
+
+\begin{minted}{python}
+  class GenericArtist(martist.Artist):
+      def __init__(self, artist:TopologicalArtist):
+          super().__init__()
+          self.artist = artist
+
+      def compose_with_tau(self, section):
+          self.section = section
+
+      def draw(self, renderer, bounds, rate):
+          for tau_local in self.section.query(bounds, rate): 
+              mu = self.artist.nu(tau_local)
+              rho = self.artist.qhat(**mu)
+              output = rho(renderer)
+  \end{minted}
 % use section* for acknowledgment

### 2023-01-19 paper/figcode/nontrivial.py

@@ -205,5 +205,81 @@ def draw_segmented_bundle(ax, plot_total_space, labels):
                                               edgecolor=fc, zorder=3)
                 axins.add_patch(at)
        
+<<<<<<< Updated upstream
         title = axins.text(.5, 0.1, f"{labels[0]} Bundle", color='k', 
                  fontsize=config['fs'], ha='center', va='bottom')
+=======
+            ys, yt = (top, bottom) if flip and xr==t2[1] else (bottom, top) 
+            at = mpatches.FancyArrowPatch(posA=(xa, ys), posB=(xa, yt),
+                                          arrowstyle='->,head_width=.15', mutation_scale=10, 
+                                          edgecolor=fc, zorder=3)
+            axins.add_patch(at)
+
+        title = axins.set_title(f"{labels[0]} Bundle", y=-.65, color='k', fontsize=config['fs'])
+        
+        
+def plot_local_trivialization(ax):
+ 
+    ax.set( ylim=(0, 1), xlim=(config['t'][0]-config['boundary_pad']*2,
+                                     config['t'][-1]+config['boundary_pad']*2))
+    
+    ax.set_title("Local Trivializations of " +r"$(E,K,\pi,F)$", 
+                 fontsize=config['tfs'], color='k', y=.9)
+
+    for i, ((x0, x1), (tc, fc, bc)) in enumerate([(config['r1'], config['dark_colors']), 
+                                   (config['r2'], config['light_colors'])]):
+        rect = mpatches.Rectangle((x0, config['ry']), height = config['rh'], 
+                                  width=(x1-x0), 
+                                  edgecolor=tc, facecolor=pc.lighten(tc,.25), 
+                                  lw=config['lw'])
+        ax.add_patch(rect)
+        ax.plot([x0, x1], [config['ly'], config['ly']], lw=config['lw'], 
+                          color=bc, marker='o', mfc='white',  markersize=4)
+
+        for x in np.linspace(x0, x1, int(np.round(x1-x0,1)*5) +1):
+            at = mpatches.FancyArrowPatch(posA=(x, config['ry']), 
+                                          posB=(x, config['ry']+ config['rh']),
+                                            arrowstyle='->,head_width=0.15', 
+                                          mutation_scale=10, edgecolor=fc)
+            ax.add_patch(at)
+            
[…]

### 2023-01-31 paper/main.tex
5.1 and 5.3 and appendix and nt

@@
 \usepackage{array}
+\usepackage{multirow}
 \usepackage{tabulary} 
@@
   \includegraphics[width=\columnwidth]{figures/local_trivalizations.png}
-  \caption{Two local trivializations, $\dtotal_{0}$ and $\dtotal_{1}$, of an arbitrary bundle \dtotal.} 
+  \caption{The local trivializations $\dtotal_{0}$ and $\dtotal_{1}$ are subspaces of $\dtotal$ where the fibers $\dfiber_{0}$ are equal, here illustrated as all pointing in the same direction.
+These local trivializations are themselves locally trivial fiber bundles.}
 \end{figure}
@@
 \begin{figure}[h!]
-  \includegraphics[width=1\columwidth]{figures/cyl_mob.png}
+  \includegraphics[width=1\columnwidth]{figures/cyl_mob.png}
   \caption{\note{figure out short caption here}\label{fig:cyl_mob_bundles}}
@@
 In the \textcolor{artist}{encoding} stage $\vchannelc$, a data bundle is treated as separable fields and each field is mapped to a measurable visual variable.
-In the encoding stage, the expected visual mappings $\equivb$ can be implemented inside the library.
+In the encoding stage, many of the expected visual mappings $\equivb$ can be implemented inside the library.
 Factoring out the encoding stage leaves the \textcolor{artist}{compositing} stage $\vmarkc$ responsible for faithfully translating those measurable visual components into a visual element.
@@
 \label{sec:construction:vtotal}
-We propose an intermediate visual fiber bundle 
-\begin{equation}
-  \vfiberc \hookrightarrow \vtotalc \xrightarrow{\pi} \dbasec
-\end{equation}
-where the space of possible visual encodings is the fiber space $\vfiber$. The space of visual sections which return visual encoding specifications is 
-\begin{equation}
-\cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} \coloneqq \big\{\vsectionc: \opensetc\rightarrow \vtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\vsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}
-\end{equation}
-As shown in \autoref{eq:artist:inout:diagram},  measurable visual components are defined as having the same continuity as the data $\dbase$.
-This means that every data record $\dsection(\dbasepoint)$ has a corresponding visual section such that $\pi(\dsection(\dbasepoint)) = \pi(\vsection(\dbasepoint))$.
-
-Although the bundle $\vtotal$ is structurally equivalent to the bundle $\dtotal$, its existence allows for separating the data that is input into the artist $\dsection$ from the data internal to the artist $\vsection$.
-This allows a visualization library to define a visual bundle space $\vtotalc$ the holds the internal representation standard for measurable visual components.
-For example, a visualization library could define a visual color fiber as an RGBA tuple $\vfiber_{color} = \mathbb{R}^{3} \times [0,1]$.
-This would then set the expectation that arbitrary color encoding functions would need to return an RGBA tuple for the library to recognize the encoding as a color.
+We encapsulate the space of measurable components reachable through the encoding stage $\vchannel$ as a visual fiber bundle $\vfiberc \hookrightarrow \vtotalc \xrightarrow{\pi} \dbasec$.
+The  restricted fiber space $\vfiber$ of the bundle acts as the specification of the internal library representation of the measurable visual components.
+The space of visual sections $\cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} \coloneqq \big\{\vsectionc: \opensetc\rightarrow \vtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\vsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}$ return a visual encoding $\vsectionc(\dbasepoint)$ corresponding to data record $\dbasepoint(\dbasepoint)$.
+Since the data bundle $dtotal$ and visual bundle $\vtotal$ have the same continuity $\pi(\dsection(\dbasepoint)) = \pi(\vsection(\dbasepoint))$, they are considered structurally equivalent such that $\dtotal=\vtotal$.
+The distinguishing characteristic of $\vtotal$ is that it is part of the construction of the artist and therefore a part of the visualization library implementation.
+We propose that reusing the fibers $\vfiber$ across components facilitates standardizing internal types across the library and that this standardization improves maintainablity (\autoref{tab:appendix:library_spec}).
 
@@
 \label{sec:construction:nu}
-The encoding functions $\vchannel: \cgamma{\dbasec}{\dtotalc} \rightarrow \cgamma{\dbasec}{\vtotalc}$ are a subset of the data to visual encoding functions $\equivc$ described in \autoref{sec:artist:equiv}.
-\begin{figure*}[h!]
-  \includegraphics*[]{encoderset.png}
-\end{figure*}
-
-such that the sections project to the same point on the base space $\pi(\dtotal) = \pi(\vchannel(\dtotal))$.
+
+As introduced in \autoref{sec:artist:equiv}, there is a set $\equivb$ of functions that map between data and corresponding visual encodings.
+We propose that for visualization library components to be structure preserving, they must implement a constrained subset of these encoding functions
+\begin{equation}
+\cgamma{\dbasec}{\dtotalc} \xrightarrow{\vchannelc} \cgamma{\dbasec}{\vtotalc}  \subset \cgamma{\dbasec}{\dtotalc} \xrightarrow{\equivc}\textcolor{set}{Hom}(\dbasec, \measurec)
+\end{equation}
+that preserve the categorical structure (operators and morphisms) of the fiber and the contininuity of the data section.
+As mentioned in \autoref{sec:construction:vtotal}, the total visual space is restricted to the space of data types internal to the library $\vfiber \subset \measure$ and sections are subsets of homsets $\Gamma(\dbase, \vtotal) \subset Hom(\dbase, \measure)$ because sections must be continuous.
+
+The encodeding functions $\vchannel$ are fiberwise transforms such that $\pi(\dtotal) = \pi(\vchannel(\dtotal))$.
 A consequence of this property is that $\vchannelc$ can be constructed as a pointwise transformation such that
@@
 which means that means that a point in a single data fiber $\delement \in \dfiber_{\dbasepointc}$ can be mapped into a corresponding point in a visual fiber $\velement \in \vfiber_{\dbasepointc}$.
-This means that an encoding function $\vchannel$ can convert a single record and may not need the whole dataset.
-
-The difference between $\dtotal$ and $\vtotal$ are semantic rather than structural; they are both maps from the topological space \dbase\ to sets of functions that return records of values.
-This means any $\vtotal$ can be redefined as $\dtotal$, which means that, as shown in \autoref{eq:construction:nu:fabrication}, any collection of $\vchannel$ functions can be composed such that they are equivalent to a $\vchannel$ that directly converts the input to the output.
+This means that an encoding function $\vchannel$ can convert a single record independent of the whole dataset.
+
+Since $\dtotal$ and $vtotal$ are structurally identical, any $\vtotal$ can be redefined as $\dtotal$; therefore, as shown in \autoref{eq:construction:nu:fabrication}, any collection of $\vchannel$ functions can be composed such that they are equivalent to a $\vchannel$ that directly converts the input to the output.
 \begin{equation}
@@
   {\renewcommand{\arraystretch}{1.2}
-\begin{   tabular}{|r|l|l|}
+\begin{tabular}{|r|l|l|}
   \hline
@@
   \hline 
-  \textcolor{functor}{lookup} & $\vindex: \gbase\rightarrow \dbase$  &add [0,1] etc...\\ 
+  \textcolor{functor}{lookup} & $\vindex: \gbase\rightarrow \dbase$  &  \\ 
   \hline
@@
 \caption{artist, verification functions, and construction $\vartist = \vmarkc \circ \nu$ introduced in \autoref{sec:artist}, and \autoref{sec:construction}}
-\label{tab:appedendix:summary:artist}
+\label{tab:appendix:summary:artist}
 }
@@
 
-These functions can be verified using the following functions:
-
+\section{Internel Library Specification}
+
+\label{tab:appendix:library_spec}
+\begin{table}[H]
+  \centering
+  \renewcommand{\arraystretch}{2}
+  \begin{tabulary}{\columnwidth}{|l|L|l|}\hline
+   \(\bm{\vchannel_{i}}\)    & \(\bm{\vsection_{i}}\)  & \(\bm{codomain(\vchannel_{i}) \subset \vfiber_{i}}\)  \\ \hline                                              
+  position                    & x, y, z, theta, r      & \(\mathbb{R}\)   \\ \hline
+  size                        & linewidth, markersize  & \(\mathbb{R}^{+}\)  \\ \hline
+  shape                       & markerstyle            & \(\{f_{0}, \ldots, f_{n}\}\)\\ \hline
+  color                       & color, facecolor, markerfacecolor, edgecolor  & \(\mathbb{R}^{4}\) \\ \hline
+  \multirow{2}{*}{texture}    & hatch      & \(\mathbb{N}^{10}\)\\\cline{2-3}
+                              & linestyle    & \((\mathbb{R}, \mathbb{R^+}^{n, n\%2=0})\) \\ \hline              
+  \end{tabulary}
+  \caption{Some of the $\vfiber$ components of the $\vtotal$ bundles in Matplotlib components}
+  \label{tab:math:artist:mpl:fiber}
+\end{table}
+
+\pagebreak
+\section{Matplotlib Compatibility}
 \label{sec:appendix:artist_shim}

### 2023-02-01 paper/main.tex
pulled up topolology to related work topological structure

@@
 
-There is an assumption in visualization that the connectivity of the elements in a dataset are preserved in the visual representation; roughly topology provides a method for encoding this connectivity \cite{wilkinsonGrammarGraphics2005}.
-Bertin gives the examples of discrete data values should be represented by discrete entities such as points, quantitative continuous data by lines, and surfaces through area marks \cite{bertinSemiologyGraphicsDiagrams2011}.
+The connectivity and relative positioning of elements in a dataset is broadly its topology\cite{wilkinsonGrammarGraphics2005} and the expectation that it be be reflected in the visual representation is present in even the earliest maps \cite{friendlyBriefHistoryData2008} and is exemplified in Bertin's recommendation that discrete data values should be represented by discrete entities such as points, quantitative continuous data by lines, and surfaces through area marks \cite{bertinSemiologyGraphicsDiagrams2011}.
+Encoding this point, interval, and surface, connectivity as mathematical topologies provides us with a consistent way to describe and discuss these and other types of connectivity and the topological structure of the corresponding visualizations.
+
+\begin{definition}
+A \textbf{topology} $\mathcal{T}$ on a space $X$ is a colletion of open sets\footnote{Open sets (open subsets) are a generalization of open intervals to n dimensional spaces.
+For example, an open ball is the set of points inside the ball and excludes points on the surface of the ball.
+\cite{weissteinOpenSet,bradleyTopologyVsTopology}} \openset\ of $X$.
+This collection has the properties that the empty set $\varnothing$ and $\dbase$ are in $\mathcal{T}$, the union of open sets in $\mathcal{T}$ is also in $\mathcal{T}$, and any finite intersection of open sets in $\mathcal{T}$ is also in $\mathscr{T}$.
+\cite{bradleyTopologyCategoricalApproach2020}
+\end{definition}
+
  In \autoref{fig:related-work:continuity:ktypes} the records of the GHCN\cite{lawrimoreGlobalHistoricalClimatology2011} weather station table can be considered disconnected from each other and therefore have a 0D continuity.
@@
 Instead of interpolation, which is how continuity was preserved in the timeseries plot, in this instance the connectivity is preserved by plotting the stations in the map such that their relative distances are preserved.
+
+\note{pull definition of topology up here and also maybe homotopy type/homomorphic spaces in the next paragraph-it'll make things less handwavy }
+
 Encoding connectivity using the formalism of topological spaces allows us to encode the connectivity of the data in a uniform manner such that we can then generalize the preservation of that connectivity in the visual representation rather than constructing a condition specifically for each type of topological structure.
@@
 We encode the topological structure of the data as the \textcolor{base}{base space} \dbase\ of the fiber bundle since the base space acts as indexing space where every point in the base space has a corresponding fiber space $\pi^{-1}(\dbasepoint) = \dfiber_{\dbasepoint}$.
-The \textcolor{base}{base space} \dbasec\ is a topological space $(\dbase,\mathcal{T}_{\dbase})$, which means it  consists of the set of points ${\dbasepoint \in \dbase}$ and topology $\mathcal{T}_{\dbase}$ \cite{munkresElementsAlgebraicTopology1984}.
-A topology $\dbase \coloneqq (\dbase,\mathcal{T}_{\dbase})$ is an axiomatic way of defining a topological structure\cite{bradleyTopologyCategoricalApproach2020}  as a collection of subsets, called open sets \footnote{Open sets (open subsets) are a generalization of open intervals to n dimensional spaces.
-For example, an open ball is the set of points inside the ball and excludes points on the surface of the ball.
-\cite{weissteinOpenSet,bradleyTopologyVsTopology}}, of that structure.
-\begin{definition}
-A \textbf{topology} $\mathcal{T}$ on a space $X$ is a collection of open sets \openset\ of $X$.
-This collection has the properties that the empty set $\varnothing$ and $\dbase$ are in $\mathcal{T}$, the union of open sets in $\mathcal{T}$ is also in $\mathcal{T}$, and any finite intersection of open sets in $\mathcal{T}$ is also in $\mathscr{T}$.
-\cite{bradleyTopologyCategoricalApproach2020}
-\end{definition}
+The \textcolor{base}{base space} \dbasec\ is a topological space $(\dbase,\mathcal{T}_{\dbase})$, which means it  consists of the set of points ${\dbasepoint \in \dbase}$ and topology $\mathcal{T}_{\dbase}$.
 We propose that the topology on \dbase\ can act as a mathematical abstraction for the \textcolor{base}{indexing space} of a dataset because the properties of a topology are the same as would be expected of an index space, namely that the space has a mathematical structure, that the space can be subsetted, and that continuity is preserved when sets of indices are merged.

### 2023-02-08 paper/main.tex
pullin top up

@@
   \caption{
-  A a topological representation of the data (in purple) is a simplified representation of a view of how the records in a dataset are connected.
-The table can be viewed as a set of independent records; therefore it has a 0D continuity represented by the disconnected dots.
-These weather records are also samples of continuous measurements at each location, which can be modeled as the time series measurements being samples on a 1d continuous interval and the geospatial station measurements as sparse samples on a 2D continuous space.
+  \note{add first panel which is table + all K type, second is scatter, etc...}
+
   \label{fig:related-work:continuity:ktypes}}
@@
 
-The connectivity and relative positioning of elements in a dataset is broadly its topology\cite{wilkinsonGrammarGraphics2005} and the expectation that it be be reflected in the visual representation is present in even the earliest maps \cite{friendlyBriefHistoryData2008} and is exemplified in Bertin's recommendation that discrete data values should be represented by discrete entities such as points, quantitative continuous data by lines, and surfaces through area marks \cite{bertinSemiologyGraphicsDiagrams2011}.
-Encoding this point, interval, and surface, connectivity as mathematical topologies provides us with a consistent way to describe and discuss these and other types of connectivity and the topological structure of the corresponding visualizations.
+The connectivity and relative positioning of elements in a dataset is broadly its topology\cite{wilkinsonGrammarGraphics2005} and the expectation that it be be reflected in the visual representation is present in even the earliest maps \cite{friendlyBriefHistoryData2008} and in Bertin's assertion that the dimensionality of the data match the class of representation \cite{bertinSemiologyGraphicsDiagrams2011}.
+In this work, we propose that using the formalism of topology provides us a uniform way to express dimensionality such that we can generalize the expectation that the data and visual dimensionalty match past points, lines, and areas.
 
 \begin{definition}
-A \textbf{topology} $\mathcal{T}$ on a space $X$ is a colletion of open sets\footnote{Open sets (open subsets) are a generalization of open intervals to n dimensional spaces.
+A \textbf{topology} $\mathcal{T}$ on a space $X$ is a collection of open sets\footnote{Open sets (open subsets) are a generalization of open intervals to n dimensional spaces.
 For example, an open ball is the set of points inside the ball and excludes points on the surface of the ball.
@@
 
- In \autoref{fig:related-work:continuity:ktypes} the records of the GHCN\cite{lawrimoreGlobalHistoricalClimatology2011} weather station table can be considered disconnected from each other and therefore have a 0D continuity.
-This means each distinct row can be reduced to a discrete point and each row will be encoded as a discrete visual element, such as a standalone scatter point.
-Each station can also be considered a time series of weather observation since temperature and pressure are sampled from a 1D continuous measurement space.
-The connectivity of these observations can be modeled as an interval and the data is encoded as a line since that visual element is also 1D continuous.
-The measurements at set of stations on any given day is a sparse sampling of a 2D geospatial grid and the connectivity can be modeled as a 2D plane.
-Instead of interpolation, which is how continuity was preserved in the timeseries plot, in this instance the connectivity is preserved by plotting the stations in the map such that their relative distances are preserved.
-
-\note{pull definition of topology up here and also maybe homotopy type/homomorphic spaces in the next paragraph-it'll make things less handwavy }
-
-Encoding connectivity using the formalism of topological spaces allows us to encode the connectivity of the data in a uniform manner such that we can then generalize the preservation of that connectivity in the visual representation rather than constructing a condition specifically for each type of topological structure.
+A topology is a partitioning of a space into sets of its subspaces and which connectivity is encoded in the topology is determined by the way in which the space is partitioned.
+For example, the GHCN\cite{lawrimoreGlobalHistoricalClimatology2011} dataset in \autoref{fig:related-work:continuity:ktypes} is a complex spatio-temporal record of weather station data.
+One way of encoding that each record is independent is to encode the underlying topology as a set of discrete points, one for each record.
+Each of these discrete points has a corresponding marker in the scatter plot.
+Another way of subsetting the data is as sets of timeseries, which can be encoded as intervals since weather station data is sampled from 1D continous natural phenomena.
+Each line plot can be thought of as a function evaluated on the underlying interval that encoding the temporal continuity.
+A third way to view the data is as a sparse sampling of a 2D geospatial grid, where the connectivity can be modeled as a 2D plane.
+Here the connectivity is preserved by plotting the stations in the map such that their relative distances are preserved.
+
+note{add summary and transition}
 

### 2023-02-08 paper/main.tex
finished out RW top section

@@
 \section{Related Work}
-We build on the extensive work codifying the structure of visualization data to develop a more generalizable method for expressing the structure that visualizations are expected to preserve.
-Structure preservation is important because it means visualzations are easier to understand \cite{norman_things_smart} and ensures that the visual representations are not distortions of the data\cite{tufteVisualDisplayQuantitative2001}.
+We build on the extensive work codifying the structure of visualization data to develop a more generalizable method for expressing the structure that tools are expected to preserve when transforming data into visualizations.
+Structure preservation ensures that visual representations are not distortions of the data\cite{tufteVisualDisplayQuantitative2001} and it makes visualizations easier to understand \cite{norman_things_smart}.
 Broadly, previous work describes structure as a combination of the topological properties of the data, the mathematical structure of the data fields, and equivalent or compatible changes to data and graphics.
@@
 
-A topology is a partitioning of a space into sets of its subspaces and which connectivity is encoded in the topology is determined by the way in which the space is partitioned.
+A topology is a partitioning of a space into its subspaces in a way where those subspaces can be composed back into the space and the choice of partitions is how we encode the connectivity of the data that we are interested in.
 For example, the GHCN\cite{lawrimoreGlobalHistoricalClimatology2011} dataset in \autoref{fig:related-work:continuity:ktypes} is a complex spatio-temporal record of weather station data.
 One way of encoding that each record is independent is to encode the underlying topology as a set of discrete points, one for each record.
-Each of these discrete points has a corresponding marker in the scatter plot.
-Another way of subsetting the data is as sets of timeseries, which can be encoded as intervals since weather station data is sampled from 1D continous natural phenomena.
-Each line plot can be thought of as a function evaluated on the underlying interval that encoding the temporal continuity.
-A third way to view the data is as a sparse sampling of a 2D geospatial grid, where the connectivity can be modeled as a 2D plane.
-Here the connectivity is preserved by plotting the stations in the map such that their relative distances are preserved.
-
-note{add summary and transition}
-
-Visual algorithms assume the topology of their input data, as described in taxonomies of visualization algorithms Chi\cite{chiTaxonomyVisualizationTechniques2000} and by Troy and M\"{o}ller \cite{toryRethinkingVisualizationHighlevel2004}.
+Each of these discrete points is also represented visually as a discrete marker in the scatter plot.
+Another way of viewing the data is as sets of timeseries since weather station data is sampled from 1D continous natural phenomena.
+We encode this topological structure as intervals and sets of sub-intervals, and  we visually represent this strucure as continuous line plots.
+A third way of viewing the data is as a sparse sampling of a 2D geospatial grid, where the connectivity is modeled as a  surface and sub regions of that surface.
+In the visualization, that connectivity is expressed as plotting the stations in the map, visually denoting that these are points in the continuous geographical space that is New York State.
+By using topology as the underlying encoding scheme, we are able to represent the structure of points, lines, surfaces, and more using the same abstractions.
+
+Since topological continuity is generalizable, we propose that encoding topology as part of a dataset type and a visualization type allows us to explicitly check which types of structure a visualization algorithm can preserve.
+Visual algorithms assume the topology of their input data, as described in taxonomies of visualization algorithms Chi\cite{chiTaxonomyVisualizationTechniques2000} and by Troy and M\"{o}ller \cite{toryRethinkingVisualizationHighlevel2004}, but generally do not verify that input structure.
 For example, a \texttt{line} algorithm often does not have a way to query whether a list of (x,y) coordinates is the distinct rows, the time series, or the list of stations in \autoref{fig:related-work:continuity:ktypes}.
 While plotting the time series as a continuous line would be correct, it would be incorrect for a visualization to indicate that the distinct rows or stations are connected in a 1D continuous manner.
-Since topological continuity is generalizable, we propose the construction of topology types such that data can explicitly state its topology type and visual algorithms can check that this type matches their assumptions.
-The typing system proposed in this paper builds on Butler's proposal of using a mathematical structure called fiber bundles as an abstract data representation in visualization \cite{butlerVectorBundleClassesForm1992, butlerVisualizationModelBased1989}.
-We extend Butler's work because he discusses how bundles can be used to encode data in a consistent manner for the topological and field structures that are common in visualization.
+Encoding the expected input structure would enable a visualization algorithm to reject data that does not explicitely state that it matches those assumptions.
+This structure can be encoded in a generalizable way by building on Butler's work on using a mathematical structure called fiber bundles as an abstract data representation in visualization \cite{butlerVectorBundleClassesForm1992, butlerVisualizationModelBased1989} where he discusses how bundles can be used to encode data in a consistent manner for the topological and field structures that are common in visualization.
 We sketch out fiber bundles in \autoref{sec:atct:fiber-bundles}, but his work provides a thorough introduction to bundles for visualization practitioners.

### 2023-02-09 paper/main.tex
rewrite of intro/and consideration of pulling up definitions

@@
 
-\IEEEPARstart{V}isualizations are expected to be a faithful translation of data, meaning that inherent structure in the data should be present in the visualization; therefore the components of visualization software libraries that translate data to graphics are also expected to be structure preserving.
-This inherent structure can be quite complex, such as when the data is high dimensional, multivariate, deeply nested, or encoded in a distributed manner.
-We propose that we can rigorously specify what structure the library components are expected to preserve using category theory.
-We also propose that these specifications can generalize to most structure types by extending previous work on encoding  inherent structure in a consistent manner using mathematical structures from algebraic topology.
+\note{The mapping between data and visualization is expected to be bidirectional -> that's what we really mean by equivariance and homomorphism  }
+\IEEEPARstart{V}isualizations are expected to be structure preserving, meaning that when data is transformed into a graphical representation enough of the structure of the data is present in the graphic that the data can be distilled solely from the graphical representation.
+Changes to structure through changing the data or the graphic in the $data\leftrightarrow graphic$ mapping are expected to be reflected in each inputs corresponding output \cite{kindlmannAlgebraicProcessVisualization2014}.Because visualizations are assumed to preserve structure, visual algorithms implemented in visualization producing tools, such as software libraries, are also expected to preserve structure; frequently these expectations are either codified implicitly, only for some aspects of the data structure, or inconsistently with respect to visualization algorithm type.
+This means that currently many visualization tools lack an explict, comprehensive, and consistent way to express what conditions must be met to gaurantee that the component produced a structure preserving visualization.
+
+We propose that we can encode the structure of the data and graphics in a uniform generalizable manner by using a mathematical structure from algebraic topology called a fiber bundle, and that we can use category theory to explicitly specify what it means for a visualization tool component to preserve structure.
 These algebraic structures can be translated into analogous programmatic types for encoding structure, while the use of category theory facilitates developing a functional design framework that describes function composition in terms of functional algebraic operations.
@@
 each column in a table is mapped to a distinct encoding.
+For example, nominal measurements are permutable, meaning that they are acted on by elements of a permutation group.
+For the visual transformation to be equivariant, that nominal measurement must be mapped to a visual encoding that is also permutable, such as distinct marker shapes or color hues.
+We evaluate whether the encoding is homomorphic when the visual encoding does not have a group structure.
+For example, partial orders are not invertible; therefore a visual encoding of partially ordered data is structure preserving when the ordering is mapped to equivalently ordered visual encodings.
+\note{add in Scales/APT/ GoG/Hullman/ETC}
+
+\subsection{General Structure Preservation}
+\note{maybe at top of section?}
 Two mathematical concepts for describing a function that preserves structure are \textit{equivariance} and \textit{homomorphism}.
@@
  \end{definition}
-For example, nominal measurements are permutable, meaning that they are acted on by elements of a permutation group.
-For the visual transformation to be equivariant, that nominal measurement must be mapped to a visual encoding that is also permutable, such as distinct marker shapes or color hues.
-We evaluate whether the encoding is homomorphic when the visual encoding does not have a group structure.
-For example, partial orders are not invertible; therefore a visual encoding of partially ordered data is structure preserving when the ordering is mapped to equivalently ordered visual encodings.
 
@@
 the data container) changes, that data maps unambiguously to visual elements \cite{ziemkiewiczEmbeddingInformationVisualization2009}, and that changing the data should correspond to changes in the visualization in a perceptually significant manner.
-In \autoref{sec:atct:sheaves} we introduce a uniform abstract data representation layer, the expectation of unambiguous mappings is expressed in \autoref{sec:atct:xi}, and the definition of equivariance introduced in \autoref{sec:artist:equivariant:artist} includes the correspondence between data and visual changes.
+A uniform abstract data representation layer is introduced in \autoref{sec:atct:sheaves}, the expectation of unambiguous mappings is expressed in \autoref{sec:atct:xi}, and the the correspondence between data and visual changes is included in the definition of equivariance introduced in \autoref{sec:artist:equivariant:artist}.
 In this work, we formally specify the conditions necessary to build structure preserving components using category theory, because it provides a rich language for expressing the structure of objects, functions between objects, and the constraints that must hold for these functions to compose \cite{wielsManagementEvolvingSpecifications1998,yorgeyMonoidsThemeVariations}.

### 2023-03-06 paper/main.tex
Updating GitHub to make Overleaf up to date

@@
   \note{add first panel which is table + all K type, second is scatter, etc...}
-
-  \label{fig:related-work:continuity:ktypes}}
+  }
+  \label{fig:related-work:continuity:ktypes}
 \end{figure}
@@
 \label{sec:related-work:equivariance}
+\note{put this b/f topology since this is the thing folks do/is known}
 As with topology, Bertin \cite{bertinSemiologyGraphicsDiagrams2011} codified the expectation that a data value and its visual encoding are expected to have compatible structure.

### 2023-03-06 paper/figcode/nontrivial.py

@@ -23,7 +23,7 @@ config.update(t = np.linspace(0, 2*np.pi, config['N']))
 config.update(r1 = np.array([config['t'][0], config['boundary']])-config['boundary_pad']/2, 
               r2 = np.array([config['boundary'], config['t'][-1]])+config['boundary_pad']/2, #trivialization E x
               ry = .35, rh =.5 , #trivilization E y
-              lw = 2, ly =  .25 #trivialization K lw and y
+              lw = 2, ly = .1 #trivialization K lw and y
              )
 
 
@@ -205,18 +205,15 @@ def draw_segmented_bundle(ax, plot_total_space, labels):
                                               edgecolor=fc, zorder=3)
                 axins.add_patch(at)
        
-<<<<<<< Updated upstream
-        title = axins.text(.5, 0.1, f"{labels[0]} Bundle", color='k', 
-                 fontsize=config['fs'], ha='center', va='bottom')
-=======
+
             ys, yt = (top, bottom) if flip and xr==t2[1] else (bottom, top) 
             at = mpatches.FancyArrowPatch(posA=(xa, ys), posB=(xa, yt),
-                                          arrowstyle='->,head_width=.15', mutation_scale=10, 
+                                          arrowstyle='->,head_width=.15', 
                                           edgecolor=fc, zorder=3)
             axins.add_patch(at)
 
-        title = axins.set_title(f"{labels[0]} Bundle", y=-.65, color='k', fontsize=config['fs'])
-        
+        title = axins.text(.5, 0.075, f"{labels[0]} Bundle", color='k', 
+                 fontsize=config['fs'], ha='center', va='bottom')
         
 def plot_local_trivialization(ax):
  
@@ -249,8 +246,11 @@ def plot_local_trivialization(ax):
                                arrowstyle=config['aws'], mutation_scale=config['ms']-5, 
                                color=config['bundle_color'])
         ax.add_artist(fp)
-        ax.text(xarr+.1, config['ry']*.75, "$\pi$", fontsize=config['fs'],
+        ax.annotate(text="$\pi$", xy=(0,.5), xycoods=fp,  fontsize=config['fs'],
                    color=config['bundle_color'], ha='left', va='center')
+        
+        #ax.text(xarr+.1, config['ry']*.75, "$\pi$", fontsize=config['fs'],
+        #           color=config['bundle_color'], ha='left', va='center')
         
[…]

### 2023-03-06 paper/main.tex
Updates from Overleaf

--- 2023-02-09 paper/main.tex
@@
 \renewcommand{\linenumberfont}{\normalfont\bfseries\small\color{lightgray}}
-\usepackage{minted}
+\usepackage[outputdir=../]{minted}
 \setminted[python]{fontsize=\scriptsize, 
@@
 % author names and IEEE memberships
-\author{Hannah~Aizenman, Thomas~Caswell, and~Michael~Grossberg,~\IEEEmembership{Member,~IEEE,}% <-this % stops a space
+\author{Hannah~Aizenman, Mikael Vejdemo-Johansson, Thomas~Caswell, and~Michael~Grossberg,~\IEEEmembership{Member,~IEEE,}% <-this % stops a space
 \IEEEcompsocitemizethanks{\IEEEcompsocthanksitem H. Aizenman and M. Grossberg are with the department of Computer Science, City College of New York. 
@@
 E-mail: haizenman@ccny.cuny.edu, mgrossberg@ccny.cuny.edu 
+\IEEEcompsocthanksitem Mikael Vejdemo-Johansson is with the department of Mathematics, CUNY College of Staten Island.
+\protect\\
+E-mail: mvj@math.csi.cuny.edu
 \IEEEcompsocthanksitem Thomas Caswell is with National Synchrotron Light Source II, Brookhaven National Lab 
@@
   \note{add first panel which is table + all K type, second is scatter, etc...}
-
+%
   \label{fig:related-work:continuity:ktypes}}

### 2023-05-08 paper/main.tex
Updates from Overleaf

--- 2023-03-06 paper/main.tex
@@
 \note{The mapping between data and visualization is expected to be bidirectional -> that's what we really mean by equivariance and homomorphism  }
-\IEEEPARstart{V}isualizations are expected to be structure preserving, meaning that when data is transformed into a graphical representation enough of the structure of the data is present in the graphic that the data can be distilled solely from the graphical representation.
+\IEEEPARstart{V}isualizations are expected to be structure preserving, meaning that when data is transformed into a graphical representation enough of the structure of the data is present in the graphic that the data can be extracted from the graphical representation.
 Changes to structure through changing the data or the graphic in the $data\leftrightarrow graphic$ mapping are expected to be reflected in each inputs corresponding output \cite{kindlmannAlgebraicProcessVisualization2014}.Because visualizations are assumed to preserve structure, visual algorithms implemented in visualization producing tools, such as software libraries, are also expected to preserve structure; frequently these expectations are either codified implicitly, only for some aspects of the data structure, or inconsistently with respect to visualization algorithm type.
@@
 \section{Related Work}
-We build on the extensive work codifying the structure of visualization data to develop a more generalizable method for expressing the structure that tools are expected to preserve when transforming data into visualizations.
-Structure preservation ensures that visual representations are not distortions of the data\cite{tufteVisualDisplayQuantitative2001} and it makes visualizations easier to understand \cite{norman_things_smart}.
-Broadly, previous work describes structure as a combination of the topological properties of the data, the mathematical structure of the data fields, and equivalent or compatible changes to data and graphics.
-We propose that a unified abstraction model can guide developers in building structure aware composable software library components with more consistent interfaces.
+We build on extensive work codifying the structure of visualization data and corresponding graphics and use the concepts of \textit{homomorphism} and \textit{equivariance} to express how that structure should be preserved.
+Structure is considered preserved when the input and output change in tandem, and these symmetries are often modeled as \textit{group actions}, which are a way for a group\footnote{A \textit{group} is a set with an associative binary operator, an identity, and an inverse operation} to transform some other set:
+
+\begin{definition}\label{def:related-work:action}\cite{nlab:action, grimaldiDiscreteCombinatorialMathematics2006}
+  An \textcolor{action}{\textbf{action}} \footnote{Throughout this paper, we augment the math with color to group conceptually similar objects and functions\cite{headMathAugmentationHow2022}, for example \textcolor{action}{actions} and \textcolor{action}{sets of actions}.
+This color coding carries through to the figures.} of \textcolor{action}{$G = (G,\circ, e)$} on $X$ is a function  $act: \textcolor{action}{G} \times X \rightarrow X$.
+An action has the properties of identity $act(\textcolor{action}{e}, x) = x$ for all  $x \in X$ and associativity $act(\textcolor{action}{g}, act(\textcolor{action}{f}, x)) = act(\textcolor{action}{f} \circ \textcolor{action}{g}, x)$ for $\textcolor{action}{f},\textcolor{action}{g} \in \textcolor{action}{G}$.
+\end{definition}
+
+Given a group $G$ that acts on both the input $X$ and the output $Y$ of a function $f: X \rightarrow Y$
+
+ \begin{definition}
+$f$ is \textbf{equivariant} when $f(act(g,x)) = act(g,f(x))$ for all $g$ in $G$ and for all $x$ in $X$\cite{nlab:equivariant, pittsNominalSetsNames2013}.
+ \end{definition}
+
+ \note{add something back in for homomorphic}
+
+The notion that visual encodings should be \textit{homomorphic} was proposed by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987} in his specification of \textit{A Presentation Tool} and the idea that visual transforms are \textit{equivariant} underlies Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}'s Algebraic Visualization Design.
+Kindlemann and Scheidegger propose three specific types of structure preservation: that the visualization should not change if the data representation (i.e.
+the data container) changes, that data maps unambiguously to visual elements \cite{ziemkiewiczEmbeddingInformationVisualization2009}, and that changing the data should correspond to changes in the visualization in a perceptually significant manner.
+A uniform abstract data representation layer is introduced in \autoref{sec:atct:sheaves}, the expectation of unambiguous mappings is expressed in \autoref{sec:atct:xi}, and the the correspondence between data and visual changes is included in the definition of equivariance introduced in \autoref{sec:artist:equivariant:artist}.
+We define structure preservation both in \note{field types and topology}
+
+\subsection{Field Type Structure} 
+\label{sec:related-work:equivariance}
+ Bertin \cite{bertinSemiologyGraphicsDiagrams2011} codified the expectation that a data value and its visual encoding are expected to have compatible structure.
+The structure on values has traditionally been described  using the measurement scales-i.e.
+nominal, ordinal, interval, ratio - which Steven's classified by their mathematical group structure\cite{stevensTheoryScalesMeasurement1946a} and others have formally expanded to include more types \cite{leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}.
+Generally the structure on each field of the dataset is preserved separately, i.e.
+each column in a table is mapped to a distinct encoding.
+For example, nominal measurements are permutable, meaning that they are acted on by elements of a permutation group.
+For the visual transformation to be equivariant, that nominal measurement must be mapped to a visual encoding that is also permutable, such as distinct marker shapes or color hues.
+We evaluate whether the encoding is homomorphic when the visual encoding does not have a group structure.
+For example, partial orders are not invertible; therefore a visual encoding of partially ordered data is structure preserving when the ordering is mapped to equivalently ordered visual encodings.
+\note{add in Scales/APT/ GoG/Hullman/ETC}
 
@@
 We sketch out fiber bundles in \autoref{sec:atct:fiber-bundles}, but his work provides a thorough introduction to bundles for visualization practitioners.
-
-\subsection{Field Type Structure} 
-\label{sec:related-work:equivariance}
-As with topology, Bertin \cite{bertinSemiologyGraphicsDiagrams2011} codified the expectation that a data value and its visual encoding are expected to have compatible structure.
-The structure on values has traditionally been described  using the measurement scales-i.e.
-nominal, ordinal, interval, ratio - which Steven's classified by their mathematical group structure\cite{stevensTheoryScalesMeasurement1946a} and others have formally expanded to include more types \cite{leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}.
-Generally the structure on each field of the dataset is preserved separately, i.e.
-each column in a table is mapped to a distinct encoding.
-For example, nominal measurements are permutable, meaning that they are acted on by elements of a permutation group.
-For the visual transformation to be equivariant, that nominal measurement must be mapped to a visual encoding that is also permutable, such as distinct marker shapes or color hues.
-We evaluate whether the encoding is homomorphic when the visual encoding does not have a group structure.
-For example, partial orders are not invertible; therefore a visual encoding of partially ordered data is structure preserving when the ordering is mapped to equivalently ordered visual encodings.
-\note{add in Scales/APT/ GoG/Hullman/ETC}
-
-\subsection{General Structure Preservation}
-\note{maybe at top of section?}
-Two mathematical concepts for describing a function that preserves structure are \textit{equivariance} and \textit{homomorphism}.
-An \textit{equivariant} function preserves symmetric group structure on its input and output, while a \textit{homomorphic} function $f$ preserves binary operations on input and output of the same algebraic type.
-Group structure and binary operations are both describing families of functions or operations that can be applied to data and visual encodings.
-A group $G$ is a set with an operator $\circ$.
-The set contains an identity element $e \in G$ and the operator $\circ$ is closed, associative and invertible.
-Applying elements of a set, such as a group, to another set, such as data $X$, is called an \textit{action}.
-
-\begin{definition}\label{def:related-work:action}\cite{nlab:action, grimaldiDiscreteCombinatorialMathematics2006}
-  An \textcolor{action}{\textbf{action}} \footnote{Throughout this paper, we augment the math with color to group conceptually similar objects and functions\cite{headMathAugmentationHow2022}, for example \textcolor{action}{actions} and \textcolor{action}{sets of actions}.
-This color coding carries through to the figures.} of \textcolor{action}{$G$} on $X$ is a function  $act: \textcolor{action}{G} \times X \rightarrow X$.
-An action has the properties of identity $act(\textcolor{action}{e}, x) = x$ for all  $x \in X$ and associativity $act(\textcolor{action}{g}, act(\textcolor{action}{f}, x)) = act(\textcolor{action}{f} \circ \textcolor{action}{g}, x)$ for $\textcolor{action}{f},\textcolor{action}{g} \in \textcolor{action}{G}$.
-\end{definition}
-
-In \autoref{sec:artist:equivariant}, actions on data and compatible actions on graphics define structure in terms of testing whether a visualization is equivariant.
-Given a group $G$ that acts on input $X$ and output $Y$ and a function $f: X \rightarrow Y$
-
- \begin{definition}
-$f$ is \textbf{equivariant} when $f(g(x)) = g(f(x))$ for all $g$ in $G$ and for all $x$ in $X$\cite{nlab:equivariant, pittsNominalSetsNames2013}.
- \end{definition}
- \begin{definition}
- $f$ is \textbf{homomorphic} when  $f(x_1 \oslash x_2) = f(x_1) \oslash f(x_2)$ for every pair $x_1, x_2$ in $X$\cite{grimaldiDiscreteCombinatorialMathematics2006}. 
- \end{definition}
-
-The notion that visual encodings should be \textit{homomorphic} was proposed by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987} in his specification of \textit{A Presentation Tool} and the idea that visual transforms are \textit{equivariant} underlies Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}'s Algebraic Visualization Design.
-Kindlemann and Scheidegger propose three specific types of structure preservation: that the visualization should not change if the data representation (i.e.
-the data container) changes, that data maps unambiguously to visual elements \cite{ziemkiewiczEmbeddingInformationVisualization2009}, and that changing the data should correspond to changes in the visualization in a perceptually significant manner.
-A uniform abstract data representation layer is introduced in \autoref{sec:atct:sheaves}, the expectation of unambiguous mappings is expressed in \autoref{sec:atct:xi}, and the the correspondence between data and visual changes is included in the definition of equivariance introduced in \autoref{sec:artist:equivariant:artist}.
-In this work, we formally specify the conditions necessary to build structure preserving components using category theory, because it provides a rich language for expressing the structure of objects, functions between objects, and the constraints that must hold for these functions to compose \cite{wielsManagementEvolvingSpecifications1998,yorgeyMonoidsThemeVariations}.
-A brief visualization oriented introduction to category theory is in Vickers et al \cite{vickersUnderstandingVisualizationFormal2013}, but they are applying category theory to semantic concerns about visualization design rather than library architecture.
 
@@
 By assuming the structure of their data, these domain specific libraries can provide more cohesive interfaces for a much more limited set of visualization algorithms than the building block libraries offer.
+\note{transition sentence}
+In this work, we formally specify the conditions necessary to build structure preserving components using category theory, because it provides a rich language for expressing the structure of objects, functions between objects, and the constraints that must hold for these functions to compose \cite{wielsManagementEvolvingSpecifications1998,yorgeyMonoidsThemeVariations}.
+A brief visualization oriented introduction to category theory is in Vickers et al \cite{vickersUnderstandingVisualizationFormal2013}, but they are applying category theory to semantic concerns about visualization design rather than library architecture.
 
@@
 We build on this work to describe how to encode the connectivity of the data as a topological space, separately encode the fields as their own topological space with a typing system, and express the mappings between these two spaces.
+
+The fiber bundle abstraction is a structure fusing some \textcolor{base}{base space} (indexing into the data) with some \textcolor{fiber}{fiber space} (acting as the data domain).
+This sort of fusion could be done by a simple direct product (or join-operation), but with the fiber bundle abstraction we get the flexibility of expressing more complicated fusions than the direct product, as long as they can be constructed from building blocks that look like direct products.
 
@@
 \begin{enumerate}
-  \item the \textcolor{fiber}{fiber space} \dfiberc\ is the preimage of the projection function $\pi$ at a point $\dbasepoint$ in the \textcolor{base}{base space} \dbasec\ such that $\dfiber_{\dbasepoint} = \pi^{-1}(\dbasepoint)$.
+  \item all fibers in the bundle are isomorphic.
+Since all fibers are isomorphic, there is a uniquely determined \textcolor{fiber}{fiber space} \dfiberc\ given by the preimage of the projection $\pi$ at any point $\dbasepoint$ in the \textcolor{base}{base space} \dbasec: $\dfiberc = \pi^{-1}(k)$.
+  %the \textcolor{fiber}{fiber space} \dfiberc\ is the preimage of the projection function $\pi$ at a point $\dbasepoint$ in the \textcolor{base}{base space} \dbasec\ such that $\dfiber_{\dbasepoint} = \pi^{-1}(\dbasepoint)$.
 All fibers in a bundle are isomorphic such that $\dfiber \cong \dfiber_{\dbasepoint}$ for all points $\dbasepoint \in \dbase$.
-  \item there is an open neighborhood $\openset_{\dbasepoint}$ surrounding each point in the base space $\dbasepoint \in \openset_{\dbasepoint} \subset \dbase$ such that the \textcolor{total}{total space} \dtotalc\ over the neighborhood, denoted $\dtotal\restriction\openset$\footnote{The symbol $\restriction$ is the restriction operator\cite{RestrictionMathematics2022} defined in \texttt{amssymb}.
+  \item each point $\dbasepointc$ in the \textcolor{base}{base space} \dbasec\ has an open neighborhood $\openset_{\dbasepointc}$ such that the \textcolor{total}{total space} \dtotalc\ over the neighborhood is locally trivial.
+\textbf{Local triviality} means $\dtotal\restriction_{\openset} = \openset\times F$.
+In this paper we use $\dtotal\restriction_{\openset} = \pi^{-1}(\openset)$ to denote the preimage of an openset, and a \textbf{local trivialization} is a specific choice of neighborhoods and their preimages.
+  %there is an open neighborhood $\openset_{\dbasepoint}$ surrounding each point in the base space $\dbasepoint \in \openset_{\dbasepoint} \subset \dbase$ such that the \textcolor{total}{total space} \dtotalc\ over the neighborhood, denoted $\dtotal\restriction\openset$\footnote{The symbol $\restriction$ is the restriction operator\cite{RestrictionMathematics2022} defined in \texttt{amssymb}.
 For example $\pi^{-1}\restriction\openset$ is the inverse function $\pi^{-1}$ defined only on the points in $\openset$.
@@
 \end{definition}
-\begin{definition} A \textcolor{section}{\textbf{section}} $\dsectionc: \dbasec \rightarrow \dtotalc$ over a fiber bundle is a smooth right inverse of $\pi$ such that $\pi(\dsection(\dbasepoint)) = \dbasepoint$ for all $\dbasepoint \in \dbase$
+\begin{definition} A \textcolor{section}{\textbf{section}} $\dsectionc: \dbasec \rightarrow \dtotalc$ over a fiber bundle is a smooth right inverse of $\pi$: $\pi(\dsection(\dbasepoint)) = \dbasepoint$ for all $\dbasepoint \in \dbase$
 \end{definition}
 
-Local triviality $\dtotal\restriction_{\openset_{k}} = \openset_{k}\times \dfiber$ means that there is always a subset of the base space over which the fibers are identical. 
 \begin{figure}[h!]
@@
 These local trivializations are themselves locally trivial fiber bundles.}
+  \label{fig:local_trivalizations}
 \end{figure}
 
-For example, in \autoref{fig:local_trivalizations}, the bundles $\dtotal_{0}=\dtotal\restriction_{[\epsilon^{-}, {\frac{2\pi}{5}}^{+}]}$ and $\dtotal_{1}=\dtotal\restriction_{[{\frac{2\pi}{5}}^{-}, \epsilon^{-}]}$ are local trivializations of an arbitrary bundle $\dtotal$.
+\note{Add motiviating note here!}
+Consider the circle.
+Points can be identified with angles (in radians, for instance), and we can imagine covering the circle with two overlapping intervals - for instance $\dbase_0=(-\varepsilon, 2\pi/5+\varepsilon)$ and $\dbase_1=(2\pi/5-\varepsilon, 2\pi+\varepsilon)$.
+We can imagine creating building blocks $\dbase_0\times[-1,1]$ and $\dbase_1\times[-2,2]$.
+
+There are two different ways to create a fiber bundle that has these building blocks as the local trivializations: on the one hand, we could imagine just a cylinder $S^1\times[-1,1]$ as a total space.
+The fibers would all be just the interval $[-1,1]$.
+Over $\dbase_0$, we could map from the local trivialization to the total space with the identity map, and over $\dbase_1$ we could map from the local trivialization to the total space by rescaling the fiber space $[-2,2]\to[-1,1]$.
+
+On the other hand, we could imagine creating a Möbius strip by flipping one end of $\dbase_1\times[-2,2]$ over before attaching it, so that while $\dbase_0\times[-1,1]$ gets mapped into the total space with the identity map, the part $(2\pi-\varepsilon,2\pi+\varepsilon)\times[-2,2]$ of $\dbase_1\times[-2,2]$ is mapped by the function $x\mapsto -x/2$ instead of the function $x\mapsto x/2$.
+
+These functions - using the isomorphisms $x\mapsto x$, $x\mapsto x/2$ and $x\mapsto -x/2$ to map from each rectangle into the cylinder compose into different maps $(\dbase_0\cap\dbase_1)\times[-2,2]\to(\dbase_0\cap\dbase_1)\times[-1,1]$ where the source is seen as a subset of $\dbase_1\times[-2,2]$ and the target is seen as a subset of $\dbase_0\times[-1,1]$.
+By specifying just these \textbf{transition maps} for all overlaps of cover elements in the local trivialization we can specify a gluing scheme that constructs the fiber bundle from these locally trivial pieces.
+
+%For example, in \autoref{fig:local_trivalizations}, the bundles $\dtotal_{0}=\dtotal\restriction_{[\epsilon^{-}, {\frac{2\pi}{5}}^{+}]}$ and $\dtotal_{1}=\dtotal\restriction_{[{\frac{2\pi}{5}}^{-}, \epsilon^{-}]}$ are local trivializations of an arbitrary bundle $\dtotal$.
 These local trivializations overlap on the intersections of the open subsets of each trivalization $[\epsilon^{-}, \epsilon^{+}]$ and $[{\frac{2\pi}{5}}^{-}, {\frac{2\pi}{5}}^{+}]$.
@@
 
-In a \textit{trivial bundle}, the fibers are identical $\dfiber = \dfiber_{k}$ for all points $\dbasepoint \in dbase$ such that all the transtion maps are also identical.
+A \textit{trivial bundle} is directly isomorphic to $\dbasec\times\dfiberc$, and for any choice of cover of $\dbasec$ by overlapping opensets, we can choose local trivializations so that all transition maps are identity maps.
+In the example in Figure~\ref{fig:cyl_mob_bundles}, we use arrows $\uparrow$ to denote fiber alignments, illustrating that by choosing simple enough trivializations, we get identity maps in the transition maps.
+
+%In a \textit{trivial bundle}, the fibers are identical $\dfiber = \dfiber_{k}$ for all points $\dbasepoint \in dbase$ such that all the transtion maps are also identical.
 For example, in the cylinder case the fibers in the overlappping regions of the trivalizations point in the same direction $\uparrow$ to illustrate that they are equal $\dfiber_0\restriction_{\openset_1\cap\openset_2} = \dfiber_{1}\restriction_{\openset_1\cap\openset_2}$.
@@
 
-In a \textit{non-trivial} bundle fibers are isomorphic $\dfiber \simeq \dfiber_{\dbasepoint}$ for all points $\dbasepoint \in \dbase$ such that there is at least one transition map that is not an identity \cite{hatcherAlgebraicTopology2002}.
+A \textit{non-trivial bundle} on the other hand is one that can not be constructed as $\dbasec\times\dfiberc$, but where the construction is more complicated.
+Then for any choice of local trivializations, there is at least one transition map that is not an identity \cite{hatcherAlgebraicTopology2002}.
+For example, in the Möbius band case, while $\dfiber_0\restriction_{(2\pi/5-\varepsilon, 2\pi/5+\varepsilon)}\to\dfiber_1\restriction_{(2\pi/5-\varepsilon, 2\pi/5+\varepsilon)}$ can be chosen to be an identity map (after transforming $\dbase_1\times[-2,2]\to\dbase_1\times[-1,1]$ by rescaling), the other transition map component $\dfiber_0\restriction_{(-\varepsilon,\varepsilon)}\to\dfiber_1\restriction_{(-\varepsilon,\varepsilon)}$ has to flip the fibers to align them between the pieces.
+
+%In a \textit{non-trivial} bundle fibers are isomorphic $\dfiber \simeq \dfiber_{\dbasepoint}$ for all points $\dbasepoint \in \dbase$ such that there is at least one transition map that is not an identity \cite{hatcherAlgebraicTopology2002}.
 For example, in the mobius band case, while $\dfiber_{0}\restriction_{[{\frac{2\pi}{5}}^{-}, {\frac{2\pi}{5}}^{+}]} = \dfiber_{0}\restriction_{[{\frac{2\pi}{5}}^{-}, {\frac{2\pi}{5}}^{+}]}$ such that the transition map is an identity map $\uparrow\rightarrow\uparrow$, the transition map must flip  $\uparrow \rightarrow \downarrow$ the fibers of $\dfiber_0\restriction_{[\epsilon^{-}, \epsilon^{+}]}$ in order for them to be equal to the fibers of $\dfiber_1\restriction_{[\epsilon^{-}, \epsilon^{+}]}$.
-\note{add summary sentence about power of triviality, probably pointing to it's power for abstraction}
 
@@
 \label{sec:atct:fb:base}
-We encode the topological structure of the data as the \textcolor{base}{base space} \dbase\ of the fiber bundle since the base space acts as indexing space where every point in the base space has a corresponding fiber space $\pi^{-1}(\dbasepoint) = \dfiber_{\dbasepoint}$.
+\note{base space of bundle for key/value pair, topology in general 'cause data + viz continuity in same way}
+Munzner~\cite{munznerChDataAbstraction} describes a key-value semantics for data representation and abstraction, where datasets can be queried with different kinds of keys depending on the dataset type - locations for values of spatial fields, times for values of timeseries, or indexes for values in flat or multidimensional tables.
+We propose that these different types of keys are instances of different underlying topological organization of the data, and that we can use the structure of the opensets in the topological space to encode the extent to which the keys can vary continuously: flat or multidimensional tables have discrete points as indices, timeseries have points on a line as indices and spatial fields (regardless of whether they are scalar or tensor fields) have points in a spatial domain as indices.
+Therefore we choose to encode the topological structure of the data as the \textcolor{base}{base space} of a fiber bundle.
+Since every point in the base space has a corresponding fiber space that contains the data values, we preserve the general notion of a key (continuity)-value (data) pairing.
+We further follow the example set by \cite{spivakSimplicialDatabases2009,spivakDatabasesAreCategories2010} and propose that the structure of the data types be formally specified as the objects of a category.
+
+%We encode the topological structure of the data as the \textcolor{base}{base space} \dbase\ of the fiber bundle since the base space acts as indexing space where every point in the base space has a corresponding fiber space $\pi^{-1}(\dbasepoint) = \dfiber_{\dbasepoint}$.
 The \textcolor{base}{base space} \dbasec\ is a topological space $(\dbase,\mathcal{T}_{\dbase})$, which means it  consists of the set of points ${\dbasepoint \in \dbase}$ and topology $\mathcal{T}_{\dbase}$.
@@
 
-The topological structure of the data can be expressed programmatically by constructing data types that encapsulate the class of topological structure-.i.e point, line, plane, network, etc.
+%The topological structure of the data can be expressed programmatically by constructing data types that encapsulate the class of topological structure-.i.e point, line, plane, network, etc.
 We propose that these data types can be formally specified as objects of a category.
+
 \begin{definition}\label{def:atct:category}
@@
 
-\section{Artist: Data to Graphic}
+\section{Codifying Structure Preservation}
 \label{sec:artist}
@@
 
-\section{Construction}
+\section{Constructing Structure Preserving Components}
 \label{sec:construction}

### 2023-05-08 paper/main.tex
main
intro take 1000

@@
 
-\note{The mapping between data and visualization is expected to be bidirectional -> that's what we really mean by equivariance and homomorphism  }
-\IEEEPARstart{V}isualizations are expected to be structure preserving, meaning that when data is transformed into a graphical representation enough of the structure of the data is present in the graphic that the data can be extracted from the graphical representation.
-Changes to structure through changing the data or the graphic in the $data\leftrightarrow graphic$ mapping are expected to be reflected in each inputs corresponding output \cite{kindlmannAlgebraicProcessVisualization2014}.Because visualizations are assumed to preserve structure, visual algorithms implemented in visualization producing tools, such as software libraries, are also expected to preserve structure; frequently these expectations are either codified implicitly, only for some aspects of the data structure, or inconsistently with respect to visualization algorithm type.
-This means that currently many visualization tools lack an explict, comprehensive, and consistent way to express what conditions must be met to gaurantee that the component produced a structure preserving visualization.
-
-We propose that we can encode the structure of the data and graphics in a uniform generalizable manner by using a mathematical structure from algebraic topology called a fiber bundle, and that we can use category theory to explicitly specify what it means for a visualization tool component to preserve structure.
-These algebraic structures can be translated into analogous programmatic types for encoding structure, while the use of category theory facilitates developing a functional design framework that describes function composition in terms of functional algebraic operations.
+\IEEEPARstart{V}isualization is a structure preserving mapping from data to graphic.
+One way of verifying that the structure of the data is present in the graphic is to verify that the data set can be distilled from the graph, i.e.
+that the values read off a bar chart are equivalent to the values used to construct the bar chart.
+Since structure preservation requires that the $data\leftrightarrow graphic$ be bidirectional, changes to either data or graphic are expected to propogate to the other side of the transformation\cite{kindlmannAlgebraicProcessVisualization2014}.
+In this paper, we propose a methodology for describing structure, verifying structure preservation, and specifying the conditions for constructing a structure preserving map between data and graphics.
+We use category theory, which is the math of structures and their relations, to explicitly specify what it means for a visualization mapping to preserve structure in an explicit, comprehensive, and consistent way.
+We encode the structure of the data and graphics in a uniform generalizable manner by using a mathematical structure from algebraic topology called a fiber bundle; doing so allows us to discuss everything from lines to complex ND spaces with the same interface.
+
+We use category theory and algebraic structures because they can be directly translated into software design specifications and our goal is to improve visulization tools.
+The algebraic structures can be translated into analogous programmatic types, while category theory facilitates developing a functional design framework that describes function composition in terms of functional algebraic operations.
 The typing system and composition inherent in a functional design encourage library developers to build complex visualization components from simple verifiable parts that have few side effects\cite{huHowFunctionalProgramming2015, hughesWhyFunctionalProgramming1989}.
@@
 \hline
- \gbasepointc\ to \dbasepointc& $ \vindexc: \opensetgc \rightarrow \opensetc$ & for $\dbasepointc \in \opensetc$ exists $\gbasepointc \in \opensetgc$ \\
+ \gbasepointc\ to \dbasepointc& $ \vindexc: \opensetgc \rightarrow \opensetc$ & for $\gbasepointc \in \opensetgc$ exists $\dbasepointc \in \opensetc$ \\
  & & s.t. $\vindexc(\gbasepointc) = \dbasepointc$\\
@@
   \textcolor{artist}{artist} & $\vartistc: \cgamma{\dbasec}{\dtotalc} \rightarrow \imartist{\gbasec}{\gtotalc}$ &  \\
-  Data to Graphi  &  $\imartist{\gbasec}{\gtotalc} \subset \cgamma{\gbasec}{\gtotalc}$ & $\vindexc(\gbasec) = \dbasec$\\
+  Data to Graphic  &  $\imartist{\gbasec}{\gtotalc} \subset \cgamma{\gbasec}{\gtotalc}$ & $\vindexc(\gbasec) = \dbasec$\\
   & & \\

### 2023-05-08 paper/main.tex
intro + rw

@@
 % correct bad hyphenation here
-\hyphenation{}
 
@@
 
-We use category theory and algebraic structures because they can be directly translated into software design specifications and our goal is to improve visulization tools.
+We use category theory and algebraic structures because they can be directly translated into software design specifications and our goal is to improve visualization tools.
 The algebraic structures can be translated into analogous programmatic types, while category theory facilitates developing a functional design framework that describes function composition in terms of functional algebraic operations.
@@
  \end{definition}
-
- \note{add something back in for homomorphic}
+In an equivariant transform, mapping the data after acting on it $f(act)$ is equivalent to acting on the data after mapping it $act(f)$ since $G$ acts on both.
+A homomorphism does not require that the input and output are of the same type; instead
+
+ \begin{definition}
+\note{must sort out exact appropriate definition of group homomorphism here}
+ \end{definition}
 
@@
 For example, partial orders are not invertible; therefore a visual encoding of partially ordered data is structure preserving when the ordering is mapped to equivalently ordered visual encodings.
-\note{add in Scales/APT/ GoG/Hullman/ETC}
+Steven's measurement scale's and the structure preservation they define is the basis of Wilkenson's scales\cite{wilkinsonGrammarGraphics2005} in Grammar of Graphics and derivative libraries, and in this work we generalize the concept of fields to support a wider array of field types.
+This also allows us to specify what consistency means when creating different views of data from a single dataset, as discussed in Qu and Hullman \cite{hullmanKeeping2018}, specifically in terms of each field type separate from the overall structure.
 
@@
 
+\pagebreak
 \section{Internel Library Specification}

### 2023-06-02 paper/main.tex
whatever from last time + consolidating union + intro to base space

@@
 \usetikzlibrary{cd} % commutative diagrams
-\usepackage{pgfplots}
+\usepackage{pgfplots} %numberline
 \DeclareUnicodeCharacter{2212}{−}
@@
 For example, the GHCN\cite{lawrimoreGlobalHistoricalClimatology2011} dataset in \autoref{fig:related-work:continuity:ktypes} is a complex spatio-temporal record of weather station data.
-One way of encoding that each record is independent is to encode the underlying topology as a set of discrete points, one for each record.
+One way of encoding that each record is independent is to encode the underlying topology as a set of 0D $\cdots$ points, one for each record.
 Each of these discrete points is also represented visually as a discrete marker in the scatter plot.
-Another way of viewing the data is as sets of timeseries since weather station data is sampled from 1D continous natural phenomena.
-We encode this topological structure as intervals and sets of sub-intervals, and  we visually represent this strucure as continuous line plots.
-A third way of viewing the data is as a sparse sampling of a 2D geospatial grid, where the connectivity is modeled as a  surface and sub regions of that surface.
+We can view the timeseries at each station as a 1D $---$ continuous line since weather station data is sampled from natural phenomena.
+We encode this topological structure as intervals, .i.e [start date, end date], and sets of sub-intervals, and we visually represent this strucure as continuous line plots.
+A third way of viewing the data is as a sparse sampling of a 2D $\blacksquare$ geospatial grid, where the connectivity is modeled as a surface and sub regions of that surface.
 In the visualization, that connectivity is expressed as plotting the stations in the map, visually denoting that these are points in the continuous geographical space that is New York State.
@@
 These local trivializations are themselves locally trivial fiber bundles.}
-  \label{fig:local_trivalizations}
+  \label{fig:atct:local_trivalizations}
 \end{figure}
@@
 Therefore we choose to encode the topological structure of the data as the \textcolor{base}{base space} of a fiber bundle.
-Since every point in the base space has a corresponding fiber space that contains the data values, we preserve the general notion of a key (continuity)-value (data) pairing.
-We further follow the example set by \cite{spivakSimplicialDatabases2009,spivakDatabasesAreCategories2010} and propose that the structure of the data types be formally specified as the objects of a category.
-
-%We encode the topological structure of the data as the \textcolor{base}{base space} \dbase\ of the fiber bundle since the base space acts as indexing space where every point in the base space has a corresponding fiber space $\pi^{-1}(\dbasepoint) = \dfiber_{\dbasepoint}$.
-The \textcolor{base}{base space} \dbasec\ is a topological space $(\dbase,\mathcal{T}_{\dbase})$, which means it  consists of the set of points ${\dbasepoint \in \dbase}$ and topology $\mathcal{T}_{\dbase}$.
-We propose that the topology on \dbase\ can act as a mathematical abstraction for the \textcolor{base}{indexing space} of a dataset because the properties of a topology are the same as would be expected of an index space, namely that the space has a mathematical structure, that the space can be subsetted, and that continuity is preserved when sets of indices are merged.
-For example, in \autoref{fig:related-work:continuity:ktypes}, a topology on the line would break the line up into subsets, and those subsets could only be joined in ways where they cover the line in the same order in which it was broken up.
-The base space of a fiber bundle is a quotient topology\cite{QuotientSpaceTopology2020,munkresElementsAlgebraicTopology1984}, meaning that it divides the topological space into the smallest possible (largest number of) open sets such that $\pi$ remains a continuous function.
-This means that the topology can be defined to have a resolution equal to the number of indicies in a dataset or at whatever resolution of continuois function is required for a given task.
-
-%The topological structure of the data can be expressed programmatically by constructing data types that encapsulate the class of topological structure-.i.e point, line, plane, network, etc.
-We propose that these data types can be formally specified as objects of a category.
+Because the base space of a fiber bundle is a quotient topology\cite{munkresElementsAlgebraicTopology1984}, it divides the topological space into the largest number of open sets such that $\pi$ remains a continuous function.
+This means that the topology can be defined to have a resolution equal to the number of indicies in a dataset such that the key (continuity)-value (data) pairing is always preserved.
+
+Following from Spivaks cataegorical abstraction of a database \cite{spivakSimplicialDatabases2009,spivakDatabasesAreCategories2010}, we also propose that the structure of the data types be formally specified as the objects of a category.
 
@@
 \end{definition}
-We encapsulate the topological structure of the data as a category $\mathcal{\dbasec}$.
+
 The standard construction of a category from a topological space is that it has open set objects $\opensetc$ and inclusion morphisms $\opensetc_i \xrightarrow{\iota} \opensetc_j$ such that $\opensetc_i \subseteq \opensetc_j$\cite{riehlCategoryTheoryContext}.
@@
 By formally specifying the properties of the topological structure datatypes as $\mathcal{\dbase}$, we can express that these are the properties that are required as part of the implementation of the data type objects.
+
+For example, the disjoint union of two bundles, as shown in \autoref{fig:atct:local_trivalizations}, is the coproduct $\dbase^{a} \subseteq_{\dbase^{c}} \dbase^{b}$ over an overlap $\dbase^{c}$ and therefore the inclusion morphism must be commmutative:
+\begin{minipage}[c]{0.25\columnwidth} 
+  \begin{equation}
+    \label{eq:atct:base_union}
+  \begin{tikzcd}
+    \dbasec^{a}\sqcup_{\dbasec^{c}}\dbasec^{b} 
+    & \dbasec_{b} \arrow[l, "\iota_b"', color=base] \\
+    \dbasec_a \arrow[u, "\iota_a", color=base] 
+    & \dbasec_c 
+    \arrow[u, "\iota_c"', color=base] 
+    \arrow[l, "\iota_c"', color=base] 
+  \end{tikzcd}
+\end{equation}
+\end{minipage}\hfil
+\begin{minipage}[c]{0.65\columnwidth}
+  \begin{tikzpicture}[commutative diagrams/every diagram]
+    \matrix[matrix of math nodes, name=m, commutative diagrams/every cell]{
+      (0,1) & (.55,1) \\
+      (0,.45) & (.40, 60)\\  
+    };
+
+    \path[commutative diagrams/.cd, every arrow, every label]
+    (m-2-2) edge[color=base] (m-1-2)
+            edge[color=base] (m-2-1)
+    (m-2-1) edge[color=base] (m-1-1)
+    (m-1-2) edge[color=base] (m-1-1);
+  \end{tikzpicture}
+\end{minipage}
+
+The coproduct in \autoref{eq:atct:base_union} shows that the way to test that two spaces have been combined correctly is to verify that every point in the subspace $\dbasepoint \in \dbase^{c}$ must be present in the spaces for which inclusion morphisms exist $\dbasepoint \in \dbase^{a}$ and $\dbasepoint \in \dbase^{b}$ such that $\dbasepoint \in \dbase^{a}\sqcup_{\dbase^{c}} \dbase_{b}$.
+This simple test that the records are joined correctly is what allows us to reliably build larger datasets out of smaller ones, such as in the case of distributed and on demand datasets.
 
@@
   \end{array}\right\}} \\
-  \textcolor{base}{\openset_2 = \left(0,1\right)} 
+  \textcolor{base}{\openset_2 = \left(0,2\pi\right)} 
   \arrow[u, "\sheafc_{1}", maps to, color=sheaf, shift right=4.5]
@@
 This allows us to express that an artist that works on a dataset can be composed of artists that work on sub parts of that dataset.
-\begin{equation}
-  \label{eq:artist:operator}
-  \begin{tikzcd}[column sep=small]
-      & \dfiberc^{a} \times_{\dfiberc^{c}} \dfiberc^{b} 
-    \arrow[ld, "{\pi_{a\times b, a}}"', color=fiber] 
-    \arrow[rd, "{\pi_{a\times b, b}}", color=fiber] & & 
-    \dtotalc^{a} \oplus_{\dfiberc^{c}} \dtotalc^{b} 
-    \arrow[ddd, "\pi"', color=total] \\
-    \dfiberc^{a} 
-    %\arrow[d, "\pi"', color=fiber] 
-    \arrow[r, "{\pi_{a,c}}", color=fiber] 
-    & \dfiberc^{c} & \dfiberc^{b} 
-    %\arrow[d, "\pi", color=fiber] 
-    \arrow[l, "{\pi_{b,c}}"', color=fiber] & \\
-    \dbasec^{a} 
-    \arrow[rd, "{\iota_{a, a+b}}"', color=base] 
-    & \dbasec^{c} 
-    \arrow[r, "{\iota_{c,b}}", color=base] 
-    \arrow[l, "{\iota_{c,a}}"', color=base] 
-    & \dbasec^{b} 
-    \arrow[ld, "{\iota_{b, a+b}}", color=base] 
-    &  \\
-    & \dbasec^{a}\sqcup_{\dbasec^c}\dbasec^{b} & & 
-    \dbasec^{a}\sqcup_{\dbasec^c}\dbasec^{b} 
-    \arrow[uuu, "{\dsectionc^{a,b}}"', dashed, bend right, color=section]
-  \end{tikzcd}
-\end{equation}
 
@@
 \label{sec:artist:addition}
-As illustrated in \autoref{eq:artist:operator}, data bundles can be combined by taking the union of base spaces $\dbase^{a} \sqcup_{\dbase^c} \dbase^b$, where $\dbase^c$ is an overlap.
-\note{amalgamated union, or data bundles of the same type can be thought of as local trivializations of a larger space and glued togetehr as discussed in \autoref{sec:atct:fiber-bundles}}
-
-When the fibers of each bundle are isomorphic $\dfiber^{a} \simeq \dfiber^{b}$, which we denote as $\dtotal$, this is analogous to adding more records to a dataset.
 We propose an addition operator that states that an artist that takes in a dataset can be constructed using artists that take as inputs subsets of the dataset
@@
 As illustrated in \autoref{eq:artist:operator},  fibers that are a cartesian product of fiberspaces $\dfiber^{a} \times_{\dfiber^c} \dfiber^{b}$, where $\dfiber^c$ is any fiber that is present in both fibers, can be projected down into component fibers.
+\begin{tikzcd}
+  \dfiberc^{a}\times_{\dfiber^{c}} \dfiberc^{b} \arrow[d, "\pi_a"] \arrow[r, "\pi_b"] & \dfiberc^{b} \arrow[d, "\pi_c"] & {(temperature, time) \times_{time} (pressure, time)} \arrow[d, "\pi_a"] \arrow[r, "\pi_b"] & {(pressure,time)} \arrow[d, "\pi_c"] \\
+  \dfiberc^{a} \arrow[r, "\pi_c"]                                                     & \dfiberc^{c}                    & {(temperature,time)} \arrow[r, "\pi_c"]                                                    & time
+  \end{tikzcd}
 In the trivial case where the base spaces are the same $\dbase^{a} = \dbase^{b} = \dbase$, this is equivalent to adding more fields to a dataset.
@@
 % BibTeX documentation can be easily obtained at:
-% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
+% http://mirror.ctan.org/biblio/bibtex/contrib/doc/https://www.overleaf.com/download/project/64063771726807fc06c1ee16/build/187faee483d-291434de4cddeda6/output/output.pdf?compileGroup=priority&clsiserverid=clsi-pre-emp-c2d-d-f-nkpv&enable_pdf_caching=true&popupDownload=truejjj
 % The IEEEtran BibTeX style support page is at:
@@
 % argument is your BibTeX string definitions and bibliography database(s)
-\bibliography{bibliography}
+\bibliography{bib
+
 % biography section 

### 2023-06-02 paper/main.tex
fiber

@@
 
-%In a \textit{non-trivial} bundle fibers are isomorphic $\dfiber \simeq \dfiber_{\dbasepoint}$ for all points $\dbasepoint \in \dbase$ such that there is at least one transition map that is not an identity \cite{hatcherAlgebraicTopology2002}.
-For example, in the mobius band case, while $\dfiber_{0}\restriction_{[{\frac{2\pi}{5}}^{-}, {\frac{2\pi}{5}}^{+}]} = \dfiber_{0}\restriction_{[{\frac{2\pi}{5}}^{-}, {\frac{2\pi}{5}}^{+}]}$ such that the transition map is an identity map $\uparrow\rightarrow\uparrow$, the transition map must flip  $\uparrow \rightarrow \downarrow$ the fibers of $\dfiber_0\restriction_{[\epsilon^{-}, \epsilon^{+}]}$ in order for them to be equal to the fibers of $\dfiber_1\restriction_{[\epsilon^{-}, \epsilon^{+}]}$.
-
 We propose that the total space of a bundle can encode the mathematical space in which a  dataset is embedded, the base space can encode the topological properties of the dataset, and the fiber space can encode the data types of the record fields of the dataset.
@@
 The $\otimes$ encapsulates both the sets associated with each fiber $\mathbb{R} \times \mathbb{R} = \mathbb{R}^{2}$ and the morphisms associated with each functor $(\dfunct_{speed}, \dfunct_{direction}) = \dfunct_{wind}$.
-Defining the field schema as a monoidal category allows us to formally express the structure of the field, as defined by its sets and functions, at different levels of composition as needed by different visualization tasks.
- \
+Combining fibers can be verified by checking that when a fiber component $\dfiber^{c}$ is present in both $\dfiber^{a}$ and $\dfiber^{b}$, it is identical when projected out of either such that the *product* diagram commutes:
+\begin{minipage}
+
+\begin{tikzcd}[c]{.25\columnwidth}
+  \dfiberc^{a}\times_{\dfiber^{c}} \dfiberc^{b} \arrow[d, "\pi_a"] \arrow[r, "\pi_b"] & \dfiberc^{b} \arrow[d, "\pi_c"] & {(temperature, time) \times_{time} (pressure, time)} \arrow[d, "\pi_a"] \arrow[r, "\pi_b"] & {(pressure,time)} \arrow[d, "\pi_c"] \\
+  \dfiberc^{a} \arrow[r, "\pi_c"]                                                     & \dfiberc^{c}                    & {(temperature,time)} \arrow[r, "\pi_c"]                                                    & time
+  \end{tikzcd}
+\end{minipage}\hfill
+\begin{minipage}[c]{.65\columwidth}
+
+\end{minipage}
+
+This means that when multifield data is decomposed into its component fields, shared records stay the same.
+For example, the red (temperature, time, pressure) record seperates into (tempereture, time) and (pressure, time) records that share the same red time.
+Furthermore this time is the same whether it is obtained from the (temperature, time) or (pressure, time) record.
+This simple test that fields are joined together correctly for the same record is what allows us to reliably combine multiple datasets together on shared properties-for example growing the weather station data from a temporal to spatial dataset by adding the weather at each  location at each time.
+
 \subsubsection{Data: Section} 
@@
 \note{fiber product vs cartesian product}
-As illustrated in \autoref{eq:artist:operator},  fibers that are a cartesian product of fiberspaces $\dfiber^{a} \times_{\dfiber^c} \dfiber^{b}$, where $\dfiber^c$ is any fiber that is present in both fibers, can be projected down into component fibers.
-\begin{tikzcd}
-  \dfiberc^{a}\times_{\dfiber^{c}} \dfiberc^{b} \arrow[d, "\pi_a"] \arrow[r, "\pi_b"] & \dfiberc^{b} \arrow[d, "\pi_c"] & {(temperature, time) \times_{time} (pressure, time)} \arrow[d, "\pi_a"] \arrow[r, "\pi_b"] & {(pressure,time)} \arrow[d, "\pi_c"] \\
-  \dfiberc^{a} \arrow[r, "\pi_c"]                                                     & \dfiberc^{c}                    & {(temperature,time)} \arrow[r, "\pi_c"]                                                    & time
-  \end{tikzcd}
+
 In the trivial case where the base spaces are the same $\dbase^{a} = \dbase^{b} = \dbase$, this is equivalent to adding more fields to a dataset.

### 2023-06-08 paper/main.tex
notes for what needs to be filled in
backup cause vscode is trying to eat my files

@@
-% bare_jrnl_compsoc, texV1.4b, 2015/08/26, Michael Shell
-\documentclass[10pt,journal,compsoc]{IEEEtran}
+\documentclass[journal]{IEEEtran}
 
 % *** MISC UTILITY PACKAGES ***
-
 \newcommand{\note}[1]{\textcolor{magenta}{#1}} 
+\usepackage{silence}
+
+\WarningsOff[fixltx2e]
+\usepackage[switch]{lineno}
+\renewcommand{\linenumberfont}{\normalfont\bfseries\small\color{lightgray}}
+
+% references
 \usepackage[nocompress]{cite}
 \usepackage[]{footmisc}
-%\usepackage[notref, notcite]{showkeys}
-\usepackage{hyperref} % autoref
+\usepackage{bookmark}
+
 % *** GRAPHICS RELATED PACKAGES ***
-\ifCLASSINFOpdf
-   \usepackage[pdftex]{graphicx}
-  % declare the path(s) where your graphic files are
+\usepackage{graphicx}
 \graphicspath{{./figures/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-   \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-   \graphicspath{{./figures/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-   \DeclareGraphicsExtensions{.eps}
-\fi
-
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
 
@@
 \usepackage{bm}
+
+% theorem environments
 \theoremstyle{definition}
@@
 \newtheorem{example}{Example}[section]
-% *** SPECIALIZED LIST PACKAGES ***
-
-\usepackage{xcolor}
-\usepackage{algorithmic}
-\usepackage[utf8]{inputenc}
-\usepackage{multicol}
+\newtheorem{axiom}{Axiom}
+\newtheorem{prop}{Proposition}
+
+% *** diagrams
 \usepackage{tikz}
@@
 \usetikzlibrary{patterns,shapes.arrows}
-
 \pgfplotsset{compat=newest}
-\newtheorem{axiom}{Axiom}
-\newtheorem{prop}{Proposition} %math?
-\usepackage[switch]{lineno}
-\renewcommand{\linenumberfont}{\normalfont\bfseries\small\color{lightgray}}
-\usepackage[outputdir=../]{minted}
+
+% *** SPECIALIZED LIST PACKAGES ***
+\usepackage{xcolor}
+\usepackage[utf8]{inputenc}
+
+\usepackage[outputdir=.]{minted}
 \setminted[python]{fontsize=\scriptsize, 
@@
 % *** ALIGNMENT PACKAGES ***
+\usepackage{multicol}
 \usepackage{array}
@@
 % *** SUBFIGURE PACKAGES ***
-\ifCLASSOPTIONcompsoc
-  \usepackage[caption=false,font=footnotesize,labelfont=sf,textfont=sf]{subfig}
-\else
-  \usepackage[caption=false,font=footnotesize]{subfig}
-\fi
+\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
 
@@
 \usepackage{xurl}
-%\usepackage{hyperref} breaks compilation?
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
 
@@
 % correct bad hyphenation here
+%\hyphenation{}
 
@@
 \maketitle
-
 \IEEEpeerreviewmaketitle
-
 \IEEEraisesectionheading{\section{Introduction}\label{sec:intro}}
 
-\IEEEPARstart{V}isualization is a structure preserving mapping from data to graphic.
+\IEEEPARstart{V}{isualization} is a structure preserving mapping from data to graphic.
 One way of verifying that the structure of the data is present in the graphic is to verify that the data set can be distilled from the graph, i.e.
 that the values read off a bar chart are equivalent to the values used to construct the bar chart.
-Since structure preservation requires that the $data\leftrightarrow graphic$ be bidirectional, changes to either data or graphic are expected to propogate to the other side of the transformation\cite{kindlmannAlgebraicProcessVisualization2014}.
+Since structure preservation requires that the $data\leftrightarrow graphic$ be bidirectional, changes to either data or graphic are expected to propagate to the other side of the transformation\cite{kindlmannAlgebraicProcessVisualization2014}.
 In this paper, we propose a methodology for describing structure, verifying structure preservation, and specifying the conditions for constructing a structure preserving map between data and graphics.
@@
  \begin{definition}
-$f$ is \textbf{equivariant} when $f(act(g,x)) = act(g,f(x))$ for all $g$ in $G$ and for all $x$ in $X$\cite{nlab:equivariant, pittsNominalSetsNames2013}.
+$f$ is \textbf{equivariant} when $f(act(g,x)) = act(g,f(x))$ for all $g$ in $G$ and for all $x$ in $X$\cite{nlab:equivariant,pittsNominalSetsNames2013}.
  \end{definition}
@@
 
-The notion that visual encodings should be \textit{homomorphic} was proposed by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987} in his specification of \textit{A Presentation Tool} and the idea that visual transforms are \textit{equivariant} underlies Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}'s Algebraic Visualization Design.
+The notion that visual encoding functions should be \textit{homomorphisms} was proposed by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987} in his specification of \textit{A Presentation Tool} and the idea that visual transforms are \textit{equivariant} underlies Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}'s Algebraic Visualization Design.
 Kindlemann and Scheidegger propose three specific types of structure preservation: that the visualization should not change if the data representation (i.e.
@@
 The structure on values has traditionally been described  using the measurement scales-i.e.
-nominal, ordinal, interval, ratio - which Steven's classified by their mathematical group structure\cite{stevensTheoryScalesMeasurement1946a} and others have formally expanded to include more types \cite{leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}.
+nominal, ordinal, interval, ratio - which Steven's classified by their mathematical group structure\cite{stevensTheoryScalesMeasurement1946} and others have formally expanded to include more types \cite{leleaFormalizationMeasurementScale1971 thomasMathematizationNotMeasurement2014}.
 Generally the structure on each field of the dataset is preserved separately, i.e.
@@
 For the visual transformation to be equivariant, that nominal measurement must be mapped to a visual encoding that is also permutable, such as distinct marker shapes or color hues.
-We evaluate whether the encoding is homomorphic when the visual encoding does not have a group structure.
+We evaluate whether the encoding is a homomorphism when the visual encoding does not have a group structure.
 For example, partial orders are not invertible; therefore a visual encoding of partially ordered data is structure preserving when the ordering is mapped to equivalently ordered visual encodings.
@@
 The connectivity and relative positioning of elements in a dataset is broadly its topology\cite{wilkinsonGrammarGraphics2005} and the expectation that it be be reflected in the visual representation is present in even the earliest maps \cite{friendlyBriefHistoryData2008} and in Bertin's assertion that the dimensionality of the data match the class of representation \cite{bertinSemiologyGraphicsDiagrams2011}.
-In this work, we propose that using the formalism of topology provides us a uniform way to express dimensionality such that we can generalize the expectation that the data and visual dimensionalty match past points, lines, and areas.
+In this work, we propose that using the formalism of topology provides us a uniform way to express dimensionality such that we can generalize the expectation that the data and visual dimensionality match past points, lines, and areas.
 
@@
 A topology is a partitioning of a space into its subspaces in a way where those subspaces can be composed back into the space and the choice of partitions is how we encode the connectivity of the data that we are interested in.
-For example, the GHCN\cite{lawrimoreGlobalHistoricalClimatology2011} dataset in \autoref{fig:related-work:continuity:ktypes} is a complex spatio-temporal record of weather station data.
+For example, the GHCN\cite{lawrimoreGlobalHistoricalClimatology2011} dataset in \autoref{fig:related-work:continuity:ktypes} is a complex spatiotemporal record of weather station data.
 One way of encoding that each record is independent is to encode the underlying topology as a set of 0D $\cdots$ points, one for each record.
 Each of these discrete points is also represented visually as a discrete marker in the scatter plot.
-We can view the timeseries at each station as a 1D $---$ continuous line since weather station data is sampled from natural phenomena.
-We encode this topological structure as intervals, .i.e [start date, end date], and sets of sub-intervals, and we visually represent this strucure as continuous line plots.
+We can view the time series at each station as a 1D $---$ continuous line since weather station data is sampled from natural phenomena.
+We encode this topological structure as intervals, .i.e [start date, end date], and sets of sub-intervals, and we visually represent this structure as continuous line plots.
 A third way of viewing the data is as a sparse sampling of a 2D $\blacksquare$ geospatial grid, where the connectivity is modeled as a surface and sub regions of that surface.
@@
 While plotting the time series as a continuous line would be correct, it would be incorrect for a visualization to indicate that the distinct rows or stations are connected in a 1D continuous manner.
-Encoding the expected input structure would enable a visualization algorithm to reject data that does not explicitely state that it matches those assumptions.
+Encoding the expected input structure would enable a visualization algorithm to reject data that does not explicitly state that it matches those assumptions.
 This structure can be encoded in a generalizable way by building on Butler's work on using a mathematical structure called fiber bundles as an abstract data representation in visualization \cite{butlerVectorBundleClassesForm1992, butlerVisualizationModelBased1989} where he discusses how bundles can be used to encode data in a consistent manner for the topological and field structures that are common in visualization.
@@
 \label{sec:atct}
-In this section, we propose a mathematical abstraction of the data input and graphic prerendered output.
+In this section, we propose a mathematical abstraction of the data input and graphic pre-rendered output.
 This mathematical abstraction has a robust language for expressing topology and fields; expresses how to verify that data continuity is preserved on subset, distributed, and streaming data representations; and formalizes the expectation of a correspondence between data and visual elements.
@@
 %For example, in \autoref{fig:local_trivalizations}, the bundles $\dtotal_{0}=\dtotal\restriction_{[\epsilon^{-}, {\frac{2\pi}{5}}^{+}]}$ and $\dtotal_{1}=\dtotal\restriction_{[{\frac{2\pi}{5}}^{-}, \epsilon^{-}]}$ are local trivializations of an arbitrary bundle $\dtotal$.
-These local trivializations overlap on the intersections of the open subsets of each trivalization $[\epsilon^{-}, \epsilon^{+}]$ and $[{\frac{2\pi}{5}}^{-}, {\frac{2\pi}{5}}^{+}]$.
+These local trivializations overlap on the intersections of the open subsets of each trivialization $[\epsilon^{-}, \epsilon^{+}]$ and $[{\frac{2\pi}{5}}^{-}, {\frac{2\pi}{5}}^{+}]$.
 A set of trivializations can be glued into a larger bundle by defining transition maps that describe how to align the fibers in the overlapping space such that the sections of the bundle remain continuous.
@@
 
-%In a \textit{trivial bundle}, the fibers are identical $\dfiber = \dfiber_{k}$ for all points $\dbasepoint \in dbase$ such that all the transtion maps are also identical.
-For example, in the cylinder case the fibers in the overlappping regions of the trivalizations point in the same direction $\uparrow$ to illustrate that they are equal $\dfiber_0\restriction_{\openset_1\cap\openset_2} = \dfiber_{1}\restriction_{\openset_1\cap\openset_2}$.
+%In a \textit{trivial bundle}, the fibers are identical $\dfiber = \dfiber_{k}$ for all points $\dbasepoint \in dbase$ such that all the transition maps are also identical.
+For example, in the cylinder case the fibers in the overlapping regions of the trivializations point in the same direction $\uparrow$ to illustrate that they are equal $\dfiber_0\restriction_{\openset_1\cap\openset_2} = \dfiber_{1}\restriction_{\openset_1\cap\openset_2}$.
 Because the fibers point in the same direction, the  transition maps at both intersections map the fiber to itself $\uparrow\rightarrow\uparrow$.
@@
 \note{base space of bundle for key/value pair, topology in general 'cause data + viz continuity in same way}
-Munzner~\cite{munznerChDataAbstraction} describes a key-value semantics for data representation and abstraction, where datasets can be queried with different kinds of keys depending on the dataset type - locations for values of spatial fields, times for values of timeseries, or indexes for values in flat or multidimensional tables.
-We propose that these different types of keys are instances of different underlying topological organization of the data, and that we can use the structure of the opensets in the topological space to encode the extent to which the keys can vary continuously: flat or multidimensional tables have discrete points as indices, timeseries have points on a line as indices and spatial fields (regardless of whether they are scalar or tensor fields) have points in a spatial domain as indices.
+Munzner~\cite{munznerChDataAbstraction} describes a key-value semantics for data representation and abstraction, where datasets can be queried with different kinds of keys depending on the dataset type - locations for values of spatial fields, times for values of time series, or indexes for values in flat or multidimensional tables.
+We propose that these different types of keys are instances of different underlying topological organization of the data, and that we can use the structure of the opensets in the topological space to encode the extent to which the keys can vary continuously: flat or multidimensional tables have discrete points as indices, time series have points on a line as indices and spatial fields (regardless of whether they are scalar or tensor fields) have points in a spatial domain as indices.
 Therefore we choose to encode the topological structure of the data as the \textcolor{base}{base space} of a fiber bundle.
 Because the base space of a fiber bundle is a quotient topology\cite{munkresElementsAlgebraicTopology1984}, it divides the topological space into the largest number of open sets such that $\pi$ remains a continuous function.
-This means that the topology can be defined to have a resolution equal to the number of indicies in a dataset such that the key (continuity)-value (data) pairing is always preserved.
-
-Following from Spivaks cataegorical abstraction of a database \cite{spivakSimplicialDatabases2009,spivakDatabasesAreCategories2010}, we also propose that the structure of the data types be formally specified as the objects of a category.
+This means that the topology can be defined to have a resolution equal to the number of indices in a dataset such that the key (continuity)-value (data) pairing is always preserved.
+
+Following from Spivak's categorical abstraction of a database \cite{spivakSimplicialDatabases2009,spivakDatabasesAreCategories2010}, we also propose that the structure of the data types be formally specified as the objects of a category.
 
@@
 The composability property expresses that inclusion is transitive, while associativity expresses that the inclusion functions can be curried in various equivalent groupings.
-By formally specifying the properties of the topological structure datatypes as $\mathcal{\dbase}$, we can express that these are the properties that are required as part of the implementation of the data type objects.
-
-For example, the disjoint union of two bundles, as shown in \autoref{fig:atct:local_trivalizations}, is the coproduct $\dbase^{a} \subseteq_{\dbase^{c}} \dbase^{b}$ over an overlap $\dbase^{c}$ and therefore the inclusion morphism must be commmutative:
+By formally specifying the properties of the topological structure data types as $\mathcal{\dbase}$, we can express that these are the properties that are required as part of the implementation of the data type objects.
+
+For example, the disjoint union of two bundles, as shown in \autoref{fig:atct:local_trivalizations}, is the coproduct $\dbase^{a} \subseteq_{\dbase^{c}} \dbase^{b}$ over an overlap $\dbase^{c}$ and therefore the inclusion morphism must be commutative:
 \begin{minipage}[c]{0.25\columnwidth} 
@@
 \end{equation}
-\end{minipage}\hfil
+\end{minipage}\hfill
+
 \begin{minipage}[c]{0.65\columnwidth}
+  \note{will be replaced by lines showing intervals}
   \begin{tikzpicture}[commutative diagrams/every diagram]
@@
 such that the preimage of a point is the corresponding data type domain $\pi^{-1}(\dbasepoint) = \dfiber_{k} = \mathscr{U}_{{\sigma}_{\dbasepoint}}$. 
-Adopting Spivak's fiber bundle construction of types allows our model to reuse types so long as the field names are distinct and that field values can be accesed by field name,  since those are sections on $\mathcal{U}_{\sigma}$.
+Adopting Spivak's fiber bundle construction of types allows our model to reuse types so long as the field names are distinct and that field values can be accessed by field name,  since those are sections on $\mathcal{U}_{\sigma}$.
 Furthermore, since domains $\mathscr{U}_{{\sigma}}$ of types are a mathematical space, multi-dimensional fields can be encoded in the same manner as single dimensional fields and fields can have different names but the same type.
@@
 We can also equip the category with any operators or relations that are part of he mathematical structure of the field type.
-For example we can equip the category with a comparison operator, which is part of the definition of the monoidal structure of a partially ordered ranking variable \cite{bruggemannRankingPrioritizationMultiindicator2011} or the group structure of Steven's ordinal measurment scale \cite{stevensTheoryScalesMeasurement1946, leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}.
+For example we can equip the category with a comparison operator, which is part of the definition of the monoidal structure of a partially ordered ranking variable \cite{bruggemannRankingPrioritizationMultiindicator2011} or the group structure of Steven's ordinal measurement scale \cite{stevensTheoryScalesMeasurement1946, leaformalizationmeasleaFormalizationMeasurementScale1971, thomasMathematizationNotMeasurement2014}.
 Steven's other scales are summarized in \autoref{tab:appendix:summary:stevens}.
@@
 The fiber category $\mathcal{\dfiberc}$ is also equipped with a bifunctor because it is a monoidal category and this functor provides a method for combining fiber types.
-The bifunctor $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$ allows us to express fields that contain multityped values.
+The bifunctor $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$ allows us to express fields that contain complexly typed values.
 For example, wind can be represented as two fields $\dfiber_{speed} \times \dfiber_{direction}$ or a composite fiber field $\dfiber_{speed} \otimes \dfiber_{direction} = \dfiber_{wind}$.
@@
 Combining fibers can be verified by checking that when a fiber component $\dfiber^{c}$ is present in both $\dfiber^{a}$ and $\dfiber^{b}$, it is identical when projected out of either such that the *product* diagram commutes:
-\begin{minipage}
-
-\begin{tikzcd}[c]{.25\columnwidth}
-  \dfiberc^{a}\times_{\dfiber^{c}} \dfiberc^{b} \arrow[d, "\pi_a"] \arrow[r, "\pi_b"] & \dfiberc^{b} \arrow[d, "\pi_c"] & {(temperature, time) \times_{time} (pressure, time)} \arrow[d, "\pi_a"] \arrow[r, "\pi_b"] & {(pressure,time)} \arrow[d, "\pi_c"] \\
-  \dfiberc^{a} \arrow[r, "\pi_c"]                                                     & \dfiberc^{c}                    & {(temperature,time)} \arrow[r, "\pi_c"]                                                    & time
+
+\begin{minipage}[c]{.25\columnwidth}
+\begin{tikzcd}
+  \dfiberc^{a}\times_{\dfiber^{c}} \dfiberc^{b} \arrow[d, "\pi_a"] \arrow[r, "\pi_b"] 
+  & \dfiberc^{b} \arrow[d, "\pi_c"] 
+  & {(temperature, time) \times_{time} (pressure, time)} \arrow[d, "\pi_a"] \arrow[r, "\pi_b"] 
+  & {(pressure,time)} \arrow[d, "\pi_c"] \\
+  \dfiberc^{a} \arrow[r, "\pi_c"] 
+  & \dfiberc^{c}
+  & {(temperature,time)} \arrow[r, "\pi_c"]
+  & time
   \end{tikzcd}
 \end{minipage}\hfill
-\begin{minipage}[c]{.65\columwidth}
-
+\begin{minipage}[c]{.65\columnwidth}
+  \note{will be little pictures like in 7 sketchs}
 \end{minipage}
 
-This means that when multifield data is decomposed into its component fields, shared records stay the same.
-For example, the red (temperature, time, pressure) record seperates into (tempereture, time) and (pressure, time) records that share the same red time.
+\note{this is really a bookkeeping thing - each field is at an index/ splitting a field shouldn't change values} This means that data with many fields is decomposed into its component fields, shared records stay the same.
+For example, the red (temperature, time, pressure) record separates into (temperature, time) and (pressure, time) records that share the same red time.
 Furthermore this time is the same whether it is obtained from the (temperature, time) or (pressure, time) record.
@@
   \caption{The table from \autoref{fig:related-work:continuity:ktypes} has a 3 dimensional fiber (name, temperature, precipitation), a 0D base space, and each row is a section.
-Each time series is a section of a bundle with a 2D fiber (time, precipitation) and a 1D base spac encoding temporal continuity.
+Each time series is a section of a bundle with a 2D fiber (time, precipitation) and a 1D base space encoding temporal continuity.
 Each location of the rain map is a section of a bundle with a 2D plane base encoding spatial continuity and a 3D fiber space encoding (latitude, longitude, precipitation)
@@
 In the bundle encoding of the table, the indexing space is arbitrarily numbered keys; in the time series bundle, the base space is the interval [0,1]; and the map is sparse samples from a continuous space $[0,1]^{2}$ In contrast to a notion of a semantic binding between indexing space and field values, as proposed by Munzner\cite{munznerVisualizationAnalysisDesign2014}, the fields describing the continuity are part of the fiber.
-For example, the time is a fiber in the timeseries and the latitude and longitude are part of the map's fiber.
+For example, the time is a fiber in the time series and the latitude and longitude are part of the map's fiber.
 This separation between connectivity and what it is called means the time or location can change units without a change to structure.
@@
   \includegraphics[width=1\columnwidth]{fb_rho.png}
-  \caption{For a 2D display, a section $\gsection$ maps from each point $\gbasepoint$ into a fiber $\gfiber$ that encodes an RGB prerender space with infinite resolution $\mathcal{R}^{2}$.
+  \caption{For a 2D display, a section $\gsection$ maps from each point $\gbasepoint$ into a fiber $\gfiber$ that encodes an RGB pre-render space with infinite resolution $\mathcal{R}^{2}$.
 Each tiny colored box is an approximation of the return value of the same section function $\gsection$ evaluated on different points $\gbasepoint \in \gbase$ in the base space.
@@
 
-In \autoref{fig:atct:fb:graphic}, the section function $\gsection$ maps into the fiber for a simplified 2D RGB infinite resolution prerender space and returns the $\{x,y,r,g,b\}$ values of a pixel in an infinite resolution space.
+In \autoref{fig:atct:fb:graphic}, the section function $\gsection$ maps into the fiber for a simplified 2D RGB infinite resolution pre-render space and returns the $\{x,y,r,g,b\}$ values of a pixel in an infinite resolution space.
 In \autoref{fig:atct:fb:graphic} these pixels are approximated as the small orange and black colored boxes.
@@
 Sheaves are an algebraic data structure that provides a way of abstractly discussing the bookkeeping that data containers must implement to keep track of the continuity of the data \cite{ghristElementaryAppliedTopology2014}.
-This abstraction facilitates representational invariance, as introduced by Kindlemannn and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}, since the container level is uniformally specified as satisfying sheaf constraints.
-These constraints generalize to data that is subsetted, distributed, streaming, and on-demand.
-
-We can mathematically encode that we expect data containers to preserve the underlying continuity of the indexing space and the mappings beween indexing space and record space using a type of function called a functor.
+This abstraction facilitates representational invariance, as introduced by Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}, since the container level is uniformly specified as satisfying sheaf constraints.
+These constraints generalize to data that is subset\note{one is not like the others}, distributed, streaming, and on-demand.
+
+We can mathematically encode that we expect data containers to preserve the underlying continuity of the indexing space and the mappings between indexing space and record space using a type of function called a functor.
 Functors are mappings between categories that preserve the domains, codomains, composition, and identities of the morphisms within the category\cite{riehlCategoryTheoryContext}.
@@
 \end{definition}
-Contravariance means that the morphisms between the input openset objects go in the opposite direction from the morphisms between the output set objects.
+A functor is contravariant when the morphisms between the input objects go in the opposite direction from the morphisms between the output objects.
 The presheaf is contravariant because the inclusion morphisms between input objects
@@
 We construct the map as going from graphic to data because that encodes the notion that every visual element traces back to the data in some way.
-As exemplified in \autoref{fig:atct:morphisms:sheaf}, we define $\vindex$ as a surjective map because it allows us to express that a union of graphic spaces $S_i$ maps to single data point $\dbasepoint$, which allows us to express visual representations of a single record that are the union of many primatives, discussed in \autoref{eq:artist:operator}, such as multipart glyphs (e.g boxplots) and combinations of plot types (e.g line with point markers).
+As exemplified in \autoref{fig:atct:morphisms:sheaf}, we define $\vindex$ as a surjective map because it allows us to express that a union of graphic spaces $S_i$ maps to single data point $\dbasepoint$, which allows us to express visual representations of a single record that are the union of many primitives, discussed in \autoref{eq:artist:operator}, such as multipart glyphs (e.g boxplots) and combinations of plot types (e.g line with point markers).
 
@@
 This allows us to describe what data we expect at each graphic index location and what graphic is expected at each data index location.
-Transport functors compose the indexing map \vindex\ with the sheave map to say that a record \dsection\ at \dbasepoint\ is at all corrresponding \gbasepoint\ and that a function \gsection\ over one point \gbasepoint\ is the same function at all points $\gbasepoint \in \gbase$ that correspond to the same record index \dbasepoint.
+Transport functors compose the indexing map \vindex\ with the sheave map to say that a record \dsection\ at \dbasepoint\ is at all corresponding \gbasepoint\ and that a function \gsection\ over one point \gbasepoint\ is the same function at all points $\gbasepoint \in \gbase$ that correspond to the same record index \dbasepoint.
 
@@
 We choose to visualize this as a circle and two line plots.
-The indexing function \vindex\ bookkeeps which parts of the circle and each curve correspond to each point on the unit circle.
+The indexing function \vindex\ book keeps \note{find better word than bookkeeping} which parts of the circle and each curve correspond to each point on the unit circle.
 The pushforward $\vindexpush$ matches each point in the data space to the specification of the graphic at that point, while the pullback $\vindexpull$ matches each point in the graphic space to the data over that point.
@@
 The index lookup $\vindex$ describes which parts of the circle and curves are generated from which points on the unit circle.
-Given this correspondance, the pullback $\vindexpull\dsection$ looks up which values are being represented in a given part of the graphic.
+Given this correspondence, the pullback $\vindexpull\dsection$ looks up which values are being represented in a given part of the graphic.
 This type of lookup is critical for interactive techniques such as brushing, linking, and tooltips\cite{beckerBrushingScatterplots1987}.
@@
  The diagram in \autoref{eq:artist:natural_transform:inclusions} shows that restricting a set of outputs of an artist to a set of graphic sections over a subspace is equivalent to restricting the inputs to data sections over the same subspace.
-Because the artist is a functor of sheaves, the artist is expected to tra nslate the data continuity to graphic continuity such that the connectivity of subsets is preserved.
+Because the artist is a functor of sheaves, the artist is expected to translate the data continuity to graphic continuity such that the connectivity of subsets is preserved.
 This bookkeeping is necessary for any visualization technique that selectively acts on different pieces of a data set; for example streaming visualizations \cite{krstajicVisualizationStreamingData2013} and panning and zooming \cite{NekrasovskiEvaluationPanZoom2006}
@@
 \noindent
-The base space transformation is a pointwise continuous map from one open set to another open set in the same base space
+The base space transformation is a point wise continuous map from one open set to another open set in the same base space
 \begin{equation}
@@
 a morphism on the fiber category $\dfunct \in Hom(\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}},\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}})$ restricted to a point $\dbasepoint^{\prime} \in \openset^{\prime}$.
-This means $\dfunct$ has to statisfy the properties of a morphism (\autoref{def:atct:category})
+This means $\dfunct$ has to satisfy the properties of a morphism (\autoref{def:atct:category})
 \begin{itemize}
@@
 The distinguishing characteristic of $\vtotal$ is that it is part of the construction of the artist and therefore a part of the visualization library implementation.
-We propose that reusing the fibers $\vfiber$ across components facilitates standardizing internal types across the library and that this standardization improves maintainablity (\autoref{tab:appendix:library_spec}).
+We propose that reusing the fibers $\vfiber$ across components facilitates standardizing internal types across the library and that this standardization improves maintainability (\autoref{tab:appendix:library_spec}).
 
@@
   \includegraphics[width=1\columnwidth]{deform_retract.png}
-  \caption{The graphic base space $\gbase$ is collapsable to the line $\dbase$ such that every band $(\dbasepoint_i, [0,1])$ on $\gbase$ maps to corresponding point $\dbasepoint_i \in \dbase$.
+  \caption{The graphic base space $\gbase$ is collapsible to the line $\dbase$ such that every band $(\dbasepoint_i, [0,1])$ on $\gbase$ maps to corresponding point $\dbasepoint_i \in \dbase$.
 The band $[0,1]$ determines the thickness of a rendered line for a given point $\dbasepoint_{i}$ by specifying how pixels corresponding to that point are colored.
@@
 \end{equation}
-that preserve the categorical structure (operators and morphisms) of the fiber and the contininuity of the data section.
+that preserve the categorical structure (operators and morphisms) of the fiber and the continuity of the data section.
 As mentioned in \autoref{sec:construction:vtotal}, the total visual space is restricted to the space of data types internal to the library $\vfiber \subset \measure$ and sections are subsets of homsets $\Gamma(\dbase, \vtotal) \subset Hom(\dbase, \measure)$ because sections must be continuous.
 
-The encodeding functions $\vchannel$ are fiberwise transforms such that $\pi(\dtotal) = \pi(\vchannel(\dtotal))$.
-A consequence of this property is that $\vchannelc$ can be constructed as a pointwise transformation such that
+The encoding functions $\vchannel$ are fiber wise transforms such that $\pi(\dtotal) = \pi(\vchannel(\dtotal))$.
+A consequence of this property is that $\vchannelc$ can be constructed as a point wise transformation such that
 \begin{equation}
@@
 \end{equation}
-As shown in \autoref{eq:construction:nu:validate}, an encoder is considered valid if there is an ismorphism between the actual outputted visual component and the expected measurable component encoding.
+As shown in \autoref{eq:construction:nu:validate}, an encoder is considered valid if there is an isomorphism between the actual outputted visual component and the expected measurable component encoding.
 An encoder is consistent if it encodes the same field in the same way even if coming from different data sources.
@@
 An encoding function $\vchannelc$ is equivariant if the change in data, as defined in \autoref{sec:artist:equivariant:data}, and change in visual components are equivariant.
-Since $\dtotal$ and $\vtotal$ are over the same base space and are pointwise, the base space change $\dfunch_{\dtotal}$ applies to both sides of the equation
+Since $\dtotal$ and $\vtotal$ are over the same base space and are point wise, the base space change $\dfunch_{\dtotal}$ applies to both sides of the equation
 \begin{equation}
@@
 As mentioned in \autoref{sec:related-work:equivariance}, this is why Bertin and many others since have advocated choosing an encoding that has a structure that matches the data structure\cite{bertinSemiologyGraphicsDiagrams2011}.
-For example choosing a quantative colormap to encode quantative data if the $\dfunctc$ operation is scaling, as in \autoref{fig:artist:equivariance}.
+For example choosing a quantitative color map to encode quantitative data if the $\dfunctc$ operation is scaling, as in \autoref{fig:artist:equivariance}.
 
@@
 
-Since encoder functions are infinitely composable, as described in \autoref{eq:construction:nu:fabrication}, a new compositor function $\vmarkc$ can be constructed by precomposing $\vchannelc$ functions with the existing $\vmarkc$.
+Since encoder functions are infinitely composable, as described in \autoref{eq:construction:nu:fabrication}, a new compositor function $\vmarkc$ can be constructed by pre=composing $\vchannelc$ functions with the existing $\vmarkc$.
 
@@
 
-Since the artist is a family of functions in the homeset between sheaves, the isomorphism allows for the specification of the transformation from data as combination of functions over different spaces such that the following diagram commutes:
+Since the artist is a family of functions in the homset between sheaves, the isomorphism allows for the specification of the transformation from data as combination of functions over different spaces such that the following diagram commutes:
 \begin{equation}
@@
 when $\vindex(\gbasepoint) = \dbasepoint$.
-This equivalence allows a  developer to connect transformations over data space, denoted with a subset $\dbase$, with transformations over graphic space $\gbase$, using $\vindexpushc$ and $\vindexpullc$ adaptors.This allows developers to for example connect transformers that transform data on a line to a color in dataspace, but build a line compositing function that dynamically resamples what is on screen in graphic space.
+This equivalence allows a  developer to connect transformations over data space, denoted with a subset $\dbase$, with transformations over graphic space $\gbase$, using $\vindexpushc$ and $\vindexpullc$ adaptors.This allows developers to for example connect transformers that transform data on a line to a color in data space, but build a line compositing function that dynamically resamples what is on screen in graphic space.
 
@@
 
-\begin{figure}[h!]
+\begin{figure}[H]
   \centering
@@
     &\textcolor{base}{point}/\textcolor{base}{openset}/\textcolor{base}{base space} & \textcolor{fiber}{fiber space} & \textcolor{total}{total space}\\
-     &  location/subset/indicies & record/fields &  dataset type\\
+     &  location/subset/indices & record/fields &  dataset type\\
     \hline
@@
 \caption{Functions $\dfunc=(\dfunchc, \dfunctc)$ for modifying data records.
-Equivalent constructions can be applied to elements in visual and graphic sheaves, and these functions are distinguised through subscripts $\dfunc_{\dtotal}$, $\dfunc_{\vtotal}$ and $\dfunc_{\gtotal}$}
+Equivalent constructions can be applied to elements in visual and graphic sheaves, and these functions are distinguished through subscripts $\dfunc_{\dtotal}$, $\dfunc_{\vtotal}$ and $\dfunc_{\gtotal}$}
 \label{tab:appendix:summary:datamod}
@@
   \end{tabular}
-  \caption{The record transformer $\dfunctc$ must satisfy the constraints listed in \autoref{tab:appendix:summary:datamod} and $\dfunctc$ must also respect the mathamatical structure of $\dfiber$.
+  \caption{The record transformer $\dfunctc$ must satisfy the constraints listed in \autoref{tab:appendix:summary:datamod} and $\dfunctc$ must also respect the mathematical structure of $\dfiber$.
 This table lists examples of $\dfunct$ preserving one of the binary operators that are part of the definition of each of the Steven's measurement scale types\cite{stevensTheoryScalesMeasurement1946}}.
@@
   \hline 
-  Encoders  & & \\
+  Encode
     & & \\
-  \hline
-  Decompositor & & \\
+           & & \\
+  \hline
+  Decompose & & \\
               && \\
@@
 \pagebreak
-\section{Internel Library Specification}
+\section{Internal Library Specification}
 
@@
 % argument is your BibTeX string definitions and bibliography database(s)
-\bibliography{bib
+\bibliography{references}
 

### 2023-06-09 paper/main.tex
paper builds + wiki/nlab in acknowledgements

@@
 % *** MISC UTILITY PACKAGES ***
+
 \newcommand{\note}[1]{\textcolor{magenta}{#1}} 
@@
 The structure on values has traditionally been described  using the measurement scales-i.e.
-nominal, ordinal, interval, ratio - which Steven's classified by their mathematical group structure\cite{stevensTheoryScalesMeasurement1946} and others have formally expanded to include more types \cite{leleaFormalizationMeasurementScale1971 thomasMathematizationNotMeasurement2014}.
+nominal, ordinal, interval, ratio - which Steven's classified by their mathematical group structure\cite{stevensTheoryScalesMeasurement1946} and others have formally expanded to include more types
+ \cite{leaFormalizationMeasurementScale1971, thomasMathematizationNotMeasurement2014}.
 Generally the structure on each field of the dataset is preserved separately, i.e.
@@
 \note{transition sentence}
-In this work, we formally specify the conditions necessary to build structure preserving components using category theory, because it provides a rich language for expressing the structure of objects, functions between objects, and the constraints that must hold for these functions to compose \cite{wielsManagementEvolvingSpecifications1998,yorgeyMonoidsThemeVariations}.
+In this work, we formally specify the conditions necessary to build structure preserving components using category theory, because it provides a rich language for expressing the structure of objects, functions between objects, and the constraints that must hold for these functions to compose \cite{wielsManagementEvolvingSpecifications1998,yorgeyMonoidsThemeVariations2012}.
 A brief visualization oriented introduction to category theory is in Vickers et al \cite{vickersUnderstandingVisualizationFormal2013}, but they are applying category theory to semantic concerns about visualization design rather than library architecture.
@@
 We can also equip the category with any operators or relations that are part of he mathematical structure of the field type.
-For example we can equip the category with a comparison operator, which is part of the definition of the monoidal structure of a partially ordered ranking variable \cite{bruggemannRankingPrioritizationMultiindicator2011} or the group structure of Steven's ordinal measurement scale \cite{stevensTheoryScalesMeasurement1946, leaformalizationmeasleaFormalizationMeasurementScale1971, thomasMathematizationNotMeasurement2014}.
+For example we can equip the category with a comparison operator, which is part of the definition of the monoidal structure of a partially ordered ranking variable \cite{bruggemannRankingPrioritizationMultiindicator2011} or the group structure of Steven's ordinal measurement scale \cite{stevensTheoryScalesMeasurement1946, leaFormalizationMeasurementScale1971, thomasMathematizationNotMeasurement2014}.
 Steven's other scales are summarized in \autoref{tab:appendix:summary:stevens}.
@@
 
-Each section of a sheaf over a point returns a single record in the fiber.
-The sheaf over an open set $\openset$ surrounding a point $\dbasepoint$ is called a \textit{stalk}\cite{StalkSheaf2019, harder2008lectures}
+Each section of a sheaf over a point returns a single record in the fiber. The sheaf over an open set $\openset$ surrounding a point $\dbasepoint$ is called a \textit{stalk}\cite{harder2008lectures}
 \begin{equation}
@@
 % use section* for acknowledgment
-\ifCLASSOPTIONcompsoc
-  % The Computer Society usually uses the plural form
-  \section*{Acknowledgments}
-\else
-  % regular IEEE prefers the singular form
+
 \section*{Acknowledgment}
-\fi
-
+\note{Acknowledge all the actual people}
 The authors would like to thank...
+
+Hannah is also very grateful to the nlab and wikipedia contributors who wrote 
+articles explaining a lot of the more abstract math on which this paper is built.
 

### 2023-06-09 paper/main.tex
numberline

@@
 \newtheorem{definition}{Definition}[section]
+\newcommand{\definitionautorefname}{Definition}
 \theoremstyle{remark}
@@
 \usepackage{tikz}
-\usetikzlibrary{cd} % commutative diagrams
+\usetikzlibrary{cd, calc, positioning}
+\usetikzlibrary{patterns,shapes.arrows}
 \usepackage{pgfplots} %numberline
-\DeclareUnicodeCharacter{2212}{−}
 \usepgfplotslibrary{groupplots,dateplot}
-\usetikzlibrary{patterns,shapes.arrows}
+\pgfplotsset{
+  compat=1.15,
+   simple axis/.style={
+    scale only axis,
+    axis x line=bottom,
+    axis y line=none,
+    axis line style={latex-latex},
+    xticklabel style = {font=\footnotesize},
+    width=40pt,
+    height=20pt,
+    xtick = {0, 180, 360},
+    xticklabels = {$0$, $\pi$, $2\pi$},
+    xmin = 0, xmax=360,
+    ymin=0, ymax=1,
+    domain=0:360,
+    range=0:1,
+    tickwidth=0,
+    }
+}
 \pgfplotsset{compat=newest}
@@
 \label{sec:related-work:continuity}
-\begin{figure}[h!]
+\begin{figure}[H]
   \includegraphics[width=1\columnwidth]{k_different_types.png}
@@
 
-\begin{figure}[h!]
+\begin{figure}[H]
   \includegraphics[width=\columnwidth]{figures/local_trivalizations.png}
@@
 
-\begin{figure}[h!]
+\begin{figure}[H]
   \includegraphics[width=1\columnwidth]{figures/cyl_mob.png}
@@
 For example, the disjoint union of two bundles, as shown in \autoref{fig:atct:local_trivalizations}, is the coproduct $\dbase^{a} \subseteq_{\dbase^{c}} \dbase^{b}$ over an overlap $\dbase^{c}$ and therefore the inclusion morphism must be commutative:
-\begin{minipage}[c]{0.25\columnwidth} 
+\begin{minipage}[c]{0.30\columnwidth} 
   \begin{equation}
@@
 \end{minipage}\hfill
-
-\begin{minipage}[c]{0.65\columnwidth}
-  \note{will be replaced by lines showing intervals}
+\begin{minipage}[c]{0.60\columnwidth}
   \begin{tikzpicture}[commutative diagrams/every diagram]
-    \matrix[matrix of math nodes, name=m, commutative diagrams/every cell]{
-      (0,1) & (.55,1) \\
-      (0,.45) & (.40, 60)\\  
-    };
+    \draw node[anchor=north west, minimum width=30pt, minimum height=10pt] (kabc) {};
+\begin{axis}[simple axis, at={(kabc)}]
+  %\addplot [thick, smooth] {sin(2*pi*x)};
+\end{axis}
+
+\draw node[anchor=north east, right=55pt of kabc, minimum width=30pt, minimum height=10pt] (kb) {};
+%\begin{axis}[simple axis,at={(kb)}]
+%  %\addplot [thick, smooth] {sin(2*pi*x)};
+%\end{axis}
+\draw[at=(kb)][draw=black](0,0) -- (360,0);
+\draw node[anchor=south west, below=of kabc, minimum width=30pt, minimum height=10pt] (ka) {};
+\begin{axis}[simple axis,at={(ka)}]
+  %\addplot [thick, smooth] {sin(2*pi*x)};
+\end{axis}
+
+\draw node[anchor=south east, below=of kb, minimum width=30pt, minimum height=10pt] (kc) {};
+\begin{axis}[simple axis, at={(kc)}]
+  %\addplot [thick, smooth] {sin(2*pi*x)};
+\end{axis}
 
 \path[commutative diagrams/.cd, every arrow, every label]
-    (m-2-2) edge[color=base] (m-1-2)
-            edge[color=base] (m-2-1)
-    (m-2-1) edge[color=base] (m-1-1)
-    (m-1-2) edge[color=base] (m-1-1);
+  (kc) edge[] (kb)
+       edge[] (ka)
+  (ka) edge[] (kabc)
+  (kb) edge[] (kabc);
+
   \end{tikzpicture}
+
 \end{minipage}
@@
 \begin{tikzcd}
+  \label{eq:atct:fiber_product}
   \dfiberc^{a}\times_{\dfiber^{c}} \dfiberc^{b} \arrow[d, "\pi_a"] \arrow[r, "\pi_b"] 
@@
 
-\begin{figure}[h!]
+\begin{figure}[H]
   \includegraphics[width=\columnwidth]{dbundle.png}
@@
 
-\begin{figure}[h!]
+\begin{figure}[H]
   \includegraphics[width=1\columnwidth]{fb_rho.png}
@@
 We construct the map as going from graphic to data because that encodes the notion that every visual element traces back to the data in some way.
-As exemplified in \autoref{fig:atct:morphisms:sheaf}, we define $\vindex$ as a surjective map because it allows us to express that a union of graphic spaces $S_i$ maps to single data point $\dbasepoint$, which allows us to express visual representations of a single record that are the union of many primitives, discussed in \autoref{eq:artist:operator}, such as multipart glyphs (e.g boxplots) and combinations of plot types (e.g line with point markers).
+As exemplified in \autoref{fig:atct:morphisms:sheaf}, we define $\vindex$ as a surjective map because it allows us to express that a union of graphic spaces $S_i$ maps to single data point $\dbasepoint$, which allows us to express visual representations of a single record that are the union of many primitives, such as multipart glyphs (e.g boxplots) and combinations of plot types (e.g line with point markers).
 
@@
 \subsubsection{Example: Graphic and Data}
-\begin{figure}[h!]
+\begin{figure}[H]
   \includegraphics*[width=1\columnwidth]{xi_sin.png}
@@
 take as input an openset object \opensetc\ or \opensetgc\ and return sets of data and graphic sections that are objects in \setb.
-As a map between these sheaf functors, the artist has to preserve the $\iota, \iota^*$ morphisms of the presheaf functor, described in \autoref{eq:atct:sheaf:inclusion} and \autoref{eq:atct:sheaf:restriction}, such that the following diagram commutes:
+As a map between these sheaf functors, the artist has to preserve the $\iota, \iota^*$ morphisms of the presheaf functor, described in \autoref{eq:atct:sheaf:inclusion} and
+\autoref{eq:atct:sheaf:restriction}, such that the following diagram commutes:
 
@@
 
-\begin{figure}[h!]
+\begin{figure}[H]
   \includegraphics[width=\columnwidth]{phi.png}
@@
 
-\begin{figure}[h!]
+\begin{figure}[H]
   \includegraphics[width=1\columnwidth]{equivariance.png}
@@
 \end{equation*}
-which following from an adjoint property of homsets \note{find citation}
-\begin{equation*}
+
+which following from an adjoint property of homsets \note{find citation and push this into a footnote or appendix maybe} 
+\begin{equation}
   Hom(\gbase, \gfiber) \times Hom(\gbase, \gfiber) = Hom(\gbase, \gfiber\times \gfiber)
-\end{equation*}
+\end{equation}
+
 which means that the artists on the subsets of fibers can be defined 
@@
 \begin{equation}
+  \label{eq:artist:join_base_fiber}
   \vartistc(\cgamma{\mathop{\sqcup}_{i} \dbasec^{i}}{\mathop{\oplus}_{i}\dtotalc^{i}}) \coloneqq \sum_{i}
@@
 
-As shown in \autoref{eq:artist:operator}, bundles over a union of base spaces can be joined as a product of the fibers.
+As shown in \autoref{eq:artist:join_base_fiber}, bundles over a union of base spaces can be joined as a product of the fibers.
 This allows us to consider all the data inputs in a complex visualization as a combined input, where some sections evaluate to null in fields for which there are no values for that point in the combined base space $\dbasepoint \in \mathop{\sqcup}_{i} \dbase^{i}$ The combined construction of the data is a method for expressing what each data input has in common with another data input-for example the data for labeling tick marks or legends-
@@
 
-\begin{figure}[h!]
+\begin{figure}[H]
 
@@
 
-\begin{figure}[h!]
+\begin{figure}[H]
   \includegraphics[width=1\columnwidth]{full_scatter.png}
@@
 \subsection{Bundle Inspired Data Containers}
-\begin{table}[h!]
+\begin{table}[H]
   \centering
@@
 For example a pandas data frame must be unpacked into its columns before it is sent into most artists and continuity is implicit in the columns being the same length rather than a tracked base space $\dbase$.
-Because it is more efficient to work with the data in column order, we often treat the data as a collection of single fiber bundles.
-This is equivalent to the total bundle, as shown in \autoref{eq:artist:operator}.
+Because it is more efficient to work with the data in column order, we often project the fiber down into individual components.
+As shown in \autoref{eq:atct:fiber_product}, we can verify that this projection is correct by checking that the values at the index are the same regardless of the level of decomposition.
 
@@
 
-\begin{table}[h!]
+\begin{table}[H]
   \centering
@@
 
-\begin{table}[h!]
+\begin{table}[H]
 \centering
@@
 
-\begin{table}[h!]
+\begin{table}[H]
   \renewcommand{\arraystretch}{1.5}
@@
 \note{color}
-\begin{table}[h!]
+\begin{table}[H]
   \centering

### 2023-08-08 paper/main.tex
presheaf and coproduct diagrams

@@
 \usepackage{tikz}
-\usetikzlibrary{cd, calc, positioning}
-\usetikzlibrary{patterns,shapes.arrows}
-\usepackage{pgfplots} %numberline
-\usepgfplotslibrary{groupplots,dateplot}
-\pgfplotsset{
-  compat=1.15,
-   simple axis/.style={
-    scale only axis,
-    axis x line=bottom,
-    axis y line=none,
-    axis line style={latex-latex},
-    xticklabel style = {font=\footnotesize},
-    width=40pt,
-    height=20pt,
-    xtick = {0, 180, 360},
-    xticklabels = {$0$, $\pi$, $2\pi$},
-    xmin = 0, xmax=360,
-    ymin=0, ymax=1,
-    domain=0:360,
-    range=0:1,
-    tickwidth=0,
-    }
-}
-\pgfplotsset{compat=newest}
+\usetikzlibrary{cd}
 
@@
 
-\IEEEPARstart{V}{isualization} is a structure preserving mapping from data to graphic.
-One way of verifying that the structure of the data is present in the graphic is to verify that the data set can be distilled from the graph, i.e.
-that the values read off a bar chart are equivalent to the values used to construct the bar chart.
-Since structure preservation requires that the $data\leftrightarrow graphic$ be bidirectional, changes to either data or graphic are expected to propagate to the other side of the transformation\cite{kindlmannAlgebraicProcessVisualization2014}.
-In this paper, we propose a methodology for describing structure, verifying structure preservation, and specifying the conditions for constructing a structure preserving map between data and graphics.
-We use category theory, which is the math of structures and their relations, to explicitly specify what it means for a visualization mapping to preserve structure in an explicit, comprehensive, and consistent way.
-We encode the structure of the data and graphics in a uniform generalizable manner by using a mathematical structure from algebraic topology called a fiber bundle; doing so allows us to discuss everything from lines to complex ND spaces with the same interface.
+\IEEEPARstart{V}{isualization} is a structure preserving mapping from data to graphic, which loosely means that the visualization retains how the data values are positioned relative to each other and the mathematical properties of those values.
+One way of verifying that the structure is preserved is to attempt to distill
+the data set from the graphic, i.e.
+check that the values read off a bar chart are equivalent to the values used to construct the chart.
+Structure preservation assumes that the $data\leftrightarrow graphic$ mapping is bidirectional, meaning changes to either data or graphic are expected to be reflected in the other\cite{kindlmannAlgebraicProcessVisualization2014}.
+
+ In this paper, we propose a methodology for describing structure, verifying structure preservation, and specifying the conditions for constructing a structure preserving map between data and graphics using topology and category theory.
+
+ We use category theory because it is the math of structures and their relations and therefore provides a robust framework for defining structure and expressing the conditions under which structure would be preserved.
+
+ We encode the structure of the data and graphics in a uniform generalizable manner ; doing so allows us to discuss everything from lines to complex ND spaces with the same interface.
 
@@
 For example, the disjoint union of two bundles, as shown in \autoref{fig:atct:local_trivalizations}, is the coproduct $\dbase^{a} \subseteq_{\dbase^{c}} \dbase^{b}$ over an overlap $\dbase^{c}$ and therefore the inclusion morphism must be commutative:
-\begin{minipage}[c]{0.30\columnwidth} 
+\begin{minipage}[c]{0.25\columnwidth} 
+  \centering
   \begin{equation}
     \label{eq:atct:base_union}
-  \begin{tikzcd}
+  \begin{tikzcd}[row sep=large]
     \dbasec^{a}\sqcup_{\dbasec^{c}}\dbasec^{b} 
@@
 \end{equation}
-\end{minipage}\hfill
-\begin{minipage}[c]{0.60\columnwidth}
-  \begin{tikzpicture}[commutative diagrams/every diagram]
-    \draw node[anchor=north west, minimum width=30pt, minimum height=10pt] (kabc) {};
-\begin{axis}[simple axis, at={(kabc)}]
-  %\addplot [thick, smooth] {sin(2*pi*x)};
-\end{axis}
-
-\draw node[anchor=north east, right=55pt of kabc, minimum width=30pt, minimum height=10pt] (kb) {};
-%\begin{axis}[simple axis,at={(kb)}]
-%  %\addplot [thick, smooth] {sin(2*pi*x)};
-%\end{axis}
-\draw[at=(kb)][draw=black](0,0) -- (360,0);
-\draw node[anchor=south west, below=of kabc, minimum width=30pt, minimum height=10pt] (ka) {};
-\begin{axis}[simple axis,at={(ka)}]
-  %\addplot [thick, smooth] {sin(2*pi*x)};
-\end{axis}
-
-\draw node[anchor=south east, below=of kb, minimum width=30pt, minimum height=10pt] (kc) {};
-\begin{axis}[simple axis, at={(kc)}]
-  %\addplot [thick, smooth] {sin(2*pi*x)};
-\end{axis}
-
-\path[commutative diagrams/.cd, every arrow, every label]
-  (kc) edge[] (kb)
-       edge[] (ka)
-  (ka) edge[] (kabc)
-  (kb) edge[] (kabc);
-
-  \end{tikzpicture}
-
 \end{minipage}
@@
 For example, lets define presheaves $\sheaf_1, \sheaf_2$. These are maps from intervals $\openset_1, \openset_2$ to a set of functions $\Gamma_1, \Gamma_2$ that are continuous over that interval: 
-\begin{equation*}
-\begin{tikzcd}
-  \textcolor{set}{\Gamma_2 = \textcolor{set}{\left\{\begin{array}{c} 
-      \textcolor{section}{constant}\\ 
-      \textcolor{section}{sin}\\
-      \textcolor{section}{cos}\\
-    \end{array}\right\}}}               
-  \arrow[rr, "\iota^*", color=set] &  & 
-  \textcolor{set}{\Gamma_1 = \left\{\begin{array}{c} 
-    \textcolor{section}{constant}\\ 
-    \textcolor{section}{sin}\\
-    \textcolor{section}{cos}\\
-    \textcolor{section}{tan}\\
-  \end{array}\right\}} \\
-  \textcolor{base}{\openset_2 = \left(0,2\pi\right)} 
-  \arrow[u, "\sheafc_{1}", maps to, color=sheaf, shift right=4.5]
-  &  & 
-  \textcolor{base}{\openset_1=\left(\frac{\pi}{2}, \frac{3\pi}{2}\right)} 
-  \arrow[ll, "\iota"', color=base] 
-  \arrow[u, "\sheafc_{2}"', maps to, color=sheaf, shift right=4.5]
-  \end{tikzcd}
-\end{equation*}
+\note{Human readable before inline not after}
+
+\begin{figure}[H]
+  \includegraphics*[width=1\columnwidth]{figures/tex/presheaf.pdf}
+  \caption*{presheafs $\sheaf_1, sheaf_2$ mapping from intervals to the trignometric functions defined over those intervals }
+\end{figure}
+
 The constraints of a presheaf functor are that since the $constant, sin, cos$ functions are defined over the interval $\left[0,1\right]$, these functions must also be continuous over the sub-interval $\left(\frac{\pi}{2}, \frac{3\pi}{2}\right)$; therefore the sections in $\Gamma_{2}$ must also be included in the set of sections over the subspace $\Gamma_{1}$.

### 2023-08-16 paper/main.tex
diagram figures + recreating data

@@
 For example, the disjoint union of two bundles, as shown in \autoref{fig:atct:local_trivalizations}, is the coproduct $\dbase^{a} \subseteq_{\dbase^{c}} \dbase^{b}$ over an overlap $\dbase^{c}$ and therefore the inclusion morphism must be commutative:
-\begin{minipage}[c]{0.25\columnwidth} 
-  \centering
-  \begin{equation}
-    \label{eq:atct:base_union}
-  \begin{tikzcd}[row sep=large]
-    \dbasec^{a}\sqcup_{\dbasec^{c}}\dbasec^{b} 
-    & \dbasec_{b} \arrow[l, "\iota_b"', color=base] \\
-    \dbasec_a \arrow[u, "\iota_a", color=base] 
-    & \dbasec_c 
-    \arrow[u, "\iota_c"', color=base] 
-    \arrow[l, "\iota_c"', color=base] 
-  \end{tikzcd}
-\end{equation}
-\end{minipage}
+\begin{figure}
+  \includegraphics*[width=1\columnwidth]{figures/tex/k_coproduct.pdf}
+\end{figure}
 
@@
 
-\begin{minipage}[c]{.25\columnwidth}
-\begin{tikzcd}
-  \label{eq:atct:fiber_product}
-  \dfiberc^{a}\times_{\dfiber^{c}} \dfiberc^{b} \arrow[d, "\pi_a"] \arrow[r, "\pi_b"] 
-  & \dfiberc^{b} \arrow[d, "\pi_c"] 
-  & {(temperature, time) \times_{time} (pressure, time)} \arrow[d, "\pi_a"] \arrow[r, "\pi_b"] 
-  & {(pressure,time)} \arrow[d, "\pi_c"] \\
-  \dfiberc^{a} \arrow[r, "\pi_c"] 
-  & \dfiberc^{c}
-  & {(temperature,time)} \arrow[r, "\pi_c"]
-  & time
-  \end{tikzcd}
-\end{minipage}\hfill
-\begin{minipage}[c]{.65\columnwidth}
-  \note{will be little pictures like in 7 sketchs}
-\end{minipage}
+\begin{figure}
+  \includegraphics*[width=1\columnwidth]{figures/tex/f_product.pdf}
+\end{figure}
 
@@
   \includegraphics*[width=1\columnwidth]{figures/tex/presheaf.pdf}
-  \caption*{presheafs $\sheaf_1, sheaf_2$ mapping from intervals to the trignometric functions defined over those intervals }
+  \caption*{presheafs $\sheaf_1, \sheaf_2$ mapping from intervals to the trignometric functions defined over those intervals }
 \end{figure}

### 2023-10-13 paper/main.tex
intro + ref

@@
 In this paper, we propose a methodology for describing structure, verifying structure preservation, and specifying the conditions for constructing a structure preserving map between data and graphics using topology and category theory.
-
 We use category theory because it is the math of structures and their relations and therefore provides a robust framework for defining structure and expressing the conditions under which structure would be preserved.
-
- We encode the structure of the data and graphics in a uniform generalizable manner ; doing so allows us to discuss everything from lines to complex ND spaces with the same interface.
+We encode the structure of the data and graphics in a uniform generalizable manner; doing so allows us to use the same framework for describing, for example, how tables and scatter plots and heatmaps
+are different views of the same underlying data or the ways in which a time series and map drawn from the same data cube are linked.
 

### 2023-11-03 paper/main.tex
put intro back together

--- 2023-08-16 paper/main.tex
@@
 
- In this paper, we propose a methodology for describing structure, verifying structure preservation, and specifying the conditions for constructing a structure preserving map between data and graphics using topology and category theory.
-
- We use category theory because it is the math of structures and their relations and therefore provides a robust framework for defining structure and expressing the conditions under which structure would be preserved.
-
- We encode the structure of the data and graphics in a uniform generalizable manner ; doing so allows us to discuss everything from lines to complex ND spaces with the same interface.
-
-We use category theory and algebraic structures because they can be directly translated into software design specifications and our goal is to improve visualization tools.
-The algebraic structures can be translated into analogous programmatic types, while category theory facilitates developing a functional design framework that describes function composition in terms of functional algebraic operations.
-The typing system and composition inherent in a functional design encourage library developers to build complex visualization components from simple verifiable parts that have few side effects\cite{huHowFunctionalProgramming2015, hughesWhyFunctionalProgramming1989}.
-These categorically specified components could be built as a standalone library and integrated into existing libraries because they are inherently self-contained.
-Furthermore we see the application of these ideas in low level visualization library design as our motivation, and we hope these ideas will help shape the reorganization of critical data visualization libraries, such as Matplotlib.
+We propose a methodology for describing structure, verifying structure preservation, and specifying the conditions for constructing a structure preserving map between data and graphics using topology and category theory.
+Using category theory and algebraic structures, we can encode the structure of data and graphics in a uniform manner.
+This uniformity allows us to develop a generalizable framework for specifying the conditions under which structure is preserved.
+For example, our methodology can be used to describe how a table and scatter plot and heatmap are different representations of the same data and track an observation from a datacube as a point along a time series and on a map and in a network.
+We developed a generalizable model because our goal is to improve data visualization software.
+The algebraic structures can be translated into analogous programmatic types, while category theory facilitates developing a functional design framework.
+Strong typing and function composition enable library developers to build complex components from simpler verifiable parts \cite{huHowFunctionalProgramming2015, hughesWhyFunctionalProgramming1989}.
+These  components can be built as a standalone library and integrated into existing libraries and we hope these ideas will help shape the reorganization of critical data visualization libraries, such as Matplotlib.
+
 The contribution of this paper is a mathematical framework for describing structure preserving visualization components in a manner that is generalizable to different types of inherent structure.

### 2023-11-22 paper/main.tex
giant xi overhaul

@@
 \newcommand{\note}[1]{\textcolor{magenta}{#1}}
+
 \usepackage{silence} 
-
 \WarningsOff[fixltx2e]
+
 \usepackage[switch]{lineno}
@@
 
-\IEEEPARstart{V}{isualization} is a structure preserving mapping from data to graphic, which loosely means that the visualization retains how the data values are positioned relative to each other and the mathematical properties of those values.
-One way of verifying that the structure is preserved is to attempt to distill
-the data set from the graphic, i.e.
-check that the values read off a bar chart are equivalent to the values used to construct the chart.
-Structure preservation assumes that the $data\leftrightarrow graphic$ mapping is bidirectional, meaning changes to either data or graphic are expected to be reflected in the other\cite{kindlmannAlgebraicProcessVisualization2014}.
-
-We propose a methodology for describing structure, verifying structure preservation, and specifying the conditions for constructing a structure preserving map between data and graphics using topology and category theory.
-Using category theory and algebraic structures, we can encode the structure of data and graphics in a uniform manner.
-This uniformity allows us to develop a generalizable framework for specifying the conditions under which structure is preserved.
-For example, our methodology can be used to describe how a table and scatter plot and heatmap are different representations of the same data and track an observation from a datacube as a point along a time series and on a map and in a network.
-We developed a generalizable model because our goal is to improve data visualization software.
-The algebraic structures can be translated into analogous programmatic types, while category theory facilitates developing a functional design framework.
+\IEEEPARstart{V}{isualization} tools implement transforms of the form $data \rightarrow graphic$ and are expected to be structure preserving.
+Loosely, that means the mathematical properties of the data and how the points are connected to each other should be inferable such that  $graphic \rightarrow data$.
+For example, values read off a bar chart have to be equivalent to the values used to construct that chart.
+Motivated by wanting to make better tools, we use algebraic topology to describe how $data$ and a corresponding $graphics$ have the same underlying structure and we use category theory to describing under what conditions the visualization tool preserves that structure such that $graphic \leftrightarrow data$.
+
+Algebraic structures allow us to describe many types of inherent structure in a uniform manner, and category theory provides a rigorous framework for specifying the behavior and properties of structure and maps between structures.
+For example, our framework can encapsulate how a table and scatter plot and heatmap are different representations of the same data and track an observation from a datacube as a point along a time series and on a map and in a network.
+The algebraic structures can be translated into programmatic types, while the categorical descriptions translate to a functional design framework.
 Strong typing and function composition enable library developers to build complex components from simpler verifiable parts \cite{huHowFunctionalProgramming2015, hughesWhyFunctionalProgramming1989}.
@@
 
-The contribution of this paper is a mathematical framework for describing structure preserving visualization components in a manner that is generalizable to different types of inherent structure.
+The contribution of this paper is a methodology for describing structure, verifying structure preservation, and specifying the conditions for constructing a structure preserving map between data and graphics.
 This framework also provides guidance for the construction and testing of structure preserving visualization library components.
@@
 \section{Related Work}
-We build on extensive work codifying the structure of visualization data and corresponding graphics and use the concepts of \textit{homomorphism} and \textit{equivariance} to express how that structure should be preserved.
-Structure is considered preserved when the input and output change in tandem, and these symmetries are often modeled as \textit{group actions}, which are a way for a group\footnote{A \textit{group} is a set with an associative binary operator, an identity, and an inverse operation} to transform some other set:
-
-\begin{definition}\label{def:related-work:action}\cite{nlab:action, grimaldiDiscreteCombinatorialMathematics2006}
-  An \textcolor{action}{\textbf{action}} \footnote{Throughout this paper, we augment the math with color to group conceptually similar objects and functions\cite{headMathAugmentationHow2022}, for example \textcolor{action}{actions} and \textcolor{action}{sets of actions}.
-This color coding carries through to the figures.} of \textcolor{action}{$G = (G,\circ, e)$} on $X$ is a function  $act: \textcolor{action}{G} \times X \rightarrow X$.
-An action has the properties of identity $act(\textcolor{action}{e}, x) = x$ for all  $x \in X$ and associativity $act(\textcolor{action}{g}, act(\textcolor{action}{f}, x)) = act(\textcolor{action}{f} \circ \textcolor{action}{g}, x)$ for $\textcolor{action}{f},\textcolor{action}{g} \in \textcolor{action}{G}$.
-\end{definition}
-
-Given a group $G$ that acts on both the input $X$ and the output $Y$ of a function $f: X \rightarrow Y$
-
- \begin{definition}
-$f$ is \textbf{equivariant} when $f(act(g,x)) = act(g,f(x))$ for all $g$ in $G$ and for all $x$ in $X$\cite{nlab:equivariant,pittsNominalSetsNames2013}.
- \end{definition}
-In an equivariant transform, mapping the data after acting on it $f(act)$ is equivalent to acting on the data after mapping it $act(f)$ since $G$ acts on both.
-A homomorphism does not require that the input and output are of the same type; instead
-
- \begin{definition}
-\note{must sort out exact appropriate definition of group homomorphism here}
- \end{definition}
-
-The notion that visual encoding functions should be \textit{homomorphisms} was proposed by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987} in his specification of \textit{A Presentation Tool} and the idea that visual transforms are \textit{equivariant} underlies Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}'s Algebraic Visualization Design.
-Kindlemann and Scheidegger propose three specific types of structure preservation: that the visualization should not change if the data representation (i.e.
-the data container) changes, that data maps unambiguously to visual elements \cite{ziemkiewiczEmbeddingInformationVisualization2009}, and that changing the data should correspond to changes in the visualization in a perceptually significant manner.
-A uniform abstract data representation layer is introduced in \autoref{sec:atct:sheaves}, the expectation of unambiguous mappings is expressed in \autoref{sec:atct:xi}, and the the correspondence between data and visual changes is included in the definition of equivariance introduced in \autoref{sec:artist:equivariant:artist}.
-We define structure preservation both in \note{field types and topology}
-
-\subsection{Field Type Structure} 
-\label{sec:related-work:equivariance}
- Bertin \cite{bertinSemiologyGraphicsDiagrams2011} codified the expectation that a data value and its visual encoding are expected to have compatible structure.
-The structure on values has traditionally been described  using the measurement scales-i.e.
-nominal, ordinal, interval, ratio - which Steven's classified by their mathematical group structure\cite{stevensTheoryScalesMeasurement1946} and others have formally expanded to include more types
- \cite{leaFormalizationMeasurementScale1971, thomasMathematizationNotMeasurement2014}.
-Generally the structure on each field of the dataset is preserved separately, i.e.
-each column in a table is mapped to a distinct encoding.
-For example, nominal measurements are permutable, meaning that they are acted on by elements of a permutation group.
-For the visual transformation to be equivariant, that nominal measurement must be mapped to a visual encoding that is also permutable, such as distinct marker shapes or color hues.
-We evaluate whether the encoding is a homomorphism when the visual encoding does not have a group structure.
-For example, partial orders are not invertible; therefore a visual encoding of partially ordered data is structure preserving when the ordering is mapped to equivalently ordered visual encodings.
-Steven's measurement scale's and the structure preservation they define is the basis of Wilkenson's scales\cite{wilkinsonGrammarGraphics2005} in Grammar of Graphics and derivative libraries, and in this work we generalize the concept of fields to support a wider array of field types.
-This also allows us to specify what consistency means when creating different views of data from a single dataset, as discussed in Qu and Hullman \cite{hullmanKeeping2018}, specifically in terms of each field type separate from the overall structure.
-
-\subsection{Topological Structure}
 \label{sec:related-work:continuity}
+
 \begin{figure}[H]
@@
 
-The connectivity and relative positioning of elements in a dataset is broadly its topology\cite{wilkinsonGrammarGraphics2005} and the expectation that it be be reflected in the visual representation is present in even the earliest maps \cite{friendlyBriefHistoryData2008} and in Bertin's assertion that the dimensionality of the data match the class of representation \cite{bertinSemiologyGraphicsDiagrams2011}.
-In this work, we propose that using the formalism of topology provides us a uniform way to express dimensionality such that we can generalize the expectation that the data and visual dimensionality match past points, lines, and areas.
+Since the earliest maps \cite{friendlyBriefHistoryData2006}, a visualization has been expected to preserve the $field$ properties and $topology$ of the corresponding dataset; a \textcolor{fiber}{\textcolor{field}} \footnote{Throughout this paper, we augment the math with color to group conceptually similar objects and functions\cite{headMathAugmentationHow2022}, for example \textcolor{action}{actions} and \textcolor{action}{sets of actions}.
+This color coding carries through to the figures.}  is a set of values of the same type, e.g.
+one column of a table or the pixels of an image, and \textcolor{base}{\textcolor{topology}} is the connectivity and relative positioning of elements in a dataset \cite{wilkinsonGrammarGraphics2005}.
+In \autoref{fig:related-work:continuity:ktypes}, the topology is a continuous 3 dimensional surface encapsulating time and position, while the fields are date, location, station, pressure, and temperature.
+Our paper provides a language to express how the scatter plot, line plot, and heatmap are preserving different structural properties of the dataset, thereby generalizing Bertin's codification of how the topology of observations must match the class of representation (i.e.
+point , line, area) and how the variation possible in a field match the variation in the corresponding visual variable \cite{bertinSemiologyGraphicsDiagrams2011} and Mackinlay's work on how structure preservation includes preserving the mathimatical relations between elements of the same and different fields \cite{mackinlayAutomatingDesignGraphical1986}.
+
+While Bertin and Mackinlay's work primarily discusses the $data \rightarrow graphic$, Kindlemann and Scheidegger's algebraic framework \cite{kindlmannAlgebraicProcessVisualization2014} also dictates a set of conditions under which the $graphic \rightarrow data$  mapping is structure preserving.
+Encapsulating their conditions, we: present a uniform abstract data representation layer in \autoref{sec:atct:sheaves} for ensuring that the visualization should not change if the data representation (i.e.
+the data container) changes, define the conditions under which data maps unambiguously to visual elements \cite{ziemkiewiczEmbeddingInformationVisualization2009} in \autoref{sec:atct:xi}, and provide a methodology for verifying that changes in data should correspond to changes in the visualization in \autoref{sec:artist:equivariant:artist} that does not necessarily require that the changes be perceptually significant.
+Furthermore, our model extends their conditions of equivariance by explicitly incorporating topology and by providing a framework for translating the theoretical ideas into buildable components in \autoref{sec:construction}.
+
+\subsection{Fields} 
+\label{sec:related-work:equivariance}
+The structure on values has traditionally been described using the measurement scales-i.e.
+nominal, ordinal, interval, ratio - which Steven's classified by their mathematical group structure\cite{stevensTheoryScalesMeasurement1946} and others have formally expanded to include more types \cite{leaFormalizationMeasurementScale1971, thomasMathematizationNotMeasurement2014}.
+Specifically, Steven's conceptualizes the structure on values as \textcolor{action}{actions} on groups \footnote{A \textit{group} is a set with an associative binary operator, an identity, and an inverse operation}:
+
+\begin{definition}\label{def:related-work:action}\cite{grimaldiDiscreteCombinatorialMathematics2006}
+  An \textcolor{action}{\textbf{action}} of \textcolor{action}{$G = (G,\circ, e)$} on $X$ is a function  $act: \textcolor{action}{G} \times X \rightarrow X$.
+An action has the properties of identity $act(\textcolor{action}{e}, x) = x$ for all  $x \in X$ and associativity $act(\textcolor{action}{g}, act(\textcolor{action}{f}, x)) = act(\textcolor{action}{f} \circ \textcolor{action}{g}, x)$ for $\textcolor{action}{f},\textcolor{action}{g} \in \textcolor{action}{G}$.
+\end{definition}
+
+where the measurements are the groups and changes to them are actions.
+Structure preservetion in fields means that the data and the visualization are equivalent representations of an underlyng phenomena such that changes to data or visualization be reflected in the other in a consistent way.
+A function that preserves structure in this manner for groups is called equivariant.
+Given a group $G$ that acts on both the input $X$ and the output $Y$ of a function $f: X \rightarrow Y$
+\begin{definition}
+ A function $f$ is \textbf{equivariant} when $f(act(g,x)) = act(g,f(x))$ for all $g$ in $G$ and for all $x$ in $X$ \cite{pittsNominalSetsNames2013}
+\end{definition}
+
+which means that a visualization is structure preserving when there exist compatible actions on the data and visualization, as proposed by Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}.
+Equivariance can sometimes be too restrictive because it can only be applied to groups and sometimes data does not have a group structure; an alternative is asserting that a visualization is structure preserving so long as it preserves the mathematical relations with and between fields, as proposed by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}, which is termed homomorphism.
+Given the function $f$ between sets $X$ and $Y$, both of which are equipped with the binary operator $\circ$
+\begin{definition}
+
+\end{definition}
+  A function $f$ is \textbf{homeomorphic} when $f(x\circ y) = f(x) \circ f(y)$ for every pair of elements $x,y$ in $X$
+\end{definition}
+which means that applying the operation $\circ$ to the input to $f$ is equivalent to applying the operation to the output, which is only possibly when the output of the map $f$ from $X$ to $Y$ has the same structure as the input.
+Our framework for structure preservation incorporates equivariance and homomorphism and allows for defining other measures of equivalent representation; it is discussed in detail in \autoref{sec:artist:equivariant:data}.
+
+\subsection{Topology}
+Borrowing strcutures from algebraic topology allows us to describe structure on individual elements in a way that holds true for the entire dataset, meaning we can use the same tools to check if structre is preserved regardless of whether the data fits in memory, is distributed, or is streaming.
+Our framework generalizesis because, informally, a topology $\mathcal{T}$ is a partitioning of a space into its subspaces in a way where those subspaces can be composed back into the space; the choice of partitions is how we encode the connectivity of the data and visualizations that we are interested in.
 
@@
 
-A topology is a partitioning of a space into its subspaces in a way where those subspaces can be composed back into the space and the choice of partitions is how we encode the connectivity of the data that we are interested in.
-For example, the GHCN\cite{lawrimoreGlobalHistoricalClimatology2011} dataset in \autoref{fig:related-work:continuity:ktypes} is a complex spatiotemporal record of weather station data.
-One way of encoding that each record is independent is to encode the underlying topology as a set of 0D $\cdots$ points, one for each record.
-Each of these discrete points is also represented visually as a discrete marker in the scatter plot.
-We can view the time series at each station as a 1D $---$ continuous line since weather station data is sampled from natural phenomena.
-We encode this topological structure as intervals, .i.e [start date, end date], and sets of sub-intervals, and we visually represent this structure as continuous line plots.
-A third way of viewing the data is as a sparse sampling of a 2D $\blacksquare$ geospatial grid, where the connectivity is modeled as a surface and sub regions of that surface.
-In the visualization, that connectivity is expressed as plotting the stations in the map, visually denoting that these are points in the continuous geographical space that is New York State.
-By using topology as the underlying encoding scheme, we are able to represent the structure of points, lines, surfaces, and more using the same abstractions.
-
-Since topological continuity is generalizable, we propose that encoding topology as part of a dataset type and a visualization type allows us to explicitly check which types of structure a visualization algorithm can preserve.
 Visual algorithms assume the topology of their input data, as described in taxonomies of visualization algorithms Chi\cite{chiTaxonomyVisualizationTechniques2000} and by Troy and M\"{o}ller \cite{toryRethinkingVisualizationHighlevel2004}, but generally do not verify that input structure.
 For example, a \texttt{line} algorithm often does not have a way to query whether a list of (x,y) coordinates is the distinct rows, the time series, or the list of stations in \autoref{fig:related-work:continuity:ktypes}.
-While plotting the time series as a continuous line would be correct, it would be incorrect for a visualization to indicate that the distinct rows or stations are connected in a 1D continuous manner.
-Encoding the expected input structure would enable a visualization algorithm to reject data that does not explicitly state that it matches those assumptions.
-This structure can be encoded in a generalizable way by building on Butler's work on using a mathematical structure called fiber bundles as an abstract data representation in visualization \cite{butlerVectorBundleClassesForm1992, butlerVisualizationModelBased1989} where he discusses how bundles can be used to encode data in a consistent manner for the topological and field structures that are common in visualization.
-We sketch out fiber bundles in \autoref{sec:atct:fiber-bundles}, but his work provides a thorough introduction to bundles for visualization practitioners.
+While plotting the time series as a continuous line would be correct, it would be incorrect for a visualization to indicate that the distinct rows or stations are connected in a 1D continuous manner because it introduces ambiguity over which part of the line maps back to the data.
+A map that by definition has continous maps between two spaces, such as data and graphics, is called a homeomorphism\cite{riehlCategoryTheoryContext}.
+
+\begin{definition}
+  A function $f$ is a $homeomorphism$ if it is bijective, continous, and has a continous inverse function f^{-1}. 
+\end{definition}
+
+To encode topology and field structure, we extend Butler's work on using a mathematical structure called fiber bundles as an abstract data representation in visualization \cite{butlerVectorBundleClassesForm1992, butlerVisualizationModelBased1989}.
+We sketch out fiber bundles in \autoref{sec:atct:fiber-bundles}, but Butler provides a thorough introduction to bundles for visualization practitioners.
 
@@
 \end{equation}
-between a graphic subspace $\opensetg \subseteq \gbase$ and data subspace $\openset \subseteq \dbase$.
-The functor $\vindex$ is surjective such that for every point $\dbasepoint \in \openset$ there is a corresponding set of points $\{\gbasepoint | \gbasepoint \in  \vindexpre(k)\}$ for all $\gbasepoint \in \opensetg$.
+between a graphic subspace $\opensetg \subseteq \gbase$ and data subspace $\openset \subseteq \dbase$. The functor $\vindex$ is surjective 
+such that we can identify the set of points in graphic space that correspond to each point in data space
+\begin{equation}
+  \label{eq:atct:xi:inverse}
+  \vindexprec(\dbasepointc) = \{\gbasepointc | \vindexc(\gbasepointc) = \dbasepointc \forall \dbasepointc \in \dbase, \gbasepointc \in \gbasec\}
+\end{equation}
+and every point in a graphic space has a corresponding point in data space.
 
@@
 For example, a section may go from every point in the graphic space to the same single point in the graphic fiber $\gsection(\gbasepoint_i) = d\; \forall \gbasepoint \in \gbase$ such that the visual output is a single inked pixel on a screen.
+
+\subsection{Homeomorphism}
+As mentioned in \autoref{sec:related-work:continuity}, preserving the topology of a visualization means that each discrete piece of differentiable visual information corresponds to a distince element of the dataset\cite{ziemkiewiczEmbeddingInformationVisualization2009} in a way where the organization of elements is preserved.
+A generalization of this condition is the idea that the graphic space can be collapased into the data indexing space, which means that the data base space is a deformation retraction of the graphic base space\cite{hatcherAlgebraicTopology2002}.
+By defining the indexing look up function \vindex, introduced in \autoref{sec:atct:sheaves}, to be
+\begin{equation}
+  \vindex: \dbase \times I \rightarrow \dbase s.t \vindex(\dbasepoint) = \dbasepoint \forall \dbasepoint \in \gbase
+\end{equation}
+
+we can assert that the data space \dbase acts as an indexing space into \gbase such that knowing the location on space yields the location on the other and any point in either base space or graphic space has a correspoinding point in the other space.
+\begin{figure}[H]
+  \includegraphics[width=1\columnwidth]{deform_retract.png}
+  \caption{The graphic base space $\gbase$ is collapsible to the line $\dbase$ such that every band $(\dbasepoint_i, [0,1])$ on $\gbase$ maps to corresponding point $\dbasepoint_i \in \dbase$.
+The band $[0,1]$ determines the thickness of a rendered line for a given point $\dbasepoint_{i}$ by specifying how pixels corresponding to that point are colored.
+\label{fig:construction:xi}}
+\end{figure}
+
+For example, as shown in \autoref{fig:construction:xi}, a line is 1D but is a 2D glyph on a screen; therefore the graphic space $\gbase$ is constructed by multiplying the base space $\dbase$ with an interval $[0,1]$.
+Because $\gbase$ is collapsible into $\dbase$, every band $(\dbasepoint_i, [0,1])$ corresponds to a point in the base space $\dbasepoint_i \in \dbase$.
+The first coordinate $\alpha=\dbasepoint_i$ provides a lookup to retrieve the associated visual variables.
+The second coordinate, which is a point in the interval $\beta=[0,1]$.
+Together they are a point $\gbasepoint=(\alpha,\beta) \in gbase$ in the graphic base space.
+This point $\gbasepoint$ is the input into the graphic section $\gsection(\gbasepoint)$ that is used to determine which pixels are colored, which in turn determines the thickness, texture, and color of the line.
 
@@
 An artist is equivariant when changes to the input and output are equivariant.
-
 As introduced in \autoref{eq:atct:morphism:base}, the base space transformation \dfunch\ is invariant because $\dsection\restriction_{\openset} = \dsection\restriction_{\dfunch(\openset^{\prime})}$.
@@
 
-This expectation of using the same encoding for the same variable is a generalization of the concept of consistency checking of multiple view encodings discussed by Zening and Hullman \cite{hullmanKeeping2018}.
+This expectation of using the same encoding for the same variable is a generalization of the concept of consistency checking of multiple view encodings discussed by Qu and Hullman \cite{hullmanKeeping2018}.
 This expectation can also be used to check that a multipart glyph is assembled correctly.
@@
 
+As mentioned in \autoref{sec:artist:homeomorphic}, we construct the data base space as a deformation retraction of the graphic space.
+On simple way of doing so is to construct the graphic base space as a constant multiple of the base space such that
+\begin{equation}
+  \underbrace{\dbasec\times[0,1]^{n}}_{\gbasec} \textcolor{functor}{\xmapsto{\hspace{1em}\vindexc\hspace{1em}}} \dbasec
+\end{equation}
+where n is a thickening of the graphic base space $\gbase$ to account for the dimensionality of the output space
+\begin{equation*}
+  n = \begin{cases}
+    dim(\gbase) - dim(\dbase) & dim(\dbase)<dim(\gbase)\\
+  0 & otherwise
+  \end{cases}
+\end{equation*}
+because otherwise the data dimensionality $\dbase$ may be too small for a graphic representation.
+For example, as shown in \autoref{fig:construction:xi}, a line is 1D but is a 2D glyph on a screen; therefore the graphic space $\gbase$ is constructed by multiplying the base space $\dbase$ with an interval $[0,1]$.
+
 \subsection{Measurable Visual Components}
@@
 We propose that reusing the fibers $\vfiber$ across components facilitates standardizing internal types across the library and that this standardization improves maintainability (\autoref{tab:appendix:library_spec}).
-
-\subsection{Map Between Graphics and Data}
-In \autoref{sec:atct:sheaves}, the functors $\vindex, \vindexpull, \vindexpushc$ are introduced to describe the expected relationship between screen and data base spaces.
-In the construction of the artist, the functor $\vindex$ is defined such that the data base space \dbase\ is a deformation retraction\cite{nlab:deformation_retract,hatcherAlgebraicTopology2002} of the graphic space $\gbase$.
-This means that there is a continuous surjective mapping from every point $\dbasepoint\in\dbase$ to a point $\dbasepoint \in dbase$.
-To simplify matters, in this paper, we construct the graphic space as a constant multiple of the base space such that
-\begin{equation}
-  \underbrace{\dbasec\times[0,1]^{n}}_{\gbasec} \textcolor{functor}{\xmapsto{\hspace{1em}\vindexc\hspace{1em}}} \dbasec
-\end{equation}
-where n is a thickening of the graphic base space $\gbase$ to account for the dimensionality of the output space
-\begin{equation*}
-  n = \begin{cases}
-    dim(\gbase) - dim(\dbase) & dim(\dbase)<dim(\gbase)\\
-  0 & otherwise
-  \end{cases}
-\end{equation*}
-because the data dimensionality $\dbase$ may be too small for a graphic representation.
-
-\begin{figure}[H]
-
-  \includegraphics[width=1\columnwidth]{deform_retract.png}
-  \caption{The graphic base space $\gbase$ is collapsible to the line $\dbase$ such that every band $(\dbasepoint_i, [0,1])$ on $\gbase$ maps to corresponding point $\dbasepoint_i \in \dbase$.
-The band $[0,1]$ determines the thickness of a rendered line for a given point $\dbasepoint_{i}$ by specifying how pixels corresponding to that point are colored.
-\label{fig:construction:xi}}
-\end{figure}
-
-For example, as shown in \autoref{fig:construction:xi}, a line is 1D but is a 2D glyph on a screen; therefore the graphic space $\gbase$ is constructed by multiplying the base space $\dbase$ with an interval $[0,1]$.
-Because $\gbase$ is collapsible into $\dbase$, every band $(\dbasepoint_i, [0,1])$ corresponds to a point in the base space $\dbasepoint_i \in \dbase$.
-The first coordinate $\alpha=\dbasepoint_i$ provides a lookup to retrieve the associated visual variables.
-The second coordinate, which is a point in the interval $\beta=[0,1]$.
-Together they are a point $\gbasepoint=(\alpha,\beta) \in gbase$ in the graphic base space.
-This point $\gbasepoint$ is the input into the graphic section $\gsection(\gbasepoint)$ that is used to determine which pixels are colored, which in turn determines the thickness, texture, and color of the line.
 

### 2023-11-22 paper/main.tex
environment

@@
 
-Since the earliest maps \cite{friendlyBriefHistoryData2006}, a visualization has been expected to preserve the $field$ properties and $topology$ of the corresponding dataset; a \textcolor{fiber}{\textcolor{field}} \footnote{Throughout this paper, we augment the math with color to group conceptually similar objects and functions\cite{headMathAugmentationHow2022}, for example \textcolor{action}{actions} and \textcolor{action}{sets of actions}.
+Since the earliest maps \cite{friendlyBriefHistoryData2006}, a visualization has been expected to preserve the $field$ properties and $topology$ of the corresponding dataset; a \textcolor{fiber}{\textbf{field}} \footnote{Throughout this paper, we augment the math with color to group conceptually related terms\cite{headMathAugmentationHow2022}, for example \textcolor{fiber}{field} and \textcolor{fiber}{fiber}.
 This color coding carries through to the figures.}  is a set of values of the same type, e.g.

### 2023-12-15 paper/main.tex
all the equality things

@@
 This color coding carries through to the figures.}  is a set of values of the same type, e.g.
-one column of a table or the pixels of an image, and \textcolor{base}{\textcolor{topology}} is the connectivity and relative positioning of elements in a dataset \cite{wilkinsonGrammarGraphics2005}.
+one column of a table or the pixels of an image, and \textcolor{base}{\textbf{topology}} is the connectivity and relative positioning of elements in a dataset \cite{wilkinsonGrammarGraphics2005}.
 In \autoref{fig:related-work:continuity:ktypes}, the topology is a continuous 3 dimensional surface encapsulating time and position, while the fields are date, location, station, pressure, and temperature.
@@
 \label{sec:related-work:equivariance}
-The structure on values has traditionally been described using the measurement scales-i.e.
+
+The structure on values has traditionally been codified using the measurement scales-i.e.
 nominal, ordinal, interval, ratio - which Steven's classified by their mathematical group structure\cite{stevensTheoryScalesMeasurement1946} and others have formally expanded to include more types \cite{leaFormalizationMeasurementScale1971, thomasMathematizationNotMeasurement2014}.
-Specifically, Steven's conceptualizes the structure on values as \textcolor{action}{actions} on groups \footnote{A \textit{group} is a set with an associative binary operator, an identity, and an inverse operation}:
+
+Specifically, Steven's conceptualizes the structure on values as \textcolor{action}{actions} on groups \footnote{A \textit{group} is a set with an associative binary operator.
+This operation must have an identity element and be closed, associative,
+and invertable}:
 
@@
 where the measurements are the groups and changes to them are actions.
-Structure preservetion in fields means that the data and the visualization are equivalent representations of an underlyng phenomena such that changes to data or visualization be reflected in the other in a consistent way.
+Structure preservetion in fields means that the data and the visualization are equivalent representations of an underlying phenomena such that changes to data or visualization be reflected in the other in a consistent way.
 A function that preserves structure in this manner for groups is called equivariant.
 Given a group $G$ that acts on both the input $X$ and the output $Y$ of a function $f: X \rightarrow Y$
-\begin{definition}
+\begin{definition}\label{def:equivariance}
  A function $f$ is \textbf{equivariant} when $f(act(g,x)) = act(g,f(x))$ for all $g$ in $G$ and for all $x$ in $X$ \cite{pittsNominalSetsNames2013}
 \end{definition}
-
-which means that a visualization is structure preserving when there exist compatible actions on the data and visualization, as proposed by Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}.
-Equivariance can sometimes be too restrictive because it can only be applied to groups and sometimes data does not have a group structure; an alternative is asserting that a visualization is structure preserving so long as it preserves the mathematical relations with and between fields, as proposed by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}, which is termed homomorphism.
-Given the function $f$ between sets $X$ and $Y$, both of which are equipped with the binary operator $\circ$
-\begin{definition}
+which means that a visualization is structure preserving when there exist compatible group actions on the data and visualization, as proposed by Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}.
+
+Some fields have structures that cannot be codified as groups, for example the ordinal 
+scale is a type of structure called a monoid\footnote{The difference between a monoid 
+and a group is that monoid operators do not have to be invertable.} For these situations, 
+we build on the claim that visualizations are $homomorphic$, as proposed by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}. Given the function $f: X \rightarrow Y$, with
+compatiable operators $(X,*)$ \note{finish this out}
+
+\begin{definition}\label{def:homomorphism}
+  A function $f$ is \textbf{homomorphic} when
+
+  $f(act_x(x)) = act_y(f(x))$ for all $x, y \in X$ if there exists a compatibility function $\varphi(act_y) = act_x$.
 
 \end{definition}
-  A function $f$ is \textbf{homeomorphic} when $f(x\circ y) = f(x) \circ f(y)$ for every pair of elements $x,y$ in $X$
-\end{definition}
-which means that applying the operation $\circ$ to the input to $f$ is equivalent to applying the operation to the output, which is only possibly when the output of the map $f$ from $X$ to $Y$ has the same structure as the input.
+which means that applying the operation $*$ to the input to $f$ is equivalent to applying the operation $\circ$ to the output.
+
 Our framework for structure preservation incorporates equivariance and homomorphism and allows for defining other measures of equivalent representation; it is discussed in detail in \autoref{sec:artist:equivariant:data}.
@@
 \subsection{Topology}
-Borrowing strcutures from algebraic topology allows us to describe structure on individual elements in a way that holds true for the entire dataset, meaning we can use the same tools to check if structre is preserved regardless of whether the data fits in memory, is distributed, or is streaming.
+Borrowing structures from algebraic topology allows us to describe structure on individual elements in a way that holds true for the entire dataset, meaning we can use the same tools to check if structre is preserved regardless of whether the data fits in memory, is distributed, or is streaming.
 Our framework generalizesis because, informally, a topology $\mathcal{T}$ is a partitioning of a space into its subspaces in a way where those subspaces can be composed back into the space; the choice of partitions is how we encode the connectivity of the data and visualizations that we are interested in.
@@
 A map that by definition has continous maps between two spaces, such as data and graphics, is called a homeomorphism\cite{riehlCategoryTheoryContext}.
-
+\note{add functional def of bijective and continous
+}
 \begin{definition}
-  A function $f$ is a $homeomorphism$ if it is bijective, continous, and has a continous inverse function f^{-1}. 
+  A function $f$ is a $homeomorphism$ if it is bijective, continous, and has a continous inverse function $f^{-1}$. 
 \end{definition}
@@
 \begin{enumerate}
-  \item a collection of \textit{objects} $X \textbf{ob}(\mathcal{C})$
+  \item a collection of \textit{objects} $X \in \textbf{ob}(\mathcal{C})$
   \item for every pair of objects $X, Y \in \textbf{ob}(\mathcal{C})$, a set of \textit{morphisms} $X \xrightarrow{f} Y \in Hom_{\mathcal{C}}(X, Y)$
@@
 
+\paragraph{Joining indexing spaces: $\oplus: \mathcal{\dbasec} \sqcup \mathcal{\dbasec} \rightarrow \mathcal{\dbasec}$}
 For example, the disjoint union of two bundles, as shown in \autoref{fig:atct:local_trivalizations}, is the coproduct $\dbase^{a} \subseteq_{\dbase^{c}} \dbase^{b}$ over an overlap $\dbase^{c}$ and therefore the inclusion morphism must be commutative:
@@
 
+\paragraph{Merging fields: $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$ }
+
 The fiber category $\mathcal{\dfiberc}$ is also equipped with a bifunctor because it is a monoidal category and this functor provides a method for combining fiber types.
-The bifunctor $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$ allows us to express fields that contain complexly typed values.
+The bifunctorallows $\otimes$ us to express fields that contain complexly typed values.
 For example, wind can be represented as two fields $\dfiber_{speed} \times \dfiber_{direction}$ or a composite fiber field $\dfiber_{speed} \otimes \dfiber_{direction} = \dfiber_{wind}$.
@@
   \end{minted}
-% use section* for acknowledgment
 
@@
 
-Hannah is also very grateful to the nlab and wikipedia contributors who wrote 
-articles explaining a lot of the more abstract math on which this paper is built.
+Hannah is also very grateful to the nlab and wikipedia contributors who wrote such 
+clear explainations of many of the topics discussed in this paper.
 

### 2024-01-12 paper/main.tex
equivariance figure

@@
 For example, values read off a bar chart have to be equivalent to the values used to construct that chart.
-Motivated by wanting to make better tools, we use algebraic topology to describe how $data$ and a corresponding $graphics$ have the same underlying structure and we use category theory to describing under what conditions the visualization tool preserves that structure such that $graphic \leftrightarrow data$.
-
-Algebraic structures allow us to describe many types of inherent structure in a uniform manner, and category theory provides a rigorous framework for specifying the behavior and properties of structure and maps between structures.
+Therefore a visualization tool is structure preserving when it preserves the mapping $data\leftrightarrow graphic$.
+
+We propose that we can better enforce this expectation in software by providing a uniform way of expressing $data$ and $graphic$ using their respective algebraic structure and by uniformally specifying the behaviors and properties of those structures and the maps between them using category theory.
 For example, our framework can encapsulate how a table and scatter plot and heatmap are different representations of the same data and track an observation from a datacube as a point along a time series and on a map and in a network.
-The algebraic structures can be translated into programmatic types, while the categorical descriptions translate to a functional design framework.
-Strong typing and function composition enable library developers to build complex components from simpler verifiable parts \cite{huHowFunctionalProgramming2015, hughesWhyFunctionalProgramming1989}.
+The algebraic structures can then be translated into programmatic types, while the categorical descriptions translate to a functional design framework.
+Strong typing and function composition enable visualization software developers to build complex components from simpler verifiable parts \cite{huHowFunctionalProgramming2015, hughesWhyFunctionalProgramming1989}.
 These  components can be built as a standalone library and integrated into existing libraries and we hope these ideas will help shape the reorganization of critical data visualization libraries, such as Matplotlib.
@@
 
-Since the earliest maps \cite{friendlyBriefHistoryData2006}, a visualization has been expected to preserve the $field$ properties and $topology$ of the corresponding dataset; a \textcolor{fiber}{\textbf{field}} \footnote{Throughout this paper, we augment the math with color to group conceptually related terms\cite{headMathAugmentationHow2022}, for example \textcolor{fiber}{field} and \textcolor{fiber}{fiber}.
-This color coding carries through to the figures.}  is a set of values of the same type, e.g.
-one column of a table or the pixels of an image, and \textcolor{base}{\textbf{topology}} is the connectivity and relative positioning of elements in a dataset \cite{wilkinsonGrammarGraphics2005}.
+Since the earliest maps \cite{friendlyBriefHistoryData2006}, a visualization has been expected to preserve the $field$ properties and $topology$ of the corresponding dataset:
+\begin{description}
+  \item[\textcolor{fiber}{\textbf{field}} \footnote{Throughout this paper, we augment the math with color to group conceptually related terms\cite{headMathAugmentationHow2022}, for example \textcolor{fiber}{field} and \textcolor{fiber}{fiber}.
+This color coding carries through to the figures.}]  is a set of values of the same type, e.g.
+one column of a table or the pixels of an image
+  \item[\textcolor{base}{\textbf{topology}}] is the connectivity and relative positioning of elements in a dataset \cite{wilkinsonGrammarGraphics2005}. 
+\end{description}
+
  In \autoref{fig:related-work:continuity:ktypes}, the topology is a continuous 3 dimensional surface encapsulating time and position, while the fields are date, location, station, pressure, and temperature.
@@
 Encapsulating their conditions, we: present a uniform abstract data representation layer in \autoref{sec:atct:sheaves} for ensuring that the visualization should not change if the data representation (i.e.
-the data container) changes, define the conditions under which data maps unambiguously to visual elements \cite{ziemkiewiczEmbeddingInformationVisualization2009} in \autoref{sec:atct:xi}, and provide a methodology for verifying that changes in data should correspond to changes in the visualization in \autoref{sec:artist:equivariant:artist} that does not necessarily require that the changes be perceptually significant.
+the data container) changes, define the conditions under which data is mapped unambiguously to visual encodings \cite{ziemkiewiczEmbeddingInformationVisualization2009} in \autoref{sec:atct:xi}, and provide a methodology for verifying that changes in data should correspond to changes in the visualization in \autoref{sec:artist:equivariant:artist} that does not necessarily require that the changes be perceptually significant.
 Furthermore, our model extends their conditions of equivariance by explicitly incorporating topology and by providing a framework for translating the theoretical ideas into buildable components in \autoref{sec:construction}.
@@
 where the measurements are the groups and changes to them are actions.
-Structure preservetion in fields means that the data and the visualization are equivalent representations of an underlying phenomena such that changes to data or visualization be reflected in the other in a consistent way.
-A function that preserves structure in this manner for groups is called equivariant.
+A function that preserves structure when the input or output is changed by a group action is called *equivariant*.
 Given a group $G$ that acts on both the input $X$ and the output $Y$ of a function $f: X \rightarrow Y$
@@
 
-Some fields have structures that cannot be codified as groups, for example the ordinal 
-scale is a type of structure called a monoid\footnote{The difference between a monoid 
-and a group is that monoid operators do not have to be invertable.} For these situations, 
-we build on the claim that visualizations are $homomorphic$, as proposed by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}. Given the function $f: X \rightarrow Y$, with
-compatiable operators $(X,*)$ \note{finish this out}
+While *equivariance* allows us to test stucture preservation under transformations to data or visualization, *homomorphism* provides a way of asserting whether the relations between data elements was preserved as proposed by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}.
+Given the function $f: X \rightarrow Y$, with operators $(X, \circ)$ and $(Y, *)$
 
 \begin{definition}\label{def:homomorphism}
-  A function $f$ is \textbf{homomorphic} when
-
-  $f(act_x(x)) = act_y(f(x))$ for all $x, y \in X$ if there exists a compatibility function $\varphi(act_y) = act_x$.
+  A function $f$ is \textbf{homomorphic} when $f(x_1 \circ x_2) = f(x_1) * f(x_2)$ and preserves identites $f(I_x) = I_y$ all $x, y \in X$
 
 \end{definition}
-which means that applying the operation $*$ to the input to $f$ is equivalent to applying the operation $\circ$ to the output.
+which means that the operators $\circ$ and $*$ are compatiable.
+For example, although the Steven's scales were conceptualized as having group structure, the ordinal scale have a monoidal structure because partial orders ($\geq, \leq$) are not invertable.
+\note{check grimalidi for citation}.
+
+As illustrated in \ref{fig:def:homomorphism}, ....
+structure preservation under ordering means that since 1 is less than 2, the visual transformation of 1 must be less than the visual transformation of two - here shown as 1 having a shorter bar than 2.
+We must define a < for visual compatable with the < for numbers.
+We handwave that for every relation we can write a compatabile closed function,  e.g.
+we can replace $\leq$ with $\min$, such that the operation is closed.
 
@@
 The toy example presented in \autoref{sec:discussion} demonstrates that it is relatively straightforward to build working visualization library components using the construction described in \autoref{sec:construction}.
-Since these components are defined with single record inputs, the can be implemented such that they are concurrent.
+Since these components are defined with single record inputs, they can be implemented such that they are concurrent.
 The cost of building a new function using these components is sometimes as small as renaming fields, meaning the new feature is relatively easy to maintain.

### 2024-05-10 paper/main.tex
working bib + reworked rw + references + environment
proof of life? + equiv pics + start of rework of sheaf unpacking

@@
 \renewcommand{\linenumberfont}{\normalfont\bfseries\small\color{lightgray}}
-
-% references
-\usepackage[nocompress]{cite}
-\usepackage[]{footmisc}
-\usepackage{bookmark}
 
@@
 
+% *** BIBLIOGRAPHY ***
+\usepackage[]{footmisc}
+\usepackage{bookmark}
+\usepackage[style=ieee, backend=biber]{biblatex}
+\addbibresource{references.bib}
+
 %\usepackage[inline]{showlabels}
@@
 % \\ is fragile and will error, could use \hfil\break instead.
-E-mail: haizenman@ccny.cuny.edu, mgrossberg@ccny.cuny.edu 
+E-mail: haizenman@.edu, mgrossberg@ccny.cuny.edu
 \IEEEcompsocthanksitem Mikael Vejdemo-Johansson is with the department of Mathematics, CUNY College of Staten Island.
@@
 
+Since the earliest maps \cite{friendlyBriefHistoryData2006}, a visualization has been expected to preserve the $field$ properties and $topology$ of the corresponding dataset:
+\begin{LaTeXdescription}
+  \item[\textcolor{fiber}{\textbf{field}} \footnote{Throughout this paper, we augment the math with color to group conceptually related terms\cite{headMathAugmentationHow2022}, for example \textcolor{fiber}{field} and \textcolor{fiber}{fiber}.
+This color coding carries through to the figures.}]  is a set of values of the same type, e.g.
+one column of a table or the pixels of an image
+  \item[\textcolor{base}{\textbf{topology}}] is the connectivity and relative positioning of elements in a dataset \cite{wilkinsonGrammarGraphics2005}.
+\end{LaTeXdescription}
+
 \begin{figure}[H]
@@
 
-Since the earliest maps \cite{friendlyBriefHistoryData2006}, a visualization has been expected to preserve the $field$ properties and $topology$ of the corresponding dataset:
-\begin{description}
-  \item[\textcolor{fiber}{\textbf{field}} \footnote{Throughout this paper, we augment the math with color to group conceptually related terms\cite{headMathAugmentationHow2022}, for example \textcolor{fiber}{field} and \textcolor{fiber}{fiber}.
-This color coding carries through to the figures.}]  is a set of values of the same type, e.g.
-one column of a table or the pixels of an image
-  \item[\textcolor{base}{\textbf{topology}}] is the connectivity and relative positioning of elements in a dataset \cite{wilkinsonGrammarGraphics2005}. 
-\end{description}
-
  In \autoref{fig:related-work:continuity:ktypes}, the topology is a continuous 3 dimensional surface encapsulating time and position, while the fields are date, location, station, pressure, and temperature.
@@
 The structure on values has traditionally been codified using the measurement scales-i.e.
-nominal, ordinal, interval, ratio - which Steven's classified by their mathematical group structure\cite{stevensTheoryScalesMeasurement1946} and others have formally expanded to include more types \cite{leaFormalizationMeasurementScale1971, thomasMathematizationNotMeasurement2014}.
-
+nominal, ordinal, interval, ratio - which Steven's classified by their mathematical structure\cite{stevensTheoryScalesMeasurement1946} and others have formally expanded to include more types \cite{leaFormalizationMeasurementScale1971, homasMathematizationNotMeasurement2014}.
+
+ \textit{equivariance} allows us to test stucture preservation under transformations to data or visualization has been proposed by Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}.
+
+ \textit{homomorphism} provides a way of asserting whether the relations between data elements was preserved as proposed by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}.
+
+\subsubsection{equivariance}
 Specifically, Steven's conceptualizes the structure on values as \textcolor{action}{actions} on groups \footnote{A \textit{group} is a set with an associative binary operator.
@@
 where the measurements are the groups and changes to them are actions.
-A function that preserves structure when the input or output is changed by a group action is called *equivariant*.
+A function that preserves structure when the input or output is changed by a group action is called \textit{equivariant}.
 Given a group $G$ that acts on both the input $X$ and the output $Y$ of a function $f: X \rightarrow Y$
@@
 
-While *equivariance* allows us to test stucture preservation under transformations to data or visualization, *homomorphism* provides a way of asserting whether the relations between data elements was preserved as proposed by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}.
+\subsection{homomorphism}
+
+For example, although the Steven's scales were conceptualized as having group structure, the ordinal scale have a monoidal structure because partial orders ($\geq, \leq$) are not invertable.
+This means we cannot use \textit{equivariance} to test for structure preservation instead we use \textit{homomorphsim} because it imposes fewer constraints on the underlying mathematical structure of the data.
+
 Given the function $f: X \rightarrow Y$, with operators $(X, \circ)$ and $(Y, *)$
@@
 \end{definition}
-which means that the operators $\circ$ and $*$ are compatiable.
-For example, although the Steven's scales were conceptualized as having group structure, the ordinal scale have a monoidal structure because partial orders ($\geq, \leq$) are not invertable.
-\note{check grimalidi for citation}.
+which means that the operators $\circ$ and $*$ are compatiable.  \note{check grimalidi for citation}.
 
@@
 
-Our framework for structure preservation incorporates equivariance and homomorphism and allows for defining other measures of equivalent representation; it is discussed in detail in \autoref{sec:artist:equivariant:data}.
+As shown in \autoref{fig:def:equivariance} and \autoref{fig:def:homomorphsim}, the basis of asserting structure preservation is that there exist compatiable actions and these actions commute; we generalize compatability and commutability as the basis for structure preservation in \autoref{sec:artist:equivariant:data}.
 
 \subsection{Topology}
+\note{lead with homomorphism}
 Borrowing structures from algebraic topology allows us to describe structure on individual elements in a way that holds true for the entire dataset, meaning we can use the same tools to check if structre is preserved regardless of whether the data fits in memory, is distributed, or is streaming.
@@
 
+%\caption{The local trivializations $\dtotal_{0}$ and $\dtotal_{1}$ are subspaces of $\dtotal$ where the fibers $\dfiber_{0}$ are equal, here illustrated as all pointing in the same direction.
+These local trivializations are themselves locally trivial fiber bundles.}
+
 \begin{definition}\label{def:fiber_bundle}
    A \textbf{fiber bundle} $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and  bundle projection map $\pi: \dtotalc \rightarrow \dbasec$ \cite{FiberBundle2020,spanier1989algebraic}.
+
     \begin{equation}
@@
     \end{equation}
+
 A continuous surjective map $\bm{\pi}$ is a \textbf{bundle projection} map when
@@
     \includegraphics[width=\columnwidth]{figures/local_trivalizations.png}
-  \caption{The local trivializations $\dtotal_{0}$ and $\dtotal_{1}$ are subspaces of $\dtotal$ where the fibers $\dfiber_{0}$ are equal, here illustrated as all pointing in the same direction.
-These local trivializations are themselves locally trivial fiber bundles.}
     \label{fig:atct:local_trivalizations}
+    \caption{Example of bundles $\dtotalc_0$ and $\dtotalc_1$, which are both trivializations
+    of a bundle $\dtotalc$ add in section here!}
 \end{figure}
+
+\autoref{fig:atct:local_trivalizations} shows an example of a two dimenensional bundle
+$\dtotal$.
+This bundle can encapsulate a set datasets where the data points are connected to each other along a unit circle $\dbasec$ = $\left[0, 2\pi\right]$ and the data values lie along an interval $dfiberc$ = $[0,1]$.
+Each dataset is encoded as a section, for example $sin$ and $cos$.
+The subspaces of $\dtotal_{0}$ and $\dtotal_{1}$ are local trivializations $\dtotal$; therefore the fibers are equal, here illustrated as all pointing in the same direction.
+These local trivializations are themselves locally trivial fiber bundles.
 
@@
 The example in this section is intentionally trivial to illustrate that the math to code translation is fairly straightforward and results in fairly self contained composable functions.
-A library applying these ideas, created by Thomas Caswell, can be found at \url{https://github.com/matplotlib/data-prototype}.
+A library applying these ideas, created by Thomas Caswell and Kyle Sunden, can be found at \url{https://github.com/matplotlib/data-prototype}.
 Further research could investigate building new systems using this model, specifically libraries for visualizing domain specific structured data and domain specific artists.
@@
 
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://mirror.ctan.org/biblio/bibtex/contrib/doc/https://www.overleaf.com/download/project/64063771726807fc06c1ee16/build/187faee483d-291434de4cddeda6/output/output.pdf?compileGroup=priority&clsiserverid=clsi-pre-emp-c2d-d-f-nkpv&enable_pdf_caching=true&popupDownload=truejjj
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-
-\bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-\bibliography{references}
+\printbibliography
 

### 2024-05-20 paper/main.tex
sketch of interactivity

@@
 The structure on values has traditionally been codified using the measurement scales-i.e.
-nominal, ordinal, interval, ratio - which Steven's classified by their mathematical structure\cite{stevensTheoryScalesMeasurement1946} and others have formally expanded to include more types \cite{leaFormalizationMeasurementScale1971, homasMathematizationNotMeasurement2014}.
+nominal, ordinal, interval, ratio - which Steven's classified by their mathematical structure\cite{stevensTheoryScalesMeasurement1946} and others have formally expanded to include more types \cite{leaFormalizationMeasurementScale1971, thomasMathematizationNotMeasurement2014}.
 
@@
 \end{definition}
+
+\note{Add motiviating note here!}
+Consider the circle.
+Points can be identified with angles (in radians, for instance), and we can imagine covering the circle with two overlapping intervals - for instance $\dbase_0=(-\varepsilon, 2\pi/5+\varepsilon)$ and $\dbase_1=(2\pi/5-\varepsilon, 2\pi+\varepsilon)$.
+We can imagine creating building blocks $\dbase_0\times[-1,1]$ and $\dbase_1\times[-2,2]$.
+
+There are two different ways to create a fiber bundle that has these building blocks as the local trivializations: on the one hand, we could imagine just a cylinder $S^1\times[-1,1]$ as a total space.
+The fibers would all be just the interval $[-1,1]$.
+Over $\dbase_0$, we could map from the local trivialization to the total space with the identity map, and over $\dbase_1$ we could map from the local trivialization to the total space by rescaling the fiber space $[-2,2]\to[-1,1]$.
+
+On the other hand, we could imagine creating a Möbius strip by flipping one end of $\dbase_1\times[-2,2]$ over before attaching it, so that while $\dbase_0\times[-1,1]$ gets mapped into the total space with the identity map, the part $(2\pi-\varepsilon,2\pi+\varepsilon)\times[-2,2]$ of $\dbase_1\times[-2,2]$ is mapped by the function $x\mapsto -x/2$ instead of the function $x\mapsto x/2$.
+
+These functions - using the isomorphisms $x\mapsto x$, $x\mapsto x/2$ and $x\mapsto -x/2$ to map from each rectangle into the cylinder compose into different maps $(\dbase_0\cap\dbase_1)\times[-2,2]\to(\dbase_0\cap\dbase_1)\times[-1,1]$ where the source is seen as a subset of $\dbase_1\times[-2,2]$ and the target is seen as a subset of $\dbase_0\times[-1,1]$.
+By specifying just these \textbf{transition maps} for all overlaps of cover elements in the local trivialization we can specify a gluing scheme that constructs the fiber bundle from these locally trivial pieces.
+
+%For example, in \autoref{fig:local_trivalizations}, the bundles $\dtotal_{0}=\dtotal\restriction_{[\epsilon^{-}, {\frac{2\pi}{5}}^{+}]}$ and $\dtotal_{1}=\dtotal\restriction_{[{\frac{2\pi}{5}}^{-}, \epsilon^{-}]}$ are local trivializations of an arbitrary bundle $\dtotal$.
+These local trivializations overlap on the intersections of the open subsets of each trivialization $[\epsilon^{-}, \epsilon^{+}]$ and $[{\frac{2\pi}{5}}^{-}, {\frac{2\pi}{5}}^{+}]$.
+A set of trivializations can be glued into a larger bundle by defining transition maps that describe how to align the fibers in the overlapping space such that the sections of the bundle remain continuous.
+\note{add transition map construction formula}
+
+\begin{figure}[H]
+  \includegraphics[width=1\columnwidth]{figures/cyl_mob.png}
+  \caption{\note{figure out short caption here}\label{fig:cyl_mob_bundles}}
+\end{figure}
+As demonstrated in \autoref{fig:cyl_mob_bundles} different types of total spaces can be decomposed into the same local trivializations; therefore the distinguishing factor between total spaces can be the transition maps for constructing the spaces from trivializations.
+
+A \textit{trivial bundle} is directly isomorphic to $\dbasec\times\dfiberc$, and for any choice of cover of $\dbasec$ by overlapping opensets, we can choose local trivializations so that all transition maps are identity maps.
+In the example in Figure~\ref{fig:cyl_mob_bundles}, we use arrows $\uparrow$ to denote fiber alignments, illustrating that by choosing simple enough trivializations, we get identity maps in the transition maps.
+
+%In a \textit{trivial bundle}, the fibers are identical $\dfiber = \dfiber_{k}$ for all points $\dbasepoint \in dbase$ such that all the transition maps are also identical.
+For example, in the cylinder case the fibers in the overlapping regions of the trivializations point in the same direction $\uparrow$ to illustrate that they are equal $\dfiber_0\restriction_{\openset_1\cap\openset_2} = \dfiber_{1}\restriction_{\openset_1\cap\openset_2}$.
+Because the fibers point in the same direction, the  transition maps at both intersections map the fiber to itself $\uparrow\rightarrow\uparrow$.
+
+A \textit{non-trivial bundle} on the other hand is one that can not be constructed as $\dbasec\times\dfiberc$, but where the construction is more complicated.
+Then for any choice of local trivializations, there is at least one transition map that is not an identity \cite{hatcherAlgebraicTopology2002}.
+For example, in the Möbius band case, while $\dfiber_0\restriction_{(2\pi/5-\varepsilon, 2\pi/5+\varepsilon)}\to\dfiber_1\restriction_{(2\pi/5-\varepsilon, 2\pi/5+\varepsilon)}$ can be chosen to be an identity map (after transforming $\dbase_1\times[-2,2]\to\dbase_1\times[-1,1]$ by rescaling), the other transition map component $\dfiber_0\restriction_{(-\varepsilon,\varepsilon)}\to\dfiber_1\restriction_{(-\varepsilon,\varepsilon)}$ has to flip the fibers to align them between the pieces.
+
+\note{ending on sectionsis probably better for our purposes!}
 \begin{definition} A \textcolor{section}{\textbf{section}} $\dsectionc: \dbasec \rightarrow \dtotalc$ over a fiber bundle is a smooth right inverse of $\pi$: $\pi(\dsection(\dbasepoint)) = \dbasepoint$ for all $\dbasepoint \in \dbase$
@@
 These local trivializations are themselves locally trivial fiber bundles.
-
-\note{Add motiviating note here!}
-Consider the circle.
-Points can be identified with angles (in radians, for instance), and we can imagine covering the circle with two overlapping intervals - for instance $\dbase_0=(-\varepsilon, 2\pi/5+\varepsilon)$ and $\dbase_1=(2\pi/5-\varepsilon, 2\pi+\varepsilon)$.
-We can imagine creating building blocks $\dbase_0\times[-1,1]$ and $\dbase_1\times[-2,2]$.
-
-There are two different ways to create a fiber bundle that has these building blocks as the local trivializations: on the one hand, we could imagine just a cylinder $S^1\times[-1,1]$ as a total space.
-The fibers would all be just the interval $[-1,1]$.
-Over $\dbase_0$, we could map from the local trivialization to the total space with the identity map, and over $\dbase_1$ we could map from the local trivialization to the total space by rescaling the fiber space $[-2,2]\to[-1,1]$.
-
-On the other hand, we could imagine creating a Möbius strip by flipping one end of $\dbase_1\times[-2,2]$ over before attaching it, so that while $\dbase_0\times[-1,1]$ gets mapped into the total space with the identity map, the part $(2\pi-\varepsilon,2\pi+\varepsilon)\times[-2,2]$ of $\dbase_1\times[-2,2]$ is mapped by the function $x\mapsto -x/2$ instead of the function $x\mapsto x/2$.
-
-These functions - using the isomorphisms $x\mapsto x$, $x\mapsto x/2$ and $x\mapsto -x/2$ to map from each rectangle into the cylinder compose into different maps $(\dbase_0\cap\dbase_1)\times[-2,2]\to(\dbase_0\cap\dbase_1)\times[-1,1]$ where the source is seen as a subset of $\dbase_1\times[-2,2]$ and the target is seen as a subset of $\dbase_0\times[-1,1]$.
-By specifying just these \textbf{transition maps} for all overlaps of cover elements in the local trivialization we can specify a gluing scheme that constructs the fiber bundle from these locally trivial pieces.
-
-%For example, in \autoref{fig:local_trivalizations}, the bundles $\dtotal_{0}=\dtotal\restriction_{[\epsilon^{-}, {\frac{2\pi}{5}}^{+}]}$ and $\dtotal_{1}=\dtotal\restriction_{[{\frac{2\pi}{5}}^{-}, \epsilon^{-}]}$ are local trivializations of an arbitrary bundle $\dtotal$.
-These local trivializations overlap on the intersections of the open subsets of each trivialization $[\epsilon^{-}, \epsilon^{+}]$ and $[{\frac{2\pi}{5}}^{-}, {\frac{2\pi}{5}}^{+}]$.
-A set of trivializations can be glued into a larger bundle by defining transition maps that describe how to align the fibers in the overlapping space such that the sections of the bundle remain continuous.
-\note{add transition map construction formula}
-
-\begin{figure}[H]
-  \includegraphics[width=1\columnwidth]{figures/cyl_mob.png}
-  \caption{\note{figure out short caption here}\label{fig:cyl_mob_bundles}}
-\end{figure}
-As demonstrated in \autoref{fig:cyl_mob_bundles} different types of total spaces can be decomposed into the same local trivializations; therefore the distinguishing factor between total spaces can be the transition maps for constructing the spaces from trivializations.
-
-A \textit{trivial bundle} is directly isomorphic to $\dbasec\times\dfiberc$, and for any choice of cover of $\dbasec$ by overlapping opensets, we can choose local trivializations so that all transition maps are identity maps.
-In the example in Figure~\ref{fig:cyl_mob_bundles}, we use arrows $\uparrow$ to denote fiber alignments, illustrating that by choosing simple enough trivializations, we get identity maps in the transition maps.
-
-%In a \textit{trivial bundle}, the fibers are identical $\dfiber = \dfiber_{k}$ for all points $\dbasepoint \in dbase$ such that all the transition maps are also identical.
-For example, in the cylinder case the fibers in the overlapping regions of the trivializations point in the same direction $\uparrow$ to illustrate that they are equal $\dfiber_0\restriction_{\openset_1\cap\openset_2} = \dfiber_{1}\restriction_{\openset_1\cap\openset_2}$.
-Because the fibers point in the same direction, the  transition maps at both intersections map the fiber to itself $\uparrow\rightarrow\uparrow$.
-
-A \textit{non-trivial bundle} on the other hand is one that can not be constructed as $\dbasec\times\dfiberc$, but where the construction is more complicated.
-Then for any choice of local trivializations, there is at least one transition map that is not an identity \cite{hatcherAlgebraicTopology2002}.
-For example, in the Möbius band case, while $\dfiber_0\restriction_{(2\pi/5-\varepsilon, 2\pi/5+\varepsilon)}\to\dfiber_1\restriction_{(2\pi/5-\varepsilon, 2\pi/5+\varepsilon)}$ can be chosen to be an identity map (after transforming $\dbase_1\times[-2,2]\to\dbase_1\times[-1,1]$ by rescaling), the other transition map component $\dfiber_0\restriction_{(-\varepsilon,\varepsilon)}\to\dfiber_1\restriction_{(-\varepsilon,\varepsilon)}$ has to flip the fibers to align them between the pieces.
 
@@
 While presheaves preserve the rules for sets of sections, sheaves add on conditions for gluing individual sections over subspaces into cohesive sections over the whole space.
-\begin{definition}\cite{bakerEuclideanSpaceMathsSheaf,spanier1989algebraic} A \textbf{sheaf} is a presheaf that satisfies the following two axioms
+\begin{definition}\label{def:atct:sheaf}\cite{bakerEuclideanSpaceMathsSheaf,spanier1989algebraic} A \textbf{sheaf} is a presheaf that satisfies the following two axioms
 \begin{itemize}
@@
 \begin{definition} Given a sheaf $\sheaf_{\gbase, \gtotal}$ on $\gbase$, the \textbf{pushforward} sheaf  $\vindexpush\sheaf_{\gbase, \gtotal}$ on $\dbase$ is defined as
-  \begin{equation*}
+  \begin{equation}
     \vindexpushc(\sheafc_{\gbasec, \gtotalc})(\opensetc)  = \sheafc_{\gbasec, \gtotalc}(\vindexc^{-1}(\opensetc))
-  \end{equation*}
+  \end{equation}
 for all opensets $\opensetc \subset \dbasec$\cite{harder2008lectures}.
@@
 \begin{equation}
+  \label{eq:atct:sheaf:pushforward_select}
   \vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexprec(\dbasepointc)}
@@
 \begin{equation}
+  \label{eq:atct:sheaf:pullback_hover}
   \vindexpullc\dsectionc(\gbasepointc) = \dsectionc(\vindexc(\gbasepointc)) = \dsectionc(\dbasepointc)
@@
 
+\note{explain why annotation is similar to brush/linking in operators section }
+
+\subsection{Animation and Interactivity}
+
+\begin{LaTeXdescription}
+  \item[pan, zoom, scroll] sheaf: locality + gluing \autoref{def:atct:sheaf}
+  \item[selection and hover] pushforward \autoref{eq:atct:sheaf:pushforward_select},  pullback \autoref{eq:atct:sheaf:pullback_hover}
+  \item[brushing, linking, annotation] composition of artists \autoref{eq:artist:join_base_fiber}
+\end{LaTeXdescription}
+
+Animation and interaction are a set of stills.
+Because the constraints are on the functions $\vartist\circ\dsection$ , satisfying the constraints on each function means that the constraint is satisfied for all visualizations $\{\vartist(\dsection(\dbasepoint)) \mid\dbasepoint \in \dbase \}$ that make up an animation or interaction.
+
 \section{Constructing Structure Preserving Components}

### 2024-05-28 paper/main.tex
clean up structure, bring types in

@@
 For example, values read off a bar chart have to be equivalent to the values used to construct that chart.
-Therefore a visualization tool is structure preserving when it preserves the mapping $data\leftrightarrow graphic$.
+Therefore a visualization tool is structure preserving when it preserves the bidirectional mapping $data\leftrightarrow graphic$.
 
@@
 
+ In \autoref{fig:related-work:continuity:ktypes}, the topology is a continuous 3 dimensional surface encapsulating time and position, while the fields are date, location, station, pressure, and temperature.
+Our paper provides a language to express how the scatter plot, line plot, and heatmap are preserving different structural properties of the dataset, thereby generalizing Bertin's codification of how the topology of observations must match the class of representation (i.e.
+point , line, area) and how the variation possible in a field match the variation in the corresponding visual variable \cite{bertinSemiologyGraphicsDiagrams2011} and Mackinlay's work on how structure preservation includes preserving the mathimatical relations between elements of the same and different fields \cite{mackinlayAutomatingDesignGraphical1986}.
+
+While Bertin and Mackinlay's work primarily discusses the $data \rightarrow graphic$, Kindlemann and Scheidegger's algebraic framework \cite{kindlmannAlgebraicProcessVisualization2014} also dictates a set of conditions under which the $graphic \rightarrow data$  mapping is structure preserving.
+Encapsulating their conditions, we: present a uniform abstract data representation layer in \autoref{sec:atct:sheaves} for ensuring that the visualization should not change if the data representation (i.e.
+the data container) changes, define the conditions under which data is mapped unambiguously to visual encodings \cite{ziemkiewiczEmbeddingInformationVisualization2009} in \autoref{sec:atct:xi}, and provide a methodology for verifying that changes in data should correspond to changes in the visualization in \autoref{sec:artist:equivariant:artist} that does not necessarily require that the changes be perceptually significant.
+Furthermore, our model extends their conditions of equivariance by explicitly incorporating topology and by providing a framework for translating the theoretical ideas into buildable components in \autoref{sec:construction}.
+
+\subsection{Fields}
+\label{sec:related-work:equivariance}
+
+Data is often described by its mathematical structure, for example the Steven's measurement scales define nominal, ordinal, interval, and ratio data by the allowed operations on each \cite{stevensTheoryScalesMeasurement1946} and other researchers have since expanded the scales to encapsulate more types of structure \cite{leaFormalizationMeasurementScale1971, thomasMathematizationNotMeasurement2014}.
+
+Loosely, the scales classify data as a set of values and the allowed transformations on that set, which can be operations, relations, or generalized as actions:
+
+\begin{definition}\label{def:related-work:action}\cite{grimaldiDiscreteCombinatorialMathematics2006}
+  An \textcolor{action}{\textbf{action}} of \textcolor{action}{$G = (G,\circ, e)$} on $X$ is a function  $act: \textcolor{action}{G} \times X \rightarrow X$.
+An action has the properties of identity $act(\textcolor{action}{e}, x) = x$ for all  $x \in X$ and associativity $act(\textcolor{action}{g}, act(\textcolor{action}{f}, x)) = act(\textcolor{action}{f} \circ \textcolor{action}{g}, x)$ for $\textcolor{action}{f},\textcolor{action}{g} \in \textcolor{action}{G}$.
+\end{definition}
+
+Computer science data types roughly describe structure in an equivalent manner;
+objects are elements of the set and class operators and methods are akin to
+actions, as described in \autoref{sec:atct:fb:fiber}.
+Given this equivalancy between code and math, we can develop software functions to test whether structure is preserved based on mathematical tests of structure preservation.
+For example, \textit{equivariance} allows us to test stucture preservation under transformations to data or visualization and has been proposed by Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014} and \textit{homomorphism} provides a way of asserting whether the relations between data elements was preserved as proposed by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}.
+
+\subsubsection{equivariance}
+Specifically, Steven's conceptualizes the structure on values as \textcolor{action}{actions} on groups \footnote{A \textit{group} is a set with an associative binary operator.
+This operation must have an identity element and be closed, associative,
+and invertable}.
+A function that preserves structure when the input or output is changed by a group action is called \textit{equivariant}.
+Given a group $G$ that acts on both the input $X$ and the output $Y$ of a function $f: X \rightarrow Y$
+
+\begin{definition}\label{def:equivariance}
+ A function $f$ is \textbf{equivariant} when $f(act(g,x)) = act(g,f(x))$ for all $g$ in $G$ and for all $x$ in $X$ \cite{pittsNominalSetsNames2013}
+\end{definition}
+which means that a visualization is structure preserving when there exist compatible group actions on the data and visualization, as proposed by Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}.
+
+\subsubsection{homomorphism}
+
+For example, although the Steven's scales were conceptualized as having group structure, the ordinal scale have a monoidal structure because partial orders ($\geq, \leq$) are not invertable.
+This means we cannot use \textit{equivariance} to test for structure preservation instead we use \textit{homomorphsim} because it imposes fewer constraints on the underlying mathematical structure of the data.
+
+Given the function $f: X \rightarrow Y$, with operators $(X, \circ)$ and $(Y, *)$
+
+\begin{definition}\label{def:homomorphism}
+  A function $f$ is \textbf{homomorphic} when $f(x_1 \circ x_2) = f(x_1) * f(x_2)$ and preserves identites $f(I_x) = I_y$ all $x, y \in X$
+
+\end{definition}
+which means that the operators $\circ$ and $*$ are compatiable.  \note{check grimalidi for citation}.
+
+As illustrated in \ref{fig:def:homomorphism}, ....
+structure preservation under ordering means that since 1 is less than 2, the visual transformation of 1 must be less than the visual transformation of two - here shown as 1 having a shorter bar than 2.
+We must define a < for visual compatable with the < for numbers.
+We handwave that for every relation we can write a compatabile closed function,  e.g.
+we can replace $\leq$ with $\min$, such that the operation is closed.
+
+As shown in \autoref{fig:def:equivariance} and \autoref{fig:def:homomorphsim}, the basis of asserting structure preservation is that there exist compatiable actions and these actions commute; we generalize compatability and commutability as the basis for structure preservation in \autoref{sec:artist:equivariant:data}.
+
+\subsection{Topology}
+
 \begin{figure}[H]
@@
 \end{figure}
-
- In \autoref{fig:related-work:continuity:ktypes}, the topology is a continuous 3 dimensional surface encapsulating time and position, while the fields are date, location, station, pressure, and temperature.
-Our paper provides a language to express how the scatter plot, line plot, and heatmap are preserving different structural properties of the dataset, thereby generalizing Bertin's codification of how the topology of observations must match the class of representation (i.e.
-point , line, area) and how the variation possible in a field match the variation in the corresponding visual variable \cite{bertinSemiologyGraphicsDiagrams2011} and Mackinlay's work on how structure preservation includes preserving the mathimatical relations between elements of the same and different fields \cite{mackinlayAutomatingDesignGraphical1986}.
-
-While Bertin and Mackinlay's work primarily discusses the $data \rightarrow graphic$, Kindlemann and Scheidegger's algebraic framework \cite{kindlmannAlgebraicProcessVisualization2014} also dictates a set of conditions under which the $graphic \rightarrow data$  mapping is structure preserving.
-Encapsulating their conditions, we: present a uniform abstract data representation layer in \autoref{sec:atct:sheaves} for ensuring that the visualization should not change if the data representation (i.e.
-the data container) changes, define the conditions under which data is mapped unambiguously to visual encodings \cite{ziemkiewiczEmbeddingInformationVisualization2009} in \autoref{sec:atct:xi}, and provide a methodology for verifying that changes in data should correspond to changes in the visualization in \autoref{sec:artist:equivariant:artist} that does not necessarily require that the changes be perceptually significant.
-Furthermore, our model extends their conditions of equivariance by explicitly incorporating topology and by providing a framework for translating the theoretical ideas into buildable components in \autoref{sec:construction}.
-
-\subsection{Fields}
-\label{sec:related-work:equivariance}
-
-The structure on values has traditionally been codified using the measurement scales-i.e.
-nominal, ordinal, interval, ratio - which Steven's classified by their mathematical structure\cite{stevensTheoryScalesMeasurement1946} and others have formally expanded to include more types \cite{leaFormalizationMeasurementScale1971, thomasMathematizationNotMeasurement2014}.
-
- \textit{equivariance} allows us to test stucture preservation under transformations to data or visualization has been proposed by Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}.
-
- \textit{homomorphism} provides a way of asserting whether the relations between data elements was preserved as proposed by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}.
-
-\subsubsection{equivariance}
-Specifically, Steven's conceptualizes the structure on values as \textcolor{action}{actions} on groups \footnote{A \textit{group} is a set with an associative binary operator.
-This operation must have an identity element and be closed, associative,
-and invertable}:
-
-\begin{definition}\label{def:related-work:action}\cite{grimaldiDiscreteCombinatorialMathematics2006}
-  An \textcolor{action}{\textbf{action}} of \textcolor{action}{$G = (G,\circ, e)$} on $X$ is a function  $act: \textcolor{action}{G} \times X \rightarrow X$.
-An action has the properties of identity $act(\textcolor{action}{e}, x) = x$ for all  $x \in X$ and associativity $act(\textcolor{action}{g}, act(\textcolor{action}{f}, x)) = act(\textcolor{action}{f} \circ \textcolor{action}{g}, x)$ for $\textcolor{action}{f},\textcolor{action}{g} \in \textcolor{action}{G}$.
-\end{definition}
-
-where the measurements are the groups and changes to them are actions.
-A function that preserves structure when the input or output is changed by a group action is called \textit{equivariant}.
-Given a group $G$ that acts on both the input $X$ and the output $Y$ of a function $f: X \rightarrow Y$
-\begin{definition}\label{def:equivariance}
- A function $f$ is \textbf{equivariant} when $f(act(g,x)) = act(g,f(x))$ for all $g$ in $G$ and for all $x$ in $X$ \cite{pittsNominalSetsNames2013}
-\end{definition}
-which means that a visualization is structure preserving when there exist compatible group actions on the data and visualization, as proposed by Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}.
-
-\subsection{homomorphism}
-
-For example, although the Steven's scales were conceptualized as having group structure, the ordinal scale have a monoidal structure because partial orders ($\geq, \leq$) are not invertable.
-This means we cannot use \textit{equivariance} to test for structure preservation instead we use \textit{homomorphsim} because it imposes fewer constraints on the underlying mathematical structure of the data.
-
-Given the function $f: X \rightarrow Y$, with operators $(X, \circ)$ and $(Y, *)$
-
-\begin{definition}\label{def:homomorphism}
-  A function $f$ is \textbf{homomorphic} when $f(x_1 \circ x_2) = f(x_1) * f(x_2)$ and preserves identites $f(I_x) = I_y$ all $x, y \in X$
-
-\end{definition}
-which means that the operators $\circ$ and $*$ are compatiable.  \note{check grimalidi for citation}.
-
-As illustrated in \ref{fig:def:homomorphism}, ....
-structure preservation under ordering means that since 1 is less than 2, the visual transformation of 1 must be less than the visual transformation of two - here shown as 1 having a shorter bar than 2.
-We must define a < for visual compatable with the < for numbers.
-We handwave that for every relation we can write a compatabile closed function,  e.g.
-we can replace $\leq$ with $\min$, such that the operation is closed.
-
-As shown in \autoref{fig:def:equivariance} and \autoref{fig:def:homomorphsim}, the basis of asserting structure preservation is that there exist compatiable actions and these actions commute; we generalize compatability and commutability as the basis for structure preservation in \autoref{sec:artist:equivariant:data}.
-
-\subsection{Topology}
 \note{lead with homomorphism}

### 2024-08-22 paper/main.tex
figures cleanup to verify they're showing what supposed to

@@
 
-\IEEEPARstart{V}{isualization} tools implement transforms of the form $data \rightarrow graphic$ and are expected to be structure preserving.
-Loosely, that means the mathematical properties of the data and how the points are connected to each other should be inferable such that  $graphic \rightarrow data$.
+\IEEEPARstart{V}isualization design guidelines, generally, describe how to choose visual encodings that preserve the structure of the data; to follow these guidelines the visualization tools that implement these $\data \rightarrow graphic$ transforms must be structure preserving.
+Loosely, preserving structure means that the athematical properties of the data and how the points are connected to each other should be inferable from the graphic such that  a $graphic \rightarrow data$ mapping can be made.
 For example, values read off a bar chart have to be equivalent to the values used to construct that chart.
@@
 
+\note{this got moved}
 In \autoref{fig:related-work:continuity:ktypes}, the topology is a continuous 3 dimensional surface encapsulating time and position, while the fields are date, location, station, pressure, and temperature.
@@
 
-Computer science data types roughly describe structure in an equivalent manner;
-objects are elements of the set and class operators and methods are akin to
-actions, as described in \autoref{sec:atct:fb:fiber}.
-Given this equivalancy between code and math, we can develop software functions to test whether structure is preserved based on mathematical tests of structure preservation.
-For example, \textit{equivariance} allows us to test stucture preservation under transformations to data or visualization and has been proposed by Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014} and \textit{homomorphism} provides a way of asserting whether the relations between data elements was preserved as proposed by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}.
+The elements of a set $X$ are roughly equivalent to instantiated objects of a class, and actions are roughly equivalent to methods on the objects; therefore mathematical definitions of structure preservation can be mapped directly to code.
+For example, \textit{equivariant} functions preserve structure under transformations to data or visualization and has been proposed by Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014} and \textit{homomorphic} maps preserve relations between data elements was preserved as proposed by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}.
 
 \subsubsection{equivariance}
+\begin{figure}
+  \includegraphics*[width=1\columnwidth]{tex/equivariant.png}
+\end{figure}
 Specifically, Steven's conceptualizes the structure on values as \textcolor{action}{actions} on groups \footnote{A \textit{group} is a set with an associative binary operator.
@@
 \subsubsection{homomorphism}
-
+\begin{figure}
+  \includegraphics[width=1\columnwidth]{tex/homomorphism.png}
+\end{figure}
 For example, although the Steven's scales were conceptualized as having group structure, the ordinal scale have a monoidal structure because partial orders ($\geq, \leq$) are not invertable.

### 2024-11-26 paper/main.tex
homeomorphism figure

@@
 
-As illustrated in \ref{fig:def:homomorphism}, ....
-structure preservation under ordering means that since 1 is less than 2, the visual transformation of 1 must be less than the visual transformation of two - here shown as 1 having a shorter bar than 2.
+As illustrated in \ref{fig:def:homomorphism}, structure preservation under ordering means that since 1 is less than 2, the visual transformation of 1 must be less than the visual transformation of two - here shown as 1 having a shorter bar than 2.
 We must define a < for visual compatable with the < for numbers.
@@
 \autoref{eq:atct:sheaf:restriction}, such that the following diagram commutes:
+\note{this needs human words - subsets of functions of the same type map to subsets of visualizations of the same type}
 

### 2024-11-26 paper/figcode/intro_plots.py

@@ -10,13 +10,13 @@ from mpl_toolkits.axes_grid1.axes_divider import make_axes_locatable
 
 #scatter position: 
 
-airports = {'ALBANY INTL AP': {'code': 'ALB', 'color': '#469990'},
+airports = {'ALBANY INTL AP': {'code': 'ALB', 'color': '#800000'},
             'BINGHAMTON': {'code': 'BGM', 'color': '#9A6324'},
             'BUFFALO': {'code': 'BUF', 'color': '#f032e6'},
             'GLENS FALLS AP': {'code': 'GFL', 'color': '#dcbeff'},
             'ISLIP-LI MACARTHUR AP': {'code': 'ISP', 'color': '#ffe119'},
             'LAGUARDIA AP': {'code': 'LGA', 'color': '#f58231'},
-            'ROCHESTER GTR INTL AP': {'code': 'ROC', 'color': '#800000'},
+            'ROCHESTER GTR INTL AP': {'code': 'ROC', 'color':'#469990'},
             'SYRACUSE HANCOCK INTL AP': {'code': 'SYR', 'color': '#aaffc3'},
             'JFK INTL AP': {'code': 'JFK', 'color': '#000075'},
             'HILO INTERNATIONAL AIRPORT': {'code': 'ITO', 'color': '#3CB44B'},

### 2024-12-03 paper/main.tex
mostly acknowledgments & rework of 2

@@
 \author{Hannah~Aizenman, Mikael Vejdemo-Johansson, Thomas~Caswell, and~Michael~Grossberg,~\IEEEmembership{Member,~IEEE,}% <-this % stops a space
-\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem H. Aizenman and M. Grossberg are with the department of Computer Science, City College of New York.
+\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem H. Aizenman is with the department of Computer Science, The Gradaute Center, CUNY.
 \protect\\
@@
 % \\ is fragile and will error, could use \hfil\break instead.
-E-mail: haizenman@.edu, mgrossberg@ccny.cuny.edu
+E-mail: haizenman@gradcenter.cuny.edu,
+\IEEEcompsocthanksitem M. Grossberg is with the department of Computer Science, City College of New York, CUNY.
+E-mail: mgrossberg@ccny.cuny.edu
 \IEEEcompsocthanksitem Mikael Vejdemo-Johansson is with the department of Mathematics, CUNY College of Staten Island.
@@
 \IEEEPARstart{V}isualization design guidelines, generally, describe how to choose visual encodings that preserve the structure of the data; to follow these guidelines the visualization tools that implement these $\data \rightarrow graphic$ transforms must be structure preserving.
-Loosely, preserving structure means that the athematical properties of the data and how the points are connected to each other should be inferable from the graphic such that  a $graphic \rightarrow data$ mapping can be made.
+Loosely, preserving structure means that the properties of the data and how the points are connected to each other should be inferable from the graphic such that a $graphic \rightarrow data$ mapping can be made.
 For example, values read off a bar chart have to be equivalent to the values used to construct that chart.
@@
 
-Since the earliest maps \cite{friendlyBriefHistoryData2006}, a visualization has been expected to preserve the $field$ properties and $topology$ of the corresponding dataset:
+In this section, we describe how structure has traditionally been discussed in visualization and mathematics and encapsulated in visualization library design because we propose that by formally incorporating mathematical definitions of equality, we provide a general framework that can account for a larger variety of fields than the existing work has, can formally expresse what it means to preserve the topology of the data, and we can do so in a uniform manner using category theory because that can lead to more consistent software interfaces in visualization software libraries \cite{wielsManagementEvolvingSpecifications1998,yorgeyMonoidsThemeVariations2012}.
+
+Generally, preserving structure means that a visualization is expected to preserve the $field$ properties and $topology$ of the corresponding dataset:
+
 \begin{LaTeXdescription}
-  \item[\textcolor{fiber}{\textbf{field}} \footnote{Throughout this paper, we augment the math with color to group conceptually related terms\cite{headMathAugmentationHow2022}, for example \textcolor{fiber}{field} and \textcolor{fiber}{fiber}.
-This color coding carries through to the figures.}]  is a set of values of the same type, e.g.
+  \item[\textcolor{fiber}{\textbf{field}} \footnote{Throughout this paper, we use color in definitions, equations, and visualizations to group conceptually related terms\cite{headMathAugmentationHow2022}, for example \textcolor{fiber}{field} and \textcolor{fiber}{fiber}.}] is a set of values of the same type, e.g.
 one column of a table or the pixels of an image
@@
 \end{LaTeXdescription}
+
+The conditions under which $data \rightarrow graphic$ is structure preserving is discussed extensively in the visualization literature, codified by Bertin\cite{bertinSemiologyGraphicsDiagrams2011} and extended to tool design by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}, and a set of conditions under which the $graphic \rightarrow data$ mapping is structure preserving is presented in Kindlemann and Scheidegger's algebraic framework \cite{kindlmannAlgebraicProcessVisualization2014}.
+Encapsulating their conditions, we present a uniform abstract data representation layer in \autoref{sec:atct:sheaves} for ensuring that the visualization should not change if the data representation (i.e.
+the data container) changes, define the conditions under which data is mapped unambiguously to visual encodings \cite{ziemkiewiczEmbeddingInformationVisualization2009} in \autoref{sec:atct:xi}, and provide a methodology for verifying that changes in data should correspond to changes in the visualization in \autoref{sec:artist:equivariant:artist} that does not necessarily require that the changes be perceptually significant.
+Furthermore, our model extends their conditions of equivariance by explicitly incorporating topology and by providing a framework for translating the theoretical ideas into buildable components in \autoref{sec:construction}.
 
@@
 point , line, area) and how the variation possible in a field match the variation in the corresponding visual variable \cite{bertinSemiologyGraphicsDiagrams2011} and Mackinlay's work on how structure preservation includes preserving the mathimatical relations between elements of the same and different fields \cite{mackinlayAutomatingDesignGraphical1986}.
-
-While Bertin and Mackinlay's work primarily discusses the $data \rightarrow graphic$, Kindlemann and Scheidegger's algebraic framework \cite{kindlmannAlgebraicProcessVisualization2014} also dictates a set of conditions under which the $graphic \rightarrow data$  mapping is structure preserving.
-Encapsulating their conditions, we: present a uniform abstract data representation layer in \autoref{sec:atct:sheaves} for ensuring that the visualization should not change if the data representation (i.e.
-the data container) changes, define the conditions under which data is mapped unambiguously to visual encodings \cite{ziemkiewiczEmbeddingInformationVisualization2009} in \autoref{sec:atct:xi}, and provide a methodology for verifying that changes in data should correspond to changes in the visualization in \autoref{sec:artist:equivariant:artist} that does not necessarily require that the changes be perceptually significant.
-Furthermore, our model extends their conditions of equivariance by explicitly incorporating topology and by providing a framework for translating the theoretical ideas into buildable components in \autoref{sec:construction}.
 
@@
 As shown in \autoref{fig:def:equivariance} and \autoref{fig:def:homomorphsim}, the basis of asserting structure preservation is that there exist compatiable actions and these actions commute; we generalize compatability and commutability as the basis for structure preservation in \autoref{sec:artist:equivariant:data}.
+
+A brief visualization oriented introduction to category theory is in Vickers et al \cite{vickersUnderstandingVisualizationFormal2013}, but they are applying category theory to semantic concerns about visualization design rather than library architecture.
 
@@
 For example, connectivity assumptions tend to be embedded in each of the visual algorithms of `building block` libraries, a term used by Wongsuphasawat \cite{wongsuphasawatNavigatingWideWorld2021} to describe libraries that provide modular components for building elements of a visualization, such as functions for making boxes or translating data values to colors.
-In building block libraries such as Matplotlib\cite{hunterMatplotlib2DGraphics2007} and D3\cite{bostockDataDrivenDocuments2011} the connectivity assumptions lead to very different interfaces; for example in Matplotlib methods for updating data and parameters for controlling aesthetics differ between plots.
-While VTK\cite{hanwellVisualizationToolkitVTK2015,geveciVTK2012} provides a language for expressing the topological properties of the data, and therefore can embed that information in its visual algorithms, it does so in a non-uniform manner.
-On the other hand, domain specific libraries are designed with the assumption of continuities that are common in the domain \cite{HeerSoftware2006}, and therefore can somewhat restrict the interface to choices that that are appropriate for the domain.
+In building block libraries such as Matplotlib\cite{hunterMatplotlib2DGraphics2007} and D3\cite{bostockDataDrivenDocuments2011} assumptions about connectivity are embedded in the interfaces such that the API is inconsistent across plot types.
+For example in Matplotlib methods for updating data and parameters for controlling aesthetics differ between (1D) line based plotting methods and (0D) marker based methods.
+While VTK\cite{hanwellVisualizationToolkitVTK2015,geveciVTK2012} provides a language for expressing the topological properties of the data, and therefore can embed that information in its visual algorithms, VTK's charts API is similar to the continuity dependent APIs of other building block libraries.
+
+Domain specific libraries are designed with the assumption of continuities that are common in the domain \cite{HeerSoftware2006}, and therefore can somewhat restrict their API to choices that that are appropriate for the domain.
 For example, a tabular topological structure of discrete rows, as illustrated in \autoref{fig:related-work:continuity:ktypes}, is assumed by
@@
 By assuming the structure of their data, these domain specific libraries can provide more cohesive interfaces for a much more limited set of visualization algorithms than the building block libraries offer.
-\note{transition sentence}
-In this work, we formally specify the conditions necessary to build structure preserving components using category theory, because it provides a rich language for expressing the structure of objects, functions between objects, and the constraints that must hold for these functions to compose \cite{wielsManagementEvolvingSpecifications1998,yorgeyMonoidsThemeVariations2012}.
-A brief visualization oriented introduction to category theory is in Vickers et al \cite{vickersUnderstandingVisualizationFormal2013}, but they are applying category theory to semantic concerns about visualization design rather than library architecture.
 
@@
 \note{Acknowledge all the actual people}
-The authors would like to thank...
-
-Hannah is also very grateful to the nlab and wikipedia contributors who wrote such
-clear explainations of many of the topics discussed in this paper.
+The authors would like to thank the anonymous reviewers who gave constructive feedback on
+an earlier version of this paper. Various Matplotlib and Napari contributors, particularly
+Juan Nunez-Iglesias , and Nicolas Kruchten have provided valuable feedback from the
+library developors perspective.
+
+Hannah is also very grateful to Nicolas for the suggestion of augmented notation and to
+the nlab and wikipedia contributors who wrote clear explanations of many of the topics
+discussed in this paper.
+
+This project has been made possible in part by grant number 2019-207333 and from the
+\note{Cycle 3 grant #} Chan Zuckerberg Initiative DAF, an advised fund of Silicon Valley Community Foundation
 

### 2024-12-03 paper/main.tex
fields

@@
 We propose that we can better enforce this expectation in software by providing a uniform way of expressing $data$ and $graphic$ using their respective algebraic structure and by uniformally specifying the behaviors and properties of those structures and the maps between them using category theory.
-For example, our framework can encapsulate how a table and scatter plot and heatmap are different representations of the same data and track an observation from a datacube as a point along a time series and on a map and in a network.
+For example, our framework can encapsulate how a table and scatter plot and heatmap are different representations of the same data and track an observation from a data cube as a point along a time series and on a map and in a network.
 The algebraic structures can then be translated into programmatic types, while the categorical descriptions translate to a functional design framework.
 Strong typing and function composition enable visualization software developers to build complex components from simpler verifiable parts \cite{huHowFunctionalProgramming2015, hughesWhyFunctionalProgramming1989}.
-These  components can be built as a standalone library and integrated into existing libraries and we hope these ideas will help shape the reorganization of critical data visualization libraries, such as Matplotlib.
+These components can be built as a standalone library and integrated into existing libraries and we hope these ideas will influence the architecture of critical data visualization libraries, such as Matplotlib.
 
@@
 \section{Related Work}
-\label{sec:related-work:continuity}
-
-In this section, we describe how structure has traditionally been discussed in visualization and mathematics and encapsulated in visualization library design because we propose that by formally incorporating mathematical definitions of equality, we provide a general framework that can account for a larger variety of fields than the existing work has, can formally expresse what it means to preserve the topology of the data, and we can do so in a uniform manner using category theory because that can lead to more consistent software interfaces in visualization software libraries \cite{wielsManagementEvolvingSpecifications1998,yorgeyMonoidsThemeVariations2012}.
-
+\label{sec:related-work}
+
+This paper builds on how structure has traditionally been discussed in visualization and mathematics and encapsulated in visualization library design to propose a uniform interface for encoding structure that supports a broader variety of fields and more rigorously define how connectivity is preserved.
 Generally, preserving structure means that a visualization is expected to preserve the $field$ properties and $topology$ of the corresponding dataset:
@@
 
-The conditions under which $data \rightarrow graphic$ is structure preserving is discussed extensively in the visualization literature, codified by Bertin\cite{bertinSemiologyGraphicsDiagrams2011} and extended to tool design by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}, and a set of conditions under which the $graphic \rightarrow data$ mapping is structure preserving is presented in Kindlemann and Scheidegger's algebraic framework \cite{kindlmannAlgebraicProcessVisualization2014}.
-Encapsulating their conditions, we present a uniform abstract data representation layer in \autoref{sec:atct:sheaves} for ensuring that the visualization should not change if the data representation (i.e.
+The conditions under which $data \rightarrow graphic$ is structure preserving is discussed extensively in the visualization literature, codified by Bertin\cite{bertinSemiologyGraphicsDiagrams2011} and extended to tool design by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}, and a set of conditions under which the $graphic \rightarrow data$ mapping is structure preserving is presented in Kindlemann and Scheidegger's algebraic visualization design (AVD) framework \cite{kindlmannAlgebraicProcessVisualization2014}.
+Encapsulating the AVD conditions, we present a uniform abstract data representation layer in \autoref{sec:atct:sheaves} for ensuring that the visualization should not change if the data representation (i.e.
 the data container) changes, define the conditions under which data is mapped unambiguously to visual encodings \cite{ziemkiewiczEmbeddingInformationVisualization2009} in \autoref{sec:atct:xi}, and provide a methodology for verifying that changes in data should correspond to changes in the visualization in \autoref{sec:artist:equivariant:artist} that does not necessarily require that the changes be perceptually significant.
-Furthermore, our model extends their conditions of equivariance by explicitly incorporating topology and by providing a framework for translating the theoretical ideas into buildable components in \autoref{sec:construction}.
-
-\note{this got moved}
-In \autoref{fig:related-work:continuity:ktypes}, the topology is a continuous 3 dimensional surface encapsulating time and position, while the fields are date, location, station, pressure, and temperature.
-Our paper provides a language to express how the scatter plot, line plot, and heatmap are preserving different structural properties of the dataset, thereby generalizing Bertin's codification of how the topology of observations must match the class of representation (i.e.
-point , line, area) and how the variation possible in a field match the variation in the corresponding visual variable \cite{bertinSemiologyGraphicsDiagrams2011} and Mackinlay's work on how structure preservation includes preserving the mathimatical relations between elements of the same and different fields \cite{mackinlayAutomatingDesignGraphical1986}.
+Furthermore, our model generalizes the AVD notion of equivariance by allowing non-group structures, explicitly incorporating topology and by providing a framework for translating the theoretical ideas into buildable components in \autoref{sec:construction}.
 
@@
 
-The elements of a set $X$ are roughly equivalent to instantiated objects of a class, and actions are roughly equivalent to methods on the objects; therefore mathematical definitions of structure preservation can be mapped directly to code.
+Elements of $X$ can be from one data field or all of them or some subset; similarly the actions act on
+the elements of $X$ and each action can be a composition of actions.
+This means actions can be used when discussing various measures of structure preservation.
 For example, \textit{equivariant} functions preserve structure under transformations to data or visualization and has been proposed by Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014} and \textit{homomorphic} maps preserve relations between data elements was preserved as proposed by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}.
 
-\subsubsection{equivariance}
-\begin{figure}
+Specifically, Steven's conceptualizes the structure on values as \textcolor{action}{actions} on groups \footnote{A \textit{group} is a set with an associative binary operator.
+This operation must have an identity element and be closed, associative, and invertable}.
+A function that preserves structure when the input or output is changed by a group action is called \textit{equivariant}.
+
+\begin{figure}[H]
   \includegraphics*[width=1\columnwidth]{tex/equivariant.png}
+  \caption{Encoding data as the bar height using an exponential transform is not equivariant because encoding the data and then scaling the bar heights yields a much taller graph then scaling the data and then encoding those heights using the same exponential transform function.}
+  \label{fig:related-work:equivariance}
 \end{figure}
-Specifically, Steven's conceptualizes the structure on values as \textcolor{action}{actions} on groups \footnote{A \textit{group} is a set with an associative binary operator.
-This operation must have an identity element and be closed, associative,
-and invertable}.
-A function that preserves structure when the input or output is changed by a group action is called \textit{equivariant}.
+
  Given a group $G$ that acts on both the input $X$ and the output $Y$ of a function $f: X \rightarrow Y$
@@
 \end{definition}
-which means that a visualization is structure preserving when there exist compatible group actions on the data and visualization, as proposed by Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}.
+which means that a visualization is structure preserving when there exist compatible group actions on the data and visualization, as discussed by Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}.
+As illustrated in the commutative diagram in \autoref{fig:related-work:equivariance}, what this means is that the visual representation is consistent whether the data is scaled and then mapped to a graphic or whether the data is mapped to a graphic that is then modified in a compatible way.
+
+Although the Steven's scales were conceptualized as having group structure, the ordinal scale has a monoidal structure because partial orders ($\geq, \leq$) are not invertable.
+This means \textit{equivariance} cannot be used to test for structure preservation.
+Instead \textit{homomorphsim} can be used because it imposes fewer constraints on the underlying mathematical structure of the data.
 
 \subsubsection{homomorphism}
-\begin{figure}
+\begin{figure}[H]
   \includegraphics[width=1\columnwidth]{tex/homomorphism.png}
+  \caption{Encoding data as bar height using an inverse transform is not homomorphic because the largest number is mapped to the smallest bar while the max function returns the largest bar.}
+  \label{fig:related-work:homomorphism}
 \end{figure}
-For example, although the Steven's scales were conceptualized as having group structure, the ordinal scale have a monoidal structure because partial orders ($\geq, \leq$) are not invertable.
-This means we cannot use \textit{equivariance} to test for structure preservation instead we use \textit{homomorphsim} because it imposes fewer constraints on the underlying mathematical structure of the data.
 
@@
 \begin{definition}\label{def:homomorphism}
-  A function $f$ is \textbf{homomorphic} when $f(x_1 \circ x_2) = f(x_1) * f(x_2)$ and preserves identites $f(I_x) = I_y$ all $x, y \in X$
+  A function $f$ is \textbf{homomorphic} when $f(x_1 \circ x_2) = f(x_1) * f(x_2)$ and preserves identites $f(I_x) = I_y$ all $x, y \in X$ \cite{grimaldiDiscreteCombinatorialMathematics2006}
 
 \end{definition}
-which means that the operators $\circ$ and $*$ are compatiable.  \note{check grimalidi for citation}.
-
-As illustrated in \ref{fig:def:homomorphism}, structure preservation under ordering means that since 1 is less than 2, the visual transformation of 1 must be less than the visual transformation of two - here shown as 1 having a shorter bar than 2.
-We must define a < for visual compatable with the < for numbers.
-We handwave that for every relation we can write a compatabile closed function,  e.g.
-we can replace $\leq$ with $\min$, such that the operation is closed.
-
-As shown in \autoref{fig:def:equivariance} and \autoref{fig:def:homomorphsim}, the basis of asserting structure preservation is that there exist compatiable actions and these actions commute; we generalize compatability and commutability as the basis for structure preservation in \autoref{sec:artist:equivariant:data}.
-
-A brief visualization oriented introduction to category theory is in Vickers et al \cite{vickersUnderstandingVisualizationFormal2013}, but they are applying category theory to semantic concerns about visualization design rather than library architecture.
+which means that the operators $\circ$ and $*$ are compatible.
+In \autoref{fig:related-work:homomorphism}, the $\geq$ operator is defined as the compatible closed functions \texttt{max} and the inverse transform
+is not homomorphic because it does not encode the maximum data value as the maximum bar value.
+
+As shown in \autoref{fig:def:equivariance} and \autoref{fig:def:homomorphsim}, a function can be homomorphic but not equivariant, such as an exponential encoding, or equivariant but not homomorphic, such as the inverse encoding.
+A function can also be homomorphic (or equivariant) with respect to one action but not with respect to another.
+The encoding transforms in visualization tools are expected to preserve the structure of whatever input they receive; therefore a methodology for codifying arbitrary structure is presented in \autoref{sec:atct:fb:fiber} and \autoref{sec:artist:equivariant:data} presents a generalization of equivariance and homomorphism for evaluating structure preservation.
 
 \subsection{Topology}
-
+\label{sec:related-work:continuity}
 \begin{figure}[H]
@@
 
+\note{this got moved}
+In \autoref{fig:related-work:continuity:ktypes}, the topology is a continuous 3 dimensional surface encapsulating time and position, while the fields are date, location, station, pressure, and temperature.
+Our paper provides a language to express how the scatter plot, line plot, and heatmap are preserving different structural properties of the dataset, thereby generalizing Bertin's codification of how the topology of observations must match the class of representation (i.e.
+point , line, area) and how the variation possible in a field match the variation in the corresponding visual variable \cite{bertinSemiologyGraphicsDiagrams2011} and Mackinlay's work on how structure preservation includes preserving the mathimatical relations between elements of the same and different fields \cite{mackinlayAutomatingDesignGraphical1986}.
+
 \subsection{Structure Preservation In Software}
@@
 While some visualization tools aim to automate the pairing of data with structure preserving visual representations, such as Tableau\cite{StoltePolaris2002,hanrahanVizQL2006,MackinlayShowme2007}, many visualization libraries leave that choice to the user.
-For example, connectivity assumptions tend to be embedded in each of the visual algorithms of `building block` libraries, a term used by Wongsuphasawat \cite{wongsuphasawatNavigatingWideWorld2021} to describe libraries that provide modular components for building elements of a visualization, such as functions for making boxes or translating data values to colors.
+For example, connectivity assumptions tend to be embedded in each of the visual algorithms of `building block` libraries, a term used by Wongsuphasawat \cite{wongsuphasawatNavigatingWideWorld2021,wongsuphasawatNavigatingWideWorld2020} to describe libraries that provide modular components for building elements of a visualization, such as functions for making boxes or translating data values to colors.
 In building block libraries such as Matplotlib\cite{hunterMatplotlib2DGraphics2007} and D3\cite{bostockDataDrivenDocuments2011} assumptions about connectivity are embedded in the interfaces such that the API is inconsistent across plot types.
@@
 By assuming the structure of their data, these domain specific libraries can provide more cohesive interfaces for a much more limited set of visualization algorithms than the building block libraries offer.
+
+The coupling of values with their allowed transformations is roughly equivalent to the data and methods of and on a software class.
+Category theory refines the mapping between math and software and its use can lead to more consistent software interfaces in visualization software libraries \cite{wielsManagementEvolvingSpecifications1998,yorgeyMonoidsThemeVariations2012}.
+A brief visualization oriented introduction to category theory is in Vickers et al \cite{vickersUnderstandingVisualizationFormal2013}, but they are applying category theory to semantic concerns about visualization design rather than library architecture.
 
@@
 
-This project has been made possible in part by grant number 2019-207333 and from the
-\note{Cycle 3 grant #} Chan Zuckerberg Initiative DAF, an advised fund of Silicon Valley Community Foundation
+This project has been made possible in part by grant number 2019-207333 and \note{Cycle 3 grant number} Chan Zuckerberg Initiative DAF, an advised fund of Silicon Valley Community Foundation
 

### 2024-12-03 paper/main.tex
relayed work

@@
 Instead \textit{homomorphsim} can be used because it imposes fewer constraints on the underlying mathematical structure of the data.
-
-\subsubsection{homomorphism}
 \begin{figure}[H]
@@
 \label{sec:related-work:continuity}
-\begin{figure}[H]
+Visual algorithms assume the topology of their input data, as described in taxonomies of visualization algorithms Chi\cite{chiTaxonomyVisualizationTechniques2000} and by Troy and M\"{o}ller \cite{toryRethinkingVisualizationHighlevel2004}, but generally do not verify that input structure.
+
+\begin{figure}[]
   \includegraphics[width=1\columnwidth]{k_different_types.png}
-  \caption{
-  \note{add first panel which is table + all K type, second is scatter, etc...}
+  \caption{This weather station data has multiple embedded continuities - points at each time and position, timeseries at each position, and maps at each time.
+The corresponding visualizations - bar chart, timeseries, and map - each preserve the continuity of the subset of the data they visualize such that there is continuous map from the visualization (continuity) to the data (continuity).}
 %
-  \label{fig:related-work:continuity:ktypes}}
+  \label{fig:related-work:continuity:ktypes}
 \end{figure}
-\note{lead with homomorphism}
-Borrowing structures from algebraic topology allows us to describe structure on individual elements in a way that holds true for the entire dataset, meaning we can use the same tools to check if structre is preserved regardless of whether the data fits in memory, is distributed, or is streaming.
-Our framework generalizesis because, informally, a topology $\mathcal{T}$ is a partitioning of a space into its subspaces in a way where those subspaces can be composed back into the space; the choice of partitions is how we encode the connectivity of the data and visualizations that we are interested in.
+
+For example, a \texttt{line} algorithm often does not have a way to query whether a list of (x,y) coordinates is the distinct rows, the time series, or the list of stations in \autoref{fig:related-work:continuity:ktypes}.
+While plotting the time series as a continuous line would be correct, it would be incorrect for a visualization to indicate that the distinct rows or stations are connected in a 1D continuous manner because it introduces ambiguity over which part of the line maps back to the data.
+A map that by definition has continous maps between two spaces, such as data and graphics, is called a \textit{homeomorphism}\cite{riehlCategoryTheoryContext}:
+
+\begin{definition}
+  A function $f$ is a $homeomorphism$ if it is bijective, continous, and has a continous inverse function $f^{-1}$.
+\end{definition}
+
+Therefore the the bar plot, line plot, and heatmap in \autoref{fig:related-work:continuity:ktypes} have a homemeomorphic relationship to the 0D points, 1D timeseries, and 2D map continuities embedded in the continuous 3 dimensional surface encapsulating time and position.
+Using homeomorphism to test whether continuity is preserved generalizes Bertin's codification of how the topology of observations must match the class of representation (i.e.
+point , line, area) \cite{bertinSemiologyGraphicsDiagrams2011} and Wilkinson's assertion that connectivity be preserved \cite{wilkinsonGrammarGraphics2005}.
+
+Describing connectivity using the language of topology allows for describing individual elements in a way that holds true for the entire dataset.
+This means that the framework described in this paper can be used regardless of whether the data fits in memory, is distributed, or is streaming, because, informally, a topology $\mathcal{T}$ is a partitioning of a space into its subspaces in a way where those subspaces can be composed back into the total space:
 
@@
 
-Visual algorithms assume the topology of their input data, as described in taxonomies of visualization algorithms Chi\cite{chiTaxonomyVisualizationTechniques2000} and by Troy and M\"{o}ller \cite{toryRethinkingVisualizationHighlevel2004}, but generally do not verify that input structure.
-For example, a \texttt{line} algorithm often does not have a way to query whether a list of (x,y) coordinates is the distinct rows, the time series, or the list of stations in \autoref{fig:related-work:continuity:ktypes}.
-While plotting the time series as a continuous line would be correct, it would be incorrect for a visualization to indicate that the distinct rows or stations are connected in a 1D continuous manner because it introduces ambiguity over which part of the line maps back to the data.
-A map that by definition has continous maps between two spaces, such as data and graphics, is called a homeomorphism\cite{riehlCategoryTheoryContext}.
-\note{add functional def of bijective and continous
-}
-\begin{definition}
-  A function $f$ is a $homeomorphism$ if it is bijective, continous, and has a continous inverse function $f^{-1}$.
-\end{definition}
-
-To encode topology and field structure, we extend Butler's work on using a mathematical structure called fiber bundles as an abstract data representation in visualization \cite{butlerVectorBundleClassesForm1992, butlerVisualizationModelBased1989}.
+This means that the data can be partitioned arbitrarily and still retain it's continuity; for example the topology of the data in \autoref{fig:fig:related-work:continuity:ktypes} could be encoded as the set of individual points (one at each time and station) or set of timeseries (one per station) or set of maps (one per time).
+In this topological model, semantic indexing as described by Munzner's key-value model of data structure act as different ways of building the partitions.
+It also makes clearer when different labeling schemes refer to the same point, for example how 0-360 and 180E-180W are two ways of labeling longitude.
+To encode topology and field structure in a way that is both uniform and generalizable, we extend Butler's work on using a mathematical structure called fiber bundles as an abstract data representation in visualization \cite{butlerVectorBundleClassesForm1992, butlerVisualizationModelBased1989}.
 We sketch out fiber bundles in \autoref{sec:atct:fiber-bundles}, but Butler provides a thorough introduction to bundles for visualization practitioners.
-
-\note{this got moved}
-In \autoref{fig:related-work:continuity:ktypes}, the topology is a continuous 3 dimensional surface encapsulating time and position, while the fields are date, location, station, pressure, and temperature.
-Our paper provides a language to express how the scatter plot, line plot, and heatmap are preserving different structural properties of the dataset, thereby generalizing Bertin's codification of how the topology of observations must match the class of representation (i.e.
-point , line, area) and how the variation possible in a field match the variation in the corresponding visual variable \cite{bertinSemiologyGraphicsDiagrams2011} and Mackinlay's work on how structure preservation includes preserving the mathimatical relations between elements of the same and different fields \cite{mackinlayAutomatingDesignGraphical1986}.
 
@@
 
-The coupling of values with their allowed transformations is roughly equivalent to the data and methods of and on a software class.
-Category theory refines the mapping between math and software and its use can lead to more consistent software interfaces in visualization software libraries \cite{wielsManagementEvolvingSpecifications1998,yorgeyMonoidsThemeVariations2012}.
+We propose that the cohesion of domain specific library APIs is obtainable using the uniform data model described in \autoref{sec:atct:fiber-bundles} while the expressivity of building block libraries can be preserved by defining explicit structure preserving constraints on the library components, as described in \autoref{sec:artist}.
+Because category theory constructions map cleanly to objects and functions, using category theory to express the structure and constraints can lead to more consistent software interfaces in visualization software libraries \cite{wielsManagementEvolvingSpecifications1998,yorgeyMonoidsThemeVariations2012}.
 A brief visualization oriented introduction to category theory is in Vickers et al \cite{vickersUnderstandingVisualizationFormal2013}, but they are applying category theory to semantic concerns about visualization design rather than library architecture.

### 2024-12-04 paper/main.tex
related work

@@
   \caption{This weather station data has multiple embedded continuities - points at each time and position, timeseries at each position, and maps at each time.
-The corresponding visualizations - bar chart, timeseries, and map - each preserve the continuity of the subset of the data they visualize such that there is continuous map from the visualization (continuity) to the data (continuity).}
+The corresponding visualizations - bar chart, timeseries, and map - each preserve the continuity of the subset of the data they visualize by not introducing or leaving out values and preserving the relative positioning of continuous values.}
 %
@@
 While plotting the time series as a continuous line would be correct, it would be incorrect for a visualization to indicate that the distinct rows or stations are connected in a 1D continuous manner because it introduces ambiguity over which part of the line maps back to the data.
-A map that by definition has continous maps between two spaces, such as data and graphics, is called a \textit{homeomorphism}\cite{riehlCategoryTheoryContext}:
+A map that by definition has continuous maps between the input and output spaces, such as data and graphics, is called a \textit{homeomorphism}\cite{riehlCategoryTheoryContext}:
 
 \begin{definition}
-  A function $f$ is a $homeomorphism$ if it is bijective, continous, and has a continous inverse function $f^{-1}$.
+  A function $f$ is a $homeomorphism$ if it is bijective, continuous, and has a continuous inverse function $f^{-1}$.
 \end{definition}
 
-Therefore the the bar plot, line plot, and heatmap in \autoref{fig:related-work:continuity:ktypes} have a homemeomorphic relationship to the 0D points, 1D timeseries, and 2D map continuities embedded in the continuous 3 dimensional surface encapsulating time and position.
-Using homeomorphism to test whether continuity is preserved generalizes Bertin's codification of how the topology of observations must match the class of representation (i.e.
-point , line, area) \cite{bertinSemiologyGraphicsDiagrams2011} and Wilkinson's assertion that connectivity be preserved \cite{wilkinsonGrammarGraphics2005}.
+The bar plot, line plot, and heatmap in \autoref{fig:related-work:continuity:ktypes} have a homemeomorphic relationship to the 0D points, 1D timeseries, and 2D surface continuities embedded in the continuous 3 dimensional surface encapsulating time and position because each point of the visualization maps back into a point in its corresponding indexing space in the cube.
+Using homeomorphism to test whether continuity is preserved formalizes Bertin's codification of how the topology of observations matches the class of representation (i.e.
+point , line, area) \cite{bertinSemiologyGraphicsDiagrams2011} and Wilkinson's assertion that connectivity must be preserved \cite{wilkinsonGrammarGraphics2005}.
 
@@
 \begin{definition}
-A \textbf{topology} $\mathcal{T}$ on a space $X$ is a collection of open sets\footnote{Open sets (open subsets) are a generalization of open intervals to n dimensional spaces.
+A \textcolor{base}{\textbf{topology}} $\mathcal{T}$ on a space $X$ is a collection of open sets\footnote{Open sets (open subsets) are a generalization of open intervals to n dimensional spaces.
 For example, an open ball is the set of points inside the ball and excludes points on the surface of the ball.
@@
 
-Domain specific libraries are designed with the assumption of continuities that are common in the domain \cite{HeerSoftware2006}, and therefore can somewhat restrict their API to choices that that are appropriate for the domain.
-For example, a tabular topological structure of discrete rows, as illustrated in \autoref{fig:related-work:continuity:ktypes}, is assumed by
-A Presentation Tool\cite{mackinlayAutomatingDesignGraphical1986, mackinlayAutomatingDesignGraphical1986} and grammar of graphics\cite{wilkinsonGrammarGraphics2005} and the ggplot\cite{wickhamGgplot2ElegantGraphics2016}, vega\cite{satyanarayanDeclarativeInteractionDesign2014}, and altair\cite{vanderplasAltairInteractiveStatistical2018} libraries built on these frameworks.
+Domain specific libraries are designed with the assumption of continuities that are common in the domain \cite{HeerSoftware2006}, and therefore can somewhat restrict their API to choices that are appropriate for the domain.
+For example, a tabular topological structure of discrete rows, as illustrated in \autoref{fig:related-work:continuity:ktypes}, is assumed by A Presentation Tool\cite{mackinlayAutomatingDesignGraphical1986, mackinlayAutomatingDesignGraphical1986} and grammar of graphics\cite{wilkinsonGrammarGraphics2005} and the ggplot\cite{wickhamGgplot2ElegantGraphics2016}, vega\cite{satyanarayanDeclarativeInteractionDesign2014}, and altair\cite{vanderplasAltairInteractiveStatistical2018} libraries built on these frameworks.
 Image libraries such as Napari\cite{nicholas_sofroniew_2021_4533308} and ImageJ\cite{schneiderNIHImageImageJ2012} and its humanities ImagePlot\cite{studiesCulturevisImageplot2021} plugin assume that the input is 2D continuous.

### 2024-12-06 paper/main.tex
intro to bundles

@@
 
-\begin{figure}[]
+\begin{figure}[H]
   \includegraphics[width=1\columnwidth]{k_different_types.png}
@@
 
-The bar plot, line plot, and heatmap in \autoref{fig:related-work:continuity:ktypes} have a homemeomorphic relationship to the 0D points, 1D timeseries, and 2D surface continuities embedded in the continuous 3 dimensional surface encapsulating time and position because each point of the visualization maps back into a point in its corresponding indexing space in the cube.
+The bar plot, line plot, and heatmap in \autoref{fig:related-work:continuity:ktypes} have a homemeomorphic relationship to the 0D ($\bullet$)\points, 1D (--) timeseries, and 2D *$\blacksquare$) surface continuities embedded in the continuous 3 dimensional surface encapsulating time and position because each point of the visualization maps back into a point in its corresponding indexing space in the cube.
 Using homeomorphism to test whether continuity is preserved formalizes Bertin's codification of how the topology of observations matches the class of representation (i.e.
@@
 
-Describing connectivity using the language of topology allows for describing individual elements in a way that holds true for the entire dataset.
-This means that the framework described in this paper can be used regardless of whether the data fits in memory, is distributed, or is streaming, because, informally, a topology $\mathcal{T}$ is a partitioning of a space into its subspaces in a way where those subspaces can be composed back into the total space:
-
-\begin{definition}
-A \textcolor{base}{\textbf{topology}} $\mathcal{T}$ on a space $X$ is a collection of open sets\footnote{Open sets (open subsets) are a generalization of open intervals to n dimensional spaces.
-For example, an open ball is the set of points inside the ball and excludes points on the surface of the ball.
-\cite{weissteinOpenSet,bradleyTopologyVsTopology}} \openset\ of $X$.
-This collection has the properties that the empty set $\varnothing$ and $\dbase$ are in $\mathcal{T}$, the union of open sets in $\mathcal{T}$ is also in $\mathcal{T}$, and any finite intersection of open sets in $\mathcal{T}$ is also in $\mathscr{T}$.
-\cite{bradleyTopologyCategoricalApproach2020}
+Describing connectivity using the language of topology allows for describing individual elements in a way that holds true whether the data fits in memory, is distributed, or is streaming.
+This is because, informally, a topology $\mathcal{T}$ on the underlying data indexing space (which is a proxy for the continuity), is a partitioning of that space such that the partitions have the same mathematical properties as each other.
+The partitions must also be composable in a continuity and property preserving way.
+There are various equivalent definitions of topology, but here we show the neighborhood axiomatization because it is most analogous the data access model of index (point) in subset (neighborhood) of all indices (mathematical space).
+
+\begin{definition}\label{def:topology}
+Given a space $X$, a point $x$ in $X$, and a function $\mathcal{N}(x)$ that assigns to $x$ subsets of $X$ , then $X$ with  $\mathcal{N}$ is a \textcolor{base}{topological space} if for each $x$ in $X$: \cite{brownronaldTopologyGroupoids2006}
+\begin{enumerate}
+  \item if $N$ is a neighborhood $N \in \mathcal{N}(x)$ of $x$ then $x \in N$
+  \item every superset of a neighborhood of x is a neighborhood of x; therefore a union of a neighborhood and an arbitrary subset of $X$ is a neighborhood of $x$
+  \item the intersection of any two neighborhoods of $x$ is a neighborhood of $x$
+  \item any neighborhood $N$ of $x$ contains a smaller neighborhood $M \subset N$ such that $N$ is also a neighborhood of each of the other points in $M$
+\end{enumerate}
 \end{definition}
 
-This means that the data can be partitioned arbitrarily and still retain it's continuity; for example the topology of the data in \autoref{fig:fig:related-work:continuity:ktypes} could be encoded as the set of individual points (one at each time and station) or set of timeseries (one per station) or set of maps (one per time).
-In this topological model, semantic indexing as described by Munzner's key-value model of data structure act as different ways of building the partitions.
+For example, in the indexing cube in \autoref{fig:fig:related-work:continuity:ktypes}, the brown surface and blue rectangle are both neighborhoods of the index for the measurement in Albany on June 09.
+The blue rectangle is also a neighborhood of
+the index for the measurement in Albany on March 09.
+The indexing cube is a neighborhood for both of these indices.
+While \autoref{def:topology} applies broadly to topological spaces, in this paper we usually model the indexing space as CW-complexes.
+CW-complexes are a class of topological spaces built by gluing together n-dimensional balls (which include points, intervals, filled circles, filled spheres, etc.) using continuous attaching maps.
+In our topological model of indexing, semantic indexing as described by Munzner's key-value model of data structure act as different ways of building the neighborhoods.
 It also makes clearer when different labeling schemes refer to the same point, for example how 0-360 and 180E-180W are two ways of labeling longitude.
@@
 \label{sec:atct}
-In this section, we propose a mathematical abstraction of the data input and graphic pre-rendered output.
+In this section, we propose a mathematical abstraction of the data input and pre-rendered graphic output.
 This mathematical abstraction has a robust language for expressing topology and fields; expresses how to verify that data continuity is preserved on subset, distributed, and streaming data representations; and formalizes the expectation of a correspondence between data and visual elements.
@@
 We model data using a mathematical representation of data that can encode topological properties, field types, and data values in a uniform manner using a structure from algebraic topology called a fiber bundle.
-We extend Butler's proposal of bundles as abstract visualization data type \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989} by incorporating Spivak's methodology for encoding named data types from his fiber bundle representation of relational databases \cite{spivakDatabasesAreCategories2010,spivakSimplicialDatabases2009}.
+We extend Butler's proposal of using bundles as an abstraction for visualization data\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989} by incorporating Spivak's methodology for encoding named data types from his fiber bundle representation of relational databases \cite{spivakDatabasesAreCategories2010,spivakSimplicialDatabases2009}.
 We build on this work to describe how to encode the connectivity of the data as a topological space, separately encode the fields as their own topological space with a typing system, and express the mappings between these two spaces.
@@
 
-%\caption{The local trivializations $\dtotal_{0}$ and $\dtotal_{1}$ are subspaces of $\dtotal$ where the fibers $\dfiber_{0}$ are equal, here illustrated as all pointing in the same direction.
-These local trivializations are themselves locally trivial fiber bundles.}
-
 \begin{definition}\label{def:fiber_bundle}
@@
 
-    \begin{equation}
-      \label{eq:atct:fb:intro}
-      \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-       \dfiberc
-        \arrow[r, hook, color=total] \&
-        \dtotalc
-        \arrow[d, "\pi"',color=total, two heads] \\
-         \&
-      \dbasec
-         \arrow[u, "\dsectionc"', bend right, pos=.5, color=section, dashed]
+   \begin{equation} \label{eq:atct:fb:intro}
+    \begin{tikzcd}
+      \dfiberc \arrow[r, hook, color=total] & \dtotalc \arrow[r, "\pi", color=total, two heads] & \dbasec
       \end{tikzcd}
@@
 
+\begin{minipage}{.5\columnwidth}
+\begin{definition}\label{def:fiber_bundle:section}
+  \\
+  A \textcolor{section}{\textbf{section}} $\dsectionc: \dbasec \rightarrow \dtotalc$ over a fiber bundle is a smooth right inverse of $\pi$: $\pi(\dsection(\dbasepoint)) = \dbasepoint$ for all $\dbasepoint \in \dbase$
+\end{definition}
+\end{minipage}
+\begin{minipage}{.4\columnwidth}
+  \begin{equation} \label{eq:atct:fb:intro-sec}
+    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
+     \dfiberc
+      \arrow[r, hook, color=total] \&
+      \dtotalc
+      \arrow[d, "\pi"',color=total, two heads] \\
+       \&
+    \dbasec
+       \arrow[u, "\dsectionc"', bend right, pos=.5, color=section, dashed]
+    \end{tikzcd}
+  \end{equation}
+\end{minipage}
+
+\begin{figure}[!h]
+       \includegraphics{fb.png}
+       \label{fig:atct:fbdef}
+       \caption{The space of all data values encoded by this fiber bundle can be modeled as a \textcolor{total}{rectangle} total space.
+Each dataset in this dataspace lies along the interval \textcolor{base}{[0,1]} base space.
+Each dataset has values along the \textcolor{fiber}{$0 \rightarrow 1$} interval fiber.
+One dataset embedded in this total space is the \textcolor{section}{sin} section over the bundle.}
+  \end{figure}
+
 \note{Add motiviating note here!}
@@
 \note{add transition map construction formula}
-
-\begin{figure}[H]
-  \includegraphics[width=1\columnwidth]{figures/cyl_mob.png}
-  \caption{\note{figure out short caption here}\label{fig:cyl_mob_bundles}}
-\end{figure}
-As demonstrated in \autoref{fig:cyl_mob_bundles} different types of total spaces can be decomposed into the same local trivializations; therefore the distinguishing factor between total spaces can be the transition maps for constructing the spaces from trivializations.
-
-A \textit{trivial bundle} is directly isomorphic to $\dbasec\times\dfiberc$, and for any choice of cover of $\dbasec$ by overlapping opensets, we can choose local trivializations so that all transition maps are identity maps.
-In the example in Figure~\ref{fig:cyl_mob_bundles}, we use arrows $\uparrow$ to denote fiber alignments, illustrating that by choosing simple enough trivializations, we get identity maps in the transition maps.
-
-%In a \textit{trivial bundle}, the fibers are identical $\dfiber = \dfiber_{k}$ for all points $\dbasepoint \in dbase$ such that all the transition maps are also identical.
-For example, in the cylinder case the fibers in the overlapping regions of the trivializations point in the same direction $\uparrow$ to illustrate that they are equal $\dfiber_0\restriction_{\openset_1\cap\openset_2} = \dfiber_{1}\restriction_{\openset_1\cap\openset_2}$.
-Because the fibers point in the same direction, the  transition maps at both intersections map the fiber to itself $\uparrow\rightarrow\uparrow$.
-
-A \textit{non-trivial bundle} on the other hand is one that can not be constructed as $\dbasec\times\dfiberc$, but where the construction is more complicated.
-Then for any choice of local trivializations, there is at least one transition map that is not an identity \cite{hatcherAlgebraicTopology2002}.
-For example, in the Möbius band case, while $\dfiber_0\restriction_{(2\pi/5-\varepsilon, 2\pi/5+\varepsilon)}\to\dfiber_1\restriction_{(2\pi/5-\varepsilon, 2\pi/5+\varepsilon)}$ can be chosen to be an identity map (after transforming $\dbase_1\times[-2,2]\to\dbase_1\times[-1,1]$ by rescaling), the other transition map component $\dfiber_0\restriction_{(-\varepsilon,\varepsilon)}\to\dfiber_1\restriction_{(-\varepsilon,\varepsilon)}$ has to flip the fibers to align them between the pieces.
-
-\note{ending on sectionsis probably better for our purposes!}
-\begin{definition} A \textcolor{section}{\textbf{section}} $\dsectionc: \dbasec \rightarrow \dtotalc$ over a fiber bundle is a smooth right inverse of $\pi$: $\pi(\dsection(\dbasepoint)) = \dbasepoint$ for all $\dbasepoint \in \dbase$
-\end{definition}
 
@@
 
-\subsubsection{Topological Structure: Base Space \dbase}
+\subsubsection{\textcolor{base}{Topological Structure: Base Space \dbase}}
 \label{sec:atct:fb:base}
+\note{all the munzner stuff got moved up to related work}
 \note{base space of bundle for key/value pair, topology in general 'cause data + viz continuity in same way}
@@
 For example, the disjoint union of two bundles, as shown in \autoref{fig:atct:local_trivalizations}, is the coproduct $\dbase^{a} \subseteq_{\dbase^{c}} \dbase^{b}$ over an overlap $\dbase^{c}$ and therefore the inclusion morphism must be commutative:
-\begin{figure}
+\begin{figure}[h]
   \includegraphics*[width=1\columnwidth]{figures/tex/k_coproduct.pdf}
@@
 
+\section{Trivial and non-trivial bundles}
+
+\begin{figure}[H]
+  \includegraphics[width=1\columnwidth]{figures/}
+  \caption{\note{figure out short caption here}\label{fig:cyl_mob_bundles}}
+\end{figure}
+In \autoref{fig:cyl_mob_bundles} the cylinder is an example of a trivial bundle and the non mobius band is an example of a non-trivial bundle.
+We explain the difference here to show that our framework is not limited to trivial bundles, but generally  the bundle type does not matter because both types can be decomposed into trivial fiber bundles.
+For example, both the cylinder and mobius band can be decomposed into the fiber bundles shown in \autoref{fig:atct:fbdef}.
+
+Generally, the distinguishing factor between a trivial bundle and a non-trivial bun transition maps for constructing the spaces from trivializations:
+\begin{description}
+  \item[\textit{trivial bundle}] is directly isomorphic to $\dbasec\times\dfiberc$, and for any choice of cover of $\dbasec$ by overlapping opensets, we can choose local trivializations so that all transition maps are identity maps.
+  \item[\textit{non-trivial bundle}] on the other hand is one that can not be constructed as $\dbasec\times\dfiberc$, but where the construction is more complicated.
+Then for any choice of local trivializations, there is at least one transition map that is not an identity \cite{hatcherAlgebraicTopology2002}.
+\end{description}
+
+In the example in Figure~\ref{fig:cyl_mob_bundles}, we use arrows $\uparrow$ to denote fiber alignments, illustrating that by choosing simple enough trivializations, we get identity maps in the transition maps.
+In the cylinder case the fibers in the overlapping regions of the trivializations point in the same direction $\uparrow$ to illustrate that they are equal $\dfiber_0\restriction_{\openset_1\cap\openset_2} = \dfiber_{1}\restriction_{\openset_1\cap\openset_2}$.
+Because the fibers point in the same direction, the  transition maps at both intersections map the fiber to itself $\uparrow\rightarrow\uparrow$.
+In the Möbius band case, while $\dfiber_0\restriction_{(2\pi/5-\varepsilon, 2\pi/5+\varepsilon)}\to\dfiber_1\restriction_{(2\pi/5-\varepsilon, 2\pi/5+\varepsilon)}$ can be chosen to be an identity map (after transforming $\dbase_1\times[-2,2]\to\dbase_1\times[-1,1]$ by rescaling), the other transition map component $\dfiber_0\restriction_{(-\varepsilon,\varepsilon)}\to\dfiber_1\restriction_{(-\varepsilon,\varepsilon)}$ has to flip the fibers to align them between the pieces.
+
 \pagebreak

### 2024-12-06 paper/figcode/nontrivial.py

@@ -6,6 +6,7 @@ import matplotlib.pyplot as plt
 import matplotlib.tri as mtri
 import matplotlib.patches as mpatches
 import matplotlib.transforms as mtransforms
+import matplotlib.ticker as mticker
 from mpl_toolkits.mplot3d import proj3d
 
 import papercolors as pc
@@ -107,34 +108,10 @@ def plot_mobius_band(ax, mask, mask_color):
 
 
 def draw_segmented_bundle(ax, plot_total_space, labels):
-
-    bundle_label = [(r"$(\;$", None ), (f"$E_{labels[1]},$", 'total'),
-                    (f"$K_{labels[1]},$",'base'), 
-                    ("$\pi$,",None), (f"$F_{labels[1]}$", 'fiber'), (")", None)] 
-    
-    
-    fig = ax.figure
-    trans = ax.transAxes
-    for s, c in bundle_label:
-        pad= " " if s.find("(")>-1 else " "
-        fs = config['tfs']+2 if ((s.find(")") >-1 or  s.find("(") >-1)) else config['tfs']
-        text = ax.text2D(0.25, 1., s+pad, transform = trans, 
-                         color=pcd.get(c, config['bundle_color']),
-                         ha='center', va='center', fontsize=fs, family='monospace')
-        text.draw(fig.canvas.get_renderer())
-        ex = text.get_window_extent()
-        ex = fig.dpi_scale_trans.inverted().transform_bbox(ex)
-        trans = text.get_transform() + \
-                mtransforms.offset_copy(mtransforms.Affine2D(), fig=fig, x=ex.width, y=0)
-      
-     
-        
-
-
+   
     for (mask, (tc, fc, bc)) in [(config['mask1'], config['light_colors']), 
                              (config['mask2'], config['dark_colors'])]:
-
-
+        
         bounds = plot_total_space(ax, mask, tc)
         
         if mask.sum()>79:
@@ -148,7 +125,7 @@ def draw_segmented_bundle(ax, plot_total_space, labels):
             ax.add_artist(fp)
             ax.text3D(config['xB'][mask][pii], config['yB'][mask][pii], 
                       config['zb']+(bounds[0]['z'][pii] - config['zb'])/2, 
-                      "$\pi$", fontsize=config['fs'],
+                      r"$\pi$", fontsize=config['fs'],
                        color=config['bundle_color'], ha='left', va='center')
 
 
[…]

### 2024-12-11 paper/main.tex
sheaf rules

@@
 
-\begin{figure}[H]
+\begin{figure}
   \includegraphics*[width=1\columnwidth]{tex/equivariant.png}
@@
 Instead \textit{homomorphsim} can be used because it imposes fewer constraints on the underlying mathematical structure of the data.
-\begin{figure}[H]
+\begin{figure}
   \includegraphics[width=1\columnwidth]{tex/homomorphism.png}
@@
   A function $f$ is \textbf{homomorphic} when $f(x_1 \circ x_2) = f(x_1) * f(x_2)$ and preserves identites $f(I_x) = I_y$ all $x, y \in X$ \cite{grimaldiDiscreteCombinatorialMathematics2006}
-
 \end{definition}
+
 which means that the operators $\circ$ and $*$ are compatible.
@@
 \label{sec:related-work:continuity}
-Visual algorithms assume the topology of their input data, as described in taxonomies of visualization algorithms Chi\cite{chiTaxonomyVisualizationTechniques2000} and by Troy and M\"{o}ller \cite{toryRethinkingVisualizationHighlevel2004}, but generally do not verify that input structure.
 
@@
 
+Visual algorithms assume the topology of their input data, as described in taxonomies of visualization algorithms Chi\cite{chiTaxonomyVisualizationTechniques2000} and by Troy and M\"{o}ller \cite{toryRethinkingVisualizationHighlevel2004}, but generally do not verify that input structure.
 For example, a \texttt{line} algorithm often does not have a way to query whether a list of (x,y) coordinates is the distinct rows, the time series, or the list of stations in \autoref{fig:related-work:continuity:ktypes}.
@@
 
-The bar plot, line plot, and heatmap in \autoref{fig:related-work:continuity:ktypes} have a homemeomorphic relationship to the 0D ($\bullet$)\points, 1D (--) timeseries, and 2D *$\blacksquare$) surface continuities embedded in the continuous 3 dimensional surface encapsulating time and position because each point of the visualization maps back into a point in its corresponding indexing space in the cube.
+The bar plot, line plot, and heatmap in \autoref{fig:related-work:continuity:ktypes} have a homemeomorphic relationship to the 0D ($\bullet$) points, 1D (--) linear, and 2D ($\blacksquare$) surface continuities embedded in the continuous 3 dimensional surface encapsulating time and position because each point of the visualization maps back into a point in its corresponding indexing space in the cube.
 Using homeomorphism to test whether continuity is preserved formalizes Bertin's codification of how the topology of observations matches the class of representation (i.e.
@@
 
-Describing connectivity using the language of topology allows for describing individual elements in a way that holds true whether the data fits in memory, is distributed, or is streaming.
-This is because, informally, a topology $\mathcal{T}$ on the underlying data indexing space (which is a proxy for the continuity), is a partitioning of that space such that the partitions have the same mathematical properties as each other.
-The partitions must also be composable in a continuity and property preserving way.
-There are various equivalent definitions of topology, but here we show the neighborhood axiomatization because it is most analogous the data access model of index (point) in subset (neighborhood) of all indices (mathematical space).
-
-\begin{definition}\label{def:topology}
-Given a space $X$, a point $x$ in $X$, and a function $\mathcal{N}(x)$ that assigns to $x$ subsets of $X$ , then $X$ with  $\mathcal{N}$ is a \textcolor{base}{topological space} if for each $x$ in $X$: \cite{brownronaldTopologyGroupoids2006}
-\begin{enumerate}
-  \item if $N$ is a neighborhood $N \in \mathcal{N}(x)$ of $x$ then $x \in N$
-  \item every superset of a neighborhood of x is a neighborhood of x; therefore a union of a neighborhood and an arbitrary subset of $X$ is a neighborhood of $x$
-  \item the intersection of any two neighborhoods of $x$ is a neighborhood of $x$
-  \item any neighborhood $N$ of $x$ contains a smaller neighborhood $M \subset N$ such that $N$ is also a neighborhood of each of the other points in $M$
-\end{enumerate}
-\end{definition}
-
-For example, in the indexing cube in \autoref{fig:fig:related-work:continuity:ktypes}, the brown surface and blue rectangle are both neighborhoods of the index for the measurement in Albany on June 09.
-The blue rectangle is also a neighborhood of
-the index for the measurement in Albany on March 09.
-The indexing cube is a neighborhood for both of these indices.
-While \autoref{def:topology} applies broadly to topological spaces, in this paper we usually model the indexing space as CW-complexes.
-CW-complexes are a class of topological spaces built by gluing together n-dimensional balls (which include points, intervals, filled circles, filled spheres, etc.) using continuous attaching maps.
-In our topological model of indexing, semantic indexing as described by Munzner's key-value model of data structure act as different ways of building the neighborhoods.
-It also makes clearer when different labeling schemes refer to the same point, for example how 0-360 and 180E-180W are two ways of labeling longitude.
 To encode topology and field structure in a way that is both uniform and generalizable, we extend Butler's work on using a mathematical structure called fiber bundles as an abstract data representation in visualization \cite{butlerVectorBundleClassesForm1992, butlerVisualizationModelBased1989}.
+Using this topological model of indexing, semantic indexing as described by Munzner's key-value model of data structure \cite{munznerWhatDataAbstraction2014} act as different ways of partitioning the underlying data continuity.
+For example, the data cube in \autoref{fig:related-work:continuity:ktypes} could be subset into sets of timeseries where the key would be $station$, or subset into maps where the key would be $date$, or subset into station records where the keys are $(date, latitude, longitude)$.
+Using a topological model rather than semantic indexing also makes clearer when different labeling schemes refer to the same point, for example how 0-360 and 180E-180W are two ways of labeling longitude or how (date, lat, lon) and (date, station) refer to the same point.
 We sketch out fiber bundles in \autoref{sec:atct:fiber-bundles}, but Butler provides a thorough introduction to bundles for visualization practitioners.
@@
 
-\section{Formal Properties of Data \& Graphics}
+\section{Uniform Abstraction for Data \& Graphics}
 \label{sec:atct}
@@
 This mathematical abstraction has a robust language for expressing topology and fields; expresses how to verify that data continuity is preserved on subset, distributed, and streaming data representations; and formalizes the expectation of a correspondence between data and visual elements.
+Using these abstractions allows us to embed information about structure in dataset types:
+
+\begin{equation}
+  \textcolor{section}{\texttt{dataset}}: \textcolor{base}{\texttt{topology}} \rightarrow \textcolor{fiber}{\texttt{fields}}
+\end{equation}
+
+which can then be checked by visualization algorithms to ensure that the assumptions of the data match the assumptions of the algorithm.
+This typing system also extends to pre-rendered graphic output, allowing us to develop the structure preservation framework in \autoref{sec:artist} that ensures that output fields are equivalent to the input fields and that the topology of the output graphic is homeomorphic to the topology of the input.
 
@@
 \label{sec:atct:fiber-bundles}
+
 We model data using a mathematical representation of data that can encode topological properties, field types, and data values in a uniform manner using a structure from algebraic topology called a fiber bundle.
 We extend Butler's proposal of using bundles as an abstraction for visualization data\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989} by incorporating Spivak's methodology for encoding named data types from his fiber bundle representation of relational databases \cite{spivakDatabasesAreCategories2010,spivakSimplicialDatabases2009}.
-We build on this work to describe how to encode the connectivity of the data as a topological space, separately encode the fields as their own topological space with a typing system, and express the mappings between these two spaces.
-
-The fiber bundle abstraction is a structure fusing some \textcolor{base}{base space} (indexing into the data) with some \textcolor{fiber}{fiber space} (acting as the data domain).
-This sort of fusion could be done by a simple direct product (or join-operation), but with the fiber bundle abstraction we get the flexibility of expressing more complicated fusions than the direct product, as long as they can be constructed from building blocks that look like direct products.
+We build on this work to describe how to encode the connectivity of the data as a topological \textcolor{base}{base space} modeling the data indexing space, encode the fields as a \textcolor{fiber}{fiber space} that acts as the data schema (domain), and express the mappings between these two spaces as \textcolor{section}{section} functions that encode datasets as mappings from indexing space to field space \texttt{dataset: topology $\rightarrow$ fields}.
 
 \begin{definition}\label{def:fiber_bundle}
-   A \textbf{fiber bundle} $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and  bundle projection map $\pi: \dtotalc \rightarrow \dbasec$ \cite{FiberBundle2020,spanier1989algebraic}.
+   A \textbf{fiber bundle} $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and  bundle projection map $\pi: \dtotalc \rightarrow \dbasec$ \cite{spanier1989algebraic}.
 
@@
   \item all fibers in the bundle are isomorphic.
-Since all fibers are isomorphic, there is a uniquely determined \textcolor{fiber}{fiber space} \dfiberc\ given by the preimage of the projection $\pi$ at any point $\dbasepoint$ in the \textcolor{base}{base space} \dbasec: $\dfiberc = \pi^{-1}(k)$.
-  %the \textcolor{fiber}{fiber space} \dfiberc\ is the preimage of the projection function $\pi$ at a point $\dbasepoint$ in the \textcolor{base}{base space} \dbasec\ such that $\dfiber_{\dbasepoint} = \pi^{-1}(\dbasepoint)$.
-All fibers in a bundle are isomorphic such that $\dfiber \cong \dfiber_{\dbasepoint}$ for all points $\dbasepoint \in \dbase$.
+Since all fibers are isomorphic $\dfiber \cong \dfiber_{\dbasepoint}$ for all points $\dbasepoint \in \dbase$, there is a uniquely determined \textcolor{fiber}{fiber space} \dfiberc\ given by the preimage of the projection $\pi$ at any point $\dbasepoint$ in the \textcolor{base}{base space} \dbasec: $\dfiberc = \pi^{-1}(k)$.
   \item each point $\dbasepointc$ in the \textcolor{base}{base space} \dbasec\ has an open neighborhood $\openset_{\dbasepointc}$ such that the \textcolor{total}{total space} \dtotalc\ over the neighborhood is locally trivial.
-\textbf{Local triviality} means $\dtotal\restriction_{\openset} = \openset\times F$.
-In this paper we use $\dtotal\restriction_{\openset} = \pi^{-1}(\openset)$ to denote the preimage of an openset, and a \textbf{local trivialization} is a specific choice of neighborhoods and their preimages.
-  %there is an open neighborhood $\openset_{\dbasepoint}$ surrounding each point in the base space $\dbasepoint \in \openset_{\dbasepoint} \subset \dbase$ such that the \textcolor{total}{total space} \dtotalc\ over the neighborhood, denoted $\dtotal\restriction\openset$\footnote{The symbol $\restriction$ is the restriction operator\cite{RestrictionMathematics2022} defined in \texttt{amssymb}.
-For example $\pi^{-1}\restriction\openset$ is the inverse function $\pi^{-1}$ defined only on the points in $\openset$.
-Since $\pi^{-1}(\dbase)=\dtotal$, we use the shorthand $\dtotal\restriction_{\openset} \coloneqq \pi^{-1}\restriction_{\openset}$}, is locally trivial.
-The condition for local triviality is that $\dtotal\restriction_{\openset} = \openset \times \dfiber = \pi^{-1}\restriction_{\openset}$
 \end{enumerate}
 \end{definition}
+
+\textbf{Local triviality} means $\dtotal\vert_{\openset} = \openset\times F$.
+In this paper we use $\dtotal\vert_{\openset} = \pi^{-1}(\openset)$ to denote the preimage of an openset\footnote{Open sets (open subsets) are a generalization of open intervals to n dimensional spaces.
+For example, an open ball is the set of points inside the ball and excludes points on the surface of the ball.
+\cite{weissteinOpenSet,bradleyTopologyVsTopology}}, and a \textbf{local trivialization} is a specific choice of neighborhoods (described in \autoref{sec:atct:fb:base}) and their preimages such that the fibers in each preimage are identical $\dfiber = \dfiber_{\dbasepoint}$ for all points $\dbasepoint \in \openset$.
+All fiber bundles can be decomposed into sets of local trivializaions that are also bundles and we can specify a gluing scheme that reconstructs the fiber bundle from locally trivial pieces  by specifying\textbf{transition maps} for all overlaps of the local trivializations; therefore in this paper we assume that the fiber bundles are trivial bundles $\dtotal = \dbase \times \dfiber$ but the model also applies to non-trivial bundles.
+For a discussion of trivial and non-trivial bundles, see \autoref{sec:appendix:bundle_triviality}.
 
@@
 
+We propose that the total space of a bundle can encode the mathematical space in which a  dataset is embedded, the base space can encode the topological properties of the dataset, the fiber space can encode the data types of the fields of the dataset, and that the datasets can be encoded as section functions from the continuity to the fiber space.
+
 \begin{figure}[!h]
-       \includegraphics{fb.png}
-       \label{fig:atct:fbdef}
+       \includegraphics[width=1\columnwidth]{fb.png}
        \caption{The space of all data values encoded by this fiber bundle can be modeled as a \textcolor{total}{rectangle} total space.
@@
 Each dataset has values along the \textcolor{fiber}{$0 \rightarrow 1$} interval fiber.
-One dataset embedded in this total space is the \textcolor{section}{sin} section over the bundle.}
+One dataset embedded in this total space is the \textcolor{section}{sin} section over the bundle.}\label{fig:atct:fb}
   \end{figure}
 
-\note{Add motiviating note here!}
-Consider the circle.
-Points can be identified with angles (in radians, for instance), and we can imagine covering the circle with two overlapping intervals - for instance $\dbase_0=(-\varepsilon, 2\pi/5+\varepsilon)$ and $\dbase_1=(2\pi/5-\varepsilon, 2\pi+\varepsilon)$.
-We can imagine creating building blocks $\dbase_0\times[-1,1]$ and $\dbase_1\times[-2,2]$.
-
-There are two different ways to create a fiber bundle that has these building blocks as the local trivializations: on the one hand, we could imagine just a cylinder $S^1\times[-1,1]$ as a total space.
-The fibers would all be just the interval $[-1,1]$.
-Over $\dbase_0$, we could map from the local trivialization to the total space with the identity map, and over $\dbase_1$ we could map from the local trivialization to the total space by rescaling the fiber space $[-2,2]\to[-1,1]$.
-
-On the other hand, we could imagine creating a Möbius strip by flipping one end of $\dbase_1\times[-2,2]$ over before attaching it, so that while $\dbase_0\times[-1,1]$ gets mapped into the total space with the identity map, the part $(2\pi-\varepsilon,2\pi+\varepsilon)\times[-2,2]$ of $\dbase_1\times[-2,2]$ is mapped by the function $x\mapsto -x/2$ instead of the function $x\mapsto x/2$.
-
-These functions - using the isomorphisms $x\mapsto x$, $x\mapsto x/2$ and $x\mapsto -x/2$ to map from each rectangle into the cylinder compose into different maps $(\dbase_0\cap\dbase_1)\times[-2,2]\to(\dbase_0\cap\dbase_1)\times[-1,1]$ where the source is seen as a subset of $\dbase_1\times[-2,2]$ and the target is seen as a subset of $\dbase_0\times[-1,1]$.
-By specifying just these \textbf{transition maps} for all overlaps of cover elements in the local trivialization we can specify a gluing scheme that constructs the fiber bundle from these locally trivial pieces.
-
-%For example, in \autoref{fig:local_trivalizations}, the bundles $\dtotal_{0}=\dtotal\restriction_{[\epsilon^{-}, {\frac{2\pi}{5}}^{+}]}$ and $\dtotal_{1}=\dtotal\restriction_{[{\frac{2\pi}{5}}^{-}, \epsilon^{-}]}$ are local trivializations of an arbitrary bundle $\dtotal$.
-These local trivializations overlap on the intersections of the open subsets of each trivialization $[\epsilon^{-}, \epsilon^{+}]$ and $[{\frac{2\pi}{5}}^{-}, {\frac{2\pi}{5}}^{+}]$.
-A set of trivializations can be glued into a larger bundle by defining transition maps that describe how to align the fibers in the overlapping space such that the sections of the bundle remain continuous.
-\note{add transition map construction formula}
-
-\begin{figure}[H]
-    \includegraphics[width=\columnwidth]{figures/local_trivalizations.png}
-    \label{fig:atct:local_trivalizations}
-    \caption{Example of bundles $\dtotalc_0$ and $\dtotalc_1$, which are both trivializations
-    of a bundle $\dtotalc$ add in section here!}
-\end{figure}
-
-\autoref{fig:atct:local_trivalizations} shows an example of a two dimenensional bundle
-$\dtotal$.
-This bundle can encapsulate a set datasets where the data points are connected to each other along a unit circle $\dbasec$ = $\left[0, 2\pi\right]$ and the data values lie along an interval $dfiberc$ = $[0,1]$.
-Each dataset is encoded as a section, for example $sin$ and $cos$.
-The subspaces of $\dtotal_{0}$ and $\dtotal_{1}$ are local trivializations $\dtotal$; therefore the fibers are equal, here illustrated as all pointing in the same direction.
-These local trivializations are themselves locally trivial fiber bundles.
-
-We propose that the total space of a bundle can encode the mathematical space in which a  dataset is embedded, the base space can encode the topological properties of the dataset, and the fiber space can encode the data types of the record fields of the dataset.
-The map $\pi$ expresses the formal binding between having a typed data field, discussed in \autoref{sec:atct:fb:fiber}, and a corresponding point in the topological structure, which is discussed in \autoref{sec:atct:fb:base}.
+For example, the fiber bundle in \autoref{fig:atct:fb} encodes the space of all continuous functions that have a domain of $[0, 2\pi]$ and range $[0,1]$.
+Using a fiber bundle abstraction encodes that the dataset has a 1D linear continuity as the base space \dbase is the interval $[0,2\pi]$ and a field type that is a \texttt{float} in the range $[0,1]$.
+Therefore the type signature of the datasets in this fiber bundle, which is called a section \dsection, would be \texttt{dataset: $[0, 2\pi] \rightarrow [0,1]$}.
+One such dataset (section) is the $\sin$ function, which as shown in \autoref{fig:atct:fb} is defined via a function \desection from a point in the base space to a corresponding point in the fiber.
+Evaluating the section function over the entire base space yields the $\sin$ curve that is composed of points intersecting each fiber over the corresponding point.
+The local trivializations shown in \autoref{fig:atct:fb} are one way of decomposing the total bundle and conversely the bundle can be constructed from the local trivializations $\dbase = \dbase_{0} \oplus \dbase_{1}$.
+As shown, the section $\sin$ spans the trivializations in the same manner that it spans the bundle; this is analogous to how a dataset may span multiple tables or be collected in one table.
+The trivializations are glued together into the bundle at the overlapping region $\frac{2\pi}{5}$ by defining the transition map $\dfiber_{1} \rightarrow \dfiber_{2}$.
+Because the fibers in \autoref{fig:atct:fb} at $\frac{2\pi}{5}$ are aligned, the transition map is an identity map that take every value in $\dfiber_1$ and maps it to the same value in $\dfiber_2$ so that the sections, such as $\sin$, remain continuous.
+
 Fiber bundles are a good abstract data representation for visualization because the field type and topological structure are unconstrained in terms of dimensionality and the only conditions that must be satisfied are that every point in the base space has a corresponding field of values and the field types must be the same for every point in the base space.
@@
 \label{sec:atct:fb:base}
-\note{all the munzner stuff got moved up to related work}
-\note{base space of bundle for key/value pair, topology in general 'cause data + viz continuity in same way}
-Munzner~\cite{munznerChDataAbstraction} describes a key-value semantics for data representation and abstraction, where datasets can be queried with different kinds of keys depending on the dataset type - locations for values of spatial fields, times for values of time series, or indexes for values in flat or multidimensional tables.
-We propose that these different types of keys are instances of different underlying topological organization of the data, and that we can use the structure of the opensets in the topological space to encode the extent to which the keys can vary continuously: flat or multidimensional tables have discrete points as indices, time series have points on a line as indices and spatial fields (regardless of whether they are scalar or tensor fields) have points in a spatial domain as indices.
-Therefore we choose to encode the topological structure of the data as the \textcolor{base}{base space} of a fiber bundle.
+
+We encode the topological structure of the data as the \textcolor{base}{base space} of a fiber bundle.
+Describing connectivity using the language of topology allows for describing individual elements in a way that holds true whether the data fits in memory, is distributed, or is streaming.
+This is because, informally, a topology $\mathcal{T}$ on the underlying data indexing space (which is a proxy for the continuity), is a partitioning of that space such that the partitions are of the same mathematical type as each other and the partitioned space.
+The partitions must also be composable in a continuity and property preserving way.
+There are various equivalent definitions of topology, but here we use the neighborhood axiomatization because it is most analogous the data access model of index (point) in subset (neighborhood) of all indices (mathematical space).
+
+\begin{definition}\label{def:topology}
+Given a space $X$, a point $x$ in $X$, and a function $\mathcal{N}(x)$ that assigns to $x$ subsets of $X$ , then $X$ with  $\mathcal{N}$ is a \textcolor{base}{topological space} if for each $x$ in $X$: \cite{brownronaldTopologyGroupoids2006}
+\begin{enumerate}
+  \item if $N$ is a neighborhood $N \in \mathcal{N}(x)$ of $x$ then $x \in N$
+  \item every superset of a neighborhood of $x$ is a neighborhood of $x$; therefore a union of a neighborhood and adjacent points in $X$ is also a neighborhood of $x$
+  \item the intersection of any two neighborhoods of $x$ is a neighborhood of $x$
+  \item any neighborhood $N$ of $x$ contains a neighborhood $M \subset N$ of $x$ such that $N$ is a neighborhood of each of the points in $M$
+\end{enumerate}
+\end{definition}
+
+Therefore a neighborhood can grow (2) or shrink (3) so long as the point $x$ remains in the neighborhood (1) and every neighborhood contains neighborhoods of points adjacent to x (4).
+For example, in the indexing cube in \autoref{fig:fig:related-work:continuity:ktypes}, the brown surface and blue rectangle are both neighborhoods of the index for the measurement in Albany on June 09.
+The blue rectangle is also a neighborhood of the index for the measurement in Albany on March 09.
+The indexing cube is a neighborhood for both of these indices.
+While \autoref{def:topology} applies broadly to topological spaces, in this paper we usually model the indexing space as CW-complexes.
+CW-complexes are a class of topological spaces built by gluing together n-dimensional balls (which include points, intervals, filled circles, filled spheres, etc.) using continuous attaching maps.
 Because the base space of a fiber bundle is a quotient topology\cite{munkresElementsAlgebraicTopology1984}, it divides the topological space into the largest number of open sets such that $\pi$ remains a continuous function.
@@
 The bifunctorallows $\otimes$ us to express fields that contain complexly typed values.
-For example, wind can be represented as two fields $\dfiber_{speed} \times \dfiber_{direction}$ or a composite fiber field $\dfiber_{speed} \otimes \dfiber_{direction} = \dfiber_{wind}$.
-The $\otimes$ encapsulates both the sets associated with each fiber $\mathbb{R} \times \mathbb{R} = \mathbb{R}^{2}$ and the morphisms associated with each functor $(\dfunct_{speed}, \dfunct_{direction}) = \dfunct_{wind}$.
-Combining fibers can be verified by checking that when a fiber component $\dfiber^{c}$ is present in both $\dfiber^{a}$ and $\dfiber^{b}$, it is identical when projected out of either such that the *product* diagram commutes:
+For example, dates can be represented as three fields $\dfiber_{year} \times \dfiber_{month} \times \dfiber_{day}$ or a composite fiber field $\dfiber_{year} \times \dfiber_{month} \times \dfiber_{day} = \dfiber_{date}$.
+The $\otimes$ encapsulates both the sets of values associated with each fiber $\{y \in \mathbb{I}\vert 1992 \leq y \leq 2025\} \times \{m \in \mathbb{I}\vert 1 \leq m \leq 12 \} \times \{d \in \mathbb{I}\vert 1 \leq d \leq 31 \}$ and the composition function $ \otimes: \dfiber_{year} \times \dfiber_{month} \times \dfiber_{day} \rightarrow \dfiber_{date}$ would include a constraint to only return dates that have the right number of days for each month.
+The bifunctor also composes the morphisms associated with each category into a morphism on the composite category $(\dfunct_{year}, \dfunct_{month}, \dfunct_{day}) = \dfunct_{date}$.
+Combining fibers correctly can be verified by checking that when a fiber component $\dfiber^{c}$ is present in both $\dfiber^{a}$ and $\dfiber^{b}$, it is identical when projected out of either such that the *product* diagram commutes:
 
@@
 
-\note{this is really a bookkeeping thing - each field is at an index/ splitting a field shouldn't change values} This means that data with many fields is decomposed into its component fields, shared records stay the same.
+This means that data with many fields is decomposed into its component fields, records maintain their integrity.
 For example, the red (temperature, time, pressure) record separates into (temperature, time) and (pressure, time) records that share the same red time.
@@
 We encode data as a \textcolor{section}{section} \dsectionc\ of a bundle because this allows us to incorporate the topology and field types in the data definition.
-We can define these section functions locally, meaning that the section is (piece-wise) continuous over a specific open subset \openset\ of \dbase\
+We define section functions locally, meaning that the section is (piece-wise) continuous over a specific open subset \openset\ of \dbase\:
 \begin{equation}
@@
 
-In a trivial bundle, the total space is the product of the fiber and base space $\dtotalc = \dbasec \times \dfiberc$.
-This allows us to define global sections $\dsection: \dbase \rightarrow \dfiber \in \Gamma(\dbase, \dfiber)$ which we translate into a data signature of the form
+When a bundle is trivial $\dtotalc = \dbasec \times \dfiberc$, we can defined a global sections $\dsection: \dbase \rightarrow \dfiber \in \Gamma(\dbase, \dfiber)$ which we translate into a data signature of the form
 \begin{equation}
@@
 where $\dsection=\texttt{dataset}$, $\dbase=\texttt{topology}$ and $\dfiber=\texttt{fields}$.
-This type signature provides a method of explicitly stating the topology and field type of the data and generalizes to almost any topology and fiber type, provided that the total space is trivial.
-
-When the total space is non-trivial, we can use the fiber bundle property of local-triviality to define local sections $\dsection\restriction_{\openset} \in \Gamma(\openset_{\dbasepoint}, \dtotal\restriction_{\openset_{\dbasepoint}})$.
+When the bundle is non-trivial, we can use the fiber bundle property of local-triviality to define local sections $\dsection\restriction_{\openset} \in \Gamma(\openset_{\dbasepoint}, \dtotal\restriction_{\openset_{\dbasepoint}})$.
 A local section is defined over an open neighborhood  $\dbasepoint \in \openset \in \dbase$, which is an open set that surrounds a point \dbasepoint.
@@
 
-\subsubsection{Example}
-
-\begin{figure}[H]
-  \includegraphics[width=\columnwidth]{dbundle.png}
-  \caption{The table from \autoref{fig:related-work:continuity:ktypes} has a 3 dimensional fiber (name, temperature, precipitation), a 0D base space, and each row is a section.
-Each time series is a section of a bundle with a 2D fiber (time, precipitation) and a 1D base space encoding temporal continuity.
-Each location of the rain map is a section of a bundle with a 2D plane base encoding spatial continuity and a 3D fiber space encoding (latitude, longitude, precipitation)
-    \label{fig:atct:trivialbundle}}
-\end{figure}
-
-In \autoref{fig:atct:trivialbundle}, the base space $\dbase$ acts as an indexing space into the fiber space $\dfiber$.
-In the bundle encoding of the table, the indexing space is arbitrarily numbered keys; in the time series bundle, the base space is the interval [0,1]; and the map is sparse samples from a continuous space $[0,1]^{2}$ In contrast to a notion of a semantic binding between indexing space and field values, as proposed by Munzner\cite{munznerVisualizationAnalysisDesign2014}, the fields describing the continuity are part of the fiber.
-For example, the time is a fiber in the time series and the latitude and longitude are part of the map's fiber.
-This separation between connectivity and what it is called means the time or location can change units without a change to structure.
-It also provides a way to express data that may seem continuous but isn't, for example independent measurements over time.
-The data in \autoref{fig:atct:trivialbundle} comes from the same dataset; therefore we know that the bundles share connectivity and fiber space.
-It is expected that shared components be translated to visual elements in a consistent manner \cite{hullmanKeeping2018}, and in \autoref{sec:artist:operators} we introduce operators for expressing which components are shared and how to verify that they have been mapped into visual elements in a consistent manner.
-
-\subsubsection{Uniform Abstract Graphic Representation}
+\subsubsection{Example: Uniform Abstract Graphic Representation}
 One of the advantages of fiber bundles is that they are general enough that we can also encode the output of a visual algorithm as a bundle.
@@
 Each tiny colored box is an approximation of the return value of the same section function $\gsection$ evaluated on different points $\gbasepoint \in \gbase$ in the base space.
-\note{add alpha and z channels and very faded out marker point}
   \label{fig:atct:fb:graphic}}
@@
 While \autoref{fig:atct:fb:graphic} illustrates a highly idealized space with no overlaps, overlaps can be managed via a fiber element $\gfiber_{z}$ for ordering.
-It is left to the renderer to choose how to blend layers based on $\gfiber_{z}$ and $\gfiber_{a}$.
+It is left to the renderer to choose how to blend $\gfiber_{z}$ and $\gfiber_{a}$ layers.
 
@@
 
+\begin{figure}
+  \includegraphics[width=1\columnwidth]{sheaf_rules.png}
+  \caption{A sheaf has the conditions that sections are equal when they match on all subsets (\textit{locality}) and that the sections can be concatenated when they match on overlaps \textit{gluing}.}
+
+\end{figure}
 While presheaves preserve the rules for sets of sections, sheaves add on conditions for gluing individual sections over subspaces into cohesive sections over the whole space.
@@
 \begin{itemize}
-  \item \textit{locality} two sections in a sheaf are equal $\dsection^{a} = \dsection^{b}$ when they evaluate to the same values over the same set of open sets $\dsection^{a}|_{\openset} =  \dsection^{b}|_{\openset}$.
-  \item \textit{gluing} the union of sections defined on specific open sets is equivalent to one big section over the union of spaces $\dsection|_{\openset_{i} \cup \openset_{j}} = \dsection^{i}|_{\openset_i} \cup \dsection^{j}|_{\openset_j}$ if these sections agree on overlaps $\dsection^{i}|_{\openset_i\cap\openset_j} =  \dsection^{j}|_{\openset_i\cap\openset_j}$
+  \item \textit{locality} two sections in a sheaf are equal $\dsection^{a} = \dsection^{b}$ when they evaluate to the same values $\dsection^{a}|_{\openset_i} =  \dsection^{b}|_{\openset_i}$ over the open cover $\bigcup_{i\in I} \openset_{i} \subset \openset$ (indexed by $I$).
+  \item \textit{gluing} the union of sections defined over subspaces $\desection^{i} \in \Gamma(\openset_i, \dtotal|_{openset_i})$ is equivalent to a section defined over the whole space $\dsection\vert_{\openset_i} = \dsection^{i}$ for all $i\in I$ if all pairs of sections agree on overlaps $\dsection^{i}|_{\openset_i\cap\openset_j} =  \dsection^{j}|_{\openset_i\cap\openset_j}$
   \end{itemize}
 \end{definition}
+
 The gluing axiom says that a distributed representation of a dataset, which is a set of local sections, is equivalent to a section over the union of the opensets of the local sections.
@@
 \section{Trivial and non-trivial bundles}
+\label{sec:appendix:bundle_triviality}
+
+Generally, the distinguishing factor between a trivial bundle and a non-trivial bundle are how they are decomposed into local trivializations:
+\begin{LaTeXdescription}
+  \item[\textit{trivial bundle}] is directly isomorphic to $\dbasec\times\dfiberc$.
+For any choice of cover of $\dbasec$ by overlapping opensets, we can choose local trivializations such that all transition maps are identity maps.
+  \item[\textit{non-trivial bundle}] can not be constructed as $\dbasec\times\dfiberc$.
+For any choice of local trivializations, there is at least one transition map that is not an identity \cite{hatcherAlgebraicTopology2002}.
+\end{LaTeXdescription}
 
 \begin{figure}[H]
-  \includegraphics[width=1\columnwidth]{figures/}
-  \caption{\note{figure out short caption here}\label{fig:cyl_mob_bundles}}
+  \includegraphics[width=1\columnwidth]{transition_maps.png}
+  \caption{The cylinder is a trivial fiber bundle; therefore it can be decomposed into local trivalizations that only need identity
+  maps to glue the trivializations together.
+The mobius band is a non-trivial bundle; therefore it can only be decomposed into trivializations where at least one transition map is not an identity map.
+}\label{fig:cyl_mob_bundles}
 \end{figure}
-In \autoref{fig:cyl_mob_bundles} the cylinder is an example of a trivial bundle and the non mobius band is an example of a non-trivial bundle.
-We explain the difference here to show that our framework is not limited to trivial bundles, but generally  the bundle type does not matter because both types can be decomposed into trivial fiber bundles.
-For example, both the cylinder and mobius band can be decomposed into the fiber bundles shown in \autoref{fig:atct:fbdef}.
-
-Generally, the distinguishing factor between a trivial bundle and a non-trivial bun transition maps for constructing the spaces from trivializations:
-\begin{description}
-  \item[\textit{trivial bundle}] is directly isomorphic to $\dbasec\times\dfiberc$, and for any choice of cover of $\dbasec$ by overlapping opensets, we can choose local trivializations so that all transition maps are identity maps.
-  \item[\textit{non-trivial bundle}] on the other hand is one that can not be constructed as $\dbasec\times\dfiberc$, but where the construction is more complicated.
-Then for any choice of local trivializations, there is at least one transition map that is not an identity \cite{hatcherAlgebraicTopology2002}.
-\end{description}
-
-In the example in Figure~\ref{fig:cyl_mob_bundles}, we use arrows $\uparrow$ to denote fiber alignments, illustrating that by choosing simple enough trivializations, we get identity maps in the transition maps.
-In the cylinder case the fibers in the overlapping regions of the trivializations point in the same direction $\uparrow$ to illustrate that they are equal $\dfiber_0\restriction_{\openset_1\cap\openset_2} = \dfiber_{1}\restriction_{\openset_1\cap\openset_2}$.
-Because the fibers point in the same direction, the  transition maps at both intersections map the fiber to itself $\uparrow\rightarrow\uparrow$.
+
+In the example in Figure~\ref{fig:cyl_mob_bundles}, we use arrows $\uparrow$ to denote fiber alignments.
+In the cylinder case the fibers all point in the same direction $\uparrow$, which illustrates that they are equal   In the Möbius band case, while the fibers in an arbitrary local trivialization are equal, the fibers are always unequal at the twist.
+
+\begin{figure}[H]
+  \includegraphics[width=1\columnwidth]{cyl_mob.png}
+  \caption{The fiber bundles over the intervals $[\epsilon^{-1}, {\frac{2\pi}{5}}^{+}]$ and $[{\frac{2\pi}{5}}^{-}, \epsilon^{-1}]$ can be composed into larger bundles via transition maps at the intersections over $\epsilon$ and $\frac{2\pi}{5}$.
+These gluing rules also apply to any sections over the bundle, such as $\sin$.}\label{fig:cyl_mob_trivializations}
+\end{figure}
+
+The cylinder and mobius band can be decomposed to the same local trivializations, for example the fiber bundles in \autoref{fig:cyl_mob_trivializaions.} In the cylinder case, the fibers in the overlapping regions of the trivializations are equal $\dfiber_0\restriction_{\openset_1\cap\openset_2} = \dfiber_{1}\restriction_{\openset_1\cap\openset_2}$; therefore the transition maps at both intersections map the fiber to itself $\uparrow\rightarrow\uparrow$.
 In the Möbius band case, while $\dfiber_0\restriction_{(2\pi/5-\varepsilon, 2\pi/5+\varepsilon)}\to\dfiber_1\restriction_{(2\pi/5-\varepsilon, 2\pi/5+\varepsilon)}$ can be chosen to be an identity map (after transforming $\dbase_1\times[-2,2]\to\dbase_1\times[-1,1]$ by rescaling), the other transition map component $\dfiber_0\restriction_{(-\varepsilon,\varepsilon)}\to\dfiber_1\restriction_{(-\varepsilon,\varepsilon)}$ has to flip the fibers to align them between the pieces.
-
-\pagebreak
+The reason this alignment is necessary is to preserve the continuity of any sections in the bundle, such as the $\sin$ curve that exists in both the trivial bundle and mobius band.
+
 \section{Internal Library Specification}
+As mentioned in \autoref{sec:construction:vtotal}, the internal types of visualization libraries can be defined using this model, which creates a consistent standard for developers writing new functions to target.
+These are the formal specifications of various aesthetic parameters in Matplotlib.
 

### 2024-12-11 paper/figcode/nontrivial.py

@@ -121,12 +121,12 @@ def draw_segmented_bundle(ax, plot_total_space, labels):
                 (config['xB'][mask][pii], config['yB'][mask][pii], config['zb']),
                 
                                    arrowstyle=config['aws'], mutation_scale=config['ms'], 
-                                   color=config['bundle_color'])
+                                   color=tc)
             ax.add_artist(fp)
             ax.text3D(config['xB'][mask][pii], config['yB'][mask][pii], 
                       config['zb']+(bounds[0]['z'][pii] - config['zb'])/2, 
                       r"$\pi$", fontsize=config['fs'],
-                       color=config['bundle_color'], ha='left', va='center')
+                       color=tc, ha='left', va='center')
 
 
         for i in [0,-1]:
@@ -147,11 +147,10 @@ def draw_segmented_bundle(ax, plot_total_space, labels):
             ax.plot(config['xB'][mask], config['yB'][mask],  config['zb'], 
                           color=bc, mfc='none', lw=2, markevery=[0, -1], 
                           marker='o', markersize=5)
-        
 
     bundle_label = [(r"$(\;$", None ), (f"$E_{labels[1]},$", 'total'),
                     (f"$K_{labels[1]},$",'base'), 
-                    (r"$\pi$,",None), (f"$F_{labels[1]}$", 'fiber'), (")", None)] 
+                    (r"$\pi$,",'total'), (f"$F_{labels[1]}$", 'fiber'), (")", None)] 
     
     s, c = bundle_label[0]
     # this should be annotation aligned

### 2024-12-13 paper/main.tex
prescheaves and sheaves

@@
 In this section, we propose a mathematical abstraction of the data input and pre-rendered graphic output.
-This mathematical abstraction has a robust language for expressing topology and fields; expresses how to verify that data continuity is preserved on subset, distributed, and streaming data representations; and formalizes the expectation of a correspondence between data and visual elements.
+This mathematical abstraction provides a uniform highly generalizable method for describing topology and fields; expresses how to verify that data continuity is preserved on subset, distributed, and streaming data representations; and formalizes the expectation of a correspondence between data and visual elements.
 Using these abstractions allows us to embed information about structure in dataset types:
@@
 \label{sec:atct:fiber-bundles}
-
 We model data using a mathematical representation of data that can encode topological properties, field types, and data values in a uniform manner using a structure from algebraic topology called a fiber bundle.
@@
 
-\textbf{Local triviality} means $\dtotal\vert_{\openset} = \openset\times F$.
+\textbf{Local triviality} means $\dtotal\vert_{\openset} = \openset\times \dfiber$.
 In this paper we use $\dtotal\vert_{\openset} = \pi^{-1}(\openset)$ to denote the preimage of an openset\footnote{Open sets (open subsets) are a generalization of open intervals to n dimensional spaces.
@@
 \cite{weissteinOpenSet,bradleyTopologyVsTopology}}, and a \textbf{local trivialization} is a specific choice of neighborhoods (described in \autoref{sec:atct:fb:base}) and their preimages such that the fibers in each preimage are identical $\dfiber = \dfiber_{\dbasepoint}$ for all points $\dbasepoint \in \openset$.
-All fiber bundles can be decomposed into sets of local trivializaions that are also bundles and we can specify a gluing scheme that reconstructs the fiber bundle from locally trivial pieces  by specifying\textbf{transition maps} for all overlaps of the local trivializations; therefore in this paper we assume that the fiber bundles are trivial bundles $\dtotal = \dbase \times \dfiber$ but the model also applies to non-trivial bundles.
-For a discussion of trivial and non-trivial bundles, see \autoref{sec:appendix:bundle_triviality}.
+All fiber bundles can be decomposed into sets of local trivializations that are also bundles and we can specify a gluing scheme that reconstructs the fiber bundle from locally trivial pieces by specifying\textbf{transition maps} for all overlaps of the local trivializations; therefore, while the framework in this paper applies to all bundles, in this paper we assume that the bundles are trivial bundles $\dtotal = \dbase \times \dfiber$ so that we can assign all fibers in a bundle the same type.
+For an example of trivial and non-trivial bundles, see \autoref{sec:appendix:bundle_triviality}.
 
@@
        \caption{The space of all data values encoded by this fiber bundle can be modeled as a \textcolor{total}{rectangle} total space.
-Each dataset in this dataspace lies along the interval \textcolor{base}{[0,1]} base space.
-Each dataset has values along the \textcolor{fiber}{$0 \rightarrow 1$} interval fiber.
+Each dataset in this dataspace lies along the interval \textcolor{base}{$[0,2\pi]$} base space.
+Each dataset has values along the \textcolor{fiber}{$-1 \rightarrow 1$} interval fiber.
 One dataset embedded in this total space is the \textcolor{section}{sin} section over the bundle.}\label{fig:atct:fb}
@@
 
-Fiber bundles are a good abstract data representation for visualization because the field type and topological structure are unconstrained in terms of dimensionality and the only conditions that must be satisfied are that every point in the base space has a corresponding field of values and the field types must be the same for every point in the base space.
-We propose that modeling data as sections provides a way to encapsulate topological and field structure in a uniform dimension and type independent manner, as discussed in \autoref{sec:atct:fb:sections}.
-
 \subsubsection{\textcolor{base}{Topological Structure: Base Space \dbase}}
@@
 The partitions must also be composable in a continuity and property preserving way.
-There are various equivalent definitions of topology, but here we use the neighborhood axiomatization because it is most analogous the data access model of index (point) in subset (neighborhood) of all indices (mathematical space).
+
+There are various equivalent definitions of topology, but here we use the neighborhood axiomatization because it is most analogous to the data access model of index (element) in subset (neighborhood) of all indices (mathematical space).
+Given a set $X$ and a function $\mathcal{N}:X\to 2^{2^X}$ that assigns to any $x\in X$ a non-empty collection of subsets $\mathcal{N}(x)$, where each element of $\mathcal{N}(x)$ is a \emph{neighborhood of $x$}, then $X$ with  $\mathcal{N}$ is a \textcolor{base}{topological space} and $\mathcal{N}$ is a neighborhood \emph{topology} if for each $x$ in $X$: \cite{brownronaldTopologyGroupoids2006}
 
 \begin{definition}\label{def:topology}
-Given a space $X$, a point $x$ in $X$, and a function $\mathcal{N}(x)$ that assigns to $x$ subsets of $X$ , then $X$ with  $\mathcal{N}$ is a \textcolor{base}{topological space} if for each $x$ in $X$: \cite{brownronaldTopologyGroupoids2006}
 \begin{enumerate}
@@
 
-Therefore a neighborhood can grow (2) or shrink (3) so long as the point $x$ remains in the neighborhood (1) and every neighborhood contains neighborhoods of points adjacent to x (4).
+Therefore a neighborhood has to contain $x$ (1), can grow (2), or shrink (3), and every neighborhood also contains smaller neighborhoods of points adjacent to x (4).
 For example, in the indexing cube in \autoref{fig:fig:related-work:continuity:ktypes}, the brown surface and blue rectangle are both neighborhoods of the index for the measurement in Albany on June 09.
@@
 
-\subsubsection{Data Field Types: Fiber Space \dfiber}
+\subsubsection{\textcolor{fiber}{Data Field Types: Fiber Space \dfiber}}
 \label{sec:atct:fb:fiber}
@@
 For example, dates can be represented as three fields $\dfiber_{year} \times \dfiber_{month} \times \dfiber_{day}$ or a composite fiber field $\dfiber_{year} \times \dfiber_{month} \times \dfiber_{day} = \dfiber_{date}$.
-The $\otimes$ encapsulates both the sets of values associated with each fiber $\{y \in \mathbb{I}\vert 1992 \leq y \leq 2025\} \times \{m \in \mathbb{I}\vert 1 \leq m \leq 12 \} \times \{d \in \mathbb{I}\vert 1 \leq d \leq 31 \}$ and the composition function $ \otimes: \dfiber_{year} \times \dfiber_{month} \times \dfiber_{day} \rightarrow \dfiber_{date}$ would include a constraint to only return dates that have the right number of days for each month.
+The $\otimes$ encapsulates both the sets of values associated with each fiber $\{y \in \mathbb{I}\vert 1992 \leq y \leq 2025\} \times \{m \in \mathbb{I}\vert 1 \leq m \leq 12 \} \times \{d \in \mathbb{I}\vert 1 \leq d \leq 31 \}$ and the composition function $ \otimes: \dfiber_{year} \times \dfiber_{month} \times \dfiber_{day} \rightarrow \dfiber_{date}$
+could include a constraint to only return dates that have the right number of days for each month.
 The bifunctor also composes the morphisms associated with each category into a morphism on the composite category $(\dfunct_{year}, \dfunct_{month}, \dfunct_{day}) = \dfunct_{date}$.
+
 Combining fibers correctly can be verified by checking that when a fiber component $\dfiber^{c}$ is present in both $\dfiber^{a}$ and $\dfiber^{b}$, it is identical when projected out of either such that the *product* diagram commutes:
@@
 
-\subsubsection{Data: Section}
+\subsubsection{\textcolor{section}{Data: Section}}
 \label{sec:atct:fb:sections}
@@
 \subsubsection{Example: Uniform Abstract Graphic Representation}
-One of the advantages of fiber bundles is that they are general enough that we can also encode the output of a visual algorithm as a bundle.
-This allows us to use the same structure to express the properties of data and the graphic that must be symmetric to the data in an equivariant (\autoref{sec:related-work:equivariance}) transformation.
+One of the reasons we use fiber bundles as an abstraction is that they are general enough that we can also encode the output of a visual algorithm as a bundle.
 We denote the output as a graphic, but the use of bundles allows us to generalize to output on any display space, such as a screen or 3D print.
@@
 The total space \gtotalc\ is an abstraction of an ideal (infinite resolution) space into which the graphic can be rendered.
-The base space \gbasec\ is a parameterization of the display area, for example the inked bounding box in cairo \cite{CairographicsOrg}.
-The fiber space \gfiberc\ is an abstraction of the renderer fields; for example a 2 dimension screen has pixels that can be parameterized $ \gfiber=\{x,\,y\,z\,r,\,g,\,b,\,a\}$.
+The base space \gbasec\ is a parameterization of the display area, for example the inked bounding box in the cairo \cite{CairographicsOrg} rendering library.
+The fiber space \gfiberc\ is an abstraction of the renderer fields; for example a 2 dimension screen has pixels that can be parameterized $ \gfiber=\{x,\,y\,z,\,r,\,g,\,b,\,a\}$.
 
@@
 The section evaluated on a single point $\gbasepoint$ returns a single graphic record, for example one pixel in an ideal resolution space.
-In our model, the unevaluated graphic section is passed to a renderer to generate graphics.
+In our model, the unevaluated graphic section is passed from the visualization library component to the renderer to generate graphics.
 
@@
   \includegraphics[width=1\columnwidth]{fb_rho.png}
-  \caption{For a 2D display, a section $\gsection$ maps from each point $\gbasepoint$ into a fiber $\gfiber$ that encodes an RGB pre-render space with infinite resolution $\mathcal{R}^{2}$.
-Each tiny colored box is an approximation of the return value of the same section function $\gsection$ evaluated on different points $\gbasepoint \in \gbase$ in the base space.
+  \caption{The scatter marker is specified by the section $\gsection$, which maps into the fiber $\gfiber$ to retrieve the values that compose into the pixel (approximated as a square) returned by the section function evaluated at each point $\gbasepoint$.
+The section evaluated over the entire space $\gsection\vert_{S}$ returns the entire scatter mark, shown here in faded form to make it easier to see the individual pixels.
   \label{fig:atct:fb:graphic}}
@@
 In \autoref{fig:atct:fb:graphic}, the section function $\gsection$ maps into the fiber for a simplified 2D RGB infinite resolution pre-render space and returns the $\{x,y,r,g,b\}$ values of a pixel in an infinite resolution space.
-In \autoref{fig:atct:fb:graphic} these pixels are approximated as the small orange and black colored boxes.
+In \autoref{fig:atct:fb:graphic} these pixels are approximated as the small colored boxes.
 Each pixel is the output of the $\gsection(\gbasepoint)$ section that intersects the box.
-The set of all pixels returned by a section evaluated on a given visual base space $\gsection|_{\gbase}$ can yield a visual element, such as a marker, line, or piece of a glyph.
+The set of all pixels returned by a section evaluated on a given visual base space $\gsection|_{\gbase}$ can yield a visual element, such as a marker, line, or piece of a glyph and in $\autoref{fig:atct:fb:graphic}$ is a blue circle with a black edge.
 While \autoref{fig:atct:fb:graphic} illustrates a highly idealized space with no overlaps, overlaps can be managed via a fiber element $\gfiber_{z}$ for ordering.
@@
 This abstraction facilitates representational invariance, as introduced by Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}, since the container level is uniformly specified as satisfying sheaf constraints.
-These constraints generalize to data that is subset\note{one is not like the others}, distributed, streaming, and on-demand.
+These constraints describe managing subsets of data and generalize to data that is distributed, streaming, and on-demand.
 
@@
 A functor is contravariant when the morphisms between the input objects go in the opposite direction from the morphisms between the output objects.
-The presheaf is contravariant because the inclusion morphisms between input objects
-\begin{equation*}
-  \label{eq:atct:sheaf:inclusion}
-  \iota: \opensetc_1 \rightarrow \opensetc_2
-\end{equation*}
-are defined such that they correspond to the partial ordering $\openset_1 \subseteq \openset_2$, but the restriction morphisms $\iota^*$ between the sets of sections
-\begin{equation*}
-  \label{eq:atct:sheaf:restriction}
-  \iota^*: \cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}} \rightarrow \cgamma{\opensetc_1}{\dtotalc\restriction_{\openset_c1}}
-\end{equation*}
-restricts the larger set to the smaller one such that all functions that are continuous over a space must be continuous over a subspace $\Gamma_2 \subseteq \Gamma_1$, where $\Gamma_{i}\coloneqq\Gamma(\openset_{i}, \dtotal\restriction_{\openset_{i}})$.
-
-For example, lets define presheaves $\sheaf_1, \sheaf_2$. These are maps from intervals $\openset_1, \openset_2$ to a set of functions $\Gamma_1, \Gamma_2$ that are continuous over that interval:
-\note{Human readable before inline not after}
+The presheaf is contravariant because the inclusion morphisms between input object $\iota: \opensetc_1 \rightarrow \opensetc_2$
+are defined such that they correspond to the partial ordering $\openset_1 \subseteq \openset_2$, but the restriction morphisms $\iota^*$ between the sets of sections $\iota^*: \cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}} \rightarrow \cgamma{\opensetc_1}{\dtotalc\restriction_{\openset_c1}}$ restricts the larger set to the smaller one such that all functions that are continuous over a space must be continuous over a subspace $\Gamma_2 \subseteq \Gamma_1$, where $\Gamma_{i}\coloneqq\Gamma(\openset_{i}, \dtotal\restriction_{\openset_{i}})$.
+
+Data containers that implement subsetting in a structure preseving way are satisfying the  presheaf constraints that subsets of the indexing space $\openset_1$ are included in any index $\openset_2$ that is a superset $\iota$ and that data defined over an indexing space must exist over any indexes inside that space $\iota^*$.
+For example, lets define presheaves $\sheaf_1, \sheaf_2$.
+These are maps from intervals $\openset_1, \openset_2$ to a set of functions $\Gamma_1, \Gamma_2$ that are continuous over that interval:
 
@@
   \includegraphics*[width=1\columnwidth]{figures/tex/presheaf.pdf}
-  \caption*{presheafs $\sheaf_1, \sheaf_2$ mapping from intervals to the trignometric functions defined over those intervals }
+  \caption{Modeling this data container as a presheaf specifies that since $\cos, \sin, and \texttt{C}$ are continuous over $\openset_{2}$, they must be continuous over $\openset_{1}$ since $\openset_1$ is a subset of and therefore must be included in $\openset_2$.
+Because $\tan$ is only defined over $\openset_{2}$, it does not need to be included in the set $\Gamma_{2}$.
+\label{fig:atct:presheaf}}
 \end{figure}
 
-The constraints of a presheaf functor are that since the $constant, sin, cos$ functions are defined over the interval $\left[0,1\right]$, these functions must also be continuous over the sub-interval $\left(\frac{\pi}{2}, \frac{3\pi}{2}\right)$; therefore the sections in $\Gamma_{2}$ must also be included in the set of sections over the subspace $\Gamma_{1}$.
-The generalization of this constraint is that data structures that contain continuous functions must support interpolating them over arbitrarily small subspaces.
-
-\begin{figure}
-  \includegraphics[width=1\columnwidth]{sheaf_rules.png}
-  \caption{A sheaf has the conditions that sections are equal when they match on all subsets (\textit{locality}) and that the sections can be concatenated when they match on overlaps \textit{gluing}.}
-
-\end{figure}
-While presheaves preserve the rules for sets of sections, sheaves add on conditions for gluing individual sections over subspaces into cohesive sections over the whole space.
+The constraints of a presheaf functor are that since the $constant, sin, cos$ functions are defined over the interval $\left[0,2\pi\right]$, these functions must also be continuous over the sub-interval $\left(\frac{\pi}{2}, \frac{3\pi}{2}\right)$; therefore the sections in $\Gamma_{2}$ must also be included in the set of sections over the subspace $\Gamma_{1}$.
+One generalization of this constraint is that data structures that contain continuous functions must support interpolating them over arbitrarily small subspaces.
+
+While presheaves preserve the rules for sets of sections, sheaves add on conditions for gluing individual sections over subspaces into cohesive sections over the whole space.These are the conditions that when satisfied ensure that a data container is managing distributed and streaming data in a structure preserving way.
+
 \begin{definition}\label{def:atct:sheaf}\cite{bakerEuclideanSpaceMathsSheaf,spanier1989algebraic} A \textbf{sheaf} is a presheaf that satisfies the following two axioms
 \begin{itemize}
-  \item \textit{locality} two sections in a sheaf are equal $\dsection^{a} = \dsection^{b}$ when they evaluate to the same values $\dsection^{a}|_{\openset_i} =  \dsection^{b}|_{\openset_i}$ over the open cover $\bigcup_{i\in I} \openset_{i} \subset \openset$ (indexed by $I$).
-  \item \textit{gluing} the union of sections defined over subspaces $\desection^{i} \in \Gamma(\openset_i, \dtotal|_{openset_i})$ is equivalent to a section defined over the whole space $\dsection\vert_{\openset_i} = \dsection^{i}$ for all $i\in I$ if all pairs of sections agree on overlaps $\dsection^{i}|_{\openset_i\cap\openset_j} =  \dsection^{j}|_{\openset_i\cap\openset_j}$
+  \item \textit{locality} two sections in a sheaf are equal $\dsection^{a} = \dsection^{b}$ when they evaluate to the same values $\dsection^{a}\vert_{\openset_i} =  \dsection^{b}\vert_{\openset_i}$ over the open cover $\bigcup_{i\in I} \openset_{i} \subset \openset$ (indexed by $I$).
+  \item \textit{gluing} the union of sections defined over subspaces $\dsection^{i} \in \Gamma(\openset_i, \dtotal|_{openset_i})$ is equivalent to a section defined over the whole space $\dsection\vert_{\openset_i} = \dsection^{i}$ for all $i\in I$ if all pairs of sections agree on overlaps $\dsection^{i}\vert_{\openset_i\cap\openset_j} =  \dsection^{j}\vert_{\openset_i\cap\openset_j}$
   \end{itemize}
@@
 The gluing axiom says that a distributed representation of a dataset, which is a set of local sections, is equivalent to a section over the union of the opensets of the local sections.
+The gluing axiom can also be used to generate the gluing rules used to construct non-trivial bundles from the set of trivial local sections.
 The locality axiom asserts that the glued section function is equivalent to a function over the union if they evaluate to the same values.
-The gluing axiom can also be used to generate the gluing rules used to construct non-trivial bundles from the set of trivial local sections.
-Generally, the sheaf asserts the expectation that the data container is implemented such that the connectivity between the opensets (indexing subspaces) is preserved.
+
+\begin{figure}[h]
+  \includegraphics[width=1\columnwidth]{tex/sheaf_rules.png}
+  \caption{A sheaf has the conditions that sections are equal when they match on all subsets (\textit{locality}) and that the sections can be concatenated when they match on overlaps \textit{gluing}.\label{fig:atct:sheaf}}
+\end{figure}
+
+For example, in \autoref{fig:atct:sheaf}, the $\tau^{a}$ and $\tau^{b}$ $\sin$ sections are equal because they match \textit{locally} on each connected subset.
+This is true whether $\sin$ is defined over parts ($\sin\vert_{\openset_i}$) or the whole space.
+If $\sin$ is defined over parts, then those parts can be \textit{glued} together.
+The concatenated $\sin$ is continuous because the pieces of the section outside the overlap are continuous with the pieces inside the overlap.
+The glued $\sin$ is also equal to the non-glued $\sin$ because they match on the opensets; therefore they are equivalent representations of the same section $\sin$ and so have the same mathematical properties.
 
@@
 The \textit{germ} is the section evaluated at a point in the stalk  $\dsection(\dbasepoint) \in \sheaf_{\dbase, \dtotal}\restriction_{\dbasepoint}$ and is the data.
-Since the stalk and the germ include the values near the limit of the point at \dbasepoint, the germ can be used to compute the mathematical derivative of the data for visualization tasks that require this information.
+Since the stalk includes the values near the limit of the point at \dbasepoint, the germ can be used to compute the mathematical derivative of the data for visualization tasks that require this information.
 
@@
 In the example in Figure~\ref{fig:cyl_mob_bundles}, we use arrows $\uparrow$ to denote fiber alignments.
-In the cylinder case the fibers all point in the same direction $\uparrow$, which illustrates that they are equal   In the Möbius band case, while the fibers in an arbitrary local trivialization are equal, the fibers are always unequal at the twist.
-
-\begin{figure}[H]
-  \includegraphics[width=1\columnwidth]{cyl_mob.png}
-  \caption{The fiber bundles over the intervals $[\epsilon^{-1}, {\frac{2\pi}{5}}^{+}]$ and $[{\frac{2\pi}{5}}^{-}, \epsilon^{-1}]$ can be composed into larger bundles via transition maps at the intersections over $\epsilon$ and $\frac{2\pi}{5}$.
-These gluing rules also apply to any sections over the bundle, such as $\sin$.}\label{fig:cyl_mob_trivializations}
-\end{figure}
-
-The cylinder and mobius band can be decomposed to the same local trivializations, for example the fiber bundles in \autoref{fig:cyl_mob_trivializaions.} In the cylinder case, the fibers in the overlapping regions of the trivializations are equal $\dfiber_0\restriction_{\openset_1\cap\openset_2} = \dfiber_{1}\restriction_{\openset_1\cap\openset_2}$; therefore the transition maps at both intersections map the fiber to itself $\uparrow\rightarrow\uparrow$.
-In the Möbius band case, while $\dfiber_0\restriction_{(2\pi/5-\varepsilon, 2\pi/5+\varepsilon)}\to\dfiber_1\restriction_{(2\pi/5-\varepsilon, 2\pi/5+\varepsilon)}$ can be chosen to be an identity map (after transforming $\dbase_1\times[-2,2]\to\dbase_1\times[-1,1]$ by rescaling), the other transition map component $\dfiber_0\restriction_{(-\varepsilon,\varepsilon)}\to\dfiber_1\restriction_{(-\varepsilon,\varepsilon)}$ has to flip the fibers to align them between the pieces.
-The reason this alignment is necessary is to preserve the continuity of any sections in the bundle, such as the $\sin$ curve that exists in both the trivial bundle and mobius band.
+In the cylinder case the fibers all point in the same direction, which illustrates that they are equal $\uparrow=\uparrow$.
+In the Möbius band case, while the fibers in an arbitrary local trivialization are equal $\uparrow=\uparrow$, the fibers at the twist are unequal but isomorphic $\uparrow \cong \downarrow$.
+The cylinder and mobius band can be decomposed to the same local trivializations, for example the fiber bundles in \autoref{fig:atct:fb} In the cylinder case, the fibers in the overlapping regions of the trivializations are equal $\dfiber_0\restriction_{\openset_1\cap\openset_2} = \dfiber_{1}\restriction_{\openset_1\cap\openset_2}$; therefore the transition maps at both intersections map the values in the fiber to themselves $\delement\rightarrow\delement$ .
+In the Möbius band case, while $\dfiber_0\restriction_{(2\pi/5-\varepsilon, 2\pi/5+\varepsilon)}\to\dfiber_1\restriction_{(2\pi/5-\varepsilon, 2\pi/5+\varepsilon)}$ can be chosen to be an identity map, the other transition map component $\dfiber_0\restriction_{(-\varepsilon,\varepsilon)}\to\dfiber_1\restriction_{(-\varepsilon,\varepsilon)}$ has to flip any section values.
+For example given  $\dfiber_{0}=\uparrow$ and $\dfiber_{1}=\downarrow$, the transition map $\delement \mapsto -\delement$ maps each point from one fiber to the other $\uparrow \mapsto \downarrow$ such that any sections remain continuous even though the fibers point in opposite directions.
 

### 2024-12-13 paper/figcode/nontrivial.py

@@ -23,7 +23,7 @@ config.update(t = np.linspace(0, 2*np.pi, config['N']))
 
 config.update(r1 = np.array([config['t'][0], config['boundary']])-config['boundary_pad']/2, 
               r2 = np.array([config['boundary'], config['t'][-1]])+config['boundary_pad']/2, #trivialization E x
-              ry = .35, rh =.5 , #trivilization E y
+              ry = .35, rh =.4 , #trivilization E y
               lw = 2, ly = .1 #trivialization K lw and y
              )
 

### 2024-12-19 paper/main.tex
new xi diagram

@@
 \usepackage{amsmath, amsthm, amsfonts,amssymb,eulervm,xspace, mathtools}
-\usepackage{stmaryrd}%mapsfrom
+\usepackage{stmaryrd}%maps from
 %\renewcommand{\restriction}{\mathord{\upharpoonright}} %restriction w/p space
@@
 
-\IEEEPARstart{V}isualization design guidelines, generally, describe how to choose visual encodings that preserve the structure of the data; to follow these guidelines the visualization tools that implement these $\data \rightarrow graphic$ transforms must be structure preserving.
+\IEEEPARstart{V}isualization design guidelines, generally, describe how to choose visual encodings that preserve the structure of the data; to follow these guidelines the visualization tools that implement these $data \rightarrow graphic$ transforms must be structure preserving.
 Loosely, preserving structure means that the properties of the data and how the points are connected to each other should be inferable from the graphic such that a $graphic \rightarrow data$ mapping can be made.
@@
 
-We propose that we can better enforce this expectation in software by providing a uniform way of expressing $data$ and $graphic$ using their respective algebraic structure and by uniformally specifying the behaviors and properties of those structures and the maps between them using category theory.
+We propose that we can better enforce this expectation in software by providing a uniform way of expressing $data$ and $graphic$ using their respective algebraic structure and by uniformly specifying the behaviors and properties of those structures and the maps between them using category theory.
 For example, our framework can encapsulate how a table and scatter plot and heatmap are different representations of the same data and track an observation from a data cube as a point along a time series and on a map and in a network.
@@
 Specifically, Steven's conceptualizes the structure on values as \textcolor{action}{actions} on groups \footnote{A \textit{group} is a set with an associative binary operator.
-This operation must have an identity element and be closed, associative, and invertable}.
+This operation must have an identity element and be closed, associative, and invertible}.
 A function that preserves structure when the input or output is changed by a group action is called \textit{equivariant}.
@@
 
-Although the Steven's scales were conceptualized as having group structure, the ordinal scale has a monoidal structure because partial orders ($\geq, \leq$) are not invertable.
+Although the Steven's scales were conceptualized as having group structure, the ordinal scale has a monoidal structure because partial orders ($\geq, \leq$) are not invertible.
 This means \textit{equivariance} cannot be used to test for structure preservation.
-Instead \textit{homomorphsim} can be used because it imposes fewer constraints on the underlying mathematical structure of the data.
+Instead \textit{homomorphism} can be used because it imposes fewer constraints on the underlying mathematical structure of the data.
 \begin{figure}
@@
 \begin{definition}\label{def:homomorphism}
-  A function $f$ is \textbf{homomorphic} when $f(x_1 \circ x_2) = f(x_1) * f(x_2)$ and preserves identites $f(I_x) = I_y$ all $x, y \in X$ \cite{grimaldiDiscreteCombinatorialMathematics2006}
+  A function $f$ is \textbf{homomorphic} when $f(x_1 \circ x_2) = f(x_1) * f(x_2)$ and preserves identities $f(I_x) = I_y$ all $x, y \in X$ \cite{grimaldiDiscreteCombinatorialMathematics2006}
 \end{definition}
@@
 
-The bar plot, line plot, and heatmap in \autoref{fig:related-work:continuity:ktypes} have a homemeomorphic relationship to the 0D ($\bullet$) points, 1D (--) linear, and 2D ($\blacksquare$) surface continuities embedded in the continuous 3 dimensional surface encapsulating time and position because each point of the visualization maps back into a point in its corresponding indexing space in the cube.
+The bar plot, line plot, and heatmap in \autoref{fig:related-work:continuity:ktypes} have a homeomorphic relationship to the 0D ($\bullet$) points, 1D (--) linear, and 2D ($\blacksquare$) surface continuities embedded in the continuous 3 dimensional surface encapsulating time and position because each point of the visualization maps back into a point in its corresponding indexing space in the cube.
 Using homeomorphism to test whether continuity is preserved formalizes Bertin's codification of how the topology of observations matches the class of representation (i.e.
@@
 \begin{definition}\label{def:fiber_bundle:section}
-  \\
-A \textcolor{section}{\textbf{section}} $\dsectionc: \dbasec \rightarrow \dtotalc$ over a fiber bundle is a smooth right inverse of $\pi$: $\pi(\dsection(\dbasepoint)) = \dbasepoint$ for all $\dbasepoint \in \dbase$
+A \textcolor{section}{\textbf{section}} $\dsectionc: \dbasec \rightarrow \dtotalc$ over a fiber bundle is a smooth right inverse of $\pi(\dsection(\dbasepoint)) = \dbasepoint$ for all $\dbasepoint \in \dbase$
 \end{definition}
@@
        \caption{The space of all data values encoded by this fiber bundle can be modeled as a \textcolor{total}{rectangle} total space.
-Each dataset in this dataspace lies along the interval \textcolor{base}{$[0,2\pi]$} base space.
+Each dataset in this data space lies along the interval \textcolor{base}{$[0,2\pi]$} base space.
 Each dataset has values along the \textcolor{fiber}{$-1 \rightarrow 1$} interval fiber.
@@
 Therefore the type signature of the datasets in this fiber bundle, which is called a section \dsection, would be \texttt{dataset: $[0, 2\pi] \rightarrow [0,1]$}.
-One such dataset (section) is the $\sin$ function, which as shown in \autoref{fig:atct:fb} is defined via a function \desection from a point in the base space to a corresponding point in the fiber.
+One such dataset (section) is the $\sin$ function, which as shown in \autoref{fig:atct:fb} is defined via a function \dsection from a point in the base space to a corresponding point in the fiber.
 Evaluating the section function over the entire base space yields the $\sin$ curve that is composed of points intersecting each fiber over the corresponding point.
@@
 \paragraph{Joining indexing spaces: $\oplus: \mathcal{\dbasec} \sqcup \mathcal{\dbasec} \rightarrow \mathcal{\dbasec}$}
-For example, the disjoint union of two bundles, as shown in \autoref{fig:atct:local_trivalizations}, is the coproduct $\dbase^{a} \subseteq_{\dbase^{c}} \dbase^{b}$ over an overlap $\dbase^{c}$ and therefore the inclusion morphism must be commutative:
+For example, the disjoint union of two bundles, as shown in \autoref{fig:atct:local_trivalizations}, is the coproduct $\dbase_{a} \sqcup_{\dbase_{c}} \dbase_{b}$ over an overlap $\dbase_{c}$ and therefore the inclusion morphism must be commutative:
 \begin{figure}[h]
-  \includegraphics*[width=1\columnwidth]{figures/tex/k_coproduct.pdf}
+  \includegraphics*[width=1\columnwidth]{figures/tex/k_coproduct.pdf}  \caption{Index spaces are combined via the \textit{coproduct} $\dbase_{a}\sqcup_{\dbase_c} \dbase_{b}$.}
+  \label{fig:atct:base}
 \end{figure}
 
-The coproduct in \autoref{eq:atct:base_union} shows that the way to test that two spaces have been combined correctly is to verify that every point in the subspace $\dbasepoint \in \dbase^{c}$ must be present in the spaces for which inclusion morphisms exist $\dbasepoint \in \dbase^{a}$ and $\dbasepoint \in \dbase^{b}$ such that $\dbasepoint \in \dbase^{a}\sqcup_{\dbase^{c}} \dbase_{b}$.
+This means that index spaces are combined on overlapping indices.
+For example, in \autoref{fig:atct:base}, the open interval $\dbase_c$ is present in both $\dbase_a$ and $\dbase_c$; therefore $\dbase_a$ and $\dbase_b$ are glued together where they overlap at $\dbase_c$.
+Two spaces have been combined correctly if every point in the overlap  $\dbasepoint \in \dbase^{c}$ is present in each of the spaces $\dbasepoint \in \dbase^{a}$ and $\dbasepoint \in \dbase^{b}$ such that it is present in the joined space $\dbasepoint \in \dbase^{a}\sqcup_{\dbase^{c}} \dbase_{b}$.
 This simple test that the records are joined correctly is what allows us to reliably build larger datasets out of smaller ones, such as in the case of distributed and on demand datasets.
@@
 The fiber category $\mathcal{\dfiberc}$ is also equipped with a bifunctor because it is a monoidal category and this functor provides a method for combining fiber types.
-The bifunctorallows $\otimes$ us to express fields that contain complexly typed values.
+The bifunctor allows $\otimes$ us to express fields that contain complexly typed values.
 For example, dates can be represented as three fields $\dfiber_{year} \times \dfiber_{month} \times \dfiber_{day}$ or a composite fiber field $\dfiber_{year} \times \dfiber_{month} \times \dfiber_{day} = \dfiber_{date}$.
@@
 
-Combining fibers correctly can be verified by checking that when a fiber component $\dfiber^{c}$ is present in both $\dfiber^{a}$ and $\dfiber^{b}$, it is identical when projected out of either such that the *product* diagram commutes:
-
-\begin{figure}
+Combining fibers correctly can be verified by checking that when a fiber component $\dfiber^{c}$ is present in both $\dfiber^{a}$ and $\dfiber^{b}$, it is identical when projected out of either such that the \textit{product} diagram commutes:
+
+\begin{figure}[H]
   \includegraphics*[width=1\columnwidth]{figures/tex/f_product.pdf}
+  \caption{Fields are combined via the \textit{product} $\dfiber_{a}\times_{\dfiber_c} \dfiber_{b}$.}
+  \label{fig:atct:fiber}
 \end{figure}
 
-This means that data with many fields is decomposed into its component fields, records maintain their integrity.
-For example, the red (temperature, time, pressure) record separates into (temperature, time) and (pressure, time) records that share the same red time.
+This means that when data with many fields is decomposed into its component fields, records maintain their integrity.
+For example in \autoref{fig:atct:fiber} the red (temperature, time, pressure) record separates into (temperature, time) and (pressure, time) records that share the same red time.
 Furthermore this time is the same whether it is obtained from the (temperature, time) or (pressure, time) record.
-This simple test that fields are joined together correctly for the same record is what allows us to reliably combine multiple datasets together on shared properties-for example growing the weather station data from a temporal to spatial dataset by adding the weather at each location at each time.
+Generally, the test for joining fields is that when a record is present in a fiber $\delement_{c} \in \pi_c(\dfiber_{a}), \delement_c \in \pi_c(\dfiber_{b})$ then it is in the joint record $\delement_{c} \in \pi_c(\dfiber_{a}\times_{\dfiber_c}\dfiber_{b})$ when is in the field $\dfiber_c$ being joined on.
+This simple test that fields are joined together correctly for the same record is what allows us to reliably combine multiple datasets together on shared properties, for example growing the weather station data in \autoref{fig:related-work:continuity:ktypes} from a temporal to spatial dataset by adding the weather at each location at each time.
 
@@
 The fiber space \gfiberc\ is an abstraction of the renderer fields; for example a 2 dimension screen has pixels that can be parameterized $ \gfiber=\{x,\,y\,z,\,r,\,g,\,b,\,a\}$.
-
 As with data, we model the graphic generating functions as sections $\gsection$ of the graphic bundle
-\begin{equation}
-  \label{eq:atct:fb_graphic_section}
-  \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \coloneqq \big\{\gsectionc: \opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} \; \bigm{\vert} \pi(\gsectionc(\gbasepointc)) = \gbasepointc\;for\, all\; \gbasepointc \in \opensetgc \big\}
+\begin{equation}\label{eq:atct:fb_graphic_section}
+  \begin{split}
+  \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} & \coloneqq \\
+  &\big\{\gsectionc: \opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} \; \bigm{\vert} \pi(\gsectionc(\gbasepointc)) = \gbasepointc\;for\, all\; \gbasepointc \in \opensetgc \big\}
+  \end{split}
 \end{equation}
@@
 Sheaves are an algebraic data structure that provides a way of abstractly discussing the bookkeeping that data containers must implement to keep track of the continuity of the data \cite{ghristElementaryAppliedTopology2014}.
-This abstraction facilitates representational invariance, as introduced by Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}, since the container level is uniformly specified as satisfying sheaf constraints.
-These constraints describe managing subsets of data and generalize to data that is distributed, streaming, and on-demand.
+This abstraction facilitates representational invariance, as introduced by Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}, since the container level is uniformly specified as satisfying presheaf and sheaf constraints.
+When a data container satisfies these constraints, the subsets and whole data space have the same mathematical properties, e.g.
+morphisms, such that the framework in this paper applies whether the data is in memory, distributed, streaming, or on-demand.
 
@@
 
-Data containers that implement subsetting in a structure preseving way are satisfying the  presheaf constraints that subsets of the indexing space $\openset_1$ are included in any index $\openset_2$ that is a superset $\iota$ and that data defined over an indexing space must exist over any indexes inside that space $\iota^*$.
+Data containers that implement managing subsets in a structure preserving way are satisfying the  presheaf constraints that subsets of the indexing space $\openset_1$ are included in any index $\openset_2$ that is a superset $\iota$ and that data defined over an indexing space must exist over any indexes inside that space $\iota^*$.
 For example, lets define presheaves $\sheaf_1, \sheaf_2$.
@@
 
-The constraints of a presheaf functor are that since the $constant, sin, cos$ functions are defined over the interval $\left[0,2\pi\right]$, these functions must also be continuous over the sub-interval $\left(\frac{\pi}{2}, \frac{3\pi}{2}\right)$; therefore the sections in $\Gamma_{2}$ must also be included in the set of sections over the subspace $\Gamma_{1}$.
+For example in \autoref{fig:atct:presheaf}, since the $constant, sin, cos$ functions are defined over the interval $\left[0,2\pi\right]$, these functions must also be continuous over the sub-interval $\left(\frac{\pi}{2}, \frac{3\pi}{2}\right)$; therefore the sections in $\Gamma_{2}$ must also be included in the set of sections over the subspace $\Gamma_{1}$.
 One generalization of this constraint is that data structures that contain continuous functions must support interpolating them over arbitrarily small subspaces.
@@
 
-For example, in \autoref{fig:atct:sheaf}, the $\tau^{a}$ and $\tau^{b}$ $\sin$ sections are equal because they match \textit{locally} on each connected subset.
+For example, in \autoref{fig:atct:sheaf}, the $\tau^{a}$ and $\tau^{b}$ $\sin$ sections are equal because they match \textit{locally} on all subsets.
 This is true whether $\sin$ is defined over parts ($\sin\vert_{\openset_i}$) or the whole space.
@@
 
-Each section of a sheaf over a point returns a single record in the fiber. The sheaf over an open set $\openset$ surrounding a point $\dbasepoint$ is called a \textit{stalk}\cite{harder2008lectures}
+While each section of a sheaf is evaluated over a point $\dsection(\dbasepoint)$ such that it returns a single record, the sheaf model also provides an abstraction when neighboring information is required.
+The sheaf over a very small region surrounding a point $\dbasepoint$ is called a \textit{stalk}\cite{harder2008lectures}
 \begin{equation}
@@
 \begin{figure}[H]
-  \includegraphics*[width=1\columnwidth]{xi_sin.png}
+  \includegraphics*[width=1\columnwidth]{xi_diagram.png}
   \caption{The data consists of the $sin$ and $cos$ functions over a unit circle base space.
@@
 \subsection{Homeomorphism}
-As mentioned in \autoref{sec:related-work:continuity}, preserving the topology of a visualization means that each discrete piece of differentiable visual information corresponds to a distince element of the dataset\cite{ziemkiewiczEmbeddingInformationVisualization2009} in a way where the organization of elements is preserved.
-A generalization of this condition is the idea that the graphic space can be collapased into the data indexing space, which means that the data base space is a deformation retraction of the graphic base space\cite{hatcherAlgebraicTopology2002}.
+As mentioned in \autoref{sec:related-work:continuity}, preserving the topology of a visualization means that each discrete piece of differentiable visual information corresponds to a distance element of the dataset\cite{ziemkiewiczEmbeddingInformationVisualization2009} in a way where the organization of elements is preserved.
+A generalization of this condition is the idea that the graphic space can be collapsed into the data indexing space, which means that the data base space is a deformation retraction of the graphic base space\cite{hatcherAlgebraicTopology2002}.
 By defining the indexing look up function \vindex, introduced in \autoref{sec:atct:sheaves}, to be
@@
 
-we can assert that the data space \dbase acts as an indexing space into \gbase such that knowing the location on space yields the location on the other and any point in either base space or graphic space has a correspoinding point in the other space.
+we can assert that the data space \dbase acts as an indexing space into \gbase such that knowing the location on space yields the location on the other and any point in either base space or graphic space has a corresponding point in the other space.
 \begin{figure}[H]
@@
   \centering
-  {\renewcommand{\arraystretch}{1.5}
+  {\renewcommand{\arraystretch}{1.2}
   \begin{tabular}{|r | c c c|}
@@
   \centering
-  {\renewcommand{\arraystretch}{1.5}
+  {\renewcommand{\arraystretch}{1.2}
   \begin{tabular}{|r | l l | }
@@
 \end{table}
-
 \begin{table}[H]
   \centering
-  {\renewcommand{\arraystretch}{1.5}
+  {\renewcommand{\arraystretch}{1.2}
+  \begin{tabular}{|r|l|}
+    \hline
+      axiom & applied to datasets and indexes \\
+    \hline
+     presheaf & given $\textcolor{base}{index_1} \subset \textcolor{base}{index_2}$:\\
+        & $\exists\;\textcolor{section}{dataset}[\textcolor{base}{index_2}]$\\
+      & $\Rightarrow \textcolor{section}{dataset}[\textcolor{base}{index_2}][\textcolor{base}{index_1}] = \textcolor{section}{dataset}[\textcolor{base}{index_1}]$\\
+     &  $\exists\;\textcolor{section}{dataset}[\textcolor{base}{index_1}] \nRightarrow \exists\;\textcolor{section}{dataset}[\textcolor{base}{index_2}]$\\
+     &\\
+    \hline
+    locality & $\textcolor{section}{dataset^1}[\textcolor{base}{i}] = \textcolor{section}{dataset^2}[\textcolor{base}{i}]\;\forall\; \textcolor{base}{i}\in \textcolor{base}{index}\;$ \\
+     & $\Rightarrow \textcolor{section}{dataset^1}=\textcolor{section}{dataset^2}$ \\
+    \hline
+    gluing & $\textcolor{base}{i}=\textcolor{base}{j}$ and $\textcolor{section}{dataset^1}[\textcolor{base}{i}] = \textcolor{section}{dataset^2}[\textcolor{base}{j}]$ \\
+     & $\textcolor{section}{dataset^3}\coloneq  \textcolor{section}{dataset^1}[\textcolor{base}{:i}]  \oplus \textcolor{section}{dataset^2}[\textcolor{base}{j:}]$ \\
+\hline
+  \end{tabular}
+  \caption{Presheaf and sheaf constraints implemented by structure preserving data containers, discussed in \autoref{sec:atct:sheaves}}
+  \label{tab:appendix:summary:sheaves}
+  }
+\end{table}
+
+\begin{table}[H]
+  \centering
+  {\renewcommand{\arraystretch}{1.2}
   \begin{tabular}{|r | l  l |}
@@
 \centering
-{\renewcommand{\arraystretch}{1.5}
+{\renewcommand{\arraystretch}{1.2}
 \begin{tabular}{|r|l|l|}
@@
 \begin{table}[H]
-  \renewcommand{\arraystretch}{1.5}
+  \renewcommand{\arraystretch}{1.2}
   \begin{tabular}{|lll|}\hline
@@
 Juan Nunez-Iglesias , and Nicolas Kruchten have provided valuable feedback from the
-library developors perspective.
+library developers perspective.
 

### 2024-12-23 paper/main.tex
reworked rho figure + clean ups to text

@@
 \begin{figure}
-  \includegraphics*[width=1\columnwidth]{tex/equivariant.png}
+  \includegraphics*[width=1\columnwidth]{equivariant.pdf}
   \caption{Encoding data as the bar height using an exponential transform is not equivariant because encoding the data and then scaling the bar heights yields a much taller graph then scaling the data and then encoding those heights using the same exponential transform function.}
@@
 \begin{figure}
-  \includegraphics[width=1\columnwidth]{tex/homomorphism.png}
+  \includegraphics[width=1\columnwidth]{homomorphism.pdf}
   \caption{Encoding data as bar height using an inverse transform is not homomorphic because the largest number is mapped to the smallest bar while the max function returns the largest bar.}
@@
 \begin{figure}[H]
-  \includegraphics[width=1\columnwidth]{k_different_types.png}
+  \includegraphics[width=1\columnwidth]{k_different_types.pdf}
   \caption{This weather station data has multiple embedded continuities - points at each time and position, timeseries at each position, and maps at each time.
@@
 
-Domain specific libraries are designed with the assumption of continuities that are common in the domain \cite{HeerSoftware2006}, and therefore can somewhat restrict their API to choices that are appropriate for the domain.
+Domain specific libraries are designed with the assumption of continuities that are common in the domain \cite{HeerSoftware2006}, and therefore can somewhat restrict their API to choices that are appropriate for that domain.
 For example, a tabular topological structure of discrete rows, as illustrated in \autoref{fig:related-work:continuity:ktypes}, is assumed by A Presentation Tool\cite{mackinlayAutomatingDesignGraphical1986, mackinlayAutomatingDesignGraphical1986} and grammar of graphics\cite{wilkinsonGrammarGraphics2005} and the ggplot\cite{wickhamGgplot2ElegantGraphics2016}, vega\cite{satyanarayanDeclarativeInteractionDesign2014}, and altair\cite{vanderplasAltairInteractiveStatistical2018} libraries built on these frameworks.
-Image libraries such as Napari\cite{nicholas_sofroniew_2021_4533308} and ImageJ\cite{schneiderNIHImageImageJ2012} and its humanities ImagePlot\cite{studiesCulturevisImageplot2021} plugin assume that the input is 2D continuous.
+Image libraries such as Napari\cite{nicholas_sofroniew_2021_4533308} and ImageJ\cite{schneiderNIHImageImageJ2012} and the ImagePlot\cite{studiesCulturevisImageplot2021} art plugin assume that the input is a 2D continuous image.
 Networking libraries such as gephi\cite{bastianGephiOpenSource2009} and networkx\cite{HagbergExploringNetwork2008} assume a graph-like structure.
-By assuming the structure of their data, these domain specific libraries can provide more cohesive interfaces for a much more limited set of visualization algorithms than the building block libraries offer.
+By assuming the structure of their data, these domain specific libraries can provide more cohesive interfaces because a limited set of visualization algorithms apply to their data and the visualizations these algorithms provide can be styled in a fixed number of ways.
 
@@
 \begin{figure}[!h]
-       \includegraphics[width=1\columnwidth]{fb.png}
+       \includegraphics[width=1\columnwidth]{fb.pdf}
        \caption{The space of all data values encoded by this fiber bundle can be modeled as a \textcolor{total}{rectangle} total space.
@@
 \paragraph{Joining indexing spaces: $\oplus: \mathcal{\dbasec} \sqcup \mathcal{\dbasec} \rightarrow \mathcal{\dbasec}$}
-For example, the disjoint union of two bundles, as shown in \autoref{fig:atct:local_trivalizations}, is the coproduct $\dbase_{a} \sqcup_{\dbase_{c}} \dbase_{b}$ over an overlap $\dbase_{c}$ and therefore the inclusion morphism must be commutative:
+Joining two bundles on their base space, for example to glue the two trivializations in \autoref{fig:atct:fb} into a bundle, is the coproduct $\dbase_{a} \sqcup_{\dbase_{c}} \dbase_{b}$ over a space $\dbase_{c}$.
+A property of the coproduct is that the inclusion morphism must be commutative, which means that index spaces are combined on overlapping indices:
 \begin{figure}[h]
@@
 \end{figure}
-
-This means that index spaces are combined on overlapping indices.
  For example, in \autoref{fig:atct:base}, the open interval $\dbase_c$ is present in both $\dbase_a$ and $\dbase_c$; therefore $\dbase_a$ and $\dbase_b$ are glued together where they overlap at $\dbase_c$.
@@
 
-Combining fibers correctly can be verified by checking that when a fiber component $\dfiber^{c}$ is present in both $\dfiber^{a}$ and $\dfiber^{b}$, it is identical when projected out of either such that the \textit{product} diagram commutes:
+Combining fibers correctly can be verified by checking that when a fiber component $\dfiber^{c}$ is present in both $\dfiber^{a}$ and $\dfiber^{b}$, it is identical when projected out of either such that the \textit{product} diagram commutes, means that when data with many fields is decomposed into its component fields, records maintain their integrity:
 
@@
 
-This means that when data with many fields is decomposed into its component fields, records maintain their integrity.
 For example in \autoref{fig:atct:fiber} the red (temperature, time, pressure) record separates into (temperature, time) and (pressure, time) records that share the same red time.
@@
 
-\subsubsection{\textcolor{section}{Data: Section}}
+\subsubsection{\textcolor{section}{Data: Section $\dsection$}}
 \label{sec:atct:fb:sections}
@@
 \begin{figure}[H]
-  \includegraphics[width=1\columnwidth]{fb_rho.png}
+  \includegraphics[width=1\columnwidth]{fb_rho.pdf}
   \caption{The scatter marker is specified by the section $\gsection$, which maps into the fiber $\gfiber$ to retrieve the values that compose into the pixel (approximated as a square) returned by the section function evaluated at each point $\gbasepoint$.
@@
 \begin{figure}[h]
-  \includegraphics[width=1\columnwidth]{tex/sheaf_rules.png}
+  \includegraphics[width=1\columnwidth]{tex/sheaf_rules.pdf}
   \caption{A sheaf has the conditions that sections are equal when they match on all subsets (\textit{locality}) and that the sections can be concatenated when they match on overlaps \textit{gluing}.\label{fig:atct:sheaf}}
@@
 \subsubsection{Example: Graphic and Data}
-\begin{figure}[H]
-  \includegraphics*[width=1\columnwidth]{xi_diagram.png}
+\begin{figure}[t]
+  \includegraphics[width=1\textwidth]{xi_diagram.pdf}
   \caption{The data consists of the $sin$ and $cos$ functions over a unit circle base space.
@@
 \end{table}
+
+\begin{table}[H]
+  {\renewcommand{\arraystretch}{1.2}
+  \begin{tabular}{|r|l|}
+    \hline
+    merge & equation & behavior\\
+    \hline
+    indices  & $\dbasec_{a} \sqcup_{\dbasec_c} \dbasec_b$ & ${\dbasepoint \in \dbase_a, \dbase_b$ | $\dbasepoint_c \in \dbase_a, \dbase_b$ for all $dbase_ \in \dbasec$ \\
+    \hline
+    fields  & $\dfiberc_{a} \times_{\dfiberc_c} \dbase_c$
+    \hline
+
+  \caption{}
+  \label{tab:appendix:summary:merge}
+  }
+\end{table}
+
 \begin{table}[H]
@@
 \begin{figure}[H]
-  \includegraphics[width=1\columnwidth]{transition_maps.png}
+  \includegraphics[width=1\columnwidth]{transition_maps.pdf}
   \caption{The cylinder is a trivial fiber bundle; therefore it can be decomposed into local trivalizations that only need identity
@@
 \note{Acknowledge all the actual people}
-The authors would like to thank the anonymous reviewers who gave constructive feedback on
-an earlier version of this paper. Various Matplotlib and Napari contributors, particularly
-Juan Nunez-Iglesias , and Nicolas Kruchten have provided valuable feedback from the
-library developers perspective.
-
-Hannah is also very grateful to Nicolas for the suggestion of augmented notation and to
-the nlab and wikipedia contributors who wrote clear explanations of many of the topics
-discussed in this paper.
+The authors would like to thank the anonymous reviewers who gave constructive feedback on an earlier version of this paper.
+The authors are also grateful to  the various Matplotlib and Napari contributors, particularly Juan Nunez-Iglesias, and Nicolas Kruchten for their valuable feedback from the library developer perspective.
+
+Hannah is also very grateful to Nicolas for the suggestion of augmented notation and to the nlab and wikipedia contributors who wrote clear explanations of many of the topics discussed in this paper.
 

### 2024-12-23 paper/main.tex
finished out section 3

@@
 The fiber space \gfiberc\ is an abstraction of the renderer fields; for example a 2 dimension screen has pixels that can be parameterized $ \gfiber=\{x,\,y\,z,\,r,\,g,\,b,\,a\}$.
-As with data, we model the graphic generating functions as sections $\gsection$ of the graphic bundle
+As with data, we model the graphic specifications as sections $\gsection$ of the graphic bundle
 \begin{equation}\label{eq:atct:fb_graphic_section}
@@
 The \textit{germ} is the section evaluated at a point in the stalk  $\dsection(\dbasepoint) \in \sheaf_{\dbase, \dtotal}\restriction_{\dbasepoint}$ and is the data.
-Since the stalk includes the values near the limit of the point at \dbasepoint, the germ can be used to compute the mathematical derivative of the data for visualization tasks that require this information.
+Since the stalk includes the values near the limit of the point at \dbasepoint\, the germ can be used to compute the mathematical derivative of the data for visualization tasks that require this information.
 
@@
 \label{sec:atct:xi}
-There is an expectation that for a visualization to be readable, the visual elements must correspond to distinct data elements\cite{ziemkiewiczEmbeddingInformationVisualization2009} and we can use the properties of sheaves to formally express this correspondence.
-We first describe the relationship between the graphic indexing space \gbase\ and the data indexing space \dbase\, which we propose is one where multiple graphic indexes map to one data index, and every index in the graphic space can be mapped to an index in the data space.
-We encode these expectations as the \textcolor{functor}{map} \vindexc, which we define to be a surjective continuous map
+When data and graphic containers are modeled as sheaves and there is a continuous map between their base spaces, the properties of sheaves can be used to describe how visual elements correspond to distinct data elements, which is a necessary condition of a visualization being readable\cite{ziemkiewiczEmbeddingInformationVisualization2009}.
+We propose that if a visualization is structure preserving, there exists a homeomorphic map \vindexc\ between the graphic indexing space \gbase\ and the data indexing space \dbase\ that maps multiple graphic indexes to one data index such that every index in the graphic space can be mapped to an index in the data space.
+We define the mapping \vindexc\ to be a surjective continuous map:
 \begin{equation}
@@
 \end{equation}
-between a graphic subspace $\opensetg \subseteq \gbase$ and data subspace $\openset \subseteq \dbase$. The functor $\vindex$ is surjective
-such that we can identify the set of points in graphic space that correspond to each point in data space
+between a graphic subspace $\opensetg \subseteq \gbase$ and data subspace $\openset \subseteq \dbase$. The set of points in graphic space that correspond to each point in data space is
 \begin{equation}
@@
 \end{equation}
-and every point in a graphic space has a corresponding point in data space.
-
-We construct the map as going from graphic to data because that encodes the notion that every visual element traces back to the data in some way.
-As exemplified in \autoref{fig:atct:morphisms:sheaf}, we define $\vindex$ as a surjective map because it allows us to express that a union of graphic spaces $S_i$ maps to single data point $\dbasepoint$, which allows us to express visual representations of a single record that are the union of many primitives, such as multipart glyphs (e.g boxplots) and combinations of plot types (e.g line with point markers).
+such that every point in a graphic space has a corresponding point in data space.
+
+We construct the map as going from graphic to data because that encodes the notion that every visual element traces back to the data in some way, but there may be more data available than what is shown in the graphic.
+As exemplified in \autoref{fig:atct:morphisms:sheaf}, defining $\vindex$ as a surjective map allows allows us to express visual representations of a single record that are the union of many primitives, each of which has a different base space $\gbase_i$.
+These visual representations include multipart glyphs (e.g boxplots), combinations of plot types (e.g line with point markers), and the same point showing up in different continuities, such as $\dsection(\dbase_point)$ in \autoref{fig:atct:morphisms:sheaf} and a station in \autoref{fig:related-work:continuity:ktypes}.
 
 \subsubsection{Data and Graphic Correspondence}
-Since we have defined a function \vindex\ between two spaces $\dbase, \gbase$, we can then construct functors that transport sheaves over each space to the other\cite{harder2008lectures}.
-This allows us to describe what data we expect at each graphic index location and what graphic is expected at each data index location.
-Transport functors compose the indexing map \vindex\ with the sheave map to say that a record \dsection\ at \dbasepoint\ is at all corresponding \gbasepoint\ and that a function \gsection\ over one point \gbasepoint\ is the same function at all points $\gbasepoint \in \gbase$ that correspond to the same record index \dbasepoint.
+Since we have defined a continuous function \vindex\ between two spaces $\dbase, \gbase$, we can construct functors that transport sheaves over each space to the other space\cite{harder2008lectures}.
+This allows us to describe what data we expect is being visualized at each graphic index location and what graphic is generated for the data at each data index location.
+Transport functors compose the indexing map \vindex\ with the sheaf map to say that a record \dsection\ evaluated at a data index \dbasepoint\ is the same record at all corresponding  graphic indices \gbasepoint\ and that a graphic specification \gsection\ over one point \gbasepoint\ is the same specification at all points $\gbasepoint \in \gbase$ that correspond to the same record index \dbasepoint.
 
 \paragraph{\textbf{Graphic Corresponding to Data}}
+
 The pushforward (direct image) sheaf establishes which graphic generating function $\gsection$ corresponds to a point $\dbasepoint \in dbase$ in the data base space.
@@
 Therefore, the continuous map $\vindex$ and transport functors $\vindexpull, \vindexpush$ allow us to express the correspondence between graphic section and data section.
+The parameterization of $\vindexpush\gsection$ in \autoref{fig:atct:morphisms:sheaf} is intended as an approximation and is akin to declarative visualization specs such as vega \cite{satyanarayanDeclarativeInteractionDesign2014} and svg \cite{quintScalable2003}.
+These specs and $\vindexpush\gsection$ also provide a renderer independent way of describing the graphic and are therefore useful for standardizing internal representation of the graphic and serializing the graphic for portability.
 
 \paragraph{\textbf{Data Corresponding to Graphic}}
-The pullback (inverse image) sheaf establishes which data record  returned by $\dsection$ corresponds to a point $\gbasepoint \in \gbase$ in the graphic base space.
-\begin{definition} \cite{harder2008lectures} Given a sheaf $\sheaf_{\dbase, \dtotal}$ on $\dbase$, the \textbf{pullback} sheaf $\vindexpullc\sheafc_{\dbasec, \dtotalc}$ on $\gbasec$ is defined as the sheaf associated to the presheaf
-  \begin{equation*}
-    \vindexpullc(\sheafc_{\dbasec, \dtotalc})(\opensetgc) = \sheafc_{\dbasec, \dtotalc}(\vindexc(\opensetgc))
-  \end{equation*}
-for $\vindex(\opensetgc) \in \dbase$.
+The pullback (inverse image) sheaf establishes which data record returned by $\dsection$ corresponds to each point $\gbasepoint \in \gbase$ in the graphic base space.
+\begin{definition} \cite{harder2008lectures} Given a sheaf $\sheaf_{\dbase, \dtotal}$ on $\dbase$, the \textbf{pullback} sheaf $\vindexpullc\sheafc_{\dbasec, \dtotalc}$ on $\gbasec$ is defined as the sheaf associated to the presheaf $\vindexpullc(\sheafc_{\dbasec, \dtotalc})(\opensetgc) = \sheafc_{\dbasec, \dtotalc}(\vindexc(\opensetgc))$ for $\vindex(\opensetgc) \in \dbase$.
 \end{definition}
@@
 As \vindex\ is surjective, there are many points $\gbasepoint \in \opensetg\subseteq\gbase$ in the graphic space that correspond to a single point $\vindex(\gbasepoint) = \dbasepoint$.
+As illustared in \autoref{fig:atct:morphisms:sheaf}, $\vindexpullc$ is critical to interactive techniques such as brushing, linking, and tooltips\cite{beckerBrushingScatterplots1987} because they depend on being able to look up which data values go to which parts of the graphic.
 
 \subsubsection{Example: Graphic and Data}
-\begin{figure}[t]
+\begin{figure*}[t]
   \includegraphics[width=1\textwidth]{xi_diagram.pdf}
-  \caption{The data consists of the $sin$ and $cos$ functions over a unit circle base space.
-We choose to visualize this as a circle and two line plots.
-The indexing function \vindex\ book keeps \note{find better word than bookkeeping} which parts of the circle and each curve correspond to each point on the unit circle.
-The pushforward $\vindexpush$ matches each point in the data space to the specification of the graphic at that point, while the pullback $\vindexpull$ matches each point in the graphic space to the data over that point.
+  \caption{The data  \dsectionc\ consists of the $sin$ and $cos$ functions over a unit circle base space, which here are circle and two line plots generated from the graphic specification \gsectionc\.
+The index map \vindexc\ keeps track of which part of the circle, sin, and cosine plots correspond to which point on the unit circle.
+For example the orange bands on $\gbase_0, \gbase_1, \gbase_2$ map to the orange point $\dbasepoint_i$.
+The pushforward $\vindexpushc$ matches each point in the data space to the specification of the graphic at that point, here illustared as an svg-like specification.
+The pullback $\vindexpull$ matches each point in the graphic space to the data over that point, and as shown often the same data point maps to multiple graphic spaces (and their corresponding pixels).
   \label{fig:atct:morphisms:sheaf}}
-\end{figure}
-
-Functors between sheaves are a way of expressing the bookkeeping involved in keeping track of which graphic section \gsection\ corresponds to which data section \dsection.
-The $(\dbasepoint_i, \gbase_i)$ pairing expressed in \autoref{eq:atct:xi} establishes that there is a correspondence between sections evaluated over $\dbasepoint_i$ and $\gbase_i$.
-This allows us to construct graphic specifications for each data index $\vindexpush\gsection$ and retrieve the data $\vindexpull\dsection$ for any graphic section generating any piece of a graphic.
-In \autoref{fig:atct:morphisms:sheaf}, the visualization is a graphic representation of a unit circle and the sin and cosine curves on that interval.
-The index lookup $\vindex$ describes which parts of the circle and curves are generated from which points on the unit circle.
-Given this correspondence, the pullback $\vindexpull\dsection$ looks up which values are being represented in a given part of the graphic.
-This type of lookup is critical for interactive techniques such as brushing, linking, and tooltips\cite{beckerBrushingScatterplots1987}.
-The pushforward $\vindexpush\gsection$ describes how a graphic is supposed to look for each point in the data space.
-The graphic parameterization in \autoref{fig:atct:morphisms:sheaf} is intended as an approximation of $\vindexpush\gsection$ and is akin to declarative visualization specs such as vega \cite{satyanarayanDeclarativeInteractionDesign2014} and svg \cite{quintScalable2003}.
-These specs and $\vindexpushc\gsection$ provide a renderer independent way of describing the graphic and are therefore useful for standardizing internal representation of the graphic and serializing the graphic for portability.
+\end{figure*}
+
+As illustrated in \autoref{fig:atct:morphisms:sheaf}, modeling the data and graphic containers as sheaves and constructing structure preserving maps between the indicies provides a method of precisely describing the which data values belong with which graphical representations.
+The $(\dbasepoint_i, \gbase_i)$ pairing expressed in \autoref{eq:atct:xi} establishes that there is a correspondence between data at an index $\dbasepoint_i$ and graphics generated at $\gbase_i$.
+In \autoref{fig:atct:morphisms:sheaf}, one example of this correspondance is the the orange bands on $\gbase_0, \gbase_1, \gbase_2$ that correspond to the orange point $\dbasepoint_i$.
+Because of this correspondance and the properties of sheaves, we can construct a graphic specifications for each data index $\vindexpush\gsection$; in \autoref{fig:atct:morphisms:sheaf}  $\vindexpush\gsection(\dbase_i)$ a psuedo-svg spec describing the piece of the circle, sin, and cosine curves generated for the data at the point \dbasepointc\.
+We can also retrieve the data $\vindexpull\dsection_{\vindexpre(\dbasepoint)}$ that is being visualized by each specific part of the graphic, which in \autoref{fig:atct:morphisms:sheaf} is the section $\dsection(\dbasepoint_i)$.
 
@@
 \begin{table}[H]
-  {\renewcommand{\arraystretch}{1.2}
-  \begin{tabular}{|r|l|}
-    \hline
-    merge & equation & behavior\\
-    \hline
-    indices  & $\dbasec_{a} \sqcup_{\dbasec_c} \dbasec_b$ & ${\dbasepoint \in \dbase_a, \dbase_b$ | $\dbasepoint_c \in \dbase_a, \dbase_b$ for all $dbase_ \in \dbasec$ \\
-    \hline
-    fields  & $\dfiberc_{a} \times_{\dfiberc_c} \dbase_c$
-    \hline
-
-  \caption{}
-  \label{tab:appendix:summary:merge}
-  }
-\end{table}
-
-\begin{table}[H]
   \centering

### 2024-12-27 paper/main.tex
artist up to equivariance

@@
 
-\begin{equation}
+\begin{equation*}
   \textcolor{section}{\texttt{dataset}}: \textcolor{base}{\texttt{topology}} \rightarrow \textcolor{fiber}{\texttt{fields}}
-\end{equation}
+\end{equation*}
 
@@
 
-When a bundle is trivial $\dtotalc = \dbasec \times \dfiberc$, we can defined a global sections $\dsection: \dbase \rightarrow \dfiber \in \Gamma(\dbase, \dfiber)$ which we translate into a data signature of the form
-\begin{equation}
-  \textcolor{section}{\texttt{dataset}}: \textcolor{base}{\texttt{topology}} \rightarrow \textcolor{fiber}{\texttt{field}}
-\end{equation}
-where $\dsection=\texttt{dataset}$, $\dbase=\texttt{topology}$ and $\dfiber=\texttt{fields}$.
+When a bundle is trivial $\dtotalc = \dbasec \times \dfiberc$, we can defined a global sections $\dsection: \dbase \rightarrow \dfiber \in \Gamma(\dbase, \dfiber)$ which we translate into a data signature of the form $\textcolor{section}{\texttt{dataset}}: \textcolor{base}{\texttt{topology}} \rightarrow \textcolor{fiber}{\texttt{field}}$ where $\dsectionc=\textcolor{section}{\texttt{dataset}}$, $\dbasec=\textcolor{base}{\texttt{topology}}$ and $\dfiberc=\textcolor{fiber}{\texttt{fields}}$.
 When the bundle is non-trivial, we can use the fiber bundle property of local-triviality to define local sections $\dsection\restriction_{\openset} \in \Gamma(\openset_{\dbasepoint}, \dtotal\restriction_{\openset_{\dbasepoint}})$.
 A local section is defined over an open neighborhood  $\dbasepoint \in \openset \in \dbase$, which is an open set that surrounds a point \dbasepoint.
-Most data sets can be encoded as a collection of local sections $\{\dsection\restriction_{\openset_{\dbasepoint}}| \dbasepoint \in \dbase \}$ and this encoding can be translated into a set of signatures
-\begin{equation}
-  \begin{split}
-\{\textcolor{section}{\texttt{data-subset}}: \textcolor{base}{\texttt{topology}} \rightarrow \textcolor{fiber}{\texttt{fields}} \\
- s.\;t.\;\textcolor{section}{\texttt{data-subset}} \subset \texttt{dataset}\}
-  \end{split}
-\end{equation}
+Most data sets can be encoded as a collection of local sections $\{\dsection\restriction_{\openset_{\dbasepoint}}| \dbasepoint \in \dbase \}$ and this encoding can be translated into a set of signatures $\{\textcolor{section}{\texttt{data-subset}}: \textcolor{base}{\texttt{topology}} \rightarrow \textcolor{fiber}{\texttt{fields}}$ s.t.
+$\textcolor{section}{\texttt{data-subset}} \subset \textcolor{section}{\texttt{dataset}}\}$.
 The subsets of the fiber bundle and the transition maps between these subsets are encoded in an atlas\cite{ghristElementaryAppliedTopology2014} and the notion of an atlas can be incorporated into the data container, as discussed in \autoref{sec:atct:sheaves}.
@@
 
-\begin{figure}[H]
+\begin{figure}[]
   \includegraphics[width=1\columnwidth]{fb_rho.pdf}
@@
 A common way of encapsulating a map from a topological space to a category of sets is as a presheaf
-\begin{definition}
+\begin{definition}\label{def:atct:presheaf}
   A \textbf{presheaf} $F:\mathcal{C}^{op} \rightarrow \setb$ is a contravariant functor from an object in an arbitrary category to an object in the category \setb\cite{nlab:presheaf, spanier1989algebraic}.
@@
 As \vindex\ is surjective, there are many points $\gbasepoint \in \opensetg\subseteq\gbase$ in the graphic space that correspond to a single point $\vindex(\gbasepoint) = \dbasepoint$.
-As illustared in \autoref{fig:atct:morphisms:sheaf}, $\vindexpullc$ is critical to interactive techniques such as brushing, linking, and tooltips\cite{beckerBrushingScatterplots1987} because they depend on being able to look up which data values go to which parts of the graphic.
+As illustrated in \autoref{fig:atct:morphisms:sheaf}, $\vindexpullc$ is critical to interactive techniques such as brushing, linking, and tooltips\cite{beckerBrushingScatterplots1987} because they depend on being able to look up which data values go to which parts of the graphic.
 
@@
 For example the orange bands on $\gbase_0, \gbase_1, \gbase_2$ map to the orange point $\dbasepoint_i$.
-The pushforward $\vindexpushc$ matches each point in the data space to the specification of the graphic at that point, here illustared as an svg-like specification.
+The pushforward $\vindexpushc$ matches each point in the data space to the specification of the graphic at that point, here illustrated as an svg-like specification.
 The pullback $\vindexpull$ matches each point in the graphic space to the data over that point, and as shown often the same data point maps to multiple graphic spaces (and their corresponding pixels).
+The artist \vartistc\ is the function
+  that transforms data into a graphic and is introduced in \autoref{sec:artist}.
   \label{fig:atct:morphisms:sheaf}}
@@
 
-As illustrated in \autoref{fig:atct:morphisms:sheaf}, modeling the data and graphic containers as sheaves and constructing structure preserving maps between the indicies provides a method of precisely describing the which data values belong with which graphical representations.
+As illustrated in \autoref{fig:atct:morphisms:sheaf}, modeling the data and graphic containers as sheaves and constructing structure preserving maps between the indices provides a method of precisely describing the which data values belong with which graphical representations.
 The $(\dbasepoint_i, \gbase_i)$ pairing expressed in \autoref{eq:atct:xi} establishes that there is a correspondence between data at an index $\dbasepoint_i$ and graphics generated at $\gbase_i$.
-In \autoref{fig:atct:morphisms:sheaf}, one example of this correspondance is the the orange bands on $\gbase_0, \gbase_1, \gbase_2$ that correspond to the orange point $\dbasepoint_i$.
-Because of this correspondance and the properties of sheaves, we can construct a graphic specifications for each data index $\vindexpush\gsection$; in \autoref{fig:atct:morphisms:sheaf}  $\vindexpush\gsection(\dbase_i)$ a psuedo-svg spec describing the piece of the circle, sin, and cosine curves generated for the data at the point \dbasepointc\.
+In \autoref{fig:atct:morphisms:sheaf}, one example of this correspondence is the the orange bands on $\gbase_0, \gbase_1, \gbase_2$ that correspond to the orange point $\dbasepoint_i$.
+Because of this correspondence and the properties of sheaves, we can construct a graphic specifications for each data index $\vindexpush\gsection$; in \autoref{fig:atct:morphisms:sheaf}  $\vindexpush\gsection(\dbase_i)$ a svg-like spec describing the piece of the circle, sin, and cosine curves generated for the data at the point \dbasepointc\.
 We can also retrieve the data $\vindexpull\dsection_{\vindexpre(\dbasepoint)}$ that is being visualized by each specific part of the graphic, which in \autoref{fig:atct:morphisms:sheaf} is the section $\dsection(\dbasepoint_i)$.
@@
 \label{sec:artist}
-In this work we propose that visualization libraries are implementing transformations from data sheaf to graphic sheaf. We call these subset of functions the artist:
+%%Representing data and graphics as fiber bundles means that the fields and topology of the dataset and graphic can be described in a uniform way.
+Implementing data containers as sheaves means that visualization components can be designed under the assumption that subsets of the data have the same properties as the whole sets.
+Maintaining an index between graphic space and data space facilitates transporting graphic and data access functions (\gsection, \dsection) to data and graphic space (\vindexpush\gsection, \vindexpull\dsection) respectively.
+-> not sure if
+Data visualization components are implementing a structure preserving mapping from a dataset to a graphic, which in the framework presented in this paper is a function from data to graphic sheaf.
+We call these structure preserving functions \textcolor{artist}{Artists}\footnote{Artists are named for the Artist class in Matplotlib.
+Artist is the parent class of all the objects that assemble and manage visual elements such as points, lines, and shapes.}:
+
 \begin{align}
   \label{eq:artist:hom_transport}
-  \vartistc:& \cgamma{\dbasec}{\dtotalc}\textcolor{artist}{\rightarrow} \cgamma{\gbasec}{\gtotalc}
+  \vartistc:& \cgamma{\dbasec}{\dtotalc}\textcolor{artist}{\rightarrow} \imartist{\gbasec}{\gtotalc}, \imartist{\gbasec}{\gtotalc} \subset \cgamma{\gbasec}{\gtotalc}
 \end{align}
- The artists can be constructed as morphisms of sheaves over the same base spaces through the application of pushforward and pullback functors; therefore they are natural transformations.
- \begin{definition} Given two functors $F,G$ with the same domain $\mathcal{C}$ and codomain $\mathcal{D}$, a \textbf{natural transformation} $\alpha: F \Rightarrow G$ is a map
-\begin{itemize}
-  \item \textit{data}: morphism $F(c) \xrightarrow{\alpha_{c}} G(c)$ for each object $c \in \mathcal{C}$
-  \item \textit{property} when $f: c_1 \rightarrow c_2$ is a morphism in $\mathcal{C}$, the components of the natural transform commute $G(f) \circ \alpha_{c_1} = \alpha_{c_2} \circ F(f)$
-\end{itemize}
-such that $\alpha = (\alpha_{c})_{c\in\mathcal{C}}$ is the set of all natural transformation components $\alpha_{c}$.\cite{bradleyWhatNaturalTransformation}
- \end{definition}
- This means that natural transforms are maps of functors that take the same input object and return objects in the same category\cite{milewskiCategoryTheoryProgrammers}.
-As illustrated in \autoref{eq:atct:sheaves:homset}, the sheaf functors
-\begin{equation}
-  \label{eq:artist:sheaf:base}
-    \begin{tikzcd}
-      \cgamma{\dbasec}{\dtotalc} &  & \dbasec \arrow[ll, "{\sheafc_{\dbasec, \dtotalc}}"', maps to, color=sheaf] \arrow[rr, "{\vindexpushc\sheafc_{\gbasec, \gtotalc}}", maps to, color=sheaf] &  & \cgamma{\dbase}{\vindexpushc\gtotalc}
-      \end{tikzcd}
-\end{equation}
-take as input an openset object \opensetc\ or \opensetgc\ and return sets of data and graphic sections that are objects in \setb.
-As a map between these sheaf functors, the artist has to preserve the $\iota, \iota^*$ morphisms of the presheaf functor, described in \autoref{eq:atct:sheaf:inclusion} and
-\autoref{eq:atct:sheaf:restriction}, such that the following diagram commutes:
-\note{this needs human words - subsets of functions of the same type map to subsets of visualizations of the same type}
-
-\begin{equation}
-  \label{eq:artist:natural_transform:inclusions}
-  \begin{tikzcd}
-    \dbasec_1 & \cgamma{\dbasec_1}{\dtotalc}
-    \arrow[dd, "\iota^*"', color=set]
-    \arrow[rr, "\vartistc_{\dbase_1}", color=artist] &  &
-    \cgamma{\dbasec_{1}}{\vindexpushc\gtotalc}
-    \arrow[dd, "\iota^*", color=set] \\
-      &  &  &  \\
-    \dbasec_{2} \arrow[uu, "\iota", hook, color=base] &
-    \cgamma{\dbasec_{2}}{\dtotalc}
-    \arrow[rr, "\vartistc_{\dbasec_2}", color=artist] &  &
-    \cgamma{\dbasec_{2}}{\vindexpushc\gtotalc}
-    \end{tikzcd}
-\end{equation}
- The diagram in \autoref{eq:artist:natural_transform:inclusions} shows that restricting a set of outputs of an artist to a set of graphic sections over a subspace is equivalent to restricting the inputs to data sections over the same subspace.
-Because the artist is a functor of sheaves, the artist is expected to translate the data continuity to graphic continuity such that the connectivity of subsets is preserved.
-This bookkeeping is necessary for any visualization technique that selectively acts on different pieces of a data set; for example streaming visualizations \cite{krstajicVisualizationStreamingData2013} and panning and zooming \cite{NekrasovskiEvaluationPanZoom2006}
-
-The output of an artist \vartist\ is a restricted subset of graphic sections
+
+The output of an artist \vartist\ is a restricted subset of graphic sections $\imartist{\gbasec}{\gtotalc}$ that are, by definition, only reachable through a structure preserving artist:
 \begin{equation}
@@
 \end{equation}
-that are, by definition, only reachable through a structure preserving artist, which we describe in \autoref{sec:artist:equivariant:artist}.
-We define this subset because the space of all sections $\cgamma{\opensetg}{\gtotal\restriction_{\openset}}$ includes sections that may not be structure preserving.
-For example, a section may go from every point in the graphic space to the same single point in the graphic fiber $\gsection(\gbasepoint_i) = d\; \forall \gbasepoint \in \gbase$ such that the visual output is a single inked pixel on a screen.
+We define this subset because the space of all sections $\cgamma{\opensetg}{\gtotal\restriction_{\openset}}$ is every possible way of setting pixels on an H base space, including non structure preserving settings such as setting all the pixels the same color.
+
+While Artists are defined as maps going from data over \dbase\ to graphic over \gbase\, they are compatible with maps defined solely over \dbase\ or \gbase\.
+This is because when a sheaf is equipped with transport functors (\vindexpush, \vindexpull), maps between sheaves over one space $Hom _{\sheaf_{\dbase}}$ are isomorphic to maps between sheaves over the other space $Hom _{\sheaf_{\gbase}}$\cite{harder2008lectures}:
+
+\begin{equation}
+  \label{eq:artist:homset}
+  \begin{tikzcd}[column sep=50]
+ \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}
+& \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}
+\arrow[l, "\vindexpushc"', color=functor]                                          \\
+\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
+\arrow[r, "\vindexpullc", color=functor]
+\arrow[u, "\textcolor{set}{Hom}_{\sheafc_{\dbasec}}", color=homset]
+\arrow[ru, "{\textcolor{set}{Hom}_{\sheafc_{\dbasec},\sheafc_{\gbasec}}}", color=homset]
+&
+\cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}
+\arrow[u, "\textcolor{set}{Hom}_{\sheafc_{\gbasec}}"', color=homset]
+\end{tikzcd}
+\end{equation}
+Both $Hom _{\sheaf_{\dbase}}$ and $Hom _{\sheaf_{\gbase}}$ are compatible, via composition with \vindex, with a mapping from data sheaf in data space to graphic sheaf in graphic space $Hom _{\sheaf_{\dbase}, \sheaf_{\gbase}}$.
+Artists are a subset of structure preserving functions in $Hom _{\sheaf_{\dbase}, \sheaf_{\gbase}}$ \textcolor{artist}{Artists}.
+For example, the artist in \autoref{fig:atct:morphisms:sheaf} is the blue arrow that converts $(i, \sin(i), cos(i))$ into the corresponding sections of the circle and lines.
+
+The commutativity of \autoref{eq:artist:homset} means that an artist over data space $\vartist_{\dbase}: \dsection \mapsto \vindexpush \gsection$, an artist over graphic space $vartist_{\gbase}: \vindexpull \dsection \mapsto \gsection$, and an artist $\vartist: \dsection \mapsto \gsection$ are equivalent such that:
+\begin{equation*}
+  \begin{split}
+  \dsection(\dbasepoint) &= \vindexpull\dsection(\gbasepoint)  \\
+   \implies
+  \vartist_{\dbase}(\dsection(\dbasepoint)) &= \vartist_{\gbase}(\vindexpull\dsection(\gbasepoint)) = \vartist(\dsection(\dbasepoint))\\
+  \implies \vindexpush\gsection(\gbasepoint) &= \gsection(\gbasepoint)
+  \end{split}
+\end{equation*}
+when $\vindex(\gbasepoint) = \dbasepoint$.
+The ability to shift base spaces using adaptors (\vindexpush, \vindexpull) allows a developer to specify structure preserving visualization components in data space, graphic space, or a mix of both depending on what is optimal for the specific use case.
+For example, mapping a line to a color in data space, but drawing the line in graphic space so that it can be dynamically resampled.
+
+Because Artists can be constructed as morphisms of sheaves over the same base spaces through the application of pushforward and pullback functors, they are natural transformations \cite{bradleyWhatNaturalTransformation}.
+\begin{definition}
+  Given two functors $F, G: \mathcal{C}\rightarrow \mathcal{D}$, a \textbf{natural transformation} $\alpha: F \rightarrow G$ is a function which assigns to each object $c$ of $\mathcal{C}$ a morphism $\alpha_c:F(input) \rightarrow G(c), G(c) \in \mathcal{D}$, in such a way that for every morphism $f:c \rightarrow c^\prime, c^\prime \in \mathcal{C}$, the morphisms in $\mathcal{D}$ commute such that $\alpha_c^{\prime}(F(f)(F(c))) = G(f)(\alpha_c(F(c))$.
+When this holds, $\alpha_{c}$ is \textit{natural} in $c$.\cite{maclaneCategoriesWorkingMathematician2013}.
+\end{definition}.
+
+This means that the structure on $\mathcal{C}$, as defined by its input objects $c \in \mathcal{C}$ and morphisms $f: c \rightarrow c^\prime$ are preserved in the output of the functors $F, G$ and by maps between functors $\alpha$.
+This preservation of the structure of the input in the output encapsulates the structure preservation discussed in \autoref{sec:related-work:equivariance} and serves as the basis for describing structure preservation in \autoref{sec:artist:equiv}.
+Because the Artist is a natural transform and natural transforms are maps of functors that take the same input object and return objects in the same category\cite{milewskiCategoryTheoryProgrammers}, the artist signature can be formulated as:
+ \begin{align*}
+  \textcolor{artist}{\texttt{Artist}}:&(\textcolor{section}{\texttt{dataset}}: \textcolor{base}{\texttt{topology}}\textcolor{section}{\rightarrow} \textcolor{fiber}{\texttt{fields}}) \\
+  \textcolor{artist}{\rightarrow} & (\textcolor{section}{\texttt{graphic}}: \textcolor{base}{\texttt{topology}}\textcolor{section}{\rightarrow} \textcolor{fiber}{\texttt{fields}})
+ \end{align*}
+where \texttt{topology} and \texttt{field} are meta classes of compatible types.
 
@@
 
-we can assert that the data space \dbase acts as an indexing space into \gbase such that knowing the location on space yields the location on the other and any point in either base space or graphic space has a corresponding point in the other space.
+we can assert that the data space \dbase\ acts as an indexing space into \gbase\ such that knowing the location in one space yields the location on the other and any point in either base space or graphic space has a corresponding point in the other space.
 \begin{figure}[H]
@@
 This point $\gbasepoint$ is the input into the graphic section $\gsection(\gbasepoint)$ that is used to determine which pixels are colored, which in turn determines the thickness, texture, and color of the line.
+
+Because the artist is a functor of sheaves, the artist preserves the inclusion and restriction conditions of a presheaf \autoref{def:atct:presheaf} and the locality and gluing conditions of a sheaf \autoref{def:atct:sheaf} when mapping between subspaces $\vindex(\opensetg) \rightarrow \openset$.
+This bookkeeping is necessary for any visualization technique that selectively acts on different pieces of a data set; for example streaming visualizations \cite{krstajicVisualizationStreamingData2013} and panning and zooming \cite{NekrasovskiEvaluationPanZoom2006}
 
@@
 \subsection{Implementing the Artist}
-
- When a sheaf is equipped with transport functors, then the functions between sheaves over one space are isomorphic to functions between sheaves over the other space\cite{harder2008lectures} such that the following diagram commutes
-
- \note{should either be oriented same as 55 and/or pushed back up to 3.3 as an intro to artist or squished a little. }
- \begin{equation}
-  \label{eq:atct:sheaves:homset}
-  \begin{tikzcd}
-    \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
-    \arrow[dd, "\textcolor{set}{Hom}_{\sheafc_{\dbasec}}"', color=homset]
-    \arrow[rrdd, "\textcolor{set}{Hom}_{\sheafc_{\dbasec},\sheafc_{\gbasec}}", color=homset]
-    \arrow[rr, "\vindexpullc", color=functor] &  &
-    \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}
-    \arrow[dd, "\textcolor{set}{Hom}_{\sheafc_{\gbasec}}", color=homset] \\
-     & & \\
-    \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} &  &
-    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}
-    \arrow[ll, "\vindexpushc"', color=functor]
-    \end{tikzcd}
-\end{equation}
-
-Since the artist is a family of functions in the homset between sheaves, the isomorphism allows for the specification of the transformation from data as combination of functions over different spaces such that the following diagram commutes:
+Since the artist is a family of functions in the homset between sheaves (\autoref{eq:atct:sheaves:homset}) the isomorphism allows for the specification of the transformation from data as a combination of functions over different spaces:
+
 \begin{equation}
@@
   \includegraphics[width=1\columnwidth]{transition_maps.pdf}
-  \caption{The cylinder is a trivial fiber bundle; therefore it can be decomposed into local trivalizations that only need identity
+  \caption{The cylinder is a trivial fiber bundle; therefore it can be decomposed into local trivializations that only need identity
   maps to glue the trivializations together.

### 2024-12-27 paper/main.tex
a presheaf is a presheaf

@@
   \item \textit{locality} two sections in a sheaf are equal $\dsection^{a} = \dsection^{b}$ when they evaluate to the same values $\dsection^{a}\vert_{\openset_i} =  \dsection^{b}\vert_{\openset_i}$ over the open cover $\bigcup_{i\in I} \openset_{i} \subset \openset$ (indexed by $I$).
-  \item \textit{gluing} the union of sections defined over subspaces $\dsection^{i} \in \Gamma(\openset_i, \dtotal|_{openset_i})$ is equivalent to a section defined over the whole space $\dsection\vert_{\openset_i} = \dsection^{i}$ for all $i\in I$ if all pairs of sections agree on overlaps $\dsection^{i}\vert_{\openset_i\cap\openset_j} =  \dsection^{j}\vert_{\openset_i\cap\openset_j}$
+  \item \textit{gluing} the union of sections defined over subspaces $\dsection^{i} \in \Gamma(\openset_i, \dtotal|_{\openset_i})$ is equivalent to a section defined over the whole space $\dsection\vert_{\openset_i} = \dsection^{i}$ for all $i\in I$ if all pairs of sections agree on overlaps $\dsection^{i}\vert_{\openset_i\cap\openset_j} =  \dsection^{j}\vert_{\openset_i\cap\openset_j}$
   \end{itemize}
@@
 \subsection{Homeomorphism}
-As mentioned in \autoref{sec:related-work:continuity}, preserving the topology of a visualization means that each discrete piece of differentiable visual information corresponds to a distance element of the dataset\cite{ziemkiewiczEmbeddingInformationVisualization2009} in a way where the organization of elements is preserved.
+As mentioned in \autoref{sec:related-work:continuity}, preserving the topology of a visualization means that each discrete piece of differentiable visual information corresponds to a dist element of the dataset\cite{ziemkiewiczEmbeddingInformationVisualization2009} in a way where the organization of elements is preserved.
 A generalization of this condition is the idea that the graphic space can be collapsed into the data indexing space, which means that the data base space is a deformation retraction of the graphic base space\cite{hatcherAlgebraicTopology2002}.
@@
 \begin{equation}
-  \vindex: \dbase \times I \rightarrow \dbase s.t \vindex(\dbasepoint) = \dbasepoint \forall \dbasepoint \in \gbase
+  \vindex: \dbase \times I \rightarrow \dbase\;s.t\;\vindex(\dbasepoint) = \dbasepoint\;\forall \dbasepoint \in \gbase
 \end{equation}
@@
 
-Because the artist is a functor of sheaves, the artist preserves the inclusion and restriction conditions of a presheaf \autoref{def:atct:presheaf} and the locality and gluing conditions of a sheaf \autoref{def:atct:sheaf} when mapping between subspaces $\vindex(\opensetg) \rightarrow \openset$.
-This bookkeeping is necessary for any visualization technique that selectively acts on different pieces of a data set; for example streaming visualizations \cite{krstajicVisualizationStreamingData2013} and panning and zooming \cite{NekrasovskiEvaluationPanZoom2006}
-
+When the data and graphic containers are implemented as sheaves, the artist preserves the sheaf constraints because it is using the sheaf constraint satisfying map $\vindex$ to maintain the correspondence between data and graphing indexing spaces.
+Given a subset of the index $\dbase_2 \xhookrightarrow{i} \dbase_1$, all the data in the larger index is included in the smaller set $\Gamma(\dbase_1, \dtotal) \xrightarrow{i^*} \Gamma(\dbase_2, \dtotal)$ such that visualization specs of the smaller set must include specs for the larger set of data values  $\vartist(\Gamma(\dbase_1, \dtotal)) \xrightarrow{i^*} \vartist(\Gamma(\dbase_2, \dtotal))$.
+While the curves in \autoref{fig:atct:presheaf} are meant as stand-ins for the data, they also serve to illustrate that the relationship is identical in the graphic space such that there may be more graphics available for data from a smaller indexing space because there may be more values on the smaller indexing space.
+The artist also preserves the \textit{locality} axiom in that identical datasets $\dsection^{a} = \dsection^{b}$ generate identical specifications $\vartist(\dsection^a) = \vartist(\dsection^b)$.
+The artist also respects \textit{gluing} in that graphics are composable, as described in \autoref{sec:artist:operators}, such that a set of artists that generate graphics for parts of a dataset are equivalent to an artist that generates a graphic for the whole dataset.
+Managing subsets of data and on demand data in a structure preserving way is necessary to support visualizing streaming data \cite{krstajicVisualizationStreamingData2013} and techniques that rely on shifting or resampling the indexing space, such as  panning and zooming \cite{NekrasovskiEvaluationPanZoom2006}.
+
+\note{stop here}
 \subsection{Equivariance}

### 2024-12-29 paper/main.tex
an artist is a natural transform on sheaves

@@
 Because Artists can be constructed as morphisms of sheaves over the same base spaces through the application of pushforward and pullback functors, they are natural transformations \cite{bradleyWhatNaturalTransformation}.
-\begin{definition}
+\begin{definition}\label{def:natural-transform}
   Given two functors $F, G: \mathcal{C}\rightarrow \mathcal{D}$, a \textbf{natural transformation} $\alpha: F \rightarrow G$ is a function which assigns to each object $c$ of $\mathcal{C}$ a morphism $\alpha_c:F(input) \rightarrow G(c), G(c) \in \mathcal{D}$, in such a way that for every morphism $f:c \rightarrow c^\prime, c^\prime \in \mathcal{C}$, the morphisms in $\mathcal{D}$ commute such that $\alpha_c^{\prime}(F(f)(F(c))) = G(f)(\alpha_c(F(c))$.
@@
 
-When the data and graphic containers are implemented as sheaves, the artist preserves the sheaf constraints because it is using the sheaf constraint satisfying map $\vindex$ to maintain the correspondence between data and graphing indexing spaces.
-Given a subset of the index $\dbase_2 \xhookrightarrow{i} \dbase_1$, all the data in the larger index is included in the smaller set $\Gamma(\dbase_1, \dtotal) \xrightarrow{i^*} \Gamma(\dbase_2, \dtotal)$ such that visualization specs of the smaller set must include specs for the larger set of data values  $\vartist(\Gamma(\dbase_1, \dtotal)) \xrightarrow{i^*} \vartist(\Gamma(\dbase_2, \dtotal))$.
-While the curves in \autoref{fig:atct:presheaf} are meant as stand-ins for the data, they also serve to illustrate that the relationship is identical in the graphic space such that there may be more graphics available for data from a smaller indexing space because there may be more values on the smaller indexing space.
-The artist also preserves the \textit{locality} axiom in that identical datasets $\dsection^{a} = \dsection^{b}$ generate identical specifications $\vartist(\dsection^a) = \vartist(\dsection^b)$.
-The artist also respects \textit{gluing} in that graphics are composable, as described in \autoref{sec:artist:operators}, such that a set of artists that generate graphics for parts of a dataset are equivalent to an artist that generates a graphic for the whole dataset.
+Specifying the artist as a natural transform (\autoref{df:natural-transform}) on sheaves means that the artist preserves the presheaf and sheaf constraints (\autoref{def:atct:presheaf}, \autoref{def:atct:sheaf}) when mapping data into graphics.
+The set of possible visualizations for a larger set of values must include all visualizations for a subset of those values  $\vartist(\Gamma(\dbase_1, \dtotal)) \xrightarrow{i^*} \vartist(\Gamma(\dbase_2, \dtotal))$ when more values may exist on the subset of the index $\dbase_2 \xhookrightarrow{i} \dbase_1$.
+While the curves in \autoref{fig:atct:presheaf} are meant as stand-ins for the data, they also serve to illustrate that the relationship is identical in the graphic space.
+The artist preserves the \textit{locality} axiom in that identical datasets $\dsection^{a} = \dsection^{b}$ generate identical specifications $\vartist(\dsection^a) = \vartist(\dsection^b)$.
+The artist respects \textit{gluing} in that graphics are composable, as described in \autoref{sec:artist:operators}, such that a set of artists that generate graphics for parts of a dataset are equivalent to an artist that generates a graphic for the whole dataset.
 Managing subsets of data and on demand data in a structure preserving way is necessary to support visualizing streaming data \cite{krstajicVisualizationStreamingData2013} and techniques that rely on shifting or resampling the indexing space, such as  panning and zooming \cite{NekrasovskiEvaluationPanZoom2006}.

### 2025-01-03 paper/main.tex
homemeomorphism

@@
 The $\otimes$ encapsulates both the sets of values associated with each fiber $\{y \in \mathbb{I}\vert 1992 \leq y \leq 2025\} \times \{m \in \mathbb{I}\vert 1 \leq m \leq 12 \} \times \{d \in \mathbb{I}\vert 1 \leq d \leq 31 \}$ and the composition function $ \otimes: \dfiber_{year} \times \dfiber_{month} \times \dfiber_{day} \rightarrow \dfiber_{date}$
-could include a constraint to only return dates that have the right number of days for each month.
+would include a constraint to only return dates that have the right number of days for each month.
 The bifunctor also composes the morphisms associated with each category into a morphism on the composite category $(\dfunct_{year}, \dfunct_{month}, \dfunct_{day}) = \dfunct_{date}$.
@@
 \subsection{Homeomorphism}
-As mentioned in \autoref{sec:related-work:continuity}, preserving the topology of a visualization means that each discrete piece of differentiable visual information corresponds to a dist element of the dataset\cite{ziemkiewiczEmbeddingInformationVisualization2009} in a way where the organization of elements is preserved.
-A generalization of this condition is the idea that the graphic space can be collapsed into the data indexing space, which means that the data base space is a deformation retraction of the graphic base space\cite{hatcherAlgebraicTopology2002}.
-By defining the indexing look up function \vindex, introduced in \autoref{sec:atct:sheaves}, to be
-\begin{equation}
-  \vindex: \dbase \times I \rightarrow \dbase\;s.t\;\vindex(\dbasepoint) = \dbasepoint\;\forall \dbasepoint \in \gbase
-\end{equation}
-
-we can assert that the data space \dbase\ acts as an indexing space into \gbase\ such that knowing the location in one space yields the location on the other and any point in either base space or graphic space has a corresponding point in the other space.
+
+As introduced in \autoref{sec:related-work:continuity}, the topology of a dataset is preserved in the visualization if there is a homeomorphic map between the graphic and data indexing space.
+Because multiple graphics can correspond to one data record and every graphic must have a corresponding record \cite{tufteVisualDisplayQuantitative2001,ziemkiewiczEmbeddingInformationVisualization2009}, we define \textcolor{functor}{indexing look up functor} \vindexc\ as a deformation retraction\cite{hatcherAlgebraicTopology2002} of the graphic indexing space \gbase\ into the data indexing space \dbase.
+This means that every point in \gbase\ can be mapped, non-uniquely, to a point in \dbase\:
+
+\begin{equation}
+  \vindexc: \dbasec \times I \rightarrow \dbasec\;s.t\;\vindexc(\gbasepointc) = \dbasepointc\;\forall \dbasepointc \in \gbasec
+\end{equation}
+The graphic indexing space is defined as the data space multiplied by an identity $\gbase = \dbase\times I$ because it allows us to encode the data index into the graphic index.
+For example, for a 2D space the graphic basepoint is defined as $\gbasepoint=(\alpha,\beta) \in \dbasepoint_i\times I$ .
+This means that the first coordinate $\alpha=\dbasepoint_i$ provides a lookup into the data space and the second coordinate is a point on the interval $\beta \in I$ and together they are input into the graphic function $\gsection(\gbasepoint)$ which returns how the pixel at $\gbasepoint$ is set.
+
 \begin{figure}[H]
-  \includegraphics[width=1\columnwidth]{deform_retract.png}
-  \caption{The graphic base space $\gbase$ is collapsible to the line $\dbase$ such that every band $(\dbasepoint_i, [0,1])$ on $\gbase$ maps to corresponding point $\dbasepoint_i \in \dbase$.
-The band $[0,1]$ determines the thickness of a rendered line for a given point $\dbasepoint_{i}$ by specifying how pixels corresponding to that point are colored.
-\label{fig:construction:xi}}
+  \includegraphics[width=1\columnwidth]{xi_zoom.pdf}
+  \caption{The function $\xi$ maps each point in each orange band on each graphic indexing space $\gbase_0, \gbase_1, \gbase_2$ to the same point $\dbasepoint_i$ in the circle $\dbase$.
+As shown in \autoref{fig:atct:morphisms:sheaf}, the parts of the graphic indexed by the orange bands are visualizing parts of the record at $\dbasepoint_i$ \label{fig:artist:xi}}
 \end{figure}
 
-For example, as shown in \autoref{fig:construction:xi}, a line is 1D but is a 2D glyph on a screen; therefore the graphic space $\gbase$ is constructed by multiplying the base space $\dbase$ with an interval $[0,1]$.
-Because $\gbase$ is collapsible into $\dbase$, every band $(\dbasepoint_i, [0,1])$ corresponds to a point in the base space $\dbasepoint_i \in \dbase$.
-The first coordinate $\alpha=\dbasepoint_i$ provides a lookup to retrieve the associated visual variables.
-The second coordinate, which is a point in the interval $\beta=[0,1]$.
-Together they are a point $\gbasepoint=(\alpha,\beta) \in gbase$ in the graphic base space.
-This point $\gbasepoint$ is the input into the graphic section $\gsection(\gbasepoint)$ that is used to determine which pixels are colored, which in turn determines the thickness, texture, and color of the line.
-
-Specifying the artist as a natural transform (\autoref{df:natural-transform}) on sheaves means that the artist preserves the presheaf and sheaf constraints (\autoref{def:atct:presheaf}, \autoref{def:atct:sheaf}) when mapping data into graphics.
-The set of possible visualizations for a larger set of values must include all visualizations for a subset of those values  $\vartist(\Gamma(\dbase_1, \dtotal)) \xrightarrow{i^*} \vartist(\Gamma(\dbase_2, \dtotal))$ when more values may exist on the subset of the index $\dbase_2 \xhookrightarrow{i} \dbase_1$.
+In \autoref{fig:artist:xi} each orange band is defined as $\frac{7\pi}{4}\times [0,1]$ such that each point on each band maps to the same point $\vindex|_{\frac{7\pi}{4}\times [0,1]} = \frac{7\pi}{2}$.
+As illustrated in \autoref{fig:atct:morphisms:sheaf}, each band acts as an indexer into its corresponding image because the section \gsection\ takes as input coordinates $(\frac{7\pi}{2}, \beta), \beta \in [0,1]$ in each orange band and returns the settings for the pixel it maps into.
+As illustrated in \autoref{fig:artist:xi}, the definition of $\vindex$ allows for some flexibility in the shape of $\gbase$.
+An annulus $\gbase_0$ is used for the green circle in \autoref{fig:atct:morphisms:sheaf} because that cyclical structure is in the graphic and in \dbase\, but rectangles $\gbase_1, \gbase_2$ are used for the $\sin$ and $\cos$ curves because they are linear.
+Despite discarding the cyclical information embedded in \dbase\, the spaces $\gbase_1$ and $\gbase_2$ are structure preserving because they satisfy the constraint on \vindex\ that every graphic element maps back into a data element \cite{ziemkiewiczEmbeddingInformationVisualization2009}.
+
+The artist is specified as a natural natural transform (\autoref{def:natural-transform}) on sheaves because it means that the artist can assume that subsets of the data behave in the same manner as the whole when the artist preserves the presheaf and sheaf constraints (\autoref{def:atct:presheaf}, \autoref{def:atct:sheaf}) when mapping data into graphics.
+Preserving the presheaf constraints means that the set of possible visualizations for a larger set of values must include all visualizations for a subset of those values  $\vartist(\Gamma(\dbase_1, \dtotal)) \xrightarrow{i^*} \vartist(\Gamma(\dbase_2, \dtotal))$ when more values may exist on the subset of the index $\dbase_2 \xhookrightarrow{i} \dbase_1$.
 While the curves in \autoref{fig:atct:presheaf} are meant as stand-ins for the data, they also serve to illustrate that the relationship is identical in the graphic space.

### 2025-01-03 paper/main.tex
typo

@@
 In \autoref{fig:artist:xi} each orange band is defined as $\frac{7\pi}{4}\times [0,1]$ such that each point on each band maps to the same point $\vindex|_{\frac{7\pi}{4}\times [0,1]} = \frac{7\pi}{2}$.
-As illustrated in \autoref{fig:atct:morphisms:sheaf}, each band acts as an indexer into its corresponding image because the section \gsection\ takes as input coordinates $(\frac{7\pi}{2}, \beta), \beta \in [0,1]$ in each orange band and returns the settings for the pixel it maps into.
+As illustrated in \autoref{fig:atct:morphisms:sheaf}, each band acts as an indexer into its corresponding image because the section \gsection\ takes as input coordinates $(\frac{7\pi}{4}, \beta), \beta \in [0,1]$ in each orange band and returns the settings for the pixel it maps into.
 As illustrated in \autoref{fig:artist:xi}, the definition of $\vindex$ allows for some flexibility in the shape of $\gbase$.
@@
 
-\note{stop here}
 \subsection{Equivariance}

### 2025-01-06 paper/main.tex
clean up nc, pull back to equations, rework intro to equivariance

@@
 As exemplified in \autoref{fig:atct:morphisms:sheaf}, defining $\vindex$ as a surjective map allows allows us to express visual representations of a single record that are the union of many primitives, each of which has a different base space $\gbase_i$.
-These visual representations include multipart glyphs (e.g boxplots), combinations of plot types (e.g line with point markers), and the same point showing up in different continuities, such as $\dsection(\dbase_point)$ in \autoref{fig:atct:morphisms:sheaf} and a station in \autoref{fig:related-work:continuity:ktypes}.
+These visual representations include multipart glyphs (e.g boxplots), combinations of plot types (e.g line with point markers), and the same point showing up in different continuities, such as $\dsection(\dbase_{point})$ in \autoref{fig:atct:morphisms:sheaf} and a station in \autoref{fig:related-work:continuity:ktypes}.
 
@@
 
-The commutativity of \autoref{eq:artist:homset} means that an artist over data space $\vartist_{\dbase}: \dsection \mapsto \vindexpush \gsection$, an artist over graphic space $vartist_{\gbase}: \vindexpull \dsection \mapsto \gsection$, and an artist $\vartist: \dsection \mapsto \gsection$ are equivalent such that:
-\begin{equation*}
-  \begin{split}
-  \dsection(\dbasepoint) &= \vindexpull\dsection(\gbasepoint)  \\
-   \implies
-  \vartist_{\dbase}(\dsection(\dbasepoint)) &= \vartist_{\gbase}(\vindexpull\dsection(\gbasepoint)) = \vartist(\dsection(\dbasepoint))\\
-  \implies \vindexpush\gsection(\gbasepoint) &= \gsection(\gbasepoint)
-  \end{split}
-\end{equation*}
-when $\vindex(\gbasepoint) = \dbasepoint$.
+The commutativity of \autoref{eq:artist:homset} means that an artist over data space $\vartist_{\dbase}: \dsection \mapsto \vindexpush \gsection$, an artist over graphic space $\vartist_{\gbase}: \vindexpull \dsection \mapsto \gsection$, and an artist $\vartist: \dsection \mapsto \gsection$ are equivalent.
+Specifically, given \autoref{eq:atct:sheaf:pullback_hover} (\vindexpull) and \autoref{eq:atct:sheaf:pushforward_select} (\vindexpush), $\vartist_{\dbase}(\dsection(\dbasepoint)) = \vartist_{\gbase}(\vindexpull\dsection(\gbasepoint)) = \vartist(\dsection(\dbasepoint)$ when $\vindex(\gbasepoint) = \dbasepoint$.
 The ability to shift base spaces using adaptors (\vindexpush, \vindexpull) allows a developer to specify structure preserving visualization components in data space, graphic space, or a mix of both depending on what is optimal for the specific use case.
-For example, mapping a line to a color in data space, but drawing the line in graphic space so that it can be dynamically resampled.
-
-Because Artists can be constructed as morphisms of sheaves over the same base spaces through the application of pushforward and pullback functors, they are natural transformations \cite{bradleyWhatNaturalTransformation}.
+For example, mapping a line to a color in data space, but drawing the line in graphic space so that it can be dynamically resampled based on panning and zooming.
+
+Because Artists can be constructed as morphisms of sheaves over either \dbase\ or \gbase\ through the application of pushforward and pullback functors (\vindexpush, \vindexpull), they are natural transformations \cite{bradleyWhatNaturalTransformation}.
 \begin{definition}\label{def:natural-transform}
@@
 This means that the structure on $\mathcal{C}$, as defined by its input objects $c \in \mathcal{C}$ and morphisms $f: c \rightarrow c^\prime$ are preserved in the output of the functors $F, G$ and by maps between functors $\alpha$.
+The artist \vartist\ is the natural transform mapping the sheaf functors $\sheaf_{data} \rightarrow \sheaf_{graphic}$, which both have the signature $\dbase^{op} (\gbase^{op})\rightarrow \texttt{set}$ \autoref{def:atct:presheaf}.
 This preservation of the structure of the input in the output encapsulates the structure preservation discussed in \autoref{sec:related-work:equivariance} and serves as the basis for describing structure preservation in \autoref{sec:artist:equiv}.
@@
 
-The artist is specified as a natural natural transform (\autoref{def:natural-transform}) on sheaves because it means that the artist can assume that subsets of the data behave in the same manner as the whole when the artist preserves the presheaf and sheaf constraints (\autoref{def:atct:presheaf}, \autoref{def:atct:sheaf}) when mapping data into graphics.
+Because the artist is specified as a natural transform (\autoref{def:natural-transform}) on sheaves, the artist expects that subsets of the dataset have the same properties as the whole dataset when the artist preserves the presheaf and sheaf constraints (\autoref{def:atct:presheaf}, \autoref{def:atct:sheaf}) when mapping data into graphics.
 Preserving the presheaf constraints means that the set of possible visualizations for a larger set of values must include all visualizations for a subset of those values  $\vartist(\Gamma(\dbase_1, \dtotal)) \xrightarrow{i^*} \vartist(\Gamma(\dbase_2, \dtotal))$ when more values may exist on the subset of the index $\dbase_2 \xhookrightarrow{i} \dbase_1$.
@@
 \label{sec:artist:equiv}
-As introduced in \autoref{sec:related-work:equivariance}, data and the corresponding visual encoding are expected to have compatible structure.
-This structure can be formally expressed as actions $\dfunc \in \dfuncset$ on the sheaf $\sheaf_{\dbase, \dtotal}$.
-We generalize from binary operations to a family of actions because that allows for expanding the set of allowable transformations on the data beyond a single operator.
-We describe the changes on the graphic side as changes in measurements \measure\, which are scaler or vector components of the rendered graphic that can be quantified, such as the color, position, shape, texture, or rotation angle of the graphic.
-The visual variables \cite{bertinIIPropertiesGraphic2011} are a subset of measurable components.
+As introduced in \autoref{sec:related-work:equivariance}, data and the corresponding visual encoding are expected to have compatible structure such that actions (\autoref{def:related-work:action}) on data and graphic are symmetric.
+We generalize from binary operations to a family of actions $\dfunc \in \dfuncset$ on the sheaf $\sheaf_{\dbase, \dtotal}$ because that allows for expanding the set of allowable transformations on the data beyond a single operator.
+We describe the changes on the graphic side as changes in measurements \measure\ of visual components.
+A superset of the Bertin visual variables \cite{\cite{bertinIIPropertiesGraphic2011}, measurable components are scaler or vector components of the rendered graphic that can be quantified, such as the color, position, shape, texture, or rotation angle of the graphic.
 For example, a measurement of a scatter marker could be its color (e.g.

### 2025-03-05 paper/main.tex
start of generalizing presheaf

@@
 \begin{definition}\label{def:atct:presheaf}
-  A \textbf{presheaf} $F:\mathcal{C}^{op} \rightarrow \setb$ is a contravariant functor from an object in an arbitrary category to an object in the category \setb\cite{nlab:presheaf, spanier1989algebraic}.
+  A \textbf{presheaf} $F:\mathcal{C}^{op} \rightarrow \setb$ is a contravariant functor from an object in an arbitrary category to an object in the category \setb\cite{spanier1989algebraic}.
 \end{definition}
 A functor is contravariant when the morphisms between the input objects go in the opposite direction from the morphisms between the output objects.
-The presheaf is contravariant because the inclusion morphisms between input object $\iota: \opensetc_1 \rightarrow \opensetc_2$
+The presheaf is contravariant when for every arbitrary morphism between input base spaces $\dfunch: \openset_1 \rightarrow \openset_2$ there exists a corresponding pullback function between the sets of sections $\dfunchpull:  \cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}} \rightarrow \cgamma{\opensetc_1}{\dtotalc\restriction_{\openset_c1}}$.
+The job
+
+%% these two paragraphs cen be condensed and moved below the figure as an example
+The presheaf is contravariant because the morphisms between input object $\iota: \opensetc_1 \rightarrow \opensetc_2$
 are defined such that they correspond to the partial ordering $\openset_1 \subseteq \openset_2$, but the restriction morphisms $\iota^*$ between the sets of sections $\iota^*: \cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}} \rightarrow \cgamma{\opensetc_1}{\dtotalc\restriction_{\openset_c1}}$ restricts the larger set to the smaller one such that all functions that are continuous over a space must be continuous over a subspace $\Gamma_2 \subseteq \Gamma_1$, where $\Gamma_{i}\coloneqq\Gamma(\openset_{i}, \dtotal\restriction_{\openset_{i}})$.
@@
 We generalize from binary operations to a family of actions $\dfunc \in \dfuncset$ on the sheaf $\sheaf_{\dbase, \dtotal}$ because that allows for expanding the set of allowable transformations on the data beyond a single operator.
-We describe the changes on the graphic side as changes in measurements \measure\ of visual components.
-A superset of the Bertin visual variables \cite{\cite{bertinIIPropertiesGraphic2011}, measurable components are scaler or vector components of the rendered graphic that can be quantified, such as the color, position, shape, texture, or rotation angle of the graphic.
+We describe the changes on the graphic side as changes in \textcolor{monoid}{measurements} \measurec\ of visual components.
+A superset of the Bertin visual variables \cite{bertinIIPropertiesGraphic2011}, measurable components are scaler or vector components of the rendered graphic that can be quantified, such as the color, position, shape, texture, or rotation angle of the graphic.
 For example, a measurement of a scatter marker could be its color (e.g.
@@
 
+\label{sec:construction:xi}
 As mentioned in \autoref{sec:artist:homeomorphic}, we construct the data base space as a deformation retraction of the graphic space.

### 2025-03-13 paper/main.tex
small clean ups/restors

--- 2025-01-06 paper/main.tex
@@
 \begin{abstract}
-The abstract goes here.
+  The contract data visualization tools make with their users is that a chart is a faithful and accurate visual representation of the numbers it is made from.
+Motivated by wanting to make better tools, we propose a methodology for fully specifying arbitrary data to visualization mappings in a manner that easily translates to code.
+We propose that fiber bundles provide a uniform interface for describing a variety of underlying data - tables, images, networks, etc.
+- in a manner that independently encodes the mathematical structure of the topology and the fields of the dataset.
+Modeling the data structures that store the datasets as sheaves provide a method for specifying visualization methods that are designed to work regardless of how the dataset is stored - whether the data is on disk, distributed, or on demand.
+Specifying the visualization library components as natural transforms of sheaves means that the constraints that the component must satisfy to be structure preserving can be specified as the set of morphisms on the data and graphic sheaves, including the structure on the topology and fields of the data.
+  Using category theory to formally express how visual elements are constructed means we can translate those expectations into code, which can then be used to enforce the expectation that a visualization tool is faithfully translating between numbers and charts.
 \end{abstract}
@@
 \begin{definition}\label{def:atct:presheaf}
-  A \textbf{presheaf} $F:\mathcal{C}^{op} \rightarrow \setb$ is a contravariant functor from an object in an arbitrary category to an object in the category \setb\cite{nlab:presheaf, spanier1989algebraic}.
+  A \textbf{presheaf} $F:\mathcal{C}^{op} \rightarrow \setb$ is a contravariant functor from an object in an arbitrary category to an object in the category \setb\cite{spanier1989algebraic}.
 \end{definition}
+
+A functor is contravariant when the morphisms between the input objects go in the opposite direction from the morphisms between the output objects.
+The presheaf is contravariant when for every arbitrary morphism between input base spaces $\dfunch: \openset_1 \rightarrow \openset_2$ there exists a corresponding pullback function between the sets of sections $\dfuncpull: \cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}} \rightarrow \cgamma{\opensetc_1}{\dtotalc\restriction_{\openset_c1}}$.
+
+%% these two paragraphs cen be condensed and moved below the figure as an example
 A functor is contravariant when the morphisms between the input objects go in the opposite direction from the morphisms between the output objects.
@@
 
+\begin{equation*}
+  \begin{tikzcd}
+    & {} \arrow[dd, "\texttt{artist}" description, Rightarrow, shift left=3.65, shorten <= .7em, shorten >= .7em, color=artist, font=\Huge]
+    & \\ {\scriptstyle\textcolor{base}{\texttt{topology}}}
+      \arrow[rr, "\texttt{graphic}"', bend right, color=sheaf, font=\huge] \arrow[rr, "\texttt{data}", bend left, color=sheaf, font=\huge] &
+    & {\scriptstyle\textcolor{base}{\texttt{topology}}\textcolor{section}{\rightarrow} \textcolor{fiber}{\texttt{fields}}} \\
+    & {}                                              &
+\end{tikzcd}
+\end{equation*}
+
+%move this down to what is math section
 \begin{align}
@@
 
-While Artists are defined as maps going from data over \dbase\ to graphic over \gbase\, they are compatible with maps defined solely over \dbase\ or \gbase\.
+While Artists are defined as maps going from data over \dbase\ to graphic over \gbase\, they are compatible with maps defined solely over \dbase\ or \gbase.
 This is because when a sheaf is equipped with transport functors (\vindexpush, \vindexpull), maps between sheaves over one space $Hom _{\sheaf_{\dbase}}$ are isomorphic to maps between sheaves over the other space $Hom _{\sheaf_{\gbase}}$\cite{harder2008lectures}:
@@
 We propose an addition operator that states that an artist that takes in a dataset can be constructed using artists that take as inputs subsets of the dataset
+\note{adjoint equation is wrong, probably multiplication b/c using $k^a$ and $K^b$ to define the piecewise, disjoint union is coproduct}
 

### 2025-03-17 short_paper/main.tex
short paper

\documentclass[11pt,titlepage]{amsart}
\usepackage{xcolor}
\usepackage{multicol}
\usepackage{notation}
\usepackage{wrapfig}
\usepackage{multirow}
\usepackage{hyperref}
\usepackage{caption}
\usepackage[anonymous]{atmcs} % for anonymous submission
% \usepackage{atmcs} % for camera-ready
\usepackage{lipsum}
\graphicspath{{C:/Users/story/Projects/team/paper/figures/}}

\begin{document}

\title[]{Topological Equivariant Artist Model}

\author{Hannah~Aizenman}
\address{The Graduate Center, CUNY \\NY, NY, USA}
\email{haizenman@gradcenter.cuny.edu}

\author{Mikael Vejdemo-Johansson}
\address{The Graduate Center, CUNY\\NY, NY, USA}
\email{mvj@math.csi.cuny.edu}

\author{Thomas Caswell}
\address{National Synchrotron Light Source II, Brookhaven National Lab\\ Upton, NY, USA}
\email{tcaswell@bnl.gov}
\author{Michael Grossberg}
\address{The Graduate Center, CUNY\\NY, NY, USA}
\email{mgrossberg@ccny.cuny.edu}

%\keywords{TODO:list keywords}

\begin{abstract}
    The contract data visualization tools make with their users is that a chart is a faithful and accurate visual representation of the numbers it is made from. Motivated by wanting to make better tools, we propose a methodology for fully specifying arbitrary data to visualization mappings in a manner that easily translates to code. We propose that fiber bundles provide a uniform interface for describing a variety of underlying data - tables, images, networks, etc. - in a manner that independently encodes the mathematical structure of the topology and the fields of the dataset. Modeling the data structures that store the datasets as sheaves provide a method for specifying visualization methods that are designed to work regardless of how the dataset is stored - whether the data is on disk, distributed, or on demand. Specifying the visualization library components as natural transforms of sheaves means that the constraints that the component must satisfy to be structure preserving can be specified as the set of morphisms on the data and graphic sheaves, including the structure on the topology and fields of the data.
    Using category theory to formally express how visual elements are constructed means we can translate those expectations into code, which can then be used to enforce the expectation that a visualization tool is faithfully translating between numbers and charts.
\end{abstract}

\maketitle

\section{Intro}
\label{sec:intro}

\begin{wrapfigure}{r}{3.5in}
\begin{center}
    \includegraphics[width=3.5in]{k_different_types.pdf}
\end{center}
\caption{This weather station data has multiple embedded continuities - points at each time and position, timeseries at each position, and maps at each time - that are expected to be homeomorphic to the continuity of their respective visualizations.}
    \label{fig:homeomorphism}
\end{wrapfigure}
Motivated by wanting to make better tools, we propose a methodology for fully specifying arbitrary data to visualization mappings in a manner that easily translates to code. Generally, preserving structure means that a visualization is expected to preserve the $field$ properties and $topology$ of the corresponding dataset, where \textcolor{fiber}{\textbf{field}} is a set of values of the same type and \textcolor{base}{\textbf{topology}} is the connectivity and relative positioning of elements in a dataset \cite{wilkinsonGrammarGraphics2005}.  Field structure is traditionally codified as the Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946}, where each scale is a set of actions on a group Topological structure is generally assumed by visual algorithms \cite{chiTaxonomyVisualizationTechniques2000, toryRethinkingVisualizationHighlevel2004}, but generally do not verify that input structure.
For example, a \texttt{line} algorithm often does not have a way to query whether a list of (x,y) coordinates is the distinct rows, the time series, or the list of stations in \autoref{fig:homeomorphism}. We propose that the bar plot, line plot, and heatmap can be verified as structure preserving because they have a homeomorphic relationship to the 0D ($\bullet$) points, 1D (--) linear, and 2D ($\blacksquare$) surface continuities embedded in the continuous 3 dimensional. Our methodology generalizes Bertin\cite{bertinSemiologyGraphicsDiagrams2011}'s codification of structure,  Mackinlay\cite{mackinlayAutomaticDesignGraphical1987} mathematical formalization, Wilkinson's set theoretic approach \cite{wilkinsonGrammarGraphics2005}, and Kindlemann and Scheidegger's algebraic visualization design (AVD) framework \cite{kindlmannAlgebraicProcessVisualization2014} by explicitly incorporating topology, supporting non-group and non-monoidal structures, and by providing a framework for translating the theoretical ideas into buildable components.

\section{Data Model}
\label{sec:data models}
We extend Butler's proposal of using fiber bundles as an abstraction for visualization data\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989} by incorporating Spivak's methodology for encoding named data types from his fiber bundle representation of relational databases \cite{spivakDatabasesAreCategories2010,spivakSimplicialDatabases2009}.

\begin{multicols}{2}
    \includegraphics[width=.5\textwidth]{fb_rho.pdf}
    \columnbreak
    \captionof{figure}{The scatter marker is specified by the section $\gsection$, which maps into the fiber $\gfiber$ to retrieve the values that compose into the pixel (approximated as a square) returned by the section function evaluated at each point $\gbasepoint$. The section evaluated over the entire space $\gsection\vert_{S}$ returns the entire scatter mark, shown here in faded form to make it easier to see the individual pixels.}\label{fig:fiber_bundle}

\end{multicols}

\begin{table}[h]
    \begin{tabular}{|r | c c c|}
    \hline
     structure/abstraction & Data & Visual & Graphic\\
    \hline
\textcolor{fiber}{fiber}/\textcolor{total}{total}\ /\textcolor{base}{base} spaces &   \multirow{2}{*}{$\dfiberc \hookrightarrow \dtotalc \xrightarrow{\pi} \dbasec$} &   \multirow{2}{*}{$\vfiberc \hookrightarrow \vtotalc \xrightarrow{\pi} \dbasec$} &  \multirow{2}{*}{$\gfiberc \hookrightarrow \gtotalc \xrightarrow{\pi} \gbasec$}\\
fields/datasets/topology&&&\\
\hline
\textcolor{base}{point}/\textcolor{base}{openset}/\textcolor{base}{base space} &  \multirow{2}{*}{$\dbasepointc \in \opensetc \subseteq \dbasec$} & \multirow{2}{*}{$\dbasepointc \in \opensetc \subseteq \dbasec$} & \multirow{2}{*}{$\gbasepointc \in \opensetgc \subseteq \gbasec$}\\
index/subset/indices & & & \\
\hline
\textcolor{fiber}{fiber space}  record/fields &  $\delementc \in \dfiberc$ & $\velementc \in \vfiberc$ & $\gelementc \in \gfiberc$ \\
\hline
\textcolor{section}{section} dataset & $\cgamma{\dbasec}{\dtotalc} \ni \dsectionc: \dbase \textcolor{section}{\rightarrow} \dfiberc$ & $\cgamma{\dbasec}{\vtotalc} \ni \vsectionc: \dbase \textcolor{section}{\rightarrow} \vfiberc$ &  $\cgamma{\gbasec}{\gtotalc} \ni \gsectionc: \gbase \textcolor{section}{\rightarrow} \gfiberc$ \\
\hline
\textcolor{sheaf}{sheaf} data container  &  \multirow{2}{*}{$\sheafc_{\dbasec, \dtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$} &  \multirow{2}{*}{$\sheafc_{\dbasec, \vtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}}$} &   \multirow{2}{*}{$\sheafc_{\gbasec, \gtotalc}: \opensetgc \rightarrow \cgamma{\opensetc}{\gtotalc\restriction_{\opensetgc}}$} \\
&&&\\
        \hline
    \end{tabular}
    \caption{Summary of how the data is abstracted using topological structures}
    \label{tab:data_abstraction}
\end{table}

The indexing space is modeled as a topology $\mathcal{T}$ on the underlying data indexing space because it is a uniform method of describing arbitrary continuity and it provides a method for managing overlapping embedded topologies, such as \autoref{fig:homeomorphism}. Adopting Spivak's fiber bundle construction of types allows our model to reuse types so long as the field names are distinct and encode multi-dimensional fields can be encoded in the same manner as single dimensional fields because domains of types are a mathematical space. We encode data as a \textcolor{section}{section} \dsectionc\ of a bundle because this allows us to incorporate the topology and field types in the data definition $\textcolor{section}{\texttt{dataset}}: \textcolor{base}{\texttt{topology}} \rightarrow \textcolor{fiber}{\texttt{field}}$. Abstracting the data containers as sheaves provides a way of encoding how data containers keep track of the continuity of the data \cite{ghristElementaryAppliedTopology2014} such that visual algorithms developed on this model apply whether the data fits in memory, is distributed, or is streaming.

\section{Artist}
\label{sec:Artist}
We call the mapping from data to graphic $\vartistc:\dsectionc\mapsto\gsectionc$ an \textcolor{artist}{Artist}\footnote{Artists are named for the Artist class in Matplotlib. Artist is the parent class of all the objects that assemble and manage visual elements such as points, lines, and shapes.} and propose that an artist is structure preserving when the artist is implemented as a natural transform on sheaves. This means that the artist satisfies the following constraints, as illustrated in \autoref{fig:artist}:

\begin{multicols}{2}
    \includegraphics[width=.5\textwidth]{artist_equiv.png}
    \captionof{figure}{}\label{fig:artist}
\columnbreak

\begin{description}
    \item[homeomorphic] there exists a deformation retract $\vindexc: \gbasec\rightarrow\dbasec$ between graphic and artist base space
    \item[topology] the artist is commutative w.r.t. the presheaf morphism $\dfunchc: \dbasepointc^{\prime}\mapsto \dbasepointc$ and pullback $\dfuncpullc \dsectionc \restriction_{\opensetc^{\prime}}: \dsectionc\restriction_{\opensetc^{\prime}} \mapsto \dsectionc \restriction_{\opensetc^{\prime} \circ \dfunchc}$ such that when an index of the record changes the values in the record are unmodified  $\dsectionc(\dbasec) = \dfuncpullc\dsectionc(\dbasepointc^{\prime}) = \dsectionc(\dfunchc(\dbasepointc^{\prime}))$
    \item[fields] the artist is commutative w.r.t. actions on the fiber $\dfunctc \in Hom(\dfiberc, \dfiberc)$, which are the functions between sections $\dfunctc: \dfuncpullc \dsectionc \restriction_{\opensetc^{\prime}} \mapsto \dfuncpullc \dsectionc^{\prime} \restriction_{\opensetc}$. Additionally, $\dfunctc$ must preserve features of $\dfiber$, such as operators, that are defined as part of the structure of $\dfiber$.

\end{description}
\end{multicols}

\begin{figure}[h]
    \includegraphics[width=7.16in]{xi_diagram.pdf}
    \caption{The graphics over the orange bands map to the data at the orange point.}
    \label{fig:xi}
\end{figure}
As illustrated in \ref{fig:xi}, the transport functors $(\vindexpush, \vindexpull)$ constructed from $\vindex$ \cite{harder2008lectures} also codify how visual elements correspond to distinct data elements, which is a necessary condition of a visualization being readable\cite{ziemkiewiczEmbeddingInformationVisualization2009}. The pushforward $\vindexpushc$ matches each point in the data space to the specification of the graphic at that point $\vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexprec(\dbasepointc)}$, while the pullback $\vindexpull$ matches each point in the graphic space to the data over that point $\vindexpullc\dsectionc(\gbasepointc) = \dsectionc(\vindexc(\gbasepointc)) = \dsectionc(\dbasepointc)$.

\begin{figure*}[h]
\includegraphics[width=\textwidth]{tex/more_records.png}
\end{figure*}

The topology and fiber categories are equipped with operators to add and extend records, respectively, Since the artist preserves these operators, artists have the property of being composable $\vartistc(\bigsqcup\limits_{i}\dbasec_i, \prod\limits_j\dfiberc_j) = \bigsqcup\limits_i\prod\limits_j\vartistc(\dbasec_i, \dfiberc_j)$

\section{Construction}
The artist can be constructed with an encoding stage $\vchannelc$ and a compositing stage $\vmarkc$ s.t $\vartistc=\vmarkc\circ\vchannelc$. In the \textcolor{artist}{encoding} stage $\vchannelc$, a data bundle is mapped to a visual variable bundle \vchannel, where the fibers of \vchannel\ are an internal library specification. The \textcolor{artist}{compositing} stage $\vmarkc$ is then responsible for assembling  the visual components into a visual element.

\section*{Note}
This is a work in progress, more information can be found at https://github.com/story645/team

\section*{Acknowledgements}
The authors would like to thank the anonymous reviewers who gave constructive feedback on an earlier version of this paper. The authors are also grateful to  the various Matplotlib and Napari contributors, particularly Juan Nunez-Iglesias, and Nicolas Kruchten for their valuable feedback from the library developer perspective. Hannah is also very grateful to Nicolas for the suggestion of augmented notation and to the nlab and wikipedia contributors who wrote clear explanations of many of the topics discussed in this paper. This project has been made possible in part by grant number 2019-207333 (EOSS Cycle 1) and EOSS Cycle 3 Chan Zuckerberg Initiative DAF, an advised fund of Silicon Valley Community Foundation

\bibliographystyle{plain}
\bibliography{references}
\end{document}

### 2025-03-27 slides/main.tex
talk
composition overhaul
equivariance + brought up related work + comments
talk version

--- 2022-03-29 slides/main.tex
@@
 \documentclass[xcolor={dvipsnames}]{beamer}
-\setbeamertemplate{footline}[frame number]
-\setbeamersize{description width=-\labelsep}
+
+%\setbeamertemplate{footline}[frame number]
+%\setbeamersize{description width=-\labelsep}
+\usetheme{SimpleDarkBlue}
 
@@
 
-\graphicspath{.figures/}
+\graphicspath{.figures/, ../papers/figures/}
 
 \usepackage[backend=bibtex, sorting=none, doi=false,isbn=false,url=false, giveninits=true]{biblatex}
-\bibliographystyle{plain}
-\bibliography{../paper/bibliography.bib}
 
-%\usetheme{ccnycrest}
+\bibliography{references}
 
@@
 
+\title{Topologically Equivariant \\ Artist Model}
+\author{Hannah Aizenman\texorpdfstring{\\}{,}  Mikael Vejdemo-Johansson \texorpdfstring{\\}{,} Thomas Caswell \texorpdfstring{\\}{,} Michael Grossberg}
+\institute{Department of Computer Science, The Graduate Center}
+\date{\today}
+
 \begin{document}
-\begin{frame}<presentation:1-|handout:0>{About}
-    Slides: https://bit.ly/team329
-    \begin{block}{Me (twitter/github: story645)}
+
+\begin{frame}<presentation:1|handout:1>
+	\titlepage
+    %Hannah Aizenman, Mikael Vejdemo-Johansson, Thomas Caswell, Michael Grossberg\\
+    %Committee: Dr. Robert Haralick, Dr. Lev Manovich, Dr. Huy Vo\\
+    %External Member: Dr. Marcus Hanwell
+\end{frame}
+
+\begin{frame}{About}
+    \begin{block}{Me}
         \begin{itemize}
-            \item nth year grad student (on my 3rd EO)
-            \item former adjunct at CCNY, former Digital Fellow
+            \item PhD candidate in Computer Science
             \item Matplotlib Community Manager \& Core Developer
+            \item  socials/github: story645
         \end{itemize}
@@
 
-    \begin{block}<presentation:2|handout:0>{Project}
+    \begin{block}{Project}
         \begin{itemize}
             \item Funded by Chan Zuckerberg Initiative EOSS 1 \& 3
-            \item paper rejected by vizweek last spring
-            \item work has since gone all in on category theory 
         \end{itemize}
@@
 
-\begin{frame}<presentation:1|handout:1>
-	\titlepage
-    Hannah Aizenman, Tom Caswell, Michael Grossberg\\
-    %Committee: Dr. Robert Haralick, Dr. Lev Manovich, Dr. Huy Vo\\
-    %External Member: Dr. Marcus Hanwell
-\end{frame}
-
 \section{Introduction}
-\begin{frame}<presentation:0|handout:0>{What are we doing?}
+\begin{frame}{Goals}
     \begin{itemize}
-        \item develop a model for describing data to graphic transformations
-        \item specify a visualization library architecture based on this model
-        \item implement functional(ish) components based on this model using ideas from functional programming
+        \item uniform API for arbitrary datasets in arbitrary data containers
+        \item generalizable methodology for expressing structure preservation
+        \item framework that translates into coding specifications
     \end{itemize}
@@
 %% maybe pull out the middle for now? move the 3 stages down to construction?
-\begin{frame}<presentation:1-|handout:0>{What do visualization libraries do?}
+\begin{frame}{What do visualization libraries do?}
     \begin{figure}
@@
 
-\begin{frame}<presentation:1|handout:0>{Matplotlib}
+\begin{frame}{Matplotlib}
     \begin{figure}
@@
 
-\begin{frame}<presentation:1-|handout:0>
-    \frametitle<presentation:1-2|handout:0>{Everything is an Artist}
-    \frametitle<presentation:3|handout:0>{Transformations Change Artists into Different Coordinates}
-    \begin{overprint}
-        \onslide<1|handout:0>\includegraphics[scale=.25]{figures/math/sphx_glr_anatomy_001_2_0x.png}
-        \onslide<2|handout:0>\includegraphics[width=\textwidth]{figures/math/artist_tree.png}
-        \onslide<3|handout:0>\includegraphics[width=\textwidth]{figures/math/transforms.png}
-    \end{overprint}
+\begin{frame}{How do we express structure?}
+
+    \begin{description}
+        \item[\textcolor{fiber}{\textbf{field}}]is a set of values of the same type, e.g. one column of a table or the pixels of an image
+        \item[\textcolor{base}{\textbf{topology}}] is the connectivity and relative positioning of elements in a dataset \cite{wilkinsonGrammarGraphics2005}.
+    \end{description}
 \end{frame}
 
-\begin{frame}[fragile]{A Simple Artist}
-\begin{minted}{python}
-    class SomeArtist(Artist):
-        'An example Artist that implements the draw method'
+\begin{frame}{Structure Preservation: Equivalent scales and transforms}
 
-        def draw(self, renderer):
-            # create some objects and use renderer to draw self here
-            renderer.draw_path(graphics_context, path, transform)
-\end{minted}
+\begin{definition}\label{def:related-work:action}\cite{grimaldiDiscreteCombinatorialMathematics2006}
+    An \textcolor{action}{\textbf{action}} of \textcolor{action}{$G = (G,\circ, e)$} on $X$ is a function  $act: \textcolor{action}{G} \times X \rightarrow X$.
+An action has the properties of identity $act(\textcolor{action}{e}, x) = x$ for all  $x \in X$ and associativity $act(\textcolor{action}{g}, act(\textcolor{action}{f}, x)) = act(\textcolor{action}{f} \circ \textcolor{action}{g}, x)$ for $\textcolor{action}{f},\textcolor{action}{g} \in \textcolor{action}{G}$.
+\end{definition}
 \end{frame}
 
-\begin{frame}<presentation:1-|handout:0>{Goals}
+\begin{frame}{Fields: Equivariance }
+
+    Given a group $G$ that acts on both the input $X$ and the output $Y$ of a function $f: X \rightarrow Y$
+    \begin{definition}\label{def:equivariance}
+     A function $f$ is \textbf{equivariant} when $f(act(g,x)) = act(g,f(x))$ for all $g$ in $G$ and for all $x$ in $X$ \cite{pittsNominalSetsNames2013}
+    \end{definition}
+\end{frame}
+\begin{frame}{Fields: Equivariance}
     \begin{figure}
-        \includegraphics[width=\linewidth]{figures/flow/artists.png}
+        \includegraphics[width=\linewidth]{../paper/figures/equivariant.pdf}
+        \label{fig:related-work:equivariance}
       \end{figure}
@@
 
-\begin{frame}<presentation:1|handout:0>{How do we express structure?} 
+\begin{frame}{Fields: Homomorphism}
+    Given the function $f: X \rightarrow Y$, with operators $(X, \circ)$ and $(Y, *)$
+
+    \begin{definition}\label{def:homomorphism}
+      A function $f$ is \textbf{homomorphic} when $f(x_1 \circ x_2) = f(x_1) * f(x_2)$ and preserves identities $f(I_x) = I_y$ all $x, y \in X$ \cite{grimaldiDiscreteCombinatorialMathematics2006}
+    \end{definition}
+\end{frame}
+
+\begin{frame}{Fields: Homomorphism}
+
     \begin{figure}
-        \includegraphics[width=1\textwidth]{figures/intro/munzner_datatypes.png}
-        \caption{Figure 2.8 in Munzner's Visualization Analysis and Design\cite{munznerVisualizationAnalysisDesign2014}}
+        \includegraphics[width=\linewidth]{../paper/figures/homomorphism.pdf}
+        \label{fig:related-work:homomorphism}
       \end{figure}
@@
 
-\begin{frame}<presentation:1|handout:0>{Continuity}
+\begin{frame}{Topology: Homeomorphism}
+
+    \begin{definition}
+        A function $f$ is a $homeomorphism$ if it is bijective, continuous, and has a continuous inverse function $f^{-1}$.
+      \end{definition}
+\end{frame}
+
+\begin{frame}{Homeomorphic topology}
     \begin{figure}
-        \includegraphics[width=\linewidth]{../paper/figures/k_different_types.png}
+        \includegraphics[width=\linewidth]{../paper/figures/k_different_types.pdf}
     \end{figure}
-    \begin{block}{Topological Properties \cite{wilkinsonGrammarGraphics2005}}
-        how elements in a dataset are organized, e.g. discrete rows in a table, networked nodes, pixels in an image, points on a line
+\end{frame}
+
+\begin{frame}{Structure Preservation}
+    \begin{itemize}
+        \item Bertin: equivalent data and visual field properties \cite{bertinSemiologyGraphicsDiagrams2011}
+        \item Mackinlay: homomorphic fields and equivalent topology \cite{mackinlayAutomaticDesignGraphical1987}
+        \item Wilkinson: homomorphic scales and  equivalent topology \cite{wilkinsonGrammarGraphics2005}
+        \item Kindleman  \& Schieidegger: equivariant scales, invariant data representation, equivalent topology \cite{kindlmannAlgebraicProcessVisualization2014}
+    \end{itemize}
+    \begin{block}{contribution}
+    explicit homeomorphic topology, equivariant actions on fields and topology
     \end{block}
@@
 
-\begin{frame}<presentation:1|handout:0>{Visual Algorithms \& Continuity}
-    \begin{figure}
-        \includegraphics[width=1\linewidth]{figures/math/plottypes.png}
-    \end{figure}
-\end{frame}
-
-\begin{frame}<presentation:1-|handout:0>{Equivariance}
-    \begin{description}
-        \item[What] Retinal Variables \& Marks: visual encodings should match properties of the data \cite{bertinSemiologyGraphicsDiagrams2011a}
-        \pause
-        \item[Why] Graphical Integrity: graphs show \textbf{only} the data\cite{tufteVisualDisplayQuantitative2001}
-        \pause 
-        \item[Why] Naturalness: easier to understand when properties match\cite{norman_things_smart}
-        \pause
-        \item[How] Expressiveness: which structure preserving mappings can a tool implement\cite{mackinlayAutomatingDesignGraphical1986}] 
-    \end{description}
-\end{frame}
-
-%flip columns
-\begin{frame}<presentation:0-|handout:0>{Theoretical Frameworks}
-    \begin{description}
-        \item[linguistically] visualization has syntax, semantics, grammar expresses how to design structure preserving visualizations
-        \cite{mackinlayAutomatingDesignGraphical1986,mackinlayAUTOMATICDESIGNGRAPHICAL1987,wilkinsonGrammarGraphics2005}
-
-        \item[algebraically] transformations on data and graphics are equivalent and symmetric \cite{kindlmannAlgebraicProcessVisualization2014}
-        \begin{columns}
-            \column{.5\textwidth}
-            \begin{description}
-                \item[D] raw data 
-                \item[R] representation (data container) 
-                \item[V] visualization
-            \end{description}
-            \column{.5\textwidth}
-            \begin{equation*}
-                \begin{tikzcd}[ampersand replacement=\&]
-                    D \arrow[d, "\alpha"'] \arrow[r, "r_1"] \& R \arrow[r, "\nu"]  \& V \arrow[d, "\omega"] \\
-                    D \arrow[r, "r_2"']                     \& R \arrow[r, "\nu"'] \& V                    
-                \end{tikzcd}
-            \end{equation*}
-        \end{columns} 
-    %\item[categorically] \textit{understanding} = \textit{read} $\circ$ \textit{render} \cite{vickersUnderstandingVisualizationFormal2013}    
-    \end{description}
-\end{frame}
-
-\begin{frame}<presentation:1-|handout:1>
-    \frametitle{Domain Specific Library: library assumes structure \cite{HeerSoftware2006}}
+\begin{frame}{Domain Specific Library: library assumes structure \cite{HeerSoftware2006}}
     \begin{table}
@@
             \includegraphics[width=.24\textwidth]{figures/intro/table.png} & \includegraphics[width=.3\textwidth]{figures/intro/landsat.png} & \includegraphics[width=.33\textwidth]{figures/math/graph.png} \\
-            ggplot\cite{wickhamGgplot2ElegantGraphics2016a}  & ImageJ\cite{schneiderNIHImageImageJ2012}& Gephi\cite{bastianGephiOpenSource2009}\\
+            ggplot\cite{wickhamGgplot2ElegantGraphics2016}  & ImageJ\cite{schneiderNIHImageImageJ2012}& Gephi\cite{bastianGephiOpenSource2009}\\
             Vega\cite{satyanarayanDeclarativeInteractionDesign2014} & ImagePlot\cite{studiesCulturevisImageplot2021} & Graphviz\cite{ellsonGraphvizOpenSource2002}\\
@@
 
-\begin{frame}<presentation:1-|handout:1>
+\begin{frame}
     \frametitle{Building Block Library\cite{wongsuphasawatNavigatingWideWorld2021}: visual algorithms assume structure \cite{toryRethinkingVisualizationHighlevel2004}}
-    \begin{figure}
-        \includegraphics[height=.3\textheight]{../paper/figures/k_different_types.png}
-    \end{figure}
 
     \begin{enumerate}
-        \item Matplotlib\cite{hunterMatplotlib2DGraphics2007} $\rightarrow$ Seaborn\cite{waskom2020seaborn}, xarray \cite{hoyer2017xarray} 
+        \item Matplotlib\cite{hunterMatplotlib2DGraphics2007} $\rightarrow$ Seaborn\cite{Waskom2021}, xarray \cite{hoyer2017xarray}
         \item D3 \cite{bostockDataDrivenDocuments2011}
@@
 
-\section{Algebraic Topology & Category Theory}
+\section{Algebraic Topology \& Category Theory}
 \begin{frame}<presentation:1|handout:1>{Design Composable Structure Preserving API}
     \begin{description}
-        \item[Fiber Bundles] "unified, dimension-independent framework" that expresses data as the mapping between continuity and fields \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}
+        \item[Fiber Bundles] Butler: "unified, dimension-independent framework" that expresses data as the mapping between continuity and fields \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}
+        \item[Simplicial Databases] Spivak: Add rich typing for fields to Bulter \cite{spivakSimplicialDatabases2009}
         \item[Category Theory Language] express constraints in specifications \cite{wielsManagementEvolvingSpecifications1998}
-        \item[Sheaves on Bundles] "algebraic data structure" for representing data over topological spaces \cite{ghristElementaryAppliedTopology2014}
+        \item[Sheaves] Ghrist "algebraic data structure" for representing data over topological spaces \cite{ghristElementaryAppliedTopology2014}
     \end{description}
@@
 
-\subsection{Fiber Bundles}
-\begin{frame}<presentation:1-|handout:1>
-    \frametitle<1|handout:1>{Fiber Bundle}
-    \frametitle<6|handout:2-3>{Data Bundle}
-    \frametitle<7|handout:4-5>{Graphic Bundle}
-    \begin{columns}
-        \column{0.31\textwidth}
+\subsection{Uniform Abstraction for Data \& Graphics}
+
+\begin{frame}<presentation:1|handout:1>{Expressive Types}
     \begin{equation*}
-            %\label{eq:related-work:continuity:fiber-bundle}
+        \textcolor{section}{\texttt{dataset}}: \textcolor{base}{\texttt{topology}} \rightarrow \textcolor{fiber}{\texttt{fields}}
+      \end{equation*}
+\end{frame}
+
+\subsubsection{Abstract data representation}
+\begin{frame}{Fiber Bundle}
+
+    \begin{definition}\label{def:fiber_bundle}
+       A \textbf{fiber bundle} $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and  bundle projection map $\pi: \dtotalc \rightarrow \dbasec$ \cite{spanier1989algebraic}.
+
+       \begin{equation} \label{eq:atct:fb:intro}
+        \begin{tikzcd}[ampersand replacement=\&]
+          \dfiberc \arrow[r, hook, color=total] \& \dtotalc \arrow[r, "\pi", color=total, two heads] \& \dbasec
+          \end{tikzcd}
+        \end{equation}
+
+    A continuous surjective map $\bm{\pi}$ is a \textbf{bundle projection} map when
+    \begin{enumerate}
+      \item all fibers in the bundle are isomorphic.
+Since all fibers are isomorphic $\dfiber \cong \dfiber_{\dbasepoint}$ for all points $\dbasepoint \in \dbase$, there is a uniquely determined \textcolor{fiber}{fiber space} \dfiberc\ given by the preimage of the projection $\pi$ at any point $\dbasepoint$ in the \textcolor{base}{base space} \dbasec: $\dfiberc = \pi^{-1}(k)$.
+      \item each point $\dbasepointc$ in the \textcolor{base}{base space} \dbasec\ has an open neighborhood $\openset_{\dbasepointc}$ such that the \textcolor{total}{total space} \dtotalc\ over the neighborhood is locally trivial.
+    \end{enumerate}
+    \end{definition}
+
+    \textbf{Local triviality} means $\dtotal\vert_{\openset} = \openset\times \dfiber$.
+\end{frame}
+
+\begin{frame}{Fiber Bundle}
+    \begin{figure}
+        \includegraphics[width=\linewidth]{../paper/figures/transition_maps.pdf}
+    \end{figure}
+\end{frame}
+
+\begin{frame}{Fiber Bundle: section}
+
+    \begin{minipage}{.5\columnwidth}
+        \begin{definition}\label{def:fiber_bundle:section}
+        A \textcolor{section}{\textbf{section}} $\dsectionc: \dbasec \rightarrow \dtotalc$ over a fiber bundle is a smooth right inverse of $\pi(\dsection(\dbasepoint)) = \dbasepoint$ for all $\dbasepoint \in \dbase$
+        \end{definition}
+        \end{minipage}
+        \begin{minipage}{.4\columnwidth}
+          \begin{equation} \label{eq:atct:fb:intro-sec}
             \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-              \onslide*<3-6|handout:1-2>{\dfiberc}
-              \onslide*<7|handout:4>{\gfiberc} 
-              \onslide*<4-|handout:1-2,4>{\arrow[r, hook]} \& 
-              \onslide*<1-6|handout:1-2>{\dtotalc} 
-              \onslide*<7|handout:4>{\gtotalc}
-              \onslide*<4-|handout:1-2,4>{\arrow[d, "\pi"']} \\
+             \dfiberc
+              \arrow[r, hook, color=total] \&
+              \dtotalc
+              \arrow[d, "\pi"',color=total, two heads] \\
                \&
-               \onslide*<2-6|handout:1-2>{\dbasec}
-               \onslide*<7|handout:4>{\gbasec} 
-               \onslide*<5-6|handout:1-2>{\arrow[u, "\dsectionc"', bend right, pos=.5]}
-               \onslide*<7|handout:4>{\arrow[u, "\gsectionc"', bend right, pos=.5]}
+            \dbasec
+               \arrow[u, "\dsectionc"', bend right, pos=.5, color=section, dashed]
+            \end{tikzcd}
+          \end{equation}
+        \end{minipage}
+\end{frame}
+
+\begin{frame}{Fiber Bundle: section}
+    \begin{figure}
+        \includegraphics[width=\textwidth]{../paper/figures/fb.pdf}
+    \end{figure}
+\end{frame}
+
+\subsubsection{Topology: base Space \dbasec}
+
+\begin{frame}{Base space: topology}
+The base space of a fiber bundle is a quotient topology\cite{munkresElementsAlgebraicTopology1984}.
+Given a set $X$ and a function $\mathcal{N}:X\to 2^{2^X}$ that assigns to any $x\in X$ a non-empty collection of subsets $\mathcal{N}(x)$, where each element of $\mathcal{N}(x)$ is a \emph{neighborhood of $x$}, then $X$ with  $\mathcal{N}$ is a \textcolor{base}{topological space} and $\mathcal{N}$ is a neighborhood \emph{topology} if for each $x$ in $X$: \cite{brownronaldTopologyGroupoids2006}
+
+\begin{definition}\label{def:topology}
+\begin{enumerate}
+  \item if $N$ is a neighborhood $N \in \mathcal{N}(x)$ of $x$ then $x \in N$
+  \item every superset of a neighborhood of $x$ is a neighborhood of $x$; therefore a union of a neighborhood and adjacent points in $X$ is also a neighborhood of $x$
+  \item the intersection of any two neighborhoods of $x$ is a neighborhood of $x$
+  \item any neighborhood $N$ of $x$ contains a neighborhood $M \subset N$ of $x$ such that $N$ is a neighborhood of each of the points in $M$
+\end{enumerate}
+\end{definition}
+\end{frame}
+
+\begin{frame}{Why neighborhood topology?}
+\begin{figure}
+\includegraphics[width=\linewidth]{../paper/figures/k_different_types.pdf}
+\end{figure}
+\end{frame}
+
+\begin{frame}{Base \texttt{TYPE} \dbasec}
+    \begin{definition}\label{def:atct:category}
+        An \textbf{category} $\mathcal{C}$ consists of the following \textit{data}:
+     \begin{enumerate}
+       \item a collection of \textit{objects} $X \in \textbf{ob}(\mathcal{C})$
+       \item for every pair of objects $X, Y \in \textbf{ob}(\mathcal{C})$, a set of \textit{morphisms} $X \xrightarrow{f} Y \in Hom_{\mathcal{C}}(X, Y)$
+       \item for every object $X$, a distinct \textit{identity morphism} $X \xrightarrow {id_x} X$ in $Hom_{\mathcal{C}}(X, X)$
+       \item a \textit{composition function} $f \in Hom_{\mathcal{C}}(X, Y) \times  g \in Hom_{\mathcal{C}}(Y, Z) \rightarrow g \circ f \in Hom_{\mathcal{C}}(X, Z)$
+     \end{enumerate}
+     such that
+     \begin{enumerate}
+       \item \textit{unitality:} for every morphism $ X \xrightarrow{f} Y$, $f \circ id_x = f = id_y \circ f$
+       \item \textit{associativity:} if any three morphisms $f, g, h$ are composable,
+         \begin{equation*}
+           \begin{tikzcd}[ampersand replacement=\&]
+             X \arrow[r, "f"] \arrow[rrr, "h\circ(g\circ f) = (h\circ g)\circ f"', bend right, dashed] \& Y  \arrow[r, "g"] \& Z \arrow[r, "h"] \& W
              \end{tikzcd}
        \end{equation*}
-     \column{0.69\textwidth}
-     \only<1-5|handout:1>{
-        \begin{description}
-            \item<1-5|handout:1>[\textcolor{total}{Total Space}]{$(\dtotalc, \mathcal{T}_{\dtotalc})$}
-            \item<2-5|handout:1>[\textcolor{base}{Base Space}]{$(\dbasec, \mathcal{T}_{\dbasec}),\; \dbasepointc \in \opensetc \subseteq \dbasec$}
-            \item<3-5|handout:1>[\textcolor{fiber}{Fiber Space}]{$\dfiberc\restriction_{\dbasepointc} = \pi^{-1}(\dbasepointc), \dfiberc = \dfiberc\restriction_{\dbasepointc}\; \forall \dbasepointc \in \dbasec$}
-        \end{description}
-        }
-    \only<6|handout:2>{
-        \begin{description}
-            \item[\textcolor{total}{Data \dtotalc}]{continuity + fields}
-            \item[\textcolor{base}{Continuity \dbasec}]{how data elements are organized (topological properties) \cite{wilkinsonGrammarGraphics2005}, index (key) space \cite{munznerWhatDataAbstraction2014})}
-            \item[\textcolor{fiber}{Fields \dfiberc}] {generalization of a schema - named and typed date fields \cite{spivakSIMPLICIALDATABASES, spivakDatabasesAreCategories2010}}
-        \end{description}
-        }
-        \only<7|handout:4>{
-            \begin{description}
-                \item[\textcolor{total}{Graphic \gtotalc}]{continuity + renderer fields}
-                \item[\textcolor{base}{Continuity \gbasec}]{parameterization of graphic area (e.g. "inked" bounding box\cite{CairographicsOrg})}
-                \item[\textcolor{fiber}{Display \gfiberc}] {renderer fields, e.g. \{xy,rgba\}, \{xy, cymk\}, \{xyz, rgba\}}
-            \end{description}
-            }
-    \end{columns}
-    \only<4-5|handout:1>{
-        \begin{alertblock}{}
-            \begin{description}
-                \only<5:handout:1>{\item[\textcolor{section}{Sections}]{$\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}$}}
-            \end{description}
-        \end{alertblock}
-    }
-    \only<4:handout:1>{
-        \begin{alertblock}{}
-    \begin{description}
-    \item[Locally Trivial] {for every point $\dbasepointc \in \dbasec$, there exists an open neighborhood $\dbasepointc \in \opensetc \subseteq \dbasec$ s.t.
-there is a homeomorphism $\pi^{-1}(\opensetc)\xrightarrow{\equivc} \opensetc \times \dfiberc$}
-    \item[(Globally) Trivial]{$\dtotalc = \dbasec \times \dfiberc$}
-\end{description}
-\end{alertblock}
-    }
-    \only<6|handout:2>{
-    \begin{alertblock}{}
-        \begin{description}
-            \item[\textcolor{section}{Data} \cite{butlerVisualizationModelBased1989,butlerVectorBundleClassesForm1992}]{$
-        \dsectionc(\dbasepointc) = \;\delementc,\; \dbasepointc\in \opensetc \subseteq \dbasec,\; \delementc \in \dfiberc\restriction_{\dbasepointc}$} 
-        \item[]{$\delementc=\{field_0:value, \cdots field_i:value, \cdots field_n:value\}$}
-        \end{description}
-    \end{alertblock}
-    }
-    \only<7|handout:4>{
-        \begin{alertblock}{}
-            \begin{description}
-                \item[\textcolor{section}{Graphic}]{$\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \coloneqq \big\{\gsectionc: \opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} \; \bigm{\vert} \pi(\gsectionc(\gbasepointc)) = \gbasepointc\;for\, all\; \gbasepointc \in \opensetgc \big\}$}
-                \item[] {$\gsectionc(\gbasepointc) = \gelementc,\; \gbasepointc\in \opensetgc\subseteq \gbasec,\; \gelementc \in \dfiberc\restriction_{\gbasepointc}$}
-                \item[] $\gelementc=\{x,\,y,\,r,\,g,\,b\}$
-
-        \end{description}    
-    \end{alertblock}
-    }
-    \only<|handout:3>{
-        \begin{figure}
-            \includegraphics[scale=.65]{figures/math/spherebundle.png}
-        \end{figure}
-    }
-    \only<|handout:5>{
-        \begin{figure}
-            \includegraphics[scale=.65]{figures/math/render.png}
-        \end{figure}
-    }
+       then they are associative such that $h\circ(g\circ f) = (h \circ g) \circ f$  \cite{lawvere2009conceptual,riehlCategoryTheoryContext,maclaneCategoriesWorkingMathematician2013,fongInvitationAppliedCategory2019}.
+       \end{enumerate}
+     \end{definition}
 \end{frame}
 
-\subsection{Category Theory Language}
-\begin{frame}<presentation:1|handout:1>[fragile]{}
-    \begin{minted}{python}
-        Artist(Data) -> Graphic
-    \end{minted}
+\begin{frame}{Add records $\rightarrow$ Add to base space}
+    \begin{figure}
+        \includegraphics[width=\textwidth]{../paper/figures/tex/k_coproduct.pdf}
+    \end{figure}
 \end{frame}
 
-\begin{frame}<presentation:1-|handout:1->
-    \frametitle<presentation:1-6|handout:1>{Category \setc}
-    \frametitle<presentation:7|handout:2>{Category \setc^{op}}
-\begin{columns}
-    \column{0.39\textwidth}
+\subsubsection{fields: fiber space \dfiberc}
+
+\begin{frame}{Field types }
+   A \textit{schema} consists of a pair $(\fnames, \sigma)$ where $\fnames$ is the set of field names and $\sigma: \fnames \rightarrow \textbf{DT}$ is a function from field name to field data type\cite{spivakSimplicialDatabases2009}
+
+   The function $\sigma$ is composed with $\pi$ such that $\pi^{-1}(\sigma(C)) \subseteq \mathscr{U}$; this composition induces a domain bundle $\pi_{\sigma}:\mathscr{U}_{\sigma} \rightarrow \fnames$ that associates a field name $c \in C$ with its corresponding domain $\pi^{-1}_{\sigma}(C) \subseteq \mathscr{U}_{\sigma}$.
+
+\end{frame}
+
+\begin{frame}{Record}
+    \begin{definition} A \textbf{record} is a function $\delement: \fnames \rightarrow \mathscr{U}_{\sigma}$ and the set of records on $\pi_{\sigma}$ is denoted $\Gamma^{\pi}(\sigma)$.
+Records must return an object of type $\sigma(\fname) \in \textbf{DT}$ for each field $c \in C$.
+    \end{definition}
+
+    Tables are sections $\dsection: \dbase \rightarrow \Gamma^{\pi}(\sigma)$ from an indexing space $\dbase$ to the set of all possible records $\Gamma^{\pi}(\sigma)$ on the schema bundle
+\end{frame}
+
+\begin{frame}{Fiber \texttt{TYPE} \dfiberc}
+    We define the \textcolor{fiber}{fiber space} \dfiberc\ to be the space of all possible data records
+    \begin{equation}
+      \dfiberc \coloneqq \{\delement: \fnames \rightarrow \mathscr{U}_{\sigma} \bigm{\vert} \pi_{\sigma}(\delement(\fname)) = \fname\;for\;all\; \fname \in \fnames \}
+    \end{equation}
+    such that the preimage of a point is the corresponding data type domain $\pi^{-1}(\dbasepoint) = \dfiber_{k} = \mathscr{U}_{{\sigma}_{\dbasepoint}}$.
+\end{frame}
+
+\begin{frame}{Fiber category \dfiberc}
+    The fiber category has a single object $\dfiberc$ of an arbitrary type and morphisms on the fiber object $\dfunctc \in Hom(\dfiberc, \dfiberc)$.
+
+    The fiber category $\mathcal{\dfiberc}$ is also equipped with a bifunctor  $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$ for combining fiber types.
+\end{frame}
+
+\begin{frame}{Fiber morphism \dfunctc}
+    \begin{table}
+        \begin{tabular}{|lll|}\hline
+            scale & operators & sample constraint \\ \hline
+            nominal & $=,\neq$ &  $\dsectionc(\dbasepointc_1) \neq \dsectionc(\dbasepointc_2)\implies \dfunctc (\dsectionc(\dbasepointc_1)) \neq\dfunctc(\dsectionc(\dbasepointc_2))$\\
+            ordinal & $<, \leq, \geq, >$ &  $\dsectionc(\dbasepointc_1) \leq \dsectionc(\dbasepointc_2) \implies \dfunctc (\dsectionc(\dbasepointc_1)) \leq \dfunctc(\dsectionc(\dbasepointc_2)$) \\
+            interval & $+, -$ &  $\dfunctc(\dsectionc(\dbasepointc) + C) = \dfunctc(\dsectionc(\dbasepointc)) + C$ \\
+            ratio & $*,/$ &  $\dfunctc(\dsectionc(\dbasepointc) * C) = \dfunctc(\dsectionc(\dbasepointc))*C $\\ \hline
+        \end{tabular}
+    \end{table}
+
+\end{frame}
+
+\begin{frame}{Constraints as definition: Dates}
+    \begin{itemize}
+        \item year:  $\{y \in \mathbb{I}\vert 1992 \leq y \leq 2025\}$
+        \item month:  $\{m \in \mathbb{I}\vert 1 \leq m \leq 12 \}$
+        \item day: $\{d \in \mathbb{I}\vert 1 \leq d \leq 31 \}$
+        \item date:  $ \otimes: \dfiber_{year} \times \dfiber_{month} \times \dfiber_{day} \rightarrow \dfiber_{date}$
+    \end{itemize}
+    where the composition $\otimes$ includes the constrain to only return dates that have the right number of days for each month.
+\end{frame}
+
+\begin{frame}{Expand records $\rightarrow$ multiply fiber space}
+
+    \begin{figure}
+        \includegraphics[width=\textwidth]{../paper/figures/tex/f_product.pdf}
+    \end{figure}
+\end{frame}
+
+\subsubsection{Dataset: section}
+
+\begin{frame}{local sections}
+    We define section functions locally:
+    \begin{equation}
+      \label{eq:atct:fb:sections}
+      \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}
+    \end{equation}
+    such that each section function $\dsection: \dbasepoint \mapsto \delement$ maps from each point $\dbasepoint \in \openset$ to a corresponding record in the fiber space $\delement \in \dfiber_{\dbasepoint}$ over that point.
+
+    This encoding can be translated into a set of signatures $\{\textcolor{section}{\texttt{data-subset}}: \textcolor{base}{\texttt{topology}} \rightarrow \textcolor{fiber}{\texttt{fields}}$ s.t.
+$\textcolor{section}{\texttt{data-subset}} \subset \textcolor{section}{\texttt{dataset}}\}$.
+\end{frame}
+
+\begin{frame}{global sections}
+    When a bundle is trivial $\dtotalc = \dbasec \times \dfiberc$, we can defined a global sections $\dsection: \dbase \rightarrow \dfiber \in \Gamma(\dbase, \dfiber)$ which we translate into a data signature of the form $\textcolor{section}{\texttt{dataset}}: \textcolor{base}{\texttt{topology}} \rightarrow \textcolor{fiber}{\texttt{field}}$ where $\dsectionc=\textcolor{section}{\texttt{dataset}}$, $\dbasec=\textcolor{base}{\texttt{topology}}$ and $\dfiberc=\textcolor{fiber}{\texttt{fields}}$
+\end{frame}
+
+\begin{frame}{Graphic bundle and section}
+    \begin{equation}
+        \label{eq:atct:fb:graphic}
+        \begin{tikzcd}[ampersand replacement=\&]
+            \gfiberc \arrow[r, hook, color=total] \& \gtotalc \arrow[r, "\pi", two heads, color=total] \& \gbasec
+        \end{tikzcd}
+    \end{equation}
+
+    \begin{equation}\label{eq:atct:fb_graphic_section}
+        \begin{split}
+        \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} & \coloneqq \\
+        &\big\{\gsectionc: \opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} \; \bigm{\vert} \pi(\gsectionc(\gbasepointc)) = \gbasepointc\;for\, all\; \gbasepointc \in \opensetgc \big\}
+        \end{split}
+      \end{equation}
+\end{frame}
+
+\begin{frame}{Graphic section}
+\begin{figure}
+    \includegraphics[width=\linewidth]{../paper/figures/fb_rho.pdf}
+\end{figure}
+\end{frame}
+
+\subsection{Abstract Data Containers}
+
+\begin{frame}{functors}
+    \begin{definition}\cite{bradleyWhatFunctorDefinitions,bradleyTopologyCategoricalApproach2020} A \textbf{functor} is a map $F: \mathcal{C} \rightarrow \mathcal{D}$, which means it is a function between objects $F: \textbf{ob}(\mathcal{C}) \mapsto \textbf{ob}(\mathcal{D})$ and that for every morphism $f \in Hom(C_1, C_2)$  there is a corresponding function $F: Hom(C1, C2) \mapsto Hom(F(C_1), F( C_2))$.
+        A \textbf{functor} must satisfy the properties
+        \begin{itemize}
+          \item \textit{identity}: $F(id_{C}(C)) = id_{D}(F(C))$
+          \item \textit{composition}: $F(g)\circ F(f) = F(g\circ f)$ for any composable morphisms $C_{1}\xrightarrow{f} C_2$, $C_2 \xrightarrow{g} C_3$
+        \end{itemize}
+        $F(C) \in \textbf{ob}(\mathcal{D})$ denotes the object to which an object $C$ is mapped, and $F(f) \in Hom(F_(C_1), F_(C_2))$ denotes the morphism that $f$ is mapped to.
+        \end{definition}
+\end{frame}
+
+\begin{frame}{topological equivariance: presheaf}
+    \begin{definition}\label{def:atct:presheaf}
+        A \textbf{presheaf} $F:\mathcal{C}^{op} \rightarrow \setb$ is a contravariant functor from an object in an arbitrary category to an object in the category \setb\cite{spanier1989algebraic}.
+    \end{definition}
+    The presheaf is contravariant when for every arbitrary morphism between input base spaces $\dfunch: \openset_1 \rightarrow \openset_2$ there exists a corresponding pullback function between the sets of sections $\dfuncpull: \cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}} \rightarrow \cgamma{\opensetc_1}{\dtotalc\restriction_{\openset_c1}}$.
+\end{frame}
+
+\begin{frame}{topological equivariance: presheaf}
+    \begin{figure}
+        \includegraphics[width=\textwidth]{../paper/figures/tex/presheaf.pdf}
+    \end{figure}
+\end{frame}
+
+\begin{frame}{Container invariance: sheaf}
+    \begin{definition}\label{def:atct:sheaf}\cite{bakerEuclideanSpaceMathsSheaf,spanier1989algebraic} A \textbf{sheaf} is a presheaf that satisfies the following two axioms
+        \begin{itemize}
+          \item \textit{locality} two sections in a sheaf are equal $\dsection^{a} = \dsection^{b}$ when they evaluate to the same values $\dsection^{a}\vert_{\openset_i} =  \dsection^{b}\vert_{\openset_i}$ over the open cover $\bigcup_{i\in I} \openset_{i} \subset \openset$ (indexed by $I$).
+          \item \textit{gluing} the union of sections defined over subspaces $\dsection^{i} \in \Gamma(\openset_i, \dtotal|_{\openset_i})$ is equivalent to a section defined over the whole space $\dsection\vert_{\openset_i} = \dsection^{i}$ for all $i\in I$ if all pairs of sections agree on overlaps $\dsection^{i}\vert_{\openset_i\cap\openset_j} =  \dsection^{j}\vert_{\openset_i\cap\openset_j}$
+          \end{itemize}
+        \end{definition}
+\end{frame}
+
+\begin{frame}{Container invariance: sheaf}
+\begin{figure}
+    \includegraphics[width=\linewidth]{../paper/figures/tex/sheaf_rules.pdf}
+\end{figure}
+\end{frame}
+
+\begin{frame}{What about needing more than one record?}
+    The sheaf over a very small region surrounding a point $\dbasepoint$ is called a \textit{stalk}\cite{harder2008lectures}
+\begin{equation}
+  \label{eq:atct:sheaf:stalk}
+    \sheaf_{\dbase, \dtotalc}\restriction_{\dbasepoint}\coloneqq \lim\limits_{\openset\ni \dbasepoint} \Gamma(\openset, \dtotal\restriction_{\openset})
+\end{equation}
+where the fiber is contained inside the stalk  $\dfiber_{\dbasepoint} \subset  \sheaf_{\dbase, \dtotal}\restriction_{\dbasepoint}$.
+The \textit{germ} is the section evaluated at a point in the stalk  $\dsection(\dbasepoint) \in \sheaf_{\dbase, \dtotal}\restriction_{\dbasepoint}$ and is the data.
+\end{frame}
+\section{Artist}
+
+\subsection{Homeomorphism}
+\begin{frame}{Homeomorphism}
+    We define the mapping \vindexc\ to be a surjective continuous map:
+    \begin{equation}
+      \label{eq:atct:xi}
+      \vindexc: \opensetgc \textcolor{functor}{\rightarrow} \opensetc
+    \end{equation}
+    between a graphic subspace $\opensetg \subseteq \gbase$ and data subspace $\openset \subseteq \dbase$.
+
+    The set of points in graphic space that correspond to each point in data space is
+\begin{equation}
+  \label{eq:atct:xi:inverse}
+  \vindexprec(\dbasepointc) = \{\gbasepointc | \vindexc(\gbasepointc) = \dbasepointc \forall \dbasepointc \in \dbase, \gbasepointc \in \gbasec\}
+\end{equation}
+such that every point in a graphic space has a corresponding point in data space.
+\end{frame}
+
+\begin{frame}{Homeomorphism}
+    \begin{figure}
+        \includegraphics[width=\textwidth]{../paper/figures/xi_zoom.pdf}
+    \end{figure}
+\end{frame}
+
+\begin{frame}{Pushforward: Graphic for Data}
+    \begin{definition} Given a sheaf $\sheaf_{\gbase, \gtotal}$ on $\gbase$, the \textbf{pushforward} sheaf  $\vindexpush\sheaf_{\gbase, \gtotal}$ on $\dbase$ is defined as
+        \begin{equation}
+          \vindexpushc(\sheafc_{\gbasec, \gtotalc})(\opensetc)  = \sheafc_{\gbasec, \gtotalc}(\vindexc^{-1}(\opensetc))
+        \end{equation}
+      for all opensets $\opensetc \subset \dbasec$\cite{harder2008lectures}.
+    \end{definition}
+This provides a way to look up which graphic corresponds with a data index
+\begin{equation}
+  \label{eq:atct:sheaf:pushforward_select}
+  \vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexprec(\dbasepointc)}
+\end{equation}
+such that $\vindexpush\gsection(\dbasepoint))(\gbasepoint) = \gsection(\gbasepoint)$ for all $\gbasepoint \in \vindexpre(\dbasepoint)$.
+\end{frame}
+
+\begin{frame}{Pullback: Data at graphic }
+    \begin{definition} \cite{harder2008lectures} Given a sheaf $\sheaf_{\dbase, \dtotal}$ on $\dbase$, the \textbf{pullback} sheaf $\vindexpullc\sheafc_{\dbasec, \dtotalc}$ on $\gbasec$ is defined as the sheaf associated to the presheaf $\vindexpullc(\sheafc_{\dbasec, \dtotalc})(\opensetgc) = \sheafc_{\dbasec, \dtotalc}(\vindexc(\opensetgc))$ for $\vindex(\opensetgc) \in \dbase$.
+    \end{definition}
+
+    This provides a way to then look up what data values correspond with a graphic index
+    \begin{equation}
+      \label{eq:atct:sheaf:pullback_hover}
+      \vindexpullc\dsectionc(\gbasepointc) = \dsectionc(\vindexc(\gbasepointc)) = \dsectionc(\dbasepointc)
+    \end{equation}
+    As \vindex\ is surjective, there are many points $\gbasepoint \in \opensetg\subseteq\gbase$ in the graphic space that correspond to a single point $\vindex(\gbasepoint) = \dbasepoint$.
+\end{frame}
+
+\begin{frame}{Homeomorphism}
+    \begin{figure}
+        \includegraphics[width=\textwidth]{../paper/figures/xi_diagram.pdf}
+    \end{figure}
+\end{frame}
+
+\begin{frame}{Artist}
+    \begin{definition}\label{def:natural-transform}
+        Given two functors $F, G: \mathcal{C}\rightarrow \mathcal{D}$, a \textbf{natural transformation} $\alpha: F \rightarrow G$ is a function which assigns to each object $c$ of $\mathcal{C}$ a morphism $\alpha_c:F(input) \rightarrow G(c), G(c) \in \mathcal{D}$, in such a way that for every morphism $f:c \rightarrow c^\prime, c^\prime \in \mathcal{C}$, the morphisms in $\mathcal{D}$ commute such that $\alpha_c^{\prime}(F(f)(F(c))) = G(f)(\alpha_c(F(c))$.
+When this holds, $\alpha_{c}$ is \textit{natural} in $c$.\cite{maclaneCategoriesWorkingMathematician2013}.
+      \end{definition}
+
+\end{frame}
+
+\begin{frame}{Artist}
+
 \begin{equation*}
     \begin{tikzcd}[ampersand replacement=\&]
-        \onslide*<1-|handout:1>{\setc_1} 
-        \onslide*<3-6|handout:1>{\arrow[r, "r_{1,2}", color=set]}
-        \onslide*<4-6|handout:1>{\arrow[rd, "r_{1,2}\circ r_{2,3}"', color=set]}
-        \onslide*<2-6|handout:1>{\arrow["id_{\setc_1}"', loop, distance=2em, in=215, out=145, color=set]} \& 
-        \onslide*<1-|handout:1>{\setc_2}
-        \onslide*<7|handout:2>{\arrow[l, color=set]}
-        \onslide*<3-6|handout:1>{\arrow[d, "r_{2,3}", color=set]} 
-        \onslide*<2-6|handout:1>{\arrow["id_{\setc_2}"', loop, distance=2em, in=125, out=55, color=set]} \\\& 
-        \onslide*<1-|handout:1>{\setc_3}
-        \onslide*<7|handout:2>{\arrow[u, color=set]}
-        \onslide*<7|handout:2>{\arrow[lu, color=set]}
-        \onslide*<2-6|handout:1>{\arrow["id_{\setc_3}"', loop, distance=2em, in=305, out=235, color=set]}              
-    \end{tikzcd}
-    \end{equation*}
-    \column{0.59\textwidth}
-    \begin{alertblock}<5-6|handout:1>{associativity} 
-        if $r_{1,2}: \setc_1 \rightarrow \setc_2$, $r_{2,3}: \setc_2 \rightarrow \setc_3$ and $r_{3,4}: \setc_3 \rightarrow \setc_4$ then $r_{3,4}\circ (r_{2,3} \circ r_{1,2}) = (r_{3,4} \circ r_{2,3}) \circ r_{1,2}$
-    \end{alertblock}
-    \begin{alertblock}<6|handout:1>{identity} 
-        for every $r_{1,2}: \setc_1 \rightarrow \setc_2$ there exists identity morphisms $r_{1,2} \circ id_{\setc_1} = r_{1,2} = id_{\setc_2} \circ r_{1,2}$
-    \end{alertblock}
-    \end{columns}
-\end{frame}
-
-\begin{frame}<presentation:0|handout:1>{Opposite Category $\textcolor{source}{\mathcal{C}^{op}}$}
-    \begin{columns}
-        \column{0.49\textwidth}
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement=\&]
-            \textcolor{fade}{C_1} \arrow[r, "f", color=fade] \arrow[rd, "g\circ f"', color=fade] \arrow["id_{C_1}"', loop, distance=2em, in=215, out=145, color=fade] \& \textcolor{fade}{C_2} \arrow[d, "g", color=fade] \arrow["id_{C_2}"', loop, distance=2em, in=125, out=55, color=fade] \\
-        \& \textcolor{fade}{C_3} \arrow["id_{C_3}"', loop, distance=2em, in=305, out=235, color=fade]              
-        \end{tikzcd}
-        \end{equation*}
-        \column{0.49\textwidth}
-        \onslide*<presentation:2|handout:1>{
-        \begin{equation*}
-        \begin{tikzcd}[ampersand replacement = \&]
-            \textcolor{source}{C_1} \arrow["id_{C_1}"', loop, distance=2em, in=215, out=145, color=source] \& \textcolor{source}{C_2} \arrow[l, "f"',color=source] \arrow["id_{C_2}"', loop, distance=2em, in=125, out=55, color=source]\\
-            \& \textcolor{source}{C_3} \arrow[u, "g"', color=source] \arrow[lu, "f \circ g", color=source] \arrow["id_{C_3}"', loop, distance=2em, in=305, out=235, color=source]
-            \end{tikzcd}
-        \end{equation*}
-        }
-        \end{columns}
-\end{frame}
-
-\begin{frame}<presentation:1-|handout:1>
-    \frametitle{Functor $\textcolor{functor}{\bm{F}}: \textcolor{source}{\mathcal{C}} \rightarrow \textcolor{target}{\mathcal{D}}$}
-\begin{columns}
-    \column{.39\textwidth}
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement = \&, ]
-            \onslide*<1-|handout:1>{\textcolor{source}{c}} 
-            \onslide*<1-|handout:1>{\arrow[d, "f"', shift right=0, color=source]} 
-            \onslide*<3-|handout:1>{\arrow[r, "F", color=functor, maps to]} \& 
-            \onslide*<2-3|handout:0>{\textcolor{target}{d}}
-            \onslide*<4-|handout:0>{\textcolor{target}{F(c)}} 
-            \onslide*<0|handout:1>{\textcolor{target}{F(c)=d}} 
-            \onslide*<2|handout:0>{\arrow[d, shift right=0, color=target]}
-            \onslide*<3-|handout:0>{\arrow[d, "F(f)", shift right=0, color=target]}
-            \onslide*<0|handout:1>{\arrow[d, "F(f)", shift right=3, color=target]} \\
-            \onslide*<1-|handout:1>{\textcolor{source}{c^{\prime}}} 
-            \onslide*<3-|handout:1>{\arrow[r, "F"', color=functor, maps to]}\& 
-            \onslide*<2-3|handout:0>{\textcolor{target}{d^{\prime}}}
-            \onslide*<4-|handout:0>{\textcolor{target}{F(c^{\prime})}}
-            \onslide*<0|handout:1>{\textcolor{target}{F(c^{\prime})= d^{\prime}}}                    
-        \end{tikzcd}
-    \end{equation*}
-    \column{.59\textwidth}
-    \begin{alertblock}<5-|handout:1>{composition}
-        \begin{align*}
-            \textcolor{functor}{F}(\textcolor{source}{g}) \circ  \textcolor{functor}{F}(\textcolor{source}{f}) = \textcolor{functor}{F} (\textcolor{source}{g}\circ \textcolor{source}{f})
-        \end{align*}
-    \end{alertblock}
-    \begin{alertblock}<6-|handout:1>{identity}
-        \begin{align*}
-            \textcolor{functor}{F}(\textcolor{source}{id_c}) = \textcolor{target}{id}_{\textcolor{functor}{F}(\textcolor{source}{c})}
-        \end{align*}
-    \end{alertblock}
-\end{columns}
-\end{frame}
-
-\subsubsection{Natural Transforms}
-\begin{frame}<presentation:1-|handout:1>
-    \frametitle{Natural Transformation $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\Rightarrow} \textcolor{functor}{G}$}
-    \adjustbox{scale=3, center}{
-        \begin{tikzcd}[ampersand replacement=\&]
-            \& {} \arrow[dd, "\alpha" , Rightarrow, shorten=1.5em, color=nattran] \&             \\
-          \textcolor{source}{\mathcal{C}} \arrow[rr, "F", bend left, color=functor] \arrow[rr, "G"', bend right, color=functor] \& \& \textcolor{target}{\mathcal{D}} \\
+      \& {} \arrow[dd, "\texttt{artist}" description, Rightarrow, shift left=3.65, shorten <= .7em, shorten >= .7em, color=artist, font=\Huge]
+      \& \\ {\scriptstyle\textcolor{base}{\texttt{topology}}}
+        \arrow[rr, "\texttt{graphic}"', bend right, color=sheaf, font=\huge] \arrow[rr, "\texttt{data}", bend left, color=sheaf, font=\huge] \&
+      \& {\scriptstyle\textcolor{base}{\texttt{topology}}\textcolor{section}{\rightarrow} \textcolor{fiber}{\texttt{fields}}} \\
       \& {}                                              \&
-          \end{tikzcd}}
-\end{frame}
-
-\begin{frame}<presentation:1-|handout:1>
-    \frametitle{Natural Transformation $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\Rightarrow} \textcolor{functor}{G}$}
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement=\&]  
-                \&  \& 
-                \textcolor{target}{d_s} \& 
-                \textcolor{source}{c} 
-                \arrow[r, "G", maps to, color=functor] 
-                \arrow[l, "F"', maps to, color=functor] \& 
-                \textcolor{target}{d_t} \\
-                \textcolor{source}{c_1} 
-                \arrow[dd, "f" description, color=source] \& \& 
-                \textcolor{target}{F(c_1)} 
-                \arrow[dd, "F(f)"', color=target] 
-                \arrow[rr, "\alpha_{c_1}", color=nattran] \& \& 
-                \textcolor{target}{G(c_1)} 
-                \arrow[dd, "G(f)", color=target] \\
-                \& \& {} 
-                \arrow[rr, "\alpha_{f}", dotted, color=nattran] \& \& {} \\
-                \textcolor{source}{c_2} \& \& 
-                \textcolor{target}{F(c_2)} 
-                \arrow[rr, "\alpha{c_2}", color=nattran] \& \& 
-                \textcolor{target}{G(c_2)} \\
-                \& \& 
-                \textcolor{target}{F} 
-                \arrow[rr, "\alpha", Rightarrow, color=nattran] \& \& 
-                \textcolor{target}{G}                        
   \end{tikzcd}
@@
 
-\subsection{Sheaves}
-
-\begin{frame}<presentation:1-|handout:1>
-    \frametitle{Presheaf: $\sheafc: \mathcal{C}^{op} \textcolor{sheaf}{\rightarrow} \setc $}
-    \begin{columns}
-    \column{0.29\textwidth}
-    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-        \onslide*<9-|handout:1->{\dfiberc}
-        \onslide*<9-|handout:1->{\arrow[r, hook]} \& 
-        \onslide*<1-|handout:1->{\dtotalc}
-        \onslide*<1-|handout:1->{\arrow[d, "\pi"']} \\
+\begin{frame}{How?}
+    \begin{equation}
+        \label{eq:artist:homset}
+        \begin{tikzcd}[ampersand replacement=\&, column sep=50]
+       \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}
+      \& \cgamma{\opensetgc},{\gtotalc\restriction_{\opensetgc}}
+      \arrow[l, "\vindexpushc"', color=functor]                                          \\
+      \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
+      \arrow[r, "\vindexpullc", color=functor]
+      \arrow[u, "\textcolor{set}{Hom}_{\sheafc_{\dbasec}}", color=homset]
+      \arrow[ru, "{\textcolor{set}{Hom}_{\sheafc_{\dbasec},\sheafc_{\gbasec}}}", color=homset]
       \&
-        \onslide*<1-|handout:1->{\dbasec}
-        \onslide*<3-|handout:1->{\arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]}
+      \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}
+      \arrow[u, "\textcolor{set}{Hom}_{\sheafc_{\gbasec}}"', color=homset]
       \end{tikzcd}
-    \column{0.69\textwidth}
-        \begin{tikzcd}[ampersand replacement = \&, column sep=small]
-            \onslide*<3,6-|handout:0>{{\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}} 
-            \onslide*<4-5|handout:0>{{\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}\in Ob(\setc)}
-            \onslide*<0|handout:1>{\Ob(\setc) \ni {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}}}
-            \&  
-            {} \& 
-            \onslide*<6-|handout:1>{{\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}}} 
-            \onslide*<7-|handout:1>{\arrow[ll, "\iota^*"', hook', color=set]} \\\& \&\\
-            \onslide*<3|handout:0>{\opensetc_1 \subset \dbasec}
-            \onslide*<4-5|handout:0>{\opensetc_1 \in Ob(\mathcal{\dbasec}^{\textcolor{base}{op}})}
-            \onslide*<6-|handout:0>{\opensetc_1}
-            \onslide*<0|handout:1>{Ob(\mathcal{\dbasec}^{\textcolor{base}{op}}) \ni \opensetc_1} 
-            \onslide*<5-|handout:0>{\arrow[uu, "\sheafc_{\dbasec, \dtotalc}", maps to, color=sheaf]}
-            \onslide*<0|handout:1>{\arrow[uu, "\sheafc_{\dbasec, \dtotalc}", maps to, color=sheaf, shift right = 10]}
-            \onslide*<7-|handout:1>{\arrow[rr, "\iota"', hook, color=base]} \&   
-            \onslide*<8-|handout:1>{\arrow[uu, "\sheafc_{\dbasec, \dtotalc}" description, maps to, color=sheaf]} \& 
-            \onslide*<6-|handout:1>{\opensetc_2} 
-            \onslide*<6-|handout:1>{\arrow[uu, "\sheafc_{\dbasec, \dtotalc}"', maps to, color=sheaf]}                   
-            \end{tikzcd}
-    \end{columns}
-    \begin{center}
-        \begin{alertblock}<8-|handout:1>{stalk}
-            \begin{align*}
-                \sheafc_{\dbasec, \dtotalc}\restriction_{\dbasepointc}\coloneqq \lim\limits_{\opensetc\ni \dbasepointc} \Gamma(\opensetc, \dtotalc\restriction_{\opensetc}) \\
-                 \dfiberc_{\dbasepointc} \subset  \sheafc_{\dbasec, \dtotalc}\restriction_{\dbasepointc}
-            \end{align*}
-        \end{alertblock}
-
-        \begin{alertblock}<9-|handout:1>{germ}
-            %germ  is tau at limit, \tau(k) \in germ
-            \begin{align*}
-                \dsectionc(\dbasepointc) \in \sheafc_{\dbasec, \dtotalc}\restriction_{\dbasepointc}
-            \end{align*}
-        \end{alertblock}
-    \end{center}
+      \end{equation}
 \end{frame}
 
-\begin{frame}<presentation:1|handout:1>{Sheaves on Bundles}
-    A sheaf is a presheaf that satisfies the following two axioms\cite{bakerMathsSheaf}\\
-    \begin{block}<2-|handout:1>{locality}
-        \onslide<3-|handout:1>{
-        given $\opensetc = \bigcup\limits_{i\in I} \opensetc_i$ and $\dsectionc^{a}, \dsectionc^{b} \in \sheafc(\opensetc)$,}
-        \onslide<4|handout:1>{\\
-        if $\dsectionc^{a}\restriction_{\opensetc_i} = \dsectionc^{b}\restriction_{\opensetc_i}$ for each $\opensetc_i \in \opensetc$}
-        \onslide<4|handout:1>{then $\dsectionc^{a} = \dsectionc^{b}$}
-    \end{block} 
-    \begin{block}<5-|handout:1>{gluing} 
-         given $\dsectionc^{i} \in \sheafc(\opensetc_i)$
-        \onslide<6-|handout:1>{s.t.
-${\dsectionc^{i}}\restriction_{\opensetc_i\cap \opensetc_j} = {\dsectionc^{j}}\restriction_{\opensetc_i \cap \opensetc_j}$ for $\opensetc_i, \opensetc_j \in \opensetc$,}
-        \onslide<7-|handout:1>{\\there exists $\dsectionc \in \sheafc(\opensetc)$ such that $\dsectionc\restriction_{\opensetc_i} = {\dsectionc^{i}}$} 
-    \end{block}      
+\begin{frame}{Artist}
+    \begin{align}
+        \label{eq:artist:hom_transport}
+        \vartistc:& \cgamma{\dbasec}{\dtotalc}\textcolor{artist}{\rightarrow} \imartist{\gbasec}{\gtotalc}, \imartist{\gbasec}{\gtotalc} \subset \cgamma{\gbasec}{\gtotalc}
+      \end{align}
+      where:
+    \begin{equation}
+        \label{eq:artist:output}
+        \imartist{\gbasec}{\gtotalc} \coloneqq\\
+        \{\gsectionc \mid\;\exists\;\dsectionc \in \cgamma{\dbasec}{\dtotalc}\;s.t.\;
+        \vartistc(\dsectionc) = \gsectionc,\; \vindexc(\gbasec) = \dbasec \}
+      \end{equation}
 \end{frame}
 
-\subsubsection{Pullback and Pushforward}
-\begin{frame}<presentation:1-|handout:1->{}
-    \frametitle<3|handout:3>{Function: $\vindexc: \gbasec \rightarrow \dbasec$}
-    \frametitle<1|handout:1>{Data}
-    \frametitle<4|handout:4>{Pullback: data to region of the visualization} 
-    \frametitle<2|handout:2>{Graphic}
-    \frametitle<5|handout:5>{Pushforward: visualization to index of data}
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-            \onslide*<2,3,5|handout:2,3,5>{\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}} 
-            \onslide*<5|handout:5>{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]} \& \& 
-            \onslide*<5|handout:5>{\cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}}  
-            \\
-            \onslide*<2,4-|handout:2,4->{\opensetgc}
-            \onslide*<3|handout:3>{\overbrace{\opensetc\times[0,1]^{m}}^{\opensetgc}}
-            \onslide*<3-|handout:3->{\arrow[rr, "\vindex", color=functor]}
-            \onslide*<2,3,5|handout:2,3,5>{\arrow[u, "\sheaf_{\gbasec, \gtotalc}", maps to, color=sheaf]}
-            \onslide*<4|handout:4>{\arrow[d, dashed, maps to, "\sheaf_{\gbasec, \vindexpullc\dtotalc}"',color=sheaf]} \&  \& 
-            \onslide*<1,3-|handout:1,3->{\opensetc}
-            \onslide*<1,3-4|handout:1,3-4>{\arrow[d, "\sheaf_{\dbasec, \dtotalc}", maps to, color=sheaf]}
-            \onslide*<5|handout:5>{\arrow[u, dashed, maps to, "\sheaf_{\dbasec, \vindexpushc\gtotalc}"', color=sheaf]} \\
-            \onslide*<4|handout:4>{\cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}} 
-             \& \& 
-            \onslide*<1,3-4|handout:1,3-4>{\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}}
-            \onslide*<4|handout:4>{\arrow[ll, "\textcolor{functor}{\vindexpull}", color=functor]} 
+\begin{frame}{Equivariant Artist}
+    \begin{equation}
+        \label{eq:artist:equivariance}
+        \begin{tikzcd}[ampersand replacement=\&, column sep=small]
+        \cgamma{\dbasec}{\dtotalc}
+        \arrow[rrr, "\vartistc", color=artist]
+        \arrow[d, "\dfuncpullc_{\dtotalc}"', color=action]
+        \& \& \&
+        \imartist{\gbasec}{\gtotalc}
+        \arrow[d, "\dfuncpullc_{\gtotalc}", dotted] \\
+        \cgamma{\dbasec^{\prime}}{\dfuncpullc_{\dtotalc}\dtotalc}
+        \arrow[dd, "\dfunctc_{\dtotalc}"', color=action] \&
+        \dbasec
+         \&
+        \gbasec
+        \arrow[l, "\vindexc"', color=functor]
+        \&
+        \imartist{\gbase^{\prime}}{\dfuncpullc_{\gtotalc}\gtotalc}
+        \arrow[dd, "\dfunctc_{\gtotalc}", dotted, color=action] \\
+        \&
+        \dbasec^{\prime}
+        \arrow[u, "\dfunchc_{\dtotalc}", color=action]
+        \&
+        \gbasec^{\prime}
+        \arrow[l, "\vindexc"', color=functor]
+        \arrow[u, "\dfunchc_{\gtotalc}"', dotted, color=action]
+        \& \\
+        \cgamma{\dbasec^{\prime}}{\dtotalc^{\prime}}
+        \arrow[rrr, "\vartistc", color=artist]
+        \& \& \&
+        \imartist{\gbasec^{\prime}}{\gtotalc^{\prime}}
         \end{tikzcd}
-    \end{equation*}
-    \only<1|handout:1>{% data sheaf
-        \begin{itemize}
-            \item $\dfiberc \hookrightarrow \dtotalc \xrightarrow{\pi} \dbasec$
-            \item $\sheafc_{\dbasec, \dtotalc}:\opensetc \mapsto \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}, \opensetc \subseteq \dbasec$
-            \item $\dsectionc: \opensetc \rightarrow \dfiberc\restriction_{\opensetc} \in \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$
-            \item $\dsectionc(\dbasepointc) = \{f_0:v_0, \cdots,\}, \dbasepointc \in \opensetc$
-        \end{itemize}
-    }
-    \only<4|handout:4>{%pulled back data sheaf
-        \begin{itemize}
-            \item $\vindexpullc\dfiberc \hookrightarrow \vindexpullc\dtotalc \xrightarrow{\pi} \gbasec$
-            \item $\vindexpullc\sheafc_{\dbasec,\dtotalc}:\opensetgc \mapsto \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}, \opensetgc \subseteq \gbasec$
-            \item $\vindexpullc\dsectionc: \opensetgc \rightarrow \vindexpullc\dfiberc\restriction_{\opensetgc} \in \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}$
-            \item $\vindexpullc\dsectionc(\gbasepointc) = \dsectionc(\vindexc(\gbasepointc)) = \dsectionc(\dbasepointc)$
-        \end{itemize}
-    }
-    \only<2|handout:2>{% graphic sheaf
-        \begin{itemize}
-            \item $\gfiberc \hookrightarrow \gtotalc \xrightarrow{\pi} \gbasec$
-            \item $\sheafc_{\gbasec, \gtotalc}:\opensetgc \mapsto \cgamma{\opensetgc}{\dtotalc\restriction_{\opensetgc}}, \opensetgc \subseteq \gbasec$
-            \item $\gsectionc: \opensetgc \rightarrow \gfiberc\restriction_{\opensetgc} \in \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}$
-            \item $\gsectionc(\gbasepointc) = \{d_0, \cdots\}, \gbasepointc \in \opensetgc$
-        \end{itemize}
-    }
-    \only<5|handout:5>{%pushed forward sheaf
-        \begin{itemize}
-            \item $\vindexpushc\gfiberc \hookrightarrow \vindexpushc\gtotalc \xrightarrow{\pi} \dbasec$
-            \item $\vindexpushc\sheafc_{\gbasec, \gtotalc}:\opensetc \mapsto \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}, \opensetc \subseteq \dbasec$
-            \item $\vindexpushc\gsectionc: \opensetc \rightarrow \vindexpushc \gfiberc\restriction_{\opensetc} \in \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}$
-            \item $\vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexpre(\dbasepointc)} = \gsectionc(\gbasepointc)\;\forall\;                                                                                                                                           \gbasepointc \in \vindexpre(\dbasepointc)$
-        \end{itemize}
-    }
+      \end{equation}
 \end{frame}
 
-\begin{frame}<presentation:1|handout:1>{Example: Scatter Plot}
+\begin{frame}{Equivariant Artist}
 \begin{figure}
-        \includegraphics[scale=.1]{figures/math/push_pull_scatter.png}
+    \includegraphics[width=\linewidth]{../paper/figures/artist_equiv.png}
 \end{figure}
@@
 
-\begin{frame}<presentation:1|handout:1>{Example: Line Plot}
+\begin{frame}{artist composition}
     \begin{figure}
-        \includegraphics[scale=.1]{figures/math/push_pull_line.png}
+        \includegraphics[width=\linewidth]{../paper/figures/composition.pdf}
     \end{figure}
@@
 
-\begin{frame}<presentation:1|handout:1>{Example: Image}
-    \begin{figure}
-        \includegraphics[scale=.1]{figures/math/push_pull_image.png}
-    \end{figure}
-\end{frame}
-
-\begin{frame}<presentation:1|handout:1>{Example: Heatmap}
-    \begin{figure}
-        \includegraphics[scale=.1]{figures/math/push_pull_heat.png}
-    \end{figure}
-\end{frame}
-
-\begin{frame}<presentation:1|handout:1>{Example: Sphere}
-    \begin{figure}
-        \includegraphics[scale=.1]{figures/math/push_pull_sphere.png}
-    \end{figure}
-\end{frame}
-
-%add transition slide->possibly call back
-
-\section{Artist}
-
-\begin{frame}<presentation:1-|handout:1->
-    \frametitle<1|handout:1>{$\textcolor{homset}{Hom}_{\gbasec}(\sheaf_{\gbasec, \vindexpullc\dtotalc}, \sheaf_{\gbasec,\gtotalc}) = \textcolor{homset}{Hom}_{\dbasec}(\sheafc_{\dbasec, \dtotalc}, \sheafc_{\dbasec, \vindexpushc\gtotalc})$}
-    \frametitle<2|handout:2>{Data Space: $\vartistc_{\dbasec}: \sheafc_{\dbasec, \dtotalc} \textcolor{artist}{\Rightarrow} \sheafc_{\dbasec, \vindexpushc\gtotalc}$}
-    \frametitle<3|handout:3>{Display Space: $\vartist_{\gbasec}: \sheafc_{\gbasec, \vindexpullc\dtotalc} \textcolor{artist}{\Rightarrow} \sheafc_{\gbasec, \gtotalc}$}
-    \frametitle<4|handout:4>{Artist: $\vartistc: \dsectionc \mapsto \gsectionc$}
-
-    \onslide*<1,4|handout:1,4>{
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement=\&]
-          \onslide*<1|handout:1>{\sheafc_{\gbasec, \gtotalc}} 
-          \onslide*<4|handout:4>{\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}}
-          \onslide*<1,4|handout:1,4>{\arrow[rr, "\textcolor{functor}{\vindexpush}", color=functor]}  
-          \& \& 
-          \onslide*<1|handout:1>{\sheafc_{\dbasec, \vindexpushc\gtotalc}}
-          \onslide*<4|handout:4>{\cgamma{\opensetc}{\vindexpushc \gtotalc\restriction_{\opensetc}}}
-          \\
-          \& \& \\
-          \onslide*<1|handout:1>{\sheafc_{\gbasec, \vindexpullc\dtotalc}}
-          \onslide*<4|handout:4>{\cgamma{\opensetgc}{\vindexpullc \dtotalc\restriction_{\opensetgc}}}
-          \onslide*<1|handout:1>{\arrow[uu, "\vartistc_{\gbasec} \in \textcolor{homset}{Hom}_{\gbasec}", color=homset]}
-          \onslide*<4|handout:4>{\arrow[uu, "\vartistc_{\gbasec}", color=artist]}
-          \&  \& 
-          \onslide*<1|handout:1>{\sheafc_{\dbasec, \dtotalc}}
-          \onslide*<4|handout:4>{\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}}
-          \onslide*<1,4|handout:1,4>{\arrow[ll, "\textcolor{functor}{\vindexpull}", color=functor]} 
-          \onslide*<1|handout:1>{\arrow[uu, "\vartistc_{\dbasec} \in \textcolor{homset}{Hom}_{\dbasec}"', color=homset]}
-          \onslide*<4|handout:4>{\arrow[uu, "\vartistc_{\dbasec}"', color=artist]}
-          \onslide*<1|handout:1>{\arrow[lluu, "\vartistc \in \textcolor{homset}{Hom}_{\dbasec,\gbasec}"' description, dashed, color=homset]}
-          \onslide*<4|handout:4>{\arrow[lluu, "\vartistc"' description, color=artist]}
-        \end{tikzcd}
-    \end{equation*}}
-
-    \onslide*<2|handout:2>{
-        \begin{equation*}
-            \begin{tikzcd}[ampersand replacement=\&]  
-                    \&  \& 
-                    \sheafc_{\dbasec, \dtotalc} 
-                    \arrow[rr, "\vartistc_{\dbasec}", Rightarrow, color=artist] \& \& 
-                    \sheafc_{\dbasec, \vindexpushc\gtotalc}  \\
-                    \opensetc_1\& \& 
-                    \cgamma{\opensetc_1}{\dtotalc\restriction_{\opensetc_1}}
-                    \arrow[dd, "\iota^*"', color=set] 
-                    \arrow[rr, "\vartistc_{\opensetc_1}", color=artist] \& \& 
-                    \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc_1}} 
-                    \arrow[dd, "\iota^*", color=set] \\
-                    \& \& {} 
-                    \arrow[rr, "\vartistc_{\sheafc}", dotted, color=artist] \& \& {} \\
-                    \arrow[uu, "\iota" description, hook, color=base] 
-                    \opensetc_2 \& \& 
-                    \cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}
-                    \arrow[rr, "\vartistc_{\opensetc_2}", color=artist] \& \& 
-                    \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc_2}} \\
-                    \& \& 
-                    \cgamma{\opensetc}{\dtotalc\restriction\opensetc} \& 
-                    \opensetc 
-                    \arrow[r, "\sheafc_{\dbasec,\vindexpushc\gtotalc}", maps to, color=sheaf] 
-                    \arrow[l, "\sheafc_{\dbasec, \dtotalc}"', maps to, color=sheaf] \& 
-                    \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction\opensetc}
-            \end{tikzcd}                    
-        \end{equation*}}
-
-    \onslide*<3|handout:3>{
-        \begin{equation*}
-            \begin{tikzcd}[ampersand replacement=\&]  
-                    \&  \& 
-                    \sheafc_{\gbasec, \vindexpullc\dtotalc} 
-                    \arrow[rr, "\vartistc_{\gbasec}", Rightarrow, color=artist] \& \& 
-                    \sheafc_{\gbasec, \gtotalc}  \\
-                    \opensetgc_1\& \& 
-                    \cgamma{\opensetgc_1}{\vindexpullc\dtotalc\restriction_{\opensetgc_1}}
-                    \arrow[dd, "\iota^*"', color=set] 
-                    \arrow[rr, "\vartistc_{\opensetgc_1}", color=artist] \& \& 
-                    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc_1}} 
-                    \arrow[dd, "\iota^*", color=set] \\
-                    \& \& {} 
-                    \arrow[rr, "\vartistc_{\sheafc}", dotted, color=artist] \& \& {} \\
-                    \arrow[uu, "\iota" description, hook, color=base] 
-                    \opensetc_2 \& \& 
-                    \cgamma{\opensetgc_2}{\vindexpullc\dtotalc\restriction_{\opensetgc_2}}
-                    \arrow[rr, "\vartistc_{\opensetgc_2}", color=artist] \& \& 
-                    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc_2}} \\
-                    \& \& 
-                    \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction\opensetgc} \& 
-                    \opensetgc 
-                    \arrow[r, "\sheafc_{\gbasec,\gtotalc}", maps to, color=sheaf] 
-                    \arrow[l, "\sheafc_{\gbasec, \vindexpullc\dtotalc}"', maps to, color=sheaf] \& 
-                    \cgamma{\opensetc}{\gtotalc\restriction\opensetgc}
-            \end{tikzcd}                    
-        \end{equation*}
-    }
-    \onslide<4|handout:4>{
-        \begin{alertblock}{pull data section \dsectionc\ over graphic space $\gbasepointc \in \gbasec$}
-            \begin{equation*}
-                \vindexpullc\dsectionc(\gbasepointc) = \dsectionc(\vindexc(\gbasepointc)) = \dsectionc(\dbasepointc)
-            \end{equation*}
-        \end{alertblock}
-        \begin{alertblock}{push graphic \gsectionc\ section over data space $\dbasepointc \in \dbasec$}
-            \begin{equation*}
-                \vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexprec(\dbasepointc)} = \gsectionc(\gbasepointc) \forall \gbasepointc \in \vindexprec(\dbasepointc)
-            \end{equation*}
-        \end{alertblock}
-        }
-\end{frame}
-
-\begin{frame}<presentation:1-|handout:1>{Reachable \gsectionc?}
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement = \&, column sep=tiny, row sep=huge]
-            \onslide*<presentation:1-|handout:1>{\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}} 
-            \&  \& 
-            \onslide*<presentation:2-|handout:1>{\imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}}} 
-            \onslide*<presentation:2-|handout:1>{\arrow[ll, "\textcolor{set}{\mathlarger{\mathlarger{\supset}}}" description, hook', color=white]} \\ 
-            \&  \& \\
-            \onslide*<presentation:1-|handout:1>{\opensetgc} 
-            \onslide*<presentation:1-|handout:1>{\arrow[uu, "\sheafc_{\gbasec,\gtotalc}", color=sheaf, maps to]} 
-            \onslide*<presentation:2-|handout:1>{\arrow[rruu, "\vartistc(\sheafc_{\dbasec,\dtotalc})" description, dashed, color=sheaf, maps to]} 
-            \&  \& 
-        \end{tikzcd}
-    \end{equation*}
-    \begin{alertblock}<presentation:2-|handout:1>{Output Subtype}
-        \begin{equation*}
-            \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} = \{\gsectionc\vert\;\exists\;\dsectionc \in \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}\;s.t.\; 
-            \vartistc(\dsectionc) = \gsectionc,\; \vindexc(\opensetgc) = \opensetc \}
-        \end{equation*}
-    \end{alertblock}
-\end{frame}
-
-\subsection{Equivariance}
-\begin{frame}<presentation:1|handout:1>{Transform Data}
-    %swap fiber & bundle, note that need section morphisms not bundle morphisms
-    \begin{block}{Fiber Category}
-        The fiber $\dfiberc$ is a  monoidal category (single object w/ bicartesian product operator on category) of an arbitrary type $\mathcal{C}$.
-The morphisms on the fiber are $\dfunctc \in Hom(\dfiberc, \dfiberc)$
-    \end{block}
-
-    \begin{block}{Fiber Bundle Category}
-        \begin{description}[style=newline]
-            \item[object] $\dfiberc \hookrightarrow \dtotalc \xrightarrow{\pi} \dbasec$
-            \item[morphisms] $\dfuncc:(\dfunchc, \dfunctc)$
+\begin{frame}{Animation and Interactivity}
+    \begin{description}
+        \item[pan, zoom, scroll] sheaf: locality + gluing \autoref{def:atct:sheaf}
+        \item[selection and hover] pushforward \autoref{eq:atct:sheaf:pushforward_select},  pullback \autoref{eq:atct:sheaf:pullback_hover}
+        \item[brushing, linking, annotation] composition of artists \autoref{eq:artist:join_base_fiber}
       \end{description}
-    \end{block}
-\end{frame}
-
-\begin{frame}<presentation:1|handout:1>{$\dfuncc = (\dfunchc, \dfunctc)$}
-    \begin{figure}
-        \includegraphics[width=0.5\textwidth]{figures/math/dfunc.png}
-    \end{figure}
-    \begin{description}
-        \item[Base Transformation:] {$\dfunchc: \opensetc^{\prime} \rightarrow \opensetc$ where $\opensetc, \opensetc^{\prime} \subseteq \dbasec$, \\
-        $\dfuncpullc\dsectionc\restriction_{\opensetc}: \dsectionc \mapsto \dsectionc\restriction_{\opensetc} \circ \dfunchc$}
-
-        \item [Fiber Transformation:]{$\dfunctc: \dfuncpullc \dtotalc_{\dbasepointc^{\prime}} \rightarrow {\dfuncpullc\dtotalc_{\dbasepointc^{\prime}}} \in Hom(\dfuncpullc\dfiberc\restriction_{\dbasepointc}, \dfuncpullc\dfiberc\restriction_{\dbasepointc}), \dbasepointc^{\prime} \in \opensetc^{\prime}$\\
-        $\dfunctc: \dfuncpullc \dsectionc\restriction_{\opensetc} \mapsto \dfuncpullc \dsectionc^{\prime}\restriction_{\opensetc}$, $\dsectionc, \dsectionc^{\prime} \in \cgamma{\opensetc^{\prime}}{\dfuncpullc\dtotalc\restriction_{\opensetc^{\prime}}}$}
-        \item [Section Transform:] $\dfuncc: \dsectionc\restriction_{\opensetc} \mapsto \dfuncpullc\dsectionc^{\prime}\restriction_{\opensetc}$
-    \end{description}
-\end{frame}
-
-\begin{frame}<presentation:1|handout:1>{Equivariant Artist}
-    $(\vartistc, \vartistc^{\prime})$ are equivariant with respect to $\dfuncc_{\dtotalc}$ if a compatible transform  $\dfuncc_{\gtotalc}$ can be defined such that 
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement=\&]
-        \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
-        \arrow[rrr, "\vartistc", color=artist] 
-        \arrow[d, "\dfuncpullc_{\dtotalc}"', color=action] 
-        \& \& \& 
-        \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
-        \arrow[d, "\dfuncpullc_{\gtotalc}", dotted] \\
-        \cgamma{\opensetc^{\prime}}{\dfuncpullc_{\dtotalc}\dtotalc\restriction_{\opensetc^{\prime}}} 
-        \arrow[dd, "\dfunctc_{\dtotalc}"', color=action] \& 
-        \opensetc 
-         \& 
-        \opensetgc 
-        \arrow[l, "\vindexc"', color=functor] 
-        \& 
-        \imartist{\opensetgc^{\prime}}{\dfuncpullc_{\gtotalc}\gtotalc\restriction_{\opensetgc^{\prime}}} 
-        \arrow[dd, "\dfunctc_{\gtotalc}", dotted, color=action] \\
-        \& 
-        \opensetc^{\prime} 
-        \arrow[u, "\dfunchc_{\dtotalc}", color=action] 
-        \& 
-        \opensetgc^{\prime} 
-        \arrow[l, "\vindexc"', color=functor] 
-        \arrow[u, "\dfunchc_{\gtotalc}"', dotted, color=action] 
-        \& \\
-        \cgamma{\opensetc^{\prime}}{\dtotalc^{\prime}\restriction_{\opensetc^{\prime}}} 
-        \arrow[rrr, "\vartistc^{\prime}", color=artist]  
-        \& \& \& 
-        \imartist{\opensetgc^{\prime}}{\gtotalc^{\prime}\restriction_{\opensetgc^{\prime}}}
-        \end{tikzcd}
-    \end{equation*}
-    For all points $\gbasepointc^{\prime}\in \gbasec^{\prime}$:
-    \begin{equation*}
-        \vartistc^{\prime}(\dfunctc_{\dtotalc}(\dsectionc(\dfunchc_{\dtotalc}(\vindexc(\gbasepointc^{\prime}))))) = \dfunctc_{\gtotalc}(\vartistc(\dsectionc(\vindexc(\dfunchc_{\gtotalc}(\gbasepointc^{\prime})))))
-    \end{equation*}
-
-\end{frame}
-
-\begin{frame}<presentation:0|handout:0>{$\vartistc= \vartistc^{\prime}$}
-    \begin{align*}
-        \vartistc(\dsectionc^{a}) = \vartistc(\dsectionc^{b}) &\centernot\implies \dsectionc^a = \dsectionc^b \\
-        \vartistc(\dsectionc^{a})  = \vartistc(\dsectionc^{b}) &\implies \vartistc(\dfuncc_{\dtotalc}(\dsectionc^{a})) = \vartistc(\dfuncc_{\dtotalc}(\dsectionc^{b}))
-    \end{align*}
 \end{frame}
@@
 
-\begin{frame}<presentation:1-|handout:->{Visual Measurement}
-    \begin{columns}
-    \column{.5\textwidth}
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement=\&]
-            \measurec 
-            \arrow[rr, "\dfunctc_{\measurec}", color=action] 
-            \& \& 
-            \measurec^{\prime}\\
-            \& \& \\
-            \opensetgc 
-            \arrow[uu, "\extractmc_{\gsectionc}"] 
-            \& \& 
-            \opensetgc^{\prime} 
-            \arrow[ll, "\dfunchc_{\gtotalc}", color=action] 
-            \arrow[uu, "\extractmc_{\dfunctc_{\gtotalc}\gsectionc}"'] 
-            \arrow[lluu, "\extractmc_{\dfuncpullc\gsectionc}" description]
-        \end{tikzcd}
-    \end{equation*}
-    \column{.5\textwidth}
-    \onslide*<presentation:2|handout:2>{
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement=\&]
-            \measurec 
-            \arrow[rr, "\dfunctc_{\measurec}", color=action] 
-            \& \& 
-            \measurec^{\prime}\\
-            \& \& \\
-            \opensetc 
-            \arrow[uu, "\equivc_{\dsectionc}"] 
-            \& \& 
-            \opensetc^{\prime} 
-            \arrow[ll, "\dfunchc_{\dtotalc}", color=action] 
-            \arrow[uu, "\equivc_{\dfunctc_{\dtotalc}\dsectionc}"'] 
-            \arrow[lluu, "\equivc_{\dfuncpullc\dsectionc}" description]
-        \end{tikzcd}
-    \end{equation*}
-    }
-    \end{columns}
-    \onslide*<presentation:1|handout:1>{
-    $\extractmc_{\gsectionc} \in Hom(\opensetgc, \measurec)$ is a mapping from graphic to visual measurement.
-$\extractmc_{\gsectionc}$ is a map from an openset $\opensetgc \subseteq \gbasec$ to a measurement $\measurec_{\opensetgc}$ that corresponds to the graphic representation at that region $\gsectionc\restriction_{\opensetgc}$ and the corresponding data $\dsectionc\restriction_{\vindexc\restriction_{\opensetgc}}$.}
-    \onslide*<presentation:2|handout:2>{
-    $\equivc_{\dsectionc} \in Hom(\opensetc, \measurec)$ is a mapping from data to visual measurement.
-$\equivc_{\dsectionc}:\dsectionc \mapsto$ is a map from an openset $\opensetc \subseteq \dbasec$ to a measurement $\measurec_{\opensetc}$ that corresponds to the data record at that region $\dsectionc\restriction_{\opensetc}$ and the corresponding graphic $\gsectionc\restriction_{\vindexprec\restriction_{\opensetc}}$
-    }
-\end{frame}
-
-\begin{frame}<presentation:1-|handout:1->
-    \frametitle<presentation:1|handout:1>{Using output $\gsectionc$ to check if $\vartistc$ is equivariant}
-    \frametitle<presentation:2|handout:2>{Using input $\dsectionc$ to check if $\vartistc$ is equivariant}
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement=\&]
-            \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
-            \arrow[rr, "\vartistc", color=artist] 
-            \arrow[d, "\dfuncc_{\dtotalc}"', color=action] 
-            \onslide*<presentation:2|handout:2>{\arrow[rrr, "\equivc", bend left, color=monoid]} \&  \& 
-            \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
-            \arrow[d, "\dfuncc_{\gtotalc}"', dotted, color=action]
-            \onslide*<presentation:1|handout:1>{\arrow[r, "\extractmc", color=monoid]}
-            \onslide*<presentation:2|handout:2>{\arrow[r,"\vindexprec \circ \extractmc", color=monoid]}
-            \& 
-            \onslide*<presentation:1|handout:1>{Hom(\opensetgc, \measurec)}
-            \onslide*<presentation:2|handout:2>{Hom(\opensetc, \measurec)} 
-            \arrow[d, "\dfuncc_{\measurec}", color=action] \\
-            \cgamma{\opensetc^{\prime}}{\dtotalc^{\prime}\restriction_{\opensetc^{\prime}}} 
-            \arrow[rr, "\vartistc", color=artist] 
-            \onslide*<presentation:2|handout:2>{\arrow[rrr, "\equivc"', bend right, color=monoid]} \&  \& 
-            \imartist{\opensetgc^{\prime}}{\gtotalc^{\prime}\restriction_{\opensetgc^{\prime}}} 
-            \onslide*<presentation:1|handout:1>{\arrow[r, "\extractmc", color=monoid]}
-            \onslide*<presentation:2|handout:2>{\arrow[r,"\vindexprec \circ \extractmc", color=monoid]} \& 
-            \onslide*<presentation:1|handout:1>{Hom(\opensetgc^{\prime}, \measurec^{\prime})}
-            \onslide*<presentation:2|handout:2>{Hom(\opensetc^{\prime}, \measurec^{\prime})}                        
-        \end{tikzcd}
-    \end{equation*}
-    \onslide*<presentation:1|handout:1>{
-        \begin{align*}
-            \vartistc^{\prime}(\dfunctc_{\dtotalc}(\dsectionc(\dfunchc_{\dtotalc}(\vindexc(\gbasepointc^{\prime}))))) &= \dfunctc_{\gtotalc}(\vartistc(\dsectionc(\vindexc(\dfunchc_{\gtotalc}(\gbasepointc^{\prime})))))\\
-            \extractmc(\dfunctc_{\gtotalc}(\gsectionc(\dfunchc_{\gtotalc})))(\gbasepointc^{\prime}) &= \dfuncc_{\measurec}(\extractmc(\gsectionc))(\gbasepointc^{\prime}) = \extractmc_{\dfunctc_{\gtotalc}\gsectionc}(\gbasepointc^{\prime})
-        \end{align*}
-    }
-    \onslide*<presentation:2|handout:2>{
-        \begin{description}
-            \item[equivariance] $\equivc(\dfunctc_{\dtotalc}(\dsectionc(\dfunchc_{\dtotalc})))(\dbasepointc^{\prime}) =\dfuncc_{\measurec}(\equivc(\dsectionc))(\dbasepointc^{\prime})= \equivc_{\dfunctc_{\dtotalc}\dsectionc}(\dbasepointc^{\prime})$ for all $\dbasepointc^{\prime} \in \dbasec^{\prime}$
-            \item[continuity] $\lim\limits_{x\to \dbasepointc}\equivc_{\dsectionc}(x) = \equivc_{\dsectionc}(\dbasepointc)$ for all $\dbasepointc \in \dbasec$ 
-        \end{description} 
-    }
-\end{frame}
-
-\begin{frame}<presentation:1|handout:1>{Complex data}
-    \begin{description}[style=newline]
-        \item[combining continuities]{given $\dbasepointc_{a} \in \dbasec_{c} \subset \dbasec_{a}$ and $\dbasepointc_{b} \in \dbasec_{c} \subset \dbasec_{b}$, if $\dbasepointc_{a} = \dbasepointc_{b}$ then $\equivc(\dsectionc(\dbasepointc_{a})) = \equivc(\dsectionc(\dbasepointc_{b}))$ for all $\dbasepointc_{a}, \dbasepointc_{b} \in \dbasec_{a}\bigsqcup\limits_{\dbasec_{c}} \dbasec_{b}$}
-        \item[shared fibers]{
-            if $p = p^{\prime}$ then $\equivc \circ q$ = $\equivc^{\prime} \circ q^{\prime}$
-        \begin{equation*}
-        \begin{tikzcd}[ampersand replacement = \&]
-        \& \dfiberc^a \times \dfiberc^b 
-        \arrow[ld, "p"', color=fiber] 
-        \arrow[rd, "q \circ \equivc"] \&     \\
-        \dfiberc^a \& \& M^a \\
-        \& \dfiberc^a \times \dfiberc^c 
-        \arrow[lu, "p^{\prime}", color=fiber] 
-        \arrow[ru, " q^{\prime} \circ\equivc^{\prime}"'] \&    
-      \end{tikzcd}
-    \end{equation*}}         
-        \item[both]{given $\dsectionc^{d} \in \sheafc_{\dbasec_{d}, \dtotalc_{d}}, \dsectionc^{e} \in \sheafc_{\dbasec_{e}, \dtotalc_{e}}$ and $\dbasepointc \in \dbasec_{d}\bigsqcup\limits_{\dbasec_{f}} \dbasec_{e}$ then $q(\equivc(\dsection^{d}(\dbasepointc))) = q^{\prime}(\equivc(\dsectionc^{e}(\dbasepointc)))$ when $p(\dfiberc^{d}\restriction_{\dbasepointc}) = p^{\prime}(\dfiberc^{e} \restriction_{\dbasepointc})$}
-    \end{description}
-\end{frame}
-
-\begin{frame}<presentation:1|handout:1>{Composable $\dfuncc = (\dfunchc, \prod\limits_{i=0}^{n}\dfunctc_i)$}
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement = \&]
-            \dtotalc_a 
-            \arrow[dd, "\dfunc a"', color=action] \& \& 
-            \dtotalc_a \times \dtotalc_b 
-            \arrow[dd, "{\dfuncc_{a,b}}" description, color=action] 
-            \arrow[ll, "\pi_a"', color=total] 
-            \arrow[rr, "\pi_b", color=total] \& \& 
-            \dtotalc_b 
-            \arrow[dd, "\dfuncc_b", color=action] \\
-            \& \& \& \& \\
-            \dtotalc_a  \& \& 
-            \dtotalc_a \times \dtotalc_b 
-            \arrow[rr, "\pi_b", color=total] 
-            \arrow[ll, "\pi_a"', color=total] \& \& 
-            \dtotalc_b
-        \end{tikzcd}
-    \end{equation*}
-    if there exists functions $\dfuncc_{a,b}: \dtotal_a \times \dtotal_b \rightarrow \dtotal_a \times \dtotal_b$, $\dfunc_{a}: \dtotal_a \rightarrow \dtotal_a$ and $\dfuncc_{b}: \dtotal_b \rightarrow \dtotal_b$ s.t.
-$\pi_a \circ \dfuncc_a = \dfuncc_{a,b} \circ \pi_a$ and $\pi_b \circ \dfuncc_b = \dfuncc_{a,b} \circ \pi_b$ then $\dfuncc_{a,b} = (\dfuncc_a, \dfuncc_b)$
-
-\end{frame}
-
 \section{Construction}
-\begin{frame}<presentation:1|handout:1>{Building an equivariant $\vartist$?}
-    \note{Add box around figure & lable it artist}
+\begin{frame}{Visual Space: Specialized data space}
+    $\vfiberc \hookrightarrow \vtotalc \xrightarrow{\pi}\dbasec$
     \begin{figure}
-        \includegraphics[width=1\textwidth]{../paper/figures/path_of_q.png}
+        \includegraphics[width=\textwidth]{figures/intro/retinal_variables.png}
     \end{figure}
@@
 
-\subsection{Measurement Specification $\vtotalc$}
+\begin{frame}{Encoding}
+    \begin{equation}
+        \label{eq:constrution:nu}
+        \vchannelc: \dfiberc_{\dbasepointc} \rightarrow \vfiberc_{\dbasepoint}
+      \end{equation}
 
-\begin{frame}<presentation:1|handout:1>{Typed Measurable Visual Components: $\vtotalc$}
-    \begin{columns}
-        \column{0.31\textwidth}
-        \begin{equation*}
-            \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-                \onslide*<1|handout:1>{\vfiberc}
-                \onslide*<1|handout:1>{\arrow[r, hook]} \& 
-                \onslide*<1|handout:1>{\vtotalc} 
-                \onslide*<1|handout:1>{\arrow[d, "\pi"']} \\
-                \& 
-                \onslide*<1|handout:1>{\dbasec} 
-                \onslide*<1|handout:1>{\arrow[u, "\vsectionc \in \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}}"', bend right, pos=.5]}
+      \begin{equation}
+        \label{eq:construction:nu:fabrication}
+        \begin{tikzcd}[ampersand replacement=\&]
+          \dfiberc_{\dbasepointc}
+          \arrow[rr, "\vchannelc", color=artist]
+          \arrow[rrrr, "\vchannelc^{\prime\prime}", dashed, bend right, color=artist] \&  \&
+          \vfiberc_{\dbasepointc}\coloneqq{\dfiberc_{\dbasepointc}^{\prime}}
+          \arrow[rr, "\vchannelc^{\prime}", color=artist] \&  \&
+          \vfiberc^{\prime}_{\dbasepointc}
         \end{tikzcd}
-        \end{equation*}
-        \column{0.69\textwidth}
-    \only<1|handout:1>{
-        \begin{description}[style=nextline]
-            \item[\textcolor{total}{Data \vtotalc}]{continuity + visual fields}
-            \item[\textcolor{base}{Continuity \dbasec}]{data continuity}
-            \item[\textcolor{fiber}{components \vfiberc}] {visual components \cite{bertinIIPropertiesGraphic2011} of a graphic, e.g. x and y position, area, color, line thickness}
-        \end{description}
-        }
-    \end{columns}
-    \pause 
-    \only<1|handout:1>{
-        \begin{alertblock}{}
-            \begin{description}[style=nextline]
-                \item[\textcolor{section}{visual components}]{$\cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} \coloneqq \big\{\vsectionc: \opensetc\rightarrow \vtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\vsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}$}
-            \end{description}
-        \end{alertblock}
-    }
+      \end{equation}
 \end{frame}
 
-\begin{frame}<presentation:1|handout:1>
-    \frametitle{Transformations in Data $\vchannelc: \dsectionc \mapsto \vsectionc$}
-    %rewrite in gamma notation
-    \begin{figure}
-        \includegraphics[.35\textwidth]{figures/math/nu_table.png}
-    \end{figure}
-    Given $\dsectionc(\dbasepointc) = \dsectionc(\dfunchc(\dbasepointc^{\prime}))$ and  $\vsectionc(\dbasepointc) = \vsectionc(\dfunchc(\dbasepointc^{\prime}))$
-    \begin{description}
-        \item[equivariance] $\dfunctc_{\vtotalc}(\vchannelc_{\opensetc}(\dsectionc))
-        = \vchannelc_{\opensetc}(\dfunctc_{\dtotalc}(\dsectionc))$, $\vchannelc_{\opensetc} = {\vchannelc^{\prime}}_{\opensetc^{\prime}}$
-        \item[continuity] $\lim\limits_{x\to \dbasepointc}(\vchannelc(\dsectionc(x))) = \vchannelc(\dsectionc(\dbasepointc))$ for all $\dbasepointc \in \dbasec$ 
-    \end{description}
+\begin{frame}{Composition}
+    \begin{equation}
+        \vmarkc: \cgamma{\dbasec}{\vtotalc} \rightarrow \cgamma{\gbasec}{\gtotalc}
+      \end{equation}
+
+      \begin{equation}
+        \label{eq:construction:q:fabrication}
+        \begin{tikzcd}[ampersand replacement=\&]
+            \cgamma{\dbasec}{\vtotalc}
+            \arrow[rr, "\vchannel", color=artist]
+            \arrow[rrrr, "\vmarkc^{\prime}", bend right, color=artist, dashed] \&  \&
+            \cgamma{\dbasec}{\vtotalc^{\prime}}
+            \arrow[rr, "\vmarkc", color=artist] \&  \& \cgamma{\gbasec}{\gtotalc}
+            \end{tikzcd}
+      \end{equation}
 \end{frame}
 
-\begin{frame}{Data to Visual Transformation $\vchannelc: \dfiberc_{\dbasepointc} \mapsto \vfiberc_{\dbasepointc}$}
-    %remove maps to 
-$\pi(\dtotalc) = \pi(\vchannelc(\dtotalc))$ and $\vchannelc$ is composable s.t 
-    \begin{columns}
-        \column{.15\textwidth}
-        \begin{equation*}
-        \begin{tikzcd}[ampersand replacement=\&]
-            \dfiberc_{\dbasepointc} 
-            \arrow[rr, "\vchannelc"] 
-            \arrow[dd, "\dfunctc_{\dtotalc}"'] 
-            \& \& 
-            \vfiberc_{\dbasepointc} 
-            \arrow[dd, "\dfunctc_{\vtotalc}"] \\
-            {} 
-            \arrow[rr, dashed]
-            \& \& {} \\
-            {\dfiberc^{\prime}}_{\dbasepointc} 
-            \arrow[rr, "\vchannelc^{\prime}"']
-            \& \& 
-            {\vfiberc^{\prime}}_{\dbasepointc}
-            \end{tikzcd}
-        \end{equation*}
-        \column{.5\textwidth} 
-        \begin{equation*}
-        \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-            \dfiberc_{\dbasepointc} 
-            \arrow[rr, "\vchannelc_{\dtotalc} ", color=artist] 
-            \arrow[rrdd, "{\vchannelc_{\dtotalc,\vtotalc}}" description, dashed, color=artist] 
-            \&  \& 
-            \vfiberc_{\dbasepointc}\coloneqq {\dfiberc_{\dbasepointc}}^{\vtotalc} 
-            \arrow[dd, "\vchannelc_{\vtotalc}", color=artist] \\
-            \& \& \\
-            \&  \& {\vfiberc_{\dbasepointc}}^{\vtotalc}
-            \end{tikzcd}
-        \end{equation*}
-    \end{columns}
-
-\end{frame}
-
-\begin{frame}<presentation:1|handout:1>\frametitle{$\vchannelc:\dfuncc_{\dtotalc} \rightarrow \dfuncc_{\vtotalc}$: Stevens' Scales \cite{stevensTheoryScalesMeasurement1946}}
-    \begin{table}[H]
-        \begin{tabularx}{\textwidth}{|l|l|X|}\toprule
-            \textbf{scale} & \textbf{group} & \textbf{constraint} \\\midrule
-            nominal & permutation &  $\text{if } \delement_1 \neq \delement_2 \text{ then } \vchannel (\delement_1) \neq\vchannel(\delement_2)$\\
-            ordinal &  monotonic & $\text{if } \delement_1 \leq \delement_2 \text{ then } \vchannel (\delement_1) \leq \vchannel(\delement_2)$\\
-            interval &  translation &  $\vchannel (\delement + c) = \vchannel(\delement) + \vchannel(c)$ \\
-            ratio &  scaling &  $\vchannel(\delement* c) = \vchannel(\delement)*\vchannel(c) $\\ \bottomrule
-        \end{tabularx}
-    \end{table}
-\end{frame}
-
-\begin{frame}<presentation:1|handout:1>
-    \frametitle{Shared Components: $\vchannelc = \prod\limits_{i=0}^{n}\vchannelc_i$}
-    \begin{equation}
-        \begin{tikzcd}[ampersand replacement=\&]
-            \dfiberc_a\times \dfiberc_b 
-            \arrow[rr, "\vchannelc", color=artist] 
-            \arrow[rd, "p_{\dfiberc}", color=fiber] \& \& \vfiberc_a\times \vfiberc_b 
-            \arrow[rd, "p_{\vfiberc}", color=fiber] \& \\    \& 
-            \dfiberc_{a} 
-            \arrow[rr, "\vchannelc_{a}", dashed, color=artist] \& \& \vfiberc_a \\
-            \dfiberc_a\times \dfiberc_c 
-            \arrow[rr, "\vchannelc^{\prime}", color=artist] 
-            \arrow[ru, "p_{\dfiberc^{\prime}}"', color=fiber] \& \& 
-            {\vfiberc_a}^{\prime}\times {\vfiberc_c}^{\prime} 
-            \arrow[ru, "p_{\vfiberc^{\prime}}"', color=fiber] \&           
-            \end{tikzcd}
-    \end{equation}
-    \begin{alertblock}{Consistent Transformations \cite{hullmanKeeping2018}}
-        if $p_{\dfiberc} = p_{\dfiberc^{\prime}}$ then $p_{\vfiberc}(\vchannelc(\dsectionc)) = p_{\vfiberc^{\prime}}(\vchannelc^{\prime}(\dsectionc^{\prime}))$ s.t.
-there exists a transformation $\vchannelc_a: \dfiberc_{a}\rightarrow \vfiberc_{a}$
-    \end{alertblock}
-\end{frame}
-
-\subsection{Combining Components into Visualizations}
-\begin{frame}<presentation:1|handout:1>{Assembly $\vmarkc$}
-    \begin{equation*}
-        \begin{tikzcd}[ampersand replacement = \&]
-            \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} 
-            \arrow[rrrr, "\vmarkc", color=artist] 
-            \arrow[ddd, "\dfunctc_{\vtotalc} \circ \dfuncpullc_{\vtotalc}"', color=action] 
-            \& \& \& \& 
-            \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
-            \arrow[ddd, "\dfunctc_{\gtotalc} \circ \dfuncpullc_{\gtotalc}", dotted, color=action]  \\
-            \& 
-            \opensetc 
-             \&  \& 
-            \opensetgc 
-            \arrow[ll, "\vindexc"', color=functor]   \& \\
-            \& 
-            \opensetc^{\prime} 
-            \arrow[u, "\dfunchc_{\dtotalc}", color=action] 
-            \&  \& 
-            \opensetgc^{\prime} 
-            \arrow[ll, "\vindexc"', color=functor] 
-            \arrow[u, "\dfunchc_{\gtotalc}"', dotted, color=action] 
-            \& \\
-            \cgamma{\opensetc^{\prime}}{\vtotalc^{\prime}\restriction_{\opensetc^{\prime}}}
-            \arrow[rrrr, "\vmarkc^{\prime}", color=artist] 
-            \& \&  \& \& 
-            \imartist{\opensetgc^{\prime}}{\gtotalc^{\prime}\restriction_{\opensetgc^{\prime}}}
-            \end{tikzcd}
-        \end{equation*}
-    \begin{alertblock}{equivariance}
-       \begin{equation*}
-        \vmarkc^{\prime}(\dfunctc_{\vtotalc}(\vsectionc(\dfunchc_{\dtotalc}(\vindexc(\gbasepointc^{\prime}))))) = \dfunctc_{\gtotalc}(\vmarkc(\vsectionc(\vindexc(\dfunchc_{\gtotalc}(\gbasepointc^{\prime})))))
-       \end{equation*}
-    \end{alertblock}
-\end{frame}
-
-\begin{frame}<presentation:1|handout:1>{Components to Graphic}
+\begin{frame}{Construction Stages}
     \begin{figure}
-        \includegraphics[scale=.15]{../paper/figures/qhat.png}
+        \includegraphics[width=\linewidth]{../paper/figures/construction_nocomp.pdf}
     \end{figure}
@@
 
-\begin{frame}<presentation:1|handout:1>{Implementation Choices: $\vartistc_{\dbasec} = \vartistc_{\gbasec}$}
+\begin{frame}{Construction Stages: composition}
     \begin{figure}
-        \includegraphics[scale=.5]{figures/math/implement_paths).png}
+        \includegraphics[width=\linewidth]{../paper/figures/construction_comp.pdf}
     \end{figure}
@@
 
-\begin{frame}{Combining Qs}
-    The codomain of all $\vmarkc$ targeting the same output space is the bundle $\gtotalc$ and $Hom(\gbasec_1, \gtotalc) + Hom(\gbasec_2, \gtotalc) = Hom(\gbasec_1 + \gbasec_2, \gtotalc)$; therefore 
-    $\sqcup \vmarkc_i(\cgamma{\opensetc_i}{\dtotalc_i\restriction_{\opensetc_i}}) = \cgamma{\sqcup \opensetgc_i}{\gtotalc\restriction_{\sqcup\opensetgc_i}}$
+\begin{frame}{Verify encoder}
+    \begin{equation}
+        \label{eq:construction:nu:validate}
+        \begin{tikzcd}[ampersand replacement = \&, column sep=4em]
+          {\dfiberc_{\dbasepointc}^{a}} \times {\dfiberc_{\dbasepointc}^{b}}
+          \arrow[d, "\pi_a"', color=fiber]
+          \arrow[r, "\vchannelc_{ab}", color=artist]
+          \arrow[rr, "\equivc_{ab}", bend left, color=monoid]  \&
+          {\vfiberc_{\dbasepointc}^{a}} \times {\vfiberc_{\dbasepointc}^{b}}
+          \arrow[d, "\pi_a", color=fiber] \&
+          \measurec_{\dbasepointc}^{ab}
+          \arrow[d, "\measurec\restriction_a", color=set] \\
+          \dfiberc_{\dbasepointc}^a
+          \arrow[r, "\vchannelc_{a}", dashed, color=artist] \&
+          \vfiberc_{\dbasepointc}^a
+          \arrow[r, "\simeq", dotted]  \&
+          \measurec_{\dbasepointc}^a   \\
+          {\dfiberc_{\dbasepointc}^{a}} \times {\dfiberc_{\dbasepointc}^{c}}
+          \arrow[u, "\pi_a", color=fiber]
+          \arrow[r, "\vchannelc_{ac}", color=artist]
+          \arrow[rr, "\equivc_{ac}", bend right, color=monoid] \&
+          {\vfiberc_{\dbasepointc}^{a}} \times {\vfiberc_{\dbasepointc}^{c}}
+          \arrow[u, "\pi_a"', color=fiber] \&
+          \measurec_{\dbasepointc}^{ac}
+          \arrow[u, "\measurec\restriction_a"', color=set]
+          \end{tikzcd}
+      \end{equation}
+\end{frame}
+\begin{frame}{Verify encoder}
     \begin{figure}
-        \includegraphics[scale=.25]{figures/math/qcom.png}
+        \includegraphics[width=\linewidth]{../paper/figures/encoder.pdf}
     \end{figure}
@@
 
-\begin{frame}{Compatible Qs}
-    \includegraphics[scale=.5]{figures/math/compatiblw_q.png}
+\begin{frame}{Verify compositor}
+    \begin{equation}
+        \label{eq:construction:q:validate}
+        \begin{tikzcd}[ampersand replacement = \&, row sep=huge]
+          \cgamma{\dbasec}{\vtotalc^{a}\times \vtotalc^{b}}
+          \arrow[rr, "\vmarkc_{ab}", color=artist]
+          \arrow[d, "\pi_a"', color=total] \&  \&  \imartistsub{ab}{\gbasec}{\gtotalc}
+          \arrow[d, "\measurec\restriction_a \circ \extractmc_{ab}", color=monoid]  \\
+         \cgamma{\dbasec}{\vtotalc^{a}}
+         \arrow[rr, "\simeq", dotted] \&  \&
+         {\textcolor{set}{Hom}(\dbasec, \measurec^{a})}  \\
+          \cgamma{\dbasec}{\vtotalc^{a}\times \vtotalc^{c}}
+          \arrow[rr, "\vmarkc_{ac}", color=artist]
+          \arrow[u, "\pi_a", color=total]  \&  \&  \imartistsub{ac}{\gbasec}{\gtotalc}
+          \arrow[u, "\measurec\restriction_a \circ \extractmc_{ac}"', color=monoid]
+         \end{tikzcd}
+      \end{equation}
+\end{frame}
+\begin{frame}{Verify compositor}
+    \begin{figure}
+        \includegraphics[width=\linewidth]{../paper/figures/compositor.pdf}
+    \end{figure}
+\end{frame}
+
+\section{Implementation}
+\begin{frame}<presentation:1|handout:1>{Implementation Choices: $\vartistc_{\dbasec} = \vartistc_{\gbasec}$}
+    \begin{equation*}
+        \begin{tikzcd}[ampersand replacement=\&, column sep=small]
+            \cgamma{\opensetgc}{\vindexpullc\dtotalc\restriction_{\opensetgc}}
+            \arrow[rr, "\vchannelc_{\gbasec}", color=artist]
+            \arrow[rrrr, "\vartistc_{\gbasec}", bend left, color=artist]
+            \&  \&
+            \cgamma{\opensetgc}{\vindexpullc\vtotalc\restriction_{\opensetgc}}
+            \arrow[rr, "\vmarkc_{\gbasec}", color=artist] \&  \&
+            \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}}
+            \arrow[dd, "\vindexpushc", color=functor] \\
+             \&  \&   \&  \& \\
+            \cgamma{\openset}{\dtotalc\restriction_{\opensetc}}
+            \arrow[rr, "\vchannelc_{\dbasec}", color=artist]
+            \arrow[rrrr, "\vartistc_{\dbasec}", bend right, color=artist]
+            \arrow[uu, "\vindexpullc", color=functor] \&  \&
+            \cgamma{\openset}{\vtotalc\restriction_{\opensetc}}
+            \arrow[rr, "\vmarkc_{\dbasec}", color=artist ]
+            \arrow[uu, "\vindexpullc", , color=functor] \&  \&
+            \imartist{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}}
+            \end{tikzcd}
+    \end{equation*}
+\end{frame}
+
+\begin{frame}[allowframebreaks]{References}
+    \printbibliography
 \end{frame}

### 2025-04-03 paper/main.tex
fixed xi figure
talk version
spivak table

@@
 
-\usepackage[outputdir=.]{minted}
+\usepackage[]{minted}
 \setminted[python]{fontsize=\scriptsize,
@@
 \usepackage{tabulary}
+\usepackage{tabularray}
 % IEEEtran contains the IEEEeqnarray family of commands
@@
 
-\title{Topological Equivariant Artist Model for Visualization Library Architecture}
+\title{Topological Equivariant Artist Model}
 % author names and IEEE memberships
@@
 
-The standard construction of a category from a topological space is that it has open set objects $\opensetc$ and inclusion morphisms $\opensetc_i \xrightarrow{\iota} \opensetc_j$ such that $\opensetc_i \subseteq \opensetc_j$\cite{riehlCategoryTheoryContext}.
-The composability property expresses that inclusion is transitive, while associativity expresses that the inclusion functions can be curried in various equivalent groupings.
+The standard construction of a category from a topological space is that it has open set objects $\opensetc$ and morphisms $\opensetc_i \xrightarrow{\hat{\phi}} \opensetc_j$\cite{riehlCategoryTheoryContext}.
+
+\note{move $\hat{\phi}$ here}
+
+The composability property expresses that the morphism is transitive, while associativity expresses that the morphisms can be curried in various equivalent groupings.
 By formally specifying the properties of the topological structure data types as $\mathcal{\dbase}$, we can express that these are the properties that are required as part of the implementation of the data type objects.
@@
 
-\subsubsection{\textcolor{fiber}{Data Field Types: Fiber Space \dfiber}}
+\subsubsection{\textcolor{fiber}{Field Data Types: Fiber Space \dfiber}}
 \label{sec:atct:fb:fiber}
@@
 Spivak shows that data typing can be expressed in a categorical framework in his fiber bundle formulation of tables in relational databases \cite{spivakDatabasesAreCategories2010,spivakSimplicialDatabases2009}.
-In this work, we adopt Spivak's definitions of \textit{type specification}, \textit{schema}, and \textit{record} because that allows us to use a dimension agnostic named typing system for the fields of our dataset that is consistent with the abstraction we are using to express the continuity.
+We adopt Spivak's definitions of \textit{type specification}, \textit{schema}, \textit{record}, and \textit{table} because that allows us to use a dimension agnostic named typing system for the dataset fields that is consistent with the abstraction we are using to express the continuity.
+
 Spivak introduces a \textit{type specification} as a bundle map $\pi: \mathscr{U} \rightarrow \textbf{DT}$.
-The base space $\textbf{DT}$ is a set of data types $T \in \textbf{DT}$ and the total space $\mathscr{U}$ is the disjoint union of the domains of each type \[\mathscr{U} = \bigsqcup_{T \in \textbf{DT}} \pi^{-1}(T)\] such that each element $x$ in the domain $\pi^{-1}(T)$ is one possible value of an object of type $T$ \cite{spivakSimplicialDatabases2009}.
+The base space $\textbf{DT}$ is a set of data types $T \in \textbf{DT}$ and the total space $\mathscr{U}$ is the disjoint union of the domains of each type $\mathscr{U} = \bigsqcup_{T \in \textbf{DT}} \pi^{-1}(T)$ such that each element $x$ in the domain $\pi^{-1}(T)$ is one possible value of an object of type $T$ \cite{spivakSimplicialDatabases2009}.
 For example, if $T=\texttt{int}$, then the image $\pi^{-1}(\texttt{int}) = \mathbb{Z} \subset \mathscr{U}$ is the set of all integers and $x=3 \in \mathbb{Z}$ is the value of one $\texttt{int}$ object.
-
-Since many fields can have the same datatype, Spivak formally defines a mapping from field name to field data type, akin to a database schema \cite{ullmanFirstCourseDatabase2008}.
-According to Spivak, a \textit{schema} consists of a pair $(\fnames, \sigma)$ where $\fnames$ is the set of field names and $\sigma: \fnames \rightarrow \textbf{DT}$ is a function from field name to field data type\cite{spivakSimplicialDatabases2009}.
+Spivak's analogue of a database \textit{schema}\cite{ullmanFirstCourseDatabase2008} is the \textit{simple schema of type $\pi$} which is defined as consisting of the pair $(\fnames, \sigma)$ where $\fnames$ is the set of field names and $\sigma: \fnames \rightarrow \textbf{DT}$ is a function from field name to field data type\cite{spivakSimplicialDatabases2009}.
 The function $\sigma$ is composed with $\pi$ such that $\pi^{-1}(\sigma(C)) \subseteq \mathscr{U}$; this composition induces a domain bundle $\pi_{\sigma}:\mathscr{U}_{\sigma} \rightarrow \fnames$ that associates a field name $c \in C$ with its corresponding domain $\pi^{-1}_{\sigma}(C) \subseteq \mathscr{U}_{\sigma}$.
-\begin{definition} A \textbf{record} is a function $\delement: \fnames \rightarrow \mathscr{U}_{\sigma}$ and the set of records on $\pi_{\sigma}$ is denoted $\Gamma^{\pi}(\sigma)$.
+He defines a \textit{record} as a function $\delement: \fnames \rightarrow \mathscr{U}_{\sigma}$ from column name to type domain and the set of records on $\pi_{\sigma}$ is denoted $\Gamma^{\pi}(\sigma)$.
 Records must return an object of type $\sigma(\fname) \in \textbf{DT}$ for each field $c \in C$.
-\end{definition}
 Spivak then describes tables as sections $\dsection: \dbase \rightarrow \Gamma^{\pi}(\sigma)$ from an indexing space $\dbase$ to the set of all possible records $\Gamma^{\pi}(\sigma)$ on the schema bundle, and his notion of a table generalizes to our notion of a data container.
 
-To build on the rich typing system provided by Spivak, we define the \textcolor{fiber}{fiber space} \dfiberc\ to be the space of all possible data records
+In summary, Spivak defines types as maps between type name (e.g.
+$temperature \in C$), data type ($\texttt{float} \in \texttt{DT}$), and type domain $\mathbb{R}\subseteq \mathscr{U_{\sigma}} \subseteq \mathscr{U}$).
+Records are maps from name to value ($\textit{temperature}:152.0^{\circ} F$) and tables are maps from index to record \cite{spivakSimplicialDatabases2009}:
+\begin{table}[H]
+  \begin{equation*}
+    \begin{tabular}{|c|c|c|}
+      \hline
+       types and schema & \textcolor{fiber}{record} & \textcolor{section}{table}\\
+      \hline
+      \begin{tikzcd}[ampersand replacement=\&, sep=huge]
+        \textcolor{total}{\mathcal{U}_\sigma} \arrow[d, "\pi_{\sigma}"', color=total, two heads] \arrow[r, dotted] \arrow[rd, dotted] \& \textcolor{total}{\mathcal{U}} \arrow[d, "\pi"', color=total, two heads] \\
+        \textcolor{base}{C} \arrow[r, "\sigma"']                                                                                   \& \textcolor{base}{\texttt{DT}}
+      \end{tikzcd}
+    &
+    \begin{tikzcd}[ampersand replacement=\&, sep=huge]
+      \textcolor{total}{\mathcal{U}_{\sigma}} \arrow[d, "\pi_{\sigma}"', color=total, two heads] \\
+      \textcolor{base}{C} \arrow[u, "\delementc"', dashed, bend right, color=fiber]
+      \end{tikzcd}
+    &
+    \begin{tikzcd}[ampersand replacement=\&, sep=huge]
+      \dfiberc\coloneqq \textcolor{total}{\Gamma^{\pi}(\sigma)} \arrow[d, "\pi"', color=total, two heads] \\
+      \dbasec \arrow[u, "\dsectionc"', dashed, bend right, color=section]
+      \end{tikzcd}\\
+      \hline
+    \end{tabular}
+    \end{equation*}
+    \caption{}
+    \label{tab:fiber_specification}
+  \end{table}
+
+Building on Spivak, we define the \textcolor{fiber}{fiber space} \dfiberc\ to be the space of all possible data records:
 \begin{equation}

### 2025-04-07 paper/main.tex
spivak section rework

@@
 
-\paragraph{Joining indexing spaces: $\oplus: \mathcal{\dbasec} \sqcup \mathcal{\dbasec} \rightarrow \mathcal{\dbasec}$}
-Joining two bundles on their base space, for example to glue the two trivializations in \autoref{fig:atct:fb} into a bundle, is the coproduct $\dbase_{a} \sqcup_{\dbase_{c}} \dbase_{b}$ over a space $\dbase_{c}$.
-A property of the coproduct is that the inclusion morphism must be commutative, which means that index spaces are combined on overlapping indices:
-\begin{figure}[h]
-  \includegraphics*[width=1\columnwidth]{figures/tex/k_coproduct.pdf}  \caption{Index spaces are combined via the \textit{coproduct} $\dbase_{a}\sqcup_{\dbase_c} \dbase_{b}$.}
-  \label{fig:atct:base}
-\end{figure}
- For example, in \autoref{fig:atct:base}, the open interval $\dbase_c$ is present in both $\dbase_a$ and $\dbase_c$; therefore $\dbase_a$ and $\dbase_b$ are glued together where they overlap at $\dbase_c$.
-Two spaces have been combined correctly if every point in the overlap  $\dbasepoint \in \dbase^{c}$ is present in each of the spaces $\dbasepoint \in \dbase^{a}$ and $\dbasepoint \in \dbase^{b}$ such that it is present in the joined space $\dbasepoint \in \dbase^{a}\sqcup_{\dbase^{c}} \dbase_{b}$.
-This simple test that the records are joined correctly is what allows us to reliably build larger datasets out of smaller ones, such as in the case of distributed and on demand datasets.
-
 \subsubsection{\textcolor{fiber}{Field Data Types: Fiber Space \dfiber}}
@@
 Spivak's analogue of a database \textit{schema}\cite{ullmanFirstCourseDatabase2008} is the \textit{simple schema of type $\pi$} which is defined as consisting of the pair $(\fnames, \sigma)$ where $\fnames$ is the set of field names and $\sigma: \fnames \rightarrow \textbf{DT}$ is a function from field name to field data type\cite{spivakSimplicialDatabases2009}.
+
 The function $\sigma$ is composed with $\pi$ such that $\pi^{-1}(\sigma(C)) \subseteq \mathscr{U}$; this composition induces a domain bundle $\pi_{\sigma}:\mathscr{U}_{\sigma} \rightarrow \fnames$ that associates a field name $c \in C$ with its corresponding domain $\pi^{-1}_{\sigma}(C) \subseteq \mathscr{U}_{\sigma}$.
@@
 Records must return an object of type $\sigma(\fname) \in \textbf{DT}$ for each field $c \in C$.
-Spivak then describes tables as sections $\dsection: \dbase \rightarrow \Gamma^{\pi}(\sigma)$ from an indexing space $\dbase$ to the set of all possible records $\Gamma^{\pi}(\sigma)$ on the schema bundle, and his notion of a table generalizes to our notion of a data container.
-
-In summary, Spivak defines types as maps between type name (e.g.
-$temperature \in C$), data type ($\texttt{float} \in \texttt{DT}$), and type domain $\mathbb{R}\subseteq \mathscr{U_{\sigma}} \subseteq \mathscr{U}$).
-Records are maps from name to value ($\textit{temperature}:152.0^{\circ} F$) and tables are maps from index to record \cite{spivakSimplicialDatabases2009}:
-\begin{table}[H]
+Spivak then describes tables as sections $\dsection: \dbase \rightarrow \Gamma^{\pi}(\sigma)$ from an indexing space $\dbase$ to the set of all possible records $\Gamma^{\pi}(\sigma)$ on the schema bundle, and his construction of a table generalizes to our notion of a dataset.
+
+In summary, Spivak defines \textit{types} as map $\sigma$ from type name (e.g.
+$temperature \in C$) to data type ($\texttt{float} \in \texttt{DT}$) and maps $\pi_{\sigma}, \pi$ from type domain ($\mathbb{R}\subseteq \mathscr{U_{\sigma}} \subseteq \mathscr{U}$) to name and type, \textit{schemas} as maps from type name to data type ($temperature: \texttt{float}$), \textit{records} as maps from name to value ($temperature:152.0^{\circ} F$) and \textit{tables} as maps from index to record \cite{spivakSimplicialDatabases2009}:
 \begin{equation*}
@@
       \textcolor{total}{\mathcal{U}_{\sigma}} \arrow[d, "\pi_{\sigma}"', color=total, two heads] \\
-      \textcolor{base}{C} \arrow[u, "\delementc"', dashed, bend right, color=fiber]
+      \textcolor{base}{C} \arrow[u, "\delement"', dashed, bend right, color=fiber]
       \end{tikzcd}
@@
 \end{equation*}
-    \caption{}
-    \label{tab:fiber_specification}
-  \end{table}
-
-Building on Spivak, we define the \textcolor{fiber}{fiber space} \dfiberc\ to be the space of all possible data records:
+
+Building on Spivak's constructions, we define the \textcolor{fiber}{fiber space} \dfiberc\ to be the space of all possible data records:
 \begin{equation}
@@
 Steven's other scales are summarized in \autoref{tab:appendix:summary:stevens}.
-
-\paragraph{Merging fields: $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$ }
 
@@
 
-Combining fibers correctly can be verified by checking that when a fiber component $\dfiber^{c}$ is present in both $\dfiber^{a}$ and $\dfiber^{b}$, it is identical when projected out of either such that the \textit{product} diagram commutes, means that when data with many fields is decomposed into its component fields, records maintain their integrity:
+\subsubsection{Adding and expanding records}
+
+A common use of category theory in software engineering is the specification of modular components \cite{wielsManagementEvolvingSpecifications1998} such that we can build systems where the structure preserved by components is preserved in the composition of the components.
+As described in \autoref{sec:atct:fb:base}, the index space is modeled as a set of open sets; therefore adding index spaces $\dbase_a$ and $\dbase_b$ is the coproduct $\dbase_{a} \sqcup_{\dbase_{c}} \dbase_{b}$ over a shared space $\dbase_{c}$.
+As described in \autoref{sec:atct:fb:fiber}, the bifunctor can be used to build composite fiber types;  in the same manner fields $\dfiber_a$ and $\dfiber_b$ can be composed by taking the fiber product $\dfiber_{a}\times_{\dfiber_{c}} \dfiber_b$ over a shared fiber $\dfiber_c$.
 
 \begin{figure}[H]
-  \includegraphics*[width=1\columnwidth]{figures/tex/f_product.pdf}
-  \caption{Fields are combined via the \textit{product} $\dfiber_{a}\times_{\dfiber_c} \dfiber_{b}$.}
-  \label{fig:atct:fiber}
+  \includegraphics*[width=1\columnwidth]{figures/tex/more_records.pdf}
+  \caption{Records are added to the dataset as a \textit{coproduct} $\dbase_{a}\sqcup_{\dbase_c} \dbase_{b}$ of index spaces.
+Fields are added to each record as a \textit{product} $\dfiber_{a}\times_{\dfiber_c} \dfiber_{b}$ of fiber spaces.}
+  \label{fig:atct:joint}
 \end{figure}
 
-For example in \autoref{fig:atct:fiber} the red (temperature, time, pressure) record separates into (temperature, time) and (pressure, time) records that share the same red time.
+For example, in \autoref{fig:atct:joint}, the open interval $\dbase_c$ is present in both $\dbase_a$ and $\dbase_c$; therefore $\dbase_a$ and $\dbase_b$ are glued together where they overlap at $\dbase_c$.
+Two spaces have been combined correctly if every point in the overlap  $\dbasepoint \in \dbase^{c}$ is present in each of the spaces $\dbasepoint \in \dbase^{a}$ and $\dbasepoint \in \dbase^{b}$ such that it is present in the joined space $\dbasepoint \in \dbase^{a}\sqcup_{\dbase^{c}} \dbase_{b}$.
+In \autoref{fig:atct:joint}, the red (temperature, time, pressure) record separates into (temperature, time) and (pressure, time) records that share the same red time.
 Furthermore this time is the same whether it is obtained from the (temperature, time) or (pressure, time) record.
 Generally, the test for joining fields is that when a record is present in a fiber $\delement_{c} \in \pi_c(\dfiber_{a}), \delement_c \in \pi_c(\dfiber_{b})$ then it is in the joint record $\delement_{c} \in \pi_c(\dfiber_{a}\times_{\dfiber_c}\dfiber_{b})$ when is in the field $\dfiber_c$ being joined on.
-This simple test that fields are joined together correctly for the same record is what allows us to reliably combine multiple datasets together on shared properties, for example growing the weather station data in \autoref{fig:related-work:continuity:ktypes} from a temporal to spatial dataset by adding the weather at each location at each time.
+Verifying that indicies are joined correctly allows us to reliably build larger datasets out of smaller ones, such as in the case of distributed and on demand datasets.
+Verifying that fields are joined together correctly for the same record allows us to reliably combine multiple datasets together on shared properties, for example growing the weather station data in \autoref{fig:related-work:continuity:ktypes} from a temporal to spatial dataset by adding the weather at each location at each time.
 
@@
 \note{large big data glued together correctly}
-A common use of category theory in software engineering is the specification of modular components \cite{wielsManagementEvolvingSpecifications1998} such that we can build systems where the structure preserved by components is preserved in the composition of the components.
  This allows us to express that an artist that works on a dataset can be composed of artists that work on sub parts of that dataset.

### 2025-06-13 paper/main.tex
other misc fixes + abstract

@@
   The contract data visualization tools make with their users is that a chart is a faithful and accurate visual representation of the numbers it is made from.
-Motivated by wanting to make better tools, we propose a methodology for fully specifying arbitrary data to visualization mappings in a manner that easily translates to code.
+Motivated by wanting to make tools that are more flexible while still rigorously structure preserving, we propose a methodology for fully specifying arbitrary data to visualization mappings in a manner that translates to code.
 We propose that fiber bundles provide a uniform interface for describing a variety of underlying data - tables, images, networks, etc.
@@
 \begin{LaTeXdescription}
-  \item[\textcolor{fiber}{\textbf{field}} \footnote{Throughout this paper, we use color in definitions, equations, and visualizations to group conceptually related terms\cite{headMathAugmentationHow2022}, for example \textcolor{fiber}{field} and \textcolor{fiber}{fiber}.}] is a set of values of the same type, e.g.
+  \item[\textcolor{fiber}{\textbf{field}}]\footnote{In this paper, color in definitions, equations, and visualizations is used to group conceptually related terms\cite{headMathAugmentationHow2022}, for example \textcolor{fiber}{field} and \textcolor{fiber}{fiber}.} is a set of values of the same type, e.g.
 one column of a table or the pixels of an image
@@
 The function $\sigma$ is composed with $\pi$ such that $\pi^{-1}(\sigma(C)) \subseteq \mathscr{U}$; this composition induces a domain bundle $\pi_{\sigma}:\mathscr{U}_{\sigma} \rightarrow \fnames$ that associates a field name $c \in C$ with its corresponding domain $\pi^{-1}_{\sigma}(C) \subseteq \mathscr{U}_{\sigma}$.
-He defines a \textit{record} as a function $\delement: \fnames \rightarrow \mathscr{U}_{\sigma}$ from column name to type domain and the set of records on $\pi_{\sigma}$ is denoted $\Gamma^{\pi}(\sigma)$.
+He defines a \textit{record} as a function $\delement: \fnames \rightarrow \mathscr{U}_{\sigma}$ from column name to type domain and the set of records on $\pi_{\sigma}$ is denoted $\Gamma^{\pi}(\sigma)$\footnote{This paper follows the convention that $\Gamma$ denotes a set.}.
 Records must return an object of type $\sigma(\fname) \in \textbf{DT}$ for each field $c \in C$.

### 2025-07-21 paper/main.tex
new laptop sync

@@
-\documentclass[journal]{IEEEtran}
+\documentclass[review]{vgtc}
 
@@
 
+% Fonts
+
+\usepackage{times}                     % we use Times as the main font
+\renewcommand*\ttdefault{txtt}         % a nicer typewriter font
+\usepackage{mathptmx}
+
 % *** SUBFIGURE PACKAGES ***
@@
 % *** PDF, URL AND HYPERLINK PACKAGES ***
-\usepackage{xurl}
+%\usepackage{xurl}
 
@@
 \usepackage[]{footmisc}
-\usepackage{bookmark}
-\usepackage[style=ieee, backend=biber]{biblatex}
-\addbibresource{references.bib}
 
@@
 
-\begin{document}
-\linenumbers
-
-\title{Topological Equivariant Artist Model}
-% author names and IEEE memberships
-\author{Hannah~Aizenman, Mikael Vejdemo-Johansson, Thomas~Caswell, and~Michael~Grossberg,~\IEEEmembership{Member,~IEEE,}% <-this % stops a space
-\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem H. Aizenman is with the department of Computer Science, The Gradaute Center, CUNY.
-\protect\\
-% note need leading \protect in front of \\ to get a newline within \thanks as
-% \\ is fragile and will error, could use \hfil\break instead.
-E-mail: haizenman@gradcenter.cuny.edu,
-\IEEEcompsocthanksitem M. Grossberg is with the department of Computer Science, City College of New York, CUNY.
-E-mail: mgrossberg@ccny.cuny.edu
-\IEEEcompsocthanksitem Mikael Vejdemo-Johansson is with the department of Mathematics, CUNY College of Staten Island.
-\protect\\
-E-mail: mvj@math.csi.cuny.edu
-\IEEEcompsocthanksitem Thomas Caswell is with National Synchrotron Light Source II, Brookhaven National Lab
-\protect \\
-E-mail: tcaswell@bnl.gov}% <-this % stops an unwanted space
-\thanks{Manuscript received X XX, XXXX; revised X XX, XXXX.}
-}
+\onlineid{0}
+
+%% declare the category of your paper, only shown in review mode
+\vgtccategory{Research}
+
+%% allow for this line if you want the electronic option to work properly
+\vgtcinsertpkg
+
+\title{Categorical Framework for Visualization Software Design}
+
+\author{Hannah Aizenman \thanks{email:haizenman@gradcenter.cuny.edu}\\%
+        \scriptsize Computer Science, The Graduate Center, CUNY
+\and Mikael Vejdemo-Johansson \thanks{e-mail: mvj@math.csi.cuny.edu}\\%
+        \scriptsize  Mathematics, CUNY College of Staten Island %
+\and Thomas Caswell \thanks{e-mail: tcaswell@bnl.gov}\\%
+  \scriptsize National Synchrotron Light Source II, Brookhaven National Lab%
+\and Michael Grossberg \thanks{e-mail: mgrossberg@ccny.cuny.edu}\\%
+  \scriptsize Computer Science, City College of New York, CUNY}
 
@@
 % in the abstract or keywords.
-\IEEEtitleabstractindextext{%
-\begin{abstract}
+
+\abstract{
   The contract data visualization tools make with their users is that a chart is a faithful and accurate visual representation of the numbers it is made from.
@@
 Using category theory to formally express how visual elements are constructed means we can translate those expectations into code, which can then be used to enforce the expectation that a visualization tool is faithfully translating between numbers and charts.
-\end{abstract}
-
-% Note that keywords are not normally used for peerreview papers.
-\begin{IEEEkeywords}
-%Computer Society, IEEE, IEEEtran, journal, \LaTeX, paper, template.
-\end{IEEEkeywords}}
-
+}
+
+\keywords{Visualization software design, visualization theory, categorical framework}
+\begin{document}
 % make the title area
+
+\firstsection{Introduction}\label{sec:intro}
+
 \maketitle
-\IEEEpeerreviewmaketitle
-\IEEEraisesectionheading{\section{Introduction}\label{sec:intro}}
-
-\IEEEPARstart{V}isualization design guidelines, generally, describe how to choose visual encodings that preserve the structure of the data; to follow these guidelines the visualization tools that implement these $data \rightarrow graphic$ transforms must be structure preserving.
+
+Visualization design guidelines, generally, describe how to choose visual encodings that preserve the structure of the data; to follow these guidelines the visualization tools that implement these $data \rightarrow graphic$ transforms must be structure preserving.
 Loosely, preserving structure means that the properties of the data and how the points are connected to each other should be inferable from the graphic such that a $graphic \rightarrow data$ mapping can be made.
@@
 Data is often described by its mathematical structure, for example the Steven's measurement scales define nominal, ordinal, interval, and ratio data by the allowed operations on each \cite{stevensTheoryScalesMeasurement1946} and other researchers have since expanded the scales to encapsulate more types of structure \cite{leaFormalizationMeasurementScale1971, thomasMathematizationNotMeasurement2014}.
-
-Loosely, the scales classify data as a set of values and the allowed transformations on that set, which can be operations, relations, or generalized as actions:
-
-\begin{definition}\label{def:related-work:action}\cite{grimaldiDiscreteCombinatorialMathematics2006}
-  An \textcolor{action}{\textbf{action}} of \textcolor{action}{$G = (G,\circ, e)$} on $X$ is a function  $act: \textcolor{action}{G} \times X \rightarrow X$.
-An action has the properties of identity $act(\textcolor{action}{e}, x) = x$ for all  $x \in X$ and associativity $act(\textcolor{action}{g}, act(\textcolor{action}{f}, x)) = act(\textcolor{action}{f} \circ \textcolor{action}{g}, x)$ for $\textcolor{action}{f},\textcolor{action}{g} \in \textcolor{action}{G}$.
-\end{definition}
-
-Elements of $X$ can be from one data field or all of them or some subset; similarly the actions act on
-the elements of $X$ and each action can be a composition of actions.
-This means actions can be used when discussing various measures of structure preservation.
-For example, \textit{equivariant} functions preserve structure under transformations to data or visualization and has been proposed by Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014} and \textit{homomorphic} maps preserve relations between data elements was preserved as proposed by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}.
-
-Specifically, Steven's conceptualizes the structure on values as \textcolor{action}{actions} on groups \footnote{A \textit{group} is a set with an associative binary operator.
-This operation must have an identity element and be closed, associative, and invertible}.
-A function that preserves structure when the input or output is changed by a group action is called \textit{equivariant}.
-
-\begin{figure}
-  \includegraphics*[width=1\columnwidth]{equivariant.pdf}
-  \caption{Encoding data as the bar height using an exponential transform is not equivariant because encoding the data and then scaling the bar heights yields a much taller graph then scaling the data and then encoding those heights using the same exponential transform function.}
-  \label{fig:related-work:equivariance}
-\end{figure}
-
- Given a group $G$ that acts on both the input $X$ and the output $Y$ of a function $f: X \rightarrow Y$
-
-\begin{definition}\label{def:equivariance}
- A function $f$ is \textbf{equivariant} when $f(act(g,x)) = act(g,f(x))$ for all $g$ in $G$ and for all $x$ in $X$ \cite{pittsNominalSetsNames2013}
-\end{definition}
-which means that a visualization is structure preserving when there exist compatible group actions on the data and visualization, as discussed by Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}.
-As illustrated in the commutative diagram in \autoref{fig:related-work:equivariance}, what this means is that the visual representation is consistent whether the data is scaled and then mapped to a graphic or whether the data is mapped to a graphic that is then modified in a compatible way.
-
+Loosely we define structure as a mathematical space $X$ and the actions on that space; our model is purposely generalizable to many different types of spaces (e.g sets, metric spaces, etc).
+Steven's conceptualizes the scales as mathematical groups with \textcolor{action}{actions}.
+A function that preserves structure when the input or output is changed by a group action is called \textit{equivariant}; a visualization is structure preserving when there exist compatible group actions on the data and visualization was proposed by Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}.
 Although the Steven's scales were conceptualized as having group structure, the ordinal scale has a monoidal structure because partial orders ($\geq, \leq$) are not invertible.
-This means \textit{equivariance} cannot be used to test for structure preservation.
-Instead \textit{homomorphism} can be used because it imposes fewer constraints on the underlying mathematical structure of the data.
-\begin{figure}
-  \includegraphics[width=1\columnwidth]{homomorphism.pdf}
-  \caption{Encoding data as bar height using an inverse transform is not homomorphic because the largest number is mapped to the smallest bar while the max function returns the largest bar.}
-  \label{fig:related-work:homomorphism}
-\end{figure}
-
-Given the function $f: X \rightarrow Y$, with operators $(X, \circ)$ and $(Y, *)$
-
-\begin{definition}\label{def:homomorphism}
-  A function $f$ is \textbf{homomorphic} when $f(x_1 \circ x_2) = f(x_1) * f(x_2)$ and preserves identities $f(I_x) = I_y$ all $x, y \in X$ \cite{grimaldiDiscreteCombinatorialMathematics2006}
-\end{definition}
-
-which means that the operators $\circ$ and $*$ are compatible.
-In \autoref{fig:related-work:homomorphism}, the $\geq$ operator is defined as the compatible closed functions \texttt{max} and the inverse transform
-is not homomorphic because it does not encode the maximum data value as the maximum bar value.
-
-As shown in \autoref{fig:def:equivariance} and \autoref{fig:def:homomorphsim}, a function can be homomorphic but not equivariant, such as an exponential encoding, or equivariant but not homomorphic, such as the inverse encoding.
+This means \textit{equivariance} cannot be used to test for structure preservation; instead \textit{homomorphism} can be used to verify whether map preserve relations between data elements, as proposed by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}.
+These tests are mutually exclusive; a function can be homomorphic but not equivariant, such as an exponential encoding, or equivariant but not homomorphic, such as the inverse encoding.
 A function can also be homomorphic (or equivariant) with respect to one action but not with respect to another.
@@
 While plotting the time series as a continuous line would be correct, it would be incorrect for a visualization to indicate that the distinct rows or stations are connected in a 1D continuous manner because it introduces ambiguity over which part of the line maps back to the data.
-A map that by definition has continuous maps between the input and output spaces, such as data and graphics, is called a \textit{homeomorphism}\cite{riehlCategoryTheoryContext}:
-
-\begin{definition}
-  A function $f$ is a $homeomorphism$ if it is bijective, continuous, and has a continuous inverse function $f^{-1}$.
-\end{definition}
-
+A map that by definition has continuous maps between the input and output spaces, such as data and graphics, is called a \textit{homeomorphism}\cite{riehlCategoryTheoryContext}.
 The bar plot, line plot, and heatmap in \autoref{fig:related-work:continuity:ktypes} have a homeomorphic relationship to the 0D ($\bullet$) points, 1D (--) linear, and 2D ($\blacksquare$) surface continuities embedded in the continuous 3 dimensional surface encapsulating time and position because each point of the visualization maps back into a point in its corresponding indexing space in the cube.
@@
 point , line, area) \cite{bertinSemiologyGraphicsDiagrams2011} and Wilkinson's assertion that connectivity must be preserved \cite{wilkinsonGrammarGraphics2005}.
+Unlike Wilkinson's
 
 To encode topology and field structure in a way that is both uniform and generalizable, we extend Butler's work on using a mathematical structure called fiber bundles as an abstract data representation in visualization \cite{butlerVectorBundleClassesForm1992, butlerVisualizationModelBased1989}.
+This allows us to use homeomorphism as the basis of equivalence, unlike Wilkinson \cite{wilkinsonGrammarGraphics2005}, because fiber bundles support planer geographic maps.
 Using this topological model of indexing, semantic indexing as described by Munzner's key-value model of data structure \cite{munznerWhatDataAbstraction2014} act as different ways of partitioning the underlying data continuity.
@@
 Using a topological model rather than semantic indexing also makes clearer when different labeling schemes refer to the same point, for example how 0-360 and 180E-180W are two ways of labeling longitude or how (date, lat, lon) and (date, station) refer to the same point.
+This is in contrast to Wilkinson modeling the continuity as metric spaces \cite{wilkinsonGrammarGraphics2005}such that a distance function is a fixed property of the continuity; in our model it distance is a property of a field over that continuity.
+
 We sketch out fiber bundles in \autoref{sec:atct:fiber-bundles}, but Butler provides a thorough introduction to bundles for visualization practitioners.
@@
 We extend Butler's proposal of using bundles as an abstraction for visualization data\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989} by incorporating Spivak's methodology for encoding named data types from his fiber bundle representation of relational databases \cite{spivakDatabasesAreCategories2010,spivakSimplicialDatabases2009}.
-We build on this work to describe how to encode the connectivity of the data as a topological \textcolor{base}{base space} modeling the data indexing space, encode the fields as a \textcolor{fiber}{fiber space} that acts as the data schema (domain), and express the mappings between these two spaces as \textcolor{section}{section} functions that encode datasets as mappings from indexing space to field space \texttt{dataset: topology $\rightarrow$ fields}.
-
-\begin{definition}\label{def:fiber_bundle}
 A \textbf{fiber bundle} $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and bundle projection map $\pi: \dtotalc \rightarrow \dbasec$ \cite{spanier1989algebraic}.
-
-   \begin{equation} \label{eq:atct:fb:intro}
-    \begin{tikzcd}
-      \dfiberc \arrow[r, hook, color=total] & \dtotalc \arrow[r, "\pi", color=total, two heads] & \dbasec
-      \end{tikzcd}
-    \end{equation}
-
-A continuous surjective map $\bm{\pi}$ is a \textbf{bundle projection} map when
-\begin{enumerate}
-  \item all fibers in the bundle are isomorphic.
-Since all fibers are isomorphic $\dfiber \cong \dfiber_{\dbasepoint}$ for all points $\dbasepoint \in \dbase$, there is a uniquely determined \textcolor{fiber}{fiber space} \dfiberc\ given by the preimage of the projection $\pi$ at any point $\dbasepoint$ in the \textcolor{base}{base space} \dbasec: $\dfiberc = \pi^{-1}(k)$.
-  \item each point $\dbasepointc$ in the \textcolor{base}{base space} \dbasec\ has an open neighborhood $\openset_{\dbasepointc}$ such that the \textcolor{total}{total space} \dtotalc\ over the neighborhood is locally trivial.
-\end{enumerate}
-\end{definition}
-
-\textbf{Local triviality} means $\dtotal\vert_{\openset} = \openset\times \dfiber$.
-In this paper we use $\dtotal\vert_{\openset} = \pi^{-1}(\openset)$ to denote the preimage of an openset\footnote{Open sets (open subsets) are a generalization of open intervals to n dimensional spaces.
-For example, an open ball is the set of points inside the ball and excludes points on the surface of the ball.
-\cite{weissteinOpenSet,bradleyTopologyVsTopology}}, and a \textbf{local trivialization} is a specific choice of neighborhoods (described in \autoref{sec:atct:fb:base}) and their preimages such that the fibers in each preimage are identical $\dfiber = \dfiber_{\dbasepoint}$ for all points $\dbasepoint \in \openset$.
-All fiber bundles can be decomposed into sets of local trivializations that are also bundles and we can specify a gluing scheme that reconstructs the fiber bundle from locally trivial pieces by specifying\textbf{transition maps} for all overlaps of the local trivializations; therefore, while the framework in this paper applies to all bundles, in this paper we assume that the bundles are trivial bundles $\dtotal = \dbase \times \dfiber$ so that we can assign all fibers in a bundle the same type.
-For an example of trivial and non-trivial bundles, see \autoref{sec:appendix:bundle_triviality}.
-
-\begin{minipage}{.5\columnwidth}
-\begin{definition}\label{def:fiber_bundle:section}
-A \textcolor{section}{\textbf{section}} $\dsectionc: \dbasec \rightarrow \dtotalc$ over a fiber bundle is a smooth right inverse of $\pi(\dsection(\dbasepoint)) = \dbasepoint$ for all $\dbasepoint \in \dbase$
-\end{definition}
-\end{minipage}
-\begin{minipage}{.4\columnwidth}
-  \begin{equation} \label{eq:atct:fb:intro-sec}
-    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
-     \dfiberc
-      \arrow[r, hook, color=total] \&
-      \dtotalc
-      \arrow[d, "\pi"',color=total, two heads] \\
-       \&
-    \dbasec
-       \arrow[u, "\dsectionc"', bend right, pos=.5, color=section, dashed]
-    \end{tikzcd}
-  \end{equation}
-\end{minipage}
-
-We propose that the total space of a bundle can encode the mathematical space in which a  dataset is embedded, the base space can encode the topological properties of the dataset, the fiber space can encode the data types of the fields of the dataset, and that the datasets can be encoded as section functions from the continuity to the fiber space.
-
-\begin{figure}[!h]
-       \includegraphics[width=1\columnwidth]{fb.pdf}
-       \caption{The space of all data values encoded by this fiber bundle can be modeled as a \textcolor{total}{rectangle} total space.
-Each dataset in this data space lies along the interval \textcolor{base}{$[0,2\pi]$} base space.
-Each dataset has values along the \textcolor{fiber}{$-1 \rightarrow 1$} interval fiber.
-One dataset embedded in this total space is the \textcolor{section}{sin} section over the bundle.}\label{fig:atct:fb}
-  \end{figure}
-
-For example, the fiber bundle in \autoref{fig:atct:fb} encodes the space of all continuous functions that have a domain of $[0, 2\pi]$ and range $[0,1]$.
-Using a fiber bundle abstraction encodes that the dataset has a 1D linear continuity as the base space \dbase is the interval $[0,2\pi]$ and a field type that is a \texttt{float} in the range $[0,1]$.
-Therefore the type signature of the datasets in this fiber bundle, which is called a section \dsection, would be \texttt{dataset: $[0, 2\pi] \rightarrow [0,1]$}.
-One such dataset (section) is the $\sin$ function, which as shown in \autoref{fig:atct:fb} is defined via a function \dsection from a point in the base space to a corresponding point in the fiber.
-Evaluating the section function over the entire base space yields the $\sin$ curve that is composed of points intersecting each fiber over the corresponding point.
-The local trivializations shown in \autoref{fig:atct:fb} are one way of decomposing the total bundle and conversely the bundle can be constructed from the local trivializations $\dbase = \dbase_{0} \oplus \dbase_{1}$.
-As shown, the section $\sin$ spans the trivializations in the same manner that it spans the bundle; this is analogous to how a dataset may span multiple tables or be collected in one table.
-The trivializations are glued together into the bundle at the overlapping region $\frac{2\pi}{5}$ by defining the transition map $\dfiber_{1} \rightarrow \dfiber_{2}$.
-Because the fibers in \autoref{fig:atct:fb} at $\frac{2\pi}{5}$ are aligned, the transition map is an identity map that take every value in $\dfiber_1$ and maps it to the same value in $\dfiber_2$ so that the sections, such as $\sin$, remain continuous.
+We build on this work to describe how to encode the connectivity of the data as a topological \textcolor{base}{base space} \dtotalc\ modeling the data indexing space, encode the fields as a \textcolor{fiber}{fiber space} \dfiberc\ that acts as the data schema (domain), and express the mappings between these two spaces as \textcolor{section}{section} \dsectionc\ functions that encode datasets as mappings from indexing space to field space \texttt{dataset: topology $\rightarrow$ fields}.
 
@@
 There are various equivalent definitions of topology, but here we use the neighborhood axiomatization because it is most analogous to the data access model of index (element) in subset (neighborhood) of all indices (mathematical space).
-Given a set $X$ and a function $\mathcal{N}:X\to 2^{2^X}$ that assigns to any $x\in X$ a non-empty collection of subsets $\mathcal{N}(x)$, where each element of $\mathcal{N}(x)$ is a \emph{neighborhood of $x$}, then $X$ with  $\mathcal{N}$ is a \textcolor{base}{topological space} and $\mathcal{N}$ is a neighborhood \emph{topology} if for each $x$ in $X$: \cite{brownronaldTopologyGroupoids2006}
-
-\begin{definition}\label{def:topology}
-\begin{enumerate}
-  \item if $N$ is a neighborhood $N \in \mathcal{N}(x)$ of $x$ then $x \in N$
-  \item every superset of a neighborhood of $x$ is a neighborhood of $x$; therefore a union of a neighborhood and adjacent points in $X$ is also a neighborhood of $x$
-  \item the intersection of any two neighborhoods of $x$ is a neighborhood of $x$
-  \item any neighborhood $N$ of $x$ contains a neighborhood $M \subset N$ of $x$ such that $N$ is a neighborhood of each of the points in $M$
-\end{enumerate}
-\end{definition}
-
-Therefore a neighborhood has to contain $x$ (1), can grow (2), or shrink (3), and every neighborhood also contains smaller neighborhoods of points adjacent to x (4).
+Roughly, a neighborhood of an element $x$ has to contain $x$ (1), can grow (2), or shrink (3), and every neighborhood also contains smaller neighborhoods of points adjacent to x (4).
 For example, in the indexing cube in \autoref{fig:fig:related-work:continuity:ktypes}, the brown surface and blue rectangle are both neighborhoods of the index for the measurement in Albany on June 09.
@@
 Following from Spivak's categorical abstraction of a database \cite{spivakSimplicialDatabases2009,spivakDatabasesAreCategories2010}, we also propose that the structure of the data types be formally specified as the objects of a category.
-
-\begin{definition}\label{def:atct:category}
-   An \textbf{category} $\mathcal{C}$ consists of the following \textit{data}:
-\begin{enumerate}
-  \item a collection of \textit{objects} $X \in \textbf{ob}(\mathcal{C})$
-  \item for every pair of objects $X, Y \in \textbf{ob}(\mathcal{C})$, a set of \textit{morphisms} $X \xrightarrow{f} Y \in Hom_{\mathcal{C}}(X, Y)$
-  \item for every object $X$, a distinct \textit{identity morphism} $X \xrightarrow {id_x} X$ in $Hom_{\mathcal{C}}(X, X)$
-  \item a \textit{composition function} $f \in Hom_{\mathcal{C}}(X, Y) \times  g \in Hom_{\mathcal{C}}(Y, Z) \rightarrow g \circ f \in Hom_{\mathcal{C}}(X, Z)$
-\end{enumerate}
-such that
-\begin{enumerate}
-  \item \textit{unitality:} for every morphism $ X \xrightarrow{f} Y$, $f \circ id_x = f = id_y \circ f$
-  \item \textit{associativity:} if any three morphisms $f, g, h$ are composable,
-    \begin{equation*}
-      \begin{tikzcd}
-        X \arrow[r, "f"] \arrow[rrr, "h\circ(g\circ f) = (h\circ g)\circ f"', bend right, dashed] & Y  \arrow[r, "g"] & Z \arrow[r, "h"] & W
-        \end{tikzcd}
-  \end{equation*}
-  then they are associative such that $h\circ(g\circ f) = (h \circ g) \circ f$  \cite{lawvere2009conceptual,riehlCategoryTheoryContext,maclaneCategoriesWorkingMathematician2013,fongInvitationAppliedCategory2019}.
-  \end{enumerate}
-\end{definition}
-
 The standard construction of a category from a topological space is that it has open set objects $\opensetc$ and morphisms $\opensetc_i \xrightarrow{\hat{\phi}} \opensetc_j$\cite{riehlCategoryTheoryContext}.
@@
 
+\section*{Supplemental Materials}
+\label{sec:supplemental_materials}
+
+Refer to the instructions for this section (\cref{sec:supplement_inst}).
+Below is an example you can follow that includes the actual supplemental material for this template:
+
+All supplemental materials are available on OSF at \url{https://doi.org/10.17605/OSF.IO/2NBSG}, released under a CC BY 4.0 license.
+In particular, they include (1) Excel files containing the data for and analyses for creating \cref{tab:vis_papers} and \cref{fig:vis_papers}, (2) figure images in multiple formats, and (3) a full version of this paper with all appendices.
+Our other code is intellectual property of a corporation---Starbucks Research---and there is no feasible way to share it publicly.
+
 \appendices
@@
 \end{table}
-
-\section{Trivial and non-trivial bundles}
-\label{sec:appendix:bundle_triviality}
-
-Generally, the distinguishing factor between a trivial bundle and a non-trivial bundle are how they are decomposed into local trivializations:
-\begin{LaTeXdescription}
-  \item[\textit{trivial bundle}] is directly isomorphic to $\dbasec\times\dfiberc$.
-For any choice of cover of $\dbasec$ by overlapping opensets, we can choose local trivializations such that all transition maps are identity maps.
-  \item[\textit{non-trivial bundle}] can not be constructed as $\dbasec\times\dfiberc$.
-For any choice of local trivializations, there is at least one transition map that is not an identity \cite{hatcherAlgebraicTopology2002}.
-\end{LaTeXdescription}
-
-\begin{figure}[H]
-  \includegraphics[width=1\columnwidth]{transition_maps.pdf}
-  \caption{The cylinder is a trivial fiber bundle; therefore it can be decomposed into local trivializations that only need identity
-  maps to glue the trivializations together.
-The mobius band is a non-trivial bundle; therefore it can only be decomposed into trivializations where at least one transition map is not an identity map.
-}\label{fig:cyl_mob_bundles}
-\end{figure}
-
-In the example in Figure~\ref{fig:cyl_mob_bundles}, we use arrows $\uparrow$ to denote fiber alignments.
-In the cylinder case the fibers all point in the same direction, which illustrates that they are equal $\uparrow=\uparrow$.
-In the Möbius band case, while the fibers in an arbitrary local trivialization are equal $\uparrow=\uparrow$, the fibers at the twist are unequal but isomorphic $\uparrow \cong \downarrow$.
-The cylinder and mobius band can be decomposed to the same local trivializations, for example the fiber bundles in \autoref{fig:atct:fb} In the cylinder case, the fibers in the overlapping regions of the trivializations are equal $\dfiber_0\restriction_{\openset_1\cap\openset_2} = \dfiber_{1}\restriction_{\openset_1\cap\openset_2}$; therefore the transition maps at both intersections map the values in the fiber to themselves $\delement\rightarrow\delement$ .
-In the Möbius band case, while $\dfiber_0\restriction_{(2\pi/5-\varepsilon, 2\pi/5+\varepsilon)}\to\dfiber_1\restriction_{(2\pi/5-\varepsilon, 2\pi/5+\varepsilon)}$ can be chosen to be an identity map, the other transition map component $\dfiber_0\restriction_{(-\varepsilon,\varepsilon)}\to\dfiber_1\restriction_{(-\varepsilon,\varepsilon)}$ has to flip any section values.
-For example given  $\dfiber_{0}=\uparrow$ and $\dfiber_{1}=\downarrow$, the transition map $\delement \mapsto -\delement$ maps each point from one fiber to the other $\uparrow \mapsto \downarrow$ such that any sections remain continuous even though the fibers point in opposite directions.
 
@@
 % references section
-
-\printbibliography
-
-% biography section
-% If you have an EPS/PDF photo (graphicx package needed) extra braces are
-% needed around the contents of the optional argument to biography to prevent
-% the LaTeX parser from getting confused when it sees the complicated
-% \includegraphics command within an optional argument. (You could create
-% your own custom macro containing the \includegraphics command to make things
-% simpler here.)
-%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
-% or if you just want to reserve a space for a photo:
-
-%\begin{IEEEbiography}{Michael Shell}
-%\end{IEEEbiography}
-
-% if you will not have a photo at all:
-\begin{IEEEbiographynophoto}{Hannah Aizenman}
-Biography text here.
-\end{IEEEbiographynophoto}
-
-\begin{IEEEbiographynophoto}{Thomas Caswell}
-  Biography text here.
-\end{IEEEbiographynophoto}
-% insert where needed to balance the two columns on the last page with
-% biographies
-%\newpage
-
-\begin{IEEEbiographynophoto}{Michael Grossberg}
-Biography text here.
-\end{IEEEbiographynophoto}
-
-% You can push biographies down or up by placing
-% a \vfill before or after them. The appropriate
-% use of \vfill depends on what kind of text is
-% on the last page and whether or not the columns
-% are being equalized.
-
-%\vfill
-
-% Can be used to pull up biographies so that the bottom of the last one
-% is flush with the other column.
-%\enlargethispage{-5in}
-
-% that's all folks
+\bibliographystyle{abbrv-doi}
+\bibliography{references}
 \end{document}

### 2025-07-21 paper/supplement.tex
supplement + new laptop

\documentclass[review]{vgtc}

% *** MISC UTILITY PACKAGES ***

\newcommand{\note}[1]{\textcolor{magenta}{#1}}

%\usepackage{silence}
%\WarningsOff[fixltx2e]

\usepackage[switch]{lineno}
\renewcommand{\linenumberfont}{\normalfont\bfseries\small\color{lightgray}}

% *** GRAPHICS RELATED PACKAGES ***
\usepackage{graphicx}
\graphicspath{{./figures/}}

% *** MATH PACKAGES ***
%% with other math-related packages, you may want to disable it.
\usepackage{amsmath, amsthm, amsfonts,amssymb,eulervm,xspace, mathtools}
\usepackage{relsize} %bigger
\usepackage{bm}

% theorem environments
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newcommand{\definitionautorefname}{Definition}
\theoremstyle{remark}
\newtheorem{example}{Example}[section]
\newtheorem{axiom}{Axiom}
\newtheorem{prop}{Proposition}

% *** diagrams
\usepackage{tikz}
\usetikzlibrary{cd}

% *** SPECIALIZED LIST PACKAGES ***
\usepackage{xcolor}
\usepackage[utf8]{inputenc}

\usepackage[]{minted}
\setminted[python]{fontsize=\scriptsize,
                   linenos,
                   numbersep=8pt,
                   frame=lines,
                   autogobble,
                   framesep=3mm,
                   breaklines=True}
% *** ALIGNMENT PACKAGES ***
\usepackage{multicol}
\usepackage{array}
\usepackage{multirow}
\usepackage{tabulary}
\usepackage{tabularray}

% Fonts
\usepackage{times}                     % we use Times as the main font
\renewcommand*\ttdefault{txtt}         % a nicer typewriter font
\usepackage{stmaryrd}%maps from
\renewcommand{\restriction}{\mathord{\upharpoonright}} %restriction w/p space
\usepackage{mathrsfs} % math script fonts
\usepackage{mathptmx}

% *** SUBFIGURE PACKAGES ***
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}

% *** FLOAT PACKAGES ***
\usepackage{dblfloatfix}

% *** PDF, URL AND HYPERLINK PACKAGES ***
\usepackage{xurl}

% *** BIBLIOGRAPHY ***
\usepackage[]{footmisc}

\usepackage[inline]{showlabels}

\usepackage{notation} %notation conventions
% correct bad hyphenation here
%\hyphenation{}

\onlineid{0}

%% declare the category of your paper, only shown in review mode
\vgtccategory{Research}

%% allow for this line if you want the electronic option to work properly
\vgtcinsertpkg
\title{Supplemental Materials: Math Glossary}

\begin{document}

\firstsection{Structure preservation}
\maketitle

\subsection{Actions}
 Loosely we define structure as a mathematical space $X$ and the actions on that space

\begin{definition}\label{def:related-work:action}\cite{grimaldiDiscreteCombinatorialMathematics2006}
  An \textcolor{action}{\textbf{action}} of \textcolor{action}{$G = (G,\circ, e)$} on $X$ is a function  $act: \textcolor{action}{G} \times X \rightarrow X$. An action has the properties of identity $act(\textcolor{action}{e}, x) = x$ for all  $x \in X$ and associativity $act(\textcolor{action}{g}, act(\textcolor{action}{f}, x)) = act(\textcolor{action}{f} \circ \textcolor{action}{g}, x)$ for $\textcolor{action}{f},\textcolor{action}{g} \in \textcolor{action}{G}$.
\end{definition}

Elements of $X$ can be from one data field or all of them or some subset; similarly the actions act on
the elements of $X$ and each action can be a composition of actions.

\subsection{Equivariance}
A mathematical {group} is a set with an associative binary operator. This operation must have an identity element and be closed, associative, and invertible, consisting of a set of values $X$ and \textcolor{action}{actions} on the set {$G = (G,\circ, e)$}.
\begin{figure}
  \includegraphics*[width=1\columnwidth]{equivariant.pdf}
  \caption{Encoding data as the bar height using an exponential transform is not equivariant because encoding the data and then scaling the bar heights yields a much taller graph then scaling the data and then encoding those heights using the same exponential transform function.}
  \label{fig:related-work:equivariance}
\end{figure}

 Given a group $G$ that acts on both the input $X$ and the output $Y$ of a function $f: X \rightarrow Y$

\begin{definition}\label{def:equivariance}
 A function $f$ is \textbf{equivariant} when $f(act(g,x)) = act(g,f(x))$ for all $g$ in $G$ and for all $x$ in $X$ \cite{pittsNominalSetsNames2013}
\end{definition}
which means that
 As illustrated in the commutative diagram in \autoref{fig:related-work:equivariance}, what this means is that the visual representation is consistent whether the data is scaled and then mapped to a graphic or whether the data is mapped to a graphic that is then modified in a compatible way.

\subsection{Homomorphism}

\begin{figure}
  \includegraphics[width=1\columnwidth]{homomorphism.pdf}
  \caption{Encoding data as bar height using an inverse transform is not homomorphic because the largest number is mapped to the smallest bar while the max function returns the largest bar.}
  \label{fig:related-work:homomorphism}
\end{figure}

Given the function $f: X \rightarrow Y$, with operators $(X, \circ)$ and $(Y, *)$

\begin{definition}\label{def:homomorphism}
  A function $f$ is \textbf{homomorphic} when $f(x_1 \circ x_2) = f(x_1) * f(x_2)$ and preserves identities $f(I_x) = I_y$ all $x, y \in X$ \cite{grimaldiDiscreteCombinatorialMathematics2006}
\end{definition}

which means that the operators $\circ$ and $*$ are compatible. In \autoref{fig:related-work:homomorphism}, the $\geq$ operator is defined as the compatible closed functions \texttt{max} and the inverse transform
is not homomorphic because it does not encode the maximum data value as the maximum bar value.

\subsection{Topology and Homeomorphism}

We encode the continuity of the datasets using their topological structure, specifically using the neighborhood definition:

 Given a set $X$ and a function $\mathcal{N}:X\to 2^{2^X}$ that assigns to any $x\in X$ a non-empty collection of subsets $\mathcal{N}(x)$, where each element of $\mathcal{N}(x)$ is a \emph{neighborhood of $x$}, then $X$ with  $\mathcal{N}$ is a \textcolor{base}{topological space} and $\mathcal{N}$ is a neighborhood \emph{topology} if for each $x$ in $X$: \cite{brownronaldTopologyGroupoids2006}

\begin{definition}\label{def:topology}
\begin{enumerate}
  \item if $N$ is a neighborhood $N \in \mathcal{N}(x)$ of $x$ then $x \in N$
  \item every superset of a neighborhood of $x$ is a neighborhood of $x$; therefore a union of a neighborhood and adjacent points in $X$ is also a neighborhood of $x$
  \item the intersection of any two neighborhoods of $x$ is a neighborhood of $x$
  \item any neighborhood $N$ of $x$ contains a neighborhood $M \subset N$ of $x$ such that $N$ is a neighborhood of each of the points in $M$
\end{enumerate}
\end{definition}

We then use this topological structure as the basis of claiming topological equivalence between a data and a graphic if there exists a homeomorphic map between the continuities of the data and graphic:

\begin{definition}
  A function $f$ is a $homeomorphism$ if it is bijective, continuous, and has a continuous inverse function $f^{-1}$.
\end{definition}

\section{Fiber Bundles}

\begin{definition}\label{def:fiber_bundle}
   A \textbf{fiber bundle} $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and  bundle projection map $\pi: \dtotalc \rightarrow \dbasec$ \cite{spanier1989algebraic}.

   \begin{equation} \label{eq:atct:fb:intro}
    \begin{tikzcd}
      \dfiberc \arrow[r, hook, color=total] & \dtotalc \arrow[r, "\pi", color=total, two heads] & \dbasec
      \end{tikzcd}
    \end{equation}

A continuous surjective map $\bm{\pi}$ is a \textbf{bundle projection} map when
\begin{enumerate}
  \item all fibers in the bundle are isomorphic. Since all fibers are isomorphic $\dfiber \cong \dfiber_{\dbasepoint}$ for all points $\dbasepoint \in \dbase$, there is a uniquely determined \textcolor{fiber}{fiber space} \dfiberc\ given by the preimage of the projection $\pi$ at any point $\dbasepoint$ in the \textcolor{base}{base space} \dbasec: $\dfiberc = \pi^{-1}(k)$.
  \item each point $\dbasepointc$ in the \textcolor{base}{base space} \dbasec\ has an open neighborhood $\openset_{\dbasepointc}$ such that the \textcolor{total}{total space} \dtotalc\ over the neighborhood is locally trivial.
\end{enumerate}
\end{definition}

\textbf{Local triviality} means $\dtotal\vert_{\openset} = \openset\times \dfiber$. In this paper we use $\dtotal\vert_{\openset} = \pi^{-1}(\openset)$ to denote the preimage of an openset\footnote{Open sets (open subsets) are a generalization of open intervals to n dimensional spaces. For example, an open ball is the set of points inside the ball and excludes points on the surface of the ball. \cite{weissteinOpenSet,bradleyTopologyVsTopology}}, and a \textbf{local trivialization} is a specific choice of neighborhoods (described in \autoref{sec:atct:fb:base}) and their preimages such that the fibers in each preimage are identical $\dfiber = \dfiber_{\dbasepoint}$ for all points $\dbasepoint \in \openset$. All fiber bundles can be decomposed into sets of local trivializations that are also bundles and we can specify a gluing scheme that reconstructs the fiber bundle from locally trivial pieces by specifying\textbf{transition maps} for all overlaps of the local trivializations; therefore, while the framework in this paper applies to all bundles, in this paper we assume that the bundles are trivial bundles $\dtotal = \dbase \times \dfiber$ so that we can assign all fibers in a bundle the same type.

\begin{minipage}{.5\columnwidth}
\begin{definition}\label{def:fiber_bundle:section}
A \textcolor{section}{\textbf{section}} $\dsectionc: \dbasec \rightarrow \dtotalc$ over a fiber bundle is a smooth right inverse of $\pi(\dsection(\dbasepoint)) = \dbasepoint$ for all $\dbasepoint \in \dbase$
\end{definition}
\end{minipage}
\begin{minipage}{.4\columnwidth}
  \begin{equation} \label{eq:atct:fb:intro-sec}
    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
     \dfiberc
      \arrow[r, hook, color=total] \&
      \dtotalc
      \arrow[d, "\pi"',color=total, two heads] \\
       \&
    \dbasec
       \arrow[u, "\dsectionc"', bend right, pos=.5, color=section, dashed]
    \end{tikzcd}
  \end{equation}
\end{minipage}

We propose that the total space of a bundle can encode the mathematical space in which a  dataset is embedded, the base space can encode the topological properties of the dataset, the fiber space can encode the data types of the fields of the dataset, and that the datasets can be encoded as section functions from the continuity to the fiber space.

\begin{figure}[!h]
       \includegraphics[width=1\columnwidth]{fb.pdf}
       \caption{The space of all data values encoded by this fiber bundle can be modeled as a \textcolor{total}{rectangle} total space. Each dataset in this data space lies along the interval \textcolor{base}{$[0,2\pi]$} base space. Each dataset has values along the \textcolor{fiber}{$-1 \rightarrow 1$} interval fiber. One dataset embedded in this total space is the \textcolor{section}{sin} section over the bundle.}\label{fig:atct:fb}
  \end{figure}

For example, the fiber bundle in \autoref{fig:atct:fb} encodes the space of all continuous functions that have a domain of $[0, 2\pi]$ and range $[0,1]$. Using a fiber bundle abstraction encodes that the dataset has a 1D linear continuity as the base space \dbase is the interval $[0,2\pi]$ and a field type that is a \texttt{float} in the range $[0,1]$. Therefore the type signature of the datasets in this fiber bundle, which is called a section \dsection, would be \texttt{dataset: $[0, 2\pi] \rightarrow [0,1]$}. One such dataset (section) is the $\sin$ function, which as shown in \autoref{fig:atct:fb} is defined via a function \dsection from a point in the base space to a corresponding point in the fiber. Evaluating the section function over the entire base space yields the $\sin$ curve that is composed of points intersecting each fiber over the corresponding point. The local trivializations shown in \autoref{fig:atct:fb} are one way of decomposing the total bundle and conversely the bundle can be constructed from the local trivializations $\dbase = \dbase_{0} \oplus \dbase_{1}$. As shown, the section $\sin$ spans the trivializations in the same manner that it spans the bundle; this is analogous to how a dataset may span multiple tables or be collected in one table. The trivializations are glued together into the bundle at the overlapping region $\frac{2\pi}{5}$ by defining the transition map $\dfiber_{1} \rightarrow \dfiber_{2}$. Because the fibers in \autoref{fig:atct:fb} at $\frac{2\pi}{5}$ are aligned, the transition map is an identity map that take every value in $\dfiber_1$ and maps it to the same value in $\dfiber_2$ so that the sections, such as $\sin$, remain continuous.

\subsection{Trivial and non-trivial bundles}
\label{sec:appendix:bundle_triviality}

Generally, the distinguishing factor between a trivial bundle and a non-trivial bundle are how they are decomposed into local trivializations:
\begin{LaTeXdescription}
  \item[\textit{trivial bundle}] is directly isomorphic to $\dbasec\times\dfiberc$. For any choice of cover of $\dbasec$ by overlapping opensets, we can choose local trivializations such that all transition maps are identity maps.
  \item[\textit{non-trivial bundle}] can not be constructed as $\dbasec\times\dfiberc$. For any choice of local trivializations, there is at least one transition map that is not an identity \cite{hatcherAlgebraicTopology2002}.
\end{LaTeXdescription}

\begin{figure}[H]
  \includegraphics[width=1\columnwidth]{transition_maps.pdf}
  \caption{The cylinder is a trivial fiber bundle; therefore it can be decomposed into local trivializations that only need identity
  maps to glue the trivializations together. The mobius band is a non-trivial bundle; therefore it can only be decomposed into trivializations where at least one transition map is not an identity map. }\label{fig:cyl_mob_bundles}
\end{figure}

In the example in Figure~\ref{fig:cyl_mob_bundles}, we use arrows $\uparrow$ to denote fiber alignments. In the cylinder case the fibers all point in the same direction, which illustrates that they are equal $\uparrow=\uparrow$. In the Möbius band case, while the fibers in an arbitrary local trivialization are equal $\uparrow=\uparrow$, the fibers at the twist are unequal but isomorphic $\uparrow \cong \downarrow$. The cylinder and mobius band can be decomposed to the same local trivializations, for example the fiber bundles in \autoref{fig:atct:fb} In the cylinder case, the fibers in the overlapping regions of the trivializations are equal $\dfiber_0\restriction_{\openset_1\cap\openset_2} = \dfiber_{1}\restriction_{\openset_1\cap\openset_2}$; therefore the transition maps at both intersections map the values in the fiber to themselves $\delement\rightarrow\delement$ . In the Möbius band case, while $\dfiber_0\restriction_{(2\pi/5-\varepsilon, 2\pi/5+\varepsilon)}\to\dfiber_1\restriction_{(2\pi/5-\varepsilon, 2\pi/5+\varepsilon)}$ can be chosen to be an identity map, the other transition map component $\dfiber_0\restriction_{(-\varepsilon,\varepsilon)}\to\dfiber_1\restriction_{(-\varepsilon,\varepsilon)}$ has to flip any section values. For example given  $\dfiber_{0}=\uparrow$ and $\dfiber_{1}=\downarrow$, the transition map $\delement \mapsto -\delement$ maps each point from one fiber to the other $\uparrow \mapsto \downarrow$ such that any sections remain continuous even though the fibers point in opposite directions.

\section{Categories}
We propose that the various structures of the bundle (base space, fiber, total space) can also be encoded as categories:

\begin{definition}\label{def:atct:category}
   An \textbf{category} $\mathcal{C}$ consists of the following \textit{data}:
\begin{enumerate}
  \item a collection of \textit{objects} $X \in \textbf{ob}(\mathcal{C})$
  \item for every pair of objects $X, Y \in \textbf{ob}(\mathcal{C})$, a set of \textit{morphisms} $X \xrightarrow{f} Y \in Hom_{\mathcal{C}}(X, Y)$
  \item for every object $X$, a distinct \textit{identity morphism} $X \xrightarrow {id_x} X$ in $Hom_{\mathcal{C}}(X, X)$
  \item a \textit{composition function} $f \in Hom_{\mathcal{C}}(X, Y) \times  g \in Hom_{\mathcal{C}}(Y, Z) \rightarrow g \circ f \in Hom_{\mathcal{C}}(X, Z)$
\end{enumerate}
such that
\begin{enumerate}
  \item \textit{unitality:} for every morphism $ X \xrightarrow{f} Y$, $f \circ id_x = f = id_y \circ f$
  \item \textit{associativity:} if any three morphisms $f, g, h$ are composable,
    \begin{equation*}
      \begin{tikzcd}
        X \arrow[r, "f"] \arrow[rrr, "h\circ(g\circ f) = (h\circ g)\circ f"', bend right, dashed] & Y  \arrow[r, "g"] & Z \arrow[r, "h"] & W
        \end{tikzcd}
  \end{equation*}
  then they are associative such that $h\circ(g\circ f) = (h \circ g) \circ f$  \cite{lawvere2009conceptual,riehlCategoryTheoryContext,maclaneCategoriesWorkingMathematician2013,fongInvitationAppliedCategory2019}.
  \end{enumerate}
\end{definition}

The composability property expresses that the morphism is transitive, while associativity expresses that the morphisms can be curried in various equivalent groupings. By formally specifying the properties of the topological structure data types as $\mathcal{\dbase}$, we can express that these are the properties that are required as part of the implementation of the data type objects.

\section{Presheaves And Sheafs}

\bibliographystyle{abbrv-doi}
\bibliography{references}

\end{document}

### 2025-07-21 paper/figcode/homeomorphism_helpers.py

@@ -1,0 +1,209 @@
+import pandas as pd
+import geopandas as gpd
+import numpy as np
+
+import metpy
+from metpy.interpolate import interpolate_to_grid
+
+import matplotlib.path as mpath
+import matplotlib.patches as mpatches
+import matplotlib.text as mtext
+import matplotlib as mpl
+import matplotlib.colors as mcolors
+import matplotlib.dates as mdates
+import matplotlib.transforms as mtransforms
+
+from mpl_toolkits.mplot3d.art3d import Poly3DCollection
+import mpl_toolkits.mplot3d.art3d as art3d
+from mpl_toolkits.mplot3d import proj3d
+
+import cmocean
+
+import intro_plots as ip
+import papercolors
+from papercolors import colordict as pcd
+
+mpl.rcParams['font.family'] = 'dejavu sans'
+mpl.rcParams['figure.dpi'] = 200
+fs={'normal': 10, 'small':9, 'footnote':8, 'script':7, 'tiny':5, 'heading':12}
+
+airports = {'map_date': '20220609',
+            'map_color': 'sienna',
+            't_lon': -73.7992,
+            't_lat': 42.7472,
+            'time_color': 'deepskyblue',
+            'point_date': '20220309',
+            'point_color': 'thistle',
+            'jfk_lon': -73.7639,
+            'jfk_lat': 40.6392,
+            'jfk_color': 'steelblue',
+            'lga_lon': -73.8803,
+            'lga_lat': 40.7794,
+            'lga_color': 'lightskyblue'}
+airports.update({'time_point_color': mcolors.to_hex(papercolors.lighten(airports['time_color'], 2))})
+
+def get_data(path):
+    df = pd.read_parquet(path)
+    dates = df['DATE'].sort_values().unique()
+    lons = df['LONGITUDE'].sort_values().unique()
+    lats = df['LATITUDE'].sort_values().unique()
+
+    df['DATE_DTYPE'] = pd.to_datetime(df['DATE'])
+    prp = df[df['NAME'].isin(['LAGUARDIA AP', 'JFK INTL AP', 'ALBANY INTL AP']) 
+             & (df['DATE_DTYPE'].dt.month==3)].pivot_table(
+                columns="NAME", index='DATE_DTYPE', values= 'PRCP')
+    
[…]

### 2025-07-25 short_paper/main.tex
Short paper edited for AccessVis submission

@@
-\documentclass[11pt,titlepage]{amsart}
+\documentclass[preprint]{vgtc}
+
+\usepackage{times}
+\renewcommand*\ttdefault{txtt}
+\usepackage{mathptmx}
+
 \usepackage{xcolor}
@@
 \usepackage{caption}
-\usepackage[anonymous]{atmcs} % for anonymous submission
-% \usepackage{atmcs} % for camera-ready
+
+\usepackage{graphicx}
+\usepackage{amsmath,amssymb}
+
 \usepackage{lipsum}
-\graphicspath{{C:/Users/story/Projects/team/paper/figures/}}
+\graphicspath{{../paper/figures/}{figs}}
 
-\begin{document}
+\onlineid{0}
+\vgtccategory{Research}
+%\vgtcinsertpkg
 
-\title[]{Topological Equivariant Artist Model}
+\title{Topological Equivariant Artist Model}
 
-\author{Hannah~Aizenman}
-\address{The Graduate Center, CUNY \\NY, NY, USA}
-\email{haizenman@gradcenter.cuny.edu}
-
-\author{Mikael Vejdemo-Johansson}
-\address{The Graduate Center, CUNY\\NY, NY, USA}
-\email{mvj@math.csi.cuny.edu}
-
-\author{Thomas Caswell}
-\address{National Synchrotron Light Source II, Brookhaven National Lab\\ Upton, NY, USA}
-\email{tcaswell@bnl.gov}
-\author{Michael Grossberg}
-\address{The Graduate Center, CUNY\\NY, NY, USA}
-\email{mgrossberg@ccny.cuny.edu}
+\author{
+Hanna Aizenman\thanks{e-mail: haizenman@gradcenter.cuny.edu} \\
+\scriptsize{CUNY Graduate Center}
+\and
+Mikael Vejdemo-Johansson\thanks{e-mail: mvj@math.csi.cuny.edu} \\
+\parbox{1.6in}{\scriptsize \centering CUNY College of Staten Island \\
+CUNY Graduate Center}
+\and
+Thomas Caswell\thanks{e-mail: tcaswell@bnl.gov} \\
+\parbox{1.8in}{\scriptsize\centering National Synchrotron Light Source II \\ Brookhaven National Lab}
+\and
+Michael Grossberg\thanks{e-mail: mgrossberg@ccny.cuny.edu} \\
+\parbox{1.6in}{\scriptsize\centering City College of New York \\ CUNY Graduate Center}
+}
 
@@
 
-\begin{abstract}
+\abstract{
     The contract data visualization tools make with their users is that a chart is a faithful and accurate visual representation of the numbers it is made from.
@@
     Using category theory to formally express how visual elements are constructed means we can translate those expectations into code, which can then be used to enforce the expectation that a visualization tool is faithfully translating between numbers and charts.
-\end{abstract}
+    The sheaf formalism is generic enough that with appropriate choices of a sheaf structure to replace the graphic sheaves, the same framework will work to model and specify non-visual modalities with identical underlying data representations, and structurally similar data to perception transformations.
+}
+
+\begin{document}
 
@@
 
-\begin{wrapfigure}{r}{3.5in}
+\begin{figure}
 \begin{center}
@@
     \label{fig:homeomorphism}
-\end{wrapfigure}
+\end{figure}
 Motivated by wanting to make better tools, we propose a methodology for fully specifying arbitrary data to visualization mappings in a manner that easily translates to code.
 Generally, preserving structure means that a visualization is expected to preserve the $field$ properties and $topology$ of the corresponding dataset, where \textcolor{fiber}{\textbf{field}} is a set of values of the same type and \textcolor{base}{\textbf{topology}} is the connectivity and relative positioning of elements in a dataset \cite{wilkinsonGrammarGraphics2005}.
-Field structure is traditionally codified as the Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946}, where each scale is a set of actions on a group Topological structure is generally assumed by visual algorithms \cite{chiTaxonomyVisualizationTechniques2000, toryRethinkingVisualizationHighlevel2004}, but generally do not verify that input structure.
+Field structure is traditionally codified as the Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946}, where each scale is a set of actions on a group.
+Topological structure is generally assumed by visual algorithms \cite{chiTaxonomyVisualizationTechniques2000, toryRethinkingVisualizationHighlevel2004}, but generally do not verify that input structure.
 For example, a \texttt{line} algorithm often does not have a way to query whether a list of (x,y) coordinates is the distinct rows, the time series, or the list of stations in \autoref{fig:homeomorphism}.
@@
 
+Through this paper we will be discussing visualization primarily, but by changing the last of the three fiber bundles we discuss in Section~\ref{sec:data models}, we could go from a visual representation to an auditory, haptic, or any other modality and retain the benefits we identify from the fiber bundle and sheaf structures.
+
 \section{Data Model}
 \label{sec:data models}
-We extend Butler's proposal of using fiber bundles as an abstraction for visualization data\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989} by incorporating Spivak's methodology for encoding named data types from his fiber bundle representation of relational databases \cite{spivakDatabasesAreCategories2010,spivakSimplicialDatabases2009}.
 
-\begin{multicols}{2}
+Spivak represents relational databases with a fiber bundle representation \cite{spivakDatabasesAreCategories2010,spivakSimplicialDatabases2009}.
+This suggests a way for us to extend Butler's proposal of using fiber bundles as an abstraction for visualization data \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989} by incorporating Spivak's approach, and extending it from data storage to represent the transformation to graphical representation.
+As we show in Table~\ref{tab:data_abstraction}, where Spivak represents a specific dataset as a fiber, a database schema with the bundle's fiber space and topological indexing through a base space, we introduce parallel analogous structures representing a visual specification and a graphic realization.
+
+\begin{figure}
+%\begin{multicols}{2}
     \includegraphics[width=.5\textwidth]{fb_rho.pdf}
-    \columnbreak
-    \captionof{figure}{The scatter marker is specified by the section $\gsection$, which maps into the fiber $\gfiber$ to retrieve the values that compose into the pixel (approximated as a square) returned by the section function evaluated at each point $\gbasepoint$.
+%    \columnbreak
+    \caption{The scatter marker is specified by the section $\gsection$, which maps into the fiber $\gfiber$ to retrieve the values that compose into the pixel (approximated as a square) returned by the section function evaluated at each point $\gbasepoint$.
 The section evaluated over the entire space $\gsection\vert_{S}$ returns the entire scatter mark, shown here in faded form to make it easier to see the individual pixels.}\label{fig:fiber_bundle}
+\end{figure}
 
-\end{multicols}
-
-\begin{table}[h]
+\begin{table*}[t]
     \begin{tabular}{|r | c c c|}
@@
     \label{tab:data_abstraction}
-\end{table}
+\end{table*}
 
-The indexing space is modeled as a topology $\mathcal{T}$ on the underlying data indexing space because it is a uniform method of describing arbitrary continuity and it provides a method for managing overlapping embedded topologies, such as \autoref{fig:homeomorphism}.
-Adopting Spivak's fiber bundle construction of types allows our model to reuse types so long as the field names are distinct and encode multi-dimensional fields can be encoded in the same manner as single dimensional fields because domains of types are a mathematical space.
-We encode data as a \textcolor{section}{section} \dsectionc\ of a bundle because this allows us to incorporate the topology and field types in the data definition $\textcolor{section}{\texttt{dataset}}: \textcolor{base}{\texttt{topology}} \rightarrow \textcolor{fiber}{\texttt{field}}$.
-Abstracting the data containers as sheaves provides a way of encoding how data containers keep track of the continuity of the data \cite{ghristElementaryAppliedTopology2014} such that visual algorithms developed on this model apply whether the data fits in memory, is distributed, or is streaming.
+By choosing a topological space $\mathcal T$ as indexing space underlying the dataset we get a uniform abstraction for describing arbitrary continuity -- time series, lists of observations, spatial fields all have the same abstract structure -- and we get a method for managing overlapping embedded topologies, such as we show in \autoref{fig:homeomorphism}.
+Spivak's fiber bundle construction of types decouples field types from field names and single-dimensional and multi-dimensional fields share a common abstraction: they all rely on some choice of mathematical object for the data types.
+A particular dataset can in this paradigm be encoded as a \textcolor{section} \dsectionc\ of a bundle, assigning to each index value an observation in the data types prescribed by the fiber space.
+This fits topology and field types into a data definition $\textcolor{section}{\texttt{dataset}}: \textcolor{base}{\texttt{topology}} \rightarrow \textcolor{fiber}{\texttt{field}}$.
+
+We abstract these data containers as sheaves, and the gluing property of sheaves provides a systematic way to track continuity of the data \cite{ghristElementaryAppliedTopology2014} and apply visual algorithms developed on this model regardless whether the data fits in memory, is distributed, or streaming.
 
@@
 \label{sec:Artist}
+
 We call the mapping from data to graphic $\vartistc:\dsectionc\mapsto\gsectionc$ an \textcolor{artist}{Artist}\footnote{Artists are named for the Artist class in Matplotlib.
-Artist is the parent class of all the objects that assemble and manage visual elements such as points, lines, and shapes.} and propose that an artist is structure preserving when the artist is implemented as a natural transform on sheaves.
+Artist is the parent class of all the objects that assemble and manage visual elements such as points, lines, and shapes.}.
+Artists are structure preserving when they are implemented as sheaf maps from the data representation sheaf to the graphics sheaf.
 This means that the artist satisfies the following constraints, as illustrated in \autoref{fig:artist}:
 
-\begin{multicols}{2}
-    \includegraphics[width=.5\textwidth]{artist_equiv.png}
-    \captionof{figure}{}\label{fig:artist}
-\columnbreak
+\begin{figure}
+    \includegraphics[width=.5\textwidth]{artist_equiv.pdf}
+    \caption{}\label{fig:artist}
+\end{figure}
 
 \begin{description}
-    \item[homeomorphic] there exists a deformation retract $\vindexc: \gbasec\rightarrow\dbasec$ between graphic and artist base space
+    \item[homotopic] there exists a deformation retract\footnote{This requirement emerges from the digital geometry nature of displays: even if the intended visual primitive may be 1-dimensional, it's screen realization must be 2-dimensional -- requiring a deformation retract most closely agrees with our intuition for drawing geometric primitives} $\vindexc: \gbasec\rightarrow\dbasec$ between graphic and artist base space
     \item[topology] the artist is commutative w.r.t.
-the presheaf morphism $\dfunchc: \dbasepointc^{\prime}\mapsto \dbasepointc$ and pullback $\dfuncpullc \dsectionc \restriction_{\opensetc^{\prime}}: \dsectionc\restriction_{\opensetc^{\prime}} \mapsto \dsectionc \restriction_{\opensetc^{\prime} \circ \dfunchc}$ such that when an index of the record changes the values in the record are unmodified  $\dsectionc(\dbasec) = \dfuncpullc\dsectionc(\dbasepointc^{\prime}) = \dsectionc(\dfunchc(\dbasepointc^{\prime}))$
+the presheaf morphism $\dfunchc: \dbasepointc^{\prime}\mapsto \dbasepointc$ and pullback $\dfuncpullc \dsectionc \restriction_{\opensetc^{\prime}}: \dsectionc\restriction_{\opensetc^{\prime}} \mapsto \dsectionc \restriction_{\opensetc^{\prime} \circ \dfunchc}$: index changes on the record leaves values in the record unmodified
+    $\dsectionc(\dbasec) = \dfuncpullc\dsectionc(\dbasepointc^{\prime}) = \dsectionc(\dfunchc(\dbasepointc^{\prime}))$
     \item[fields] the artist is commutative w.r.t.
-actions on the fiber $\dfunctc \in Hom(\dfiberc, \dfiberc)$, which are the functions between sections $\dfunctc: \dfuncpullc \dsectionc \restriction_{\opensetc^{\prime}} \mapsto \dfuncpullc \dsectionc^{\prime} \restriction_{\opensetc}$.
-Additionally, $\dfunctc$ must preserve features of $\dfiber$, such as operators, that are defined as part of the structure of $\dfiber$.
+actions on the fiber.
+An action $\dfunctc: \dfiberc\to\dfiberc$ induces a function between sections $\dfunctc: \dfuncpullc \dsectionc \restriction_{\opensetc^{\prime}} \mapsto \dfuncpullc \dsectionc^{\prime} \restriction_{\opensetc}$.
+We also require $\dfunctc$ to respect the structure, such as operators, of $\dfiber$.
+\end{description}
 
-\end{description}
-\end{multicols}
+With this underlying topological structure to describe the visualization pipeline, traditional constructions from topology and category theory generate natural operators on the data structures.
+Pushouts and pullbacks, of central importance in category theory and topology, emerge here as operators to add and extend records, respectively.
+Since (in categorical language) sheaf maps preserve limits and colimits, this observation shows that artists are composable: $\vartistc(\bigsqcup\limits_{i}\dbasec_i, \prod\limits_j\dfiberc_j) = \bigsqcup\limits_i\prod\limits_j\vartistc(\dbasec_i, \dfiberc_j)$
 
-\begin{figure}[h]
+\begin{figure*}[h]
+\includegraphics[width=\textwidth]{more_records.pdf}
+\caption{Pullback (left) and pushout (right) structure of the operations that add records (increasing the indexing base space) or expand records (merging different fiber spaces with a database join-like operation).}\label{fig:more-records}
+\end{figure*}
+
+\begin{figure*}[h]
     \includegraphics[width=7.16in]{xi_diagram.pdf}
-    \caption{The graphics over the orange bands map to the data at the orange point.}
+    \caption{A deformation retract $\vindexc$ that relates the graphics footprint $S_0$ of an abstract circle to the circle itself gives rise to a pair of additional maps.
+The pushforward $\vindexpushc$ matches a point in the graphical realization to the specification of the graphic at that point, while the pullback $\vindexpullc$ matches each point in the graphic space to the data giving rise to that point.
+    }
     \label{fig:xi}
-\end{figure}
-As illustrated in \ref{fig:xi}, the transport functors $(\vindexpush, \vindexpull)$ constructed from $\vindex$ \cite{harder2008lectures} also codify how visual elements correspond to distinct data elements, which is a necessary condition of a visualization being readable\cite{ziemkiewiczEmbeddingInformationVisualization2009}.
+\end{figure*}
+
+As illustrated in \autoref{fig:xi}, this deformation retract $\vindexc$ gives rise to two so-called transport functors $(\vindexpush, \vindexpull)$ -- known as the pushforward $\vindexpushc$ and pullback\footnote{different use of the word pullback than in the preceding paragraph -- an unfortunate collision of terminology} $\vindexpullc$.
+These also codify how visual elements correspond to distinct data elements, which is a necessary condition of a visualization being readable\cite{ziemkiewiczEmbeddingInformationVisualization2009}.
 The pushforward $\vindexpushc$ matches each point in the data space to the specification of the graphic at that point $\vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexprec(\dbasepointc)}$, while the pullback $\vindexpull$ matches each point in the graphic space to the data over that point $\vindexpullc\dsectionc(\gbasepointc) = \dsectionc(\vindexc(\gbasepointc)) = \dsectionc(\dbasepointc)$.
 
-\begin{figure*}[h]
-\includegraphics[width=\textwidth]{tex/more_records.png}
-\end{figure*}
+\section{Construction}
 
-The topology and fiber categories are equipped with operators to add and extend records, respectively, Since the artist preserves these operators, artists have the property of being composable $\vartistc(\bigsqcup\limits_{i}\dbasec_i, \prod\limits_j\dfiberc_j) = \bigsqcup\limits_i\prod\limits_j\vartistc(\dbasec_i, \dfiberc_j)$
+Artists naturally factorize into a two-stage process, starting with an encoding stage $\vchannelc$ followed by a compositing stage $\vmarkc$.
+In the \textcolor{artist}{encoding} stage $\vchannelc$, a data bundle is mapped to a visual variable bundle \vchannel, where the fibers of \vchannel\ are an internal library specification of graphics.
+The \textcolor{artist}{compositing} stage $\vmarkc$ then assembles the visual components into a visual element.
 
-\section{Construction}
-The artist can be constructed with an encoding stage $\vchannelc$ and a compositing stage $\vmarkc$ s.t $\vartistc=\vmarkc\circ\vchannelc$.
-In the \textcolor{artist}{encoding} stage $\vchannelc$, a data bundle is mapped to a visual variable bundle \vchannel, where the fibers of \vchannel\ are an internal library specification.
-The \textcolor{artist}{compositing} stage $\vmarkc$ is then responsible for assembling  the visual components into a visual element.
+\section{Conclusion}
+
+We have demonstrated a mathematical formalism for representing a data visualization pipeline, in which the mathematical structures both enforce well known principles of data visualization and suggest factorizations and decompositions resulting in a highly modularizable infrastructure, where artists can be constructed from composable primitive.
+This decomposability leads directly to a test framework ambition: if we can split $A=\vmarkc\circ\vchannelc$, and each factor $\vchannelc$, $\vmarkc$ can be further decomposed into composable primitives, we can focus testing on each of these building blocks reducing testing for correctness of the process as a whole to correctness of each building block, and correctness of their chosen combinations.
+
+Furthermore, while the graphics sheaves here deal with visual properties -- pixel positions and colors, et.c.
+-- one could easily imagine picking a different choice for the fiber $\gfiberc$ in the realization sheaf that tracks information about haptic or auditory expressions, resulting in an identical abstraction, with many overlapping building blocks, for multimodal data representations.
 

### 2025-07-28 short_paper/main.tex
captions + alt text + cleanup

@@
 \usepackage{mathptmx}
+
+% *** MATH PACKAGES ***
+%% with other math-related packages, you may want to disable it.
+\usepackage{amsmath, amsthm, amsfonts,amssymb,eulervm,xspace, mathtools}
+\usepackage{stmaryrd}%maps from
+%\renewcommand{\restriction}{\mathord{\upharpoonright}} %restriction w/p space
+\usepackage{mathrsfs} % math script fonts
+\usepackage{relsize} %bigger
+\usepackage{bm}
 
@@
 \usepackage{amsmath,amssymb}
-
-\usepackage{lipsum}
+\usepackage{mathtools}
+% *** diagrams
+\usepackage{tikz}
+\usetikzlibrary{cd}
+% *** ALIGNMENT PACKAGES ***
+\usepackage{tabu}                      % only used for the table example
+\usepackage{tabulary}
 \graphicspath{{../paper/figures/}{figs}}
@@
 
-\begin{figure}
+Motivated by wanting to make better tools, we propose a methodology for fully specifying arbitrary data to visualization mappings in a manner that easily translates to code.
+Furthermore, this methodology provides a consistent interface for accessibility tools to query for information about the state of the visualization.
+Generally, preserving structure means that a visualization is expected to preserve the $field$ properties and $topology$ of the corresponding dataset, where \textcolor{fiber}{\textbf{field}} is a set of values of the same type and \textcolor{base}{\textbf{topology}} is the connectivity and relative positioning of elements in a dataset \cite{wilkinsonGrammarGraphics2005}.
+Field structure is traditionally codified as the Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946}, where each scale is a set of actions on a group.
+Topological structure is generally assumed by visual algorithms \cite{chiTaxonomyVisualizationTechniques2000, toryRethinkingVisualizationHighlevel2004}, but generally these algorithms do not verify that input structure.
+\begin{figure}[!h]
 \begin{center}
-    \includegraphics[width=3.5in]{k_different_types.pdf}
+    \includegraphics[width=3.5in, alt={Data cube illustrating index space with map, timeseries, and bar charts visualizing slices of the data.}]{k_different_types.pdf}
 \end{center}
@@
 \end{figure}
-Motivated by wanting to make better tools, we propose a methodology for fully specifying arbitrary data to visualization mappings in a manner that easily translates to code.
-Generally, preserving structure means that a visualization is expected to preserve the $field$ properties and $topology$ of the corresponding dataset, where \textcolor{fiber}{\textbf{field}} is a set of values of the same type and \textcolor{base}{\textbf{topology}} is the connectivity and relative positioning of elements in a dataset \cite{wilkinsonGrammarGraphics2005}.
-Field structure is traditionally codified as the Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946}, where each scale is a set of actions on a group.
-Topological structure is generally assumed by visual algorithms \cite{chiTaxonomyVisualizationTechniques2000, toryRethinkingVisualizationHighlevel2004}, but generally do not verify that input structure.
+
 For example, a \texttt{line} algorithm often does not have a way to query whether a list of (x,y) coordinates is the distinct rows, the time series, or the list of stations in \autoref{fig:homeomorphism}.
 We propose that the bar plot, line plot, and heatmap can be verified as structure preserving because they have a homeomorphic relationship to the 0D ($\bullet$) points, 1D (--) linear, and 2D ($\blacksquare$) surface continuities embedded in the continuous 3 dimensional.
-Our methodology generalizes Bertin\cite{bertinSemiologyGraphicsDiagrams2011}'s codification of structure,  Mackinlay\cite{mackinlayAutomaticDesignGraphical1987} mathematical formalization, Wilkinson's set theoretic approach \cite{wilkinsonGrammarGraphics2005}, and Kindlemann and Scheidegger's algebraic visualization design (AVD) framework \cite{kindlmannAlgebraicProcessVisualization2014} by explicitly incorporating topology, supporting non-group and non-monoidal structures, and by providing a framework for translating the theoretical ideas into buildable components.
+Our methodology generalizes Bertin\cite{bertinSemiologyGraphicsDiagrams2011}'s codification of structure,  Mackinlay's\cite{mackinlayAutomaticDesignGraphical1987} mathematical formalization, Wilkinson's set theoretic approach \cite{wilkinsonGrammarGraphics2005}, and Kindlemann and Scheidegger's algebraic visualization design (AVD) framework \cite{kindlmannAlgebraicProcessVisualization2014} by explicitly incorporating topology, supporting non-group and non-monoidal structures, and providing a framework for translating the theoretical ideas into buildable components.
 
-Through this paper we will be discussing visualization primarily, but by changing the last of the three fiber bundles we discuss in Section~\ref{sec:data models}, we could go from a visual representation to an auditory, haptic, or any other modality and retain the benefits we identify from the fiber bundle and sheaf structures.
+Throughout this paper we will be discussing visualization primarily, but by changing the last of the three fiber bundles (the graphic bundle) we discuss in Section~\ref{sec:data models}, we could go from a visual representation to an auditory, haptic, or any other modality and retain the benefits we identify from the fiber bundle and sheaf structures.
 
@@
 
-Spivak represents relational databases with a fiber bundle representation \cite{spivakDatabasesAreCategories2010,spivakSimplicialDatabases2009}.
-This suggests a way for us to extend Butler's proposal of using fiber bundles as an abstraction for visualization data \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989} by incorporating Spivak's approach, and extending it from data storage to represent the transformation to graphical representation.
-As we show in Table~\ref{tab:data_abstraction}, where Spivak represents a specific dataset as a fiber, a database schema with the bundle's fiber space and topological indexing through a base space, we introduce parallel analogous structures representing a visual specification and a graphic realization.
+\begin{table}[!h]
+  \caption{Summary of how the data is abstracted using topological structures}
+  \label{tab:data_abstraction}
+  \centering
+  \scriptsize
+  \begin{tabu}{|r | l l l l|}
+    \hline
+    &\textcolor{base}{point}/\textcolor{base}{openset}/\textcolor{base}{base} & \textcolor{fiber}{fiber} & \textcolor{total}{total} & Fiber bundle\\
+     &  index/subset/indices & record/fields &  data type & \\
+    \hline
+   Data & $\dbasepointc \in \opensetc \subseteq \dbasec$ & $\delementc \in \dfiberc$ & \dtotalc &  $\dfiberc \hookrightarrow \dtotalc \rightarrow \dbasec$  \\
+   Visual & $\dbasepointc \in \opensetc \subseteq \dbasec$  & $\velementc \in \vfiberc$ & \vtotalc & $\vfiberc \hookrightarrow \vtotalc \rightarrow \dbasec$\\
+   Graphic & $\gbasepointc \in \opensetgc \subseteq \gbasec$ & $\gelementc \in \gfiberc$ & \gtotalc &$\gfiberc \hookrightarrow \gtotalc \rightarrow \gbasec$ \\
+   \hline
+  \end{tabu}
+\end{table}
 
-\begin{figure}
+We model data as a fiber bundle because it can encode topological properties, field types, and data values in a uniform manner.
+We extend Butler's proposal of using fiber bundles as an abstraction for visualization data \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989} by incorporating Spivak's fiber bundle representation of relational databases \cite{spivakDatabasesAreCategories2010,spivakSimplicialDatabases2009}.
+As we show in Table~\ref{tab:data_abstraction} and Table~\ref{tab:data_sheaves}, where Spivak represents a database schema with the bundle's fiber space \dfiber\, topological indexing through a base space \dbase\, and a specific dataset as a section of the bundle \dsection\, and we introduce parallel analogous structures representing visual and graphic specifications.
+
+\begin{figure}[!h]
     %\begin{multicols}{2}
-    \includegraphics[width=.5\textwidth]{fb_rho.pdf}
+    \includegraphics[width=.5\textwidth, alt={illustration of how a pixel in a scatter plot is the return value of a function from the graphic indexing space to the fiber space that specifies the display space}]{fb_rho.pdf}
 %    \columnbreak
-    \caption{The scatter marker is specified by the section $\gsection$, which maps into the fiber $\gfiber$ to retrieve the values that compose into the pixel (approximated as a square) returned by the section function evaluated at each point $\gbasepoint$.
+    \caption{The scatter marker is specified by the section $\gsection$, which maps from a point $\gbasepoint \in \gbase$ into the fiber $\gfiber$ to retrieve the values that compose into the pixel (approximated as a square).
 The section evaluated over the entire space $\gsection\vert_{S}$ returns the entire scatter mark, shown here in faded form to make it easier to see the individual pixels.}\label{fig:fiber_bundle}
 \end{figure}
-
-\begin{table*}[t]
-    \begin{tabular}{|r | c c c|}
-    \hline
-     structure/abstraction & Data & Visual & Graphic\\
-    \hline
-\textcolor{fiber}{fiber}/\textcolor{total}{total}\ /\textcolor{base}{base} spaces &   \multirow{2}{*}{$\dfiberc \hookrightarrow \dtotalc \xrightarrow{\pi} \dbasec$} &   \multirow{2}{*}{$\vfiberc \hookrightarrow \vtotalc \xrightarrow{\pi} \dbasec$} &  \multirow{2}{*}{$\gfiberc \hookrightarrow \gtotalc \xrightarrow{\pi} \gbasec$}\\
-fields/datasets/topology&&&\\
-\hline
-\textcolor{base}{point}/\textcolor{base}{openset}/\textcolor{base}{base space} &  \multirow{2}{*}{$\dbasepointc \in \opensetc \subseteq \dbasec$} & \multirow{2}{*}{$\dbasepointc \in \opensetc \subseteq \dbasec$} & \multirow{2}{*}{$\gbasepointc \in \opensetgc \subseteq \gbasec$}\\
-index/subset/indices & & & \\
-\hline
-\textcolor{fiber}{fiber space}  record/fields &  $\delementc \in \dfiberc$ & $\velementc \in \vfiberc$ & $\gelementc \in \gfiberc$ \\
-\hline
-\textcolor{section}{section} dataset & $\cgamma{\dbasec}{\dtotalc} \ni \dsectionc: \dbase \textcolor{section}{\rightarrow} \dfiberc$ & $\cgamma{\dbasec}{\vtotalc} \ni \vsectionc: \dbase \textcolor{section}{\rightarrow} \vfiberc$ &  $\cgamma{\gbasec}{\gtotalc} \ni \gsectionc: \gbase \textcolor{section}{\rightarrow} \gfiberc$ \\
-\hline
-\textcolor{sheaf}{sheaf} data container  &  \multirow{2}{*}{$\sheafc_{\dbasec, \dtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$} &  \multirow{2}{*}{$\sheafc_{\dbasec, \vtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}}$} &   \multirow{2}{*}{$\sheafc_{\gbasec, \gtotalc}: \opensetgc \rightarrow \cgamma{\opensetc}{\gtotalc\restriction_{\opensetgc}}$} \\
-&&&\\
-        \hline
-    \end{tabular}
-    \caption{Summary of how the data is abstracted using topological structures}
-    \label{tab:data_abstraction}
-\end{table*}
 
@@
 This fits topology and field types into a data definition $\textcolor{section}{\texttt{dataset}}: \textcolor{base}{\texttt{topology}} \rightarrow \textcolor{fiber}{\texttt{field}}$.
+We abstract these data containers as sheaves $\sheafc$, as summarized in \autoref{tab:data_sheaves} and the gluing property of sheaves provides a systematic way to track continuity of the data \cite{ghristElementaryAppliedTopology2014} and apply visual algorithms developed on this model regardless whether the data fits in memory, is distributed, or streaming.
 
-We abstract these data containers as sheaves, and the gluing property of sheaves provides a systematic way to track continuity of the data \cite{ghristElementaryAppliedTopology2014} and apply visual algorithms developed on this model regardless whether the data fits in memory, is distributed, or streaming.
+\begin{table}
+  \caption{Functions that associate topological subspaces with records}
+  \label{tab:data_sheaves}
+  \centering
+   \scriptsize%
+  \begin{tabu}{|r | l l | }
+    \hline
+     & \textcolor{section}{section} & \textcolor{sheaf}{sheaf} \\
+     & record at location & set of possible records for subset \\
+     \hline
+  Data & $ \cgamma{\dbasec}{\dtotalc} \ni \dsectionc: \dbase \textcolor{section}{\rightarrow} \dfiberc$ & $\sheafc_{\dbasec, \dtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$\\
+  Visual &  $\cgamma{\dbasec}{\vtotalc} \ni \vsectionc: \dbase \textcolor{section}{\rightarrow} \vfiberc$ & $\sheafc_{\dbasec, \vtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}}$\\
+  Graphic &    $\cgamma{\gbasec}{\gtotalc} \ni \gsectionc: \gbase \textcolor{section}{\rightarrow} \gfiberc$ &  $\sheafc_{\gbasec, \gtotalc}: \opensetgc \rightarrow \cgamma{\opensetc}{\gtotalc\restriction_{\opensetgc}}$ \\
+  \hline
+  \end{tabu}
+\end{table}
 
@@
 
-We call the mapping from data to graphic $\vartistc:\dsectionc\mapsto\gsectionc$ an \textcolor{artist}{Artist}\footnote{Artists are named for the Artist class in Matplotlib.
-Artist is the parent class of all the objects that assemble and manage visual elements such as points, lines, and shapes.}.
-Artists are structure preserving when they are implemented as sheaf maps from the data representation sheaf to the graphics sheaf.
+We call the mapping from data to graphic $\vartistc:\dsectionc\mapsto\gsectionc$ an \textcolor{artist}{Artist}.
+Artists are named for the artist class in Matplotlib because the Artist class (and its children) assemble and manage the visual
+elements produced in Matplotlib\cite{hunterArchitectureOpenSource}.
+We propose that Artists (the transform and the library implementation) are structure preserving when they are implemented as sheaf maps from the data representation sheaf to the graphics sheaf.
 This means that the artist satisfies the following constraints, as illustrated in \autoref{fig:artist}:
 
-\begin{figure}
-    \includegraphics[width=.5\textwidth]{artist_equiv.pdf}
-    \caption{}\label{fig:artist}
+\begin{figure}[!h]
+    \includegraphics[width=.5\textwidth, alt={Diagram of a dirichelet distribution evaluated on a triangle, transposed, and binned and the corresponding visualization output by the artist function.}]{artist_equiv.pdf}
+    \caption{The transposition of the Dirichilet distribution is preserved in the visualization because the Artist implements presheaf constraints and the binning of the Dirichelet data is preserved in the visualization because the artist is equivariant.}\label{fig:artist}
 \end{figure}
@@
 \begin{description}
-    \item[homotopic] there exists a deformation retract\footnote{This requirement emerges from the digital geometry nature of displays: even if the intended visual primitive may be 1-dimensional, it's screen realization must be 2-dimensional -- requiring a deformation retract most closely agrees with our intuition for drawing geometric primitives} $\vindexc: \gbasec\rightarrow\dbasec$ between graphic and artist base space
+    \item[homotopic] there exists a deformation retract\footnote{A deformation retract is required because it ensures a continuous mapping between graphic and data continuity when a screen is 2-dimensional and the data continuity is <2D]} $\vindexc: \gbasec\rightarrow\dbasec$ between graphic and artist base space
     \item[topology] the artist is commutative w.r.t.
-the presheaf morphism $\dfunchc: \dbasepointc^{\prime}\mapsto \dbasepointc$ and pullback $\dfuncpullc \dsectionc \restriction_{\opensetc^{\prime}}: \dsectionc\restriction_{\opensetc^{\prime}} \mapsto \dsectionc \restriction_{\opensetc^{\prime} \circ \dfunchc}$: index changes on the record leaves values in the record unmodified
+the presheaf morphism $\dfunchc: \dbasepointc^{\prime}\mapsto \dbasepointc$ and pullback $\dfuncpullc \dsectionc \restriction_{\opensetc^{\prime}}: \dsectionc\restriction_{\opensetc^{\prime}} \mapsto \dsectionc \restriction_{\opensetc^{\prime} \circ \dfunchc}$: index changes on the record leave values in the record unmodified
     $\dsectionc(\dbasec) = \dfuncpullc\dsectionc(\dbasepointc^{\prime}) = \dsectionc(\dfunchc(\dbasepointc^{\prime}))$
@@
 An action $\dfunctc: \dfiberc\to\dfiberc$ induces a function between sections $\dfunctc: \dfuncpullc \dsectionc \restriction_{\opensetc^{\prime}} \mapsto \dfuncpullc \dsectionc^{\prime} \restriction_{\opensetc}$.
-We also require $\dfunctc$ to respect the structure, such as operators, of $\dfiber$.
+We also require $\dfunctc$ to respect the structure, such as operators, of $\dfiberc$.
 \end{description}
@@
 With this underlying topological structure to describe the visualization pipeline, traditional constructions from topology and category theory generate natural operators on the data structures.
+\begin{figure}[!h]
+\includegraphics[width=3.5in, alt={Records are added by taking the union of base spaces. Fields are added by taking the cartesian product of fiber spaces.}]{more_records.pdf}
+\caption{Pullback (left) and pushout (right) structure of the operations that add records (increasing the indexing base space) or expand records (merging different fiber spaces with a database join-like operation).}\label{fig:more-records}
+\end{figure}
+
 Pushouts and pullbacks, of central importance in category theory and topology, emerge here as operators to add and extend records, respectively.
@@
 \begin{figure*}[h]
-\includegraphics[width=\textwidth]{more_records.pdf}
-\caption{Pullback (left) and pushout (right) structure of the operations that add records (increasing the indexing base space) or expand records (merging different fiber spaces with a database join-like operation).}\label{fig:more-records}
-\end{figure*}
-
-\begin{figure*}[h]
-    \includegraphics[width=7.16in]{xi_diagram.pdf}
+    \includegraphics[width=7.16in, alt={An annulus indexing space deformation retracts into a circle.
+For any point i, cos(i), sin(i), there is a specification to draw a piece of a circle, sine plot, and cosine plot, a  patch of pixels of a rendered circle, sin plot, and cosine plot.
+Hovering over any of those pixels should get back the data i, cos(i), sin(i)}]{xi_diagram.pdf}
     \caption{A deformation retract $\vindexc$ that relates the graphics footprint $S_0$ of an abstract circle to the circle itself gives rise to a pair of additional maps.
@@
 As illustrated in \autoref{fig:xi}, this deformation retract $\vindexc$ gives rise to two so-called transport functors $(\vindexpush, \vindexpull)$ -- known as the pushforward $\vindexpushc$ and pullback\footnote{different use of the word pullback than in the preceding paragraph -- an unfortunate collision of terminology} $\vindexpullc$.
-These also codify how visual elements correspond to distinct data elements, which is a necessary condition of a visualization being readable\cite{ziemkiewiczEmbeddingInformationVisualization2009}.
+These functors also codify how visual elements correspond to distinct data elements, which is a necessary condition of a visualization being readable\cite{ziemkiewiczEmbeddingInformationVisualization2009}.
 The pushforward $\vindexpushc$ matches each point in the data space to the specification of the graphic at that point $\vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexprec(\dbasepointc)}$, while the pullback $\vindexpull$ matches each point in the graphic space to the data over that point $\vindexpullc\dsectionc(\gbasepointc) = \dsectionc(\vindexc(\gbasepointc)) = \dsectionc(\dbasepointc)$.
+If the map $\vindex$ was explicitly implemented with the corresponding sheaves, the information in $\dsection, \vindexpull\dsection,\gsection, \vindexpush\gsection$, as illustrated in \autoref{fig:xi}, could be exposed to an accessibility tool (like a screen reader).
+This functionality could, for example, compensate for a major accessibility limitation of the current Matplotlib architecture, which is that the interactive widgets and GUIs are rendered as PNGs (so that Matplotlib does not have to maintain a GUI library) which means that screen readers cannot query the individual elements.
 
 \section{Construction}
+\begin{figure}[!h]
+  \includegraphics[width=1\columnwidth, alt={cos(i) and sin(i) gets encoded as x and y values that for a patch of a green circle.
+That the visual mark is a green line can either be encoded in the compositor function or passed to the compositor function as part of the encoding spec.
+Green and line can be pulled back to represent constant parameters of the data.}]{construction.pdf}
+  \caption{The compositor $\vmark$ can be fully specified $\vmark_{green,line}$ such that the only input is the encoding of the data values.
+Alternatively, the compositor $\vmark$ supports more customizability by encoding  a constant parameters in the visual bundle, which pulls back to constant parameters in the data bundle.}\label{fig:construction:q}
+\end{figure}
 
 Artists naturally factorize into a two-stage process, starting with an encoding stage $\vchannelc$ followed by a compositing stage $\vmarkc$.
-In the \textcolor{artist}{encoding} stage $\vchannelc$, a data bundle is mapped to a visual variable bundle \vchannel, where the fibers of \vchannel\ are an internal library specification of graphics.
+In the \textcolor{artist}{encoding} stage $\vchannelc$, a data bundle is mapped to a visual variable bundle \vchannel.
+Since $\dtotal$ and $vtotal$ are structurally identical, any $\vtotal$ can be redefined as $\dtotal$; therefore, as shown in \autoref{eq:construction:nu:fabrication}, any collection of $\vchannel$ functions can be composed such that they are equivalent to a $\vchannel$ that directly converts the input to the output.
+\begin{equation}
+  \label{eq:construction:nu:fabrication}
+  \begin{tikzcd}
+    \dfiberc_{\dbasepointc}
+    \arrow[rr, "\vchannelc", color=artist]
+    \arrow[rrrr, "\vchannelc^{\prime\prime}", dashed, bend right, color=artist] &  &
+    \vfiberc_{\dbasepointc}\coloneqq{\dfiberc_{\dbasepointc}^{\prime}}
+    \arrow[rr, "\vchannelc^{\prime}", color=artist] &  &
+    \vfiberc^{\prime}_{\dbasepointc}
+  \end{tikzcd}
+\end{equation}
+
+ As with artists, $\vchannel$ are maps of sections such that operators can also act on transformers $\vchannel$, meaning that encoders can be added $\vchannel_{a+b} = \vchannel_{a} + \vchannel_{b}$ and multiplied d $\vchannel_{a\times b} = \vchannel_{a}  \vchannel_{b}$.
+
+\begin{table}[!h]
+  \centering
+  \begin{tabulary}{\columnwidth}{|l|L|l|}\hline
+   \(\bm{\vchannel_{i}}\)    & \(\bm{\vsection_{i}}\)  & \(\bm{codomain(\vchannel_{i}) \subset \vfiber_{i}}\)  \\ \hline
+  position                    & x, y, z, theta, r      & \(\mathbb{R}\)   \\ \hline
+  size                        & linewidth, markersize  & \(\mathbb{R}^{+}\)  \\ \hline
+  shape                       & markerstyle            & \(\{f_{0}, \ldots, f_{n}\}\)\\ \hline
+  color                       & color, facecolor, markerfacecolor, edgecolor  & \(\mathbb{R}^{4}\) \\ \hline
+  \multirow{2}{*}{texture}    & hatch      & \(\mathbb{N}^{10}\)\\\cline{2-3}
+                              & linestyle    & \((\mathbb{R}, \mathbb{R^+}^{n, n\%2=0})\) \\ \hline
+  \end{tabulary}
+  \caption{Some of the $\vfiber$ components of the $\vtotal$ bundles in Matplotlib components}
+  \label{tab:math:artist:mpl:fiber}
+\end{table}
+
+The fibers of \vchannel\ are an internal library specification of graphics.
+For example, \autoref{tab:math:artist:mpl:fiber} shows a standardized internal representation of the visual parameters used by Matplotlib functions.
+One of the ways Matplotlib facilitates accessibility is by allowing the user to customize most of the aesthetics - colors, fonts, textures.
+Standardizing the internal parameters provides for a consistent interface for external accessibility tools, such as libraries of accessible stylesheets or tools that may want to query the Matplotlib Artist for context of the current state.
+
 The \textcolor{artist}{compositing} stage $\vmarkc$ then assembles the visual components into a visual element.
+Since encoder functions are infinitely composable, as described in \autoref{eq:construction:nu:fabrication}, a new compositor function $\vmarkc$ can be constructed by pre=composing $\vchannelc$ functions with the existing $\vmarkc$.
+\begin{equation}
+  \label{eq:construction:q:fabrication}
+  \begin{tikzcd}
+      \cgamma{\dbasec}{\vtotalc}
+      \arrow[rr, "\vchannel", color=artist]
+      \arrow[rrrr, "\vmarkc^{\prime}", bend right, color=artist, dashed] &  & \cgamma{\dbasec}{\vtotalc^{\prime}}
+      \arrow[rr, "\vmarkc", color=artist] &  & \cgamma{\gbasec}{\gtotalc}
+      \end{tikzcd}
+\end{equation}
+The composition in \autoref{eq:construction:q:fabrication} means that different measurable components can yield the same visual elements.
+Operators can also act on compositors $\vmark$ such that $\vmark_{a+b} = \vmark_{a} + \vmark_{b}$ and multiplied d $\vmark_{a\times b} = \vmark_{a}  \vmark_{b}$.
+
+\begin{figure}[!h]
+  \includegraphics[width=\columnwidth, alt={The data value JFK INTL AP is paired with different variables (pressure and temperature) to verify that JFK is consistently encoded as the same color while pressure and temperature are encoded as different marker styles.
+The markerstyles and colors are then tested to ensure that the compositor builds a square and a circle that are the same color.}]{nu_qu.pdf}
+  \caption{The composability of the model allows for separation of concern in testing: the encoding functions $vchannel$ are tested to ensure that they convert the data into the internal specification, while the compositors are tested to ensure that the visual specifications are encoded consistently.}\label{fig:construction:verification}
+\end{figure}
+As illustrated in \autoref{fig:construction:verification}, the factorization of the artist into the encoding and composition stages provides a framework for testing each independently.
+For example, encoders build in this framework would be designed to ensure that a transformation from a non colorblind safe colormap into a colorblind safe colormap would ensure that the colormaps are equivalent such that the transformation is not introducing a change in structure.
+Similarly, the composition stage can be tested to ensure that parameters are applied consistently across visual elements and, as with encoding, more accessible elements can be swapped in without introducing new structure.
 
@@
 
-We have demonstrated a mathematical formalism for representing a data visualization pipeline, in which the mathematical structures both enforce well known principles of data visualization and suggest factorizations and decompositions resulting in a highly modularizable infrastructure, where artists can be constructed from composable primitive.
+We have demonstrated a mathematical formalism for representing the architecture of components in a data visualization library, in which the mathematical structures both enforce well known principles of data visualization and suggest factorizations and decompositions the result in highly modularizable infrastructure, such that artists can be constructed from composable primitive.
 This decomposability leads directly to a test framework ambition: if we can split $A=\vmarkc\circ\vchannelc$, and each factor $\vchannelc$, $\vmarkc$ can be further decomposed into composable primitives, we can focus testing on each of these building blocks reducing testing for correctness of the process as a whole to correctness of each building block, and correctness of their chosen combinations.
-
+The composability of this framework allows for staged adaptation such that it can be incorporated into an existing library, as exemplified by \url{https://github.com/matplotlib/data-prototype}.
 Furthermore, while the graphics sheaves here deal with visual properties -- pixel positions and colors, et.c.
@@
 
-\section*{Note}
-This is a work in progress, more information can be found at https://github.com/story645/team
+\section*{Supplemental Materials}
+\label{sec:supplemental_materials}
+This paper reuses figures from a longer version of this paper; therefore all code for the figures can be found in \url{https://github.com/story645/team/paper/figcode}.
+Additionally, there is an attached supplement with a glossary of math terms used in this paper.
 

### 2025-07-29 paper/supplement.tex
accessviz submission

@@
 \usepackage{bm}
-
+\usepackage{stmaryrd}%maps from
+%\renewcommand{\restriction}{\mathord{\upharpoonright}} %restriction w/p space
+\usepackage{mathrsfs} % math script fonts
 % theorem environments
@@
 \usepackage{mathptmx}
-
+\DeclareMathDelimiter{(}{\mathopen} {operators}{"28}{largesymbols}{"00}
+\DeclareMathDelimiter{)}{\mathclose}{operators}{"29}{largesymbols}{"01}
 % *** SUBFIGURE PACKAGES ***
@@
 
-\usepackage[inline]{showlabels}
+%\usepackage[inline]{showlabels} %show equation labels
 
@@
 \begin{document}
-
-\firstsection{Structure preservation}
 \maketitle
 
+\firstsection{Categories}
+\label{sec:category}
+\begin{definition}\label{def:atct:category}
+   An \textbf{category} $\mathcal{C}$ consists of the following \textit{data}:
+\begin{enumerate}
+  \item a collection of \textit{objects} $X \in \textbf{ob}(\mathcal{C})$
+  \item for every pair of objects $X, Y \in \textbf{ob}(\mathcal{C})$, a set of \textit{morphisms} $X \xrightarrow{f} Y \in Hom_{\mathcal{C}}(X, Y)$
+  \item for every object $X$, a distinct \textit{identity morphism} $X \xrightarrow {id_x} X$ in $Hom_{\mathcal{C}}(X, X)$
+  \item a \textit{composition function} $f \in Hom_{\mathcal{C}}(X, Y) \times  g \in Hom_{\mathcal{C}}(Y, Z) \rightarrow g \circ f \in Hom_{\mathcal{C}}(X, Z)$
+\end{enumerate}
+such that
+\begin{enumerate}
+  \item \textit{unitality:} for every morphism $ X \xrightarrow{f} Y$, $f \circ id_x = f = id_y \circ f$
+  \item \textit{associativity:} if any three morphisms $f, g, h$ are composable,
+    \begin{equation*}
+      \begin{tikzcd}
+        X \arrow[r, "f"] \arrow[rrr, "h\circ(g\circ f) = (h\circ g)\circ f"', bend right, dashed] & Y  \arrow[r, "g"] & Z \arrow[r, "h"] & W
+        \end{tikzcd}
+  \end{equation*}
+  then they are associative such that $h\circ(g\circ f) = (h \circ g) \circ f$  \cite{lawvere2009conceptual,riehlCategoryTheoryContext,maclaneCategoriesWorkingMathematician2013,fongInvitationAppliedCategory2019}.
+  \end{enumerate}
+\end{definition}
+
+The composability property expresses that the morphism is transitive, while associativity expresses that the morphisms can be curried in various equivalent groupings.
+By formally specifying the properties of the topological structure data types as $\mathcal{\dbase}$, we can express that these are the properties that are required as part of the implementation of the data type objects.
+
+\subsection{Functor}
+\begin{definition}\cite{bradleyWhatFunctorDefinitions,bradleyTopologyCategoricalApproach2020} A \textbf{functor} is a map $F: \mathcal{C} \rightarrow \mathcal{D}$, which means it is a function between objects $F: \textbf{ob}(\mathcal{C}) \mapsto \textbf{ob}(\mathcal{D})$ and that for every morphism $f \in Hom(C_1, C_2)$  there is a corresponding function $F: Hom(C1, C2) \mapsto Hom(F(C_1), F( C_2))$.
+A \textbf{functor} must satisfy the properties
+\begin{itemize}
+  \item \textit{identity}: $F(id_{C}(C)) = id_{D}(F(C))$
+  \item \textit{composition}: $F(g)\circ F(f) = F(g\circ f)$ for any composable morphisms $C_{1}\xrightarrow{f} C_2$, $C_2 \xrightarrow{g} C_3$
+\end{itemize}
+$F(C) \in \textbf{ob}(\mathcal{D})$ denotes the object to which an object $C$ is mapped, and $F(f) \in Hom(F_(C_1), F_(C_2))$ denotes the morphism that $f$ is mapped to.
+\end{definition}
+
+\subsection{Natural transforms}
+\begin{definition}\label{def:natural-transform}
+  Given two functors $F, G: \mathcal{C}\rightarrow \mathcal{D}$, a \textbf{natural transformation} $\alpha: F \rightarrow G$ is a function which assigns to each object $c$ of $\mathcal{C}$ a morphism $\alpha_c:F(input) \rightarrow G(c), G(c) \in \mathcal{D}$, in such a way that for every morphism $f:c \rightarrow c^\prime, c^\prime \in \mathcal{C}$, the morphisms in $\mathcal{D}$ commute such that $\alpha_c^{\prime}(F(f)(F(c))) = G(f)(\alpha_c(F(c))$.
+When this holds, $\alpha_{c}$ is \textit{natural} in $c$.\cite{maclaneCategoriesWorkingMathematician2013}.
+\end{definition}.
+
+\section{Structure}
+
 \subsection{Actions}
- Loosely we define structure as a mathematical space $X$ and the actions on that space
+Since our proposed methods work well for a range of possible settings, we abstract the notion of data type into some object in a category with some monoid action on the object.
+Here, we deliberately refrain from stating whether the structure is a group, a vector space, a partial order, or some other structure - just that there is some category the data comes from.
+Here, by a monoid action of a monoid $M$ on a set $X$ we mean a function $act: G\times X \rightarrow X$:
 
-\begin{definition}\label{def:related-work:action}\cite{grimaldiDiscreteCombinatorialMathematics2006}
-  An \textcolor{action}{\textbf{action}} of \textcolor{action}{$G = (G,\circ, e)$} on $X$ is a function  $act: \textcolor{action}{G} \times X \rightarrow X$.
-An action has the properties of identity $act(\textcolor{action}{e}, x) = x$ for all  $x \in X$ and associativity $act(\textcolor{action}{g}, act(\textcolor{action}{f}, x)) = act(\textcolor{action}{f} \circ \textcolor{action}{g}, x)$ for $\textcolor{action}{f},\textcolor{action}{g} \in \textcolor{action}{G}$.
+\begin{definition}\label{def:related-work:action}
+  An \textcolor{action}{\textbf{action}} is a function  $act: \textcolor{action}{G} \times X \rightarrow X$.
+An action has the properties of identity $act(\textcolor{action}{e}, x) = x$ for all  $x \in X$ and associativity $act(\textcolor{action}{g}, act(\textcolor{action}{f}, x)) = act(\textcolor{action}{f} \circ \textcolor{action}{g}, x)$ for $\textcolor{action}{f},\textcolor{action}{g} \in \textcolor{action}{G}$.\cite{grimaldiDiscreteCombinatorialMathematics2006}
 \end{definition}
@@
 
-\subsection{Equivariance}
-A mathematical {group} is a set with an associative binary operator.
-This operation must have an identity element and be closed, associative, and invertible, consisting of a set of values $X$ and \textcolor{action}{actions} on the set {$G = (G,\circ, e)$}.
-\begin{figure}
-  \includegraphics*[width=1\columnwidth]{equivariant.pdf}
-  \caption{Encoding data as the bar height using an exponential transform is not equivariant because encoding the data and then scaling the bar heights yields a much taller graph then scaling the data and then encoding those heights using the same exponential transform function.}
-  \label{fig:related-work:equivariance}
-\end{figure}
-
- Given a group $G$ that acts on both the input $X$ and the output $Y$ of a function $f: X \rightarrow Y$
-
-\begin{definition}\label{def:equivariance}
- A function $f$ is \textbf{equivariant} when $f(act(g,x)) = act(g,f(x))$ for all $g$ in $G$ and for all $x$ in $X$ \cite{pittsNominalSetsNames2013}
-\end{definition}
-which means that
- As illustrated in the commutative diagram in \autoref{fig:related-work:equivariance}, what this means is that the visual representation is consistent whether the data is scaled and then mapped to a graphic or whether the data is mapped to a graphic that is then modified in a compatible way.
-
-\subsection{Homomorphism}
-
-\begin{figure}
-  \includegraphics[width=1\columnwidth]{homomorphism.pdf}
-  \caption{Encoding data as bar height using an inverse transform is not homomorphic because the largest number is mapped to the smallest bar while the max function returns the largest bar.}
-  \label{fig:related-work:homomorphism}
-\end{figure}
-
-Given the function $f: X \rightarrow Y$, with operators $(X, \circ)$ and $(Y, *)$
-
-\begin{definition}\label{def:homomorphism}
-  A function $f$ is \textbf{homomorphic} when $f(x_1 \circ x_2) = f(x_1) * f(x_2)$ and preserves identities $f(I_x) = I_y$ all $x, y \in X$ \cite{grimaldiDiscreteCombinatorialMathematics2006}
-\end{definition}
-
-which means that the operators $\circ$ and $*$ are compatible.
-In \autoref{fig:related-work:homomorphism}, the $\geq$ operator is defined as the compatible closed functions \texttt{max} and the inverse transform
-is not homomorphic because it does not encode the maximum data value as the maximum bar value.
-
-\subsection{Topology and Homeomorphism}
-
-We encode the continuity of the datasets using their topological structure, specifically using the neighborhood definition:
-
+\subsection{Topology}
 Given a set $X$ and a function $\mathcal{N}:X\to 2^{2^X}$ that assigns to any $x\in X$ a non-empty collection of subsets $\mathcal{N}(x)$, where each element of $\mathcal{N}(x)$ is a \emph{neighborhood of $x$}, then $X$ with  $\mathcal{N}$ is a \textcolor{base}{topological space} and $\mathcal{N}$ is a neighborhood \emph{topology} if for each $x$ in $X$: \cite{brownronaldTopologyGroupoids2006}
@@
 
-We then use this topological structure as the basis of claiming topological equivalence between a data and a graphic if there exists a homeomorphic map between the continuities of the data and graphic:
+\section{Structure preservation}
+
+\subsection{Equivariance}
+A mathematical {group} is a set with an associative binary operator.
+This operation must have an identity element and be closed, associative, and invertible, consisting of a set of values $X$ and \textcolor{action}{actions} on the set {$G = (G,\circ, e)$}.
+
+\begin{definition}\label{def:equivariance}
+  Given a group $G$ that acts on both $X$ and $Y$, we say that a function $f: X \rightarrow Y$ is \textbf{equivariant} when $f(act(g,x)) = act(g,f(x))$ for all $g$ in $G$ and for all $x$ in $X$ \cite{pittsNominalSetsNames2013}
+\end{definition}
+
+\begin{figure}[H]
+  \includegraphics*[width=1\columnwidth]{equivariant.pdf}
+  \caption{Encoding data as the bar height using an exponential transform is not equivariant because encoding the data and then scaling the bar heights yields a much taller graph then scaling the data and then encoding those heights using the same exponential transform function.}
+  \label{fig:related-work:equivariance}
+\end{figure}
+
+ As illustrated in the commutative diagram in \autoref{fig:related-work:equivariance}, what this means is that the visual representation is consistent whether the data is scaled ($act(g,x)$) and then mapped ($f$) to a graphic or whether the data is mapped to a graphic that is then modified in a compatible way.
+
+\subsection{Homomorphism}
+Given the function $f: X \rightarrow Y$, with operators $(X, \circ)$ and $(Y, *)$
+\begin{definition}\label{def:homomorphism}
+  A function $f$ is \textbf{homomorphic} when $f(x_1 \circ x_2) = f(x_1) * f(x_2)$ and preserves identities $f(I_x) = I_y$ for all $x, y \in X$ \cite{grimaldiDiscreteCombinatorialMathematics2006}
+\end{definition}
+which means that the operators $\circ$ and $*$ are compatible.
+
+\begin{figure}[H]
+  \includegraphics[width=1\columnwidth]{homomorphism.pdf}
+  \caption{Encoding data as bar height ($f$) using an inverse transform $\circ$ is not homomorphic because the largest number is mapped to the smallest bar while the max function ($*$)returns the largest bar.}
+  \label{fig:related-work:homomorphism}
+\end{figure}
+
+In \autoref{fig:related-work:homomorphism}, the $\geq$ operator ($\circ$) is defined as the compatible closed functions \texttt{max}($*$) and the inverse transform is not homomorphic because it does not encode the maximum data value as the maximum bar value.
+
+\subsection{Homeomorphism}
 
 \begin{definition}
-  A function $f$ is a $homeomorphism$ if it is bijective, continuous, and has a continuous inverse function $f^{-1}$.
+  A function $f$ is a \textit{homeomorphism} if it is bijective, continuous, and has a continuous inverse function $f^{-1}$.
 \end{definition}
+
+A function between topological spaces is continuous if the inverse image of any open sets is open.
+If two spaces are homeomorphic, they have identical topological properties and are considered topologically the same.
+
+\subsection{Deformation retraction}
+\begin{definition}
+  A \textit{deformation retraction} of a space $X$ onto a subspace A is a family of maps $f_{t}: X \rightarrow X, t \in I$ such that $f_0 =\mathbb{1}$ (identity), $f_{1}(X)=A$, and $f_{t}|A=\mathbb{1}$ for all $t$.
+The family $f_{t}$ should be continuous in the sense that the associated map $X \times I \rightarrow X, (x,t)
+ \mapsto f_{t}(x)$ is continuous. \cite{hatcherAlgebraicTopology2002} \end{definition}
+
+\begin{figure}[H]
+  \includegraphics[width=1\columnwidth]{xi_zoom.pdf}
+  \caption{The deformation retraction $\xi$ maps each point in each orange band on each space $\gbase_0, \gbase_1, \gbase_2$ to the same point $\dbasepoint_i$ in the circle $\dbase$.\label{fig:artist:xi}}
+\end{figure}
+
+For example, the map $\vindex$ in \autoref{fig:artist:xi} is a deformation retraction between the spaces \gbase\ and \dbase\:
+\begin{align}
+  \vindexc\textcolor{functor}{:} \underbrace{\dbasec\times[0,1]^{n}}_{\gbasec} \textcolor{functor}{\mapsto} \dbasec
+&&
+  n = \begin{cases}
+    dim(\gbasec) - dim(\dbasec) & dim(\dbasec)<dim(\gbasec)\\
+  0 & otherwise
+  \end{cases}
+\end{align}
 
@@
 
-\begin{figure}[!h]
+\begin{figure}[H]
        \includegraphics[width=1\columnwidth]{fb.pdf}
@@
 Generally, the distinguishing factor between a trivial bundle and a non-trivial bundle are how they are decomposed into local trivializations:
-\begin{LaTeXdescription}
+\begin{description}
   \item[\textit{trivial bundle}] is directly isomorphic to $\dbasec\times\dfiberc$.
@@
 For any choice of local trivializations, there is at least one transition map that is not an identity \cite{hatcherAlgebraicTopology2002}.
-\end{LaTeXdescription}
+\end{description}
 
@@
 
-\section{Categories}
-We propose that the various structures of the bundle (base space, fiber, total space) can also be encoded as categories:
-
-\begin{definition}\label{def:atct:category}
-   An \textbf{category} $\mathcal{C}$ consists of the following \textit{data}:
-\begin{enumerate}
-  \item a collection of \textit{objects} $X \in \textbf{ob}(\mathcal{C})$
-  \item for every pair of objects $X, Y \in \textbf{ob}(\mathcal{C})$, a set of \textit{morphisms} $X \xrightarrow{f} Y \in Hom_{\mathcal{C}}(X, Y)$
-  \item for every object $X$, a distinct \textit{identity morphism} $X \xrightarrow {id_x} X$ in $Hom_{\mathcal{C}}(X, X)$
-  \item a \textit{composition function} $f \in Hom_{\mathcal{C}}(X, Y) \times  g \in Hom_{\mathcal{C}}(Y, Z) \rightarrow g \circ f \in Hom_{\mathcal{C}}(X, Z)$
-\end{enumerate}
-such that
-\begin{enumerate}
-  \item \textit{unitality:} for every morphism $ X \xrightarrow{f} Y$, $f \circ id_x = f = id_y \circ f$
-  \item \textit{associativity:} if any three morphisms $f, g, h$ are composable,
-    \begin{equation*}
-      \begin{tikzcd}
-        X \arrow[r, "f"] \arrow[rrr, "h\circ(g\circ f) = (h\circ g)\circ f"', bend right, dashed] & Y  \arrow[r, "g"] & Z \arrow[r, "h"] & W
-        \end{tikzcd}
-  \end{equation*}
-  then they are associative such that $h\circ(g\circ f) = (h \circ g) \circ f$  \cite{lawvere2009conceptual,riehlCategoryTheoryContext,maclaneCategoriesWorkingMathematician2013,fongInvitationAppliedCategory2019}.
-  \end{enumerate}
+\section{Presheaves and Sheaves}
+A common way of encapsulating a map from an arbitrary category to a category of sets is as a presheaf and sheaves add on conditions for gluing individual sections over subspaces into cohesive sections over the whole space.
+\subsection{Presheaf}
+\begin{definition}\label{def:atct:presheaf}
+  A \textbf{presheaf} $F:\mathcal{C}^{op} \rightarrow \setb$ is a contravariant functor from an object in an arbitrary category to an object in the category \setb\cite{spanier1989algebraic}.
 \end{definition}
 
-The composability property expresses that the morphism is transitive, while associativity expresses that the morphisms can be curried in various equivalent groupings.
-By formally specifying the properties of the topological structure data types as $\mathcal{\dbase}$, we can express that these are the properties that are required as part of the implementation of the data type objects.
+A functor is contravariant when the morphisms between the input objects go in the opposite direction from the morphisms between the output objects.
+The presheaf is contravariant when for every arbitrary morphism between input base spaces $\dfunch: \openset_1 \rightarrow \openset_2$ there exists a corresponding pullback function between the sets of sections $\dfuncpull: \cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}} \rightarrow \cgamma{\opensetc_1}{\dtotalc\restriction_{\openset_c1}}$.
 
-\section{Presheaves And Sheafs}
+\begin{figure}[H]
+  \includegraphics*[width=1\columnwidth]{figures/tex/presheaf.pdf}
+  \caption{Modeling this data container as a presheaf specifies that since $\cos, \sin, and \texttt{C}$ are continuous over $\openset_{2}$, they must be continuous over $\openset_{1}$ since $\openset_1$ is a subset of and therefore must be included in $\openset_2$.
+Because $\tan$ is only defined over $\openset_{2}$, it does not need to be included in the set $\Gamma_{2}$.
+\label{fig:atct:presheaf}}
+\end{figure}
+
+For example in \autoref{fig:atct:presheaf}, the transform $\dfunch$ maps indexing points on the subspace $\openset_{1}$ to points on the larger subspace $\openset_{2}$.
+A constraint on presheaves is that because $\dfunch$ exists, there must exist a corresponding pullback map $\dfuncpull$ that changes the indexes of the data indexed by $\openset_{2}$ to the corresponding index in $\openset_{1}$.
+Since the $constant, sin, cos$ functions are defined over the interval $\left[0,2\pi\right]$, these functions must also be continuous over the sub-interval $\left(\frac{\pi}{2}, \frac{3\pi}{2}\right)$; therefore the sections in $\Gamma_{2}$ must also be included in the set of sections over the subspace $\Gamma_{1}$.
+Sections (data) can also be present in subindexes without being present over the whole indexing space, as illustrated here by the presence of $tan$ over $\openset_{1}$.
+
+\subsection{Sheaf}
+
+\begin{definition}\label{def:atct:sheaf}\cite{bakerEuclideanSpaceMathsSheaf,spanier1989algebraic} A \textbf{sheaf} is a presheaf that satisfies the following two axioms
+\begin{itemize}
+  \item \textit{locality} two sections in a sheaf are equal $\dsection^{a} = \dsection^{b}$ when they evaluate to the same values $\dsection^{a}\vert_{\openset_i} =  \dsection^{b}\vert_{\openset_i}$ over the open cover $\bigcup_{i\in I} \openset_{i} \subset \openset$ (indexed by $I$).
+  \item \textit{gluing} the union of sections defined over subspaces $\dsection^{i} \in \Gamma(\openset_i, \dtotal|_{\openset_i})$ is equivalent to a section defined over the whole space $\dsection\vert_{\openset_i} = \dsection^{i}$ for all $i\in I$ if all pairs of sections agree on overlaps $\dsection^{i}\vert_{\openset_i\cap\openset_j} =  \dsection^{j}\vert_{\openset_i\cap\openset_j}$
+  \end{itemize}
+\end{definition}
+
+The gluing axiom says that a distributed representation of a dataset, which is a set of local sections, is equivalent to a section over the union of the opensets of the local sections.
+The gluing axiom can also be used to generate the gluing rules used to construct non-trivial bundles from the set of trivial local sections.
+The locality axiom asserts that the glued section function is equivalent to a function over the union if they evaluate to the same values.
+
+\begin{figure}[H]
+  \includegraphics[width=1\columnwidth]{tex/sheaf_rules.pdf}
+  \caption{A sheaf has the conditions that sections are equal when they match on all subsets (\textit{locality}) and that the sections can be concatenated when they match on overlaps \textit{gluing}.\label{fig:atct:sheaf}}
+\end{figure}
+
+For example, in \autoref{fig:atct:sheaf}, the $\tau^{a}$ and $\tau^{b}$ $\sin$ sections are equal because they match \textit{locally} on all subsets.
+This is true whether $\sin$ is defined over parts ($\sin\vert_{\openset_i}$) or the whole space.
+If $\sin$ is defined over parts, then those parts can be \textit{glued} together.
+The concatenated $\sin$ is continuous because the pieces of the section outside the overlap are continuous with the pieces inside the overlap.
+The glued $\sin$ is also equal to the non-glued $\sin$ because they match on the opensets; therefore they are equivalent representations of the same section $\sin$ and so have the same mathematical properties.
+
+\subsection{Germ}
+While each section of a sheaf is evaluated over a point $\dsection(\dbasepoint)$ such that it returns a single record, the sheaf model also provides an abstraction when neighboring information is required.
+The sheaf over a very small region surrounding a point $\dbasepoint$ is called a \textit{stalk}\cite{harder2008lectures}
+\begin{equation}
+  \label{eq:atct:sheaf:stalk}
+    \sheaf_{\dbase, \dtotalc}\restriction_{\dbasepoint}\coloneqq \lim\limits_{\openset\ni \dbasepoint} \Gamma(\openset, \dtotal\restriction_{\openset})
+\end{equation}
+where the fiber is contained inside the stalk  $\dfiber_{\dbasepoint} \subset  \sheaf_{\dbase, \dtotal}\restriction_{\dbasepoint}$.
+The \textit{germ} is the section evaluated at a point in the stalk  $\dsection(\dbasepoint) \in \sheaf_{\dbase, \dtotal}\restriction_{\dbasepoint}$.
+Since the stalk includes the values near the limit of the point at \dbasepoint\, the germ can be used to compute mathematical derivatives.
 

### 2025-07-29 short_paper/main.tex
accessviz submission

@@
 \author{
-Hanna Aizenman\thanks{e-mail: haizenman@gradcenter.cuny.edu} \\
+Hannah Aizenman\thanks{e-mail: haizenman@gradcenter.cuny.edu} \\
 \scriptsize{CUNY Graduate Center}
@@
     The contract data visualization tools make with their users is that a chart is a faithful and accurate visual representation of the numbers it is made from.
-Motivated by wanting to make better tools, we propose a methodology for fully specifying arbitrary data to visualization mappings in a manner that easily translates to code.
+Motivated by wanting to make better tools, we propose a methodology for fully specifying arbitrary data to visualization mappings in a manner that easily translates to code and provides a consistent interface for accessibility tools to query for information about the state of the visualization.
 We propose that fiber bundles provide a uniform interface for describing a variety of underlying data - tables, images, networks, etc.

### 2025-09-05 slides/main.tex
added new fiber discussion

@@
 
-\graphicspath{.figures/, ../papers/figures/}
+\graphicspath{.figures/, ../paper/figures/}
 
@@
 
-\bibliography{references}
+\bibliography{../paper/references}
 
@@
 
-\begin{frame}{Field types }
+\begin{frame}{Fiber space: Fields}
+\begin{equation*}
+    \resizebox{\textwidth}{!}{
+    \begin{tabular}{|c|c|c|}
+      \hline
+       types and schema & \textcolor{fiber}{record} & \textcolor{section}{table}\\
+      \hline
+      \begin{tikzcd}[ampersand replacement=\&, sep=huge]
+        \textcolor{total}{\mathcal{U}_\sigma} \arrow[d, "\pi_{\sigma}"', color=total, two heads] \arrow[r, dotted] \arrow[rd, dotted] \& \textcolor{total}{\mathcal{U}} \arrow[d, "\pi"', color=total, two heads] \\
+        \textcolor{base}{C} \arrow[r, "\sigma"']                                                                                   \& \textcolor{base}{\texttt{DT}}
+      \end{tikzcd}
+    &
+    \begin{tikzcd}[ampersand replacement=\&, sep=huge]
+      \textcolor{total}{\mathcal{U}_{\sigma}} \arrow[d, "\pi_{\sigma}"', color=total, two heads] \\
+      \textcolor{base}{C} \arrow[u, "\delement"', dashed, bend right, color=fiber]
+      \end{tikzcd}
+    &
+    \begin{tikzcd}[ampersand replacement=\&, sep=huge]
+      \dfiberc\coloneqq \textcolor{total}{\Gamma^{\pi}(\sigma)} \arrow[d, "\pi"', color=total, two heads] \\
+      \dbasec \arrow[u, "\dsectionc"', dashed, bend right, color=section]
+      \end{tikzcd}\\
+      \hline
+    \end{tabular}
+    }
+\end{equation*}
+
+\begin{description}
+    \item[$\textcolor{total}{\mathcal{U}}$] Space of all possible values, e.g. $\mathbb{R} \times \mathbb{R}^{+}$
+    \item[$\textcolor{base}{C}$] Field names, e.g. $TEMP, PRCP$
+    \item[$\textcolor{base}{\texttt{DT}}$] Data Type, e.g. $\texttt{float}$, $\texttt{PosFloat}$
+\end{description} 
+\end{frame}
+
+\begin{frame}{Types and Schema }
    A \textit{schema} consists of a pair $(\fnames, \sigma)$ where $\fnames$ is the set of field names and $\sigma: \fnames \rightarrow \textbf{DT}$ is a function from field name to field data type\cite{spivakSimplicialDatabases2009}
@@
 \begin{figure}
-    \includegraphics[width=\linewidth]{../paper/figures/artist_equiv.png}
+    \includegraphics[scale=.95, keepaspectratio]{../paper/figures/artist_equiv.pdf}
 \end{figure}
@@
     \begin{figure}
-        \includegraphics[width=\linewidth]{../paper/figures/composition.pdf}
+        \includegraphics[scale=.95, keepaspectratio]{../paper/figures/composition.pdf}
     \end{figure}

### 2025-11-27 paper/main.tex
accessviz submission
fold back in equivariance and homomorphism

@@
 
-\usepackage{silence}
-\WarningsOff[fixltx2e]
+%\usepackage{silence}
+%\WarningsOff[fixltx2e]
 
@@
 %% with other math-related packages, you may want to disable it.
+\usepackage{mathptmx}
 \usepackage{amsmath, amsthm, amsfonts,amssymb,eulervm,xspace, mathtools}
@@
 
-\usepackage[]{minted}
+\usepackage[cache=false]{minted}
 \setminted[python]{fontsize=\scriptsize,
@@
 
-We propose that we can better enforce this expectation in software by providing a uniform way of expressing $data$ and $graphic$ using their respective algebraic structure and by uniformly specifying the behaviors and properties of those structures and the maps between them using category theory.
+We propose that structures from category theory and algebraic topology can be used as a sort of psuedo-code for uniformly specifying the behaviors and properties of $data$ and $graphic$ and rigorously expressing the structure preserving constraints on maps between them.
 For example, our framework can encapsulate how a table and scatter plot and heatmap are different representations of the same data and track an observation from a data cube as a point along a time series and on a map and in a network.
-The algebraic structures can then be translated into programmatic types, while the categorical descriptions translate to a functional design framework.
-Strong typing and function composition enable visualization software developers to build complex components from simpler verifiable parts \cite{huHowFunctionalProgramming2015, hughesWhyFunctionalProgramming1989}.
+By design, this psuedo-code translates to strong programmatic types and a functional design framework, which enable visualization software developers to build complex components from simpler verifiable parts \cite{huHowFunctionalProgramming2015, hughesWhyFunctionalProgramming1989}.
 These components can be built as a standalone library and integrated into existing libraries and we hope these ideas will influence the architecture of critical data visualization libraries, such as Matplotlib.
@@
 
-\begin{LaTeXdescription}
+\begin{description}
   \item[\textcolor{fiber}{\textbf{field}}]\footnote{In this paper, color in definitions, equations, and visualizations is used to group conceptually related terms\cite{headMathAugmentationHow2022}, for example \textcolor{fiber}{field} and \textcolor{fiber}{fiber}.} is a set of values of the same type, e.g.
@@
   \item[\textcolor{base}{\textbf{topology}}] is the connectivity and relative positioning of elements in a dataset \cite{wilkinsonGrammarGraphics2005}.
-\end{LaTeXdescription}
+\end{description}
 
@@
 
-Data is often described by its mathematical structure, for example the Steven's measurement scales define nominal, ordinal, interval, and ratio data by the allowed operations on each \cite{stevensTheoryScalesMeasurement1946} and other researchers have since expanded the scales to encapsulate more types of structure \cite{leaFormalizationMeasurementScale1971, thomasMathematizationNotMeasurement2014}.
-Loosely we define structure as a mathematical space $X$ and the actions on that space; our model is purposely generalizable to many different types of spaces (e.g sets, metric spaces, etc).
-Steven's conceptualizes the scales as mathematical groups with \textcolor{action}{actions}.
-A function that preserves structure when the input or output is changed by a group action is called \textit{equivariant}; a visualization is structure preserving when there exist compatible group actions on the data and visualization was proposed by Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}.
+Data is often described by its mathematical structure, for example the Steven's measurement scales define nominal, ordinal, interval, and ratio data by the transformations allowed on the data \cite{stevensTheoryScalesMeasurement1946}.
+Other researchers have since expanded the scales to encapsulate more types of structure \cite{leaFormalizationMeasurementScale1971, thomasMathematizationNotMeasurement2014}, and in \autoref{sec:atct:fb:fiber} we generalize further by proposing that data can be encoded as mathematical categories.
+
+Loosely, structure is defined as a collection of values $X$ and transformations $G$ on those values. These transformations can be operations, relations, or generalized as actions:
+\begin{definition}\label{def:related-work:action}
+  An \textcolor{action}{\textbf{action}} is a function  $act: \textcolor{action}{G} \times X \rightarrow X$.
+An action has the properties of identity $act(\textcolor{action}{e}, x) = x$ for all  $x \in X$ and associativity $act(\textcolor{action}{g}, act(\textcolor{action}{f}, x)) = act(\textcolor{action}{f} \circ \textcolor{action}{g}, x)$ for $\textcolor{action}{f},\textcolor{action}{g} \in \textcolor{action}{G}$.\cite{grimaldiDiscreteCombinatorialMathematics2006}
+\end{definition}
+Elements of $X$ can be from one data field or all of them or some subset; similarly the actions act on the elements of $X$ and each action can be a composition of actions.
+
+A function that preserves structure when the input or output is changed by a group\footnote{A mathematical {group} is a set with an associative binary operator.
+This operation must have an identity element and be closed, associative, and invertible, consisting of a set of values $X$ and \textcolor{action}{actions} on the set {$G = (G,\circ, e)$}.} action is called \textit{equivariant}.
+Given a group $G$ that acts on both $X$ and $Y$,
+\begin{definition}\label{def:equivariance}
+ a function $f: X \rightarrow Y$ is \textbf{equivariant} when $f(act(g,x)) = act(g,f(x))$ for all $g$ in $G$ and for all $x$ in $X$ \cite{pittsNominalSetsNames2013}
+\end{definition}
+This means that a visualization is structure preserving when there exist compatible group actions on the data and visualization, as proposed by Kindlemann and Scheidegger\cite{kindlmannAlgebraicProcessVisualization2014}.
+
 Although the Steven's scales were conceptualized as having group structure, the ordinal scale has a monoidal structure because partial orders ($\geq, \leq$) are not invertible.
-This means \textit{equivariance} cannot be used to test for structure preservation; instead \textit{homomorphism} can be used to verify whether map preserve relations between data elements, as proposed by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}.
-These tests are mutually exclusive; a function can be homomorphic but not equivariant, such as an exponential encoding, or equivariant but not homomorphic, such as the inverse encoding.
-A function can also be homomorphic (or equivariant) with respect to one action but not with respect to another.
-The encoding transforms in visualization tools are expected to preserve the structure of whatever input they receive; therefore a methodology for codifying arbitrary structure is presented in \autoref{sec:atct:fb:fiber} and \autoref{sec:artist:equivariant:data} presents a generalization of equivariance and homomorphism for evaluating structure preservation.
+This means \textit{equivariance} cannot be used to test for structure preservation; instead \textit{homomorphism} can be used.
+Given the function $f: X \rightarrow Y$, with operators $(X, \circ)$ and $(Y, *)$
+\begin{definition}\label{def:homomorphism}
+  A function $f$ is \textbf{homomorphic} when $f(x_1 \circ x_2) = f(x_1) * f(x_2)$ and preserves identities $f(I_x) = I_y$ for all $x, y \in X$ \cite{grimaldiDiscreteCombinatorialMathematics2006}
+\end{definition}
+This means that a visualization is structure preserving when there
+exists compatible operators (here $\circ$ and $*$) such that the relations between data elements is preserved in the mapping, 
+as proposed by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}.
+
+\begin{figure}[H]
+  \includegraphics[width=1\columnwidth]{equiv_homo.pdf}
+  \caption{Encoding data as the bar height using an exponential transform is not equivariant because encoding the data and then scaling the bar heights yields a much taller graph then scaling the data and then encoding those heights using the same exponential transform function.
+Encoding data as bar height  using an inverse transform is not homomorphic because the largest number is mapped to the smallest bar while the max function returns the largest bar.}
+  \label{fig:related-work:field-preservation}
+\end{figure}
+
+As illustrated in \autoref{fig:related-work:field-preservation}, a visualization is structure preserving when the functions commute such that the graphic is the same whether the data is acted on ($
+\dfunct$) and then mapped ($A$) to a graphic or whether the data is mapped to a graphic that is then modified in a compatible way.
+As shown in \autoref{fig:related-work:field-preservation}, a function can be homomorphic but not equivariant, such as an exponential encoding, or equivariant but not homomorphic, such as the inverse encoding.
+As shown, a function can also be homomorphic (or equivariant) with respect to one action but not with respect to another.
+We generalize the expectation of commutativity to codify structure preservation in \autoref{sec:artist:equivariant:data}.
 
@@
 
-\begin{LaTeXdescription}
+\begin{description}
   \item[pan, zoom, scroll] sheaf: locality + gluing \autoref{def:atct:sheaf}
@@
   \item[brushing, linking, annotation] composition of artists \autoref{eq:artist:join_base_fiber}
-\end{LaTeXdescription}
+\end{description}
 
@@
 
-Refer to the instructions for this section (\cref{sec:supplement_inst}).
-Below is an example you can follow that includes the actual supplemental material for this template:
-
-All supplemental materials are available on OSF at \url{https://doi.org/10.17605/OSF.IO/2NBSG}, released under a CC BY 4.0 license.
-In particular, they include (1) Excel files containing the data for and analyses for creating \cref{tab:vis_papers} and \cref{fig:vis_papers}, (2) figure images in multiple formats, and (3) a full version of this paper with all appendices.
-Our other code is intellectual property of a corporation---Starbucks Research---and there is no feasible way to share it publicly.
-
-\appendices
+\acknowledgments{
+The authors would like to thank the anonymous reviewers who gave constructive feedback on an earlier version of this paper.
+The authors are also grateful to  the various Matplotlib and Napari contributors, particularly Juan Nunez-Iglesias, and Nicolas Kruchten for their valuable feedback from the library developer perspective.
+
+Hannah is also very grateful to Nicolas for the suggestion of augmented notation and to the nlab and wikipedia contributors who wrote clear explanations of many of the topics discussed in this paper.
+
+This project has been made possible in part by grant number 2019-207333 and \note{Cycle 3 grant number} Chan Zuckerberg Initiative DAF, an advised fund of Silicon Valley Community Foundation
+}
+
+% trigger a \newpage just before the given reference
+% number - used to balance the columns on the last page
+% adjust value as needed - may need to be readjusted if
+% the document is modified later
+%\IEEEtriggeratref{8}
+% The "triggered" command can be changed if desired:
+%\IEEEtriggercmd{\enlargethispage{-5in}}
+
+% references section
+\bibliographystyle{abbrv-doi}
+\bibliography{references}
+
+\appendix
 
 \section{Summary}
-\label{sec:appndix:summary}
+\label{sec:appendix:summary}
 The topological spaces and functions introduced throughout this paper are summarized here for reference.
@@
 
-\section*{Acknowledgment}
-\note{Acknowledge all the actual people}
-The authors would like to thank the anonymous reviewers who gave constructive feedback on an earlier version of this paper.
-The authors are also grateful to  the various Matplotlib and Napari contributors, particularly Juan Nunez-Iglesias, and Nicolas Kruchten for their valuable feedback from the library developer perspective.
-
-Hannah is also very grateful to Nicolas for the suggestion of augmented notation and to the nlab and wikipedia contributors who wrote clear explanations of many of the topics discussed in this paper.
-
-This project has been made possible in part by grant number 2019-207333 and \note{Cycle 3 grant number} Chan Zuckerberg Initiative DAF, an advised fund of Silicon Valley Community Foundation
-
-% Can use something like this to put references on a page
-% by themselves when using endfloat and the captionsoff option.
-\ifCLASSOPTIONcaptionsoff
-  \newpage
-\fi
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-% references section
-\bibliographystyle{abbrv-doi}
-\bibliography{references}
 \end{document}

### 2025-11-27 paper/supplement.tex
fold back in equivariance and homomorphism

@@
 % Fonts
-\usepackage{times}                     % we use Times as the main font
+% Fonts: use T1 encoding and a Times-like text+math package for pdfLaTeX
+\usepackage[T1]{fontenc}
+\usepackage{newtxtext,newtxmath}
 \renewcommand*\ttdefault{txtt}         % a nicer typewriter font
@@
 \usepackage{mathrsfs} % math script fonts
-\usepackage{mathptmx}
-\DeclareMathDelimiter{(}{\mathopen} {operators}{"28}{largesymbols}{"00}
-\DeclareMathDelimiter{)}{\mathclose}{operators}{"29}{largesymbols}{"01}
 % *** SUBFIGURE PACKAGES ***
@@
 \subsection{Natural transforms}
+\begin{equation*} {\renewcommand{\arraystretch}{1.2}
+  \begin{tabular}{|c|c|}
+    \hline
+    $\mathcal{C}$ & $\mathcal{D}$ \\
+    \hline
+    \begin{tikzcd}[ampersand replacement=\&, sep=huge]
+        \textcolor{base}{c} \arrow[d, "f", color=action] \\ \textcolor{base}{c^{\prime}}      
+    \end{tikzcd}
+    & 
+    \begin{tikzcd}[ampersand replacement=\&, sep=huge]
+        \textcolor{sheaf}{F(c)} \arrow[d, "F(f)", color=action] \arrow[r, "\alpha_{c}", color=artist] \& 
+        \textcolor{sheaf}{G(c)} \arrow[d, "G(f)", color=action] \\
+        \textcolor{sheaf}{F(c^{\prime})} \arrow[r, "\alpha_{c^{\prime}}", color=artist] \& \textcolor{sheaf}{G(c^{\prime})} 
+    \end{tikzcd}\\
+    \hline
+  \end{tabular}}
+\end{equation*}
+
 \begin{definition}\label{def:natural-transform}
-  Given two functors $F, G: \mathcal{C}\rightarrow \mathcal{D}$, a \textbf{natural transformation} $\alpha: F \rightarrow G$ is a function which assigns to each object $c$ of $\mathcal{C}$ a morphism $\alpha_c:F(input) \rightarrow G(c), G(c) \in \mathcal{D}$, in such a way that for every morphism $f:c \rightarrow c^\prime, c^\prime \in \mathcal{C}$, the morphisms in $\mathcal{D}$ commute such that $\alpha_c^{\prime}(F(f)(F(c))) = G(f)(\alpha_c(F(c))$.
-When this holds, $\alpha_{c}$ is \textit{natural} in $c$.\cite{maclaneCategoriesWorkingMathematician2013}.
+  Given two functors $F, G: \mathcal{C}\rightarrow \mathcal{D}$, a \textbf{natural transformation} $\alpha: F \rightarrow G$ is a function which assigns to each object $c$ of $\mathcal{C}$ a morphism $\alpha_c:F(c) \rightarrow G(c)$ of $\mathcal{D}$, in such a way that for every morphism $f:c \rightarrow c^\prime$ in $\mathcal{C}$, the morphisms $F(f), G(f)$ in $\mathcal{D}$ commute such that $\alpha_c^{\prime}(F(f)(F(c))) = G(f)(\alpha_c(F(c)))$.
+When this holds, $\alpha_{c}$ can be said to be \textit{natural} in $\mathcal{C}$.\cite{maclaneCategoriesWorkingMathematician2013}.
 \end{definition}.
@@
 \subsection{Presheaf}
+
 \begin{definition}\label{def:atct:presheaf}
@@
 \end{definition}
+
+\begin{equation*}
+\begin{tikzcd}
+\cgamma{\opensetc_2}{\dtotalc|_{\opensetc_2}} \arrow[r, "
+\hat{\phi}^*", color=action, dashed] & \cgamma{\opensetc_1}{\dtotalc|_{\opensetc_1}}                \\
+\opensetc_2 \arrow[u, "\sheaf_{2}", maps to, color=sheaf]                  & \opensetc_1 \arrow[l, "\hat{
+  \phi}"', color=action] \arrow[u, "\sheaf_{1}", maps to, color=sheaf]
+\end{tikzcd}
+\end{equation*}
 
@@
   \label{eq:atct:sheaf:stalk}
-    \sheaf_{\dbase, \dtotalc}\restriction_{\dbasepoint}\coloneqq \lim\limits_{\openset\ni \dbasepoint} \Gamma(\openset, \dtotal\restriction_{\openset})
+    \sheafc_{\dbasec, \dtotalc}\restriction_{\dbasepointc}\coloneqq \lim\limits_{\opensetc\ni \dbasepointc} \Gamma(\opensetc, \dtotalc\restriction_{\opensetc})
 \end{equation}

### 2025-11-27 paper/figcode/homeomorphism_helpers.py

@@ -162,11 +162,12 @@ def make_ts(df, ax):
     ts = get_timeseries(df, airports['t_lat'], airports['t_lon'])
     
     ln,  = ax.plot('DATE_DTYPE', "PRCP", data=ts, color = airports['time_color'])
-    ax.axvline(pd.to_datetime(airports['point_date']), color=airports['point_color'], alpha=.75, zorder=-1)
+    point_date = pd.to_datetime(airports['point_date'])
+    ax.axvline(point_date, color=airports['point_color'], alpha=.75, zorder=-1)
     ax.set_xlim(xmin=ts['DATE_DTYPE'].min(), xmax=ts['DATE_DTYPE'].max())
     ax.xaxis.set_major_locator(mdates.MonthLocator())
     ax.xaxis.set_major_formatter(mdates.DateFormatter("%b"))
-    ax.annotate("2002", (1, -0.1), xycoords='axes fraction', 
+    ax.annotate(point_date.year, (1, -0.1), xycoords='axes fraction', 
                 fontsize=fs['footnote'], va='top', ha='center')
     return ln 
     

### 2025-11-27 paper/main.tex
homemeomorphism

@@
 While plotting the time series as a continuous line would be correct, it would be incorrect for a visualization to indicate that the distinct rows or stations are connected in a 1D continuous manner because it introduces ambiguity over which part of the line maps back to the data.
-A map that by definition has continuous maps between the input and output spaces, such as data and graphics, is called a \textit{homeomorphism}\cite{riehlCategoryTheoryContext}.
+This ambiguity cannot exist when the map between data and graphic is a homeomorphism
+\begin{definition}
+  A function $f$ is a \textit{homeomorphism} if it is bijective, continuous, and has a continuous inverse function $f^{-1}$.
+\end{definition}
+which means that every element of the graphic maps back to a data element.
 The bar plot, line plot, and heatmap in \autoref{fig:related-work:continuity:ktypes} have a homeomorphic relationship to the 0D ($\bullet$) points, 1D (--) linear, and 2D ($\blacksquare$) surface continuities embedded in the continuous 3 dimensional surface encapsulating time and position because each point of the visualization maps back into a point in its corresponding indexing space in the cube.

### 2025-11-27 paper/main.tex
fb & sheaves

@@
 While plotting the time series as a continuous line would be correct, it would be incorrect for a visualization to indicate that the distinct rows or stations are connected in a 1D continuous manner because it introduces ambiguity over which part of the line maps back to the data.
-This ambiguity cannot exist when the map between data and graphic is a homeomorphism
+This ambiguity cannot exist when the map between data and graphic is a homeomorphism,
 \begin{definition}
@@
 This is in contrast to Wilkinson modeling the continuity as metric spaces \cite{wilkinsonGrammarGraphics2005}such that a distance function is a fixed property of the continuity; in our model it distance is a property of a field over that continuity.
-
 We sketch out fiber bundles in \autoref{sec:atct:fiber-bundles}, but Butler provides a thorough introduction to bundles for visualization practitioners.
@@
 
+\begin{definition}\label{def:fiber_bundle}
+   A \textbf{fiber bundle} $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and  bundle projection map $\pi: \dtotalc \rightarrow \dbasec$ \cite{spanier1989algebraic}.
+
+   \begin{equation} \label{eq:atct:fb:intro}
+    \begin{tikzcd}
+      \dfiberc \arrow[r, hook, color=total] & \dtotalc \arrow[r, "\pi", color=total, two heads] & \dbasec
+      \end{tikzcd}
+    \end{equation}
+
+A continuous surjective map $\bm{\pi}$ is a \textbf{bundle projection} map when
+\begin{enumerate}
+  \item all fibers in the bundle are isomorphic.
+Since all fibers are isomorphic $\dfiber \cong \dfiber_{\dbasepoint}$ for all points $\dbasepoint \in \dbase$, there is a uniquely determined \textcolor{fiber}{fiber space} \dfiberc\ given by the preimage of the projection $\pi$ at any point $\dbasepoint$ in the \textcolor{base}{base space} \dbasec: $\dfiberc = \pi^{-1}(k)$.
+  \item each point $\dbasepointc$ in the \textcolor{base}{base space} \dbasec\ has an open neighborhood $\openset_{\dbasepointc}$ such that the \textcolor{total}{total space} \dtotalc\ over the neighborhood is locally trivial.
+\end{enumerate}
+\end{definition}
+
+\textbf{Local triviality} means $\dtotal\vert_{\openset} = \openset\times \dfiber$.
+In this paper we use $\dtotal\vert_{\openset} = \pi^{-1}(\openset)$ to denote the preimage of an openset\footnote{Open sets (open subsets) are a generalization of open intervals to n dimensional spaces.
+For example, an open ball is the set of points inside the ball and excludes points on the surface of the ball.
+\cite{weissteinOpenSet,bradleyTopologyVsTopology}}, and a \textbf{local trivialization} is a specific choice of neighborhoods (described in \autoref{sec:atct:fb:base}) and their preimages such that the fibers in each preimage are identical $\dfiber = \dfiber_{\dbasepoint}$ for all points $\dbasepoint \in \openset$.
+All fiber bundles can be decomposed into sets of local trivializations that are also bundles and we can specify a gluing scheme that reconstructs the fiber bundle from locally trivial pieces by specifying\textbf{transition maps} for all overlaps of the local trivializations; therefore, while the framework in this paper applies to all bundles, in this paper we assume that the bundles are trivial bundles $\dtotal = \dbase \times \dfiber$ so that we can assign all fibers in a bundle the same type.
+
+\begin{minipage}{.5\columnwidth}
+\begin{definition}\label{def:fiber_bundle:section}
+A \textcolor{section}{\textbf{section}} $\dsectionc: \dbasec \rightarrow \dtotalc$ over a fiber bundle is a smooth right inverse of $\pi(\dsection(\dbasepoint)) = \dbasepoint$ for all $\dbasepoint \in \dbase$
+\end{definition}
+\end{minipage}
+\begin{minipage}{.4\columnwidth}
+  \begin{equation} \label{eq:atct:fb:intro-sec}
+    \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
+     \dfiberc
+      \arrow[r, hook, color=total] \&
+      \dtotalc
+      \arrow[d, "\pi"',color=total, two heads] \\
+       \&
+    \dbasec
+       \arrow[u, "\dsectionc"', bend right, pos=.5, color=section, dashed]
+    \end{tikzcd}
+  \end{equation}
+\end{minipage}
+
+We propose that the total space of a bundle can encode the mathematical space in which a  dataset is embedded, the base space can encode the topological properties of the dataset, the fiber space can encode the data types of the fields of the dataset, and that the datasets can be encoded as section functions from the continuity to the fiber space.
+
+\begin{figure}[H]
+       \includegraphics[width=1\columnwidth]{fb.pdf}
+       \caption{The space of all data values encoded by this fiber bundle can be modeled as a \textcolor{total}{rectangle} total space.
+Each dataset in this data space lies along the interval \textcolor{base}{$[0,2\pi]$} base space.
+Each dataset has values along the \textcolor{fiber}{$-1 \rightarrow 1$} interval fiber.
+One dataset embedded in this total space is the \textcolor{section}{sin} section over the bundle.}\label{fig:atct:fb}
+  \end{figure}
+
+For example, the fiber bundle in \autoref{fig:atct:fb} encodes the space of all continuous functions that have a domain of $[0, 2\pi]$ and range $[0,1]$.
+Using a fiber bundle abstraction encodes that the dataset has a 1D linear continuity as the base space \dbase is the interval $[0,2\pi]$ and a field type that is a \texttt{float} in the range $[0,1]$.
+Therefore the type signature of the datasets in this fiber bundle, which is called a section \dsection, would be \texttt{dataset: $[0, 2\pi] \rightarrow [0,1]$}.
+One such dataset (section) is the $\sin$ function, which as shown in \autoref{fig:atct:fb} is defined via a function \dsection from a point in the base space to a corresponding point in the fiber.
+Evaluating the section function over the entire base space yields the $\sin$ curve that is composed of points intersecting each fiber over the corresponding point.
+The local trivializations shown in \autoref{fig:atct:fb} are one way of decomposing the total bundle and conversely the bundle can be constructed from the local trivializations $\dbase = \dbase_{0} \oplus \dbase_{1}$.
+As shown, the section $\sin$ spans the trivializations in the same manner that it spans the bundle; this is analogous to how a dataset may span multiple tables or be collected in one table.
+The trivializations are glued together into the bundle at the overlapping region $\frac{2\pi}{5}$ by defining the transition map $\dfiber_{1} \rightarrow \dfiber_{2}$.
+Because the fibers in \autoref{fig:atct:fb} at $\frac{2\pi}{5}$ are aligned, the transition map is an identity map that take every value in $\dfiber_1$ and maps it to the same value in $\dfiber_2$ so that the sections, such as $\sin$, remain continuous.
+
+\paragraph{Trivial and non-trivial bundles}
+\label{sec:appendix:bundle_triviality}
+
+Generally, the distinguishing factor between a trivial bundle and a non-trivial bundle are how they are decomposed into local trivializations:
+\begin{description}
+  \item[\textit{trivial bundle}] is directly isomorphic to $\dbasec\times\dfiberc$.
+For any choice of cover of $\dbasec$ by overlapping opensets, we can choose local trivializations such that all transition maps are identity maps.
+  \item[\textit{non-trivial bundle}] can not be constructed as $\dbasec\times\dfiberc$.
+For any choice of local trivializations, there is at least one transition map that is not an identity \cite{hatcherAlgebraicTopology2002}.
+\end{description}
+
+\begin{figure}[H]
+  \includegraphics[width=1\columnwidth]{transition_maps.pdf}
+  \caption{The cylinder is a trivial fiber bundle; therefore it can be decomposed into local trivializations that only need identity
+  maps to glue the trivializations together.
+The mobius band is a non-trivial bundle; therefore it can only be decomposed into trivializations where at least one transition map is not an identity map.
+}\label{fig:cyl_mob_bundles}
+\end{figure}
+
+In the example in Figure~\ref{fig:cyl_mob_bundles}, we use arrows $\uparrow$ to denote fiber alignments.
+In the cylinder case the fibers all point in the same direction, which illustrates that they are equal $\uparrow=\uparrow$.
+In the Möbius band case, while the fibers in an arbitrary local trivialization are equal $\uparrow=\uparrow$, the fibers at the twist are unequal but isomorphic $\uparrow \cong \downarrow$.
+The cylinder and mobius band can be decomposed to the same local trivializations, for example the fiber bundles in \autoref{fig:atct:fb} In the cylinder case, the fibers in the overlapping regions of the trivializations are equal $\dfiber_0\restriction_{\openset_1\cap\openset_2} = \dfiber_{1}\restriction_{\openset_1\cap\openset_2}$; therefore the transition maps at both intersections map the values in the fiber to themselves $\delement\rightarrow\delement$ .
+In the Möbius band case, while $\dfiber_0\restriction_{(2\pi/5-\varepsilon, 2\pi/5+\varepsilon)}\to\dfiber_1\restriction_{(2\pi/5-\varepsilon, 2\pi/5+\varepsilon)}$ can be chosen to be an identity map, the other transition map component $\dfiber_0\restriction_{(-\varepsilon,\varepsilon)}\to\dfiber_1\restriction_{(-\varepsilon,\varepsilon)}$ has to flip any section values.
+For example given  $\dfiber_{0}=\uparrow$ and $\dfiber_{1}=\downarrow$, the transition map $\delement \mapsto -\delement$ maps each point from one fiber to the other $\uparrow \mapsto \downarrow$ such that any sections remain continuous even though the fibers point in opposite directions.
+
 \subsubsection{\textcolor{base}{Topological Structure: Base Space \dbase}}
@@
 There are various equivalent definitions of topology, but here we use the neighborhood axiomatization because it is most analogous to the data access model of index (element) in subset (neighborhood) of all indices (mathematical space).
-Roughly, a neighborhood of an element $x$ has to contain $x$ (1), can grow (2), or shrink (3), and every neighborhood also contains smaller neighborhoods of points adjacent to x (4).
+Given a set $X$ and a function $\mathcal{N}:X\to 2^{2^X}$ that assigns to any $x\in X$ a non-empty collection of subsets $\mathcal{N}(x)$, where each element of $\mathcal{N}(x)$ is a \emph{neighborhood of $x$}, then $X$ with  $\mathcal{N}$ is a \textcolor{base}{topological space} and $\mathcal{N}$ is a neighborhood \emph{topology} if for each $x$ in $X$: \cite{brownronaldTopologyGroupoids2006}
+
+\begin{definition}\label{def:topology}
+\begin{enumerate}
+  \item if $N$ is a neighborhood $N \in \mathcal{N}(x)$ of $x$ then $x \in N$
+  \item every superset of a neighborhood of $x$ is a neighborhood of $x$; therefore a union of a neighborhood and adjacent points in $X$ is also a neighborhood of $x$
+  \item the intersection of any two neighborhoods of $x$ is a neighborhood of $x$
+  \item any neighborhood $N$ of $x$ contains a neighborhood $M \subset N$ of $x$ such that $N$ is a neighborhood of each of the points in $M$
+\end{enumerate}
+\end{definition}
+Therefore, a neighborhood of an element $x$ has to contain $x$ (1), can grow (2), or shrink (3), and every neighborhood also contains smaller neighborhoods of points adjacent to x (4).
 For example, in the indexing cube in \autoref{fig:fig:related-work:continuity:ktypes}, the brown surface and blue rectangle are both neighborhoods of the index for the measurement in Albany on June 09.
